### 2008-08-05 writing/outline.tex
first commit of the master paaper

\documentclass[a4paper,10pt]{IEEEconf}
\usepackage[latin1]{inputenc}
\usepackage[american]{babel}
%\usepackage[T1]{fontenc}
\usepackage[dvips,pdftex]{graphicx}
%\usepackage{multicol}
\usepackage{subfigure}
\usepackage{named}
%\usepackage{fancybox}
%\usepackage{natbib}
\bibliographystyle{named}
\usepackage[dvips,pdftex]{hyperref}
%opening
\title{ Symbol Recognition using Particle Swarm Algorithms}
%\title{A Study on Using Particle Swarm Algorithms in Symbol Recognition }
\author{Maha El Meseery, Dr. Mahmoud Fakhr El Din, Dr. Samyah Mashaly,
\\ Dr. Magdah Fayek, Dr. Nevin Darwish}
\begin{document}
\maketitle
\begin{abstract}
Sketch recognition is defined by the process of identifying the symbols the user draws using a single or multiple strokes. Users draw strokes using a pen and the system immediately interprets their strokes into objects that can be easily manipulated. This paper uses Particle Swarm Algorithm(PSO) to divide strokes a user draws into meaningful geometric primitives. These geometric primitives are grouped to formulate symbols which will be further identified. The results showed that using PSO improves segmentation results which guide the symbol recognition phase. Final recognition have improved after the PSO algorithm using a SVM classifier.  
 %Hand drawn sketches are the natural method of exchanging thoughts between engineers and humans in general. Nevertheless, researchers barely give attention to improve current gesture and sketch recognition systems. In a sketch understanding system, users need to feel the freedom of paper and pencil along with the flexibility of computer manipulation. 
\end{abstract}
\section{Introduction}
%A picture speak a thousand words is also true for sketches. 
Scientists in general and engineers specifically express thoughts and designs using sketches. Engineers use sketches to exchange designs as a natural method of communication rather than writing or speaking. Currently, engineers use paper and pencil design in early stages of design. Design engineers need a powerful symbol recognition and manipulation system to help them operate as freely as with papers. Increasing the interaction between the computer and the users in sketch and CAD systems was recently the reason for the emerging of few advanced sketch recognition systems. Sketch recognition system converts the drawings and informal graphs the user draws into objects and symbols can be used in further interaction with the system. \\
Most researchers in sketch recognition divided the system into four main steps preprocessing, segmentation, symbol recognition and finally sketch understanding. The preprocessing process saves the location of points from the hardware and collects basic information about the stroke data. After removing noise and calculating primary data the system proceeds to divide the stroke into a set of simple geometrical primitives.  The second step of the system is clustering strokes and segments to formulate symbols that can be recognized by the system. Finally, priori information about the sketch is used to support identification of symbols by excluding any possible invalid symbols. \\
%the basic symbols are identified in consistent with the whole sketch. \\
%from the user information from the users converts the strokes the user draw into segments which is recognized into basic symbols that can be easily manipulated. Finally, the basic symbols are interpreted in consistent with the whole sketch. \\
The next section will relate previous work in the field. The remaining of the paper will be as following; Section \ref{PSO} will explain the general particle swarm algorithm. Section \ref{Sysdisc} will describe the system block diagram and then section \ref{Prepross} will explain the preprocessing step, section \ref{seg} will describe the segmentation algorithm in details. Finally, section \ref{sec:Experiments} will describe the experiments that was preformed.\\
\section{Related Work}
%%%What is sketch recognitions. Why we need sketch recognition systems. 
The field of symbol recognition has gained interest in the last few years. A wide variety of techniques were used either on segmentation process or symbol recognition.\\ 
To segment strokes, \cite{meanshift10} introduced a feature area for each primitive and then compute the segmentation error for different types of primitives based on this area. %\citeauthor{HeloiseBeautification} used a dynamic programming algorithm to solve the segmentation problem \cite{HeloiseBeautification}. 
Complex and hybrid algorithm were also introduced as in  \cite{earlyprocess} where they generated different sets of segmentation based on both curvature and speed dominant points then they choose the segmentation with the least error from the hybrid set. A dynamic programming algorithm was also used to obtain optimum segmentation of the input strokes in \cite{GeometryAndDomain102}. Genetic algorithm was also used by \cite{CruveDivisionSwarm} to optimally divide digital curves into lines and curves. \cite{PolygonApproximationPSO} used PSO to convert digital curves into polygons.  \\
  %Some tried to detect dominant or corner points based on data generated from speed   \cite{earlyprocess} and mostly used curvature of the stroke drawn \cite{earlyprocess,meanshift10}.

%Particle swarm algorithms was used in \cite{} to fit a set of points into ellipse or higher order curves.
%The first set of features introduced to be used on gesture recognition was introduced by Rubine in \cite{gestureexample12}. later,  Spatial relations and features where used in  \cite{cognitivesketch18,SRGraph57}  . \citeauthor{GeometryAndDomain102} introduced ink density feature which they used.  The use of Zernike moments  was used  as the only features to the SVM classifiers in \cite{HeloiseBeautification}.  \cite{SketchRead2007} used a set of features varied from relative spatial features and some compositional features.  \\

%After segmentation researchers varies on the methods in which to continue.
Many different approaches have been investigated to achieve final recognition of symbols. Graph searching and template matching where used in \cite{mulitstroke5,GraphBased07}, where  a graph of the drawn symbol is generated and matched to stored set of graphs templates to reach a classification. Geometrical and spatial descriptive language was used in \cite{SketchRead2007,Ladder30} where a parser like architecture was used to recognize symbols  . \citeauthor{imagetrainable48} used trainable image template to classify different set of symbols \cite{imagetrainable48} .  The use of SVM and HMM classification was implemented in \cite{SVM300,SmartSketch56} and \cite{HMM53}.\\% Other techniques as  [          ] was used  in []. \\  
 %In the next section the particle swarm algorithm will be explained. After that the system block description and detailed algorithms will be described. We will then proceed to describe the experiments preformed and the achieved results.  
%\subsection{Sketch Recognitions systems}
%Details of Sketch systems. 
%Related work and research in this area. 
%Writing on the pros and cons of current systems.  
%Write on the difficulties of the sketch recognition systems. 
\subsection{Particle Swarm Algorithm}
\label{PSO}
%What is particle swarm algorithm and how it was used in related researches. 
In the Particle Swarm Algorithm (PSO) each agent represents a particle from the solution space. The agents move the particle with direction and velocity based on the general swarm equation [\ref{eq:Swarm}]. Equation [\ref{eq:Swarm}] shows how velocity and direction of each particle is computed where $r_1$ \& $r_2$ are random variable and $c_1$ \& $c_2$ are the swarm system variables. Each iteration the global best $g_{best}$ particle and  the agent local best $l_{best}$ particle are evaluated based on the maximum fitness functions of  all particles. The solution is found after a maximum number of iteration or after an error threshold is achieved. Equation \ref{eq:descrite} where t is a threshold and $r_{3}$  is a random variable, is used to change general swarm algorithm into binary particle which will handle particle values of either 0 or 1. .  
\begin{equation}
%\[
v_{ij}  = v_{ij}  + c_1 r_1 (lbest_{ij}  - p_{ij} ) + c_2 r_2 (gbest_{ij}  - p_{ij} )
%\]
\label{eq:Swarm}
\end{equation}
\begin{equation}
%\[
p_{ij}=p_{ij}+v_{ij},
%\]
\end{equation}
\begin{equation}
   P(i)\Leftarrow 
\{
\begin{array}{c} 
1 \quad \quad if\quad r_{3}>t  \\

0 \quad \quad if\quad r_{3}<t 
\label{eq:descrite}
\end{array}\}
\end{equation}
\section{System Description}
\label{Sysdisc}
%The stroke is the path of points from the instant the pen is down till it is up.
As shown in fig. \ref{fig:Blockdiagram} the system is divided into four main blocks. Preprocessing, segmentation, clustering, feature extraction and finally classification. \\
 %Due to sloppiness of users and hardware glitches the system preprocesses the data before further processing. 
 The preprocessing block is responsible of capturing the input data and removing the noise from it. The system then proceeds to estimate a set of possible dominant points to help in the segmentation process.  \\% The time differences, direction, speed and curvature of all points of the stroke is computed along with some other geometrical data. The system then proceed to estimate a set of possible dominant points to help in segmentation process.  \\
  %then passes through the segmentation process which will try to 
 %The goal of the segmentation stage is to divide stroke into segments of either curves or lines.
  The segmentation process is divided into two steps ellipse fitting and  curve segmentation. In the first step, the system tries to fit the stroke into an ellipse. If the system fails it passes the stroke into two PSO segmentation algorithms. The two algorithms will generate two segmentations, the system will choose the segmentation that has the minimum segmentation error.%  generate the segmentations,  the minimum error will be the chosen segmentation. % is the ellipse fitting
 %the stroke first If the ellipse detection fails the stroke is passed to the segmentation algorithms which will pass it to the two PSO algorithms described below the segmentation with the minimum error will be the chosen segmentation.  The segmentation is then added to the set of un-recognized segments in the system. \\% This part is repeated for each stroke.

After the user finish sketching the symbol by using any number of strokes, the set of unrecognized segments is passed to the feature extraction and classifier.% The system compute composite set of features some are statistical other are spatial features based on the type primitives. Finally, the strokes are classified into the corresponding classes.% the classifier identifies the strokes into a symbol from the set of known symbols.\\
\begin{figure}[]
	\centering
		\includegraphics[scale=0.35]{images/Blockdiagram.pdf}
	\caption[  ]{The system Block Diagram}
	\label{fig:Blockdiagram}
\end{figure}
\subsection{Preprocessing}
\label{Prepross}
%Some data is extracted from the points in the stroke.
 It was noted that as the user draws a shape the pen will slow down near corners and will pick up speed when drawing a straight lines. Therefore, the speed and time difference data were widely used in sketch understanding systems \cite{earlyprocess}.  The direction and curvature data are used to decide the points with high angular changes along the path of points\cite{meanshift10}.  \\

 In the presented system, we computed time difference, direction, speed and curvature of each point along the stroke. For speed calculation $v=\Delta t/\Delta s$ where t is the time difference between two points and s is the length between them. The direction is calculated as the angle between two vectors and curvature is the change in direction with respect to length i.e. $c= \Delta d/\Delta s$. \\
 %system compute speed, distance data for all points in the stroke. Curvature is computed using estimation used in [] where direction computed is angle between two lines. Curvature is 
 %$\Delta d/  \Delta S $ where d is the difference direction of point and s is difference in distance between points. 
 All the calculations are performed in real time while the user draws his strokes.
 So the resultant computation complexity of computing of all these computation is O(n) where n is number of points. \\
 Figure \ref{fig:speed2} shows the computed data for the stroke drawn in fig. \ref{fig:orignalStroke}.  If you look at the data curves you will see that dominant points are characterized by lower speed values and higher curvature and direction values. \\% the lower speed points correspond to vertex and dominant point's location. The higher the direction and curvature data correspond to location with higher change in curvature witch promote the location for vertices. \\
After the system computes all the speed, time difference and curvature data it proceeds to detect the points with low velocity and high curvatures. Using simple differentiations to detect local extreme points resulted in false points due to the non smooth curves. So instead the following procedure was used, firstly the mean of the curve is used as threshold for lower values then the high values are further investigated to find the extreme points. Hence, the mean of the direction curve is calculated and used as threshold. This threshold is used to separate the curve into regions; each region is defined as the part between two intersection points of the threshold line and the curve. Only regions higher than the threshold value are processed when computing the maximum of each region. The point that corresponds to those maximum values are labeled possible dominant points.   \\
   %If we tried to differentiate the curve the result will be false threshold values we divide the data into regions of data higher or lower that the threshold. This will let us only look for high data only. The maximum of each region is then computed and reported as a possible vertex point. 
\begin{figure}[]
	\centering
		\includegraphics[scale=0.3]{images/orignalStroke.pdf}
	\caption{Stroke Sample}
	\label{fig:orignalStroke}
\end{figure}
The system repeat this process to curvature, time difference and speed curves. All the points labeled as possible dominant points are saved into a single array. They are used as input and guiding data for the next stage of segmentation. 
\begin{figure*}[]
\begin{minipage}[b]{0.5\linewidth}
	\centering

		\subfigure[ The Speed of data] {\includegraphics[scale=0.15]{images/speed2.pdf}}

		\subfigure[ Direction ] {\includegraphics[scale=0.15]{images/direction2.pdf}}
					\end{minipage}
					\begin{minipage}[b]{0.5\linewidth}
			\subfigure[Curvature] {\includegraphics[scale=0.15]{images/curvature2.pdf}}

			\subfigure[ Time Difference] {\includegraphics[scale=0.15]{images/time2.pdf}}
			\end{minipage}

	\caption[Data Curves]%{The data of the stroke}

	\label{fig:speed2}
\end{figure*}
%Description of input data. 
%how to calculate speed, curvature, area, bounding box.
%how to remove noise. 
%Finally how to compute primarily dominant points. 
\subsection{Segmentation}
\label{seg}
After computing the primary data the system now tries to segment the stroke into a set of primitives. The segmentation algorithm first tries to detect if the stroke is an ellipse. If the stroke proved to be an ellipse then the segmentation process ends and the system proceeds to the next step. Otherwise, the stroke is passed into two particle swarm algorithms that will divide the stroke to either lines or lines and curves. The algorithms takes the stroke points along with the possible dominant points computed then produce a set of dominant points which are connected with either lines or curves (see fig. \ref{fig:Blockdiagram}).  The next section will describe the ellipse detection algorithm and both the two particle swarm algorithms used to divide the stroke. \\
%Paragraphs describe the segmentation algorithm.  
\subsubsection{ Ellipse Detection } 
%Describe the ellipse detection algorithms 
The process starts with computing the center of the stroke bounding box. The bounding box center point is used as the first estimation of the center of the ellipse. The axes of the ellipse are estimated as the w/2 and h/2 of the stroke bounding box. The least square fitting algorithm is used to minimize the fitting error of the ellipse equation (see \ref{eq:circleFit} where N is number of points in the stroke, $a,b$ are the length of ellipse axes, $x_0$ \& $y_0$ are the coordinates of the center point, $x_i$ \& $y_i$ are the coordinates of point $i$ in the stroke). After only few loops the error and a confidence values  are computed to check if the storke can be labled as ellipse or not.\\ %is computed to check if stroke is an ellipse. \\
%We found that if the confidante is above threshold then the probability of ellipse is highest otherwise the stroke is passed to the next section to get the divisions of stroke and test its error.  
\begin{equation}
E = \sum\limits_{i = 0}^N {\frac{{(x_i - x_0 )}}{{a^2 }}^2  + \frac{{(y_i - y_0 )}}{{b^2 }}^2  - 1} 
\label{eq:circleFit}
\end{equation}
\subsubsection{ Discrete particle swarm algorithm}
There are two PSO algorithm described in this section. The system generates segmentation from both algorithms and then chooses the segmentation with the minimum error value. The problem definition is the same in both algorithms but they differ in the method they compute fitness and error functions. %formation is nearly the same in both. 
\subsubsection{ problem definition}
The input stroke with N points can be represented by set $S = \left\{ {x_1 ,x_2  \ldots x_N } \right\}$ where $x_i$ is the location of the point $i$ . The swarm algorithms consist of M agents which are represented by the set 
$A  = \left\{ {P_i \left| {i = 1,2 \cdots M} \right.} \right\}$ where $P_i$ is a single solution particle from the solution space. Each particle decodes the problem with binary array with the same length N as the input stroke.  
% The particles of the swarm represent a single solution of the solution space. For this problem the, each particle %will give a different segmentation for the input stroke. Firstly, we will define the stroke with N points by the set %S where.  We define the arc %An array with the same length as the number of points of the strokes.
So the system will represent each particle $P_i$ by $P_i = \left\{ {p_{ij} \left| {j = 1,2 \cdots N} \right.} \right\}$ where $p_{ij}$ will have only two values 0 or 1 where $p_{ij}=1$ will mean that point $j$ is a dominant point. 
% a!402B!
%$A = \left\{ {p_i \left| {i = 1,2 \cdots N} \right.} \right\}$
% a!402A!
%\begin{equation}
%A = \left\{ {p_i \left| {i = 1,2 \cdots N} \right.} \right\}.
%\end{equation}
\subsubsection{The algorithm}
The fitness function and error calculation are different in each algorithm. For the first algorithm, 
let's define the arc $\widehat{x_ix_j}$ as the consecutive set of points from point $x_i,x_{i+1} \cdots,x_j$. The line
$\overline{x_i x_j} $ as the straight line connecting point $x_i$ to point $x_j$. The approximation error is computed by the equation \ref{eq:ErrorSwarm1}  where M is the number of dominant points in this solution.  The error $ e ( \widehat{x_ix_j},\overline{x_i x_j})$ is computed as the sum of squared perpendicular distance from every point along the arc $\widehat{x_ix_j}$ to the line $\overline{x_i x_j}$.  \\
\begin{equation}
E=\sum\nolimits_{i = 0}^M e ( \widehat{x_ix_{i+1}},\overline{x_i x_{i+1}})
\label{eq:ErrorSwarm1}
\end{equation}
\begin{equation}
\max fitness(p_i ) = \left\{ {\begin{array}{*{20}c}
   { - E/\varepsilon N} & {ifE > \varepsilon ,}  \\
   {D/\sum\limits_{j = 1}^N {p_{ij} } } & {otherwise}  \\
\end{array}} \right.
\label{eq:fitnessSwarm1}
\end{equation}%\]
The fitness is computed using equation \ref{eq:fitnessSwarm1} where N is the number of points in the stroke, D is the number of point in the solution that was previously labeled as a possible dominant point and E is the computed error and $\varepsilon$ is the error threshold.   As you can see when the error is larger than the threshold is given a -ve value to lower the value of solution otherwise the system will favor the lower number of vertices.\\ %we will want to lower the number of vertices. \\
%if we say that s
%Alg1:
%Alg2: 
The second algorithm has the same problem formulation but different fitness and error functions. It was previously introduced in  \cite{CruveDivisionSwarm} where but genetic programming was used as the optimizing algorithm.  The error of both circle and line estimation are computed for each segment, the approximation with the lower error value will be the chosen approximation of this segment. The sum of the approximation error of all segments is the total error of the particle.  The error is computed by equation \ref{eq:errorSwarm2} where M is the number of segments in the solution, $D_i$ is the minimum approximation error of curve and line approximations $min(d_c,d_l)$ as computed by \cite{CruveDivisionSwarm}.  The fitness is computed by the equation \ref{eq:fitnessSwarm2} where E is the error and M is number of segments and k is a parameter tweaked to get minimum number of segments. \\
\begin{equation}
E=\sum\nolimits_{i = 0}^M e(D_i) 
\label{eq:errorSwarm2}
\end{equation}
\begin{equation}
\max fitness(P_i ) = \frac{1}{{E \times M^k }}
\label{eq:fitnessSwarm2}
\end{equation}
%Describe how we used particle swarm in curve polygon approximation using first algorithm.  
%Describe how we used PSO in to divide stroke into segments and curves. 
%[These two algorithm was modified from the algorithm in paper \cite{PolygonApproximationPSO,CruveDivisionSwarm}].
\cite{PolygonApproximationPSO} used a merge and divide algorithm after each loop of the swarm system to refine the solution but we used another enhancement method. After each loop in the swarm algorithm, each particle loops on the set of selected dominant points to enhance the solution. Each dominant point is checked to find if it was labeled before as a possible dominant (computed as in section \ref{Prepross}), if not the point is moved to the nearest labeled point.    
%check the new solution and check the new vertices if a vertex points is a possible dominant point it is left as a vertex. Otherwise, the system will search in the local neighborhood for the nearest possible dominant point to move these vertices to. After each change the error is computed and if new error is less that the old then the new set is approved. \\
\subsection{Recognition}
After the user draw all strokes of the symbol he has to wait 10 seconds or press finish button beside the drawing area. The set of unrecognized strokes is grouped together along with their segmentation as input to the feature extraction process. The system uses a composite set of features include Rubine feature set \cite{gestureexample12}, Zernike moments \cite{HeloiseBeautification}, ink density \cite{GeometryAndDomain102} and some structural information like number of perpendiculars lines , number of parallel lines and types of primitives in each symbol. After computing the features the symbol is introduced to the classifier. The system used a support vector machine (SVM) classifier with Gaussian kernel (RBF kernel) \cite{libsvm}.\\%Some features like number of parallel, perpendicular and intersecting lines are computed from the segmentation other like density, centroid and Zernike moments are computed from the raw stroke points. 
 %trained with the 13 categories in the system. The same training set was used by \cite{HeloiseBeautification} in their symbol recognition system. \\ %  The table describe the features used in the system. 
%\begin{table}
%\begin{minipage}[t][1\totalheight]{0.3\columnwidth}
%
%\begin{tabular}{|l|l|l|}
%\hline 
%Type & Features & Notes\tabularnewline
%\hline 
%\hline Structural & No. of parallel lines & \tabularnewline
%\hline & No. of perpendicular lines & \tabularnewline
%\hline & No. of intersections & \tabularnewline
%\hline Statistical & Mean and standard deviation of the Curvature data   & \tabularnewline
%\hline & Mean of the speed data & \tabularnewline
%\hline 
%Rubine & Set of features \\for gesture recognition & Introduced by Rubine in \cite{} \tabularnewline
%\hline Zernike &  A Zernike\\ moments  &  Used by \cite{zernike61} \tabularnewline
%\hline Composite & width to\\ height ratio  & \tabularnewline
%\hline  & ink density  & Used by \cite{} \tabularnewline
%\hline
%\end{tabular}		
%\end{minipage}
%	\caption{Feature Table}
%	\label{tab:FeatureTable}
%\end{table}
  % Rubine el al.\cite{gestureexample12}  used a set of statistical features,  There are a lot of features used in the sketch recognition literature
%How to handle multi-stroked symbols. 
%The clustering algorithm.  
%The classifier used in recognition
%Describe the features used in recognition [ref to \cite{zernike61}]
\section{Experiments}\label{sec:Experiments}
For testing the system used a data set collected by \citeauthor{HeloiseBeautification}. The data are drawn by 16 users each of them have drawn from 30 to 50 samples for each shape. Figure \ref{fig:symbolSet} shows the shapes used in the data set. We divided the dataset equally into training set and test set.  The results displayed are the recognition accuracy resulted from the classifier after recognizing the symbols. \\
%We divided the dataset into four sets each set contains four different users. Furthermore, each set is divided into train and test set where the trained set where used to train the classifiers and the test set where used for testing.  The algorithms are then tested using the four sets and the results are averaged.
\begin{figure}[]\centering
\fbox{  \parbox{4cm}{% 
		\includegraphics[scale=0.4]{images/symbolSet.pdf}	}}
	\caption{The Symbol Set}
	\label{fig:symbolSet}
\end{figure}
\begin{figure}[]
	\centering
		\includegraphics[scale=0.3]{images/test1.pdf}
	\caption{The results of first experiment}
	\label{fig:test1}
\end{figure}
%In the figure \ref{fig:test1}, the accuracy means that final hit rate of all symbols used in the test set.by the  four datasets . 
We performed three experiments to test the system firstly we tested recognition accuracy of shapes in the data set with both algorithms. We also implemented the segmentation algorithm described in \cite{earlyprocess} to use as reference to our swarm algorithms.  As you see in the fig. \ref{fig:test1} shows the accuracy achieved by each algorithm. The results shows that both PSO algorithm achieve better result than other algorithms.  The swarm algorithms were tested with and without the ellipse detection module. The ellipse detection module appear to be superior to results with only the PSO algorithms.  \\

The second experiment we implemented was to test the effect of symbol complexity and type on the recognition rate. Figure \ref{fig:test2} shows what the accuracy of each symbol, it clearly noted that symbols that have only line segments achieve higher accuracy rate than other symbols. Figure \ref{fig:test2} also shows that algorithm 1 alone achieve better performance than algorithm 2 in the symbols that consist of lines only. The combining of both algorithm improves the recognition rate of other symbols.  \\
\begin{figure*}[]
	\centering
		\includegraphics[scale=0.5]{images/test2.pdf}
	\caption{The results of second experiment}
	\label{fig:test2}
\end{figure*}
  The final experiment we performed was to test the efficiency svm classifier. We implemented the linear discriminator classifier described in \cite{gestureexample12}(see figure \ref{fig:test3}).  %The result shows that SVM is more scalable than the linear discriminator for this problem. %The test focus on the size of data needed to test each classifier and the achieved accuracy.  %For this experiment we also shows the classification time for both classifiers 
\begin{figure}[]
	\centering
		\includegraphics[scale=0.4]{images/test1.png}
	\caption{The results of third experiment}
	\label{fig:test3}
\end{figure}
 %Details description of results got from comparing various algorithms  
 %Describe data set used for experiments ( No of data set, No. of categories, No of samples).
 %Describe other Algorithms [Alg 3 in paper \cite{earlyprocess}]used for comparing with the one described in the paper. 
\section{Conclusion}
This paper presented a new approach to sketch recognition using PSO. It was noted that the PSO in general improve the accuracy of the final symbol recognition in the system. The use of both speed and curvature data helps in improvement of the PSO algorithm over the original algorithms \cite{CruveDivisionSwarm,PolygonApproximationPSO}. The tradeoff between accuracy achieved and time complexity must be further investigated to achieve better results.  
%As you can see form the experiments PSO have proved a spuriously than other system. 
\section{Future Work}
The next step in this research is to complete the clustering algorithm for fully automated sketch recognition. The clustering must be performed without the user explicit involvement. Other area of modification can be the features extraction and classifier. Adding more spatial features will improve classifications.  %Each new stroke is checked if it can be a symbol or is a part of already drawn un completed symbol. 
%\bibliography{Mybibliography}
\bibliography{../../neededfiles/Bibliographies/Mybibliography2}
\end{document}

### 2008-08-18 writing/outline.tex
after my first changes till the system discription

@@
 %\title{A Study on Using Particle Swarm Algorithms in Symbol Recognition }
-\author{Maha El Meseery, Dr. Mahmoud Fakhr El Din, Dr. Samyah Mashaly,
-\\ Dr. Magdah Fayek, Dr. Nevin Darwish}
+\author{Maha El Meseery,  Mahmoud Fakhr El Din,  Samyah Mashaly,
+\\  Magdah Fayek,  Nevin Darwish}
 \begin{document}
@@
 \begin{abstract}
-Sketch recognition is defined by the process of identifying the symbols the user draws using a single or multiple strokes.
+Sketch recognition is defined as the process of identifying the symbols the user draws using a single or multiple strokes.
 Users draw strokes using a pen and the system immediately interprets their strokes into objects that can be easily manipulated.
-This paper uses Particle Swarm Algorithm(PSO) to divide strokes a user draws into meaningful geometric primitives.
+This paper uses Particle Swarm Algorithm (PSO) to divide the strokes the user draws into meaningful geometric primitives.
 These geometric primitives are grouped to formulate symbols which will be further identified.
 The results showed that using PSO improves segmentation results which guide the symbol recognition phase.
-Final recognition have improved after the PSO algorithm using a SVM classifier.
+Final recognition has improved accuracy after the PSO algorithm using a SVM classifier.
  %Hand drawn sketches are the natural method of exchanging thoughts between engineers and humans in general.
@@
 %A picture speak a thousand words is also true for sketches. 
-Scientists in general and engineers specifically express thoughts and designs using sketches.
+Scientists generally and engineers specifically express thoughts and designs using sketches.
 Engineers use sketches to exchange designs as a natural method of communication rather than writing or speaking.
 Currently, engineers use paper and pencil design in early stages of design.
-Design engineers need a powerful symbol recognition and manipulation system to help them operate as freely as with papers.
-Increasing the interaction between the computer and the users in sketch and CAD systems was recently the reason for the emerging of few advanced sketch recognition systems.
+Design engineers need a powerful computer based system with symbol recognition to help them design, manipulate and store sketchs more effectively than with only papers.
+Increasing the interaction between the computer and the users in sketch and CAD systems has been recently the reason for the emerging of few advanced sketch recognition systems.
 Sketch recognition system converts the drawings and informal graphs the user draws into objects and symbols can be used in further interaction with the system.
-\\
-Most researchers in sketch recognition divided the system into four main steps preprocessing, segmentation, symbol recognition and finally sketch understanding.
-The preprocessing process saves the location of points from the hardware and collects basic information about the stroke data.
-After removing noise and calculating primary data the system proceeds to divide the stroke into a set of simple geometrical primitives.
-The second step of the system is clustering strokes and segments to formulate symbols that can be recognized by the system.
-Finally, priori information about the sketch is used to support identification of symbols by excluding any possible invalid symbols.
-\\
+
+Sketch recognition researchers divide the system into four main steps preprocessing, segmentation, symbol recognition and finally sketch understanding.
+The preprocessing process saves the location of points from the hardware and collects basic information about the stroke data then proceed to remove noise and primary data calculation.
+In segmentation step, the system  divide the stroke into a set of simple geometrical primitives or segments.
+The third step of the system cluster strokes and segments to formulate symbols that can be recognized by a classifier system.
+Finally, the sketch understanding step uses priori information about the sketch to support identification of symbols by excluding any possible invalid symbols.
 %the basic symbols are identified in consistent with the whole sketch. \\
@@
 \\
-The next section will relate previous work in the field.
+
+Section \ref{Sec:RelatedWork} provides brief description to related work.
+The remaining of this papers will be as following: Section \ref{PSO} will explain the general particle swarm algorithm.
+Section \ref{Sysdisc} will describe the system in details.
+Section \ref{sec:Experiments} will describe the experiments that was preformed.
+Finally, section \ref{ConclusionandFutureWork} will state the conculusion and future work.
+%The next section will relate previous work in the field.
 The remaining of the paper will be as following; Section \ref{PSO} will explain the general particle swarm algorithm.
@@
 Finally, section \ref{sec:Experiments} will describe the experiments that was preformed.\\
-\section{Related Work}
+\section{Literature Review}
+\label{Sec:RelatedWork}
 %%%What is sketch recognitions. Why we need sketch recognition systems. 
-The field of symbol recognition has gained interest in the last few years. A wide variety of techniques were used either on segmentation process or symbol recognition.\\ 
-To segment strokes, \cite{meanshift10} introduced a feature area for each primitive and then compute the segmentation error for different types of primitives based on this area.
+The field of symbol recognition has gained interest in the last few years. A wide variety of techniques were used either on segmentation process or symbol recognition.
+
+%To segment strokes
+In order to segment strokes, \cite{meanshift10} introduced a feature area for each primitive and then computed the segmentation error for different types of primitives based on this area.
 %\citeauthor{HeloiseBeautification} used a dynamic programming algorithm to solve the segmentation problem \cite{HeloiseBeautification}.
-Complex and hybrid algorithm were also introduced as in  \cite{earlyprocess} where they generated different sets of segmentation based on both curvature and speed dominant points then they choose the segmentation with the least error from the hybrid set.
-A dynamic programming algorithm was also used to obtain optimum segmentation of the input strokes in \cite{GeometryAndDomain102}.
+
+Complex and hybrid algorithm were also introduced as in  \cite{earlyprocess} where they generated different sets of segmentation based on both curvature and speed dominant points then followed by a choice a segmentation with the least error from the hybrid set.
+Additionally, a dynamic programming algorithm was used to obtain optimum segmentation of the input strokes in \cite{GeometryAndDomain102}.
 Genetic algorithm was also used by \cite{CruveDivisionSwarm} to optimally divide digital curves into lines and curves.
-\cite{PolygonApproximationPSO} used PSO to convert digital curves into polygons.
-\\
-  %Some tried to detect dominant or corner points based on data generated from speed   \cite{earlyprocess} and mostly used curvature of the stroke drawn \cite{earlyprocess,meanshift10}.
+\citeauthor{PolygonApproximationPSO} used PSO to convert digital curves into polygons\cite{PolygonApproximationPSO}.
 
-%Particle swarm algorithms was used in \cite{} to fit a set of points into ellipse or higher order curves.
-%The first set of features introduced to be used on gesture recognition was introduced by Rubine in \cite{gestureexample12}.
-later,  Spatial relations and features where used in  \cite{cognitivesketch18,SRGraph57}  .
-\citeauthor{GeometryAndDomain102} introduced ink density feature which they used.
-The use of Zernike moments  was used  as the only features to the SVM classifiers in \cite{HeloiseBeautification}.
-\cite{SketchRead2007} used a set of features varied from relative spatial features and some compositional features.
-\\
-
-%After segmentation researchers varies on the methods in which to continue.
 Many different approaches have been investigated to achieve final recognition of symbols.
-Graph searching and template matching where used in \cite{mulitstroke5,GraphBased07}, where  a graph of the drawn symbol is generated and matched to stored set of graphs templates to reach a classification.
-Geometrical and spatial descriptive language was used in \cite{SketchRead2007,Ladder30} where a parser like architecture was used to recognize symbols  .
-\citeauthor{imagetrainable48} used trainable image template to classify different set of symbols \cite{imagetrainable48} .
-The use of SVM and HMM classification was implemented in \cite{SVM300,SmartSketch56} and \cite{HMM53}.\\% Other techniques as  [          ] was used  in [].
-\\
+Graph searching and template matching where used in \cite{mulitstroke5,GraphBased07}, where a graph of the drawn symbol is generated and matched with a stored set of graphs templates to reach a classification decission.
+Geometrical and spatial descriptive language was used in \cite{SketchRead2007,Ladder30} where an architecture simillar to compiler was used to recognize symbols.
+\citeauthor{imagetrainable48} used trainable image template to classify different set of symbols \cite{imagetrainable48}.
+The use of SVM and HMM classification was implemented in\cite{SVM300,SmartSketch56,HMM53}.
+% Other techniques as  [          ] was used  in []. \\  
  %In the next section the particle swarm algorithm will be explained.
@@
 %Write on the difficulties of the sketch recognition systems. 
-\subsection{Particle Swarm Algorithm}
+  %Some tried to detect dominant or corner points based on data generated from speed   \cite{earlyprocess} and mostly used curvature of the stroke drawn \cite{earlyprocess,meanshift10}. 
+%Particle swarm algorithms was used in \cite{} to fit a set of points into ellipse or higher order curves.
+%The first set of features introduced to be used on gesture recognition was introduced by Rubine in \cite{gestureexample12}.
+later,  Spatial relations and features where used in  \cite{cognitivesketch18,SRGraph57}  .
+\citeauthor{GeometryAndDomain102} introduced ink density feature which they used.
+The use of Zernike moments  was used  as the only features to the SVM classifiers in \cite{HeloiseBeautification}.
+\cite{SketchRead2007} used a set of features varied from relative spatial features and some compositional features.
+\\
+%After segmentation researchers varies on the methods in which to continue.
+\section{Particle Swarm Algorithm}
 \label{PSO}
 %What is particle swarm algorithm and how it was used in related researches. 
-In the Particle Swarm Algorithm (PSO) each agent represents a particle from the solution space.
-The agents move the particle with direction and velocity based on the general swarm equation [\ref{eq:Swarm}].
-Equation [\ref{eq:Swarm}] shows how velocity and direction of each particle is computed where $r_1$ \& $r_2$ are random variable and $c_1$ \& $c_2$ are the swarm system variables.
-Each iteration the global best $g_{best}$ particle and  the agent local best $l_{best}$ particle are evaluated based on the maximum fitness functions of  all particles.
-The solution is found after a maximum number of iteration or after an error threshold is achieved.
-Equation \ref{eq:descrite} where t is a threshold and $r_{3}$  is a random variable, is used to change general swarm algorithm into binary particle which will handle particle values of either 0 or 1.
-.
+The main idea of  Particle Swarm Algorithm (PSO) is to represent each agent with a particle from the solution space.
+Each agent moves the particle with a direction and velocity based on the following equation [\ref{eq:Swarm}].
 \begin{equation}
@@
 \end{equation}
+ Equation [\ref{eq:Swarm}] shows how velocity and direction of each particle is computed where $r_1$ \& $r_2$ are random variable and $c_1$ \& $c_2$ are the swarm system variables.
+Each iteration the global best $g_{best}$ particle and  the agent local best $l_{best}$ particle are evaluated based on the maximum fitness functions of  all particles.
+The solution is found after a maximum number of iteration or after an error threshold is achieved.
  \begin{equation}
@@
 \end{equation}
+  Equation \ref{eq:descrite} where t is a threshold and $r_{3}$  is a random variable, is used to change general swarm algorithm into binary particle which will handle particle values of either 0 or 1.
+.
 \section{System Description}
@@
 %The stroke is the path of points from the instant the pen is down till it is up.
-As shown in fig. \ref{fig:Blockdiagram} the system is divided into four main blocks. Preprocessing, segmentation, clustering, feature extraction and finally classification. \\
+Figure. \ref{fig:Blockdiagram} shows that the main system blocks are 1) preprocessing, 2) segmentation, 3) clustering and feature extraction, and finally 4) classification.
+
  %Due to sloppiness of users and hardware glitches the system preprocesses the data before further processing. 
- The preprocessing block is responsible of capturing the input data and removing the noise from it.
+The preprocessing is responsible of capturing the input data and removing the noise from it.
 The system then proceeds to estimate a set of possible dominant points to help in the segmentation process.
-\\% The time differences, direction, speed and curvature of all points of the stroke is computed along with some other geometrical data.
+% The time differences, direction, speed and curvature of all points of the stroke is computed along with some other geometrical data.
 The system then proceed to estimate a set of possible dominant points to help in segmentation process.
@@
   %then passes through the segmentation process which will try to
+
  %The goal of the segmentation stage is to divide stroke into segments of either curves or lines.
-  The segmentation process is divided into two steps ellipse fitting and  curve segmentation.
-In the first step, the system tries to fit the stroke into an ellipse.
-If the system fails it passes the stroke into two PSO segmentation algorithms.
+The segmentation is divided into two steps ellipse fitting and  curve segmentation.
+In the first step, the ellipse fitting process tries to fit the stroke into an ellipse.
+If the system fails it passes the stroke to the second step; curve segmentation which consist of two PSO segmentation algorithms.
 The two algorithms will generate two segmentations, the system will choose the segmentation that has the minimum segmentation error.%  generate the segmentations,  the minimum error will be the chosen segmentation.
 % is the ellipse fitting
+
  %the stroke first If the ellipse detection fails the stroke is passed to the segmentation algorithms which will pass it to the two PSO algorithms described below the segmentation with the minimum error will be the chosen segmentation.
@@
 \\% This part is repeated for each stroke.
+ The clustering alogrithms starts to group segments togther after the segmentation step.
+The system let the user draws the symbol by using any number of strokes, a set of unrecognized segments is passed to the clustering algoirthm to generate a symbol and compute a feature vector for it.
+The final step is a SVM classifier that will use the features computed to classify the segments into one of the previously trained classes.
+%  Or to determine the symbol of the given segments from the set of preciously trained symbols.
 
-After the user finish sketching the symbol by using any number of strokes, the set of unrecognized segments is passed to the feature extraction and classifier.% The system compute composite set of features some are statistical other are spatial features based on the type primitives.
+ %with a symbol from the training set.% The system compute composite set of features some are statistical other are spatial features based on the type primitives.
 Finally, the strokes are classified into the corresponding classes.% the classifier identifies the strokes into a symbol from the set of known symbols.\\
@@
 %Some data is extracted from the points in the stroke.
- It was noted that as the user draws a shape the pen will slow down near corners and will pick up speed when drawing a straight lines.
+ It is noted that as the user draws a shape the pen will slow down near corners and picks up speed when drawing a straight lines.
 Therefore, the speed and time difference data were widely used in sketch understanding systems \cite{earlyprocess}.
-The direction and curvature data are used to decide the points with high angular changes along the path of points\cite{meanshift10}.
-\\
+The direction and curvature data are used to determine the points with high angular changes along the path of points \cite{meanshift10}.
 
  In the presented system, we computed time difference, direction, speed and curvature of each point along the stroke.
-For speed calculation $v=\Delta t/\Delta s$ where t is the time difference between two points and s is the length between them.
+The following equation $v=\Delta t/\Delta s$ is used for speed calculation  where $t$ is the time difference between two points and $s$ is the length between them.
 The direction is calculated as the angle between two vectors and curvature is the change in direction with respect to length i.e.
 $c= \Delta d/\Delta s$.
-\\
+
  %system compute speed, distance data for all points in the stroke. Curvature is computed using estimation used in [] where direction computed is angle between two lines. Curvature is 
  %$\Delta d/  \Delta S $ where d is the difference direction of point and s is difference in distance between points. 
- All the calculations are performed in real time while the user draws his strokes.
- So the resultant computation complexity of computing of all these computation is O(n) where n is number of points. \\
+ All these calculations are performed in real time while the user draws the strokes.
+ The complexity of compution is $O(n)$ where $n$ is number of points. \\
  Figure \ref{fig:speed2} shows the computed data for the stroke drawn in fig.
@@
 	\centering
-		\includegraphics[scale=0.4]{images/test1.png}
+		\includegraphics[scale=0.4]{images/test1.pdf}
 	\caption{The results of third experiment}
@@
  %Describe other Algorithms [Alg 3 in paper \cite{earlyprocess}]used for comparing with the one described in the paper. 
-\section{Conclusion}
+\section{Conclusion and Future Work}
+\label{ConclusionandFutureWork}
 This paper presented a new approach to sketch recognition using PSO.
@@
 %As you can see form the experiments PSO have proved a spuriously than other system. 
-\section{Future Work}
+%\section{Future Work}
 The next step in this research is to complete the clustering algorithm for fully automated sketch recognition.

### 2008-09-02 trunk/draft/ChapterIntroduction.tex
First add of the master draft

%\usepackage{graphicx}
\chapter {Introduciton}
\pagenumbering{arabic}

Computer research nowadays aimed at facilitating computer human interaction in every way. Pen-based interface gives the user a pencil-paper like feeling that enhances interaction more than the current used keyboard-mouse computer interfaces. However, until now there are hardly any complete pen based computer systems. Generally, Current Pen-Based systems make use of the Pen or stylus to perform the same role as the mouse / keyboard. The minority of such systems are equipped with handwriting or digit recognition that is typically restricted for specific language.

 Despite the fact that people frequently express their ideas through drawing them there isn't a system that can fully understand those sketches. Lately, only a few researchers were concerned with the problem although it is highly important.  
Firstly, we need to know;
 \textit { What are the main goals of Sketch understanding systems?}

The main goal of sketch understating systems is to build a system that will interpret the sketches the user draw. The systems ought to be natural as possible to provide user with paper and pencil impression. Furthermore, interpreted sketches are expected to transform into objects, which can me manipulated, modeled and/or simulated.

Sketch here stand for any hand drawn drawing the user draws using digital stylus or similar device. Normally, sketch are composed of smaller segments we call strokes. The time between pen-up and pen-down events identify each stroke.  The strokes are mainly the path of points, which the user path through between the pen-down and pen-up events while using the stylus or similar device.

Sketch recognition is procedure that will input a sequence of points from user then segment them into clusters of strokes and recognize those clusters to primitive geometric objects and shapes. However, Sketch understanding is transforming the sketch from a group of primitive shapes into comprehendible identified symbols consistent with each other.
% the entire sketched symbols.

\section{Importance of sketch recognition}
\label{sec:ImportanceOfSketchRecognition}

People normally sketch thoughts rather than expressing them linguistically. For instance engineers mostly drafts their early designs using paper and pencil before using CAD tools.  People in general and engineers specially make use of various sketches during design process as instance organizational charts, informal drafting and formal design sketches \cite{}. The sketching process itself go thorough numerous stages. Firstly, early stages with rough drafting designs or notations which are mainly done using paper and pencil. After those primary rough sketches are edited and altered until they are stabilized and finally they are transformed into the computer.

 Paper and Pencil are preferred in design as they offer a remarkable ease in editing and altering the primary rough sketches. Current CAD systems don't give the designer the ease of altering the early designs. Designer had to sketch primary designs on paper and pencil next, the finally design is transferred into computer using a menu CAD system. The process is time consuming and tiring. The need for fast design and development process is driven by the fast growing industries and technology.

 A sketch recognition and understanding system will provide a powerful tool to the designers to draw sketch as they usually draw using paper and pencil with the advantages of computer program. Designer needs a normal sketching environment that will provide him facilities in paper interface and the modifying and simulation that is in current CAD systems.

\section{Problems and challenges}
\label{sec:ProblemsAndChallenges}

 Understanding hand drawn sketches may be trivial for humans but it is challenging problem when speaking about computer-based system. Most of the research made on the sketch understanding is achieved for one particular domain or for strictly a set of pre defined symbols. Further, others restrict the user the freedom in drawing which loose the main goals of paper-like-interface systems.  The only few systems that have achieved the desired degree of freedom for the users are extremely computationally expensive.

The research in the area had faced few problems that will need to be handled to implement an entirely sketch understanding system. The user draws strokes in a spontaneous and ambiguous manner that intensifies the challenges. The first problem is segmenting strokes from the way the user draw it into primitive geometric. Ink parsing is the problem dealing with clustering interpolated strokes and grouping them into symbols. After that, recognizing the symbols from the space the system can identify is the final problem. All those problems must be solved on a real time constrain to maintain an online sketching system if it was a requirement.

The next sections will explain and investigate each of those problems in details.

\subsection{Segmentations}
\label{sec:Segmentations}

Segmentation is dividing the strokes draws into the geometric primitive's as lines, arcs, and ellipses. The strokes as mentioned before are the path of points between pen-down and pen-up events while using a stylus. The processing of those points is done after that to firstly to detect the critical points or vertices. After that the segment is classified as either line, arc, ellipse or circle. The main problem in this process is in detecting the critical and vertices which are points divide strokes into lines and arc. The user draw the lines and arc in an undefined manner which make it difficult for the program to deicide where the user had finish a line and started a new arc or vice verse; Fig [\ref {fig:strokeinterpolation}].

\begin{figure}

		\includegraphics[scale=0.8]{../../neededfiles/Figures/strokeinterpolation.eps}
	\caption[Segmentation error]{The figure shows an arc that the user draw which can be segmented in different ways as a single line or as an arc where a) The stroke users draw  b)interpretation as a single line segment  b) interpretation as an arc.}
	\label{fig:strokeinterpolation}
\end{figure}

\subsection{Ink parsing}
\label{sec:InkParsing}

 The clustering of segmented strokes into group of strokes before passing into the next stage this process is called ink parsing. It is simply deciding when a symbols ends and a new symbol begins.  Ink parsing is most important and difficult challenge as grouping of strokes can be done in many ways and configuration, which increase exponentially with the number of strokes. In addition, the ink parsing will affect the next steps of recognizing. The user does not draw the symbol in same order. Moreover, users can draw a sequence of strokes in symbols then leave it to draw another then return again to add segment to symbols so temporal data is helpful but not conclusive.  The user may also draw the same symbols using different orders and methods for example the user can draw a rectangle using single stroke or using two, three or even four strokes. All those factors gather with the sloppy drawing of the user to increase the complexity of the problem.

\subsection{Ambiguities}
\label{sec:Ambiguities}

The major problem of sketch understanding is the high level of ambiguities in the drawing. The user draws strokes that the computer cannot decide if it is arc or line. The nature of hand drawn sketch it self is highly ambiguous even for human eye you may see two drawn you cannot decide if it is square or rectangle.
 The ambiguities in sketch understanding are found at many levels in the system. Firstly, at the segmentation phase where the same user stroke can be interpreted as a ellipse and a circle. Then the symbol recognition as a symbol can have more than one interpolation. Due to the user sloppy drawing the symbols can have more than one interpolation which varies with the domain drawn in it similar problem was faced in handwriting recognition see Fig [\ref{fig:VorU}] but it was handled with knowing information about the context information which may not be the case in sketch recognition.

\begin{figure}	
	\centering
		\includegraphics[scale=0.85]{../../neededfiles/Figures/VorU.eps}
	\caption[Ambiguities Handwriting ]{The figure shows that the same letter can be interpreted differently in various contexts; first as V in drive or as U in Cut}
\label{fig:VorU}
\end{figure}

The ambiguity problem must be first recognized ambiguities or varicose interpolations of same symbol. After that, resolve those ambiguities is detecting the best correct interpolation of the symbol.  To solve this problem the program must be able to detect the amount of information needed to decide which interpolation of symbols is the best one.

\subsection{Search space}
\label{sec:Searchspace}

The user drawn symbols can be having more than one interpolation. In addition, those interpolations increase exponentially with the number of strokes the user draw and as the sketch expands the number of interpolation escalate. This intensifies the search space that the program search to resolve the ambiguities and find best interpolation.

\subsection{Real time}
\label{sec:Realtime}

There are mainly three types of sketch understanding system offline \cite{Vibratory8}, online and interactive. Both online and interactive recognition like \cite{incrmentintention41, SmartSketch56} need the processing after each stroke to be real time so the user does not feel the processing.

### 2008-09-02 trunk/draft/SketchMaster.tex
First add of the master draft

%==============================================================================
%
%      File:  Master.TEX  (Stored in TEX$LATEX: as UNHTHESIS.TEMPLATE)
%  Language:  LaTeX - Document Preparation System
%      Date:  26/Dec/2005
%    Author:  Maha El Meseery
%             
%
%   This document  is my master thesis on sketch Recognition and understanding
%
%
%        % latex mythesis
%
%------------------------------------------------------------------------------

%    14 Aug 02 Updated document to latex 2e. Also, changed wording 
%              to be consistent with an X windows system.

%    14 Aug 02 Added \includeonly command in comments.

% During the editing process you will not want to view and print the
% whole document. Uncomment the command below so that only the
% chapters that you want to work on are visible.
%\includeonly{ChapterIntroduction}
%\documentstyle[11pt,unhthesis]{report}
\documentclass[11pt,doublespace]{Sketchthesis}

% Extra modules can be added to your document to extend the things
% the Latex can Do. One of these is psfig which allows you to print
% postscript files. You tell Latex to read these packages using the
% \usepackage command:

\usepackage{psfig}

\usepackage{natbib}
\bibliographystyle{plain}
\hyphenation{gno-mon-ly}

\usepackage {graphicx}
\usepackage{subfigure}

%------------------------------------------------------------------------------
%  Preliminary Pages - Fill in the `blanks' noted.
%------------------------------------------------------------------------------

\begin{document}
                                                        % TITLE PAGE
                                                        %======================
\title{Sketch recogniton and understanding}             % My theses title
\author{Maha Mohammed Nabeel El Meseery}                               % Y name
\prevdegrees{B.S.,Computer  engineering(2003)}       		% Your old degree
\major{Computer Engineering}                            % My current major
\degree{Master of Science}                              % My new degree
\degreemonth{Dec}                                      % When awarded. to be added when known
\degreeyear{2005}                                      %
%th today or first day of writing  December 26 2005 
\thesisdate[]                                     % Date of document.
\DOCUMENTtype{Survey}                                   % or DISSERTATION
\Documenttype{Survey}                                   % or Dissertation
\documenttype{Survey}                                   % or dissertation
\maketitle
                                                        % COPYRIGHT PAGE
                                                        %======================
\copyrightyear{2005}                                    % Delete these
\makecopyright                                          % if no copyright
                                                        % page.

                                                        % APPROVAL PAGE
                                                        %======================
%\supervisor{A. B. Smith}{Associate Professor of xxx}    % <- One of these
%\committee{A. B. Jones}{Associate Professor of xxx}     % <- As many as
%\committee{A. B. Black}{Faculty in Residence in xxx}    %    you need...
%\makeapproval                                           %

                                                        % DEDICATION PAGE
                                                        %======================
%\begin{dedication}                                      % Delete these
												                                  % if no dedication
%\end{dedication}                                        % page.

                                                        % ACK. PAGE
                                                        %======================
%\begin{acknowledgments}                                 % Delete these if
 %\end{acknowledgments}                                   %

                                                        % FOREWORD PAGE
%\begin{foreword}                                        %======================
    											                             % Delete these if
%\end{foreword}                                          % no foreword page.

                                                        % OTHER PAGES
                                                        %======================
\tableofcontents                                        % Always needed...
%\listoftables                                           % Delete if no tables.
\listoffigures                                          % Delete if no figures.

%------------------------------------------------------------------------------
%  Document body - Place text into individual include files, such as
%                  as "chap1.tex", or replace each "\include{}" statement
%                  with your actual document text.
%------------------------------------------------------------------------------
\begin{abstractpage}          
% ABSTRACT PAGE
%==============================
% Creates the abstract page.
%   Your text goes here. Just
%   follow the rules given in
%   the LaTeX manual on how to
%   enter/format text using
%   LaTeX commands.

% didn't finish it all Need some addint

The need for new level of human computer interaction had lead the researchers to attempt improve the current interfaces. Even though the area of research progressed quickly in speech and handwriting systems gesture and sketch understanding systems are still starting to get noticed. This survey will highlight the current research that is ongoing on the sketch recognition and sketch understanding systems.

\end{abstractpage}                              % This ends the page.

                                                % CHAPTERS
                                                %==============================
\include{ChapterIntroduction}                              % The best way to organize your
\include{chapterSurvey}                              % document is probably to keep
                                                % in its own file.  
%\include {Mybibliography}

\bibliography{../../neededfiles/Bibliographies/Mybibliography.bib}     
                                                % THE APPENDIX
                                                %==============================
%\appendix                                       % Include appendix files as
  %\begin{singlespace}                           % needed or delete this block
  %\include{appendix_a}                          % of lines if no appendix.
  %\include{appendix_b}                          %
 % \include{appendix_c}                          %
%\end{singlespace}                               %

\end{document}                                  % Ends the entire document...

### 2008-09-02 trunk/draft/chapterSurvey.tex
First add of the master draft

\chapter{Literature Survey}

\section{Introduction }
\label{sec:Introduction}

The researchers had divided the sketch recognition problem into image based and AI Based problem. The next section will discuss the methods used by current sketch understanding system. Firstly, in section \ref{sec:FirstStepSymbolRecognition} Symbol recognition research will be investigated and divided into image based and AI based methods.

 In section \ref{sec:SecondStep:SketchUnderstanding}, the understanding techniques will be demonstrated. Understanding mainly will handle knowledge representation methods, handling uncertainty and finally how researcher in sketch reduced the sketch search space.

\section{First Step: Symbol Recognition}
\label{sec:FirstStepSymbolRecognition}

To understand a drawing we need to recognize every symbol in it. Hand-drawn sketches need more processing before it is passed to recognitions algorithms.

 Firstly, we have to determine how symbols are defined or grouped. Some System implies some constrain on user to defined symbol. In\cite{gestureexample12,aideddesgin22,sketchinginterfaces2,sketchinginterfaces2,cognitivesketch18,ewhitboard9}for example the user is constrained to draw each symbol using only one stroke which is not only unnatural but also constrains the symbols that can be drawn. More recent systems force the user to state when symbol is finished by pausing a period of time or clicking a certain button in the interface. Only few systems permit the user to freely draw strokes and symbols in any order \cite{Vibratory8,visualpattern43,Cali63,statisticalparsing26,physicalmeaning6}. Those systems must employ a method to detect when a user starts drawing a symbol and when he end.

Currently symbol recognition is both image based and feature based or AI-based. In image based recognition symbol is recognized based on there features, structure, some image computed feature . The AI methods in recognizing symbols mainly imply more time and processing as for the search and rules applied but claims to be more accurate .

\subsection{Pre processing }
\label {sec:preprocessing}
 A stroke is all the points that lie along the path of the user pen from the moment of his pen down to pen up events. Not all those points are important even more, there may be redundancies unwanted points. Therefore, those strokes will need some preprocessing.

The stroke that the user draws are mainly sequences of line segments or/and curves which the any system need to divide them to recognize them easily. The main problem is to do so without constraining the user drawing style. The majority of researchers first use a preprocessing process to identify those segments and the critical points from the user strokes.  
Beside points location information (x, y positions) researcher made uses of the extra information from using digital ink. Some as \cite {mulitstroke5,polygonfeedback31}  used user speed information to detect critical points where it is stated that user slow down at those critical points. Other used curvature and directions data. There was also mentioning using the pressure information.  Thought some \cite {MilitaryCOA37,computationalmodel16,polygonfeedback31}  debated on the importance of the extra information and weather the temporal and speed information is significant.  \cite{polygonfeedback31}  used the timestamp for each point and compute time difference between every two successive point to help him finding corners.
\cite{Overview36} uses the speed and direction information but they use on it a hybrid average based filtering to filter noise and unwanted false data from appearing in the data. Finally, they use that information to get a poly line approximating.

\begin{figure}
	\centering
		\includegraphics[scale=0.7]{../../neededfiles/Figures/dirspeedgraph.eps}
	\caption[Direction and Speed]{direction and speed graphs of the rectangle stroke the minima in the graph show the location of the corners of the rectangle.}
	\label{fig:dirspeedgraph}
\end{figure}

\cite{Mathpad46}Handle rotations invariant using polar transformation then shifting to make it more reliable when template matching. However, they state that the only problem to there system is that the points near the centroid affect the polar transformations which will effect the matching. (To solve it they say that loosing the weight of the pixel near the centroid will improve classification).

The next sections will demonstrate firstly the noise removal algorithms and techniques. Then next sections will show the most used critical points detection algorithms along with the segment division algorithms.

  In\cite{sketchRead29} they divided user strokes into geometric primitives based on some critical points they computes.  After that, they either cluster segment to form a symbol   or directly apply some kind of basic symbol recognition leaving the compound object for next processing step .

From the image-based view the symbol is a sequences of line segments or/and curves. To do so without constraining the user drawing style we have to easily divide the user drawn strokes.

%\begin{figure}
%	\includegraphics[scale=0.6]{../../neededfiles/Figures/shapesegmentation.eps}
%	\caption[segmented shapes]{example of Segmented shapes}
%	\label{fig:shapesegmentation}
%\end{figure}

\subsubsection{Removing noise}
\label{sec:RemovingNoise}

Due to the sloppiness of the user, his stroke may need a lot of noise removal before any algorithm can extract its features. For example, when a user draw sketches he may over trace the stroke more than one time. Human perception can easily differentiate what exactly is drawn. Nevertheless, for an AI program this is more difficult problem. The user may also overshoot see \textit{Fig\ref{fig:Overshootandundershoot}} or undershoot while drawing intersecting segments. 
\begin{figure}
	\centering
		\includegraphics[scale=0.7]{../../neededfiles/Figures/Overshootandundershoot.eps}
	\caption[Neat and Sloppy symbols] {The Figure show a sloppy strokes with under shoot and overshoot problems and the neat version of the symbol.}
	\label{fig:Overshootandundershoot}
\end{figure}

 \cite {overtraced24} tried to solve the over traced strokes. they used a thinning algorithm to move points perpendicularly towards the regression line from the formed from fitting line of neighbour's into line or circle or ellipse which is calculated with least square fit for the data. The next step is selecting near neighbours and connection then which gives a serious of linked edges which a greedy link edge algorithm this produce a single stroke instead of over traced see \textit{Fig [\ref {fig:overstroked}]}.

\begin{figure}
	\centering
		\includegraphics[scale=0.7]{../../neededfiles/Figures/overstroked.eps}
	\caption{Example of an overstroked stroke}
	\label{fig:overstroked}
\end{figure} 
Some systems as in \cite{SmartSketch56} where the noise reduction is done in two steps the first step is the polygonal approximation to remove non critical points in the user stroke. The process is mainly removing the redundant points that are along the path of stroke. Due to the shaky nature of user stroke, there are additional end-points at the stroke segments. Secondly, the next step will be applying an Agglomerate point filtering.  Agglomerate filtering is according to \cite {statesurvey35} removing the undesired extra points that is produced due to the unsteady operations produced while lifting and touching tablet.  The results of those processes are passed to final end-point refinement to merge or remove any redundant points.

In \cite{meanshift10,domainindependent17}, Bo Yu uses mean shift procedure as a noise remover and preprocessing step to the recognition step. The mean shift procedure is simply shifting the point to the average of its neighborhoods so smoothing out the noise without changing the global structure of data.  This process is clamed to help find critical points easily after that.

In \cite {domainindependent17}    approximation step uses the speed, curvature, and direction of strokes drawn by user to determine the location of vertices to gain more resistance to noise that data may contain.

\subsubsection{Critical points and corners Detection}
\label{sec:CriticalPointsAndCornersDetection}

Locating the critical points is provided to the next process of the systems which in most cases segment detection algorithm as in \cite{Simusketch51} . Other systems after detecting the points the system directly uses those information to template matching.

Generally, the critical points along with the segment points of the path from one critical point to another are used to detect labeling of that segment. 
Line detection or multi line segments algorithm is found in \cite{polygonfeedback31} . Mostly after detecting the vertices if the segment cannot be labeled as arc it is defined as a line. Other mainly defines a threshold value of the slope if it increase than that it is counted as an arc.

Systems like \cite{physicalmeaning6} detect critical point for each stroke then the final point is merged with the first if they are in range defined by threshold. However, stroke may also intersect with other strokes drawn before therefore \cite{physicalmeaning6}  introduced a structure to detect such approximates he first define every stroke drawn as series point and labeled as line, arc or area. Then after labeling most strokes the system check for proximity on bases of line -line, line-end point, or line- area.

In   \cite {meanshift10,domainindependent17}   approximation step uses the speed, curvature, and direction of strokes drawn by user to determine the location of vertices to gain more resistance to noise that data may contain. They firstly analysis the curvature and speed data and use the mean of data as the threshold then use that threshold to separate data into regions and finally locating the maxima and minima in each region . The previous step generates a set of vertices and its certainties they use to choose best fit.

\subsubsection {Primitive stroke segmentation and clustering}
\label{sec:Primitivestrokesegmentationandclustering}

The most segmentation method is least square method in \cite {overtraced24}. Other systems  approximate a line segmented using the information from the direction curve where the curve is fitted with horizontal line. firstly the system compute least square fit for direction curve or the stroke point computed and check if deviation points is small if they pass then hypothesize al line segment  from the first point to the last one.

\begin{figure}
	\centering
		\includegraphics[scale=0.7]{../../neededfiles/Figures/rectanglerecog.eps}
	\caption{rectangle recognition}
	\label{fig:rectanglerecog}
\end{figure}

%//Other like \cite {tobeadd}  uses the relaxation process to try to relax parameter on a classification .//
 In \cite {meanshift10,domainindependent17}  Bo first uses the direction graph to check if the stroke can be approximated as a line or no. If the stroke can't be approximated as a line it is checked to be a circle or ellipse. the procedure is simply assume a circle classification then check this assumption with the user drawn stroke given a min error range.  
\cite {Fluid25} Trying to combine fast morphing and basic recognition for the sketch drawn as the user draw it. It tries to move the points of strokes drawn by the user draw into one of the known shapes (ex: box, lines and circle). It uses a relaxation and least square method for circles and uses the string of forces for finding best-fit box.

The noise removal and critical point detection is only a preprocessing step for the symbol classification algorithms. Some system implements a segment segmentation and clustering to help in the classification problem.

After the strokes are segmented \cite {sketchunderstanding1,HierarchicalParsing7,Simusketch51}  proposed a merging techniques that will merge strokes together into symbols to help classification. They use a method combine between divide and merge. They firstly divide the drawn sketches into section based on marker technique they used. After that they merge strokes if there bounding boxes intersects and then  merge generated cluster till no more merge can occur on one segment.

%Intension exteraciton in paper Intention54
% Check to add

%/*Mathpad46
%methion  check stroke clustering.  Simusketch51  and in 8 
%
%Bo \cite {tobeadd}   The ellipse or

\subsubsection{Image and Statistical Features calculation}
\label{sec:ImageandStatisticalFeaturescalculation}

Most system calculates the general image properties for the strokes. Most used features are bounding box, width, size, points along path. 
\cite {meanshift10,domainindependent17}  introduced a new feature which they named "Feature Area which" they used in primitive classification introduced before. They first make and assumption for classification and compute the difference. For example if they make assumption as a circle they first calculate the center of the circle then its radius based on the bounding box of the stroke. The area difference between the computed circle and the current stroke is the feature area. see \textit{Fig\ref{fig:featurearearc}}.

\begin{figure}

		\subfigure[Feature area of line] {\includegraphics[scale=0.7]{../../neededfiles/Figures/featureareline.eps}}

	\subfigure[Feature area of an arc]{	\includegraphics[scale=0.7]{../../neededfiles/Figures/featurearearc.eps}	}
			\caption[Feature area]{the shaded are shows the difference between the computed feature area and the assumption area. }
	\label{fig:featurearearc}
\end{figure}

 System that uses statistical classifiers use other features for each stroke. Features like center of gravity\cite{gestureexample12,sketchunderstanding1}, maximum length line, Zenerik moments \cite{zernike61}, number of lines, curvature angles, minimum enclosing rectangle\cite {incrmentintention41} , convex hull were calculated in \cite {Cali63}.  Also \cite {geometrydomain49} introduced a new feature which he named ink density that he used in the classification of the glyphs see \textit{Fig \ref{fig:inkdensity}}. 
\begin{figure}
	\centering
		\includegraphics[scale=0.7]{../../neededfiles/Figures/inkdensity.eps}
	\caption[Statistical Feature Example]{The figure shows the Ink Density feature that \cite{geometrydomain49} introduced. }
	\label{fig:inkdensity}
\end{figure}
All spatial relationships needed in the graph or template matching is computed after statistical features are calculated. Relations like intersections, containment, perpendicular, right-of, left-of and below are the most used relations concerning constructing graphs for symbols\cite {geometrydomain49,sketchinginterfaces2} .%stretchdiagrammer4,     i dont know where is this refrence now

\subsection{Image Based Classification}
\label{sec:Image Based Classification}

In the following sections, systems generally try to recognize the strokes user draws and give it a label based on there image features. The information collected and features computed in the previous step.
 \subsubsection{Low Level Analysis}
\label{sec:Low Level Analysis}

%
%Detecting object using low level analysis only 
%\cite {mulitstroke5} first detecting vertices and control points then fit these segment into line and arc using best fit and least square error classification.

%Check the DB

% description[4] after that it passes through the curve approximation into the Bezier curves
%
%\cite {statesurvey35} we need to fit and filter the classified shapes that is done using on of the following least square is the most frequently method used some use fuzzy logic to make circular fitting some use b-splines or ellipse and circle fitting. some after that do some  shape beatification or regularization.

The most used method in low level recognition is best fit and least square methods where they was used in most systems. Other systmes like \cite{polygonfeedback31} prefered to implement its own polygon recognition algorithm.

Bo et al. In \cite{meanshift10} introduced an algorithm to recognize line and arcs.  The system used the direction graph and fitted stroke points with a non horizontal line. So the recognizing is done using the following algorithm which is preformed before the mean shift procedure so not the destroy the direction graph .Firstly, if direction graph can be approximated using a line then it is recognized as a line. Otherwise the algorithm Bisect the line from the first and last  point of stroke and donate it as "M". Then calculate the intersection point of the line "M" and original stroke and donate it as "A" . Now using the three points first , last points of stroke and point A find the center of circle which passes through all of them which will donate the candidate arc. Finally , calculate the radius of the circle and the mean of distance  between center and each point of stroke and  compute the feature area of the stroke and the circle arc see \textit{Fig \ref{fig:featurearearc}}.
%//

\subsubsection{Template matching}
\label{sec:Template matching}

Template matching is the most used method to recognize strokes and symbols into classes.  Image based template stored on the database is matched with the one user draw to find the correct classification of the strokes drawn.  In most cases, the stroke template is stored as image bitmap \textit{Fig \ref{fig:template}}in the database when the user draws the symbol it is matched using classifiers. \cite {imagetrainable48,HierarchicalParsing7}  used four kind of classifiers to detect the matching template. He used two classifiers to measure similarity and two classifiers to measure dissimilarity then combine them to find best match of drawn stroke from the database.

\begin{figure}
	\centering
		\includegraphics[scale=0.7]{../../neededfiles/Figures/template.eps}
	\caption[Template matching]{Examples of image templates used to match symbols in \cite{imagetrainable48}. }
	\label{fig:template}
\end{figure}

To refine the matching procedure and create more robust classification \cite {imagetrainable48}  first transform the strokes into polar coordinate then match the transformed stroke with the templates in the database. They claim that enhance the classification procedure as the polar coordinate system due to its definition make the stroke more robust to transformation and rotation.

\cite{templatefrag21}  Introduce new way to match stroke temples. He store template as a sequence of lines and ellipse.  When the user draws a stroke the system, segment it into a sequence of lines and ellipse. The system then matches the stored sequences with computed one to find correct classification.

The main problem of template matching is that in most system the templates are hard coded inside the system and cannot be modified at runtime.

\subsubsection{Graph matching}
\label{sec:Graph matching}

Representing the symbol that the user draws as a graph transforms the classification problem into a graph-matching problem. In general, each user stroke is represented as a node of the graph and the relations between those nodes are represented as arc. To classify the symbols the graph constructed must be matched with the graphs in the database.

\cite{physicalmeaning6} uses spatial relationships for there graph matching. They first calculate spatial proximity relationships. Relations like intersections, containment, perpendicular, right-of, left-of and below are the most used relations concerning constructing graphs for symbols.

Other systems  first analysis the spatial relationship between glyphs (sequence of strokes) using a library of spatial relations and information calculated in the preprocessing step.  After that, they filter out important relation-using hierarchy of rules and type of geometrical constrain. All redundant constrains and relations are removed the final graphs is matched with the database. If more than one candidate is available the algorithm choose the best fit but if no matches found constrains are relaxed and strokes are passed again to classifier.

 Proposed a structure representation of the strokes in \cite{sketchunderstanding1,HierarchicalParsing7} . She added ratio of proximity of the relations between strokes. In addition, she used an error-driven, stochastic search algorithm for graph matching.

\subsubsection{Statistical methods}
\label{sec:Statisticalmethods}

The application of Statistical classifiers in recognizing shapes are widely used in sketch recognition. Statistical classifier requires training time before the recognition actually happen. Then first step of classification is features computation that is done in the previous step. Finally, the recognition of symbols will be based on the classifier and the training set used.

One of the first systems in sketch recognition is \citep{gestureexample12,aideddesgin22} algorithm to recognize ink gestures. He used Gaussian statistical classifier to classify any stroke user draw.  In \cite {gestureexample12}  he firstly compute statistical information about the user stroke then pass it to a Gaussian classifier that had been trained using 10-20 samples. The output of the classifier is classification of the stroke from one of the known classes.

Bayesian classifiers was used in \cite{Vibratory8} as a stroke classifier to recognize different shapes. The recognition is done using feature based statistical Gaussian distribution and Bayesian classifier techniques. The trainer use nine features extracted from segmented symbols (ex number of line segments, number of L intersections, ) all feature has a tolerance within. After all nine feature are computed a Gaussian statically model is constructed.

Some systems employ more than one statistical classifier to get better results for the recognition. For example \cite{zernike61} uses Neural network, SVM (support vector machine )and MMD. \cite{polygonfeedback31}
%the closed shaper recognizer which is done using classifiers (SVM , Rule based and ANN ) to make a 
%classification of triangle, rectangle or ellipse

\subsubsection{Other hybrid methods}
\label{sec:Other hybrid methods}

Due to the complexity of the sketch recognition problem, more than one system proposed a hybrid method to solve those problems. They combine more than one approach to recognize symbols the user draws.
\cite{sketchunderstanding1,Simusketch51}  used four kind of classification image, graph matching, statistical and structure based classifiers. Each symbol will contain all graph, statistical, structure information and a (Wight, recognizer) tuples to determine which recognizer has the most influence in recognizing this symbol see \textit{Fig\ref{fig:structuralscemanticdeff}}.
\begin{figure}
	\centering
		\includegraphics[scale=0.5]{../../neededfiles/Figures/structuralscemanticdeff.eps}
	\caption[Structure definitions]{Example of the structure definition of sin symbol in \cite{sketchunderstanding1}.}
	\label{fig:structuralscemanticdeff}
\end{figure}

\cite {HMM53}  firstly normalize the stroke and compute all the statistical features from it. After that, a noise reduction algorithm is used on the strokes which will be translated to produce size and rotation invariance data. Those data is passed to a multistage hybrid recognizer. The first step of his multistage classifier is dynamic programming procedure to compute the optimal degree of similarity and differentiation of symbols. Secondly, a statistical feature ranking is employed on the data produced. Finally, the system will reject all unlikely classification and produce best fit for this stroke using a dynamic programming using domain and context knowledge as input.

\cite{statisticalparsing26} is a statically parser make use of the stored rules and probability of drawn stroked and uses Bayesian classifiers to do the classifications.

\subsection{AI based Classification}
\label{sec:AI based Classification}

Most system mentioned before use image or statistical feature based on the points the user draw. The next few sections will present methods that will use AI approached to classify the symbols the user draw.  Systems using AI approached imply a higher level of understanding of the strokes the user draws. Some construct a graph schema of the symbols from strokes the user draw. Other tried generating a visual grammar to identify the symbols that can be drawn.

\subsubsection{Graph Searching }
\label{sec:Graph Searching }

\textit{Figure \ref{fig:squarescematic}} shows a semantic graph that is constructed by the trainable recognizer in  \cite{mulitstroke5} for a rectangle. The trainable recognizer construct the semantic network at train time with then match the symbol user draw with the networks in the database. To match the networks quickly the system assume the user draws the symbol with the same order as in training set. A deep heuristic graph search will be employed if there was no match found.

\begin{figure}
	\centering
		\includegraphics[scale=0.7]{../../neededfiles/Figures/squarescematic.eps}
	\caption{square scemantic graph}
	\label{fig:squarescematic}
\end{figure}

\cite{SRGraph57}  Uses two kind of graphs relations a spatial relationship graph (SRG) which is build on defining the spatial relationship between the source object and candidate object on the sketch. The relations are defined using Spatial Adjacency Relation (SAR) that falls into relation previously computed in system.   To make the graph fully descriptive the system computes Relative Position Relation of SAR relations and adds these relations into the SRG graph.

\subsubsection{Rule based }
\label{sec:Rule based }

The general algorithm of rule based system as used in \cite{interpretationmechanical50,SmartSketch56}  based on a knowledge of relation ship between nodes (strokes) a role is defined using those relation ships. The rule based system check the strokes the user draw against all strokes to detect if there a rule match the drawn symbol. When a rule matches the consequence of this rule, apply on the input symbol.

The rule system used in \cite{sketchinginterfaces2}  uses the (UI user interface) domain knowledge to do its basic recognition. It is divided into two step first is testing if the rules applies on the input symbol. Second step, is ranking the rules that apply on the symbol based on the highest confidence that was given to this input symbol.

Some system like in \cite{SmartSketch56}  used the rules to help in noise removal. Furthermore, rules based system are used to employ some reasoning to the recognized symbols with the help of context knowledge \cite {interpretationmechanical50}.

\subsubsection{HMM }
\label{sec:HMM }

 The main idea of HMM is representing knowledge and recognition in serious of states each state has its input conditions. Only a node that has the input conditions of the destination state that can transit it from one state to another. The final state is the recognition state of a given class. For example to recognize a rectangle the first state should be first line detected, second will be second line and the condition is the recognition of either a perpendicular line or parallel line to first one to another.

In sketch, recognition HMM was used to recognize symbols to different classes. Each class recognized was modeled using a single HMM model. The system in \cite {HMM53} encode the geometrical primitive that construct the symbol the user draws. After the encoding the codes are passed to HMM based classifier. Each object is represented by a HMM model and to manage multiple classes they combine matching score from different HMM models.

\subsubsection{Graph language and visual language describing }
\label{sec:Graph language and visual language describing }

Researchers tried to generate a visual language that can describe hand drawn sketches. In such systems, sketch recognition will be a question of parsing the sketch document into recognizer that parse strokes according to the constructed grammar.

Researchers proposed more than one visual grammar to use it in sketch recognition. \cite{statisticalparsing26} proposed a language parser that will parse sketched strokes using grammars rules that was constructed . The system first construct a parser that will parse drawn strokes into tokens then formulate those tokens using the grammar rules stored. Parsing is done in two steps; Firstly, the system should accept input of strokes then it use a statistical parser to generate the parse tree. The statically parser make use of the stored rules and probability of drawn stroked and uses Bayesian classifiers to produce the parse tree.

%\citep*{XPGParser59} i need author   
Introduced a grammar they called an extended positional grammar (XPG) that is based on XpLR parsing methodology \cite{XPGParser59} . The XPG imagine the symbol with a set of attribute as a sentence that can be parsed using the physical, syntactic and semantic information in sketch. They divided the grammar into two kind the first is named ink grammar, it define the symbol of the language as a composition of primitive geometric shapes. The second is language grammar that they use to identify the sketch as a composition of shapes defined by the ink grammar relation. The parsing itself is done using a incrementally bidirectional parsers. They also add a sketch editor to the system to help user to add new symbols and generate its XPG grammar for it \textit{(see Fig[\ref{fig:symboleditor}])}.

\begin{figure}
	\centering
		\includegraphics[scale=0.7]{../../neededfiles/Figures/symboleditor.eps}
	\caption[symbol editor]{The Figure shows the symbol editor that\cite{XPGParser59} uses to generate grammar rules.}
	\label{fig:symboleditor}
\end{figure}

Relational grammar was also used to help in reducing the ambiguity with the help of fuzzy logic\cite {visualpattern43} . The fuzzy set is used to resolve the visual relations of the sketched stroke with others strokes. Spatial relations like (left of, right of) are computed and then passed to the fuzzy sets. The fuzzy set associate a degree of certainty to the shape defined. 
%EfficientAbstract39 also used the grammar language ladder
\cite{Ladder30,GenericHMM28}  Used visual description language which will generate the information of the structure and constrain of each object classified. The system automatically generates java code for the recognizer for each object. The generated code cycles through the strokes drawn to check if any of the recognizers generated had interpreted any symbols partially or completely. HMM network is used to partially segment the strokes drawn to help the recognition process.

Similar system was constructed in \cite {Ladder30} , where language grammar is written to descript complex shapes ( see \textit{Fig\ref{fig:arrowladderdef1}}) form primitive geometrical, constrains and triggers. The grammar is then passed to "jess" a java rule bases system that produce java code used in recognition of  drawn strokes.

\begin{figure}
	\centering
	\begin {center}
		\subfigure{An Arrow}
			\centering

			{\label {fig:arrow1}\includegraphics[scale=0.75]{../../neededfiles/Figures/arrowladderdef1.eps}}
		\subfigure[Arrow grammar definition]{\includegraphics[scale=0.88]{../../neededfiles/Figures/arrowladderdef2.eps}}
		\end {center}
	\caption[Arrow definition]{The figures show and regular arrow with 2 shafts and head and its definition in Ladder language \cite{Ladder30}}
	\label{fig:arrowladderdef1}
\end{figure}

%\subsubsection{Fuzzy classifiers }
%\label{sec:FuzzyClassifiers}
%Fuzzy classifiers is used to help in identifying degree of certainty given to certain interpolation the

\subsection{Hybrid Systems}
\label{sec:Hybrid Systems}

Given that the sketch recognition of hand drawn strokes is exceptionally challenging problem the researcher tried to take advantage of additional of combining approaches to achieve higher recognition rate.

Most of the problems of the sketch recognition are not still addressed. To do so, The sketch recognition system must maintain reliably interactive and extensible interface.  The users minor change can cause the recognizers to fail. The user sloppy drawing with over or/and under shoot strokes  \textit{Fig\ref{fig:overstroked}}. Over processing of data and interaction with background are still major problems in the sketched documents.

To solve some of these problems,\cite{threeproblmes23}  revealed a two stages recognizer that will firstly construct a data graph in which strokes will be added. After constructing the full data graph the recognizer compute all possible solutions and reject any inconsistent alternatives. Second stage, Employs a topology structure based on the geometrical matching using a CSP making use of sub graphs.

As mentioned before,\cite{visualpattern43,Cali63} combine fuzzy logic sets with language grammar to create more robust classifiers. Grammar rules are used to describe the spatial relationships of the strokes. Fuzzy sets are used to handle the uncertainties of the relation and classification. Therefore, by combining Fuzzy set and spatial grammar rules the system was able quantify the spatial and statistical attribute calculated before in \ref{sec:Statisticalmethods} from the drawn symbols.

Also the use of HMM with generated visual grammar was introduced in the system by \cite{HMM53} where they use HMM as a stroke segmentor and pass the output to the visual parser. furthermore, \cite{SmartSketch56} used more than one recognizers to recognize closed shapes rule based approach with hypothesis and test strategy, Neural network and SVM.

\paragraph{Dynamic programming}
\label{sec:DynamicProgramming}
 was used by \cite {Mathpad46,templatefrag21}  to calculate the optimal degree of different and similarity of symbols. The system also uses a dynamic programming procedure to find a best fit from the given set of classifications.

\section{Second Step: Sketch Understanding}
\label{sec:SecondStep:SketchUnderstanding}

Sketch understanding is concerned with converting a set of recognized symbols into a whole document that is fully consistent with each other. Mainly the recognized symbols are a list of best candidates of that symbol for which the sketch understanding process must choose best fit based on the document structure.

After recognizing the symbols drawn, Understanding the sketch document involve additional processing. In spite of various recognition systems that had reached a satisfactory symbol recognition. The sketch understanding problems are at a standstill.

The recognized symbols must be verified with the rest of the document to fully understand the whole sketch. Consistency, ambiguities and uncertainty of the recognized symbols are the major issues that rise in any sketch understanding system.  
Most recent sketch understanding system uses more than just the context knowledge that is always available. Domain knowledge \cite {geometrydomain49}, Speech and verbal description of drawn sketch was used in\cite {speechMulti27}.Some verbal description of the drawn symbol helps in creating the casual model for each symbol.

In the following sections, the ambiguities and uncertainty of sketch understanding systems will be discussed. Followed by how those systems represented knowledge that they use to formulate the understanding process.

\subsection{Handling uncertainty and ambiguities }
\label{sec:Handlinguncertaintyandambiguities }

The main tasks of the sketch understanding are to handle higher-level ambiguities of the sketch drawn and to manage uncertainty of the recognized symbols. After the user draw a symbol using a single or multi-stroke symbol the recognizer labels each symbol with a known class either low level interpolation or high level. Low level interpolation means simple geometrical shapes and high level means more complex domain specific. Most current recognizers return a list of acceptable label for each symbol. The role of the sketch understating is converting such list of acceptable labels into one interpolation based on the surrounding symbols. They do so with preserving the sketch consistency of interpretation, which means that, the whole document has the same interpretation.

The next section the various methods that was recently employ to completely understand sketched documents.

\subsubsection{Rank and prune based on classifiers}
\label{sec:Rankandprunebasedonclassifiers}

The most used method in understanding the whole document and reducing the inconsistent solutions is the typical method of ranking the solutions and then pruning unwanted ones. 
Initially the symbols nodes are initialized by all possible pairs of interpolation of the two symbols then the Consistency checker eliminate all inconsistence pairs. \cite {tobeadd} \cite {geometrydomain49}  firstly recognize all possible interpolation then prune and reason using knowledge to get a score for each interpolation. The final step is resolving and choosing the best result.

 \cite {Alvarado2002Framework11,HierarchicalParsing7}  First uses a bottom-top search to look for the interpolations of the strokes drawn. A top bottom search is used to prune out incomplete interpolations that was generated in the previous step.  The top bottom firstly checks what symbol may be missing in the current document context after that it prune inconsistent interpolations.
\subsubsection{Constrain satisfaction methods}
\label{sec:Constrainsatisfactionmethods}

Constrain satisfaction is the most used method for converting a group of recognized symbols into an understood document. Most recognizers produce a ranked set of possible interpolations. The problem is to constrain theses interpolations to get a total  understanding of the sketch document. Some like in \cite {tobeadd}  use constrain satisfaction methods to apply rules and constrains to match interpolations and rules. Constrains must be carefully written to check that they are not rough or severe. \cite {tobeadd}  checked the constrains applied on this system and if they produce no match the system relax the constrain and runs again with more loss constrains till it find the correct interpolations.

\begin{figure}
		\includegraphics[scale=0.7]{../../neededfiles/Figures/CSPphysical.eps}
	\caption[Physical Constrains on Symbols]{ The figure shows two nodes and the constrain applied on both nodes to detect the inconsistency in the interpretations \cite{physicalmeaning6}. }
	\label{fig:CSPphysical}
\end{figure}

 \cite {physicalmeaning6}  firstly try connection between symbols firstly using bounding boxes of the symbol to locate the intersecting symbols. The intersected bounding boxes are checked again for ten different types of connection approximately for each terminal of the symbol. After identifying which symbols are connected the problem is identify which interpolation make those connected symbols has physical meaning. This is a "constrain satisfaction" problem the system used two processing steps to solve it. Firstly, the system checks the consistency of node where the node in graph is connection between two adjacent symbols. That means that each two connected symbol are checked and every inconsistent interpolation is rejected and deleted. This generate a local consistent sub graph but there may be a global inconsistent which take us to the second step where the algorithm search the arc consistency. The arcs of the consistency graph are any pair of nodes that have a symbol in common.  Each consistency arc is checked for local and global consistency. This step is applied iteratively till all consistency are cancelled and only one global interpolation is reached \textit{(see Fig[ \ref{fig:CSPphysical}])}.

 The use of constrain satisfaction has also helped in minimizing the search \cite {tobeadd} . The strokes are incrementally processed using the Constrain system which check for constrain between image feature. Each new stroke either a new partial interpolation is created or an old one is extended and cloned provided that no constrained is violated.

\subsubsection{Bayesians networks}
\label{sec:Bayesiansnetworks}

Since the problem of deciding the best interpretation of a set of valid ones is a certainty problem researchers tried to solve it as a one. Bayesians networks were used to represent certainties in sketch documents\cite{SketchRead2007} .Bayesians classifiers were also used to help the understanding problem in various ways.\cite{statisticalparsing26}  Used Bayesians classifiers to help the statistical parser to recognize the best candidate based on the system rules.

\cite{Alvarado2002Framework11,SketchRead2007} is a complete sketch-understating framework that used top bottom and bottom up search of interpolation of the strokes. The system represented uncertainty using Bayesians networks. Each compound object is represented by a network in which the object is the root \textit{see Fig[\ref {fig:anddefnetwork}]}.
 \begin{figure}
		\includegraphics[scale=0.7]{../../neededfiles/Figures/anddefnetwork.eps}
	\caption[Bayesian network]{Bayesian network of an And Gate as defined in \cite{Alvarado2002Framework11}}
	\label{fig:anddefnetwork}
\end{figure}

\subsubsection{Other Methods}
\label{sec:OtherMethods}

Other Method were used in \cite{agent47,sketchinginterfaces2}to understand the sketch documents . For example \cite{sketchinginterfaces2} defined  visual and domain rules that applies on the sketch to finally understand the whole document. \cite{agent47} used software agents to resolve the conflict in curve matching algorithms he used. In addition to agents that resolve the conflict he also used agents to search for applicable features on the strokes. The agent will either report a feature on the strokes drawn or disappear till new strokes are drawn.

\subsection{Representing knowledge }
\label{sec:Representingknowledge }

Representing knowledge is the one of the challenges that faced the researcher in any AI program. In sketch recognition and understanding, the knowledge can be from more than one resource. The real problem is not to dump the program in too much data knowledge that is unstructured or unused in processing. To do so, means that the system is exhausted in undesired and unfulfilled processing. Processing time in sketch understanding system must be as close to real-time as possible to preserve the user interaction with the system. Therefore, the knowledge must be represented in special method that insures high speed and accuracy of understanding of the drawn sketch.

Systems tried to make uses of more than one kind of knowledge to help in there recognitions and understandings. Spatial and visual knowledge of the drawn symbols are widely used in the sketch systems that relies on grammar languages \cite{GenericHMM28,statisticalparsing26,SketchRead2007} . Speech and multimodal systems helped the recognition of the symbols and reduced the search space.  Domain and grammar knowledge were also used in \cite {geometrydomain49}  .

Black board architecture \cite {blackboardur} was widely used to represent the various source and forms of data in the sketch understanding systems. In \cite {Alvarado2002Framework11}  black board was used to represent incremental knowledge which increase by each symbol drawn in the sketch. The black board architecture is simply a modeling of group of people trying to solving a problem. When any one of them find new information he write it down on the black board soon all information will be in the black board for all solver to see and add.

\subsubsection{	Spatial and Visual knowledge}
\label{sec:SpatialandVisualknowlege}

The visual and spatial information of the drawn symbols with respect to other symbols in the sketch helps in identifying the relation of these symbols with each other.  Those information was used in \cite {Cali63,visualpattern43,incrmentintention41,geometrydomain49,marksacts40,sketchinginterfaces2,captureknowledge19}  to help the recognition and understanding of sketch drawn.

\cite{SRGraph57}  Introduced a new spatial relationship graph to represent those spatial relation of one symbol to the other. The Spatial relation graph is build from the sketch drawn using primitive spatial adjacency relation and relative position relation.  The produced graphs are then matched using a graph-matching algorithm with the database of graph to recognize best fit.

The spatial and visual information are used in \cite{XPGParser59} to help in building a parser that will parse the sketch document into symbols and objects.

\subsubsection{Domain and Context knowledge}
\label{sec:DomainandContextknowledge}

\begin{figure}
	\centering
		\includegraphics[scale=0.8]{../../neededfiles/Figures/domainexamples.eps}
	\caption[domain examples]{Examples of sketched documents from different domains}
	\label{fig:domainexamples}
\end{figure}

Users can draw sketches in different domains \textit{see Fig [\ref{fig:domainexamples}]}and Until now there have not been a sketch understanding system that can identify drawing in any sketched domain. To do so, the system must be multi-domain system and employ a kind of domain identification or classification.  Current systems uses domain and context knowledge to assist in the understanding procedure\cite {geometrydomain49,RetargetableInteractive58,sketchunderstanding1, HierarchicalParsing7,Mathpad46} .

ASSIST \cite{AlvaradoFreedom42} uses domain knowledge, context memory and user feedback to help in understanding procedure. Domain knowledge to get a set of possible interpretations then context knowledge is used to prune out inconsistent finally the user feedback is used to correct error interoperations.

The \cite{incrmentintention41} store domain knowledge in a memory like architecture where It stores the domain knowledge in long-term memory were strokes are firstly checked for possible interpolations. Context knowledge and previous interpolations are stored in short-term memory to help more accurate recognition rate. The context and domain knowledge are also used to feed rule-based systems that uses visual language rules in making the final document recognitions \cite{interpretationmechanical50}.

%&&&&&&&&&&&&&&&&oksearch 
\subsubsection{Using speech and various other knowledge}
\label{sec:Usingspeechandvariousotherknowledge}

Multimodal systems are system that uses more than one source of input devices as input to the system. Theses systems help in providing sketch understanding with more information's and data that will help in the recognition procedure. Speech and verbal explanation of drawn symbols are most used data in multimodal system to aid sketch understanding \cite {cognitivesketch18,speechMulti27,computationalmodel16} .

\subsection{Space Search}
\label{sec:SpaceSearch}

The solutions of any sketch understanding are not found until the problem of minimizing and find an efficient way to search the solution space is solved.  The solution space increase exponentially with each stroke the user draws dependably on the system. Either ways the space search problem must be addressed to properly solve the problem of sketch understanding.

\cite {EfficientAbstract39}  illustrate a new search space minimization where the recognizer can perform one of the checking for constrains between images feature. Check for partial interpretation by assigning an image feature to a model feature or the user draws delaying until more strokes. Every new stroke is drawn either a new partial interpolation is created an old interpolation is extended and cloned provided that no constrains or rule is violated.

### 2008-09-02 draft/ConFut.tex
First add of the master draft

\chapter{Discusion and Conclusion}

\section{  Conclusion}
\label{sec:ConclusionConclusion }

\section{Future Work}
\label{sec:FutureWork}

### 2008-09-02 draft/System.tex
First add of the master draft

\chapter{Proposed System for Sketch Understanding}
\section{Aims of the System}
\label{sec:AimsOfTheSystem}
The previous chapter shows the current trends in sketch recognition. This thesis introduce a solution for the problem that give optimize solution for the segmentation problem.

\section{An Overview of the System}
\label{sec:AnOverviewOfTheSystem}

\section{System Components}
\label{sec:SystemComponents}
% the system and discription of this

\chapter{Preprocessing and Segmentation}
\label{sec:Preprocessingandsegmentation}

\section{Preprocessing}
\label{sec:Preprocessing}

\subsection{Curvature Calculation}
\label{sec:CurvatureCalculation}

\subsection{Critical Point Detection}
\label{sec:CriticalPointDetection}

\section{Segmentation}
\label{sec:Segmentation}
\subsection{Swarm Introduction}
\label{sec:SwarmIntroduction}

\subsection{Ellipse Detection }
\label{sec:EllipseDetection}

\subsection{Swarm Segmentation}
\label{sec:SwarmSegmentation}
\subsubsection{Problem Formulation }
\label{sec:ProblemFormulation}

\subsubsection{Fitness Computation }
\label{sec:FitnessComputation}
\subsubsection{Polygon Division Algorithm}
\label{sec:PolygonDivisionAlgorithm}

\subsubsection{Curve Divisions Algorithm}
\label{sec:PolygonDivisionAlgorithm}

\section{BenchMarck algorithm}
\label{sec:BenchMarckAlgorithm}

\chapter{Symbol Recognition}
\label{sec:SymbolRecognition}
\section{Stroke Clustring Algoirthm }
\label{sec:ClustringAlgoirthm}
\section{ Feature Extraction}%%Feature Set
\label{sec:FeatureExtraction}

%%%%%%%%%%%%%%%%%%%55555555
%Symbol Recognition 16
%4.1 Preprocessing and segmentation . . . . . . . . . . . . . . . . . . . . . . . . . . . 16
%4.2 Feature-Based Statistical Symbol Recognition . . . . . . . . . . . . . . . . . . . . 18
%4.2.1 Feature Set . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . 19
%4.2.2 Training . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . 21
%4.2.3 Recognition . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . 23
%4.2.4 User Study . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . 24
%4.3 Graph-Based Symbol Recognition . . . . . . . . . . . . . . . . . . . . . . . . . . 28
%4.3.1 Previous Work . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . 28
%4.3.2 Statistical Graph Definitions and Error-driven Stochastic Graph Mathcing . 29
%4.4 Image Based Symbol Recognition . . . . . . . . . . . . . . . . . . . . . . . . . . 30
%4.4.1 User Study . . . . . . . . . . . . . .
%3.3 System Components 43
%4 Segmenting a Freehand Sketch
%4.1 Interface 51
%4.2 Spline representation 52
%4.3 Points of potential feature separation 59
%4.4 Corner detection 59
%4.5 Intersection detection 61
%4.6 Clustering of interesting points 65
%4.7 Feature Extraction

### 2008-09-02 trunk/draft/Results.tex
First add of the master draft

\chapter{Results}

\section{Experiments Setup}
\label{sec:ExperimentsSetup}
\section{Datasets}
\label{sec:Datasets}
\section{Evaluations techniques}
\label{sec:EvaluationsTechniques}

\section{Segmentation Algoirthms Results}
\label{sec:SegmentationAlgoirthmsResults}

\subsection{Swarm Algorithms}
\label{sec:Swarm Algorithms}

\section{Clustering Algorithm Results}
\label{sec:ClusteringAlgorithmResults}

\section{Conclusion}
\label{sec:Conclusion}

\section{ BenchMark Results }
\label{sec:BenchMarkResults}

### 2008-11-23 trunk/writing/outline.tex
the paper before i edit after dina requested changes 2

--- 2008-08-18 writing/outline.tex
@@
 %opening
-\title{ Symbol Recognition using Particle Swarm Algorithms}
+\title{Symbol Recognition using Particle Swarm Algorithms}
 %\title{A Study on Using Particle Swarm Algorithms in Symbol Recognition }
@@
 \begin{abstract}
-Sketch recognition is defined as the process of identifying the symbols the user draws using a single or multiple strokes.
-Users draw strokes using a pen and the system immediately interprets their strokes into objects that can be easily manipulated.
+Sketch recognition is defined as the process of identifying symbols that users draw using single or multiple strokes.
+Users draw strokes using a pen and the system should immediately interprets their strokes into objects that can be easily manipulated.
 This paper uses Particle Swarm Algorithm (PSO) to divide the strokes the user draws into meaningful geometric primitives.
-These geometric primitives are grouped to formulate symbols which will be further identified.
-The results showed that using PSO improves segmentation results which guide the symbol recognition phase.
-Final recognition has improved accuracy after the PSO algorithm using a SVM classifier.
+These geometric primitives are grouped to formulate symbols which are further identified.
+The results shows that using PSO improves segmentation results which guide the symbol recognition phase.
+This paper uses SVM classifier which  further improves the final recognition accuracy.
+%u Final recognition has improved accuracy after the PSO algorithm using a SVM classifier.
  %Hand drawn sketches are the natural method of exchanging thoughts between engineers and humans in general.
@@
 Engineers use sketches to exchange designs as a natural method of communication rather than writing or speaking.
-Currently, engineers use paper and pencil design in early stages of design.
-Design engineers need a powerful computer based system with symbol recognition to help them design, manipulate and store sketchs more effectively than with only papers.
-Increasing the interaction between the computer and the users in sketch and CAD systems has been recently the reason for the emerging of few advanced sketch recognition systems.
-Sketch recognition system converts the drawings and informal graphs the user draws into objects and symbols can be used in further interaction with the system.
+Engineers use paper and pencil design in early stages of design.
+Design engineers need a powerful computer based system with symbol recognition to help them design, manipulate and store sketches more effectively than using only papers.
+Increasing the interaction between computers and users in sketch and CAD systems has been the reason for the emerging of few advanced sketch recognition systems.
+Sketch recognition system converts the drawings and informal graphs user draws into objects and symbols that can be used in further interaction with the system.
 
-Sketch recognition researchers divide the system into four main steps preprocessing, segmentation, symbol recognition and finally sketch understanding.
-The preprocessing process saves the location of points from the hardware and collects basic information about the stroke data then proceed to remove noise and primary data calculation.
-In segmentation step, the system  divide the stroke into a set of simple geometrical primitives or segments.
-The third step of the system cluster strokes and segments to formulate symbols that can be recognized by a classifier system.
-Finally, the sketch understanding step uses priori information about the sketch to support identification of symbols by excluding any possible invalid symbols.
+Sketch recognition is divided into four main steps preprocessing, segmentation, symbol recognition and finally sketch understanding.
+The preprocessing process saves user input stroke points from the hardware and collects basic information about the stroke then proceeds to remove noise and primary data calculation.
+In segmentation phase, the system  divides the stroke into a set of simple geometrical primitives or segments.
+The  third phase, sketch recognition, is clustering strokes and segments to formulate symbols that can be recognized by a classifier system.
+Finally, the sketch understanding phase uses priori information about the sketch to support identification of symbols by excluding any possible invalid symbols.
 %the basic symbols are identified in consistent with the whole sketch. \\
@@
 \\
+This paper introduces a sketch recognition system.
+The presented system uses particle swarm algorithm(PSO) to optimially segment users strokes.
+Users can draw symbols using any number of strokes in any order they like.
+The system segments stroke user draws then cluster symbols together then passes them to the SVM symbol recognizier.
+The experiments in section \ref{sec:Experiments} shows that the proposed system improves the final symbol recognition accurracy.
 
-Section \ref{Sec:RelatedWork} provides brief description to related work.
-The remaining of this papers will be as following: Section \ref{PSO} will explain the general particle swarm algorithm.
-Section \ref{Sysdisc} will describe the system in details.
-Section \ref{sec:Experiments} will describe the experiments that was preformed.
-Finally, section \ref{ConclusionandFutureWork} will state the conculusion and future work.
+The remaining of this papers is organized as following; section \ref{Sec:RelatedWork} provides a brief literature review.
+Section \ref{PSO} explains the general particle swarm algorithm which is used in the presented system.
+The proposed system block diagram along with a detailed system discribtion is presented in Section \ref{Sysdisc}.
+The experiments and results achieved are explained in Section \ref{sec:Experiments}.
+Finally, section \ref{ConclusionandFutureWork} states the conclusion and future work.
 %The next section will relate previous work in the field.
@@
 The field of symbol recognition has gained interest in the last few years. A wide variety of techniques were used either on segmentation process or symbol recognition.
-
 %To segment strokes
-In order to segment strokes, \cite{meanshift10} introduced a feature area for each primitive and then computed the segmentation error for different types of primitives based on this area.
+%In order to segment strokes, \cite{meanshift10} introduced a feature area for each primitive and then computed the segmentation error for different types of primitives based on this area.
 %\citeauthor{HeloiseBeautification} used a dynamic programming algorithm to solve the segmentation problem \cite{HeloiseBeautification}.
-
-Complex and hybrid algorithm were also introduced as in  \cite{earlyprocess} where they generated different sets of segmentation based on both curvature and speed dominant points then followed by a choice a segmentation with the least error from the hybrid set.
-Additionally, a dynamic programming algorithm was used to obtain optimum segmentation of the input strokes in \cite{GeometryAndDomain102}.
-Genetic algorithm was also used by \cite{CruveDivisionSwarm} to optimally divide digital curves into lines and curves.
-\citeauthor{PolygonApproximationPSO} used PSO to convert digital curves into polygons\cite{PolygonApproximationPSO}.
+A hybrid algorithm was introduced in \cite{earlyprocess} where they generated different sets of segmentation based on both curvature and speed dominant points, this is followed by choosing a segmentation with the least error from generated hybrid set.
+Their system has a draw back as it only recognizes simple specific shapes with a set of low level recognizers.
+\cite{meanshift10} introduced a feature area for each primitive and then computed the segmentation error for different types of primitives based on the computed feature area.
+Their system achieved good accuracy in simple shapes (square, ellipse,...) but did not perform well in more complex shapes\cite{meanshift10,Paleosketch08}.
+\citeauthor{Paleosketch08} \cite{Paleosketch08} introduced a set of low level recognizers that were reported to achieve 98\%  accuracy but their system simillar to all low level recognizers only identify a small set of simple shapes.
 
 Many different approaches have been investigated to achieve final recognition of symbols.
-Graph searching and template matching where used in \cite{mulitstroke5,GraphBased07}, where a graph of the drawn symbol is generated and matched with a stored set of graphs templates to reach a classification decission.
-Geometrical and spatial descriptive language was used in \cite{SketchRead2007,Ladder30} where an architecture simillar to compiler was used to recognize symbols.
-\citeauthor{imagetrainable48} used trainable image template to classify different set of symbols \cite{imagetrainable48}.
-The use of SVM and HMM classification was implemented in\cite{SVM300,SmartSketch56,HMM53}.
+Graph searching and template matching where used in \cite{GraphBased07} where a graph of the drawn symbol is generated and matched with a stored set of graphs templates to reach a classification decision.
+These systems may generate good performance but they are very computationally expensive.
+Geometrical and spatial descriptive language was used in \cite{SketchRead2007,Ladder30} where an architecture similar to compiler was used to recognize symbols.
+%%SVM and HMM classification was implemented in\cite{SVM300,HMM53} to correctly classify symbols.
+
+Genetic algorithm was used by \cite{CruveDivisionSwarm} to optimally divide digital curves into lines and curves.
+\citeauthor{CruveDivisionSwarm} uses digital curves scanned from paper as input to the system and did not take advantage of the curvature or local geometric properties of the digital curve.
+\citeauthor{PolygonApproximationPSO} \cite{PolygonApproximationPSO} used PSO to convert digital curves into polygons, our system adoptes \citeauthor{PolygonApproximationPSO} system but improves it by adding curvature and other local information while segmenting strokes to achieve better segmentations.
+The next section presents the general particle swarm algorithm.
+The proposed swarm segmentation algorithms are peresented in section \ref{subsubsec:Discreteparticleswarmalgorithm}
 % Other techniques as  [ ] was used  in []. \\  
@@
 The main idea of  Particle Swarm Algorithm (PSO) is to represent each agent with a particle from the solution space.
-Each agent moves the particle with a direction and velocity based on the following equation [\ref{eq:Swarm}].
+Each agent moves the particle with a direction and velocity based on equation \ref{eq:Swarm}.
+ %Equation [\ref{eq:Swarm}] shows how velocity and direction of each particle are computed
  \begin{equation}
-%\[
 v_{ij}  = v_{ij}  + c_1 r_1 (lbest_{ij}  - p_{ij} ) + c_2 r_2 (gbest_{ij}  - p_{ij} )
-%\]
 \label{eq:Swarm}
 \end{equation}
+ where $r_1$ \& $r_2$ are random variables and $c_1$ \& $c_2$ are the swarm system variables.
 \begin{equation}
@@
 p_{ij}=p_{ij}+v_{ij},
-%\]
+%\
 \end{equation}
- Equation [\ref{eq:Swarm}] shows how velocity and direction of each particle is computed where $r_1$ \& $r_2$ are random variable and $c_1$ \& $c_2$ are the swarm system variables.
+where $p_{ij}$ represent the $jth$ particle in the $ith$ agent and $v_{ij}$ is the velocity of the $jth$ particle in the $ith$ agent.
 Each iteration the global best $g_{best}$ particle and the agent local best $l_{best}$ particle are evaluated based on the maximum fitness functions of all particles.
-The solution is found after a maximum number of iteration or after an error threshold is achieved.
+The solution is found after acheiving a specific number of iteration or after an error threshold is achieved.
+The following equation \ref{eq:descrite}  
 \begin{equation}
@@
 \end{equation}
-  Equation \ref{eq:descrite} where t is a threshold and $r_{3}$  is a random variable, is used to change general swarm algorithm into binary particle which will handle particle values of either 0 or 1.
-.
+ (where $t$ is a threshold and $r_{3}$ is a random variable), is used in order to change the general swarm algorithm into binary particle which handles particle values of either $0$ or $1$.
+
 \section{System Description}
@@
 %The stroke is the path of points from the instant the pen is down till it is up.
-Figure. \ref{fig:Blockdiagram} shows that the main system blocks are 1) preprocessing, 2) segmentation, 3) clustering and feature extraction, and finally 4) classification.
+Figure. \ref{fig:Blockdiagram} shows that the main system blocks are 1) preprocessing, 2) segmentation, 3) clustering, 4) feature extraction, and finally 5) classification.
 
- %Due to sloppiness of users and hardware glitches the system preprocesses the data before further processing. 
+ %Due to sloppiness of users and hardware glitches the system preprocess the data before further processing. 
 The preprocessing is responsible of capturing the input data and removing the noise from it.
@@
 \\% This part is repeated for each stroke.
- The clustering alogrithms starts to group segments togther after the segmentation step.
-The system let the user draws the symbol by using any number of strokes, a set of unrecognized segments is passed to the clustering algoirthm to generate a symbol and compute a feature vector for it.
+ The clustering algorithms starts to group segments together after the segmentation step.
+The system let the user draws the symbol by using any number of strokes, a set of unrecognized segments is passed to the clustering algorithm to generate a symbol and compute a feature vector for it.
+The system uses a hybrid set of statistical, geometric and spatial features.
 The final step is a SVM classifier that will use the features computed to classify the segments into one of the previously trained classes.
 %  Or to determine the symbol of the given segments from the set of preciously trained symbols.
-
  %with a symbol from the training set.% The system compute composite set of features some are statistical other are spatial features based on the type primitives.
 Finally, the strokes are classified into the corresponding classes.% the classifier identifies the strokes into a symbol from the set of known symbols.\\
+
 \begin{figure}[]
+%\begin{minipage}[b]{0.5\linewidth}
 	\centering
-		\includegraphics[scale=0.35]{images/Blockdiagram.pdf}
-	\caption[  ]{The system Block Diagram}
+		\includegraphics[scale=0.85]{images/block2.pdf}%Blockdiagram.pdf}
+	\caption{The system Block Diagram.} The figure shows the main system blocks. Preprocessing, segmentation, clustering,  features extraction and classification.
+
 	\label{fig:Blockdiagram}
+%	\end{minipage}
 \end{figure}
+
 \subsection{Preprocessing}
@@
  %$\Delta d/  \Delta S $ where d is the difference direction of point and s is difference in distance between points. 
- All these calculations are performed in real time while the user draws the strokes.
- The complexity of compution is $O(n)$ where $n$ is number of points. \\
+ All these calculations are performed in real time while the user draws the strokes. The complexity of computation is $O(n)$ where $n$ is number of points. \\
  Figure \ref{fig:speed2} shows the computed data for the stroke drawn in fig.
@@
 Using simple differentiations to detect local extreme points resulted in false points due to the non smooth curves.
-So instead the following procedure was used, firstly the mean of the curve is used as threshold for lower values then the high values are further investigated to find the extreme points.
-Hence, the mean of the direction curve is calculated and used as threshold.
+Hence, the system adopted a process presented by \cite{earlyprocess}, where the mean
+%So instead the following procedure was used, firstly the mean of the curve is used as threshold for lower values then the high values are further investigated to find the extreme points.
+Hence, the mean
+of the curve is calculated and used as threshold.
 This threshold is used to separate the curve into regions; each region is defined as the part between two intersection points of the threshold line and the curve.
-Only regions higher than the threshold value are processed when computing the maximum of each region.
-The point that corresponds to those maximum values are labeled possible dominant points.
+Regions higher than the threshold value are processed to find the maximum of each region.
+The points that correspond to those maximum values are labeled as \textit{possible dominant points}.
 \\
@@
 		\includegraphics[scale=0.3]{images/orignalStroke.pdf}
-	\caption{Stroke Sample}
+	\caption{Stroke Sample} An example of the stroke drawn in the system.
 	\label{fig:orignalStroke}
@@
 All the points labeled as possible dominant points are saved into a single array.
-They are used as input and guiding data for the next stage of segmentation.
+They are used as input to guide the next stage of segmentation.
 \begin{figure*}[]
-\begin{minipage}[b]{0.5\linewidth}
+%\begin{minipage}[b]{0.9\linewidth}
 	\centering
 
-		\subfigure[ The Speed of data] {\includegraphics[scale=0.15]{images/speed2.pdf}}
-
-		\subfigure[ Direction ] {\includegraphics[scale=0.15]{images/direction2.pdf}}
-					\end{minipage}
-					\begin{minipage}[b]{0.5\linewidth}
-			\subfigure[Curvature] {\includegraphics[scale=0.15]{images/curvature2.pdf}}
-
-			\subfigure[ Time Difference] {\includegraphics[scale=0.15]{images/time2.pdf}}
-			\end{minipage}
-
-	\caption[Data Curves]%{The data of the stroke}
+		\subfigure[ The Speed of data] {\includegraphics[scale=0.25]{images/speed2.pdf}}	
+	%	\subfigure[ Direction ] {\includegraphics[scale=0.15]{images/direction2.pdf}}
+					%\end{minipage}
+					\hspace{20 mm}
+					%\begin{minipage}[b]{0.5\linewidth}
+			\subfigure[The Curvature] {\includegraphics[scale=0.29]{images/curvature2.pdf}}
+	%		\subfigure[ Time Difference] {\includegraphics[scale=0.15]{images/time2.pdf}}
+%			\end{minipage}
+	\caption{Data Curves} The curve in a) shows the speed of the points in stroke in fig. \ref{fig:orignalStroke}. b) Shows the curvature curve for the same stroke.  %{The data of the stroke}
 
@@
 \label{seg}
-After computing the primary data the system now tries to segment the stroke into a set of primitives.
+After computing the primary data the system tries to segment the stroke into a set of primitives.
 The segmentation algorithm first tries to detect if the stroke is an ellipse.
@@
 \ref{fig:Blockdiagram}).
-The next section will describe the ellipse detection algorithm and both the two particle swarm algorithms used to divide the stroke.
-\\
+The next section will describe the ellipse detection algorithm and both the two particle swarm algorithms used to divide the stroke.\\
 %Paragraphs describe the segmentation algorithm.  
-\subsubsection{ Ellipse Detection } 
+\subsubsection{Ellipse Detection} 
 %Describe the ellipse detection algorithms 
 The process starts with computing the center of the stroke bounding box.
-The bounding box center point is used as the first estimation of the center of the ellipse.
-The axes of the ellipse are estimated as the w/2 and h/2 of the stroke bounding box.
-The least square fitting algorithm is used to minimize the fitting error of the ellipse equation (see \ref{eq:circleFit} where N is number of points in the stroke, $a,b$ are the length of ellipse axes, $x_0$ \& $y_0$ are the coordinates of the center point, $x_i$ \& $y_i$ are the coordinates of point $i$ in the stroke).
-After only few loops the error and a confidence values  are computed to check if the storke can be labled as ellipse or not.\\ %is computed to check if stroke is an ellipse.
-\\
-%We found that if the confidante is above threshold then the probability of ellipse is highest otherwise the stroke is passed to the next section to get the divisions of stroke and test its error.  
+The bounding box center point is used as the first estimation of the ellipse center.
+The axes of the ellipse are estimated as $width/2$ and $height/2$ of the stroke bounding box.
+The least square fitting algorithm is used to minimize the fitting error of the ellipse equation see equation \ref{eq:circleFit}.
 \begin{equation}
@@
 \label{eq:circleFit}
-\end{equation}
-\subsubsection{ Discrete particle swarm algorithm}
+\end{equation} where N is number of points in the stroke, $a,b$ are the length of ellipse axes, $x_0$ \& $y_0$ are the coordinates of the center point, $x_i$ \& $y_i$ are the coordinates of point $i$ in the stroke.
+After only few loops the error and a confidence values  are computed to check if the stroke can be labeled as ellipse or not.\\ %is computed to check if stroke is an ellipse.
+\\
+%We found that if the confidante is above threshold then the probability of ellipse is highest otherwise the stroke is passed to the next section to get the divisions of stroke and test its error.  
+\subsubsection{Discrete particle swarm algorithm}
+\label{subsubsec:Discreteparticleswarmalgorithm}
 There are two PSO algorithm described in this section.
@@
 %formation is nearly the same in both.
-\subsubsection{ problem definition}
+\subsubsection{Problem definition}
 The input stroke with N points can be represented by set $S = \left\{ {x_1 ,x_2  \ldots x_N } \right\}$ where $x_i$ is the location of the point $i$ .
@@
 The fitness function and error calculation are different in each algorithm. For the first algorithm, 
-let's define the arc $\widehat{x_ix_j}$ as the consecutive set of points from point $x_i,x_{i+1} \cdots,x_j$. The line
-$\overline{x_i x_j} $ as the straight line connecting point $x_i$ to point $x_j$.
-The approximation error is computed by the equation \ref{eq:ErrorSwarm1}  where M is the number of dominant points in this solution.
-The error $ e ( \widehat{x_ix_j},\overline{x_i x_j})$ is computed as the sum of squared perpendicular distance from every point along the arc $\widehat{x_ix_j}$ to the line $\overline{x_i x_j}$.
-\\
+let's define the arc $\widehat{x_ix_j}$ as the consecutive set of points from point$x_i$ to point $x_{j}$ as in $x_i,x_{i+1} \cdots,x_j$.
+The line $\overline{x_i x_j} $ as the straight line connecting point $x_i$ to point $x_j$.
+The approximation error is computed by the equation \ref{eq:ErrorSwarm1}
 \begin{equation}
@@
 \label{eq:ErrorSwarm1}
-\end{equation}
-\begin{equation}
+\end{equation} where M is the number of dominant points in this solution.
+The error $e ( \widehat{x_ix_j},\overline{x_i x_j})$ is computed as the sum of squared perpendicular distance from every point along the arc $\widehat{x_ix_j}$ to the line $\overline{x_i x_j}$.
+\\
+The fitness is computed using equation \ref{eq:fitnessSwarm1}\begin{equation}
 \max fitness(p_i ) = \left\{ {\begin{array}{*{20}c}
@@
 \label{eq:fitnessSwarm1}
-\end{equation}%\]
-The fitness is computed using equation \ref{eq:fitnessSwarm1} where N is the number of points in the stroke, D is the number of point in the solution that was previously labeled as a possible dominant point and E is the computed error and $\varepsilon$ is the error threshold.
-As you can see when the error is larger than the threshold is given a -ve value to lower the value of solution otherwise the system will favor the lower number of vertices.\\ %we will want to lower the number of vertices.
+\end{equation} where $N$ is the number of points in the stroke, $D$ is the number of point in the solution that was previously labeled as a possible dominant point and $E$ is the computed error and $\varepsilon$ is the error threshold.
+Notice that when the error is larger than the threshold the fitness is given a -ve value to lower the value of solution otherwise the system will favor the lower number of vertices.\\ %we will want to lower the number of vertices.
 \\
@@
 %Alg1:
-%Alg2: 
+%Alg2: o
 The second algorithm has the same problem formulation but different fitness and error functions.
-It was previously introduced in  \cite{CruveDivisionSwarm} where but genetic programming was used as the optimizing algorithm.
+It was previously introduced in  \cite{CruveDivisionSwarm} but a genetic programming was used as the optimizing algorithm.
 The error of both circle and line estimation are computed for each segment, the approximation with the lower error value will be the chosen approximation of this segment.
 The sum of the approximation error of all segments is the total error of the particle.
-The error is computed by equation \ref{eq:errorSwarm2} where M is the number of segments in the solution, $D_i$ is the minimum approximation error of curve and line approximations $min(d_c,d_l)$ as computed by \cite{CruveDivisionSwarm}.
-The fitness is computed by the equation \ref{eq:fitnessSwarm2} where E is the error and M is number of segments and k is a parameter tweaked to get minimum number of segments.
-\\
-\begin{equation}
+The error is computed by equation \ref{eq:errorSwarm2} \begin{equation}
 E=\sum\nolimits_{i = 0}^M e(D_i) 
 \label{eq:errorSwarm2}
-\end{equation}
-\begin{equation}
+\end{equation}where $M$ is the number of segments in the solution, $D_i$ is the minimum approximation error of curve and line approximations $min(d_c,d_l)$ as computed by \cite{CruveDivisionSwarm}.
+The fitness is computed by the equation \ref{eq:fitnessSwarm2} \begin{equation}
 \max fitness(P_i ) = \frac{1}{{E \times M^k }}
 \label{eq:fitnessSwarm2}
-\end{equation}
+\end{equation} where $E$ is the error and $M$ is number of segments and $k$ is a parameter tweaked to get minimum number of segments. \\
 %Describe how we used particle swarm in curve polygon approximation using first algorithm.  
@@
 %[These two algorithm was modified from the algorithm in paper \cite{PolygonApproximationPSO,CruveDivisionSwarm}].
-\cite{PolygonApproximationPSO} used a merge and divide algorithm after each loop of the swarm system to refine the solution but we used another enhancement method.
-After each loop in the swarm algorithm, each particle loops on the set of selected dominant points to enhance the solution.
-Each dominant point is checked to find if it was labeled before as a possible dominant (computed as in section \ref{Prepross}), if not the point is moved to the nearest labeled point.
+\citeauthor{PolygonApproximationPSO} \cite{PolygonApproximationPSO} used a merge and divide algorithm after each loop of the swarm system to refine the solution but this system is introducing another enhancement method.
+After each loop of the swarm algorithm, each particle is refined using the following procedure.
+For each particle %loops on the set of selected dominant points to enhance the solution.
+each dominant point is checked to find if it was labeled before as a \textit{possible dominant point} (computed as in section \ref{Prepross}), if it is not labeled the point is moved to the nearest labeled point.
+After that the particles are tested to make sure that the distance between every two successive dominate point is larger than the constant $min_D$.
+If two points are found that they are nearer than $min_D$ then based on the decrease in the error of the solution one of the points will either be removed or moved to the nearest possible dominate point.
 %check the new solution and check the new vertices if a vertex points is a possible dominant point it is left as a vertex.
@@
 \subsection{Recognition}
-After the user draw all strokes of the symbol he has to wait 10 seconds or press finish button beside the drawing area.
+After the user draws all strokes of the symbol he has to wait 10 seconds or press finish button beside the drawing area.
 The set of unrecognized strokes is grouped together along with their segmentation as input to the feature extraction process.
-The system uses a composite set of features include Rubine feature set \cite{gestureexample12}, Zernike moments \cite{HeloiseBeautification}, ink density \cite{GeometryAndDomain102} and some structural information like number of perpendiculars lines , number of parallel lines and types of primitives in each symbol.
+The system uses a composite set of features include Rubine feature set \cite{gestureexample12}, Zernike moments \cite{HeloiseBeautification}, ink density \cite{GeometryAndDomain102} and some structural and spatial information like number of perpendiculars lines , number of parallel lines and types of primitives in each symbol.
 After computing the features the symbol is introduced to the classifier.
@@
 \begin{figure}[]\centering
-\fbox{  \parbox{4cm}{% 
+\fbox{  \parbox{3cm}{% 
 		\includegraphics[scale=0.4]{images/symbolSet.pdf}	}}
-	\caption{The Symbol Set}
+	\caption{The Symbol Set} Ths figure shows the some of the symbols used in the dataset. 
 	\label{fig:symbolSet}
@@
 	\centering
-		\includegraphics[scale=0.3]{images/test1.pdf}
-	\caption{The results of first experiment}
+		\includegraphics[scale=0.5]{images/test1.jpg}
+	\caption{Algorithm comparison} The results of comparing the algorithms. The graphs displays the recognition rate of symbols. 
 	\label{fig:test1}
@@
 %In the figure \ref{fig:test1}, the accuracy means that final hit rate of all symbols used in the test set.by the  four datasets . 
-We performed three experiments to test the system firstly we tested recognition accuracy of shapes in the data set with both algorithms.
+We performed two experiments to test the system firstly we tested recognition accuracy of shapes in the data set with both algorithms.
 We also implemented the segmentation algorithm described in \cite{earlyprocess} to use as reference to our swarm algorithms.
@@
 The swarm algorithms were tested with and without the ellipse detection module.
-The ellipse detection module appear to be superior to results with only the PSO algorithms.
-\\
+The ellipse detection module appear to be superior to results with only the PSO algorithms.\\
 
@@
 Figure \ref{fig:test2} also shows that algorithm 1 alone achieve better performance than algorithm 2 in the symbols that consist of lines only.
-The combining of both algorithm improves the recognition rate of other symbols.
+This is natural as the algorithm divides strokes into line segments only but algorithm 2 can divide  strokes into lines and curves based on the minimum error of the segment itself.
+Algorithms 3 can give good performance as long as the symbols consist of lines and curves, if the stroke consist of curve only  the algorithm may lead to wrong segmentation result.
+This is because the system divides the stroke first to line segments then tries to decide if each segment can be represent better as a curve unlike algorithm 2 where the curve segments are tested while choosing the best segmentation.
+Combining both algorithm 1 and algorithm 2 improved the recognition rate of all symbols as the system now have the advantages of both algorithms.
 \\
@@
 	\centering
-		\includegraphics[scale=0.5]{images/test2.pdf}
-	\caption{The results of second experiment}
+		\includegraphics[scale=0.8]{images/test2.jpg}
+	\caption{Symbols Comparison} The graph shows the recognition rate of each symbol using different algorithms. 
 	\label{fig:test2}
 \end{figure*}
-  The final experiment we performed was to test the efficiency svm classifier.
+We could not test the result of the segmentation algorithm directly due to the fact that the correct segmentation is highly ambiguous.
+It is only based on what the user intended to draw while sketching the symbol.
+Figure \ref{fig:SampleSeg} shows a sample of the output of the segmentation block.
+  %The final experiment we performed was to test the efficiency svm classifier.
 We implemented the linear discriminator classifier described in \cite{gestureexample12}(see figure \ref{fig:test3}).
@@
 %For this experiment we also shows the classification time for both classifiers
+
 \begin{figure}[]
 	\centering
-		\includegraphics[scale=0.4]{images/test1.pdf}
-	\caption{The results of third experiment}
-	\label{fig:test3}
+
+				\subfigure[User strokes] {\includegraphics[scale=0.7]{images/strokebefore.jpg}}
+		 	\subfigure[Segmented strokes] {\includegraphics[scale=0.7]{images/strokeafter.jpg}}
+		 		\subfigure[Multiple symbols] {\includegraphics[scale=0.7]{images/results3.jpg}}
+
+	\caption{Segmentation outputs} The figure shows sample of the output of the segmentation phase.
+Figure a) shows the stroke after the user draws it.
+The red points are points labeled as \textit{possible dominate points}.
+Figure b) shows the symbol after being segmented.
+Figure c) shows the segmentation of different and overlabed strokes.
+Notice the system mark the center of the ellipse that was identified.
+	\label{fig:SampleSeg}
 \end{figure}
+%\begin{figure}[]
+%	\centering
+%		\includegraphics[scale=0.4,clip=true, viewport=0 380 700 1280]{images/test1.pdf}
+%	\caption{The results of third experiment}
+%	\label{fig:test3}
+%\end{figure}
  %Details description of results got from comparing various algorithms  
@@
 This paper presented a new approach to sketch recognition using PSO.
-It was noted that the PSO in general improve the accuracy of the final symbol recognition in the system.
-The use of both speed and curvature data helps in improvement of the PSO algorithm over the original algorithms \cite{CruveDivisionSwarm,PolygonApproximationPSO}.
-The tradeoff between accuracy achieved and time complexity must be further investigated to achieve better results.
+The system used both speed and curvature data which helped in improving the PSO algorithms over the original algorithms \cite{CruveDivisionSwarm,PolygonApproximationPSO}.
+It was noted that the PSO in general generated an optimized stroke segmentation which improved the final recognition rate.
+The trade off between accuracy achieved and time complexity must be further investigated to achieve better results.
+The use of statistical moments and structural features improved the recognition rate.
+The system was tested on 13 different symbol and achieved satisfying results.
+The system dose not depend on low level recognizer but rather on a set of high level features which make the system easily expandable with aspect to symbols recognized.
+\\
 %As you can see form the experiments PSO have proved a spuriously than other system. 
@@
 The clustering must be performed without the user explicit involvement.
-Other area of modification can be the features extraction and classifier.
-Adding more spatial features will improve classifications.
+Other area of modification will the features extraction methods.
+Introducing more spatial and geometrical features will improve classifications.
 %Each new stroke is checked if it can be a symbol or is a part of already drawn un completed symbol.

### 2009-01-06 trunk/writing/outline.tex
before dr. samaia changes

@@
 These geometric primitives are grouped to formulate symbols which are further identified.
-The results shows that using PSO improves segmentation results which guide the symbol recognition phase.
+The results show that using PSO improves segmentation results which guide the symbol recognition phase.
 This paper uses SVM classifier which further improves the final recognition accuracy.
@@
 Design engineers need a powerful computer based system with symbol recognition to help them design, manipulate and store sketches more effectively than using only papers.
-Increasing the interaction between computers and users in sketch and CAD systems has been the reason for the emerging of few advanced sketch recognition systems.
+Increasing the interaction between computers and users in sketch and CAD systems has been the reason for the emerging of few advanced sketch recognition systems \cite{CAD}.
 Sketch recognition system converts the drawings and informal graphs user draws into objects and symbols that can be used in further interaction with the system.
@@
 This paper introduces a sketch recognition system.
-The presented system uses particle swarm algorithm(PSO) to optimially segment users strokes.
+The presented system uses' particle swarm algorithm (PSO) to optimally segment users strokes.
 Users can draw symbols using any number of strokes in any order they like.
-The system segments stroke user draws then cluster symbols together then passes them to the SVM symbol recognizier.
-The experiments in section \ref{sec:Experiments} shows that the proposed system improves the final symbol recognition accurracy.
+The system segments stroke user draws then cluster symbols together then passes them to the SVM symbol recognizer.
+The experiments in section \ref{sec:Experiments} shows that the proposed system improves the final symbol recognition accuracy.
 
@@
 Section \ref{PSO} explains the general particle swarm algorithm which is used in the presented system.
-The proposed system block diagram along with a detailed system discribtion is presented in Section \ref{Sysdisc}.
+The proposed system block diagram along with a detailed system description is presented in Section \ref{Sysdisc}.
 The experiments and results achieved are explained in Section \ref{sec:Experiments}.
@@
 Their system achieved good accuracy in simple shapes (square, ellipse,...) but did not perform well in more complex shapes\cite{meanshift10,Paleosketch08}.
-\citeauthor{Paleosketch08} \cite{Paleosketch08} introduced a set of low level recognizers that were reported to achieve 98\%  accuracy but their system simillar to all low level recognizers only identify a small set of simple shapes.
+\citeauthor{Paleosketch08} \cite{Paleosketch08} introduced a set of low level recognizers that were reported to achieve 98\%  accuracy but their system similar to all low level recognizers only identify a small set of simple shapes.
 
@@
 Geometrical and spatial descriptive language was used in \cite{SketchRead2007,Ladder30} where an architecture similar to compiler was used to recognize symbols.
+The drawback of this architecture is that the recognition system can only parse symbols that were hard coded using the descriptive language.
+New symbols are not recognized unless the description of the symbol is provided or generated using another system.
 %%SVM and HMM classification was implemented in\cite{SVM300,HMM53} to correctly classify symbols.
@@
 \citeauthor{CruveDivisionSwarm} uses digital curves scanned from paper as input to the system and did not take advantage of the curvature or local geometric properties of the digital curve.
-\citeauthor{PolygonApproximationPSO} \cite{PolygonApproximationPSO} used PSO to convert digital curves into polygons, our system adoptes \citeauthor{PolygonApproximationPSO} system but improves it by adding curvature and other local information while segmenting strokes to achieve better segmentations.
-The next section presents the general particle swarm algorithm.
-The proposed swarm segmentation algorithms are peresented in section \ref{subsubsec:Discreteparticleswarmalgorithm}
+\citeauthor{PolygonApproximationPSO} \cite{PolygonApproximationPSO} used PSO to convert digital curves into polygons, our system adopts \citeauthor{PolygonApproximationPSO} method but improves it by adding curvature and other local information while segmenting strokes to achieve better segmentations.
+The next section presents the general particle swarm algorithm which is used in our proposed segmentation algorithms.
+% are presented in section \ref{subsubsec:Discreteparticleswarmalgorithm}
 % Other techniques as  [ ] was used  in []. \\  
@@
 where $p_{ij}$ represent the $jth$ particle in the $ith$ agent and $v_{ij}$ is the velocity of the $jth$ particle in the $ith$ agent.
-Each iteration the global best $g_{best}$ particle and the agent local best $l_{best}$ particle are evaluated based on the maximum fitness functions of all particles.
-The solution is found after acheiving a specific number of iteration or after an error threshold is achieved.
+After each iteration the global best $g_{best}$ particle and the agent local best $l_{best}$ particle are evaluated based on the maximum fitness functions of all particles.
+The solution is found after achieving a specific number of iteration or after an error threshold is achieved.
 The following equation \ref{eq:descrite}  
@@
 
-\section{System Description}
+\section{System Details}
 \label{Sysdisc}
 %The stroke is the path of points from the instant the pen is down till it is up.
-Figure. \ref{fig:Blockdiagram} shows that the main system blocks are 1) preprocessing, 2) segmentation, 3) clustering, 4) feature extraction, and finally 5) classification.
+Figure \ref{fig:Blockdiagram} shows that the main system blocks are 1) preprocessing, 2) segmentation, 3) clustering, 4) feature extraction, and finally 5) classification.
 
  %Due to sloppiness of users and hardware glitches the system preprocess the data before further processing. 
-The preprocessing is responsible of capturing the input data and removing the noise from it.
+The preprocessing phase is responsible of capturing the input data and removing the noise from it.
 The system then proceeds to estimate a set of possible dominant points to help in the segmentation process.
@@
 In the first step, the ellipse fitting process tries to fit the stroke into an ellipse.
-If the system fails it passes the stroke to the second step; curve segmentation which consist of two PSO segmentation algorithms.
-The two algorithms will generate two segmentations, the system will choose the segmentation that has the minimum segmentation error.%  generate the segmentations,  the minimum error will be the chosen segmentation.
+If the system fails, it passes the stroke to the second step; curve segmentation which consists of two PSO segmentation algorithms.
+The two algorithms generate two segmentations.
+The system then chooses the segmentation that has the minimum segmentation error.%generate the segmentations,  the minimum error will be the chosen segmentation.
 % is the ellipse fitting
@@
 \\% This part is repeated for each stroke.
- The clustering algorithms starts to group segments together after the segmentation step.
-The system let the user draws the symbol by using any number of strokes, a set of unrecognized segments is passed to the clustering algorithm to generate a symbol and compute a feature vector for it.
-The system uses a hybrid set of statistical, geometric and spatial features.
-The final step is a SVM classifier that will use the features computed to classify the segments into one of the previously trained classes.
-%  Or to determine the symbol of the given segments from the set of preciously trained symbols.
+ The clustering algorithm starts to group segments together after the segmentation step.
+The system allows the user draws the symbol by using any number of strokes; a set of unrecognized segments is passed to the clustering algorithm to generate a symbol and compute a feature vector for it.
+A set of hybrid features of statistical, geometric and spatial features is used to generate the symbol feature vector.
+The final classification step, SVM classifier uses the feature vector computed to classify the segments into one of the previously trained classes.
+The following sections provides a detail description of each step.%  Or to determine the symbol of the given segments from the set of preciously trained symbols.
  %with a symbol from the training set.% The system compute composite set of features some are statistical other are spatial features based on the type primitives.
@@
 	\centering
-		\includegraphics[scale=0.85]{images/block2.pdf}%Blockdiagram.pdf}
+		\includegraphics[scale=0.75]{images/block2.pdf}%Blockdiagram.pdf}
 	\caption{The system Block Diagram.} The figure shows the main system blocks. Preprocessing, segmentation, clustering, features extraction and classification.
@@
 %Some data is extracted from the points in the stroke.
- It is noted that as the user draws a shape the pen will slow down near corners and picks up speed when drawing a straight lines.
+ It is noted that as the user draws a shape, the pen slows down near corners and picks up speed when drawing a straight lines.
 Therefore, the speed and time difference data were widely used in sketch understanding systems \cite{earlyprocess}.
 The direction and curvature data are used to determine the points with high angular changes along the path of points \cite{meanshift10}.
-
+ %speed is calculated as 
  In the presented system, we computed time difference, direction, speed and curvature of each point along the stroke.
-The following equation $v=\Delta t/\Delta s$ is used for speed calculation where $t$ is the time difference between two points and $s$ is the length between them.
-The direction is calculated as the angle between two vectors and curvature is the change in direction with respect to length i.e.
+The speed is calculated as  $v=\Delta t/\Delta s$ where $t$ is the time difference between two points and $s$ is the length between them.
+The direction is calculated as the angle between two vectors and curvature is considered as the change in direction with respect to length i.e.
 $c= \Delta d/\Delta s$.
@@
  %$\Delta d/  \Delta S $ where d is the difference direction of point and s is difference in distance between points. 
- All these calculations are performed in real time while the user draws the strokes. The complexity of computation is $O(n)$ where $n$ is number of points. \\
- Figure \ref{fig:speed2} shows the computed data for the stroke drawn in fig.
-\ref{fig:orignalStroke}.
-If you look at the data curves you will see that dominant points are characterized by lower speed values and higher curvature and direction values.
+ It should be noted that, all these calculations are performed in real time while the user draws the strokes. The complexity of computation is $O(n)$ where $n$ is number of points. \\
+ Figure \ref{fig:speed2} shows the computed data for the stroke drawn in fig. \ref{fig:orignalStroke}. %If you look at the data curves you will see by looking at the data curves s
+ It is noted that the dominant points are characterized by lower speed values and higher curvature and direction values (fig.
+\ref{fig:speed2}).
 \\% the lower speed points correspond to vertex and dominant point's location.
@@
 Hence, the mean
-of the curve is calculated and used as threshold.
-This threshold is used to separate the curve into regions; each region is defined as the part between two intersection points of the threshold line and the curve.
-Regions higher than the threshold value are processed to find the maximum of each region.
-The points that correspond to those maximum values are labeled as \textit{possible dominant points}.
+of the curve is calculated and used as threshold \textit{th}.
+The threshold \textit{th} is used to separate the curve into regions  $Region_i$; each region $Region_i$ is defined as the part between two intersection points of the threshold line and the curve.
+Regions higher than the threshold \textit{th} are processed to find the maximum point $Max(Region_i)$ of each region $Region_i$.
+The stroke points $p_i(x,y)$ that correspond to those maximum values are labeled as \textit{possible dominant points} $P_{pd}$.
 \\
@@
 	\centering
-		\includegraphics[scale=0.3]{images/orignalStroke.pdf}
+		\includegraphics[scale=0.5]{images/orignalStroke.pdf}
 	\caption{Stroke Sample} An example of the stroke drawn in the system.
@@
 \end{figure}
-The system repeat this process to curvature, time difference and speed curves.
-All the points labeled as possible dominant points are saved into a single array.
-They are used as input to guide the next stage of segmentation.
+The system repeats this process to curvature, time difference and speed curves.
+All the points labeled as possible dominant points $P_{pd}$ are saved into a single array.
+They are then used as input to guide the next segmentation phase.
 \begin{figure*}[]
@@
 After computing the primary data the system tries to segment the stroke into a set of primitives.
-The segmentation algorithm first tries to detect if the stroke is an ellipse.
+The segmentation algorithm first attempts to detect if the stroke is an ellipse.
 If the stroke proved to be an ellipse then the segmentation process ends and the system proceeds to the next step.
-Otherwise, the stroke is passed into two particle swarm algorithms that will divide the stroke to either lines or lines and curves.
-The algorithms takes the stroke points along with the possible dominant points computed then produce a set of dominant points which are connected with either lines or curves (see fig.
+Otherwise, the stroke is passed into two particle swarm algorithms that divides the stroke to either lines or lines and curves.
+The algorithms take the stroke points along with the possible dominant points $P_{pd}$ computed then produce a set of dominant points which are connected with either lines or curves (see fig.
 \ref{fig:Blockdiagram}).
-The next section will describe the ellipse detection algorithm and both the two particle swarm algorithms used to divide the stroke.\\
+The next section describes the ellipse detection algorithm and the two particle swarm algorithms used to divide the stroke.\\
 %Paragraphs describe the segmentation algorithm.  
@@
 %Describe the ellipse detection algorithms 
-The process starts with computing the center of the stroke bounding box.
+The process starts by computing the center of the stroke bounding box.
 The bounding box center point is used as the first estimation of the ellipse center.
 The axes of the ellipse are estimated as $width/2$ and $height/2$ of the stroke bounding box.
-The least square fitting algorithm is used to minimize the fitting error of the ellipse equation see equation \ref{eq:circleFit}.
+The least square fitting algorithm is used to minimize the fitting error of the ellipse equation eq.
+\ref{eq:circleFit}.
 \begin{equation}
@@
 \label{eq:circleFit}
-\end{equation} where N is number of points in the stroke, $a,b$ are the length of ellipse axes, $x_0$ \& $y_0$ are the coordinates of the center point, $x_i$ \& $y_i$ are the coordinates of point $i$ in the stroke.
-After only few loops the error and a confidence values  are computed to check if the stroke can be labeled as ellipse or not.\\ %is computed to check if stroke is an ellipse.
+\end{equation} where $N$ is the number of points in the stroke, $a,b$ are the length of ellipse axes, $x_0$ , $y_0$ are the coordinates of the center point and $x_i$ \& $y_i$ are the coordinates of point $i$ in the stroke.
+After few loops, the error and a confidence values  are computed to check if the stroke can be labeled as ellipse or not.\\ %is computed to check if stroke is an ellipse.
 \\
 %We found that if the confidante is above threshold then the probability of ellipse is highest otherwise the stroke is passed to the next section to get the divisions of stroke and test its error.  
-\subsubsection{Discrete particle swarm algorithm}
+\subsubsection{Discrete particle swarm algorithm (\textit{DPSO})}
 \label{subsubsec:Discreteparticleswarmalgorithm}
-There are two PSO algorithm described in this section.
-The system generates segmentation from both algorithms and then chooses the segmentation with the minimum error value.
-The problem definition is the same in both algorithms but they differ in the method they compute fitness and error functions.
-%formation is nearly the same in both.
+Two DPSO algorithms are used to generate segmentations for each stroke the user draws.
+The segmentation with the minimum error value is chosen as the stroke segmentation.
+The problem definition is the same in both algorithms but they differ in the method they use to compute fitness and error functions.
 \subsubsection{Problem definition}
-The input stroke with N points can be represented by set $S = \left\{ {x_1 ,x_2  \ldots x_N } \right\}$ where $x_i$ is the location of the point $i$ .
-The swarm algorithms consist of M agents which are represented by the set
+The input stroke with $N$ points can be represented by set $S = \left\{ {x_1 ,x_2  \ldots x_N } \right\}$ where $x_i$ is the location of the point $i$.
+The swarm algorithms consist of $M$ agents which are represented by the set
 $A  = \left\{ {P_i \left| {i = 1,2 \cdots M} \right.} \right\}$ where $P_i$ is a single solution particle from the solution space.
-Each particle decodes the problem with binary array with the same length N as the input stroke.
+Each particle decodes the problem with binary array with the same length $N$ as the input stroke.
 % The particles of the swarm represent a single solution of the solution space.
@@
 We define the arc %An array with the same length as the number of points of the strokes.
-So the system will represent each particle $P_i$ by $P_i = \left\{ {p_{ij} \left| {j = 1,2 \cdots N} \right.} \right\}$ where $p_{ij}$ will have only two values 0 or 1 where $p_{ij}=1$ will mean that point $j$ is a dominant point.
+Therefore, the system represents each particle $P_i$ by $P_i = \left\{ {p_{ij} \left| {j = 1,2 \cdots N} \right.} \right\}$ where $p_{ij}$ has only two values 0 or 1.
+When $p_{ij}=1$ means that point $j$ is a dominant point.
 % a!402B!
@@
 \subsubsection{The algorithm}
-The fitness function and error calculation are different in each algorithm. For the first algorithm, 
-let's define the arc $\widehat{x_ix_j}$ as the consecutive set of points from point$x_i$ to point $x_{j}$ as in $x_i,x_{i+1} \cdots,x_j$.
+The fitness function and error calculation are different in each of the two \textit{DPSO} algorithms. For the first algorithm \textsl{AlgS1}, 
+ the arc $\widehat{x_ix_j}$ is defined as the consecutive set of points from point$x_i$ to point $x_{j}$ as in $x_i,x_{i+1} \cdots,x_j$.
 The line $\overline{x_i x_j} $ as the straight line connecting point $x_i$ to point $x_j$.
@@
 \label{eq:ErrorSwarm1}
-\end{equation} where M is the number of dominant points in this solution.
+\end{equation} where $M$ is the number of dominant points in this solution.
 The error $e ( \widehat{x_ix_j},\overline{x_i x_j})$ is computed as the sum of squared perpendicular distance from every point along the arc $\widehat{x_ix_j}$ to the line $\overline{x_i x_j}$.
@@
 \label{eq:fitnessSwarm1}
-\end{equation} where $N$ is the number of points in the stroke, $D$ is the number of point in the solution that was previously labeled as a possible dominant point and $E$ is the computed error and $\varepsilon$ is the error threshold.
-Notice that when the error is larger than the threshold the fitness is given a -ve value to lower the value of solution otherwise the system will favor the lower number of vertices.\\ %we will want to lower the number of vertices.
+\end{equation} where $N$ is the number of points in the stroke, $D$ is the number of points in the solution that was previously labeled as a possible dominant point ($P_{pd}$), $E$ is the computed error and $\varepsilon$ is the error threshold.
+It should be noticed that when the error is larger than the threshold $E$ the fitness is given a -ve value to lower the value of solution.
+Otherwise the system favors the lower number of vertices.\\ %we will want to lower the number of vertices.
 \\
@@
 %Alg2: o
-The second algorithm has the same problem formulation but different fitness and error functions.
+The second algorithm \textsl{AlgS2} has the same problem formulation but different fitness and error functions.
 It was previously introduced in \cite{CruveDivisionSwarm} but a genetic programming was used as the optimizing algorithm.
-The error of both circle and line estimation are computed for each segment, the approximation with the lower error value will be the chosen approximation of this segment.
+The error of both circle and line estimation are computed for each segment $S_i=\widehat{x_ix_j}$, the approximation with the lower error value is the chosen approximation of this segment $S_i$.
 The sum of the approximation error of all segments is the total error of the particle.
@@
 \label{eq:fitnessSwarm2}
-\end{equation} where $E$ is the error and $M$ is number of segments and $k$ is a parameter tweaked to get minimum number of segments. \\
+\end{equation} where $E$ is the error and $M$ is the number of segments and $k$ is a parameter tweaked to get minimum number of segments. \\
 %Describe how we used particle swarm in curve polygon approximation using first algorithm.  
@@
 %[These two algorithm was modified from the algorithm in paper \cite{PolygonApproximationPSO,CruveDivisionSwarm}].
-\citeauthor{PolygonApproximationPSO} \cite{PolygonApproximationPSO} used a merge and divide algorithm after each loop of the swarm system to refine the solution but this system is introducing another enhancement method.
-After each loop of the swarm algorithm, each particle is refined using the following procedure.
-For each particle %loops on the set of selected dominant points to enhance the solution.
-each dominant point is checked to find if it was labeled before as a \textit{possible dominant point} (computed as in section \ref{Prepross}), if it is not labeled the point is moved to the nearest labeled point.
+\citeauthor{PolygonApproximationPSO} \cite{PolygonApproximationPSO} used a merge and divide algorithm after each loop of the swarm system to refine the solution but this paper is introducing another enhancement method.
+After each loop of the swarm algorithm (\textsl{AlgS1} and \textsl{AlgS2}), each particle is refined using the following procedure.
+For each particle $P_{ij}$ %loops on the set of selected dominant points to enhance the solution.
+each dominant point is checked to find if it was labeled before as a \textit{possible dominant point} $P_{pd}$ (computed as in section \ref{Prepross}), if it is not labeled the point is moved to the nearest labeled point.
 After that the particles are tested to make sure that the distance between every two successive dominate point is larger than the constant $min_D$.
-If two points are found that they are nearer than $min_D$ then based on the decrease in the error of the solution one of the points will either be removed or moved to the nearest possible dominate point.
+If two points are found that they are nearer than $min_D$ then based on the decrease in the error $E$ in eq.
+\ref{eq:ErrorSwarm1} \& \ref{eq:errorSwarm2} one of the points are either removed or moved to the nearest possible dominate point.
 %check the new solution and check the new vertices if a vertex points is a possible dominant point it is left as a vertex.
@@
 \subsection{Recognition}
-After the user draws all strokes of the symbol he has to wait 10 seconds or press finish button beside the drawing area.
+After the user draws all strokes of the symbol, he/she has to wait 10 seconds or press finish button beside the drawing area.
 The set of unrecognized strokes is grouped together along with their segmentation as input to the feature extraction process.
-The system uses a composite set of features include Rubine feature set \cite{gestureexample12}, Zernike moments \cite{HeloiseBeautification}, ink density \cite{GeometryAndDomain102} and some structural and spatial information like number of perpendiculars lines , number of parallel lines and types of primitives in each symbol.
+A composite set of features is used to generate a single feature vector.
+The features used consist of Rubine feature set \cite{gestureexample12},  Zernike moments of order 21 \cite{HeloiseBeautification}, ink density \cite{GeometryAndDomain102} as well as some structural and spatial information like number of perpendiculars lines ,number of parallel lines and types of primitives in each symbol.
 After computing the features the symbol is introduced to the classifier.
-The system used a support vector machine (SVM) classifier with Gaussian kernel (RBF kernel) \cite{libsvm}.\\%Some features like number of parallel, perpendicular and intersecting lines are computed from the segmentation other like density, centroid and Zernike moments are computed from the raw stroke points.
+The system uses a support vector machine (SVM) classifier with Gaussian kernel (RBF kernel)\cite{libsvm}.\\%Some features like number of parallel, perpendicular and intersecting lines are computed from the segmentation other like density, centroid and Zernike moments are computed from the raw stroke points.
  %trained with the 13 categories in the system.
@@
 \section{Experiments}\label{sec:Experiments}
-For testing the system used a data set collected by \citeauthor{HeloiseBeautification}.
-The data are drawn by 16 users each of them have drawn from 30 to 50 samples for each shape.
-Figure \ref{fig:symbolSet} shows the shapes used in the data set.
-We divided the dataset equally into training set and test set.
+A dataset collected by \citeauthor{HeloiseBeautification} is used to test the presented system\cite{HeloiseBeautification}.
+The data are drawn by 16 users each of them was asked to draw the 13 shapes from 30 to 50 times.
+Figure \ref{fig:symbolSet} shows a set of the shapes used in the data set.
+We divide the dataset into four sets each set contains four different users.
+We divide the dataset equally into training set and test set.
+To generate trusted results, four different splits for the training and test data are generated from the dataset.
+The results displayed are the recognition accuracy resulted from the classifier after recognizing the symbols.
+In the following section, the recognition accuracy means the average recognition accuracy in the four data splits.
+%The result presented in this section are the average result of the four splits.
+%then result are stored The results displayed are the recognition accuracy resulted from the classifier after recognizing the symbols.
+%Furthermore, each set is divided into train and test set where the trained set where used to train the classifiers and the test set where used for testing.
+The algorithms are then tested using the four sets and the results are averaged.
+%We divide the dataset equally into training set and test set.
 The results displayed are the recognition accuracy resulted from the classifier after recognizing the symbols.
@@
 		\includegraphics[scale=0.4]{images/symbolSet.pdf}	}}
-	\caption{The Symbol Set} Ths figure shows the some of the symbols used in the dataset. 
+	\caption{The Symbol Set} Ths figure shows a set of the symbols used in the dataset. 
 	\label{fig:symbolSet}
@@
 %In the figure \ref{fig:test1}, the accuracy means that final hit rate of all symbols used in the test set.by the  four datasets . 
-We performed two experiments to test the system firstly we tested recognition accuracy of shapes in the data set with both algorithms.
-We also implemented the segmentation algorithm described in \cite{earlyprocess} to use as reference to our swarm algorithms.
-As you see in the fig.
-\ref{fig:test1} shows the accuracy achieved by each algorithm.
-The results shows that both PSO algorithm achieve better result than other algorithms.
+We performed two experiments to test the system firstly we tested recognition accuracy of shapes in the data set with both algorithms (\textsl{AlgS1} and \textsl{AlgS2}).
+We also implemented the segmentation algorithm described in \cite{earlyprocess} (\textsl{Alg3}) to use as reference to our swarm algorithms.
+Figure \ref{fig:test1} shows the accuracy achieved by each algorithm.
+The results shows that both \textit{DPSO} algorithms achieve better result than other algorithms.
 The swarm algorithms were tested with and without the ellipse detection module.
-The ellipse detection module appear to be superior to results with only the PSO algorithms.\\
+The ellipse detection module appears to be superior to results with only the \textit{DPSO} algorithms.\\
 
-The second experiment we implemented was to test the effect of symbol complexity and type on the recognition rate.
-Figure \ref{fig:test2} shows what the accuracy of each symbol, it clearly noted that symbols that have only line segments achieve higher accuracy rate than other symbols.
-Figure \ref{fig:test2} also shows that algorithm 1 alone achieve better performance than algorithm 2 in the symbols that consist of lines only.
-This is natural as the algorithm divides strokes into line segments only but algorithm 2 can divide  strokes into lines and curves based on the minimum error of the segment itself.
-Algorithms 3 can give good performance as long as the symbols consist of lines and curves, if the stroke consist of curve only  the algorithm may lead to wrong segmentation result.
-This is because the system divides the stroke first to line segments then tries to decide if each segment can be represent better as a curve unlike algorithm 2 where the curve segments are tested while choosing the best segmentation.
-Combining both algorithm 1 and algorithm 2 improved the recognition rate of all symbols as the system now have the advantages of both algorithms.
+The second experiment we implemented is testing the effect of symbol complexity and type on the recognition rate.
+Figure \ref{fig:test2} shows the accuracy of each symbol, it is clearly noted that symbols that have only line segments achieve higher accuracy rate than other symbols.
+Figure \ref{fig:test2} also indicates that algorithm \textsl{AlgS1} alone achieve better performance than algorithm   \textsl{AlgS2} in the symbols that consist only of lines.
+This is natural as algorithm  \textsl{AlgS1} divides strokes into line segments only but \textsl{AlgS2} is able to divide strokes into lines and curves based on the minimum error of the segment itself.
+Algorithm \textsl{Alg3} gives good performance as long as the symbols consist of lines and curves, if the stroke consists of curve only  the algorithm may lead to wrong segmentation result.
+This is because the system divides the stroke first to line segments then tries to decide if each segment can be represent better as a curve unlike algorithm \textsl{AlgS2} where the curve segments are tested while choosing the best segmentation.
+Combining both algorithm \textsl{AlgS1} and \textsl{AlgS2} improved the recognition rate of all symbols because of combining the advantages of both algorithms.
+The penalty for this increased performance is in the computational time needed to run both the swarm algorithms.
 \\
@@
 \end{figure*}
-We could not test the result of the segmentation algorithm directly due to the fact that the correct segmentation is highly ambiguous.
-It is only based on what the user intended to draw while sketching the symbol.
+%We could not test the result of the segmentation algorithm directly due to the fact that the correct segmentation is highly ambiguous.
+It is only based on what the user intends to draw while sketching the symbol.
 Figure \ref{fig:SampleSeg} shows a sample of the output of the segmentation block.
@@
 %For this experiment we also shows the classification time for both classifiers
-
-\begin{figure}[]
-	\centering
-
-				\subfigure[User strokes] {\includegraphics[scale=0.7]{images/strokebefore.jpg}}
-		 	\subfigure[Segmented strokes] {\includegraphics[scale=0.7]{images/strokeafter.jpg}}
-		 		\subfigure[Multiple symbols] {\includegraphics[scale=0.7]{images/results3.jpg}}
-
-	\caption{Segmentation outputs} The figure shows sample of the output of the segmentation phase.
+%\begin{figure}[]
+%	\centering
+% 
+%				\subfigure[User strokes] {\includegraphics[scale=0.7]{images/strokebefore.jpg}}
+%		 	\subfigure[Segmented strokes] {\includegraphics[scale=0.7]{images/strokeafter.jpg}}
+%		 		\subfigure[Multiple symbols] {\includegraphics[scale=0.7]{images/results3.jpg}}
+%		
+%	\caption{Segmentation outputs} The figure shows sample of the output of the segmentation phase.
 Figure a) shows the stroke after the user draws it.
@@
 Figure b) shows the symbol after being segmented.
-Figure c) shows the segmentation of different and overlabed strokes.
+Figure c) shows the segmentation of different and overlapped strokes.
 Notice the system mark the center of the ellipse that was identified.
-	\label{fig:SampleSeg}
-\end{figure}
+%	\label{fig:SampleSeg}
+%\end{figure}
 %\begin{figure}[]
@@
 This paper presented a new approach to sketch recognition using PSO.
-The system used both speed and curvature data which helped in improving the PSO algorithms over the original algorithms \cite{CruveDivisionSwarm,PolygonApproximationPSO}.
-It was noted that the PSO in general generated an optimized stroke segmentation which improved the final recognition rate.
-The trade off between accuracy achieved and time complexity must be further investigated to achieve better results.
-The use of statistical moments and structural features improved the recognition rate.
-The system was tested on 13 different symbol and achieved satisfying results.
-The system dose not depend on low level recognizer but rather on a set of high level features which make the system easily expandable with aspect to symbols recognized.
+The system uses both speed and curvature data which helps in improving the \textit{DPSO} algorithms over the original algorithms \cite{CruveDivisionSwarm,PolygonApproximationPSO}.
+It is noted that the \textit{DPSO}in general generates an optimized stroke segmentation which improves the final recognition rate.
+The tradeoff between accuracy achieved and time complexity must be further investigated to achieve better results.
+The use of statistical moments and structural features improves the recognition rate.
+The system was tested on 13 different symbols and achieved satisfying results.
+The system dose not depends on low level recognizer but rather on a set of high level features.
+This makes the system easily expandable with aspect to symbols recognized.
 \\
 %As you can see form the experiments PSO have proved a spuriously than other system. 
-%\section{Future Work}
-The next step in this research is to complete the clustering algorithm for fully automated sketch recognition.
+%\section{Future Work}The next step in 
+A possible extension of this research is to complete the clustering algorithm for fully automated sketch recognition.
 The clustering must be performed without the user explicit involvement.
-Other area of modification will the features extraction methods.
-Introducing more spatial and geometrical features will improve classifications.
+Other area of enhancements is the features extraction methods.
+Introducing more spatial and geometrical features is believed to improve classifications.
 %Each new stroke is checked if it can be a symbol or is a part of already drawn un completed symbol.

### 2009-01-25 smallpaper/SmallPaper.tex
adding small version before review

--- 2009-01-06 trunk/writing/outline.tex
@@
 %opening
-\title{Symbol Recognition using Particle Swarm Algorithms}
+\title{Sketch Recognition using Particle Swarm Algorithms}
 %\title{A Study on Using Particle Swarm Algorithms in Symbol Recognition }
-\author{Maha El Meseery,  Mahmoud Fakhr El Din,  Samyah Mashaly,
-\\  Magdah Fayek,  Nevin Darwish}
+\author{Maha El Meseery,  Mahmoud Fakhr El Din,  Samia Mashali,
+\\  Magda Fayek and Nevin Darwish}
 \begin{document}
@@
 The results show that using PSO improves segmentation results which guide the symbol recognition phase.
-This paper uses SVM classifier which further improves the final recognition accuracy.
+This paper uses Support Vector Machines(SVM) classifier which further improves the final recognition accuracy.
 %u Final recognition has improved accuracy after the PSO algorithm using a SVM classifier.
@@
 Engineers use sketches to exchange designs as a natural method of communication rather than writing or speaking.
-Engineers use paper and pencil design in early stages of design.
+%Engineers use paper and pencil design in early stages of design.
 Design engineers need a powerful computer based system with symbol recognition to help them design, manipulate and store sketches more effectively than using only papers.
 Increasing the interaction between computers and users in sketch and CAD systems has been the reason for the emerging of few advanced sketch recognition systems \cite{CAD}.
-Sketch recognition system converts the drawings and informal graphs user draws into objects and symbols that can be used in further interaction with the system.
-
+%Sketch recognition system converts the drawings and informal graphs user draws into objects and symbols that can be used in further interaction with the system.
 Sketch recognition is divided into four main steps preprocessing, segmentation, symbol recognition and finally sketch understanding.
@@
 Finally, the sketch understanding phase uses priori information about the sketch to support identification of symbols by excluding any possible invalid symbols.
+
 %the basic symbols are identified in consistent with the whole sketch. \\
@@
 This paper introduces a sketch recognition system.
-The presented system uses' particle swarm algorithm (PSO) to optimally segment users strokes.
+The presented system uses particle swarm algorithm (PSO) to optimally segment users strokes.
 Users can draw symbols using any number of strokes in any order they like.
@@
 The experiments in section \ref{sec:Experiments} shows that the proposed system improves the final symbol recognition accuracy.
-
-The remaining of this papers is organized as following; section \ref{Sec:RelatedWork} provides a brief literature review.
+%The remaining of this papers is organized as following; section \ref{Sec:RelatedWork} provides a brief literature review.
 Section \ref{PSO} explains the general particle swarm algorithm which is used in the presented system.
@@
 Their system has a draw back as it only recognizes simple specific shapes with a set of low level recognizers.
-\cite{meanshift10} introduced a feature area for each primitive and then computed the segmentation error for different types of primitives based on the computed feature area.
-Their system achieved good accuracy in simple shapes (square, ellipse,...) but did not perform well in more complex shapes\cite{meanshift10,Paleosketch08}.
+\citeauthor{meanshift10} introduced a feature area for each primitive and then computed the segmentation error for different types of primitives based on the computed feature area.
+Their system achieved good accuracy in simple shapes (square, ellipse,...) but did not perform well in more complex shapes\cite{meanshift10}.
 \citeauthor{Paleosketch08} \cite{Paleosketch08} introduced a set of low level recognizers that were reported to achieve 98\%  accuracy but their system similar to all low level recognizers only identify a small set of simple shapes.
 
-Many different approaches have been investigated to achieve final recognition of symbols.
+%Many different approaches have been investigated to achieve final recognition of symbols.
 Graph searching and template matching where used in \cite{GraphBased07} where a graph of the drawn symbol is generated and matched with a stored set of graphs templates to reach a classification decision.
@@
 \citeauthor{CruveDivisionSwarm} uses digital curves scanned from paper as input to the system and did not take advantage of the curvature or local geometric properties of the digital curve.
-\citeauthor{PolygonApproximationPSO} \cite{PolygonApproximationPSO} used PSO to convert digital curves into polygons, our system adopts \citeauthor{PolygonApproximationPSO} method but improves it by adding curvature and other local information while segmenting strokes to achieve better segmentations.
-The next section presents the general particle swarm algorithm which is used in our proposed segmentation algorithms.
+\citeauthor{PolygonApproximationPSO} \cite{PolygonApproximationPSO} used PSO to convert digital curves into polygons, our system adopts \citeauthor{PolygonApproximationPSO} method but improves it by adding curvature and other local information while segmenting strokes to achieve better segmentations.% The next section presents the general particle swarm algorithm which is used in our proposed segmentation algorithms.
 % are presented in section \ref{subsubsec:Discreteparticleswarmalgorithm}
@@
 %After segmentation researchers varies on the methods in which to continue.
-\section{Particle Swarm Algorithm}
-\label{PSO}
-%What is particle swarm algorithm and how it was used in related researches. 
-The main idea of Particle Swarm Algorithm (PSO) is to represent each agent with a particle from the solution space.
+%\section{Particle Swarm Algorithm}
+%\label{PSO}
+%%What is particle swarm algorithm and how it was used in related researches. 
+%The main idea of Particle Swarm Algorithm (PSO) is to represent each agent with a particle from the solution space.
 Each agent moves the particle with a direction and velocity based on equation \ref{eq:Swarm}.
- %Equation [\ref{eq:Swarm}] shows how velocity and direction of each particle are computed
- \begin{equation}
-v_{ij}  = v_{ij}  + c_1 r_1 (lbest_{ij}  - p_{ij} ) + c_2 r_2 (gbest_{ij}  - p_{ij} )
-\label{eq:Swarm}
-\end{equation}
- where $r_1$ \& $r_2$ are random variables and $c_1$ \& $c_2$ are the swarm system variables.
-\begin{equation}
-%\[
-p_{ij}=p_{ij}+v_{ij},
-%\
-\end{equation}
-where $p_{ij}$ represent the $jth$ particle in the $ith$ agent and $v_{ij}$ is the velocity of the $jth$ particle in the $ith$ agent.
+% %Equation [\ref{eq:Swarm}] shows how velocity and direction of each particle are computed
+% \begin{equation}
+%v_{ij}  = v_{ij}  + c_1 r_1 (lbest_{ij}  - p_{ij} ) + c_2 r_2 (gbest_{ij}  - p_{ij} )
+%\label{eq:Swarm}
+%\end{equation}
+% where $r_1$ \& $r_2$ are random variables and $c_1$ \& $c_2$ are the swarm system variables.
+%\begin{equation}
+%%\[
+%p_{ij}=p_{ij}+v_{ij},
+%%\
+%\end{equation}
+%where $p_{ij}$ represent the $jth$ particle in the $ith$ agent and $v_{ij}$ is the velocity of the $jth$ particle in the $ith$ agent.
 After each iteration the global best $g_{best}$ particle and the agent local best $l_{best}$ particle are evaluated based on the maximum fitness functions of all particles.
 The solution is found after achieving a specific number of iteration or after an error threshold is achieved.
-The following equation \ref{eq:descrite}  
-\begin{equation}
-   P(i)\Leftarrow 
-\{
-\begin{array}{c} 
-1 \quad \quad if\quad r_{3}>t  \\
-
-0 \quad \quad if\quad r_{3}<t 
-\label{eq:descrite}
-\end{array}\}
-\end{equation}
- (where $t$ is a threshold and $r_{3}$ is a random variable), is used in order to change the general swarm algorithm into binary particle which handles particle values of either $0$ or $1$.
-
+%The following equation \ref{eq:descrite}  
+%\begin{equation}
+%   P(i)\Leftarrow 
+%\{
+%\begin{array}{c} 
+%1 \quad \quad if\quad r_{3}>t  \\
+%
+%0 \quad \quad if\quad r_{3}<t 
+%\label{eq:descrite}
+%\end{array}\}
+%\end{equation}
+% (where $t$ is a threshold and $r_{3}$ is a random variable), is used in order to change the general swarm algorithm into binary particle which handles particle values of either $0$ or $1$. 
+%    
 \section{System Details}
@@
 %The stroke is the path of points from the instant the pen is down till it is up.
-Figure \ref{fig:Blockdiagram} shows that the main system blocks are 1) preprocessing, 2) segmentation, 3) clustering, 4) feature extraction, and finally 5) classification.
+%Figure \ref{fig:Blockdiagram} 
+The following sketch recognitions systems is divided into three main system blocks are 1) preprocessing, 2) segmentation and 3) Recognition.
+The following sections provides a detail description of each step.
 
  %Due to sloppiness of users and hardware glitches the system preprocess the data before further processing. 
-The preprocessing phase is responsible of capturing the input data and removing the noise from it.
+%The preprocessing phase is responsible of capturing the input data and removing the noise from it.
 The system then proceeds to estimate a set of possible dominant points to help in the segmentation process.
@@
  %The goal of the segmentation stage is to divide stroke into segments of either curves or lines.
-The segmentation is divided into two steps ellipse fitting and curve segmentation.
-In the first step, the ellipse fitting process tries to fit the stroke into an ellipse.
+%The segmentation is divided into two steps ellipse fitting and curve segmentation.
+%In the first step, the ellipse fitting process tries to fit the stroke into an ellipse.
 If the system fails, it passes the stroke to the second step; curve segmentation which consists of two PSO segmentation algorithms.
@@
 \\% This part is repeated for each stroke.
- The clustering algorithm starts to group segments together after the segmentation step.
+ %The clustering algorithm starts to group segments together after the segmentation step.
 The system allows the user draws the symbol by using any number of strokes; a set of unrecognized segments is passed to the clustering algorithm to generate a symbol and compute a feature vector for it.
@@
 The final classification step, SVM classifier uses the feature vector computed to classify the segments into one of the previously trained classes.
-The following sections provides a detail description of each step.%  Or to determine the symbol of the given segments from the set of preciously trained symbols.
+  %  Or to determine the symbol of the given segments from the set of preciously trained symbols. 
  %with a symbol from the training set.% The system compute composite set of features some are statistical other are spatial features based on the type primitives.
 Finally, the strokes are classified into the corresponding classes.% the classifier identifies the strokes into a symbol from the set of known symbols.\\
-
-\begin{figure}[]
-%\begin{minipage}[b]{0.5\linewidth}
-	\centering
-		\includegraphics[scale=0.75]{images/block2.pdf}%Blockdiagram.pdf}
-	\caption{The system Block Diagram.} The figure shows the main system blocks. Preprocessing, segmentation, clustering, features extraction and classification.
-
-	\label{fig:Blockdiagram}
-%	\end{minipage}
-\end{figure}
-
+%\begin{figure}[]
+%%\begin{minipage}[b]{0.5\linewidth}
+%	\centering
+%		\includegraphics[scale=0.75]{images/block2.pdf}%Blockdiagram.pdf}
+%	\caption{The system Block Diagram.} The figure shows the main system blocks. Preprocessing, segmentation, clustering, features extraction and classification. 
+%	
+%	\label{fig:Blockdiagram}
+%%	\end{minipage}
+%\end{figure}
 \subsection{Preprocessing}
@@
 %Some data is extracted from the points in the stroke.
- It is noted that as the user draws a shape, the pen slows down near corners and picks up speed when drawing a straight lines.
-Therefore, the speed and time difference data were widely used in sketch understanding systems \cite{earlyprocess}.
+ %It is noted that as the user draws a shape, the pen slows down near corners and picks up speed when drawing a straight lines. Therefore, the 
+ Speed and time difference data were widely used in sketch understanding systems \cite{earlyprocess}.
 The direction and curvature data are used to determine the points with high angular changes along the path of points \cite{meanshift10}.
  %speed is calculated as
+
  In the presented system, we computed time difference, direction, speed and curvature of each point along the stroke.
@@
  %$\Delta d/  \Delta S $ where d is the difference direction of point and s is difference in distance between points. 
- It should be noted that, all these calculations are performed in real time while the user draws the strokes. The complexity of computation is $O(n)$ where $n$ is number of points. \\
- Figure \ref{fig:speed2} shows the computed data for the stroke drawn in fig. \ref{fig:orignalStroke}. %If you look at the data curves you will see by looking at the data curves s
- It is noted that the dominant points are characterized by lower speed values and higher curvature and direction values (fig.
+ %It should be noted that, all these calculations are performed in real time while the user draws the strokes. The complexity of computation is $O(n)$ where $n$ is number of points.
+
+ Figure \ref{fig:speed2} shows the computed speed data for the stroke drawn in Fig. \ref{fig:orignalStroke}. %If you look at the data curves you will see by looking at the data curves s
+ It is noted that the dominant points are characterized by lower speed values and higher curvature and direction values(Fig.
 \ref{fig:speed2}).
-\\% the lower speed points correspond to vertex and dominant point's location.
+% the lower speed points correspond to vertex and dominant point's location.
 The higher the direction and curvature data correspond to location with higher change in curvature witch promote the location for vertices.
@@
 After the system computes all the speed, time difference and curvature data it proceeds to detect the points with low velocity and high curvatures.
-Using simple differentiations to detect local extreme points resulted in false points due to the non smooth curves.
-Hence, the system adopted a process presented by \cite{earlyprocess}, where the mean
+The system adopted the \textsl{average thresholding} process presented by \cite{earlyprocess}.
+This process is used to curvature, time difference and speed curves to generate a set of possible dominate points$P_{pd}$.
+These points are saved into a single array to be used as input to guide the next segmentation phase.
+%All the points labeled as possible dominant points $P_{pd}$ are saved into a single array.
+They are then used as input to guide the next segmentation phase.%Using simple differentiations to detect local extreme points resulted in false points due to the non smooth curves.
+Hence, the system adopted the \textsl{average threshold} process presented by \cite{earlyprocess}.
+This process is used to curvature, time difference and speed curves.
+All the points labeled as possible dominant points $P_{pd}$ are saved into a single array.
+They are then used as input to guide the next segmentation phase.
+%, where the mean
 %So instead the following procedure was used, firstly the mean of the curve is used as threshold for lower values then the high values are further investigated to find the extreme points.
 Hence, the mean
-of the curve is calculated and used as threshold \textit{th}.
+%of the curve is calculated and used as threshold \textit{th}.
 The threshold \textit{th} is used to separate the curve into regions  $Region_i$; each region $Region_i$ is defined as the part between two intersection points of the threshold line and the curve.
@@
 The maximum of each region is then computed and reported as a possible vertex point.
-\begin{figure}[]
-	\centering
-		\includegraphics[scale=0.5]{images/orignalStroke.pdf}
-	\caption{Stroke Sample} An example of the stroke drawn in the system.
-	\label{fig:orignalStroke}
-\end{figure}
-The system repeats this process to curvature, time difference and speed curves.
+%\begin{figure}[]
+%	\centering
+%	
+%		
+%	\caption{Stroke Sample} An example of the stroke drawn in the system.
+%
+%\end{figure}
+%The system repeats this process to curvature, time difference and speed curves.
 All the points labeled as possible dominant points $P_{pd}$ are saved into a single array.
 They are then used as input to guide the next segmentation phase.
-\begin{figure*}[]
+\begin{figure}[]
 %\begin{minipage}[b]{0.9\linewidth}
 	\centering
-
-		\subfigure[ The Speed of data] {\includegraphics[scale=0.25]{images/speed2.pdf}}	
+		\subfigure[ The Speed of data] {	\label{fig:speed2}\includegraphics[scale=0.2]{images/speed22.PNG}}	
+			\subfigure[Stroke Sample] {		\label{fig:orignalStroke} \includegraphics[scale=0.3]{images/orignalStroke.pdf}}
 	%	\subfigure[ Direction ] {\includegraphics[scale=0.15]{images/direction2.pdf}}
 					%\end{minipage}
-					\hspace{20 mm}
+		%			\hspace{20 mm}
 					%\begin{minipage}[b]{0.5\linewidth}
-			\subfigure[The Curvature] {\includegraphics[scale=0.29]{images/curvature2.pdf}}
+		%	\subfigure[The Curvature] {\includegraphics[scale=0.29]{images/curvature2.pdf}}
 	%		\subfigure[ Time Difference] {\includegraphics[scale=0.15]{images/time2.pdf}}
 %			\end{minipage}
-	\caption{Data Curves} The curve in a) shows the speed of the points in stroke in fig. \ref{fig:orignalStroke}. b) Shows the curvature curve for the same stroke.  %{The data of the stroke}
-
-	\label{fig:speed2}
-\end{figure*}
+	\caption{Data Curves} An example of the stroke drawn in the system with the curve that shows the speed of the points in the stroke.% in fig.
+\ref{fig:orignalStroke}.% b) Shows the curvature curve for the same stroke.
+%{The data of the stroke}
+\end{figure}
 %Description of input data. 
@@
 After computing the primary data the system tries to segment the stroke into a set of primitives.
-The segmentation algorithm first attempts to detect if the stroke is an ellipse.
-If the stroke proved to be an ellipse then the segmentation process ends and the system proceeds to the next step.
+The segmentation algorithm first attempts to fit the stroke points into a curve or an ellipse uisng a minimum square error fitting algorithm.
+If the stroke proved to be an ellipse arc then the segmentation process ends and the system proceeds to the next step.
 Otherwise, the stroke is passed into two particle swarm algorithms that divides the stroke to either lines or lines and curves.
-The algorithms take the stroke points along with the possible dominant points $P_{pd}$ computed then produce a set of dominant points which are connected with either lines or curves (see fig.
+The algorithms take the stroke points along with the possible dominant points $P_{pd}$ computed then produce a set of dominant points which are connected with either lines or curves (see Fig.
 \ref{fig:Blockdiagram}).
-The next section describes the ellipse detection algorithm and the two particle swarm algorithms used to divide the stroke.\\
+The next section describes the two particle swarm algorithms used to divide the stroke.
+% The segmentation algorithm first attempts to detect if the stroke is an ellipse or part of an ellipse.
+If the stroke proved to be an ellipse arc then the segmentation process ends and the system proceeds to the next step.
+Otherwise,
 %Paragraphs describe the segmentation algorithm.  
-\subsubsection{Ellipse Detection} 
-%Describe the ellipse detection algorithms 
-The process starts by computing the center of the stroke bounding box.
+%\subsubsection{Ellipse Detection} 
+%%Describe the ellipse detection algorithms 
+%The process starts by computing the center of the stroke bounding box.
 The bounding box center point is used as the first estimation of the ellipse center.
 The axes of the ellipse are estimated as $width/2$ and $height/2$ of the stroke bounding box.
-The least square fitting algorithm is used to minimize the fitting error of the ellipse equation eq.
-\ref{eq:circleFit}.
-\begin{equation}
-E = \sum\limits_{i = 0}^N {\frac{{(x_i - x_0 )}}{{a^2 }}^2  + \frac{{(y_i - y_0 )}}{{b^2 }}^2  - 1} 
-\label{eq:circleFit}
-\end{equation} where $N$ is the number of points in the stroke, $a,b$ are the length of ellipse axes, $x_0$ , $y_0$ are the coordinates of the center point and $x_i$ \& $y_i$ are the coordinates of point $i$ in the stroke.
-After few loops, the error and a confidence values  are computed to check if the stroke can be labeled as ellipse or not.\\ %is computed to check if stroke is an ellipse.
-\\
+The least square fitting algorithm is used to minimize the fitting error of the ellipse equation.
+%After few loops, the error and a confidence values  are computed to check if the stroke can be labeled as ellipse or not.
+% eq. \ref{eq:circleFit}.
+%\begin{equation}
+%E = \sum\limits_{i = 0}^N {\frac{{(x_i - x_0 )}}{{a^2 }}^2  + \frac{{(y_i - y_0 )}}{{b^2 }}^2  - 1} 
+%\label{eq:circleFit}
+%\end{equation} where $N$ is the number of points in the stroke, $a,b$ are the length of ellipse axes, $x_0$ , $y_0$ are the coordinates of the center point and $x_i$ \& $y_i$ are the coordinates of point $i$ in the stroke.
+%is computed to check if stroke is an ellipse. \\
 %We found that if the confidante is above threshold then the probability of ellipse is highest otherwise the stroke is passed to the next section to get the divisions of stroke and test its error.  
@@
 The problem definition is the same in both algorithms but they differ in the method they use to compute fitness and error functions.
-\subsubsection{Problem definition}
+\paragraph{Problem definition}
 The input stroke with $N$ points can be represented by set $S = \left\{ {x_1 ,x_2  \ldots x_N } \right\}$ where $x_i$ is the location of the point $i$.
@@
 When $p_{ij}=1$ means that point $j$ is a dominant point.
-% a!402B!
-%$A = \left\{ {p_i \left| {i = 1,2 \cdots N} \right.} \right\}$
-% a!402A!
-%\begin{equation}
-%A = \left\{ {p_i \left| {i = 1,2 \cdots N} \right.} \right\}.
-%\end{equation}
-\subsubsection{The algorithm}
+\paragraph{The fitness functions}
+%\subsubsection{The algorithm}
 The fitness function and error calculation are different in each of the two \textit{DPSO} algorithms. For the first algorithm \textsl{AlgS1}, 
@@
 It should be noticed that when the error is larger than the threshold $E$ the fitness is given a -ve value to lower the value of solution.
-Otherwise the system favors the lower number of vertices.\\ %we will want to lower the number of vertices.
-\\
-%if we say that s
-%Alg1:
-%Alg2: o
+Otherwise the system favors the lower number of vertices.
 The second algorithm \textsl{AlgS2} has the same problem formulation but different fitness and error functions.
@@
 \label{eq:fitnessSwarm2}
-\end{equation} where $E$ is the error and $M$ is the number of segments and $k$ is a parameter tweaked to get minimum number of segments. \\
+\end{equation} where $E$ is the error and $M$ is the number of segments and $k$ is a parameter tweaked to get minimum number of segments. 
 %Describe how we used particle swarm in curve polygon approximation using first algorithm.  
@@
 %[These two algorithm was modified from the algorithm in paper \cite{PolygonApproximationPSO,CruveDivisionSwarm}].
-\citeauthor{PolygonApproximationPSO} \cite{PolygonApproximationPSO} used a merge and divide algorithm after each loop of the swarm system to refine the solution but this paper is introducing another enhancement method.
+%\citeauthor{PolygonApproximationPSO} \cite{PolygonApproximationPSO} used a merge and divide algorithm after each loop of the swarm system to refine the solution but this paper is introducing another enhancement method.
 After each loop of the swarm algorithm (\textsl{AlgS1} and \textsl{AlgS2}), each particle is refined using the following procedure.
@@
 \subsection{Recognition}
-After the user draws all strokes of the symbol, he/she has to wait 10 seconds or press finish button beside the drawing area.
-The set of unrecognized strokes is grouped together along with their segmentation as input to the feature extraction process.
+After the user draws all strokes of the symbol, the set of unrecognized strokes is grouped together along with their segmentation as input to the feature extraction process.
 A composite set of features is used to generate a single feature vector.
-The features used consist of Rubine feature set \cite{gestureexample12},  Zernike moments of order 21 \cite{HeloiseBeautification}, ink density \cite{GeometryAndDomain102} as well as some structural and spatial information like number of perpendiculars lines ,number of parallel lines and types of primitives in each symbol.
+The features used consist of Rubine feature set,  Zernike moments of order 21 \cite{HeloiseBeautification}, ink density \cite{GeometryAndDomain102} as well as some structural and spatial information like number of perpendiculars lines ,number of parallel lines and types of primitives in each symbol.
 After computing the features the symbol is introduced to the classifier.
-The system uses a support vector machine (SVM) classifier with Gaussian kernel (RBF kernel)\cite{libsvm}.\\%Some features like number of parallel, perpendicular and intersecting lines are computed from the segmentation other like density, centroid and Zernike moments are computed from the raw stroke points.
+The system uses a support vector machine (SVM) classifier with Gaussian kernel (RBF kernel).%Some features like number of parallel, perpendicular and intersecting lines are computed from the segmentation other like density, centroid and Zernike moments are computed from the raw stroke points.
  %trained with the 13 categories in the system.
@@
 We divide the dataset into four sets each set contains four different users.
-We divide the dataset equally into training set and test set.
+Then we divide the dataset equally into training set and test set.
 To generate trusted results, four different splits for the training and test data are generated from the dataset.
@@
 \begin{figure}[]\centering
-\fbox{  \parbox{3cm}{% 
-		\includegraphics[scale=0.4]{images/symbolSet.pdf}	}}
+\fbox{ \parbox{3.5cm}{% 
+		\includegraphics[scale=0.4]{images/symbolSet.PNG}	}}
 	\caption{The Symbol Set} Ths figure shows a set of the symbols used in the dataset. 
@@
 \end{figure}
-\begin{figure}[]
+\begin{figure*}[]
 	\centering
-		\includegraphics[scale=0.5]{images/test1.jpg}
-	\caption{Algorithm comparison} The results of comparing the algorithms. The graphs displays the recognition rate of symbols. 
-	\label{fig:test1}
-\end{figure}
+		\subfigure[Algorithm comparison]{\includegraphics[width=5cm]{images/result1.jpg} 	\label{fig:test1}}
+		\hfill
+  \subfigure [Symbols Comparison ] {\includegraphics[width=10.5cm]{images/result2.jpg} \label{fig:test2}}
+	\caption{Experiments results:}  a)The results of comparing the algorithms.
+The graphs displays the recognition rate of symbols.
+b) The graph shows the recognition rate of each symbol using different algorithms.
+\end{figure*}
 %In the figure \ref{fig:test1}, the accuracy means that final hit rate of all symbols used in the test set.by the  four datasets . 
 We performed two experiments to test the system firstly we tested recognition accuracy of shapes in the data set with both algorithms (\textsl{AlgS1} and \textsl{AlgS2}).
-We also implemented the segmentation algorithm described in \cite{earlyprocess} (\textsl{Alg3}) to use as reference to our swarm algorithms.
+%We also implemented the segmentation algorithm described in \cite{earlyprocess} (\textsl{Alg3}) to use as reference to our swarm algorithms.
   Figure \ref{fig:test1} shows the accuracy achieved by each algorithm.
 The results shows that both \textit{DPSO} algorithms achieve better result than other algorithms.
-The swarm algorithms were tested with and without the ellipse detection module.
-The ellipse detection module appears to be superior to results with only the \textit{DPSO} algorithms.\\
-
+The swarm algorithms were tested with and without the ellipse fitting module.
+The ellipse detection module appears to be superior to results with only the \textit{DPSO} algorithms.
 The second experiment we implemented is testing the effect of symbol complexity and type on the recognition rate.
@@
 This is natural as algorithm  \textsl{AlgS1} divides strokes into line segments only but \textsl{AlgS2} is able to divide strokes into lines and curves based on the minimum error of the segment itself.
-Algorithm \textsl{Alg3} gives good performance as long as the symbols consist of lines and curves, if the stroke consists of curve only  the algorithm may lead to wrong segmentation result.
+%Algorithm \textsl{Alg3} gives good performance as long as the symbols consist of lines and curves, if the stroke consists of curve only  the algorithm may lead to wrong segmentation result.
 This is because the system divides the stroke first to line segments then tries to decide if each segment can be represent better as a curve unlike algorithm \textsl{AlgS2} where the curve segments are tested while choosing the best segmentation.
@@
 \\
-\begin{figure*}[]
-	\centering
-		\includegraphics[scale=0.8]{images/test2.jpg}
-	\caption{Symbols Comparison} The graph shows the recognition rate of each symbol using different algorithms. 
-	\label{fig:test2}
-\end{figure*}
+%\begin{figure}[]
+%	\centering
+%		\includegraphics[scale=0.4]{images/test2.jpg}
+%	\caption{Symbols Comparison} The graph shows the recognition rate of each symbol using different algorithms. 
+%	\label{fig:test2}
+%\end{figure}
 %We could not test the result of the segmentation algorithm directly due to the fact that the correct segmentation is highly ambiguous.
@@
 This makes the system easily expandable with aspect to symbols recognized.
-\\
+
 %As you can see form the experiments PSO have proved a spuriously than other system. 

### 2009-02-11 trunk/smallpaper/SmallPaper.tex
before dr. magdah changes..

--- 2009-01-25 smallpaper/SmallPaper.tex
@@
-\documentclass[a4paper,10pt]{IEEEconf}
-\usepackage[latin1]{inputenc}
-\usepackage[american]{babel}
+\documentclass{article}%{IEEEconf}[a4paper,10pt]
+\usepackage{spconf}
+\usepackage{amsmath}
+%\usepackage[latin1]{inputenc}
+%\usepackage[american]{babel}
 %\usepackage[T1]{fontenc}
@@
 %\usepackage{natbib}
-\bibliographystyle{named}
+
+%\bibliographystyle{named}
 \usepackage[dvips,pdftex]{hyperref}
@@
 %\title{A Study on Using Particle Swarm Algorithms in Symbol Recognition }
-\author{Maha El Meseery,  Mahmoud Fakhr El Din,  Samia Mashali,
-\\  Magda Fayek and Nevin Darwish}
+%\author{Maha El Meseery ;  Mahmoud Fakhr El Din ;  Samia Mashali ;  Magda Fayek and Nevin Darwish }
+%\name{Maha El Meseery ;  Mahmoud Fakhr El Din ;  Samia Mashali ;  Magda Fayek and Nevin Darwish }
+
+\twoauthors
+  {Maha El Meseery, Mahmoud Fakhr El Din, Samia Mashali}
+  %{}
+	{Signals Processing Group \\ Computers and Systems Department \\
+  Electronic Research Institute\\Cairo, Egypt\\
+	melmeseery@eri.sci.eg, samia@eri.sci.eg}
+  {Magda Fayek, Nevin Darwish}%\sthanks{The fourth author performed the work while at ...}
+  %{}
+	{Department of Computer Engineering\\
+	Faculty of Engineering \\Cairo University\\Cairo, Egypt\\
+	 magda@gmail.com,ndarwish@ieee.org}
+
+%\address{ Electronic research inistiute}
 \begin{document}
@@
 \end{abstract}
+\begin{keywords}
+Handwriting recognition, pattern recognition ,Document image processing
+\end{keywords}
 \section{Introduction}
 %A picture speak a thousand words is also true for sketches. 
+The field of symbol recognition has gained interest in the last few years.
+A wide variety of techniques were used either on segmentation process or symbol recognition.
 Scientists generally and engineers specifically express thoughts and designs using sketches.
@@
 Design engineers need a powerful computer based system with symbol recognition to help them design, manipulate and store sketches more effectively than using only papers.
-Increasing the interaction between computers and users in sketch and CAD systems has been the reason for the emerging of few advanced sketch recognition systems \cite{CAD}.
+Increasing the interaction between computers and users in sketch and CAD systems has been the reason for the emerging of few advanced sketch recognition systems.
 %Sketch recognition system converts the drawings and informal graphs user draws into objects and symbols that can be used in further interaction with the system.
-Sketch recognition is divided into four main steps preprocessing, segmentation, symbol recognition and finally sketch understanding.
+Sketch recognition is divided into three main steps preprocessing, segmentation and symbol recognition.
 The preprocessing process saves user input stroke points from the hardware and collects basic information about the stroke then proceeds to remove noise and primary data calculation.
@@
 The system segments stroke user draws then cluster symbols together then passes them to the SVM symbol recognizer.
-The experiments in section \ref{sec:Experiments} shows that the proposed system improves the final symbol recognition accuracy.
+The experiments in section \ref{sec:Experiments} shows that the proposed system can improve the final symbol recognition accuracy.
 %The remaining of this papers is organized as following; section \ref{Sec:RelatedWork} provides a brief literature review.
@@
 Finally, section \ref{sec:Experiments} will describe the experiments that was preformed.\\
-\section{Literature Review}
-\label{Sec:RelatedWork}
+%\section{Literature Review}
+%\label{Sec:RelatedWork}
 %%%What is sketch recognitions. Why we need sketch recognition systems.
-The field of symbol recognition has gained interest in the last few years. A wide variety of techniques were used either on segmentation process or symbol recognition.
+
 %To segment strokes
@@
 A hybrid algorithm was introduced in \cite{earlyprocess} where they generated different sets of segmentation based on both curvature and speed dominant points, this is followed by choosing a segmentation with the least error from generated hybrid set.
-Their system has a draw back as it only recognizes simple specific shapes with a set of low level recognizers.
-\citeauthor{meanshift10} introduced a feature area for each primitive and then computed the segmentation error for different types of primitives based on the computed feature area.
+Their system has a draw back as it only recognizes specific simple geometric shapes with a set of low level recognizers.
+Yu introduced a feature area for each primitive and then computed the segmentation error for different types of primitives based on the computed feature area.
 Their system achieved good accuracy in simple shapes (square, ellipse,...) but did not perform well in more complex shapes\cite{meanshift10}.
-\citeauthor{Paleosketch08} \cite{Paleosketch08} introduced a set of low level recognizers that were reported to achieve 98\%  accuracy but their system similar to all low level recognizers only identify a small set of simple shapes.
+Paulson and Hammond\cite{Paleosketch08} introduced a set of low level recognizers that were reported to achieve 98\%  accuracy but their system similar to all low level recognizers only identify a small set of simple shapes.
 
@@
 Genetic algorithm was used by \cite{CruveDivisionSwarm} to optimally divide digital curves into lines and curves.
-\citeauthor{CruveDivisionSwarm} uses digital curves scanned from paper as input to the system and did not take advantage of the curvature or local geometric properties of the digital curve.
-\citeauthor{PolygonApproximationPSO} \cite{PolygonApproximationPSO} used PSO to convert digital curves into polygons, our system adopts \citeauthor{PolygonApproximationPSO} method but improves it by adding curvature and other local information while segmenting strokes to achieve better segmentations.% The next section presents the general particle swarm algorithm which is used in our proposed segmentation algorithms.
+Chen et al.
+uses digital curves scanned from paper as input to the system and did not take advantage of the curvature or local geometric properties of the digital curve.
+Yin \cite{PolygonApproximationPSO} used PSO to convert digital curves into polygons, our system adopts Yin's method and improves it by adding curvature and other local information while segmenting strokes to achieve better segmentations.% The next section presents the general particle swarm algorithm which is used in our proposed segmentation algorithms.
 % are presented in section \ref{subsubsec:Discreteparticleswarmalgorithm}
@@
 %Figure \ref{fig:Blockdiagram} 
-The following sketch recognitions systems is divided into three main system blocks are 1) preprocessing, 2) segmentation and 3) Recognition.
+The following sketch recognitions systems is divided into three main system steps are 1) preprocessing, 2) segmentation and 3) Recognition.
 The following sections provides a detail description of each step.
-
  %Due to sloppiness of users and hardware glitches the system preprocess the data before further processing. 
@@
   %then passes through the segmentation process which will try to 
-
  %The goal of the segmentation stage is to divide stroke into segments of either curves or lines.
@@
 % is the ellipse fitting
-
  %the stroke first If the ellipse detection fails the stroke is passed to the segmentation algorithms which will pass it to the two PSO algorithms described below the segmentation with the minimum error will be the chosen segmentation.
@@
  %speed is calculated as 
-
  In the presented system, we computed time difference, direction, speed and curvature of each point along the stroke.
@@
 $c= \Delta d/\Delta s$.
-
  %system compute speed, distance data for all points in the stroke. Curvature is computed using estimation used in [] where direction computed is angle between two lines. Curvature is 
@@
 After the system computes all the speed, time difference and curvature data it proceeds to detect the points with low velocity and high curvatures.
+Using simple differentiations to detect local extreme points resulted in false points due to the non smooth curves.
+Hence, the system adopted a process presented by \cite{earlyprocess}, where the mean of the curve is calculated and used as threshold \textit{th}.
+The threshold \textit{th} is used to separate the curve into regions  $Region_i$; each region $Region_i$ is defined as the part between two intersection points of the threshold line and the curve.
+Regions higher than the threshold \textit{th} are processed to find the maximum point $Max(Region_i)$ of each region $Region_i$.
+The stroke points $p_i(x,y)$ that correspond to those maximum values are labeled as \textit{possible dominant points} $P_{pd}$.
+The system repeats this process to curvature, time difference and speed curves.
+All the points labeled as possible dominant points $P_{pd}$ are saved into a single array.
+%They are then used as input to guide the next segmentation phase.
+%After the system computes all the speed, time difference and curvature data it proceeds to detect the points with low velocity and high curvatures.
 The system adopted the \textsl{average thresholding} process presented by \cite{earlyprocess}.
@@
 Otherwise, the stroke is passed into two particle swarm algorithms that divides the stroke to either lines or lines and curves.
-The algorithms take the stroke points along with the possible dominant points $P_{pd}$ computed then produce a set of dominant points which are connected with either lines or curves (see Fig.
-\ref{fig:Blockdiagram}).
-The next section describes the two particle swarm algorithms used to divide the stroke.
+The algorithms take the stroke points along with the possible dominant points $P_{pd}$ computed then produce a set of dominant points which are connected with either lines or curves.
+The next section describes the ellipse detection algorithm and the two particle swarm algorithms used to divide the stroke.\\ %The next section describes the two particle swarm algorithms used to divide the stroke.
+%(see Fig.
+\ref{fig:Blockdiagram})
 % The segmentation algorithm first attempts to detect if the stroke is an ellipse or part of an ellipse.
@@
 %Paragraphs describe the segmentation algorithm.  
-%\subsubsection{Ellipse Detection} 
+ \begin{figure}
+	\centering
+		\includegraphics[scale=0.7]{images/segblock.jpg}
+	\caption{Segmentation Block}
+	\label{fig:segblock}
+\end{figure}
+\subsubsection{Ellipse Detection} 
 %%Describe the ellipse detection algorithms 
-%The process starts by computing the center of the stroke bounding box.
+The process starts by computing the center of the stroke bounding box.
 The bounding box center point is used as the first estimation of the ellipse center.
@@
 The problem definition is the same in both algorithms but they differ in the method they use to compute fitness and error functions.
-\paragraph{Problem definition}
-The input stroke with $N$ points can be represented by set $S = \left\{ {x_1 ,x_2  \ldots x_N } \right\}$ where $x_i$ is the location of the point $i$.
-The swarm algorithms consist of $M$ agents which are represented by the set
-$A  = \left\{ {P_i \left| {i = 1,2 \cdots M} \right.} \right\}$ where $P_i$ is a single solution particle from the solution space.
+\\
+%\paragraph{Problem definition}
+\textsl{Problem definition:} The input stroke with $N$ points can be represented by set $S = \left\{ {x_1 ,x_2  \ldots x_N }\right\}$ where $x_i$ is the location of the point $i$.
+The swarm algorithms consist of $M$ agents which are represented by the set  $A = \left\{ {P_i \left| {i = 1,2 \cdots M} \right.} \right\}$ where $P_i$ is a single solution particle from the solution space.
 Each particle decodes the problem with binary array with the same length $N$ as the input stroke.
@@
 When $p_{ij}=1$ means that point $j$ is a dominant point.
-\paragraph{The fitness functions}
+\\
+\begin{figure}
+	\centering
+			\subfigure[Particle Encoding]{	\includegraphics[scale=0.5]{images/pso1.jpg}}
+					\subfigure[Posible domninate point]{	\includegraphics[scale=0.5]{images/ppd.jpg}}
+	\caption{}
+	\label{fig:pso1}
+\end{figure}
+%\paragraph{The fitness functions}
 %\subsubsection{The algorithm}
-The fitness function and error calculation are different in each of the two \textit{DPSO} algorithms. For the first algorithm \textsl{AlgS1}, 
+\textsl{Fitness function:} The fitness function and error calculation are different in each of the two \textit{DPSO} algorithms. For the first algorithm \textsl{AlgS1}, 
  the arc $\widehat{x_ix_j}$ is defined as the consecutive set of points from point$x_i$ to point $x_{j}$ as in $x_i,x_{i+1} \cdots,x_j$.
@@
 \end{equation} where $N$ is the number of points in the stroke, $D$ is the number of points in the solution that was previously labeled as a possible dominant point ($P_{pd}$), $E$ is the computed error and $\varepsilon$ is the error threshold.
-It should be noticed that when the error is larger than the threshold $E$ the fitness is given a -ve value to lower the value of solution.
+It should be noticed that when the error is larger than the threshold $E$ the fitness is given a -ve value to lower the fitness value of the solution.
 Otherwise the system favors the lower number of vertices.
@@
 %Describe the features used in recognition [ref to \cite{zernike61}]
-\section{Experiments}\label{sec:Experiments}
-A dataset collected by \citeauthor{HeloiseBeautification} is used to test the presented system\cite{HeloiseBeautification}.
+\section{Experiments}
+\label{sec:Experiments}
+The dataset collected by Hse and Newton is used to test the presented system\cite{HeloiseBeautification}.
 The data are drawn by 16 users each of them was asked to draw the 13 shapes from 30 to 50 times.
@@
 The algorithms are then tested using the four sets and the results are averaged.
-\begin{figure}[]\centering
-\fbox{ \parbox{3.5cm}{% 
-		\includegraphics[scale=0.4]{images/symbolSet.PNG}	}}
-	\caption{The Symbol Set} Ths figure shows a set of the symbols used in the dataset. 
-	\label{fig:symbolSet}
-\end{figure}
+%\begin{figure}[]\centering
+%	\caption{The Symbol Set} 
+%	\label{fig:symbolSet}
+%\end{figure}
 \begin{figure*}[]
 	\centering
+	\begin{minipage}[b]{4cm}
+	\subfigure [The Symbol Set] {\fbox{ \parbox{2cm}{% 
+		\includegraphics[scale=0.2]{images/symbolSet.PNG}	}}}
+
 		\subfigure[Algorithm comparison]{\includegraphics[width=5cm]{images/result1.jpg} 	\label{fig:test1}}
+		\end{minipage}
 		\hfill
   \subfigure [Symbols Comparison ] {\includegraphics[width=10.5cm]{images/result2.jpg} \label{fig:test2}}
-	\caption{Experiments results:}  a)The results of comparing the algorithms.
+	\caption{Experiments results:} a) The figure shows a set of the symbols used in the dataset.
+b)The results of comparing the algorithms.
 The graphs displays the recognition rate of symbols.
-b) The graph shows the recognition rate of each symbol using different algorithms.
+c) The graph shows the recognition rate of each symbol using different algorithms.
 \end{figure*}
 %In the figure \ref{fig:test1}, the accuracy means that final hit rate of all symbols used in the test set.by the  four datasets .
+
 We performed two experiments to test the system firstly we tested recognition accuracy of shapes in the data set with both algorithms (\textsl{AlgS1} and \textsl{AlgS2}).
@@
 The ellipse detection module appears to be superior to results with only the \textit{DPSO} algorithms.
-The second experiment we implemented is testing the effect of symbol complexity and type on the recognition rate.
+
+The second experiment we implemented was testing the effect of symbol complexity and type on the recognition rate.
 Figure \ref{fig:test2} shows the accuracy of each symbol, it is clearly noted that symbols that have only line segments achieve higher accuracy rate than other symbols.
@@
 %Each new stroke is checked if it can be a symbol or is a part of already drawn un completed symbol.
+
 %\bibliography{Mybibliography}
+\bibliographystyle{IEEEbib}
 \bibliography{../../neededfiles/Bibliographies/Mybibliography2}

### 2009-02-18 draft/SketchMaster.tex
now i am reviewing

--- 2008-09-02 trunk/draft/SketchMaster.tex
@@
 \usepackage{psfig}
-
+%\usepackage{named}
 \usepackage{natbib}
@@
 \usepackage{subfigure}
+\usepackage{longtable}
+\usepackage{array}
+%\usepackage[T1]{fontenc}
+\usepackage[latin9]{inputenc}
 
@@
                                                         %======================
-\title{Sketch recogniton and understanding}             % My theses title
+\title{Sketch Recogniton using Particle Swarm Algorithm}             % My theses title
 \author{Maha Mohammed Nabeel El Meseery}                               % Y name
-\prevdegrees{B.S.,Computer  engineering(2003)}       		% Your old degree
+\prevdegrees{B.S.,Computer  Engineering(2003)}       		% Your old degree
 \major{Computer Engineering}                            % My current major
 \degree{Master of Science}                              % My new degree
-\degreemonth{Dec}                                      % When awarded. to be added when known
-\degreeyear{2005}                                      %
+\degreemonth{}                                      % When awarded. to be added when known
+\degreeyear{}                                      %
 %th today or first day of writing  December 26 2005 
-\thesisdate[]                                     % Date of document.
+%\thesisdate[]                                     % Date of document.
 \DOCUMENTtype{Survey}                                   % or DISSERTATION
@@
 \maketitle
+
+\providecommand{\tabularnewline}{\\}
                                                         % COPYRIGHT PAGE
                                                         %======================
-\copyrightyear{2005}                                    % Delete these
+\copyrightyear{2008}                                    % Delete these
 \makecopyright                                          % if no copyright
@@
 Even though the area of research progressed quickly in speech and handwriting systems gesture and sketch understanding systems are still starting to get noticed.
-This survey will highlight the current research that is ongoing on the sketch recognition and sketch understanding systems.
+
+Sketch recognition is defined as the process of identifying the symbols the user draws using a single or multiple strokes.
+Users draw strokes using a pen and the system immediately interprets their strokes into objects that can be easily manipulated.
+This research uses Particle Swarm Algorithm (PSO) to divide the strokes the user draws into meaningful geometric primitives.
+These geometric primitives are grouped to formulate symbols which will be further identified.
+The research focus on the effect of PSO segmentation to the sketch recognition systems.
+
+Results showed that using PSO improves segmentation results which guide the symbol recognition phase. Final recognition has improved accuracy after the PSO algorithm using a SVM classifier.  
+ %Hand drawn sketches are the natural method of exchanging thoughts between engineers and humans in general.
+Nevertheless, researchers barely give attention to improve current gesture and sketch recognition systems.
+In a sketch understanding system, users need to feel the freedom of paper and pencil along with the flexibility of computer manipulation.
 
@@
 \include{ChapterIntroduction}                              % The best way to organize your
-\include{chapterSurvey}                              % document is probably to keep
+%\include{chapterSurvey}                              % document is probably to keep
+%\include{System}
+%\include{Results}
+%\include{ConFut}
                                                 % in its own file.  
@@
 
-\bibliography{../../neededfiles/Bibliographies/Mybibliography.bib}     
+\bibliography{../../neededfiles/Bibliographies/Mybibliography}     
                                                 % THE APPENDIX

### 2009-02-18 trunk/draft/ChapterIntroduction.tex
now i am reviewing

@@
 %\usepackage{graphicx}
-\chapter {Introduciton}
+\chapter {Introduction}
 \pagenumbering{arabic}
@@
 Computer research nowadays aimed at facilitating computer human interaction in every way.
-Pen-based interface gives the user a pencil-paper like feeling that enhances interaction more than the current used keyboard-mouse computer interfaces.
+Pen-based interfaces give the user a pencil-paper like feeling that enhances interaction more than the current used keyboard-mouse computer interfaces.
 However, until now there are hardly any complete pen based computer systems.
 Generally, Current Pen-Based systems make use of the Pen or stylus to perform the same role as the mouse / keyboard.
-The minority of such systems are equipped with handwriting or digit recognition that is typically restricted for specific language.
+A minority of such systems are equipped with handwriting or digit recognition that is typically restricted for specific language.
 
- Despite the fact that people frequently express their ideas through drawing them there isn't a system that can fully understand those sketches.
-Lately, only a few researchers were concerned with the problem although it is highly important.
-Firstly, we need to know;
- \textit { What are the main goals of Sketch understanding systems?}
+\section{Sketch Recognition}
+Sketch recognition is defined as the process of identifying symbols that users draw using single or multiple strokes.
+Users draw strokes using a pen and the system should immediately interprets their strokes into objects that can be easily manipulated.
 
-The main goal of sketch understating systems is to build a system that will interpret the sketches the user draw.
-The systems ought to be natural as possible to provide user with paper and pencil impression.
-Furthermore, interpreted sketches are expected to transform into objects, which can me manipulated, modeled and/or simulated.
-
-Sketch here stand for any hand drawn drawing the user draws using digital stylus or similar device.
+Sketchs here stand for any hand drawn drawing the user draws using digital stylus or similar device.
 Normally, sketch are composed of smaller segments we call strokes.
@@
 
-Sketch recognition is procedure that will input a sequence of points from user then segment them into clusters of strokes and recognize those clusters to primitive geometric objects and shapes.
+%Sketch recognition is procedure that will input a sequence of points from user then segment them into clusters of strokes and recognize those clusters to primitive geometric objects and shapes.
 However, Sketch understanding is transforming the sketch from a group of primitive shapes into comprehendible identified symbols consistent with each other.
 % the entire sketched symbols.
+%Sketch recognition is defined as the process of identifying symbols that users draw using single or multiple strokes.
+Users draw strokes using a pen and the system should immediately interprets their strokes into objects that can be easily manipulated.
+This paper uses Particle Swarm Algorithm (PSO) to divide the strokes the user draws into meaningful geometric primitives.
+These geometric primitives are grouped to formulate symbols which are further identified.
+The results show that using PSO improves segmentation results which guide the symbol recognition phase.
+This paper uses Support Vector Machine(SVM) classifier which further improves the final recognition accuracy.
+%u Final recognition has improved accuracy after the PSO algorithm using a SVM classifier.
+ %Hand drawn sketches are the natural method of exchanging thoughts between engineers and humans in general.
+Nevertheless, researchers barely give attention to improve current gesture and sketch recognition systems.
+In a sketch understanding system, users need to feel the freedom of paper and pencil along with the flexibility of computer manipulation.
+
+ %A picture speak a thousand words is also true for sketches. 
+%Scientists generally and engineers specifically express thoughts and designs using sketches.
+Engineers use sketches to exchange designs as a natural method of communication rather than writing or speaking.
+Engineers use paper and pencil design in early stages of design.
+Design engineers need a powerful computer based system with symbol recognition to help them design, manipulate and store sketches more effectively than using only papers.
+Increasing the interaction between computers and users in sketch and CAD systems has been the reason for the emerging of few advanced sketch recognition systems \cite{CAD}.
+Sketch recognition system converts the drawings and informal graphs user draws into objects and symbols that can be used in further interaction with the system.
+
+%Sketch recognition is divided into four main steps preprocessing, segmentation, symbol recognition and finally sketch understanding.
+The preprocessing process saves user input stroke points from the hardware and collects basic information about the stroke then proceeds to remove noise and primary data calculation.
+In segmentation phase, the system divides the stroke into a set of simple geometrical primitives or segments.
+The third phase, sketch recognition, is clustering strokes and segments to formulate symbols that can be recognized by a classifier system.
+Finally, the sketch understanding phase uses priori information about the sketch to support identification of symbols by excluding any possible invalid symbols.
+%the basic symbols are identified in consistent with the whole sketch. \\
+%from the user information from the users converts the strokes the user draw into segments which is recognized into basic symbols that can be easily manipulated.
+Finally, the basic symbols are interpreted in consistent with the whole sketch.
+\\
+%This paper introduces a sketch recognition system.
+The presented system uses' particle swarm algorithm (PSO) to optimally segment users strokes.
+Users can draw symbols using any number of strokes in any order they like.
+The system segments stroke user draws then cluster symbols together then passes them to the SVM symbol recognizer.
+The experiments in section \ref{sec:Experiments} shows that the proposed system improves the final symbol recognition accuracy.
+
+%The remaining of this papers is organized as follows; section \ref{Sec:RelatedWork} provides a brief literature review.
+Section \ref{PSO} explains the general particle swarm algorithm.
+The proposed system block diagram along with a detailed system description is presented in Section \ref{Sysdisc}.
+The experiments and results achieved are explained in Section \ref{sec:Experiments}.
+Finally, section \ref{ConclusionandFutureWork} states the conclusion and future work.
+%The next section will relate previous work in the field.
+The remaining of the paper will be as following; Section \ref{PSO} will explain the general particle swarm algorithm which is used in the presented system.
+Section \ref{Sysdisc} will describe the system block diagram and then section \ref{Prepross} will explain the preprocessing step, section \ref{seg} will describe the segmentation algorithm in details.
+Finally, section \ref{sec:Experiments} will describe the experiments that was preformed.\\
 
@@
 
-People normally sketch thoughts rather than expressing them linguistically.
+Scientists generally and engineers specifically express thoughts and designs using sketches.
+Engineers use sketches to exchange designs as a natural method of communication rather than writing or speaking.
+Engineers use paper and pencil design in early stages of design.
+The sketching process itself go thorough numerous stages.
+Firstly, early stages with rough drafting designs or notations which are mainly done using paper and pencil.
+After those primary rough sketches are edited and altered until they are stabilized and finally they are transformed into the computer.
+
+%Design engineers need a powerful computer based system with symbol recognition to help them design, manipulate and store sketches more effectively than using only papers.
+Increasing the interaction between computers and users in sketch and CAD systems has been the reason for the emerging of few advanced sketch recognition systems \cite{CAD}.
+Sketch recognition system converts the drawings and informal graphs user draws into objects and symbols that can be used in further interaction with the system.
+
+%People normally sketch thoughts rather than expressing them linguistically.
 For instance engineers mostly drafts their early designs using paper and pencil before using CAD tools.
@@
  Paper and Pencil are preferred in design as they offer a remarkable ease in editing and altering the primary rough sketches.
-Current CAD systems don't give the designer the ease of altering the early designs.
+Current CAD systems do not give the designer the ease of altering the early designs.
 Designer had to sketch primary designs on paper and pencil next, the finally design is transferred into computer using a menu CAD system.
@@
 
-\section{Problems and challenges}
+\section{Applicability }
+% talk about application like cad systems , software engineering uml diagrams ....
+% meeting idea brainstorming , mind mapping string ..
+% communication program  
+% sketching programs ...
+Application of sketch recognition are ranging from engineering device simulation and design.
+Drawing and sketching programs are most likely to benefit from integrating a sketch recognition system into there design environments.
+Other applications include object retrieval, map navigations and various engineering design systems.
+
+\section{Research Scope}
+%rewirte this part.............................
+%  research will invistgated till sketch recognition in domain of uml diagrams. 
+%This research will include investigating the best method to recognize sketches.
+The scope of the research will include recognizing various sketches.
+The methods used in this research are comparing current systems and enhancing them and adobpiting them into simple geometrical symbols.
+Using well established AI methods specially \textit{Swarm Intelligence} to solve the problem.
+
+\section{Problem Definition}
 \label{sec:ProblemsAndChallenges}
+
+%Sketch recognition is divided into three main steps preprocessing, segmentation and symbol recognition.
+The preprocessing phase captures user input stroke points and collects basic information about the stroke then proceeds to remove noise and compute basic statistical and geometrical information.
+In segmentation phase, the strokes are divided into a set of simple geometrical primitives or segments.
+In the third phase, sketch recognition, strokes and segments are clustered to formulate symbols that can be recognized by a classifier system.
 
@@
 
-\subsection{Search space}
-\label{sec:Searchspace}
-
-The user drawn symbols can be having more than one interpolation.
+\section{Thesis Overview}
+This thesis is arranged such that chapter \ref{} presents a survey of different components of the
+sketch recognition. In chapter \ref{}, more light is shed on the feature filter technique for dimensionality
+reduction. This is in conjunction with a detailed description of the proposed methods to
+enhance the feature filtering. Comparative studies among the proposed methods and stateof-
+arts methods are presented in chapter 4. Finally, chapter 5 presents the conclusions of
+this thesis and suggested future work.
+%\subsubsection{Search space}
+%\label{sec:Searchspace}
+%
+%
+%The user drawn symbols can be having more than one interpolation.
 In addition, those interpolations increase exponentially with the number of strokes the user draw and as the sketch expands the number of interpolation escalate.
 This intensifies the search space that the program search to resolve the ambiguities and find best interpolation.
+%
+%\sublsubsection{Real time}
+%\label{sec:Realtime}
+%
+%
+%There are mainly three types of sketch understanding system offline \cite{Vibratory8}, online and interactive.
+Both online and interactive recognition like \cite{incrmentintention41, SmartSketch56} need the processing after each stroke to be real time so the user does not feel the processing.
 
-\subsection{Real time}
-\label{sec:Realtime}
+%A picture speak a thousand words is also true for sketches. 
+%Scientists generally and engineers specifically express thoughts and designs using sketches.
+Engineers use sketches to exchange designs as a natural method of communication rather than writing or speaking.
+Engineers use paper and pencil design in early stages of design.
+Design engineers need a powerful computer based system with symbol recognition to help them design, manipulate and store sketches more effectively than using only papers.
+Increasing the interaction between computers and users in sketch and CAD systems has been the reason for the emerging of few advanced sketch recognition systems \cite{CAD}.
+Sketch recognition system converts the drawings and informal graphs user draws into objects and symbols that can be used in further interaction with the system.
 
-There are mainly three types of sketch understanding system offline \cite{Vibratory8}, online and interactive.
-Both online and interactive recognition like \cite{incrmentintention41, SmartSketch56} need the processing after each stroke to be real time so the user does not feel the processing.
+%Sketch recognition is divided into four main steps preprocessing, segmentation, symbol recognition and finally sketch understanding.
+The preprocessing process saves user input stroke points from the hardware and collects basic information about the stroke then proceeds to remove noise and primary data calculation.
+In segmentation phase, the system divides the stroke into a set of simple geometrical primitives or segments.
+The third phase, sketch recognition, is clustering strokes and segments to formulate symbols that can be recognized by a classifier system.
+Finally, the sketch understanding phase uses priori information about the sketch to support identification of symbols by excluding any possible invalid symbols.
+%the basic symbols are identified in consistent with the whole sketch. \\
+%from the user information from the users converts the strokes the user draw into segments which is recognized into basic symbols that can be easily manipulated.
+Finally, the basic symbols are interpreted in consistent with the whole sketch.
+\\

### 2009-02-18 trunk/draft/Results.tex
now i am reviewing

@@
 \chapter{Results}
+
+This chapter discribe and illustrate the results generated from the system.
+In the chapter, section \ref{sec:ExperimentsSetup} will describe the experiments performed on each block of the system.
+section \ref{sec:Datasets} will describe the datasets used.
+Section \ref{sec:ResultsDetails} will demonstrate and analysis the result generated.
 
@@
 \label{sec:ExperimentsSetup}
+The system was tested using various experiments.
+The preprocessing procedures and computation was tested by its effects on the segmentation algorithm and was reported in section (\ref{sec:ResultsDetails}).
+The variable used to tune the swarm algorithm were tested to choose the best parameters to use(see section \ref{sec:PSO}).
+Segmentation algorithms test was reported and noted in section \ref{sec:SegmentationAlgorithms}.
+
 \section{Datasets}
 \label{sec:Datasets}
-\section{Evaluations techniques}
-\label{sec:EvaluationsTechniques}
+For classifier and algorithm testing the system used a data set collected by \citeauthor{HeloiseBeautification}.
+The data are drawn by 16 users each of them have drawn from 30 to 50 samples for each shape.
+Figure \ref{fig:symbolSet} shows the shapes used in the data set.
+We divided the dataset equally into training set and test set.
+The results displayed are the recognition accuracy resulted from the classifier after recognizing the symbols.
+\\
+%We divided the dataset into four sets each set contains four different users.
+Furthermore, each set is divided into train and test set where the trained set where used to train the classifiers and the test set where used for testing.
+The algorithms are then tested using the four sets and the results are averaged.
+\begin{figure}[]\centering
+\fbox{  \parbox{8cm}{% 
+		\includegraphics[scale=0.7]{images/symbolSet.eps}	}
+		}
+	\caption{The Symbol Set}
+	\label{fig:symbolSet}
+\end{figure}
 
-\section{Segmentation Algoirthms Results}
-\label{sec:SegmentationAlgoirthmsResults}
+%\section{Evaluations techniques}
+%\label{sec:EvaluationsTechniques}
+%\section {Curvature estimation results}
+%\label{sec:Curvatureestimationresults}
 
-\subsection{Swarm Algorithms}
-\label{sec:Swarm Algorithms}
+\section{Results}
+\label{sec:ResultsDetails}
 
-\section{Clustering Algorithm Results}
-\label{sec:ClusteringAlgorithmResults}
+\subsection{PSO Algorithm}
+\label{sec:PSO}
+The fig.
+\ref{fig:dataerrorvsiteration.jpg} display the effect of the size of swarm population on the number of vertex reported while segmenting the stroke.\footnote{The fewer the vertex the better the segmentation} The graphs shows that as the number of iteration increase there is a decrease in error until the error reach a saturation and any increase in number of iteration dose not affect the error calculated.
+Similar behavior is noticed in fig.
+\ref{fig:datapopulation.jpg}.
+These curves and test result in choosing the swarm parameters $c_1,c_2$ and maximum number of iteration and the swarm population.
+The final values are for maximum number of iteration = 150 with 15 particle for better compensation in the time domain.
 
-\section{Conclusion}
-\label{sec:Conclusion}
+\begin{figure}
+	\centering
+		\includegraphics[scale=0.8]{images/dataerrorvsiteration.jpg.eps}
+	\caption{Error vs. iterations }
+	\label{fig:dataerrorvsiteration.jpg}
+\end{figure}
+\begin{figure}
+	\centering
+		\includegraphics[scale=0.8]{images/datapopulation.jpg.eps}
+	\caption{Swarm Population }
+	\label{fig:datapopulation.jpg}
+\end{figure}
 
-\section{ BenchMark Results }
-\label{sec:BenchMarkResults}
+\subsection{Segmentation Algorithms}
+\label{sec:SegmentationAlgorithms}
+The result of the segmentation algorithm can be viewed in the figs. \ref{fig:results1.JPG}. These result shows the originals stroke and the segmentation that the system generated for the stroke. 
+\begin{figure}
+	\centering
+		\includegraphics{images/results1.JPG.eps}
+	\caption{Results Segmentations}
+	\label{fig:results1.JPG}
+\end{figure}
+\begin{figure}
+	\centering
+		\includegraphics{images/results2.JPG.eps}
+	\caption{Results Segmentaion }
+	\label{fig:results2.JPG}
+\end{figure}
+\begin{figure}
+	\centering
+		\includegraphics{images/results3.JPG.eps}
+	\caption{Results }
+	\label{fig:results3.JPG}
+\end{figure}
+\begin{figure}
+	\centering
+		\includegraphics{images/results4.JPG.eps}
+	\caption{Resluts of the segmentations}
+	\label{fig:results4.JPG}
+\end{figure}
+
+The experiment we implemented was to test the effect of symbol complexity and type on the recognition rate.
+Figure \ref{fig:test2} shows what the accuracy of each symbol, it clearly noted that symbols that have only line segments achieve higher accuracy rate than other symbols.
+Figure \ref{fig:test2} also shows that algorithm 1 alone achieve better performance than algorithm 2 in the symbols that consist of lines only.
+The combining of both algorithm improves the recognition rate of other symbols.
+\\
+\begin{figure*}[]
+	\centering
+		\includegraphics[scale=0.5]{images/test2.eps}
+	\caption{The results of second experiment}
+	\label{fig:test2}
+\end{figure*}
+ Another experiment we performed was to test the efficiency svm classifier.
+We implemented the linear discriminator classifier described in \cite{gestureexample12}(see figure \ref{fig:test3}).
+%The result shows that SVM is more scalable than the linear discriminator for this problem.
+%The test focus on the size of data needed to test each classifier and the achieved accuracy.
+%For this experiment we also shows the classification time for both classifiers
+\begin{figure}[]
+	\centering
+		\includegraphics[scale=0.4]{images/test1.eps}
+	\caption{The results of third experiment}
+	\label{fig:test3}
+\end{figure}
+ %Details description of results got from comparing various algorithms  
+ %Describe data set used for experiments ( No of data set, No. of categories, No of samples).
+ %Describe other Algorithms [Alg 3 in paper \cite{earlyprocess}]used for comparing with the one described in the paper.
+
+ %In the figure \ref{fig:test1}, the accuracy means that final hit rate of all symbols used in the test set.by the  four datasets . 
+We performed three experiments to test the system firstly we tested recognition accuracy of shapes in the data set with both algorithms.
+We also implemented the segmentation algorithm described in \cite{earlyprocess} to use as reference to our swarm algorithms.
+As you see in the fig.
+\ref{fig:test1} shows the accuracy achieved by each algorithm.
+The results shows that both PSO algorithm achieve better result than other algorithms.
+The swarm algorithms were tested with and without the ellipse detection module.
+The ellipse detection module appear to be superior to results with only the PSO algorithms.
+\\
+\begin{figure}[]
+	\centering
+		\includegraphics[scale=0.3]{images/test1.eps}
+	\caption{The results of first experiment}
+	\label{fig:test1}
+\end{figure}
+
+%\subsection{Clustering Algorithm Results}
+%\label{sec:ClusteringAlgorithmResults}
+
+\subsection {Recognition Algorithms }
+\label{sec:recognitionAlgorithms}
+To test the recognition different set of features are used to select the features that will define symbols better to finally achieve better recognition rate.
+We tested the system using different types of features.
+The features explained in section \ref{sec:FeatureExtraction} was divided into different sets; statistical, structural, Rubine features and zernik moments along with composite features.
+The fig.
+\ref{fig:testFeatures} shows the result of each set of features when used with the SVM classifier.
+The results shows that the Rubine features although may achieve good results in single stroke symbols, have a bad performance when used with multi stroke symbols.
+\begin{figure}[]
+	\centering
+		\includegraphics[scale=0.3]{images/test4.eps}
+	\caption{The results of Features experiment}
+	\label{fig:testFeatures}
+\end{figure}
+
+We test the system using different classifier to see what the effect of the classification step on the final recognition.
+We also wanted to test the train and test time in both classifiers.
+We used SVM (RBF) classifiers and the Gaussian Linear classifier used in \cite{}.
+The fig.
+\ref{} shows the result of both classifiers with the main three segmentation algorithms.
+The results shows that the SVM classifier achieve better results.
+To measure the scalability of the system\footnote{What happens when adding a new symbols to the set of known symbols.} the classifiers are test using different number of symbols (fig.
+\ref{fig:sampleperCat}).
+Figures \ref{fig:sampleperCat} demonstrate the scalability of both classifiers by demonstrating the recognition rate in both classifiers with different number of symbols.
+Figures \ref{fig:sampleperCat},\ref{fig:sampleperCat} shows the time and number of symbols needed for training versus each classifier.
+
+\begin{figure}
+	\centering
+		\includegraphics{images/sampleperCat.eps}
+	\caption{Samples per Category }
+	\label{fig:sampleperCat}
+\end{figure}
+
+\begin{figure}
+	\centering
+		\includegraphics[scale=0.7]{images/test4.eps}
+	\caption{Tests}
+	\label{fig:test4}
+\end{figure}
+
+\section{Summary}
+\label{sec:ResultSummary}
+
+The result illustrated in this chapter shows the work done on the system and experiments performed.
+The results shows that the ellipse detection and PSO swarm algorithms gives better performance than other known systems.
+
+%\section{ BenchMark Results }
+%\label{sec:BenchMarkResults}

### 2009-02-18 draft/ConFut.tex
now i am reviewing

@@
-\chapter{Discusion and Conclusion}
+\chapter{Discussion and Conclusion}
 
-\section{  Conclusion}
+\section{Conclusion}
 \label{sec:ConclusionConclusion }
+
+This paper presented a new approach to sketch recognition using PSO.
+It was noted that the PSO in general improve the accuracy of the final symbol recognition in the system.
+The use of both speed and curvature data helps in improvement of the PSO algorithm over the original algorithms \cite{CruveDivisionSwarm,PolygonApproximationPSO}.
+The trade off between accuracy achieved and time complexity must be further investigated to achieve better results.
+The system introduced an efficient method to sketch recognition using the PSO.
+The test performed confirmed that PSO achieve better performance and optimization than other algorithms.
+The number and the complexity of symbols recognized by the system varied from simple to complex but did not affect the final recognition of the system.
+The only draw back of the system that it takes more time with comparison with similar algorithms.
+
+\newpage
 
@@
 \label{sec:FutureWork}
+
+%As you can see form the experiments PSO have proved a spuriously than other system. 
+%\section{Future Work}
+The next step in this research is to complete the clustering algorithm for fully automated sketch recognition.
+The clustering must be performed without the user explicit involvement.
+The currently used algorithm must be modified to accept error reporting from users.
+The segmentation algorithm may be more powerful if it can segments the stroke into more types of primitives other than the line and curve.
+Other area of modification can be the features extraction and classifier.
+Adding more spatial features and structure features will improve classifications.
+After segmenting the stroke the system can generate a structural graph from segments and there spatial relations.
+This graph will be matched with the set of known symbol template to make the identification.
+Those result along with the SVM classifier results will enhance the system as a whole and achieve better recognition rate.
+%Each new stroke is checked if it can be a symbol or is a part of already drawn un completed symbol.

### 2009-02-18 draft/System.tex
now i am reviewing

@@
 \label{sec:AimsOfTheSystem}
-The previous chapter shows the current trends in sketch recognition. This thesis introduce a solution for the problem that give optimize solution for the segmentation problem.
+The previous chapter shows the current trends in sketch recognition.
+This thesis introduce a solution for the problem that give optimize solution for the segmentation problem.
+Particle swarm optimization is used to correctly segment strokes into curves and lines.
+The current implemented system either segment strokes based on simple curvature data or using some more complex algorithm.
+The performance of most algorithms are based on user style.
+
+% Stroke segmentation is one of the most complex problems in the sketch recognition system. It is dividing the   
+%In stroke segmentation problem is  Particle swarm optimization is used as the
 
@@
 \label{sec:AnOverviewOfTheSystem}
+   This research solve sketch segmentation optimization problem using particle swarm algorithm PSO.
+Due to sloppiness of users and hardware glitches the system process the data before segmentation to generate a set of possible corner or critical points.
+The clustering algorithm then tries to group the segments together to form a single symbol.
+A set of features is extracted from the symbols to be used by the classifier.
+%The stroke is the path of points from the instant the pen is down till it is up.
+The block digram in fig.
+\ref{fig:Blockdiagram} shows that the main system blocks are 1) preprocessing, 2) segmentation, 3) clustering, 4) feature extraction, and finally 5) classification.
+The next section will describe each component in details.
+ %Due to sloppiness of users and hardware glitches the system processes the data before further
+\begin{figure}[]
+	\centering
+		\includegraphics[scale=0.6]{images/Blockdiagram.eps}
+	\caption[  ]{The system Block Diagram}
+	\label{fig:Blockdiagram}
+\end{figure}
 
@@
 \label{sec:SystemComponents}
-% the system and discription of this
 
-\chapter{Preprocessing and Segmentation}
-\label{sec:Preprocessingandsegmentation}
+The \textbf{preprocessing} is responsible of capturing the input data and removing the noise from it.
+The preprocessing block as in fig.
+\ref{fig:Blockdiagram}   is consistent of  some calculation based on    the time differences, direction, speed and curvature of all points of the stroke.
+After computing these data is the local minimum and local maximum of each information is extracted from the data computed.
+The remaining of the geometrical and statistical data is then evaluated from the stroke.
+The system then proceeds to estimate a set of possible dominant points to help in the segmentation process.
+\\
+
+The  next step is \textbf{segmentation}, the goal of the segmentation stage is to divide stroke into segments of either curves or lines.
+The segmentation process is divided into two steps ellipse fitting and curve segmentation.
+In the first step, the ellipse fitting process tries to fit the stroke into an ellipse.
+If the system fails it passes the stroke to the second step; curve segmentation which consist of two PSO segmentation algorithms.
+The two algorithms will generate two segmentations, the system will choose the segmentation that has the minimum segmentation error.%  generate the segmentations,  the minimum error will be the chosen segmentation.
+% is the ellipse fitting
+
+ %the stroke first If the ellipse detection fails the stroke is passed to the segmentation algorithms which will pass it to the two PSO algorithms described below the segmentation with the minimum error will be the chosen segmentation.
+The segmentation is then added to the set of un-recognized segments in the system.
+\\% This part is repeated for each stroke.
+ The clustering algorithms starts to group segments together after the segmentation step.
+The system let the user draws the symbol by using any number of strokes, a set of unrecognized segments is passed to the clustering algorithm to generate a symbol and compute a feature vector for it.
+
+ Feature Extraction is the fourth block in the system.
+he system compute composite set of features some are statistical other are spatial features based on the type primitives.%The system extract segment based and statistical based features from the set of segments that the clustering algorithms produce.
+
+   The final step is a SVM classifier that will use the features computed to classify the segments into one of the previously trained classes.
+%  Or to determine the symbol of the given segments from the set of preciously trained symbols.
+
+ %with a symbol from the training set.% The system compute composite set of features some are statistical other are spatial features based on the type primitives.
+Finally, the strokes are classified into the corresponding classes.% the classifier identifies the strokes into a symbol from the set of known symbols.\\
+
+\chapter{System Details}
+\label{sec:SystemDetails}
 
@@
 \label{sec:Preprocessing}
+%Some data is extracted from the points in the stroke.
+ It is noted that as the user draws a shape the pen will slow down near corners and picks up speed when drawing a straight lines.
+Therefore, the speed and time difference data were widely used in sketch understanding systems \cite{earlyprocess}.
+The direction and curvature data are used to determine the points with high angular changes along the path of points \cite{meanshift10}.
+ It was noticed that when the users draw  corners the pen slows down and picks up speed when users draw straight lines.
+This observation helps detecting corners points as the points with lower speeds and high curvatures.
+The system compute speed, curvature, time difference and direction data then generate a set of possible corners that will be used in the segmentation as an initial solution.
 
@@
 
+ In the presented system, we computed time difference, direction, speed and curvature of each point along the stroke.
+The following equation $v=\Delta t/\Delta s$ is used for speed calculation  where $t$ is the time difference between two points and $s$ is the length between them.
+The time difference is calculated as $\Delta t = t_{i+1} - t_i$ where $t_i$ is the time of the point $i$ and $t_{i+1}$ is the time of the point $i+1$.
+
+\begin{equation}
+\label{eq:direction}
+	\vartheta  = \cos ^{ - 1} \left( {\frac{{\overrightarrow {p_1 p_2 }  \cdot \overrightarrow {p_2 p_3 } }}{{\left\| {\overrightarrow {p_1 p_2 } } \right\| \times \left\| {\overrightarrow {p_2 p_3 } } \right\|}}} \right)
+\end{equation}
+
+  The direction is calculated as the angle between two vectors as in equation \ref{eq:direction} where $\overrightarrow {p_i p_j }$ is the vector from point $pi$ to point $pj$ and $\|{\overrightarrow {p_i p_j }}\|$ is the normal of the  vector $\overrightarrow{p_i p_j }$  .
+The curvature estimated using more than one method so in this research the result of each curvatures estimation is reported in \ref{sec:Curvatureestimationresults} and the final curvatures definition used that define curvatures as the change in direction with respect to length i.e.
+$c= \Delta d/\Delta s$.
+
+ %system compute speed, distance data for all points in the stroke. Curvature is computed using estimation used in [] where direction computed is angle between two lines. Curvature is 
+ %$\Delta d/  \Delta S $ where d is the difference direction of point and s is difference in distance between points. 
+ All these calculations are performed in real time while the user draws the strokes.
+The complexity of computation is $O(n)$ where $n$ is number of points.
+Figure \ref{fig:speed2Distance} shows the computed data for the stroke drawn in fig.
+\ref{fig:orignalStroke}.
+If you look at the data curves you will see that dominant points are characterized by lower speed values and higher curvature and direction values.
+The systm also compute the average, minimum and maximum values of speed and curvature curves.
+\\% the lower speed points correspond to vertex and dominant point's location.
+The higher the direction and curvature data correspond to location with higher change in curvature witch promote the location for vertices's.
+\\
+%Description of input data. 
+%how to calculate speed, curvature, area, bounding box.
+%how to remove noise. 
+%Finally how to compute primarily dominant points.
+
+\begin{figure}
+	\centering
+		\includegraphics[scale=0.6]{images/speed2.eps}
+	\caption{Speed Vs. Distance Graph}
+	\label{fig:speed2Distance}
+\end{figure}
+
+\begin{figure}[]
+	\centering
+		\includegraphics[scale=0.8]{images/orignalStroke.eps}
+	\caption{Stroke Sample}
+	\label{fig:orignalStroke}
+\end{figure}
+
 \subsection{Critical Point Detection}
 \label{sec:CriticalPointDetection}
+After the system computes all the speed, time difference and curvature data it proceeds to detect the points with low velocity and high curvatures.
+Using simple differentiations to detect local extreme points resulted in false points due to the non smooth curves.
+So instead the following procedure was used, firstly the mean and maximum values of the curve is computed from the actual data.
+
+The mean is used as threshold to divide curves values into low and high values.
+This threshold is used to separate the curve into regions; each region is defined as the part between two intersection points of the threshold line and the curve.
+Thus, their will be regions with value higher than the threshold   and regions with values lower than the threshold.
+Only regions higher than the threshold value are examined when computing the maximum of each region.
+The point that corresponds to those maximum values are labeled possible dominant points.
+
+   %After the threshold is applied the curve is divided into regions.
+Each region is defined by a part in the data curve between tow intersection point of the threshold line and the curve.
+% to find the extreme points.
+Hence, the mean of the direction curve is calculated and used as threshold.
+\\
+   %If we tried to differentiate the curve the result will be false threshold values we divide the data into regions of data higher or lower that the threshold.
+This will let us only look for high data only.
+The maximum of each region is then computed and reported as a possible vertex point.
+
+The system repeat this process to curvature, time difference and speed curves (see fig.
+\ref{fig:speed2}).
+For each curve the points labeled as possible dominant points are saved into a single array .
+The points that are repeatedly labeled are given a higher score than the other points.
+This score is then used to sort the possible points array.
+The array that contains possible points is used as input for the next stage of segmentation.
+\begin{figure}[]
+%\begin{minipage}[b]{0.8\linewidth}
+	\centering
+
+	%	\subfigure[ The Speed of data] {\includegraphics[scale=0.35]{images/speed2.eps}}
+
+		\subfigure[ Direction ] {\includegraphics[scale=0.35]{images/direction2.eps}}
+					%\end{minipage}
+					%\begin{minipage}[b]{0.8\linewidth}
+			\subfigure[Curvature] {\includegraphics[scale=0.35]{images/curvature2.eps}}
+			\subfigure[ Time Difference] {\includegraphics[scale=0.35]{images/time2.eps}}
+			%\end{minipage}
+
+	\caption[Data Curves]%{The data of the stroke}
+
+	\label{fig:speed2}
+\end{figure}
 
@@
 \label{sec:Segmentation}
-\subsection{Swarm Introduction}
-\label{sec:SwarmIntroduction}
+The segmentation process tries to divide the stroke into a set of primitives.
+The segmentation algorithm first tries to detect if the stroke is an ellipse.
+If the stroke proved to be an ellipse then the segmentation process ends and the system proceeds to the next step.
+Otherwise, the stroke is passed into two particle swarm algorithms that will divide the stroke to either lines or lines and curves (see the block diagram in fig.
+\ref{fig:Blockdiagram} ).
+The algorithms takes the stroke points along with the possible dominant points computed then produce a set of dominant points which are connected with either lines or curves (see fig.
+\ref{fig:Blockdiagram}).
+The next sections will describe the ellipse detection algorithm and both the two particle swarm algorithms used to divide the stroke.
+%After computing the primary data the system now tries  \\
+%Paragraphs describe the segmentation algorithm.
 
@@
 \label{sec:EllipseDetection}
+%Describe the ellipse detection algorithms 
+This process will try to fit the stroke point into an ellipse, it starts with computing the center of the stroke bounding box.
+The bounding box center point is used as the first estimation of the center of the ellipse.
+The axes of the ellipse are estimated as the $w/2$ and $h/2$ of the stroke bounding box where $w$ is width and $h$ is hight of the bonding box.
+
+\begin{equation}
+E = \sum\limits_{i = 0}^N {\frac{{(x_i - x_0 )}}{{a^2 }}^2  + \frac{{(y_i - y_0 )}}{{b^2 }}^2  - 1} 
+\label{eq:circleFit}
+\end{equation}
+
+ The least square fitting algorithm is used to minimize the fitting error of the ellipse equation (see \ref{eq:circleFit} where N is number of points in the stroke, $a,b$ are the length of ellipse axes, $x_0$ \& $y_0$ are the coordinates of the center point, $x_i$ \& $y_i$ are the coordinates of point $i$ in the stroke).
+A list of new values for $x_0$ , $y_0$ ,$a$ and $b$ are generated randomly from the older values with small increment and the minimum error are evalated from this list.
+for After only few loops the error and a confidence values  are computed.
+ \begin{equation}
+E = E_{ellipse}+E_{area}
+\label{eq:circleError}
+\end{equation}
+ \begin{equation}
+E_{area}  = \sqrt {\left( Area(stroke) - Area(ellipse)\right)^2 }
+\label{eq:ErrorArea}
+\end{equation}
+
+   The confidence value computed by eq.
+( \ref{eq:circleError}) where $E_{ellipse}$ is the error computed by eq.
+(\ref{eq:circleFit})  and $E_{area}$ is computed as the difference between the actual area of the stroke and the area of the estimated ellipse ( see eq.\ref{eq:ErrorArea}).
+The confidence value is used to label the stroke as ellipse or un segmented stroke.
+If the stroke as unsegmented the next process will be to pass the stroke into curve segmentation algorithm.
+
+   The next section will describe firstly the particle swarm algorithm (PSO) then it will proceed to detail the remaining of the segmentation algorithms.
+
+   %to check if the stroke can be labled as ellipse or not.\\ %is computed to check if stroke is an ellipse. \\
+%We found that if the confidante is above threshold then the probability of ellipse is highest otherwise the stroke is passed to the next section to get the divisions of stroke and test its error.
+
+\subsection{Particle Swarm Algorithm Introduction}
+\label{sec:ParticleSwarmAlgorithm}
+%\section{Particle Swarm Algorithm}
+%\label{PSO}
+%What is particle swarm algorithm and how it was used in related researches. 
+The main idea of  Particle Swarm Algorithm (PSO) is to represent each agent with a particle from the solution space.
+Each agent moves the particle with a direction and velocity based on the following equation [\ref{eq:Swarm}].
+\begin{equation}
+%\[
+v_{ij}  = v_{ij}  + c_1 r_1 (lbest_{ij}  - p_{ij} ) + c_2 r_2 (gbest_{ij}  - p_{ij} )
+%\]
+\label{eq:Swarm}
+\end{equation}
+\begin{equation}
+%\[
+p_{ij}=p_{ij}+v_{ij},
+%\]
+\end{equation}
+ Equation [\ref{eq:Swarm}] shows how velocity and direction of each particle is computed where $r_1$ \& $r_2$ are random variable and $c_1$ \& $c_2$ are the swarm system variables.
+Each iteration the global best $g_{best}$ particle and  the agent local best $l_{best}$ particle are evaluated based on the maximum fitness functions of  all particles.
+The solution is found after a maximum number of iteration or after an error threshold is achieved.
+ \begin{equation}
+   P(i)\Leftarrow 
+\{
+\begin{array}{c} 
+1 \quad \quad if\quad r_{3}>t  \\
+
+0 \quad \quad if\quad r_{3}<t 
+\label{eq:descrite}
+\end{array}\}
+\end{equation}
+  Equation \ref{eq:descrite} where t is a threshold and $r_{3}$  is a random variable, is used to change general swarm algorithm into binary particle which will handle particle values of either 0 or 1.
 
@@
 \label{sec:SwarmSegmentation}
-\subsubsection{Problem Formulation }
+
+In this process the system generate two stroke segmentation using two PSO segmentation algorithm.
+The system generates segmentation from both algorithms and then chooses the segmentation with the minimum error value.
+The problem definition is the same in both algorithms but they differ in the fitness function and the error functions.
+%formation is nearly the same in both.
+\subsubsection{Problem Formulation}
 \label{sec:ProblemFormulation}
+The input stroke with $N$ points can be represented by set $S = \left\{ {x_1 ,x_2  \ldots x_N } \right\}$ where $x_i$ is the location of the point $i$ .
+The swarm algorithms consist of $M$ agents which are represented by the set
+$A  = \left\{ {P_i \left| {i = 1,2 \cdots M} \right.} \right\}$ where $P_i$ is a single solution particle from the solution space.
+Each particle decodes the problem with binary array with the same length $N$ as the input stroke .
+\begin{figure}
+	\centering
+		\includegraphics{images/CodingSwarm.eps}
+	\caption{An Example Stroke and the Coding}
+	\label{fig:CodingSwarm}
+\end{figure}
 
-\subsubsection{Fitness Computation }
-\label{sec:FitnessComputation}
+% The particles of the swarm represent a single solution of the solution space.
+For this problem the, each particle %will give a different segmentation for the input stroke.
+Firstly, we will define the stroke with N points by the set %S where.
+We define the arc %An array with the same length as the number of points of the strokes.
+So the system will represent each particle $P_i$ by $P_i = \left\{ {p_{ij} \left| {j = 1,2 \cdots N} \right.} \right\}$ where $p_{ij}$ will have only two values 0 or 1 where $p_{ij}=1$ will mean that point $j$ is a dominant point(see fig.
+\ref{fig:CodingSwarm} ).
+Thus , the particle represent which points are chosen from the set of points in the stroke to be used as dominate points for segmentation.
+The goal of the PSO algorithm is to find the solution $P_i$ that generate the minimum set of dominate points that will define the stroke with minimum segmentation error.
+In other words, we want to find the fewer number of points (1's) location in particle array which will gives minimum error.
+
+The fitness function and error calculation are different in each algorithm. The first fitness and error function is described below.
+
 \subsubsection{Polygon Division Algorithm}
 \label{sec:PolygonDivisionAlgorithm}
+For the first algorithm, the algorithm tries to segment the stroke as a polygon approximation.
+The final solution will be set of line segment that best define the input stroke.
+Given the inputs stroke let's define the set of points in the stroke as  $S = \left\{ {x_1 ,x_2  \ldots x_N } \right\}$ which is the set of consecutive points from start of stroke till the end.
+Then define the arc $\widehat{x_ix_j}$ as the consecutive set of points from point $x_i,x_{i+1} \cdots,x_j$.
+The line
+$\overline{x_i x_j} $ as the straight line connecting point $x_i$ to point $x_j$.
+The approximation error is computed by the equation \ref{eq:ErrorSwarm1}  where M is the number of dominant points in this solution.
+The error $ e ( \widehat{x_ix_j},\overline{x_i x_j})$ is computed as the sum of squared perpendicular distance from every point along the arc $\widehat{x_ix_j}$ to the line $\overline{x_i x_j}$.
+\\
+\begin{equation}
+E=\sum\nolimits_{i = 0}^M e ( \widehat{x_ix_{i+1}},\overline{x_i x_{i+1}})
+\label{eq:ErrorSwarm1}
+\end{equation}
+\begin{equation}
+\max fitness(p_i ) = \left\{ {\begin{array}{*{20}c}
+   { - E/\varepsilon N} & {ifE > \varepsilon ,}  \\
+   {D/\sum\limits_{j = 1}^N {p_{ij} } } & {otherwise}  \\
+\end{array}} \right.
+\label{eq:fitnessSwarm1}
+\end{equation}%\]
 
+The fitness is computed using equation \ref{eq:fitnessSwarm1} where N is the number of points in the stroke, D is the number of point in the solution that was previously labeled as a possible dominant point and E is the computed error and $\varepsilon$ is the error threshold.
+If the solution produced an error that exceed the error threshold the fitness function assign a -ve value for the solution to express it's in feasibility, otherwise the inverse of the number of vertex produced is used to access the solution fitness in terms of minimum number of vertices's.
+This fitness function optimize two goal the first is to move the solution into the feasible solution space with accebtable error bounds and second is to fly the particle to a new position which may result in a polygon with a fewer number of vertices's.
+
+ %The error is chosen to in favor of larger than the threshold is given a -ve value to lower the value of solution otherwise the system will favor the lower number of vertices's.\\ %we will want to lower the number of vertices's.
+\\
+%if we say that s
+%Alg1:
+%Alg2: 
 \subsubsection{Curve Divisions Algorithm}
@@
 
+The second algorithm has the same problem formulation but different fitness and error functions.
+It was previously introduced in  \cite{CruveDivisionSwarm} where but genetic programming was used as the optimizing algorithm.
+The particle is represented using the arrays $P_i = \left\{ {p_{ij} \left| {j = 1,2 \cdots N} \right.} \right\}$ where $p_{ij}$ will have only two values 0 or 1 as mentioned before is section \ref{sec:PolygonDivisionAlgorithm}.
+Lets denote the segments $\widehat{x_ix_j}$ as the consecutive set of points from point $x_i,x_{i+1} \cdots,x_j$.
+The goal of the algorithm is to fit the segments between the solution vertices's into straight lines or a circular arc.
+\paragraph{Fitting segments into straight line.}
+\label{sec:FittingSegmentIntoStraightLine}
+According to \cite{CruveDivisionSwarm} each segment $\widehat{x_ix_j}$ given a set of points $x_i,x_{i+1} \cdots,x_j$ can be fitted into the line: $y=kx+c$ where $k$ and $c$ are the slope and the intercept of the line respectively.
+Equation \ref{eq:Linek} and \ref{} are used to fit the segment $\widehat{x_ix_j}$ into the straight line where N is the number of points in the segment and $(x_i, y_i)$ are the coordinate of the point $i$.
+
+\begin{equation}
+\label{eq:Linek}
+k = \frac{{N\sum\limits_{i = 1}^N {x_i y_i }  - \sum\limits_{i = 1}^N {x_i } \sum\limits_{i = 1}^N {y_i } }}{{N\sum\limits_{i = 1}^N {x_i^2 }  - \left( {\sum\limits_{i = 1}^N {x_i } } \right)^2 }}
+\end{equation}
+\begin{equation}
+\label{eq:LineC}
+c = \frac{{\sum\limits_{i = 1}^N {x_i^2 } \sum\limits_{i = 1}^N {y_i }  - \sum\limits_{i = 1}^N {x_i } \sum\limits_{i = 1}^N {x_i y_i } }}{{N\sum\limits_{i = 1}^N {x_i^2 }  - \left( {\sum\limits_{i = 1}^N {x_i } } \right)^2 }}
+\end{equation}
+\begin{equation}
+\label{eq:ds}
+ d_s  = \frac{{\sum\limits_{i = 1}^N {\left| {(kx_i  + c) - y_i } \right|} }}{{N\sqrt {k^2  + 1} }}
+\end{equation}
+
+The distance($d_s$) (see equation \ref{eq:ds}) is the average distance from segment points $(x_i, y_i)$ to the  line estimated.
+This distance ($d_s$) is used to estimate the error of the line approximation which we will dnote as ($E_s$).
+
+\paragraph{Fitting segments into circular arc.}
+\label{sec:FittingSegmentcirculararc}
+The segment points can also be fitted into the circular arc $(x - a)^2  + (y - b)^2  = R^2, $ where ($a,b$) are the coordinate of the center of the circle and $R$ is the radius of the circle.
+They can be estimated using the following set of equations:
+
+\begin{equation}
+a = \frac{{b_{1}a_{22} - b_{2}a_{12}}}{\Delta},
+\end{equation}
+\begin{equation}
+	b = \frac{{b_{2}a_{11} - b_{1}a_{21}}}{\Delta},
+\end{equation}
+\begin{equation}
+R = \sqrt {\frac{1}{N}(\sum\limits_{i = 1}^N {x_i^2 }  - 2\sum\limits_{i = 1}^N {x_i a}  + Na^2  + \sum\limits_{i = 1}^N {y_i^2  - 2} \sum\limits_{i = 1}^N {y_i b + Nb^2 } )} ,
+\end{equation}
+
+Where 
+\begin{equation}
+a_{11}  = 2\left[ {\left( {\sum\limits_{i = 1}^N {x_i } } \right) - N\sum\limits_{i = 1}^N {x_i^2 } } \right]
+\end{equation}
+\begin{equation}
+a_{12}  = a_{21}  = 2\left( {\sum\limits_{i = 1}^N {x_i \sum\limits_{i = 1}^N {y_i } }  - N\sum\limits_{i = 1}^N {x_i y_i^2 } } \right),
+\end{equation}
+\begin{equation}
+ a_{22}  = 2\left[ {\left( {\sum\limits_{i = 0}^N {y_i } } \right)^2  - N\sum\limits_{i = 0}^N {y_i^2 } } \right]
+\end{equation}
+\begin{equation}
+b_1  = \sum\limits_{i = 0}^N {x_i^2 } \sum\limits_{i = 0}^N {x_i }  - N\sum\limits_{i = 0}^N {x_i^3 }  + \sum\limits_{i = 0}^N {x_i } \sum\limits_{i = 0}^N {y_i^2 }  - N\sum\limits_{i = 0}^N {x_i y_i^2 } ,
+\end{equation}
+\begin{equation}
+b_2  = \sum\limits_{i = 0}^N {x_i^2 } \sum\limits_{i = 0}^N {y_i }  - N\sum\limits_{i = 0}^N {y_i^3 }  + \sum\limits_{i = 0}^N {y_i } \sum\limits_{i = 0}^N {y_i^2 }  - N\sum\limits_{i = 0}^N {x_i^2 y_i } ,
+\end{equation}
+
+\begin{equation}
+\Delta  = a_{11}b_{22}-a_{12}a_{21}.
+\end{equation}
+ The error of the circle estimated is measured using the distance ($d_c$) which is calculated using the following equation:
+ \begin{equation}
+ \label{eq:circleE}
+d_c  = \frac{{\sum\limits_{i = 0}^N {\left| {\sqrt {(x_i  - a)^2  + (y_i  - b)^2 }  - R} \right|} }}{N}
+\end{equation}
+Thus, eq.(\ref{eq:circleE}) ($d_c$) define the average distance ($d_c$) from the segment points ($x_i,y_i$) to the estimated fitted circle ($a,b$) and $R$.
+
+\paragraph{Determinig The Segment Type.} 
+\label{sec:DeterminigSegmentType}
+If the average distance ($d_s$) equalls zero that means that the segments point is exactly the straight line approximated.
+If the distance ($d_c$) equals zero means that the segments point lies exactly on the circular arc estimated.
+But the user rarely draws an exact line or circle, so the distances are used as the error estimations.
+The distance will determine the type of the current segment, in other words if ($d_s$) is smaller than ($d_c$) then the segment is labeled as a line segment otherwise it is labeled as a circular arc.
+\\
+
+After each segment type is determined the the particle segmentation error is computed by the equation equation \ref{eq:errorSwarm2} where M is the number of segments in the solution, $D_i$ is the  approximation error of each segment as $min(d_c,d_s)$ as computed in eq.(\ref{eq:circleE})  and eq.
+(\ref{eq:ds})  \cite{CruveDivisionSwarm}.
+The fitness is computed by the equation \ref{eq:fitnessSwarm2} where E is the error and M is number of segments and k is a parameter tweaked to get minimum number of segments.
+
+ %and line estimation are computed for each segment, the approximation with the lower error value will be the chosen approximation of this segment.
+The sum of the approximation error of all segments is the total error of the particle.
+The error is computed by   \\
+
+\begin{equation}
+E=\sum\nolimits_{i = 0}^M e(D_i) 
+\label{eq:errorSwarm2}
+\end{equation}
+\begin{equation}
+\max fitness(P_i ) = \frac{1}{{E \times M^k }}
+\label{eq:fitnessSwarm2}
+\end{equation}
+
+%Describe how we used particle swarm in curve polygon approximation using first algorithm.  
+%Describe how we used PSO in to divide stroke into segments and curves. 
+%[These two algorithm was modified from the algorithm in paper \cite{PolygonApproximationPSO,CruveDivisionSwarm}].
+\cite{PolygonApproximationPSO} used a merge and divide algorithm after each loop of the swarm system to refine the solution but we used another enhancement method.
+After each loop in the swarm algorithm, each particle loops on the set of selected dominant points to enhance the solution.
+Each dominant point is checked to find if it was labeled before as a possible dominant (computed as in section \ref{Prepross}), if not the point is moved to the nearest labeled point.
+%check the new solution and check the new vertices's if a vertex points is a possible dominant point it is left as a vertex.
+Otherwise, the system will search in the local neighborhood for the nearest possible dominant point to move these vertices's to.
+After each change the error is computed and if new error is less that the old then the new set is approved.
+\\
 \section{BenchMarck algorithm}
 \label{sec:BenchMarckAlgorithm}
+In this research, a bench mark algorith is used to compare to results obtained by the PSO algorithms.
+The algorithm was first implemented by  \citeauthor{earlyprocess} in \cite{earlyprocess}.
+The \citeauthor{earlyprocess} algorithms is divided into the following steps.
+\begin{enumerate}
+	\item compute the curvatures and the speed of the stroke points. 
+	\item Threshold the curvature and speed curves using a percentage of the maximum value in the curve. 
+	\item Mark the values higher than the threshold in an new arrays $L_c$ for list of corner points generated from the curvature curve. The list $L_s$ contains the points generated from the speed curve.
+	\item Compute a confidence on each list based on the type of curve. Then sort the lists $L_s$and $L_c$ by the confidence value. 
+	\item Find the intersection points in both list and generate the a Sample solution $h_{curr}$. 
+	\item Use $h_{curr}$ as initial solution. Loop doing the following:
 
-\chapter{Symbol Recognition}
-\label{sec:SymbolRecognition}
+		\begin{enumerate}
+			\item Generate a new solution $h_{s}$ using $h_{curr}$ and  the first point from the $L_s$ list .
+			\item Generate a new solution $h_{c}$ using $h_{curr}$ and  the new point from the $L_c$ list.
+			\item Test the solutions $h_{s},h_c$ to find the minium solution.
+			\item  If $h_s$ then add the solution a list of hybrid solutions $Solutions_{set}$ and remove the first point from the list $L_s$ then set $h_s$ as $h_{curr}$. 
+			\item If $h_c$ then add the solution a list of hybrid solutions $Solutions_{set}$ and remove the first point from the list $L_c$ then set $h_c$ as $h_{curr}$. 
+		\end{enumerate}
+		\item Choose the final solution from the list of solutions $Solutions_{set}$ as a trade off between  minimum error and number of corners.  
+		\item Check each segment in the solution if the solution is can be estimated as a line segment if not it is fitted using a higher level Bezire curve. 
+\end{enumerate}
+More details about the algorithm is in \cite{earlyprocess}.
+
 \section{Stroke Clustring Algoirthm }
 \label{sec:ClustringAlgoirthm}
+
+When the user finish drawing a stroke the segmentation algorithm generate a list of segments that represent this stroke.
+The clustering algorithms group the segments generated from the segmentation algorithm.
+These segments are grouped into a symbol and tested if they can be classified into one of the known symbols.
+If no identification is achieved the segments are added into a list of unclassified segments.
+When a user draws another stroke the segmented stroke is added to the list and the process is repeated until all the segments are classified to a known symbol.\\
+
+After the user draw all strokes of the symbol he has to wait 10 seconds or press finish button beside the drawing area.
+The set of unrecognized strokes is grouped together along with their segmentation as input to the feature extraction process.
+
 \section{ Feature Extraction}%%Feature Set
 \label{sec:FeatureExtraction}
+
+The system uses a composite set of features include Rubine feature set \cite{gestureexample12}, Zernike moments \cite{HeloiseBeautification}, ink density \cite{GeometryAndDomain102} and some structural information like number of perpendiculars lines , number of parallel lines and types of primitives in each symbol.
+The table \ref{tab:FeatureTable} gives details on all the features used in the system.
+ %trained with the 13 categories in the system.
+The same training set was used by \cite{HeloiseBeautification} in their symbol recognition system.
+\\ %  The table describe the features used in the system.
+%\begin{table}
+%\begin{minipage}[t][1\totalheight]{0.3\columnwidth}
+%
+%\begin{tabular}{|l|l|l|l|}
+%\hline 
+%Type & index & Features & Notes\tabularnewline
+%\hline 
+%\hline Structural & F1. & No. of parallel lines & \tabularnewline
+%\hline & F2. &  No. of perpendicular lines & \tabularnewline
+%\hline & F3. &  No. of intersections between any two segments. & \tabularnewline
+%
+%\hline Statistical& F4.and F5. & Mean and standard deviation of the Curvature data   & \tabularnewline
+%\hline & Mean of the speed data & \tabularnewline
+%\hline 
+%Rubine & Set of features \\for gesture recognition & Introduced by Rubine in \cite{} \tabularnewline
+%\hline Zernike &  A Zernike\\ moments  &  Used by \cite{zernike61} \tabularnewline
+%\hline Composite & width to\\ height ratio  & \tabularnewline
+%\hline  & ink density  & Used by \cite{} \tabularnewline
+%\hline
+%\end{tabular}		
+%\end{minipage}
+%	\caption{Feature Table}
+%	\label{tab:FeatureTable}
+%\end{table}
+
+%\begin{table}
+%\begin{tabular*}{0.5\textwidth}{|l|l|l|l|}
+%\begin{tabular}{|l|l|p{6cm}|p{5cm}|}
+%\hline
+%Type & Index & Features & Notes \\ 
+%\hline
+%Structural & \multicolumn{1}{c|}{} &  &  \\ 
+%\hline
+% & F1. & No. of parallel lines &  \\ 
+%\cline{2-4}
+% & F2. & No. of perpendicular lines &  \\ 
+%\cline{2-4}
+% & F3. & No. of intersection between of any two segmetns  &  \\ 
+%\cline{2-4}
+% & F4.  & No of primitives  &  \\ 
+%\cline{2-4}
+% & F5.  & No. of line segments &  \\ 
+%\cline{2-4}
+% & F6 & No. of ellipse, circular arcs.  &  \\ 
+%\hline
+%Statistical &  &  &  \\ 
+%\hline
+% & F7.  &  &  \\ 
+%\cline{2-4}
+% & F8. &  &  \\ 
+%\hline
+%Rubine Features &  & From features F9. till F13  & Introduced by in \cite{gestureexample12} \\ 
+%\hline
+% & F9. & Cosine of starting angle &  \\ 
+%\cline{2-4}
+% & F10. & Sine of starting angle &  \\ 
+%\cline{2-4}
+% & F11.  & Length of diagonal of bounding box 
+% & (gives an idea of the size of the bounding box) \\ 
+%\cline{2-4}
+% & F12. & Angle of diagonal
+%  & gives an idea of the shape of the bounding box (long, tall, square)
+%  \\ 
+%\cline{2-4}
+% & F13. & Distance from start to end
+% &  \\ 
+%\cline{2-4}
+% & F14. & Cosine of ending angle
+% &  \\ 
+%\cline{2-4}
+% & F15. & Sine of ending angle  &  \\ 
+%\cline{2-4}
+% & F16. & Total stroke length
+% &  \\ 
+%\cline{2-4}
+% & F & Change in Rotation( Arctan )& gives the directional angle \\ 
+%\cline{2-4}
+% & F & Absolute rotation 
+% &  \\ 
+%\cline{2-4}
+% & F & Rotation squared 
+% &  \\ 
+%\cline{2-4}
+% & F & The maximum speed reached (squared)
+% &  \\ 
+%\cline{2-4}
+% & F & Total time of stroke
+%  &  \\ 
+%\hline
+%Zernike moments &  &  &  \\ 
+%\hline
+% & F & Zernike moments to the n order &  used by \cite{zernike61} \\ 
+%Composite  &  &  &  \\ 
+%\hline
+% &  &  &  \\ 
+%\cline{2-4}
+% &  &  &  \\ 
+%\cline{2-4}
+% &  &  &  \\ 
+%\hline
+%\end{tabular}
+
+%\newpage
+
+%\begin{table}
+\begin{longtable}{|l|l|p{6cm}|p{5cm}|}
+	%\caption{Feature Table}
+	%\label{tab:FeatureTable}
+\hline 
+Type  & Index  & Features  & Notes \tabularnewline
+\endhead
+\hline 
+Type  & Index  & Features  & Notes \tabularnewline
+\endfirsthead
+\hline 
+Structural  & \multicolumn{1}{c||}{} &  & \tabularnewline
+\hline 
+ & F1.  & No. of parallel lines  & \tabularnewline
+\hline 
+ & F2.  & No. of perpendicular lines  & \tabularnewline
+\hline 
+ & F3.  & No. of intersection between of any two segmetns  & \tabularnewline
+\hline 
+ & F4.  & No of primitives  & \tabularnewline
+\hline 
+ & F5.  & No. of line segments  & \tabularnewline
+\hline 
+ & F6  & No. of ellipse, circular arcs.  & \tabularnewline
+\hline 
+Statistical  &  &  & \tabularnewline
+\hline 
+ & F7.  &  & \tabularnewline
+\hline 
+ & F8.  &  & \tabularnewline
+\hline 
+Rubine Features  &  & From features F9. till F13  & Introduced by in \cite{gestureexample12} \tabularnewline
+\hline 
+ & F9.  & Cosine of starting angle  & \tabularnewline
+\hline 
+ & F10.  & Sine of starting angle  & \tabularnewline
+\hline 
+ & F11.  & Length of diagonal of bounding box  & (gives an idea of the size of the bounding box) \tabularnewline
+\hline 
+ & F12.  & Angle of diagonal  & gives an idea of the shape of the bounding box (long, tall, square) \tabularnewline
+\hline 
+ & F13.  & Distance from start to end  & \tabularnewline
+\hline 
+ & F14.  & Cosine of ending angle  & \tabularnewline
+\hline 
+ & F15.  & Sine of ending angle  & \tabularnewline
+\hline 
+ & F16.  & Total stroke length  & \tabularnewline
+\hline 
+ & F  & Change in Rotation( Arctan ) & gives the directional angle \tabularnewline
+\hline 
+ & F  & Absolute rotation  & \tabularnewline
+\hline 
+ & F  & Rotation squared  & \tabularnewline
+\hline 
+ & F  & The maximum speed reached (squared)  & \tabularnewline
+\hline 
+ & F  & Total time of stroke  & \tabularnewline
+%\newpage
+\hline 
+Zernike moments  &  &  & \tabularnewline
+\hline 
+ & F  & Zernike moments to the n order  & used by \cite{zernike61} \tabularnewline
+Composite  &  &  & \tabularnewline
+\hline 
+ &  &  & \tabularnewline
+\hline 
+ &  &  & \tabularnewline
+\hline 
+ &  &  & \tabularnewline
+\hline
+\end{longtable}
+%\end{table}
+  % Rubine el al.\cite{gestureexample12}  used a set of statistical features,  There are a lot of features used in the sketch recognition literature
+%How to handle multi-stroked symbols. 
+%The clustering algorithm.  
+%The classifier used in recognition
+%Describe the features used in recognition [ref to \cite{zernike61}]
+
+\section{Classification}%%Feature Set
+\label{sec:Classification}
+ After computing the features the symbol is introduced to the classifier.
+The system used a support vector machine (SVM) classifier with Gaussian kernel (RBF kernel) \cite{libsvm}.
+The system used (object versus all) OVA classifiers structures.
+A validation algorithm used to choose the Kernel variables parameters  $c,\gamma$ to be used in the SVM classifiers.
+The Rubine algorithm (Linear Gaussian classifier ) was implemented for comparing the system.
+The results section \ref{} demonstrates the results of both classifiers.
+\\%Some features like number of parallel, perpendicular and intersecting lines are computed from the segmentation other like density, centroid and Zernike moments are computed from the raw stroke points.
 
@@
 %4.3.1 Previous Work . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . 28
-%4.3.2 Statistical Graph Definitions and Error-driven Stochastic Graph Mathcing . 29
+%4.3.2 Statistical Graph Definitions and Error-driven Stochastic Graph Matching . 29
 %4.4 Image Based Symbol Recognition . . . . . . . . . . . . . . . . . . . . . . . . . . 30

### 2009-02-18 draft/ChapterIntroduction.tex
before removing the comment

--- 2009-02-18 trunk/draft/ChapterIntroduction.tex
@@
 Sketchs here stand for any hand drawn drawing the user draws using digital stylus or similar device.
-Normally, sketch are composed of smaller segments we call strokes.
+Normally, sketchs are composed of a set of strokes.
 The time between pen-up and pen-down events identify each stroke.
@@
 Current CAD systems do not give the designer the ease of altering the early designs.
-Designer had to sketch primary designs on paper and pencil next, the finally design is transferred into computer using a menu CAD system.
+Designer had to sketch primary designs on paper and pencil next, the final design is transferred into computer using a CAD system menus.
 The process is time consuming and tiring.
@@
 
- A sketch recognition and understanding system will provide a powerful tool to the designers to draw sketch as they usually draw using paper and pencil with the advantages of computer program.
+ A sketch recognition and understanding system will provide a powerful tool to the designers to draw sketch as they usually draw using paper and pencil with the advantages of a computer program.
 Designer needs a normal sketching environment that will provide him facilities in paper interface and the modifying and simulation that is in current CAD systems.
@@
 
-\section{Research Scope}
-%rewirte this part.............................
-%  research will invistgated till sketch recognition in domain of uml diagrams. 
-%This research will include investigating the best method to recognize sketches.
-The scope of the research will include recognizing various sketches.
-The methods used in this research are comparing current systems and enhancing them and adobpiting them into simple geometrical symbols.
-Using well established AI methods specially \textit{Swarm Intelligence} to solve the problem.
-
-\section{Problem Definition}
+\section{Problem Definition and challenges}
 \label{sec:ProblemsAndChallenges}
-
-%Sketch recognition is divided into three main steps preprocessing, segmentation and symbol recognition.
-The preprocessing phase captures user input stroke points and collects basic information about the stroke then proceeds to remove noise and compute basic statistical and geometrical information.
-In segmentation phase, the strokes are divided into a set of simple geometrical primitives or segments.
-In the third phase, sketch recognition, strokes and segments are clustered to formulate symbols that can be recognized by a classifier system.
 
@@
 
-The research in the area had faced few problems that will need to be handled to implement an entirely sketch understanding system.
+Sketch recognition is divided into three steps preprocessing, segmentation and symbol recognition.
+The preprocessing phase captures user input stroke points and collects basic information about the stroke then proceeds to remove noise and compute basic statistical and geometrical information.
+In segmentation phase, the strokes are divided into a set of simple geometrical primitives or segments.
+In the third phase, sketch recognition, strokes and segments are clustered to formulate symbols that can be recognized by a classifier system.
+
+The research in the area had faced few problems that will need to be handled to implement an entirely sketch recognition system.
 The user draws strokes in a spontaneous and ambiguous manner that intensifies the challenges.
-The first problem is segmenting strokes from the way the user draw it into primitive geometric.
-Ink parsing is the problem dealing with clustering interpolated strokes and grouping them into symbols.
-After that, recognizing the symbols from the space the system can identify is the final problem.
-All those problems must be solved on a real time constrain to maintain an online sketching system if it was a requirement.
-
-The next sections will explain and investigate each of those problems in details.
-
-\subsection{Segmentations}
-\label{sec:Segmentations}
-
+One of the problems is segmenting strokes from the way the user draw it into primitive geometric.
+Another major problem is the high level of ambiguities in the drawing.
+The next sections discripe each problem in details
+\subsection {Segmentation}
 Segmentation is dividing the strokes draws into the geometric primitive's as lines, arcs, and ellipses.
@@
 
-\subsection{Ink parsing}
-\label{sec:InkParsing}
-
- The clustering of segmented strokes into group of strokes before passing into the next stage this process is called ink parsing.
-It is simply deciding when a symbols ends and a new symbol begins.
-Ink parsing is most important and difficult challenge as grouping of strokes can be done in many ways and configuration, which increase exponentially with the number of strokes.
-In addition, the ink parsing will affect the next steps of recognizing.
-The user does not draw the symbol in same order.
-Moreover, users can draw a sequence of strokes in symbols then leave it to draw another then return again to add segment to symbols so temporal data is helpful but not conclusive.
-The user may also draw the same symbols using different orders and methods for example the user can draw a rectangle using single stroke or using two, three or even four strokes.
-All those factors gather with the sloppy drawing of the user to increase the complexity of the problem.
-
-\subsection{Ambiguities}
-\label{sec:Ambiguities}
-
-The major problem of sketch understanding is the high level of ambiguities in the drawing.
+\subsection{Ambiguity}
 The user draws strokes that the computer cannot decide if it is arc or line.
 The nature of hand drawn sketch it self is highly ambiguous even for human eye you may see two drawn you cannot decide if it is square or rectangle.
+
  The ambiguities in sketch understanding are found at many levels in the system.
 Firstly, at the segmentation phase where the same user stroke can be interpreted as a ellipse and a circle.
-Then the symbol recognition as a symbol can have more than one interpolation.
+Symbol ambiguity is antoher problem as a symbol can have more than one interpolation.
 Due to the user sloppy drawing the symbols can have more than one interpolation which varies with the domain drawn in it similar problem was faced in handwriting recognition see Fig [\ref{fig:VorU}] but it was handled with knowing information about the context information which may not be the case in sketch recognition.
@@
 
-The ambiguity problem must be first recognized ambiguities or varicose interpolations of same symbol.
-After that, resolve those ambiguities is detecting the best correct interpolation of the symbol.
-To solve this problem the program must be able to detect the amount of information needed to decide which interpolation of symbols is the best one.
+\section{Research Scope}
+%rewirte this part.............................
+%  research will invistgated till sketch recognition in domain of uml diagrams. 
+%This research will include investigating the best method to recognize sketches.
+The scope of the research will include recognizing various sketches.
+The methods used in this research are comparing current systems and enhancing them and adobpiting them into simple geometrical symbols.
+Using well established AI methods specially \textit{Swarm Intelligence} to solve the problem.
+ This research investigats the best method to recognize sketches.
+The scope of the research includes recognizing various symbols drawn using single and multipule strokes.
+The research foucs on using various Particle Swarm Optimization (PSO) methods to segments strokes into goemetrical primitives.
+Sketch recognition is achieved by  Support Vector Machines (SVM) classifier.
+A various set of features are investigated to reach the best set of features.
 

### 2009-03-03 draft/System.tex
After the finish of chapter 1,,3,4

--- 2008-09-02 draft/System.tex
@@
-\chapter{Proposed System for Sketch Understanding}
+\chapter{The Proposed System}
 \section{Aims of the System}
 \label{sec:AimsOfTheSystem}
-The previous chapter shows the current trends in sketch recognition. This thesis introduce a solution for the problem that give optimize solution for the segmentation problem.
+The previous chapter shows the current trends in sketch recognition.
+This thesis introduces a solution for sketch recognition problem.
+The research gives an optimize solution for the sketch segmentation problem.
+Particle Swarm Optimization (PSO) is used to correctly segment strokes into curves and lines.
+The current implemented systems either segment strokes based on simple curvature\cite{meanshift10,earlySketchbased4} or speed information\cite{earlySketchbased4}.
+%the performance of most algorithms is based on user style.
+
+% Stroke segmentation is one of the most complex problems in the sketch recognition system. It is dividing the   
+%In stroke segmentation problem is Particle swarm optimization is used as the
 
@@
 \label{sec:AnOverviewOfTheSystem}
+   This research solves sketch segmentation problem using particle swarm algorithm PSO.
+Due to sloppiness of users and hardware glitches the captured points must be processed to remove noise and redundancy before the system proceeds to compute the speed and curvature information.
+After that a set of possible corner or critical points is computed to guide the segmentation procedure.
+A set of features is extracted from the segmented strokes to be used by the classifier.
+%The stroke is the path of points from the instant the pen is down till it is up.
+The block diagram in Figure \ref{fig:Blockdiagram} shows that the main system blocks are 1) preprocessing, 2) segmentation, 3) feature extraction, and finally 4) classification.
+The next section will describe each component in details.
+ %Due to sloppiness of users and hardware glitches the system processes the data before further
+\begin{figure}[]
+	\centering
 
-\section{System Components}
+\begin{center}
+	\includegraphics[scale=0.75]{images/AllBlockDiagram.eps}
+	\caption[The System Block Diagram]{The block diagram of the system}
+	\label{fig:Blockdiagram}
+\end{center}
+\end{figure}
+
+\subsection{System Components}
 \label{sec:SystemComponents}
-% the system and discription of this
 
-\chapter{Preprocessing and Segmentation}
-\label{sec:Preprocessingandsegmentation}
+\textbf{Preprocessing} step is responsible of capturing the input data and removing the noise from it.
+The preprocessing block as in Figure \ref{fig:Blockdiagram} consist of computing the time differences, direction, speed and curvature of every point in the stroke.
+After computing this information the local minimum and local maximum of each curve is extracted from the data computed.
+This process estimate a set of a set of \textit{Possible Dominant Points $P_{pd}$} that guide the segmentation process.
+Later, other geometrical and statistical computations are evaluated from the stroke.
+%The system then proceeds to estimate to help in the segmentation process.
+\\
+
+The next step is \textbf{Segmentation}, the goal of the segmentation stage is to divide strokes into segments of either curves or lines.
+As shown in Figure \ref{fig:Blockdiagram} first an attempt is made to fit the stroke points into a curve or an ellipse using a minimum square error fitting algorithm \cite{chernov-2003}.
+If the stroke proved to be an elliptical arc then the segmentation process ends and the system proceeds to the next step.
+Otherwise, the stroke is passed to two particle swarm algorithms that divide the stroke to either a polygon or a set of lines and curves.
+The algorithms takes the stroke points along with the possible dominant points $P_{pd}$ computed during preprocessing then produce a set of dominant points which are connected by either lines or curves.
+The two algorithms will generate two segmentations, the system will choose the segmentation that has the minimum segmentation error.%
+% First an attempts the segmentation process is divided into two steps ellipse fitting and curve segmentation.
+In the first step, the ellipse fitting process tries to fit the stroke into an ellipse.
+If the system fails it passes the stroke to the second step; curve segmentation which consists of two PSO segmentation algorithms.
+The two algorithms will generate two segmentations, the system will choose the segmentation that has the minimum segmentation error.%  generate the segmentations,  the minimum error will be the chosen segmentation.
+% is the ellipse fitting
+
+ %the stroke first If the ellipse detection fails the stroke is passed to the segmentation algorithms which will pass it to the two PSO algorithms described below the segmentation with the minimum error will be the chosen segmentation.
+The segmentation is then added to the set of un-recognized segments in the system.
+\\% this part is repeated for each stroke.
+ %the clustering algorithm starts to group segments together after the segmentation step.
+The system let the user draws the symbol by using any number of strokes, a set of unrecognized segments is passed to the clustering algorithm to generate a symbol and compute a feature vector for it.
+
+ In \textbf{Feature Extraction} step, a feature vector consists of a set of statistical, geometrical and spatial features is extracted from the segmented strokes.
+The feature vector is used as input for the next step.
+%the system computes composite set of features some are statistical other are spatial features based on the type primitives.% The system extract segment based and statistical based features from the set of segments that the clustering algorithms produce.
+
+   The final step is \textbf{Classification} using a SVM classifier.
+The classifier attempts to classify the symbol into one the previously trained classes.% c  that will use the features computed to classify the segments into one of the previously trained classes.
+% Or to determine the symbol of the given segments from the set of preciously trained symbols.
+
+ %with a symbol from the training set.% The system compute composite set of features some are statistical other are spatial features based on the type primitives.
+Finally, the strokes are classified into the corresponding classes.% the classifier identifies the strokes into a symbol from the set of known symbols.\\
+
+\chapter{System Details}
+\label{sec:SystemDetails}
 
@@
 \label{sec:Preprocessing}
+%Some data is extracted from the points in the stroke.
+The preprocessing step captures the points from the pointing devices then compute the information needed to determine a set of  \textit{Possible Dominant Points $P_{pd}$} which guides the segmentation step.
+It is noted as a person draws a shape the pen slows down near corners and picks up speed when drawing straight lines.
+Therefore, the speed information is widely used to identify the shape corners and edges \cite{earlyprocess}.
+The curvature information is used to determine the points with high angular changes along the path of points \cite{meanshift10}.
+Those observations help detecting the dominant points as the points with lower speeds and high curvatures.
 
-\subsection{Curvature Calculation}
+Also the time difference between samples was used in \cite{polygonfeedback31} as it provides more distinct maxima than the speed information.
+Agar et al.
+\cite{polygonfeedback31} mentions that the pointing device (for example the mouse) sampling rate is the reason for this phenomena.
+The points are sampled at regular time intervals while the pointing device is moving.
+There are no points while the pointing device is stationary.
+This leads to a nearly constant time difference between samples while the pen is moving and large difference while the pen is stationary.
+Contrary to speed information, where the users draw with variant speeds the speed information has a lot of noise in the data.
+Figure \ref{fig:speed2Distance} shows the time difference and the speed graphs for the stroke drawn in Figure \ref{fig:orignalStroke}.
+Similarly, direction information is used as it provides better distinctive maxima for the corners than the curvature information \cite{meanshift10}.
+Figure \ref{fig:curvatures} shows the direction and curvature graphs for the stroke drawn in Figure \ref{fig:orignalStroke}
+
+ %The system compute speed, curvature, time difference and direction data then generate a set of possible corners that will be used in the segmentation as an initial solution. 
+\subsection{Preliminary Calculation}
 \label{sec:CurvatureCalculation}
+
+ In this research, time difference, direction, speed and curvature of each point along the stroke was computed.
+The experiments in section \ref{} proved that computing all information's provide better segmentations results than only using one of the graphs.
+The time difference is calculated as $\Delta t = t_{i+1} - t_i$ where $t_i$ is the time of the point $i$ and $t_{i+1}$ is the time of the point $i+1$.
+The speed is calcuated using equation $v=\Delta s/\Delta t$ where $t$ is the time difference between two points and $s$ is the length between them.
+
+\begin{equation}
+\label{eq:direction}
+	\vartheta  = \cos ^{ - 1} \left( {\frac{{\overrightarrow {p_1 p_2 }  \cdot \overrightarrow {p_2 p_3 } }}{{\left\| {\overrightarrow {p_1 p_2 } } \right\| \times \left\| {\overrightarrow {p_2 p_3 } } \right\|}}} \right)
+\end{equation}
+
+  The direction is calculated as the angle between two vectors as in equation \ref{eq:direction} where $\overrightarrow {p_i p_j }$ is the vector from point $pi$ to point $pj$ and $\|{\overrightarrow {p_i p_j }}\|$ is the normal of the  vector $\overrightarrow{p_i p_j }$.
+The curvature is estimated using more than one method the result of each curvatures estimation is reported in section \ref{sec:Curvatureestimationresults}.
+The best curvature estimation is used, the estimation define curvature as the change in direction with respect to length i.e.
+$c= \Delta d/\Delta s$.
+
+ %system compute speed, distance data for all points in the stroke. Curvature is computed using estimation used in [] where direction computed is angle between two lines. Curvature is 
+ %$\Delta d/  \Delta S $ where d is the difference direction of point and s is difference in distance between points. 
+ All these calculations are performed in real time while the user draws the strokes.
+The complexity of computation is $O(n)$ where $n$ is number of points.
+Figures \ref{fig:speed2Distance} \&  \ref{fig:curvatures} show the computed information's for the stroke drawn in Figure \ref{fig:orignalStroke}.% After computing the graphs the system  compute the average, minimum and maximum values of all curves.
+\\% the lower speed points correspond to vertex and dominant point's location.
+The higher the direction and curvature data correspond to location with higher change in curvature witch promote the location for vertices'.
+\\
+%Description of input data. 
+%how to calculate speed, curvature, area, bounding box.
+%how to remove noise. 
+%Finally how to compute primarily dominant points.
+
+\begin{figure}
+	\centering
+			\subfigure[ Speed Graph]{	\includegraphics[scale=0.5]{images/speed2.eps}}
+			\hfill
+			\subfigure[ Time Difference Graph] {\includegraphics[scale=0.5]{images/time2.eps}}
+	\caption{Speed and Time Difference Graphs}
+	\label{fig:speed2Distance}
+\end{figure}
+
+\begin{figure}[]
+	\centering
+		\includegraphics[scale=0.8]{images/orignalStroke.eps}
+	\caption{Input Stroke} Example of an input stroke to the system. 
+	\label{fig:orignalStroke}
+\end{figure}
+
+\begin{figure}[]
+%\begin{minipage}[b]{0.8\linewidth}
+	\centering
+
+	%	\subfigure[ The Speed of data] {\includegraphics[scale=0.35]{images/speed2.eps}}
+
+		\subfigure[ Direction Graph ] {\includegraphics[scale=0.35]{images/direction2.eps}}
+					%\end{minipage}
+					%\begin{minipage}[b]{0.8\linewidth}
+			\subfigure[Curvature Graph] {\includegraphics[scale=0.35]{images/curvature2.eps}}
+
+			%\end{minipage}
+
+	\caption{Curvature and Direction Graphs}%{The data of the stroke}
+
+	\label{fig:curvatures}
+\end{figure}
 
@@
 \label{sec:CriticalPointDetection}
+After the system computes all the speed, time difference and curvature data it proceeds to detect the points with low velocity and high curvatures.
+Using simple differentiation to detect local extreme points resulted in false points due to the non smooth curves.
+Hence, the system adopted a process presented by \cite{earlyprocess}, where the mean of the curve is calculated.
+Then a threshold \textit{th} is used to separate the curve into regions $Region_i$; each region $Region_i$ is defined as a range of points, where the curve values are either above or below the threshold \textit{th}.
+Those regions are further processed to find the maximum point $Max(Region_i)$ of each region $Region_i$.
+The stroke points $p_i(x,y)$ that correspond to those maximum values are labeled as \textit{possible dominant points} $P_{pd}$.
+Figure \ref{fig:MaxRegioi} shows an example of $Region_i$ of the speed curve and the dominate points $P_{pd}$ which correspond the minimum of point in this region $Min(Region_i)$.
+% (as shown are redundant)
+\begin{figure}
+	\centering
+		\includegraphics[scale=0.5]{images/speedreg.eps}
+	\caption{Possible Dominant Points Calculations}
+	\label{fig:MaxRegioi}
+\end{figure}
+
+   %After the threshold is applied the curve is divided into regions.
+Each region is defined by a part in the data curve between tow intersection point of the threshold line and the curve.
+% to find the extreme points.
+Hence, the mean of the direction curve is calculated and used as threshold.
+\\
+   %If we tried to differentiate the curve the result will be false threshold values we divide the data into regions of data higher or lower that the threshold.
+This will let us only look for high data only.
+The maximum of each region is then computed and reported as a possible vertex point.
+
+The system repeats this process to curvature, direction, time difference and speed curves.
+For each curve the points labeled as possible dominant points $P_{pd}$ are saved into a single array.
+The points that are repeatedly labeled are given a higher score than the other points.
+This score is then used to sort the  $P_{pd}$ points array.
+The array that contains possible points $P_{pd}$ is used as input for the next segmentation step.
+Figure \ref{fig:LabelsPPD} shows the particles labeled as Possible dominate point $P_{pd}$ by the preprocessing, it is noticed the redundancy of some $P_{pd}$ points.
+
+\begin{figure}
+	\centering
+		\includegraphics{images/ppd.eps}
+	\caption{Posible domninate points}% a) Possible dominate point b) Particle encoding  
+	\label{fig:LabelsPPD}
+\end{figure}
 
@@
 \label{sec:Segmentation}
-\subsection{Swarm Introduction}
-\label{sec:SwarmIntroduction}
+The segmentation tries to divide the stroke into a set of primitives.
+As shown in Figure \ref{fig:Blockdiagram} first an attempt is made to fit the stroke points into a curve or an ellipse using a minimum square error fitting algorithm \cite{chernov-2003}.
+If the stroke proved to be an ellipse arc then the segmentation process ends and the system proceeds to the next step.
+Otherwise, the stroke is passed to two particle swarm algorithms that divide the stroke to either lines or lines and curves.
+The algorithms take the stroke points along with the possible dominant points $P_{pd}$ computed during preprocessing then produce a set of dominant points which are connected by either lines or curves.
+The next section describes the ellipse detection algorithm and the two particle swarm algorithms used to divide the stroke.
+%If the stroke proved to be an ellipse then the segmentation process ends and the system proceeds to the next step.
+Otherwise, the stroke is passed into two particle swarm algorithms that will divide the stroke to either lines or lines and curves (see the block diagram in fig.
+\ref{fig:Blockdiagram} ).
+%The algorithms takes the stroke points along with the possible dominant points computed then produce a set of dominant points which are connected with either lines or curves (see fig.
+\ref{fig:Blockdiagram}).
+The next sections will describe the ellipse detection algorithm and both the two particle swarm algorithms used to divide the stroke.
+%After computing the primary data the system now tries \\
+%Paragraphs describe the segmentation algorithm.
 
-\subsection{Ellipse Detection }
+\subsection{Ellipse Fitting }
 \label{sec:EllipseDetection}
+%Describe the ellipse detection algorithms 
+This process tries to fit the stroke points into an ellipse arc; it starts with computing the center of the stroke bounding box.
+The bounding box center point is used as the first estimation of the center of the ellipse.
+The axes of the ellipse are estimated as the $w/2$ and $h/2$ of the stroke bounding box where $w$ is width and $h$ is height of the bonding box.
+The least square fitting algorithm \cite{chernov-2003} is used to minimize the fitting error of the ellipse Equation (\ref{eq:circleFit})
 
+\begin{equation}
+E = \sum\limits_{i = 0}^N {\frac{{(x_i - x_0 )}}{{a^2 }}^2  + \frac{{(y_i - y_0 )}}{{b^2 }}^2  - 1} 
+\label{eq:circleFit}
+\end{equation}
+
+ where $N$ is number of points in the stroke, $a,b$ are the length of ellipse axes, $x_0$ \& $y_0$ are the coordinates of the center point, $x_i$ \& $y_i$ are the coordinates of point $i$ in the stroke.
+A list of new values for $x_0$ , $y_0$ ,$a$ and $b$ are generated randomly from the older values with small increments after each loop.
+After few loops, the final fit error and the ellipse confidence value are computed.
+The confidence value $Conf$ computed by Equation (\ref{eq:circleError})
+ \begin{equation}
+Conf = E_{ellipse}+E_{area}
+\label{eq:circleError}
+\end{equation}
+ \begin{equation}
+E_{area}  = \sqrt {\left( Area(stroke) - Area(ellipse)\right)^2 }
+\label{eq:ErrorArea}
+\end{equation}
+
+   where $E_{ellipse}$ is the error computed by Equation(\ref{eq:circleFit}) and $E_{area}$ is computed as the difference between the actual area of the stroke and the area of the estimated ellipse Equation(\ref{eq:ErrorArea}).
+If confidence value $Conf$ is less than threshold $th_{Ellipse}$\footnote{The threshold used is $th_{Ellipse}=0.7$ } then the stroke is segmented as an ellipse otherwise the system proceeds proceed to the two PSO segmentation algorithms.
+% The confidence value is used to label the stroke as ellipse or un segmented stroke.
+If the stroke as un-segmented the next process will be to pass the stroke into curve segmentation algorithm.
+
+   The next section describes the \textit{Discrete Particle Swarm Algorithm (DPSO)} then it proceeds to detail the two DPSO segmentation algorithms.
+
+   %to check if the stroke can be labeled as ellipse or not.\\ %is computed to check if stroke is an ellipse. \\
+%We found that if the confidante is above threshold then the probability of ellipse is highest otherwise the stroke is passed to the next section to get the divisions of stroke and test its error.
+
+\subsection{Discrete Particle Swarm Algorithm}
+\label{sec:ParticleSwarmAlgorithm}
+%\section{Particle Swarm Algorithm}
+%\label{PSO}
+%What is particle swarm algorithm and how it was used in related researches. 
+The main idea of \textit{Particle Swarm Algorithm (PSO)} is to represent each agent with a particle from the solution space \cite{PSOFirst}.
+Each agent moves the particle with a direction and velocity $v_{ij}$ based on equations \ref{eq:Swarm} \& \ref{eq:Swarm1}.
+\begin{equation}
+%\[
+p_{ij}=p_{ij}+v_{ij},
+%\
+\label{eq:Swarm1}
+\end{equation}
+where $p_{ij}$ represent the $jth$ particle in the $ith$ agent and $v_{ij}$ is the velocity of the $jth$ particle in the $ith$ agent.
+ %Equation [\ref{eq:Swarm}] shows how velocity and direction of each particle are computed
+ \begin{equation}
+v_{ij}  = v_{ij}  + c_1 r_1 (lbest_{ij}  - p_{ij} ) + c_2 r_2 (gbest_{ij}  - p_{ij} )
+\label{eq:Swarm}
+\end{equation}
+ where $lbest_{ij}$ is the local best particle, $gbest_{ij}$ is the global best particle, $r_1$ \& $r_2$ are random variables and $c_1$ \& $c_2$ are the swarm system variables.
+ After each iteration the global best $g_{best}$ particle and the agent local best $l_{best}$ particle are evaluated based on the maximum fitness functions of all particles in the solution space.
+The solution is found after achieving a specific number of iteration or after an error threshold is achieved.
+Equation \ref{eq:descrite}  
+\begin{equation}
+   P(i)\Leftarrow 
+\{
+\begin{array}{c} 
+1 \quad \quad if\quad r_{3}>p_{i}  \\
+
+0 \quad \quad if\quad r_{3}<p_{i} 
+\label{eq:descrite}
+\end{array}\}
+\end{equation}
+ where $p_{ij}$ is the numerical values of the particle and $r_{3}$ is a random variable, is used to change the general swarm algorithm into binary particle (\textit{Discrete Particle Swarm Algorithm DPSO}) which handles particle values of either $0$ or $1$ \cite{PSODisceret}.
 \subsection{Swarm Segmentation}
 \label{sec:SwarmSegmentation}
-\subsubsection{Problem Formulation }
+Two DPSO algorithms generate two different segmentations for each stroke.
+Each DPSO algorithms tries to find the best particle in the solution space.
+Both algorithms has the same problem formulation but different fitness functions.
+After generating two segmentations the system chooses the best segmentation from the outputs of the DPSO.
+The details of the problem formulation and fitness functions are in the following sections.
+%In this process the system generate two stroke segmentation using two PSO segmentation algorithm.
+The system generates segmentation from both algorithms and then chooses the segmentation with the minimum error value.
+The problem definition is the same in both algorithms but they differ in the fitness function and the error functions.
+%formation is nearly the same in both.
+\subsubsection{Problem Formulation}
 \label{sec:ProblemFormulation}
+The input stroke with $N$ points can be represented by set $S = \left\{ {x_1 ,x_2  \ldots x_N } \right\}$ where $x_i$ is the location of the point $i$ .
+The swarm algorithms consist of $M$ agents which are represented by the set
+$A  = \left\{ {P_i \left| {i = 1,2 \cdots M} \right.} \right\}$ where $P_i$ is a single solution particle from the solution space.
+Each particle decodes the problem with binary array with the same length $N$ as the input stroke (Figure \ref{fig:CodingSwarm}).
+\begin{figure}
+	\centering
+		\includegraphics{images/CodingSwarm.eps}
+	\caption{An Example Stroke and the Coding}
+	\label{fig:CodingSwarm}
+\end{figure}
+% The particles of the swarm represent a single solution of the solution space.
+For this problem the, each particle %will give a different segmentation for the input stroke.
+Firstly, we will define the stroke with N points by the set %S where.
+We define the arc %An array with the same length as the number of points of the strokes.
+The system represents each particle $P_i$ by $P_i = \left\{ {p_{ij} \left| {j = 1,2 \cdots N} \right.} \right\}$ where $p_{ij}$ is either 0 or 1 where $p_{ij}=1$ means that point $j$ is a dominant point(see Figure \ref{fig:CodingSwarm}.
+Thus , the particle represents points that are chosen to be used as dominant points for this segmentation.
+The goal of the DPSO algorithm is to find the solution $P_i$ that generates the minimum set of dominate points that defines the stroke with minimum segmentation error.
+In other words, the system tries to find the fewer number of points (1's) location in the particle array which gives minimum segmentation error.
 
-\subsubsection{Fitness Computation }
-\label{sec:FitnessComputation}
-\subsubsection{Polygon Division Algorithm}
+The fitness function and error calculations are different in each algorithm. The first fitness and error function are described below.
+
+\subsubsection{Polygon Division Algorithm \textsl{AlgS1}}
+\label{sec:PolygonDivisionAlgorithm}
+The first algorithm, the algorithm tries to segment the stroke as a polygon.
+The final solution is a set of line segments that best define the input stroke.
+Given the inputs stroke let's define the set of points in the stroke as  $S = \left\{ {x_1 ,x_2  \ldots x_N } \right\}$ which is the set of consecutive points from start of stroke till the end.
+The arc $\widehat{x_ix_j}$ is defined as the consecutive set of points from point $x_i,x_{i+1} \cdots,x_j$.
+The line
+$\overline{x_i x_j} $ is the straight line connecting point $x_i$ to point $x_j$. The approximation error is computed by the equation \ref{eq:ErrorSwarm1} 
+\begin{equation}
+E=\sum\nolimits_{i = 0}^M e ( \widehat{x_ix_{i+1}},\overline{x_i x_{i+1}})
+\label{eq:ErrorSwarm1}
+\end{equation}
+ where $M$ is the number of dominant points in this solution.
+The error $ e ( \widehat{x_ix_j},\overline{x_i x_j})$ is computed as the sum of squared perpendicular distance from every point along the arc $\widehat{x_ix_j}$ to the line $\overline{x_i x_j}$ \cite{PolygonApproximationPSO}.
+Figure \ref{fig:DPSOERROR} shows a graphical representation of the error.
+The fitness is computed using equation \ref{eq:fitnessSwarm1}
+\begin{equation}
+\max fitness(p_i ) = \left\{ {\begin{array}{*{20}c}
+   { - E/\varepsilon N} & {ifE > \varepsilon ,}  \\
+   {D/\sum\limits_{j = 1}^N {p_{ij} } } & {otherwise}  \\
+\end{array}} \right.
+\label{eq:fitnessSwarm1}
+\end{equation}%\]
+where $N$ is the number of points in the stroke, $D$ is the number of point in the solution that was previously labeled as $P_{pd}$  and $E$ is the computed error and $\varepsilon$ is the error threshold.
+If the solution produced an error that exceed the error threshold $\varepsilon$ the fitness function assign a -ve value for the solution to express it's in feasibility, otherwise the inverse of the number of vertex produced is used to access the solution fitness in terms of minimum number of vertices'.
+This fitness function optimize two goals the first is to move the solution into the feasible solution space with acceptable error bounds and second is to fly the particle to a new position which may result in a polygon with a fewer number of vertices'\cite{PolygonApproximationPSO}.
+\begin{figure}
+	\centering
+  \includegraphics[scale=0.8]{images/pso1.eps}			
+	\caption{ALS1 Error}% a) Possible dominate point b) Particle encoding  
+	\label{fig:DPSOERROR}
+\end{figure}
+ %The error is chosen to in favor of larger than the threshold is given a -ve value to lower the value of solution otherwise the system will favor the lower number of vertices'.\\ %we will want to lower the number of vertices'.
+\\
+%if we say that s
+%Alg1:
+%Alg2: 
+\subsubsection{Curve Divisions Algorithm \textsl{AlgS2}}
 \label{sec:PolygonDivisionAlgorithm}
 
-\subsubsection{Curve Divisions Algorithm}
-\label{sec:PolygonDivisionAlgorithm}
+The second algorithm has the same problem formulation but different fitness and error functions.
+It was previously introduced in  \cite{CruveDivisionSwarm} where genetic programming was used as the optimizing algorithm.
+The particle is represented using the arrays $P_i = \left\{ {p_{ij} \left| {j = 1,2 \cdots N} \right.} \right\}$ where $p_{ij}$ is either 0 or 1.
+Lets denote the segments $\widehat{x_ix_j}$ as the consecutive set of points from point $x_i,x_{i+1} \cdots,x_j$.
+The algorithms attempts to fit each segment $\widehat{x_ix_j}$ to both line and circle.
+The type of the segment is determined according to the minimum error of each fit.
+The final error of the solutions is the summation of error of all segments.
+%The goal of the algorithm is to fit the segments between the solution vertices' into straight lines or a circular arc.
+\paragraph{Fitting segments into straight line.}
+\label{sec:FittingSegmentIntoStraightLine}
+According to \cite{CruveDivisionSwarm} each segment $\widehat{x_ix_j}$ given a set of points $x_i,x_{i+1} \cdots,x_j$ can be fitted into the line: $y=kx+c$ where $k$ and $c$ are the slope and the intercept of the line respectively.
+Equation \ref{eq:Linek} and \ref{eq:LineC}
+\begin{equation}
+\label{eq:Linek}
+k = \frac{{N\sum\limits_{i = 1}^N {x_i y_i }  - \sum\limits_{i = 1}^N {x_i } \sum\limits_{i = 1}^N {y_i } }}{{N\sum\limits_{i = 1}^N {x_i^2 }  - \left( {\sum\limits_{i = 1}^N {x_i } } \right)^2 }}
+\end{equation}
+\begin{equation}
+\label{eq:LineC}
+c = \frac{{\sum\limits_{i = 1}^N {x_i^2 } \sum\limits_{i = 1}^N {y_i }  - \sum\limits_{i = 1}^N {x_i } \sum\limits_{i = 1}^N {x_i y_i } }}{{N\sum\limits_{i = 1}^N {x_i^2 }  - \left( {\sum\limits_{i = 1}^N {x_i } } \right)^2 }}
+\end{equation}
+, where $N$ is the number of points in the segment and $(x_i, y_i)$ are the coordinate of the point $i$, are used to fit the segment $\widehat{x_ix_j}$ into the straight line.
+\begin{equation}
+\label{eq:ds}
+ d_s  = \frac{{\sum\limits_{i = 1}^N {\left| {(kx_i  + c) - y_i } \right|} }}{{N\sqrt {k^2  + 1} }}
+\end{equation}
+The distance ($d_s$) (Equation \ref{eq:ds}) is the average distance from segment points $(x_i, y_i)$ to the estimated line. This distance $d_s$ is used as the error of the line approximation.
+\paragraph{Fitting segments into circular arc.}
+\label{sec:FittingSegmentcirculararc}
+Each segment is also fitted into a circular arc $(x - a)^2  + (y - b)^2  = R^2, $ where ($a,b$) are the coordinate of the center of the circle and $R$ is the radius of the circle.
+They is estimated using the following set of equations:
 
-\section{BenchMarck algorithm}
+\begin{equation}
+a = \frac{{b_{1}a_{22} - b_{2}a_{12}}}{\Delta},
+\end{equation}
+\begin{equation}
+	b = \frac{{b_{2}a_{11} - b_{1}a_{21}}}{\Delta},
+\end{equation}
+\begin{equation}
+R = \sqrt {\frac{1}{N}(\sum\limits_{i = 1}^N {x_i^2 }  - 2\sum\limits_{i = 1}^N {x_i a}  + Na^2  + \sum\limits_{i = 1}^N {y_i^2  - 2} \sum\limits_{i = 1}^N {y_i b + Nb^2 } )} ,
+\end{equation}
+
+Where 
+\begin{equation}
+a_{11}  = 2\left[ {\left( {\sum\limits_{i = 1}^N {x_i } } \right) - N\sum\limits_{i = 1}^N {x_i^2 } } \right]
+\end{equation}
+\begin{equation}
+a_{12}  = a_{21}  = 2\left( {\sum\limits_{i = 1}^N {x_i \sum\limits_{i = 1}^N {y_i } }  - N\sum\limits_{i = 1}^N {x_i y_i^2 } } \right),
+\end{equation}
+\begin{equation}
+ a_{22}  = 2\left[ {\left( {\sum\limits_{i = 0}^N {y_i } } \right)^2  - N\sum\limits_{i = 0}^N {y_i^2 } } \right]
+\end{equation}
+\begin{equation}
+b_1  = \sum\limits_{i = 0}^N {x_i^2 } \sum\limits_{i = 0}^N {x_i }  - N\sum\limits_{i = 0}^N {x_i^3 }  + \sum\limits_{i = 0}^N {x_i } \sum\limits_{i = 0}^N {y_i^2 }  - N\sum\limits_{i = 0}^N {x_i y_i^2 } ,
+\end{equation}
+\begin{equation}
+b_2  = \sum\limits_{i = 0}^N {x_i^2 } \sum\limits_{i = 0}^N {y_i }  - N\sum\limits_{i = 0}^N {y_i^3 }  + \sum\limits_{i = 0}^N {y_i } \sum\limits_{i = 0}^N {y_i^2 }  - N\sum\limits_{i = 0}^N {x_i^2 y_i } ,
+\end{equation}
+
+\begin{equation}
+\Delta  = a_{11}b_{22}-a_{12}a_{21}.
+\end{equation}
+ The error of the circle estimation is measured using the distance $d_c$ which is calculated using the equation \ref{eq:circleE}:
+ \begin{equation}
+ \label{eq:circleE}
+d_c  = \frac{{\sum\limits_{i = 0}^N {\left| {\sqrt {(x_i  - a)^2  + (y_i  - b)^2 }  - R} \right|} }}{N}
+\end{equation}
+Thus, Equation \ref{eq:circleE} $d_c$ defines the average distance $d_c$ from the segment points $x_i,y_i$ to the estimated circle ($a,b$) and $R$.
+
+\paragraph{Determinig The Segment Type.} 
+\label{sec:DeterminigSegmentType}
+If the average distance $d_s$ equals zero this means that the segment$\widehat{x_ix_j}$ points is exactly the straight line approximated.
+If the distance ($d_c$) equals zero means that the segments point lies exactly on the circular arc estimated.
+But the user rarely draws an exact line or circle, so the distances are used as the fitting error.
+The minimum distance determines the type of the current segment, in other words if ($d_s$) is smaller than ($d_c$) then the segment is labeled as a line segment otherwise it is labeled as a circular arc.
+
+After each segment type is determined the the particle segmentation error is computed by Equation \ref{eq:errorSwarm2} 
+\begin{equation}
+E=\sum\nolimits_{i = 0}^M e(D_i) 
+\label{eq:errorSwarm2}
+\end{equation}
+where $M$ is the number of segments in the solution, $D_i$ is the  approximation error of each segment as $min(d_c,d_s)$ as computed in Equation\ref{eq:circleE}  and Equation \ref{eq:ds} \cite{CruveDivisionSwarm}.
+The fitness is computed by the equation \ref{eq:fitnessSwarm2}
+\begin{equation}
+\max fitness(P_i ) = \frac{1}{{E \times M^k }}
+\label{eq:fitnessSwarm2}
+\end{equation} where $E$ is the error and $M$ is number of segments and $k$ is a parameter tweaked to get minimum number of segments. $k$ is selected to be 0.5\cite{CruveDivisionSwarm}.
+
+\subsubsection{Solution Refine procedures } 
+\cite{PolygonApproximationPSO} used a merge and divide algorithm after each loop of the swarm system to refine the solution but we used another enhancement method.
+After each loop in the swarm algorithm, each particle loops on the set of selected dominant points to enhance the solution.
+Each dominant point is checked to find if it was labeled before as a possible dominant (computed as in section \ref{sec:Preprocessing}), if not the point is moved to the nearest labeled point.
+   After each loop of the swarm algorithm (\textsl{AlgS1} and \textsl{AlgS2}), each particle is refined using the following procedure.
+For each particle $P_i$ each dominant point $P_{ij}$ is checked to find if it was labeled before as a \textit{possible dominant point} $P_{pd}$ (computed as in section \ref{sec:Preprocessing}).
+If it was not labeled the point $P_{ij}$ is moved to the nearest labeled point.
+This ensures that all of the points generated by the DPSO are possible dominate points $P_{pd}$.
+After that the particles are tested to make sure that the distance between every two successive dominate point is larger than the constant $min_D$.
+If two points are nearer than $min_D$\footnote{Constant selected to be 10 pixels} then one of the points is removed.
+\subsection{Choosing the best fit}
+The two PSO algorithms produce two segmentation solutions.
+After that the system evaluates each solution to finally select the best segmentation.
+This evaluation is based on the area of both the segmentation and the input stroke.
+Equation \ref{eq:bestFit1} and \ref{eq:bestFit2} are used to evaluate each segmentation.
+
+\begin{equation}
+\label{eq:bestFit1}
+E_{area}=A_{st}-A_{sg}
+ \end{equation}
+ \begin{equation}
+\label{eq:bestFit2}
+E_{tot}=E_{area}+E_{alg}
+ \end{equation}
+ where $A_{st}$ is the area of the input stroke, $A_{sg}$ is area of computed segmenation, $E_{alg}$ is the approximation error of the segmentation.
+The segmentation which crosspond to the minimum $E_{tot}$ is the best segmentation.
+
+ \section{Benchmark algorithm}
+
 \label{sec:BenchMarckAlgorithm}
+In this research, a bench mark algorithm is used as reference to the results obtained by the DPSO algorithms.
+The algorithm was first implemented by \citeauthor{earlyprocess} in \cite{earlyprocess}.
+The algorithm is divided into the following steps:
+\begin{enumerate}
+	\item Compute the curvature and the speed for each point in the stroke. 
+	\item   A percentage of maximum value of the speed and curvature is used as threshold.
+		\item Mark the values higher than the threshold in a new arrays $L_c$ as the list of corner points generated from the curvature curve.
+The list $L_s$ contains the points generated from the speed curve.
+	\item Compute a confidence on each list based on the type of curve. Then sort the lists $L_s$and $L_c$ by the confidence value. 
+	\item Find the intersection points in both list and generate the a Sample solution $h_{curr}$. 
+	\item Use $h_{curr}$ as initial solution. Loop doing the following:
 
-\chapter{Symbol Recognition}
-\label{sec:SymbolRecognition}
-\section{Stroke Clustring Algoirthm }
-\label{sec:ClustringAlgoirthm}
+		\begin{enumerate}
+			\item Generate a new solution $h_{s}$ using $h_{curr}$ and  the first point in the $L_s$ list .
+			\item Generate a new solution $h_{c}$ using $h_{curr}$ and  the first point in the $L_c$ list.
+			\item Test the solutions $h_{s},h_c$ to find the minimum solution.
+			\item  If $h_s$ is the minimum then add the solution a list of hybrid solutions $Solutions_{set}$ and remove the first point from the list $L_s$ then set $h_s$ as $h_{curr}$. 
+			\item If $h_c$ then add the solution a list of hybrid solutions $Solutions_{set}$ and remove the first point from the list $L_c$ then set $h_c$ as $h_{curr}$. 
+		\end{enumerate}
+		\item Choose the final solution from the list of solutions $Solutions_{set}$ as a tradeoff between  minimum error and number of corners.  
+		\item Check each segment in the solution if the solution is can be estimated as a line segment if not it is fitted using a higher level Bezire curve. 
+\end{enumerate}
+More details about the algorithm is in \cite{earlyprocess}.
+
+%\section{Stroke Clustring Algoirthm }
+%\label{sec:ClustringAlgoirthm}
+%
+%When the user finishes drawing a stroke the segmentation algorithm generate a list of segments that represent this stroke.
+The system then groups the segments generated from the segmentation algorithm.
+%These segments are grouped into a symbol and tested if they can be classified into one of the known symbols.
+If no identification is achieved the segments are added into a list of unclassified segments.
+When a user draws another stroke the segmented stroke is added to the list and the process is repeated until all the segments are classified to a known symbol.\\
+%
+%After the user draw all strokes of the symbol he has to wait 10 seconds or press finish button beside the drawing area.
+The set of unrecognized strokes is grouped together along with their segmentation as input to the feature extraction process.
+
 \section{ Feature Extraction}%%Feature Set
 \label{sec:FeatureExtraction}
+
+The system uses a composite set of features include Rubine feature set \cite{gestureexample12}, Zernike moments \cite{HeloiseBeautification}, ink density \cite{GeometryAndDomain102} and some structural information like number of perpendiculars lines , number of parallel lines and types of primitives in each symbol.
+The table \ref{tab:FeatureTable} gives details on all the features used in the system.
+\begin{longtable}{|l|l|p{6cm}|p{5cm}|}
+	%\caption{Feature Table}
+	%\label{tab:FeatureTable}
+\hline 
+Type  & Index  & Features  & Notes \tabularnewline
+\endhead
+\hline 
+Type  & Index  & Features  & Notes \tabularnewline
+\endfirsthead
+\hline 
+Structural & \multicolumn{1}{c||}{} &  & \tabularnewline
+\hline 
+ & F1.  & No. of parallel lines  & \tabularnewline
+\hline 
+ & F2.  & No. of perpendicular lines  & \tabularnewline
+\hline 
+ & F3.  & No. of intersection between of any two segmetns  & \tabularnewline
+\hline 
+ & F4.  & No of primitives & \tabularnewline
+\hline 
+ & F5.  & No. of line segments & \tabularnewline
+\hline 
+ & F6  & No. of ellipse, circular arcs.  & \tabularnewline
+\hline 
+Statistical &  &  & \tabularnewline
+\hline 
+ & F7.  &  & \tabularnewline
+\hline 
+ & F8.  &  & \tabularnewline
+\hline 
+Rubine Features  &  & From features F9. till F13  & Introduced by in \cite{gestureexample12} \tabularnewline
+\hline 
+ & F9.  & Cosine of starting angle  & \tabularnewline
+\hline 
+ & F10.  & Sine of starting angle & \tabularnewline
+\hline 
+ & F11.  & Length of diagonal of bounding box & (gives an idea of the size of the bounding box) \tabularnewline
+\hline 
+ & F12.  & Angle of diagonal & gives an idea of the shape of the bounding box (long, tall, square) \tabularnewline
+\hline 
+ & F13.  & Distance from start to end  & \tabularnewline
+\hline 
+ & F14.  & Cosine of ending angle  & \tabularnewline
+\hline 
+ & F15.  & Sine of ending angle  & \tabularnewline
+\hline 
+ & F16.  & Total stroke length  & \tabularnewline
+\hline 
+ & F17  & Change in Rotation( Arctan ) & gives the directional angle \tabularnewline
+\hline 
+ & F18  & Absolute rotation  & \tabularnewline
+\hline 
+ & F19  & Rotation squared  & \tabularnewline
+\hline 
+ & F20  & The maximum speed reached (squared)  & \tabularnewline
+\hline 
+ & F21  & Total time of stroke  & \tabularnewline
+%\newpage
+\hline 
+Zernike moments  &  &  & \tabularnewline
+\hline 
+ & F22  & Zernike moments to the order  $n$  & used by \cite{zernike61} \tabularnewline
+Composite  &  &  & \tabularnewline
+\hline 
+ &  &  & \tabularnewline
+\hline 
+ &  &  & \tabularnewline
+\hline 
+ &  &  & \tabularnewline
+\hline
+\end{longtable}
+%\end{table}
+  % Rubine el al.\cite{gestureexample12}  used a set of statistical features,  There are a lot of features used in the sketch recognition literature
+%How to handle multi-stroked symbols. 
+%The clustering algorithm.  
+%The classifier used in recognition
+%Describe the features used in recognition [ref to \cite{zernike61}]
+
+ Section \ref{sec:recognitionAlgorithms} presents a comparative result of the different feature sets from these features.
+
+\section{Classification}%%Feature Set
+\label{sec:Classification}
+ The system uses support vector machine (SVM) classifier with Gaussian kernel (RBF kernel) \cite{libsvm}.
+The system uses OVO classifiers structure (object versus object).
+A validation algorithm is used to choose the Kernel variables parameters $c, \gamma$.
+The Rubine algorithm (Linear Gaussian classifier) was implemented for comparing the system.
+The results Section \ref{sec:recognitionAlgorithms} demonstrates the results of both classifiers.
+\\%Some features like number of parallel, perpendicular and intersecting lines are computed from the segmentation other like density, centroid and Zernike moments are computed from the raw stroke points.
 

### 2009-03-03 trunk/draft/ChapterIntroduction.tex
After the finish of chapter 1,,3,4

--- 2008-09-02 trunk/draft/ChapterIntroduction.tex
@@
 Computer research nowadays aimed at facilitating computer human interaction in every way.
-Pen-based interface gives the user a pencil-paper like feeling that enhances interaction more than the current used keyboard-mouse computer interfaces.
+Pen-based interfaces give the user a pencil-paper like feeling that enhances interaction more than the current used keyboard-mouse computer interfaces.
 However, until now there are hardly any complete pen based computer systems.
 Generally, Current Pen-Based systems make use of the Pen or stylus to perform the same role as the mouse / keyboard.
-The minority of such systems are equipped with handwriting or digit recognition that is typically restricted for specific language.
+A minority of such systems are equipped with handwriting or digit recognition that is typically restricted for specific language.
 
- Despite the fact that people frequently express their ideas through drawing them there isn't a system that can fully understand those sketches.
-Lately, only a few researchers were concerned with the problem although it is highly important.
-Firstly, we need to know;
- \textit { What are the main goals of Sketch understanding systems?}
+\section{Sketch Recognition}
+Sketch recognition is defined as the process of identifying symbols that users draw using single or multiple strokes.
+Users draw strokes using a pen and the system should immediately interprets their strokes into objects that can be easily manipulated.
 
-The main goal of sketch understating systems is to build a system that will interpret the sketches the user draw.
-The systems ought to be natural as possible to provide user with paper and pencil impression.
-Furthermore, interpreted sketches are expected to transform into objects, which can me manipulated, modeled and/or simulated.
-
-Sketch here stand for any hand drawn drawing the user draws using digital stylus or similar device.
-Normally, sketch are composed of smaller segments we call strokes.
+Sketches here stand for any hand drawn drawing the user draws using digital stylus or similar device.
+Normally, sketches are composed of a set of strokes.
 The time between pen-up and pen-down events identify each stroke.
 The strokes are mainly the path of points, which the user path through between the pen-down and pen-up events while using the stylus or similar device.
-
-Sketch recognition is procedure that will input a sequence of points from user then segment them into clusters of strokes and recognize those clusters to primitive geometric objects and shapes.
-However, Sketch understanding is transforming the sketch from a group of primitive shapes into comprehendible identified symbols consistent with each other.
-% the entire sketched symbols.
 
@@
 
-People normally sketch thoughts rather than expressing them linguistically.
-For instance engineers mostly drafts their early designs using paper and pencil before using CAD tools.
-People in general and engineers specially make use of various sketches during design process as instance organizational charts, informal drafting and formal design sketches \cite{}.
-The sketching process itself go thorough numerous stages.
+Scientists generally and engineers specifically express thoughts and designs using sketches.
+Engineers use sketches to exchange designs as a natural method of communication rather than writing or speaking.
+Engineers use paper and pencil design in early stages of design.
+The sketching process itself go through  numerous stages.
 Firstly, early stages with rough drafting designs or notations which are mainly done using paper and pencil.
@@
  Paper and Pencil are preferred in design as they offer a remarkable ease in editing and altering the primary rough sketches.
-Current CAD systems don't give the designer the ease of altering the early designs.
-Designer had to sketch primary designs on paper and pencil next, the finally design is transferred into computer using a menu CAD system.
+Current CAD systems do not give the designer the ease of altering the early designs.
+Designer had to sketch primary designs on paper and pencil next, the final design is transferred into computer using a CAD system menus.
 The process is time consuming and tiring.
@@
 
- A sketch recognition and understanding system will provide a powerful tool to the designers to draw sketch as they usually draw using paper and pencil with the advantages of computer program.
+ A sketch recognition and understanding system will provide a powerful tool to the designers to draw sketch as they usually draw using paper and pencil with the advantages of a computer program.
 Designer needs a normal sketching environment that will provide him facilities in paper interface and the modifying and simulation that is in current CAD systems.
 
-\section{Problems and challenges}
+\section{Applicability }
+% talk about application like cad systems, software engineering uml diagrams ....
+% meeting idea brainstorming , mind mapping string ..
+% communication program  
+% sketching programs ...
+Applications of sketch recognition are ranging from engineering device simulation and design.
+Drawing and sketching programs are most likely to benefit from integrating a sketch recognition system into their design environments.
+Other applications include object retrieval, map navigations and various engineering design systems.
+
+\section{Problem Definition and challenges}
 \label{sec:ProblemsAndChallenges}
@@
  Understanding hand drawn sketches may be trivial for humans but it is challenging problem when speaking about computer-based system.
-Most of the research made on the sketch understanding is achieved for one particular domain or for strictly a set of pre defined symbols.
+Most of the research made on the sketch recognition is achieved for one particular domain or for strictly a set of pre defined symbols.
 Further, others restrict the user the freedom in drawing which loose the main goals of paper-like-interface systems.
@@
 
-The research in the area had faced few problems that will need to be handled to implement an entirely sketch understanding system.
+Sketch recognition is divided into three steps preprocessing, segmentation and symbol recognition.
+The preprocessing phase captures user input stroke points and collects basic information about the stroke then proceeds to remove noise and compute basic statistical and geometrical information.
+In segmentation phase, the strokes are divided into a set of simple geometrical primitives or segments.
+In the third phase, sketch recognition, strokes and segments are clustered to formulate symbols that can be recognized by a classifier system.
+
+The research in the area had faced few problems that will need to be handled to implement an entirely sketch recognition system.
 The user draws strokes in a spontaneous and ambiguous manner that intensifies the challenges.
-The first problem is segmenting strokes from the way the user draw it into primitive geometric.
-Ink parsing is the problem dealing with clustering interpolated strokes and grouping them into symbols.
-After that, recognizing the symbols from the space the system can identify is the final problem.
-All those problems must be solved on a real time constrain to maintain an online sketching system if it was a requirement.
-
-The next sections will explain and investigate each of those problems in details.
-
-\subsection{Segmentations}
-\label{sec:Segmentations}
-
+One of the problems is segmenting strokes from the way the user draw it into geometrical primitive segments.
+Another major problem is the high level of ambiguities in the drawing.
+The next sections describe each problem in details.
+\subsection {Segmentation}
 Segmentation is dividing the strokes draws into the geometric primitive's as lines, arcs, and ellipses.
 The strokes as mentioned before are the path of points between pen-down and pen-up events while using a stylus.
-The processing of those points is done after that to firstly to detect the critical points or vertices.
-After that the segment is classified as either line, arc, ellipse or circle.
-The main problem in this process is in detecting the critical and vertices which are points divide strokes into lines and arc.
-The user draw the lines and arc in an undefined manner which make it difficult for the program to deicide where the user had finish a line and started a new arc or vice verse; Fig [\ref {fig:strokeinterpolation}].
-
+After that, those points are processed to detect the critical points or vertices which divide the stroke into segments.
+Later, each segment is classified as line, arc, ellipse or circle.
+The main problem in this process is in detecting the critical and vertices which divide strokes into lines and arc.
+The user draws the lines and arc in an undefined manner which make it difficult for the program to decide where the user had finish a line and started a new arc or vice verse; Fig [\ref {fig:strokeinterpolation}].
 \begin{figure}
 
+\begin{center}
 		\includegraphics[scale=0.8]{../../neededfiles/Figures/strokeinterpolation.eps}
@@
 	\label{fig:strokeinterpolation}
+\end{center}
 \end{figure}
 
-\subsection{Ink parsing}
-\label{sec:InkParsing}
+\subsection{Ambiguity}
+The user draws strokes that the computer cannot decide if it is arc or line.
+The nature of hand drawn sketch itself is highly ambiguous even for human eye you may see two drawn you cannot decide if it is square or rectangle.
 
- The clustering of segmented strokes into group of strokes before passing into the next stage this process is called ink parsing.
-It is simply deciding when a symbols ends and a new symbol begins.
-Ink parsing is most important and difficult challenge as grouping of strokes can be done in many ways and configuration, which increase exponentially with the number of strokes.
-In addition, the ink parsing will affect the next steps of recognizing.
-The user does not draw the symbol in same order.
-Moreover, users can draw a sequence of strokes in symbols then leave it to draw another then return again to add segment to symbols so temporal data is helpful but not conclusive.
-The user may also draw the same symbols using different orders and methods for example the user can draw a rectangle using single stroke or using two, three or even four strokes.
-All those factors gather with the sloppy drawing of the user to increase the complexity of the problem.
-
-\subsection{Ambiguities}
-\label{sec:Ambiguities}
-
-The major problem of sketch understanding is the high level of ambiguities in the drawing.
-The user draws strokes that the computer cannot decide if it is arc or line.
-The nature of hand drawn sketch it self is highly ambiguous even for human eye you may see two drawn you cannot decide if it is square or rectangle.
  The ambiguities in sketch understanding are found at many levels in the system.
-Firstly, at the segmentation phase where the same user stroke can be interpreted as a ellipse and a circle.
-Then the symbol recognition as a symbol can have more than one interpolation.
-Due to the user sloppy drawing the symbols can have more than one interpolation which varies with the domain drawn in it similar problem was faced in handwriting recognition see Fig [\ref{fig:VorU}] but it was handled with knowing information about the context information which may not be the case in sketch recognition.
+Firstly, in the segmentation phase, a single stroke can be interpreted as an ellipse or as a circle.
+Symbol ambiguity is another problem as a symbol can have more than one interpolation.
+Due to the user sloppy drawing the symbols can have more than one interpolation which varies with the domain drawn in it similar problem was faced in handwriting recognition (Figure [\ref{fig:VorU}]).
+The problem could be solved by providing context information which could not be provided in most sketches.
 
@@
 		\includegraphics[scale=0.85]{../../neededfiles/Figures/VorU.eps}
-	\caption[Ambiguities Handwriting ]{The figure shows that the same letter can be interpreted differently in various contexts; first as V in drive or as U in Cut}
+	\caption[Ambiguities Handwriting] { The same letter can be interpreted differently in various contexts; first as V in drive or as U in Cut}
 \label{fig:VorU}
@@
 
-The ambiguity problem must be first recognized ambiguities or varicose interpolations of same symbol.
-After that, resolve those ambiguities is detecting the best correct interpolation of the symbol.
-To solve this problem the program must be able to detect the amount of information needed to decide which interpolation of symbols is the best one.
+\section{Research Scope}
+%rewirte this part.............................
+%  research will investigated till sketch recognition in domain of uml diagrams. 
+%this research will include investigating the best method to recognize sketches.
+The scope of the research will include recognizing various sketches.
+The methods used in this research are comparing current systems and enhancing them and adopting them into simple geometrical symbols.
+Using well established AI methods specially \textit{Swarm Intelligence} to solve the problem.
+ This research investigates the best method to recognize sketches.
+The scope of the research includes recognizing various symbols drawn using single and multiple strokes.
+The research compares various segmentation methods.
+This research used two Particle Swarm Optimization (PSO) \cite{PSOFirst} methods to segments strokes into geometrical primitives.
+The research also investigates various spatial, geometrical and statistical features set used in sketch recognition.
+Support Vector Machines (SVM) classifier\cite{svmintroduce} is used to classify  the sketch  into one of symbols from the set of previously trained symbols.
+This research assumes the user draws one symbols at a time. The user can draw can draw this symbol in one stroke or in multiple stroke without any restriction on order, orientation or style.
 
-\subsection{Search space}
-\label{sec:Searchspace}
-
-The user drawn symbols can be having more than one interpolation.
-In addition, those interpolations increase exponentially with the number of strokes the user draw and as the sketch expands the number of interpolation escalate.
-This intensifies the search space that the program search to resolve the ambiguities and find best interpolation.
-
-\subsection{Real time}
-\label{sec:Realtime}
-
-There are mainly three types of sketch understanding system offline \cite{Vibratory8}, online and interactive.
-Both online and interactive recognition like \cite{incrmentintention41, SmartSketch56} need the processing after each stroke to be real time so the user does not feel the processing.
+\section{Thesis Overview}
+This thesis is arranged such that chapter 2 presents a survey of different components of the sketch recognition.
+Chapter 3 describes the proposed system and its main building blocks.
+The details description of each component is presented in chapter 4.
+The comparative study of the different techniques and their detailed results in various experiments are detailed in chapter 5.
+Finally, chapter 6 presents the conclusions of this thesis and suggested future work.
+%This thesis is arranged such that chapter \ref{} presents a survey of different components of the sketch recognition.
+Chapter \ref{} describe the proposed system and its main building blocks.
+The details description of each component is presented in Chapter \ref{}.
+The comparative study of different techniques with the detailed results of various experiments are detailed in chapter\ref{}.
+Finally, chapter \ref{} presents the conclusions of this thesis and suggested future work.

### 2009-03-03 trunk/draft/SketchMaster.tex
After the finish of chapter 1,,3,4

@@
 \usepackage{psfig}
-
+\usepackage{named}
 \usepackage{natbib}
@@
                                                         %======================
-\title{Sketch recogniton and understanding}             % My theses title
+\title{Sketch Recogniton using Particle Swarm Algorithm}             % My theses title
 \author{Maha Mohammed Nabeel El Meseery}                               % Y name
-\prevdegrees{B.S.,Computer  engineering(2003)}       		% Your old degree
+\prevdegrees{B.S.,Computer  Engineering(2009)}       		% Your old degree
 \major{Computer Engineering}                            % My current major
 \degree{Master of Science}                              % My new degree
-\degreemonth{Dec}                                      % When awarded. to be added when known
-\degreeyear{2005}                                      %
+\degreemonth{}                                      % When awarded. to be added when known
+\degreeyear{}                                      %
 %th today or first day of writing  December 26 2005 
-\thesisdate[]                                     % Date of document.
+%\thesisdate[]                                     % Date of document.
 \DOCUMENTtype{Survey}                                   % or DISSERTATION
@@
 \maketitle
+
+\providecommand{\tabularnewline}{\\}
                                                         % COPYRIGHT PAGE
                                                         %======================
-\copyrightyear{2005}                                    % Delete these
+\copyrightyear{2009}                                    % Delete these
 \makecopyright                                          % if no copyright
-                                                        % page.
-
-                                                        % APPROVAL PAGE
-                                                        %======================
+                                                    % APPROVAL                                                         %======================
 %\supervisor{A. B. Smith}{Associate Professor of xxx}    % <- One of these
@@
 \include{ChapterIntroduction}                              % The best way to organize your
-\include{chapterSurvey}                              % document is probably to keep
+%\include{chapterSurvey}                              % document is probably to keep
+\include{System}
+\include{Results}
+%\include{ConFut}
                                                 % in its own file.  

### 2009-03-03 paper2/Paper2.tex
paper 2

\documentclass{article}
\usepackage{spconf}
\usepackage{amsmath}
\usepackage[dvips,pdftex]{graphicx}
\usepackage{subfigure}
\usepackage{named}
\usepackage[dvips,pdftex]{hyperref}
\title{Sketch Recognition using Particle Swarm Algorithms}
\twoauthors
  {Maha El Meseery, Mahmoud Fakhr El Din\sthanks{Associate Professor}, Samia Mashali\sthanks{Professor}}
	{Signals Processing Group \\ Computers and Systems Department \\
  Electronic Research Institute\\Cairo, Egypt\\
	melmeseery@eri.sci.eg, samia@eri.sci.eg}
  {Magda Fayek\sthanks{Associate Professor}, Nevin Darwish\sthanks{Professor}}
	{ Computer Engineering Department\\
	Faculty of Engineering \\Cairo University\\Cairo, Egypt\\
	magdafayek@gmail.com,ndarwish@ieee.org}
\begin{document}
\maketitle
\begin{abstract}
Sketch recognition is defined as the process of identifying symbols that users draw using single or multiple strokes. Users draw strokes using a pen and the system immediately interprets their strokes as objects that can be easily manipulated. This paper uses Particle Swarm Algorithm (PSO) to divide the strokes the user draws into meaningful geometric primitives. These geometric primitives are grouped to formulate symbols which are further identified. The results show that using PSO improves segmentation results which guide the symbol recognition phase. This paper uses Support Vector Machines (SVM) classifier which further improves the final recognition accuracy.  
\end{abstract}
\section{Introduction}
The field of symbol recognition has gained interest in the last few years. Scientists generally and engineers specifically express thoughts and designs using sketches. Engineers use sketches to exchange designs as a natural method of communication rather than writing or speaking. Design engineers need a powerful computer based system with symbol recognition to help them design, manipulate and store sketches more effectively than using only papers. Increasing the interaction between computers and users in sketch and CAD systems has been the reason for the emerging of a few advanced sketch recognition systems.

Sketch recognition is divided into three main steps preprocessing, segmentation and symbol recognition. The preprocessing phase captures user input stroke points and collects basic information about the stroke then proceeds to remove noise and compute basic statistical and geometrical information. In segmentation phase, the strokes are divided into a set of simple geometrical primitives or segments. In the third phase, sketch recognition, strokes and segments are clustered to formulate symbols that can be recognized by a classifier system.

A wide variety of techniques were suggested for sketch recognition.  A hybrid algorithm was introduced in \cite{earlyprocess} where they generated different sets of segments based on both curvature and speed dominant points, followed by choosing a segmentation with the least error from a generated hybrid set. Their system is however limited to recognizing only specific simple geometric shapes with a set of low level recognizers. Each low level recognizer is designed to recognize only one geometric shape using spatial and geometric information extracted from the stroke.  Yu \cite{meanshift10}introduced a feature area for each primitive and then computed the segmentation error for different types of primitives based on the computed feature area. His system achieved good accuracy in simple shapes (square, ellipse,...) but did not perform well in more complex shapes. Paulson and Hammond\cite{Paleosketch08} introduced a set of low level recognizers that were reported to achieve 98\%  accuracy but their system, similar to all low level recognizers, only identify a small set of simple shapes. A genetic algorithm was used by \cite{CruveDivisionSwarm} to optimally divide digital curves into lines and curves. Chen et al.\cite{CruveDivisionSwarm} uses digital curves scanned from paper as input to the system and did not take advantage of the curvature or local geometric properties of the digital curve.

This paper introduces a new sketch recognition system, which uses particle swarm algorithm (PSO) to optimally segment user's strokes. Users can draw symbols using any number of strokes in any order they like. The system segments each stroke using two different PSO algorithms. Segmentation is based on dividing the digital curve into polygon\cite{PolygonApproximationPSO}. The algorithm enhances its performance by adding curvature and other local information while segmenting strokes to achieve better results. To handle complex strokes, our system uses a second segmentation PSO algorithm to segment strokes into a set of lines and circular arcs. After segmentation, a symbol features vector composed of different statistical, geometrical and spatial features is deduced. The final classification step, SVM classifier uses the computed feature vector to classify the segments into one of the previously trained classes.

 The remaining of the paper is as following; Section \ref{Sysdisc} describes the system in details.  Sections \ref{Prepross}, \ref{seg} and \ref{sec:Recognition} provide details on preprocessing, segmentation and recognition process respectively. Section \ref{sec:Experiments} describes the experiments that were preformed. We conclude in Section \ref{ConclusionandFutureWork}.
\section{The proposed system}
\label{Sysdisc}
 The suggested sketch recognitions system is divided into three main steps 1) preprocessing, 2) segmentation and 3) Recognition. The following section provides a detailed description of each step.
\subsection{Preprocessing}
\label{Prepross}
 Speed and time difference information were widely used in sketch understanding systems \cite{earlyprocess}. It was noted that time difference provides more distinct maxima than the speed information. Agar \cite{polygonfeedback31} mentions that the pointing device (for example the mouse) sampling rate is the reason for this phenomena. 
 In the presented system, we computed time difference, direction, speed and curvature of each point along a stroke. The speed is calculated as $v=\Delta s/\Delta t$ where $t$ is the time difference between two points and $s$ is the length between them. The direction $d$ is calculated as the angle between two vectors and curvature is considered as the change in direction with respect to length i.e. $c= \Delta d/\Delta s$.  Dominant points are characterized by lower speed values and higher curvature and direction values.

After the system computes all the speed, time difference and curvature information it proceeds to detect the points with low velocity and high curvatures. Using simple differentiation to detect local extreme points resulted in false points due to the non smooth curves. Hence, the system adopted a process presented by \cite{earlyprocess}, where the mean of the curve is calculated. Then a threshold \textit{th} is used to separate the curve into regions $Region_i$; each region $Region_i$ is defined as a range of points, where the curve values are either above or below the threshold \textit{th}. Those regions are further processed to find the maximum point $Max(Region_i)$ of each region $Region_i$. The stroke points $p_i(x,y)$ that correspond to those maximum values are labeled as \textit{possible dominant points} $P_{pd}$. The system repeats this process for curvature, time difference and speed information. All the points labeled as possible dominant points $P_{pd}$ are saved into a single array. Figure \ref{fig:LabelsPPD} shows the particles labeled as Possible dominate point $P_{pd}$ by the preprocessing, it is noticed the redundancy of some $P_{pd}$ points. % (as shown are redundant)
\begin{figure}
	\centering
		\subfigure[Posible domninate point]{ 	\label{fig:LabelsPPD}	\includegraphics[scale=0.5]{images/ppd.jpg}}
		\hfill
			\subfigure[DPSO algorithm encoding]{	\label{fig:LabelsPSO} \includegraphics[scale=0.5]{images/pso1.jpg}}			
	\caption{Segmentation}% a) Possible dominate point b) Particle encoding  
	\label{fig:pso1}
\end{figure}
\subsection{Segmentation}
\label{seg}
In the segmentation stage the stroke is divided into a set of primitives. As shown in Figure \ref{fig:segblock} first an attempt is made to fit the stroke points into a curve or an ellipse using a minimum square error fitting algorithm \cite{chernov-2003}. If the stroke proved to be an ellipse arc then the segmentation process ends and the system proceeds to the next step. Otherwise, the stroke is passed to two particle swarm algorithms that divide the stroke to either lines or lines and curves. The algorithms takes the stroke points along with the possible dominant points $P_{pd}$ computed during preprocessing then produce a set of dominant points which are connected by either lines or curves. The next section describes the ellipse detection algorithm and the two particle swarm algorithms used to divide the stroke.
 \begin{figure}
	\centering
		\includegraphics[scale=0.55]{images/segblock.jpg}
	\caption{Segmentation Block}
	\label{fig:segblock}
\end{figure}
\subsubsection{Ellipse Detection} 
The process starts by computing the center of the stroke bounding box. The bounding box center point is used as the first estimation of the ellipse center. The axes of the ellipse are estimated as $width/2$ and $height/2$ of the stroke bounding box. The least square fitting algorithm is used to minimize the fitting error of the ellipse equation \cite{chernov-2003}.  
\subsubsection{Discrete particle swarm algorithm (\textit{DPSO})}
\label{subsubsec:Discreteparticleswarmalgorithm}
Two DPSO algorithms are used to generate segmentations for each stroke the user draws. For each generated segmentation evaluated an error is determined. The segmentation with the minimum error value is chosen as the best stroke segmentation (Figure \ref{fig:segblock}). The problem definition is the same in both algorithms but they differ in the method they use to compute fitness and error functions.

\textsl{Problem definition:} The input stroke with $N$ points can be represented by set $S = \left\{ {x_1 ,x_2  \ldots x_N }\right\}$ where $x_i$ is the location of the point $i$. The swarm algorithms consist of $M$ agents which are represented by the set  $A = \left\{ {P_i \left| {i = 1,2 \cdots M} \right.} \right\}$ where $P_i$ is a single solution particle from the solution space. Each particle decodes the problem using a binary array with the same length $N$ as the input stroke.

Therefore, the system represents each particle $P_i$ by $P_i = \left\{ {p_{ij} \left| {j = 1,2 \cdots N} \right.} \right\}$ where $p_{ij}$ has only two values 0 or 1. $p_{ij}=1$ means that point $j$ is a dominant point. Figure\ref{fig:LabelsPSO} shows a particle encoding in the DPSO system.

\textsl{Fitness function:} The fitness function and error calculation are different in each of the two \textit{DPSO} algorithms.

For the \textit{first algorithm \textsl{AlgS1}}, 
 the arc $\widehat{x_ix_j}$ is defined as the consecutive set of points from point$x_i$ to point $x_{j}$ as in $x_i,x_{i+1} \cdots,x_j$. The line $\overline{x_i x_j} $ as the straight line connecting point $x_i$ to point $x_j$. The approximation error is computed by the equation \ref{eq:ErrorSwarm1}. The graphical meaning of the error is shown in Figure\ref{fig:LabelsPSO}.
\begin{equation}
E=\sum\nolimits_{i = 0}^M e ( \widehat{x_ix_{i+1}},\overline{x_i x_{i+1}})
\label{eq:ErrorSwarm1}
\end{equation} where $M$ is the number of dominant points in this solution as generated by the swarm algorithm. The error $e ( \widehat{x_ix_j},\overline{x_i x_j})$ is computed as the sum of squared perpendicular distance from every point along the arc $\widehat{x_ix_j}$ to the line $\overline{x_i x_j}$. The fitness is computed using equation \ref{eq:fitnessSwarm1}\begin{equation}
\max fitness(p_i ) = \left\{ {\begin{array}{*{20}c}
   { - E/\varepsilon N} & {ifE > \varepsilon ,}  \\
   {D/\sum\limits_{j = 1}^N {p_{ij} } } & {otherwise}  \\
\end{array}} \right.
\label{eq:fitnessSwarm1}
\end{equation} where $N$ is the number of points in the stroke, $D$ is the number of points in the solution that was previously labeled as a possible dominant point ($P_{pd}$), $E$ is the computed error and $\varepsilon$ is the error threshold. It should be noticed that when the error is larger than the threshold $\varepsilon$ the fitness is given a -ve value to lower the fitness value of the solution. Otherwise the system favors the lower number of vertices.

The \textit{second algorithm \textsl{AlgS2}} has the same problem formulation but different fitness and error functions are used. An attempt is made to fit each segment into line or circular arc, the error of both circle and line fit estimations are computed for each segment $S_i=\widehat{x_ix_j}$, the approximation with the lower error value is the chosen approximation of this segment $S_i$\cite{CruveDivisionSwarm}. The sum of the approximation error of all segments is the total error of the particle.  The total error of the particle is computed by equation \ref{eq:errorSwarm2}
 \begin{equation}
E=\sum\nolimits_{i = 0}^M e(D_i) 
\label{eq:errorSwarm2}
\end{equation}where $M$ is the number of segments in the solution as generated by the swarm algorithm, $D_i$ is the minimum approximation error of curve and line approximations $min(d_c,d_l)$ as computed by \cite{CruveDivisionSwarm}.  The fitness is computed by the equation \ref{eq:fitnessSwarm2} \begin{equation}
\max fitness(P_i ) = \frac{1}{{E \times M^k }}
\label{eq:fitnessSwarm2}
\end{equation} where $E$ is the error and $M$ is the number of segments and $k$ is a parameter tweaked to get minimum number of segments. The larger k is, the more effect the number of segments will have. For our system, k is selected to be 0.5\cite{CruveDivisionSwarm}.

After each loop of the swarm algorithm (\textsl{AlgS1} and \textsl{AlgS2}), each particle is refined using the following procedure. For each particle $P_i$ each dominant point $P_{ij}$ is checked to find if it was labeled before as a \textit{possible dominant point} $P_{pd}$ (computed as in section \ref{Prepross}). If it was not labeled the point $P_{ij}$ is moved to the nearest labeled point. This ensures that all of the points generated by the DPSO are possible dominate points $P_{pd}$. After that the particles are tested to make sure that the distance between every two successive dominate point is larger than the constant $min_D$. If two points are nearer than $min_D$ then one of the points is removed. 
\subsection{Recognition}
\label{sec:Recognition}
After the user draws all strokes of the symbol, the set of un-recognized strokes is grouped together along with their segmentation as input to the feature extraction process. A composite set of features is used to generate a single feature vector. The features used consist of Rubine feature set,  Zernike moments of order 21 \cite{HeloiseBeautification}, ink density as well as some structural and spatial information like number of perpendiculars lines ,number of parallel lines and number of different types of primitives in each symbol. After computing the features the symbol is introduced to the (SVM) classifier. 
\section{Experimental}
\label{sec:Experiments}
The dataset collected by Hse and Newton\cite{HeloiseBeautification} is used to test the presented system. The data are drawn by 16 users each of them was asked to draw the 13 shapes from 30 to 50 times. Figure \ref{fig:symbolSet} shows a set of the shapes used in the data set. The dataset was divided into training set and test set. To generate trusted results, four different splits for the training and test data are generated from the dataset. The results displayed are the recognition accuracy resulted from the classifier after recognizing the symbols. The accuracy is computed as the number of correctly recognized samples divided by the total number of samples in the test. %In the following section, the recognition accuracy means the average recognition accuracy in the four data splits.  
 \begin{figure*}[]
	\centering
	\begin{minipage}[b]{5cm}	
\begin{center}
\subfigure [The Symbol Set] {\label{fig:symbolSet}
\fbox{ \parbox{2cm}{% 
		\includegraphics[scale=0.2]{images/symbolSet.PNG}	}}}

		\subfigure[Algorithm comparison]{\includegraphics[width=5cm]{images/test1smallpaper.jpg} 	\label{fig:test1}}
		\end{center}
		\end{minipage}
		\hfill
  \subfigure [Symbols Comparison] {\includegraphics[width=9cm]{images/test2smallpaper.jpg} \label{fig:test2}}
	\caption{Experiments results :} a) the figure shows a set of the symbols used in the dataset.   b)The results of comparing the algorithms. The graph displays the recognition rate of symbols. c) The graph shows the recognition rate of each symbol using different algorithms. 
\end{figure*} 
We performed two experiments to test the system. Firstly we tested recognition accuracy of shapes in the data set with both algorithms (\textsl{AlgS1} and \textsl{AlgS2}). Figure \ref{fig:test1} shows the accuracy achieved by each algorithm. The two swarm algorithms were tested with and without the ellipse fitting module. Including the ellipse detection module improves the results with both the \textit{DPSO} algorithms.% The results show that both \textit{DPSO} algorithms achieve better result than other algorithms.

The second experiment we implemented was to test the effect of symbol complexity and type on the recognition rate. Figure \ref{fig:test2} shows the accuracy of each symbol, it is clearly noted that symbols that have only line segments achieve higher accuracy rate than other symbols. Figure \ref{fig:test2} also indicates that algorithm \textsl{AlgS1} alone achieve better performance than algorithm   \textsl{AlgS2} in the symbols that consist only of lines. This is natural as algorithm \textsl{AlgS1} divides strokes into line segments only but \textsl{AlgS2} is able to divide strokes into lines and curves based on the minimum error of the segment itself.  
 Combining both algorithm \textsl{AlgS1} and \textsl{AlgS2} improved the recognition rate of all symbols. The penalty for this increased performance is in the computational time needed to run both swarm algorithms. 
\section{Conclusion and Future Work}
\label{ConclusionandFutureWork}
This paper presented a new approach to sketch recognition using PSO. The system uses both speed and curvature information which help improving the \textit{DPSO}. It is noted that the \textit{DPSO} in general generates an optimized stroke segmentation which improves the final recognition rate.  The tradeoff between accuracy achieved and time complexity must be further investigated to achieve better results. The use of statistical moments and structural features improves the recognition rate. The system was tested on 13 different symbols and achieved an accuracy of 91.5\%. The system does not depend on low level recognizer but rather on a set of high level features. This makes the system easily expandable to more and more symbols.

 A possible extension of this research is to complete the clustering algorithm for fully automated sketch recognition. Currently, the system process one symbol at a time, an enchantment would be to allow users to draw more than one symbol at a time and incorporate a method for separating symbols. Other area of enhancements is the features extraction methods. Introducing more spatial and geometrical features is believed to improve classifications.  
\bibliographystyle{IEEEbib}
\bibliography{../../neededfiles/Bibliographies/Mybibliography2}
\end{document}

### 2009-03-21 paper2/Paper2.tex
sent to dr. mahmoud

@@
 \usepackage[dvips,pdftex]{hyperref}
-\title{Sketch Recognition using Particle Swarm Algorithms}
+\title{Sketch Recognition using Particle Swarm Optimizations}
 \twoauthors
@@
 Sketch recognition is defined as the process of identifying symbols that users draw using single or multiple strokes.
-Users draw strokes using a pen and the system immediately interprets their strokes as objects that can be easily manipulated.
-This paper uses Particle Swarm Algorithm (PSO) to divide the strokes the user draws into meaningful geometric primitives.
+Usually, users draw strokes using pens.
+The sketch recognition system immediately interprets their strokes as objects that can be easily manipulated.
+This paper uses Particle Swarm Optimization (PSO) to divide the strokes the user draws into meaningful geometric primitives.
 These geometric primitives are grouped to formulate symbols which are further identified.
 The results show that using PSO improves segmentation results which guide the symbol recognition phase.
-This paper uses Support Vector Machines (SVM) classifier which further improves the final recognition accuracy.
+As for recognition we use Support Vector Machines (SVM) classifier which further improves the final recognition accuracy.
 \end{abstract}
@@
 Engineers use sketches to exchange designs as a natural method of communication rather than writing or speaking.
-Design engineers need a powerful computer based system with symbol recognition to help them design, manipulate and store sketches more effectively than using only papers.
-Increasing the interaction between computers and users in sketch and CAD systems has been the reason for the emerging of a few advanced sketch recognition systems.
+Design engineers need a powerful computer based system with symbol recognition capabilities to help them design, manipulate and store sketches more effectively than using only papers.
+Increasing the interaction between computers and users in sketch and CAD systems has been the main reason for the emerging of a few advanced sketch recognition systems.
 
-Sketch recognition is divided into three main steps preprocessing, segmentation and symbol recognition.
-The preprocessing phase captures user input stroke points and collects basic information about the stroke then proceeds to remove noise and compute basic statistical and geometrical information.
-In segmentation phase, the strokes are divided into a set of simple geometrical primitives or segments.
-In the third phase, sketch recognition, strokes and segments are clustered to formulate symbols that can be recognized by a classifier system.
+Sketch recognition is divided into three main steps a) preprocessing, b) segmentation and c) symbol recognition.
+The preprocessing phase captures user input stroke points and collects basic information about the stroke.
+This is followed by proceeding to remove noise and computing basic statistical and geometrical information.
+In the segmentation phase, the strokes are divided into a set of simple geometrical primitives or segments.
+In the third phase, sketch recognition, strokes and segments are clustered to formulate symbols that can be recognized by a classifier.
 
-A wide variety of techniques were suggested for sketch recognition.
-A hybrid algorithm was introduced in \cite{earlyprocess} where they generated different sets of segments based on both curvature and speed dominant points, followed by choosing a segmentation with the least error from a generated hybrid set.
-Their system is however limited to recognizing only specific simple geometric shapes with a set of low level recognizers.
-Each low level recognizer is designed to recognize only one geometric shape using spatial and geometric information extracted from the stroke.
-Yu \cite{meanshift10}introduced a feature area for each primitive and then computed the segmentation error for different types of primitives based on the computed feature area.
-His system achieved good accuracy in simple shapes (square, ellipse,...) but did not perform well in more complex shapes.
-Paulson and Hammond\cite{Paleosketch08} introduced a set of low level recognizers that were reported to achieve 98\%  accuracy but their system, similar to all low level recognizers, only identify a small set of simple shapes.
-A genetic algorithm was used by \cite{CruveDivisionSwarm} to optimally divide digital curves into lines and curves.
-Chen et al.\cite{CruveDivisionSwarm} uses digital curves scanned from paper as input to the system and did not take advantage of the curvature or local geometric properties of the digital curve.
-
-This paper introduces a new sketch recognition system, which uses particle swarm algorithm (PSO) to optimally segment user's strokes.
+This paper introduces a new sketch recognition system, which uses particle swarm optimization (PSO) to optimally segment user's strokes.
 Users can draw symbols using any number of strokes in any order they like.
 The system segments each stroke using two different PSO algorithms.
-Segmentation is based on dividing the digital curve into polygon\cite{PolygonApproximationPSO}.
-The algorithm enhances its performance by adding curvature and other local information while segmenting strokes to achieve better results.
+The first segmentation algorithm is based on dividing the digital curve into polygons\cite{PolygonApproximationPSO}.
+Our algorithm enhances its performance by adding curvature and other local information while segmenting strokes to achieve better results.
 To handle complex strokes, our system uses a second segmentation PSO algorithm to segment strokes into a set of lines and circular arcs.
-After segmentation, a symbol features vector composed of different statistical, geometrical and spatial features is deduced.
-The final classification step, SVM classifier uses the computed feature vector to classify the segments into one of the previously trained classes.
+After segmentation, a feature vector composed of different statistical, geometrical and composite features is deduced.
+The final classification step uses a SVM classifier to classify symbols into one of the previously trained classes based on the features vector computed.
+% uses the computed feature vector to classify the segments into one of the previously trained classes.
 
- The remaining of the paper is as following; Section \ref{Sysdisc} describes the system in details.
+ The remaining of the paper is as follows; the literature review is in Section \ref{sec:review}.
+Section \ref{sec:ParticleSwarmAlgorithm} introduces Particle Swarm Optimization.
+Section \ref{Sysdisc} describes the system in details.
 Sections \ref{Prepross}, \ref{seg} and \ref{sec:Recognition} provide details on preprocessing, segmentation and recognition process respectively.
@@
 We conclude in Section \ref{ConclusionandFutureWork}.
-\section{The proposed system}
+
+\section{Literature Review}
+\label{sec:review}
+A wide variety of techniques were suggested for sketch recognition.
+A hybrid algorithm was introduced in \cite{earlyprocess} where different sets of segments are generated based on both curvature and speed dominant points, followed by choosing a segmentation with the least error from a generated hybrid set.
+However, this system is limited to recognizing only specific simple geometric shapes with a set of low level recognizers.
+Each low level recognizer is designed to recognize only one geometric shape using spatial and geometric information extracted from input stroke.
+Yu \cite{meanshift10} introduced a \textit{feature area} for each primitive and then computed the segmentation error for different types of primitives based on the computed \textit{feature area}.
+His system achieved good accuracy in simple shapes (square, ellipse,...etc) but did not perform well in complex shapes.
+Paulson and Hammond \cite{Paleosketch08} introduced a set of low level recognizers that were reported to achieve 98\%   accuracy.
+However, their system similar to all low level recognizers in the way that it identifies a small set of simple shapes.% A genetic algorithm was used by \cite{CruveDivisionSwarm} to optimally divide digital curves into lines and curves.
+Chen et al.\cite{CruveDivisionSwarm} uses digital curves scanned from paper as input to the system and did not take advantage of the curvature or local geometric properties of the digital curve.
+
+A genetic algorithm was used by \cite{CruveDivisionSwarm} to optimally divide digital curves into lines and curves.
+Chen et al.
+\cite{CruveDivisionSwarm} uses digital curves scanned from paper as input to the system.
+Also, they did not take advantage of the curvature or local geometric properties of the digital curve.
+Yin \cite{PolygonApproximationPSO} used PSO to convert digital curves into polygons, our system adopts  Yin \cite{PolygonApproximationPSO} method and tries to improve it by adding curvature and other local information while segmenting strokes to achieve better segmentations.
+The next section presents the general particle swarm algorithm which is used in both our segmentation algorithms.
+% are presented in section \ref{subsubsec:Discreteparticleswarmalgorithm}
+
+\section{Discrete Particle Swarm Optimization}
+\label{sec:ParticleSwarmAlgorithm}
+%\section{Particle Swarm Algorithm}
+%\label{PSO}
+%What is particle swarm algorithm and how it was used in related researches. 
+The main idea of \textit{Particle Swarm Algorithm (PSO)} is to represent each agent with a particle from the solution space \cite{PSOFirst}.
+Each agent moves the particle with a direction and velocity $v_{ij}$ based on equations \ref{eq:Swarm} \& \ref{eq:Swarm1}.
+\begin{equation}
+%\[
+p_{ij}=p_{ij}+v_{ij},
+%\
+\label{eq:Swarm1}
+\end{equation}
+where $p_{ij}$ represent the $jth$ particle in the $ith$ agent and $v_{ij}$ is the velocity of the $jth$ particle in the $ith$ agent.
+ %Equation [\ref{eq:Swarm}] shows how velocity and direction of each particle are computed
+ \begin{equation}
+v_{ij}  = v_{ij}  + c_1 r_1 (lbest_{ij}  - p_{ij} ) + c_2 r_2 (gbest_{ij}  - p_{ij} )
+\label{eq:Swarm}
+\end{equation}
+ where $lbest_{ij}$ is the local best particle, $gbest_{ij}$ is the global best particle, $r_1$ \& $r_2$ are random variables and $c_1$ \& $c_2$ are the swarm system variables.
+ After each iteration the global best $g_{best}$ particle and the agent local best $l_{best}$ particle are evaluated based on the maximum fitness functions of all particles in the solution space.
+The solution is found after achieving a specific number of iteration or after an error threshold is achieved.
+Equation \ref{eq:descrite} is used to change the general swarm algorithm into binary particle (\textit{Discrete Particle Swarm Algorithm DPSO}) which handles particle values of either $0$ or $1$ \cite{PSODisceret}.
+\begin{equation}
+   P(i)\Leftarrow 
+\{
+\begin{array}{c} 
+1 \quad \quad if\quad r_{3}>p_{i}  \\
+
+0 \quad \quad if\quad r_{3}<p_{i} 
+\label{eq:descrite}
+\end{array}\}
+\end{equation}
+ where $p_{ij}$ is the numerical values of the particle and $r_{3}$ is a random variable.
+
+\section{The Proposed System}
 \label{Sysdisc}
- The suggested sketch recognitions system is divided into three main steps 1) preprocessing, 2) segmentation and 3) Recognition. The following section provides a detailed description of each step.
+ The suggested sketch recognition system is divided into three main steps 1) preprocessing, 2) segmentation and 3) Recognition. The following section provides a detailed description of each step.
+
 \subsection{Preprocessing}
-\label{Prepross}
- Speed and time difference information were widely used in sketch understanding systems \cite{earlyprocess}.
+\label{Prepross}% Agar \cite{polygonfeedback31} mentions that the pointing device (for example the mouse) sampling rate is the reason for this phenomena.
+ Speed and time difference information were widely used in sketch recognition systems \cite{earlyprocess}.
 It was noted that time difference provides more distinct maxima than the speed information.
-Agar \cite{polygonfeedback31} mentions that the pointing device (for example the mouse) sampling rate is the reason for this phenomena.
+Agar et al.
+\cite{polygonfeedback31} mentions that the pointing device (for example the mouse) sampling rate is the reason for this phenomena.
+The points are sampled at regular time intervals while the pointing device is moving.
+There are no points while the pointing device is stationary.
+This leads to a nearly constant time difference between samples while the pen is moving and large difference while the pen is stationary.
+Contrary to speed information, where the users draw with variant speeds.
+The speed information has a lot of noise in the data.
+Figure \ref{fig:speed2Distance} shows the time difference and the speed graphs for the stroke drawn in Figure \ref{fig:orignalStroke}.
+Similarly, direction information is used as it provides better distinctive maxima for the corners than the curvature information \cite{meanshift10}.% Figure \ref{fig:curvatures} shows the direction and curvature graphs for the stroke drawn in Figure \ref{fig:orignalStroke}
+ \begin{figure}
+	\centering
+			\subfigure[ Speed Graph]{	\includegraphics[scale=0.3]{images/vel3.png}}
+			\hfill
+			\subfigure[ Time Difference Graph] {\includegraphics[scale=0.3]{images/td3.png}}
+	\caption{Speed and Time Difference Graphs}
+	\label{fig:speed2Distance}
+\end{figure}
+
+\begin{figure}[]
+	\centering
+		\includegraphics[scale=0.5]{images/stroke3.jpg}
+	\caption{Input Stroke} Example of an input stroke to the system. 
+	\label{fig:orignalStroke}
+\end{figure}
+
  In the presented system, we computed time difference, direction, speed and curvature of each point along a stroke.
+The system computes all these information to ensure that no possible dominate point is overlooked in this phase.
 The speed is calculated as $v=\Delta s/\Delta t$ where $t$ is the time difference between two points and $s$ is the length between them.
@@
 $c= \Delta d/\Delta s$.
-Dominant points are characterized by lower speed values and higher curvature and direction values.
+Dominant points are characterized by lower speed values, higher curvature and direction values.
 
-After the system computes all the speed, time difference and curvature information it proceeds to detect the points with low velocity and high curvatures.
-Using simple differentiation to detect local extreme points resulted in false points due to the non smooth curves.
+After the system computes the speed, time difference and curvature information it proceeds to detect the points with low velocity and high curvature.
+However, using simple differentiation to detect local extreme points resulted in false points due to the non smooth curves.
 Hence, the system adopted a process presented by \cite{earlyprocess}, where the mean of the curve is calculated.
-Then a threshold \textit{th} is used to separate the curve into regions $Region_i$; each region $Region_i$ is defined as a range of points, where the curve values are either above or below the threshold \textit{th}.
+Then a threshold \textit{th} is used to separate the curve into regions ($Region_i$); each region $Region_i$ is defined as a range of points, where the curve values are either above or below the threshold \textit{th}.
 Those regions are further processed to find the maximum point $Max(Region_i)$ of each region $Region_i$.
@@
 	\centering
-		\subfigure[Posible domninate point]{ 	\label{fig:LabelsPPD}	\includegraphics[scale=0.5]{images/ppd.jpg}}
+		\subfigure[Possible dominate point]{ 	\label{fig:LabelsPPD}	\includegraphics[scale=0.7]{images/ppd.jpg}}
 		\hfill
-			\subfigure[DPSO algorithm encoding]{	\label{fig:LabelsPSO} \includegraphics[scale=0.5]{images/pso1.jpg}}			
+			\subfigure[DPSO algorithm encoding]{	\label{fig:LabelsPSO} \includegraphics[scale=0.7]{images/pso1.jpg}}			
 	\caption{Segmentation}% a) Possible dominate point b) Particle encoding  
@@
 \end{figure}
+
 \subsection{Segmentation}
 \label{seg}
-In the segmentation stage the stroke is divided into a set of primitives.
+In the segmentation stage the input stroke is divided into a set of primitives.
 As shown in Figure \ref{fig:segblock} first an attempt is made to fit the stroke points into a curve or an ellipse using a minimum square error fitting algorithm \cite{chernov-2003}.
+The ellipse fitting step helps to prevent the system from over segmenting the stroke into multiple lines or curves if the input stroke is an ellipse.
 If the stroke proved to be an ellipse arc then the segmentation process ends and the system proceeds to the next step.
 Otherwise, the stroke is passed to two particle swarm algorithms that divide the stroke to either lines or lines and curves.
-The algorithms takes the stroke points along with the possible dominant points $P_{pd}$ computed during preprocessing then produce a set of dominant points which are connected by either lines or curves.
+The input to these algorithms are the stroke points along with the possible dominant points $P_{pd}$ computed during preprocessing.
+The output is a set of dominant points which are connected by either lines or curves.
 The next section describes the ellipse detection algorithm and the two particle swarm algorithms used to divide the stroke.
+
+%///AllBlockDiagram.jpg  you can change the block to the main block. 
  \begin{figure}
 	\centering
-		\includegraphics[scale=0.55]{images/segblock.jpg}
+		\includegraphics[scale=0.7]{images/blockSmall.jpg}
 	\caption{Segmentation Block}
@@
 \subsubsection{Ellipse Detection} 
-The process starts by computing the center of the stroke bounding box.
+%The process starts by computing the center of the stroke bounding box.
 The bounding box center point is used as the first estimation of the ellipse center.
@@
 The least square fitting algorithm is used to minimize the fitting error of the ellipse equation \cite{chernov-2003}.
+This process tries to fit the stroke points into an ellipse arc; it starts with computing the center of the stroke bounding box.
+The bounding box center point is used as the first estimation of the center of the ellipse.
+The axes of the ellipse are estimated as the $width/2$ and $height/2$ of the stroke bounding box where $width$ is width and $height$ is height of the bonding box.
+The least square fitting algorithm \cite{chernov-2003} is used to minimize the fitting error of the ellipse Equation (\ref{eq:circleFit})
+
+\begin{equation}
+E = \sum\limits_{i = 0}^N {\frac{{(x_i - x_0 )}}{{a^2 }}^2  + \frac{{(y_i - y_0 )}}{{b^2 }}^2  - 1} 
+\label{eq:circleFit}
+\end{equation}
+
+ where $N$ is number of points in the stroke, $a,b$ are the length of ellipse axes, $x_0$ \& $y_0$ are the coordinates of the center point, $x_i$ \& $y_i$ are the coordinates of point $i$ in the stroke.
+A list of new values for $x_0$ , $y_0$ ,$a$ and $b$ are generated randomly from the older values with small increments after each loop.
+After few loops, the final fit error and the ellipse confidence value are computed.
+The confidence value $Conf$ computed by Equation (\ref{eq:circleError})
+ \begin{equation}
+Conf = E_{ellipse}+E_{area}
+\label{eq:circleError}
+\end{equation}
+ \begin{equation}
+E_{area}  = \sqrt {\left( Area(stroke) - Area(ellipse)\right)^2 }
+\label{eq:ErrorArea}
+\end{equation}
+where $E_{ellipse}$ is the error computed by Equation(\ref{eq:circleFit}) and $E_{area}$ is computed as the difference between the actual area of the stroke and the area of the estimated ellipse Equation(\ref{eq:ErrorArea}).
+If confidence value $Conf$ is less than threshold $th_{Ellipse}$\footnote{The threshold used is $th_{Ellipse}=0.7$ } then the stroke is segmented as an ellipse otherwise the system proceeds to the two PSO segmentation algorithms.
+
 \subsubsection{Discrete particle swarm algorithm (\textit{DPSO})}
@@
 Two DPSO algorithms are used to generate segmentations for each stroke the user draws.
-For each generated segmentation evaluated an error is determined.
+For each segmentation generated an error is evaluated.
 The segmentation with the minimum error value is chosen as the best stroke segmentation (Figure \ref{fig:segblock}).
 The problem definition is the same in both algorithms but they differ in the method they use to compute fitness and error functions.
-
-\textsl{Problem definition:} The input stroke with $N$ points can be represented by set $S = \left\{ {x_1 ,x_2  \ldots x_N }\right\}$ where $x_i$ is the location of the point $i$.
+\begin{description}
+	\item[ Problem definition:] The input stroke $S$  with $N$ points can be represented by set $S = \left\{ {x_1 ,x_2  \ldots x_N }\right\}$ where $x_i$ is the location of the point $i$.
 The swarm algorithms consist of $M$ agents which are represented by the set $A = \left\{ {P_i \left| {i = 1,2 \cdots M} \right.} \right\}$ where $P_i$ is a single solution particle from the solution space.
@@
 
-Therefore, the system represents each particle $P_i$ by $P_i = \left\{ {p_{ij} \left| {j = 1,2 \cdots N} \right.} \right\}$ where $p_{ij}$ has only two values 0 or 1.
-$p_{ij}=1$ means that point $j$ is a dominant point.
+Therefore, the system represents each particle $P_i$ by $P_i = \left\{ {p_{ij} \left| {j = 1,2 \cdots N} \right.} \right\}$ where $p_{ij}$ has only two values a)1 ($p_{ij}=1$); means that this point ($j$) is a dominate point, or b) 0($p_{ij}=0$) means this point ($j$) is not a dominate point.
 Figure\ref{fig:LabelsPSO} shows a particle encoding in the DPSO system.
 
-\textsl{Fitness function:} The fitness function and error calculation are different in each of the two \textit{DPSO} algorithms.
+	\item[Fitness function:] The fitness function and error calculation are different in each of the two \textit{DPSO} algorithms.
 
-For the \textit{first algorithm \textsl{AlgS1}}, 
- the arc $\widehat{x_ix_j}$ is defined as the consecutive set of points from point$x_i$ to point $x_{j}$ as in $x_i,x_{i+1} \cdots,x_j$.
-The line $\overline{x_i x_j} $ as the straight line connecting point $x_i$ to point $x_j$.
-The approximation error is computed by the equation \ref{eq:ErrorSwarm1}.
-The graphical meaning of the error is shown in Figure\ref{fig:LabelsPSO}.
+\textit{The first algorithm \textsl{AlgS1}}\\
+The approximation error is computed by the equation \ref{eq:ErrorSwarm1}. The graphical meaning of the error is shown in Figure\ref{fig:LabelsPSO}.
 \begin{equation}
@@
 \label{eq:ErrorSwarm1}
-\end{equation} where $M$ is the number of dominant points in this solution as generated by the swarm algorithm.
+\end{equation} where the arc $\widehat{x_ix_j}$ is defined as the consecutive set of points from point $x_i$ to point $x_{j}$ as in $x_i,x_{i+1} \cdots,x_j$.
+The line $\overline{x_i x_j} $ as the straight line connecting point $x_i$ to point $x_j$, $M$ is the number of dominant points in this solution as generated by the swarm algorithm.
 The error $e ( \widehat{x_ix_j},\overline{x_i x_j})$ is computed as the sum of squared perpendicular distance from every point along the arc $\widehat{x_ix_j}$ to the line $\overline{x_i x_j}$.
-The fitness is computed using equation \ref{eq:fitnessSwarm1}\begin{equation}
+The fitness is computed using equation %\ref{eq:fitnessSwarm1}
+\begin{equation}
 \max fitness(p_i ) = \left\{ {\begin{array}{*{20}c}
@@
 
-The \textit{second algorithm \textsl{AlgS2}} has the same problem formulation but different fitness and error functions are used.
-An attempt is made to fit each segment into line or circular arc, the error of both circle and line fit estimations are computed for each segment $S_i=\widehat{x_ix_j}$, the approximation with the lower error value is the chosen approximation of this segment $S_i$\cite{CruveDivisionSwarm}.
-The sum of the approximation error of all segments is the total error of the particle.
-The total error of the particle is computed by equation \ref{eq:errorSwarm2}
+ \textit{The second algorithm \textsl{AlgS2}} \\
+The algorithm has the same problem formulation but different fitness and error functions are used.
+An attempt is made to fit each segment into line or circular arc.
+The errors of both circle and line fit estimations are computed for each segment $S_i=\widehat{x_ix_j}$.
+The approximation with the lower error value is chosen as the final approximation of this segment $S_i$\cite{CruveDivisionSwarm}.
+The sum of the approximation errors of all segments is the total error of the particle.
+The total error of the particle is computed by equation %\ref{eq:errorSwarm2}
  \begin{equation}
@@
 \label{eq:errorSwarm2}
-\end{equation}where $M$ is the number of segments in the solution as generated by the swarm algorithm, $D_i$ is the minimum approximation error of curve and line approximations $min(d_c,d_l)$ as computed by \cite{CruveDivisionSwarm}.
-The fitness is computed by the equation \ref{eq:fitnessSwarm2} \begin{equation}
+\end{equation}where $M$ is the number of segments in the solution as generated by the swarm algorithm, $D_i$ is the minimum approximation error of curve and line approximations $min(d_c,d_l)$ where $d_c$ is the circle approximation error and $d_l$ is the line approximation error as computed by \cite{CruveDivisionSwarm}.
+The fitness is computed by the equation %\ref{eq:fitnessSwarm2}
+\begin{equation}
 \max fitness(P_i ) = \frac{1}{{E \times M^k }}
@@
 \end{equation} where $E$ is the error and $M$ is the number of segments and $k$ is a parameter tweaked to get minimum number of segments.
-The larger k is, the more effect the number of segments will have.
-For our system, k is selected to be 0.5\cite{CruveDivisionSwarm}.
+The larger $k$ is, the more effect the number of segments will have.
+For our system, $k$ is selected to be 0.5\cite{CruveDivisionSwarm}.
+\end{description}
 
-After each loop of the swarm algorithm (\textsl{AlgS1} and \textsl{AlgS2}), each particle is refined using the following procedure.
-For each particle $P_i$ each dominant point $P_{ij}$ is checked to find if it was labeled before as a \textit{possible dominant point} $P_{pd}$ (computed as in section \ref{Prepross}).
+After each iteration of the swarm algorithm (\textsl{AlgS1} and \textsl{AlgS2}), each particle is refined using the following procedures: 
+\begin{enumerate}
+	\item For each particle $P_i$ each dominant point $P_{ij}$ is checked to find if it was labeled before as a \textit{possible dominant point} $P_{pd}$ (computed as in section \ref{Prepross}).
 If it was not labeled the point $P_{ij}$ is moved to the nearest labeled point.
 This ensures that all of the points generated by the DPSO are possible dominate points $P_{pd}$.
-After that the particles are tested to make sure that the distance between every two successive dominate point is larger than the constant $min_D$.
-If two points are nearer than $min_D$ then one of the points is removed.
+	\item The particles are tested to make sure that the distance between every two successive dominate point is larger than the constant $min_D$.
+Otherwise, one of these points are removed based on the error computed by the segmentation algorithm.
+	\item Each two successive segments are test to determine if they can be merged into a single segment.
+If merging the two segments have the same or smaller the error, the two segments are merged.
+This step ensures that the final segmentation of the system has minimum number of segments.
+\end{enumerate}
+%1) 2)  % If two points are nearer than $min_D$ then one of the points is removed. 
 \subsection{Recognition}
@@
 A composite set of features is used to generate a single feature vector.
-The features used consist of Rubine feature set,  Zernike moments of order 21 \cite{HeloiseBeautification}, ink density as well as some structural and spatial information like number of perpendiculars lines ,number of parallel lines and number of different types of primitives in each symbol.
-After computing the features the symbol is introduced to the (SVM) classifier.
+The following list describes each feature in details.
+\begin{description}
+ \item [Structural and geometrical Features(FS1)] Features defines the structure of the geometrical symbol.  
+\begin{description}
+	 \item[No. Of segments] Number of segments in the symbol. 
+		\item[No. Of primitives] Number of primitives in the symbol. The feature helps when identifying             symbols with mixed geometric primitives like cylinders and callouts.  
+		\item [No. of curves] Number of curves or ellipses in the symbol. 
+		\item [No. of lines] Number of lines in the symbol. 
+		\item [No. of perpendicular lines] Number of perpendicular lines. 
+		\item [No. of parallel lines] Number of parallel lines. 
+		\item [No. of intersections] Number of intersection between lines and curves. 
+		\item [Size Ratio] Ratio between width to height of the symbol.
+		%\item[No. of holes] Number of holes in the symbols.
+
+\end{description}
+\item [Rubine Feature Set (FS2)] Features introduced by Rubine\cite{gestureexample12} for single stroke gestures. 
+\item[Statistical Features (FS3)]   
+\begin{description}
+	\item[Zernike moments]  Zernike moments of order \cite{HeloiseBeautification}. 
+\end{description}
+	\item[Composite Features (FS4)] Features that are composites of different geometrical and statistical symbol. 
+	\begin{description}
+
+	\item [Ink density] Compute the density of points inside its bounding box\cite{GeometryAndDomain102}.   
+ 	\item[Convex Hull Area] Area of convex hull with respect to area of bounding box of symbol.
+	\item[Convex Hull Perimeter] Perimeter of convex hull with respect to total length of symbol.
+		\item [Mean Centroidal radius] The Mean of the centroidal radius which is the distance from each point in the symbol to the center of gravity.
+
+	\item [Mean Time difference] The mean of the time difference between each two successive points in the symbol. %Different strokes are appended to construct a single path.
+	%\item [Ra]
+  \end{description}
+
+\end{description}
+%The features used consist of Rubine feature set,  of order 21 , ink density as well as some structural and spatial information like number of perpendiculars lines ,number of parallel lines and number of different types of primitives in each symbol.
+After computing the features the symbol is introduced to the classifier.
+The system uses Support Vector Machine (SVM) classifier with Linear kernel \cite{libsvm}.
+An OVO classifier structure (object versus object) is used to handle multi-class.% A validation algorithm is used to choose the Kernel variables parameters $c, \gamma$.
+
 \section{Experimental}
@@
 The dataset collected by Hse and Newton\cite{HeloiseBeautification} is used to test the presented system.
-The data are drawn by 16 users each of them was asked to draw the 13 shapes from 30 to 50 times.
+The data are drawn by 16 users each of them was asked to draw 13 shapes from 30 to 50 times.
 Figure \ref{fig:symbolSet} shows a set of the shapes used in the data set.
 The dataset was divided into training set and test set.
-To generate trusted results, four different splits for the training and test data are generated from the dataset.
-The results displayed are the recognition accuracy resulted from the classifier after recognizing the symbols.
+Four different splits for the training and test data are generated from the dataset.
+The results displayed are the average recognition accuracy of the four splits resulted from the classifier after recognizing the symbols.
 The accuracy is computed as the number of correctly recognized samples divided by the total number of samples in the test.
 %In the following section, the recognition accuracy means the average recognition accuracy in the four data splits.
- \begin{figure*}[]
+\begin{figure}
 \centering 
-	\begin{minipage}[b]{5cm}	
-\begin{center}
-\subfigure [The Symbol Set] {\label{fig:symbolSet}
-\fbox{ \parbox{2cm}{% 
-		\includegraphics[scale=0.2]{images/symbolSet.PNG}	}}}
+\label{fig:symbolSet}
+\fbox{ \parbox{5cm}{% 
+		\includegraphics[scale=0.5]{images/symbolSet.PNG}	}}
+		\caption{The Symbol Set} Symbols used in the dataset.
+\end{figure}
+\subsection{Algorithms comparisons}
+\label{sec:AlgExp}
+A set of test were performed on the system to enhance both the swarm algorithms outputs.
+The effect of number of agents, the error thresholds and various other paramters was investigated.
+The result of these tests are shown in Figure \ref{fig:swarmtesting} and   \ref{fig:swarmtesting2}.
+Figure \ref{fig:swarmtesting} shows that as the number of agents increases the error value decrease (error of the segmentation algorithm).
+The final number of agents and maximum swarm iteration used in the system was based trade off between error achieved versus the computation time.
+Other algorithm parameters like $c_1$,$c_2$ that were mentioned in Section\ref{sec:ParticleSwarmAlgorithm} are tested using similar tests (Figure \ref{fig:swarmtesting} and \ref{fig:swarmtesting2}).
 
-		\subfigure[Algorithm comparison]{\includegraphics[width=5cm]{images/test1smallpaper.jpg} 	\label{fig:test1}}
-		\end{center}
-		\end{minipage}
-		\hfill
-  \subfigure [Symbols Comparison] {\includegraphics[width=9cm]{images/test2smallpaper.jpg} \label{fig:test2}}
-	\caption{Experiments results :} a) the figure shows a set of the symbols used in the dataset.
-b)The results of comparing the algorithms.
-The graph displays the recognition rate of symbols.
-c) The graph shows the recognition rate of each symbol using different algorithms.
+ \begin{figure}
+	\centering		
+	 \includegraphics[scale=0.3]{images/swarmTesting.jpg}
+	 	\caption{Swarm testing} Segmentation Error Vs. Number of agents
+	 	\label{fig:swarmtesting}
+	%\caption{Experiments results :} a)    b) c) 
+\end{figure}
+
+\begin{figure*}
+	\centering		
+	 \includegraphics[width=18cm]{images/swarmTesting2.jpg}
+	 	\caption{Swarm testing} Segmentation Error Vs. PSO parameters.
+	 	\label{fig:swarmtesting2}
+	%\caption{Experiments results :} a)    b) c)
+
 \end{figure*} 
-We performed two experiments to test the system.
+We performed several experiments to evaluate the presented recognition system.
 Firstly we tested recognition accuracy of shapes in the data set with both algorithms (\textsl{AlgS1} and \textsl{AlgS2}).
+We also implemented the segmentation algorithm described in \cite{earlyprocess} (\textsl{Alg3}) to use as reference to compare it with our swarm algorithms.
 Figure \ref{fig:test1} shows the accuracy achieved by each algorithm.
 The two swarm algorithms were tested with and without the ellipse fitting module.
-Including the ellipse detection module improves the results with both the \textit{DPSO} algorithms.% The results show that both \textit{DPSO} algorithms achieve better result than other algorithms.
+The ellipse detection module appears to be superior.
+The result shows that combining both  (\textsl{AlgS1} and \textsl{AlgS2}) out perform any single algorithm.
+%show that combining algoirthmsachieve higher accuracy than any single algoirthm.
+%improves the results.% with both the \textit{DPSO} algorithms.% The results show that both \textit{DPSO} algorithms achieve better result than other algorithms.
 
-The second experiment we implemented was to test the effect of symbol complexity and type on the recognition rate.
-Figure \ref{fig:test2} shows the accuracy of each symbol, it is clearly noted that symbols that have only line segments achieve higher accuracy rate than other symbols.
-Figure \ref{fig:test2} also indicates that algorithm \textsl{AlgS1} alone achieve better performance than algorithm   \textsl{AlgS2} in the symbols that consist only of lines.
-This is natural as algorithm \textsl{AlgS1} divides strokes into line segments only but \textsl{AlgS2} is able to divide strokes into lines and curves based on the minimum error of the segment itself.
+ \begin{figure}
+	\centering		
+	 \includegraphics[scale=0.4]{images/testAlg.jpg}
+	 	\caption{Algorithm comparison} The recognition rate of different algorithms. 
+	 	\label{fig:test1}
+	%\caption{Experiments results :} a)    b) c) 
+\end{figure}
+
+\subsection{Shape Complexity Analysis}
+\label{sec:ShapeComplexityExperiments}
+
+\begin{figure*}
+	\centering
+		\includegraphics[scale=0.5]{images/testsym.jpg}
+	\caption{Symbols Comparison} The effect of symbol complexity.  %The graph shows the recognition rate of each symbol using different algorithms. 
+	\label{fig:test2}
+\end{figure*}  
+The second experiment we implemented was to investigate the effect of symbol complexity and type on the recognition rate.
+Figure \ref{fig:test2} shows the achieved accuracy of each symbol by our algorithms compared to \cite{earlyprocess} (\textsl{Alg3}).
+It is clearly noted that symbols that have only line segments achieve higher accuracy rate than other symbols.
+The results indicate that algorithm \textsl{AlgS1} achieve better performance than algorithm \textsl{AlgS2} in the symbols that consist only of lines.
+This is understandable because algorithm \textsl{AlgS1} divides strokes into line segments only but \textsl{AlgS2} is able to divide strokes into lines and curves based on the minimum error of the segment itself.
+Algorithm \textsl{Alg3} gives good performance as long as the symbols consist of combination of lines and curves, if the stroke consists of only curve or lines the algorithm may lead to wrong segmentation result.
+This is because the system divides the stroke first to line segments then tries to decide if each segment can be represent better as a curve unlike algorithm \textsl{AlgS2} where the curve segments are tested while choosing the best segmentation.
 Combining both algorithm \textsl{AlgS1} and \textsl{AlgS2} improved the recognition rate of all symbols.
-The penalty for this increased performance is in the computational time needed to run both swarm algorithms.
+The penalty for this improved performance is the computational time required to run both swarm algorithms.
+Table \ref{tab:ConfusionMatrix} shows the confusion matrix of symbols.
+The table shows that errors are only between two sets of symbols moon with triangles and trapezoid with triangles.
+This observation is understandable as the symbols are visually similar but the system must be able to differentiate between them.
+This lead us to the next experiment to choose best set of feature to recognize symbols.
+\begin{table*}
+	\centering
+	%\small
+%	\scalebox{0.7}{
+		%	\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|c|c|c|c|}\hline 
+ \include{conf}
+ 		%\end{tabular}
+ 	%}
+	\caption{Confusion Matrix}
+	\label{tab:ConfusionMatrix}
+	%\end{minipage}
+\end{table*}
+\subsection{Features Analysis}
+\label{sec:featexp}
+Different feature sets are tested to determine the best features that can used in sketch and symbol recognition.
+Figure\ref{fig:testFeaturesAll} shows the result of different feature sets from the basic four sets (\textbf{FS1,FS2,FS3,FS4}).
+Result shows that (\textbf{FS2}) Rubine features \cite{gestureexample12} gives the worst results when used alone without any other features.
+This is because they are mainly computed for single stroke gestures and fare bad in multi-stroke symbols \cite{compareFeaturSVM}.
+Features \textbf{(FS4)} gives good results but it is improved by adding structural features \textbf{(FS1)}.
+ \begin{figure*}
+	\centering
+		\includegraphics[scale=0.5]{images/FeaturesCompare.jpg}
+	\caption{Feature Comparison} The effect of different features on accuracy.  %The graph shows the recognition rate of each symbol using different algorithms. 
+	\label{fig:testFeaturesAll}
+\end{figure*}
+
+ We could not test the result of the segmentation algorithm directly due to the fact that the correct segmentation is highly ambiguous.
+It is only based on what the user intends to draw while sketching the symbol.
+Figure \ref{fig:SampleSeg} shows a sample of the output of the segmentation block.
+The result of segmentation block is tested by their effect on the final recognition results using the features that will be highly affected using different algorithms like \textbf{FS1} and \textbf{FS4} (Section\ref{sec:Recognition}).
+Figure \ref{fig:testFeatonly} shows the result of the different algorithms using only \textbf{FS1 }features.
+\begin{figure}
+	\centering
+		\includegraphics[scale=0.4]{images/FeaturesCompare.jpg}
+	\caption{} %The effect of symbol complexity.  %The graph shows the recognition rate of each symbol using different algorithms. 
+	\label{fig:testFeatonly}
+\end{figure}
+
+\begin{figure}[]
+	\centering
+				\subfigure[User strokes] {\includegraphics[scale=0.7]{images/strokebefore.jpg}}
+		 	\subfigure[Segmented strokes] {\includegraphics[scale=0.7]{images/strokeafter.jpg}}
+		 		\subfigure[Multiple symbols] {\includegraphics[scale=0.7]{images/results3.jpg}}
+
+	\caption{Segmentation outputs}  The output segmentation of different input strokes.
+%The figure shows sample of the output of the segmentation phase.
+Figure a) shows the stroke after the user draws it.
+The red points are points labeled as \textit{possible dominate points}.
+Figure b) shows the symbol after being segmented.
+Figure c) shows the segmentation of different and overlapped strokes.
+Notice the system mark the center of the ellipse that was identified.
+	\label{fig:SampleSeg}
+\end{figure}
+
 \section{Conclusion and Future Work}
@@
 The use of statistical moments and structural features improves the recognition rate.
-The system was tested on 13 different symbols and achieved an accuracy of 91.5\%.
+The system was tested on 13 different symbols and achieved an accuracy of 93.5\%.
 The system does not depend on low level recognizer but rather on a set of high level features.

### 2009-04-04 draft/Results.tex
before change of reuslt on data of 4 appril

--- 2009-02-18 trunk/draft/Results.tex
@@
 
-This chapter discribe and illustrate the results generated from the system.
-In the chapter, section \ref{sec:ExperimentsSetup} will describe the experiments performed on each block of the system.
-section \ref{sec:Datasets} will describe the datasets used.
-Section \ref{sec:ResultsDetails} will demonstrate and analysis the result generated.
+This chapter discribes and illustrates the experiments performed on the system.
+In the chapter, section \ref{sec:ExperimentsSetup} describes the experimental setup.
+Section \ref{sec:Datasets} describes the datasets used.
+Section \ref{sec:ResultsDetails} demonstrates and analyzes the result generated.
 
@@
 The system was tested using various experiments.
-The preprocessing procedures and computation was tested by its effects on the segmentation algorithm and was reported in section (\ref{sec:ResultsDetails}).
-The variable used to tune the swarm algorithm were tested to choose the best parameters to use(see section \ref{sec:PSO}).
-Segmentation algorithms test was reported and noted in section \ref{sec:SegmentationAlgorithms}.
+The preprocessing procedures and computation is tested by its effects on the segmentation algorithm and was reported in section (\ref{sec:ResultsDetails}).
+The effect of number of agents, the error thresholds and various other paramters was investigated(see section \ref{sec:PSO}).
+We performed several experiments to evaluate the presented recognition system, those experiments are reported and noted in section \ref{sec:SegmentationAlgorithms}.
 
+ The accuracy is computed as the number of correctly recognized samples divided by the total number of samples in the test. 
+%The variable used to tune the swarm algorithm were tested to choose the best parameters to use
 \section{Datasets}
 \label{sec:Datasets}
-For classifier and algorithm testing the system used a data set collected by \citeauthor{HeloiseBeautification}.
+Two different data set were used, 
+%For classifier and algorithm testing the system used a data set collected by
+\begin{description}
+	\item [Heloise Dataset] Data set collected by \citeauthor{HeloiseBeautification}.
 The data are drawn by 16 users each of them have drawn from 30 to 50 samples for each shape.
 Figure \ref{fig:symbolSet} shows the shapes used in the data set.
-We divided the dataset equally into training set and test set.
-The results displayed are the recognition accuracy resulted from the classifier after recognizing the symbols.
-\\
+The dataset was divided into training set and test set.
+Four different splits for the training and test data are generated from the dataset.
+The results displayed are the average recognition accuracy of the four splits resulted from the classifier after recognizing the symbols.
+		\item [MultiDomain Dataset] Dataset collected manually.
+The dataset contains symbols from three diffrent domains ( UML, mechanical and electrical engineering symbols).
+Data are collected from 4 different users each of them have drawn from 20 to 40 samples of each shapes.
+Figure \ref{fig:MultisymbolSet} shows some of the shapes used in the dataset.
+The dataset was divided into training set and test set.
+Four different splits for the training and test data are generated from the dataset.
+The results displayed are the average recognition accuracy of the four splits resulted from the classifier after recognizing the symbols.
+\end{description}
+ %We divided the dataset equally into training set and test set.  The results displayed are the recognition accuracy resulted from the classifier after recognizing the symbols. \\
 %We divided the dataset into four sets each set contains four different users.
@@
 	\label{fig:symbolSet}
+\end{figure}
+
+\begin{figure}[]\centering
+\fbox{  \parbox{8cm}{% 
+		\includegraphics[scale=0.7]{images/symbolSet.eps}	}
+		}
+	\caption{The MultiDomain Dataset }
+	\label{fig:MultisymbolSet}
 \end{figure}

### 2009-04-04 draft/Results.tex
going to edit the survey

@@
-\chapter{Results}
+\chapter{Evaluation}
 
@@
 Section \ref{sec:Datasets} describes the datasets used.
-Section \ref{sec:ResultsDetails} demonstrates and analyzes the result generated.
+Section \ref{sec:ResultsDetails} demonstrates and analyzes the results.
 
-\section{Experiments Setup}
-\label{sec:ExperimentsSetup}
-The system was tested using various experiments.
-The preprocessing procedures and computation is tested by its effects on the segmentation algorithm and was reported in section (\ref{sec:ResultsDetails}).
-The effect of number of agents, the error thresholds and various other paramters was investigated(see section \ref{sec:PSO}).
-We performed several experiments to evaluate the presented recognition system, those experiments are reported and noted in section \ref{sec:SegmentationAlgorithms}.
+The system has two main parts the DPSO stroke segmenation and symbol recogntion.
+Each part is evaluated sperately.
+The preprocessing procedures and computation is evaluated using their effect on the segmentation algorithm.
+These results are reported in section \ref{sec:ResultsDetails}.
+The effect of number of agents, the error thresholds and various other PSO paramters are investigated in Section \ref{sec:PSO}.
+The results presented in Section \ref{sec:ResultsDetails} and \ref{sec:PSO}.
 
- The accuracy is computed as the number of correctly recognized samples divided by the total number of samples in the test. 
+The recognition system evalutation is presented in Section \ref{sec:SegmentationAlgorithms}.
+The recogntion rate (computed as the number of correctly recognized samples divided by the total number of samples in the test) for each segmentation algorithm compared to another segmenation algorithm are reported in \ref{sec:recognitionAlgorithms}.
+Next, the scallability of our recognition system is evaluated and reported in Section \ref{sec:recognitionAlgorithms}.
+Finally, a feature set comparision is reported in Section \ref{sec:featuresComparisions}.
+
+%The accuracy is computed as the number of correctly recognized samples divided by the total number of samples in the test.
+
+%The recognition system evalutation is presented in Section \ref{sec:SegmentationAlgorithms}. The
+
 %The variable used to tune the swarm algorithm were tested to choose the best parameters to use
@@
 \begin{description}
-	\item [Heloise Dataset] Data set collected by \citeauthor{HeloiseBeautification}.
+	\item [Heloise Dataset] Data set collected by \cite{HeloiseBeautification} and was previously used by \cite{Oltmans07} as benchmarck for sketch recognition problem.
 The data are drawn by 16 users each of them have drawn from 30 to 50 samples for each shape.
@@
 Four different splits for the training and test data are generated from the dataset.
-The results displayed are the average recognition accuracy of the four splits resulted from the classifier after recognizing the symbols.
-		\item [MultiDomain Dataset] Dataset collected manually.
+%The results displayed are the average recognition accuracy of the four splits resulted from the classifier after recognizing the symbols.
+		\item [MultiDomain Dataset] A dataset we collected manually from different users.
 The dataset contains symbols from three diffrent domains ( UML, mechanical and electrical engineering symbols).
@@
 Four different splits for the training and test data are generated from the dataset.
-The results displayed are the average recognition accuracy of the four splits resulted from the classifier after recognizing the symbols.
+%The results displayed are the average recognition accuracy of the four splits resulted from the classifier after recognizing the symbols.
 \end{description}
@@
 %\label{sec:Curvatureestimationresults}
+\section{Performance Measures }
+\label{sec:PerformanceMeasures}
+
+\begin{description}
+	\item[Segmentation Error] The segmentation error is computed from each stroke after segmentation as computed from Equation \ref{} in Section \ref{}. 
+	\item [Recognition accuracy] We measure recognition performance for each algorithm by determining the number of correctly identified symbols in the whole dataset.
+Note that All the results are shown as the average of the four split of the dataset used.
+\end{description}
 
@@
 \label{sec:PSO}
-The fig.
+
+Figrue \ref{fig:swarmtesting0} shows the Average segmentation error using different preliminary calculation (Section \ref{sec:CurvatureCalculation}).
+The figure shows that using all computations (time difference, direction, speed and curvature) for each stroke decrease the Segmentation error.
+Figure \ref{} shows the average number of Possible dominate point $P_{pd}$ with respect to information used in the system.
+The results shows that the using all information increases the number of Possible dominate point $P_{pd}$ which in segmenation stage decreses the \textit{Segmenation Error}.
+
+ The effect of number of agents, the error thresholds and various other paramters was investigated.
+The result of these tests are shown in Figure \ref{fig:swarmtesting} and  \ref{fig:swarmtesting2}.
+Figure \ref{fig:swarmtesting} shows that as the number of agents increases the \textit{Segmenation Error} value decrease (error of the segmentation algorithm).
+Figure \ref{fig:swarmtesting1} shows simillar result for the maximum swarm iterations.
+Other algorithm parameters like $c_1$,$c_2$ that were mentioned in Section\ref{sec:ParticleSwarmAlgorithm} are tested using similar tests (Figure \ref{fig:swarmtesting} and \ref{fig:swarmtesting2}).
+The final values are for maximum number of iteration = 60 with 15 particles.
+The final number of agents and maximum swarm iteration used in the system was based trade off between error achieved versus the computation time.
+
+ \begin{figure}
+	\centering		
+	 \includegraphics[scale=0.3]{images/swarmTesting}
+	 	\caption{Swarm testing} Segmentation Error Vs. Number of agents
+	 	\label{fig:swarmtesting}
+	%\caption{Experiments results :} a)    b) c) 
+\end{figure}
+
+\begin{figure*}
+	\centering		
+	 \includegraphics[width=18cm]{images/swarmTesting2}
+	 	\caption{Swarm testing} Segmentation Error Vs. PSO parameters.
+	 	\label{fig:swarmtesting2}
+	%\caption{Experiments results :} a)    b) c)
+
+\end{figure*}
+
+%The fig.
 \ref{fig:dataerrorvsiteration.jpg} display the effect of the size of swarm population on the number of vertex reported while segmenting the stroke.\footnote{The fewer the vertex the better the segmentation} The graphs shows that as the number of iteration increase there is a decrease in error until the error reach a saturation and any increase in number of iteration dose not affect the error calculated.
@@
 
-\begin{figure}
-	\centering
-		\includegraphics[scale=0.8]{images/dataerrorvsiteration.jpg.eps}
-	\caption{Error vs. iterations }
-	\label{fig:dataerrorvsiteration.jpg}
-\end{figure}
-\begin{figure}
-	\centering
-		\includegraphics[scale=0.8]{images/datapopulation.jpg.eps}
-	\caption{Swarm Population }
-	\label{fig:datapopulation.jpg}
-\end{figure}
+%\begin{figure}
+%	\centering
+%		\includegraphics[scale=0.8]{images/dataerrorvsiteration.jpg.eps}
+%	\caption{Error vs. iterations }
+%	\label{fig:dataerrorvsiteration.jpg}
+%\end{figure}
+%\begin{figure}
+%	\centering
+%		\includegraphics[scale=0.8]{images/datapopulation.jpg.eps}
+%	\caption{Swarm Population }
+%	\label{fig:datapopulation.jpg}
+%\end{figure}
 
@@
 \label{sec:SegmentationAlgorithms}
-The result of the segmentation algorithm can be viewed in the figs. \ref{fig:results1.JPG}. These result shows the originals stroke and the segmentation that the system generated for the stroke. 
+The result of the segmentation algorithm can be viewed in the Figure \ref{fig:results1.JPG}. These result shows the originals stroke and the segmentation that the system generated for the stroke. 
 \begin{figure}
@@
 \end{figure}
+
+\subsection {Recognition Algorithms}
+\label{sec:RecognitionAlgorithms}%}
+To test the recognition different set of features are used to select the features that will define symbols better to finally achieve better recognition rate.
+We tested the system using different types of features.
+The features explained in section \ref{sec:FeatureExtraction} was divided into different sets; statistical, structural, Rubine features and zernik moments along with composite features.
+The fig.
+\ref{fig:testFeatures} shows the result of each set of features when used with the SVM classifier.
+The results shows that the Rubine features although may achieve good results in single stroke symbols, have a bad performance when used with multi stroke symbols.
+\begin{figure}[]
+	\centering
+		\includegraphics[scale=0.3]{images/test4.eps}
+	\caption{The results of Features experiment}
+	\label{fig:testFeatures}
+\end{figure}
+
+We test the system using different classifier to see what the effect of the classification step on the final recognition.
+We also wanted to test the train and test time in both classifiers.
+We used SVM (RBF) classifiers and the Gaussian Linear classifier used in \cite{}.
+The fig.
+\ref{} shows the result of both classifiers with the main three segmentation algorithms.
+The results shows that the SVM classifier achieve better results.
+To measure the scalability of the system\footnote{What happens when adding a new symbols to the set of known symbols.} the classifiers are test using different number of symbols (fig.
+\ref{fig:sampleperCat}).
+Figures \ref{fig:sampleperCat} demonstrate the scalability of both classifiers by demonstrating the recognition rate in both classifiers with different number of symbols.
+Figures \ref{fig:sampleperCat},\ref{fig:sampleperCat} shows the time and number of symbols needed for training versus each classifier.
+
+\begin{figure}
+	\centering
+		\includegraphics{images/sampleperCat.eps}
+	\caption{Samples per Category }
+	\label{fig:sampleperCat}
+\end{figure}
+
+\begin{figure}
+	\centering
+		\includegraphics[scale=0.7]{images/test4.eps}
+	\caption{Tests}
+	\label{fig:test4}
+\end{figure}
+
+\subsubsection {Effect of sysmbol complixity}
 
@@
 
-\subsection {Recognition Algorithms }
-\label{sec:recognitionAlgorithms}
-To test the recognition different set of features are used to select the features that will define symbols better to finally achieve better recognition rate.
-We tested the system using different types of features.
-The features explained in section \ref{sec:FeatureExtraction} was divided into different sets; statistical, structural, Rubine features and zernik moments along with composite features.
-The fig.
-\ref{fig:testFeatures} shows the result of each set of features when used with the SVM classifier.
-The results shows that the Rubine features although may achieve good results in single stroke symbols, have a bad performance when used with multi stroke symbols.
-\begin{figure}[]
-	\centering
-		\includegraphics[scale=0.3]{images/test4.eps}
-	\caption{The results of Features experiment}
-	\label{fig:testFeatures}
-\end{figure}
-
-We test the system using different classifier to see what the effect of the classification step on the final recognition.
-We also wanted to test the train and test time in both classifiers.
-We used SVM (RBF) classifiers and the Gaussian Linear classifier used in \cite{}.
-The fig.
-\ref{} shows the result of both classifiers with the main three segmentation algorithms.
-The results shows that the SVM classifier achieve better results.
-To measure the scalability of the system\footnote{What happens when adding a new symbols to the set of known symbols.} the classifiers are test using different number of symbols (fig.
-\ref{fig:sampleperCat}).
-Figures \ref{fig:sampleperCat} demonstrate the scalability of both classifiers by demonstrating the recognition rate in both classifiers with different number of symbols.
-Figures \ref{fig:sampleperCat},\ref{fig:sampleperCat} shows the time and number of symbols needed for training versus each classifier.
-
-\begin{figure}
-	\centering
-		\includegraphics{images/sampleperCat.eps}
-	\caption{Samples per Category }
-	\label{fig:sampleperCat}
-\end{figure}
-
-\begin{figure}
-	\centering
-		\includegraphics[scale=0.7]{images/test4.eps}
-	\caption{Tests}
-	\label{fig:test4}
-\end{figure}
+\subsubsection{Features Comparisions}
+\label{sec:featuresComparisions}
 

### 2009-04-06 trunk/draft/Results.tex
6 appril

@@
-\chapter{Results}
+\chapter{Evaluation}
 
-This chapter discribe and illustrate the results generated from the system.
-In the chapter, section \ref{sec:ExperimentsSetup} will describe the experiments performed on each block of the system.
-section \ref{sec:Datasets} will describe the datasets used.
-Section \ref{sec:ResultsDetails} will demonstrate and analysis the result generated.
+This chapter discribes and illustrates the experiments performed on the system.
+In the chapter, section \ref{sec:ExperimentsSetup} describes the experimental setup.
+Section \ref{sec:Datasets} describes the datasets used.
+Section \ref{sec:ResultsDetails} demonstrates and analyzes the results.
 
-\section{Experiments Setup}
-\label{sec:ExperimentsSetup}
-The system was tested using various experiments.
-The preprocessing procedures and computation was tested by its effects on the segmentation algorithm and was reported in section (\ref{sec:ResultsDetails}).
-The variable used to tune the swarm algorithm were tested to choose the best parameters to use(see section \ref{sec:PSO}).
-Segmentation algorithms test was reported and noted in section \ref{sec:SegmentationAlgorithms}.
+The system has two main parts the DPSO stroke segmenation and symbol recogntion.
+Each part is evaluated sperately.
+The preprocessing procedures and computation is evaluated using their effect on the segmentation algorithm.
+These results are reported in section \ref{sec:ResultsDetails}.
+The effect of number of agents, the error thresholds and various other PSO paramters are investigated in Section \ref{sec:PSO}.
+The results presented in Section \ref{sec:ResultsDetails} and \ref{sec:PSO}.
 
+The recognition system evalutation is presented in Section \ref{sec:SegmentationAlgorithms}.
+The recogntion rate (computed as the number of correctly recognized samples divided by the total number of samples in the test) for each segmentation algorithm compared to another segmenation algorithm are reported in \ref{sec:recognitionAlgorithms}.
+Next, the scallability of our recognition system is evaluated and reported in Section \ref{sec:recognitionAlgorithms}.
+Finally, a feature set comparision is reported in Section \ref{sec:featuresComparisions}.
+
+%The accuracy is computed as the number of correctly recognized samples divided by the total number of samples in the test.
+
+%The recognition system evalutation is presented in Section \ref{sec:SegmentationAlgorithms}. The
+
+%The variable used to tune the swarm algorithm were tested to choose the best parameters to use
 \section{Datasets}
 \label{sec:Datasets}
-For classifier and algorithm testing the system used a data set collected by \citeauthor{HeloiseBeautification}.
+Two different data set were used, 
+%For classifier and algorithm testing the system used a data set collected by
+\begin{description}
+	\item [Heloise Dataset] Data set collected by \cite{HeloiseBeautification} and was previously used by \cite{Oltmans07} as benchmarck for sketch recognition problem.
 The data are drawn by 16 users each of them have drawn from 30 to 50 samples for each shape.
 Figure \ref{fig:symbolSet} shows the shapes used in the data set.
-We divided the dataset equally into training set and test set.
-The results displayed are the recognition accuracy resulted from the classifier after recognizing the symbols.
-\\
+The dataset was divided into training set and test set.
+Four different splits for the training and test data are generated from the dataset.
+%The results displayed are the average recognition accuracy of the four splits resulted from the classifier after recognizing the symbols.
+		\item [MultiDomain Dataset] A dataset we collected manually from different users.
+The dataset contains symbols from three diffrent domains ( UML, mechanical and electrical engineering symbols).
+Data are collected from 4 different users each of them have drawn from 20 to 40 samples of each shapes.
+Figure \ref{fig:MultisymbolSet} shows some of the shapes used in the dataset.
+The dataset was divided into training set and test set.
+Four different splits for the training and test data are generated from the dataset.
+%The results displayed are the average recognition accuracy of the four splits resulted from the classifier after recognizing the symbols.
+\end{description}
+ %We divided the dataset equally into training set and test set.  The results displayed are the recognition accuracy resulted from the classifier after recognizing the symbols. \\
 %We divided the dataset into four sets each set contains four different users.
@@
 
+\begin{figure}[]\centering
+\fbox{  \parbox{8cm}{% 
+		\includegraphics[scale=0.7]{images/symbolSet.eps}	}
+		}
+	\caption{The MultiDomain Dataset }
+	\label{fig:MultisymbolSet}
+\end{figure}
+
 %\section{Evaluations techniques}
@@
 \label{sec:PSO}
-The fig.
+
+Figrue \ref{fig:swarmtesting0} shows the \textit{Average Segmentation Error} using different preliminary calculation (Section \ref{sec:CurvatureCalculation}).
+The figure shows that using all computations (time difference, direction, speed and curvature) for each stroke decrease the Segmentation error.
+Figure \ref{fig:swarmtestingpso} shows the average number of Possible dominate point $P_{pd}$ with respect to information used in the system.
+The results shows that the using all information increases the number of Possible dominate point $P_{pd}$ which in segmenation stage decreses the \textit{Segmenation Error}.
+
+ \begin{figure}
+	\centering		
+	 \includegraphics[scale=0.3]{images/swarmTesting}
+	 	\caption{Swarm testing} Segmentation Error Vs. Number of agents
+	 	\label{fig:swarmtesting0}
+	%\caption{Experiments results :} a)    b) c) 
+\end{figure}
+
+\begin{figure*}
+	\centering		
+	 \includegraphics[width=18cm]{images/swarmTesting2}
+	 	\caption{Swarm testing} Segmentation Error Vs. PSO parameters.
+	 	\label{fig:swarmtestingpso}
+	 	\end{figure*}
+
+ The effect of number of agents, the error thresholds and various other paramters was investigated.
+The result of these tests are shown in Figure \ref{fig:swarmtesting} and  \ref{fig:swarmtesting2}.
+Figure \ref{fig:swarmtesting} shows that as the number of agents increases the \textit{Segmenation Error} value decrease (error of the segmentation algorithm).
+Figure \ref{fig:swarmtesting1} shows simillar result for the maximum swarm iterations.
+Other algorithm parameters like $c_1$,$c_2$ that were mentioned in Section\ref{sec:ParticleSwarmAlgorithm} are tested using similar tests (Figure \ref{fig:swarmtesting} and \ref{fig:swarmtesting2}).
+The final values are for maximum number of iteration = 60 with 15 particles.
+The final number of agents and maximum swarm iteration used in the system was based trade off between error achieved versus the computation time.
+
+ \begin{figure}
+	\centering		
+	 \includegraphics[scale=0.3]{images/swarmTesting}
+	 	\caption{Swarm testing} Segmentation Error Vs. Number of agents
+	 	\label{fig:swarmtesting}
+	%\caption{Experiments results :} a)    b) c) 
+\end{figure}
+
+\begin{figure*}
+	\centering		
+	 \includegraphics[width=18cm]{images/swarmTesting2}
+	 	\caption{Swarm testing} Segmentation Error Vs. PSO parameters.
+	 	\label{fig:swarmtesting2}
+	%\caption{Experiments results :} a)    b) c)
+
+\end{figure*}
+
+%The fig.
 \ref{fig:dataerrorvsiteration.jpg} display the effect of the size of swarm population on the number of vertex reported while segmenting the stroke.\footnote{The fewer the vertex the better the segmentation} The graphs shows that as the number of iteration increase there is a decrease in error until the error reach a saturation and any increase in number of iteration dose not affect the error calculated.

### 2009-04-06 trunk/draft/chapterSurvey.tex
6 appril

@@
 
-The researchers had divided the sketch recognition problem into image based and AI Based problem.
+A wide variety of techniques were suggested for sketch recognition.
+Currently symbol recognition is solved as an either image based and feature based or AI-based problem.
+In image based recognition symbols are recognized based on there goemetrical, structure or other spatial features.
+%The AI based recognition treat the problem as a methods in recognizing symbols mainly imply more time and processing as for the search and rules applied but claims to be more accurate.
+
+%The researchers had divided the sketch recognition problem into Image based and AI Based problem.
 The next section will discuss the methods used by current sketch understanding system.
@@
 
- In section \ref{sec:SecondStep:SketchUnderstanding}, the understanding techniques will be demonstrated.
+ %In section \ref{sec:SecondStep:SketchUnderstanding}, the understanding techniques will be demonstrated.
 Understanding mainly will handle knowledge representation methods, handling uncertainty and finally how researcher in sketch reduced the sketch search space.
 
-\section{First Step: Symbol Recognition}
-\label{sec:FirstStepSymbolRecognition}
+%\section{Symbol Recognition}
+%\label{sec:FirstStepSymbolRecognition}
+ Only few systems permit the user to freely draw strokes and symbols in any order \cite{Vibratory8,visualpattern43,Cali63,statisticalparsing26,physicalmeaning6}.
+Those systems must employ a method to identify the strokes that belongs to the same symbol.
+In\cite{gestureexample12,aideddesgin22,sketchinginterfaces2,sketchinginterfaces2,cognitivesketch18,ewhitboard9}for example the user is constrained to draw each symbol using only one stroke which is not only unnatural but also constrains the type fo symbols that can be drawn.
+Others, constrain the order or style of drawing a symbol\cite{}.
 
-To understand a drawing we need to recognize every symbol in it. Hand-drawn sketches need more processing before it is passed to recognitions algorithms.
+%   A hybrid algorithm was introduced in \cite{earlyprocess} where different sets of segments are generated based on both curvature and speed dominant points, followed by choosing a segmentation with the least error from a generated hybrid set.
+However, this system is limited to recognizing only specific simple geometric shapes with a set of low level recognizers.
+Each low level recognizer is designed to recognize only one geometric shape using spatial and geometric information extracted from input stroke.
+Yu \cite{meanshift10} introduced a \textit{feature area} for each primitive and then computed the segmentation error for different types of primitives based on the computed \textit{feature area}.
+His system achieved good accuracy in simple shapes (square, ellipse,...etc) but did not perform well in complex shapes.
+Paulson and Hammond \cite{Paleosketch08} introduced a set of low level recognizers that were reported to achieve 98\%   accuracy.
+However, their system similar to all low level recognizers in the way that it identifies a small set of simple shapes.% A genetic algorithm was used by \cite{CruveDivisionSwarm} to optimally divide digital curves into lines and curves.
+Chen et al.\cite{CruveDivisionSwarm} uses digital curves scanned from paper as input to the system and did not take advantage of the curvature or local geometric properties of the digital curve.
 
- Firstly, we have to determine how symbols are defined or grouped.
-Some System implies some constrain on user to defined symbol.
-In\cite{gestureexample12,aideddesgin22,sketchinginterfaces2,sketchinginterfaces2,cognitivesketch18,ewhitboard9}for example the user is constrained to draw each symbol using only one stroke which is not only unnatural but also constrains the symbols that can be drawn.
-More recent systems force the user to state when symbol is finished by pausing a period of time or clicking a certain button in the interface.
-Only few systems permit the user to freely draw strokes and symbols in any order \cite{Vibratory8,visualpattern43,Cali63,statisticalparsing26,physicalmeaning6}.
-Those systems must employ a method to detect when a user starts drawing a symbol and when he end.
+%A genetic algorithm was used by \cite{CruveDivisionSwarm} to optimally divide digital curves into lines and curves.
+Chen et al.
+\cite{CruveDivisionSwarm} uses digital curves scanned from paper as input to the system.
+Also, they did not take advantage of the curvature or local geometric properties of the digital curve.
+Yin \cite{PolygonApproximationPSO} used PSO to convert digital curves into polygons, our system adopts  Yin \cite{PolygonApproximationPSO} method and tries to improve it by adding curvature and other local information while segmenting strokes to achieve better segmentations.
+The next section presents the general particle swarm algorithm which is used in both our segmentation algorithms.
+% are presented in section \ref{subsubsec:Discreteparticleswarmalgorithm}
 
-Currently symbol recognition is both image based and feature based or AI-based.
-In image based recognition symbol is recognized based on there features, structure, some image computed feature .
-The AI methods in recognizing symbols mainly imply more time and processing as for the search and rules applied but claims to be more accurate .
+%To understand a drawing we need to recognize every symbol in it. Hand-drawn sketches need more processing before it is passed to recognitions algorithms.
+
+Sketch recognition task is divided into three main component: 1) preporcessing, 2) stroke segmenation and symbol or guesture recognition. The next sections presents a survey of each component
+ %Firstly, we have to determine how symbols are defined or grouped. Some System implies some constrain on user to defined symbol.
 
@@
  A stroke is all the points that lie along the path of the user pen from the moment of his pen down to pen up events.
-Not all those points are important even more, there may be redundancies unwanted points.
+Not all those points are important even more, there may be redundance points.
 Therefore, those strokes will need some preprocessing.
@@
 However, they state that the only problem to there system is that the points near the centroid affect the polar transformations which will effect the matching.
-(To solve it they say that loosing the weight of the pixel near the centroid will improve classification).
+%(To solve it they say that loosing the weight of the pixel near the centroid will improve classification).
 
@@
 
-\section{Second Step: Sketch Understanding}
-\label{sec:SecondStep:SketchUnderstanding}
-
-Sketch understanding is concerned with converting a set of recognized symbols into a whole document that is fully consistent with each other.
+%\section{Second Step: Sketch Understanding}
+%\label{sec:SecondStep:SketchUnderstanding}
+%
+%
+%Sketch understanding is concerned with converting a set of recognized symbols into a whole document that is fully consistent with each other.
 Mainly the recognized symbols are a list of best candidates of that symbol for which the sketch understanding process must choose best fit based on the document structure.
-
-After recognizing the symbols drawn, Understanding the sketch document involve additional processing.
+%
+%
+%After recognizing the symbols drawn, Understanding the sketch document involve additional processing.
 In spite of various recognition systems that had reached a satisfactory symbol recognition.
 The sketch understanding problems are at a standstill.
-
-The recognized symbols must be verified with the rest of the document to fully understand the whole sketch.
+%
+%
+%The recognized symbols must be verified with the rest of the document to fully understand the whole sketch.
 Consistency, ambiguities and uncertainty of the recognized symbols are the major issues that rise in any sketch understanding system.
-Most recent sketch understanding system uses more than just the context knowledge that is always available.
+%Most recent sketch understanding system uses more than just the context knowledge that is always available.
 Domain knowledge \cite {geometrydomain49}, Speech and verbal description of drawn sketch was used in\cite {speechMulti27}.Some verbal description of the drawn symbol helps in creating the casual model for each symbol.
-
-In the following sections, the ambiguities and uncertainty of sketch understanding systems will be discussed.
+%
+%In the following sections, the ambiguities and uncertainty of sketch understanding systems will be discussed.
 Followed by how those systems represented knowledge that they use to formulate the understanding process.
-
-\subsection{Handling uncertainty and ambiguities }
-\label{sec:Handlinguncertaintyandambiguities }
-
-The main tasks of the sketch understanding are to handle higher-level ambiguities of the sketch drawn and to manage uncertainty of the recognized symbols.
+%
+%\subsection{Handling uncertainty and ambiguities }
+%\label{sec:Handlinguncertaintyandambiguities }
+%
+%The main tasks of the sketch understanding are to handle higher-level ambiguities of the sketch drawn and to manage uncertainty of the recognized symbols.
 After the user draw a symbol using a single or multi-stroke symbol the recognizer labels each symbol with a known class either low level interpolation or high level.
@@
 They do so with preserving the sketch consistency of interpretation, which means that, the whole document has the same interpretation.
-
-The next section the various methods that was recently employ to completely understand sketched documents.
-
-\subsubsection{Rank and prune based on classifiers}
-\label{sec:Rankandprunebasedonclassifiers}
-
-The most used method in understanding the whole document and reducing the inconsistent solutions is the typical method of ranking the solutions and then pruning unwanted ones. 
-Initially the symbols nodes are initialized by all possible pairs of interpolation of the two symbols then the Consistency checker eliminate all inconsistence pairs.
+%
+%The next section the various methods that was recently employ to completely understand sketched documents. 
+%
+%\subsubsection{Rank and prune based on classifiers}
+%\label{sec:Rankandprunebasedonclassifiers}
+%
+%The most used method in understanding the whole document and reducing the inconsistent solutions is the typical method of ranking the solutions and then pruning unwanted ones. 
+%Initially the symbols nodes are initialized by all possible pairs of interpolation of the two symbols then the Consistency checker eliminate all inconsistence pairs.
 \cite {tobeadd} \cite {geometrydomain49}  firstly recognize all possible interpolation then prune and reason using knowledge to get a score for each interpolation.
 The final step is resolving and choosing the best result.
-
- \cite {Alvarado2002Framework11,HierarchicalParsing7}  First uses a bottom-top search to look for the interpolations of the strokes drawn.
+%
+% \cite {Alvarado2002Framework11,HierarchicalParsing7}  First uses a bottom-top search to look for the interpolations of the strokes drawn.
 A top bottom search is used to prune out incomplete interpolations that was generated in the previous step.
 The top bottom firstly checks what symbol may be missing in the current document context after that it prune inconsistent interpolations.
-\subsubsection{Constrain satisfaction methods}
-\label{sec:Constrainsatisfactionmethods}
-
-Constrain satisfaction is the most used method for converting a group of recognized symbols into an understood document.
+%\subsubsection{Constrain satisfaction methods}
+%\label{sec:Constrainsatisfactionmethods}
+%
+%Constrain satisfaction is the most used method for converting a group of recognized symbols into an understood document.
 Most recognizers produce a ranked set of possible interpolations.
@@
 \cite {tobeadd}  checked the constrains applied on this system and if they produce no match the system relax the constrain and runs again with more loss constrains till it find the correct interpolations.
-
-\begin{figure}
-		\includegraphics[scale=0.7]{../../neededfiles/Figures/CSPphysical.eps}
-	\caption[Physical Constrains on Symbols]{ The figure shows two nodes and the constrain applied on both nodes to detect the inconsistency in the interpretations \cite{physicalmeaning6}. }
-	\label{fig:CSPphysical}
-\end{figure}
-
- \cite {physicalmeaning6}  firstly try connection between symbols firstly using bounding boxes of the symbol to locate the intersecting symbols.
+%
+%\begin{figure}
+%		\includegraphics[scale=0.7]{../../neededfiles/Figures/CSPphysical.eps}
+%	\caption[Physical Constrains on Symbols]{ The figure shows two nodes and the constrain applied on both nodes to detect the inconsistency in the interpretations \cite{physicalmeaning6}. }
+%	\label{fig:CSPphysical}
+%\end{figure}
+%
+% \cite {physicalmeaning6}  firstly try connection between symbols firstly using bounding boxes of the symbol to locate the intersecting symbols.
 The intersected bounding boxes are checked again for ten different types of connection approximately for each terminal of the symbol.
@@
 This step is applied iteratively till all consistency are cancelled and only one global interpolation is reached \textit{(see Fig[ \ref{fig:CSPphysical}])}.
-
- The use of constrain satisfaction has also helped in minimizing the search \cite {tobeadd} .
+% 
+% The use of constrain satisfaction has also helped in minimizing the search \cite {tobeadd} .
 The strokes are incrementally processed using the Constrain system which check for constrain between image feature.
 Each new stroke either a new partial interpolation is created or an old one is extended and cloned provided that no constrained is violated.
-
-\subsubsection{Bayesians networks}
-\label{sec:Bayesiansnetworks}
-
-Since the problem of deciding the best interpretation of a set of valid ones is a certainty problem researchers tried to solve it as a one.
+%
+%
+%\subsubsection{Bayesians networks}
+%\label{sec:Bayesiansnetworks}
+%
+%
+%Since the problem of deciding the best interpretation of a set of valid ones is a certainty problem researchers tried to solve it as a one.
 Bayesians networks were used to represent certainties in sketch documents\cite{SketchRead2007} .Bayesians classifiers were also used to help the understanding problem in various ways.\cite{statisticalparsing26}  Used Bayesians classifiers to help the statistical parser to recognize the best candidate based on the system rules.
-
-\cite{Alvarado2002Framework11,SketchRead2007} is a complete sketch-understating framework that used top bottom and bottom up search of interpolation of the strokes.
+%
+%\cite{Alvarado2002Framework11,SketchRead2007} is a complete sketch-understating framework that used top bottom and bottom up search of interpolation of the strokes.
 The system represented uncertainty using Bayesians networks.
 Each compound object is represented by a network in which the object is the root \textit{see Fig[\ref {fig:anddefnetwork}]}.
- \begin{figure}
-		\includegraphics[scale=0.7]{../../neededfiles/Figures/anddefnetwork.eps}
-	\caption[Bayesian network]{Bayesian network of an And Gate as defined in \cite{Alvarado2002Framework11}}
-	\label{fig:anddefnetwork}
-\end{figure}
-
-\subsubsection{Other Methods}
-\label{sec:OtherMethods}
-
-Other Method were used in \cite{agent47,sketchinginterfaces2}to understand the sketch documents .
+% \begin{figure}
+%		\includegraphics[scale=0.7]{../../neededfiles/Figures/anddefnetwork.eps}
+%	\caption[Bayesian network]{Bayesian network of an And Gate as defined in \cite{Alvarado2002Framework11}}
+%	\label{fig:anddefnetwork}
+%\end{figure}
+%
+%\subsubsection{Other Methods}
+%\label{sec:OtherMethods}
+%
+%Other Method were used in \cite{agent47,sketchinginterfaces2}to understand the sketch documents .
 For example \cite{sketchinginterfaces2} defined  visual and domain rules that applies on the sketch to finally understand the whole document.
@@
 The agent will either report a feature on the strokes drawn or disappear till new strokes are drawn.
-
-\subsection{Representing knowledge }
-\label{sec:Representingknowledge }
-
-Representing knowledge is the one of the challenges that faced the researcher in any AI program.
+%
+%\subsection{Representing knowledge }
+%\label{sec:Representingknowledge }
+%
+%
+%Representing knowledge is the one of the challenges that faced the researcher in any AI program.
 In sketch recognition and understanding, the knowledge can be from more than one resource.
@@
 Therefore, the knowledge must be represented in special method that insures high speed and accuracy of understanding of the drawn sketch.
-
-Systems tried to make uses of more than one kind of knowledge to help in there recognitions and understandings.
+%
+%Systems tried to make uses of more than one kind of knowledge to help in there recognitions and understandings.
 Spatial and visual knowledge of the drawn symbols are widely used in the sketch systems that relies on grammar languages \cite{GenericHMM28,statisticalparsing26,SketchRead2007} .
@@
 Domain and grammar knowledge were also used in \cite {geometrydomain49}  .
-
-Black board architecture \cite {blackboardur} was widely used to represent the various source and forms of data in the sketch understanding systems.
+%
+%
+%Black board architecture \cite {blackboardur} was widely used to represent the various source and forms of data in the sketch understanding systems.
 In \cite {Alvarado2002Framework11}  black board was used to represent incremental knowledge which increase by each symbol drawn in the sketch.
@@
 When any one of them find new information he write it down on the black board soon all information will be in the black board for all solver to see and add.
-
-\subsubsection{	Spatial and Visual knowledge}
-\label{sec:SpatialandVisualknowlege}
-
-The visual and spatial information of the drawn symbols with respect to other symbols in the sketch helps in identifying the relation of these symbols with each other.
+%
+%\subsubsection{	Spatial and Visual knowledge}
+%\label{sec:SpatialandVisualknowlege}
+%
+%The visual and spatial information of the drawn symbols with respect to other symbols in the sketch helps in identifying the relation of these symbols with each other.
 Those information was used in \cite {Cali63,visualpattern43,incrmentintention41,geometrydomain49,marksacts40,sketchinginterfaces2,captureknowledge19}  to help the recognition and understanding of sketch drawn.
-
-\cite{SRGraph57}  Introduced a new spatial relationship graph to represent those spatial relation of one symbol to the other.
+%
+%\cite{SRGraph57}  Introduced a new spatial relationship graph to represent those spatial relation of one symbol to the other.
 The Spatial relation graph is build from the sketch drawn using primitive spatial adjacency relation and relative position relation.
 The produced graphs are then matched using a graph-matching algorithm with the database of graph to recognize best fit.
-
-The spatial and visual information are used in \cite{XPGParser59} to help in building a parser that will parse the sketch document into symbols and objects.
-
-\subsubsection{Domain and Context knowledge}
-\label{sec:DomainandContextknowledge}
-
-\begin{figure}
-	\centering
-		\includegraphics[scale=0.8]{../../neededfiles/Figures/domainexamples.eps}
-	\caption[domain examples]{Examples of sketched documents from different domains}
-	\label{fig:domainexamples}
-\end{figure}
-
-Users can draw sketches in different domains \textit{see Fig [\ref{fig:domainexamples}]}and Until now there have not been a sketch understanding system that can identify drawing in any sketched domain.
+%
+%The spatial and visual information are used in \cite{XPGParser59} to help in building a parser that will parse the sketch document into symbols and objects. 
+%
+%\subsubsection{Domain and Context knowledge}
+%\label{sec:DomainandContextknowledge}
+%
+%\begin{figure}
+%	\centering
+%		\includegraphics[scale=0.8]{../../neededfiles/Figures/domainexamples.eps}
+%	\caption[domain examples]{Examples of sketched documents from different domains}
+%	\label{fig:domainexamples}
+%\end{figure}
+%
+%Users can draw sketches in different domains \textit{see Fig [\ref{fig:domainexamples}]}and Until now there have not been a sketch understanding system that can identify drawing in any sketched domain.
 To do so, the system must be multi-domain system and employ a kind of domain identification or classification.
 Current systems uses domain and context knowledge to assist in the understanding procedure\cite {geometrydomain49,RetargetableInteractive58,sketchunderstanding1, HierarchicalParsing7,Mathpad46} .
-
-ASSIST \cite{AlvaradoFreedom42} uses domain knowledge, context memory and user feedback to help in understanding procedure.
+%
+%
+%ASSIST \cite{AlvaradoFreedom42} uses domain knowledge, context memory and user feedback to help in understanding procedure.
 Domain knowledge to get a set of possible interpretations then context knowledge is used to prune out inconsistent finally the user feedback is used to correct error interoperations.
-
-The \cite{incrmentintention41} store domain knowledge in a memory like architecture where It stores the domain knowledge in long-term memory were strokes are firstly checked for possible interpolations.
+%
+%The \cite{incrmentintention41} store domain knowledge in a memory like architecture where It stores the domain knowledge in long-term memory were strokes are firstly checked for possible interpolations.
 Context knowledge and previous interpolations are stored in short-term memory to help more accurate recognition rate.
 The context and domain knowledge are also used to feed rule-based systems that uses visual language rules in making the final document recognitions \cite{interpretationmechanical50}.
-
+%
+%
 %&&&&&&&&&&&&&&&&oksearch 
-\subsubsection{Using speech and various other knowledge}
-\label{sec:Usingspeechandvariousotherknowledge}
-
-Multimodal systems are system that uses more than one source of input devices as input to the system.
+%\subsubsection{Using speech and various other knowledge}
+%\label{sec:Usingspeechandvariousotherknowledge}
+%
+%Multimodal systems are system that uses more than one source of input devices as input to the system.
 Theses systems help in providing sketch understanding with more information's and data that will help in the recognition procedure.
 Speech and verbal explanation of drawn symbols are most used data in multimodal system to aid sketch understanding \cite {cognitivesketch18,speechMulti27,computationalmodel16} .
+%
 
-\subsection{Space Search}
-\label{sec:SpaceSearch}
-
-The solutions of any sketch understanding are not found until the problem of minimizing and find an efficient way to search the solution space is solved.
+%\subsection{Space Search}
+%\label{sec:SpaceSearch}
+%
+%The solutions of any sketch understanding are not found until the problem of minimizing and find an efficient way to search the solution space is solved.
 The solution space increase exponentially with each stroke the user draws dependably on the system.
 Either ways the space search problem must be addressed to properly solve the problem of sketch understanding.
-
-\cite {EfficientAbstract39}  illustrate a new search space minimization where the recognizer can perform one of the checking for constrains between images feature.
+%
+%\cite {EfficientAbstract39}  illustrate a new search space minimization where the recognizer can perform one of the checking for constrains between images feature.
 Check for partial interpretation by assigning an image feature to a model feature or the user draws delaying until more strokes.

### 2009-04-07 draft/System.tex
april 7

@@
 
-\chapter{System Details}
-\label{sec:SystemDetails}
+%\chapter{System Details}
+%\label{sec:SystemDetails}
 
@@
    where $E_{ellipse}$ is the error computed by Equation(\ref{eq:circleFit}) and $E_{area}$ is computed as the difference between the actual area of the stroke and the area of the estimated ellipse Equation(\ref{eq:ErrorArea}).
-If confidence value $Conf$ is less than threshold $th_{Ellipse}$\footnote{The threshold used is $th_{Ellipse}=0.7$ } then the stroke is segmented as an ellipse otherwise the system proceeds proceed to the two PSO segmentation algorithms.
+If confidence value $Conf$ is less than threshold $th_{Ellipse}$\footnote{The threshold used is $th_{Ellipse}=0.7$ } then the stroke is segmented as an ellipse otherwise the system proceeds proceed to the two DPSO segmentation algorithms.
 % The confidence value is used to label the stroke as ellipse or un segmented stroke.
@@
 
-   The next section describes the \textit{Discrete Particle Swarm Algorithm (DPSO)} then it proceeds to detail the two DPSO segmentation algorithms.
+ %  The next section describes the \textit{Discrete Particle Swarm Algorithm (DPSO)} then it proceeds to detail the two DPSO segmentation algorithms.
 
@@
 
-\subsection{Discrete Particle Swarm Algorithm}
-\label{sec:ParticleSwarmAlgorithm}
-%\section{Particle Swarm Algorithm}
-%\label{PSO}
-%What is particle swarm algorithm and how it was used in related researches. 
-The main idea of \textit{Particle Swarm Algorithm (PSO)} is to represent each agent with a particle from the solution space \cite{PSOFirst}.
+%\subsection{Discrete Particle Swarm Algorithm}
+%\label{sec:ParticleSwarmAlgorithm}
+%%\section{Particle Swarm Algorithm}
+%%\label{PSO}
+%%What is particle swarm algorithm and how it was used in related researches. 
+%The main idea of \textit{Particle Swarm Algorithm (PSO)} is to represent each agent with a particle from the solution space \cite{PSOFirst}.
 Each agent moves the particle with a direction and velocity $v_{ij}$ based on equations \ref{eq:Swarm} \& \ref{eq:Swarm1}.
-\begin{equation}
-%\[
-p_{ij}=p_{ij}+v_{ij},
-%\
-\label{eq:Swarm1}
-\end{equation}
-where $p_{ij}$ represent the $jth$ particle in the $ith$ agent and $v_{ij}$ is the velocity of the $jth$ particle in the $ith$ agent.
- %Equation [\ref{eq:Swarm}] shows how velocity and direction of each particle are computed
- \begin{equation}
-v_{ij}  = v_{ij}  + c_1 r_1 (lbest_{ij}  - p_{ij} ) + c_2 r_2 (gbest_{ij}  - p_{ij} )
-\label{eq:Swarm}
-\end{equation}
- where $lbest_{ij}$ is the local best particle, $gbest_{ij}$ is the global best particle, $r_1$ \& $r_2$ are random variables and $c_1$ \& $c_2$ are the swarm system variables.
- After each iteration the global best $g_{best}$ particle and the agent local best $l_{best}$ particle are evaluated based on the maximum fitness functions of all particles in the solution space.
+%\begin{equation}
+%%\[
+%p_{ij}=p_{ij}+v_{ij},
+%%\
+%\label{eq:Swarm1}
+%\end{equation}
+%where $p_{ij}$ represent the $jth$ particle in the $ith$ agent and $v_{ij}$ is the velocity of the $jth$ particle in the $ith$ agent.
+% %Equation [\ref{eq:Swarm}] shows how velocity and direction of each particle are computed
+% \begin{equation}
+%v_{ij}  = v_{ij}  + c_1 r_1 (lbest_{ij}  - p_{ij} ) + c_2 r_2 (gbest_{ij}  - p_{ij} )
+%\label{eq:Swarm}
+%\end{equation}
+% where $lbest_{ij}$ is the local best particle, $gbest_{ij}$ is the global best particle, $r_1$ \& $r_2$ are random variables and $c_1$ \& $c_2$ are the swarm system variables.
+% After each iteration the global best $g_{best}$ particle and the agent local best $l_{best}$ particle are evaluated based on the maximum fitness functions of all particles in the solution space.
 The solution is found after achieving a specific number of iteration or after an error threshold is achieved.
-Equation \ref{eq:descrite}  
-\begin{equation}
-   P(i)\Leftarrow 
-\{
-\begin{array}{c} 
-1 \quad \quad if\quad r_{3}>p_{i}  \\
-
-0 \quad \quad if\quad r_{3}<p_{i} 
-\label{eq:descrite}
-\end{array}\}
-\end{equation}
- where $p_{ij}$ is the numerical values of the particle and $r_{3}$ is a random variable, is used to change the general swarm algorithm into binary particle (\textit{Discrete Particle Swarm Algorithm DPSO}) which handles particle values of either $0$ or $1$ \cite{PSODisceret}.
+%Equation \ref{eq:descrite}  
+%\begin{equation}
+%   P(i)\Leftarrow 
+%\{
+%\begin{array}{c} 
+%1 \quad \quad if\quad r_{3}>p_{i}  \\
+%
+%0 \quad \quad if\quad r_{3}<p_{i} 
+%\label{eq:descrite}
+%\end{array}\}
+%\end{equation}
+% where $p_{ij}$ is the numerical values of the particle and $r_{3}$ is a random variable, is used to change the general swarm algorithm into binary particle (\textit{Discrete Particle Swarm Algorithm DPSO}) which handles particle values of either $0$ or $1$ \cite{PSODisceret}.
 \subsection{Swarm Segmentation}

### 2009-04-07 draft/chapterSurvey.tex
april 7

--- 2009-04-06 trunk/draft/chapterSurvey.tex
@@
 
-\section{Introduction }
-\label{sec:Introduction}
+%\section{Introduction }
+%\label{sec:Introduction}
+The history of pen and sketch based computing goes back to 1963 with Ivan Sutherland's SKETCHPAD \cite{}, which used a light pen to draw on the monitor to create circuit diagrams.
+Rubine \cite{} Guesture recognition system.
 
+In this chapter, we will compare our system with some of the systems that support free-hand sketching.
+In general, these systems lack either one or more of the following properties that we believe a sketching system should have:
+\begin{enumerate}
+	\item It should be possible to draw arbitrary shapes with a single stroke, (i.e., without requiring the user to draw objects in pieces).
+\item It should be possible using multiple strokes without a specific order, style or orientation. 
+\item The system should do automatic stroke segmenation. The user should not ave to specify vertex positions manually. 
+\item The sketching should be as natural as possible. 
+\end{enumerate}
 A wide variety of techniques were suggested for sketch recognition.
@@
 
-\subsection{Pre processing }
+\section{Pre processing and Segmentation}
 \label {sec:preprocessing}
- A stroke is all the points that lie along the path of the user pen from the moment of his pen down to pen up events.
+% A stroke is all the points that lie along the path of the user pen from the moment of his pen down to pen up events.
 Not all those points are important even more, there may be redundance points.
@@
 
-\cite{Mathpad46}Handle rotations invariant using polar transformation then shifting to make it more reliable when template matching.
+%\cite{Mathpad46} Handle rotations invariant using polar transformation then shifting to make it more reliable when template matching.
 However, they state that the only problem to there system is that the points near the centroid affect the polar transformations which will effect the matching.
@@
 
-From the image-based view the symbol is a sequences of line segments or/and curves. To do so without constraining the user drawing style we have to easily divide the user drawn strokes.
+%From the image-based view the symbol is a sequences of line segments or/and curves. To do so without constraining the user drawing style we have to easily divide the user drawn strokes.
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%This secton will be removed only mention the mean shift to be replaced by 
+In \cite{meanshift10,domainindependent17}, Bo Yu uses mean shift procedure as a noise remover and preprocessing step to the recognition step.
+The mean shift procedure is simply shifting the point to the average of its neighborhoods so smoothing out the noise without changing the global structure of data.
+This process is clamed to help find critical points easily after that.
+%% now add the othero one in that used gaussian filter ..
+and sayit smooths the cruves....
+
+ \cite {overtraced24} tried to solve the over traced strokes.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsubsection{Removing noise}
@@
 
+\section{Sketch recogntion task}
+\label{sec:sketchRecogntionTask}
 \subsection{Image Based Classification}
@@
 Every new stroke is drawn either a new partial interpolation is created an old interpolation is extended and cloned provided that no constrains or rule is violated.
+
+\chapter{Particle Swram Optimization}
+\label{sec:ParticleSwramOptimization}
+
+\section{Particle Swram Optimization Review}
+\label{sec:ParticleSwramOptimizationReview}
+
+%review on the pso alogrithm alone.................
+
+\section{Discrete Particle Swarm Algorithm}
+\label{sec:ParticleSwarmAlgorithm}
+%\section{Particle Swarm Algorithm}
+%\label{PSO}
+%What is particle swarm algorithm and how it was used in related researches. 
+The main idea of \textit{Particle Swarm Algorithm (PSO)} is to represent each agent with a particle from the solution space \cite{PSOFirst}.
+Each agent moves the particle with a direction and velocity $v_{ij}$ based on equations \ref{eq:Swarm} \& \ref{eq:Swarm1}.
+\begin{equation}
+%\[
+p_{ij}=p_{ij}+v_{ij},
+%\
+\label{eq:Swarm1}
+\end{equation}
+where $p_{ij}$ represent the $jth$ particle in the $ith$ agent and $v_{ij}$ is the velocity of the $jth$ particle in the $ith$ agent.
+ %Equation [\ref{eq:Swarm}] shows how velocity and direction of each particle are computed
+ \begin{equation}
+v_{ij}  = v_{ij}  + c_1 r_1 (lbest_{ij}  - p_{ij} ) + c_2 r_2 (gbest_{ij}  - p_{ij} )
+\label{eq:Swarm}
+\end{equation}
+ where $lbest_{ij}$ is the local best particle, $gbest_{ij}$ is the global best particle, $r_1$ \& $r_2$ are random variables and $c_1$ \& $c_2$ are the swarm system variables.
+ After each iteration the global best $g_{best}$ particle and the agent local best $l_{best}$ particle are evaluated based on the maximum fitness functions of all particles in the solution space.
+The solution is found after achieving a specific number of iteration or after an error threshold is achieved.
+Equation \ref{eq:descrite}  
+\begin{equation}
+   P(i)\Leftarrow 
+\{
+\begin{array}{c} 
+1 \quad \quad if\quad r_{3}>p_{i}  \\
+
+0 \quad \quad if\quad r_{3}<p_{i} 
+\label{eq:descrite}
+\end{array}\}
+\end{equation}
+ where $p_{ij}$ is the numerical values of the particle and $r_{3}$ is a random variable, is used to change the general swarm algorithm into binary particle (\textit{Discrete Particle Swarm Algorithm DPSO}) which handles particle values of either $0$ or $1$ \cite{PSODisceret}.

### 2009-04-09 trunk/draft/System.tex
now it 9 - appril and after club day

--- 2009-04-07 draft/System.tex
@@
 \chapter{The Proposed System}
+\label{sec:proposedSystem}
 \section{Aims of the System}
 \label{sec:AimsOfTheSystem}
-The previous chapter shows the current trends in sketch recognition.
-This thesis introduces a solution for sketch recognition problem.
-The research gives an optimize solution for the sketch segmentation problem.
+The previous chapter showed the current trends in sketch recognition.
+This thesis introduces a new approach for the sketch recognition problem.
+Particle Swarm Optimization (PSO) is used to correctly segment strokes into curves and lines.
 Particle Swarm Optimization (PSO) is used to correctly segment strokes into curves and lines.
@@
 
-\section{An Overview of the System}
+\section{System Overview}
 \label{sec:AnOverviewOfTheSystem}
@@
 %The stroke is the path of points from the instant the pen is down till it is up.
-The block diagram in Figure \ref{fig:Blockdiagram} shows that the main system blocks are 1) preprocessing, 2) segmentation, 3) feature extraction, and finally 4) classification.
-The next section will describe each component in details.
+The block diagram in Figure \ref{fig:Blockdiagram} shows the main system blocks. The next section describes each block in details.
  %Due to sloppiness of users and hardware glitches the system processes the data before further
@@
 
-\subsection{System Components}
-\label{sec:SystemComponents}
+%\subsection{System Components}
+%\label{sec:SystemComponents}
 
-\textbf{Preprocessing} step is responsible of capturing the input data and removing the noise from it.
-The preprocessing block as in Figure \ref{fig:Blockdiagram} consist of computing the time differences, direction, speed and curvature of every point in the stroke.
-After computing this information the local minimum and local maximum of each curve is extracted from the data computed.
-This process estimate a set of a set of \textit{Possible Dominant Points $P_{pd}$} that guide the segmentation process.
+The \textbf{preprocessing} step is  is responsible for capturing the input data and removing the noise from it.
+As shown in Figure \ref{fig:Blockdiagram} preprocessing consists of computing the time differences, direction, speed and curvature of every point in the stroke.
+After computing this information the local minimum and local maximum of each curve is extracted.
+This process estimates a set of a set of Possible Dominant Points $P_{pd}$ that guide the segmentation process.
 Later, other geometrical and statistical computations are evaluated from the stroke.
@@
 
+\begin{figure}[]
+	\centering
+		\includegraphics[scale=0.8]{images/orignalStroke.eps}
+	\caption{Input Stroke} Example of an input stroke to the system. 
+	\label{fig:orignalStroke}
+\end{figure}
+
 \begin{figure}
@@
 	\label{fig:speed2Distance}
-\end{figure}
-
-\begin{figure}[]
-	\centering
-		\includegraphics[scale=0.8]{images/orignalStroke.eps}
-	\caption{Input Stroke} Example of an input stroke to the system. 
-	\label{fig:orignalStroke}
 \end{figure}
@@
 \label{sec:CriticalPointDetection}
-After the system computes all the speed, time difference and curvature data it proceeds to detect the points with low velocity and high curvatures.
+After computing speed, time difference and curvature information the system proceeds to detect the points with low velocity and high curvatures.
 Using simple differentiation to detect local extreme points resulted in false points due to the non smooth curves.
 Hence, the system adopted a process presented by \cite{earlyprocess}, where the mean of the curve is calculated.
-Then a threshold \textit{th} is used to separate the curve into regions $Region_i$; each region $Region_i$ is defined as a range of points, where the curve values are either above or below the threshold \textit{th}.
+Then a threshold \textit{th} is used to separate the curve into regions; each region $Region_i$ is defined as a range of points, where the curve values are either above or below the threshold \textit{th}.
 Those regions are further processed to find the maximum point $Max(Region_i)$ of each region $Region_i$.
 The stroke points $p_i(x,y)$ that correspond to those maximum values are labeled as \textit{possible dominant points} $P_{pd}$.
-Figure \ref{fig:MaxRegioi} shows an example of $Region_i$ of the speed curve and the dominate points $P_{pd}$ which correspond the minimum of point in this region $Min(Region_i)$.
+Figure \ref{fig:MaxRegioi} shows an example of $Region_i$ of the speed curve and the dominate points $P_{pd}$ which correspond to the minimum of point in this region $Min(Region_i)$.
 % (as shown are redundant)
@@
 The array that contains possible points $P_{pd}$ is used as input for the next segmentation step.
-Figure \ref{fig:LabelsPPD} shows the particles labeled as Possible dominate point $P_{pd}$ by the preprocessing, it is noticed the redundancy of some $P_{pd}$ points.
+Figure \ref{fig:LabelsPPD} shows the particles labeled as Possible dominate points $P_{pd}$ by the preprocessing, it is noticed the redundancy of some $P_{pd}$ points.
 
@@
    where $E_{ellipse}$ is the error computed by Equation(\ref{eq:circleFit}) and $E_{area}$ is computed as the difference between the actual area of the stroke and the area of the estimated ellipse Equation(\ref{eq:ErrorArea}).
-If confidence value $Conf$ is less than threshold $th_{Ellipse}$\footnote{The threshold used is $th_{Ellipse}=0.7$ } then the stroke is segmented as an ellipse otherwise the system proceeds proceed to the two DPSO segmentation algorithms.
+If confidence value $Conf$ is less than threshold $th_{Ellipse}$\footnote{ Using trial and error the best threshold used is $th_{Ellipse}=0.7$ } the stroke is segmented as an ellipse otherwise the system proceeds to the two DPSO segmentation algorithms.
 % The confidence value is used to label the stroke as ellipse or un segmented stroke.
@@
 After generating two segmentations the system chooses the best segmentation from the outputs of the DPSO.
-The details of the problem formulation and fitness functions are in the following sections.
+The details of the problem formulation and fitness functions are given in the following sections.
 %In this process the system generate two stroke segmentation using two PSO segmentation algorithm.
@@
 \end{equation}
- The error of the circle estimation is measured using the distance $d_c$ which is calculated using the equation \ref{eq:circleE}:
+ The error of the circle estimation is measured using the distance $d_c$ which is calculated using equation \ref{eq:circleE}:
  \begin{equation}
@@
 \subsubsection{Solution Refine procedures } 
-\cite{PolygonApproximationPSO} used a merge and divide algorithm after each loop of the swarm system to refine the solution but we used another enhancement method.
+%\cite{PolygonApproximationPSO} used a merge and divide algorithm after each loop of the swarm system to refine the solution but we used another enhancement method.
  After each loop in the swarm algorithm, each particle loops on the set of selected dominant points to enhance the solution.
@@
 
- \section{Benchmark algorithm}
+ \section{Hybird Segmentation Algorithm}
 
 \label{sec:BenchMarckAlgorithm}
-In this research, a bench mark algorithm is used as reference to the results obtained by the DPSO algorithms.
-The algorithm was first implemented by \citeauthor{earlyprocess} in \cite{earlyprocess}.
+In this research, we developed another algorithm as reference to the results obtained by the DPSO algorithms.
+The algorithm was first implemented by  Sezgin et al.
+in \cite{earlyprocess}.
 The algorithm is divided into the following steps:
@@
 \end{enumerate}
-More details about the algorithm is in \cite{earlyprocess}.
+More details about the algorithm are in \cite{earlyprocess}.
 
@@
 The table \ref{tab:FeatureTable} gives details on all the features used in the system.
-\begin{longtable}{|l|l|p{6cm}|p{5cm}|}
-	%\caption{Feature Table}
-	%\label{tab:FeatureTable}
-\hline 
-Type  & Index  & Features  & Notes \tabularnewline
-\endhead
-\hline 
-Type  & Index  & Features  & Notes \tabularnewline
-\endfirsthead
-\hline 
-Structural & \multicolumn{1}{c||}{} &  & \tabularnewline
-\hline 
- & F1.  & No. of parallel lines  & \tabularnewline
-\hline 
- & F2.  & No. of perpendicular lines  & \tabularnewline
-\hline 
- & F3.  & No. of intersection between of any two segmetns  & \tabularnewline
-\hline 
- & F4.  & No of primitives & \tabularnewline
-\hline 
- & F5.  & No. of line segments & \tabularnewline
-\hline 
- & F6  & No. of ellipse, circular arcs.  & \tabularnewline
-\hline 
-Statistical &  &  & \tabularnewline
-\hline 
- & F7.  &  & \tabularnewline
-\hline 
- & F8.  &  & \tabularnewline
-\hline 
-Rubine Features  &  & From features F9. till F13  & Introduced by in \cite{gestureexample12} \tabularnewline
-\hline 
- & F9.  & Cosine of starting angle  & \tabularnewline
-\hline 
- & F10.  & Sine of starting angle & \tabularnewline
-\hline 
- & F11.  & Length of diagonal of bounding box & (gives an idea of the size of the bounding box) \tabularnewline
-\hline 
- & F12.  & Angle of diagonal & gives an idea of the shape of the bounding box (long, tall, square) \tabularnewline
-\hline 
- & F13.  & Distance from start to end  & \tabularnewline
-\hline 
- & F14.  & Cosine of ending angle  & \tabularnewline
-\hline 
- & F15.  & Sine of ending angle  & \tabularnewline
-\hline 
- & F16.  & Total stroke length  & \tabularnewline
-\hline 
- & F17  & Change in Rotation( Arctan ) & gives the directional angle \tabularnewline
-\hline 
- & F18  & Absolute rotation  & \tabularnewline
-\hline 
- & F19  & Rotation squared  & \tabularnewline
-\hline 
- & F20  & The maximum speed reached (squared)  & \tabularnewline
-\hline 
- & F21  & Total time of stroke  & \tabularnewline
-%\newpage
-\hline 
-Zernike moments  &  &  & \tabularnewline
-\hline 
- & F22  & Zernike moments to the order  $n$  & used by \cite{zernike61} \tabularnewline
-Composite  &  &  & \tabularnewline
-\hline 
- &  &  & \tabularnewline
-\hline 
- &  &  & \tabularnewline
-\hline 
- &  &  & \tabularnewline
-\hline
-\end{longtable}
+
+\begin{description}
+ \item [Structural and geometrical Features(FS1)] Features defines the structure of the geometrical symbol.  
+\begin{description}
+	 \item[No. Of segments] Number of segments in the symbol. 
+		\item[No. Of primitives] Number of primitives in the symbol. The feature helps when identifying             symbols with mixed geometric primitives like cylinders and callouts.  
+		\item [No. of curves] Number of curves or ellipses in the symbol. 
+		\item [No. of lines] Number of lines in the symbol. 
+		\item [No. of perpendicular lines] Number of perpendicular lines. 
+		\item [No. of parallel lines] Number of parallel lines. 
+		\item [No. of intersections] Number of intersection between lines and curves. 
+		\item [Size Ratio] Ratio between width to height of the symbol.
+		%\item[No. of holes] Number of holes in the symbols.
+
+\end{description}
+\item [Rubine Feature Set (FS2)] Features introduced by Rubine\cite{gestureexample12} for single stroke gestures.
+To compute the features we append all segments points into single path $ink_{path}$.
+The features is computed on the generated path $ink_{path}$.%then compute the features based on this path.
+\begin{enumerate}
+	\item Cosine of starting angle.
+	\item Sine of starting angle.
+	\item Length of diagonal of bounding box.It gives an idea of the size of the bounding box).
+	\item Angle of diagonal. It gives an idea of the shape of the bounding box (long, tall, square).
+	\item Distance from start to end.  
+	\item Total stroke length
+	\item Change in Rotation( Arctan ). It  gives the directional angle.
+	\item Absolute rotation 
+	\item Rotation squared 
+	\item The maximum speed reached (squared) 
+	\item  Total time of stroke 
+\end{enumerate}
+
+\item[Statistical Features (FS3)]   
+\begin{description}
+	\item[Zernike moments]  Zernike moments of order \cite{HeloiseBeautification}. 
+\end{description}
+	\item[Composite Features (FS4)] Features that are composites of different geometrical and statistical symbol. 
+	\begin{description}
+
+	\item [Ink density] Compute the density of points inside its bounding box\cite{GeometryAndDomain102}.   
+ 	\item[Convex Hull Area] Area of convex hull with respect to area of bounding box of symbol.
+	\item[Convex Hull Perimeter] Perimeter of convex hull with respect to total length of symbol.
+		\item [Mean Centroidal radius] The Mean of the centroidal radius which is the distance from each point in the symbol to the center of gravity.
+
+	\item [Mean Time difference] The mean of the time difference between each two successive points in the symbol. %Different strokes are appended to construct a single path.
+	%\item [Ra]
+  \end{description}
+
+\end{description}
+%
+%\begin{longtable}{|l|l|p{6cm}|p{5cm}|}
+%	%\caption{Feature Table}
+%	%\label{tab:FeatureTable}
+%\hline 
+%Type  & Index  & Features  & Notes \tabularnewline
+%\endhead
+%\hline 
+%Type  & Index  & Features  & Notes \tabularnewline
+%\endfirsthead
+%\hline 
+%Structural & \multicolumn{1}{c||}{} &  & \tabularnewline
+%\hline 
+% & F1.  & No. of parallel lines  & \tabularnewline
+%\hline 
+% & F2.  & No. of perpendicular lines  & \tabularnewline
+%\hline 
+% & F3.  & No. of intersection between of any two segmetns  & \tabularnewline
+%\hline 
+% & F4.  & No of primitives & \tabularnewline
+%\hline 
+% & F5.  & No. of line segments & \tabularnewline
+%\hline 
+% & F6  & No. of ellipse, circular arcs.  & \tabularnewline
+%\hline 
+%Statistical &  &  & \tabularnewline
+%\hline 
+% & F7.  &  & \tabularnewline
+%\hline 
+% & F8.  &  & \tabularnewline
+%\hline 
+%Rubine Features  &  & From features F9. till F13  & Introduced by in \cite{gestureexample12} \tabularnewline
+%\hline 
+% & F9.  & Cosine of starting angle  & \tabularnewline
+%\hline 
+% & F10.  & Sine of starting angle & \tabularnewline
+%\hline 
+% & F11.  & Length of diagonal of bounding box & (gives an idea of the size of the bounding box) \tabularnewline
+%\hline 
+% & F12.  & Angle of diagonal & gives an idea of the shape of the bounding box (long, tall, square) \tabularnewline
+%\hline 
+% & F13.  & Distance from start to end  & \tabularnewline
+%\hline 
+% & F14.  & Cosine of ending angle  & \tabularnewline
+%\hline 
+% & F15.  & Sine of ending angle  & \tabularnewline
+%\hline 
+% & F16.  & Total stroke length  & \tabularnewline
+%\hline 
+% & F17  & Change in Rotation( Arctan ) & gives the directional angle \tabularnewline
+%\hline 
+% & F18  & Absolute rotation  & \tabularnewline
+%\hline 
+% & F19  & Rotation squared  & \tabularnewline
+%\hline 
+% & F20  & The maximum speed reached (squared)  & \tabularnewline
+%\hline 
+% & F21  & Total time of stroke  & \tabularnewline
+%%\newpage
+%\hline 
+%Zernike moments  &  &  & \tabularnewline
+%\hline 
+% & F22  & Zernike moments to the order  $n$  & used by \cite{zernike61} \tabularnewline
+%Composite  &  &  & \tabularnewline
+%\hline 
+% &  &  & \tabularnewline
+%\hline 
+% &  &  & \tabularnewline
+%\hline 
+% &  &  & \tabularnewline
+%\hline
+%\end{longtable}
 %\end{table}

### 2009-04-09 draft/ChapterIntroduction.tex
now it 9 - appril and after club day

--- 2008-09-02 trunk/draft/ChapterIntroduction.tex
@@
 %\usepackage{graphicx}
-\chapter {Introduciton}
+\chapter {Introduction}
 \pagenumbering{arabic}
-
-Computer research nowadays aimed at facilitating computer human interaction in every way.
-Pen-based interface gives the user a pencil-paper like feeling that enhances interaction more than the current used keyboard-mouse computer interfaces.
+Computer research nowadays aims at facilitating computer human interaction in every way.
+Pen-based interfaces give the user a pencil-paper like feeling that enhances interaction more than the current used keyboard-mouse computer interfaces.
 However, until now there are hardly any complete pen based computer systems.
 Generally, Current Pen-Based systems make use of the Pen or stylus to perform the same role as the mouse / keyboard.
-The minority of such systems are equipped with handwriting or digit recognition that is typically restricted for specific language.
+A minority of such systems are equipped with handwriting or digit recognition that is typically restricted for specific language.
 
- Despite the fact that people frequently express their ideas through drawing them there isn't a system that can fully understand those sketches.
-Lately, only a few researchers were concerned with the problem although it is highly important.
-Firstly, we need to know;
- \textit { What are the main goals of Sketch understanding systems?}
+\section{Sketch Recognition}
+Sketch recognition is defined as the process of identifying symbols that users draw using single or multiple strokes.
+Users draw strokes using a pen and the system should immediately interprets their strokes into objects that can be easily manipulated.
 
-The main goal of sketch understating systems is to build a system that will interpret the sketches the user draw.
-The systems ought to be natural as possible to provide user with paper and pencil impression.
-Furthermore, interpreted sketches are expected to transform into objects, which can me manipulated, modeled and/or simulated.
-
-Sketch here stand for any hand drawn drawing the user draws using digital stylus or similar device.
-Normally, sketch are composed of smaller segments we call strokes.
+Sketches here stand for any hand drawing the user draws using digital stylus or similar device.
+Normally, sketches are composed of a set of strokes.
 The time between pen-up and pen-down events identify each stroke.
@@
 
-Sketch recognition is procedure that will input a sequence of points from user then segment them into clusters of strokes and recognize those clusters to primitive geometric objects and shapes.
-However, Sketch understanding is transforming the sketch from a group of primitive shapes into comprehendible identified symbols consistent with each other.
-% the entire sketched symbols.
-
 \section{Importance of sketch recognition}
 \label{sec:ImportanceOfSketchRecognition}
-
-People normally sketch thoughts rather than expressing them linguistically.
-For instance engineers mostly drafts their early designs using paper and pencil before using CAD tools.
-People in general and engineers specially make use of various sketches during design process as instance organizational charts, informal drafting and formal design sketches \cite{}.
-The sketching process itself go thorough numerous stages.
-Firstly, early stages with rough drafting designs or notations which are mainly done using paper and pencil.
-After those primary rough sketches are edited and altered until they are stabilized and finally they are transformed into the computer.
+Scientists generally, and engineers specifically express thoughts and designs using sketches.
+Engineers use sketches to exchange designs as a natural method of communication rather than writing or speaking.
+Engineers use paper and pencil design in early stages of design.
+The sketching process itself goes through  numerous stages.
+Firstly, early stages with rough drafting designs or notations are mainly done using paper and pencil.
+These primary rough sketches are edited and altered until they are stabilized and finally they are transformed into the computer.
 
  Paper and Pencil are preferred in design as they offer a remarkable ease in editing and altering the primary rough sketches.
-Current CAD systems don't give the designer the ease of altering the early designs.
-Designer had to sketch primary designs on paper and pencil next, the finally design is transferred into computer using a menu CAD system.
+Current CAD systems do not give the designer the ease of altering the early designs.
+Designer had to sketch primary designs on paper and pencil, next the final design is transferred into computer using a CAD system menus.
 The process is time consuming and tiring.
@@
 
- A sketch recognition and understanding system will provide a powerful tool to the designers to draw sketch as they usually draw using paper and pencil with the advantages of computer program.
-Designer needs a normal sketching environment that will provide him facilities in paper interface and the modifying and simulation that is in current CAD systems.
+ A sketch recognition and understanding system will provide a powerful tool to the designers to draw sketch as they usually draw using paper and pencil with the advantages of a computer program.
+Designers need a normal sketching environment that will provide him facilities in paper interface and the modifying and simulation that is in current CAD systems.
 
-\section{Problems and challenges}
+\section{Applicability }
+% talk about application like cad systems, software engineering uml diagrams ....
+% meeting idea brainstorming , mind mapping string ..
+% communication program  
+% sketching programs ...
+Applications of sketch recognition rang from engineering device simulation to design.
+Drawing and sketching programs are most likely to benefit from integrating a sketch recognition system into their design environments.
+Other applications include object retrieval, map navigations and various engineering design systems.
+
+\section{Problem Definition and challenges}
 \label{sec:ProblemsAndChallenges}
 
- Understanding hand drawn sketches may be trivial for humans but it is challenging problem when speaking about computer-based system.
-Most of the research made on the sketch understanding is achieved for one particular domain or for strictly a set of pre defined symbols.
-Further, others restrict the user the freedom in drawing which loose the main goals of paper-like-interface systems.
-The only few systems that have achieved the desired degree of freedom for the users are extremely computationally expensive.
+Understanding hand drawn sketches may be trivial for humans but it is a challenging problem when speaking about computer-based system.
+Most of the research made on the sketch recognition is achieved for one particular domain or for strictly a set of pre-defined symbols\cite{Vibratory8,physicalmeaning6}.
+Furthermore, others restrict the user the freedom in drawing which loose the main goals of paper-like-interface systems\cite{gestureexample12,aideddesgin22,sketchinginterfaces2}.
+The only few systems that have achieved the desired degree of freedom for the users are extremely computationally expensive\cite{EfficientAbstract39,SketchRead2007}.
 
-The research in the area had faced few problems that will need to be handled to implement an entirely sketch understanding system.
-The user draws strokes in a spontaneous and ambiguous manner that intensifies the challenges.
-The first problem is segmenting strokes from the way the user draw it into primitive geometric.
-Ink parsing is the problem dealing with clustering interpolated strokes and grouping them into symbols.
-After that, recognizing the symbols from the space the system can identify is the final problem.
-All those problems must be solved on a real time constrain to maintain an online sketching system if it was a requirement.
+Sketch recognition is divided into three steps: preprocessing, segmentation and symbol recognition.
+The preprocessing phase captures user input stroke points and collects basic information about the stroke then proceeds to remove noise and compute basic statistical and geometrical information.
+In the segmentation phase, the strokes are divided into a set of simple geometrical primitives or segments.
+In the third phase, sketch recognition, strokes and segments are clustered to formulate symbols that can be recognized by a classifier system.
+There are several problems in this research area that need to be handled to implement an integrated sketch recognition system.
+The fact that the user draws strokes in a spontaneous and ambiguous manner that intensifies the challenges.
+One of the main problems is segmenting spontaneous strokes drawn by the user into geometrical primitive segments.
+Another major problem is the high level of ambiguities in the drawing.
+The next sections describe each problem in details.
 
-The next sections will explain and investigate each of those problems in details.
-
-\subsection{Segmentations}
-\label{sec:Segmentations}
-
-Segmentation is dividing the strokes draws into the geometric primitive's as lines, arcs, and ellipses.
-The strokes as mentioned before are the path of points between pen-down and pen-up events while using a stylus.
-The processing of those points is done after that to firstly to detect the critical points or vertices.
-After that the segment is classified as either line, arc, ellipse or circle.
-The main problem in this process is in detecting the critical and vertices which are points divide strokes into lines and arc.
-The user draw the lines and arc in an undefined manner which make it difficult for the program to deicide where the user had finish a line and started a new arc or vice verse; Fig [\ref {fig:strokeinterpolation}].
-
+\subsection {Segmentation}
+The goal of segmentation is dividing the strokes into the geometric primitive's as lines, arcs, and ellipses.
+The strokes, as mentioned before are the path of points between pen-down and pen-up events while using a stylus.
+Furthermore, after identifying a stroke, it should be processed to detect the critical points or vertices which divide the stroke into segments.
+Later, each segment is classified as line, arc, ellipse or circle.
+The main problem in this process is in detecting the critical points and vertices which divide strokes into lines and arcs.
+The user draws the lines and arcs in an undefined manner which make it difficult for the program to decide where the user had finished a line and started a new arc or vice verse;  Figure \ref {fig:strokeinterpolation}.
 \begin{figure}
 
+\begin{center}
 		\includegraphics[scale=0.8]{../../neededfiles/Figures/strokeinterpolation.eps}
-	\caption[Segmentation error]{The figure shows an arc that the user draw which can be segmented in different ways as a single line or as an arc where a) The stroke users draw  b)interpretation as a single line segment  b) interpretation as an arc.}
+	\caption[Segmentation error]{Different ways for segmenting a line drawn by user. a) The stroke users draw b)interpretation as a single line segment c) interpretation as an arc.}
+	%The figure shows an arc that the user draw which can be segmented in different ways as a single line or as an arc where a) The stroke users draw  b)interpretation as a single line segment b) interpretation as an arc.}
 	\label{fig:strokeinterpolation}
+\end{center}
 \end{figure}
 
-\subsection{Ink parsing}
-\label{sec:InkParsing}
+\subsection{Ambiguity}
+Figure 1 shows an example of ambiguity, where the strokes drawn by the user can be classified in several ways.
+The nature of hand drawn sketch itself is highly ambiguous even for human eye.
+You may see two drawings you cannot decide if they are a square or a rectangle.
 
- The clustering of segmented strokes into group of strokes before passing into the next stage this process is called ink parsing.
-It is simply deciding when a symbols ends and a new symbol begins.
-Ink parsing is most important and difficult challenge as grouping of strokes can be done in many ways and configuration, which increase exponentially with the number of strokes.
-In addition, the ink parsing will affect the next steps of recognizing.
-The user does not draw the symbol in same order.
-Moreover, users can draw a sequence of strokes in symbols then leave it to draw another then return again to add segment to symbols so temporal data is helpful but not conclusive.
-The user may also draw the same symbols using different orders and methods for example the user can draw a rectangle using single stroke or using two, three or even four strokes.
-All those factors gather with the sloppy drawing of the user to increase the complexity of the problem.
-
-\subsection{Ambiguities}
-\label{sec:Ambiguities}
-
-The major problem of sketch understanding is the high level of ambiguities in the drawing.
-The user draws strokes that the computer cannot decide if it is arc or line.
-The nature of hand drawn sketch it self is highly ambiguous even for human eye you may see two drawn you cannot decide if it is square or rectangle.
 The ambiguities in sketch understanding are found at many levels in the system.
-Firstly, at the segmentation phase where the same user stroke can be interpreted as a ellipse and a circle.
-Then the symbol recognition as a symbol can have more than one interpolation.
-Due to the user sloppy drawing the symbols can have more than one interpolation which varies with the domain drawn in it similar problem was faced in handwriting recognition see Fig [\ref{fig:VorU}] but it was handled with knowing information about the context information which may not be the case in sketch recognition.
-
+Firstly, in the segmentation phase, a single stroke can be interpreted as an ellipse or as a circle.
+Symbol ambiguity is another problem as a symbol can have more than one interpolation.
+Due to the user's sloppy drawing the symbols can have more than one interpolation which varies with the domain drawn in it.
+A similar problem was faced in handwriting recognition (Figure \ref{fig:VorU}).
+The problem could be solved by providing context or domain information which could not be provided in most sketches.
 \begin{figure}	
@@
 
-The ambiguity problem must be first recognized ambiguities or varicose interpolations of same symbol.
-After that, resolve those ambiguities is detecting the best correct interpolation of the symbol.
-To solve this problem the program must be able to detect the amount of information needed to decide which interpolation of symbols is the best one.
+\section{Research Scope}
+%rewirte this part.............................
+%  research will investigated till sketch recognition in domain of uml diagrams. 
+%this research will include investigating the best method to recognize sketches.
+The scope of the research will include recognizing various sketches.
+The methods used in this research are comparing current systems and enhancing them and adopting them into simple geometrical symbols.
+Using well established AI methods specially \textit{Swarm Intelligence} to solve the problem.
+This research  presents a new method for sketch recognition.
+The scope of the research includes recognizing various symbols drawn using single and multiple strokes.
+The research compares various segmentation methods.
+This research used two Particle Swarm Optimization (PSO) \cite{PSOFirst} methods to segment strokes into geometrical primitives.
+The research also investigates various spatial, geometrical and statistical feature sets used in sketch recognition.
+A support Vector Machines (SVM) classifier \cite{svmintroduce} is used to classify the sketch into one of symbols from the set of previously trained symbols.
+This research targets the recognition of one symbol at a time.
+The symbol may be drawn in one or more stroke without any restriction on order, orientation or style.
 
-\subsection{Search space}
-\label{sec:Searchspace}
+%This research assumes the user draws one symbols at a time. The user can draw can draw this symbol in one stroke or in multiple stroke without any restriction on order, orientation or style.
 
-The user drawn symbols can be having more than one interpolation.
-In addition, those interpolations increase exponentially with the number of strokes the user draw and as the sketch expands the number of interpolation escalate.
-This intensifies the search space that the program search to resolve the ambiguities and find best interpolation.
-
-\subsection{Real time}
-\label{sec:Realtime}
-
-There are mainly three types of sketch understanding system offline \cite{Vibratory8}, online and interactive.
-Both online and interactive recognition like \cite{incrmentintention41, SmartSketch56} need the processing after each stroke to be real time so the user does not feel the processing.
+\section{Thesis Overview}
+This thesis is arranged as follows: chapter \ref{sec:Survey} presents a survey of different components of the sketch recognition.
+Chapter \ref{sec:ParticleSwramOptimization} gives a review on swarm intelligance and particle swarm optimization.
+Chapter \ref{sec:proposedSystem} describes the proposed system and its main building blocks.
+The comparative study of the different techniques and their detailed results in various experiments are given in chapter \ref{sec:Evaluation}.
+Finally, chapter \ref{sec:DiscussionConclusion} presents the conclusions of this thesis and suggested future work.
+%This thesis is arranged such that chapter \ref{} presents a survey of different components of the sketch recognition.
+Chapter \ref{} describe the proposed system and its main building blocks.
+The details description of each component is presented in Chapter \ref{}.
+The comparative study of different techniques with the detailed results of various experiments are detailed in chapter\ref{}.
+Finally, chapter \ref{} presents the conclusions of this thesis and suggested future work.

### 2009-04-09 draft/chapterSurvey.tex
now it 9 - appril and after club day

--- 2008-09-02 trunk/draft/chapterSurvey.tex
@@
 \chapter{Literature Survey}
+\label{sec:survey}
+% this will contians only the finished work .... and my comments to view the old survey see the file
+%draft_chaptersurvey
 
-\section{Introduction }
-\label{sec:Introduction}
-
-The researchers had divided the sketch recognition problem into image based and AI Based problem.
-The next section will discuss the methods used by current sketch understanding system.
-Firstly, in section \ref{sec:FirstStepSymbolRecognition} Symbol recognition research will be investigated and divided into image based and AI based methods.
-
- In section \ref{sec:SecondStep:SketchUnderstanding}, the understanding techniques will be demonstrated.
-Understanding mainly will handle knowledge representation methods, handling uncertainty and finally how researcher in sketch reduced the sketch search space.
-
-\section{First Step: Symbol Recognition}
-\label{sec:FirstStepSymbolRecognition}
-
-To understand a drawing we need to recognize every symbol in it. Hand-drawn sketches need more processing before it is passed to recognitions algorithms.
-
- Firstly, we have to determine how symbols are defined or grouped.
-Some System implies some constrain on user to defined symbol.
-In\cite{gestureexample12,aideddesgin22,sketchinginterfaces2,sketchinginterfaces2,cognitivesketch18,ewhitboard9}for example the user is constrained to draw each symbol using only one stroke which is not only unnatural but also constrains the symbols that can be drawn.
-More recent systems force the user to state when symbol is finished by pausing a period of time or clicking a certain button in the interface.
-Only few systems permit the user to freely draw strokes and symbols in any order \cite{Vibratory8,visualpattern43,Cali63,statisticalparsing26,physicalmeaning6}.
-Those systems must employ a method to detect when a user starts drawing a symbol and when he end.
-
-Currently symbol recognition is both image based and feature based or AI-based.
-In image based recognition symbol is recognized based on there features, structure, some image computed feature .
-The AI methods in recognizing symbols mainly imply more time and processing as for the search and rules applied but claims to be more accurate .
-
-\subsection{Pre processing }
-\label {sec:preprocessing}
- A stroke is all the points that lie along the path of the user pen from the moment of his pen down to pen up events.
-Not all those points are important even more, there may be redundancies unwanted points.
-Therefore, those strokes will need some preprocessing.
-
-The stroke that the user draws are mainly sequences of line segments or/and curves which the any system need to divide them to recognize them easily.
-The main problem is to do so without constraining the user drawing style.
-The majority of researchers first use a preprocessing process to identify those segments and the critical points from the user strokes.
-Beside points location information (x, y positions) researcher made uses of the extra information from using digital ink.
-Some as \cite {mulitstroke5,polygonfeedback31}  used user speed information to detect critical points where it is stated that user slow down at those critical points.
-Other used curvature and directions data.
-There was also mentioning using the pressure information.
-Thought some \cite {MilitaryCOA37,computationalmodel16,polygonfeedback31}  debated on the importance of the extra information and weather the temporal and speed information is significant.
-\cite{polygonfeedback31}  used the timestamp for each point and compute time difference between every two successive point to help him finding corners.
-\cite{Overview36} uses the speed and direction information but they use on it a hybrid average based filtering to filter noise and unwanted false data from appearing in the data.
-Finally, they use that information to get a poly line approximating.
+%\section{Introduction }
+%\label{sec:Introduction}
+The history of pen and sketch based computing goes back to 1963 with Ivan Sutherland's SKETCHPAD \cite{sutherlandsketch}, which used a light pen to draw on the monitor to create circuit diagrams.
+Rubine \cite{gestureexample12} Gesture recognition system.
+% i need to write a history then small introudction to the survey...
+%  i have to state that the next  sections will generally group segmentation and symbol recognition task. 
+\section{Preprocessing and Segmentation}
+\label{sec:preprocessing}
+%this section should describe how others did the first preprocessing of the problem. i have to summarize what i wrote before. 
+ % first paragrah about how stroke is a messing thing and naturally it is ambigues. 
+  In a system that supports free hand sketching, sketching should be as natural as possible.
+This means that the user can draw symbols using single or multiple strokes without constrains.
+The goal of a sketch system is making the user feel that using the using the system is nearly the same as using pencil and paper.
+But sketchs are incomplete drawings that are sloppy and messy, for example figures \ref{fig:Overshootandundershoot} and \ref{fig:overstroked} show examples of messy sketches drawn by users.
+This led researchers to perform sort of preprocessing before starting the sketch recognition task itself.
+The next sections explain different preprocessing techniques.
 
@@
 	\centering
-		\includegraphics[scale=0.7]{../../neededfiles/Figures/dirspeedgraph.eps}
-	\caption[Direction and Speed]{direction and speed graphs of the rectangle stroke the minima in the graph show the location of the corners of the rectangle.}
-	\label{fig:dirspeedgraph}
-\end{figure}
-
-\cite{Mathpad46}Handle rotations invariant using polar transformation then shifting to make it more reliable when template matching.
-However, they state that the only problem to there system is that the points near the centroid affect the polar transformations which will effect the matching.
-(To solve it they say that loosing the weight of the pixel near the centroid will improve classification).
-
-The next sections will demonstrate firstly the noise removal algorithms and techniques.
-Then next sections will show the most used critical points detection algorithms along with the segment division algorithms.
-
-  In\cite{sketchRead29} they divided user strokes into geometric primitives based on some critical points they computes.
-After that, they either cluster segment to form a symbol   or directly apply some kind of basic symbol recognition leaving the compound object for next processing step .
-
-From the image-based view the symbol is a sequences of line segments or/and curves. To do so without constraining the user drawing style we have to easily divide the user drawn strokes.
-
-%\begin{figure}
-%	\includegraphics[scale=0.6]{../../neededfiles/Figures/shapesegmentation.eps}
-%	\caption[segmented shapes]{example of Segmented shapes}
-%	\label{fig:shapesegmentation}
-%\end{figure}
-
-\subsubsection{Removing noise}
-\label{sec:RemovingNoise}
-
-Due to the sloppiness of the user, his stroke may need a lot of noise removal before any algorithm can extract its features.
-For example, when a user draw sketches he may over trace the stroke more than one time.
-Human perception can easily differentiate what exactly is drawn.
-Nevertheless, for an AI program this is more difficult problem.
-The user may also overshoot see \textit{Fig\ref{fig:Overshootandundershoot}} or undershoot while drawing intersecting segments.
-\begin{figure}
-	\centering
 		\includegraphics[scale=0.7]{../../neededfiles/Figures/Overshootandundershoot.eps}
-	\caption[Neat and Sloppy symbols] {The Figure show a sloppy strokes with under shoot and overshoot problems and the neat version of the symbol.}
+	\caption[Neat and Sloppy symbols] {A sloppy and neat versions of symbol.}
 	\label{fig:Overshootandundershoot}
 \end{figure}
-
- \cite {overtraced24} tried to solve the over traced strokes.
-they used a thinning algorithm to move points perpendicularly towards the regression line from the formed from fitting line of neighbour's into line or circle or ellipse which is calculated with least square fit for the data.
-The next step is selecting near neighbours and connection then which gives a serious of linked edges which a greedy link edge algorithm this produce a single stroke instead of over traced see \textit{Fig [\ref {fig:overstroked}]}.
 
@@
 		\includegraphics[scale=0.7]{../../neededfiles/Figures/overstroked.eps}
-	\caption{Example of an overstroked stroke}
+	\caption{Example of an over traced stroke}
 	\label{fig:overstroked}
 \end{figure}
-Some systems as in \cite{SmartSketch56} where the noise reduction is done in two steps the first step is the polygonal approximation to remove non critical points in the user stroke.
-The process is mainly removing the redundant points that are along the path of stroke.
-Due to the shaky nature of user stroke, there are additional end-points at the stroke segments.
-Secondly, the next step will be applying an Agglomerate point filtering.
-Agglomerate filtering is according to \cite {statesurvey35} removing the undesired extra points that is produced due to the unsteady operations produced while lifting and touching tablet.
-The results of those processes are passed to final end-point refinement to merge or remove any redundant points.
 
+ %The preprocessing can be noise removing, stroke beautification or segmentation.  %Because of this 
+ Sezgin and Davis \cite{overtraced24} tried to solve the over traced strokes (see Figure \ref {fig:overstroked})using a thinning algorithm.
 In \cite{meanshift10,domainindependent17}, Bo Yu uses mean shift procedure as a noise remover and preprocessing step to the recognition step.
 The mean shift procedure is simply shifting the point to the average of its neighborhoods so smoothing out the noise without changing the global structure of data.
-This process is clamed to help find critical points easily after that.
+Gaussian filters was used to smooth curves \cite {Phoenix88}.
+These procedures may seem to remove noises in the sketch but it often treat vertices as noise which obscure vertex locations.
+ %% now the third paragraph.....
+Beside points location information (x, y positions) researcher made uses of the extra information from using digital ink.
+Some as \cite{mulitstroke5,polygonfeedback31} used user speed information to detect critical points where it is stated that user slow down at those critical points.
+Other used curvature and directions data.
+%Others \cite {MilitaryCOA37,computationalmodel16,polygonfeedback31}, debated on the importance of the extra information and weather the temporal and speed information is significant.
 
-In \cite {domainindependent17}    approximation step uses the speed, curvature, and direction of strokes drawn by user to determine the location of vertices to gain more resistance to noise that data may contain.
+ Agar et al.
+\cite{polygonfeedback31} used the timestamp for each point and compute time difference between every two successive point to help him finding corners as it provides more distinct maxima than the speed information.
+He mentions that the pointing device (for example the mouse) sampling rate is the reason for this phenomena.
+The points are sampled at regular time intervals while the pointing device is moving.
+There are no points while the pointing device is stationary.
+This leads to a nearly constant time difference between samples while the pen is moving and large difference while the pen is stationary.
+Contrary to speed information, where the users draw with variant speeds the speed information has a lot of noise in the data.
 
-\subsubsection{Critical points and corners Detection}
-\label{sec:CriticalPointsAndCornersDetection}
+Sezgin et al.\cite{earlyprocess} used both speed and direction information.
+Their system gives better results than using only speed or direction information.
+This is because the speed information tend to be noisy when drawing large symbols, on the other side when using only direction information the some corner are missed.
+Our system uses speed, direction and time difference information in its preliminary calculation step.
+%both speed and direction information help detecting vertices when the
 
-Locating the critical points is provided to the next process of the systems which in most cases segment detection algorithm as in \cite{Simusketch51} .
-Other systems after detecting the points the system directly uses those information to template matching.
+ A genetic algorithm was used by \cite{CruveDivisionSwarm} to optimally divide digital curves into lines and curves.
+Chen et al.
+\cite{CruveDivisionSwarm} uses digital curves scanned from paper as input to the system.
+Also, they did not take advantage of the curvature or local geometric properties of the digital curve.
+Yin \cite{PolygonApproximationPSO} used PSO to convert digital curves into polygons, our system adopts  Yin \cite{PolygonApproximationPSO} method and tries to improve it by adding curvature and other local information while segmenting strokes to achieve better segmentations.
+%The next section presents the general particle swarm algorithm which is used in both our segmentation algorithms.
+% are presented in section \ref{subsubsec:Discreteparticleswarmalgorithm}
 
-Generally, the critical points along with the segment points of the path from one critical point to another are used to detect labeling of that segment. 
-Line detection or multi line segments algorithm is found in \cite{polygonfeedback31} .
-Mostly after detecting the vertices if the segment cannot be labeled as arc it is defined as a line.
-Other mainly defines a threshold value of the slope if it increase than that it is counted as an arc.
+  % third should speech on how systems define critical points (this paragrah should speak on the data collected (speed crital points, ...
 
-Systems like \cite{physicalmeaning6} detect critical point for each stroke then the final point is merged with the first if they are in range defined by threshold.
-However, stroke may also intersect with other strokes drawn before therefore \cite{physicalmeaning6}  introduced a structure to detect such approximates he first define every stroke drawn as series point and labeled as line, arc or area.
-Then after labeling most strokes the system check for proximity on bases of line -line, line-end point, or line- area.
+ After critical point detection most sketch system proceeds to label segments into simple geometric primitives (i.e line, arc, etc...).
+A hybrid algorithm was introduced in \cite{earlyprocess} where different sets of segments are generated based on both curvature and speed dominant points, followed by choosing a segmentation with the least error from a generated hybrid set.
+However, this system is limited to recognizing only specific simple geometric shapes with a set of low level recognizers.
+Each low level recognizer is designed to recognize only one geometric shape using spatial and geometric information extracted from input stroke.
+Yu \cite{meanshift10} introduced a \textit{feature area} for each primitive and then computed the segmentation error for different types of primitives based on the computed \textit{feature area}.
+Figure \ref{fig:featurearearc} shows the feature area of a line and circle in Yu system, the shaded area is the difference between the computed feature area and the assumption area.
+His system achieved good accuracy in simple shapes (square, ellipse,...etc) but did not perform well in complex shapes.
 
-In   \cite {meanshift10,domainindependent17}   approximation step uses the speed, curvature, and direction of strokes drawn by user to determine the location of vertices to gain more resistance to noise that data may contain.
-They firstly analysis the curvature and speed data and use the mean of data as the threshold then use that threshold to separate data into regions and finally locating the maxima and minima in each region .
-The previous step generates a set of vertices and its certainties they use to choose best fit.
+ Paulson and Hammond \cite{Paleosketch08} introduced a set of low level recognizers that were reported to achieve 98\% accuracy.
+However, their system similar to all low level recognizers in the way that it identifies a small set of simple shapes.% A genetic algorithm was used by \cite{CruveDivisionSwarm} to optimally divide digital curves into lines and curves.
+Chen et al.\cite{CruveDivisionSwarm} uses digital curves scanned from paper as input to the system and did not take advantage of the curvature or local geometric properties of the digital curve.
+\cite{Fluid25} Trying to combine fast morphing and basic recognition for the sketch drawn as the user draw it.
+It tries to move the points of strokes drawn by the user draw into one of the known shapes (ex: box, lines and circle).
+It uses a relaxation and least square method for circles and uses the string of forces for finding best-fit box.
+However, their system can only identify circle and rectangles drawn using a single stroke.
+\begin{figure}
+	\centering
+	\subfigure[Feature area of line an arc]{
+		\includegraphics{images/featurearearc.eps}}
+		\subfigure[Feature area of line]{
+				\includegraphics{images/featureareline.eps}}
+	\caption{Feature area}
+	\label{fig:featurearearc}
+\end{figure}
 
-\subsubsection {Primitive stroke segmentation and clustering}
-\label{sec:Primitivestrokesegmentationandclustering}
+\section{Symbol Recogntion}
+\label{sec:symbolrecogntion}
 
-The most segmentation method is least square method in \cite {overtraced24}.
-Other systems  approximate a line segmented using the information from the direction curve where the curve is fitted with horizontal line.
-firstly the system compute least square fit for direction curve or the stroke point computed and check if deviation points is small if they pass then hypothesize al line segment  from the first point to the last one.
+In this section, we review three different types of symbol recognition methods.
+Symbol recognition can be based on: strokes, global shape properties, and appearance \cite{Oltmans07}.
+The role of which each stroke plays in the structure of the sketched symbol is used to determine shapes in stroke based methods.
+In global shape properties approach, systems investigate general shape properties and its underlying strokes.
+Using shape properties relaxes the assumption and role of each stroke in the symbol but it does not truly represent the appearance of the symbol.
+Unlike appearance based methods where individual strokes are disregarded and they focus on the overall appearance of shapes.
+We choose to use the appearance based method but added some geometrical and structure properties used in most stroke based approaches.
+Our method tries to avoid many of the problems faced by approaches based on individual strokes or global shape properties while trying to gain the additional information that the stroke based system use.
+In the next sections, we look into three different symbol recognition approaches, describe their advantages and challenges.
+
+\subsection{Stroke Based Recognition}
+\label{sec:StrokebasedRecognition}
+Stroke based approach is based on the assumption that each stroke has a specific role in representing the symbol.
+Stroke based methods examine each stroke as the user draw it to determine its role in the symbol.
+These methods are suited to interactive interfaces since users expected that system displays interpretation of the symbol after the stroke is drawn or after drawing a group of strokes.
+The majority of the research in sketch recognition has focused on stroke-based methods because of the corresponding emphasis on interactive interfaces.
+Most stroke based systems either represent symbols as gestures or   hierarchical shape.
+
+\subsubsection{Guesture Recognition}
+\label{sec:GuestureRecognition}
+
+Single stroke gestures recognition was the focus of the early work in symbol recognition \cite{gestureexample12,aideddesgin22,sketchinginterfaces2}.
+Gestures are pen strokes drawn by the user that can immediately recognized by the system.
+The result was either creating an action (e.g.
+deleting a shape) or drawing a shape on the screen.
+These systems are based on how the user draw the shape rather than how it looks.
+
+One of the first systems in gestures recognition is Rubine's algorithm \cite{gestureexample12} .
+He used Gaussian statistical classifier to classify gestures.
+Simple features as length of stroke, smoothness of stroke, the angle that the first part of the gesture is drawn at and properties of the bounding box of the stroke are used to train the classifier to recognize  symbols.
+In our system, we combine Rubine features with other to include the information on how the user draws sketch while recognizing the symbols.
+In \cite{Long00}  Long shows how similarities between gestures can be analyzed.
+He used this information to identify gestures that are likely to confuse the recognition system.
+This helped interface designers to choose gestures that are easily recognized.
+
+The main problem with gesture based recognition system that they impose several constrains on how the user can draw.
+Gestures must be drawn using a specific order and direction to be recognized correctly.
+Moreover, in most cases gestures do not necessarily look like the symbols they represent (e.g.
+to represent a rectangle the user draws only the left edge of it).
+Because of these reasons, gestures recognition systems are not suitable for free hand drawn sketches.
+
+\subsubsection{Hierarchical Shape Descriptions}
+
+A Hierarchical shape descriptions was widely used in sketch systems \cite{sketchunderstanding1,HierarchicalParsing7,Ladder30,AlvaradoFreedom42,napkinGross}.
+Symbols are represented using hierarchical levels.
+The lower levels represent simple geometric primitives such as lines, arc and ellipses.
+Intermediate levels shapes are composed of lower level parts and constrain or connections between them.
+For example, a triangle is described as three lines connected at their endpoints.
+Higher level shapes are constructed from intermediate and lower level shapes, for example, a diode is described as a line on the top of a triangle head (acute angle).
+
+ Systems that employ such representation must first preprocess strokes to break them into geometric primitives (see section \ref{sec:preprocessing} for a review on preprocessing methods).
+Recognition can then be treated as sub graph matching problem or constrain satisfaction problem.
+
+%Although not based on a formal structural model of shapes, another probabilistic approach by Szummer and Qi in [37] uses conditional random fields (CRFs) to propagate information about the labeling of one stroke fragment to its neighbors.
+This allows their system to assign labels to stroke fragments that cannot easily be identified in isolation and to regroup over-segmented fragments.
+This helps mitigate the difficulty of determining the exact granularity at which to perform fragmentation by using the context of surrounding fragments and a global interpretation of the sketch.
+They have applied their algorithm to the binary labeling problem of distinguishing boxes from connectors in organizational charts with good results, even in several highly ambiguous cases.
+\paragraph{Sub-graph matching:}
+\label{sec:GraphSearching}
+
+Representing the symbol that the user draws as a graph transforms the classification problem into a graph-matching problem.
+In general, each user geometrical primitive is represented as a node of the graph and the relations between those nodes are represented as arc.
+Figure \ref{fig:squarescematic} shows a graph that is constructed by \cite{mulitstroke5} for a rectangle.
+To classify the symbols the graph constructed must be matched with the graphs in the database.
 
@@
 	\centering
-		\includegraphics[scale=0.7]{../../neededfiles/Figures/rectanglerecog.eps}
-	\caption{rectangle recognition}
-	\label{fig:rectanglerecog}
+		\includegraphics[scale=0.7]{../../neededfiles/Figures/squarescematic.eps}
+	\caption{square scemantic graph}
+	\label{fig:squarescematic}
 \end{figure}
 
-%//Other like \cite {tobeadd}  uses the relaxation process to try to relax parameter on a classification .//
- In \cite {meanshift10,domainindependent17}  Bo first uses the direction graph to check if the stroke can be approximated as a line or no.
-If the stroke can't be approximated as a line it is checked to be a circle or ellipse.
-the procedure is simply assume a circle classification then check this assumption with the user drawn stroke given a min error range.
-\cite {Fluid25} Trying to combine fast morphing and basic recognition for the sketch drawn as the user draw it.
-It tries to move the points of strokes drawn by the user draw into one of the known shapes (ex: box, lines and circle).
-It uses a relaxation and least square method for circles and uses the string of forces for finding best-fit box.
+In the worst case, sub-graph matching is an exponential problem and thus is expensive to compute.
+Researcher tried to reduce complexity by restricting the search to fragments that are spatially or temporally grouped, or by using other assumption.
+Some of the common assumptions are that each stroke is only part of one symbol and that user must draw one shape before drawing the next\cite{physicalmeaning6}.
+These assumptions add constrains on how user can draw sketches and thus violates freely drawn sketches.
+Moreover, the matching process in highly affected by segmentation results.
+Over or under fragmentation of stroke into geometric primitive complicates the matching process.
+If a stroke is broken into too many fragments then the there will be extra component that will not map to any part of the symbol.
+If the stroke is divided enough then there will be component of the shape that are not filled while matching.
+The different ways of segmenting the sketches worsen the already high cost graph matching.
 
-The noise removal and critical point detection is only a preprocessing step for the symbol classification algorithms.
-Some system implements a segment segmentation and clustering to help in the classification problem.
+ To solve this problem,  \cite{mulitstroke5} provided two methods to match graphs, a simpler one using restrictive assumptions and if it fails the system uses a longer more computationally expensive search.
+.
+Figure \ref{fig:squarescematic} shows a semantic graph that is constructed by the trainable recognizer in   \cite{mulitstroke5} for a rectangle.
+The trainable recognizer construct the semantic network at train time with then match the symbol user draw with the networks in the database.
+To match the networks quickly the system assumes the user draws the symbol with the same order as in training set.
+A deep heuristic graph search will employed only if no match was found.
+% there was no match found.
 
-After the strokes are segmented \cite {sketchunderstanding1,HierarchicalParsing7,Simusketch51}  proposed a merging techniques that will merge strokes together into symbols to help classification.
-They use a method combine between divide and merge.
-They firstly divide the drawn sketches into section based on marker technique they used.
-After that they merge strokes if there bounding boxes intersects and then  merge generated cluster till no more merge can occur on one segment.
+ Several other systems tried to model the matching process as a probabilistic problem \cite{HMM53,SketchRead2007}.
+These systems allow matching of partially satisfied shapes to propagate down to re interpret low level fragmentation hypothesis.
+In SKETCREAD \cite{SketchRead2007}, a dynamically generated Bayesian network fragments represents the shape hypotheses.
+Higher level structure of shape fragments can cause re interpretation of the geometric primitives.
+For example, if one head of an arrow is initially labeled as an arc and the system fragments the shape as an arrow then the belief that this stroke is an arc is decrease and the belief that it is a line increase.
+Finally, the arrow structure is completed and it is fully recognized.
 
-%Intension exteraciton in paper Intention54
-% Check to add
+\subsection{Global Shape Properties Based Recognition}
+\label{sec:GlobalFeaturesBasedRecognition}
 
-%/*Mathpad46
-%methion  check stroke clustering.  Simusketch51  and in 8 
-%
-%Bo \cite {tobeadd}   The ellipse or
+To step back from properties of individual strokes several researcher tried to use shape properties on the whole shape to classify shapes. The properties used
 
-\subsubsection{Image and Statistical Features calculation}
-\label{sec:ImageandStatisticalFeaturescalculation}
+A number of approaches have stepped back from the properties of individual strokes to
+classify shapes based on a set of properties calculated on the whole shape. Properties
+that attempt to summarize the information in the entire shape are called global
+features.
+In [6] a carefully crafted set of filters based on global features, such as the ratio
+of their bounding box area to convex hull area and the ratio of the perimeter to the
+area, were used to progressively eliminate possible interpretations for a stroke until a
+suitable interpretation was found. For example, rectangles can be distinguished from
+triangles by looking at the ratio between the area of the convex hull and the area of
+the rectangular bounding box. The ratio for rectangles will be near 1.0 and the ratio
+for triangles will be near 0.5.
+A similar set of features was used by Fonseca et. al. in [14]. In that system,
+a carefully selected set of features was used by a number of rules and fuzzy logic
+to perform classifcation. Their system could also be trained to learn to distinguish
+between shape classes using a Naïve Bayes model based on the features.
+A more general approach using Zernike moments was demonstrated by Hse et. al.
+in [21]. Instead of using a hand tuned collection of properties they make use of the
+magnitudes of Zernike moments. Zernike moments are a class of orthogonal moments
+which describe the distribution of points in the input shape. Higher order moments
+represent finer levels of detail in the point distribution. The magnitudes of these
+moments, calculated up to a given order, form a feature vector which can be used to
+train an SVM. The magnitudes of Zernike moments have been shown to be invariant
+to both rotation and reflection of the input shape. A comparison of Zernike moments
+to our match vector representation is presented in detail in Section 5.2.2.
+These types of systems are based on global properties of shapes. Shape classes are
+identified by determining which properties each class tends to have. The properties do
+not typically depend on the number or order of the strokes and are thus based on the
+appearance of the shape and not how it was drawn. However, they do not represent
+the individual details of the shape. For example, it would be impossible to distinguish
+between the symbol for a current source (circle containing an arrow) and an ac-source
+(circle containing a "tilde") using just the ratio of convex hull area to bounding box
+area. This approach cannot distinguishing between shapes that difier by small details,
+and cannot deal with shapes that allow substantial conceptual variation. Handling
+these requires an approach that can represent the shapes at multiple levels of detail
 
-Most system calculates the general image properties for the strokes. Most used features are bounding box, width, size, points along path. 
-\cite {meanshift10,domainindependent17}  introduced a new feature which they named "Feature Area which" they used in primitive classification introduced before.
-They first make and assumption for classification and compute the difference.
-For example if they make assumption as a circle they first calculate the center of the circle then its radius based on the bounding box of the stroke.
-The area difference between the computed circle and the current stroke is the feature area.
-see \textit{Fig\ref{fig:featurearearc}}.
+The application of Statistical classifiers in recognizing shapes are widely used in sketch recognition.
+Statistical classifier requires training time before the recognition actually happen.
+Then first step of classification is features computation that is done in the previous step.
+Finally, the recognition of symbols will be based on the classifier and the training set used.
 
-\begin{figure}
+One of the first systems in sketch recognition is \citep{gestureexample12,aideddesgin22} algorithm to recognize ink gestures.
+He used Gaussian statistical classifier to classify any stroke user draw.
+In \cite {gestureexample12}  he firstly compute statistical information about the user stroke then pass it to a Gaussian classifier that had been trained using 10-20 samples.
+The output of the classifier is classification of the stroke from one of the known classes.
 
-		\subfigure[Feature area of line] {\includegraphics[scale=0.7]{../../neededfiles/Figures/featureareline.eps}}
+Bayesian classifiers was used in \cite{Vibratory8} as a stroke classifier to recognize different shapes.
+The recognition is done using feature based statistical Gaussian distribution and Bayesian classifier techniques.
+The trainer use nine features extracted from segmented symbols (ex number of line segments, number of L intersections, ) all feature has a tolerance within.
+After all nine feature are computed a Gaussian statically model is constructed.
 
-	\subfigure[Feature area of an arc]{	\includegraphics[scale=0.7]{../../neededfiles/Figures/featurearearc.eps}	}
-			\caption[Feature area]{the shaded are shows the difference between the computed feature area and the assumption area. }
-	\label{fig:featurearearc}
-\end{figure}
+Some systems employ more than one statistical classifier to get better results for the recognition.
+For example \cite{zernike61} uses Neural network, SVM (support vector machine )and MMD.
+\cite{polygonfeedback31}
+%the closed shaper recognizer which is done using classifiers (SVM , Rule based and ANN ) to make a 
+%classification of triangle, rectangle or ellipse
 
- System that uses statistical classifiers use other features for each stroke.
-Features like center of gravity\cite{gestureexample12,sketchunderstanding1}, maximum length line, Zenerik moments \cite{zernike61}, number of lines, curvature angles, minimum enclosing rectangle\cite {incrmentintention41} , convex hull were calculated in \cite {Cali63}.
-Also \cite {geometrydomain49} introduced a new feature which he named ink density that he used in the classification of the glyphs see \textit{Fig \ref{fig:inkdensity}}.
-\begin{figure}
-	\centering
-		\includegraphics[scale=0.7]{../../neededfiles/Figures/inkdensity.eps}
-	\caption[Statistical Feature Example]{The figure shows the Ink Density feature that \cite{geometrydomain49} introduced. }
-	\label{fig:inkdensity}
-\end{figure}
-All spatial relationships needed in the graph or template matching is computed after statistical features are calculated.
-Relations like intersections, containment, perpendicular, right-of, left-of and below are the most used relations concerning constructing graphs for symbols\cite {geometrydomain49,sketchinginterfaces2} .%stretchdiagrammer4,     i dont know where is this refrence now
-
-\subsection{Image Based Classification}
-\label{sec:Image Based Classification}
-
-In the following sections, systems generally try to recognize the strokes user draws and give it a label based on there image features.
-The information collected and features computed in the previous step.
- \subsubsection{Low Level Analysis}
-\label{sec:Low Level Analysis}
-
-%
-%Detecting object using low level analysis only 
-%\cite {mulitstroke5} first detecting vertices and control points then fit these segment into line and arc using best fit and least square error classification.
-
-%Check the DB
-
-% description[4] after that it passes through the curve approximation into the Bezier curves
-%
-%\cite {statesurvey35} we need to fit and filter the classified shapes that is done using on of the following least square is the most frequently method used some use fuzzy logic to make circular fitting some use b-splines or ellipse and circle fitting.
-some after that do some  shape beatification or regularization.
-
-The most used method in low level recognition is best fit and least square methods where they was used in most systems.
-Other systmes like \cite{polygonfeedback31} prefered to implement its own polygon recognition algorithm.
-
-Bo et al.
-In \cite{meanshift10} introduced an algorithm to recognize line and arcs.
-The system used the direction graph and fitted stroke points with a non horizontal line.
-So the recognizing is done using the following algorithm which is preformed before the mean shift procedure so not the destroy the direction graph .Firstly, if direction graph can be approximated using a line then it is recognized as a line.
-Otherwise the algorithm Bisect the line from the first and last  point of stroke and donate it as "M".
-Then calculate the intersection point of the line "M" and original stroke and donate it as "A" .
-Now using the three points first , last points of stroke and point A find the center of circle which passes through all of them which will donate the candidate arc.
-Finally , calculate the radius of the circle and the mean of distance  between center and each point of stroke and  compute the feature area of the stroke and the circle arc see \textit{Fig \ref{fig:featurearearc}}.
-%//
-
-\subsubsection{Template matching}
-\label{sec:Template matching}
-
+\subsection{Apperance Based Recognition}
+\label{sec:ApperanceBasedRecognition}
 Template matching is the most used method to recognize strokes and symbols into classes.
@@
 
-\subsubsection{Graph matching}
-\label{sec:Graph matching}
-
-Representing the symbol that the user draws as a graph transforms the classification problem into a graph-matching problem.
-In general, each user stroke is represented as a node of the graph and the relations between those nodes are represented as arc.
-To classify the symbols the graph constructed must be matched with the graphs in the database.
-
-\cite{physicalmeaning6} uses spatial relationships for there graph matching.
-They first calculate spatial proximity relationships.
-Relations like intersections, containment, perpendicular, right-of, left-of and below are the most used relations concerning constructing graphs for symbols.
-
-Other systems  first analysis the spatial relationship between glyphs (sequence of strokes) using a library of spatial relations and information calculated in the preprocessing step.
-After that, they filter out important relation-using hierarchy of rules and type of geometrical constrain.
-All redundant constrains and relations are removed the final graphs is matched with the database.
-If more than one candidate is available the algorithm choose the best fit but if no matches found constrains are relaxed and strokes are passed again to classifier.
-
- Proposed a structure representation of the strokes in \cite{sketchunderstanding1,HierarchicalParsing7} .
-She added ratio of proximity of the relations between strokes.
-In addition, she used an error-driven, stochastic search algorithm for graph matching.
-
-\subsubsection{Statistical methods}
-\label{sec:Statisticalmethods}
-
-The application of Statistical classifiers in recognizing shapes are widely used in sketch recognition.
-Statistical classifier requires training time before the recognition actually happen.
-Then first step of classification is features computation that is done in the previous step.
-Finally, the recognition of symbols will be based on the classifier and the training set used.
-
-One of the first systems in sketch recognition is \citep{gestureexample12,aideddesgin22} algorithm to recognize ink gestures.
-He used Gaussian statistical classifier to classify any stroke user draw.
-In \cite {gestureexample12}  he firstly compute statistical information about the user stroke then pass it to a Gaussian classifier that had been trained using 10-20 samples.
-The output of the classifier is classification of the stroke from one of the known classes.
-
-Bayesian classifiers was used in \cite{Vibratory8} as a stroke classifier to recognize different shapes.
-The recognition is done using feature based statistical Gaussian distribution and Bayesian classifier techniques.
-The trainer use nine features extracted from segmented symbols (ex number of line segments, number of L intersections, ) all feature has a tolerance within.
-After all nine feature are computed a Gaussian statically model is constructed.
-
-Some systems employ more than one statistical classifier to get better results for the recognition.
-For example \cite{zernike61} uses Neural network, SVM (support vector machine )and MMD.
-\cite{polygonfeedback31}
-%the closed shaper recognizer which is done using classifiers (SVM , Rule based and ANN ) to make a 
-%classification of triangle, rectangle or ellipse
-
-\subsubsection{Other hybrid methods}
-\label{sec:Other hybrid methods}
-
-Due to the complexity of the sketch recognition problem, more than one system proposed a hybrid method to solve those problems. They combine more than one approach to recognize symbols the user draws.
-\cite{sketchunderstanding1,Simusketch51}  used four kind of classification image, graph matching, statistical and structure based classifiers.
-Each symbol will contain all graph, statistical, structure information and a (Wight, recognizer) tuples to determine which recognizer has the most influence in recognizing this symbol see \textit{Fig\ref{fig:structuralscemanticdeff}}.
-\begin{figure}
-	\centering
-		\includegraphics[scale=0.5]{../../neededfiles/Figures/structuralscemanticdeff.eps}
-	\caption[Structure definitions]{Example of the structure definition of sin symbol in \cite{sketchunderstanding1}.}
-	\label{fig:structuralscemanticdeff}
-\end{figure}
-
-\cite {HMM53}  firstly normalize the stroke and compute all the statistical features from it.
-After that, a noise reduction algorithm is used on the strokes which will be translated to produce size and rotation invariance data.
-Those data is passed to a multistage hybrid recognizer.
-The first step of his multistage classifier is dynamic programming procedure to compute the optimal degree of similarity and differentiation of symbols.
-Secondly, a statistical feature ranking is employed on the data produced.
-Finally, the system will reject all unlikely classification and produce best fit for this stroke using a dynamic programming using domain and context knowledge as input.
-
-\cite{statisticalparsing26} is a statically parser make use of the stored rules and probability of drawn stroked and uses Bayesian classifiers to do the classifications.
-
-\subsection{AI based Classification}
-\label{sec:AI based Classification}
-
-Most system mentioned before use image or statistical feature based on the points the user draw.
-The next few sections will present methods that will use AI approached to classify the symbols the user draw.
-Systems using AI approached imply a higher level of understanding of the strokes the user draws.
-Some construct a graph schema of the symbols from strokes the user draw.
-Other tried generating a visual grammar to identify the symbols that can be drawn.
-
-\subsubsection{Graph Searching }
-\label{sec:Graph Searching }
-
-\textit{Figure \ref{fig:squarescematic}} shows a semantic graph that is constructed by the trainable recognizer in  \cite{mulitstroke5} for a rectangle.
-The trainable recognizer construct the semantic network at train time with then match the symbol user draw with the networks in the database.
-To match the networks quickly the system assume the user draws the symbol with the same order as in training set.
-A deep heuristic graph search will be employed if there was no match found.
-
-\begin{figure}
-	\centering
-		\includegraphics[scale=0.7]{../../neededfiles/Figures/squarescematic.eps}
-	\caption{square scemantic graph}
-	\label{fig:squarescematic}
-\end{figure}
-
-\cite{SRGraph57}  Uses two kind of graphs relations a spatial relationship graph (SRG) which is build on defining the spatial relationship between the source object and candidate object on the sketch.
-The relations are defined using Spatial Adjacency Relation (SAR) that falls into relation previously computed in system.
-To make the graph fully descriptive the system computes Relative Position Relation of SAR relations and adds these relations into the SRG graph.
-
-\subsubsection{Rule based }
-\label{sec:Rule based }
-
-The general algorithm of rule based system as used in \cite{interpretationmechanical50,SmartSketch56}  based on a knowledge of relation ship between nodes (strokes) a role is defined using those relation ships.
-The rule based system check the strokes the user draw against all strokes to detect if there a rule match the drawn symbol.
-When a rule matches the consequence of this rule, apply on the input symbol.
-
-The rule system used in \cite{sketchinginterfaces2}  uses the (UI user interface) domain knowledge to do its basic recognition.
-It is divided into two step first is testing if the rules applies on the input symbol.
-Second step, is ranking the rules that apply on the symbol based on the highest confidence that was given to this input symbol.
-
-Some system like in \cite{SmartSketch56}  used the rules to help in noise removal.
-Furthermore, rules based system are used to employ some reasoning to the recognized symbols with the help of context knowledge \cite {interpretationmechanical50}.
-
-\subsubsection{HMM }
-\label{sec:HMM }
-
- The main idea of HMM is representing knowledge and recognition in serious of states each state has its input conditions.
-Only a node that has the input conditions of the destination state that can transit it from one state to another.
-The final state is the recognition state of a given class.
-For example to recognize a rectangle the first state should be first line detected, second will be second line and the condition is the recognition of either a perpendicular line or parallel line to first one to another.
-
-In sketch, recognition HMM was used to recognize symbols to different classes.
-Each class recognized was modeled using a single HMM model.
-The system in \cite {HMM53} encode the geometrical primitive that construct the symbol the user draws.
-After the encoding the codes are passed to HMM based classifier.
-Each object is represented by a HMM model and to manage multiple classes they combine matching score from different HMM models.
-
-\subsubsection{Graph language and visual language describing }
+\paragraph{Graph language and visual language describing }
 \label{sec:Graph language and visual language describing }
@@
 
+\subsection{Other hybrid methods}
+\label{sec:Otherhybridmethods}
+
+Due to the complexity of the sketch recognition problem, more than one system proposed a hybrid method to solve those problems. They combine more than one approach to recognize symbols the user draws.
+\cite{sketchunderstanding1,Simusketch51}  used four kind of classification image, graph matching, statistical and structure based classifiers.
+Each symbol will contain all graph, statistical, structure information and a (Wight, recognizer) tuples to determine which recognizer has the most influence in recognizing this symbol see \textit{Fig\ref{fig:structuralscemanticdeff}}.
+\begin{figure}
+	\centering
+		\includegraphics[scale=0.5]{../../neededfiles/Figures/structuralscemanticdeff.eps}
+	\caption[Structure definitions]{Example of the structure definition of sin symbol in \cite{sketchunderstanding1}.}
+	\label{fig:structuralscemanticdeff}
+\end{figure}
+
+\cite {HMM53}  firstly normalize the stroke and compute all the statistical features from it.
+After that, a noise reduction algorithm is used on the strokes which will be translated to produce size and rotation invariance data.
+Those data is passed to a multistage hybrid recognizer.
+The first step of his multistage classifier is dynamic programming procedure to compute the optimal degree of similarity and differentiation of symbols.
+Secondly, a statistical feature ranking is employed on the data produced.
+Finally, the system will reject all unlikely classification and produce best fit for this stroke using a dynamic programming using domain and context knowledge as input.
+
+\cite{statisticalparsing26} is a statically parser make use of the stored rules and probability of drawn stroked and uses Bayesian classifiers to do the classifications.
+
 %\subsubsection{Fuzzy classifiers }
@@
 
-\subsection{Hybrid Systems}
-\label{sec:Hybrid Systems}
+%\subsection{Hybrid Systems}
+%\label{sec:Hybrid Systems}
 
@@
 The system also uses a dynamic programming procedure to find a best fit from the given set of classifications.
+ \chapter{Particle Swram Optimization}
+\label{sec:ParticleSwramOptimization}
 
-\section{Second Step: Sketch Understanding}
-\label{sec:SecondStep:SketchUnderstanding}
+\section{Particle Swram Optimization Review}
+\label{sec:ParticleSwramOptimizationReview}
 
-Sketch understanding is concerned with converting a set of recognized symbols into a whole document that is fully consistent with each other.
-Mainly the recognized symbols are a list of best candidates of that symbol for which the sketch understanding process must choose best fit based on the document structure.
+%review on the pso alogrithm alone.................
 
-After recognizing the symbols drawn, Understanding the sketch document involve additional processing.
-In spite of various recognition systems that had reached a satisfactory symbol recognition.
-The sketch understanding problems are at a standstill.
+\section{Discrete Particle Swarm Algorithm}
+\label{sec:ParticleSwarmAlgorithm}
+%\section{Particle Swarm Algorithm}
+%\label{PSO}
+%What is particle swarm algorithm and how it was used in related researches. 
+The main idea of \textit{Particle Swarm Algorithm (PSO)} is to represent each agent with a particle from the solution space \cite{PSOFirst}.
+Each agent moves the particle with a direction and velocity $v_{ij}$ based on equations \ref{eq:Swarm} \& \ref{eq:Swarm1}.
+\begin{equation}
+%\[
+p_{ij}=p_{ij}+v_{ij},
+%\
+\label{eq:Swarm1}
+\end{equation}
+where $p_{ij}$ represent the $jth$ particle in the $ith$ agent and $v_{ij}$ is the velocity of the $jth$ particle in the $ith$ agent.
+ %Equation [\ref{eq:Swarm}] shows how velocity and direction of each particle are computed
+ \begin{equation}
+v_{ij}  = v_{ij}  + c_1 r_1 (lbest_{ij}  - p_{ij} ) + c_2 r_2 (gbest_{ij}  - p_{ij} )
+\label{eq:Swarm}
+\end{equation}
+ where $lbest_{ij}$ is the local best particle, $gbest_{ij}$ is the global best particle, $r_1$ \& $r_2$ are random variables and $c_1$ \& $c_2$ are the swarm system variables.
+ After each iteration the global best $g_{best}$ particle and the agent local best $l_{best}$ particle are evaluated based on the maximum fitness functions of all particles in the solution space.
+The solution is found after achieving a specific number of iteration or after an error threshold is achieved.
+Equation \ref{eq:descrite} is used to change the general swarm algorithm into binary particle (\textit{Discrete Particle Swarm Algorithm DPSO}) which handles particle values of either $0$ or $1$ \cite{PSODisceret}.
+ \begin{equation}
+   P(i)\Leftarrow 
+\{
+\begin{array}{c} 
+1 \quad \quad if\quad r_{3}>p_{i}  \\
 
-The recognized symbols must be verified with the rest of the document to fully understand the whole sketch.
-Consistency, ambiguities and uncertainty of the recognized symbols are the major issues that rise in any sketch understanding system.
-Most recent sketch understanding system uses more than just the context knowledge that is always available.
-Domain knowledge \cite {geometrydomain49}, Speech and verbal description of drawn sketch was used in\cite {speechMulti27}.Some verbal description of the drawn symbol helps in creating the casual model for each symbol.
-
-In the following sections, the ambiguities and uncertainty of sketch understanding systems will be discussed.
-Followed by how those systems represented knowledge that they use to formulate the understanding process.
-
-\subsection{Handling uncertainty and ambiguities }
-\label{sec:Handlinguncertaintyandambiguities }
-
-The main tasks of the sketch understanding are to handle higher-level ambiguities of the sketch drawn and to manage uncertainty of the recognized symbols.
-After the user draw a symbol using a single or multi-stroke symbol the recognizer labels each symbol with a known class either low level interpolation or high level.
-Low level interpolation means simple geometrical shapes and high level means more complex domain specific.
-Most current recognizers return a list of acceptable label for each symbol.
-The role of the sketch understating is converting such list of acceptable labels into one interpolation based on the surrounding symbols.
-They do so with preserving the sketch consistency of interpretation, which means that, the whole document has the same interpretation.
-
-The next section the various methods that was recently employ to completely understand sketched documents.
-
-\subsubsection{Rank and prune based on classifiers}
-\label{sec:Rankandprunebasedonclassifiers}
-
-The most used method in understanding the whole document and reducing the inconsistent solutions is the typical method of ranking the solutions and then pruning unwanted ones. 
-Initially the symbols nodes are initialized by all possible pairs of interpolation of the two symbols then the Consistency checker eliminate all inconsistence pairs.
-\cite {tobeadd} \cite {geometrydomain49}  firstly recognize all possible interpolation then prune and reason using knowledge to get a score for each interpolation.
-The final step is resolving and choosing the best result.
-
- \cite {Alvarado2002Framework11,HierarchicalParsing7}  First uses a bottom-top search to look for the interpolations of the strokes drawn.
-A top bottom search is used to prune out incomplete interpolations that was generated in the previous step.
-The top bottom firstly checks what symbol may be missing in the current document context after that it prune inconsistent interpolations.
-\subsubsection{Constrain satisfaction methods}
-\label{sec:Constrainsatisfactionmethods}
-
-Constrain satisfaction is the most used method for converting a group of recognized symbols into an understood document.
-Most recognizers produce a ranked set of possible interpolations.
-The problem is to constrain theses interpolations to get a total  understanding of the sketch document.
-Some like in \cite {tobeadd}  use constrain satisfaction methods to apply rules and constrains to match interpolations and rules.
-Constrains must be carefully written to check that they are not rough or severe.
-\cite {tobeadd}  checked the constrains applied on this system and if they produce no match the system relax the constrain and runs again with more loss constrains till it find the correct interpolations.
-
-\begin{figure}
-		\includegraphics[scale=0.7]{../../neededfiles/Figures/CSPphysical.eps}
-	\caption[Physical Constrains on Symbols]{ The figure shows two nodes and the constrain applied on both nodes to detect the inconsistency in the interpretations \cite{physicalmeaning6}. }
-	\label{fig:CSPphysical}
-\end{figure}
-
- \cite {physicalmeaning6}  firstly try connection between symbols firstly using bounding boxes of the symbol to locate the intersecting symbols.
-The intersected bounding boxes are checked again for ten different types of connection approximately for each terminal of the symbol.
-After identifying which symbols are connected the problem is identify which interpolation make those connected symbols has physical meaning.
-This is a "constrain satisfaction" problem the system used two processing steps to solve it.
-Firstly, the system checks the consistency of node where the node in graph is connection between two adjacent symbols.
-That means that each two connected symbol are checked and every inconsistent interpolation is rejected and deleted.
-This generate a local consistent sub graph but there may be a global inconsistent which take us to the second step where the algorithm search the arc consistency.
-The arcs of the consistency graph are any pair of nodes that have a symbol in common.
-Each consistency arc is checked for local and global consistency.
-This step is applied iteratively till all consistency are cancelled and only one global interpolation is reached \textit{(see Fig[ \ref{fig:CSPphysical}])}.
-
- The use of constrain satisfaction has also helped in minimizing the search \cite {tobeadd} .
-The strokes are incrementally processed using the Constrain system which check for constrain between image feature.
-Each new stroke either a new partial interpolation is created or an old one is extended and cloned provided that no constrained is violated.
-
-\subsubsection{Bayesians networks}
-\label{sec:Bayesiansnetworks}
-
-Since the problem of deciding the best interpretation of a set of valid ones is a certainty problem researchers tried to solve it as a one.
-Bayesians networks were used to represent certainties in sketch documents\cite{SketchRead2007} .Bayesians classifiers were also used to help the understanding problem in various ways.\cite{statisticalparsing26}  Used Bayesians classifiers to help the statistical parser to recognize the best candidate based on the system rules.
-
-\cite{Alvarado2002Framework11,SketchRead2007} is a complete sketch-understating framework that used top bottom and bottom up search of interpolation of the strokes.
-The system represented uncertainty using Bayesians networks.
-Each compound object is represented by a network in which the object is the root \textit{see Fig[\ref {fig:anddefnetwork}]}.
- \begin{figure}
-		\includegraphics[scale=0.7]{../../neededfiles/Figures/anddefnetwork.eps}
-	\caption[Bayesian network]{Bayesian network of an And Gate as defined in \cite{Alvarado2002Framework11}}
-	\label{fig:anddefnetwork}
-\end{figure}
-
-\subsubsection{Other Methods}
-\label{sec:OtherMethods}
-
-Other Method were used in \cite{agent47,sketchinginterfaces2}to understand the sketch documents .
-For example \cite{sketchinginterfaces2} defined  visual and domain rules that applies on the sketch to finally understand the whole document.
-\cite{agent47} used software agents to resolve the conflict in curve matching algorithms he used.
-In addition to agents that resolve the conflict he also used agents to search for applicable features on the strokes.
-The agent will either report a feature on the strokes drawn or disappear till new strokes are drawn.
-
-\subsection{Representing knowledge }
-\label{sec:Representingknowledge }
-
-Representing knowledge is the one of the challenges that faced the researcher in any AI program.
-In sketch recognition and understanding, the knowledge can be from more than one resource.
-The real problem is not to dump the program in too much data knowledge that is unstructured or unused in processing.
-To do so, means that the system is exhausted in undesired and unfulfilled processing.
-Processing time in sketch understanding system must be as close to real-time as possible to preserve the user interaction with the system.
-Therefore, the knowledge must be represented in special method that insures high speed and accuracy of understanding of the drawn sketch.
-
-Systems tried to make uses of more than one kind of knowledge to help in there recognitions and understandings.
-Spatial and visual knowledge of the drawn symbols are widely used in the sketch systems that relies on grammar languages \cite{GenericHMM28,statisticalparsing26,SketchRead2007} .
-Speech and multimodal systems helped the recognition of the symbols and reduced the search space.
-Domain and grammar knowledge were also used in \cite {geometrydomain49}  .
-
-Black board architecture \cite {blackboardur} was widely used to represent the various source and forms of data in the sketch understanding systems.
-In \cite {Alvarado2002Framework11}  black board was used to represent incremental knowledge which increase by each symbol drawn in the sketch.
-The black board architecture is simply a modeling of group of people trying to solving a problem.
-When any one of them find new information he write it down on the black board soon all information will be in the black board for all solver to see and add.
-
-\subsubsection{	Spatial and Visual knowledge}
-\label{sec:SpatialandVisualknowlege}
-
-The visual and spatial information of the drawn symbols with respect to other symbols in the sketch helps in identifying the relation of these symbols with each other.
-Those information was used in \cite {Cali63,visualpattern43,incrmentintention41,geometrydomain49,marksacts40,sketchinginterfaces2,captureknowledge19}  to help the recognition and understanding of sketch drawn.
-
-\cite{SRGraph57}  Introduced a new spatial relationship graph to represent those spatial relation of one symbol to the other.
-The Spatial relation graph is build from the sketch drawn using primitive spatial adjacency relation and relative position relation.
-The produced graphs are then matched using a graph-matching algorithm with the database of graph to recognize best fit.
-
-The spatial and visual information are used in \cite{XPGParser59} to help in building a parser that will parse the sketch document into symbols and objects.
-
-\subsubsection{Domain and Context knowledge}
-\label{sec:DomainandContextknowledge}
-
-\begin{figure}
-	\centering
-		\includegraphics[scale=0.8]{../../neededfiles/Figures/domainexamples.eps}
-	\caption[domain examples]{Examples of sketched documents from different domains}
-	\label{fig:domainexamples}
-\end{figure}
-
-Users can draw sketches in different domains \textit{see Fig [\ref{fig:domainexamples}]}and Until now there have not been a sketch understanding system that can identify drawing in any sketched domain.
-To do so, the system must be multi-domain system and employ a kind of domain identification or classification.
-Current systems uses domain and context knowledge to assist in the understanding procedure\cite {geometrydomain49,RetargetableInteractive58,sketchunderstanding1, HierarchicalParsing7,Mathpad46} .
-
-ASSIST \cite{AlvaradoFreedom42} uses domain knowledge, context memory and user feedback to help in understanding procedure.
-Domain knowledge to get a set of possible interpretations then context knowledge is used to prune out inconsistent finally the user feedback is used to correct error interoperations.
-
-The \cite{incrmentintention41} store domain knowledge in a memory like architecture where It stores the domain knowledge in long-term memory were strokes are firstly checked for possible interpolations.
-Context knowledge and previous interpolations are stored in short-term memory to help more accurate recognition rate.
-The context and domain knowledge are also used to feed rule-based systems that uses visual language rules in making the final document recognitions \cite{interpretationmechanical50}.
-
-%&&&&&&&&&&&&&&&&oksearch 
-\subsubsection{Using speech and various other knowledge}
-\label{sec:Usingspeechandvariousotherknowledge}
-
-Multimodal systems are system that uses more than one source of input devices as input to the system.
-Theses systems help in providing sketch understanding with more information's and data that will help in the recognition procedure.
-Speech and verbal explanation of drawn symbols are most used data in multimodal system to aid sketch understanding \cite {cognitivesketch18,speechMulti27,computationalmodel16} .
-
-\subsection{Space Search}
-\label{sec:SpaceSearch}
-
-The solutions of any sketch understanding are not found until the problem of minimizing and find an efficient way to search the solution space is solved.
-The solution space increase exponentially with each stroke the user draws dependably on the system.
-Either ways the space search problem must be addressed to properly solve the problem of sketch understanding.
-
-\cite {EfficientAbstract39}  illustrate a new search space minimization where the recognizer can perform one of the checking for constrains between images feature.
-Check for partial interpretation by assigning an image feature to a model feature or the user draws delaying until more strokes.
-Every new stroke is drawn either a new partial interpolation is created an old interpolation is extended and cloned provided that no constrains or rule is violated.
+0 \quad \quad if\quad r_{3}<p_{i} 
+\label{eq:descrite}
+\end{array}\}
+\end{equation}
+ where $p_{ij}$ is the numerical values of the particle and $r_{3}$ is a random variable.

### 2009-04-11 draft/Results.tex
now is 11 april

@@
 \chapter{Evaluation}
-
+\label{sec:Evaluation}
 This chapter discribes and illustrates the experiments performed on the system.
@@
 %The accuracy is computed as the number of correctly recognized samples divided by the total number of samples in the test. 
-
 %The recognition system evalutation is presented in Section \ref{sec:SegmentationAlgorithms}. The  
-
 %The variable used to tune the swarm algorithm were tested to choose the best parameters to use
@@
 
-Figrue \ref{fig:swarmtesting0} shows the Average segmentation error using different preliminary calculation (Section \ref{sec:CurvatureCalculation}).
+Figrue \ref{fig:swarmtesting0} shows the \textit{Average Segmentation Error} using different preliminary calculation (Section \ref{sec:CurvatureCalculation}).
 The figure shows that using all computations (time difference, direction, speed and curvature) for each stroke decrease the Segmentation error.
-Figure \ref{} shows the average number of Possible dominate point $P_{pd}$ with respect to information used in the system.
+Figure \ref{fig:swarmtestingpso} shows the average number of Possible dominate point $P_{pd}$ with respect to information used in the system.
 The results shows that the using all information increases the number of Possible dominate point $P_{pd}$ which in segmenation stage decreses the \textit{Segmenation Error}.
+
+ \begin{figure}
+	\centering		
+	 \includegraphics[scale=0.3]{images/swarmTesting}
+	 	\caption{Swarm testing} Segmentation Error Vs. Number of agents
+	 	\label{fig:swarmtesting0}
+	%\caption{Experiments results :} a)    b) c) 
+\end{figure}
+
+\begin{figure*}
+	\centering		
+	 \includegraphics[width=18cm]{images/swarmTesting2}
+	 	\caption{Swarm testing} Segmentation Error Vs. PSO parameters.
+	 	\label{fig:swarmtestingpso}
+	 	\end{figure*}
 

### 2009-04-11 draft/chapterSurvey.tex
now is 11 april

@@
 %draft_chaptersurvey 
-
 %\section{Introduction }
@@
 
-To step back from properties of individual strokes several researcher tried to use shape properties on the whole shape to classify shapes. The properties used
+drawing back from properties of individual strokes several researcher tried to use shape properties on the whole shape to classify shapes.
+They used global features that tries to define the entire shape properties.
 
-A number of approaches have stepped back from the properties of individual strokes to
-classify shapes based on a set of properties calculated on the whole shape. Properties
-that attempt to summarize the information in the entire shape are called global
-features.
-In [6] a carefully crafted set of filters based on global features, such as the ratio
-of their bounding box area to convex hull area and the ratio of the perimeter to the
-area, were used to progressively eliminate possible interpretations for a stroke until a
-suitable interpretation was found. For example, rectangles can be distinguished from
-triangles by looking at the ratio between the area of the convex hull and the area of
-the rectangular bounding box. The ratio for rectangles will be near 1.0 and the ratio
-for triangles will be near 0.5.
-A similar set of features was used by Fonseca et. al. in [14]. In that system,
-a carefully selected set of features was used by a number of rules and fuzzy logic
-to perform classifcation. Their system could also be trained to learn to distinguish
-between shape classes using a Naïve Bayes model based on the features.
-A more general approach using Zernike moments was demonstrated by Hse et. al.
-in [21]. Instead of using a hand tuned collection of properties they make use of the
-magnitudes of Zernike moments. Zernike moments are a class of orthogonal moments
-which describe the distribution of points in the input shape. Higher order moments
-represent finer levels of detail in the point distribution. The magnitudes of these
-moments, calculated up to a given order, form a feature vector which can be used to
-train an SVM. The magnitudes of Zernike moments have been shown to be invariant
-to both rotation and reflection of the input shape. A comparison of Zernike moments
-to our match vector representation is presented in detail in Section 5.2.2.
-These types of systems are based on global properties of shapes. Shape classes are
-identified by determining which properties each class tends to have. The properties do
-not typically depend on the number or order of the strokes and are thus based on the
-appearance of the shape and not how it was drawn. However, they do not represent
-the individual details of the shape. For example, it would be impossible to distinguish
-between the symbol for a current source (circle containing an arrow) and an ac-source
-(circle containing a "tilde") using just the ratio of convex hull area to bounding box
-area. This approach cannot distinguishing between shapes that difier by small details,
-and cannot deal with shapes that allow substantial conceptual variation. Handling
-these requires an approach that can represent the shapes at multiple levels of detail
+ Several systems \cite{DiagramOfflineConvexHull,MulitStrokeConvexHull} presented features like ratio of bounding box to convex hull area and the ratio of the preimeter to the area.
+These global feature can be used to distinguish between classes.
+For example, a rectangle can be distinguish from a triangle by looking to the raito of area of bounding box to the area of the convex hull.
+The triangle ratio will be much lower than that of the rectangle.
 
-The application of Statistical classifiers in recognizing shapes are widely used in sketch recognition.
-Statistical classifier requires training time before the recognition actually happen.
-Then first step of classification is features computation that is done in the previous step.
-Finally, the recognition of symbols will be based on the classifier and the training set used.
+ Fonseca et al.\cite{Cali63} created a system using similar set of features, their features were used by a number of fuzzy logic rules to identify symbols.
+They also used Naïve Bayes model that was trained to classify shapes based on those features.
 
-One of the first systems in sketch recognition is \citep{gestureexample12,aideddesgin22} algorithm to recognize ink gestures.
-He used Gaussian statistical classifier to classify any stroke user draw.
-In \cite {gestureexample12}  he firstly compute statistical information about the user stroke then pass it to a Gaussian classifier that had been trained using 10-20 samples.
-The output of the classifier is classification of the stroke from one of the known classes.
+%Zernike moments were used by Hse et al. in \cite{}. 
+Hse et al.
+\cite{HeloiseBeautification,zernike61} proposed using the magnitudes of Zernike moments as a shape property.
+Zernike moemnts are orthognals moments which define the distribution of point in the input space.
+The higher the order of the moment the finer the detail level the point distribution represents.
+Hse et al.
+\cite{HeloiseBeautification,zernike61} used a SVM classifier to train the features vector calculated from the moments.
+Early evluation of Zernike moemnts showed that they are invariant to both reflection and rotation of the input shape \cite{ }.
+In our system we used Zernike moemnts as part of our feature vector.
+A detailed comparison between Zernike moments and other features are demonstrated in Section \ref{}.
 
-Bayesian classifiers was used in \cite{Vibratory8} as a stroke classifier to recognize different shapes.
-The recognition is done using feature based statistical Gaussian distribution and Bayesian classifier techniques.
-The trainer use nine features extracted from segmented symbols (ex number of line segments, number of L intersections, ) all feature has a tolerance within.
-After all nine feature are computed a Gaussian statically model is constructed.
-
-Some systems employ more than one statistical classifier to get better results for the recognition.
-For example \cite{zernike61} uses Neural network, SVM (support vector machine )and MMD.
-\cite{polygonfeedback31}
-%the closed shaper recognizer which is done using classifiers (SVM , Rule based and ANN ) to make a 
-%classification of triangle, rectangle or ellipse
+ Systems based on Global shape properties distinguish shapes by learning the different properties each shape class tend to have.
+These properties dose not depend on the order or number of strokes and thus are only based on apperance of the shape not how the user draw it.
+However, individual details of the shape are impossible to distinguish using these properties.
+For example, it is impoissible to use  to  differentiate between a current source symbole (circle contiaing and arrow Figure \ref{}) and a clock (circle containing a square signal Figure \ref{}).
+Finear details of the symbols are not represented in this approach, thus shapes that have smalls details or only significant conceptual variation are not distiguished using this method.
+These symbols requires a system to represent different levels of shape details.
 
@@
 \label{sec:ApperanceBasedRecognition}
-Template matching is the most used method to recognize strokes and symbols into classes.
-Image based template stored on the database is matched with the one user draw to find the correct classification of the strokes drawn.
-In most cases, the stroke template is stored as image bitmap \textit{Fig \ref{fig:template}}in the database when the user draws the symbol it is matched using classifiers.
-\cite {imagetrainable48,HierarchicalParsing7}  used four kind of classifiers to detect the matching template.
-He used two classifiers to measure similarity and two classifiers to measure dissimilarity then combine them to find best match of drawn stroke from the database.
 
-\begin{figure}
-	\centering
-		\includegraphics[scale=0.7]{../../neededfiles/Figures/template.eps}
-	\caption[Template matching]{Examples of image templates used to match symbols in \cite{imagetrainable48}. }
-	\label{fig:template}
-\end{figure}
+In the third approach, the recognition foucus on appearance of shapes contrary to individual strokes or global shape properties.
+Kara \cite{imagetrainable48} used this approach while presenting a system that matches the input shape into a one of the  prototypes in the database.
+Figure \ref{fig:template} shows an example of prototypes of some of the shapes in the system.
+Input shapes are first pose and scale normalized then four image similarity measure are used to match input shapes to the database prototypes.
+He used two measures (Hausdor distance )  to measure similarity and two coeficients (Yule, Tanimoto) to measure dissimilarity.
+His approach allows the system to recongized a shape using only a single training sample as it is based on the shape prototype.
+However, the shape protoytpe must represent all  transformation and variation the input shape can have.
+Moreover, they restrict the image representation to 48x48 pixel due to computationa reasons.
+This restriction leads to elimintating some of the fine details that may be needed to distinguish some shape classes.
+There results on isolated symbols seems promising but the system have not beet tested on full sketch that tends to have high overlapping between symbols.
 
-To refine the matching procedure and create more robust classification \cite {imagetrainable48}  first transform the strokes into polar coordinate then match the transformed stroke with the templates in the database.
-They claim that enhance the classification procedure as the polar coordinate system due to its definition make the stroke more robust to transformation and rotation.
+\subsection{Discussion}
+\label{sec:discuss}
 
-\cite{templatefrag21}  Introduce new way to match stroke temples.
-He store template as a sequence of lines and ellipse.
-When the user draws a stroke the system, segment it into a sequence of lines and ellipse.
-The system then matches the stored sequences with computed one to find correct classification.
+In all of these system only \cite{4} using stroke based recognitiion have been implemented to the fully creat a system that free draw sketch and understand them.
+This is because the emphasis on the interactive side of the sketch sytem which result in large difficultes in processing free hand sketches.
+Methods based on gesture and global stroke properties produce systems that are dependent on assumptions about how strokes can be groubed.
+These systems assume that users will finish drawing a shape before moving to the next one \cite{6,11,14,15}.
+This let them to temporally group strokes and then recognize them based on their strokes or apperance properties.
+\cite{5} argued that this assumption dose not hold in natural sketchs.
 
-The main problem of template matching is that in most system the templates are hard coded inside the system and cannot be modified at runtime.
+Recogntion based on strokes using hierarchical shape description is applied in \cite{}.
+The principle problem in these systems is the computatnal complexity of the matching between shape description and geometeric primitves produced in the segmentaiton process.
+Natural sketch are dificult to segment and group to create reliable fragment of geometeric primitives.
+This is partially due to the increase of noise and other phenomena such as overlapped, overtraced strokes and multiple shapes drawn using
 
-\paragraph{Graph language and visual language describing }
-\label{sec:Graph language and visual language describing }
+Natural sketches are dificult to reliably fragment into geometric primitives, because of the increase in noise and other phenomena such as interspersed drawing of parts from multiple shapes, multiple
+shapes drawn with single strokes, and overtraced strokes. Because these systems are
+dependent on the fragmentation process, they are either faced with performing recognition
+with unreliable fragmentation results or must consider many possible ways to
+fragment the sketch. Avoiding the fragmentation issue was one of our key motivating
+factors in taking an appearance-based approach to recognition.
+The appearance based methods described above have not been applied, to our
+knowledge, to complete, freely drawn sketches. They have only been applied to
+isolated shapes or synthetic images generated from isolated shapes. Although their
+focus on the visual nature of the sketches appears promising for handling freely drawn
+sketches they have not yet been evaluated on them.
 
-Researchers tried to generate a visual language that can describe hand drawn sketches.
-In such systems, sketch recognition will be a question of parsing the sketch document into recognizer that parse strokes according to the constructed grammar.
-
-Researchers proposed more than one visual grammar to use it in sketch recognition.
-\cite{statisticalparsing26} proposed a language parser that will parse sketched strokes using grammars rules that was constructed .
-The system first construct a parser that will parse drawn strokes into tokens then formulate those tokens using the grammar rules stored.
-Parsing is done in two steps; Firstly, the system should accept input of strokes then it use a statistical parser to generate the parse tree.
-The statically parser make use of the stored rules and probability of drawn stroked and uses Bayesian classifiers to produce the parse tree.
-
-%\citep*{XPGParser59} i need author   
-Introduced a grammar they called an extended positional grammar (XPG) that is based on XpLR parsing methodology \cite{XPGParser59} .
-The XPG imagine the symbol with a set of attribute as a sentence that can be parsed using the physical, syntactic and semantic information in sketch.
-They divided the grammar into two kind the first is named ink grammar, it define the symbol of the language as a composition of primitive geometric shapes.
-The second is language grammar that they use to identify the sketch as a composition of shapes defined by the ink grammar relation.
-The parsing itself is done using a incrementally bidirectional parsers.
-They also add a sketch editor to the system to help user to add new symbols and generate its XPG grammar for it \textit{(see Fig[\ref{fig:symboleditor}])}.
-
-\begin{figure}
-	\centering
-		\includegraphics[scale=0.7]{../../neededfiles/Figures/symboleditor.eps}
-	\caption[symbol editor]{The Figure shows the symbol editor that\cite{XPGParser59} uses to generate grammar rules.}
-	\label{fig:symboleditor}
-\end{figure}
-
-Relational grammar was also used to help in reducing the ambiguity with the help of fuzzy logic\cite {visualpattern43} .
-The fuzzy set is used to resolve the visual relations of the sketched stroke with others strokes.
-Spatial relations like (left of, right of) are computed and then passed to the fuzzy sets.
-The fuzzy set associate a degree of certainty to the shape defined.
-%EfficientAbstract39 also used the grammar language ladder
-\cite{Ladder30,GenericHMM28}  Used visual description language which will generate the information of the structure and constrain of each object classified.
-The system automatically generates java code for the recognizer for each object.
-The generated code cycles through the strokes drawn to check if any of the recognizers generated had interpreted any symbols partially or completely.
-HMM network is used to partially segment the strokes drawn to help the recognition process.
-
-Similar system was constructed in \cite {Ladder30} , where language grammar is written to descript complex shapes ( see \textit{Fig\ref{fig:arrowladderdef1}}) form primitive geometrical, constrains and triggers.
-The grammar is then passed to "jess" a java rule bases system that produce java code used in recognition of  drawn strokes.
-
-\begin{figure}
-	\centering
-	\begin {center}
-		\subfigure{An Arrow}
-			\centering
-
-			{\label {fig:arrow1}\includegraphics[scale=0.75]{../../neededfiles/Figures/arrowladderdef1.eps}}
-		\subfigure[Arrow grammar definition]{\includegraphics[scale=0.88]{../../neededfiles/Figures/arrowladderdef2.eps}}
-		\end {center}
-	\caption[Arrow definition]{The figures show and regular arrow with 2 shafts and head and its definition in Ladder language \cite{Ladder30}}
-	\label{fig:arrowladderdef1}
-\end{figure}
-
-\subsection{Other hybrid methods}
-\label{sec:Otherhybridmethods}
-
-Due to the complexity of the sketch recognition problem, more than one system proposed a hybrid method to solve those problems. They combine more than one approach to recognize symbols the user draws.
-\cite{sketchunderstanding1,Simusketch51}  used four kind of classification image, graph matching, statistical and structure based classifiers.
-Each symbol will contain all graph, statistical, structure information and a (Wight, recognizer) tuples to determine which recognizer has the most influence in recognizing this symbol see \textit{Fig\ref{fig:structuralscemanticdeff}}.
-\begin{figure}
-	\centering
-		\includegraphics[scale=0.5]{../../neededfiles/Figures/structuralscemanticdeff.eps}
-	\caption[Structure definitions]{Example of the structure definition of sin symbol in \cite{sketchunderstanding1}.}
-	\label{fig:structuralscemanticdeff}
-\end{figure}
-
-\cite {HMM53}  firstly normalize the stroke and compute all the statistical features from it.
-After that, a noise reduction algorithm is used on the strokes which will be translated to produce size and rotation invariance data.
-Those data is passed to a multistage hybrid recognizer.
-The first step of his multistage classifier is dynamic programming procedure to compute the optimal degree of similarity and differentiation of symbols.
-Secondly, a statistical feature ranking is employed on the data produced.
-Finally, the system will reject all unlikely classification and produce best fit for this stroke using a dynamic programming using domain and context knowledge as input.
-
-\cite{statisticalparsing26} is a statically parser make use of the stored rules and probability of drawn stroked and uses Bayesian classifiers to do the classifications.
-
-%\subsubsection{Fuzzy classifiers }
-%\label{sec:FuzzyClassifiers}
-%Fuzzy classifiers is used to help in identifying degree of certainty given to certain interpolation the
-
-%\subsection{Hybrid Systems}
-%\label{sec:Hybrid Systems}
-
-Given that the sketch recognition of hand drawn strokes is exceptionally challenging problem the researcher tried to take advantage of additional of combining approaches to achieve higher recognition rate.
-
-Most of the problems of the sketch recognition are not still addressed.
-To do so, The sketch recognition system must maintain reliably interactive and extensible interface.
-The users minor change can cause the recognizers to fail.
-The user sloppy drawing with over or/and under shoot strokes  \textit{Fig\ref{fig:overstroked}}.
-Over processing of data and interaction with background are still major problems in the sketched documents.
-
-To solve some of these problems,\cite{threeproblmes23}  revealed a two stages recognizer that will firstly construct a data graph in which strokes will be added.
-After constructing the full data graph the recognizer compute all possible solutions and reject any inconsistent alternatives.
-Second stage, Employs a topology structure based on the geometrical matching using a CSP making use of sub graphs.
-
-As mentioned before,\cite{visualpattern43,Cali63} combine fuzzy logic sets with language grammar to create more robust classifiers.
-Grammar rules are used to describe the spatial relationships of the strokes.
-Fuzzy sets are used to handle the uncertainties of the relation and classification.
-Therefore, by combining Fuzzy set and spatial grammar rules the system was able quantify the spatial and statistical attribute calculated before in \ref{sec:Statisticalmethods} from the drawn symbols.
-
-Also the use of HMM with generated visual grammar was introduced in the system by \cite{HMM53} where they use HMM as a stroke segmentor and pass the output to the visual parser.
-furthermore, \cite{SmartSketch56} used more than one recognizers to recognize closed shapes rule based approach with hypothesis and test strategy, Neural network and SVM.
-
-\paragraph{Dynamic programming}
-\label{sec:DynamicProgramming}
- was used by \cite {Mathpad46,templatefrag21}  to calculate the optimal degree of different and similarity of symbols.
-The system also uses a dynamic programming procedure to find a best fit from the given set of classifications.
  \chapter{Particle Swram Optimization}

### 2009-04-12 trunk/draft/Results.tex
now it 9 - appril and after club day
12-4

@@
 The effect of number of agents, the error thresholds and various other PSO paramters are investigated in Section \ref{sec:PSO}.
-The results presented in Section \ref{sec:ResultsDetails} and \ref{sec:PSO}.
+%The results presented in Section \ref{sec:ResultsDetails} and \ref{sec:PSO}.
 
@@
 %The accuracy is computed as the number of correctly recognized samples divided by the total number of samples in the test. 
+%The recognition system evalutation is presented in Section \ref{sec:SegmentationAlgorithms}. The  
+%The variable used to tune the swarm algorithm were tested to choose the best parameters to use
 
-%The recognition system evalutation is presented in Section \ref{sec:SegmentationAlgorithms}. The
-
-%The variable used to tune the swarm algorithm were tested to choose the best parameters to use
 \section{Datasets}
@@
 Four different splits for the training and test data are generated from the dataset.
-%The results displayed are the average recognition accuracy of the four splits resulted from the classifier after recognizing the symbols.
+In worth noting that \cite{HeloiseBeautification} and \cite{Oltmans07} achieved 96.7\% and  94.4\%  on the dataset respecitvily.
+%, 94.4\%  , which
+is comparable to the 96.7%  %The results displayed are the average recognition accuracy of the four splits resulted from the classifier after recognizing the symbols.
 		\item [MultiDomain Dataset] A dataset we collected manually from different users.
@@
 Data are collected from 4 different users each of them have drawn from 20 to 40 samples of each shapes.
-Figure \ref{fig:MultisymbolSet} shows some of the shapes used in the dataset.
+Figures \ref{fig:LogicSet} and \ref{fig:ElectSet} show some of the shapes used in the dataset.
+Symbols from different domains were gathered to evaluate the system thourhgtout different types of shapes.
+Useres were requested to draw symbols without any restriction on style,size or orientations.
 The dataset was divided into training set and test set.
@@
 \begin{figure}[]\centering
+\fbox{  \parbox{6cm}{% 
+		\includegraphics[scale=0.7]{images/logicSymbols.eps}	}
+		}
+	\caption{Example of Logic Symbols}
+	\label{fig:LogicSet}
+\end{figure}
+\begin{figure}[]\centering
 \fbox{  \parbox{8cm}{% 
-		\includegraphics[scale=0.7]{images/symbolSet.eps}	}
+		\includegraphics[scale=0.7]{images/SymbolsElectricalr.eps}	}
 		}
-	\caption{The MultiDomain Dataset }
-	\label{fig:MultisymbolSet}
+	\caption{Example of Electrical Symbols}
+	\label{fig:ElectSet}
 \end{figure}
@@
 \begin{description}
-	\item[Segmentation Error] The segmentation error is computed from each stroke after segmentation as computed from Equation \ref{} in Section \ref{}. 
+	\item[Segmentation Error] The segmentation error is computed from each stroke after segmentation as computed from Equation \ref{eq:bestFit1} and \ref{eq:bestFit2}. 
 	\item [Recognition accuracy] We measure recognition performance for each algorithm by determining the number of correctly identified symbols in the whole dataset.

### 2009-04-12 trunk/draft/chapterSurvey.tex
12-4

--- 2008-09-02 trunk/draft/chapterSurvey.tex
@@
 \chapter{Literature Survey}
+\label{sec:survey}
+% this will contians only the finished work .... and my comments to view the old survey see the file
+%draft_chaptersurvey 
+%\section{Introduction }
+%\label{sec:Introduction}
+The history of pen and sketch based computing goes back to 1963 with Ivan Sutherland's SKETCHPAD \cite{sutherlandsketch}, which used a light pen to draw on the monitor to create circuit diagrams.
+Current technology and the increae in computuing power have increased the abilities of such interfaces and allowed the implementaion of fully sketching systems.
+Over the years, many other sketching based system have been developed, elthough throught this history most system foucused on building the interactive interfaces to assist users while entering information to the computer.
+To demonstrate our system inspiration we will first present current research in the area of sketch recognition.
+Fristly, preprosessing and segmentaiton methods will be discussed in details.
+Later, symbol recognition methods will be presented and analysied in Section \ref{sec:symbolrecogntion}.
 
-\section{Introduction }
-\label{sec:Introduction}
-
-The researchers had divided the sketch recognition problem into image based and AI Based problem.
-The next section will discuss the methods used by current sketch understanding system.
-Firstly, in section \ref{sec:FirstStepSymbolRecognition} Symbol recognition research will be investigated and divided into image based and AI based methods.
-
- In section \ref{sec:SecondStep:SketchUnderstanding}, the understanding techniques will be demonstrated.
-Understanding mainly will handle knowledge representation methods, handling uncertainty and finally how researcher in sketch reduced the sketch search space.
-
-\section{First Step: Symbol Recognition}
-\label{sec:FirstStepSymbolRecognition}
-
-To understand a drawing we need to recognize every symbol in it. Hand-drawn sketches need more processing before it is passed to recognitions algorithms.
-
- Firstly, we have to determine how symbols are defined or grouped.
-Some System implies some constrain on user to defined symbol.
-In\cite{gestureexample12,aideddesgin22,sketchinginterfaces2,sketchinginterfaces2,cognitivesketch18,ewhitboard9}for example the user is constrained to draw each symbol using only one stroke which is not only unnatural but also constrains the symbols that can be drawn.
-More recent systems force the user to state when symbol is finished by pausing a period of time or clicking a certain button in the interface.
-Only few systems permit the user to freely draw strokes and symbols in any order \cite{Vibratory8,visualpattern43,Cali63,statisticalparsing26,physicalmeaning6}.
-Those systems must employ a method to detect when a user starts drawing a symbol and when he end.
-
-Currently symbol recognition is both image based and feature based or AI-based.
-In image based recognition symbol is recognized based on there features, structure, some image computed feature .
-The AI methods in recognizing symbols mainly imply more time and processing as for the search and rules applied but claims to be more accurate .
-
-\subsection{Pre processing }
-\label {sec:preprocessing}
- A stroke is all the points that lie along the path of the user pen from the moment of his pen down to pen up events.
-Not all those points are important even more, there may be redundancies unwanted points.
-Therefore, those strokes will need some preprocessing.
-
-The stroke that the user draws are mainly sequences of line segments or/and curves which the any system need to divide them to recognize them easily.
-The main problem is to do so without constraining the user drawing style.
-The majority of researchers first use a preprocessing process to identify those segments and the critical points from the user strokes.
-Beside points location information (x, y positions) researcher made uses of the extra information from using digital ink.
-Some as \cite {mulitstroke5,polygonfeedback31}  used user speed information to detect critical points where it is stated that user slow down at those critical points.
-Other used curvature and directions data.
-There was also mentioning using the pressure information.
-Thought some \cite {MilitaryCOA37,computationalmodel16,polygonfeedback31}  debated on the importance of the extra information and weather the temporal and speed information is significant.
-\cite{polygonfeedback31}  used the timestamp for each point and compute time difference between every two successive point to help him finding corners.
-\cite{Overview36} uses the speed and direction information but they use on it a hybrid average based filtering to filter noise and unwanted false data from appearing in the data.
-Finally, they use that information to get a poly line approximating.
+% i need to write a history then small introudction to the survey...
+%  i have to state that the next  sections will generally group segmentation and symbol recognition task. 
+\section{Preprocessing and Segmentation}
+\label{sec:preprocessing}
+%this section should describe how others did the first preprocessing of the problem. i have to summarize what i wrote before. 
+ % first paragrah about how stroke is a messing thing and naturally it is ambigues. 
+  In a system that supports free hand sketching, sketching should be as natural as possible.
+This means that the user can draw symbols using single or multiple strokes without constrains.
+The goal of a sketch system is making the user feel that using the using the system is nearly the same as using pencil and paper.
+But sketchs are incomplete drawings that are sloppy and messy, for example figures \ref{fig:Overshootandundershoot} and \ref{fig:overstroked} show examples of messy sketches drawn by users.
+This led researchers to perform sort of preprocessing before starting the sketch recognition task itself.
+The next sections explain different preprocessing techniques.
 
@@
 		\includegraphics[scale=0.7]{../../neededfiles/Figures/overstroked.eps}
-	\caption{Example of an overstroked stroke}
+	\caption{Example of an over traced stroke}
 	\label{fig:overstroked}
 \end{figure}
-Some systems as in \cite{SmartSketch56} where the noise reduction is done in two steps the first step is the polygonal approximation to remove non critical points in the user stroke.
-The process is mainly removing the redundant points that are along the path of stroke.
-Due to the shaky nature of user stroke, there are additional end-points at the stroke segments.
-Secondly, the next step will be applying an Agglomerate point filtering.
-Agglomerate filtering is according to \cite {statesurvey35} removing the undesired extra points that is produced due to the unsteady operations produced while lifting and touching tablet.
-The results of those processes are passed to final end-point refinement to merge or remove any redundant points.
 
+ %The preprocessing can be noise removing, stroke beautification or segmentation.  %Because of this 
+ Sezgin and Davis \cite{overtraced24} tried to solve the over traced strokes (see Figure \ref {fig:overstroked})using a thinning algorithm.
 In \cite{meanshift10,domainindependent17}, Bo Yu uses mean shift procedure as a noise remover and preprocessing step to the recognition step.
 The mean shift procedure is simply shifting the point to the average of its neighborhoods so smoothing out the noise without changing the global structure of data.
-This process is clamed to help find critical points easily after that.
+Gaussian filters was used to smooth curves \cite {Phoenix88}.
+These procedures may seem to remove noises in the sketch but it often treat vertices as noise which obscure vertex locations.
+ %% now the third paragraph.....
+Beside points location information (x, y positions) researcher made uses of the extra information from using digital ink.
+Some as \cite{mulitstroke5,polygonfeedback31} used user speed information to detect critical points where it is stated that user slow down at those critical points.
+Other used curvature and directions data.
+%Others \cite {MilitaryCOA37,computationalmodel16,polygonfeedback31}, debated on the importance of the extra information and weather the temporal and speed information is significant.
 
-In \cite {domainindependent17}    approximation step uses the speed, curvature, and direction of strokes drawn by user to determine the location of vertices to gain more resistance to noise that data may contain.
+ Agar et al.
+\cite{polygonfeedback31} used the timestamp for each point and compute time difference between every two successive point to help him finding corners as it provides more distinct maxima than the speed information.
+He mentions that the pointing device (for example the mouse) sampling rate is the reason for this phenomena.
+The points are sampled at regular time intervals while the pointing device is moving.
+There are no points while the pointing device is stationary.
+This leads to a nearly constant time difference between samples while the pen is moving and large difference while the pen is stationary.
+Contrary to speed information, where the users draw with variant speeds the speed information has a lot of noise in the data.
 
-\subsubsection{Critical points and corners Detection}
-\label{sec:CriticalPointsAndCornersDetection}
+Sezgin et al.\cite{earlyprocess} used both speed and direction information.
+Their system gives better results than using only speed or direction information.
+This is because the speed information tend to be noisy when drawing large symbols, on the other side when using only direction information the some corner are missed.
+Our system uses speed, direction and time difference information in its preliminary calculation step.
+%both speed and direction information help detecting vertices when the
 
-Locating the critical points is provided to the next process of the systems which in most cases segment detection algorithm as in \cite{Simusketch51} .
-Other systems after detecting the points the system directly uses those information to template matching.
+ A genetic algorithm was used by \cite{CruveDivisionSwarm} to optimally divide digital curves into lines and curves.
+Chen et al.
+\cite{CruveDivisionSwarm} uses digital curves scanned from paper as input to the system.
+Also, they did not take advantage of the curvature or local geometric properties of the digital curve.
+Yin \cite{PolygonApproximationPSO} used PSO to convert digital curves into polygons, our system adopts  Yin \cite{PolygonApproximationPSO} method and tries to improve it by adding curvature and other local information while segmenting strokes to achieve better segmentations.
+%The next section presents the general particle swarm algorithm which is used in both our segmentation algorithms.
+% are presented in section \ref{subsubsec:Discreteparticleswarmalgorithm}
 
-Generally, the critical points along with the segment points of the path from one critical point to another are used to detect labeling of that segment. 
-Line detection or multi line segments algorithm is found in \cite{polygonfeedback31} .
-Mostly after detecting the vertices if the segment cannot be labeled as arc it is defined as a line.
-Other mainly defines a threshold value of the slope if it increase than that it is counted as an arc.
+  % third should speech on how systems define critical points (this paragrah should speak on the data collected (speed crital points, ...
 
-Systems like \cite{physicalmeaning6} detect critical point for each stroke then the final point is merged with the first if they are in range defined by threshold.
-However, stroke may also intersect with other strokes drawn before therefore \cite{physicalmeaning6}  introduced a structure to detect such approximates he first define every stroke drawn as series point and labeled as line, arc or area.
-Then after labeling most strokes the system check for proximity on bases of line -line, line-end point, or line- area.
+ After critical point detection most sketch system proceeds to label segments into simple geometric primitives (i.e line, arc, etc...).
+A hybrid algorithm was introduced in \cite{earlyprocess} where different sets of segments are generated based on both curvature and speed dominant points, followed by choosing a segmentation with the least error from a generated hybrid set.
+However, this system is limited to recognizing only specific simple geometric shapes with a set of low level recognizers.
+Each low level recognizer is designed to recognize only one geometric shape using spatial and geometric information extracted from input stroke.
+Yu \cite{meanshift10} introduced a \textit{feature area} for each primitive and then computed the segmentation error for different types of primitives based on the computed \textit{feature area}.
+Figure \ref{fig:featurearearc} shows the feature area of a line and circle in Yu system, the shaded area is the difference between the computed feature area and the assumption area.
+His system achieved good accuracy in simple shapes (square, ellipse,...etc) but did not perform well in complex shapes.
 
-In   \cite {meanshift10,domainindependent17}   approximation step uses the speed, curvature, and direction of strokes drawn by user to determine the location of vertices to gain more resistance to noise that data may contain.
-They firstly analysis the curvature and speed data and use the mean of data as the threshold then use that threshold to separate data into regions and finally locating the maxima and minima in each region .
-The previous step generates a set of vertices and its certainties they use to choose best fit.
-
-\subsubsection {Primitive stroke segmentation and clustering}
-\label{sec:Primitivestrokesegmentationandclustering}
-
-The most segmentation method is least square method in \cite {overtraced24}.
-Other systems  approximate a line segmented using the information from the direction curve where the curve is fitted with horizontal line.
-firstly the system compute least square fit for direction curve or the stroke point computed and check if deviation points is small if they pass then hypothesize al line segment  from the first point to the last one.
-
+ Paulson and Hammond \cite{Paleosketch08} introduced a set of low level recognizers that were reported to achieve 98\% accuracy.
+However, their system similar to all low level recognizers in the way that it identifies a small set of simple shapes.% A genetic algorithm was used by \cite{CruveDivisionSwarm} to optimally divide digital curves into lines and curves.
+Chen et al.\cite{CruveDivisionSwarm} uses digital curves scanned from paper as input to the system and did not take advantage of the curvature or local geometric properties of the digital curve.
+\cite{Fluid25} Trying to combine fast morphing and basic recognition for the sketch drawn as the user draw it.
+It tries to move the points of strokes drawn by the user draw into one of the known shapes (ex: box, lines and circle).
+It uses a relaxation and least square method for circles and uses the string of forces for finding best-fit box.
+However, their system can only identify circle and rectangles drawn using a single stroke.
 \begin{figure}
 	\centering
-		\includegraphics[scale=0.7]{../../neededfiles/Figures/rectanglerecog.eps}
-	\caption{rectangle recognition}
-	\label{fig:rectanglerecog}
-\end{figure}
-
-%//Other like \cite {tobeadd}  uses the relaxation process to try to relax parameter on a classification .//
- In \cite {meanshift10,domainindependent17}  Bo first uses the direction graph to check if the stroke can be approximated as a line or no.
-If the stroke can't be approximated as a line it is checked to be a circle or ellipse.
-the procedure is simply assume a circle classification then check this assumption with the user drawn stroke given a min error range.
-\cite {Fluid25} Trying to combine fast morphing and basic recognition for the sketch drawn as the user draw it.
-It tries to move the points of strokes drawn by the user draw into one of the known shapes (ex: box, lines and circle).
-It uses a relaxation and least square method for circles and uses the string of forces for finding best-fit box.
-
-The noise removal and critical point detection is only a preprocessing step for the symbol classification algorithms.
-Some system implements a segment segmentation and clustering to help in the classification problem.
-
-After the strokes are segmented \cite {sketchunderstanding1,HierarchicalParsing7,Simusketch51}  proposed a merging techniques that will merge strokes together into symbols to help classification.
-They use a method combine between divide and merge.
-They firstly divide the drawn sketches into section based on marker technique they used.
-After that they merge strokes if there bounding boxes intersects and then  merge generated cluster till no more merge can occur on one segment.
-
-%Intension exteraciton in paper Intention54
-% Check to add
-
-%/*Mathpad46
-%methion  check stroke clustering.  Simusketch51  and in 8 
-%
-%Bo \cite {tobeadd}   The ellipse or
-
-\subsubsection{Image and Statistical Features calculation}
-\label{sec:ImageandStatisticalFeaturescalculation}
-
-Most system calculates the general image properties for the strokes. Most used features are bounding box, width, size, points along path. 
-\cite {meanshift10,domainindependent17}  introduced a new feature which they named "Feature Area which" they used in primitive classification introduced before.
-They first make and assumption for classification and compute the difference.
-For example if they make assumption as a circle they first calculate the center of the circle then its radius based on the bounding box of the stroke.
-The area difference between the computed circle and the current stroke is the feature area.
-see \textit{Fig\ref{fig:featurearearc}}.
-
-\begin{figure}
-
-		\subfigure[Feature area of line] {\includegraphics[scale=0.7]{../../neededfiles/Figures/featureareline.eps}}
-
-	\subfigure[Feature area of an arc]{	\includegraphics[scale=0.7]{../../neededfiles/Figures/featurearearc.eps}	}
-			\caption[Feature area]{the shaded are shows the difference between the computed feature area and the assumption area. }
+	\subfigure[Feature area of line an arc]{
+		\includegraphics{images/featurearearc.eps}}
+		\subfigure[Feature area of line]{
+				\includegraphics{images/featureareline.eps}}
+	\caption{Feature area}
 	\label{fig:featurearearc}
@@
 
- System that uses statistical classifiers use other features for each stroke.
-Features like center of gravity\cite{gestureexample12,sketchunderstanding1}, maximum length line, Zenerik moments \cite{zernike61}, number of lines, curvature angles, minimum enclosing rectangle\cite {incrmentintention41} , convex hull were calculated in \cite {Cali63}.
-Also \cite {geometrydomain49} introduced a new feature which he named ink density that he used in the classification of the glyphs see \textit{Fig \ref{fig:inkdensity}}.
-\begin{figure}
-	\centering
-		\includegraphics[scale=0.7]{../../neededfiles/Figures/inkdensity.eps}
-	\caption[Statistical Feature Example]{The figure shows the Ink Density feature that \cite{geometrydomain49} introduced. }
-	\label{fig:inkdensity}
-\end{figure}
-All spatial relationships needed in the graph or template matching is computed after statistical features are calculated.
-Relations like intersections, containment, perpendicular, right-of, left-of and below are the most used relations concerning constructing graphs for symbols\cite {geometrydomain49,sketchinginterfaces2} .%stretchdiagrammer4,     i dont know where is this refrence now
+\section{Symbol Recogntion}
+\label{sec:symbolrecogntion}
 
-\subsection{Image Based Classification}
-\label{sec:Image Based Classification}
+In this section, we review three different types of symbol recognition methods.
+Symbol recognition can be based on: strokes, global shape properties, and appearance \cite{Oltmans07}.
+The role of which each stroke plays in the structure of the sketched symbol is used to determine shapes in stroke based methods.
+In global shape properties approach, systems investigate general shape properties and its underlying strokes.
+Using shape properties relaxes the assumption and role of each stroke in the symbol but it does not truly represent the appearance of the symbol.
+Unlike appearance based methods where individual strokes are disregarded and they focus on the overall appearance of shapes.
+We choose to use the appearance based method but added some geometrical and structure properties used in most stroke based approaches.
+Our method tries to avoid many of the problems faced by approaches based on individual strokes or global shape properties while trying to gain the additional information that the stroke based system use.
+In the next sections, we look into three different symbol recognition approaches, describe their advantages and challenges.
 
-In the following sections, systems generally try to recognize the strokes user draws and give it a label based on there image features.
-The information collected and features computed in the previous step.
- \subsubsection{Low Level Analysis}
-\label{sec:Low Level Analysis}
+\subsection{Stroke Based Recognition}
+\label{sec:StrokebasedRecognition}
+Stroke based approach is based on the assumption that each stroke has a specific role in representing the symbol.
+Stroke based methods examine each stroke as the user draw it to determine its role in the symbol.
+These methods are suited to interactive interfaces since users expected that system displays interpretation of the symbol after the stroke is drawn or after drawing a group of strokes.
+The majority of the research in sketch recognition has focused on stroke-based methods because of the corresponding emphasis on interactive interfaces.
+Most stroke based systems either represent symbols as gestures or   hierarchical shape.
 
-%
-%Detecting object using low level analysis only 
-%\cite {mulitstroke5} first detecting vertices and control points then fit these segment into line and arc using best fit and least square error classification.
+\subsubsection{Guesture Recognition}
+\label{sec:GuestureRecognition}
 
-%Check the DB
+Single stroke gestures recognition was the focus of the early work in symbol recognition \cite{gestureexample12,aideddesgin22,sketchinginterfaces2}.
+Gestures are pen strokes drawn by the user that can immediately recognized by the system.
+The result was either creating an action (e.g.
+deleting a shape) or drawing a shape on the screen.
+These systems are based on how the user draw the shape rather than how it looks.
 
-% description[4] after that it passes through the curve approximation into the Bezier curves
-%
-%\cite {statesurvey35} we need to fit and filter the classified shapes that is done using on of the following least square is the most frequently method used some use fuzzy logic to make circular fitting some use b-splines or ellipse and circle fitting.
-some after that do some  shape beatification or regularization.
+One of the first systems in gestures recognition is Rubine's algorithm \cite{gestureexample12} .
+He used Gaussian statistical classifier to classify gestures.
+Simple features as length of stroke, smoothness of stroke, the angle that the first part of the gesture is drawn at and properties of the bounding box of the stroke are used to train the classifier to recognize  symbols.
+In our system, we combine Rubine features with others to include the information on how the user draws sketch while recognizing the symbols.
+In \cite{Long00}  Long shows how similarities between gestures can be analyzed.
+He used this information to identify gestures that are likely to confuse the recognition system.
+This helped interface designers to choose gestures that are easily recognized.
 
-The most used method in low level recognition is best fit and least square methods where they was used in most systems.
-Other systmes like \cite{polygonfeedback31} prefered to implement its own polygon recognition algorithm.
+The main problem with gesture based recognition system that they impose several constrains on how the user can draw.
+Gestures must be drawn using a specific order and direction to be recognized correctly.
+Moreover, in most cases gestures do not necessarily look like the symbols they represent (e.g.
+to represent a rectangle the user draws only the left edge of it).
+Because of these reasons, gestures recognition systems are not suitable for free hand drawn sketches.
 
-Bo et al.
-In \cite{meanshift10} introduced an algorithm to recognize line and arcs.
-The system used the direction graph and fitted stroke points with a non horizontal line.
-So the recognizing is done using the following algorithm which is preformed before the mean shift procedure so not the destroy the direction graph .Firstly, if direction graph can be approximated using a line then it is recognized as a line.
-Otherwise the algorithm Bisect the line from the first and last  point of stroke and donate it as "M".
-Then calculate the intersection point of the line "M" and original stroke and donate it as "A" .
-Now using the three points first , last points of stroke and point A find the center of circle which passes through all of them which will donate the candidate arc.
-Finally , calculate the radius of the circle and the mean of distance  between center and each point of stroke and  compute the feature area of the stroke and the circle arc see \textit{Fig \ref{fig:featurearearc}}.
-%//
+\subsubsection{Hierarchical Shape Descriptions}
 
-\subsubsection{Template matching}
-\label{sec:Template matching}
+A Hierarchical shape descriptions was widely used in sketch systems \cite{sketchunderstanding1,HierarchicalParsing7,Ladder30,AlvaradoFreedom42,napkinGross}.
+Symbols are represented using hierarchical levels.
+The lower levels represent simple geometric primitives such as lines, arc and ellipses.
+Intermediate levels shapes are composed of lower level parts and constrain or connections between them.
+For example, a triangle is described as three lines connected at their endpoints.
+Higher level shapes are constructed from intermediate and lower level shapes, for example, a diode is described as a line on the top of a triangle head (acute angle).
 
-Template matching is the most used method to recognize strokes and symbols into classes.
-Image based template stored on the database is matched with the one user draw to find the correct classification of the strokes drawn.
-In most cases, the stroke template is stored as image bitmap \textit{Fig \ref{fig:template}}in the database when the user draws the symbol it is matched using classifiers.
-\cite {imagetrainable48,HierarchicalParsing7}  used four kind of classifiers to detect the matching template.
-He used two classifiers to measure similarity and two classifiers to measure dissimilarity then combine them to find best match of drawn stroke from the database.
+ Systems that employ such representation must first preprocess strokes to break them into geometric primitives (see section \ref{sec:preprocessing} for a review on preprocessing methods).
+Recognition can then be treated as sub graph matching problem.
 
-\begin{figure}
-	\centering
-		\includegraphics[scale=0.7]{../../neededfiles/Figures/template.eps}
-	\caption[Template matching]{Examples of image templates used to match symbols in \cite{imagetrainable48}. }
-	\label{fig:template}
-\end{figure}
-
-To refine the matching procedure and create more robust classification \cite {imagetrainable48}  first transform the strokes into polar coordinate then match the transformed stroke with the templates in the database.
-They claim that enhance the classification procedure as the polar coordinate system due to its definition make the stroke more robust to transformation and rotation.
-
-\cite{templatefrag21}  Introduce new way to match stroke temples.
-He store template as a sequence of lines and ellipse.
-When the user draws a stroke the system, segment it into a sequence of lines and ellipse.
-The system then matches the stored sequences with computed one to find correct classification.
-
-The main problem of template matching is that in most system the templates are hard coded inside the system and cannot be modified at runtime.
-
-\subsubsection{Graph matching}
-\label{sec:Graph matching}
+%Although not based on a formal structural model of shapes, another probabilistic approach by Szummer and Qi in [37] uses conditional random fields (CRFs) to propagate information about the labeling of one stroke fragment to its neighbors.
+This allows their system to assign labels to stroke fragments that cannot easily be identified in isolation and to regroup over-segmented fragments.
+This helps mitigate the difficulty of determining the exact granularity at which to perform fragmentation by using the context of surrounding fragments and a global interpretation of the sketch.
+They have applied their algorithm to the binary labeling problem of distinguishing boxes from connectors in organizational charts with good results, even in several highly ambiguous cases.
+\paragraph{Sub-graph matching:}
+\label{sec:GraphSearching}
 
 Representing the symbol that the user draws as a graph transforms the classification problem into a graph-matching problem.
-In general, each user stroke is represented as a node of the graph and the relations between those nodes are represented as arc.
+In general, each user geometrical primitive is represented as a node of the graph and the relations between those nodes are represented as arc.
+Figure \ref{fig:squarescematic} shows a graph that is constructed by \cite{mulitstroke5} for a rectangle.
 To classify the symbols the graph constructed must be matched with the graphs in the database.
-
-\cite{physicalmeaning6} uses spatial relationships for there graph matching.
-They first calculate spatial proximity relationships.
-Relations like intersections, containment, perpendicular, right-of, left-of and below are the most used relations concerning constructing graphs for symbols.
-
-Other systems  first analysis the spatial relationship between glyphs (sequence of strokes) using a library of spatial relations and information calculated in the preprocessing step.
-After that, they filter out important relation-using hierarchy of rules and type of geometrical constrain.
-All redundant constrains and relations are removed the final graphs is matched with the database.
-If more than one candidate is available the algorithm choose the best fit but if no matches found constrains are relaxed and strokes are passed again to classifier.
-
- Proposed a structure representation of the strokes in \cite{sketchunderstanding1,HierarchicalParsing7} .
-She added ratio of proximity of the relations between strokes.
-In addition, she used an error-driven, stochastic search algorithm for graph matching.
-
-\subsubsection{Statistical methods}
-\label{sec:Statisticalmethods}
-
-The application of Statistical classifiers in recognizing shapes are widely used in sketch recognition.
-Statistical classifier requires training time before the recognition actually happen.
-Then first step of classification is features computation that is done in the previous step.
-Finally, the recognition of symbols will be based on the classifier and the training set used.
-
-One of the first systems in sketch recognition is \citep{gestureexample12,aideddesgin22} algorithm to recognize ink gestures.
-He used Gaussian statistical classifier to classify any stroke user draw.
-In \cite {gestureexample12}  he firstly compute statistical information about the user stroke then pass it to a Gaussian classifier that had been trained using 10-20 samples.
-The output of the classifier is classification of the stroke from one of the known classes.
-
-Bayesian classifiers was used in \cite{Vibratory8} as a stroke classifier to recognize different shapes.
-The recognition is done using feature based statistical Gaussian distribution and Bayesian classifier techniques.
-The trainer use nine features extracted from segmented symbols (ex number of line segments, number of L intersections, ) all feature has a tolerance within.
-After all nine feature are computed a Gaussian statically model is constructed.
-
-Some systems employ more than one statistical classifier to get better results for the recognition.
-For example \cite{zernike61} uses Neural network, SVM (support vector machine )and MMD.
-\cite{polygonfeedback31}
-%the closed shaper recognizer which is done using classifiers (SVM , Rule based and ANN ) to make a 
-%classification of triangle, rectangle or ellipse
-
-\subsubsection{Other hybrid methods}
-\label{sec:Other hybrid methods}
-
-Due to the complexity of the sketch recognition problem, more than one system proposed a hybrid method to solve those problems. They combine more than one approach to recognize symbols the user draws.
-\cite{sketchunderstanding1,Simusketch51}  used four kind of classification image, graph matching, statistical and structure based classifiers.
-Each symbol will contain all graph, statistical, structure information and a (Wight, recognizer) tuples to determine which recognizer has the most influence in recognizing this symbol see \textit{Fig\ref{fig:structuralscemanticdeff}}.
-\begin{figure}
-	\centering
-		\includegraphics[scale=0.5]{../../neededfiles/Figures/structuralscemanticdeff.eps}
-	\caption[Structure definitions]{Example of the structure definition of sin symbol in \cite{sketchunderstanding1}.}
-	\label{fig:structuralscemanticdeff}
-\end{figure}
-
-\cite {HMM53}  firstly normalize the stroke and compute all the statistical features from it.
-After that, a noise reduction algorithm is used on the strokes which will be translated to produce size and rotation invariance data.
-Those data is passed to a multistage hybrid recognizer.
-The first step of his multistage classifier is dynamic programming procedure to compute the optimal degree of similarity and differentiation of symbols.
-Secondly, a statistical feature ranking is employed on the data produced.
-Finally, the system will reject all unlikely classification and produce best fit for this stroke using a dynamic programming using domain and context knowledge as input.
-
-\cite{statisticalparsing26} is a statically parser make use of the stored rules and probability of drawn stroked and uses Bayesian classifiers to do the classifications.
-
-\subsection{AI based Classification}
-\label{sec:AI based Classification}
-
-Most system mentioned before use image or statistical feature based on the points the user draw.
-The next few sections will present methods that will use AI approached to classify the symbols the user draw.
-Systems using AI approached imply a higher level of understanding of the strokes the user draws.
-Some construct a graph schema of the symbols from strokes the user draw.
-Other tried generating a visual grammar to identify the symbols that can be drawn.
-
-\subsubsection{Graph Searching }
-\label{sec:Graph Searching }
-
-\textit{Figure \ref{fig:squarescematic}} shows a semantic graph that is constructed by the trainable recognizer in  \cite{mulitstroke5} for a rectangle.
-The trainable recognizer construct the semantic network at train time with then match the symbol user draw with the networks in the database.
-To match the networks quickly the system assume the user draws the symbol with the same order as in training set.
-A deep heuristic graph search will be employed if there was no match found.
 
@@
 
-\cite{SRGraph57}  Uses two kind of graphs relations a spatial relationship graph (SRG) which is build on defining the spatial relationship between the source object and candidate object on the sketch.
-The relations are defined using Spatial Adjacency Relation (SAR) that falls into relation previously computed in system.
-To make the graph fully descriptive the system computes Relative Position Relation of SAR relations and adds these relations into the SRG graph.
+In the worst case, sub-graph matching is an exponential problem and thus is expensive to compute.
+Researcher tried to reduce complexity by restricting the search to fragments that are spatially or temporally grouped, or by using other assumption.
+Some of the common assumptions are that each stroke is only part of one symbol and that user must draw one shape before drawing the next\cite{physicalmeaning6}.
+These assumptions add constrains on how user can draw sketches and thus violates freely drawn sketches.
+Moreover, the matching process in highly affected by segmentation results.
+Over or under fragmentation of stroke into geometric primitive complicates the matching process.
+If a stroke is broken into too many fragments then there will be extra components that will not map to any part of the symbol.
+If the stroke is divided enough then there will be component of the shape that are not filled while matching.
+The different ways of segmenting the sketches worsen the already high cost graph matching.
 
-\subsubsection{Rule based }
-\label{sec:Rule based }
+ To solve this problem,  \cite{mulitstroke5} provided two methods to match graphs, a simpler one using restrictive assumptions and if it fails the system uses a longer more computationally expensive search.
+.
+Figure \ref{fig:squarescematic} shows a semantic graph that is constructed by the trainable recognizer in   \cite{mulitstroke5} for a rectangle.
+The trainable recognizer construct the semantic network at train time with then match the symbol user draw with the networks in the database.
+To match the networks quickly the system assumes the user draws the symbol with the same order as in training set.
+A deep heuristic graph search will employed only if no match was found.
+% there was no match found.
 
-The general algorithm of rule based system as used in \cite{interpretationmechanical50,SmartSketch56}  based on a knowledge of relation ship between nodes (strokes) a role is defined using those relation ships.
-The rule based system check the strokes the user draw against all strokes to detect if there a rule match the drawn symbol.
-When a rule matches the consequence of this rule, apply on the input symbol.
+ Several other systems tried to model the matching process as a probabilistic problem \cite{HMM53,SketchRead2007}.
+These systems allow matching of partially satisfied shapes to propagate down to re interpret low level fragmentation hypothesis.
+In SKETCREAD \cite{SketchRead2007}, a dynamically generated Bayesian network fragments represents the shape hypotheses.
+Higher level structure of shape fragments can cause re interpretation of the geometric primitives.
+For example, if one head of an arrow is initially labeled as an arc and the system fragments the shape as an arrow then the belief that this stroke is an arc is decrease and the belief that it is a line increase.
+Finally, the arrow structure is completed and it is fully recognized.
 
-The rule system used in \cite{sketchinginterfaces2}  uses the (UI user interface) domain knowledge to do its basic recognition.
-It is divided into two step first is testing if the rules applies on the input symbol.
-Second step, is ranking the rules that apply on the symbol based on the highest confidence that was given to this input symbol.
+ \paragraph{Visual Describtive Languages:}
+ \label{sec:VisualDescribtiveLanguages}
 
-Some system like in \cite{SmartSketch56}  used the rules to help in noise removal.
-Furthermore, rules based system are used to employ some reasoning to the recognized symbols with the help of context knowledge \cite {interpretationmechanical50}.
+ Avoiding graph matching problems, research tried to use languages that discribe shapes instead of graphs.
+Visual describtive languages represents the shape into tokens of geometric primitives or relations between smaller  geometric primitives.
+Figures \ref{fig:arrow1} and \ref{fig:arrow2} show a regular arrow with 2 shafts and head and its definition in Ladder language \cite{Ladder30}.
+Shape recognition is then a question of parsing the sketch document into a recognizer that parse strokes according to the constructed grammar.
 
-\subsubsection{HMM }
-\label{sec:HMM }
-
- The main idea of HMM is representing knowledge and recognition in serious of states each state has its input conditions.
-Only a node that has the input conditions of the destination state that can transit it from one state to another.
-The final state is the recognition state of a given class.
-For example to recognize a rectangle the first state should be first line detected, second will be second line and the condition is the recognition of either a perpendicular line or parallel line to first one to another.
-
-In sketch, recognition HMM was used to recognize symbols to different classes.
-Each class recognized was modeled using a single HMM model.
-The system in \cite {HMM53} encode the geometrical primitive that construct the symbol the user draws.
-After the encoding the codes are passed to HMM based classifier.
-Each object is represented by a HMM model and to manage multiple classes they combine matching score from different HMM models.
-
-\subsubsection{Graph language and visual language describing }
-\label{sec:Graph language and visual language describing }
-
-Researchers tried to generate a visual language that can describe hand drawn sketches.
-In such systems, sketch recognition will be a question of parsing the sketch document into recognizer that parse strokes according to the constructed grammar.
-
-Researchers proposed more than one visual grammar to use it in sketch recognition.
-\cite{statisticalparsing26} proposed a language parser that will parse sketched strokes using grammars rules that was constructed .
-The system first construct a parser that will parse drawn strokes into tokens then formulate those tokens using the grammar rules stored.
-Parsing is done in two steps; Firstly, the system should accept input of strokes then it use a statistical parser to generate the parse tree.
-The statically parser make use of the stored rules and probability of drawn stroked and uses Bayesian classifiers to produce the parse tree.
-
-%\citep*{XPGParser59} i need author   
-Introduced a grammar they called an extended positional grammar (XPG) that is based on XpLR parsing methodology \cite{XPGParser59} .
-The XPG imagine the symbol with a set of attribute as a sentence that can be parsed using the physical, syntactic and semantic information in sketch.
-They divided the grammar into two kind the first is named ink grammar, it define the symbol of the language as a composition of primitive geometric shapes.
-The second is language grammar that they use to identify the sketch as a composition of shapes defined by the ink grammar relation.
-The parsing itself is done using a incrementally bidirectional parsers.
-They also add a sketch editor to the system to help user to add new symbols and generate its XPG grammar for it \textit{(see Fig[\ref{fig:symboleditor}])}.
-
-\begin{figure}
-	\centering
-		\includegraphics[scale=0.7]{../../neededfiles/Figures/symboleditor.eps}
-	\caption[symbol editor]{The Figure shows the symbol editor that\cite{XPGParser59} uses to generate grammar rules.}
-	\label{fig:symboleditor}
-\end{figure}
-
-Relational grammar was also used to help in reducing the ambiguity with the help of fuzzy logic\cite {visualpattern43} .
-The fuzzy set is used to resolve the visual relations of the sketched stroke with others strokes.
-Spatial relations like (left of, right of) are computed and then passed to the fuzzy sets.
-The fuzzy set associate a degree of certainty to the shape defined.
-%EfficientAbstract39 also used the grammar language ladder
+ Several visual grammer were introduced \cite{statisticalparsing26,Ladder30,GenericHMM28}.
 \cite{Ladder30,GenericHMM28}  Used visual description language which will generate the information of the structure and constrain of each object classified.
@@
 The generated code cycles through the strokes drawn to check if any of the recognizers generated had interpreted any symbols partially or completely.
-HMM network is used to partially segment the strokes drawn to help the recognition process.
-
-Similar system was constructed in \cite {Ladder30} , where language grammar is written to descript complex shapes ( see \textit{Fig\ref{fig:arrowladderdef1}}) form primitive geometrical, constrains and triggers.
-The grammar is then passed to "jess" a java rule bases system that produce java code used in recognition of  drawn strokes.
-
+These systems are less computaitonal expensive than graph matching approach but they also operate based on the output of the segmentation process.
+Any error in the segmentation process lead to huge error while parsing the whole sketch.
+Moreover, symbols cannot share strokes between them which infleunce and restrict how the user draw symbols.
+ % complexity of the graph matching problem 
 \begin{figure}
 	\centering
-	\begin {center}
-		\subfigure{An Arrow}
-			\centering
-
-			{\label {fig:arrow1}\includegraphics[scale=0.75]{../../neededfiles/Figures/arrowladderdef1.eps}}
-		\subfigure[Arrow grammar definition]{\includegraphics[scale=0.88]{../../neededfiles/Figures/arrowladderdef2.eps}}
-		\end {center}
-	\caption[Arrow definition]{The figures show and regular arrow with 2 shafts and head and its definition in Ladder language \cite{Ladder30}}
+		\subfigure[An Arrow]
+			{\label{fig:arrow1}\includegraphics[scale=0.75]{../../neededfiles/Figures/arrowladderdef1.eps}}
+			\hfill
+		\subfigure[Arrow grammar definition]
+		 {	\label{fig:arrow2}\includegraphics[scale=0.88]{../../neededfiles/Figures/arrowladderdef2.eps}}
+	\caption[Arrow definition]
 	\label{fig:arrowladderdef1}
@@
 
-%\subsubsection{Fuzzy classifiers }
-%\label{sec:FuzzyClassifiers}
-%Fuzzy classifiers is used to help in identifying degree of certainty given to certain interpolation the
+\subsection{Global Shape Properties Based Recognition}
+\label{sec:GlobalFeaturesBasedRecognition}
 
-\subsection{Hybrid Systems}
-\label{sec:Hybrid Systems}
+drawing back from properties of individual strokes several researcher tried to use shape properties on the whole shape to classify shapes.
+They used global features that tries to define the entire shape properties.
 
-Given that the sketch recognition of hand drawn strokes is exceptionally challenging problem the researcher tried to take advantage of additional of combining approaches to achieve higher recognition rate.
+ Several systems \cite{DiagramOfflineConvexHull,MulitStrokeConvexHull} presented features like ratio of bounding box to convex hull area and the ratio of the preimeter to the area.
+These global feature can be used to distinguish between classes.
+For example, a rectangle can be distinguish from a triangle by looking to the raito of area of bounding box to the area of the convex hull.
+The triangle ratio will be much lower than that of the rectangle.
 
-Most of the problems of the sketch recognition are not still addressed.
-To do so, The sketch recognition system must maintain reliably interactive and extensible interface.
-The users minor change can cause the recognizers to fail.
-The user sloppy drawing with over or/and under shoot strokes  \textit{Fig\ref{fig:overstroked}}.
-Over processing of data and interaction with background are still major problems in the sketched documents.
+ Fonseca et al.\cite{Cali63} created a system using similar set of features, their features were used by a number of fuzzy logic rules to identify symbols.
+They also used Naïve Bayes model that was trained to classify shapes based on those features.
 
-To solve some of these problems,\cite{threeproblmes23}  revealed a two stages recognizer that will firstly construct a data graph in which strokes will be added.
-After constructing the full data graph the recognizer compute all possible solutions and reject any inconsistent alternatives.
-Second stage, Employs a topology structure based on the geometrical matching using a CSP making use of sub graphs.
+%Zernike moments were used by Hse et al. in \cite{}. 
+Hse et al.
+\cite{HeloiseBeautification,zernike61} proposed using the magnitudes of Zernike moments as a shape property.
+Zernike moemnts are orthognals moments which define the distribution of point in the input space.
+The higher the order of the moment the finer the detail level the point distribution represents.
+Hse et al.
+\cite{HeloiseBeautification,zernike61} used a SVM classifier to train the features vector calculated from the moments.
+Early evluation of Zernike moemnts showed that they are invariant to both reflection and rotation of the input shape \cite{ZerMomentOrthogonal}.
+In our system we used Zernike moemnts as part of our feature vector.
+A detailed comparison between Zernike moments and other features are demonstrated in Section \ref{sec:featuresComparisions}.
 
-As mentioned before,\cite{visualpattern43,Cali63} combine fuzzy logic sets with language grammar to create more robust classifiers.
-Grammar rules are used to describe the spatial relationships of the strokes.
-Fuzzy sets are used to handle the uncertainties of the relation and classification.
-Therefore, by combining Fuzzy set and spatial grammar rules the system was able quantify the spatial and statistical attribute calculated before in \ref{sec:Statisticalmethods} from the drawn symbols.
+ \begin{figure}[]
+	\centering
+		\subfigure{Ac Source}{
 
-Also the use of HMM with generated visual grammar was introduced in the system by \cite{HMM53} where they use HMM as a stroke segmentor and pass the output to the visual parser.
-furthermore, \cite{SmartSketch56} used more than one recognizers to recognize closed shapes rule based approach with hypothesis and test strategy, Neural network and SVM.
+		\includegraphics{images/acSource.eps}		\label{fig:acSource}}
 
-\paragraph{Dynamic programming}
-\label{sec:DynamicProgramming}
- was used by \cite {Mathpad46,templatefrag21}  to calculate the optimal degree of different and similarity of symbols.
-The system also uses a dynamic programming procedure to find a best fit from the given set of classifications.
+ 		\subfigure{Current Source}{
 
-\section{Second Step: Sketch Understanding}
-\label{sec:SecondStep:SketchUnderstanding}
-
-Sketch understanding is concerned with converting a set of recognized symbols into a whole document that is fully consistent with each other.
-Mainly the recognized symbols are a list of best candidates of that symbol for which the sketch understanding process must choose best fit based on the document structure.
-
-After recognizing the symbols drawn, Understanding the sketch document involve additional processing.
-In spite of various recognition systems that had reached a satisfactory symbol recognition.
-The sketch understanding problems are at a standstill.
-
-The recognized symbols must be verified with the rest of the document to fully understand the whole sketch.
-Consistency, ambiguities and uncertainty of the recognized symbols are the major issues that rise in any sketch understanding system.
-Most recent sketch understanding system uses more than just the context knowledge that is always available.
-Domain knowledge \cite {geometrydomain49}, Speech and verbal description of drawn sketch was used in\cite {speechMulti27}.Some verbal description of the drawn symbol helps in creating the casual model for each symbol.
-
-In the following sections, the ambiguities and uncertainty of sketch understanding systems will be discussed.
-Followed by how those systems represented knowledge that they use to formulate the understanding process.
-
-\subsection{Handling uncertainty and ambiguities }
-\label{sec:Handlinguncertaintyandambiguities }
-
-The main tasks of the sketch understanding are to handle higher-level ambiguities of the sketch drawn and to manage uncertainty of the recognized symbols.
-After the user draw a symbol using a single or multi-stroke symbol the recognizer labels each symbol with a known class either low level interpolation or high level.
-Low level interpolation means simple geometrical shapes and high level means more complex domain specific.
-Most current recognizers return a list of acceptable label for each symbol.
-The role of the sketch understating is converting such list of acceptable labels into one interpolation based on the surrounding symbols.
-They do so with preserving the sketch consistency of interpretation, which means that, the whole document has the same interpretation.
-
-The next section the various methods that was recently employ to completely understand sketched documents.
-
-\subsubsection{Rank and prune based on classifiers}
-\label{sec:Rankandprunebasedonclassifiers}
-
-The most used method in understanding the whole document and reducing the inconsistent solutions is the typical method of ranking the solutions and then pruning unwanted ones. 
-Initially the symbols nodes are initialized by all possible pairs of interpolation of the two symbols then the Consistency checker eliminate all inconsistence pairs.
-\cite {tobeadd} \cite {geometrydomain49}  firstly recognize all possible interpolation then prune and reason using knowledge to get a score for each interpolation.
-The final step is resolving and choosing the best result.
-
- \cite {Alvarado2002Framework11,HierarchicalParsing7}  First uses a bottom-top search to look for the interpolations of the strokes drawn.
-A top bottom search is used to prune out incomplete interpolations that was generated in the previous step.
-The top bottom firstly checks what symbol may be missing in the current document context after that it prune inconsistent interpolations.
-\subsubsection{Constrain satisfaction methods}
-\label{sec:Constrainsatisfactionmethods}
-
-Constrain satisfaction is the most used method for converting a group of recognized symbols into an understood document.
-Most recognizers produce a ranked set of possible interpolations.
-The problem is to constrain theses interpolations to get a total  understanding of the sketch document.
-Some like in \cite {tobeadd}  use constrain satisfaction methods to apply rules and constrains to match interpolations and rules.
-Constrains must be carefully written to check that they are not rough or severe.
-\cite {tobeadd}  checked the constrains applied on this system and if they produce no match the system relax the constrain and runs again with more loss constrains till it find the correct interpolations.
-
-\begin{figure}
-		\includegraphics[scale=0.7]{../../neededfiles/Figures/CSPphysical.eps}
-	\caption[Physical Constrains on Symbols]{ The figure shows two nodes and the constrain applied on both nodes to detect the inconsistency in the interpretations \cite{physicalmeaning6}. }
-	\label{fig:CSPphysical}
+		\includegraphics{images/CurrentSource.eps} \label{fig:CurrentSource}}
+	\caption{Example of current and ac sources}
+	\label{fig:ACandCurrentSources}
 \end{figure}
 
- \cite {physicalmeaning6}  firstly try connection between symbols firstly using bounding boxes of the symbol to locate the intersecting symbols.
-The intersected bounding boxes are checked again for ten different types of connection approximately for each terminal of the symbol.
-After identifying which symbols are connected the problem is identify which interpolation make those connected symbols has physical meaning.
-This is a "constrain satisfaction" problem the system used two processing steps to solve it.
-Firstly, the system checks the consistency of node where the node in graph is connection between two adjacent symbols.
-That means that each two connected symbol are checked and every inconsistent interpolation is rejected and deleted.
-This generate a local consistent sub graph but there may be a global inconsistent which take us to the second step where the algorithm search the arc consistency.
-The arcs of the consistency graph are any pair of nodes that have a symbol in common.
-Each consistency arc is checked for local and global consistency.
-This step is applied iteratively till all consistency are cancelled and only one global interpolation is reached \textit{(see Fig[ \ref{fig:CSPphysical}])}.
+ Systems based on Global shape properties distinguish shapes by learning the different properties each shape class tend to have.
+These properties dose not depend on the order or number of strokes and thus are only based on apperance of the shape not how the user draw it.
+However, individual details of the shape are impossible to distinguish using these properties.
+For example, it is impoissible to use  to  differentiate between a current source symbole (circle contiaing and arrow Figure \ref{fig:CurrentSource}) and a Ac Source (circle containing a tilate Figure \ref{fig:acSource}).
+Finear details of the symbols are not represented in this approach, thus shapes that have smalls details or only significant conceptual variation are not distiguished using this method.
+These symbols requires a system to represent different levels of shape details.
 
- The use of constrain satisfaction has also helped in minimizing the search \cite {tobeadd} .
-The strokes are incrementally processed using the Constrain system which check for constrain between image feature.
-Each new stroke either a new partial interpolation is created or an old one is extended and cloned provided that no constrained is violated.
+\subsection{Apperance Based Recognition}
+\label{sec:ApperanceBasedRecognition}
 
-\subsubsection{Bayesians networks}
-\label{sec:Bayesiansnetworks}
+In the third approach, the recognition foucus on appearance of shapes contrary to individual strokes or global shape properties.
+Kara \cite{imagetrainable48} used this approach while presenting a system that matches the input shape into a one of the  prototypes in the database.
+Figure \ref{fig:template} shows an example of prototypes of some of the shapes in the system.
+Input shapes are first pose and scale normalized then four image similarity measure are used to match input shapes to the database prototypes.
+He used two measures (Hausdor distance)  to measure similarity and two coeficients (Yule, Tanimoto) to measure dissimilarity.
+His approach allows the system to recongized a shape using only a single training sample as it is based on the shape prototype.
+However, the shape protoytpe must represent all  transformation and variation the input shape can have.
+Moreover, they restrict the image representation to 48x48 pixel due to computationa reasons.
+This restriction leads to elimintating some of the fine details that may be needed to distinguish some shape classes.
+There results on isolated symbols seems promising but the system have not beet tested on full sketch that tends to have high overlapping between symbols.
 
-Since the problem of deciding the best interpretation of a set of valid ones is a certainty problem researchers tried to solve it as a one.
-Bayesians networks were used to represent certainties in sketch documents\cite{SketchRead2007} .Bayesians classifiers were also used to help the understanding problem in various ways.\cite{statisticalparsing26}  Used Bayesians classifiers to help the statistical parser to recognize the best candidate based on the system rules.
+\subsection{Discussion}
+\label{sec:discuss}
 
-\cite{Alvarado2002Framework11,SketchRead2007} is a complete sketch-understating framework that used top bottom and bottom up search of interpolation of the strokes.
-The system represented uncertainty using Bayesians networks.
-Each compound object is represented by a network in which the object is the root \textit{see Fig[\ref {fig:anddefnetwork}]}.
- \begin{figure}
-		\includegraphics[scale=0.7]{../../neededfiles/Figures/anddefnetwork.eps}
-	\caption[Bayesian network]{Bayesian network of an And Gate as defined in \cite{Alvarado2002Framework11}}
-	\label{fig:anddefnetwork}
-\end{figure}
+In all of these system only \cite{SketchRead2007} using stroke based recognitiion have been implemented to the fully creat a system that free draw sketch and understand them.
+This is because the emphasis on the interactive side of the sketch sytem which result in large difficultes in processing free hand sketches.
+Methods based on gesture and global stroke properties produce systems that are dependent on assumptions about how strokes can be groubed.
+These systems assume that users will finish drawing a shape before moving to the next one \cite{6multistrokeEvaluation,Cali63,geometrydomain49}.
+This let them to temporally group strokes and then recognize them based on their strokes or apperance properties.
+\cite{AlvaradoDigital} argued that this assumption dose not hold in natural sketchs.
 
-\subsubsection{Other Methods}
-\label{sec:OtherMethods}
+Recogntion based on strokes using hierarchical shape description is applied in \cite{SketchRead2007}.
+The principle problem in these systems is the computatnal complexity of the matching between shape description and geometeric primitves produced in the segmentaiton process.
+Natural sketch are dificult to segment and group to create reliable fragment of geometeric primitives.
+This is partially due to the increase of noise and other phenomena such as overlapped, overtraced strokes and sharing of strokes between different symbols.
+Since these systems are depenable on the segmenation results, they either performe recognition using undependable segmentation results or consider different ways to segment the input sketch.
 
-Other Method were used in \cite{agent47,sketchinginterfaces2}to understand the sketch documents .
-For example \cite{sketchinginterfaces2} defined  visual and domain rules that applies on the sketch to finally understand the whole document.
-\cite{agent47} used software agents to resolve the conflict in curve matching algorithms he used.
-In addition to agents that resolve the conflict he also used agents to search for applicable features on the strokes.
-The agent will either report a feature on the strokes drawn or disappear till new strokes are drawn.
+To avoid most problems in such system, our system uses hybird of both global shape and hierarchical discribtion methods.
+Currenlty, our system recognize on sysmbols at a time.
+Each symbol is segmented and hierarchical features are computed.
+Features  that are represent the global discription of the shape are also generated and appended to the hierarchical featusres.
+The recognition is done using SVM Linear classifier as it proved to be one of the best learning algorithms.
 
-\subsection{Representing knowledge }
-\label{sec:Representingknowledge }
+%Natural sketches are dificult to reliably fragment into geometric primitives, because of the increase in noise and other phenomena such as interspersed drawing of parts from multiple shapes, multiple
+%shapes drawn with single strokes, and overtraced strokes
 
-Representing knowledge is the one of the challenges that faced the researcher in any AI program.
-In sketch recognition and understanding, the knowledge can be from more than one resource.
-The real problem is not to dump the program in too much data knowledge that is unstructured or unused in processing.
-To do so, means that the system is exhausted in undesired and unfulfilled processing.
-Processing time in sketch understanding system must be as close to real-time as possible to preserve the user interaction with the system.
-Therefore, the knowledge must be represented in special method that insures high speed and accuracy of understanding of the drawn sketch.
+ %The appearance based methods described above have not been applied, to our knowledge, to complete, freely drawn sketches.
+They have only been applied to isolated shapes or synthetic images generated from isolated shapes.
+Although their focus on the visual nature of the sketches appears promising for handling freely drawn sketches they have not yet been evaluated on them.
 
-Systems tried to make uses of more than one kind of knowledge to help in there recognitions and understandings.
-Spatial and visual knowledge of the drawn symbols are widely used in the sketch systems that relies on grammar languages \cite{GenericHMM28,statisticalparsing26,SketchRead2007} .
-Speech and multimodal systems helped the recognition of the symbols and reduced the search space.
-Domain and grammar knowledge were also used in \cite {geometrydomain49}  .
+ \chapter{Particle Swram Optimization}
+\label{sec:ParticleSwramOptimization}
 
-Black board architecture \cite {blackboardur} was widely used to represent the various source and forms of data in the sketch understanding systems.
-In \cite {Alvarado2002Framework11}  black board was used to represent incremental knowledge which increase by each symbol drawn in the sketch.
-The black board architecture is simply a modeling of group of people trying to solving a problem.
-When any one of them find new information he write it down on the black board soon all information will be in the black board for all solver to see and add.
+\section{Particle Swram Optimization Review}
+\label{sec:ParticleSwramOptimizationReview}
 
-\subsubsection{	Spatial and Visual knowledge}
-\label{sec:SpatialandVisualknowlege}
+%review on the pso alogrithm alone.................
 
-The visual and spatial information of the drawn symbols with respect to other symbols in the sketch helps in identifying the relation of these symbols with each other.
-Those information was used in \cite {Cali63,visualpattern43,incrmentintention41,geometrydomain49,marksacts40,sketchinginterfaces2,captureknowledge19}  to help the recognition and understanding of sketch drawn.
+\section{Discrete Particle Swarm Algorithm}
+\label{sec:ParticleSwarmAlgorithm}
+%\section{Particle Swarm Algorithm}
+%\label{PSO}
+%What is particle swarm algorithm and how it was used in related researches. 
+The main idea of \textit{Particle Swarm Algorithm (PSO)} is to represent each agent with a particle from the solution space \cite{PSOFirst}.
+Each agent moves the particle with a direction and velocity $v_{ij}$ based on equations \ref{eq:Swarm} \& \ref{eq:Swarm1}.
+\begin{equation}
+%\[
+p_{ij}=p_{ij}+v_{ij},
+%\
+\label{eq:Swarm1}
+\end{equation}
+where $p_{ij}$ represent the $jth$ particle in the $ith$ agent and $v_{ij}$ is the velocity of the $jth$ particle in the $ith$ agent.
+ %Equation [\ref{eq:Swarm}] shows how velocity and direction of each particle are computed
+ \begin{equation}
+v_{ij}  = v_{ij}  + c_1 r_1 (lbest_{ij}  - p_{ij} ) + c_2 r_2 (gbest_{ij}  - p_{ij} )
+\label{eq:Swarm}
+\end{equation}
+ where $lbest_{ij}$ is the local best particle, $gbest_{ij}$ is the global best particle, $r_1$ \& $r_2$ are random variables and $c_1$ \& $c_2$ are the swarm system variables.
+ After each iteration the global best $g_{best}$ particle and the agent local best $l_{best}$ particle are evaluated based on the maximum fitness functions of all particles in the solution space.
+The solution is found after achieving a specific number of iteration or after an error threshold is achieved.
+Equation \ref{eq:descrite} is used to change the general swarm algorithm into binary particle (\textit{Discrete Particle Swarm Algorithm DPSO}) which handles particle values of either $0$ or $1$ \cite{PSODisceret}.
+ \begin{equation}
+   P(i)\Leftarrow 
+\{
+\begin{array}{c} 
+1 \quad \quad if\quad r_{3}>p_{i}  \\
 
-\cite{SRGraph57}  Introduced a new spatial relationship graph to represent those spatial relation of one symbol to the other.
-The Spatial relation graph is build from the sketch drawn using primitive spatial adjacency relation and relative position relation.
-The produced graphs are then matched using a graph-matching algorithm with the database of graph to recognize best fit.
-
-The spatial and visual information are used in \cite{XPGParser59} to help in building a parser that will parse the sketch document into symbols and objects.
-
-\subsubsection{Domain and Context knowledge}
-\label{sec:DomainandContextknowledge}
-
-\begin{figure}
-	\centering
-		\includegraphics[scale=0.8]{../../neededfiles/Figures/domainexamples.eps}
-	\caption[domain examples]{Examples of sketched documents from different domains}
-	\label{fig:domainexamples}
-\end{figure}
-
-Users can draw sketches in different domains \textit{see Fig [\ref{fig:domainexamples}]}and Until now there have not been a sketch understanding system that can identify drawing in any sketched domain.
-To do so, the system must be multi-domain system and employ a kind of domain identification or classification.
-Current systems uses domain and context knowledge to assist in the understanding procedure\cite {geometrydomain49,RetargetableInteractive58,sketchunderstanding1, HierarchicalParsing7,Mathpad46} .
-
-ASSIST \cite{AlvaradoFreedom42} uses domain knowledge, context memory and user feedback to help in understanding procedure.
-Domain knowledge to get a set of possible interpretations then context knowledge is used to prune out inconsistent finally the user feedback is used to correct error interoperations.
-
-The \cite{incrmentintention41} store domain knowledge in a memory like architecture where It stores the domain knowledge in long-term memory were strokes are firstly checked for possible interpolations.
-Context knowledge and previous interpolations are stored in short-term memory to help more accurate recognition rate.
-The context and domain knowledge are also used to feed rule-based systems that uses visual language rules in making the final document recognitions \cite{interpretationmechanical50}.
-
-%&&&&&&&&&&&&&&&&oksearch 
-\subsubsection{Using speech and various other knowledge}
-\label{sec:Usingspeechandvariousotherknowledge}
-
-Multimodal systems are system that uses more than one source of input devices as input to the system.
-Theses systems help in providing sketch understanding with more information's and data that will help in the recognition procedure.
-Speech and verbal explanation of drawn symbols are most used data in multimodal system to aid sketch understanding \cite {cognitivesketch18,speechMulti27,computationalmodel16} .
-
-\subsection{Space Search}
-\label{sec:SpaceSearch}
-
-The solutions of any sketch understanding are not found until the problem of minimizing and find an efficient way to search the solution space is solved.
-The solution space increase exponentially with each stroke the user draws dependably on the system.
-Either ways the space search problem must be addressed to properly solve the problem of sketch understanding.
-
-\cite {EfficientAbstract39}  illustrate a new search space minimization where the recognizer can perform one of the checking for constrains between images feature.
-Check for partial interpretation by assigning an image feature to a model feature or the user draws delaying until more strokes.
-Every new stroke is drawn either a new partial interpolation is created an old interpolation is extended and cloned provided that no constrains or rule is violated.
+0 \quad \quad if\quad r_{3}<p_{i} 
+\label{eq:descrite}
+\end{array}\}
+\end{equation}
+ where $p_{ij}$ is the numerical values of the particle and $r_{3}$ is a random variable.

### 2009-04-12 draft/ChapterIntroduction.tex
12-4

@@
 Later, each segment is classified as line, arc, ellipse or circle.
-The main problem in this process is in detecting the critical points and vertices which divide strokes into lines and arcs.
+The main problem in this process is detecting the critical points and vertices which divide strokes into lines and arcs.
 The user draws the lines and arcs in an undefined manner which make it difficult for the program to decide where the user had finished a line and started a new arc or vice verse;  Figure \ref {fig:strokeinterpolation}.
@@
 The research also investigates various spatial, geometrical and statistical feature sets used in sketch recognition.
-A support Vector Machines (SVM) classifier \cite{svmintroduce} is used to classify the sketch into one of symbols from the set of previously trained symbols.
+A Support Vector Machines (SVM) classifier \cite{svmintroduce} is used to classify the sketch into one of symbols from the set of previously trained symbols.
 This research targets the recognition of one symbol at a time.

### 2009-04-12 draft/System.tex
12-4

@@
 \chapter{The Proposed System}
+\label{sec:proposedSystem}
 \section{Aims of the System}
 \label{sec:AimsOfTheSystem}
-The previous chapter shows the current trends in sketch recognition.
-This thesis introduces a solution for sketch recognition problem.
-The research gives an optimize solution for the sketch segmentation problem.
+%The previous chapter showed the current trends in sketch recognition. 
+This thesis introduces a new approach for the sketch recognition problem.
 Particle Swarm Optimization (PSO) is used to correctly segment strokes into curves and lines.
@@
 
-\section{An Overview of the System}
+\section{System Overview}
 \label{sec:AnOverviewOfTheSystem}
-   This research solves sketch segmentation problem using particle swarm algorithm PSO.
+   This research solves sketch segmentation problem using particle swarm algorithm (PSO).
 Due to sloppiness of users and hardware glitches the captured points must be processed to remove noise and redundancy before the system proceeds to compute the speed and curvature information.
@@
 	\includegraphics[scale=0.75]{images/AllBlockDiagram.eps}
-	\caption[The System Block Diagram]{The block diagram of the system}
+	\caption[The System Block Diagram]{Block diagram of the system}
 	\label{fig:Blockdiagram}
@@
 
-\subsection{System Components}
-\label{sec:SystemComponents}
+%\subsection{System Components}
+%\label{sec:SystemComponents}
 
-\textbf{Preprocessing} step is responsible of capturing the input data and removing the noise from it.
-The preprocessing block as in Figure \ref{fig:Blockdiagram} consist of computing the time differences, direction, speed and curvature of every point in the stroke.
-After computing this information the local minimum and local maximum of each curve is extracted from the data computed.
-This process estimate a set of a set of \textit{Possible Dominant Points $P_{pd}$} that guide the segmentation process.
+The \textbf{preprocessing} step is  responsible for capturing the input data and removing the noise from it.
+As shown in Figure \ref{fig:Blockdiagram} preprocessing consists of computing the time differences, direction, speed and curvature of every point in the stroke.
+After computing this information the local minimum and local maximum of each curve is extracted.
+This process estimates a set of a set of Possible Dominant Points $P_{pd}$ that guide the segmentation process.
 Later, other geometrical and statistical computations are evaluated from the stroke.
@@
 
-After each segment type is determined the the particle segmentation error is computed by Equation \ref{eq:errorSwarm2} 
+After each segment type is determined the particle segmentation error is computed by Equation \ref{eq:errorSwarm2} 
 \begin{equation}
@@
 If two points are nearer than $min_D$\footnote{Constant selected to be 10 pixels} then one of the points is removed.
+
 \subsection{Choosing the best fit}
@@
 
- \section{Benchmark algorithm}
+\section{Hybird Segmentation Algorithm}
+\label{sec:BenchMarckAlgorithm}
 
-\label{sec:BenchMarckAlgorithm}
-In this research, a bench mark algorithm is used as reference to the results obtained by the DPSO algorithms.
-The algorithm was first implemented by \citeauthor{earlyprocess} in \cite{earlyprocess}.
+In this research, we developed another algorithm as reference to the results obtained by the DPSO algorithms.
+The algorithm was first implemented by  Sezgin et al.
+in \cite{earlyprocess}.
 The algorithm is divided into the following steps:
@@
 The system uses a composite set of features include Rubine feature set \cite{gestureexample12}, Zernike moments \cite{HeloiseBeautification}, ink density \cite{GeometryAndDomain102} and some structural information like number of perpendiculars lines , number of parallel lines and types of primitives in each symbol.
-The table \ref{tab:FeatureTable} gives details on all the features used in the system.
-\begin{longtable}{|l|l|p{6cm}|p{5cm}|}
-	%\caption{Feature Table}
-	%\label{tab:FeatureTable}
-\hline 
-Type  & Index  & Features  & Notes \tabularnewline
-\endhead
-\hline 
-Type  & Index  & Features  & Notes \tabularnewline
-\endfirsthead
-\hline 
-Structural & \multicolumn{1}{c||}{} &  & \tabularnewline
-\hline 
- & F1.  & No. of parallel lines  & \tabularnewline
-\hline 
- & F2.  & No. of perpendicular lines  & \tabularnewline
-\hline 
- & F3.  & No. of intersection between of any two segmetns  & \tabularnewline
-\hline 
- & F4.  & No of primitives & \tabularnewline
-\hline 
- & F5.  & No. of line segments & \tabularnewline
-\hline 
- & F6  & No. of ellipse, circular arcs.  & \tabularnewline
-\hline 
-Statistical &  &  & \tabularnewline
-\hline 
- & F7.  &  & \tabularnewline
-\hline 
- & F8.  &  & \tabularnewline
-\hline 
-Rubine Features  &  & From features F9. till F13  & Introduced by in \cite{gestureexample12} \tabularnewline
-\hline 
- & F9.  & Cosine of starting angle  & \tabularnewline
-\hline 
- & F10.  & Sine of starting angle & \tabularnewline
-\hline 
- & F11.  & Length of diagonal of bounding box & (gives an idea of the size of the bounding box) \tabularnewline
-\hline 
- & F12.  & Angle of diagonal & gives an idea of the shape of the bounding box (long, tall, square) \tabularnewline
-\hline 
- & F13.  & Distance from start to end  & \tabularnewline
-\hline 
- & F14.  & Cosine of ending angle  & \tabularnewline
-\hline 
- & F15.  & Sine of ending angle  & \tabularnewline
-\hline 
- & F16.  & Total stroke length  & \tabularnewline
-\hline 
- & F17  & Change in Rotation( Arctan ) & gives the directional angle \tabularnewline
-\hline 
- & F18  & Absolute rotation  & \tabularnewline
-\hline 
- & F19  & Rotation squared  & \tabularnewline
-\hline 
- & F20  & The maximum speed reached (squared)  & \tabularnewline
-\hline 
- & F21  & Total time of stroke  & \tabularnewline
-%\newpage
-\hline 
-Zernike moments  &  &  & \tabularnewline
-\hline 
- & F22  & Zernike moments to the order  $n$  & used by \cite{zernike61} \tabularnewline
-Composite  &  &  & \tabularnewline
-\hline 
- &  &  & \tabularnewline
-\hline 
- &  &  & \tabularnewline
-\hline 
- &  &  & \tabularnewline
-\hline
-\end{longtable}
+The following list gives details on all the features used in the system.
+
+\begin{description}
+ \item [Structural and geometrical Features(FS1)] Features defines the structure of the geometrical symbol.  
+\begin{description}
+	 \item[No. Of segments] Number of segments in the symbol. 
+		\item[No. Of primitives] Number of primitives in the symbol. The feature helps when identifying             symbols with mixed geometric primitives like cylinders and callouts.  
+		\item [No. of curves] Number of curves or ellipses in the symbol. 
+		\item [No. of lines] Number of lines in the symbol. 
+		\item [No. of perpendicular lines] Number of perpendicular lines. 
+		\item [No. of parallel lines] Number of parallel lines. 
+		\item [No. of intersections] Number of intersection between lines and curves. 
+		\item [Size Ratio] Ratio between width to height of the symbol.
+		%\item[No. of holes] Number of holes in the symbols.
+
+\end{description}
+\item [Rubine Feature Set (FS2)] Features introduced by Rubine\cite{gestureexample12} for single stroke gestures.
+To compute the features we append all segments points into single path $ink_{path}$.
+The features is computed on the generated path $ink_{path}$.%then compute the features based on this path.
+\begin{enumerate}
+	\item Cosine of starting angle.
+	\item Sine of starting angle.
+	\item Length of diagonal of bounding box.It gives an idea of the size of the bounding box).
+	\item Angle of diagonal. It gives an idea of the shape of the bounding box (long, tall, square).
+	\item Distance from start to end.  
+	\item Total stroke length
+	\item Change in Rotation( Arctan ). It  gives the directional angle.
+	\item Absolute rotation 
+	\item Rotation squared 
+	\item The maximum speed reached (squared) 
+	\item  Total time of stroke 
+\end{enumerate}
+
+\item[Statistical Features (FS3)]   
+\begin{description}
+	\item[Zernike moments] Zernike moments of order n\footnote{ We tried different values of n. Results show the analysis of each order.} \cite{HeloiseBeautification}. 
+\end{description}
+	\item[Composite Features (FS4)] Features that are composites of different geometrical and statistical symbol. 
+	\begin{description}
+
+	\item [Ink density] Compute the density of points inside its bounding box\cite{GeometryAndDomain102}.   
+ 	\item[Convex Hull Area] Area of convex hull with respect to area of bounding box of symbol.
+	\item[Convex Hull Perimeter] Perimeter of convex hull with respect to total length of symbol.
+		\item [Mean Centroidal radius] The Mean of the centroidal radius which is the distance from each point in the symbol to the center of gravity.
+
+	\item [Mean Time difference] The mean of the time difference between each two successive points in the symbol. %Different strokes are appended to construct a single path.
+	%\item [Ra]
+  \end{description}
+
+\end{description}
+%
+%\begin{longtable}{|l|l|p{6cm}|p{5cm}|}
+%	%\caption{Feature Table}
+%	%\label{tab:FeatureTable}
+%\hline 
+%Type  & Index  & Features  & Notes \tabularnewline
+%\endhead
+%\hline 
+%Type  & Index  & Features  & Notes \tabularnewline
+%\endfirsthead
+%\hline 
+%Structural & \multicolumn{1}{c||}{} &  & \tabularnewline
+%\hline 
+% & F1.  & No. of parallel lines  & \tabularnewline
+%\hline 
+% & F2.  & No. of perpendicular lines  & \tabularnewline
+%\hline 
+% & F3.  & No. of intersection between of any two segmetns  & \tabularnewline
+%\hline 
+% & F4.  & No of primitives & \tabularnewline
+%\hline 
+% & F5.  & No. of line segments & \tabularnewline
+%\hline 
+% & F6  & No. of ellipse, circular arcs.  & \tabularnewline
+%\hline 
+%Statistical &  &  & \tabularnewline
+%\hline 
+% & F7.  &  & \tabularnewline
+%\hline 
+% & F8.  &  & \tabularnewline
+%\hline 
+%Rubine Features  &  & From features F9. till F13  & Introduced by in \cite{gestureexample12} \tabularnewline
+%\hline 
+% & F9.  & Cosine of starting angle  & \tabularnewline
+%\hline 
+% & F10.  & Sine of starting angle & \tabularnewline
+%\hline 
+% & F11.  & Length of diagonal of bounding box & (gives an idea of the size of the bounding box) \tabularnewline
+%\hline 
+% & F12.  & Angle of diagonal & gives an idea of the shape of the bounding box (long, tall, square) \tabularnewline
+%\hline 
+% & F13.  & Distance from start to end  & \tabularnewline
+%\hline 
+% & F14.  & Cosine of ending angle  & \tabularnewline
+%\hline 
+% & F15.  & Sine of ending angle  & \tabularnewline
+%\hline 
+% & F16.  & Total stroke length  & \tabularnewline
+%\hline 
+% & F17  & Change in Rotation( Arctan ) & gives the directional angle \tabularnewline
+%\hline 
+% & F18  & Absolute rotation  & \tabularnewline
+%\hline 
+% & F19  & Rotation squared  & \tabularnewline
+%\hline 
+% & F20  & The maximum speed reached (squared)  & \tabularnewline
+%\hline 
+% & F21  & Total time of stroke  & \tabularnewline
+%%\newpage
+%\hline 
+%Zernike moments  &  &  & \tabularnewline
+%\hline 
+% & F22  & Zernike moments to the order  $n$  & used by \cite{zernike61} \tabularnewline
+%Composite  &  &  & \tabularnewline
+%\hline 
+% &  &  & \tabularnewline
+%\hline 
+% &  &  & \tabularnewline
+%\hline 
+% &  &  & \tabularnewline
+%\hline
+%\end{longtable}
 %\end{table}

### 2009-04-12 draft/chapterSurvey.tex
12-4

@@
 The history of pen and sketch based computing goes back to 1963 with Ivan Sutherland's SKETCHPAD \cite{sutherlandsketch}, which used a light pen to draw on the monitor to create circuit diagrams.
-Rubine \cite{gestureexample12} Gesture recognition system.
+Current technology and the increae in computuing power have increased the abilities of such interfaces and allowed the implementaion of fully sketching systems.
+Over the years, many other sketching based system have been developed, elthough throught this history most system foucused on building the interactive interfaces to assist users while entering information to the computer.
+To demonstrate our system inspiration we will first present current research in the area of sketch recognition.
+Fristly, preprosessing and segmentaiton methods will be discussed in details.
+Later, symbol recognition methods will be presented and analysied in Section \ref{sec:symbolrecogntion}.
+
 % i need to write a history then small introudction to the survey...
@@
 Simple features as length of stroke, smoothness of stroke, the angle that the first part of the gesture is drawn at and properties of the bounding box of the stroke are used to train the classifier to recognize  symbols.
-In our system, we combine Rubine features with other to include the information on how the user draws sketch while recognizing the symbols.
+In our system, we combine Rubine features with others to include the information on how the user draws sketch while recognizing the symbols.
 In \cite{Long00}  Long shows how similarities between gestures can be analyzed.
@@
  Systems that employ such representation must first preprocess strokes to break them into geometric primitives (see section \ref{sec:preprocessing} for a review on preprocessing methods).
-Recognition can then be treated as sub graph matching problem or constrain satisfaction problem.
+Recognition can then be treated as sub graph matching problem.
 
@@
 Over or under fragmentation of stroke into geometric primitive complicates the matching process.
-If a stroke is broken into too many fragments then the there will be extra component that will not map to any part of the symbol.
+If a stroke is broken into too many fragments then there will be extra components that will not map to any part of the symbol.
 If the stroke is divided enough then there will be component of the shape that are not filled while matching.
@@
 Finally, the arrow structure is completed and it is fully recognized.
+
+ \paragraph{Visual Describtive Languages:}
+ \label{sec:VisualDescribtiveLanguages}
+
+ Avoiding graph matching problems, research tried to use languages that discribe shapes instead of graphs.
+Visual describtive languages represents the shape into tokens of geometric primitives or relations between smaller  geometric primitives.
+Figures \ref{fig:arrow1} and \ref{fig:arrow2} show a regular arrow with 2 shafts and head and its definition in Ladder language \cite{Ladder30}.
+Shape recognition is then a question of parsing the sketch document into a recognizer that parse strokes according to the constructed grammar.
+
+ Several visual grammer were introduced \cite{statisticalparsing26,Ladder30,GenericHMM28}.
+\cite{Ladder30,GenericHMM28}  Used visual description language which will generate the information of the structure and constrain of each object classified.
+The system  automatically generates java code for the recognizer for each object.
+The generated code cycles through the strokes drawn to check if any of the recognizers generated had interpreted any symbols partially or completely.
+These systems are less computaitonal expensive than graph matching approach but they also operate based on the output of the segmentation process.
+Any error in the segmentation process lead to huge error while parsing the whole sketch.
+Moreover, symbols cannot share strokes between them which infleunce and restrict how the user draw symbols.
+ % complexity of the graph matching problem 
+\begin{figure}
+	\centering
+		\subfigure[An Arrow]
+			{\label{fig:arrow1}\includegraphics[scale=0.75]{../../neededfiles/Figures/arrowladderdef1.eps}}
+			\hfill
+		\subfigure[Arrow grammar definition]
+		 {	\label{fig:arrow2}\includegraphics[scale=0.88]{../../neededfiles/Figures/arrowladderdef2.eps}}
+	\caption[Arrow definition]
+	\label{fig:arrowladderdef1}
+\end{figure}
 
@@
 \cite{HeloiseBeautification,zernike61} used a SVM classifier to train the features vector calculated from the moments.
-Early evluation of Zernike moemnts showed that they are invariant to both reflection and rotation of the input shape \cite{ }.
+Early evluation of Zernike moemnts showed that they are invariant to both reflection and rotation of the input shape \cite{ZerMomentOrthogonal}.
 In our system we used Zernike moemnts as part of our feature vector.
-A detailed comparison between Zernike moments and other features are demonstrated in Section \ref{}.
+A detailed comparison between Zernike moments and other features are demonstrated in Section \ref{sec:featuresComparisions}.
+
+ \begin{figure}[]
+	\centering
+		\subfigure{Ac Source}{
+
+		\includegraphics{images/acSource.eps}		\label{fig:acSource}}
+
+ 		\subfigure{Current Source}{
+
+		\includegraphics{images/CurrentSource.eps} \label{fig:CurrentSource}}
+	\caption{Example of current and ac sources}
+	\label{fig:ACandCurrentSources}
+\end{figure}
 
@@
 However, individual details of the shape are impossible to distinguish using these properties.
-For example, it is impoissible to use  to  differentiate between a current source symbole (circle contiaing and arrow Figure \ref{}) and a clock (circle containing a square signal Figure \ref{}).
+For example, it is impoissible to use  to  differentiate between a current source symbole (circle contiaing and arrow Figure \ref{fig:CurrentSource}) and a Ac Source (circle containing a tilate Figure \ref{fig:acSource}).
 Finear details of the symbols are not represented in this approach, thus shapes that have smalls details or only significant conceptual variation are not distiguished using this method.
@@
 Input shapes are first pose and scale normalized then four image similarity measure are used to match input shapes to the database prototypes.
-He used two measures (Hausdor distance )  to measure similarity and two coeficients (Yule, Tanimoto) to measure dissimilarity.
+He used two measures (Hausdor distance)  to measure similarity and two coeficients (Yule, Tanimoto) to measure dissimilarity.
 His approach allows the system to recongized a shape using only a single training sample as it is based on the shape prototype.
@@
 
-In all of these system only \cite{4} using stroke based recognitiion have been implemented to the fully creat a system that free draw sketch and understand them.
+In all of these system only \cite{SketchRead2007} using stroke based recognitiion have been implemented to the fully creat a system that free draw sketch and understand them.
 This is because the emphasis on the interactive side of the sketch sytem which result in large difficultes in processing free hand sketches.
 Methods based on gesture and global stroke properties produce systems that are dependent on assumptions about how strokes can be groubed.
-These systems assume that users will finish drawing a shape before moving to the next one \cite{6,11,14,15}.
+These systems assume that users will finish drawing a shape before moving to the next one \cite{6multistrokeEvaluation,Cali63,geometrydomain49}.
 This let them to temporally group strokes and then recognize them based on their strokes or apperance properties.
-\cite{5} argued that this assumption dose not hold in natural sketchs.
+\cite{AlvaradoDigital} argued that this assumption dose not hold in natural sketchs.
 
-Recogntion based on strokes using hierarchical shape description is applied in \cite{}.
+Recogntion based on strokes using hierarchical shape description is applied in \cite{SketchRead2007}.
 The principle problem in these systems is the computatnal complexity of the matching between shape description and geometeric primitves produced in the segmentaiton process.
 Natural sketch are dificult to segment and group to create reliable fragment of geometeric primitives.
-This is partially due to the increase of noise and other phenomena such as overlapped, overtraced strokes and multiple shapes drawn using
+This is partially due to the increase of noise and other phenomena such as overlapped, overtraced strokes and sharing of strokes between different symbols.
+Since these systems are depenable on the segmenation results, they either performe recognition using undependable segmentation results or consider different ways to segment the input sketch.
 
-Natural sketches are dificult to reliably fragment into geometric primitives, because of the increase in noise and other phenomena such as interspersed drawing of parts from multiple shapes, multiple
-shapes drawn with single strokes, and overtraced strokes. Because these systems are
-dependent on the fragmentation process, they are either faced with performing recognition
-with unreliable fragmentation results or must consider many possible ways to
-fragment the sketch. Avoiding the fragmentation issue was one of our key motivating
-factors in taking an appearance-based approach to recognition.
-The appearance based methods described above have not been applied, to our
-knowledge, to complete, freely drawn sketches. They have only been applied to
-isolated shapes or synthetic images generated from isolated shapes. Although their
-focus on the visual nature of the sketches appears promising for handling freely drawn
-sketches they have not yet been evaluated on them.
+To avoid most problems in such system, our system uses hybird of both global shape and hierarchical discribtion methods.
+Currenlty, our system recognize on sysmbols at a time.
+Each symbol is segmented and hierarchical features are computed.
+Features  that are represent the global discription of the shape are also generated and appended to the hierarchical featusres.
+The recognition is done using SVM Linear classifier as it proved to be one of the best learning algorithms.
+
+%Natural sketches are dificult to reliably fragment into geometric primitives, because of the increase in noise and other phenomena such as interspersed drawing of parts from multiple shapes, multiple
+%shapes drawn with single strokes, and overtraced strokes
+
+ %The appearance based methods described above have not been applied, to our knowledge, to complete, freely drawn sketches.
+They have only been applied to isolated shapes or synthetic images generated from isolated shapes.
+Although their focus on the visual nature of the sketches appears promising for handling freely drawn sketches they have not yet been evaluated on them.
 
@@
 %review on the pso alogrithm alone.................
+%A literature overview of the simulation done on swarm behavior and there application in swarm robots.
+The survey  describe system that made biological modeling of birds, ants behavior.
+Birds was modeled in "BIOD" and Flack systems which only modeled the distributed behaviors of flock of birds.
+Kube and zhang simulated the behavior of ants in his system.
+Tersopooulos et al.
+developed artificial fish in a 3d physical world that emulate real fish behavior.
+\\
+
+The survey then mention swarm robots and the literature over it.
+Beoi proposed a swarm cellar of autonomously  robots.
+others  introduced a behavior based swarms  Beni also proposed a sign post system where robots are sample but act under the influence of "sign post robot" which change state of the robot by passing by a "sign post robot".
+other studied the affect of communication between robot and it was found that  as inter robot comunication improve more complex behavior can be achieved.
+\subsection*{A Hybrid network for feature extraction}
+The paper uses the SIDS to generate correct$x_{i}$ and $x_{j}$t mapping of features in images.
+its is divided into two steps.
+First FCS feature classifier system based on simple one layer neural networks.
+Each cell compute a micro function of object to be classified.
+The system use an n tuple feature classification  .
+The second is a feature search system that  use to search the n tuple mapping if feature is present in this mapping.
+
+%\section{Paper:89  [2005]
+}
+\subsection*{Dynamic probabilistic particle swarms}
+PSO is a swarm optimization where the particle are moved around the search space using direction and velocity equation.
+Each particle have location equation x and V velocity equation to compute next location in the solution space.
+The basic algorithm use constants or static tuning parameters with the help of particle best and global best to influence the equations of particle transformation.
+The paper present a dynamic probabilistic based on more probability and  information than the original algorithm.\\
+
+There are two proposals first is trimmed uniform probabilities where new positions of particle is computed on biases of expected position of old best location of the particle and global best.
+The second is Gaussian  probabilities where the influence of moving particles are give based on the probability of how near the particle to the solution X.
+the propability is computed using gaussian distribution and subsituted to influence in the transformation equation of the particels.
+
+%\section{Paper:90  [2000]}
+\subsection*{Artificial Ant colonies in digital image habitats a mass behavior effect study on pattern recognition}
+The paper describe how ant can be used in image processing environments.
+The system will need to have describe how ant move around the image using the transition rules from one location to another.
+The transition rule $w(\sigma)=\left( 1+\frac{\sigma}{1+\sigma\delta}\right)^{\beta} $ Function depend on the pheromone density on the environment of the pixel and neighborhood pixels.
+The system also include a definition of the rate $\eta$ that the ant emit pheromonal and the rate \textit{k }the rate pheromal evaporate.
+%\section{Paper:91  [2004]}
+\subsection*{An algorithm for swarm based color image segmentation}
+The Ant system was used to segment color images using a system that is described in this paper.
+The algorithm main goal was to segment pixel of similar color into segments.
+The paper describe in details the data Structure and the swarm structure of the system.\\
+
+The Data is located in 2D matrices which represent the environment [pixel info] and the segmented map [what ants write data in] each with the same size as the original image.
+There is also a Color map array that is used as a label for the segment map.\\
+
+The swarm system is divided into to main agents Scots and worker.
+The Scots are the agent that have a sense of color based on an equation and error bound.
+The Scot loop over color label form the color map move around the image randomly till he found its found a similar color.
+When it found a similar color it write to the segment map the  label in the pixel location then check the neighbors to see if they have a similar color.
+If not then the Scot is moved around into different color and location.
+The worker work on the segment map where intailly  it has a fixed color liable and its moves randomly on location till it found a segmented location with same label as itself.
+The worker check the neighborhood location if the Scots has label it or not.
+if it is labeled with different color then the larger of the two labeled will still be present else  moves randomly till found an interesting neighborhood.
 

### 2009-04-12 draft/chapterSurvey.tex
1 am

@@
 
-%Natural sketches are dificult to reliably fragment into geometric primitives, because of the increase in noise and other phenomena such as interspersed drawing of parts from multiple shapes, multiple
-%shapes drawn with single strokes, and overtraced strokes
-
- %The appearance based methods described above have not been applied, to our knowledge, to complete, freely drawn sketches.
-They have only been applied to isolated shapes or synthetic images generated from isolated shapes.
-Although their focus on the visual nature of the sketches appears promising for handling freely drawn sketches they have not yet been evaluated on them.
-
 \chapter{Particle Swram Optimization}
@@
 
-\section{Particle Swram Optimization Review}
+\section{Swram Optimization Review}
 \label{sec:ParticleSwramOptimizationReview}
+
+%This seems to emerge of what is called 'collective intelligence' from often large simple agents.  
+% start with swarm in general  .. say about flock , ant and usin in image processing 
+% then speeak about PSO and  its firstly generated   (see article)
+%mention that is best than ga (see thesis)
+
+Modeling how a biological swarm has been recentlsy studied by biologist and computer scientists.
+Scientist in the field or artifical life need to understand how such "social anaimals" interact, achieve goals and evolove.
+Moreover, swarm behaviours that achieves some kind of swram intelligance is increasly intersting for engineers as it they can be applied in optimization.
+Looking from a higher view, agents in a swarm are cooperating to achieve some proposeful behaviour and achieve some common goal.
+For example, ant groups looking for food source.
+Large gourps of simple independent agents seems to generate this behaviour in what is called 'collective intelligence'.
+Each agent has a set of simple rules to follow, and accourding to the interaction of the entrie group the objective is met.
+%with other neighbourhood agents.
+
+Reynolds \cite{swarmflock} created 'boid' which simulated the motion of a flock of birds.
+Each boid is a simple agent that navigate according to his perception and a set of basic rules.
+Kube and Zhange \cite{antsimulation} introduced a simulation model for ants behvaiour and used in cooperative transportation.
+Ants leaves a trail of its movement, so when an ant find food it cannot cary alone, it can easily trace back to the colony and then a group of ants can track the trail and bring the food back.
+Kube and Zhange \cite{antsimulation} system seems to corectly models ants behavour with a methods of task modeling.
 
@@
 
-The survey then mention swarm robots and the literature over it.
+%The survey then mention swarm robots and the literature over it.
 Beoi proposed a swarm cellar of autonomously  robots.
@@
 \subsection*{A Hybrid network for feature extraction}
-The paper uses the SIDS to generate correct$x_{i}$ and $x_{j}$t mapping of features in images.
+%The paper uses the SIDS to generate correct$x_{i}$ and $x_{j}$t mapping of features in images.
 its is divided into two steps.
@@
 }
-\subsection*{Dynamic probabilistic particle swarms}
-PSO is a swarm optimization where the particle are moved around the search space using direction and velocity equation.
+%\subsection*{Dynamic probabilistic particle swarms}
+%PSO is a swarm optimization where the particle are moved around the search space using direction and velocity equation.
 Each particle have location equation x and V velocity equation to compute next location in the solution space.
@@
 The paper present a dynamic probabilistic based on more probability and  information than the original algorithm.\\
-
-There are two proposals first is trimmed uniform probabilities where new positions of particle is computed on biases of expected position of old best location of the particle and global best.
+%
+%
+%There are two proposals first is trimmed uniform probabilities where new positions of particle is computed on biases of expected position of old best location of the particle and global best.
 The second is Gaussian  probabilities where the influence of moving particles are give based on the probability of how near the particle to the solution X.
 the propability is computed using gaussian distribution and subsituted to influence in the transformation equation of the particels.
-
-%\section{Paper:90  [2000]}
-\subsection*{Artificial Ant colonies in digital image habitats a mass behavior effect study on pattern recognition}
-The paper describe how ant can be used in image processing environments.
+%
+%%\section{Paper:90  [2000]}
+%\subsection*{Artificial Ant colonies in digital image habitats a mass behavior effect study on pattern recognition}
+%The paper describe how ant can be used in image processing environments.
 The system will need to have describe how ant move around the image using the transition rules from one location to another.
@@
 The system also include a definition of the rate $\eta$ that the ant emit pheromonal and the rate \textit{k }the rate pheromal evaporate.
-%\section{Paper:91  [2004]}
-\subsection*{An algorithm for swarm based color image segmentation}
-The Ant system was used to segment color images using a system that is described in this paper.
+%%\section{Paper:91  [2004]}
+%\subsection*{An algorithm for swarm based color image segmentation}
+%The Ant system was used to segment color images using a system that is described in this paper.
 The algorithm main goal was to segment pixel of similar color into segments.
 The paper describe in details the data Structure and the swarm structure of the system.\\
-
-The Data is located in 2D matrices which represent the environment [pixel info] and the segmented map [what ants write data in] each with the same size as the original image.
+%
+%The Data is located in 2D matrices which represent the environment [pixel info] and the segmented map [what ants write data in] each with the same size as the original image.
 There is also a Color map array that is used as a label for the segment map.\\
-
-The swarm system is divided into to main agents Scots and worker.
+%
+%The swarm system is divided into to main agents Scots and worker.
 The Scots are the agent that have a sense of color based on an equation and error bound.
@@
 if it is labeled with different color then the larger of the two labeled will still be present else  moves randomly till found an interesting neighborhood.
+%
 

### 2009-04-12 trunk/draft/chapterSurvey.tex

@@
 
-\section{Particle Swram Optimization Review}
+\section{Swram Optimization Review}
 \label{sec:ParticleSwramOptimizationReview}
 
+%This seems to emerge of what is called 'collective intelligence' from often large simple agents.  
+% start with swarm in general  .. say about flock , ant and usin in image processing 
+% then speeak about PSO and  its firstly generated   (see article)
+%mention that is best than ga (see thesis)
+
+Modeling how a biological swarm has been recentlsy studied by biologist and computer scientists.
+Scientist in the field or artifical life need to understand how such "social anaimals" interact, achieve goals and evolove.
+Moreover, swarm behaviours that achieves some kind of swram intelligance is increasly intersting for engineers as it they can be applied in optimization.
+Looking from a higher view, agents in a swarm are cooperating to achieve some proposeful behaviour and achieve some common goal.
+For example, ant groups looking for food source.
+Large gourps of simple independent agents seems to generate this behaviour in what is called 'collective intelligence'.
+Each agent has a set of simple rules to follow, and accourding to the interaction of the entrie group the objective is met\cite{swarmGeneral}.
+%with other neighbourhood agents.
+
+Reynolds \cite{swarmflock} created 'boid' which simulated the motion of a flock of birds.
+Each boid is a simple agent that navigate according to his perception and a set of basic rules.
+Kube and Zhange \cite{antsimulation} introduced a simulation model for ants behvaiour and used in cooperative transportation.
+Ants leaves a trail of its movement, so when an ant find food it cannot cary alone, it can easily trace back to the colony and then a group of ants can track the trail and bring the food back.
+Kube and Zhange \cite{antsimulation,swarmGeneral} system seems to corectly models ants behavour with a methods of task modeling.
+
+Image processing is one of the displines that used Swarm Intelligance.
+Ant colony was widely used to segment objects in \cite{}.
+Since 1995 when Kennedy and Eberhart \cite{PSOFirst,PSO2} introduced particle swarm alogrithm (PSO).
+The algorithm had been used in various image processing sytem as an optimization algorithm\cite{SwarmImageProcessing,Swarmimageprocessing2}.
+
 %review on the pso alogrithm alone.................
+
+%A literature overview of the simulation done on swarm behavior and there application in swarm robots.
+The survey  describe system that made biological modeling of birds, ants behavior.
+Birds was modeled in "BIOD" and Flack systems which only modeled the distributed behaviors of flock of birds.
+Kube and zhang simulated the behavior of ants in his system.
+Tersopooulos et al.
+developed artificial fish in a 3d physical world that emulate real fish behavior.
+\\
+
+%The survey then mention swarm robots and the literature over it.
+Beoi proposed a swarm cellar of autonomously  robots.
+others  introduced a behavior based swarms  Beni also proposed a sign post system where robots are sample but act under the influence of "sign post robot" which change state of the robot by passing by a "sign post robot".
+other studied the affect of communication between robot and it was found that  as inter robot comunication improve more complex behavior can be achieved.
+\subsection*{A Hybrid network for feature extraction}
+%The paper uses the SIDS to generate correct$x_{i}$ and $x_{j}$t mapping of features in images.
+its is divided into two steps.
+First FCS feature classifier system based on simple one layer neural networks.
+Each cell compute a micro function of object to be classified.
+The system use an n tuple feature classification  .
+The second is a feature search system that  use to search the n tuple mapping if feature is present in this mapping.
+
+%\section{Paper:89  [2005]
+}
+%\subsection*{Dynamic probabilistic particle swarms}
+%PSO is a swarm optimization where the particle are moved around the search space using direction and velocity equation.
+Each particle have location equation x and V velocity equation to compute next location in the solution space.
+The basic algorithm use constants or static tuning parameters with the help of particle best and global best to influence the equations of particle transformation.
+The paper present a dynamic probabilistic based on more probability and  information than the original algorithm.\\
+%
+%
+%There are two proposals first is trimmed uniform probabilities where new positions of particle is computed on biases of expected position of old best location of the particle and global best.
+The second is Gaussian  probabilities where the influence of moving particles are give based on the probability of how near the particle to the solution X.
+the propability is computed using gaussian distribution and subsituted to influence in the transformation equation of the particels.
+%
+%%\section{Paper:90  [2000]}
+%\subsection*{Artificial Ant colonies in digital image habitats a mass behavior effect study on pattern recognition}
+%The paper describe how ant can be used in image processing environments.
+The system will need to have describe how ant move around the image using the transition rules from one location to another.
+The transition rule $w(\sigma)=\left( 1+\frac{\sigma}{1+\sigma\delta}\right)^{\beta} $ Function depend on the pheromone density on the environment of the pixel and neighborhood pixels.
+The system also include a definition of the rate $\eta$ that the ant emit pheromonal and the rate \textit{k }the rate pheromal evaporate.
+%%\section{Paper:91  [2004]}
+%\subsection*{An algorithm for swarm based color image segmentation}
+%The Ant system was used to segment color images using a system that is described in this paper.
+The algorithm main goal was to segment pixel of similar color into segments.
+The paper describe in details the data Structure and the swarm structure of the system.\\
+%
+%The Data is located in 2D matrices which represent the environment [pixel info] and the segmented map [what ants write data in] each with the same size as the original image.
+There is also a Color map array that is used as a label for the segment map.\\
+%
+%The swarm system is divided into to main agents Scots and worker.
+The Scots are the agent that have a sense of color based on an equation and error bound.
+The Scot loop over color label form the color map move around the image randomly till he found its found a similar color.
+When it found a similar color it write to the segment map the  label in the pixel location then check the neighbors to see if they have a similar color.
+If not then the Scot is moved around into different color and location.
+The worker work on the segment map where intailly  it has a fixed color liable and its moves randomly on location till it found a segmented location with same label as itself.
+The worker check the neighborhood location if the Scots has label it or not.
+if it is labeled with different color then the larger of the two labeled will still be present else  moves randomly till found an interesting neighborhood.
+%
 

### 2009-04-12 trunk/draft/System.tex
now is 11 april
before i change the results

@@
 \label{sec:AimsOfTheSystem}
-The previous chapter showed the current trends in sketch recognition.
+%The previous chapter showed the current trends in sketch recognition. 
 This thesis introduces a new approach for the sketch recognition problem.
-Particle Swarm Optimization (PSO) is used to correctly segment strokes into curves and lines.
 Particle Swarm Optimization (PSO) is used to correctly segment strokes into curves and lines.
@@
 \label{sec:AnOverviewOfTheSystem}
-   This research solves sketch segmentation problem using particle swarm algorithm PSO.
+   This research solves sketch segmentation problem using particle swarm algorithm (PSO).
 Due to sloppiness of users and hardware glitches the captured points must be processed to remove noise and redundancy before the system proceeds to compute the speed and curvature information.
@@
 
-The \textbf{preprocessing} step is  is responsible for capturing the input data and removing the noise from it.
+The \textbf{preprocessing} step is  responsible for capturing the input data and removing the noise from it.
 As shown in Figure \ref{fig:Blockdiagram} preprocessing consists of computing the time differences, direction, speed and curvature of every point in the stroke.
@@
 \section{Hybird Segmentation Algorithm}
+\label{sec:BenchMarckAlgorithm}
 
-\label{sec:BenchMarckAlgorithm}
 In this research, we developed another algorithm as reference to the results obtained by the DPSO algorithms.
@@
 	\item Find the intersection points in both list and generate the a Sample solution $h_{curr}$. 
-	\item Use $h_{curr}$ as initial solution. Loop doing the following:
+	\item Use $h_{curr}$ as initial solution. Loop on the following:
 
@@
 The system uses a composite set of features include Rubine feature set \cite{gestureexample12}, Zernike moments \cite{HeloiseBeautification}, ink density \cite{GeometryAndDomain102} and some structural information like number of perpendiculars lines , number of parallel lines and types of primitives in each symbol.
-The table \ref{tab:FeatureTable} gives details on all the features used in the system.
+The following list gives details on all the features used in the system.
 
@@
 \begin{description}
-	 \item[No. Of segments] Number of segments in the symbol. 
-		\item[No. Of primitives] Number of primitives in the symbol. The feature helps when identifying             symbols with mixed geometric primitives like cylinders and callouts.  
-		\item [No. of curves] Number of curves or ellipses in the symbol. 
-		\item [No. of lines] Number of lines in the symbol. 
-		\item [No. of perpendicular lines] Number of perpendicular lines. 
-		\item [No. of parallel lines] Number of parallel lines. 
-		\item [No. of intersections] Number of intersection between lines and curves. 
+	 \item[Segments] Number of segments in the symbol.
+	 \item [Strokes] Number of strokes or partical strokes that created the symbols.  
+		\item[Primitives] Number of primitives in the symbol. The feature helps when identifying             symbols with mixed geometric primitives like cylinders and callouts.  
+		\item [Curves] Number of curves or ellipses in the symbol. 
+		\item [Lines] Number of lines in the symbol. 
+		\item [Perpendicular lines] Number of perpendicular lines. 
+		\item [Parallel lines] Number of parallel lines. 
+		\item [Intersections] Number of intersection between lines and curves. 
+		\item [T intersections] Number of T intersections. 
+		\item [L intersections] Number of L intersections. 
+		\item [X intersections] Number of X intersections.
 		\item [Size Ratio] Ratio between width to height of the symbol.
+
 		%\item[No. of holes] Number of holes in the symbols.
@@
 \begin{description}
-	\item[Zernike moments]  Zernike moments of order \cite{HeloiseBeautification}. 
+	\item[Zernike moments] Zernike moments of order n\footnote{ We tried different values of n. Results show the analysis of each order.} \cite{HeloiseBeautification}. 
 \end{description}

### 2009-04-12 trunk/draft/chapterSurvey.tex
before i change the results

--- 2009-04-12 trunk/draft/chapterSurvey.tex
@@
 Visual describtive languages represents the shape into tokens of geometric primitives or relations between smaller  geometric primitives.
-Figures \ref{fig:arrow1} and \ref{fig:arrow2} show a regular arrow with 2 shafts and head and its definition in Ladder language \cite{Ladder30}.
+Figures \ref{fig:arrow1} and \ref{fig:arr2} show a regular arrow with 2 shafts and head and its definition in Ladder language \cite{Ladder30}.
 Shape recognition is then a question of parsing the sketch document into a recognizer that parse strokes according to the constructed grammar.
@@
 		\subfigure[Arrow grammar definition]
-		 {	\label{fig:arrow2}\includegraphics[scale=0.88]{../../neededfiles/Figures/arrowladderdef2.eps}}
-	\caption[Arrow definition]
+		 {\label{fig:arr2}\includegraphics[scale=0.88]{../../neededfiles/Figures/arrowladderdef2.eps}}
+	\caption[Arrow definition]{}
 		\label{fig:arrowladderdef1}
@@
 
-%Zernike moments were used by Hse et al. in \cite{}. 
+%Zernike moments were used by Hse et al. in \cite{ToADD}. 
 Hse et al.
@@
 Methods based on gesture and global stroke properties produce systems that are dependent on assumptions about how strokes can be groubed.
-These systems assume that users will finish drawing a shape before moving to the next one \cite{6multistrokeEvaluation,Cali63,geometrydomain49}.
+These systems assume that users will finish drawing a shape before moving to the next one \cite{multistrokeEvaluation,Cali63,geometrydomain49}.
 This let them to temporally group strokes and then recognize them based on their strokes or apperance properties.
@@
 
-\section{Particle Swram Optimization Review}
+\section{Swram Optimization Review}
 \label{sec:ParticleSwramOptimizationReview}
 
-%review on the pso alogrithm alone.................
+%This seems to emerge of what is called 'collective intelligence' from often large simple agents.  
+% start with swarm in general  .. say about flock , ant and usin in image processing 
+% then speeak about PSO and  its firstly generated   (see article)
+%mention that is best than ga (see thesis)
+
+Modeling how a biological swarm has been recentlsy studied by biologist and computer scientists.
+Scientist in the field or artifical life need to understand how such "social anaimals" interact, achieve goals and evolove.
+Moreover, swarm behaviours that achieves some kind of swram intelligance is increasly intersting for engineers as it they can be applied in optimization.
+Looking from a higher view, agents in a swarm are cooperating to achieve some proposeful behaviour and achieve some common goal.
+For example, ant groups looking for food source.
+Large gourps of simple independent agents seems to generate this behaviour in what is called 'collective intelligence'.
+Each agent has a set of simple rules to follow, and accourding to the interaction of the entrie group the objective is met\cite{swarmGeneral}.
+%with other neighbourhood agents.
+
+Reynolds \cite{swarmflock} created 'boid' which simulated the motion of a flock of birds.
+Each boid is a simple agent that navigate according to his perception and a set of basic rules.
+Kube and Zhange \cite{antsimulation} introduced a simulation model for ants behvaiour and used in cooperative transportation.
+Ants leaves a trail of its movement, so when an ant find food it cannot cary alone, it can easily trace back to the colony and then a group of ants can track the trail and bring the food back.
+Kube and Zhange \cite{antsimulation,swarmGeneral} system seems to corectly models ants behavour with a methods of task modeling.
+
+Image processing is one of the displines that used Swarm Intelligance.
+Ant colony was widely used to segment objects in \cite{antSegmentation,antthreshold,SwarmMedical}.
+Since 1995 when Kennedy and Eberhart \cite{PSOFirst,PSO2} introduced particle swarm alogrithm (PSO) the algorithm had been used in various image processing sytem as an optimization  algorithm \cite{SwarmImageProcessing,Swarmimageprocessing2,Colorquantization,SwarmMedical}.
+PSO algorithm was used in \cite{ToADD} to classify pixel to simplify the object segmentaiton process.
+
+ToADD
 

### 2009-04-12 draft/Results.tex
before i change the results

@@
 %The variable used to tune the swarm algorithm were tested to choose the best parameters to use
+
 \section{Datasets}
@@
 Four different splits for the training and test data are generated from the dataset.
-%The results displayed are the average recognition accuracy of the four splits resulted from the classifier after recognizing the symbols.
+In worth noting that \cite{HeloiseBeautification} and \cite{Oltmans07} achieved 96.7\% and  94.4\%  on this dataset respecitvily.
+%, 94.4\%  , which
+%is comparable to the 96.7%  %The results displayed are the average recognition accuracy of the four splits resulted from the classifier after recognizing the symbols.
 		\item [MultiDomain Dataset] A dataset we collected manually from different users.

### 2009-04-12 trunk/draft/Results.tex

@@
 The recogntion rate (computed as the number of correctly recognized samples divided by the total number of samples in the test) for each segmentation algorithm compared to another segmenation algorithm are reported in \ref{sec:recognitionAlgorithms}.
-Next, the scallability of our recognition system is evaluated and reported in Section \ref{sec:recognitionAlgorithms}.
-Finally, a feature set comparision is reported in Section \ref{sec:featuresComparisions}.
-
-%The accuracy is computed as the number of correctly recognized samples divided by the total number of samples in the test. 
-%The recognition system evalutation is presented in Section \ref{sec:SegmentationAlgorithms}. The  
-%The variable used to tune the swarm algorithm were tested to choose the best parameters to use
+Next, a comparison between different shapes and datasets and it influence on the recogntion rate is presented in Section \ref{sec:EffectsofSymbolComplixity}.
+Finally, a  comparision between different feature vectors is reported in Section \ref{sec:featuresComparisions}.
 
@@
 \begin{description}
-	\item [Heloise Dataset] Data set collected by \cite{HeloiseBeautification} and was previously used by \cite{Oltmans07} as benchmarck for sketch recognition problem.
+	\item [Hse Dataset] (denoted as Hs-DB) Data set collected by \cite{HeloiseBeautification} and was previously used by \cite{Oltmans07} as benchmarck for sketch recognition problem.
 The data are drawn by 16 users each of them have drawn from 30 to 50 samples for each shape.
@@
 Four different splits for the training and test data are generated from the dataset.
-In worth noting that \cite{HeloiseBeautification} and \cite{Oltmans07} achieved 96.7\% and  94.4\%  on the dataset respecitvily.
+In worth noting that \cite{HeloiseBeautification} and \cite{Oltmans07} achieved 96.7\% and  94.4\%  on this dataset respecitvily.
 %, 94.4\%  , which
-is comparable to the 96.7%  %The results displayed are the average recognition accuracy of the four splits resulted from the classifier after recognizing the symbols.
-		\item [MultiDomain Dataset] A dataset we collected manually from different users.
+%is comparable to the 96.7%  %The results displayed are the average recognition accuracy of the four splits resulted from the classifier after recognizing the symbols.
+		\item [MultiDomain Dataset] (denoted as MD-DB) A dataset we collected manually from different users.
 The dataset contains symbols from three diffrent domains ( UML, mechanical and electrical engineering symbols).
@@
 \begin{description}
-	\item[Segmentation Error] The segmentation error is computed from each stroke after segmentation as computed from Equation \ref{eq:bestFit1} and \ref{eq:bestFit2}. 
+	\item[Segmentation Error] The segmentation error is computed from each stroke after segmentation as computed from in $E_{alg}$ in Equation \ref{eq:bestFit2}.
+The error represents the sum of Square Root Error of the length from the point on the stroke to the estimated curve or line.
+
 	\item [Recognition accuracy] We measure recognition performance for each algorithm by determining the number of correctly identified symbols in the whole dataset.
@@
 
-Figrue \ref{fig:swarmtesting0} shows the \textit{Average Segmentation Error} using different preliminary calculation (Section \ref{sec:CurvatureCalculation}).
+Figrue \ref{fig:SegErrPreCal} shows the \textit{Average Segmentation Error} using different preliminary calculation (Section \ref{sec:CurvatureCalculation}).
 The figure shows that using all computations (time difference, direction, speed and curvature) for each stroke decrease the Segmentation error.
-Figure \ref{fig:swarmtestingpso} shows the average number of Possible dominate point $P_{pd}$ with respect to information used in the system.
+Figure \ref{fig:PossibledpPreCalculation} shows the average number of Possible dominant point $P_{pd}$ with respect to information used in the system.
 The results shows that the using all information increases the number of Possible dominate point $P_{pd}$ which in segmenation stage decreses the \textit{Segmenation Error}.
+It is worth noting that when using all information the number of Possible dominant point $P_{pd}$ is not the same as the sum of using every information alone.
+This is because we found out that there was some redundant points found using more than one information.
 
@@
 	\centering
-	 \includegraphics[scale=0.3]{images/swarmTesting}
-	 	\caption{Swarm testing} Segmentation Error Vs. Number of agents
-	 	\label{fig:swarmtesting0}
-	%\caption{Experiments results :} a)    b) c) 
+		\includegraphics[scale=0.5]{results/SegErrPreCal.eps}
+	\caption{preliminary calculation effect}
+	\label{fig:SegErrPreCal}
 \end{figure}
 
-\begin{figure*}
+\begin{figure}
 	\centering
-	 \includegraphics[width=18cm]{images/swarmTesting2}
-	 	\caption{Swarm testing} Segmentation Error Vs. PSO parameters.
-	 	\label{fig:swarmtestingpso}
-	 	\end{figure*}
+		\includegraphics[scale=0.5]{results/PossibledpPreCalculation.eps}
+	\caption{Possible Dominant Point Count}
+	\label{fig:PossibledpPreCalculation}
+\end{figure}
 
  The effect of number of agents, the error thresholds and various other paramters was investigated.
-The result of these tests are shown in Figure \ref{fig:swarmtesting} and  \ref{fig:swarmtesting2}.
-Figure \ref{fig:swarmtesting} shows that as the number of agents increases the \textit{Segmenation Error} value decrease (error of the segmentation algorithm).
-Figure \ref{fig:swarmtesting1} shows simillar result for the maximum swarm iterations.
+The result of these tests are shown in Figure \ref{fig:swarmtestingS1} and  \ref{fig:swarmtestingS2}.
+Figure \ref{fig:swarmtesting} shows that as the number of agents increases the \textit{Segmenation Error} value decrease.
+Simillar results for the maximum swarm iterations are shown in Figure \ref{fig:swarmtesting1} .
 Other algorithm parameters like $c_1$,$c_2$ that were mentioned in Section\ref{sec:ParticleSwarmAlgorithm} are tested using similar tests (Figure \ref{fig:swarmtesting} and \ref{fig:swarmtesting2}).
-The final values are for maximum number of iteration = 60 with 15 particles.
+The final values are for maximum number of iteration = 80 with 15 particles.
 The final number of agents and maximum swarm iteration used in the system was based trade off between error achieved versus the computation time.
@@
 	\centering		
-	 \includegraphics[scale=0.3]{images/swarmTesting}
-	 	\caption{Swarm testing} Segmentation Error Vs. Number of agents
-	 	\label{fig:swarmtesting}
+	 \includegraphics[scale=0.3]{results/swarmalg1.eps}
+	 	\caption{Segmentation Error Vs. Maximum iterations For ALgS1}
+	 	\label{fig:swarmtestingS1}
 	%\caption{Experiments results :} a)    b) c) 
@@
 
-\begin{figure*}
+ \begin{figure}
 	\centering		
-	 \includegraphics[width=18cm]{images/swarmTesting2}
-	 	\caption{Swarm testing} Segmentation Error Vs. PSO parameters.
-	 	\label{fig:swarmtesting2}
+	 \includegraphics[scale=0.3]{results/swarmalg2.eps}
+	 	\caption{Segmentation Error Vs. Maximum iterations For ALgS2}
+	 	\label{fig:swarmtestingS2}
 	%\caption{Experiments results :} a)    b) c) 
-
-\end{figure*}
+\end{figure}
 
@@
 \label{sec:SegmentationAlgorithms}
-The result of the segmentation algorithm can be viewed in the Figure \ref{fig:results1.JPG}. These result shows the originals stroke and the segmentation that the system generated for the stroke. 
-\begin{figure}
-	\centering
-		\includegraphics{images/results1.JPG.eps}
-	\caption{Results Segmentations}
-	\label{fig:results1.JPG}
-\end{figure}
-\begin{figure}
-	\centering
-		\includegraphics{images/results2.JPG.eps}
-	\caption{Results Segmentaion }
-	\label{fig:results2.JPG}
-\end{figure}
-\begin{figure}
-	\centering
-		\includegraphics{images/results3.JPG.eps}
-	\caption{Results }
-	\label{fig:results3.JPG}
-\end{figure}
-\begin{figure}
-	\centering
-		\includegraphics{images/results4.JPG.eps}
-	\caption{Resluts of the segmentations}
-	\label{fig:results4.JPG}
-\end{figure}
-
-\subsection {Recognition Algorithms}
-\label{sec:RecognitionAlgorithms}%}
-To test the recognition different set of features are used to select the features that will define symbols better to finally achieve better recognition rate.
-We tested the system using different types of features.
-The features explained in section \ref{sec:FeatureExtraction} was divided into different sets; statistical, structural, Rubine features and zernik moments along with composite features.
-The fig.
-\ref{fig:testFeatures} shows the result of each set of features when used with the SVM classifier.
-The results shows that the Rubine features although may achieve good results in single stroke symbols, have a bad performance when used with multi stroke symbols.
-\begin{figure}[]
-	\centering
-		\includegraphics[scale=0.3]{images/test4.eps}
-	\caption{The results of Features experiment}
-	\label{fig:testFeatures}
-\end{figure}
-
-We test the system using different classifier to see what the effect of the classification step on the final recognition.
-We also wanted to test the train and test time in both classifiers.
-We used SVM (RBF) classifiers and the Gaussian Linear classifier used in \cite{}.
-The fig.
-\ref{} shows the result of both classifiers with the main three segmentation algorithms.
-The results shows that the SVM classifier achieve better results.
-To measure the scalability of the system\footnote{What happens when adding a new symbols to the set of known symbols.} the classifiers are test using different number of symbols (fig.
-\ref{fig:sampleperCat}).
-Figures \ref{fig:sampleperCat} demonstrate the scalability of both classifiers by demonstrating the recognition rate in both classifiers with different number of symbols.
-Figures \ref{fig:sampleperCat},\ref{fig:sampleperCat} shows the time and number of symbols needed for training versus each classifier.
+The result of the segmentation algorithm can be viewed in the Figure \ref{fig:results1}.
+These result shows the originals stroke and the segmentation that the system generated for the stroke.
+The figure \ref{fig:results2} shows the result of \textsl{AlgS1} and \textsl{AlgS2} for the same input strokes.
+Figure \ref{fig:results3} shows the ouptut of \textsl{Alg3} to the same set of input strokes.
+It clrearly shows that \textsl{AlgS2} have better segmentation results than both \textsl{AlgS1} and \textsl{Alg3} but since these segmentation are subjective to what the user intended to draw the next section will focus on the effect of the segmenation algorithm on the recognition accuracy.
 
 \begin{figure}
+
 	\centering
-		\includegraphics{images/sampleperCat.eps}
-	\caption{Samples per Category }
-	\label{fig:sampleperCat}
+				\subfigure{\includegraphics[scale=0.55]{results/result1.eps}}
+					\subfigure{	\includegraphics[scale=0.6]{results/result1logic.eps}}
+	\caption{Segmentation Results of \textsl{AlgS1}}
+	\label{fig:results1}
+
 \end{figure}
@@
 	\centering
-		\includegraphics[scale=0.7]{images/test4.eps}
-	\caption{Tests}
-	\label{fig:test4}
+			\subfigure{	\includegraphics[scale=0.55]{results/result2.eps}}
+			\includegraphics[scale=0.6]{results/result2logic.eps}
+	\caption{Segmentation Results of \textsl{AlgS2}}
+	\label{fig:results2}
+\end{figure}
+\begin{figure}
+	\centering
+		\subfigure{
+		\includegraphics[scale=0.55]{results/result3.eps}}
+			\subfigure{\includegraphics[scale=0.55]{results/result3logic.eps}}
+
+	\caption{Segmentation Results of \textsl{Alg3} }
+	\label{fig:results3}
 \end{figure}
 
-\subsubsection {Effect of sysmbol complixity}
+\subsection {Recognition Algorithms}
+\label{sec:RecognitionAlgorithms}
 
-The experiment we implemented was to test the effect of symbol complexity and type on the recognition rate.
-Figure \ref{fig:test2} shows what the accuracy of each symbol, it clearly noted that symbols that have only line segments achieve higher accuracy rate than other symbols.
-Figure \ref{fig:test2} also shows that algorithm 1 alone achieve better performance than algorithm 2 in the symbols that consist of lines only.
-The combining of both algorithm improves the recognition rate of other symbols.
-\\
-\begin{figure*}[]
+e performed several experiments to evaluate the presented recognition system.
+Firstly we tested recognition accuracy of shapes in the data set with both algorithms (\textsl{AlgS1} and \textsl{AlgS2}).
+We also implemented the segmentation algorithm described in \cite{earlyprocess} (\textsl{Alg3}) to use as reference to compare it with our swarm algorithms.
+Figure \ref{fig:test1} shows the accuracy achieved by each algorithm.
+The two swarm algorithms were tested with and without the ellipse fitting module.
+The ellipse detection module appears to be superior.
+The result shows that combining both  (\textsl{AlgS1} and \textsl{AlgS2}) out perform any single algorithm.
+%show that combining algoirthmsachieve higher accuracy than any single algoirthm.
+%improves the results.% with both the \textit{DPSO} algorithms.% The results show that both \textit{DPSO} algorithms achieve better result than other algorithms.
+
+ \begin{figure}
 	\centering		
-		\includegraphics[scale=0.5]{images/test2.eps}
-	\caption{The results of second experiment}
+	 \includegraphics[scale=0.4]{results/testAlg.eps}
+	 	\caption{Algorithm comparison} The recognition rate of different algorithms. 
+	 	\label{fig:test1}
+	%\caption{Experiments results :} a)    b) c) 
+\end{figure}
+
+\subsubsection {Effects of Symbol Complixity }
+ \label{sec:EffectsofSymbolComplixity}
+
+\begin{figure*}
+	\centering
+		\includegraphics[scale=0.5]{results/testsym.eps}
+	\caption{Symbols Comparison} The effect of symbol complexity.  %The graph shows the recognition rate of each symbol using different algorithms. 
 	\label{fig:test2}
 \end{figure*}  
- Another experiment we performed was to test the efficiency svm classifier.
-We implemented the linear discriminator classifier described in \cite{gestureexample12}(see figure \ref{fig:test3}).
-%The result shows that SVM is more scalable than the linear discriminator for this problem.
-%The test focus on the size of data needed to test each classifier and the achieved accuracy.
-%For this experiment we also shows the classification time for both classifiers
-\begin{figure}[]
+The second experiment we implemented was to investigate the effect of symbol complexity and type on the recognition rate.
+Figure \ref{fig:test2} shows the achieved accuracy of each symbol by our algorithms compared to \cite{earlyprocess} (\textsl{Alg3}).
+It is clearly noted that symbols that have only line segments achieve higher accuracy rate than other symbols.
+The results indicate that algorithm \textsl{AlgS1} achieve better performance than algorithm \textsl{AlgS2} in the symbols that consist only of lines.
+This is understandable because algorithm \textsl{AlgS1} divides strokes into line segments only but \textsl{AlgS2} is able to divide strokes into lines and curves based on the minimum error of the segment itself.
+Algorithm \textsl{Alg3} gives good performance as long as the symbols consist of combination of lines and curves, if the stroke consists of only curve or lines the algorithm may lead to wrong segmentation result.
+This is because the system divides the stroke first to line segments then tries to decide if each segment can be represent better as a curve unlike algorithm \textsl{AlgS2} where the curve segments are tested while choosing the best segmentation.
+Combining both algorithm \textsl{AlgS1} and \textsl{AlgS2} improved the recognition rate of all symbols.
+The penalty for this improved performance is the computational time required to run both swarm algorithms.
+Table \ref{tab:ConfusionMatrix} shows the confusion matrix of symbols.
+The table shows that errors are only between two sets of symbols moon with triangles and trapezoid with triangles.
+This observation is understandable as the symbols are visually similar but the system must be able to differentiate between them.
+This lead us to the next experiment to choose best set of feature to recognize symbols.
+ \begin{table*}
 	\centering
-		\includegraphics[scale=0.4]{images/test1.eps}
-	\caption{The results of third experiment}
-	\label{fig:test3}
-\end{figure}
- %Details description of results got from comparing various algorithms  
- %Describe data set used for experiments ( No of data set, No. of categories, No of samples).
- %Describe other Algorithms [Alg 3 in paper \cite{earlyprocess}]used for comparing with the one described in the paper.
-
- %In the figure \ref{fig:test1}, the accuracy means that final hit rate of all symbols used in the test set.by the  four datasets . 
-We performed three experiments to test the system firstly we tested recognition accuracy of shapes in the data set with both algorithms.
-We also implemented the segmentation algorithm described in \cite{earlyprocess} to use as reference to our swarm algorithms.
-As you see in the fig.
-\ref{fig:test1} shows the accuracy achieved by each algorithm.
-The results shows that both PSO algorithm achieve better result than other algorithms.
-The swarm algorithms were tested with and without the ellipse detection module.
-The ellipse detection module appear to be superior to results with only the PSO algorithms.
-\\
-\begin{figure}[]
-	\centering
-		\includegraphics[scale=0.3]{images/test1.eps}
-	\caption{The results of first experiment}
-	\label{fig:test1}
-\end{figure}
-
-%\subsection{Clustering Algorithm Results}
-%\label{sec:ClusteringAlgorithmResults}
+	%\small
+%	\scalebox{0.7}{
+		%	\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|c|c|c|c|}\hline 
+	\scalebox{0.7}{
+		\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|c|c|c|c|}\hline 
+ Categories 	&ellipse	&heart	&trapezoid	&pentagon	&arch	&hexagon	&square	&triangle	&cube	&cylinder	&parallelogram	&moon	&callout\\ \hline
+ellipse	&155	&9	&0	&0	&0	&5	&0	&0	&0	&0	&0	&6	&0\\ \hline
+heart	&0	&109	&1	&1	&25	&0	&0	&1	&0	&0	&2	&1	&1\\ \hline
+trapezoid	&0	&0	&103	&5	&2	&0	&0	&2	&0	&0	&34	&0	&0\\ \hline
+pentagon	&0	&0	&20	&121	&0	&20	&6	&2	&0	&0	&1	&0	&0\\ \hline
+arch	&0	&1	&13	&0	&86	&0	&0	&2	&3	&0	&2	&0	&0\\ \hline
+hexagon	&0	&0	&0	&0	&0	&98	&0	&0	&0	&0	&0	&0	&0\\ \hline
+square	&0	&0	&1	&0	&2	&0	&107	&14	&0	&0	&8	&1	&0\\ \hline
+triangle	&0	&0	&7	&0	&1	&0	&0	&89	&0	&0	&12	&4	&0\\ \hline
+cube	&0	&0	&0	&0	&0	&0	&0	&0	&147	&0	&0	&0	&0\\ \hline
+cylinder	&0	&0	&0	&6	&2	&8	&16	&0	&4	&156	&1	&10	&0\\ \hline
+parallelogram	&0	&0	&4	&0	&0	&0	&0	&0	&0	&0	&91	&0	&0\\ \hline
+moon	&0	&3	&3	&2	&17	&0	&9	&17	&0	&0	&2	&133	&11\\ \hline
+callout	&1	&34	&0	&22	&21	&22	&16	&25	&0	&0	&0	&0	&143\\ \hline
+		\end{tabular}
+			}
+ 		%\end{tabular}
+ 	%}
+	\caption{Confusion Matrix}
+	\label{tab:ConfusionMatrix}
+	%\end{minipage}
+\end{table*}
 
@@
 \label{sec:featuresComparisions}
+Different feature sets are tested to determine the best features that can used in sketch and symbol recognition.
+Figure\ref{fig:testFeaturesAllHS} shows the result of different feature sets from the basic four sets (\textbf{FS1,FS2,FS3,FS4}).
+Result shows that (\textbf{FS2}) Rubine features \cite{gestureexample12} gives the worst results when used alone without any other features.
+This is because they are mainly computed for single stroke gestures and fare bad in multi-stroke symbols \cite{compareFeaturSVM}.
+Features \textbf{(FS4)} gives good results but it is improved by adding structural features \textbf{(FS1)}.
+Figure \ref{fig:testFeaturesAllMD} shows the simillar results on \textsl{MD-DB} multidomain database.
+The figure shows that on this dataset \textbf{FS1} performe better than on \textsl{HS-DB}, this is a result that shapes in the \textsl{MD-DB} has more structural difference than shapes in \textsl{HS-DB}.
+ \begin{figure}
+	\centering
+		\includegraphics[scale=0.5]{results/testFeat.eps}
+	\caption{Feature Comparison} The effect of different features on accuracy on \textsl{Hs-DB}.  %The graph shows the recognition rate of each symbol using different algorithms. 
+	\label{fig:testFeaturesAllHS}
+\end{figure}
+
+ \begin{figure}
+	\centering
+		\includegraphics[scale=0.5]{results/testF1.eps}
+	\caption{Feature Comparison} The effect of different features on accuracy on \textsl{MD-DB}.  %The graph shows the recognition rate of each symbol using different algorithms. 
+	\label{fig:testFeaturesAllMD}
+\end{figure}
 

### 2009-04-12 draft/System.tex

@@
 This evaluation is based on the area of both the segmentation and the input stroke.
-Equation \ref{eq:bestFit1} and \ref{eq:bestFit2} are used to evaluate each segmentation.
-
+The segmentation error($E_{alg}$ in Equation \ref{eq:bestFit2})represents the sum of Square Root Error of the length from the point on the stroke to the estimated curve or line.
+In \textsl{AlgS2}, this error is the same as the computed error in Equation \ref{eq:ErrorSwarm1} but For the second algorithm \textsl{AlgS2} it is different than the computed error in Equation \ref{eq:errorSwarm2}.
+The additional computation is necessary to  compare the two segmentation fairly.
 \begin{equation}
@@
  \end{equation}
- where $A_{st}$ is the area of the input stroke, $A_{sg}$ is area of computed segmenation, $E_{alg}$ is the approximation error of the segmentation.
+ where $A_{st}$ is the area of the input stroke, $A_{sg}$ is area of computed segmenation, $E_{alg}$ is the Segmentation Error.
 The segmentation which crosspond to the minimum $E_{tot}$ is the best segmentation.
 
-\section{Hybird Segmentation Algorithm}
+\section{Hybird Segmentation Algorithm \textsl{Alg3}}
 \label{sec:BenchMarckAlgorithm}

### 2009-04-12 trunk/draft/chapterSurvey.tex
after correctly splled now look at any thing else

@@
 The history of pen and sketch based computing goes back to 1963 with Ivan Sutherland's SKETCHPAD \cite{sutherlandsketch}, which used a light pen to draw on the monitor to create circuit diagrams.
-Current technology and the increae in computuing power have increased the abilities of such interfaces and allowed the implementaion of fully sketching systems.
-Over the years, many other sketching based system have been developed, elthough throught this history most system foucused on building the interactive interfaces to assist users while entering information to the computer.
+Current technology and the increase in computing power have increased the abilities of such interfaces and allowed the implementation of fully sketching systems.
+Over the years, many other sketching based system have been developed, although through this history most system focused on building the interactive interfaces to assist users while entering information to the computer.
 To demonstrate our system inspiration we will first present current research in the area of sketch recognition.
-Fristly, preprosessing and segmentaiton methods will be discussed in details.
-Later, symbol recognition methods will be presented and analysied in Section \ref{sec:symbolrecogntion}.
+Firstly, preprocessing and segmentation methods will be discussed in details.
+Later, symbol recognition methods will be presented and analyzed in Section \ref{sec:symbolrecogntion}.
 
@@
 The goal of a sketch system is making the user feel that using the using the system is nearly the same as using pencil and paper.
-But sketchs are incomplete drawings that are sloppy and messy, for example figures \ref{fig:Overshootandundershoot} and \ref{fig:overstroked} show examples of messy sketches drawn by users.
+But sketches are incomplete drawings that are sloppy and messy, for example figures \ref{fig:Overshootandundershoot} and \ref{fig:overstroked} show examples of messy sketches drawn by users.
 This led researchers to perform sort of preprocessing before starting the sketch recognition task itself.
@@
  Agar et al.
-\cite{polygonfeedback31} used the timestamp for each point and compute time difference between every two successive point to help him finding corners as it provides more distinct maxima than the speed information.
+\cite{polygonfeedback31} used the time stamp for each point and compute time difference between every two successive point to help him finding corners as it provides more distinct maximum than the speed information.
 He mentions that the pointing device (for example the mouse) sampling rate is the reason for this phenomena.
@@
 
-\section{Symbol Recogntion}
+\section{Symbol Recognition}
 \label{sec:symbolrecogntion}
@@
 
-\subsubsection{Guesture Recognition}
+\subsubsection{Gesture Recognition}
 \label{sec:GuestureRecognition}
@@
 
- Systems that employ such representation must first preprocess strokes to break them into geometric primitives (see section \ref{sec:preprocessing} for a review on preprocessing methods).
+ Systems that employ such representation must first pre-process strokes to break them into geometric primitives (see section \ref{sec:preprocessing} for a review on preprocessing methods).
 Recognition can then be treated as sub graph matching problem.
@@
 		\includegraphics[scale=0.7]{../../neededfiles/Figures/squarescematic.eps}
-	\caption{square scemantic graph}
+	\caption{square semantic graph}
 	\label{fig:squarescematic}
@@
 
- \paragraph{Visual Describtive Languages:}
+ \paragraph{Visual Descriptive Languages:}
  \label{sec:VisualDescribtiveLanguages}
 
- Avoiding graph matching problems, research tried to use languages that discribe shapes instead of graphs.
-Visual describtive languages represents the shape into tokens of geometric primitives or relations between smaller  geometric primitives.
+ Avoiding graph matching problems, research tried to use languages that describe shapes instead of graphs.
+Visual descriptive languages represents the shape into tokens of geometric primitives or relations between smaller  geometric primitives.
 Figures \ref{fig:arrow1} and \ref{fig:arr2} show a regular arrow with 2 shafts and head and its definition in Ladder language \cite{Ladder30}.
@@
 
- Several visual grammer were introduced \cite{statisticalparsing26,Ladder30,GenericHMM28}.
+ Several visual grammar were introduced \cite{statisticalparsing26,Ladder30,GenericHMM28}.
 \cite{Ladder30,GenericHMM28}  Used visual description language which will generate the information of the structure and constrain of each object classified.
-The system  automatically generates java code for the recognizer for each object.
+The system  automatically generates Java code for the recognizer for each object.
 The generated code cycles through the strokes drawn to check if any of the recognizers generated had interpreted any symbols partially or completely.
-These systems are less computaitonal expensive than graph matching approach but they also operate based on the output of the segmentation process.
+These systems are less computational expensive than graph matching approach but they also operate based on the output of the segmentation process.
 Any error in the segmentation process lead to huge error while parsing the whole sketch.
-Moreover, symbols cannot share strokes between them which infleunce and restrict how the user draw symbols.
+Moreover, symbols cannot share strokes between them which influence and restrict how the user draw symbols.
  % complexity of the graph matching problem 
@@
 
- Several systems \cite{DiagramOfflineConvexHull,MulitStrokeConvexHull} presented features like ratio of bounding box to convex hull area and the ratio of the preimeter to the area.
+ Several systems \cite{DiagramOfflineConvexHull,MulitStrokeConvexHull} presented features like ratio of bounding box to convex hull area and the ratio of the perimeter to the area.
 These global feature can be used to distinguish between classes.
-For example, a rectangle can be distinguish from a triangle by looking to the raito of area of bounding box to the area of the convex hull.
+For example, a rectangle can be distinguish from a triangle by looking to the ratio of area of bounding box to the area of the convex hull.
 The triangle ratio will be much lower than that of the rectangle.
@@
 \cite{HeloiseBeautification,zernike61} proposed using the magnitudes of Zernike moments as a shape property.
-Zernike moemnts are orthognals moments which define the distribution of point in the input space.
+Zernike moments are orthogonal moments which define the distribution of point in the input space.
 The higher the order of the moment the finer the detail level the point distribution represents.
@@
 \cite{HeloiseBeautification,zernike61} used a SVM classifier to train the features vector calculated from the moments.
-Early evluation of Zernike moemnts showed that they are invariant to both reflection and rotation of the input shape \cite{ZerMomentOrthogonal}.
-In our system we used Zernike moemnts as part of our feature vector.
+Early evaluation of Zernike moments showed that they are invariant to both reflection and rotation of the input shape \cite{ZerMomentOrthogonal}.
+In our system we used Zernike moments as part of our feature vector.
 A detailed comparison between Zernike moments and other features are demonstrated in Section \ref{sec:featuresComparisions}.
@@
  Systems based on Global shape properties distinguish shapes by learning the different properties each shape class tend to have.
-These properties dose not depend on the order or number of strokes and thus are only based on apperance of the shape not how the user draw it.
+These properties dose not depend on the order or number of strokes and thus are only based on appearance of the shape not how the user draw it.
 However, individual details of the shape are impossible to distinguish using these properties.
-For example, it is impoissible to use  to  differentiate between a current source symbole (circle contiaing and arrow Figure \ref{fig:CurrentSource}) and a Ac Source (circle containing a tilate Figure \ref{fig:acSource}).
-Finear details of the symbols are not represented in this approach, thus shapes that have smalls details or only significant conceptual variation are not distiguished using this method.
+For example, it is impossible to use  to  differentiate between a current source symbol (circle containing and arrow Figure \ref{fig:CurrentSource}) and a Ac Source (circle containing a tilate Figure \ref{fig:acSource}).
+Finer details of the symbols are not represented in this approach, thus shapes that have smalls details or only significant conceptual variation are not distinguished using this method.
 These symbols requires a system to represent different levels of shape details.
@@
 
-In the third approach, the recognition foucus on appearance of shapes contrary to individual strokes or global shape properties.
+In the third approach, the recognition focus on appearance of shapes contrary to individual strokes or global shape properties.
 Kara \cite{imagetrainable48} used this approach while presenting a system that matches the input shape into a one of the  prototypes in the database.
@@
 Input shapes are first pose and scale normalized then four image similarity measure are used to match input shapes to the database prototypes.
-He used two measures (Hausdor distance)  to measure similarity and two coeficients (Yule, Tanimoto) to measure dissimilarity.
-His approach allows the system to recongized a shape using only a single training sample as it is based on the shape prototype.
-However, the shape protoytpe must represent all  transformation and variation the input shape can have.
-Moreover, they restrict the image representation to 48x48 pixel due to computationa reasons.
-This restriction leads to elimintating some of the fine details that may be needed to distinguish some shape classes.
+He used two measures (Hausdor distance)  to measure similarity and two coefficients (Yule, Tanimoto) to measure dissimilarity.
+His approach allows the system to recognized a shape using only a single training sample as it is based on the shape prototype.
+However, the shape prototype must represent all  transformation and variation the input shape can have.
+Moreover, they restrict the image representation to 48x48 pixel due to computational reasons.
+This restriction leads to eliminating some of the fine details that may be needed to distinguish some shape classes.
 There results on isolated symbols seems promising but the system have not beet tested on full sketch that tends to have high overlapping between symbols.
@@
 
-In all of these system only \cite{SketchRead2007} using stroke based recognitiion have been implemented to the fully creat a system that free draw sketch and understand them.
-This is because the emphasis on the interactive side of the sketch sytem which result in large difficultes in processing free hand sketches.
-Methods based on gesture and global stroke properties produce systems that are dependent on assumptions about how strokes can be groubed.
+In all of these system only \cite{SketchRead2007} using stroke based recognition have been implemented to the fully creat a system that free draw sketch and understand them.
+This is because the emphasis on the interactive side of the sketch sytem which result in large difficulties in processing free hand sketches.
+Methods based on gesture and global stroke properties produce systems that are dependent on assumptions about how strokes can be grouped.
 These systems assume that users will finish drawing a shape before moving to the next one \cite{multistrokeEvaluation,Cali63,geometrydomain49}.
-This let them to temporally group strokes and then recognize them based on their strokes or apperance properties.
-\cite{AlvaradoDigital} argued that this assumption dose not hold in natural sketchs.
+This let them to temporally group strokes and then recognize them based on their strokes or appearance properties.
+\cite{AlvaradoDigital} argued that this assumption dose not hold in natural sketches.
 
-Recogntion based on strokes using hierarchical shape description is applied in \cite{SketchRead2007}.
-The principle problem in these systems is the computatnal complexity of the matching between shape description and geometeric primitves produced in the segmentaiton process.
-Natural sketch are dificult to segment and group to create reliable fragment of geometeric primitives.
-This is partially due to the increase of noise and other phenomena such as overlapped, overtraced strokes and sharing of strokes between different symbols.
-Since these systems are depenable on the segmenation results, they either performe recognition using undependable segmentation results or consider different ways to segment the input sketch.
+Recognition based on strokes using hierarchical shape description is applied in \cite{SketchRead2007}.
+The principle problem in these systems is the computational complexity of the matching between shape description and geometric primitives produced in the segmentation process.
+Natural sketch are difficult to segment and group to create reliable fragment of geometeric primitives.
+This is partially due to the increase of noise and other phenomena such as overlapped, over traced strokes and sharing of strokes between different symbols.
+Since these systems are dependable on the segmentation results, they either performe recognition using undependable segmentation results or consider different ways to segment the input sketch.
 
-To avoid most problems in such system, our system uses hybird of both global shape and hierarchical discribtion methods.
-Currenlty, our system recognize on sysmbols at a time.
+To avoid most problems in such system, our system uses hybrid of both global shape and hierarchical description methods.
+Currently, our system recognize on symbols at a time.
 Each symbol is segmented and hierarchical features are computed.
-Features  that are represent the global discription of the shape are also generated and appended to the hierarchical featusres.
+Features  that are represent the global description of the shape are also generated and appended to the hierarchical features.
 The recognition is done using SVM Linear classifier as it proved to be one of the best learning algorithms.
 
-\chapter{Particle Swram Optimization}
+\chapter{Particle Swarm Optimization}
 \label{sec:ParticleSwramOptimization}
 
-\section{Swram Optimization Review}
+\section{Swarm Optimization Review}
 \label{sec:ParticleSwramOptimizationReview}
@@
 
-Modeling how a biological swarm has been recentlsy studied by biologist and computer scientists.
-Scientist in the field or artifical life need to understand how such "social anaimals" interact, achieve goals and evolove.
-Moreover, swarm behaviours that achieves some kind of swram intelligance is increasly intersting for engineers as it they can be applied in optimization.
-Looking from a higher view, agents in a swarm are cooperating to achieve some proposeful behaviour and achieve some common goal.
+Modeling how a biological swarm has been recently studied by biologist and computer scientists.
+Scientist in the field or artificial life need to understand how such "social animals" interact, achieve goals and evolve.
+Moreover, swarm behaviors that achieves some kind of swarm intelligence is increasingly interesting for engineers as it they can be applied in optimization.
+Looking from a higher view, agents in a swarm are cooperating to achieve some purposeful behavior and achieve some common goal.
 For example, ant groups looking for food source.
-Large gourps of simple independent agents seems to generate this behaviour in what is called 'collective intelligence'.
-Each agent has a set of simple rules to follow, and accourding to the interaction of the entrie group the objective is met\cite{swarmGeneral}.
+Large groups of simple independent agents seems to generate this behavior in what is called 'collective intelligence'.
+Each agent has a set of simple rules to follow, and according to the interaction of the entire group the objective is met\cite{swarmGeneral}.
 %with other neighbourhood agents.
@@
 Each boid is a simple agent that navigate according to his perception and a set of basic rules.
-Kube and Zhange \cite{antsimulation} introduced a simulation model for ants behvaiour and used in cooperative transportation.
-Ants leaves a trail of its movement, so when an ant find food it cannot cary alone, it can easily trace back to the colony and then a group of ants can track the trail and bring the food back.
-Kube and Zhange \cite{antsimulation,swarmGeneral} system seems to corectly models ants behavour with a methods of task modeling.
+Kube and Zhange \cite{antsimulation} introduced a simulation model for ants behavior and used in cooperative transportation.
+Ants leaves a trail of its movement, so when an ant find food it cannot carry alone, it can easily trace back to the colony and then a group of ants can track the trail and bring the food back.
+Kube and Zhange \cite{antsimulation,swarmGeneral} system seems to correctly models ants behavior with a methods of task modeling.
 
-Image processing is one of the displines that used Swarm Intelligance.
+Image processing is one of the disciplines that used Swarm Intelligence.
 Ant colony was widely used to segment objects in \cite{antSegmentation,antthreshold,SwarmMedical}.
-Since 1995 when Kennedy and Eberhart \cite{PSOFirst,PSO2} introduced particle swarm alogrithm (PSO) the algorithm had been used in various image processing sytem as an optimization  algorithm \cite{SwarmImageProcessing,Swarmimageprocessing2,Colorquantization,SwarmMedical}.
-PSO algorithm was used in \cite{ToADD} to classify pixel to simplify the object segmentaiton process.
+Since 1995 when Kennedy and Eberhart \cite{PSOFirst,PSO2} introduced particle swarm algorithm (PSO) the algorithm had been used in various image processing system as an optimization  algorithm \cite{SwarmImageProcessing,Swarmimageprocessing2,Colorquantization,SwarmMedical}.
+PSO algorithm was used in \cite{ToADD} to classify pixel to simplify the object segmentation process.
 

### 2009-04-12 draft/Results.tex
after correctly splled now look at any thing else

@@
 \label{sec:Evaluation}
-This chapter discribes and illustrates the experiments performed on the system.
+This chapter describes and illustrates the experiments performed on the system.
 In the chapter, section \ref{sec:ExperimentsSetup} describes the experimental setup.
@@
 
-The system has two main parts the DPSO stroke segmenation and symbol recogntion.
-Each part is evaluated sperately.
+The system has two main parts the DPSO stroke segmentation and symbol recognition.
+Each part is evaluated separately.
 The preprocessing procedures and computation is evaluated using their effect on the segmentation algorithm.
 These results are reported in section \ref{sec:ResultsDetails}.
-The effect of number of agents, the error thresholds and various other PSO paramters are investigated in Section \ref{sec:PSO}.
+The effect of number of agents, the error thresholds and various other PSO parameters are investigated in Section \ref{sec:PSO}.
 %The results presented in Section \ref{sec:ResultsDetails} and \ref{sec:PSO}.
 
-The recognition system evalutation is presented in Section \ref{sec:SegmentationAlgorithms}.
-The recogntion rate (computed as the number of correctly recognized samples divided by the total number of samples in the test) for each segmentation algorithm compared to another segmenation algorithm are reported in \ref{sec:recognitionAlgorithms}.
-Next, the scallability of our recognition system is evaluated and reported in Section \ref{sec:recognitionAlgorithms}.
-Finally, a feature set comparision is reported in Section \ref{sec:featuresComparisions}.
-
-%The accuracy is computed as the number of correctly recognized samples divided by the total number of samples in the test. 
-%The recognition system evalutation is presented in Section \ref{sec:SegmentationAlgorithms}. The  
-%The variable used to tune the swarm algorithm were tested to choose the best parameters to use
+The recognition system evaluation is presented in Section \ref{sec:SegmentationAlgorithms}.
+The recognition rate (computed as the number of correctly recognized samples divided by the total number of samples in the test) for each segmentation algorithm compared to another segmentation algorithm are reported in \ref{sec:recognitionAlgorithms}.
+Next, a comparison between different shapes and datasets and it influence on the recognition rate is presented in Section \ref{sec:EffectsofSymbolComplixity}.
+Finally, a  comparison between different feature vectors is reported in Section \ref{sec:featuresComparisions}.
 
@@
 \begin{description}
-	\item [Heloise Dataset] Data set collected by \cite{HeloiseBeautification} and was previously used by \cite{Oltmans07} as benchmarck for sketch recognition problem.
+	\item [Hse Dataset] (denoted as Hs-DB) Data set collected by \cite{HeloiseBeautification} and was previously used by \cite{Oltmans07} as benchmark for sketch recognition problem.
 The data are drawn by 16 users each of them have drawn from 30 to 50 samples for each shape.
@@
 Four different splits for the training and test data are generated from the dataset.
-In worth noting that \cite{HeloiseBeautification} and \cite{Oltmans07} achieved 96.7\% and  94.4\%  on this dataset respecitvily.
+In worth noting that \cite{HeloiseBeautification} and \cite{Oltmans07} achieved 96.7\% and  94.4\%  on this dataset respectively.
 %, 94.4\%  , which
 %is comparable to the 96.7%  %The results displayed are the average recognition accuracy of the four splits resulted from the classifier after recognizing the symbols.
-		\item [MultiDomain Dataset] A dataset we collected manually from different users.
-The dataset contains symbols from three diffrent domains ( UML, mechanical and electrical engineering symbols).
+		\item [Multi Domain Dataset] (denoted as MD-DB) A dataset we collected manually from different users.
+The dataset contains symbols from three different domains ( UML, mechanical and electrical engineering symbols).
 Data are collected from 4 different users each of them have drawn from 20 to 40 samples of each shapes.
 Figures \ref{fig:LogicSet} and \ref{fig:ElectSet} show some of the shapes used in the dataset.
-Symbols from different domains were gathered to evaluate the system thourhgtout different types of shapes.
-Useres were requested to draw symbols without any restriction on style,size or orientations.
+Symbols from different domains were gathered to evaluate the system throughout different types of shapes.
+Users were requested to draw symbols without any restriction on style,size or orientations.
 The dataset was divided into training set and test set.
@@
 
-Figrue \ref{fig:swarmtesting0} shows the \textit{Average Segmentation Error} using different preliminary calculation (Section \ref{sec:CurvatureCalculation}).
+Figure \ref{fig:SegErrPreCal} shows the \textit{Average Segmentation Error} using different preliminary calculation (Section \ref{sec:CurvatureCalculation}).
 The figure shows that using all computations (time difference, direction, speed and curvature) for each stroke decrease the Segmentation error.
-Figure \ref{fig:swarmtestingpso} shows the average number of Possible dominate point $P_{pd}$ with respect to information used in the system.
-The results shows that the using all information increases the number of Possible dominate point $P_{pd}$ which in segmenation stage decreses the \textit{Segmenation Error}.
+Figure \ref{fig:PossibledpPreCalculation} shows the average number of Possible dominant point $P_{pd}$ with respect to information used in the system.
+The results shows that the using all information increases the number of Possible dominate point $P_{pd}$ which in segmentation stage decreases the \textit{Segmentation Error}.
+It is worth noting that when using all information the number of Possible dominant point $P_{pd}$ is not the same as the sum of using every information alone.
+This is because we found out that there was some redundant points found using more than one information.
 
@@
 	\centering
-	 \includegraphics[scale=0.3]{images/swarmTesting}
-	 	\caption{Swarm testing} Segmentation Error Vs. Number of agents
-	 	\label{fig:swarmtesting0}
-	%\caption{Experiments results :} a)    b) c) 
+		\includegraphics[scale=0.5]{results/SegErrPreCal.eps}
+	\caption{preliminary calculation effect}
+	\label{fig:SegErrPreCal}
 \end{figure}
 
-\begin{figure*}
+\begin{figure}
 	\centering
-	 \includegraphics[width=18cm]{images/swarmTesting2}
-	 	\caption{Swarm testing} Segmentation Error Vs. PSO parameters.
-	 	\label{fig:swarmtestingpso}
-	 	\end{figure*}
+		\includegraphics[scale=0.5]{results/PossibledpPreCalculation.eps}
+	\caption{Possible Dominant Point Count}
+	\label{fig:PossibledpPreCalculation}
+\end{figure}
 
- The effect of number of agents, the error thresholds and various other paramters was investigated.
-The result of these tests are shown in Figure \ref{fig:swarmtesting} and  \ref{fig:swarmtesting2}.
-Figure \ref{fig:swarmtesting} shows that as the number of agents increases the \textit{Segmenation Error} value decrease (error of the segmentation algorithm).
-Figure \ref{fig:swarmtesting1} shows simillar result for the maximum swarm iterations.
+ The effect of number of agents, the error thresholds and various other parameters was investigated.
+The result of these tests are shown in Figure \ref{fig:swarmtestingS1} and  \ref{fig:swarmtestingS2}.
+Figure \ref{fig:swarmtesting} shows that as the number of agents increases the \textit{Segmentation Error} value decrease.
+Similar results for the maximum swarm iterations are shown in Figure \ref{fig:swarmtesting1} .
 Other algorithm parameters like $c_1$,$c_2$ that were mentioned in Section\ref{sec:ParticleSwarmAlgorithm} are tested using similar tests (Figure \ref{fig:swarmtesting} and \ref{fig:swarmtesting2}).
-The final values are for maximum number of iteration = 60 with 15 particles.
+The final values are for maximum number of iteration = 80 with 15 particles.
 The final number of agents and maximum swarm iteration used in the system was based trade off between error achieved versus the computation time.
@@
 \label{sec:SegmentationAlgorithms}
-The result of the segmentation algorithm can be viewed in the Figure \ref{fig:results1.JPG}. These result shows the originals stroke and the segmentation that the system generated for the stroke. 
-\begin{figure}
-	\centering
-		\includegraphics{images/results1.JPG.eps}
-	\caption{Results Segmentations}
-	\label{fig:results1.JPG}
-\end{figure}
-\begin{figure}
-	\centering
-		\includegraphics{images/results2.JPG.eps}
-	\caption{Results Segmentaion }
-	\label{fig:results2.JPG}
-\end{figure}
-\begin{figure}
-	\centering
-		\includegraphics{images/results3.JPG.eps}
-	\caption{Results }
-	\label{fig:results3.JPG}
-\end{figure}
-\begin{figure}
-	\centering
-		\includegraphics{images/results4.JPG.eps}
-	\caption{Resluts of the segmentations}
-	\label{fig:results4.JPG}
-\end{figure}
-
-\subsection {Recognition Algorithms}
-\label{sec:RecognitionAlgorithms}%}
-To test the recognition different set of features are used to select the features that will define symbols better to finally achieve better recognition rate.
-We tested the system using different types of features.
-The features explained in section \ref{sec:FeatureExtraction} was divided into different sets; statistical, structural, Rubine features and zernik moments along with composite features.
-The fig.
-\ref{fig:testFeatures} shows the result of each set of features when used with the SVM classifier.
-The results shows that the Rubine features although may achieve good results in single stroke symbols, have a bad performance when used with multi stroke symbols.
-\begin{figure}[]
-	\centering
-		\includegraphics[scale=0.3]{images/test4.eps}
-	\caption{The results of Features experiment}
-	\label{fig:testFeatures}
-\end{figure}
-
-We test the system using different classifier to see what the effect of the classification step on the final recognition.
-We also wanted to test the train and test time in both classifiers.
-We used SVM (RBF) classifiers and the Gaussian Linear classifier used in \cite{}.
-The fig.
-\ref{} shows the result of both classifiers with the main three segmentation algorithms.
-The results shows that the SVM classifier achieve better results.
-To measure the scalability of the system\footnote{What happens when adding a new symbols to the set of known symbols.} the classifiers are test using different number of symbols (fig.
-\ref{fig:sampleperCat}).
-Figures \ref{fig:sampleperCat} demonstrate the scalability of both classifiers by demonstrating the recognition rate in both classifiers with different number of symbols.
-Figures \ref{fig:sampleperCat},\ref{fig:sampleperCat} shows the time and number of symbols needed for training versus each classifier.
+The result of the segmentation algorithm can be viewed in the Figure \ref{fig:results1}.
+These result shows the originals stroke and the segmentation that the system generated for the stroke.
+The figure \ref{fig:results2} shows the result of \textsl{AlgS1} and \textsl{AlgS2} for the same input strokes.
+Figure \ref{fig:results3} shows the output of \textsl{Alg3} to the same set of input strokes.
+It clearly shows that \textsl{AlgS2} have better segmentation results than both \textsl{AlgS1} and \textsl{Alg3} but since these segmentation are subjective to what the user intended to draw the next section will focus on the effect of the segmentation algorithm on the recognition accuracy.
 
 \begin{figure}
+
 	\centering
-		\includegraphics{images/sampleperCat.eps}
-	\caption{Samples per Category }
-	\label{fig:sampleperCat}
+				\subfigure{\includegraphics[scale=0.55]{results/result1.eps}}
+					\subfigure{	\includegraphics[scale=0.6]{results/result1logic.eps}}
+	\caption{Segmentation Results of \textsl{AlgS1}}
+	\label{fig:results1}
+
 \end{figure}
@@
 	\centering
-		\includegraphics[scale=0.7]{images/test4.eps}
-	\caption{Tests}
-	\label{fig:test4}
+			\subfigure{	\includegraphics[scale=0.55]{results/result2.eps}}
+			\includegraphics[scale=0.6]{results/result2logic.eps}
+	\caption{Segmentation Results of \textsl{AlgS2}}
+	\label{fig:results2}
+\end{figure}
+\begin{figure}
+	\centering
+		\subfigure{
+		\includegraphics[scale=0.55]{results/result3.eps}}
+			\subfigure{\includegraphics[scale=0.55]{results/result3logic.eps}}
+
+	\caption{Segmentation Results of \textsl{Alg3} }
+	\label{fig:results3}
 \end{figure}
 
-\subsubsection {Effect of sysmbol complixity}
+\subsection {Recognition Algorithms}
+\label{sec:RecognitionAlgorithms}
 
-The experiment we implemented was to test the effect of symbol complexity and type on the recognition rate.
-Figure \ref{fig:test2} shows what the accuracy of each symbol, it clearly noted that symbols that have only line segments achieve higher accuracy rate than other symbols.
-Figure \ref{fig:test2} also shows that algorithm 1 alone achieve better performance than algorithm 2 in the symbols that consist of lines only.
-The combining of both algorithm improves the recognition rate of other symbols.
-\\
-\begin{figure*}[]
+e performed several experiments to evaluate the presented recognition system.
+Firstly we tested recognition accuracy of shapes in the data set with both algorithms (\textsl{AlgS1} and \textsl{AlgS2}).
+We also implemented the segmentation algorithm described in \cite{earlyprocess} (\textsl{Alg3}) to use as reference to compare it with our swarm algorithms.
+Figure \ref{fig:test1} shows the accuracy achieved by each algorithm.
+The two swarm algorithms were tested with and without the ellipse fitting module.
+The ellipse detection module appears to be superior.
+The result shows that combining both  (\textsl{AlgS1} and \textsl{AlgS2}) out perform any single algorithm.
+%show that combining algoirthmsachieve higher accuracy than any single algoirthm.
+%improves the results.% with both the \textit{DPSO} algorithms.% The results show that both \textit{DPSO} algorithms achieve better result than other algorithms.
+
+ \begin{figure}
 	\centering		
-		\includegraphics[scale=0.5]{images/test2.eps}
-	\caption{The results of second experiment}
+	 \includegraphics[scale=0.4]{results/testAlg.eps}
+	 	\caption{Algorithm comparison} The recognition rate of different algorithms. 
+	 	\label{fig:test1}
+	%\caption{Experiments results :} a)    b) c) 
+\end{figure}
+
+\subsubsection {Effects of Symbol Complexity }
+ \label{sec:EffectsofSymbolComplixity}
+
+\begin{figure*}
+	\centering
+		\includegraphics[scale=0.5]{results/testsym.eps}
+	\caption{Symbols Comparison} The effect of symbol complexity.  %The graph shows the recognition rate of each symbol using different algorithms. 
 	\label{fig:test2}
 \end{figure*}  
- Another experiment we performed was to test the efficiency svm classifier.
-We implemented the linear discriminator classifier described in \cite{gestureexample12}(see figure \ref{fig:test3}).
-%The result shows that SVM is more scalable than the linear discriminator for this problem.
-%The test focus on the size of data needed to test each classifier and the achieved accuracy.
-%For this experiment we also shows the classification time for both classifiers
-\begin{figure}[]
+The second experiment we implemented was to investigate the effect of symbol complexity and type on the recognition rate.
+Figure \ref{fig:test2} shows the achieved accuracy of each symbol by our algorithms compared to \cite{earlyprocess} (\textsl{Alg3}).
+It is clearly noted that symbols that have only line segments achieve higher accuracy rate than other symbols.
+The results indicate that algorithm \textsl{AlgS1} achieve better performance than algorithm \textsl{AlgS2} in the symbols that consist only of lines.
+This is understandable because algorithm \textsl{AlgS1} divides strokes into line segments only but \textsl{AlgS2} is able to divide strokes into lines and curves based on the minimum error of the segment itself.
+Algorithm \textsl{Alg3} gives good performance as long as the symbols consist of combination of lines and curves, if the stroke consists of only curve or lines the algorithm may lead to wrong segmentation result.
+This is because the system divides the stroke first to line segments then tries to decide if each segment can be represent better as a curve unlike algorithm \textsl{AlgS2} where the curve segments are tested while choosing the best segmentation.
+Combining both algorithm \textsl{AlgS1} and \textsl{AlgS2} improved the recognition rate of all symbols.
+The penalty for this improved performance is the computational time required to run both swarm algorithms.
+Table \ref{tab:ConfusionMatrix} shows the confusion matrix of symbols.
+The table shows that errors are only between two sets of symbols moon with triangles and trapezoid with triangles.
+This observation is understandable as the symbols are visually similar but the system must be able to differentiate between them.
+This lead us to the next experiment to choose best set of feature to recognize symbols.
+ \begin{table*}
 	\centering
-		\includegraphics[scale=0.4]{images/test1.eps}
-	\caption{The results of third experiment}
-	\label{fig:test3}
-\end{figure}
- %Details description of results got from comparing various algorithms  
- %Describe data set used for experiments ( No of data set, No. of categories, No of samples).
- %Describe other Algorithms [Alg 3 in paper \cite{earlyprocess}]used for comparing with the one described in the paper.
+	%\small
+%	\scalebox{0.7}{
+		%	\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|c|c|c|c|}\hline 
+	\scalebox{0.7}{
+		\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|c|c|c|c|}\hline 
+ Categories 	&ellipse	&heart	&trapezoid	&pentagon	&arch	&hexagon	&square	&triangle	&cube	&cylinder	&parallelogram	&moon	&callout\\ \hline
+ellipse	&155	&9	&0	&0	&0	&5	&0	&0	&0	&0	&0	&6	&0\\ \hline
+heart	&0	&109	&1	&1	&25	&0	&0	&1	&0	&0	&2	&1	&1\\ \hline
+trapezoid	&0	&0	&103	&5	&2	&0	&0	&2	&0	&0	&34	&0	&0\\ \hline
+pentagon	&0	&0	&20	&121	&0	&20	&6	&2	&0	&0	&1	&0	&0\\ \hline
+arch	&0	&1	&13	&0	&86	&0	&0	&2	&3	&0	&2	&0	&0\\ \hline
+hexagon	&0	&0	&0	&0	&0	&98	&0	&0	&0	&0	&0	&0	&0\\ \hline
+square	&0	&0	&1	&0	&2	&0	&107	&14	&0	&0	&8	&1	&0\\ \hline
+triangle	&0	&0	&7	&0	&1	&0	&0	&89	&0	&0	&12	&4	&0\\ \hline
+cube	&0	&0	&0	&0	&0	&0	&0	&0	&147	&0	&0	&0	&0\\ \hline
+cylinder	&0	&0	&0	&6	&2	&8	&16	&0	&4	&156	&1	&10	&0\\ \hline
+parallelogram	&0	&0	&4	&0	&0	&0	&0	&0	&0	&0	&91	&0	&0\\ \hline
+moon	&0	&3	&3	&2	&17	&0	&9	&17	&0	&0	&2	&133	&11\\ \hline
+callout	&1	&34	&0	&22	&21	&22	&16	&25	&0	&0	&0	&0	&143\\ \hline
+		\end{tabular}
+			}
+ 		%\end{tabular}
+ 	%}
+	\caption{Confusion Matrix}
+	\label{tab:ConfusionMatrix}
+	%\end{minipage}
+\end{table*}
 
- %In the figure \ref{fig:test1}, the accuracy means that final hit rate of all symbols used in the test set.by the  four datasets . 
-We performed three experiments to test the system firstly we tested recognition accuracy of shapes in the data set with both algorithms.
-We also implemented the segmentation algorithm described in \cite{earlyprocess} to use as reference to our swarm algorithms.
-As you see in the fig.
-\ref{fig:test1} shows the accuracy achieved by each algorithm.
-The results shows that both PSO algorithm achieve better result than other algorithms.
-The swarm algorithms were tested with and without the ellipse detection module.
-The ellipse detection module appear to be superior to results with only the PSO algorithms.
-\\
-\begin{figure}[]
+\subsubsection{Features Comparisons}
+\label{sec:featuresComparisions}
+Different feature sets are tested to determine the best features that can used in sketch and symbol recognition.
+Figure\ref{fig:testFeaturesAllHS} shows the result of different feature sets from the basic four sets (\textbf{FS1,FS2,FS3,FS4}).
+Result shows that (\textbf{FS2}) Rubine features \cite{gestureexample12} gives the worst results when used alone without any other features.
+This is because they are mainly computed for single stroke gestures and fare bad in multi-stroke symbols \cite{compareFeaturSVM}.
+Features \textbf{(FS4)} gives good results but it is improved by adding structural features \textbf{(FS1)}.
+Figure \ref{fig:testFeaturesAllMD} shows the similar results on \textsl{MD-DB} multi domain database.
+The figure shows that on this dataset \textbf{FS1} perform better than on \textsl{HS-DB}, this is a result that shapes in the \textsl{MD-DB} has more structural difference than shapes in \textsl{HS-DB}.
+ \begin{figure}
 	\centering
-		\includegraphics[scale=0.3]{images/test1.eps}
-	\caption{The results of first experiment}
-	\label{fig:test1}
+		\includegraphics[scale=0.5]{results/testFeat.eps}
+	\caption{Feature Comparison} The effect of different features on accuracy on \textsl{Hs-DB}.  %The graph shows the recognition rate of each symbol using different algorithms. 
+	\label{fig:testFeaturesAllHS}
 \end{figure}
 
-%\subsection{Clustering Algorithm Results}
-%\label{sec:ClusteringAlgorithmResults}
-
-\subsubsection{Features Comparisions}
-\label{sec:featuresComparisions}
+ \begin{figure}
+	\centering
+		\includegraphics[scale=0.5]{results/testF1.eps}
+	\caption{Feature Comparison} The effect of different features on accuracy on \textsl{MD-DB}.  %The graph shows the recognition rate of each symbol using different algorithms. 
+	\label{fig:testFeaturesAllMD}
+\end{figure}
 

### 2009-04-12 draft/System.tex
after correctly splled now look at any thing else

@@
  In this research, time difference, direction, speed and curvature of each point along the stroke was computed.
-The experiments in section \ref{} proved that computing all information's provide better segmentations results than only using one of the graphs.
+The experiments in section \ref{sec:PSO} proved that computing all information's provide better segmentations results than only using one of the graphs.
 The time difference is calculated as $\Delta t = t_{i+1} - t_i$ where $t_i$ is the time of the point $i$ and $t_{i+1}$ is the time of the point $i+1$.
-The speed is calcuated using equation $v=\Delta s/\Delta t$ where $t$ is the time difference between two points and $s$ is the length between them.
+The speed is calculated using equation $v=\Delta s/\Delta t$ where $t$ is the time difference between two points and $s$ is the length between them.
 
@@
 The stroke points $p_i(x,y)$ that correspond to those maximum values are labeled as \textit{possible dominant points} $P_{pd}$.
-Figure \ref{fig:MaxRegioi} shows an example of $Region_i$ of the speed curve and the dominate points $P_{pd}$ which correspond to the minimum of point in this region $Min(Region_i)$.
+Figure \ref{fig:MaxRegioi} shows an example of $Region_i$ of the speed curve and thedominantpoints $P_{pd}$ which correspond to the minimum of point in this region $Min(Region_i)$.
 % (as shown are redundant)
@@
 The array that contains possible points $P_{pd}$ is used as input for the next segmentation step.
-Figure \ref{fig:LabelsPPD} shows the particles labeled as Possible dominate points $P_{pd}$ by the preprocessing, it is noticed the redundancy of some $P_{pd}$ points.
+Figure \ref{fig:LabelsPPD} shows the particles labeled as Possible dominant points $P_{pd}$ by the preprocessing, it is noticed the redundancy of some $P_{pd}$ points.
 
@@
 		\includegraphics{images/ppd.eps}
-	\caption{Posible domninate points}% a) Possible dominate point b) Particle encoding  
+	\caption{Possible dominant points}% a) Possibledominantpoint b) Particle encoding  
 	\label{fig:LabelsPPD}
@@
 Thus , the particle represents points that are chosen to be used as dominant points for this segmentation.
-The goal of the DPSO algorithm is to find the solution $P_i$ that generates the minimum set of dominate points that defines the stroke with minimum segmentation error.
+The goal of the DPSO algorithm is to find the solution $P_i$ that generates the minimum set ofdominantpoints that defines the stroke with minimum segmentation error.
 In other words, the system tries to find the fewer number of points (1's) location in the particle array which gives minimum segmentation error.
@@
   \includegraphics[scale=0.8]{images/pso1.eps}			
-	\caption{ALS1 Error}% a) Possible dominate point b) Particle encoding  
+	\caption{AlgS1 Error}% a) Possibledominantpoint b) Particle encoding  
 	\label{fig:DPSOERROR}
@@
 
-\paragraph{Determinig The Segment Type.} 
+\paragraph{Determining The Segment Type.} 
 \label{sec:DeterminigSegmentType}
@@
 If it was not labeled the point $P_{ij}$ is moved to the nearest labeled point.
-This ensures that all of the points generated by the DPSO are possible dominate points $P_{pd}$.
-After that the particles are tested to make sure that the distance between every two successive dominate point is larger than the constant $min_D$.
+This ensures that all of the points generated by the DPSO are possible dominant points $P_{pd}$.
+After that the particles are tested to make sure that the distance between every two successive dominant point is larger than the constant $min_D$.
 If two points are nearer than $min_D$\footnote{Constant selected to be 10 pixels} then one of the points is removed.
@@
  \end{equation}
- where $A_{st}$ is the area of the input stroke, $A_{sg}$ is area of computed segmenation, $E_{alg}$ is the Segmentation Error.
-The segmentation which crosspond to the minimum $E_{tot}$ is the best segmentation.
+ where $A_{st}$ is the area of the input stroke, $A_{sg}$ is area of computed segmentation, $E_{alg}$ is the Segmentation Error.
+The segmentation which cross ponds to the minimum $E_{tot}$ is the best segmentation.
 
-\section{Hybird Segmentation Algorithm \textsl{Alg3}}
+\section{Hybrid Segmentation Algorithm \textsl{Alg3}}
 \label{sec:BenchMarckAlgorithm}
@@
 	 \item[Segments] Number of segments in the symbol.
-	 \item [Strokes] Number of strokes or partical strokes that created the symbols.  
+	 \item [Strokes] Number of strokes or partial strokes that created the symbols.  
 		\item[Primitives] Number of primitives in the symbol. The feature helps when identifying             symbols with mixed geometric primitives like cylinders and callouts.  
@@
 \label{sec:Classification}
- The system uses support vector machine (SVM) classifier with Gaussian kernel (RBF kernel) \cite{libsvm}.
+ The system uses support vector machine (SVM) classifier with Linear kernel (Linear kernel) \cite{libsvm}.
 The system uses OVO classifiers structure (object versus object).
-A validation algorithm is used to choose the Kernel variables parameters $c, \gamma$.
+A validation algorithm is used to choose the Kernel variables parameters $c$ The results Section \ref{sec:recognitionAlgorithms} demonstrates the results of both classifiers.
+\\%Some features like number of parallel, perpendicular and intersecting lines are computed from the segmentation other like density, centroid and Zernike moments are computed from the raw stroke points.
+.
 The Rubine algorithm (Linear Gaussian classifier) was implemented for comparing the system.
-The results Section \ref{sec:recognitionAlgorithms} demonstrates the results of both classifiers.
-\\%Some features like number of parallel, perpendicular and intersecting lines are computed from the segmentation other like density, centroid and Zernike moments are computed from the raw stroke points.
 

### 2009-04-12 draft/ChapterIntroduction.tex
after correctly splled now look at any thing else

@@
 There are several problems in this research area that need to be handled to implement an integrated sketch recognition system.
-The fact that the user draws strokes in a spontaneous and ambiguous manner that intensifies the challenges.
+The fact that the user draws strokes in a spontaneous and ambiguous manner intensifies the challenges.
 One of the main problems is segmenting spontaneous strokes drawn by the user into geometrical primitive segments.
@@
 This thesis is arranged as follows: chapter \ref{sec:Survey} presents a survey of different components of the sketch recognition.
-Chapter \ref{sec:ParticleSwramOptimization} gives a review on swarm intelligance and particle swarm optimization.
+Chapter \ref{sec:ParticleSwramOptimization} gives a review on swarm intelligence and particle swarm optimization.
 Chapter \ref{sec:proposedSystem} describes the proposed system and its main building blocks.
@@
 Finally, chapter \ref{sec:DiscussionConclusion} presents the conclusions of this thesis and suggested future work.
-%This thesis is arranged such that chapter \ref{} presents a survey of different components of the sketch recognition.
-Chapter \ref{} describe the proposed system and its main building blocks.
-The details description of each component is presented in Chapter \ref{}.
-The comparative study of different techniques with the detailed results of various experiments are detailed in chapter\ref{}.
-Finally, chapter \ref{} presents the conclusions of this thesis and suggested future work.

### 2009-04-12 draft/ConFut.tex
now it 9 - appril and after club day
after grammer correct now will print to review

@@
 \chapter{Discussion and Conclusion}
+\label{sec:DiscussionConclusion}
 
 \section{Conclusion}
-\label{sec:ConclusionConclusion }
+\label{sec:ConclusionConclusion}
 
@@
 The use of both speed and curvature data helps in improvement of the PSO algorithm over the original algorithms \cite{CruveDivisionSwarm,PolygonApproximationPSO}.
-The trade off between accuracy achieved and time complexity must be further investigated to achieve better results.
+The tradeoff between accuracy achieved and time complexity must be further investigated to achieve better results.
 The system introduced an efficient method to sketch recognition using the PSO.
@@
 The number and the complexity of symbols recognized by the system varied from simple to complex but did not affect the final recognition of the system.
-The only draw back of the system that it takes more time with comparison with similar algorithms.
+One drawback of the system is that it takes more time to segment strokes in comparison with other similar algorithms.
+Another drawback is that until now we only applied the system to single symbols.
+Full free hand sketch test has not been applied on the system.
+%sketched
 
@@
 This graph will be matched with the set of known symbol template to make the identification.
-Those result along with the SVM classifier results will enhance the system as a whole and achieve better recognition rate.
+Those results along with the SVM classifier results will enhance the system as a whole and achieve better recognition rate.
 %Each new stroke is checked if it can be a symbol or is a part of already drawn un completed symbol.

### 2009-04-12 draft/Results.tex
after grammer correct now will print to review

@@
 Symbols from different domains were gathered to evaluate the system throughout different types of shapes.
-Users were requested to draw symbols without any restriction on style,size or orientations.
+Users were requested to draw symbols without any restriction on style, size or orientations.
 The dataset was divided into training set and test set.
@@
 It is worth noting that when using all information the number of Possible dominant point $P_{pd}$ is not the same as the sum of using every information alone.
-This is because we found out that there was some redundant points found using more than one information.
+This is because we found out that there was some redundant points found using more than one information source.
 
@@
 The result of these tests are shown in Figure \ref{fig:swarmtestingS1} and \ref{fig:swarmtestingS2}.
-Figure \ref{fig:swarmtesting} shows that as the number of agents increases the \textit{Segmentation Error} value decrease.
-Similar results for the maximum swarm iterations are shown in Figure \ref{fig:swarmtesting1} .
+Figure \ref{fig:swarmtesting} shows that as the number of agents increase the \textit{Segmentation Error} value decrease.
+Similar results for the maximum swarm iterations are shown in Figure \ref{fig:swarmtesting1}.
 Other algorithm parameters like $c_1$,$c_2$ that were mentioned in Section\ref{sec:ParticleSwarmAlgorithm} are tested using similar tests (Figure \ref{fig:swarmtesting} and \ref{fig:swarmtesting2}).
 The final values are for maximum number of iteration = 80 with 15 particles.
-The final number of agents and maximum swarm iteration used in the system was based trade off between error achieved versus the computation time.
+The final number of agents and maximum swarm iteration used in the system was based trade off between errors achieved versus the computation time.
 
@@
 %The fig.
-\ref{fig:dataerrorvsiteration.jpg} display the effect of the size of swarm population on the number of vertex reported while segmenting the stroke.\footnote{The fewer the vertex the better the segmentation} The graphs shows that as the number of iteration increase there is a decrease in error until the error reach a saturation and any increase in number of iteration dose not affect the error calculated.
+\ref{fig:dataerrorvsiteration.jpg} display the effect of the size of swarm population on the number of vertex reported while segmenting the stroke.\footnote{The fewer the vertex the better the segmentation} The graphs shows that as the number of iteration increase there is a decrease in error until the error reach a saturation and any increase in number of iteration does not affect the error calculated.
 Similar behavior is noticed in fig.
@@
 The result of the segmentation algorithm can be viewed in the Figure \ref{fig:results1}.
-These result shows the originals stroke and the segmentation that the system generated for the stroke.
-The figure \ref{fig:results2} shows the result of \textsl{AlgS1} and \textsl{AlgS2} for the same input strokes.
+This result shows the originals stroke and the segmentation that the system generated for the stroke.
+Figure \ref{fig:results2} shows the result of \textsl{AlgS1} and \textsl{AlgS2} for the same input strokes.
 Figure \ref{fig:results3} shows the output of \textsl{Alg3} to the same set of input strokes.
@@
 
-e performed several experiments to evaluate the presented recognition system.
+We performed several experiments to evaluate the presented recognition system.
 Firstly we tested recognition accuracy of shapes in the data set with both algorithms (\textsl{AlgS1} and \textsl{AlgS2}).
@@
 The result illustrated in this chapter shows the work done on the system and experiments performed.
-The results shows that the ellipse detection and PSO swarm algorithms gives better performance than other known systems.
+The results show that the ellipse detection and PSO swarm algorithms gives better performance than other known systems.
 

### 2009-04-12 trunk/draft/System.tex
after grammer correct now will print to review

@@
 
- In \textbf{Feature Extraction} step, a feature vector consists of a set of statistical, geometrical and spatial features is extracted from the segmented strokes.
+ In \textbf{Feature Extraction} step, a feature vector consists of a set of statistical, geometrical and spatial features are extracted from the segmented strokes.
 The feature vector is used as input for the next step.
@@
 The stroke points $p_i(x,y)$ that correspond to those maximum values are labeled as \textit{possible dominant points} $P_{pd}$.
-Figure \ref{fig:MaxRegioi} shows an example of $Region_i$ of the speed curve and the dominate points $P_{pd}$ which correspond to the minimum of point in this region $Min(Region_i)$.
+Figure \ref{fig:MaxRegioi} shows an example of $Region_i$ of the speed curve and the dominant points $P_{pd}$ which correspond to the minimum of point in this region $Min(Region_i)$.
 % (as shown are redundant)
@@
 		\includegraphics{images/ppd.eps}
-	\caption{Posible domninate points}% a) Possible dominate point b) Particle encoding  
+	\caption{Possible dominant points}% a) Possible dominant point b) Particle encoding  
 	\label{fig:LabelsPPD}
@@
 Thus , the particle represents points that are chosen to be used as dominant points for this segmentation.
-The goal of the DPSO algorithm is to find the solution $P_i$ that generates the minimum set of dominate points that defines the stroke with minimum segmentation error.
+The goal of the DPSO algorithm is to find the solution $P_i$ that generates the minimum set of dominant points that defines the stroke with minimum segmentation error.
 In other words, the system tries to find the fewer number of points (1's) location in the particle array which gives minimum segmentation error.
@@
   \includegraphics[scale=0.8]{images/pso1.eps}			
-	\caption{ALS1 Error}% a) Possible dominate point b) Particle encoding  
+	\caption{AlgS1 Error}% a) Possible dominant point b) Particle encoding  
 	\label{fig:DPSOERROR}
@@
 
-\subsubsection{Solution Refine procedures } 
+\subsubsection{Solution Refine procedures} 
 %\cite{PolygonApproximationPSO} used a merge and divide algorithm after each loop of the swarm system to refine the solution but we used another enhancement method.
@@
  \end{equation}
- where $A_{st}$ is the area of the input stroke, $A_{sg}$ is area of computed segmenation, $E_{alg}$ is the approximation error of the segmentation.
-The segmentation which crosspond to the minimum $E_{tot}$ is the best segmentation.
+ where $A_{st}$ is the area of the input stroke, $A_{sg}$ is area of computed segmentation, $E_{alg}$ is the Segmentation Error.
+The segmentation which cross ponds to the minimum $E_{tot}$ is the best segmentation.
 
-\section{Hybird Segmentation Algorithm}
+\section{Hybrid Segmentation Algorithm \textsl{Alg3}}
 \label{sec:BenchMarckAlgorithm}
@@
 
-%\section{Stroke Clustring Algoirthm }
+%\section{Stroke Clustring Algorithms }
 %\label{sec:ClustringAlgoirthm}
@@
 \begin{description}
- \item [Structural and geometrical Features(FS1)] Features defines the structure of the geometrical symbol.  
+ \item [Structural and geometrical Features (FS1)] Features defines the structure of the geometrical symbol.  
 \begin{description}
 	 \item[Segments] Number of segments in the symbol.
-	 \item [Strokes] Number of strokes or partical strokes that created the symbols.  
+	 \item [Strokes] Number of strokes or partial strokes that created the symbols.  
 		\item[Primitives] Number of primitives in the symbol. The feature helps when identifying             symbols with mixed geometric primitives like cylinders and callouts.  
@@
 	\item Sine of starting angle.
-	\item Length of diagonal of bounding box.It gives an idea of the size of the bounding box).
+	\item Length of diagonal of bounding box. It gives an idea of the size of the bounding box).
 	\item Angle of diagonal. It gives an idea of the shape of the bounding box (long, tall, square).

### 2009-04-12 trunk/draft/chapterSurvey.tex
after grammer correct now will print to review

@@
 
-% i need to write a history then small introudction to the survey...
+% i need to write a history then small introduction to the survey...
 %  i have to state that the next  sections will generally group segmentation and symbol recognition task. 
@@
 \label{sec:preprocessing}
-%this section should describe how others did the first preprocessing of the problem. i have to summarize what i wrote before. 
- % first paragrah about how stroke is a messing thing and naturally it is ambigues. 
   In a system that supports free hand sketching, sketching should be as natural as possible.
@@
 In our system, we combine Rubine features with others to include the information on how the user draws sketch while recognizing the symbols.
-In \cite{Long00}  Long shows how similarities between gestures can be analyzed.
+In \cite{Long00}, Long shows how similarities between gestures can be analyzed.
 He used this information to identify gestures that are likely to confuse the recognition system.
@@
  Avoiding graph matching problems, research tried to use languages that describe shapes instead of graphs.
-Visual descriptive languages represents the shape into tokens of geometric primitives or relations between smaller  geometric primitives.
+Visual descriptive languages represent the shape into tokens of geometric primitives or relations between smaller geometric primitives.
 Figures \ref{fig:arrow1} and \ref{fig:arr2} show a regular arrow with 2 shafts and head and its definition in Ladder language \cite{Ladder30}.
@@
 These systems are less computational expensive than graph matching approach but they also operate based on the output of the segmentation process.
-Any error in the segmentation process lead to huge error while parsing the whole sketch.
-Moreover, symbols cannot share strokes between them which influence and restrict how the user draw symbols.
+Any error in the segmentation process leads to huge error while parsing the whole sketch.
+Moreover, symbols cannot share strokes between them which influence and restrict how the user draws symbols.
  % complexity of the graph matching problem 
@@
 
-drawing back from properties of individual strokes several researcher tried to use shape properties on the whole shape to classify shapes.
-They used global features that tries to define the entire shape properties.
+Drawing back from properties of individual strokes several researcher tried to use shape properties on the whole shape to classify shapes.
+They used global features that try to define the entire shape properties.
 
  Several systems \cite{DiagramOfflineConvexHull,MulitStrokeConvexHull} presented features like ratio of bounding box to convex hull area and the ratio of the perimeter to the area.
-These global feature can be used to distinguish between classes.
-For example, a rectangle can be distinguish from a triangle by looking to the ratio of area of bounding box to the area of the convex hull.
+These global features can be used to distinguish between classes.
+For example, a rectangle can be distinguished from a triangle by looking to the ratio of area of bounding box to the area of the convex hull.
 The triangle ratio will be much lower than that of the rectangle.
@@
 
- Systems based on Global shape properties distinguish shapes by learning the different properties each shape class tend to have.
-These properties dose not depend on the order or number of strokes and thus are only based on appearance of the shape not how the user draw it.
+ Systems based on Global shape properties distinguish shapes by learning the different properties each shape class tends to have.
+These properties do not depend on the order or number of strokes and thus are only based on appearance of the shape not how the user draws it.
 However, individual details of the shape are impossible to distinguish using these properties.
-For example, it is impossible to use  to  differentiate between a current source symbol (circle containing and arrow Figure \ref{fig:CurrentSource}) and a Ac Source (circle containing a tilate Figure \ref{fig:acSource}).
-Finer details of the symbols are not represented in this approach, thus shapes that have smalls details or only significant conceptual variation are not distinguished using this method.
-These symbols requires a system to represent different levels of shape details.
+For example, it is impossible to use to differentiate between a current source symbol (circle containing and arrow Figure \ref{fig:CurrentSource}) and a Ac Source (circle containing a wave Figure \ref{fig:acSource}).
+Finer details of the symbols are not represented in this approach, thus shapes that have small's details or only significant conceptual variation are not distinguished using this method.
+These symbols require a system to represent different levels of shape details.
 
@@
 
-In the third approach, the recognition focus on appearance of shapes contrary to individual strokes or global shape properties.
+In the third approach, the recognition focuses on appearance of shapes contrary to individual strokes or global shape properties.
 Kara \cite{imagetrainable48} used this approach while presenting a system that matches the input shape into a one of the prototypes in the database.
 Figure \ref{fig:template} shows an example of prototypes of some of the shapes in the system.
-Input shapes are first pose and scale normalized then four image similarity measure are used to match input shapes to the database prototypes.
+Input shapes are first pose and scale normalized then four image similarity measures are used to match input shapes to the database prototypes.
 He used two measures (Hausdor distance) to measure similarity and two coefficients (Yule, Tanimoto) to measure dissimilarity.
-His approach allows the system to recognized a shape using only a single training sample as it is based on the shape prototype.
+His approach allows the system to recognize a shape using only a single training sample as it is based on the shape prototype.
 However, the shape prototype must represent all transformation and variation the input shape can have.
-Moreover, they restrict the image representation to 48x48 pixel due to computational reasons.
+Moreover, they restrict the image representation to 48x48 pixels due to computational reasons.
 This restriction leads to eliminating some of the fine details that may be needed to distinguish some shape classes.
-There results on isolated symbols seems promising but the system have not beet tested on full sketch that tends to have high overlapping between symbols.
+There results on isolated symbols seem promising but the system has not been tested on full sketch that tends to have high overlapping between symbols.
 
@@
 
-In all of these system only \cite{SketchRead2007} using stroke based recognition have been implemented to the fully creat a system that free draw sketch and understand them.
-This is because the emphasis on the interactive side of the sketch sytem which result in large difficulties in processing free hand sketches.
+In all of these system only \cite{SketchRead2007} using stroke based recognition have been implemented to the fully create a system that free draw sketch and understand them.
+This is because the emphasis on the interactive side of the sketch system which result in large difficulties in processing free hand sketches.
 Methods based on gesture and global stroke properties produce systems that are dependent on assumptions about how strokes can be grouped.
 These systems assume that users will finish drawing a shape before moving to the next one \cite{multistrokeEvaluation,Cali63,geometrydomain49}.
-This let them to temporally group strokes and then recognize them based on their strokes or appearance properties.
-\cite{AlvaradoDigital} argued that this assumption dose not hold in natural sketches.
+This let them to temporally group strokes and then recognizes them based on their strokes or appearance properties.
+\cite{AlvaradoDigital} argued that this assumption does not hold in natural sketches.
 
@@
 The principle problem in these systems is the computational complexity of the matching between shape description and geometric primitives produced in the segmentation process.
-Natural sketch are difficult to segment and group to create reliable fragment of geometeric primitives.
+Natural sketch are difficult to segment and group to create reliable fragment of geometric primitives.
 This is partially due to the increase of noise and other phenomena such as overlapped, over traced strokes and sharing of strokes between different symbols.
-Since these systems are dependable on the segmentation results, they either performe recognition using undependable segmentation results or consider different ways to segment the input sketch.
+Since these systems are dependable on the segmentation results, they either perform recognition using undependable segmentation results or consider different ways to segment the input sketch.
 
@@
 Each symbol is segmented and hierarchical features are computed.
-Features  that are represent the global description of the shape are also generated and appended to the hierarchical features.
+Features that are representing the global description of the shape are also generated and appended to the hierarchical features.
 The recognition is done using SVM Linear classifier as it proved to be one of the best learning algorithms.
@@
 Scientist in the field or artificial life need to understand how such "social animals" interact, achieve goals and evolve.
-Moreover, swarm behaviors that achieves some kind of swarm intelligence is increasingly interesting for engineers as it they can be applied in optimization.
+Moreover, a swarm behavior that achieves some kind of swarm intelligence is increasingly interesting for engineers as it they can be applied in optimization.
 Looking from a higher view, agents in a swarm are cooperating to achieve some purposeful behavior and achieve some common goal.
-For example, ant groups looking for food source.
-Large groups of simple independent agents seems to generate this behavior in what is called 'collective intelligence'.
+A good example of this cooperation is ant groups looking for food source.
+A large group of simple independent agents seems to generate this behavior in what is called 'collective intelligence'.
 Each agent has a set of simple rules to follow, and according to the interaction of the entire group the objective is met\cite{swarmGeneral}.
@@
 Reynolds \cite{swarmflock} created 'boid' which simulated the motion of a flock of birds.
-Each boid is a simple agent that navigate according to his perception and a set of basic rules.
-Kube and Zhange \cite{antsimulation} introduced a simulation model for ants behavior and used in cooperative transportation.
+Each boid is a simple agent that navigates according to his perception and a set of basic rules.
+Kube and Zhange \cite{antsimulation} introduced a simulation model for ant's behavior and used in cooperative transportation.
 Ants leaves a trail of its movement, so when an ant find food it cannot carry alone, it can easily trace back to the colony and then a group of ants can track the trail and bring the food back.
@@
 PSO algorithm was used in \cite{ToADD} to classify pixel to simplify the object segmentation process.
-
-ToADD
 

### 2009-04-14 draft/ChapterIntroduction.tex
i went to dr. magdah but she did not look at the thesis

@@
 
-\section{Importance of sketch recognition}
+\section{Importance of Sketch Recognition}
 \label{sec:ImportanceOfSketchRecognition}
@@
 
-\section{Problem Definition and challenges}
+\section{Problem Definition and Challenges}
 \label{sec:ProblemsAndChallenges}
@@
 		\includegraphics[scale=0.8]{../../neededfiles/Figures/strokeinterpolation.eps}
-	\caption[Segmentation error]{Different ways for segmenting a line drawn by user. a) The stroke users draw b)interpretation as a single line segment c) interpretation as an arc.}
+	\caption[Segmentation Error]{Different ways for segmenting a stroke drawn by user. a) The stroke users draw b)interpretation as two line segments c) interpretation as an arc.}
 	%The figure shows an arc that the user draw which can be segmented in different ways as a single line or as an arc where a) The stroke users draw  b)interpretation as a single line segment b) interpretation as an arc.}
@@
 \subsection{Ambiguity}
-Figure 1 shows an example of ambiguity, where the strokes drawn by the user can be classified in several ways.
+Figure \ref{fig:VorU} shows an example of ambiguity, where the strokes drawn by the user can be classified in several ways.
 The nature of hand drawn sketch itself is highly ambiguous even for human eye.
@@
 		\includegraphics[scale=0.85]{../../neededfiles/Figures/VorU.eps}
-	\caption[Ambiguities Handwriting] { The same letter can be interpreted differently in various contexts; first as V in drive or as U in Cut}
+	\caption[Handwriting Ambiguities] { The same letter can be interpreted differently in various contexts; first as V in drive or as U in Cut}
 \label{fig:VorU}
@@
 \section{Thesis Overview}
-This thesis is arranged as follows: chapter \ref{sec:Survey} presents a survey of different components of the sketch recognition.
+This thesis is arranged as follows: chapter \ref{sec:survey} presents a survey of different components of the sketch recognition.
 Chapter \ref{sec:ParticleSwramOptimization} gives a review on swarm intelligence and particle swarm optimization.

### 2009-04-14 draft/chapterSurvey.tex
i went to dr. magdah but she did not look at the thesis

--- 2009-04-11 draft/chapterSurvey.tex
@@
 The history of pen and sketch based computing goes back to 1963 with Ivan Sutherland's SKETCHPAD \cite{sutherlandsketch}, which used a light pen to draw on the monitor to create circuit diagrams.
-Rubine \cite{gestureexample12} Gesture recognition system.
-% i need to write a history then small introudction to the survey...
+Current technology and the increase in computing power have increased the abilities of such interfaces and allowed the implementation of fully sketching systems.
+Over the years, many other sketching based systems have been developed.
+Through this history most system focused on building the interactive interfaces to assist users while entering information to the computer.
+To demonstrate our system inspiration we will first present current research in the area of sketch recognition.
+In Section \ref{sec:preprocessing}, preprocessing and segmentation methods will be discussed in details.
+Later, symbol recognition methods will be presented and analyzed in Section \ref{sec:symbolrecogntion}.
+
+% i need to write a history then small introduction to the survey...
 %  i have to state that the next  sections will generally group segmentation and symbol recognition task. 
@@
 \label{sec:preprocessing}
-%this section should describe how others did the first preprocessing of the problem. i have to summarize what i wrote before. 
- % first paragrah about how stroke is a messing thing and naturally it is ambigues. 
   In a system that supports free hand sketching, sketching should be as natural as possible.
 This means that the user can draw symbols using single or multiple strokes without constrains.
-The goal of a sketch system is making the user feel that using the using the system is nearly the same as using pencil and paper.
-But sketchs are incomplete drawings that are sloppy and messy, for example figures \ref{fig:Overshootandundershoot} and \ref{fig:overstroked} show examples of messy sketches drawn by users.
+The goal of a sketch system is making the user feel that using the system is nearly the same as using a pencil and paper.
+But sketches are incomplete drawings that are sloppy and messy, for example figures \ref{fig:Overshootandundershoot} and \ref{fig:overstroked} show examples of messy sketches drawn by users.
 This led researchers to perform sort of preprocessing before starting the sketch recognition task itself.
@@
 		\includegraphics[scale=0.7]{../../neededfiles/Figures/Overshootandundershoot.eps}
-	\caption[Neat and Sloppy symbols] {A sloppy and neat versions of symbol.}
+	\caption[Neat and Sloppy Symbols] {A sloppy and neat versions of symbol.}
 	\label{fig:Overshootandundershoot}
@@
  Sezgin and Davis \cite{overtraced24} tried to solve the over traced strokes (see Figure \ref {fig:overstroked})using a thinning algorithm.
-In \cite{meanshift10,domainindependent17}, Bo Yu uses mean shift procedure as a noise remover and preprocessing step to the recognition step.
+In \cite{meanshift10,domainindependent17}, Yu uses mean shift procedure as a noise remover and preprocessing step to the recognition step.
 The mean shift procedure is simply shifting the point to the average of its neighborhoods so smoothing out the noise without changing the global structure of data.
@@
 These procedures may seem to remove noises in the sketch but it often treat vertices as noise which obscure vertex locations.
+\\
  %% now the third paragraph..... 
 Beside points location information (x, y positions) researcher made uses of the extra information from using digital ink.
-Some as \cite{mulitstroke5,polygonfeedback31} used user speed information to detect critical points where it is stated that user slow down at those critical points.
+Some as \cite{mulitstroke5,polygonfeedback31} used stroke speed information to detect critical points where it is stated that user slow down at those critical points.
 Other used curvature and directions data.
@@
  Agar et al.
-\cite{polygonfeedback31} used the timestamp for each point and compute time difference between every two successive point to help him finding corners as it provides more distinct maxima than the speed information.
+\cite{polygonfeedback31} used the timestamp for each point and compute time difference between every two successive point to help him find corners as he argued that it provides more distinct maximum than the speed information.
 He mentions that the pointing device (for example the mouse) sampling rate is the reason for this phenomena.
@@
 This leads to a nearly constant time difference between samples while the pen is moving and large difference while the pen is stationary.
-Contrary to speed information, where the users draw with variant speeds the speed information has a lot of noise in the data.
+Contrary to speed information, where the user draws with variant speeds which tends to make the speed information include a lot of noise in the data.
 
@@
 Our system uses speed, direction and time difference information in its preliminary calculation step.
+This ensures that all the possible vertices are considered when the segmentation algorithm segments strokes.
 %both speed and direction information help detecting vertices when the
+
+  % third should speech on how systems define critical points (this paragrah should speak on the data collected (speed crital points, ...
+
+ After critical point detection most sketch system proceeds to label segments into simple geometric primitives (i.e line, arc, etc...).
+A hybrid algorithm was introduced in \cite{earlyprocess} where different sets of segments are generated based on both curvature and speed dominant points, followed by choosing a segmentation with the least error from a generated hybrid set.
+However, this system is limited to recognizing only specific simple geometric shapes with a set of low level recognizers.
+Each low level recognizer is designed to recognize only one geometric shape using spatial and geometric information extracted from input stroke.
 
@@
 
-  % third should speech on how systems define critical points (this paragrah should speak on the data collected (speed crital points, ...
-
- After critical point detection most sketch system proceeds to label segments into simple geometric primitives (i.e line, arc, etc...).
-A hybrid algorithm was introduced in \cite{earlyprocess} where different sets of segments are generated based on both curvature and speed dominant points, followed by choosing a segmentation with the least error from a generated hybrid set.
-However, this system is limited to recognizing only specific simple geometric shapes with a set of low level recognizers.
-Each low level recognizer is designed to recognize only one geometric shape using spatial and geometric information extracted from input stroke.
 Yu \cite{meanshift10} introduced a \textit{feature area} for each primitive and then computed the segmentation error for different types of primitives based on the computed \textit{feature area}.
@@
 Chen et al.\cite{CruveDivisionSwarm} uses digital curves scanned from paper as input to the system and did not take advantage of the curvature or local geometric properties of the digital curve.
-\cite{Fluid25} Trying to combine fast morphing and basic recognition for the sketch drawn as the user draw it.
+
+Arvo \cite{Fluid25} Tried to combine fast morphing and basic recognition for the sketch drawn as the user draw it.
 It tries to move the points of strokes drawn by the user draw into one of the known shapes (ex: box, lines and circle).
@@
 				\includegraphics{images/featureareline.eps}}
-	\caption{Feature area}
+	\caption{Feature Area}
 	\label{fig:featurearearc}
@@
 
-\section{Symbol Recogntion}
+\section{Symbol Recognition}
 \label{sec:symbolrecogntion}
 
-In this section, we review three different types of symbol recognition methods.
+%In this section, we review three different types of symbol recognition methods.
  Symbol recognition can be based on: strokes, global shape properties, and appearance \cite{Oltmans07}.
@@
 Unlike appearance based methods where individual strokes are disregarded and they focus on the overall appearance of shapes.
-We choose to use the appearance based method but added some geometrical and structure properties used in most stroke based approaches.
+We choose to use the method based on shape property but added some geometrical and structure properties used in most stroke based approaches.
 Our method tries to avoid many of the problems faced by approaches based on individual strokes or global shape properties while trying to gain the additional information that the stroke based system use.
@@
 Simple features as length of stroke, smoothness of stroke, the angle that the first part of the gesture is drawn at and properties of the bounding box of the stroke are used to train the classifier to recognize symbols.
-In our system, we combine Rubine features with other to include the information on how the user draws sketch while recognizing the symbols.
-In \cite{Long00}  Long shows how similarities between gestures can be analyzed.
+In our system, we combine Rubine features with others to include the information on how the user draws sketch while recognizing the symbols.
+In \cite{Long00}, Long shows how similarities between gestures can be analyzed.
 He used this information to identify gestures that are likely to confuse the recognition system.
@@
 
-The main problem with gesture based recognition system that they impose several constrains on how the user can draw.
+The main problem with gesture based recognition systems that they impose several constrains on how the user can draw.
 Gestures must be drawn using a specific order and direction to be recognized correctly.
@@
 
- Systems that employ such representation must first preprocess strokes to break them into geometric primitives (see section \ref{sec:preprocessing} for a review on preprocessing methods).
-Recognition can then be treated as sub graph matching problem or constrain satisfaction problem.
+ Systems that employ such representation must first pre-process strokes to break them into geometric primitives (see section \ref{sec:preprocessing} for a review on preprocessing methods).
+Recognition can then be treated as sub graph matching problem.
 
@@
 		\includegraphics[scale=0.7]{../../neededfiles/Figures/squarescematic.eps}
-	\caption{square scemantic graph}
+	\caption{Semantic Graph of a Square }
 	\label{fig:squarescematic}
@@
 In the worst case, sub-graph matching is an exponential problem and thus is expensive to compute.
-Researcher tried to reduce complexity by restricting the search to fragments that are spatially or temporally grouped, or by using other assumption.
+Researcher tried to reduce complexity by restricting the search to fragments that are spatially or temporally grouped, or by using other assumptions.
 Some of the common assumptions are that each stroke is only part of one symbol and that user must draw one shape before drawing the next\cite{physicalmeaning6}.
@@
 These systems allow matching of partially satisfied shapes to propagate down to re interpret low level fragmentation hypothesis.
-In SKETCREAD \cite{SketchRead2007}, a dynamically generated Bayesian network fragments represents the shape hypotheses.
+In SKETCREAD \cite{SketchRead2007}, a dynamically generated Bayesian network represents the shape fragments hypotheses.
 Higher level structure of shape fragments can cause re interpretation of the geometric primitives.
@@
 
+ \paragraph{Visual Descriptive Languages:}
+ \label{sec:VisualDescribtiveLanguages}
+
+ Avoiding graph matching problems, research tried to use languages that describe shapes instead of graphs.
+Visual descriptive languages represent the shape with a set of  tokens representing geometric primitives or spatial relations between smaller geometric primitives.
+Figures \ref{fig:arrow1} and \ref{fig:arr2} show a regular arrow with 2 shafts and head and its definition in Ladder language \cite{Ladder30}.
+Shape recognition is then a question of parsing the sketch document into a recognizer that parse strokes according to the constructed grammar.
+
+ Several visual grammar were introduced \cite{statisticalparsing26,Ladder30,GenericHMM28}.
+\cite{Ladder30,GenericHMM28}  Used visual description language which will generate the information of the structure and constrain of each object classified.
+The system automatically generates Java code for the recognizer for each object.
+The generated code cycles through the strokes drawn to check if any of the recognizers generated had interpreted any symbols partially or completely.
+These systems are less computational expensive than graph matching approach but they also operate based on the output of the segmentation process.
+Any error in the segmentation process leads to huge error while parsing the whole sketch.
+Moreover, symbols cannot share strokes between them which influence and restrict how the user draws symbols.
+ % complexity of the graph matching problem 
+\begin{figure}
+	\centering
+		\subfigure[An Arrow]
+			{\label{fig:arrow1}\includegraphics[scale=0.75]{../../neededfiles/Figures/arrowladderdef1.eps}}
+			\hfill
+		\subfigure[Arrow grammar definition]
+		 {\label{fig:arr2}\includegraphics[scale=0.88]{../../neededfiles/Figures/arrowladderdef2.eps}}
+	\caption[Arrow Definition]{}
+		\label{fig:arrowladderdef1}
+\end{figure}
+
 \subsection{Global Shape Properties Based Recognition}
@@
 
-drawing back from properties of individual strokes several researcher tried to use shape properties on the whole shape to classify shapes.
-They used global features that tries to define the entire shape properties.
+Drawing back from properties of individual strokes several researcher tried to use shape properties on the whole shape to classify shapes.
+They used global features that try to define the entire shape properties.
 
- Several systems \cite{DiagramOfflineConvexHull,MulitStrokeConvexHull} presented features like ratio of bounding box to convex hull area and the ratio of the preimeter to the area.
-These global feature can be used to distinguish between classes.
-For example, a rectangle can be distinguish from a triangle by looking to the raito of area of bounding box to the area of the convex hull.
+ Several systems \cite{DiagramOfflineConvexHull,MulitStrokeConvexHull} presented features like ratio of bounding box to convex hull area and the ratio of the perimeter to the area.
+These global features can be used to distinguish between classes.
+For example, a rectangle can be distinguished from a triangle by looking to the ratio of area of bounding box to the area of the convex hull.
 The triangle ratio will be much lower than that of the rectangle.
@@
 
-%Zernike moments were used by Hse et al. in \cite{}. 
+%Zernike moments were used by Hse et al. in \cite{ToADD}. 
 Hse et al.
 \cite{HeloiseBeautification,zernike61} proposed using the magnitudes of Zernike moments as a shape property.
-Zernike moemnts are orthognals moments which define the distribution of point in the input space.
+Zernike moments are orthogonal moments which define the distribution of point in the input space.
 The higher the order of the moment the finer the detail level the point distribution represents.
@@
 \cite{HeloiseBeautification,zernike61} used a SVM classifier to train the features vector calculated from the moments.
-Early evluation of Zernike moemnts showed that they are invariant to both reflection and rotation of the input shape \cite{ }.
-In our system we used Zernike moemnts as part of our feature vector.
-A detailed comparison between Zernike moments and other features are demonstrated in Section \ref{}.
+Early evaluation of Zernike moments showed that they are invariant to both reflection and rotation of the input shape \cite{ZerMomentOrthogonal}.
+In our system we used Zernike moments as part of our feature vector.
+A detailed comparison between Zernike moments and other features are demonstrated in Section \ref{sec:featuresComparisions}.
 
- Systems based on Global shape properties distinguish shapes by learning the different properties each shape class tend to have.
-These properties dose not depend on the order or number of strokes and thus are only based on apperance of the shape not how the user draw it.
+ \begin{figure}[]
+	\centering
+		\subfigure{Ac Source}{
+
+		\includegraphics[scale=0.6]{images/acSource.eps}		\label{fig:acSource}}
+
+ 		\subfigure{Current Source}{
+
+		\includegraphics[scale=0.6]{images/CurrentSource.eps} \label{fig:CurrentSource}}
+	\caption{Example of Current and AC Sources}
+	\label{fig:ACandCurrentSources}
+\end{figure}
+
+ Systems based on Global shape properties distinguish shapes by learning the different properties each shape class tends to have.
+These properties do not depend on the order or number of strokes and thus are only based on appearance of the shape not how the user draws it.
 However, individual details of the shape are impossible to distinguish using these properties.
-For example, it is impoissible to use  to  differentiate between a current source symbole (circle contiaing and arrow Figure \ref{}) and a clock (circle containing a square signal Figure \ref{}).
-Finear details of the symbols are not represented in this approach, thus shapes that have smalls details or only significant conceptual variation are not distiguished using this method.
-These symbols requires a system to represent different levels of shape details.
+For example, it is impossible to use to differentiate between a current source symbol (circle containing and arrow Figure \ref{fig:CurrentSource}) and a Ac Source (circle containing a wave Figure \ref{fig:acSource}).
+Finer details of the symbols are not represented in this approach, thus shapes that have small's details or only significant conceptual variation are not distinguished using this method.
+These symbols require a system to represent different levels of shape details.
 
-\subsection{Apperance Based Recognition}
+\subsection{Appearance Based Recognition}
 \label{sec:ApperanceBasedRecognition}
 
-In the third approach, the recognition foucus on appearance of shapes contrary to individual strokes or global shape properties.
+In the third approach, the recognition focuses on appearance of shapes contrary to individual strokes or global shape properties.
 Kara \cite{imagetrainable48} used this approach while presenting a system that matches the input shape into a one of the prototypes in the database.
 Figure \ref{fig:template} shows an example of prototypes of some of the shapes in the system.
-Input shapes are first pose and scale normalized then four image similarity measure are used to match input shapes to the database prototypes.
-He used two measures (Hausdor distance )  to measure similarity and two coeficients (Yule, Tanimoto) to measure dissimilarity.
-His approach allows the system to recongized a shape using only a single training sample as it is based on the shape prototype.
-However, the shape protoytpe must represent all  transformation and variation the input shape can have.
-Moreover, they restrict the image representation to 48x48 pixel due to computationa reasons.
-This restriction leads to elimintating some of the fine details that may be needed to distinguish some shape classes.
-There results on isolated symbols seems promising but the system have not beet tested on full sketch that tends to have high overlapping between symbols.
+Input shapes are first pose and scale normalized then four image similarity measures are used to match input shapes to the database prototypes.
+He used two measures (Hausdor distance) to measure similarity and two coefficients (Yule, Tanimoto) to measure dissimilarity.
+His approach allows the system to recognize a shape using only a single training sample as it is based on the shape prototype.
+However, the shape prototype must represent all transformation and variation the input shape can have.
+Moreover, they restrict the image representation to 48x48 pixels due to computational reasons.
+This restriction leads to eliminating some of the fine details that may be needed to distinguish some shape classes.
+There results on isolated symbols seem promising but the system has not been tested on full sketch that tends to have high overlapping between symbols.
+\begin{figure}
+	\centering
+		\includegraphics[scale=0.7]{images/template.eps}
+	\caption{Prototypes of Shapes}
+	\label{fig:template}
+\end{figure}
 
@@
 
-In all of these system only \cite{4} using stroke based recognitiion have been implemented to the fully creat a system that free draw sketch and understand them.
-This is because the emphasis on the interactive side of the sketch sytem which result in large difficultes in processing free hand sketches.
-Methods based on gesture and global stroke properties produce systems that are dependent on assumptions about how strokes can be groubed.
-These systems assume that users will finish drawing a shape before moving to the next one \cite{6,11,14,15}.
-This let them to temporally group strokes and then recognize them based on their strokes or apperance properties.
-\cite{5} argued that this assumption dose not hold in natural sketchs.
+In all of these system only \cite{SketchRead2007} using stroke based recognition have been implemented to the fully create a system that free draw sketch and understand them.
+This is because the emphasis on the interactive side of the sketch system which result in large difficulties in processing free hand sketches.
+Methods based on gesture and global stroke properties produce systems that are dependent on assumptions about how strokes can be grouped.
+These systems assume that users will finish drawing a shape before moving to the next one \cite{multistrokeEvaluation,Cali63,geometrydomain49}.
+This let them to temporally group strokes and then recognizes them based on their strokes or global shape properties.
+\cite{AlvaradoDigital} argued that this assumption does not hold in natural sketches.
 
-Recogntion based on strokes using hierarchical shape description is applied in \cite{}.
-The principle problem in these systems is the computatnal complexity of the matching between shape description and geometeric primitves produced in the segmentaiton process.
-Natural sketch are dificult to segment and group to create reliable fragment of geometeric primitives.
-This is partially due to the increase of noise and other phenomena such as overlapped, overtraced strokes and multiple shapes drawn using
+Recognition based on strokes using hierarchical shape description is applied in \cite{SketchRead2007}.
+The principle problem in these systems is the computational complexity of the matching between shape description and geometric primitives produced in the segmentation process.
+Natural sketch are difficult to segment and group to create reliable fragment of geometric primitives.
+This is partially due to the increase of noise and other phenomena such as overlapped, over traced strokes and sharing of strokes between different symbols.
+Since these systems are dependable on the segmentation results, they either perform recognition using undependable segmentation results or consider different ways to segment the input sketch.
 
-Natural sketches are dificult to reliably fragment into geometric primitives, because of the increase in noise and other phenomena such as interspersed drawing of parts from multiple shapes, multiple
-shapes drawn with single strokes, and overtraced strokes. Because these systems are
-dependent on the fragmentation process, they are either faced with performing recognition
-with unreliable fragmentation results or must consider many possible ways to
-fragment the sketch. Avoiding the fragmentation issue was one of our key motivating
-factors in taking an appearance-based approach to recognition.
-The appearance based methods described above have not been applied, to our
-knowledge, to complete, freely drawn sketches. They have only been applied to
-isolated shapes or synthetic images generated from isolated shapes. Although their
-focus on the visual nature of the sketches appears promising for handling freely drawn
-sketches they have not yet been evaluated on them.
+To avoid most problems in such system, our system uses hybrid of both global shape and hierarchical description methods.
+Currently, our system recognize independent symbols.
+Each symbol is segmented and hierarchical features are computed.
+Features that are representing the global description of the shape are also generated and appended to the hierarchical features.
+Recognition is done using SVM Linear classifier as it proved to be one of the best learning algorithms.
 
- \chapter{Particle Swram Optimization}
+\chapter{Particle Swarm Optimization}
 \label{sec:ParticleSwramOptimization}
 
-\section{Particle Swram Optimization Review}
+\section{Swarm Optimization Review}
 \label{sec:ParticleSwramOptimizationReview}
 
-%review on the pso alogrithm alone.................
+%This seems to emerge of what is called 'collective intelligence' from often large simple agents.  
+% start with swarm in general  .. say about flock , ant and usin in image processing 
+% then speeak about PSO and  its firstly generated   (see article)
+%mention that is best than ga (see thesis)
+
+Modeling how a biological swarm has been recently studied by biologist and computer scientists.
+Scientists in the field of artificial life try to understand how such "social animals" interact, achieve goals and evolve.
+Moreover, a swarm behavior that achieves some kind of swarm intelligence is increasingly interesting for engineers as they can apply it in optimization.
+Looking from a higher view, agents in a swarm are cooperating to achieve some purposeful behavior and achieve some common goal.
+A good example of this cooperation is ant groups looking for food source.
+A large group of simple independent agents seems to generate this behavior in what is called 'collective intelligence'.
+Each agent has a set of simple rules to follow, and according to the interaction of the entire group the objective is met\cite{swarmGeneral}.
+%with other neighbourhood agents.
+
+Reynolds \cite{swarmflock} created 'boid' which simulated the motion of a flock of birds.
+Each boid is a simple agent that navigates according to his perception and a set of basic rules.
+Kube and Zhange \cite{antsimulation} introduced a simulation model for ant's behavior and used in cooperative transportation.
+Ants leaves a trail of its movement, so when an ant find food it cannot carry alone, it can easily trace back to the colony and then a group of ants can track the trail and bring the food back.
+Kube and Zhange \cite{antsimulation,swarmGeneral} system seems to correctly models ants behavior with a methods of task modeling.
+
+Image processing is one of the disciplines that used Swarm Intelligence.
+Ant colony was widely used to segment objects in \cite{antSegmentation,antthreshold,SwarmMedical}.
+Since 1995 when Kennedy and Eberhart \cite{PSOFirst,PSO2} introduced particle swarm algorithm (PSO) the algorithm had been used in various image processing system as an optimization  algorithm \cite{SwarmImageProcessing,Swarmimageprocessing2,Colorquantization,SwarmMedical}.
+PSO algorithm was used in \cite{ToADD} to classify pixel to simplify the object segmentation process.
 

### 2009-04-14 trunk/paper2/Paper2.tex
edit pdp part

--- 2009-03-21 paper2/Paper2.tex
@@
  Speed and time difference information were widely used in sketch recognition systems \cite{earlyprocess}.
+Figure \ref{fig:speed2Distance} shows the time difference and the speed graphs for the stroke drawn in Figure \ref{fig:orignalStroke}.
 It was noted that time difference provides more distinct maxima than the speed information.
@@
 \cite{polygonfeedback31} mentions that the pointing device (for example the mouse) sampling rate is the reason for this phenomena.
-The points are sampled at regular time intervals while the pointing device is moving.
-There are no points while the pointing device is stationary.
-This leads to a nearly constant time difference between samples while the pen is moving and large difference while the pen is stationary.
-Contrary to speed information, where the users draw with variant speeds.
-The speed information has a lot of noise in the data.
-Figure \ref{fig:speed2Distance} shows the time difference and the speed graphs for the stroke drawn in Figure \ref{fig:orignalStroke}.
+The points are sampled at regular time intervals as long as the pointing device is moving.
+There are no detected points when the pointing device is stationary.
+This leads to a nearly constant time difference between samples while the pen is moving and large difference while the pen is stationary (Figure \ref{fig:timediff}).
+Contrary to speed information, where the users draw with variant speeds, the speed information has a lot of noise in the data.
 Similarly, direction information is used as it provides better distinctive maxima for the corners than the curvature information \cite{meanshift10}.% Figure \ref{fig:curvatures} shows the direction and curvature graphs for the stroke drawn in Figure \ref{fig:orignalStroke}
@@
 	\centering
-			\subfigure[ Speed Graph]{	\includegraphics[scale=0.3]{images/vel3.png}}
+			\subfigure[ Speed Graph]{	\includegraphics[scale=0.45]{images/vel3.jpg}}
 			\hfill
-			\subfigure[ Time Difference Graph] {\includegraphics[scale=0.3]{images/td3.png}}
+			\subfigure[ Time Difference Graph] {\label{fig:timediff}\includegraphics[scale=0.45]{images/td3.jpg}}
 	\caption{Speed and Time Difference Graphs}
@@
 \end{figure}
-
 \begin{figure}[]
@@
 \end{figure}
-
- In the presented system, we computed time difference, direction, speed and curvature of each point along a stroke.
-The system computes all these information to ensure that no possible dominate point is overlooked in this phase.
+Dominant points are characterized by lower speed values, higher curvature and direction values.
+However, the dominant points from the curvature information may not be the same points as the dominant points extracted from the speed information.
+Hence, we compute time difference, direction, speed and curvature of each point along a stroke to ensure that no dominant point is overlooked in this phase.
 The speed is calculated as $v=\Delta s/\Delta t$ where $t$ is the time difference between two points and $s$ is the length between them.
@@
 $c= \Delta d/\Delta s$.
-Dominant points are characterized by lower speed values, higher curvature and direction values.
 
@@
 However, using simple differentiation to detect local extreme points resulted in false points due to the non smooth curves.
-Hence, the system adopted a process presented by \cite{earlyprocess}, where the mean of the curve is calculated.
-Then a threshold \textit{th} is used to separate the curve into regions ($Region_i$); each region $Region_i$ is defined as a range of points, where the curve values are either above or below the threshold \textit{th}.
-Those regions are further processed to find the maximum point $Max(Region_i)$ of each region $Region_i$.
+Hence, the system adopted a process presented by \cite{earlyprocess}, where the mean of each of the four calcuated information  is calculated.
+This mean is taken as the threshold \textit{th} which is used to separate the curve into regions ($R_i$) (Figure\ref{fig:timediff}); each region $R_i$ is defined as a range of points, where the curve values are either above or below the threshold \textit{th}.
+Those regions are further processed to find the maximum point $Max(R_i)$ of each region $R_i$.
 The stroke points $p_i(x,y)$ that correspond to those maximum values are labeled as \textit{possible dominant points} $P_{pd}$.
@@
 All the points labeled as possible dominant points $P_{pd}$ are saved into a single array.
-Figure \ref{fig:LabelsPPD} shows the particles labeled as Possible dominate point $P_{pd}$ by the preprocessing, it is noticed the redundancy of some $P_{pd}$ points.
+Figure \ref{fig:LabelsPPD} shows the particles labeled as Possible dominate point $P_{pd}$ as the preprocessing.
+It is clear that some of the $P_{pd}$ points are redundant as specified in the preprocessing stage.
 % (as shown are redundant)
@@
 	\centering
-		\includegraphics[scale=0.5]{images/FeaturesCompare.jpg}
+		\includegraphics[scale=0.5]{images/testFeat.jpg}
 	\caption{Feature Comparison} The effect of different features on accuracy.  %The graph shows the recognition rate of each symbol using different algorithms. 
@@
 	\centering
-		\includegraphics[scale=0.4]{images/FeaturesCompare.jpg}
+		\includegraphics[scale=0.4]{images/testF1.jpg}
 	\caption{} %The effect of symbol complexity.  %The graph shows the recognition rate of each symbol using different algorithms. 
@@
 \bibliographystyle{IEEEbib}
-\bibliography{../../neededfiles/Bibliographies/Mybibliography2}
+\bibliography{../../neededfiles/Bibliographies/Mybibliography}
 \end{document}

### 2009-04-14 paper2/Paper2.tex
The final modification that was given to doc
midway my edditing

@@
  Speed and time difference information were widely used in sketch recognition systems \cite{earlyprocess}.
+Figure \ref{fig:speed2Distance} shows the time difference and the speed graphs for the stroke drawn in Figure \ref{fig:orignalStroke}.
 It was noted that time difference provides more distinct maxima than the speed information.
 Agar et al.
-\cite{polygonfeedback31} mentions that the pointing device (for example the mouse) sampling rate is the reason for this phenomena.
-The points are sampled at regular time intervals while the pointing device is moving.
-There are no points while the pointing device is stationary.
-This leads to a nearly constant time difference between samples while the pen is moving and large difference while the pen is stationary.
-Contrary to speed information, where the users draw with variant speeds.
-The speed information has a lot of noise in the data.
-Figure \ref{fig:speed2Distance} shows the time difference and the speed graphs for the stroke drawn in Figure \ref{fig:orignalStroke}.
+\cite{polygonfeedback31} mention that the pointing device (for example the mouse) sampling rate is the reason for this phenomena.
+The points are sampled at regular time intervals as long as the pointing device is moving.
+Therefore, there are no new detected points when the pointing device is stationary.
+This leads to a nearly constant time difference between samples while the pen is moving and large difference while the pen is stationary (Figure \ref{fig:timediff}).
+Contrary to speed information, where the users draw with variant speeds, the speed information has a lot of noise in the data.
 Similarly, direction information is used as it provides better distinctive maxima for the corners than the curvature information \cite{meanshift10}.% Figure \ref{fig:curvatures} shows the direction and curvature graphs for the stroke drawn in Figure \ref{fig:orignalStroke}
@@
 	\centering
-			\subfigure[ Speed Graph]{	\includegraphics[scale=0.3]{images/vel3.png}}
+			\subfigure[ Speed Graph]{	\includegraphics[scale=0.4]{images/vel3.jpg}}
 			\hfill
-			\subfigure[ Time Difference Graph] {\includegraphics[scale=0.3]{images/td3.png}}
+			\subfigure[ Time Difference Graph] {\label{fig:timediff}\includegraphics[scale=0.4]{images/td3.jpg}}
 	\caption{Speed and Time Difference Graphs}
@@
 If the stroke proved to be an ellipse arc then the segmentation process ends and the system proceeds to the next step.
-Otherwise, the stroke is passed to two particle swarm algorithms that divide the stroke to either lines or lines and curves.
-The input to these algorithms are the stroke points along with the possible dominant points $P_{pd}$ computed during preprocessing.
+Otherwise, the stroke is passed to two furthur segmentation algorithms that divide the stroke to either lines or lines and curves.
+The next section describes in details the ellipse detection algorithm and the two PSO algorithms used to divide the stroke.
+% The input to these algorithms are the stroke points along with the possible dominant points $P_{pd}$ computed during preprocessing.
 The output is a set of dominant points which are connected by either lines or curves.
-The next section describes the ellipse detection algorithm and the two particle swarm algorithms used to divide the stroke.
 
@@
 	\centering
-		\includegraphics[scale=0.7]{images/blockSmall.jpg}
+		\includegraphics[scale=0.5]{images/blockSmall.jpg}
 	\caption{Segmentation Block}
@@
 \end{figure}
-\subsubsection{Ellipse Detection} 
+\subsection{Ellipse fitting} 
 %The process starts by computing the center of the stroke bounding box.
@@
 The least square fitting algorithm is used to minimize the fitting error of the ellipse equation \cite{chernov-2003}.
+t
 This process tries to fit the stroke points into an ellipse arc; it starts with computing the center of the stroke bounding box.
@@
 A list of new values for $x_0$ , $y_0$ ,$a$ and $b$ are generated randomly from the older values with small increments after each loop.
-After few loops, the final fit error and the ellipse confidence value are computed.
-The confidence value $Conf$ computed by Equation (\ref{eq:circleError})
+After few iteration, the final fit error of the estimated ellipse is reported.
+Another measure is used to compute the efficiency of the final estimated ellipse.
+Equation \ref{eq:circleError} ensures that the drawn percentage of ellipse is considered.
+This eliminates fitting a  line into very large ellipse but leaves small ellipses to be fittted as a pratial or full ellipse.
+
  \begin{equation}
-Conf = E_{ellipse}+E_{area}
+eff= (P_{percent}/E)
 \label{eq:circleError}
@@
  \begin{equation}
-E_{area}  = \sqrt {\left( Area(stroke) - Area(ellipse)\right)^2 }
+P_{percent}  = L_{stroke} /P_{ellipse} 
 \label{eq:ErrorArea}
 \end{equation}
-where $E_{ellipse}$ is the error computed by Equation(\ref{eq:circleFit}) and $E_{area}$ is computed as the difference between the actual area of the stroke and the area of the estimated ellipse Equation(\ref{eq:ErrorArea}).
-If confidence value $Conf$ is less than threshold $th_{Ellipse}$\footnote{The threshold used is $th_{Ellipse}=0.7$ } then the stroke is segmented as an ellipse otherwise the system proceeds to the two PSO segmentation algorithms.
+where $E$ is the error computed by Equation(\ref{eq:circleFit}), $L_{stroke}$ total length of stroke and $P_{ellipse} $ is the perimeter of the estimated ellipse.
+If $eff$ is more than threshold $th_{Ellipse}$\footnote{By trial and error best threshold found was $th_{Ellipse}=0.2$} then the stroke is segmented as an ellipse otherwise the system proceeds to the next step.
 
-\subsubsection{Discrete particle swarm algorithm (\textit{DPSO})}
+\subsection{Non ellipse fitting algorithm}
 \label{subsubsec:Discreteparticleswarmalgorithm}
-Two DPSO algorithms are used to generate segmentations for each stroke the user draws.
+Two DPSO algorithms are used to generate different segmentations for each stroke the user draws.
 For each segmentation generated an error is evaluated.
@@
 	\item[Fitness function:] The fitness function and error calculation are different in each of the two \textit{DPSO} algorithms. 
-
-\textit{The first algorithm \textsl{AlgS1}}\\
+	\end{description}
+\subsubsection{\textit{Polygon approximation\textsl{AlgS1}}}
+%\\
 The approximation error is computed by the equation \ref{eq:ErrorSwarm1}. The graphical meaning of the error is shown in Figure\ref{fig:LabelsPSO}.
@@
 
- \textit{The second algorithm \textsl{AlgS2}} \\
+ \subsubsection{\textit{Hybird Fitting \textsl{AlgS2}}}
 The algorithm has the same problem formulation but different fitness and error functions are used.
-An attempt is made to fit each segment into line or circular arc.
+An attempt is made to fit each segment into a line or circular arc.
 The errors of both circle and line fit estimations are computed for each segment $S_i=\widehat{x_ix_j}$.
@@
 For our system, $k$ is selected to be 0.5\cite{CruveDivisionSwarm}.
-\end{description}
 
@@
 If it was not labeled the point $P_{ij}$ is moved to the nearest labeled point.
-This ensures that all of the points generated by the DPSO are possible dominate points $P_{pd}$.
-	\item The particles are tested to make sure that the distance between every two successive dominate point is larger than the constant $min_D$.
-Otherwise, one of these points are removed based on the error computed by the segmentation algorithm.
+This ensures that all of the points generated by the DPSO are possible dominant points $P_{pd}$.
+	\item The particles are tested to make sure that the distance between every two successive dominate point is larger than $min_D$, where $min_D$ is 5\% of the total length of the stroke.
+Otherwise, one of these points are removed based on the error cased by its removal.
 	\item Each two successive segments are test to determine if they can be merged into a single segment.
@@
 The following list describes each feature in details.
-\begin{description}
- \item [Structural and geometrical Features(FS1)] Features defines the structure of the geometrical symbol.  
-\begin{description}
-	 \item[No. Of segments] Number of segments in the symbol. 
-		\item[No. Of primitives] Number of primitives in the symbol. The feature helps when identifying             symbols with mixed geometric primitives like cylinders and callouts.  
-		\item [No. of curves] Number of curves or ellipses in the symbol. 
-		\item [No. of lines] Number of lines in the symbol. 
-		\item [No. of perpendicular lines] Number of perpendicular lines. 
-		\item [No. of parallel lines] Number of parallel lines. 
-		\item [No. of intersections] Number of intersection between lines and curves. 
-		\item [Size Ratio] Ratio between width to height of the symbol.
+%\begin{description}
+\subsubsection{Structural and geometrical Features(FS1)}
+  Features defines the structure of the geometrical symbol.  
+%\begin{description}
+%	 \item[No. Of segments] Number of segments in the symbol. 
+%		\item[No. Of primitives] Number of primitives in the symbol. The feature helps when identifying             symbols with mixed geometric primitives like cylinders and callouts.  
+%		\item [No. of curves] Number of curves or ellipses in the symbol. 
+%		\item [No. of lines] Number of lines in the symbol. 
+%		\item [No. of perpendicular lines] Number of perpendicular lines. 
+%		\item [No. of parallel lines] Number of parallel lines. 
+%		\item [No. of intersections] Number of intersection between lines and curves. 
+%		\item [Size Ratio] Ratio between width to height of the symbol.
+%		%\item[No. of holes] Number of holes in the symbols. 
+%
+%\end{description}
+ \begin{itemize}
+	 \item \emph{Segments:} Number of segments in the symbol.
+	 \item \emph{Strokes:} Number of strokes or partial strokes that created the symbols.  
+		\item  \emph{Primitives:} Number of primitives in the symbol. The feature helps when identifying             symbols with mixed geometric primitives like cylinders and callouts.  
+		\item \emph{Curves:} Number of curves or ellipses in the symbol. 
+		\item \emph{Lines:} Number of lines in the symbol. 
+		\item \emph{Perpendicular} lines Number of perpendicular lines. 
+		\item \emph{Parallel} lines Number of parallel lines. 
+		\item \emph{Intersections} Number of intersection between lines and curves. 
+		\item \emph{T intersections} Number of T intersections. 
+		\item \emph{L intersections} Number of L intersections. 
+		\item \emph{X intersections} Number of X intersections.
+
 		%\item[No. of holes] Number of holes in the symbols. 
+\end{itemize}
+\subsubsection{Rubine Feature Set (FS2)}
+  Features introduced by Rubine\cite{gestureexample12} for single stroke gestures. Feature such as length of stroke, smoothness of symbol and properties of the bounding box of the symbol.
 
-\end{description}
-\item [Rubine Feature Set (FS2)] Features introduced by Rubine\cite{gestureexample12} for single stroke gestures. 
-\item[Statistical Features (FS3)]   
-\begin{description}
-	\item[Zernike moments]  Zernike moments of order \cite{HeloiseBeautification}. 
-\end{description}
-	\item[Composite Features (FS4)] Features that are composites of different geometrical and statistical symbol. 
-	\begin{description}
+\subsubsection{Statistical Features (FS3)}  
+\begin{itemize}
+	\item \emph{Zernike moments } Zernike moments of order \cite{HeloiseBeautification}. 
+\end{itemize}
+\subsubsection{Composite Features (FS4)}
+ Features that are composites of different geometrical and statistical symbol. 
+	\begin{itemize}
+\item \emph{Size Ratio} Ratio between width to height of the symbol.
+	\item \emph{Ink density} Compute the density of points inside its bounding box\cite{GeometryAndDomain102}.   
+ 	\item \emph{Convex Hull Area} Area of convex hull with respect to area of bounding box of symbol.
+	\item \emph{Convex Hull Perimeter} Perimeter of convex hull with respect to total length of symbol.
+		\item \emph{Mean Centroidal radius} The Mean of the centroidal radius which is the distance from each point in the symbol to the center of gravity.
 
-	\item [Ink density] Compute the density of points inside its bounding box\cite{GeometryAndDomain102}.   
- 	\item[Convex Hull Area] Area of convex hull with respect to area of bounding box of symbol.
-	\item[Convex Hull Perimeter] Perimeter of convex hull with respect to total length of symbol.
-		\item [Mean Centroidal radius] The Mean of the centroidal radius which is the distance from each point in the symbol to the center of gravity.
+	\item \emph{Mean Time difference} The mean of the time difference between each two successive points in the symbol. %Different strokes are appended to construct a single path.
+	%\item [Ra]
+  \end{itemize}
 
-	\item [Mean Time difference] The mean of the time difference between each two successive points in the symbol. %Different strokes are appended to construct a single path.
-	%\item [Ra]
-  \end{description}
-
-\end{description}
+%\end{description}
 %The features used consist of Rubine feature set,  of order 21 , ink density as well as some structural and spatial information like number of perpendiculars lines ,number of parallel lines and number of different types of primitives in each symbol.
@@
 Four different splits for the training and test data are generated from the dataset.
-The results displayed are the average recognition accuracy of the four splits resulted from the classifier after recognizing the symbols.
+The results displayed are the average recognition accuracy of the four splits.
 The accuracy is computed as the number of correctly recognized samples divided by the total number of samples in the test.

### 2009-04-14 trunk/paper2/Paper2.tex
after my edit today 15 midnight

@@
 Agar et al.
-\cite{polygonfeedback31} mentions that the pointing device (for example the mouse) sampling rate is the reason for this phenomena.
+\cite{polygonfeedback31} mention that the pointing device (for example the mouse) sampling rate is the reason for this phenomena.
 The points are sampled at regular time intervals as long as the pointing device is moving.
-There are no detected points when the pointing device is stationary.
+Therefore, there are no new detected points when the pointing device is stationary.
 This leads to a nearly constant time difference between samples while the pen is moving and large difference while the pen is stationary (Figure \ref{fig:timediff}).
@@
 	\centering
-			\subfigure[ Speed Graph]{	\includegraphics[scale=0.45]{images/vel3.jpg}}
+			\subfigure[ Speed Graph]{	\includegraphics[scale=0.4]{images/vel3.jpg}}
 			\hfill
-			\subfigure[ Time Difference Graph] {\label{fig:timediff}\includegraphics[scale=0.45]{images/td3.jpg}}
+			\subfigure[ Time Difference Graph] {\label{fig:timediff}\includegraphics[scale=0.4]{images/td3.jpg}}
 	\caption{Speed and Time Difference Graphs}
@@
 \label{sec:AlgExp}
-A set of test were performed on the system to enhance both the swarm algorithms outputs.
-The effect of number of agents, the error thresholds and various other paramters was investigated.
+A set of tests were performed on the system to enhance non ellipse fitting algorithm.
+The effect of number of agents, the error thresholds and various other paramters were investigated.
 The result of these tests are shown in Figure \ref{fig:swarmtesting} and   \ref{fig:swarmtesting2}.
 Figure \ref{fig:swarmtesting} shows that as the number of agents increases the error value decrease (error of the segmentation algorithm).
-The final number of agents and maximum swarm iteration used in the system was based trade off between error achieved versus the computation time.
-Other algorithm parameters like $c_1$,$c_2$ that were mentioned in Section\ref{sec:ParticleSwarmAlgorithm} are tested using similar tests (Figure \ref{fig:swarmtesting} and \ref{fig:swarmtesting2}).
+The final number of agents and maximum swarm iteration used in the system was based on a trade off between error achieved versus the computation time.
+Other algorithm parameters such as $c_1$,$c_2$ that were mentioned in Section\ref{sec:ParticleSwarmAlgorithm} are tested using similar tests (Figure \ref{fig:swarmtesting} and \ref{fig:swarmtesting2}).
 
@@
 \end{figure*}  
-
  We could not test the result of the segmentation algorithm directly due to the fact that the correct segmentation is highly ambiguous.
 It is only based on what the user intends to draw while sketching the symbol.
-Figure \ref{fig:SampleSeg} shows a sample of the output of the segmentation block.
+ The main features affected by teh segmentation stage are the geometeric features computed in\textbf{FS1 }.
+Hence those are the features we used to estimate the efficency of the segmentation algorithm.
+The recognition accuracy using only the gometrical features is reported in Figure \ref{fig:testFeatonly} .
+The figure shows that \textsl{AlgS2} achieve better results.
+%effect of the segmentation algorithm
+ %Figure \ref{fig:SampleSeg} shows a sample of the output of the segmentation block.
 The result of segmentation block is tested by their effect on the final recognition results using the features that will be highly affected using different algorithms like \textbf{FS1} and \textbf{FS4} (Section\ref{sec:Recognition}).
@@
 The use of statistical moments and structural features improves the recognition rate.
-The system was tested on 13 different symbols and achieved an accuracy of 93.5\%.
+The system was tested on 13 different symbols and achieved an overall accuracy of 93.5\%.
 The system does not depend on low level recognizer but rather on a set of high level features.
@@
  A possible extension of this research is to complete the clustering algorithm for fully automated sketch recognition.
-Currently, the system process one symbol at a time, an enchantment would be to allow users to draw more than one symbol at a time and incorporate a method for separating symbols.
-Other area of enhancements is the features extraction methods.
+Currently, the system processes one symbol at a time, an enhancement would be to allow users to draw more than one symbol at a time and incorporate a method for separating symbols.
+Another area of enhancements is the features extraction methods.
 Introducing more spatial and geometrical features is believed to improve classifications.

### 2009-04-16 trunk/paper2/Paper2.tex
after the my edit of introduction and review

@@
   {Magda Fayek\sthanks{Associate Professor}, Nevin Darwish\sthanks{Professor}}
-	{ Computer Engineering Department\\
+	{Computer Engineering Department\\
 	Faculty of Engineering \\Cairo University\\Cairo, Egypt\\
@@
 \end{abstract}
+
 \section{Introduction}
-The field of symbol recognition has gained interest in the last few years.
-Scientists generally and engineers specifically express thoughts and designs using sketches.
-Engineers use sketches to exchange designs as a natural method of communication rather than writing or speaking.
-Design engineers need a powerful computer based system with symbol recognition capabilities to help them design, manipulate and store sketches more effectively than using only papers.
-Increasing the interaction between computers and users in sketch and CAD systems has been the main reason for the emerging of a few advanced sketch recognition systems.
 
-Sketch recognition is divided into three main steps a) preprocessing, b) segmentation and c) symbol recognition.
-The preprocessing phase captures user input stroke points and collects basic information about the stroke.
-This is followed by proceeding to remove noise and computing basic statistical and geometrical information.
-In the segmentation phase, the strokes are divided into a set of simple geometrical primitives or segments.
-In the third phase, sketch recognition, strokes and segments are clustered to formulate symbols that can be recognized by a classifier.
+Computer research nowadays aims at facilitating computer human interaction in every way.
+Pen-based interfaces give the user a pencil-paper like feeling that enhances interaction more than the current used keyboard-mouse computer interfaces.
+However, Using sketches as input to a computer system requires the system to understand the drawn sketches.
+In a truly freely sketch system this is not a trivial problem due to the sloppy and messy way users draw sketches.
+Sketches here stand for any hand drawing the user draws using digital stylus or similar device.
+Normally, sketches are composed of a set of strokes.
+The time between pen-up and pen-down events identify each stroke.
+The strokes are mainly the path of points, which the user path through between the pen-down and pen-up events while using the stylus or similar device.
 
-This paper introduces a new sketch recognition system, which uses particle swarm optimization (PSO) to optimally segment user's strokes.
+Sketch recognition can be based on: strokes, global shape properties, and appearance \cite{Oltmans07}.
+The role of which each stroke plays in the structure of the sketched symbol is used to determine shapes in stroke based methods.
+Most stroke based systems either represent symbols as gestures or as a  hierarchical shape.
+Gestures are pen strokes drawn by the user that can immediately recognized by the system.
+A Hierarchical shape descriptions was widely used in sketch systems \cite{HierarchicalParsing7,Ladder30,AlvaradoFreedom42,SketchRead2007}.
+Symbols are represented using hierarchical levels of simpler geometric shapes or primitives and spatial relations between them.
+Systems that employ such representation must first pre-process strokes to break them into geometric primitives.
+Recognition can then be treated as sub graph matching problem.
+
+ In global shape properties approach, systems investigate general shape properties and its underlying strokes.
+Using shape properties (i.e.
+ratio of area of convex hull to area of shape) relaxes the assumption and role of each stroke in the symbol but it does not truly represent the appearance of the symbol\cite{DiagramOfflineConvexHull}.
+Unlike appearance based methods where individual strokes are disregarded and they focus on the overall appearance of shapes.
+In general, the  appearance based methods use image processing techniques to recognize sketches that the users draw.
+
+ Methods based on gesture and global shape properties produce systems that are dependent on assumptions about how strokes can be grouped.
+These systems assume that users will finish drawing a shape before moving to the next one \cite{Cali63,geometrydomain49}.
+This let them to temporally group strokes and then recognizes them based on their strokes or global shape properties.
+\cite{AlvaradoDigital} argued that this assumption does not hold in natural sketches.
+
+ Recognition based on strokes using hierarchical shape description is applied in \cite{SketchRead2007}.
+The principle problem in these systems is the computational complexity of the matching between shape description and geometric primitives produced in the segmentation process.
+Natural sketch are difficult to segment and group to create reliable fragment of geometric primitives.
+This is partially due to the increase of noise and other phenomena such as overlapped, over traced strokes and sharing of strokes between different symbols.
+Since these systems are dependable on the segmentation results, they either perform recognition using undependable segmentation results or consider different ways to segment the input sketch.
+
+In this paper we choose to implement a hybrid method between shape property and stroke based approaches.
+Our method tries to avoid many of the problems faced by approaches based on individual strokes or global shape properties while trying to gain the additional information that the stroke based system use.
+Firstly, the system uses particle swarm optimization (PSO) to optimally segment user's strokes.
 Users can draw symbols using any number of strokes in any order they like.
@@
 The final classification step uses a SVM classifier to classify symbols into one of the previously trained classes based on the features vector computed.
-% uses the computed feature vector to classify the segments into one of the previously trained classes.
 
- The remaining of the paper is as follows; the literature review is in Section \ref{sec:review}.
+ %The preprocessing phase captures user input stroke points and collects basic information about the stroke.
+In the segmentation phase, the strokes are divided into a set of simple geometrical primitives or segments.
+In the third phase, sketch recognition, strokes and segments are clustered to formulate symbols that can be recognized by a classifier.
+
+%This paper introduces a new sketch recognition system, which uses  % uses the computed feature vector to classify the segments into one of the previously trained classes.
+
+ The remaining of the paper is as follows; the literature review on stroke segmentation is in Section \ref{sec:review}.
 Section \ref{sec:ParticleSwarmAlgorithm} introduces Particle Swarm Optimization.
@@
 \label{sec:review}
-A wide variety of techniques were suggested for sketch recognition.
+
+Yu \cite{meanshift10} introduced a \textit{feature area} for each primitive and then computed the segmentation error for different types of primitives based on the computed \textit{feature area}.
+His system achieved good accuracy in simple shapes (square, ellipse,...etc) but did not perform well in complex shapes.
 A hybrid algorithm was introduced in \cite{earlyprocess} where different sets of segments are generated based on both curvature and speed dominant points, followed by choosing a segmentation with the least error from a generated hybrid set.
@@
 Each low level recognizer is designed to recognize only one geometric shape using spatial and geometric information extracted from input stroke.
-Yu \cite{meanshift10} introduced a \textit{feature area} for each primitive and then computed the segmentation error for different types of primitives based on the computed \textit{feature area}.
-His system achieved good accuracy in simple shapes (square, ellipse,...etc) but did not perform well in complex shapes.
-Paulson and Hammond \cite{Paleosketch08} introduced a set of low level recognizers that were reported to achieve 98\%   accuracy.
-However, their system similar to all low level recognizers in the way that it identifies a small set of simple shapes.% A genetic algorithm was used by \cite{CruveDivisionSwarm} to optimally divide digital curves into lines and curves.
-Chen et al.\cite{CruveDivisionSwarm} uses digital curves scanned from paper as input to the system and did not take advantage of the curvature or local geometric properties of the digital curve.
 
@@
 The next section presents the general particle swarm algorithm which is used in both our segmentation algorithms.
+
  % are presented in section \ref{subsubsec:Discreteparticleswarmalgorithm}
@@
 However, using simple differentiation to detect local extreme points resulted in false points due to the non smooth curves.
-Hence, the system adopted a process presented by \cite{earlyprocess}, where the mean of each of the four calcuated information  is calculated.
+Hence, the system adopted a process presented by \cite{earlyprocess}, where the mean of each of the four calculated information  is calculated.
 This mean is taken as the threshold \textit{th} which is used to separate the curve into regions ($R_i$) (Figure\ref{fig:timediff}); each region $R_i$ is defined as a range of points, where the curve values are either above or below the threshold \textit{th}.
@@
 In the segmentation stage the input stroke is divided into a set of primitives.
-As shown in Figure \ref{fig:segblock} first an attempt is made to fit the stroke points into a curve or an ellipse using a minimum square error fitting algorithm \cite{chernov-2003}.
+As shown in Figure \ref{fig:segblock} first an attempt is made to fit the stroke points into a curve or an ellipse using a minimum square error fitting algorithm \cite{ellipsefit}.
 The ellipse fitting step helps to prevent the system from over segmenting the stroke into multiple lines or curves if the input stroke is an ellipse.
 If the stroke proved to be an ellipse arc then the segmentation process ends and the system proceeds to the next step.
-Otherwise, the stroke is passed to two furthur segmentation algorithms that divide the stroke to either lines or lines and curves.
-The next section describes in details the ellipse detection algorithm and the two PSO algorithms used to divide the stroke.
+Otherwise, the stroke is passed to two further segmentation algorithms that divide the stroke to either lines or lines and curves.
+After two different segmentations are generated, the system chooses the segmentation with the minimum error.
+The following sections describe in details the ellipse detection algorithm and the two segmentation algorithms used to divide the stroke.
 % The input to these algorithms are the stroke points along with the possible dominant points $P_{pd}$ computed during preprocessing.
@@
 	\centering
-		\includegraphics[scale=0.5]{images/blockSmall.jpg}
+		\includegraphics[scale=0.48]{images/blockSmall.jpg}
 	\caption{Segmentation Block}
@@
 The least square fitting algorithm is used to minimize the fitting error of the ellipse equation \cite{chernov-2003}.
-t
 This process tries to fit the stroke points into an ellipse arc; it starts with computing the center of the stroke bounding box.
@@
 Equation \ref{eq:circleError} ensures that the drawn percentage of ellipse is considered.
-This eliminates fitting a  line into very large ellipse but leaves small ellipses to be fittted as a pratial or full ellipse.
+This eliminates fitting a  line into very large ellipse but leaves small ellipses to be fitted as a partial or full ellipse.
 
@@
 
- \subsubsection{\textit{Hybird Fitting \textsl{AlgS2}}}
+ \subsubsection{\textit{Hybrid Fitting \textsl{AlgS2}}}
 The algorithm has the same problem formulation but different fitness and error functions are used.
@@
 A set of tests were performed on the system to enhance non ellipse fitting algorithm.
-The effect of number of agents, the error thresholds and various other paramters were investigated.
+The effect of number of agents, the error thresholds and various other parameters were investigated.
 The result of these tests are shown in Figure \ref{fig:swarmtesting} and   \ref{fig:swarmtesting2}.
@@
 It is only based on what the user intends to draw while sketching the symbol.
- The main features affected by teh segmentation stage are the geometeric features computed in\textbf{FS1 }.
-Hence those are the features we used to estimate the efficency of the segmentation algorithm.
-The recognition accuracy using only the gometrical features is reported in Figure \ref{fig:testFeatonly} .
+ The main features affected by the segmentation stage are the geometric features computed in\textbf{FS1 }.
+Hence those are the features we used to estimate the efficiency of the segmentation algorithm.
+The recognition accuracy using only the geometrical features is reported in Figure \ref{fig:testFeatonly} .
 The figure shows that \textsl{AlgS2} achieve better results.
@@
 The tradeoff between accuracy achieved and time complexity must be further investigated to achieve better results.
-The use of statistical moments and structural features improves the recognition rate.
+The final recognition is based on both global shape and hierarchical stroke properties.
+Adding stroke and segment geometrical features improved the final recognition.
+Global shape and moment properties was used to prevent the need to use graph matching which is computationally expensive.
 The system was tested on 13 different symbols and achieved an overall accuracy of 93.5\%.
@@
 Introducing more spatial and geometrical features is believed to improve classifications.
+Features that represents the appearance of the symbols are likely to improve the recognition that have similar geometrical structure (i.e.
+Nand and And gate in logic symbols).
 \bibliographystyle{IEEEbib}

### 2009-04-18 trunk/draft/chapterSurvey.tex
today is 18 appril

@@
 Current technology and the increase in computing power have increased the abilities of such interfaces and allowed the implementation of fully sketching systems.
-Over the years, many other sketching based system have been developed, although through this history most system focused on building the interactive interfaces to assist users while entering information to the computer.
+Over the years, many other sketching based systems have been developed.
+Through this history most system focused on building the interactive interfaces to assist users while entering information to the computer.
 To demonstrate our system inspiration we will first present current research in the area of sketch recognition.
-Firstly, preprocessing and segmentation methods will be discussed in details.
+In Section \ref{sec:preprocessing}, preprocessing and segmentation methods will be discussed in details.
 Later, symbol recognition methods will be presented and analyzed in Section \ref{sec:symbolrecogntion}.
@@
 This means that the user can draw symbols using single or multiple strokes without constrains.
-The goal of a sketch system is making the user feel that using the using the system is nearly the same as using pencil and paper.
+The goal of a sketch system is making the user feel that using the system is nearly the same as using a pencil and paper.
 But sketches are incomplete drawings that are sloppy and messy, for example figures \ref{fig:Overshootandundershoot} and \ref{fig:overstroked} show examples of messy sketches drawn by users.
@@
  Agar et al.
-\cite{polygonfeedback31} used the time stamp for each point and compute time difference between every two successive point to help him finding corners as it provides more distinct maximum than the speed information.
+\cite{polygonfeedback31} used the timestamp for each point and compute time difference between every two successive point to help him find corners as he argued that it provides more distinct maximum than the speed information.
 He mentions that the pointing device (for example the mouse) sampling rate is the reason for this phenomena.
@@
 
-In this section, we review three different types of symbol recognition methods.
+%In this section, we review three different types of symbol recognition methods.
  Symbol recognition can be based on: strokes, global shape properties, and appearance \cite{Oltmans07}.
@@
 		\includegraphics[scale=0.7]{../../neededfiles/Figures/squarescematic.eps}
-	\caption{square semantic graph}
+	\caption{Semantic Graph of a Square }
 	\label{fig:squarescematic}
@@
  Avoiding graph matching problems, research tried to use languages that describe shapes instead of graphs.
-Visual descriptive languages represent the shape into tokens of geometric primitives or relations between smaller geometric primitives.
+Visual descriptive languages represent the shape with a set of  tokens representing geometric primitives or spatial relations between smaller geometric primitives.
 Figures \ref{fig:arrow1} and \ref{fig:arr2} show a regular arrow with 2 shafts and head and its definition in Ladder language \cite{Ladder30}.
@@
 		 {\label{fig:arr2}\includegraphics[scale=0.88]{../../neededfiles/Figures/arrowladderdef2.eps}}
-	\caption[Arrow definition]{}
+	\caption[Arrow Definition]{}
 		\label{fig:arrowladderdef1}
@@
 
-		\includegraphics{images/acSource.eps}		\label{fig:acSource}}
+		\includegraphics[scale=0.6]{images/acSource.eps}		\label{fig:acSource}}
 
@@
 
-		\includegraphics{images/CurrentSource.eps} \label{fig:CurrentSource}}
-	\caption{Example of current and ac sources}
+		\includegraphics[scale=0.6]{images/CurrentSource.eps} \label{fig:CurrentSource}}
+	\caption{Example of Current and AC Sources}
 	\label{fig:ACandCurrentSources}
@@
 
-\subsection{Apperance Based Recognition}
+\subsection{Appearance Based Recognition}
 \label{sec:ApperanceBasedRecognition}
@@
 There results on isolated symbols seem promising but the system has not been tested on full sketch that tends to have high overlapping between symbols.
+\begin{figure}
+	\centering
+		\includegraphics[scale=0.7]{images/template.eps}
+	\caption{Prototypes of Shapes}
+	\label{fig:template}
+\end{figure}
 
@@
 These systems assume that users will finish drawing a shape before moving to the next one \cite{multistrokeEvaluation,Cali63,geometrydomain49}.
-This let them to temporally group strokes and then recognizes them based on their strokes or appearance properties.
+This let them to temporally group strokes and then recognizes them based on their strokes or global shape properties.
 \cite{AlvaradoDigital} argued that this assumption does not hold in natural sketches.
@@
 To avoid most problems in such system, our system uses hybrid of both global shape and hierarchical description methods.
-Currently, our system recognize on symbols at a time.
+Currently, our system recognize independent symbols.
 Each symbol is segmented and hierarchical features are computed.
 Features that are representing the global description of the shape are also generated and appended to the hierarchical features.
-The recognition is done using SVM Linear classifier as it proved to be one of the best learning algorithms.
+Recognition is done using SVM Linear classifier as it proved to be one of the best learning algorithms.
 
@@
 Modeling how a biological swarm has been recently studied by biologist and computer scientists.
-Scientist in the field or artificial life need to understand how such "social animals" interact, achieve goals and evolve.
-Moreover, a swarm behavior that achieves some kind of swarm intelligence is increasingly interesting for engineers as it they can be applied in optimization.
+Scientists in the field of artificial life try to understand how such "social animals" interact, achieve goals and evolve.
+Moreover, a swarm behavior that achieves some kind of swarm intelligence is increasingly interesting for engineers as they can apply it in optimization.
 Looking from a higher view, agents in a swarm are cooperating to achieve some purposeful behavior and achieve some common goal.

### 2009-04-19 draft/Results.tex
results after change i mad for paper2

@@
 	\item [Hse Dataset] (denoted as Hs-DB) Data set collected by \cite{HeloiseBeautification} and was previously used by \cite{Oltmans07} as benchmark for sketch recognition problem.
-The data are drawn by 16 users each of them have drawn from 30 to 50 samples for each shape.
+The data are drawn by 16 users each of them have drawn from 50 to 60 samples for each shape.
 Figure \ref{fig:symbolSet} shows the shapes used in the data set.
 The dataset was divided into training set and test set.
-Four different splits for the training and test data are generated from the dataset.
+Five different splits for the training and test data are generated from the dataset.
 In worth noting that \cite{HeloiseBeautification} and \cite{Oltmans07} achieved 96.7\% and  94.4\%  on this dataset respectively.
@@
 		\item [Multi Domain Dataset] (denoted as MD-DB) A dataset we collected manually from different users.
-The dataset contains symbols from three different domains ( UML, mechanical and electrical engineering symbols).
-Data are collected from 4 different users each of them have drawn from 20 to 40 samples of each shapes.
+The dataset contains symbols from two different domains Electrical Engineering symbols (EL-DB) and logic design symbols (LD-DB).
+Data are collected from 7 different users each of them have drawn from 20 to 40 samples of each shapes.
 Figures \ref{fig:LogicSet} and \ref{fig:ElectSet} show some of the shapes used in the dataset.
@@
 The dataset was divided into training set and test set.
-Four different splits for the training and test data are generated from the dataset.
+Five different splits for the training and test data are generated from the dataset.
 %The results displayed are the average recognition accuracy of the four splits resulted from the classifier after recognizing the symbols.
 \end{description}
+
+Table \ref{tab:datasets} shows a  comparison between the three datasets used in our experiments.
+
  %We divided the dataset equally into training set and test set.  The results displayed are the recognition accuracy resulted from the classifier after recognizing the symbols. \\
@@
 \fbox{  \parbox{6cm}{% 
-		\includegraphics[scale=0.7]{images/logicSymbols.eps}	}
+		\includegraphics[scale=0.7]{paperImages/logicSet.jpg.eps}	}
 		}
@@
 \begin{figure}[]\centering
-\fbox{  \parbox{8cm}{% 
-		\includegraphics[scale=0.7]{images/SymbolsElectricalr.eps}	}
+\fbox{  \parbox{8.2cm}{% 
+		\includegraphics[scale=0.7]{paperImages/EelectImage.jpg.eps}	}
 		}
@@
 \end{figure}
+\begin{table}
+\begin{center}
+\scalebox{0.9}{
+ \begin{tabular}{|p{4cm}|p{2.5cm}|p{2.5cm}|p{2.5cm}|}
+ \hline
+ & Hs-DB &  EL-DB & LD-DB \\ \hline
+No of samples  &  7791 & 2764 & 1859 \\ \hline
+No of categories  & 13 & 15 & 8 \\ \hline
+Avg. Samples per Category& 600 & 184 & 232 \\\hline 
+No of users  & 16 & 7 & 7 \\  \hline
+Balanced & Yes&No & No \\ \hline
+Splits  & 5 random splits  & 5 random splits  & 5 random splits  \\ \hline
+Examples & Figure \ref{fig:symbolSet}  & Figure \ref{fig:ELsymbolSet}  & Figure \ref{fig:LogicsymbolSet}   \\ \hline
+\end{tabular}
+}
+\caption[Datasets Comparisons]{Datasets Comparisons} 
+\label{tab:datasets}
+
+\end{center}
+\end{table}
 
@@
 Similar results for the maximum swarm iterations are shown in Figure \ref{fig:swarmtesting1}.
-Other algorithm parameters like $c_1$,$c_2$ that were mentioned in Section\ref{sec:ParticleSwarmAlgorithm} are tested using similar tests (Figure \ref{fig:swarmtesting} and \ref{fig:swarmtesting2}).
+Other algorithm parameters like $c_1$,$c_2$ that were mentioned in Section\ref{sec:ParticleSwarmAlgorithm} are tested using similar tests (Figure \ref{fig:swarmparamterw.jpg} and \ref{fig:swarmtesting2}).
 The final values are for maximum number of iteration = 80 with 15 particles.
@@
 	\centering		
-	 \includegraphics[scale=0.3]{results/swarmalg1.eps}
-	 	\caption{Segmentation Error Vs. Maximum iterations For ALgS1}
-	 	\label{fig:swarmtestingS1}
+	 \includegraphics[scale=0.5]{paperImages/swarmAgentTest.jpg.eps}
+	 	\caption{Segmentation Error Vs. Number of agents}
+	 	\label{fig:swarmtesting}
 	%\caption{Experiments results :} a)    b) c) 
@@
 	\centering		
-	 \includegraphics[scale=0.3]{results/swarmalg2.eps}
-	 	\caption{Segmentation Error Vs. Maximum iterations For ALgS2}
-	 	\label{fig:swarmtestingS2}
+	 \includegraphics[scale=0.5]{paperImages/swarmParamterC.jpg.eps}
+	 	\caption{Segmentation Error Vs. PSO parameters C1}
+	 	\label{fig:swarmtesting2}
 	%\caption{Experiments results :} a)    b) c)
+
+\end{figure}
+
+\begin{figure}
+	\centering
+		\includegraphics[scale=0.6]{paperImages/swarmparamterw.jpg.eps}
+	\caption{Segmentation Error Vs. PSO parameters W.}
+	\label{fig:swarmparamterw.jpg}
 \end{figure}
@@
 We also implemented the segmentation algorithm described in \cite{earlyprocess} (\textsl{Alg3}) to use as reference to compare it with our swarm algorithms.
-Figure \ref{fig:test1} shows the accuracy achieved by each algorithm.
+Figure \ref{fig:test1},\ref{fig:testEL} and \ref{fig:testLD}  shows the accuracy achieved by each algorithm.
 The two swarm algorithms were tested with and without the ellipse fitting module.
@@
 	\centering		
-	 \includegraphics[scale=0.4]{results/testAlg.eps}
-	 	\caption{Algorithm comparison} The recognition rate of different algorithms. 
+	 \includegraphics[scale=0.4]{paperImages/testAlg.jpg.eps}
+	 	\caption{Algorithm comparison on HS-DB} The recognition rate of different algorithms. 
 	 	\label{fig:test1}
+	%\caption{Experiments results :} a)    b) c) 
+\end{figure}
+
+\begin{figure}
+	\centering		
+	 \includegraphics[scale=0.4]{paperImages/LDAlg.jpg.eps}
+	 	\caption{Algorithm comparison on LD-DB} The recognition rate of different algorithms. 
+	 	\label{fig:testLD}
+	%\caption{Experiments results :} a)    b) c) 
+\end{figure}
+
+\begin{figure}
+	\centering		
+	 \includegraphics[scale=0.1]{paperImages/LDAlg.jpg.eps}
+	 	\caption{Algorithm comparison on EL-DB} The recognition rate of different algorithms. 
+	 	\label{fig:testEL}
 	%\caption{Experiments results :} a)    b) c) 
@@
 
-\begin{figure*}
+%\begin{figure}
+%	\centering
+%		\includegraphics[scale=0.5]{results/testsym.eps}
+%	\caption{Symbols Comparison} The effect of symbol complexity.  %The graph shows the recognition rate of each symbol using different algorithms. 
+%	\label{fig:test2}
+%\end{figure}
+
+\begin{figure}
 	\centering
-		\includegraphics[scale=0.5]{results/testsym.eps}
-	\caption{Symbols Comparison} The effect of symbol complexity.  %The graph shows the recognition rate of each symbol using different algorithms. 
+		\includegraphics[scale=0.35]{paperImages/testsym.jpg.eps}
+	\caption{Symbols Comparison on HS-DB} The effect of symbol complexity.  %The graph shows the recognition rate of each symbol using different algorithms. 
 	\label{fig:test2}
-\end{figure*}  
+\end{figure}
+
+\begin{figure}
+	\centering
+		\includegraphics[scale=0.5]{paperImages/LDsymbols.jpg.eps}
+	\caption{Symbols Comparison on LD-DB symbols}.  %The graph shows the recognition rate of each symbol using different algorithms. 
+	\label{fig:LDtest2}
+\end{figure}
+
+\begin{figure}
+	\centering
+		\includegraphics[scale=0.1]{paperImages/LDsymbols.jpg.eps}
+	\caption{Symbols Comparison on EL-DB symbols}.  %The graph shows the recognition rate of each symbol using different algorithms. 
+	\label{fig:ELtest2}
+\end{figure}
+
 The second experiment we implemented was to investigate the effect of symbol complexity and type on the recognition rate.
@@
 	\centering
-		\includegraphics[scale=0.5]{results/testFeat.eps}
-	\caption{Feature Comparison} The effect of different features on accuracy on \textsl{Hs-DB}.  %The graph shows the recognition rate of each symbol using different algorithms. 
-	\label{fig:testFeaturesAllHS}
+		\includegraphics[scale=0.5]{paperImages/featuresdifference.jpg.eps}
+	\caption{Feature Comparison on Hs-DB} The effect of different features on accuracy.  %The graph shows the recognition rate of each symbol using different algorithms. 
+	\label{fig:testFeaturesAll}
 \end{figure}
@@
 	\centering
-		\includegraphics[scale=0.5]{results/testF1.eps}
-	\caption{Feature Comparison} The effect of different features on accuracy on \textsl{MD-DB}.  %The graph shows the recognition rate of each symbol using different algorithms. 
-	\label{fig:testFeaturesAllMD}
+		\includegraphics[scale=0.5]{paperImages/featuresLD.jpg.eps}
+	\caption{Feature Comparison on LD-DB} The effect of different features on accuracy.  %The graph shows the recognition rate of each symbol using different algorithms. 
+	\label{fig:testFeaturesAll}
 \end{figure}
+
+ \begin{figure}
+	\centering
+		\includegraphics[scale=0.5]{paperImages/featuresdifference.jpg.eps}
+	\caption{Feature Comparison on EL-DB} The effect of different features on accuracy.  %The graph shows the recognition rate of each symbol using different algorithms. 
+	\label{fig:testFeaturesAll}
+\end{figure}
+
+ The main features affected by the segmentation stage are the geometric features computed in\textbf{FS1 }.
+Hence those are the features we used to estimate the efficiency of the segmentation algorithm.
+The recognition accuracy using only the geometrical features is reported in Figure \ref{fig:testFeatonly} .
+The figure shows that \textsl{AlgS2} achieve better results.
+%effect of the segmentation algorithm
+ %Figure \ref{fig:SampleSeg} shows a sample of the output of the segmentation block.
+The result of segmentation block is tested by their effect on the final recognition results using the features that will be highly affected using different algorithms like \textbf{FS1} and \textbf{FS4} (Section\ref{sec:Recognition}).
+Figure \ref{fig:testFeatonly} shows the result of the different algorithms using only \textbf{FS1 }features.
+ \begin{figure}
+	\centering
+		\includegraphics[scale=0.4]{paperImages/featureFS1only.jpg.eps}
+	\caption{The effect of segmentaion on Hs-DB accuracy} %The effect of symbol complexity.  %The graph shows the recognition rate of each symbol using different algorithms. 
+	\label{fig:testFeatonly}
+\end{figure}  
+ \begin{figure}
+	\centering
+		\includegraphics[scale=0.4]{paperImages/featureFS1only.jpg.eps}
+	\caption{The effect of segmentaion on LD-DB accuracy} %The effect of symbol complexity.  %The graph shows the recognition rate of each symbol using different algorithms. 
+	\label{fig:LDtestFeatonly}
+\end{figure}  
+ \begin{figure}
+	\centering
+		\includegraphics[scale=0.4]{paperImages/featureFS1only.jpg.eps}
+	\caption{The effect of segmentaion on EL-DB accuracy} %The effect of symbol complexity.  %The graph shows the recognition rate of each symbol using different algorithms. 
+	\label{fig:ELtestFeatonly}
+\end{figure}  
+%
+% \begin{figure}
+%	\centering
+%		\includegraphics[scale=0.5]{results/testFeat.eps}
+%	\caption{Feature Comparison} The effect of different features on accuracy on \textsl{Hs-DB}.  %The graph shows the recognition rate of each symbol using different algorithms. 
+%	\label{fig:testFeaturesAllHS}
+%\end{figure}  
+%
+% \begin{figure}
+%	\centering
+%		\includegraphics[scale=0.5]{results/testF1.eps}
+%	\caption{Feature Comparison} The effect of different features on accuracy on \textsl{MD-DB}.  %The graph shows the recognition rate of each symbol using different algorithms. 
+%	\label{fig:testFeaturesAllMD}
+%\end{figure}
 
@@
 The result illustrated in this chapter shows the work done on the system and experiments performed.
-The results show that the ellipse detection and PSO swarm algorithms gives better performance than other known systems.
+We showed result of three different dataset.
+The system performed better than most known algoirthm on the EL-DB and LD-DB.
+This is due to the focus on strcutre and stroke based properties gained from the sytstem after correctly segmenting the input strokes.
+The result also showed that even using \textsl{AlgS2} only as the segmentaion aglgoirthm the system perform better than other segmenation algorithm.
+Using different databases shows that the more the strucutre difference in the sysmobl the better recogniton rate.
+LD-DB proved to have the lowest recogntion rate due to the high simillarity between its symbols.
+% better than
 

### 2009-04-19 draft/System.tex
results after change i mad for paper2

@@
 The stroke points $p_i(x,y)$ that correspond to those maximum values are labeled as \textit{possible dominant points} $P_{pd}$.
-Figure \ref{fig:MaxRegioi} shows an example of $Region_i$ of the speed curve and thedominantpoints $P_{pd}$ which correspond to the minimum of point in this region $Min(Region_i)$.
+Figure \ref{fig:MaxRegioi} shows an example of $Region_i$ of the speed curve and the dominant points $P_{pd}$ which correspond to the minimum of point in this region $Min(Region_i)$.
 % (as shown are redundant)
@@
 		\includegraphics{images/ppd.eps}
-	\caption{Possible dominant points}% a) Possibledominantpoint b) Particle encoding  
+	\caption{Possible dominant points}% a) Possible dominant point b) Particle encoding  
 	\label{fig:LabelsPPD}
@@
 \label{sec:EllipseDetection}
-%Describe the ellipse detection algorithms 
+
+%The process starts by computing the center of the stroke bounding box.
+The bounding box center point is used as the first estimation of the ellipse center.
+The axes of the ellipse are estimated as $width/2$ and $height/2$ of the stroke bounding box.
+The least square fitting algorithm is used to minimize the fitting error of the ellipse equation \cite{chernov-2003}.
 This process tries to fit the stroke points into an ellipse arc; it starts with computing the center of the stroke bounding box.
 The bounding box center point is used as the first estimation of the center of the ellipse.
-The axes of the ellipse are estimated as the $w/2$ and $h/2$ of the stroke bounding box where $w$ is width and $h$ is height of the bonding box.
+The axes of the ellipse are estimated as the $width/2$ and $height/2$ of the stroke bounding box where $width$ is width and $height$ is height of the bonding box.
 The least square fitting algorithm \cite{chernov-2003} is used to minimize the fitting error of the ellipse Equation (\ref{eq:circleFit})
@@
 A list of new values for $x_0$ , $y_0$ ,$a$ and $b$ are generated randomly from the older values with small increments after each loop.
-After few loops, the final fit error and the ellipse confidence value are computed.
-The confidence value $Conf$ computed by Equation (\ref{eq:circleError})
+After few iteration, the final fit error of the estimated ellipse is reported.
+Another measure is used to compute the efficiency of the final estimated ellipse.
+Equation \ref{eq:circleError} ensures that the drawn percentage of ellipse is considered.
+This eliminates fitting a  line into very large ellipse but leaves small ellipses to be fitted as a partial or full ellipse.
+
   \begin{equation}
-Conf = E_{ellipse}+E_{area}
+eff = \left\{ {\begin{array}{*{20}c}
+   {\frac{{P_{percent} }}{E}} & {ifP_{percent}  < 0.5}  \\
+   {\frac{{P_{percent} }}{{E \times 2}}} & {ifP_{percent}  \ge 0.5}  \\
+\end{array}} \right.
+%eff= \begin{cases} 
+% \frac{P_{percent}}{E},&\mbox{if} P_{percent}\mbox{< 0.5} \\
+% \frac{P_{percent}}{2\times E},&\mbox{if} P_{percent}\mbox{\geq0.5} 
+%\end{cases}
 \label{eq:circleError}
@@
  \begin{equation}
-E_{area}  = \sqrt {\left( Area(stroke) - Area(ellipse)\right)^2 }
+P_{percent}  = L_{stroke} /P_{ellipse} 
 \label{eq:ErrorArea}
 \end{equation}
+where $E$ is the error computed by Equation(\ref{eq:circleFit}), $L_{stroke}$ total length of stroke and $P_{ellipse}$ is the perimeter of the estimated ellipse.
+Figure \ref{fig:ellipseFitExamples} shows different $eff$, $P_{ellipse}$  and  $E$ values of strokes drawn by users.
+If $eff$ is more than threshold $th_{Ellipse}$\footnote{By trial and error best threshold found was $th_{Ellipse}=0.25$} then the stroke is segmented as an ellipse otherwise the system proceeds to the next step.
 
-   where $E_{ellipse}$ is the error computed by Equation(\ref{eq:circleFit}) and $E_{area}$ is computed as the difference between the actual area of the stroke and the area of the estimated ellipse Equation(\ref{eq:ErrorArea}).
+\begin{figure}
+	\centering
+		\includegraphics[scale=0.7]{images/ellipseFitError.eps}
+	\caption{Ellipse Fitting Error}
+	\label{fig:ellipseFitExamples}
+\end{figure}
+
+%Describe the ellipse detection algorithms 
+%This process tries to fit the stroke points into an ellipse arc; it starts with computing the center of the stroke bounding box.
+The bounding box center point is used as the first estimation of the center of the ellipse.
+The axes of the ellipse are estimated as the $w/2$ and $h/2$ of the stroke bounding box where $w$ is width and $h$ is height of the bonding box.
+The least square fitting algorithm \cite{chernov-2003} is used to minimize the fitting error of the ellipse Equation (\ref{eq:circleFit})
+%
+%\begin{equation}
+%E = \sum\limits_{i = 0}^N {\frac{{(x_i - x_0 )}}{{a^2 }}^2  + \frac{{(y_i - y_0 )}}{{b^2 }}^2  - 1} 
+%\label{eq:circleFit}
+%\end{equation}
+%
+% where $N$ is number of points in the stroke, $a,b$ are the length of ellipse axes, $x_0$ \& $y_0$ are the coordinates of the center point, $x_i$ \& $y_i$ are the coordinates of point $i$ in the stroke.
+A list of new values for $x_0$ , $y_0$ ,$a$ and $b$ are generated randomly from the older values with small increments after each loop.
+After few loops, the final fit error and the ellipse confidence value are computed.
+The confidence value $Conf$ computed by Equation (\ref{eq:circleError})
+% \begin{equation}
+%Conf = E_{ellipse}+E_{area}
+%\label{eq:circleError}
+%\end{equation}
+% \begin{equation}
+%E_{area}  = \sqrt {\left( Area(stroke) - Area(ellipse)\right)^2 }
+%\label{eq:ErrorArea}
+%\end{equation}
+%
+%   where $E_{ellipse}$ is the error computed by Equation(\ref{eq:circleFit}) and $E_{area}$ is computed as the difference between the actual area of the stroke and the area of the estimated ellipse Equation(\ref{eq:ErrorArea}).
 If confidence value $Conf$ is less than threshold $th_{Ellipse}$\footnote{ Using trial and error the best threshold used is $th_{Ellipse}=0.7$ } the stroke is segmented as an ellipse otherwise the system proceeds to the two DPSO segmentation algorithms.
@@
 % where $p_{ij}$ is the numerical values of the particle and $r_{3}$ is a random variable, is used to change the general swarm algorithm into binary particle (\textit{Discrete Particle Swarm Algorithm DPSO}) which handles particle values of either $0$ or $1$ \cite{PSODisceret}.
-\subsection{Swarm Segmentation}
+\subsection{Non Ellipse Fitting}
 \label{sec:SwarmSegmentation}
@@
 Thus , the particle represents points that are chosen to be used as dominant points for this segmentation.
-The goal of the DPSO algorithm is to find the solution $P_i$ that generates the minimum set ofdominantpoints that defines the stroke with minimum segmentation error.
+The goal of the DPSO algorithm is to find the solution $P_i$ that generates the minimum set of dominant points that defines the stroke with minimum segmentation error.
 In other words, the system tries to find the fewer number of points (1's) location in the particle array which gives minimum segmentation error.
@@
   \includegraphics[scale=0.8]{images/pso1.eps}			
-	\caption{AlgS1 Error}% a) Possibledominantpoint b) Particle encoding  
+	\caption{AlgS1 Error}% a) Possible dominant point b) Particle encoding  
 	\label{fig:DPSOERROR}
@@
 %Alg2: 
-\subsubsection{Curve Divisions Algorithm \textsl{AlgS2}}
+\subsubsection{Hybird Curve Algorithm \textsl{AlgS2}}
 \label{sec:PolygonDivisionAlgorithm}
@@
 After that the particles are tested to make sure that the distance between every two successive dominant point is larger than the constant $min_D$.
-If two points are nearer than $min_D$\footnote{Constant selected to be 10 pixels} then one of the points is removed.
+If two points are nearer than $min_D$\footnote{ 5\% of the total stroke length} then one of the points is removed.
 
-\subsection{Choosing the best fit}
+\subsection{Choosing the Best Fit}
 The two PSO algorithms produce two segmentation solutions.
@@
 
-\section{ Feature Extraction}%%Feature Set
+\section{Feature Extraction Methods}%%Feature Set
 \label{sec:FeatureExtraction}
 
-The system uses a composite set of features include Rubine feature set \cite{gestureexample12}, Zernike moments \cite{HeloiseBeautification}, ink density \cite{GeometryAndDomain102} and some structural information like number of perpendiculars lines , number of parallel lines and types of primitives in each symbol.
+The system uses a composite set of features which include global shape properties like Rubine feature set \cite{gestureexample12},  ink density \cite{GeometryAndDomain102}, some appearance based properties like Zernike moments \cite{HeloiseBeautification}, and some stroke based structural information like number of perpendiculars lines , number of parallel lines and types of primitives in each symbol.
 The following list gives details on all the features used in the system.
@@
 \begin{description}
- \item [Structural and geometrical Features(FS1)] Features defines the structure of the geometrical symbol.  
+ \item [Structural and geometrical Features (FS1)] Features defines the structure of the geometrical symbol.  
 \begin{description}
@@
 		\item [X intersections] Number of X intersections.
-		\item [Size Ratio] Ratio between width to height of the symbol.
 
@@
 \item [Rubine Feature Set (FS2)] Features introduced by Rubine\cite{gestureexample12} for single stroke gestures.
+These features represent the global shape of the symbol drawn.
 To compute the features we append all segments points into single path $ink_{path}$.
@@
 \end{description}
-	\item[Composite Features (FS4)] Features that are composites of different geometrical and statistical symbol. 
+	\item[Composite Features (FS4)] Features that are composites of different geometrical and statistical symbol. These features represent the global shape of the symbol drawn.
 	\begin{description}
-
+	\item [Size Ratio] Ratio between width to height of the symbol.
 	\item [Ink density] Compute the density of points inside its bounding box\cite{GeometryAndDomain102}.   

### 2009-04-19 trunk/paper2/Paper2.tex
today is 19 appril

@@
 \begin{itemize}
-	\item \emph{Zernike moments } Zernike moments of order \cite{HeloiseBeautification}. 
+	\item \emph{Zernike moments } Zernike moments of order n (10 or 20).  \cite{HeloiseBeautification}. 
 \end{itemize}
@@
 \label{sec:Experiments}
-The dataset collected by Hse and Newton\cite{HeloiseBeautification} is used to test the presented system.
+In our experiment we used three different datasets.
+First dataset was collected by Hse and Newton\cite{HeloiseBeautification} (Hs-DB).
+The data are drawn by 16 users each of them was asked to draw 13 shapes from 30 to 50 times.
+Figure \ref{fig:symbolSet} shows a set of the shapes used in the data set.
+We collected the second and third datasets.
+Figure \ref{fig:ELsymbolSet} and \ref{fig:LogicsymbolSet} shows the symbols gathered from 7 different users in electrical (EL-DB) and logic design (LD-DB) domain respectively.
+Table \ref{tab:datasets} shows a  comparison between the three datasets used in our experiments.
+The datasets were divided into training set and test set.
+Five different splits  for the training and test data are generated from each dataset.
+The results displayed are the average recognition accuracy of the five splits.
+The accuracy is computed as the number of correctly recognized samples divided by the total number of samples in the test.
+
+\begin{figure}
+\centering 
+\label{fig:symbolSet}
+\fbox{ \parbox{5cm}{% 
+		\includegraphics[scale=0.5]{images/symbolSet.PNG}	}}
+		\caption{The Hs-DB Symbol Set}.
+\end{figure}
+\begin{figure}
+\centering 
+\label{fig:LogicsymbolSet}
+
+		\includegraphics[scale=0.5]{images/logicSet.jpg}	 
+		\caption{The Digital Design Symbols Dataset(LD-DB)} Symbols used in the dataset.
+\end{figure}
+\begin{figure}
+\centering 
+\label{fig:ELsymbolSet}
+\fbox{ \parbox{4.5cm}{% 
+		\includegraphics[scale=0.5]{images/EelectImage.jpg}	}}
+		\caption{The Electrical Symbols Dataset(EL-DB)} .
+\end{figure}
+
+\begin{table}
+\begin{center}
+\scalebox{0.9}{
+ \begin{tabular}{|p{2cm}|p{1.5cm}|p{1.5cm}|p{1.5cm}|}
+ \hline
+ & Hs-DB &  EL-DB & LD-DB \\ \hline
+No of samples  &  7791 & 2764 & 1859 \\ \hline
+No of categories  & 13 & 15 & 8 \\ \hline
+Avg. Samples per Category& 600 & 184 & 232 \\\hline 
+No of users  & 16 & 7 & 7 \\  \hline
+Balanced & Yes&No & No \\ \hline
+Splits  & 5 random splits  & 5 random splits  & 5 random splits  \\ \hline
+Examples & Figure \ref{fig:symbolSet}  & Figure \ref{fig:ELsymbolSet}  & Figure \ref{fig:LogicsymbolSet}   \\ \hline
+\end{tabular}
+}
+\caption[Datasets Comparisons]{Datasets Comparisons} 
+\label{tab:datasets}
+
+\end{center}
+\end{table}
+
+%The dataset collected by Hse and Newton\cite{HeloiseBeautification} is used to test the presented system.
 The data are drawn by 16 users each of them was asked to draw 13 shapes from 30 to 50 times.
@@
 %In the following section, the recognition accuracy means the average recognition accuracy in the four data splits.
-\begin{figure}
-\centering 
-\label{fig:symbolSet}
-\fbox{ \parbox{5cm}{% 
-		\includegraphics[scale=0.5]{images/symbolSet.PNG}	}}
-		\caption{The Symbol Set} Symbols used in the dataset.
-\end{figure}
+
 \subsection{Algorithms comparisons}
@@
 	\centering		
-	 \includegraphics[scale=0.3]{images/swarmTesting.jpg}
+	 \includegraphics[scale=0.3]{images/swarmAgentTest.jpg.eps}
 	 	\caption{Swarm testing} Segmentation Error Vs. Number of agents
@@
 
-\begin{figure*}
+\begin{figure}
 	\centering		
-	 \includegraphics[width=18cm]{images/swarmTesting2.jpg}
+	 \includegraphics[scale=0.4]{images/swarmParamterC.jpg.eps}
 	 	\caption{Swarm testing} Segmentation Error Vs. PSO parameters.
@@
 
-\end{figure*} 
+\end{figure}
+
 We performed several experiments to evaluate the presented recognition system.
@@
 We also implemented the segmentation algorithm described in \cite{earlyprocess} (\textsl{Alg3}) to use as reference to compare it with our swarm algorithms.
-Figure \ref{fig:test1} shows the accuracy achieved by each algorithm.
+Figure \ref{fig:test1} and \ref{fig:testLD} shows the accuracy achieved by each algorithm.
 The two swarm algorithms were tested with and without the ellipse fitting module.
@@
 	 \includegraphics[scale=0.4]{images/testAlg.jpg}
-	 	\caption{Algorithm comparison} The recognition rate of different algorithms. 
+	 	\caption{Algorithm comparison on HS-DB} The recognition rate of different algorithms. 
 	 	\label{fig:test1}
@@
 
+\begin{figure}
+	\centering		
+	 \includegraphics[scale=0.4]{images/LDAlg.jpg}
+	 	\caption{Algorithm comparison on LD-DB} The recognition rate of different algorithms. 
+	 	\label{fig:testLD}
+	%\caption{Experiments results :} a)    b) c) 
+\end{figure} 
 \subsection{Shape Complexity Analysis}
@@
 		\includegraphics[scale=0.5]{images/testsym.jpg}
-	\caption{Symbols Comparison} The effect of symbol complexity.  %The graph shows the recognition rate of each symbol using different algorithms. 
+	\caption{Symbols Comparison on HS-DB} The effect of symbol complexity.  %The graph shows the recognition rate of each symbol using different algorithms. 
 	\label{fig:test2}
 \end{figure*}
+
+\begin{figure*}
+	\centering
+		\includegraphics[scale=0.5]{images/LDsymbols.jpg}
+	\caption{Symbols Comparison on LD-DB symbols}.  %The graph shows the recognition rate of each symbol using different algorithms. 
+	\label{fig:LDtest2}
+\end{figure*}  
 The second experiment we implemented was to investigate the effect of symbol complexity and type on the recognition rate.
-Figure \ref{fig:test2} shows the achieved accuracy of each symbol by our algorithms compared to \cite{earlyprocess} (\textsl{Alg3}).
+Figure \ref{fig:test2} and \ref{fig:LDtest2} shows the achieved accuracy of each symbol by our algorithms compared to \cite{earlyprocess} (\textsl{Alg3}).
 It is clearly noted that symbols that have only line segments achieve higher accuracy rate than other symbols.
@@
 This is understandable because algorithm \textsl{AlgS1} divides strokes into line segments only but \textsl{AlgS2} is able to divide strokes into lines and curves based on the minimum error of the segment itself.
+Figure \ref{fig:LDtest2} shows that in logic design symbols \textsl{AlgS2} achieved better performance than \textsl{Alg3} \textsl{AlgS1} this emphases that complex shapes that are combination of lines and arc are better segmented using \textsl{AlgS2}.
 Algorithm \textsl{Alg3} gives good performance as long as the symbols consist of combination of lines and curves, if the stroke consists of only curve or lines the algorithm may lead to wrong segmentation result.
@@
  	%}
-	\caption{Confusion Matrix}
+	\caption{Confusion Matrix of Hs-DB}
 	\label{tab:ConfusionMatrix}
@@
 \end{table*}
+% \begin{table*}
+% 	\centering
+% 	%\small
+%	\scalebox{0.7}{
+% 		%	\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|c|c|c|c|}\hline 
+%  \include{conf2}
+%  		%\end{tabular}
+%  	%}
+% 	\caption{Confusion Matrix of DL-DB}
+% 	\label{tab:ConfusionMatrix}
+% 	%\end{minipage}
+% \end{table*}
 \subsection{Features Analysis}
@@
 	\centering
-		\includegraphics[scale=0.5]{images/testFeat.jpg}
+		\includegraphics[scale=0.5]{images/featuresdifference.jpg}
 	\caption{Feature Comparison} The effect of different features on accuracy.  %The graph shows the recognition rate of each symbol using different algorithms. 
@@
 	\centering
-		\includegraphics[scale=0.4]{images/testF1.jpg}
+		\includegraphics[scale=0.4]{images/featureFS1only.jpg}
 	\caption{} %The effect of symbol complexity.  %The graph shows the recognition rate of each symbol using different algorithms. 
@@
 Global shape and moment properties was used to prevent the need to use graph matching which is computationally expensive.
-The system was tested on 13 different symbols and achieved an overall accuracy of 93.5\%.
+The system was tested on 13 different symbols and achieved an overall accuracy of 89.5\%.
 The system does not depend on low level recognizer but rather on a set of high level features.

### 2009-04-20 trunk/draft/Results.tex
second commit

@@
 
-The system has two main parts the DPSO stroke segmenation and symbol recogntion.
-Each part is evaluated sperately.
+The system has two main parts the DPSO stroke segmentation and symbol recognition.
+Each part is evaluated separately.
 The preprocessing procedures and computation is evaluated using their effect on the segmentation algorithm.
 These results are reported in section \ref{sec:ResultsDetails}.
-The effect of number of agents, the error thresholds and various other PSO paramters are investigated in Section \ref{sec:PSO}.
+The effect of number of agents, the error thresholds and various other PSO parameters are investigated in Section \ref{sec:PSO}.
 %The results presented in Section \ref{sec:ResultsDetails} and \ref{sec:PSO}.
 
-The recognition system evalutation is presented in Section \ref{sec:SegmentationAlgorithms}.
-The recogntion rate (computed as the number of correctly recognized samples divided by the total number of samples in the test) for each segmentation algorithm compared to another segmenation algorithm are reported in \ref{sec:recognitionAlgorithms}.
-Next, a comparison between different shapes and datasets and it influence on the recogntion rate is presented in Section \ref{sec:EffectsofSymbolComplixity}.
-Finally, a  comparision between different feature vectors is reported in Section \ref{sec:featuresComparisions}.
+The recognition system evaluation is presented in Section \ref{sec:SegmentationAlgorithms}.
+The recognition rate (computed as the number of correctly recognized samples divided by the total number of samples in the test) for each segmentation algorithm compared to another segmentation algorithm are reported in \ref{sec:recognitionAlgorithms}.
+Next, a comparison between different shapes and datasets and it influence on the recognition rate is presented in Section \ref{sec:EffectsofSymbolComplixity}.
+Finally, a comparison between different feature vectors is reported in Section \ref{sec:featuresComparisions}.
 
@@
 \begin{description}
-	\item [Hse Dataset] (denoted as Hs-DB) Data set collected by \cite{HeloiseBeautification} and was previously used by \cite{Oltmans07} as benchmarck for sketch recognition problem.
-The data are drawn by 16 users each of them have drawn from 30 to 50 samples for each shape.
+	\item [Hse Dataset] (denoted as Hs-DB) Data set collected by \cite{HeloiseBeautification} and was previously used by \cite{Oltmans07} as benchmark for sketch recognition problem.
+The data are drawn by 16 users each of them have drawn from 50 to 60 samples for each shape.
 Figure \ref{fig:symbolSet} shows the shapes used in the data set.
-The dataset was divided into training set and test set.
-Four different splits for the training and test data are generated from the dataset.
-In worth noting that \cite{HeloiseBeautification} and \cite{Oltmans07} achieved 96.7\% and  94.4\%  on this dataset respecitvily.
+The dataset was divided into 75\% training set and 25\% test set.
+Five different splits for the training and test data are generated from the dataset.
+In worth noting that \cite{HeloiseBeautification} and \cite{Oltmans07} achieved 96.7\% and  94.4\%  on this dataset respectively.
 %, 94.4\%  , which
 %is comparable to the 96.7%  %The results displayed are the average recognition accuracy of the four splits resulted from the classifier after recognizing the symbols.
-		\item [MultiDomain Dataset] (denoted as MD-DB) A dataset we collected manually from different users.
-The dataset contains symbols from three diffrent domains ( UML, mechanical and electrical engineering symbols).
-Data are collected from 4 different users each of them have drawn from 20 to 40 samples of each shapes.
+		\item [Multi Domain Dataset] (denoted as MD-DB) A dataset we collected manually from different users.
+The dataset contains symbols from two different domains Electrical Engineering symbols (EL-DB) and logic design symbols (LD-DB).
+Data are collected from 7 different users each of them have drawn from 20 to 40 samples of each shapes.
 Figures \ref{fig:LogicSet} and \ref{fig:ElectSet} show some of the shapes used in the dataset.
-Symbols from different domains were gathered to evaluate the system thourhgtout different types of shapes.
-Useres were requested to draw symbols without any restriction on style,size or orientations.
-The dataset was divided into training set and test set.
-Four different splits for the training and test data are generated from the dataset.
-%The results displayed are the average recognition accuracy of the four splits resulted from the classifier after recognizing the symbols.
+Symbols from different domains were gathered to evaluate the system throughout different types of shapes.
+Users were requested to draw symbols without any restriction on style, size or orientations.
+The dataset was divided into 75\% training set and 25\% test set.
+Five different splits for the training and test data are generated from the dataset.
+%In worth noting that \cite{Oltmans07} gathered simillar dataset and acheived 89\% on it   %The results displayed are the average recognition accuracy of the four splits resulted from the classifier after recognizing the symbols.
 \end{description}
+
+Table \ref{tab:datasets} shows a  comparison between the three datasets used in our experiments.
+
  %We divided the dataset equally into training set and test set.  The results displayed are the recognition accuracy resulted from the classifier after recognizing the symbols. \\
@@
 \begin{figure}[]\centering
-\fbox{  \parbox{6cm}{% 
-		\includegraphics[scale=0.7]{images/logicSymbols.eps}	}
-		}
+
+		\includegraphics[scale=0.9]{paperImages/logicSet.jpg.eps}
+
 	\caption{Example of Logic Symbols}
@@
 \begin{figure}[]\centering
-\fbox{  \parbox{8cm}{% 
-		\includegraphics[scale=0.7]{images/SymbolsElectricalr.eps}	}
-		}
+
+		\includegraphics[scale=0.9]{paperImages/EelectImage.jpg.eps}
+
 	\caption{Example of Electrical Symbols}
@@
 \end{figure}
+\begin{table}
+\begin{center}
+\scalebox{0.9}{
+ \begin{tabular}{|p{5cm}|p{2.5cm}|p{2.5cm}|p{2.5cm}|}
+ \hline
+ & Hs-DB &  EL-DB & LD-DB \\ \hline
+No of samples  &  7791 & 2764 & 1859 \\ \hline
+No of categories  & 13 & 15 & 8 \\ \hline
+Avg. Samples per Category& 600 & 184 & 232 \\\hline 
+No of users  & 16 & 7 & 7 \\  \hline
+Balanced & Yes&No & No \\ \hline
+Splits  & 5 random splits  & 5 random splits  & 5 random splits  \\ \hline
+Examples & Figure \ref{fig:symbolSet}  & Figure \ref{fig:ELsymbolSet}  & Figure \ref{fig:LogicsymbolSet}   \\ \hline
+\end{tabular}
+}
+\caption[Datasets Comparisons]{Datasets Comparisons} 
+\label{tab:datasets}
+
+\end{center}
+\end{table}
 
@@
 
-Figrue \ref{fig:SegErrPreCal} shows the \textit{Average Segmentation Error} using different preliminary calculation (Section \ref{sec:CurvatureCalculation}).
+%Figure \ref{fig:SegErrPreCal} shows the \textit{Average Segmentation Error} using different preliminary calculation (Section \ref{sec:CurvatureCalculation}).
 The figure shows that using all computations (time difference, direction, speed and curvature) for each stroke decrease the Segmentation error.
 Figure \ref{fig:PossibledpPreCalculation} shows the average number of Possible dominant point $P_{pd}$ with respect to information used in the system.
-The results shows that the using all information increases the number of Possible dominate point $P_{pd}$ which in segmenation stage decreses the \textit{Segmenation Error}.
+The results shows that the using all information increases the number of Possible dominate point $P_{pd}$ which in segmentation stage decreases the \textit{Segmentation Error}.
 It is worth noting that when using all information the number of Possible dominant point $P_{pd}$ is not the same as the sum of using every information alone.
-This is because we found out that there was some redundant points found using more than one information.
+This is because we found out that there was some redundant points found using more than one information source.
+
+%\begin{figure}
+%	\centering
+%		\includegraphics[scale=0.5]{results/SegErrPreCal.eps}
+%	\caption{preliminary calculation effect}
+%	\label{fig:SegErrPreCal}
+%\end{figure}
+% 
+% 
+%\begin{figure}
+%	\centering
+%		\includegraphics[scale=0.5]{results/PossibledpPreCalculation.eps}
+%	\caption{Possible Dominant Point Count}
+%	\label{fig:PossibledpPreCalculation}
+%\end{figure}
+
+ The effect of number of agents, the error thresholds and various other parameters was investigated.
+The result of these tests are shown in Figure \ref{fig:swarmtestingS1} and \ref{fig:swarmtestingS2}.
+Figure \ref{fig:swarmtesting} shows that as the number of agents increase the \textit{Segmentation Error} value decrease.
+Similar results for the maximum swarm iterations are shown in Figure \ref{fig:swarmtesting1}.
+Other algorithm parameters like $c_1$,$c_2$ that were mentioned in Section\ref{sec:ParticleSwarmAlgorithm} are tested using similar tests (Figure \ref{fig:swarmparamterw.jpg} and \ref{fig:swarmtesting2}).
+The final values are for maximum number of iteration = 80 with 15 particles.
+The final number of agents and maximum swarm iteration used in the system was based trade off between errors achieved versus the computation time.
 
@@
 	\centering		
-		\includegraphics[scale=0.5]{results/SegErrPreCal.eps}
-	\caption{preliminary calculation effect}
-	\label{fig:SegErrPreCal}
-\end{figure}
-
-\begin{figure}
-	\centering
-		\includegraphics[scale=0.5]{results/PossibledpPreCalculation.eps}
-	\caption{Possible Dominant Point Count}
-	\label{fig:PossibledpPreCalculation}
-\end{figure}
-
- The effect of number of agents, the error thresholds and various other paramters was investigated.
-The result of these tests are shown in Figure \ref{fig:swarmtestingS1} and  \ref{fig:swarmtestingS2}.
-Figure \ref{fig:swarmtesting} shows that as the number of agents increases the \textit{Segmenation Error} value decrease.
-Simillar results for the maximum swarm iterations are shown in Figure \ref{fig:swarmtesting1} .
-Other algorithm parameters like $c_1$,$c_2$ that were mentioned in Section\ref{sec:ParticleSwarmAlgorithm} are tested using similar tests (Figure \ref{fig:swarmtesting} and \ref{fig:swarmtesting2}).
-The final values are for maximum number of iteration = 80 with 15 particles.
-The final number of agents and maximum swarm iteration used in the system was based trade off between error achieved versus the computation time.
-
- \begin{figure}
-	\centering		
-	 \includegraphics[scale=0.3]{results/swarmalg1.eps}
-	 	\caption{Segmentation Error Vs. Maximum iterations For ALgS1}
-	 	\label{fig:swarmtestingS1}
+	 \includegraphics[scale=0.5]{paperImages/swarmAgentTest.jpg.eps}
+	 	\caption{Segmentation Error Vs. Number of agents}
+	 	\label{fig:swarmtesting}
 	%\caption{Experiments results :} a)    b) c) 
@@
 The result of the segmentation algorithm can be viewed in the Figure \ref{fig:results1}.
-These result shows the originals stroke and the segmentation that the system generated for the stroke.
-The figure \ref{fig:results2} shows the result of \textsl{AlgS1} and \textsl{AlgS2} for the same input strokes.
-Figure \ref{fig:results3} shows the ouptut of \textsl{Alg3} to the same set of input strokes.
-It clrearly shows that \textsl{AlgS2} have better segmentation results than both \textsl{AlgS1} and \textsl{Alg3} but since these segmentation are subjective to what the user intended to draw the next section will focus on the effect of the segmenation algorithm on the recognition accuracy.
+This result shows the originals stroke and the segmentation that the system generated for the stroke.
+Figure \ref{fig:results2} shows the result of \textsl{AlgS1} and \textsl{AlgS2} for the same input strokes.
+Figure \ref{fig:results3} shows the output of \textsl{Alg3} to the same set of input strokes.
+It clearly shows that \textsl{AlgS2} have better segmentation results than both \textsl{AlgS1} and \textsl{Alg3} but since these segmentation are subjective to what the user intended to draw the next section will focus on the evaluating the segmentation algorithms using the effect on the recognition accuracy.
 
@@
 	\centering
-				\subfigure{\includegraphics[scale=0.55]{results/result1.eps}}
+				\subfigure{\includegraphics[scale=0.45]{results/result1.eps}}
 					\subfigure{	\includegraphics[scale=0.6]{results/result1logic.eps}}
@@
 	\centering
-			\subfigure{	\includegraphics[scale=0.55]{results/result2.eps}}
+			\subfigure{	\includegraphics[scale=0.45]{results/result2.eps}}
 			\includegraphics[scale=0.6]{results/result2logic.eps}
@@
 		\subfigure{
-		\includegraphics[scale=0.55]{results/result3.eps}}
+		\includegraphics[scale=0.45]{results/result3.eps}}
 			\subfigure{\includegraphics[scale=0.55]{results/result3logic.eps}}
@@
 
-\subsection {Recognition Algorithms}
+\subsection {Recognition system}
 \label{sec:RecognitionAlgorithms}
 
-e performed several experiments to evaluate the presented recognition system.
+several expriments is presented to test the different segmentation algoirthms on the three data set used in our system.
+Algorithms (\textsl{AlgS1}, \textsl{AlgS2} and \textsl{Alg3} (in section \ref{sec:BenchMarckAlgorithm})) were tested individualy with and without the ellipse fitting module.
+The final system (Choosing the Minimum error of (\textsl{AlgS1}, \textsl{AlgS2}) was evaluated on the three dataset(HS, EL and LD datasets ).
+Figure \ref{fig:test1},\ref{fig:testEL} and \ref{fig:testLD} shows the accuracy achieved by each algorithm.
+Theses expriments were done using the whole dataset as 75\% of the data as testset and 25\% as training.
+The results shows that \textsl{AlgS2} with the ellipse fitting modeule gives better recognition rate accross all datasets (Figure \ref{fig:test1},\ref{fig:testEL} and \ref{fig:testLD} ).
+Eventhough, \textsl{AlgS1} gives better result than \textsl{Alg3} in both EL-DB and HS-DB it perform poorly in LD-DB due to the high number of curves in every symbol in this dataset.
+These result are proved using different number of samples per category as shown in Figures \ref{fig:LDRes}, \ref{fig:ElRes}.
+% i need to organize this as following. 
+% number 1. differnt algorithm on different data 
+% which algortihm is better oon which data 
+% on ld and el  the alg2  is better,
+
+%We performed several experiments to evaluate the presented recognition system.
 Firstly we tested recognition accuracy of shapes in the data set with both algorithms (\textsl{AlgS1} and \textsl{AlgS2}).
-We also implemented the segmentation algorithm described in \cite{earlyprocess} (\textsl{Alg3}) to use as reference to compare it with our swarm algorithms.
-Figure \ref{fig:test1} shows the accuracy achieved by each algorithm.
+We also implemented the segmentation algorithm described in \cite{earlyprocess} (\textsl{Alg3} in section \ref{sec:BenchMarckAlgorithm} )  to use as reference to compare it with our swarm algorithms.
+Figure \ref{fig:test1},\ref{fig:testEL} and \ref{fig:testLD}  shows the accuracy achieved by each algorithm.
 The two swarm algorithms were tested with and without the ellipse fitting module.
-The ellipse detection module appears to be superior.
+The ellipse fitting module appears to be superior.
 The result shows that combining both (\textsl{AlgS1} and \textsl{AlgS2}) out perform any single algorithm.
@@
 %improves the results.% with both the \textit{DPSO} algorithms.% The results show that both \textit{DPSO} algorithms achieve better result than other algorithms. 
+\begin{figure}
+	\centering		
+	 \includegraphics[scale=0.4]{paperImages/testAlg.jpg.eps}
+	 	\caption{Algorithm comparison on HS-DB} The recognition rate of different algorithms. 
+	 	\label{fig:test1}
+	%\caption{Experiments results :} a)    b) c) 
+\end{figure} 
+\begin{figure}
+	\centering		
+	 \includegraphics[scale=0.8]{results/AlgLD.eps}
+	 	\caption{Algorithm comparison on LD-DB} The recognition rate of different algorithms. 
+	 	\label{fig:testLD}
+	%\caption{Experiments results :} a)    b) c) 
+\end{figure}
 
@@
 	\centering		
-	 \includegraphics[scale=0.4]{results/testAlg.eps}
-	 	\caption{Algorithm comparison} The recognition rate of different algorithms. 
-	 	\label{fig:test1}
+	 \includegraphics[scale=0.8]{results/AlgEL.eps}
+	 	\caption{Algorithm comparison on EL-DB} The recognition rate of different algorithms. 
+	 	\label{fig:testEL}
 	%\caption{Experiments results :} a)    b) c) 
@@
 
-\subsubsection {Effects of Symbol Complixity }
+\begin{figure}
+	\centering
+		\includegraphics[scale=0.6]{results/LDRes.eps}
+	\caption{Result Algorithms on LD-DB}
+	\label{fig:LDRes}
+
+\end{figure}
+\begin{figure}
+	\centering
+		\includegraphics[scale=0.6]{results/ElRes.eps}
+	\caption{Results of Algorithms on EL-DB}
+	\label{fig:ElRes}
+\end{figure}
+
+\subsubsection {Evaluating Different Datasets}
  \label{sec:EffectsofSymbolComplixity}
 
-\begin{figure*}
+ The effect of symbol complixity in a dataset were testing in the following expriment.
+These expriments test how many samples will be required to finaly obtain good performance from the recognizier system.
+In the following expriments, we tested the system using different number of sample per category.
+ Figure \ref{fig:Alg1Ds} and \ref{fig:Alg2Ds} shows that Hs-DB requires more number of sample for traininig to achieve same result as in other data sets.
+This is due that it contains less signal and conceptual variation that the other two datasets.
+Figure \ref{fig:samplesofHS-DB.jpg} shows sample of data collected from one user in Hs-Db it is clear that it has reasnabole amount of variation between users but it is limited.
+Recognizing symbols that contains high variation in drawing style using fewer symbols in training prove the efficency of our system.
+ \begin{figure}
 	\centering
-		\includegraphics[scale=0.5]{results/testsym.eps}
-	\caption{Symbols Comparison} The effect of symbol complexity.  %The graph shows the recognition rate of each symbol using different algorithms. 
-	\label{fig:test2}
-\end{figure*}  
-The second experiment we implemented was to investigate the effect of symbol complexity and type on the recognition rate.
-Figure \ref{fig:test2} shows the achieved accuracy of each symbol by our algorithms compared to \cite{earlyprocess} (\textsl{Alg3}).
-It is clearly noted that symbols that have only line segments achieve higher accuracy rate than other symbols.
-The results indicate that algorithm \textsl{AlgS1} achieve better performance than algorithm \textsl{AlgS2} in the symbols that consist only of lines.
-This is understandable because algorithm \textsl{AlgS1} divides strokes into line segments only but \textsl{AlgS2} is able to divide strokes into lines and curves based on the minimum error of the segment itself.
-Algorithm \textsl{Alg3} gives good performance as long as the symbols consist of combination of lines and curves, if the stroke consists of only curve or lines the algorithm may lead to wrong segmentation result.
-This is because the system divides the stroke first to line segments then tries to decide if each segment can be represent better as a curve unlike algorithm \textsl{AlgS2} where the curve segments are tested while choosing the best segmentation.
-Combining both algorithm \textsl{AlgS1} and \textsl{AlgS2} improved the recognition rate of all symbols.
-The penalty for this improved performance is the computational time required to run both swarm algorithms.
-Table \ref{tab:ConfusionMatrix} shows the confusion matrix of symbols.
-The table shows that errors are only between two sets of symbols moon with triangles and trapezoid with triangles.
-This observation is understandable as the symbols are visually similar but the system must be able to differentiate between them.
-This lead us to the next experiment to choose best set of feature to recognize symbols.
+		\includegraphics{results/Alg1Ds.eps}
+	\caption{AlgS1 on different datasets}
+	\label{fig:Alg1Ds}
+\end{figure}
+\begin{figure}
+	\centering
+		\includegraphics{results/Alg2Ds.eps}
+	\caption{AlgS2 on different Datasets }
+	\label{fig:Alg2Ds}
+\end{figure}
+
+  Figure \ref{} and \ref{} shows examples of correctly recognizied symbol using our system.
+The figure shows that system recognizied some hard to recognize example in both EL-DB and DL-DB.
+An example of symbols that was not correctly recognized are in Figure \ref{}.
+Most of the wrong classification were due to how messsy the user draw symbols and how simillary they are conceptual.
+Table \ref{} show the confusion matrix of EL-DB which shows that most error were between capcatior, batter and crystall which differe only on number of parallel lines.
+Table \ref{tab:ConfusionMatrix} (confusion matrix) shows the types of error in Hs-DB which shows that the system was confused between parallegram and trapizoid, moon and archs.
+Those error were due to users drawing styles as some user may draw a parallegram as a trapizoid.
+
+\begin{figure}
+	\includegraphics[scale=0.5]{results/samplesofHS-DB.jpg.eps}
+	\caption{Hs-DB shapes drawn by single user}
+	\label{fig:samplesofHS-DB.jpg}
+\end{figure}
  \begin{table*}
@@
 
-\subsubsection{Features Comparisions}
+%\begin{figure}
+%	\centering
+%		\includegraphics[scale=0.5]{results/testsym.eps}
+%	\caption{Symbols Comparison} The effect of symbol complexity.  %The graph shows the recognition rate of each symbol using different algorithms. 
+%	\label{fig:test2}
+%\end{figure}
+
+Another experiment we implemented was to investigate the effect of symbol complexity and type on the recognition rate.
+Figure \ref{fig:test2} shows the achieved accuracy of each symbol by our algorithms compared to \cite{earlyprocess} (\textsl{Alg3}).
+It is clearly noted that symbols that have only line segments achieve higher accuracy rate than other symbols.
+The results indicate that algorithm \textsl{AlgS1} achieve better performance than algorithm \textsl{AlgS2} in the symbols that consist only of lines.
+This is understandable because algorithm \textsl{AlgS1} divides strokes into line segments only but \textsl{AlgS2} is able to divide strokes into lines and curves based on the minimum error of the segment itself.
+Algorithm \textsl{Alg3} gives good performance as long as the symbols consist of combination of lines and curves, if the stroke consists of only curve or lines the algorithm may lead to wrong segmentation result.
+This is because the system divides the stroke first to line segments then tries to decide if each segment can be represent better as a curve unlike algorithm \textsl{AlgS2} where the curve segments are tested while choosing the best segmentation.
+Combining both algorithm \textsl{AlgS1} and \textsl{AlgS2} improved the recognition rate of all symbols.
+The penalty for this improved performance is the computational time required to run both swarm algorithms.
+Table \ref{tab:ConfusionMatrix} shows the confusion matrix of symbols.
+The table shows that errors are only between two sets of symbols moon with triangles and trapezoid with triangles.
+This observation is understandable as the symbols are visually similar but the system must be able to differentiate between them.
+This lead us to the next experiment to choose best set of feature to recognize symbols.
+
+ \begin{figure}
+	\centering
+		\includegraphics[scale=0.35]{paperImages/testsym.jpg.eps}
+	\caption{Symbols Comparison on HS-DB} The effect of symbol complexity.  %The graph shows the recognition rate of each symbol using different algorithms. 
+	\label{fig:test2}
+\end{figure}
+
+\begin{figure}
+	\centering
+		\includegraphics[scale=0.5]{paperImages/LDsymbols.jpg.eps}
+	\caption{Symbols Comparison on LD-DB symbols}.  %The graph shows the recognition rate of each symbol using different algorithms. 
+	\label{fig:LDtest2}
+\end{figure}
+
+\begin{figure}
+	\centering
+		\includegraphics[scale=0.1]{paperImages/LDsymbols.jpg.eps}
+	\caption{Symbols Comparison on EL-DB symbols}.  %The graph shows the recognition rate of each symbol using different algorithms. 
+	\label{fig:ELtest2}
+\end{figure}
+
+\subsubsection{Features Comparisons}
 \label{sec:featuresComparisions}
@@
 Features \textbf{(FS4)} gives good results but it is improved by adding structural features \textbf{(FS1)}.
-Figure \ref{fig:testFeaturesAllMD} shows the simillar results on \textsl{MD-DB} multidomain database.
-The figure shows that on this dataset \textbf{FS1} performe better than on \textsl{HS-DB}, this is a result that shapes in the \textsl{MD-DB} has more structural difference than shapes in \textsl{HS-DB}.
+Figure \ref{fig:testFeaturesAllMD} shows the similar results on \textsl{EL-DB} and \textsl{LD-DB} dataset.
+The figures show that on both datasets \textbf{FS1} perform better than on \textsl{HS-DB}, this is a result that shapes in the \textsl{EL-DB} and \textsl{LD-DB} has more structural and conceptual differences than shapes in \textsl{HS-DB}.
+In \cite{HeloiseBeautification} used Zernike moment(FS3) and mentioned that order higher than 8 yield to little or no improvment.
+In our system we used much higher order to reach the this saturation level.
+This is due to the level of details in the EL-DB and DL-DB sets.
+% achieved  HS-DB a
  \begin{figure}
 	\centering
-		\includegraphics[scale=0.5]{results/testFeat.eps}
-	\caption{Feature Comparison} The effect of different features on accuracy on \textsl{Hs-DB}.  %The graph shows the recognition rate of each symbol using different algorithms. 
-	\label{fig:testFeaturesAllHS}
+		\includegraphics[scale=0.5]{paperImages/featuresdifference.jpg.eps}
+	\caption{Feature Comparison on Hs-DB} The effect of different features on accuracy.  %The graph shows the recognition rate of each symbol using different algorithms. 
+	\label{fig:testFeaturesAll}
 \end{figure}
@@
 	\centering
-		\includegraphics[scale=0.5]{results/testF1.eps}
-	\caption{Feature Comparison} The effect of different features on accuracy on \textsl{MD-DB}.  %The graph shows the recognition rate of each symbol using different algorithms. 
-	\label{fig:testFeaturesAllMD}
+		\includegraphics[scale=0.5]{paperImages/featuresLD.jpg.eps}
+	\caption{Feature Comparison on LD-DB} The effect of different features on accuracy.  %The graph shows the recognition rate of each symbol using different algorithms. 
+	\label{fig:testFeaturesAll}
 \end{figure}
+
+ \begin{figure}
+	\centering
+		\includegraphics[scale=0.5]{paperImages/featuresdifference.jpg.eps}
+	\caption{Feature Comparison on EL-DB} The effect of different features on accuracy.  %The graph shows the recognition rate of each symbol using different algorithms. 
+	\label{fig:testFeaturesAll}
+\end{figure}
+
+ The main features affected by the segmentation stage are the geometric features computed in\textbf{FS1}.
+Hence those are the features we used to estimate the efficiency of the segmentation algorithm.
+The recognition accuracy using only the geometrical features is reported in Figure \ref{fig:testFeatonly} .
+The figure shows that \textsl{AlgS2} achieve better results.
+%effect of the segmentation algorithm
+ %Figure \ref{fig:SampleSeg} shows a sample of the output of the segmentation block.
+The result of segmentation block is tested by their effect on the final recognition results using the features that will be highly affected using different algorithms like \textbf{FS1} and \textbf{FS4} (Section\ref{sec:Recognition}).
+Figure \ref{fig:testFeatonly} shows the result of the different algorithms using only \textbf{FS1 }features.
+ \begin{figure}
+	\centering
+		\includegraphics[scale=0.4]{paperImages/featureFS1only.jpg.eps}
+	\caption{The effect of segmentaion on Hs-DB accuracy} %The effect of symbol complexity.  %The graph shows the recognition rate of each symbol using different algorithms. 
+	\label{fig:testFeatonly}
+\end{figure}  
+ \begin{figure}
+	\centering
+		\includegraphics[scale=0.4]{paperImages/featureFS1only.jpg.eps}
+	\caption{The effect of segmentaion on LD-DB accuracy} %The effect of symbol complexity.  %The graph shows the recognition rate of each symbol using different algorithms. 
+	\label{fig:LDtestFeatonly}
+\end{figure}  
+ \begin{figure}
+	\centering
+		\includegraphics[scale=0.4]{paperImages/featureFS1only.jpg.eps}
+	\caption{The effect of segmentaion on EL-DB accuracy} %The effect of symbol complexity.  %The graph shows the recognition rate of each symbol using different algorithms. 
+	\label{fig:ELtestFeatonly}
+\end{figure}  
+%
+% \begin{figure}
+%	\centering
+%		\includegraphics[scale=0.5]{results/testFeat.eps}
+%	\caption{Feature Comparison} The effect of different features on accuracy on \textsl{Hs-DB}.  %The graph shows the recognition rate of each symbol using different algorithms. 
+%	\label{fig:testFeaturesAllHS}
+%\end{figure}  
+%
+% \begin{figure}
+%	\centering
+%		\includegraphics[scale=0.5]{results/testF1.eps}
+%	\caption{Feature Comparison} The effect of different features on accuracy on \textsl{MD-DB}.  %The graph shows the recognition rate of each symbol using different algorithms. 
+%	\label{fig:testFeaturesAllMD}
+%\end{figure}
 
@@
 The result illustrated in this chapter shows the work done on the system and experiments performed.
-The results shows that the ellipse detection and PSO swarm algorithms gives better performance than other known systems.
+We showed result of three different dataset.
+The system performed better than most known algoirthm on the EL-DB and LD-DB.
+This is due to the focus on strcutre and stroke based properties gained from the sytstem after correctly segmenting the input strokes.
+The result also showed that even using \textsl{AlgS2} only as the segmentaion aglgoirthm the system perform better than other segmenation algorithm.
+Using different databases shows that the more the strucutre difference in the sysmobl the better recogniton rate.
+LD-DB proved to have the lowest recogntion rate due to the high simillarity between its symbols.
+% better than
 

### 2009-04-20 trunk/draft/System.tex
i went to dr. magdah but she did not look at the thesis
second commit

@@
 
-  The direction is calculated as the angle between two vectors as in equation \ref{eq:direction} where $\overrightarrow {p_i p_j }$ is the vector from point $pi$ to point $pj$ and $\|{\overrightarrow {p_i p_j }}\|$ is the normal of the  vector $\overrightarrow{p_i p_j }$.
+\begin{figure}
+	\centering
+		\includegraphics{images/curvatureCal.jpg.eps}
+	\caption{Direction Calculation}
+	\label{fig:curvatureCal.jpg}
+\end{figure}
+
+  The direction is calculated as the angle between vectors $Q=$ and the x-xaxis as in equation \ref{eq:direction} where $\overrightarrow {p_i p_j }$ is the vector from point $pi$ to point $pj$ and $\|{\overrightarrow {p_i p_j }}\|$ is the normal of the  vector $\overrightarrow{p_i p_j }$.
 The curvature is estimated using more than one method the result of each curvatures estimation is reported in section \ref{sec:Curvatureestimationresults}.

### 2009-04-21 trunk/draft/Results.tex
final before sending to doc

@@
 \begin{description}
-	\item[Segmentation Error] The segmentation error is computed from each stroke after segmentation as computed from in $E_{alg}$ in Equation \ref{eq:bestFit2}.
+	\item[Segmentation Error] The segmentation error is computed from each stroke after segmentation as computed in Section \ref{sec:bestFit}.
 The error represents the sum of Square Root Error of the length from the point on the stroke to the estimated curve or line.
@@
  The effect of number of agents, the error thresholds and various other parameters was investigated.
-The result of these tests are shown in Figure \ref{fig:swarmtestingS1} and \ref{fig:swarmtestingS2}.
+The result of these tests are shown in Figure \ref{fig:swarmtesting} and \ref{fig:swarmtesting2}.
 Figure \ref{fig:swarmtesting} shows that as the number of agents increase the \textit{Segmentation Error} value decrease.
-Similar results for the maximum swarm iterations are shown in Figure \ref{fig:swarmtesting1}.
+Similar results for the maximum swarm iterations are shown in Figure \ref{fig:swarmtesting}.
 Other algorithm parameters like $c_1$,$c_2$ that were mentioned in Section\ref{sec:ParticleSwarmAlgorithm} are tested using similar tests (Figure \ref{fig:swarmparamterw.jpg} and \ref{fig:swarmtesting2}).

### 2009-04-21 trunk/draft/System.tex
final before sending to doc

@@
 
-\begin{equation}
-\label{eq:direction}
-	\vartheta  = \cos ^{ - 1} \left( {\frac{{\overrightarrow {p_1 p_2 }  \cdot \overrightarrow {p_2 p_3 } }}{{\left\| {\overrightarrow {p_1 p_2 } } \right\| \times \left\| {\overrightarrow {p_2 p_3 } } \right\|}}} \right)
-\end{equation}
-
 \begin{figure}
@@
 
-  The direction is calculated as the angle between vectors $Q=$ and the x-xaxis as in equation \ref{eq:direction} where $\overrightarrow {p_i p_j }$ is the vector from point $pi$ to point $pj$ and $\|{\overrightarrow {p_i p_j }}\|$ is the normal of the  vector $\overrightarrow{p_i p_j }$.
-The curvature is estimated using more than one method the result of each curvatures estimation is reported in section \ref{sec:Curvatureestimationresults}.
-The best curvature estimation is used, the estimation define curvature as the change in direction with respect to length i.e.
+  The direction is calculated as the angle between vectors $Q_i$ and the $x-axis$ (vector $O_i$) as in equation \ref{eq:direction} (Figure \ref{fig:curvatureCal.jpg} explains direction calculations).We used a n estimation of curvature that define curvature as the change in direction with respect to length i.e.
 $c= \Delta d/\Delta s$.
+
+ \begin{equation}
+\label{eq:direction}
+	\vartheta  = \cos ^{ - 1} \left( {\frac{{\overrightarrow {Q_i}  \cdot \overrightarrow {O_i } }}{{\left\| {\overrightarrow {Q_i} } \right\| \times \left\| {\overrightarrow {O_i} } \right\|}}} \right)
+\end{equation}
+where $Q_i=\overrightarrow {P_{i - 1} P_i }$ is the vector from the point  $P_{i - 1}$ to point $P_i$ and $O_i$ represent the $x-axis$ at this point (Figure \ref{fig:curvatureCal.jpg}).
+$\|{\overrightarrow {Q_i }}\|$ is the normal of the  vector $\overrightarrow{Q_i}$
 
@@
 The complexity of computation is $O(n)$ where $n$ is number of points.
-Figures \ref{fig:speed2Distance} \&  \ref{fig:curvatures} show the computed information's for the stroke drawn in Figure \ref{fig:orignalStroke}.% After computing the graphs the system  compute the average, minimum and maximum values of all curves.
+Figures \ref{fig:speed2Distance} \&  \ref{fig:curvatures} show the computed information's for the stroke drawn in Figure \ref{fig:AnotherorignalStroke}.% After computing the graphs the system  compute the average, minimum and maximum values of all curves.
 \\% the lower speed points correspond to vertex and dominant point's location.
@@
 	\centering
-		\includegraphics[scale=0.8]{images/orignalStroke.eps}
+		\includegraphics[scale=0.8]{images/stroke3.jpg.eps}
 	\caption{Input Stroke} Example of an input stroke to the system. 
@@
 	\centering
-			\subfigure[ Speed Graph]{	\includegraphics[scale=0.5]{images/speed2.eps}}
+			\subfigure[ Speed Graph]{	\includegraphics[scale=0.5]{images/vel3orginal.png.eps}}
 			\hfill
-			\subfigure[ Time Difference Graph] {\includegraphics[scale=0.5]{images/time2.eps}}
+			\subfigure[ Time Difference Graph] {\includegraphics[scale=0.5]{images/td3.jpg.eps}}
 	\caption{Speed and Time Difference Graphs}
 	\label{fig:speed2Distance}
+\end{figure}
+
+\begin{figure}[]
+	\centering
+		\includegraphics[scale=0.8]{images/orignalStroke.eps}
+	\caption{Input Stroke} Example of an input stroke to the system. 
+	\label{fig:AnotherorignalStroke}
 \end{figure}
@@
 The stroke points $p_i(x,y)$ that correspond to those maximum values are labeled as \textit{possible dominant points} $P_{pd}$.
-Figure \ref{fig:MaxRegioi} shows an example of $Region_i$ of the speed curve and the dominant points $P_{pd}$ which correspond to the minimum of point in this region $Min(Region_i)$.
+Figure \ref{fig:MaxRegioi} and \ref{fig:speed2Distance} shows an example of $Region_i$ of the speed curve and the dominant points $P_{pd}$ which correspond to the minimum of point in this region $Min(Region_i)$.
 % (as shown are redundant)
@@
 	\centering
-		\includegraphics[scale=0.5]{images/speedreg.eps}
+		\includegraphics[scale=0.5]{images/vel3.jpg.eps}
 	\caption{Possible Dominant Points Calculations}
@@
 \subsection{Choosing the Best Fit}
+\label{sec:bestFit}
 The two PSO algorithms produce two segmentation solutions.
@@
 This evaluation is based on the area of both the segmentation and the input stroke.
-The segmentation error($E_{alg}$ in Equation \ref{eq:bestFit2})represents the sum of Square Root Error of the length from the point on the stroke to the estimated curve or line.
+The segmentation error($E_{alg}$ )represents the sum of Square Root Error of the length from the point on the stroke to the estimated curve or line.
 In \textsl{AlgS2}, this error is the same as the computed error in Equation \ref{eq:ErrorSwarm1} but For the second algorithm \textsl{AlgS2} it is different than the computed error in Equation \ref{eq:errorSwarm2}.
 The additional computation is necessary to  compare the two segmentation fairly.
-\begin{equation}
-\label{eq:bestFit1}
-E_{area}=A_{st}-A_{sg}
- \end{equation}
- \begin{equation}
-\label{eq:bestFit2}
-E_{tot}=E_{area}+E_{alg}
- \end{equation}
- where $A_{st}$ is the area of the input stroke, $A_{sg}$ is area of computed segmentation, $E_{alg}$ is the Segmentation Error.
-The segmentation which cross ponds to the minimum $E_{tot}$ is the best segmentation.
+The system choose the minimum segmentaiton as the final segmentation of this stroke.
+% The segmentation which crossponds to the minimum $E_{tot}$ is the best segmentation.  
+%\begin{equation}
+%\label{eq:bestFit1}
+%E_{area}=A_{st}-A_{sg}
+% \end{equation}
+% \begin{equation}
+% where $A_{st}$ is the area of the input stroke, $A_{sg}$ is area of computed segmentation, $E_{alg}$ is the Segmentation Error.
 
@@
 The following list gives details on all the features used in the system.
-
-\begin{description}
- \item [Structural and geometrical Features (FS1)] Features defines the structure of the geometrical symbol.  
-\begin{description}
-	 \item[Segments] Number of segments in the symbol.
-	 \item [Strokes] Number of strokes or partial strokes that created the symbols.  
-		\item[Primitives] Number of primitives in the symbol. The feature helps when identifying             symbols with mixed geometric primitives like cylinders and callouts.  
-		\item [Curves] Number of curves or ellipses in the symbol. 
-		\item [Lines] Number of lines in the symbol. 
-		\item [Perpendicular lines] Number of perpendicular lines. 
-		\item [Parallel lines] Number of parallel lines. 
-		\item [Intersections] Number of intersection between lines and curves. 
-		\item [T intersections] Number of T intersections. 
-		\item [L intersections] Number of L intersections. 
-		\item [X intersections] Number of X intersections.
+\subsection{Structural and geometrical Features(FS1)}
+Features defines the structure of the geometrical symbol.  
+%\begin{description}
+% \item [Structural and geometrical Features (FS1)] 
+  \begin{itemize}
+	 \item \emph{Segments:} Number of segments in the symbol.
+	 \item \emph{Strokes:} Number of strokes or partial strokes that created the symbols.  
+		\item  \emph{Primitives:} Number of primitives in the symbol. The feature helps when identifying             symbols with mixed geometric primitives like cylinders and callouts.  
+		\item \emph{Curves:} Number of curves or ellipses in the symbol. 
+		\item \emph{Lines:} Number of lines in the symbol. 
+		\item \emph{Perpendicular} lines Number of perpendicular lines. 
+		\item \emph{Parallel} lines Number of parallel lines. 
+		\item \emph{Intersections} Number of intersection between lines and curves. 
+		\item \emph{T intersections} Number of T intersections. 
+		\item \emph{L intersections} Number of L intersections. 
+		\item \emph{X intersections} Number of X intersections.
 
 		%\item[No. of holes] Number of holes in the symbols. 
-
-\end{description}
-\item [Rubine Feature Set (FS2)] Features introduced by Rubine\cite{gestureexample12} for single stroke gestures.
+\end{itemize}
+%\begin{description}
+%	 \item[Segments] Number of segments in the symbol.
+%	 \item [Strokes] Number of strokes or partial strokes that created the symbols.  
+%		\item[Primitives] Number of primitives in the symbol. The feature helps when identifying             symbols with mixed geometric primitives like cylinders and callouts.  
+%		\item [Curves] Number of curves or ellipses in the symbol. 
+%		\item [Lines] Number of lines in the symbol. 
+%		\item [Perpendicular lines] Number of perpendicular lines. 
+%		\item [Parallel lines] Number of parallel lines. 
+%		\item [Intersections] Number of intersection between lines and curves. 
+%		\item [T intersections] Number of T intersections. 
+%		\item [L intersections] Number of L intersections. 
+%		\item [X intersections] Number of X intersections.
+%		
+%		
+%		%\item[No. of holes] Number of holes in the symbols. 
+%
+%\end{description}
+\subsection{Rubine Feature Set (FS2)} Features introduced by Rubine\cite{gestureexample12} for single stroke gestures.
 These features represent the global shape of the symbol drawn.
@@
 \end{enumerate}
+\subsection{Statistical Features (FS3)} 
+\begin{enumerate}
+	\item \textsl{Zernike moments:} Zernike moments of order n\footnote{ We tried different values of n. Results show the analysis of each order.} \cite{HeloiseBeautification}. 
+\end{enumerate}
+\subsection{Global shape properties set (FS4)}
+ Features that are composites of different geometrical and statistical symbol. These features represent the global shape of the symbol drawn.
 
-\item[Statistical Features (FS3)]   
-\begin{description}
-	\item[Zernike moments] Zernike moments of order n\footnote{ We tried different values of n. Results show the analysis of each order.} \cite{HeloiseBeautification}. 
-\end{description}
-	\item[Composite Features (FS4)] Features that are composites of different geometrical and statistical symbol. These features represent the global shape of the symbol drawn.
-	\begin{description}
-	\item [Size Ratio] Ratio between width to height of the symbol.
-	\item [Ink density] Compute the density of points inside its bounding box\cite{GeometryAndDomain102}.   
- 	\item[Convex Hull Area] Area of convex hull with respect to area of bounding box of symbol.
-	\item[Convex Hull Perimeter] Perimeter of convex hull with respect to total length of symbol.
-		\item [Mean Centroidal radius] The Mean of the centroidal radius which is the distance from each point in the symbol to the center of gravity.
+ 	\begin{itemize}
+\item \emph{Size Ratio} Ratio between width to height of the symbol.
+	\item \emph{Ink density} Compute the density of points inside its bounding box\cite{GeometryAndDomain102}.   
+ 	\item \emph{Convex Hull Area} Area of convex hull with respect to area of bounding box of symbol.
+	\item \emph{Convex Hull Perimeter} Perimeter of convex hull with respect to total length of symbol.
+		\item \emph{Mean Centroidal radius} The Mean of the centroidal radius which is the distance from each point in the symbol to the center of gravity.
 
-	\item [Mean Time difference] The mean of the time difference between each two successive points in the symbol. %Different strokes are appended to construct a single path.
+	\item \emph{Mean Time difference} The mean of the time difference between each two successive points in the symbol. %Different strokes are appended to construct a single path.
 	%\item [Ra]
-  \end{description}
+  \end{itemize}
 
-\end{description}
-%
-%\begin{longtable}{|l|l|p{6cm}|p{5cm}|}
-%	%\caption{Feature Table}
-%	%\label{tab:FeatureTable}
-%\hline 
-%Type  & Index  & Features  & Notes \tabularnewline
-%\endhead
-%\hline 
-%Type  & Index  & Features  & Notes \tabularnewline
-%\endfirsthead
-%\hline 
-%Structural & \multicolumn{1}{c||}{} &  & \tabularnewline
-%\hline 
-% & F1.  & No. of parallel lines  & \tabularnewline
-%\hline 
-% & F2.  & No. of perpendicular lines  & \tabularnewline
-%\hline 
-% & F3.  & No. of intersection between of any two segmetns  & \tabularnewline
-%\hline 
-% & F4.  & No of primitives & \tabularnewline
-%\hline 
-% & F5.  & No. of line segments & \tabularnewline
-%\hline 
-% & F6  & No. of ellipse, circular arcs.  & \tabularnewline
-%\hline 
-%Statistical &  &  & \tabularnewline
-%\hline 
-% & F7.  &  & \tabularnewline
-%\hline 
-% & F8.  &  & \tabularnewline
-%\hline 
-%Rubine Features  &  & From features F9. till F13  & Introduced by in \cite{gestureexample12} \tabularnewline
-%\hline 
-% & F9.  & Cosine of starting angle  & \tabularnewline
-%\hline 
-% & F10.  & Sine of starting angle & \tabularnewline
-%\hline 
-% & F11.  & Length of diagonal of bounding box & (gives an idea of the size of the bounding box) \tabularnewline
-%\hline 
-% & F12.  & Angle of diagonal & gives an idea of the shape of the bounding box (long, tall, square) \tabularnewline
-%\hline 
-% & F13.  & Distance from start to end  & \tabularnewline
-%\hline 
-% & F14.  & Cosine of ending angle  & \tabularnewline
-%\hline 
-% & F15.  & Sine of ending angle  & \tabularnewline
-%\hline 
-% & F16.  & Total stroke length  & \tabularnewline
-%\hline 
-% & F17  & Change in Rotation( Arctan ) & gives the directional angle \tabularnewline
-%\hline 
-% & F18  & Absolute rotation  & \tabularnewline
-%\hline 
-% & F19  & Rotation squared  & \tabularnewline
-%\hline 
-% & F20  & The maximum speed reached (squared)  & \tabularnewline
-%\hline 
-% & F21  & Total time of stroke  & \tabularnewline
-%%\newpage
-%\hline 
-%Zernike moments  &  &  & \tabularnewline
-%\hline 
-% & F22  & Zernike moments to the order  $n$  & used by \cite{zernike61} \tabularnewline
-%Composite  &  &  & \tabularnewline
-%\hline 
-% &  &  & \tabularnewline
-%\hline 
-% &  &  & \tabularnewline
-%\hline 
-% &  &  & \tabularnewline
-%\hline
-%\end{longtable}
-%\end{table}
-  % Rubine el al.\cite{gestureexample12}  used a set of statistical features,  There are a lot of features used in the sketch recognition literature
-%How to handle multi-stroked symbols. 
-%The clustering algorithm.  
-%The classifier used in recognition
-%Describe the features used in recognition [ref to \cite{zernike61}]
-
- Section \ref{sec:recognitionAlgorithms} presents a comparative result of the different feature sets from these features.
+ Section \ref{sec:featuresComparisions} presents a comparative result of the different feature sets from these features.
 
@@
  The system uses support vector machine (SVM) classifier with Linear kernel (Linear kernel) \cite{libsvm}.
-The system uses OVO classifiers structure (object versus object).
-A validation algorithm is used to choose the Kernel variables parameters $c$ The results Section \ref{sec:recognitionAlgorithms} demonstrates the results of both classifiers.
-\\%Some features like number of parallel, perpendicular and intersecting lines are computed from the segmentation other like density, centroid and Zernike moments are computed from the raw stroke points.
-.
-The Rubine algorithm (Linear Gaussian classifier) was implemented for comparing the system.
-
-%%%%%%%%%%%%%%%%%%%55555555
-%Symbol Recognition 16
-%4.1 Preprocessing and segmentation . . . . . . . . . . . . . . . . . . . . . . . . . . . 16
-%4.2 Feature-Based Statistical Symbol Recognition . . . . . . . . . . . . . . . . . . . . 18
-%4.2.1 Feature Set . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . 19
-%4.2.2 Training . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . 21
-%4.2.3 Recognition . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . 23
-%4.2.4 User Study . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . 24
-%4.3 Graph-Based Symbol Recognition . . . . . . . . . . . . . . . . . . . . . . . . . . 28
-%4.3.1 Previous Work . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . 28
-%4.3.2 Statistical Graph Definitions and Error-driven Stochastic Graph Matching . 29
-%4.4 Image Based Symbol Recognition . . . . . . . . . . . . . . . . . . . . . . . . . . 30
-%4.4.1 User Study . . . . . . . . . . . . . .
-%3.3 System Components 43
-%4 Segmenting a Freehand Sketch
-%4.1 Interface 51
-%4.2 Spline representation 52
-%4.3 Points of potential feature separation 59
-%4.4 Corner detection 59
-%4.5 Intersection detection 61
-%4.6 Clustering of interesting points 65
-%4.7 Feature Extraction
+SVM is a binary classifier that maximize the margin between the decision boundary and the support vectors.
+To enable SVM to learn multiclassifiers we used one versus one strategy for combiinig set of binary classifer.
+This stratgy trains $ \frac{n(n-1)}{{2}}$ binary classifier each distinguish between only two shapes.
+The final result is based on a number of votes each shape has from all classifiers.
+We used the a library implementation of LibSVM \cite{libsvm} and its implementation of the one-vs-one strategy.

### 2009-04-21 trunk/draft/Results.tex
after spelling correction s

@@
 
-several expriments is presented to test the different segmentation algoirthms on the three data set used in our system.
-Algorithms (\textsl{AlgS1}, \textsl{AlgS2} and \textsl{Alg3} (in section \ref{sec:BenchMarckAlgorithm})) were tested individualy with and without the ellipse fitting module.
+several experiments is presented to test the different segmentation algorithms on the three data set used in our system.
+Algorithms (\textsl{AlgS1}, \textsl{AlgS2} and \textsl{Alg3} (in section \ref{sec:BenchMarckAlgorithm})) were tested individually with and without the ellipse fitting module.
 The final system (Choosing the Minimum error of (\textsl{AlgS1}, \textsl{AlgS2}) was evaluated on the three dataset(HS, EL and LD datasets ).
 Figure \ref{fig:test1},\ref{fig:testEL} and \ref{fig:testLD} shows the accuracy achieved by each algorithm.
-Theses expriments were done using the whole dataset as 75\% of the data as testset and 25\% as training.
-The results shows that \textsl{AlgS2} with the ellipse fitting modeule gives better recognition rate accross all datasets (Figure \ref{fig:test1},\ref{fig:testEL} and \ref{fig:testLD} ).
-Eventhough, \textsl{AlgS1} gives better result than \textsl{Alg3} in both EL-DB and HS-DB it perform poorly in LD-DB due to the high number of curves in every symbol in this dataset.
+Theses experiments were done using the whole dataset as 75\% of the data as test set and 25\% as training.
+The results shows that \textsl{AlgS2} with the ellipse fitting module gives better recognition rate across all datasets (Figure \ref{fig:test1},\ref{fig:testEL} and \ref{fig:testLD} ).
+Even though, \textsl{AlgS1} gives better result than \textsl{Alg3} in both EL-DB and HS-DB it perform poorly in LD-DB due to the high number of curves in every symbol in this dataset.
 These result are proved using different number of samples per category as shown in Figures \ref{fig:LDRes}, \ref{fig:ElRes}.
@@
 
- The effect of symbol complixity in a dataset were testing in the following expriment.
-These expriments test how many samples will be required to finaly obtain good performance from the recognizier system.
-In the following expriments, we tested the system using different number of sample per category.
- Figure \ref{fig:Alg1Ds} and \ref{fig:Alg2Ds} shows that Hs-DB requires more number of sample for traininig to achieve same result as in other data sets.
+ The effect of symbol complexity in a dataset were testing in the following experiment.
+These experiments test how many samples will be required to finally obtain good performance from the recognizer system.
+In the following experiments, we tested the system using different number of sample per category.
+ Figure \ref{fig:Alg1Ds} and \ref{fig:Alg2Ds} shows that Hs-DB requires more number of sample for training to achieve same result as in other data sets.
 This is due that it contains less signal and conceptual variation that the other two datasets.
-Figure \ref{fig:samplesofHS-DB.jpg} shows sample of data collected from one user in Hs-Db it is clear that it has reasnabole amount of variation between users but it is limited.
-Recognizing symbols that contains high variation in drawing style using fewer symbols in training prove the efficency of our system.
+Figure \ref{fig:samplesofHS-DB.jpg} shows sample of data collected from one user in Hs-Db it is clear that it has reasonable amount of variation between users but it is limited.
+Recognizing symbols that contains high variation in drawing style using fewer symbols in training prove the efficiency of our system.
  \begin{figure}
@@
 
-  Figure \ref{} and \ref{} shows examples of correctly recognizied symbol using our system.
-The figure shows that system recognizied some hard to recognize example in both EL-DB and DL-DB.
+  Figure \ref{} and \ref{} shows examples of correctly recognized symbol using our system.
+The figure shows that system recognized some hard to recognize example in both EL-DB and DL-DB.
 An example of symbols that was not correctly recognized are in Figure \ref{}.
-Most of the wrong classification were due to how messsy the user draw symbols and how simillary they are conceptual.
-Table \ref{} show the confusion matrix of EL-DB which shows that most error were between capcatior, batter and crystall which differe only on number of parallel lines.
-Table \ref{tab:ConfusionMatrix} (confusion matrix) shows the types of error in Hs-DB which shows that the system was confused between parallegram and trapizoid, moon and archs.
-Those error were due to users drawing styles as some user may draw a parallegram as a trapizoid.
+Most of the wrong classification were due to how messy the user draw symbols and how similarly they are conceptual.
+Table \ref{} show the confusion matrix of EL-DB which shows that most error were between capacitor, batter and crystal which differ only on number of parallel lines.
+Table \ref{tab:ConfusionMatrix} (confusion matrix) shows the types of error in Hs-DB which shows that the system was confused between parallelogram and trapezoid, moon and arch.
+Those error were due to users drawing styles as some user may draw a parallelogram as a trapezoid.
 
@@
 The figures show that on both datasets \textbf{FS1} perform better than on \textsl{HS-DB}, this is a result that shapes in the \textsl{EL-DB} and \textsl{LD-DB} has more structural and conceptual differences than shapes in \textsl{HS-DB}.
-In \cite{HeloiseBeautification} used Zernike moment(FS3) and mentioned that order higher than 8 yield to little or no improvment.
+In \cite{HeloiseBeautification} used Zernike moment(FS3) and mentioned that order higher than 8 yield to little or no improvement.
 In our system we used much higher order to reach the this saturation level.
@@
 		\includegraphics[scale=0.4]{paperImages/featureFS1only.jpg.eps}
-	\caption{The effect of segmentaion on Hs-DB accuracy} %The effect of symbol complexity.  %The graph shows the recognition rate of each symbol using different algorithms. 
+	\caption{The effect of segmentation on Hs-DB accuracy} %The effect of symbol complexity.  %The graph shows the recognition rate of each symbol using different algorithms. 
 	\label{fig:testFeatonly}
@@
 		\includegraphics[scale=0.4]{paperImages/featureFS1only.jpg.eps}
-	\caption{The effect of segmentaion on LD-DB accuracy} %The effect of symbol complexity.  %The graph shows the recognition rate of each symbol using different algorithms. 
+	\caption{The effect of segmentation on LD-DB accuracy} %The effect of symbol complexity.  %The graph shows the recognition rate of each symbol using different algorithms. 
 	\label{fig:LDtestFeatonly}
@@
 		\includegraphics[scale=0.4]{paperImages/featureFS1only.jpg.eps}
-	\caption{The effect of segmentaion on EL-DB accuracy} %The effect of symbol complexity.  %The graph shows the recognition rate of each symbol using different algorithms. 
+	\caption{The effect of segmentation on EL-DB accuracy} %The effect of symbol complexity.  %The graph shows the recognition rate of each symbol using different algorithms. 
 	\label{fig:ELtestFeatonly}
@@
 We showed result of three different dataset.
-The system performed better than most known algoirthm on the EL-DB and LD-DB.
-This is due to the focus on strcutre and stroke based properties gained from the sytstem after correctly segmenting the input strokes.
-The result also showed that even using \textsl{AlgS2} only as the segmentaion aglgoirthm the system perform better than other segmenation algorithm.
-Using different databases shows that the more the strucutre difference in the sysmobl the better recogniton rate.
-LD-DB proved to have the lowest recogntion rate due to the high simillarity between its symbols.
+The system performed better than most known algorithm on the EL-DB and LD-DB.
+This is due to the focus on structure and stroke based properties gained from the system after correctly segmenting the input strokes.
+The result also showed that even using \textsl{AlgS2} only as the segmentation algorithm the system perform better than other segmentation algorithm.
+Using different databases shows that the more the structure difference in the symbols the better recognition rate.
+LD-DB proved to have the lowest recognition rate due to the high similarity between its symbols.
 % better than

### 2009-04-21 draft/System.tex
after spelling correction s

@@
 
+\begin{figure}
+	\centering
+		\includegraphics{images/curvatureCal.jpg.eps}
+	\caption{Direction Calculation}
+	\label{fig:curvatureCal.jpg}
+\end{figure}
+
+  The direction is calculated as the angle between vectors $Q_i$ and the $x-axis$ (vector $O_i$) as in equation \ref{eq:direction} (Figure \ref{fig:curvatureCal.jpg} explains direction calculations).We used a n estimation of curvature that define curvature as the change in direction with respect to length i.e.
+$c= \Delta d/\Delta s$.
+
  \begin{equation}
 \label{eq:direction}
-	\vartheta  = \cos ^{ - 1} \left( {\frac{{\overrightarrow {p_1 p_2 }  \cdot \overrightarrow {p_2 p_3 } }}{{\left\| {\overrightarrow {p_1 p_2 } } \right\| \times \left\| {\overrightarrow {p_2 p_3 } } \right\|}}} \right)
+	\vartheta  = \cos ^{ - 1} \left( {\frac{{\overrightarrow {Q_i}  \cdot \overrightarrow {Px_i } }}{{\left\| {\overrightarrow {Q_i} } \right\| \times \left\| {\overrightarrow {Px_i} } \right\|}}} \right)
 \end{equation}
-
-  The direction is calculated as the angle between two vectors as in equation \ref{eq:direction} where $\overrightarrow {p_i p_j }$ is the vector from point $pi$ to point $pj$ and $\|{\overrightarrow {p_i p_j }}\|$ is the normal of the  vector $\overrightarrow{p_i p_j }$.
-The curvature is estimated using more than one method the result of each curvatures estimation is reported in section \ref{sec:Curvatureestimationresults}.
-The best curvature estimation is used, the estimation define curvature as the change in direction with respect to length i.e.
-$c= \Delta d/\Delta s$.
+where $Q_i=\overrightarrow {P_{i - 1} P_i }$ is the vector from the point  $P_{i - 1}$ to point $P_i$ and $Px_i$ represent the $x-axis$ at this point (Figure \ref{fig:curvatureCal.jpg}).
+$\|{\overrightarrow {Q_i }}\|$ is the normal of the  vector $\overrightarrow{Q_i}$
 
@@
 %Alg2: 
-\subsubsection{Hybird Curve Algorithm \textsl{AlgS2}}
+\subsubsection{Hybrid Curve Algorithm \textsl{AlgS2}}
 \label{sec:PolygonDivisionAlgorithm}
@@
 This evaluation is based on the area of both the segmentation and the input stroke.
-The segmentation error($E_{alg}$ in Equation \ref{eq:bestFit2})represents the sum of Square Root Error of the length from the point on the stroke to the estimated curve or line.
+The segmentation error($E_{alg}$ )represents the sum of Square Root Error of the length from the point on the stroke to the estimated curve or line.
 In \textsl{AlgS2}, this error is the same as the computed error in Equation \ref{eq:ErrorSwarm1} but For the second algorithm \textsl{AlgS2} it is different than the computed error in Equation \ref{eq:errorSwarm2}.
 The additional computation is necessary to  compare the two segmentation fairly.
-\begin{equation}
-\label{eq:bestFit1}
-E_{area}=A_{st}-A_{sg}
- \end{equation}
- \begin{equation}
-\label{eq:bestFit2}
-E_{tot}=E_{area}+E_{alg}
- \end{equation}
- where $A_{st}$ is the area of the input stroke, $A_{sg}$ is area of computed segmentation, $E_{alg}$ is the Segmentation Error.
-The segmentation which cross ponds to the minimum $E_{tot}$ is the best segmentation.
+The system choose the minimum segmentation as the final segmentation of this stroke.
+% The segmentation which crossponds to the minimum $E_{tot}$ is the best segmentation.  
+%\begin{equation}
+%\label{eq:bestFit1}
+%E_{area}=A_{st}-A_{sg}
+% \end{equation}
+% \begin{equation}
+% where $A_{st}$ is the area of the input stroke, $A_{sg}$ is area of computed segmentation, $E_{alg}$ is the Segmentation Error.
 
@@
  The system uses support vector machine (SVM) classifier with Linear kernel (Linear kernel) \cite{libsvm}.
-The system uses OVO classifiers structure (object versus object).
-A validation algorithm is used to choose the Kernel variables parameters $c$ The results Section \ref{sec:recognitionAlgorithms} demonstrates the results of both classifiers.
-\\%Some features like number of parallel, perpendicular and intersecting lines are computed from the segmentation other like density, centroid and Zernike moments are computed from the raw stroke points.
-.
-The Rubine algorithm (Linear Gaussian classifier) was implemented for comparing the system.
-
-%%%%%%%%%%%%%%%%%%%55555555
-%Symbol Recognition 16
-%4.1 Preprocessing and segmentation . . . . . . . . . . . . . . . . . . . . . . . . . . . 16
-%4.2 Feature-Based Statistical Symbol Recognition . . . . . . . . . . . . . . . . . . . . 18
-%4.2.1 Feature Set . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . 19
-%4.2.2 Training . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . 21
-%4.2.3 Recognition . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . 23
-%4.2.4 User Study . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . 24
-%4.3 Graph-Based Symbol Recognition . . . . . . . . . . . . . . . . . . . . . . . . . . 28
-%4.3.1 Previous Work . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . 28
-%4.3.2 Statistical Graph Definitions and Error-driven Stochastic Graph Matching . 29
-%4.4 Image Based Symbol Recognition . . . . . . . . . . . . . . . . . . . . . . . . . . 30
-%4.4.1 User Study . . . . . . . . . . . . . .
-%3.3 System Components 43
-%4 Segmenting a Freehand Sketch
-%4.1 Interface 51
-%4.2 Spline representation 52
-%4.3 Points of potential feature separation 59
-%4.4 Corner detection 59
-%4.5 Intersection detection 61
-%4.6 Clustering of interesting points 65
-%4.7 Feature Extraction
+SVM is a binary classifier that maximize the margin between the decision boundary and the support vectors.
+To enable SVM to learn multi classes we used one versus one strategy for combining set of binary classifier.
+This strategy trains $ \frac{n(n-1)}{{2}}$ binary classifier each distinguish between only two shapes.
+The final result is based on a number of votes each shape has from all classifiers.
+We used the a library implementation of LibSVM \cite{libsvm} and its implementation of the one-vs-one strategy.

### 2009-04-21 trunk/draft/Results.tex
after the grammer mistake check

@@
 \label{sec:Datasets}
-Two different data set were used, 
+Three different data set were used, table \ref{tab:datasets} shows a comparison between the three datasets used in our experiments.
+
 %For classifier and algorithm testing the system used a data set collected by
@@
 Five different splits for the training and test data are generated from the dataset.
-%In worth noting that \cite{Oltmans07} gathered simillar dataset and acheived 89\% on it   %The results displayed are the average recognition accuracy of the four splits resulted from the classifier after recognizing the symbols.
+%In worth noting that \cite{Oltmans07} gathered similar dataset and achieved 89\% on it   %The results displayed are the average recognition accuracy of the four splits resulted from the classifier after recognizing the symbols.
 \end{description}
-
-Table \ref{tab:datasets} shows a  comparison between the three datasets used in our experiments.
 
@@
 The ellipse fitting module appears to be superior.
-The result shows that combining both (\textsl{AlgS1} and \textsl{AlgS2}) out perform any single algorithm.
-%show that combining algoirthmsachieve higher accuracy than any single algoirthm.
+The result shows that combining both algorithms (\textsl{AlgS1} and \textsl{AlgS2}) outperform any single algorithm.
+%show that combining algorithms achieve higher accuracy than any single algoirthm.
 %improves the results.% with both the \textit{DPSO} algorithms.% The results show that both \textit{DPSO} algorithms achieve better result than other algorithms. 
@@
 
- The effect of symbol complexity in a dataset were testing in the following experiment.
+ The effects of symbol complexity in a dataset were testing in the following experiment.
 These experiments test how many samples will be required to finally obtain good performance from the recognizer system.
@@
 An example of symbols that was not correctly recognized are in Figure \ref{}.
-Most of the wrong classification were due to how messy the user draw symbols and how similarly they are conceptual.
+Most of the wrong classifications were due to how messy the user draw symbols and how similarly they are conceptual.
 Table \ref{} show the confusion matrix of EL-DB which shows that most error were between capacitor, batter and crystal which differ only on number of parallel lines.
 Table \ref{tab:ConfusionMatrix} (confusion matrix) shows the types of error in Hs-DB which shows that the system was confused between parallelogram and trapezoid, moon and arch.
-Those error were due to users drawing styles as some user may draw a parallelogram as a trapezoid.
+Those errors were due to users drawing styles as some user may draw a parallelogram as a trapezoid.
 
@@
 This observation is understandable as the symbols are visually similar but the system must be able to differentiate between them.
-This lead us to the next experiment to choose best set of feature to recognize symbols.
+This led us to the next experiment to choose best set of feature to recognize symbols.
 
@@
 The figures show that on both datasets \textbf{FS1} perform better than on \textsl{HS-DB}, this is a result that shapes in the \textsl{EL-DB} and \textsl{LD-DB} has more structural and conceptual differences than shapes in \textsl{HS-DB}.
-In \cite{HeloiseBeautification} used Zernike moment(FS3) and mentioned that order higher than 8 yield to little or no improvement.
-In our system we used much higher order to reach the this saturation level.
+In \cite{HeloiseBeautification} used Zernike moment (FS3) and mentioned that order higher than 8 yield to little or no improvement.
+In our system we used much higher order to reach this saturation level.
 This is due to the level of details in the EL-DB and DL-DB sets.
@@
 The result also showed that even using \textsl{AlgS2} only as the segmentation algorithm the system perform better than other segmentation algorithm.
-Using different databases shows that the more the structure difference in the symbols the better recognition rate.
+Using different databases shows that the more the structure differences in the symbols the better recognition rate.
 LD-DB proved to have the lowest recognition rate due to the high similarity between its symbols.

### 2009-04-21 draft/System.tex
after the grammer mistake check

@@
 The system uses support vector machine (SVM) classifier with Linear kernel (Linear kernel) \cite{libsvm}.
-SVM is a binary classifier that maximize the margin between the decision boundary and the support vectors.
+SVM is a binary classifier that maximizes the margin between the decision boundary and the support vectors.
 To enable SVM to learn multi classes we used one versus one strategy for combining set of binary classifier.
@@
 The final result is based on a number of votes each shape has from all classifiers.
-We used the a library implementation of LibSVM \cite{libsvm} and its implementation of the one-vs-one strategy.
+We used the library implementation of LibSVM \cite{libsvm} and its implementation of the one-vs-one strategy.

### 2009-04-21 trunk/draft/ConFut.tex

--- 2009-04-12 draft/ConFut.tex
@@
 
-This paper presented a new approach to sketch recognition using PSO.
+This research presented a new approach to sketch recognition using PSO.
 It was noted that the PSO in general improve the accuracy of the final symbol recognition in the system.
-The use of both speed and curvature data helps in improvement of the PSO algorithm over the original algorithms \cite{CruveDivisionSwarm,PolygonApproximationPSO}.
+The use of different preliminary information (speed, time difference and curvature)  helps to improve of the PSO algorithm over the original algorithms \cite{CruveDivisionSwarm,PolygonApproximationPSO}.
 The tradeoff between accuracy achieved and time complexity must be further investigated to achieve better results.
 The system introduced an efficient method to sketch recognition using the PSO.
-The test performed confirmed that PSO achieve better performance and optimization than other algorithms.
+The features computed was a hybrid of a stroke based features which gives more information on the geometrical and structure of the shape tested.
+
+The system was tested on three different datasets, a simple presentation symbols, electrical symbols and logic design dataset.
 The number and the complexity of symbols recognized by the system varied from simple to complex but did not affect the final recognition of the system.
-One drawback of the system is that it takes more time to segment strokes in comparison with other similar algorithms.
+Test performed confirmed that PSO achieve better performance and optimization than other algorithms.
+In general \textsl{ALS2} proved to generate the best result on different datasets.
+Algorithm \textsl{ALS2} achieved better segmentation result that any other algorithm in shapes that are combination of curves and lines as in digital design dataset.
+Algorithm \textsl{ALS1} achieve high performance in electrical symbols but gives poor result in logic design set.
+The result is justified because logic design dataset contains symbols that are differentiated by number of curves in OR-gate and XOr gate, therefore representing the symbol as lines will remove important information.
+
+ On the other hand, the system still has drawback as it lack in handling over traced strokes.
 Another drawback is that until now we only applied the system to single symbols.
@@
 %sketched
+%drawback of the system is that it takes more time to segment strokes in comparison with other similar algorithms.
 

### 2009-04-22 draft/ConFut.tex
before dr. nevin change

@@
 
-This paper presented a new approach to sketch recognition using PSO.
+This research presented a new approach to sketch recognition using PSO.
 It was noted that the PSO in general improve the accuracy of the final symbol recognition in the system.
-The use of both speed and curvature data helps in improvement of the PSO algorithm over the original algorithms \cite{CruveDivisionSwarm,PolygonApproximationPSO}.
+The use of different preliminary information (speed, time difference and curvature)  helps to improve of the PSO algorithm over the original algorithms \cite{CruveDivisionSwarm,PolygonApproximationPSO}.
 The tradeoff between accuracy achieved and time complexity must be further investigated to achieve better results.
 The system introduced an efficient method to sketch recognition using the PSO.
-The test performed confirmed that PSO achieve better performance and optimization than other algorithms.
+The features computed was a hybrid of a stroke based features which gives more information on the geometrical and structure of the shape tested.
+
+The system was tested on three different datasets, simple presentation symbols, electrical symbols and logic design dataset.
 The number and the complexity of symbols recognized by the system varied from simple to complex but did not affect the final recognition of the system.
-One drawback of the system is that it takes more time to segment strokes in comparison with other similar algorithms.
+Test performed confirmed that PSO achieve better performance and optimization than other algorithms.
+In general \textsl{ALS2} proved to generate the best result on different datasets.
+Algorithm \textsl{ALS2} achieved better segmentation result that any other algorithm in shapes that are combination of curves and lines as in digital design dataset.
+Algorithm \textsl{ALS1} achieve high performance in electrical symbols but gives poor result in logic design set.
+The result is justified because logic design dataset contains symbols that are differentiated by number of curves in OR-gate and XOr gate, therefore representing the symbol as lines will remove important information.
+
+The system uses different set of feature that represents shape geometrical and global properties.
+Literature review showed that system that relies only on global shape properties fails to identify the shapes that have small geometrical difference (ie.
+Xor gate and Nor gate).
+Using geometrical and stroke based features with the global shape properties helps to improve the final recognition rate.
+
+ On the other hand, the system still has drawback as it lack in handling over traced strokes.
 Another drawback is that until now we only applied the system to single symbols.
@@
 The next step in this research is to complete the clustering algorithm for fully automated sketch recognition.
-The clustering must be performed without the user explicit involvement.
-The currently used algorithm must be modified to accept error reporting from users.
+This will enable testing the sytsem for more than single symbol at a time.
 The segmentation algorithm may be more powerful if it can segments the stroke into more types of primitives other than the line and curve.
 Other area of modification can be the features extraction and classifier.
-Adding more spatial features and structure features will improve classifications.
+Adding more spatial and structure features will improve classifications.
 After segmenting the stroke the system can generate a structural graph from segments and there spatial relations.

### 2009-04-22 draft/Results.tex
before dr. nevin change

@@
 This chapter describes and illustrates the experiments performed on the system.
-In the chapter, section \ref{sec:ExperimentsSetup} describes the experimental setup.
-Section \ref{sec:Datasets} describes the datasets used.
+In the chapter, Section \ref{sec:Datasets} describes the datasets used.
 Section \ref{sec:ResultsDetails} demonstrates and analyzes the results.
@@
 The recognition system evaluation is presented in Section \ref{sec:SegmentationAlgorithms}.
-The recognition rate (computed as the number of correctly recognized samples divided by the total number of samples in the test) for each segmentation algorithm compared to another segmentation algorithm are reported in \ref{sec:recognitionAlgorithms}.
+The recognition rate (computed as the number of correctly recognized samples divided by the total number of samples in the test) for each segmentation algorithm compared to another segmentation algorithm are reported in \ref{sec:RecognitionAlgorithms}.
 Next, a comparison between different shapes and datasets and it influence on the recognition rate is presented in Section \ref{sec:EffectsofSymbolComplixity}.
@@
 Figure \ref{fig:symbolSet} shows the shapes used in the data set.
-The dataset was divided into training set and test set.
+The dataset was divided into 75\% training set and 25\% test set.
 Five different splits for the training and test data are generated from the dataset.
@@
 Users were requested to draw symbols without any restriction on style, size or orientations.
-The dataset was divided into training set and test set.
+The dataset was divided into 75\% training set and 25\% test set.
 Five different splits for the training and test data are generated from the dataset.
-%The results displayed are the average recognition accuracy of the four splits resulted from the classifier after recognizing the symbols.
+%In worth noting that \cite{Oltmans07} gathered similar dataset and achieved 89\% on it   %The results displayed are the average recognition accuracy of the four splits resulted from the classifier after recognizing the symbols.
 \end{description}
-
-Table \ref{tab:datasets} shows a  comparison between the three datasets used in our experiments.
 
@@
 \begin{figure}[]\centering
-\fbox{  \parbox{6cm}{% 
-		\includegraphics[scale=0.7]{paperImages/logicSet.jpg.eps}	}
-		}
+
+		\includegraphics[scale=0.9]{paperImages/logicSet.jpg.eps}
+
 	\caption{Example of Logic Symbols}
@@
 \begin{figure}[]\centering
-\fbox{  \parbox{8.2cm}{% 
-		\includegraphics[scale=0.7]{paperImages/EelectImage.jpg.eps}	}
-		}
+
+		\includegraphics[scale=0.9]{paperImages/EelectImage.jpg.eps}
+
 	\caption{Example of Electrical Symbols}
@@
 \scalebox{0.9}{
- \begin{tabular}{|p{4cm}|p{2.5cm}|p{2.5cm}|p{2.5cm}|}
+ \begin{tabular}{|p{5cm}|p{2.5cm}|p{2.5cm}|p{2.5cm}|}
  \hline
@@
 Splits  & 5 random splits  & 5 random splits  & 5 random splits  \\ \hline
-Examples & Figure \ref{fig:symbolSet}  & Figure \ref{fig:ELsymbolSet}  & Figure \ref{fig:LogicsymbolSet}   \\ \hline
+Examples & Figure \ref{fig:symbolSet}  & Figure \ref{fig:ElectSet}  & Figure \ref{fig:LogicSet}   \\ \hline
 \end{tabular}
@@
 	\item [Recognition accuracy] We measure recognition performance for each algorithm by determining the number of correctly identified symbols in the whole dataset.
-Note that All the results are shown as the average of the four split of the dataset used.
+Note that All the results are shown as the average of the five split of the dataset used.
 \end{description}
@@
 
-Figure \ref{fig:SegErrPreCal} shows the \textit{Average Segmentation Error} using different preliminary calculation (Section \ref{sec:CurvatureCalculation}).
+%Figure \ref{fig:SegErrPreCal} shows the \textit{Average Segmentation Error} using different preliminary calculation (Section \ref{sec:CurvatureCalculation}).
 The figure shows that using all computations (time difference, direction, speed and curvature) for each stroke decrease the Segmentation error.
@@
 
-\begin{figure}
-	\centering
-		\includegraphics[scale=0.5]{results/SegErrPreCal.eps}
-	\caption{preliminary calculation effect}
-	\label{fig:SegErrPreCal}
-\end{figure}
-
-\begin{figure}
-	\centering
-		\includegraphics[scale=0.5]{results/PossibledpPreCalculation.eps}
-	\caption{Possible Dominant Point Count}
-	\label{fig:PossibledpPreCalculation}
-\end{figure}
+%\begin{figure}
+%	\centering
+%		\includegraphics[scale=0.5]{results/SegErrPreCal.eps}
+%	\caption{preliminary calculation effect}
+%	\label{fig:SegErrPreCal}
+%\end{figure}
+% 
+% 
+%\begin{figure}
+%	\centering
+%		\includegraphics[scale=0.5]{results/PossibledpPreCalculation.eps}
+%	\caption{Possible Dominant Point Count}
+%	\label{fig:PossibledpPreCalculation}
+%\end{figure}
 
  The effect of number of agents, the error thresholds and various other parameters was investigated.
-The result of these tests are shown in Figure \ref{fig:swarmtestingS1} and \ref{fig:swarmtestingS2}.
+The result of these tests are shown in Figure \ref{fig:swarmtesting} and \ref{fig:swarmtesting2}.
 Figure \ref{fig:swarmtesting} shows that as the number of agents increase the \textit{Segmentation Error} value decrease.
-Similar results for the maximum swarm iterations are shown in Figure \ref{fig:swarmtesting1}.
+Similar results for the maximum swarm iterations are shown in Figure \ref{fig:swarmtesting}.
 Other algorithm parameters like $c_1$,$c_2$ that were mentioned in Section\ref{sec:ParticleSwarmAlgorithm} are tested using similar tests (Figure \ref{fig:swarmparamterw.jpg} and \ref{fig:swarmtesting2}).
@@
 Figure \ref{fig:results3} shows the output of \textsl{Alg3} to the same set of input strokes.
-It clearly shows that \textsl{AlgS2} have better segmentation results than both \textsl{AlgS1} and \textsl{Alg3} but since these segmentation are subjective to what the user intended to draw the next section will focus on the effect of the segmentation algorithm on the recognition accuracy.
+It clearly shows that \textsl{AlgS2} have better segmentation results than both \textsl{AlgS1} and \textsl{Alg3} but since these segmentation are subjective to what the user intended to draw the next section will focus on the evaluating the segmentation algorithms using the effect on the recognition accuracy.
 
@@
 
-\subsection {Recognition Algorithms}
+\subsection {Recognition system}
 \label{sec:RecognitionAlgorithms}
 
-We performed several experiments to evaluate the presented recognition system.
+several experiments is presented to test the different segmentation algorithms on the three data set used in our system.
+Algorithms (\textsl{AlgS1}, \textsl{AlgS2} and \textsl{Alg3} (in section \ref{sec:BenchMarckAlgorithm})) were tested individually with and without the ellipse fitting module.
+The final system (Choosing the Minimum error of (\textsl{AlgS1}, \textsl{AlgS2}) was evaluated on the three dataset(HS, EL and LD datasets ).
+Figure \ref{fig:test1},\ref{fig:testEL} and \ref{fig:testLD} shows the accuracy achieved by each algorithm.
+Theses experiments were done using the whole dataset as 75\% of the data as test set and 25\% as training.
+The results shows that \textsl{AlgS2} with the ellipse fitting module gives better recognition rate across all datasets (Figure \ref{fig:test1},\ref{fig:testEL} and \ref{fig:testLD} ).
+Even though, \textsl{AlgS1} gives better result than \textsl{Alg3} in both EL-DB and HS-DB it perform poorly in LD-DB due to the high number of curves in every symbol in this dataset.
+These result are proved using different number of samples per category as shown in Figures \ref{fig:LDRes}, \ref{fig:ElRes}.
+% i need to organize this as following. 
+% number 1. differnt algorithm on different data 
+% which algortihm is better oon which data 
+% on ld and el  the alg2  is better,
+
+%We performed several experiments to evaluate the presented recognition system.
 Firstly we tested recognition accuracy of shapes in the data set with both algorithms (\textsl{AlgS1} and \textsl{AlgS2}).
-We also implemented the segmentation algorithm described in \cite{earlyprocess} (\textsl{Alg3}) to use as reference to compare it with our swarm algorithms.
+We also implemented the segmentation algorithm described in \cite{earlyprocess} (\textsl{Alg3} in section \ref{sec:BenchMarckAlgorithm} )  to use as reference to compare it with our swarm algorithms.
 Figure \ref{fig:test1},\ref{fig:testEL} and \ref{fig:testLD}  shows the accuracy achieved by each algorithm.
 The two swarm algorithms were tested with and without the ellipse fitting module.
-The ellipse detection module appears to be superior.
-The result shows that combining both  (\textsl{AlgS1} and \textsl{AlgS2}) out perform any single algorithm.
-%show that combining algoirthmsachieve higher accuracy than any single algoirthm.
+The ellipse fitting module appears to be superior.
+The result shows that combining both algorithms (\textsl{AlgS1} and \textsl{AlgS2}) outperform any single algorithm.
+%show that combining algorithms achieve higher accuracy than any single algoirthm.
 %improves the results.% with both the \textit{DPSO} algorithms.% The results show that both \textit{DPSO} algorithms achieve better result than other algorithms. 
-
 \begin{figure}
@@
 \end{figure} 
-
 \begin{figure}
 	\centering		
-	 \includegraphics[scale=0.4]{paperImages/LDAlg.jpg.eps}
+	 \includegraphics[scale=0.8]{results/AlgLD.eps}
 	 	\caption{Algorithm comparison on LD-DB} The recognition rate of different algorithms. 
@@
 	\centering		
-	 \includegraphics[scale=0.1]{paperImages/LDAlg.jpg.eps}
+	 \includegraphics[scale=0.8]{results/AlgEL.eps}
 	 	\caption{Algorithm comparison on EL-DB} The recognition rate of different algorithms. 
@@
 
-\subsubsection {Effects of Symbol Complexity }
+\begin{figure}
+	\centering
+		\includegraphics[scale=0.6]{results/LDRes.eps}
+	\caption{Result Algorithms on LD-DB}
+	\label{fig:LDRes}
+\end{figure}
+\begin{figure}
+	\centering
+		\includegraphics[scale=0.6]{results/ElRes.eps}
+	\caption{Results of Algorithms on EL-DB}
+	\label{fig:ElRes}
+\end{figure}
+
+\subsubsection {Evaluating Different Datasets}
  \label{sec:EffectsofSymbolComplixity}
 
-%\begin{figure}
-%	\centering
-%		\includegraphics[scale=0.5]{results/testsym.eps}
-%	\caption{Symbols Comparison} The effect of symbol complexity.  %The graph shows the recognition rate of each symbol using different algorithms. 
-%	\label{fig:test2}
-%\end{figure}
-
+ The effects of symbol complexity in a dataset were testing in the following experiment.
+These experiments test how many samples will be required to finally obtain good performance from the recognizer system.
+In the following experiments, we tested the system using different number of sample per category.
+ Figure \ref{fig:Alg1Ds} and \ref{fig:Alg2Ds} shows that Hs-DB requires more number of sample for training to achieve same result as in other data sets.
+This is due that it contains less signal and conceptual variation that the other two datasets.
+Figure \ref{fig:samplesofHS-DB.jpg} shows sample of data collected from one user in Hs-Db it is clear that it has reasonable amount of variation between users but it is limited.
+Recognizing symbols that contains high variation in drawing style using fewer symbols in training prove the efficiency of our system.
  \begin{figure}
 	\centering
-		\includegraphics[scale=0.35]{paperImages/testsym.jpg.eps}
-	\caption{Symbols Comparison on HS-DB} The effect of symbol complexity.  %The graph shows the recognition rate of each symbol using different algorithms. 
-	\label{fig:test2}
+		\includegraphics{results/Alg1Ds.eps}
+	\caption{AlgS1 on different datasets}
+	\label{fig:Alg1Ds}
+\end{figure}
+\begin{figure}
+	\centering
+		\includegraphics{results/Alg2Ds.eps}
+	\caption{AlgS2 on different Datasets }
+	\label{fig:Alg2Ds}
+\end{figure}
+ \begin{figure}
+	\centering
+	\includegraphics[scale=0.5]{results/samplesofHS-DB.jpg.eps}
+	\caption{Hs-DB shapes drawn by single user}
+	\label{fig:samplesofHS-DB.jpg}
+\end{figure}
+  Figure \ref{fig:correct2.jpg} shows examples of correctly recognized symbol using our system.
+The figure shows that system recognized some hard to recognize example in both EL-DB and DL-DB.
+An example of symbols that was not correctly recognized are in Figure \ref{fig:erros}.
+Most of the wrong classifications were due to how messy the user draw symbols and how similarly they are conceptual.
+Table \ref{tab:ConfusionMatrix} (confusion matrix) shows the types of error in Hs-DB which shows that the system was confused between parallelogram and trapezoid, moon and arch.
+Those errors were due to users drawing styles as some user may draw a parallelogram as a trapezoid.
+
+  %55Table \ref{} show the confusion matrix of EL-DB which shows that most error were between capacitor, batter and crystal which differ only on number of parallel lines. 
+\begin{figure}
+	\centering
+		\includegraphics{results/erros.jpg.eps}
+	\caption{Error in Recognition }
+	\label{fig:erros}
+\end{figure}
+\begin{figure}
+	\centering
+		\includegraphics{results/correct2.jpg.eps}
+	\caption{Symbols Correctly Recognized}
+	\label{fig:correct2.jpg}
 \end{figure}
 
-\begin{figure}
-	\centering
-		\includegraphics[scale=0.5]{paperImages/LDsymbols.jpg.eps}
-	\caption{Symbols Comparison on LD-DB symbols}.  %The graph shows the recognition rate of each symbol using different algorithms. 
-	\label{fig:LDtest2}
-\end{figure}
-
-\begin{figure}
-	\centering
-		\includegraphics[scale=0.1]{paperImages/LDsymbols.jpg.eps}
-	\caption{Symbols Comparison on EL-DB symbols}.  %The graph shows the recognition rate of each symbol using different algorithms. 
-	\label{fig:ELtest2}
-\end{figure}
-
-The second experiment we implemented was to investigate the effect of symbol complexity and type on the recognition rate.
-Figure \ref{fig:test2} shows the achieved accuracy of each symbol by our algorithms compared to \cite{earlyprocess} (\textsl{Alg3}).
-It is clearly noted that symbols that have only line segments achieve higher accuracy rate than other symbols.
-The results indicate that algorithm \textsl{AlgS1} achieve better performance than algorithm \textsl{AlgS2} in the symbols that consist only of lines.
-This is understandable because algorithm \textsl{AlgS1} divides strokes into line segments only but \textsl{AlgS2} is able to divide strokes into lines and curves based on the minimum error of the segment itself.
-Algorithm \textsl{Alg3} gives good performance as long as the symbols consist of combination of lines and curves, if the stroke consists of only curve or lines the algorithm may lead to wrong segmentation result.
-This is because the system divides the stroke first to line segments then tries to decide if each segment can be represent better as a curve unlike algorithm \textsl{AlgS2} where the curve segments are tested while choosing the best segmentation.
-Combining both algorithm \textsl{AlgS1} and \textsl{AlgS2} improved the recognition rate of all symbols.
-The penalty for this improved performance is the computational time required to run both swarm algorithms.
-Table \ref{tab:ConfusionMatrix} shows the confusion matrix of symbols.
-The table shows that errors are only between two sets of symbols moon with triangles and trapezoid with triangles.
-This observation is understandable as the symbols are visually similar but the system must be able to differentiate between them.
-This lead us to the next experiment to choose best set of feature to recognize symbols.
  \begin{table*}
@@
 
+%\begin{figure}
+%	\centering
+%		\includegraphics[scale=0.5]{results/testsym.eps}
+%	\caption{Symbols Comparison} The effect of symbol complexity.  %The graph shows the recognition rate of each symbol using different algorithms. 
+%	\label{fig:test2}
+%\end{figure}
+
+Another experiment we implemented was to investigate the effect of symbol complexity and type on the recognition rate.
+Figure \ref{fig:test2} shows the achieved accuracy of each symbol by our algorithms compared to \cite{earlyprocess} (\textsl{Alg3}).
+It is clearly noted that symbols that have only line segments achieve higher accuracy rate than other symbols.
+The results indicate that algorithm \textsl{AlgS1} achieve better performance than algorithm \textsl{AlgS2} in the symbols that consist only of lines.
+This is understandable because algorithm \textsl{AlgS1} divides strokes into line segments only but \textsl{AlgS2} is able to divide strokes into lines and curves based on the minimum error of the segment itself.
+Algorithm \textsl{Alg3} gives good performance as long as the symbols consist of combination of lines and curves, if the stroke consists of only curve or lines the algorithm may lead to wrong segmentation result.
+This is because the system divides the stroke first to line segments then tries to decide if each segment can be represent better as a curve unlike algorithm \textsl{AlgS2} where the curve segments are tested while choosing the best segmentation.
+Combining both algorithm \textsl{AlgS1} and \textsl{AlgS2} improved the recognition rate of all symbols.
+The penalty for this improved performance is the computational time required to run both swarm algorithms.
+Table \ref{tab:ConfusionMatrix} shows the confusion matrix of symbols.
+The table shows that errors are only between two sets of symbols moon with triangles and trapezoid with triangles.
+This observation is understandable as the symbols are visually similar but the system must be able to differentiate between them.
+This led us to the next experiment to choose best set of feature to recognize symbols.
+
+ \begin{figure}
+	\centering
+		\includegraphics[scale=0.35]{paperImages/testsym.jpg.eps}
+	\caption{Symbols Comparison on HS-DB} The effect of symbol complexity.  %The graph shows the recognition rate of each symbol using different algorithms. 
+	\label{fig:test2}
+\end{figure}
+
+\begin{figure}
+	\centering
+		\includegraphics[scale=0.5]{paperImages/LDsymbols.jpg.eps}
+	\caption{Symbols Comparison on LD-DB symbols}.  %The graph shows the recognition rate of each symbol using different algorithms. 
+	\label{fig:LDtest2}
+\end{figure}
+
+\begin{figure}
+	\centering
+		\includegraphics[scale=0.5]{results/ELSymbols.eps}
+	\caption{Symbols Comparison on EL-DB symbols}.  %The graph shows the recognition rate of each symbol using different algorithms. 
+	\label{fig:ELtest2}
+\end{figure}
+
 \subsubsection{Features Comparisons}
@@
 Features \textbf{(FS4)} gives good results but it is improved by adding structural features \textbf{(FS1)}.
-Figure \ref{fig:testFeaturesAllMD} shows the similar results on \textsl{MD-DB} multi domain database.
-The figure shows that on this dataset \textbf{FS1} perform better than on \textsl{HS-DB}, this is a result that shapes in the \textsl{MD-DB} has more structural difference than shapes in \textsl{HS-DB}.
+Figure \ref{fig:testFeaturesAllMD} shows the similar results on \textsl{EL-DB} and \textsl{LD-DB} dataset.
+The figures show that on both datasets \textbf{FS1} perform better than on \textsl{HS-DB}, this is a result that shapes in the \textsl{EL-DB} and \textsl{LD-DB} has more structural and conceptual differences than shapes in \textsl{HS-DB}.
+In \cite{HeloiseBeautification} used Zernike moment (FS3) and mentioned that order higher than 8 yield to little or no improvement.
+In our system we used much higher order to reach this saturation level.
+This is due to the level of details in the EL-DB and DL-DB sets.
+% achieved  HS-DB a
  \begin{figure}
@@
 	\centering
-		\includegraphics[scale=0.5]{paperImages/featuresdifference.jpg.eps}
+		\includegraphics[scale=0.5]{results/FeatEL.eps}
 	\caption{Feature Comparison on EL-DB} The effect of different features on accuracy.  %The graph shows the recognition rate of each symbol using different algorithms. 
@@
 
- The main features affected by the segmentation stage are the geometric features computed in\textbf{FS1 }.
+ The main features affected by the segmentation stage are the geometric features computed in\textbf{FS1}.
 Hence those are the features we used to estimate the efficiency of the segmentation algorithm.
@@
 	\centering
-		\includegraphics[scale=0.4]{paperImages/featureFS1only.jpg.eps}
-	\caption{The effect of segmentaion on LD-DB accuracy} %The effect of symbol complexity.  %The graph shows the recognition rate of each symbol using different algorithms. 
+		\includegraphics[scale=0.3]{paperImages/featureFS1only.jpg.eps}
+	\caption{The effect of segmentation on LD-DB accuracy} %The effect of symbol complexity.  %The graph shows the recognition rate of each symbol using different algorithms. 
 	\label{fig:LDtestFeatonly}
@@
 	\centering
-		\includegraphics[scale=0.4]{paperImages/featureFS1only.jpg.eps}
-	\caption{The effect of segmentaion on EL-DB accuracy} %The effect of symbol complexity.  %The graph shows the recognition rate of each symbol using different algorithms. 
+		\includegraphics[scale=0.8]{results/F1EL.eps}
+	\caption{The effect of segmentation on EL-DB accuracy} %The effect of symbol complexity.  %The graph shows the recognition rate of each symbol using different algorithms. 
 	\label{fig:ELtestFeatonly}
@@
 We showed result of three different dataset.
-The system performed better than most known algoirthm on the EL-DB and LD-DB.
-This is due to the focus on strcutre and stroke based properties gained from the sytstem after correctly segmenting the input strokes.
-The result also showed that even using \textsl{AlgS2} only as the segmentaion aglgoirthm the system perform better than other segmenation algorithm.
-Using different databases shows that the more the strucutre difference in the sysmobl the better recogniton rate.
-LD-DB proved to have the lowest recogntion rate due to the high simillarity between its symbols.
+The system performed better than most known algorithm on the EL-DB and LD-DB.
+This is due to the focus on structure and stroke based properties gained from the system after correctly segmenting the input strokes.
+The result also showed that even using \textsl{AlgS2} only as the segmentation algorithm the system perform better than other segmentation algorithm.
+Using different databases shows that the more the structure differences in the symbols the better recognition rate.
+LD-DB proved to have the lowest recognition rate due to the high similarity between its symbols.
 % better than

### 2009-04-22 trunk/draft/SketchMaster.tex
changes for the doc review before making the reults
before change of reuslt on data of 4 appril
after correctly splled now look at any thing else
before dr. nevin change

@@
                                                         %======================
-\title{Sketch Recogniton using Particle Swarm Algorithm}             % My theses title
+\title{Sketch Recognition using Particle Swarm Optimization Algorithm}             % My theses title
 \author{Maha Mohammed Nabeel El Meseery}                               % Y name
-\prevdegrees{B.S.,Computer  Engineering(2009)}       		% Your old degree
+\prevdegrees{B.S.,Computer  Engineering}       		% Your old degree
 \major{Computer Engineering}                            % My current major
@@
 %\thesisdate[]                                     % Date of document.
-\DOCUMENTtype{Survey}                                   % or DISSERTATION
-\Documenttype{Survey}                                   % or Dissertation
-\documenttype{Survey}                                   % or dissertation
+\DOCUMENTtype{Thesis}                                   % or DISSERTATION
+\Documenttype{Thesis}                                   % or Dissertation
+\documenttype{Thesis}                                   % or dissertation
 \maketitle
@@
                                                         %======================
-\copyrightyear{2009}                                    % Delete these
-\makecopyright                                          % if no copyright
+%\copyrightyear{2009}                                    % Delete these
+%\makecopyright                                          % if no copyright
                                                     % APPROVAL                                                         %======================
-%\supervisor{A. B. Smith}{Associate Professor of xxx}    % <- One of these
-%\committee{A. B. Jones}{Associate Professor of xxx}     % <- As many as
-%\committee{A. B. Black}{Faculty in Residence in xxx}    %    you need...
+\thispagestyle{empty}               % Don't do headers or footers.
+\newpage                                                             
+\begin{center}     
+\LARGE  
+\textbf{Sketch Recognition using Particle Swarm Optimization Algorithm}\\
+\normalsize by      \\ 
+\Large  \textbf{ Maha mohamed Nabeel El Meseery}\\
+\vspace*{0.4in}        
+\normalsize
+A Thesis Submitted To the\\
+Faculty of Engineering, Cairo University\\
+in Partial Fulfillment of the\\
+Requirements for the Degree of\\
+MASTER OF SCIENCE
+in\\
+COMPUTER ENGINEERING\\
+
+\end{center}
+
+    Approved by\\                                           
+Examining Committee:\\
+ \makebox[6in]{\hrulefill}\\ 
+Professor Nevin Darwish, Thesis Main Advisor\\
+ \makebox[6in]{\hrulefill}\\ 
+Professor Samia Mashali,Thesis Advisor \\
+ \makebox[6in]{\hrulefill}\\  
+Professor Magda Fayek,Thesis Advisor\\ 
+ \makebox[6in]{\hrulefill}\\ 
+Professor  Ashraf Abd El Wahab, Member\\
+ \makebox[6in]{\hrulefill}\\ 
+Professor Mohamed Zaki Abd ElMegeed, Member\\
+ \makebox[6in]{\hrulefill}\\ 
+  % \\   you need...\\
 %\makeapproval                                           %
@@
                                                         %======================
-%\begin{acknowledgments}                                 % Delete these if
- %\end{acknowledgments}                                   %
+\begin{acknowledgments}                                 % Delete these if
+All praise is due to Allah Who guided me to this. I could not truly have been led aright if
+Allah had not guided me.
+
+I would like to convey my sincere appreciation and graditude to my supervisors; Dr.
+Nevin Darwish, Dr.
+Magdah Fayek and Dr.
+Samia Mashali.
+I would also be very gratuful for dr.
+Mahmoud fakhr el din for his assitance with my thesis.
+Thanks for all the time devoted in making constructive comments and suggestions for improving this thesis.
+I thank my family and friends for thier moral and support.
+I specifically thank, dina said, noha mansour and marwah shafee.
+Thanks, for being always there when I needed you the most.
+I hope to express my gratiude to all of you for supproting me and praying for me through all these years.
+ %Finally, I am very grateful to my dear parents, my family, and my friends whom I consider
+%as my sisters. I would like to thank particularly Radwa Aboudina, Fatma Nada, Maha
+%Nabil, Haidi Badr, and Marwa Kamal. Thank you all for being always there when I needed
+%you most. Thank you for believing in me and supporting me through all these years. I think
+%without your support and your prayers, none of this work would be accomplished.
+%ii
+ \end{acknowledgments}                                   %
 
@@
 % didn't finish it all Need some addint
-
 The need for new level of human computer interaction had lead the researchers to attempt improve the current interfaces.
-Even though the area of research progressed quickly in speech and handwriting systems gesture and sketch understanding systems are still starting to get noticed.
+Even though the area of research progressed quickly in speech and handwriting systems, gesture and sketch understanding systems are still starting to get noticed.
 
@@
 This research uses Particle Swarm Algorithm (PSO) to divide the strokes the user draws into meaningful geometric primitives.
-These geometric primitives are grouped to formulate symbols which will be further identified.
-The research focus on the effect of PSO segmentation to the sketch recognition systems.
+These geometric primitives are grouped to formulate symbols which are further identified.
+Two algorithm are used to divide strokes, the first algorithm \textsl{ALgS1} divides stroke to polygons.
+The second one \textit{AlgS2}, divides stroke to segments of either lines or curves.
+The research focuses on the effect of PSO segmentations to the sketch recognition systems.
+The final recognition is achieved using a set of geometrical and global shape properties features.
+A SVM (support vector machine) classifier is used to correctly identify shapes.
 
-Results showed that using PSO improves segmentation results which guide the symbol recognition phase. Final recognition has improved accuracy after the PSO algorithm using a SVM classifier.  
- %Hand drawn sketches are the natural method of exchanging thoughts between engineers and humans in general.
-Nevertheless, researchers barely give attention to improve current gesture and sketch recognition systems.
-In a sketch understanding system, users need to feel the freedom of paper and pencil along with the flexibility of computer manipulation.
+Experiments were conducted on three different datasets; Hs-DB a benchmark dataset which contains simple presentation symbols, a LD-DB a set of logic design symbols and EL-DB Electrical symbols.
+Results show that using \textit{ALgS2} to segment stroke improves segmentation than known algorithms.
+\textit{ALgS1} achieved good performance on EL-DB and Hs-DB but performed poorly on LD-DB due to high number of curves in every symbol.
+%The segmentation improvement is tested by final symbol recognition accuracy.
 
@@
 \include{ChapterIntroduction}                              % The best way to organize your
-%\include{chapterSurvey}                              % document is probably to keep
+\include{chapterSurvey}                              % document is probably to keep
 \include{System}
 \include{Results}
-%\include{ConFut}
+\include{ConFut}
                                                 % in its own file.  
 %\include {Mybibliography}
-
 \bibliography{../../neededfiles/Bibliographies/Mybibliography}     

### 2009-04-22 trunk/draft/Results.tex
aftetr my change

@@
 These results are reported in section \ref{sec:ResultsDetails}.
-The effect of number of agents, the error thresholds and various other PSO parameters are investigated in Section \ref{sec:PSO}.
+The effect of number of particles, the error thresholds and various other PSO parameters are investigated in Section \ref{sec:PSO}.
 %The results presented in Section \ref{sec:ResultsDetails} and \ref{sec:PSO}.
@@
 		}
-	\caption{The Symbol Set}
+	\caption[Hs-DB Symbol Set]{The symboles Set in \cite{HeloiseBeautification} } 
 	\label{fig:symbolSet}
@@
 
-	\caption{Example of Logic Symbols}
+	\caption[Symbols in LD-DB]{Example of Logic Symbols}
 	\label{fig:LogicSet}
@@
 
-	\caption{Example of Electrical Symbols}
+	\caption[Symbols in EL-DB]{Example of Electrical Symbols}
 	\label{fig:ElectSet}
@@
 
- The effect of number of agents, the error thresholds and various other parameters was investigated.
+ The effect of number of particles, the error thresholds and various other parameters was investigated.
 The result of these tests are shown in Figure \ref{fig:swarmtesting} and \ref{fig:swarmtesting2}.
-Figure \ref{fig:swarmtesting} shows that as the number of agents increase the \textit{Segmentation Error} value decrease.
+Figure \ref{fig:swarmtesting} shows that as the number of particles increase the \textit{Segmentation Error} value decrease.
 Similar results for the maximum swarm iterations are shown in Figure \ref{fig:swarmtesting}.
@@
 The final values are for maximum number of iteration = 80 with 15 particles.
-The final number of agents and maximum swarm iteration used in the system was based trade off between errors achieved versus the computation time.
+The final number of particles and maximum swarm iteration used in the system was based trade off between errors achieved versus the computation time.
 
@@
 	 \includegraphics[scale=0.5]{paperImages/swarmAgentTest.jpg.eps}
-	 	\caption{Segmentation Error Vs. Number of agents}
+	 	\caption{Segmentation Error Vs. Number of particles}
 	 	\label{fig:swarmtesting}
@@
 	\label{fig:LDRes}
-
 \end{figure}
@@
 \end{figure}
-
-  Figure \ref{} and \ref{} shows examples of correctly recognized symbol using our system.
-The figure shows that system recognized some hard to recognize example in both EL-DB and DL-DB.
-An example of symbols that was not correctly recognized are in Figure \ref{}.
-Most of the wrong classifications were due to how messy the user draw symbols and how similarly they are conceptual.
-Table \ref{} show the confusion matrix of EL-DB which shows that most error were between capacitor, batter and crystal which differ only on number of parallel lines.
-Table \ref{tab:ConfusionMatrix} (confusion matrix) shows the types of error in Hs-DB which shows that the system was confused between parallelogram and trapezoid, moon and arch.
-Those errors were due to users drawing styles as some user may draw a parallelogram as a trapezoid.
-
  \begin{figure}
+	\centering
 	\includegraphics[scale=0.5]{results/samplesofHS-DB.jpg.eps}
@@
 \end{figure}
+  Figure \ref{fig:correct2.jpg} shows examples of correctly recognized symbol using our system.
+The figure shows that system recognized some hard to recognize example in both EL-DB and DL-DB.
+An example of symbols that was not correctly recognized are in Figure \ref{fig:erros}.
+Most of the wrong classifications were due to how messy the user draw symbols and how similarly they are conceptual.
+Table \ref{tab:ConfusionMatrix} (confusion matrix) shows the types of error in Hs-DB which shows that the system was confused between parallelogram and trapezoid, moon and arch.
+Those errors were due to users drawing styles as some user may draw a parallelogram as a trapezoid.
+
+  %55Table \ref{} show the confusion matrix of EL-DB which shows that most error were between capacitor, batter and crystal which differ only on number of parallel lines. 
+\begin{figure}
+	\centering
+		\includegraphics{results/erros.jpg.eps}
+	\caption{Error in Recognition }
+	\label{fig:erros}
+\end{figure}
+\begin{figure}
+	\centering
+		\includegraphics{results/correct2.jpg.eps}
+	\caption{Symbols Correctly Recognized}
+	\label{fig:correct2.jpg}
+\end{figure}
+
  \begin{table*}
@@
 	\centering
-		\includegraphics[scale=0.1]{paperImages/LDsymbols.jpg.eps}
+		\includegraphics[scale=0.5]{results/ELSymbols.eps}
 	\caption{Symbols Comparison on EL-DB symbols}.  %The graph shows the recognition rate of each symbol using different algorithms. 
@@
 	\centering
-		\includegraphics[scale=0.4]{paperImages/featureFS1only.jpg.eps}
+		\includegraphics[scale=0.3]{paperImages/featureFS1only.jpg.eps}
 	\caption{The effect of segmentation on LD-DB accuracy} %The effect of symbol complexity.  %The graph shows the recognition rate of each symbol using different algorithms. 
@@
 	\centering
-		\includegraphics[scale=0.4]{paperImages/featureFS1only.jpg.eps}
+		\includegraphics[scale=0.8]{results/F1EL.eps}
 	\caption{The effect of segmentation on EL-DB accuracy} %The effect of symbol complexity.  %The graph shows the recognition rate of each symbol using different algorithms. 
@@
 We showed result of three different dataset.
-The system performed better than most known algorithm on the EL-DB and LD-DB.
+The system performed better than the benchmark algorithm on the EL-DB and LD-DB.
 This is due to the focus on structure and stroke based properties gained from the system after correctly segmenting the input strokes.
-The result also showed that even using \textsl{AlgS2} only as the segmentation algorithm the system perform better than other segmentation algorithm.
+The result also showed that even using \textsl{AlgS2} only as the segmentation algorithm the system perform better than other segmentation algorithms.
 Using different databases shows that the more the structure differences in the symbols the better recognition rate.

### 2009-04-22 trunk/draft/chapterSurvey.tex
aftetr my change

@@
 		\includegraphics[scale=0.7]{../../neededfiles/Figures/Overshootandundershoot.eps}
-	\caption[Neat and Sloppy Symbols] {A sloppy and neat versions of symbol.}
+	\caption[Neat and Sloppy Symbols] {A sloppy and neat versions of symbol \cite{threeproblmes23}.}
 	\label{fig:Overshootandundershoot}
@@
 		\includegraphics[scale=0.7]{../../neededfiles/Figures/overstroked.eps}
-	\caption{Example of an over traced stroke}
+	\caption[Example of an over traced stroke]{Example of an over traced stroke \cite{overtraced24}}
 	\label{fig:overstroked}
@@
 				\includegraphics{images/featureareline.eps}}
-	\caption{Feature Area}
+	\caption[Feature Area]{Feature Area \cite{meanshift10}}
 	\label{fig:featurearearc}
@@
 		\includegraphics[scale=0.7]{../../neededfiles/Figures/squarescematic.eps}
-	\caption{Semantic Graph of a Square }
+	\caption[Semantic Graph of a Square ]{Semantic Graph of a Square \cite{mulitstroke5}}
 	\label{fig:squarescematic}
@@
 			\hfill
-		\subfigure[Arrow grammar definition]
+		\subfigure[Arrow grammar definition]%{\cite{Ladder30}}
 		 {\label{fig:arr2}\includegraphics[scale=0.88]{../../neededfiles/Figures/arrowladderdef2.eps}}
-	\caption[Arrow Definition]{}
+	\caption[Arrow Definition]{\cite{Ladder30}}
 		\label{fig:arrowladderdef1}
@@
 		\includegraphics[scale=0.7]{images/template.eps}
-	\caption{Prototypes of Shapes}
+	\caption[Prototypes of Shapes]{Prototypes of Shapes \cite{imagetrainable48}}
 	\label{fig:template}
@@
 \label{sec:ParticleSwramOptimizationReview}
-
 %This seems to emerge of what is called 'collective intelligence' from often large simple agents.  

### 2009-04-22 draft/System.tex
aftetr my change

@@
 		\includegraphics[scale=0.8]{images/stroke3.jpg.eps}
-	\caption{Input Stroke} Example of an input stroke to the system. 
+	\caption{Example of Input Stroke} Example of an input stroke to the system. 
 	\label{fig:orignalStroke}
@@
 		\includegraphics[scale=0.8]{images/orignalStroke.eps}
-	\caption{Input Stroke} Example of an input stroke to the system. 
+	\caption{Another Example of Input Stroke} Example of an input stroke to the system. 
 	\label{fig:AnotherorignalStroke}
@@
 The input stroke with $N$ points can be represented by set $S = \left\{ {x_1 ,x_2  \ldots x_N } \right\}$ where $x_i$ is the location of the point $i$ .
-The swarm algorithms consist of $M$ agents which are represented by the set
+The swarm algorithms consist of $M$ particles which are represented by the set
 $A  = \left\{ {P_i \left| {i = 1,2 \cdots M} \right.} \right\}$ where $P_i$ is a single solution particle from the solution space.

### 2009-04-22 draft/Results.tex
images and stuff

@@
 	\centering		
-	 \includegraphics[scale=0.5]{paperImages/swarmAgentTest.jpg.eps}
-	 	\caption{Segmentation Error Vs. Number of agents}
+	 \includegraphics[scale=0.5]{paperImages/swarmAgentTest.eps}
+	 	\caption{Segmentation Error Vs. Number of particles}
 	 	\label{fig:swarmtesting}
 	%\caption{Experiments results :} a)    b) c) 
+\end{figure} 
+\begin{figure}
+	\centering
+		\includegraphics{paperImages/swarmIterationTest.eps}
+	\caption{Segmentation Error Vs. Number of Swarm iteration}
+	\label{fig:swarmIterationTest}
 \end{figure}
@@
 	\includegraphics[scale=0.5]{results/samplesofHS-DB.jpg.eps}
-	\caption{Hs-DB shapes drawn by single user}
+	\caption[Hs-DB shapes drawn by single user]{Hs-DB shapes drawn by single user \cite{HeloiseBeautification} }
 	\label{fig:samplesofHS-DB.jpg}

### 2009-04-22 draft/chapterSurvey.tex
images and stuff

@@
 Since 1995 when Kennedy and Eberhart \cite{PSOFirst,PSO2} introduced particle swarm algorithm (PSO) the algorithm had been used in various image processing system as an optimization  algorithm \cite{SwarmImageProcessing,Swarmimageprocessing2,Colorquantization,SwarmMedical}.
-PSO algorithm was used in \cite{ToADD} to classify pixel to simplify the object segmentation process.
+PSO algorithm was used in \cite{PSOIMageClassification} to classify pixel to simplify the object segmentation process.
 

### 2009-04-22 draft/ChapterIntroduction.tex
aftetr my change

@@
 		\includegraphics[scale=0.8]{../../neededfiles/Figures/strokeinterpolation.eps}
-	\caption[Segmentation Error]{Different ways for segmenting a stroke drawn by user. a) The stroke users draw b)interpretation as two line segments c) interpretation as an arc.}
+	\caption[Segmentation Error]{Different ways for segmenting a stroke drawn by user.
+a) The stroke users draw b)interpretation as two line segments c) interpretation as an arc.
+This example is adapted from \cite{earlyprocess}}
 	%The figure shows an arc that the user draw which can be segmented in different ways as a single line or as an arc where a) The stroke users draw  b)interpretation as a single line segment b) interpretation as an arc.}
@@
 		\includegraphics[scale=0.85]{../../neededfiles/Figures/VorU.eps}
-	\caption[Handwriting Ambiguities] { The same letter can be interpreted differently in various contexts; first as V in drive or as U in Cut}
+	\caption[Handwriting Ambiguities] { The same letter can be interpreted differently in various contexts; first as V in drive or as U in Cut. This example is adapted from \cite{Ambiguity}.}
 \label{fig:VorU}

### 2009-04-22 draft/Results.tex
sss

@@
 The dataset was divided into 75\% training set and 25\% test set.
-Five different splits for the training and test data are generated from the dataset.
-In worth noting that \cite{HeloiseBeautification} and \cite{Oltmans07} achieved 96.7\% and  94.4\%  on this dataset respectively.
+Five different splits for the training and test data are generated from the dataset.% In worth noting that \cite{HeloiseBeautification} and \cite{Oltmans07} achieved 96.7\% and  94.4\%  on this dataset respectively.
 %, 94.4\%  , which
@@
  Categories 	&ellipse	&heart	&trapezoid	&pentagon	&arch	&hexagon	&square	&triangle	&cube	&cylinder	&parallelogram	&moon	&callout\\ \hline 
-ellipse	&155	&9	&0	&0	&0	&5	&0	&0	&0	&0	&0	&6	&0\\ \hline
-heart	&0	&109	&1	&1	&25	&0	&0	&1	&0	&0	&2	&1	&1\\ \hline
-trapezoid	&0	&0	&103	&5	&2	&0	&0	&2	&0	&0	&34	&0	&0\\ \hline
-pentagon	&0	&0	&20	&121	&0	&20	&6	&2	&0	&0	&1	&0	&0\\ \hline
-arch	&0	&1	&13	&0	&86	&0	&0	&2	&3	&0	&2	&0	&0\\ \hline
-hexagon	&0	&0	&0	&0	&0	&98	&0	&0	&0	&0	&0	&0	&0\\ \hline
-square	&0	&0	&1	&0	&2	&0	&107	&14	&0	&0	&8	&1	&0\\ \hline
-triangle	&0	&0	&7	&0	&1	&0	&0	&89	&0	&0	&12	&4	&0\\ \hline
-cube	&0	&0	&0	&0	&0	&0	&0	&0	&147	&0	&0	&0	&0\\ \hline
-cylinder	&0	&0	&0	&6	&2	&8	&16	&0	&4	&156	&1	&10	&0\\ \hline
-parallelogram	&0	&0	&4	&0	&0	&0	&0	&0	&0	&0	&91	&0	&0\\ \hline
-moon	&0	&3	&3	&2	&17	&0	&9	&17	&0	&0	&2	&133	&11\\ \hline
-callout	&1	&34	&0	&22	&21	&22	&16	&25	&0	&0	&0	&0	&143\\ \hline
+ellipse	&150	&20	&0	&0	&0	&5	&1	&0	&0	&0	&0	&1	&0\\  \hline 
+heart	&2	&130	&0	&1	&10	&1	&1	&0	&0	&0	&1	&4	&0\\  \hline 
+trapezoid	&0	&1	&147	&5	&1	&0	&1	&3	&2	&0	&37	&0	&0\\  \hline 
+pentagon	&0	&0	&1	&139	&0	&7	&0	&0	&0	&0	&1	&0	&0\\   \hline 
+arch	&0	&0	&1	&1	&125	&0	&4	&0	&3	&0	&0	&15	&1\\   \hline 
+hexagon	&0	&0	&0	&2	&0	&135	&0	&0	&0	&0	&0	&0	&0\\  \hline 
+square	&0	&0	&0	&0	&0	&0	&138	&0	&0	&0	&0	&1	&0\\   \hline 
+triangle	&0	&0	&1	&2	&5	&0	&1	&148	&0	&0	&4	&7	&0\\  \hline 
+cube	&0	&0	&0	&1	&0	&0	&0	&0	&146	&5	&0	&0	&0\\   \hline 
+cylinder	&0	&0	&0	&1	&0	&1	&0	&0	&0	&147	&0	&0	&0\\  \hline 
+parallelogram	&0	&0	&0	&0	&0	&0	&2	&0	&2	&0	&107	&0	&0\\  \hline 
+moon	&0	&1	&0	&0	&10	&2	&3	&0	&0	&0	&0	&125	&5\\   \hline 
+callout	&0	&0	&0	&0	&0	&0	&0	&0	&0	&0	&0	&0	&145\\  \hline 
  		\end{tabular}
@@
  	%}
-	\caption{Confusion Matrix}
+	\caption{Confusion Matrix of HS-DB}
 	\label{tab:ConfusionMatrix}
@@
 Different feature sets are tested to determine the best features that can used in sketch and symbol recognition.
-Figure\ref{fig:testFeaturesAllHS} shows the result of different feature sets from the basic four sets (\textbf{FS1,FS2,FS3,FS4}).
+Figure\ref{fig:testFeaturesAll} shows the result of different feature sets from the basic four sets (\textbf{FS1,FS2,FS3,FS4}).
 Result shows that (\textbf{FS2}) Rubine features \cite{gestureexample12} gives the worst results when used alone without any other features.
@@
 Features \textbf{(FS4)} gives good results but it is improved by adding structural features \textbf{(FS1)}.
-Figure \ref{fig:testFeaturesAllMD} shows the similar results on \textsl{EL-DB} and \textsl{LD-DB} dataset.
+Figures \ref{fig:testFeaturesLDDB} and \ref{fig:ELtestFeaturesAll} show similar results on \textsl{EL-DB} and \textsl{LD-DB} dataset.
 The figures show that on both datasets \textbf{FS1} perform better than on \textsl{HS-DB}, this is a result that shapes in the \textsl{EL-DB} and \textsl{LD-DB} has more structural and conceptual differences than shapes in \textsl{HS-DB}.
@@
 This is due to the level of details in the EL-DB and DL-DB sets.
+Result shows that combining feature set FS4 and FS1 gives results compared to moments in both datasets \textsl{EL-DB} and \textsl{LD-DB}.
 % achieved  HS-DB a
@@
 	\caption{Feature Comparison on LD-DB} The effect of different features on accuracy.  %The graph shows the recognition rate of each symbol using different algorithms. 
-	\label{fig:testFeaturesAll}
+	\label{fig:testFeaturesLDDB}
 \end{figure}
@@
 	\caption{Feature Comparison on EL-DB} The effect of different features on accuracy.  %The graph shows the recognition rate of each symbol using different algorithms. 
-	\label{fig:testFeaturesAll}
+	\label{fig:ELtestFeaturesAll}
 \end{figure}
@@
 \end{figure}  
- \begin{figure}
-	\centering
-		\includegraphics[scale=0.3]{paperImages/featureFS1only.jpg.eps}
-	\caption{The effect of segmentation on LD-DB accuracy} %The effect of symbol complexity.  %The graph shows the recognition rate of each symbol using different algorithms. 
-	\label{fig:LDtestFeatonly}
-\end{figure}  
+% \begin{figure}
+%	\centering
+%		\includegraphics[scale=0.3]{paperImages/featureFS1only.jpg.eps}
+%	\caption{The effect of segmentation on LD-DB accuracy} %The effect of symbol complexity.  %The graph shows the recognition rate of each symbol using different algorithms. 
+%	\label{fig:LDtestFeatonly}
+%\end{figure}  
  \begin{figure}

### 2009-04-27 draft/SketchMaster.tex
going to edit the survey
before i change the results

@@
 \usepackage{psfig}
-%\usepackage{named}
+\usepackage{named}
 \usepackage{natbib}
@@
                                                         %======================
-\title{Sketch Recogniton using Particle Swarm Algorithm}             % My theses title
+\title{Sketch Recognition and Understanding}             % My theses title  using Particle Swarm Optimization Algorithm
 \author{Maha Mohammed Nabeel El Meseery}                               % Y name
-\prevdegrees{B.S.,Computer  Engineering(2003)}       		% Your old degree
+\prevdegrees{B.S.,Computer  Engineering}       		% Your old degree
 \major{Computer Engineering}                            % My current major
@@
                                                         %======================
-\copyrightyear{2008}                                    % Delete these
-\makecopyright                                          % if no copyright
-                                                        % page.
-
-                                                        % APPROVAL PAGE
-                                                        %======================
-%\supervisor{A. B. Smith}{Associate Professor of xxx}    % <- One of these
-%\committee{A. B. Jones}{Associate Professor of xxx}     % <- As many as
-%\committee{A. B. Black}{Faculty in Residence in xxx}    %    you need...
+%\copyrightyear{2009}                                    % Delete these
+%\makecopyright                                          % if no copyright
+                                                    % APPROVAL                                                         %======================
+%\thispagestyle{empty}               % Don't do headers or footers.
+%\newpage                                                             
+%\begin{center}     
+%\LARGE  
+%\textbf{Sketch Recognition using Particle Swarm Optimization Algorithm}\\
+%\normalsize by      \\ 
+%\Large  \textbf{ Maha mohamed Nabeel El Meseery}\\
+%\vspace*{0.4in}        
+%\normalsize
+%A Thesis Submitted To the\\
+%Faculty of Engineering, Cairo University\\
+%in Partial Fulfillment of the\\
+%Requirements for the Degree of\\
+%MASTER OF SCIENCE
+%in\\
+%COMPUTER ENGINEERING\\
+%
+%\end{center}
+%          
+%    Approved by\\                                           
+%Examining Committee:\\
+% \makebox[6in]{\hrulefill}\\ 
+%Professor Nevin Darwish, Thesis Main Advisor\\
+% \makebox[6in]{\hrulefill}\\ 
+%Professor Samia Mashali,Thesis Advisor \\
+% \makebox[6in]{\hrulefill}\\  
+%Assoc. Professor Magda Fayek,Thesis Advisor\\ 
+% \makebox[6in]{\hrulefill}\\ 
+%Professor  Ashraf Abd El Wahab, Member\\
+% \makebox[6in]{\hrulefill}\\ 
+%Professor Mohamed Zaki Abd ElMegeed, Member\\
+% \makebox[6in]{\hrulefill}\\ 
+  % \\   you need...\\
 %\makeapproval                                           %
@@
                                                         %======================
-%\begin{acknowledgments}                                 % Delete these if
- %\end{acknowledgments}                                   %
+\begin{acknowledgments}                                 % Delete these if
+All praise is due to Allah Who guided me to this. I could not truly have been led aright if
+Allah had not guided me.
+
+I would like to convey my sincere appreciation and graditude to my supervisors; Dr.
+Nevin Darwish, Dr.
+Magdah Fayek and Dr.
+Samia Mashali.
+I would also be very gratuful for Dr.
+Mahmoud Fakhr El Din for his assitance with my thesis.
+Thanks for all the time devoted in making constructive comments and suggestions for improving this thesis.
+I thank my family and friends for thier moral and support.
+I specifically thank, dina said, noha mansour and marwah shafee.
+Thanks, for being always there when I needed you the most.
+I hope to express my gratiude to all of you for supproting me and praying for me through all these years.
+ %Finally, I am very grateful to my dear parents, my family, and my friends whom I consider
+%as my sisters. I would like to thank particularly Radwa Aboudina, Fatma Nada, Maha
+%Nabil, Haidi Badr, and Marwa Kamal. Thank you all for being always there when I needed
+%you most. Thank you for believing in me and supporting me through all these years. I think
+%without your support and your prayers, none of this work would be accomplished.
+%ii
+ \end{acknowledgments}                                   %
 
@@
 This research uses Particle Swarm Algorithm (PSO) to divide the strokes the user draws into meaningful geometric primitives.
-These geometric primitives are grouped to formulate symbols which will be further identified.
-The research focus on the effect of PSO segmentation to the sketch recognition systems.
+These geometric primitives are grouped to formulate symbols which are further identified.
+Two algorithm are used to divide strokes, the first algorithm \textit{ALgS1} divides stroke to polygons.
+The second one \textit{AlgS2}, divides stroke to segments of either lines or curves.
+The research focuses on the effect of PSO segmentations to the sketch recognition systems.
+The final recognition is achieved using a set of geometrical and global shape properties features.
+A SVM (support vector machine) classifier is used to correctly identify shapes.
 
-Results showed that using PSO improves segmentation results which guide the symbol recognition phase. Final recognition has improved accuracy after the PSO algorithm using a SVM classifier.  
- %Hand drawn sketches are the natural method of exchanging thoughts between engineers and humans in general.
-Nevertheless, researchers barely give attention to improve current gesture and sketch recognition systems.
-In a sketch understanding system, users need to feel the freedom of paper and pencil along with the flexibility of computer manipulation.
+Experiments were conducted on three different datasets; Hs-DB a benchmark dataset which contains simple presentation symbols, a LD-DB a set of logic design symbols and EL-DB Electrical symbols.
+Results show that using \textit{ALgS2} to segment stroke improves segmentation than known algorithms.
+\textit{ALgS1} achieved good performance on EL-DB and Hs-DB but performed poorly on LD-DB due to high number of curves in every symbol.
+%The segmentation improvement is tested by final symbol recognition accuracy.
 
@@
 \include{ChapterIntroduction}                              % The best way to organize your
-%\include{chapterSurvey}                              % document is probably to keep
-%\include{System}
-%\include{Results}
-%\include{ConFut}
+\include{chapterSurvey}                              % document is probably to keep
+\include{System}
+\include{Results}
+\include{ConFut}
                                                 % in its own file.  
 %\include {Mybibliography}
-
 \bibliography{../../neededfiles/Bibliographies/Mybibliography}     

### 2009-04-27 draft/SketchMaster.tex

@@
                                                         %======================
-\begin{acknowledgments}                                 % Delete these if
-All praise is due to Allah Who guided me to this. I could not truly have been led aright if
-Allah had not guided me.
+\begin{acknowledgments}     
+                            % Delete these if
+All praise is due to Allah Who guided me to this. I could not truly have been led aright if Allah had not guided me.
 
-I would like to convey my sincere appreciation and graditude to my supervisors; Dr.
+I would like to convey my sincere appreciation and gratitude to my supervisors; Dr.
 Nevin Darwish, Dr.
@@
 Samia Mashali.
-I would also be very gratuful for Dr.
-Mahmoud Fakhr El Din for his assitance with my thesis.
+I would also be very grateful for Dr.
+Mahmoud Fakhr El Din for his assistance with my thesis.
 Thanks for all the time devoted in making constructive comments and suggestions for improving this thesis.
-I thank my family and friends for thier moral and support.
+
+ I thank my family and friends for their moral and support.
 I specifically thank, dina said, noha mansour and marwah shafee.
 Thanks, for being always there when I needed you the most.
-I hope to express my gratiude to all of you for supproting me and praying for me through all these years.
+I hope to express my gratitude to all of you for supporting me and praying for me through all these years.
  %Finally, I am very grateful to my dear parents, my family, and my friends whom I consider

### 2009-04-27 trunk/draft/SketchMaster.tex
sss
change to report

@@
 %\documentstyle[11pt,unhthesis]{report}
-\documentclass[11pt,doublespace]{Sketchthesis}
+\documentclass[11pt,doublespace]{report}%{Sketchthesis}
 
@@
                                                         %======================
-\title{Sketch Recognition using Particle Swarm Optimization Algorithm}             % My theses title
-\author{Maha Mohammed Nabeel El Meseery}                               % Y name
-\prevdegrees{B.S.,Computer  Engineering}       		% Your old degree
-\major{Computer Engineering}                            % My current major
-\degree{Master of Science}                              % My new degree
-\degreemonth{}                                      % When awarded. to be added when known
-\degreeyear{}                                      %
-%th today or first day of writing  December 26 2005 
-%\thesisdate[]                                     % Date of document.
-\DOCUMENTtype{Thesis}                                   % or DISSERTATION
-\Documenttype{Thesis}                                   % or Dissertation
-\documenttype{Thesis}                                   % or dissertation
+          % My theses title  using Particle Swarm Optimization Algorithm
+%\author{Maha Mohammed Nabeel El Meseery}                               % Y name
+%\prevdegrees{B.S.,Computer  Engineering}       		% Your old degree
+%\major{Computer Engineering}                            % My current major
+%\degree{Master of Science}                              % My new degree
+%\degreemonth{}                                      % When awarded. to be added when known
+%\degreeyear{}                                      %
+%%th today or first day of writing  December 26 2005 
+%%\thesisdate[]                                     % Date of document.
+%\DOCUMENTtype{Thesis}                                   % or DISSERTATION
+%\Documenttype{Thesis}                                   % or Dissertation
+%\documenttype{Thesis}
+
+   \vspace*{0.375in}
+               {\Large\bf\title{Sketch Recognition and Understanding}  }
+                                 \vspace*{0.375in}
+
+                                \vspace*{\fill}\par
+                                      BY
+                                \vspace*{\fill}\par
+                               {\Large \author{Maha Mohammed Nabeel El Meseery }}
+                                     \par
+                               B.Sc.,Computer  Engineering
+                                \vspace*{\fill}\par
+                                    Thesis
+                                \vspace*{\fill}\par
+                           Submitted to the Faculty of Engineering Cairo University\\
+                         in partial fulfillment of \\
+                   the requirements for the degree\ of
+                                \vspace*{\fill}\par
+                                   Master of Science
+                                     \par
+                                      in
+                                     \par
+                                   Master of Science
+                                \vspace*{\fill}\par
+
+                                                        % or dissertation
 \maketitle
@@
                                                     % APPROVAL                                                         %======================
-\thispagestyle{empty}               % Don't do headers or footers.
-\newpage                                                             
-\begin{center}     
-\LARGE  
-\textbf{Sketch Recognition using Particle Swarm Optimization Algorithm}\\
-\normalsize by      \\ 
-\Large  \textbf{ Maha mohamed Nabeel El Meseery}\\
-\vspace*{0.4in}        
-\normalsize
-A Thesis Submitted To the\\
-Faculty of Engineering, Cairo University\\
-in Partial Fulfillment of the\\
-Requirements for the Degree of\\
-MASTER OF SCIENCE
-in\\
-COMPUTER ENGINEERING\\
-
-\end{center}
-
-    Approved by\\                                           
-Examining Committee:\\
- \makebox[6in]{\hrulefill}\\ 
-Professor Nevin Darwish, Thesis Main Advisor\\
- \makebox[6in]{\hrulefill}\\ 
-Professor Samia Mashali,Thesis Advisor \\
- \makebox[6in]{\hrulefill}\\  
-Professor Magda Fayek,Thesis Advisor\\ 
- \makebox[6in]{\hrulefill}\\ 
-Professor  Ashraf Abd El Wahab, Member\\
- \makebox[6in]{\hrulefill}\\ 
-Professor Mohamed Zaki Abd ElMegeed, Member\\
- \makebox[6in]{\hrulefill}\\ 
+%\thispagestyle{empty}               % Don't do headers or footers.
+%\newpage                                                             
+%\begin{center}     
+%\LARGE  
+%\textbf{Sketch Recognition using Particle Swarm Optimization Algorithm}\\
+%\normalsize by      \\ 
+%\Large  \textbf{ Maha mohamed Nabeel El Meseery}\\
+%\vspace*{0.4in}        
+%\normalsize
+%A Thesis Submitted To the\\
+%Faculty of Engineering, Cairo University\\
+%in Partial Fulfillment of the\\
+%Requirements for the Degree of\\
+%MASTER OF SCIENCE
+%in\\
+%COMPUTER ENGINEERING\\
+%
+%\end{center}
+%          
+%    Approved by\\                                           
+%Examining Committee:\\
+% \makebox[6in]{\hrulefill}\\ 
+%Professor Nevin Darwish, Thesis Main Advisor\\
+% \makebox[6in]{\hrulefill}\\ 
+%Professor Samia Mashali,Thesis Advisor \\
+% \makebox[6in]{\hrulefill}\\  
+%Assoc. Professor Magda Fayek,Thesis Advisor\\ 
+% \makebox[6in]{\hrulefill}\\ 
+%Professor  Ashraf Abd El Wahab, Member\\
+% \makebox[6in]{\hrulefill}\\ 
+%Professor Mohamed Zaki Abd ElMegeed, Member\\
+% \makebox[6in]{\hrulefill}\\ 
   % \\   you need...\\
@@
                                                         %======================
-\begin{acknowledgments}                                 % Delete these if
-All praise is due to Allah Who guided me to this. I could not truly have been led aright if
-Allah had not guided me.
+ \def\acknowledgments
+   {\vspace*{0.5in}
+    \thispagestyle{plain}
+    \chapter*{Acknowledgments}
+    \addcontentsline{toc}{section}{Acknowledgments}
 
-I would like to convey my sincere appreciation and graditude to my supervisors; Dr.
+   }
+
+\def\endacknowledgments{\par}
+
+\begin{acknowledgments}     
+                            % Delete these if
+All praise is due to Allah Who guided me to this. I could not truly have been led aright if Allah had not guided me.
+
+I would like to convey my sincere appreciation and gratitude to my supervisors; Dr.
 Nevin Darwish, Dr.
@@
 Samia Mashali.
-I would also be very gratuful for dr.
-Mahmoud fakhr el din for his assitance with my thesis.
+I would also be very grateful for Dr.
+Mahmoud Fakhr El Din for his assistance with my thesis.
 Thanks for all the time devoted in making constructive comments and suggestions for improving this thesis.
-I thank my family and friends for thier moral and support.
+
+ I thank my family and friends for their moral and support.
 I specifically thank, dina said, noha mansour and marwah shafee.
 Thanks, for being always there when I needed you the most.
-I hope to express my gratiude to all of you for supproting me and praying for me through all these years.
+I hope to express my gratitude to all of you for supporting me and praying for me through all these years.
  %Finally, I am very grateful to my dear parents, my family, and my friends whom I consider
@@
 %------------------------------------------------------------------------------
-\begin{abstractpage}          
+\begin{abstract}          
 % ABSTRACT PAGE
@@
 These geometric primitives are grouped to formulate symbols which are further identified.
-Two algorithm are used to divide strokes, the first algorithm \textsl{ALgS1} divides stroke to polygons.
+Two algorithm are used to divide strokes, the first algorithm \textit{ALgS1} divides stroke to polygons.
 The second one \textit{AlgS2}, divides stroke to segments of either lines or curves.
@@
 
-\end{abstractpage}                              % This ends the page.
+\end{abstract}                              % This ends the page.
 

### 2009-04-27 trunk/draft/chapterSurvey.tex
change to report

@@
 %In this section, we review three different types of symbol recognition methods.
- Symbol recognition can be based on: strokes, global shape properties, and appearance \cite{Oltmans07}.
+ Oltmans\cite{Oltmans07} mentions that symbol recognition can be based on: strokes, global shape properties, and appearance or image based\cite{Oltmans07}.
 The role of which each stroke plays in the structure of the sketched symbol is used to determine shapes in stroke based methods.
-In global shape properties approach, systems investigate general shape properties and its underlying strokes.
-Using shape properties relaxes the assumption and role of each stroke in the symbol but it does not truly represent the appearance of the symbol.
-Unlike appearance based methods where individual strokes are disregarded and they focus on the overall appearance of shapes.
+Image or appearance based methods focuses focus on the overall appearance of shapes rather than individual strokes.
+Unlike global shape properties approach, systems investigate general shape properties.
+Using shape properties loosens the assumption and role of each stroke in the symbol but it does not truly represent the symbol visually.
+
  We choose to use the method based on shape property but added some geometrical and structure properties used in most stroke based approaches.
@@
 In our system, we combine Rubine features with others to include the information on how the user draws sketch while recognizing the symbols.
-In \cite{Long00}, Long shows how similarities between gestures can be analyzed.
+%In \cite{Long00}, Long shows how similarities between gestures can be analyzed.
 He used this information to identify gestures that are likely to confuse the recognition system.
@@
 to represent a rectangle the user draws only the left edge of it).
-Because of these reasons, gestures recognition systems are not suitable for free hand drawn sketches.
+Because of these restrictions, gestures recognition systems are not suitable for free hand drawn sketches.
 
@@
 
- Systems that employ such representation must first pre-process strokes to break them into geometric primitives (see section \ref{sec:preprocessing} for a review on preprocessing methods).
-Recognition can then be treated as sub graph matching problem.
+ Systems that employ such representation must first pre-process strokes to break them into geometric primitives ( section \ref{sec:preprocessing} reviews on preprocessing methods).
+Recognition can then be treated as sub graph matching problem or constrain satisfaction problem.
 
@@
 		 {\label{fig:arr2}\includegraphics[scale=0.88]{../../neededfiles/Figures/arrowladderdef2.eps}}
-	\caption[Arrow Definition]{\cite{Ladder30}}
+	\caption[Arrow Definition]{Arrow Definition in \cite{Ladder30}}
 		\label{fig:arrowladderdef1}

### 2009-04-27 trunk/draft/ChapterIntroduction.tex
return to old but adding supervisor to first page

@@
 \pagenumbering{arabic}
-
-Computer research nowadays aimed at facilitating computer human interaction in every way.
+Computer research nowadays aims at facilitating computer human interaction in every way.
 Pen-based interfaces give the user a pencil-paper like feeling that enhances interaction more than the current used keyboard-mouse computer interfaces.
@@
 
-Sketches here stand for any hand drawn drawing the user draws using digital stylus or similar device.
+Sketches here stand for any hand drawing the user draws using digital stylus or similar device.
 Normally, sketches are composed of a set of strokes.
@@
 
-\section{Importance of sketch recognition}
+\section{Importance of Sketch Recognition}
 \label{sec:ImportanceOfSketchRecognition}
-
-Scientists generally and engineers specifically express thoughts and designs using sketches.
+Scientists generally, and engineers specifically express thoughts and designs using sketches.
 Engineers use sketches to exchange designs as a natural method of communication rather than writing or speaking.
 Engineers use paper and pencil design in early stages of design.
-The sketching process itself go through  numerous stages.
-Firstly, early stages with rough drafting designs or notations which are mainly done using paper and pencil.
-After those primary rough sketches are edited and altered until they are stabilized and finally they are transformed into the computer.
+The sketching process itself goes through numerous stages.
+Firstly, early stages with rough drafting designs or notations are mainly done using paper and pencil.
+These primary rough sketches are edited and altered until they are stabilized and finally they are transformed into the computer.
 
@@
 Current CAD systems do not give the designer the ease of altering the early designs.
-Designer had to sketch primary designs on paper and pencil next, the final design is transferred into computer using a CAD system menus.
+Designer had to sketch primary designs on paper and pencil, next the final design is transferred into computer using a CAD system menus.
 The process is time consuming and tiring.
@@
  A sketch recognition and understanding system will provide a powerful tool to the designers to draw sketch as they usually draw using paper and pencil with the advantages of a computer program.
-Designer needs a normal sketching environment that will provide him facilities in paper interface and the modifying and simulation that is in current CAD systems.
+Designers need a normal sketching environment that will provide him facilities in paper interface and the modifying and simulation that is in current CAD systems.
 
@@
 % sketching programs ...
-Applications of sketch recognition are ranging from engineering device simulation and design.
+Applications of sketch recognition rang from engineering device simulation to design.
 Drawing and sketching programs are most likely to benefit from integrating a sketch recognition system into their design environments.
@@
 
-\section{Problem Definition and challenges}
+\section{Problem Definition and Challenges}
 \label{sec:ProblemsAndChallenges}
 
- Understanding hand drawn sketches may be trivial for humans but it is challenging problem when speaking about computer-based system.
-Most of the research made on the sketch recognition is achieved for one particular domain or for strictly a set of pre defined symbols.
-Further, others restrict the user the freedom in drawing which loose the main goals of paper-like-interface systems.
-The only few systems that have achieved the desired degree of freedom for the users are extremely computationally expensive.
+Understanding hand drawn sketches may be trivial for humans but it is a challenging problem when speaking about computer-based system.
+Most of the research made on the sketch recognition is achieved for one particular domain or for strictly a set of pre-defined symbols\cite{Vibratory8,physicalmeaning6}.
+Furthermore, others restrict the user the freedom in drawing which loose the main goals of paper-like-interface systems\cite{gestureexample12,aideddesgin22,sketchinginterfaces2}.
+The only few systems that have achieved the desired degree of freedom for the users are extremely computationally expensive\cite{EfficientAbstract39,SketchRead2007}.
 
-Sketch recognition is divided into three steps preprocessing, segmentation and symbol recognition.
+Sketch recognition is divided into three steps: preprocessing, segmentation and symbol recognition.
 The preprocessing phase captures user input stroke points and collects basic information about the stroke then proceeds to remove noise and compute basic statistical and geometrical information.
-In segmentation phase, the strokes are divided into a set of simple geometrical primitives or segments.
+In the segmentation phase, the strokes are divided into a set of simple geometrical primitives or segments.
 In the third phase, sketch recognition, strokes and segments are clustered to formulate symbols that can be recognized by a classifier system.
-
-The research in the area had faced few problems that will need to be handled to implement an entirely sketch recognition system.
-The user draws strokes in a spontaneous and ambiguous manner that intensifies the challenges.
-One of the problems is segmenting strokes from the way the user draw it into geometrical primitive segments.
+There are several problems in this research area that need to be handled to implement an integrated sketch recognition system.
+The fact that the user draws strokes in a spontaneous and ambiguous manner intensifies the challenges.
+One of the main problems is segmenting spontaneous strokes drawn by the user into geometrical primitive segments.
 Another major problem is the high level of ambiguities in the drawing.
 The next sections describe each problem in details.
+
 \subsection {Segmentation}
-Segmentation is dividing the strokes draws into the geometric primitive's as lines, arcs, and ellipses.
-The strokes as mentioned before are the path of points between pen-down and pen-up events while using a stylus.
-After that, those points are processed to detect the critical points or vertices which divide the stroke into segments.
+The goal of segmentation is dividing the strokes into the geometric primitive's as lines, arcs, and ellipses.
+The strokes, as mentioned before are the path of points between pen-down and pen-up events while using a stylus.
+Furthermore, after identifying a stroke, it should be processed to detect the critical points or vertices which divide the stroke into segments.
 Later, each segment is classified as line, arc, ellipse or circle.
-The main problem in this process is in detecting the critical and vertices which divide strokes into lines and arc.
-The user draws the lines and arc in an undefined manner which make it difficult for the program to decide where the user had finish a line and started a new arc or vice verse; Fig [\ref {fig:strokeinterpolation}].
+The main problem in this process is detecting the critical points and vertices which divide strokes into lines and arcs.
+The user draws the lines and arcs in an undefined manner which make it difficult for the program to decide where the user had finished a line and started a new arc or vice verse;  Figure \ref {fig:strokeinterpolation}.
 \begin{figure}
@@
 \begin{center}
-		\includegraphics[scale=0.8]{../../neededfiles/Figures/strokeinterpolation.eps}
-	\caption[Segmentation error]{The figure shows an arc that the user draw which can be segmented in different ways as a single line or as an arc where a) The stroke users draw  b)interpretation as a single line segment b) interpretation as an arc.}
+		\includegraphics[scale=0.6]{../../neededfiles/Figures/strokeinterpolation.eps}
+	\caption[Segmentation Error]{Different ways for segmenting a stroke drawn by user.
+a) The stroke users draw b)interpretation as two line segments c) interpretation as an arc.
+This example is adapted from \cite{earlyprocess}}
+	%The figure shows an arc that the user draw which can be segmented in different ways as a single line or as an arc where a) The stroke users draw  b)interpretation as a single line segment b) interpretation as an arc.}
 	\label{fig:strokeinterpolation}
@@
 \subsection{Ambiguity}
-The user draws strokes that the computer cannot decide if it is arc or line.
-The nature of hand drawn sketch itself is highly ambiguous even for human eye you may see two drawn you cannot decide if it is square or rectangle.
+Figure \ref{fig:VorU} shows an example of ambiguity, where the strokes drawn by the user can be classified in several ways.
+The nature of hand drawn sketch itself is highly ambiguous even for human eye.
+You may see two drawings you cannot decide if they are a square or a rectangle.
 
@@
 Symbol ambiguity is another problem as a symbol can have more than one interpolation.
-Due to the user sloppy drawing the symbols can have more than one interpolation which varies with the domain drawn in it similar problem was faced in handwriting recognition (Figure [\ref{fig:VorU}]).
-The problem could be solved by providing context information which could not be provided in most sketches.
-
+Due to the user's sloppy drawing the symbols can have more than one interpolation which varies with the domain drawn in it.
+A similar problem was faced in handwriting recognition (Figure \ref{fig:VorU}).
+The problem could be solved by providing context or domain information which could not be provided in most sketches.
 \begin{figure}	
 	\centering
-		\includegraphics[scale=0.85]{../../neededfiles/Figures/VorU.eps}
-	\caption[Ambiguities Handwriting] { The same letter can be interpreted differently in various contexts; first as V in drive or as U in Cut}
+		\includegraphics[scale=0.65]{../../neededfiles/Figures/VorU.eps}
+	\caption[Handwriting Ambiguities] { The same letter can be interpreted differently in various contexts; first as V in drive or as U in Cut. This example is adapted from \cite{Ambiguity}.}
 \label{fig:VorU}
@@
 Using well established AI methods specially \textit{Swarm Intelligence} to solve the problem.
- This research investigates the best method to recognize sketches.
+This research presents a new method for sketch recognition.
 The scope of the research includes recognizing various symbols drawn using single and multiple strokes.
 The research compares various segmentation methods.
-This research used two Particle Swarm Optimization (PSO) \cite{PSOFirst} methods to segments strokes into geometrical primitives.
-The research also investigates various spatial, geometrical and statistical features set used in sketch recognition.
-Support Vector Machines (SVM) classifier\cite{svmintroduce} is used to classify  the sketch  into one of symbols from the set of previously trained symbols.
-This research assumes the user draws one symbols at a time. The user can draw can draw this symbol in one stroke or in multiple stroke without any restriction on order, orientation or style.
+This research used two Particle Swarm Optimization (PSO) \cite{PSOFirst} methods to segment strokes into geometrical primitives.
+The research also investigates various spatial, geometrical and statistical feature sets used in sketch recognition.
+A Support Vector Machines (SVM) classifier \cite{svmintroduce} is used to classify the sketch into one of symbols from the set of previously trained symbols.
+This research targets the recognition of one symbol at a time.
+The symbol may be drawn in one or more stroke without any restriction on order, orientation or style.
+
+%This research assumes the user draws one symbols at a time. The user can draw can draw this symbol in one stroke or in multiple stroke without any restriction on order, orientation or style.
 
 \section{Thesis Overview}
-This thesis is arranged such that chapter 2 presents a survey of different components of the sketch recognition.
-Chapter 3 describes the proposed system and its main building blocks.
-The details description of each component is presented in chapter 4.
-The comparative study of the different techniques and their detailed results in various experiments are detailed in chapter 5.
-Finally, chapter 6 presents the conclusions of this thesis and suggested future work.
-%This thesis is arranged such that chapter \ref{} presents a survey of different components of the sketch recognition.
-Chapter \ref{} describe the proposed system and its main building blocks.
-The details description of each component is presented in Chapter \ref{}.
-The comparative study of different techniques with the detailed results of various experiments are detailed in chapter\ref{}.
-Finally, chapter \ref{} presents the conclusions of this thesis and suggested future work.
+This thesis is arranged as follows: chapter \ref{sec:survey} presents a survey of different components of the sketch recognition.
+Chapter \ref{sec:ParticleSwramOptimization} gives a review on swarm intelligence and particle swarm optimization.
+Chapter \ref{sec:proposedSystem} describes the proposed system and its main building blocks.
+The comparative study of the different techniques and their detailed results in various experiments are given in chapter \ref{sec:Evaluation}.
+Finally, chapter \ref{sec:DiscussionConclusion} presents the conclusions of this thesis and suggested future work.

### 2009-04-27 trunk/draft/SketchMaster.tex
return to old but adding supervisor to first page

@@
 %\documentstyle[11pt,unhthesis]{report}
-\documentclass[11pt,doublespace]{report}%{Sketchthesis}
+\documentclass[11pt,doublespace]{Sketchthesis}
 
@@
                                                         %======================
-          % My theses title  using Particle Swarm Optimization Algorithm
-%\author{Maha Mohammed Nabeel El Meseery}                               % Y name
-%\prevdegrees{B.S.,Computer  Engineering}       		% Your old degree
-%\major{Computer Engineering}                            % My current major
-%\degree{Master of Science}                              % My new degree
-%\degreemonth{}                                      % When awarded. to be added when known
-%\degreeyear{}                                      %
-%%th today or first day of writing  December 26 2005 
-%%\thesisdate[]                                     % Date of document.
-%\DOCUMENTtype{Thesis}                                   % or DISSERTATION
-%\Documenttype{Thesis}                                   % or Dissertation
-%\documenttype{Thesis}
+\title{Sketch Recognition and Understanding}             % My theses title  using Particle Swarm Optimization Algorithm
+\author{Maha Mohammed Nabeel El Meseery}                               % Y name
+\prevdegrees{B.S.,Computer  Engineering}       		% Your old degree
+\major{Computer Engineering}                            % My current major
+\degree{Master of Science}                              % My new degree
+\degreemonth{}                                      % When awarded. to be added when known
+\degreeyear{}                                      %
+\superad{Professor Nevin Darwish\\Professor Samia Mashali\\Assoc. Professor Magda Fayek\\}
+%th today or first day of writing  December 26 2005 
+%\thesisdate[]                                     % Date of document.
+\DOCUMENTtype{Thesis}                                   % or DISSERTATION
+\Documenttype{Thesis}                                   % or Dissertation
+\documenttype{Thesis}                                   % or dissertation
 
-   \vspace*{0.375in}
-               {\Large\bf\title{Sketch Recognition and Understanding}  }
-                                 \vspace*{0.375in}
-
-                                \vspace*{\fill}\par
-                                      BY
-                                \vspace*{\fill}\par
-                               {\Large \author{Maha Mohammed Nabeel El Meseery }}
-                                     \par
-                               B.Sc.,Computer  Engineering
-                                \vspace*{\fill}\par
-                                    Thesis
-                                \vspace*{\fill}\par
-                           Submitted to the Faculty of Engineering Cairo University\\
-                         in partial fulfillment of \\
-                   the requirements for the degree\ of
-                                \vspace*{\fill}\par
-                                   Master of Science
-                                     \par
-                                      in
-                                     \par
-                                   Master of Science
-                                \vspace*{\fill}\par
-
-                                                        % or dissertation
 \maketitle
 
-\providecommand{\tabularnewline}{\\}
                                                         % COPYRIGHT PAGE
@@
                                                         %======================
- \def\acknowledgments
-   {\vspace*{0.5in}
-    \thispagestyle{plain}
-    \chapter*{Acknowledgments}
-    \addcontentsline{toc}{section}{Acknowledgments}
-
-   }
-
-\def\endacknowledgments{\par}
-
 \begin{acknowledgments}     
@@
 %------------------------------------------------------------------------------
-\begin{abstract}          
+\begin{abstractpage}          
 % ABSTRACT PAGE
@@
 
-\end{abstract}                              % This ends the page.
+\end{abstractpage}                              % This ends the page.
 

### 2009-04-27 trunk/draft/chapterSurvey.tex
mid way edit.

@@
 %mention that is best than ga (see thesis) 
-
+\subsection*{Swarm Intelligence History}
 Modeling how a biological swarm has been recently studied by biologist and computer scientists.
@@
 
+\subsection*{Swarm Intelligence Applications}
+\label{sec:SwarmApp}
 Image processing is one of the disciplines that used Swarm Intelligence.
-Ant colony was widely used to segment objects in \cite{antSegmentation,antthreshold,SwarmMedical}.
-Since 1995 when Kennedy and Eberhart \cite{PSOFirst,PSO2} introduced particle swarm algorithm (PSO) the algorithm had been used in various image processing system as an optimization  algorithm \cite{SwarmImageProcessing,Swarmimageprocessing2,Colorquantization,SwarmMedical}.
+Simulation to ant colony was widely used to segment objects in \cite{antSegmentation,antthreshold,SwarmMedical}.
+Since 1995, when Kennedy and Eberhart \cite{PSOFirst,PSO2} introduced Particle Swarm Optimization Algorithm (PSO) the algorithm had been used in various image processing system as an optimization  algorithm \cite{SwarmImageProcessing,Swarmimageprocessing2,Colorquantization,SwarmMedical}.
 PSO algorithm was used in \cite{PSOIMageClassification} to classify pixel to simplify the object segmentation process.
+The next section discribes in details the Particle Swarm Optimization and its advantages and disatvantages.
 
@@
 \label{sec:ParticleSwarmAlgorithm}
+
+ \subsection*{Optimization Algorithms}
+ In optimization problems, the algorithms attempts to find a feasable solution which satisfy all constrain of this problem.
+An optimal solution is the best feasable solution in the solution space.
+PSO is a population based stochastic search algorithm\cite{PSOFirst,PSO2}.
+It models the  % which has the best or  select a set of paramters that satisfy all constrains
+ The Particle Swarm Algorithm is an adaptive algorithm
+based on a social-psychological metaphor [1]. A
+population of individuals adapt by returning stochastically
+toward previously successful regions in the search
+space, and are influenced by the successes of their topological
+neighbors. Most particle swarms are based on two
+sociometric principles. Particles fly through the solution
+space and are influenced by both the best particle in the
+particle population and the best solution that a current
+particle has discovered so far. The best particle in the
+population is typically denoted by (global best), while the
+best position that has been visited by the current particle
+is denoted by (local best). The (global best) individual
+conceptually connects all members of the population to
+one another. That is, each particle is influenced by the
+very best performance of any member in the entire population.
+The (local best) individual is conceptually seen
+as the ability for particles to remember past personal
+successes.
+
+ Kennedy and Eberhart \cite{PSOFirst,PSO2} introduced particle swarm algorithm (PSO) as a general optimization algorithm.
+
+\subsection*{Discrite Particle Swarm Equations(DPSO)}
 %\section{Particle Swarm Algorithm}
@@
 %What is particle swarm algorithm and how it was used in related researches. 
-The main idea of \textit{Particle Swarm Algorithm (PSO)} is to represent each agent with a particle from the solution space \cite{PSOFirst}.
-Each agent moves the particle with a direction and velocity $v_{ij}$ based on equations \ref{eq:Swarm} \& \ref{eq:Swarm1}.
+The main idea of \textit{Particle Swarm Algorithm (PSO)} is to represent each solution with a particle from the solution space \cite{PSOFirst}.
+Each particle moves the parameter with a direction and velocity $v_{ij}$ based on equations \ref{eq:Swarm} \& \ref{eq:Swarm1}.
 \begin{equation}
@@
 \end{equation}
-where $p_{ij}$ represent the $jth$ particle in the $ith$ agent and $v_{ij}$ is the velocity of the $jth$ particle in the $ith$ agent.
+where $p_{ij}$ represent the $jth$ parameter in the $ith$ particle and $v_{ij}$ is the velocity of the $jth$ particle in the $ith$ particle.
  %Equation [\ref{eq:Swarm}] shows how velocity and direction of each particle are computed
  \begin{equation}
-v_{ij}  = v_{ij}  + c_1 r_1 (lbest_{ij}  - p_{ij} ) + c_2 r_2 (gbest_{ij}  - p_{ij} )
+v_{ij}  = v_{ij} W  + c_1 r_1 (lbest_{ij}  - p_{ij} ) + c_2 r_2 (gbest_{ij}  - p_{ij} )
 \label{eq:Swarm}
 \end{equation}
- where $lbest_{ij}$ is the local best particle, $gbest_{ij}$ is the global best particle, $r_1$ \& $r_2$ are random variables and $c_1$ \& $c_2$ are the swarm system variables.
+ where $W$ is the inertia weight paramter which controls the trade off between exploration and exploitation, $lbest_{ij}$ is the local best particle, $gbest_{ij}$ is the global best particle, $r_1$ \& $r_2$ are random variables and $c_1$ \& $c_2$ are the swarm acceleration paramters.
  After each iteration the global best $g_{best}$ particle and the agent local best $l_{best}$ particle are evaluated based on the maximum fitness functions of all particles in the solution space.

### 2009-04-27 trunk/draft/chapterSurvey.tex

@@
 
-\subsection*{Swarm Intelligence Applications}
-\label{sec:SwarmApp}
-Image processing is one of the disciplines that used Swarm Intelligence.
-Simulation to ant colony was widely used to segment objects in \cite{antSegmentation,antthreshold,SwarmMedical}.
-Since 1995, when Kennedy and Eberhart \cite{PSOFirst,PSO2} introduced Particle Swarm Optimization Algorithm (PSO) the algorithm had been used in various image processing system as an optimization  algorithm \cite{SwarmImageProcessing,Swarmimageprocessing2,Colorquantization,SwarmMedical}.
-PSO algorithm was used in \cite{PSOIMageClassification} to classify pixel to simplify the object segmentation process.
-The next section discribes in details the Particle Swarm Optimization and its advantages and disatvantages.
+The underliny research in swarm intelligence led to optimization algorithm that simulate the animal social behvaiors. 
+Ant Colony Optimization (ACO) algoirithm simulate how ants search for optimize path for food \cite{}.
+Real ants leave evidence of their exploriation while they are searching for food.
+This trail is used by other ants to locate shortest distance between the colony and food location.
+ACO use similar type of cooperation between agents to find optimal solution.
+The synchronous bird flock movement was simulated in \cite{PSOFirst,PSO2}.
+Kennedy and Eberhart \cite{PSOFirst,PSO2} simulated the rules of bird flock movement and introduced the Particle Swarm Optimization Algorithm (PSO).
+
+% Kennedy and Eberhart \cite{PSOFirst,PSO2} introduced Particle Swarm Optimization Algorithm (PSO). The algorithm simulate the moving
 
@@
 PSO is a population based stochastic search algorithm\cite{PSOFirst,PSO2}.
-It models the  % which has the best or  select a set of paramters that satisfy all constrains
+It simulate the  % which has the best or  select a set of paramters that satisfy all constrains
+
+ Studies to simulate the underly
+ing rules of bird social behavior led to a new approach for optimizing realvalued
+functions 	Particle Swarm Optimization PSO Kennedy and Eberhart 
+
+ This new approach combines notions of evolutionary computation and
+swarm theory Kennedy 
+
+PSO provides a populationbased search procedure where each individual is
+abstracted as a 	particle that 
+ies around in a multidimensional search space
+The best positions encountered by a particle and its neighbor determine the parti
+cles trajectory along with other PSO parameters In other words a PSO system
+attempts to balance exploration and exploitation by combining local and global
+search methods
+
  The Particle Swarm Algorithm is an adaptive algorithm
@@
  where $p_{ij}$ is the numerical values of the particle and $r_{3}$ is a random variable.
+
+\subsection*{Swarm Intelligence Applications}
+\label{sec:SwarmApp}
+Image processing is one of the disciplines that used Swarm Intelligence.
+Simulation to ant colony was widely used to segment objects in \cite{antSegmentation,antthreshold,SwarmMedical}.
+Since 1995, when Kennedy and Eberhart \cite{PSOFirst,PSO2} introduced Particle Swarm Optimization Algorithm (PSO) the algorithm had been used in various image processing system as an optimization  algorithm \cite{SwarmImageProcessing,Swarmimageprocessing2,Colorquantization,SwarmMedical}.
+PSO algorithm was used in \cite{PSOIMageClassification} to classify pixel to simplify the object segmentation process.
+The next section discribes in details the Particle Swarm Optimization and its advantages and disatvantages.
+
+\subsection*{Advantages}
+Particle swarm Optimization is gar

### 2009-04-28 trunk/draft/chapterSurvey.tex
its is now 2 in the morining and i am tiere

--- 2009-04-27 trunk/draft/chapterSurvey.tex
@@
 Recognition is done using SVM Linear classifier as it proved to be one of the best learning algorithms.
-
+%*********************************************************************************************5&&&&&&&&&&&&&&&&&&&&
+%*********************************************************************************************5&&&&&&&&&&&&&&&&&&&&
+%*********************************************************************************************5&&&&&&&&&&&&&&&&&&&&
+%*********************************************************************************************5&&&&&&&&&&&&&&&&&&&&
+%*********************************************************************************************5&&&&&&&&&&&&&&&&&&&&
 \chapter{Particle Swarm Optimization}
@@
 %mention that is best than ga (see thesis) 
-
+%\subsection*{Swarm Intelligence History}
 Modeling how a biological swarm has been recently studied by biologist and computer scientists.
@@
 
-Image processing is one of the disciplines that used Swarm Intelligence.
-Ant colony was widely used to segment objects in \cite{antSegmentation,antthreshold,SwarmMedical}.
-Since 1995 when Kennedy and Eberhart \cite{PSOFirst,PSO2} introduced particle swarm algorithm (PSO) the algorithm had been used in various image processing system as an optimization  algorithm \cite{SwarmImageProcessing,Swarmimageprocessing2,Colorquantization,SwarmMedical}.
+The underliny research in swarm intelligence led to optimization algorithm that simulate the animal social behvaiors. 
+Ant Colony Optimization (ACO) algoirithm simulate how ants search for optimize path for food \cite{ACOFirst}.
+Real ants leave evidence of their exploriation while they are searching for food.
+This trail is used by other ants to locate shortest distance between the colony and food location.
+ACO use similar type of cooperation between agents to find optimal solution.
+The synchronous bird flock movement was simulated in \cite{PSOFirst,PSO2}.
+Kennedy and Eberhart \cite{PSOFirst,PSO2} simulated the rules of bird flock movement and introduced the Particle Swarm Optimization Algorithm (PSO).
+
+   The next section discribes in details the Particle Swarm Optimization and its advantages and disatvantages. 
+% Kennedy and Eberhart \cite{PSOFirst,PSO2} introduced Particle Swarm Optimization Algorithm (PSO).
+The algorithm simulate the moving  Kennedy and Eberhart \cite{PSOFirst,PSO2} introduced particle swarm algorithm (PSO) as a general optimization algorithm.
+ \subsection*{Particle Swarm Algorithm}
+ In optimization problems, the algorithm attempts to find a feasable solution which satisfy all constrain of this problem.
+An optimal solution is the best feasable solution in the solution space.
+Several methods were used to solve optimiation problems.
+Stochastic search is based on the assumption that good solution are near each other.
+Algoirthms like simluating aneeling and hill climbing used stochastic to find best solution.
+The problem in these methods is they may be traped into a local best and fail to find the global best solution.
+
+  Evolutionary algorithms (e.g.
+Genatic Algoirthms) explores solution using a population which is evalutated using fitness functions.
+The algorithms initially start with a randomly choosen population which is used to generate new population using various operators (e.g.
+cross over, mutation...
+).
+After each iteration the population is evaluated using a fitness function and only the strongest are used to generate new population.
+The use of population helps to avoid stochastic search problem and explores the solution space to find better global optimal solutions, but unfortunately they may not alway converge to a global best solution\cite{PSOpattern}.
+
+ Particle Swarm Algorithm\cite{PSOFirst}is a population based stochastic search where each solution is simulated as a  particle that flies around in a multidimensional search space.
+A fitness function is used to evaluate the goodness of each particle in solution space.
+But unlike evolutionly algorithm the new population is the PSO is influenced by the social cooperation of the particle rather than survival to the fitest.
+Each particle saves a history of the best location in the swarm and its own best location.
+After each iteration, the particles moves to a new location based on the global best location and it own history of best location.
+This use of pervious particle locations gives PSO advantage over Evolutionary algoirthms where no similar information is used.
+
+ PSO can be viewed as an algorithm between stochastic search and Evolutionary algorithms\cite{PSOpattern}.
+Comparison between PSO and Genetic algorithm was perofrmed in \cite{}.
+Expriments proved that PSO out performe GA in most cases specially in problem with high dimensinalty.
+ \subsection*{PSO neighbourhood topology}  
+There is various toplogyies used in connecting a swarm in PSO. The most used is the totaly connected, star and ring topology.
+
+\subsection*{Particle Swarm Applications}
+\label{sec:SwarmApp}
+ Since 1995, when Kennedy and Eberhart \cite{PSOFirst,PSO2} introduced Particle Swarm Optimization Algorithm (PSO) the algorithm had been used in various systems.
+The algorithm was used to train Neural network in \cite{} and proved to perform better than GA algoirthm.
+In Image processing decipline PSO was used an optimization algorithm \cite{SwarmImageProcessing,Swarmimageprocessing2,Colorquantization,SwarmMedical}.
 PSO algorithm was used in \cite{PSOIMageClassification} to classify pixel to simplify the object segmentation process.
+
+\subsection*{Drawbacks}
+The main drawback of the particle swarm algorithm is the premanture coverge problem.
+This happens when global best cannot be improved after a few swarm iterations.
+This is because the use of global best makes pso converge to one place which is between local best and global best.
+Another resons is because of the high amount of information between particle makes the solution less diversy which limit the explored solution space.
+This can be solved by using high interita $w$ or
+\subsection*{Improvements}
 
@@
 \label{sec:ParticleSwarmAlgorithm}
+
+  %It simulate the  % which has the best or  select a set of paramters that satisfy all constrains
+
+\subsection*{Discrite Particle Swarm Equations(DPSO)}
 %\section{Particle Swarm Algorithm}
@@
 %What is particle swarm algorithm and how it was used in related researches. 
-The main idea of \textit{Particle Swarm Algorithm (PSO)} is to represent each agent with a particle from the solution space \cite{PSOFirst}.
-Each agent moves the particle with a direction and velocity $v_{ij}$ based on equations \ref{eq:Swarm} \& \ref{eq:Swarm1}.
+The main idea of \textit{Particle Swarm Algorithm (PSO)} is to represent each solution with a particle from the solution space \cite{PSOFirst}.
+Each particle moves with a direction and velocity $v_{ij}$ based on equations \ref{eq:Swarm} \& \ref{eq:Swarm1}.
 \begin{equation}
@@
 \end{equation}
-where $p_{ij}$ represent the $jth$ particle in the $ith$ agent and $v_{ij}$ is the velocity of the $jth$ particle in the $ith$ agent.
+where $p_{ij}$ represent the $jth$ dimention in the $ith$ particle and $v_{ij}$ is the velocity of the $jth$ dimention in the $ith$ particle.
  %Equation [\ref{eq:Swarm}] shows how velocity and direction of each particle are computed
  \begin{equation}
-v_{ij}  = v_{ij}  + c_1 r_1 (lbest_{ij}  - p_{ij} ) + c_2 r_2 (gbest_{ij}  - p_{ij} )
+v_{ij}  = v_{ij} W  + c_1 r_1 (lbest_{ij}  - p_{ij} ) + c_2 r_2 (gbest_{ij}  - p_{ij} )
 \label{eq:Swarm}
 \end{equation}
- where $lbest_{ij}$ is the local best particle, $gbest_{ij}$ is the global best particle, $r_1$ \& $r_2$ are random variables and $c_1$ \& $c_2$ are the swarm system variables.
+ where $W$ is the inertia weight paramter which controls the trade off between exploration and exploitation, $lbest_{ij}$ is the local best particle, $gbest_{ij}$ is the global best particle, $r_1$ \& $r_2$ are random variables and $c_1$ \& $c_2$ are the swarm acceleration paramters.
  After each iteration the global best $g_{best}$ particle and the agent local best $l_{best}$ particle are evaluated based on the maximum fitness functions of all particles in the solution space.
+The velocity is bounded to maximum of $V_{max}$ using Equation\ref{eq:vmax} to prevent exploding which leads to premature convergence.
 The solution is found after achieving a specific number of iteration or after an error threshold is achieved.
@@
 \end{array}\}
-\end{equation}
- where $p_{ij}$ is the numerical values of the particle and $r_{3}$ is a random variable.
+\end{equation} where $p_{ij}$ is the numerical values of the particle and $r_{3}$ is a random variable. 
+  \begin{equation}
+V_{ij}  = \frac{{V_{ij} }}{{2V_{\max } }} + 0.5
+\label{eq:vmax}
+\end{equation} where $V_{\max }$ is the velocity threshold.
+
+\section*{The  Algorithm}
+%\label{sec:ParticleSwarmAlgorithm} 
+\begin{verbatim}
+ Initialize 
+    Generate N particles with M dimentions (P_i...P_N) 
+    Generate velocities for each particle for M dimention randomly. 
+ Repeat till Maximum number of iteration OR Fitness 
+         For each particle P_i do the following 
+                Compute fitness  
+                Determine Lbest  
+                Check if need to change Gbest. 
+                Compute the Velocity for each dimention using eq. 
+                Check if maximum Velocity reach and limit it 
+                update particle values 
+          end . 
+        Improve all particle positions 
+  end
+
+  Final solution reach in Gbest.
+
+\end{verbatim}

### 2009-04-28 draft/chapterSurvey.tex
before edit in swarm
before fager

@@
 \label{sec:ParticleSwramOptimizationReview}
-%This seems to emerge of what is called 'collective intelligence' from often large simple agents.  
-% start with swarm in general  .. say about flock , ant and usin in image processing 
-% then speeak about PSO and  its firstly generated   (see article)
-%mention that is best than ga (see thesis)
 
@@
 
-Image processing is one of the disciplines that used Swarm Intelligence.
-Ant colony was widely used to segment objects in \cite{antSegmentation,antthreshold,SwarmMedical}.
-Since 1995 when Kennedy and Eberhart \cite{PSOFirst,PSO2} introduced particle swarm algorithm (PSO) the algorithm had been used in various image processing system as an optimization  algorithm \cite{SwarmImageProcessing,Swarmimageprocessing2,Colorquantization,SwarmMedical}.
+The underliny research in swarm intelligence led to optimization algorithm that simulate the animal social behvaiors. 
+Ant Colony Optimization (ACO) algoirithm simulate how ants search for optimize path for food \cite{ACOFirst}.
+Real ants leave evidence of their exploriation while they are searching for food.
+This trail is used by other ants to locate shortest distance between the colony and food location.
+ACO use similar type of cooperation between agents to find optimal solution.
+The synchronous bird flock movement was simulated in \cite{PSOFirst,PSO2}.
+Kennedy and Eberhart \cite{PSOFirst,PSO2} simulated the rules of bird flock movement and introduced the Particle Swarm Optimization Algorithm (PSO).
+
+   The next section discribes in details the Particle Swarm Optimization and its advantages and disatvantages. 
+% Kennedy and Eberhart \cite{PSOFirst,PSO2} introduced Particle Swarm Optimization Algorithm (PSO).
+The algorithm simulate the moving  Kennedy and Eberhart \cite{PSOFirst,PSO2} introduced particle swarm algorithm (PSO) as a general optimization algorithm.
+ \subsection*{Particle Swarm Algorithm}
+ In optimization problems, the algorithm attempts to find a feasable solution which satisfy all constrain of this problem.
+An optimal solution is the best feasable solution in the solution space.
+Several methods were used to solve optimiation problems.
+Stochastic search is based on the assumption that good solution are near each other.
+Algoirthms like simluating aneeling and hill climbing used stochastic to find best solution.
+The problem in these methods is they may be traped into a local best and fail to find the global best solution.
+
+  Evolutionary algorithms (e.g.
+Genatic Algoirthms) explores solution using a population which is evalutated using fitness functions.
+The algorithms initially start with a randomly choosen population which is used to generate new population using various operators (e.g.
+cross over, mutation...
+).
+After each iteration the population is evaluated using a fitness function and only the strongest are used to generate new population.
+The use of population helps to avoid stochastic search problem and explores the solution space to find better global optimal solutions, but unfortunately they may not alway converge to a global best solution\cite{PSOpattern}.
+
+ Particle Swarm Algorithm\cite{PSOFirst}is a population based stochastic search where each solution is simulated as a  particle that flies around in a multidimensional search space.
+A fitness function is used to evaluate the goodness of each particle in solution space.
+But unlike evolutionly algorithm the new population is the PSO is influenced by the social cooperation of the particle rather than survival to the fitest.
+Each particle saves a history of the best location in the swarm and its own best location.
+After each iteration, the particles moves to a new location based on the global best location and it own history of best location.
+This use of pervious particle locations gives PSO advantage over Evolutionary algoirthms where no similar information is used.
+
+ PSO can be viewed as an algorithm between stochastic search and Evolutionary algorithms\cite{PSOpattern}.
+In general the PSO is a continous optimization problem but expriments showed it performed as GA is discrite and binary optimizaition problems.
+The comparison between PSO and Genetic algorithm was perofrmed in \cite{PSOGAComp} and  proved that PSO generate faster, better solutions than GA specially in problems with high dimensinalty.
+
+ A hybrid of PSO and other algorithms was investigated in various systems.
+\cite{} used a hybird of GA and PSO to breed solutions using GA then feed them to PSO.
+They domenstrated that using PSO-GA and PSO outperform using GA as an optimization algorithm.
+
+ \subsection*{PSO neighbourhood topology}  
+In general the PSO algorithm have fully connected particle which means that every particle know the global best and local best of each other particle.
+Other forms of cooperation and organizatoin were suggested in various systems.
+\cite{} suggested to use a grid like topology where particle are connected to four other particles sround it.
+A ring  and star topology were widley used in \cite{}.
+In \cite{hiericalPSO}, Janson and Middendorf proposed to use a dynamic and hierrchinal neighbourhood structure.
+They showed that using such structure helps in mentainaing diversity over the solution space in a dynamic problem.
+% a PSO using a dynamic and hierarchical neighbourhood structure to handle dynamic optimization problems.
+They demonstrated that such a structure is useful for maintaining some particle diversity in a dynamic environment.
+
+ Other forms of cooperations were also investigated in \cite{facotrsSwarm} where the problem is divided among particles and each particle solves part of the problem.
+\cite{facotrsSwarm} discuss different behvaiorus that governs cooperation particles in the swarm system.
+%were used  There is various toplogyies used in connecting a swarm in PSO.
+The most used is the totaly connected, star and ring topology.
+
+\subsection*{Drawbacks}
+The main drawback of the particle swarm algorithm is the premanture coverge problem.
+This happens when global best cannot be improved after a few swarm iterations.
+This is because the use of global best makes pso converge to one place which is between local best and global best.
+Another resons is because of the high amount of information between particle makes the solution less diversy which limit the explored solution space.
+This can be solved by using high interita $w$ or using another improvement that generate new random populations.
+
+Another problem of PSO is that performance is dependent on the problem and the PSO paramter selected for this specific problem.
+This means that any change in the PSO paramter change the efficency of PSO solution.
+
+\subsection*{Improvements to PSO}
+
+change so that the optima can still be tracked.
+One of the early works on using PSO for dynamic optimization
+was by Eberhart and Shi in [13], where they used
+an inertia weight version of PSO to track the optimum of
+a 3-dimensional unimodal parabolic function which changes
+its maxima every 100 iterations. It was found under certain
+circumstances that the PSO’s performance was comparable
+or better than that of previously published evolutionary algorithms
+[1, 2].
+For detection, Carlisle and Dozier used a sentry particle
+which is randomly chosen at each iteration [10]. The sentry
+particle gets evaluated before each iteration and compares
+its fitness with its previous fitness value. If the two values
+are different, indicating the environment has changed, ie.,
+the optimum has moved, then the whole population gets
+alerted and several possible responses can then be triggered.
+Hu and Eberhart proposed to re-evaluate  pg and a second
+best particle to detect if a change has occurred.
+Various response strategies have also been proposed. Carlisle
+and Dozier proposed to re-evaluate all pi of all particles when
+a change has been detected [11]. If the re-evaluated pi is less
+fit than its current position xi, then this pi is replaced by
+its associated xi, because only these two evaluations can
+be considered up-to-date. Hu and Eberhart studied the effects
+of re-randomizing various proportions of the swarm to
+maintain some degree of diversity in order to better track
+the optima after a change [14]. However, this approach suffers
+from possible information loss since the re-randomized
+portion of the population does not retain any information
+that might be useful from the past iterations.
+In order to maintain better particle diversity throughout
+a run, Blackwell and Bentley introduced charged swarms
+where mutually repelling charged particles orbit a nucleus of
+neutral particles (conventional PSO particles) [3]. Whereas
+the charged particles allow the swarm to better adapt to
+changes in the environment, the neutral particles play the
+role of continuing to converge towards the optimum in finer
+details.
+Inspired by multi-population EA approaches such as the
+multinational GA [22] and self-organizing scouts [7], Blackwell
+and Branke proposed an interacting multi-swarm PSO
+as a further improvement to the charged swarms [4]. The
+multi-swarm PSO aims at maintaining multiple swarm populations
+on different peaks. Multiple swarms are prevented
+from converging to the same optimum by randomizing the
+worse of two swarms that come too close. The multi-swarm
+PSO also replaces the charged particles by quantum particles
+whose position is solely based on a probability function
+centered around the swarm attractor. The resulting multiquantum
+swarms outperform charged and standard PSO on
+the moving peaks function. This multi-swarm approach is
+particularly attractive because of its improved adaptability
+in a more complex multimodal dynamic environment where
+multiple peaks exist and need to be tracked.
+With a similar aim to locate and track multiple peaks
+in a dynamic environment, Parrott and Li in [19, 20] proposed
+a species-based PSO (SPSO) incorporating a speciation
+algorithm first proposed by Petrowski [21]. The SPSO
+uses a local “species seed” which provides the local  pg to
+particles whose positions in the search space are within a
+user-specified radius of the seed. This encourages swarms
+to converge onto multiple local optima instead of a single
+global optimum, hence developing multiple sub-populations
+in parallel. In addition, the dynamic SPSO uses a parameter
+pmax to limit the number of particles allowed in a species
+(or swarm), with the excess particles reinitialized at random
+positions in the total search space.
+
+\subsection*{Particle Swarm Applications}
+\label{sec:SwarmApp}
+ Since 1995, when Kennedy and Eberhart \cite{PSOFirst,PSO2} introduced Particle Swarm Optimization Algorithm (PSO) the algorithm had been used in various systems.
+The algorithm was used to train Neural network in \cite{} and proved to perform better than GA algoirthm.
+In Image processing decipline PSO was used an optimization algorithm \cite{SwarmImageProcessing,Swarmimageprocessing2,Colorquantization,SwarmMedical}.
 PSO algorithm was used in \cite{PSOIMageClassification} to classify pixel to simplify the object segmentation process.
@@
 \end{array}\}
-\end{equation}
- where $p_{ij}$ is the numerical values of the particle and $r_{3}$ is a random variable.
+\end{equation} where $p_{ij}$ is the numerical values of the particle and $r_{3}$ is a random variable. 
+  \begin{equation}
+V_{ij}  = \frac{{V_{ij} }}{{2V_{\max } }} + 0.5
+\label{eq:vmax}
+\end{equation} where $V_{\max }$ is the velocity threshold.
+
+\subsection*{The  Algorithm}
+%\label{sec:ParticleSwarmAlgorithm} 
+\begin{verbatim}
+ Initialize 
+    Generate N particles with M dimentions (P_i...P_N) 
+    Generate velocities for each particle for M dimention randomly. 
+ Repeat till Maximum number of iteration OR Fitness 
+         For each particle P_i do the following 
+                Compute fitness  
+                Determine Lbest  
+                Check if need to change Gbest. 
+                Compute the Velocity for each dimention using eq. 
+                Check if maximum Velocity reach and limit it 
+                update particle values 
+          end . 
+        Improve all particle positions 
+  end
+
+  Final solution reach in Gbest.
+
+\end{verbatim}

### 2009-04-28 draft/chapterSurvey.tex
before grammer before i pary

@@
 They domenstrated that using PSO-GA and PSO outperform using GA as an optimization algorithm.
-
   \subsection*{PSO neighbourhood topology}  
@@
 
-change so that the optima can still be tracked.
-One of the early works on using PSO for dynamic optimization
-was by Eberhart and Shi in [13], where they used
-an inertia weight version of PSO to track the optimum of
-a 3-dimensional unimodal parabolic function which changes
-its maxima every 100 iterations. It was found under certain
-circumstances that the PSO’s performance was comparable
-or better than that of previously published evolutionary algorithms
-[1, 2].
-For detection, Carlisle and Dozier used a sentry particle
-which is randomly chosen at each iteration [10]. The sentry
-particle gets evaluated before each iteration and compares
-its fitness with its previous fitness value. If the two values
-are different, indicating the environment has changed, ie.,
-the optimum has moved, then the whole population gets
-alerted and several possible responses can then be triggered.
-Hu and Eberhart proposed to re-evaluate  pg and a second
-best particle to detect if a change has occurred.
-Various response strategies have also been proposed. Carlisle
-and Dozier proposed to re-evaluate all pi of all particles when
-a change has been detected [11]. If the re-evaluated pi is less
-fit than its current position xi, then this pi is replaced by
-its associated xi, because only these two evaluations can
-be considered up-to-date. Hu and Eberhart studied the effects
-of re-randomizing various proportions of the swarm to
-maintain some degree of diversity in order to better track
-the optima after a change [14]. However, this approach suffers
-from possible information loss since the re-randomized
-portion of the population does not retain any information
-that might be useful from the past iterations.
-In order to maintain better particle diversity throughout
-a run, Blackwell and Bentley introduced charged swarms
-where mutually repelling charged particles orbit a nucleus of
-neutral particles (conventional PSO particles) [3]. Whereas
-the charged particles allow the swarm to better adapt to
-changes in the environment, the neutral particles play the
-role of continuing to converge towards the optimum in finer
-details.
-Inspired by multi-population EA approaches such as the
-multinational GA [22] and self-organizing scouts [7], Blackwell
-and Branke proposed an interacting multi-swarm PSO
-as a further improvement to the charged swarms [4]. The
-multi-swarm PSO aims at maintaining multiple swarm populations
-on different peaks. Multiple swarms are prevented
-from converging to the same optimum by randomizing the
-worse of two swarms that come too close. The multi-swarm
-PSO also replaces the charged particles by quantum particles
-whose position is solely based on a probability function
-centered around the swarm attractor. The resulting multiquantum
-swarms outperform charged and standard PSO on
-the moving peaks function. This multi-swarm approach is
-particularly attractive because of its improved adaptability
-in a more complex multimodal dynamic environment where
-multiple peaks exist and need to be tracked.
-With a similar aim to locate and track multiple peaks
-in a dynamic environment, Parrott and Li in [19, 20] proposed
-a species-based PSO (SPSO) incorporating a speciation
-algorithm first proposed by Petrowski [21]. The SPSO
-uses a local “species seed” which provides the local  pg to
-particles whose positions in the search space are within a
-user-specified radius of the seed. This encourages swarms
-to converge onto multiple local optima instead of a single
-global optimum, hence developing multiple sub-populations
-in parallel. In addition, the dynamic SPSO uses a parameter
-pmax to limit the number of particles allowed in a species
-(or swarm), with the excess particles reinitialized at random
-positions in the total search space.
+There was various improvement used to improve the PSO algoirhtm.
+They moslty try to solve the problems premature coverge problem and increase the diversity of the particle population.
+Early work on dynamic optimization was introduced by Eberhard and shi \cite{paramterSelection}, where they added the inertia wegiht version of PSO to find the optimim of 3 dimentional function which change every 100 iteration.
 
@@
  Since 1995, when Kennedy and Eberhart \cite{PSOFirst,PSO2} introduced Particle Swarm Optimization Algorithm (PSO) the algorithm had been used in various systems.
-The algorithm was used to train Neural network in \cite{} and proved to perform better than GA algoirthm.
+The algorithm was used to train Neural network in \cite{PSOandNeuarl} and proved to perform better than GA algoirthm.
 In Image processing decipline PSO was used an optimization algorithm \cite{SwarmImageProcessing,Swarmimageprocessing2,Colorquantization,SwarmMedical}.

### 2009-04-28 trunk/draft/chapterSurvey.tex
finished

@@
 
-\section{Swarm Optimization Review}
+\section{Swarm Intelligence Review}
 \label{sec:ParticleSwramOptimizationReview}
-%This seems to emerge of what is called 'collective intelligence' from often large simple agents.  
-% start with swarm in general  .. say about flock , ant and usin in image processing 
-% then speeak about PSO and  its firstly generated   (see article)
-%mention that is best than ga (see thesis) 
-%\subsection*{Swarm Intelligence History}
+
  Modeling how a biological swarm has been recently studied by biologist and computer scientists.
@@
 Each agent has a set of simple rules to follow, and according to the interaction of the entire group the objective is met\cite{swarmGeneral}.
-%with other neighbourhood agents.
+%with other neighborhood agents.
 
@@
 
-The underliny research in swarm intelligence led to optimization algorithm that simulate the animal social behvaiors. 
-Ant Colony Optimization (ACO) algoirithm simulate how ants search for optimize path for food \cite{ACOFirst}.
-Real ants leave evidence of their exploriation while they are searching for food.
+The underling research in swarm intelligence led to optimization algorithm that simulates the animal social behaviors.
+Ant Colony Optimization (ACO) algorithm simulate how ants search for optimize path for food \cite{ACOFirst}.
+Real ants leave evidence of their exploration while they are searching for food.
 This trail is used by other ants to locate shortest distance between the colony and food location.
@@
 
-   The next section discribes in details the Particle Swarm Optimization and its advantages and disatvantages. 
+   The next section describes in details the Particle Swarm Optimization and its advantages and disadvantages. 
 % Kennedy and Eberhart \cite{PSOFirst,PSO2} introduced Particle Swarm Optimization Algorithm (PSO).
 The algorithm simulate the moving Kennedy and Eberhart \cite{PSOFirst,PSO2} introduced particle swarm algorithm (PSO) as a general optimization algorithm.
- \subsection*{Particle Swarm Algorithm}
- In optimization problems, the algorithm attempts to find a feasable solution which satisfy all constrain of this problem.
-An optimal solution is the best feasable solution in the solution space.
-Several methods were used to solve optimiation problems.
+ \subsection*{Particle Swarm Optimization}
+ In optimization problems, the algorithm attempts to find a feasible solution which satisfies all constrain of this problem.
+An optimal solution is the best feasible solution in the solution space.
+Several methods were used to solve optimization problems.
 Stochastic search is based on the assumption that good solution are near each other.
-Algoirthms like simluating aneeling and hill climbing used stochastic to find best solution.
-The problem in these methods is they may be traped into a local best and fail to find the global best solution.
+Algorithms like simulating annealing and hill climbing used stochastic to find best solution.
+The problem in these methods is they may be trapped into a local best and fail to find the global best solution.
 
   Evolutionary algorithms (e.g.
-Genatic Algoirthms) explores solution using a population which is evalutated using fitness functions.
-The algorithms initially start with a randomly choosen population which is used to generate new population using various operators (e.g.
+Genatic Algoirthms) explore solution using a population which is evaluated using fitness functions.
+The algorithms initially start with a randomly chosen population which is used to generate new population using various operators (e.g.
 cross over, mutation...
 ).
-After each iteration the population is evaluated using a fitness function and only the strongest are used to generate new population.
-The use of population helps to avoid stochastic search problem and explores the solution space to find better global optimal solutions, but unfortunately they may not alway converge to a global best solution\cite{PSOpattern}.
+After each iteration, the population is evaluated using a fitness function and only the strongest are used to generate new population.
+The use of population helps to avoid stochastic search problem and explores the solution space to find better global optimal solutions, but unfortunately they may not always converge to a global best solution\cite{PSOpattern}.
 
@@
 A fitness function is used to evaluate the goodness of each particle in solution space.
-But unlike evolutionly algorithm the new population is the PSO is influenced by the social cooperation of the particle rather than survival to the fitest.
+But unlike evolutionally algorithm the new population is the PSO is influenced by the social cooperation of the particle rather than survival to the fittest.
 Each particle saves a history of the best location in the swarm and its own best location.
-After each iteration, the particles moves to a new location based on the global best location and it own history of best location.
-This use of pervious particle locations gives PSO advantage over Evolutionary algoirthms where no similar information is used.
+At the end each iteration, the particles moves to a new location based on the global best location and it own history of best location.
+This use of pervious particle locations gives PSO advantage over Evolutionary algorithms where no similar information is used.
 
  PSO can be viewed as an algorithm between stochastic search and Evolutionary algorithms\cite{PSOpattern}.
-Comparison between PSO and Genetic algorithm was perofrmed in \cite{}.
-Expriments proved that PSO out performe GA in most cases specially in problem with high dimensinalty.
- \subsection*{PSO neighbourhood topology}  
-There is various toplogyies used in connecting a swarm in PSO. The most used is the totaly connected, star and ring topology.
+In general the PSO is a continuous optimization problem but experiments showed it performed as GA is discrete and binary optimization problems.
+The comparison between PSO and Genetic algorithm was performed in \cite{PSOGAComp} and proved that PSO generate faster, better solutions than GA especially in problems with high dimensionality.
+
+ A hybrid of PSO and other algorithms was investigated in various systems.
+Veeramachaneni \cite{kaylanGAPSO} used a hybrid of GA and PSO to breed solutions using GA then feed them to PSO.
+They demonstrated that using PSO-GA and PSO outperform using GA as an optimization algorithm.
+
+\subsection*{PSO Neighborhood Topology}  
+In general the PSO algorithm have fully connected particle which means that every particle know the global best and local best of each other particle.
+Other forms of cooperation and organization were suggested in various systems.
+\cite{grid} suggested using a grid like topology where particle are connected to four other particles around it.
+A ring and star topology were widely used in \cite{neighbourrhood}.
+The advantages over using different topology than the global best is that they slow down the swarm evolving (they need more time to propagate the best solution) which enable more space exploration and thus helps in eliminating the premature convergence problem.
+In \cite{hiericalPSO}, Janson and Middendorf proposed to use a dynamic and hierarchical neighborhood structure.
+They showed that using such structure helps in maintaining diversity over the solution space in a dynamic problem.
+% a PSO using a dynamic and hierarchical neighborhood structure to handle dynamic optimization problems.
+They demonstrated that such a structure is useful for maintaining some particle diversity in a dynamic environment.
+
+ Other forms of cooperation were also investigated in \cite{facotrsSwarm} where the problem is divided among particles and each particle solves part of the problem.
+\cite{facotrsSwarm} discuss a different behavior that governs cooperation particles in the swarm system.
+%were used There is various topologies used in connecting a swarm in PSO.
+The most used is the totally connected, star and ring topology.
+
+\subsection*{Drawbacks}
+
+The main drawback of the particle swarm algorithm is the premature convergence problem.
+This happens when global best cannot be improved after a few swarm iterations.
+This is because the use of global best makes PSO converge to one place which is between local best and global best.
+Another reason is because of the high amount of information between particle makes the solution less diverse which limit the explored solution space.
+This problem can be solved by using various methods; one of these methods is choosing another topology rather than the global best topology.
+Another solution is using adaptive inertia $w$ (Section \ref{sec:ParticleSwarmAlgorithm}).
+Other methods of solving problem were introduced as improvement to the PSO algorithm.
+Another problem of PSO is that performance is dependent on the problem and the PSO parameter selected for this specific problem.
+This means that any change in the PSO parameter change the efficiency of PSO solution.
+The solution to this problem is proposed by using self adaptive parameters.
+
+\subsection*{Improvements to PSO}
+
+There was various improvement used to improve the PSO algorithm.
+They mostly try to solve the problems premature converge problem and increase the diversity of the particle population.
+Early work on dynamic optimization was introduced by Eberhard and shi \cite{paramterSelection}, where they added the inertia weight version of PSO to find the optimum of 3 dimensional function which change every 100 iteration.
+Others used constrained PSO, where the general PSO Equation \ref{eq:Swarm} is constrained.
+Guaranteed Convergence PSO is introduced in \cite{GCPSO} where a system is guaranteed to converge to at least a local optimal solution.
+A repetition of this algorithm is used to insure finding of an optimal global solution.
+
+Other hybrid system that combines PSO with other techniques was introduced.
+Bayesian network and probability estimation were used to move particle to better solutions.
+Fuzzy rules were used to improve the general PSO algorithms in \cite{shapeFuzzy}.
 
@@
 \label{sec:SwarmApp}
- Since 1995, when Kennedy and Eberhart \cite{PSOFirst,PSO2} introduced Particle Swarm Optimization Algorithm (PSO) the algorithm had been used in various systems.
-The algorithm was used to train Neural network in \cite{} and proved to perform better than GA algoirthm.
-In Image processing decipline PSO was used an optimization algorithm \cite{SwarmImageProcessing,Swarmimageprocessing2,Colorquantization,SwarmMedical}.
+
+ Since 1995, when Kennedy and Eberhart \cite{PSOFirst,PSO2} introduced Particle Swarm Optimization Algorithm (PSO) the algorithm had been used in every discipline.
+The algorithm was used to train Neural network in \cite{PSOandNeuarl} and proved to perform better than GA algorithm.
+In Image processing discipline PSO was used an optimization algorithm in \cite{SwarmImageProcessing,Swarmimageprocessing2,Colorquantization,SwarmMedical}.
 PSO algorithm was used in \cite{PSOIMageClassification} to classify pixel to simplify the object segmentation process.
-
-\subsection*{Drawbacks}
-The main drawback of the particle swarm algorithm is the premanture coverge problem.
-This happens when global best cannot be improved after a few swarm iterations.
-This is because the use of global best makes pso converge to one place which is between local best and global best.
-Another resons is because of the high amount of information between particle makes the solution less diversy which limit the explored solution space.
-This can be solved by using high interita $w$ or
-\subsection*{Improvements}
+Other system used in shape representation like \cite{shapeFuzzy}.
+Yin \cite{PolygonApproximationPSO} used Discrete PSO to optimally convert digital curves into polygons.
 
@@
 
-  %It simulate the  % which has the best or  select a set of paramters that satisfy all constrains
+  %It simulate the  % which has the best or  select a set of parameters that satisfy all constrains
 
@@
 %What is particle swarm algorithm and how it was used in related researches. 
-The main idea of \textit{Particle Swarm Algorithm (PSO)} is to represent each solution with a particle from the solution space \cite{PSOFirst}.
+The main idea of \textit{Particle Swarm Algorithm (PSO)} is to represent each solution with a $N$ dimension particle from the solution space \cite{PSOFirst}.
 Each particle moves with a direction and velocity $v_{ij}$ based on equations \ref{eq:Swarm} \& \ref{eq:Swarm1}.
@@
 \end{equation}
-where $p_{ij}$ represent the $jth$ dimention in the $ith$ particle and $v_{ij}$ is the velocity of the $jth$ dimention in the $ith$ particle.
+where $p_{ij}$ represent the $jth$ dimension in the $ith$ particle and $v_{ij}$ is the velocity of the $jth$ dimension in the $ith$ particle.
  %Equation [\ref{eq:Swarm}] shows how velocity and direction of each particle are computed
@@
 \end{equation}
- where $W$ is the inertia weight paramter which controls the trade off between exploration and exploitation, $lbest_{ij}$ is the local best particle, $gbest_{ij}$ is the global best particle, $r_1$ \& $r_2$ are random variables and $c_1$ \& $c_2$ are the swarm acceleration paramters.
+ where $W$ is the inertia weight parameter which controls the tradeoff between exploration and exploitation, $lbest_{ij}$ is the local best particle, $gbest_{ij}$ is the global best particle, $r_1$ \& $r_2$ are random variables and $c_1$ \& $c_2$ are the swarm acceleration parameters.
  After each iteration the global best $g_{best}$ particle and the agent local best $l_{best}$ particle are evaluated based on the maximum fitness functions of all particles in the solution space.
+
  The velocity is bounded to maximum of $V_{max}$ using Equation\ref{eq:vmax} to prevent exploding which leads to premature convergence.
@@
 
-\section*{The  Algorithm}
-%\label{sec:ParticleSwarmAlgorithm} 
-\begin{verbatim}
- Initialize 
-    Generate N particles with M dimentions (P_i...P_N) 
-    Generate velocities for each particle for M dimention randomly. 
- Repeat till Maximum number of iteration OR Fitness 
-         For each particle P_i do the following 
-                Compute fitness  
-                Determine Lbest  
-                Check if need to change Gbest. 
-                Compute the Velocity for each dimention using eq. 
-                Check if maximum Velocity reach and limit it 
-                update particle values 
-          end . 
-        Improve all particle positions 
-  end
-
-  Final solution reach in Gbest.
-
-\end{verbatim}
+%\subsection{The Algorithm}
+%%\label{sec:ParticleSwarmAlgorithm} 
+%\begin{verbatim}
+% Initialize 
+%    Generate N particles with M dimensions (P_i...P_N) 
+%    Generate velocities for each particle for M dimension randomly. 
+% Repeat till Maximum number of iteration OR Fitness 
+%         For each particle P_i do the following 
+%                Compute fitness  
+%                Determine Lbest  
+%                Check if need to change Gbest. 
+%                Compute the Velocity for each dimension using eq. 
+%                Check if maximum Velocity reach and limit it 
+%                update particle values 
+%          end . 
+%        Improve all particle positions 
+%  end 
+%  
+%  Final solution reaches in Gbest.       
+%        
+%        
+%\end{verbatim}

### 2009-05-07 trunk/paper2/Paper2.tex
saving before chainging the abstract6

@@
 	\centering		
-	 \includegraphics[scale=0.3]{images/swarmAgentTest.jpg.eps}
+	 \includegraphics[scale=0.3]{images/swarmAgentTest.jpg}
 	 	\caption{Swarm testing} Segmentation Error Vs. Number of agents
@@
 	\centering		
-	 \includegraphics[scale=0.4]{images/swarmParamterC.jpg.eps}
+	 \includegraphics[scale=0.4]{images/swarmParamterC.jpg}
 	 	\caption{Swarm testing} Segmentation Error Vs. PSO parameters.
@@
 \end{figure}
+
+\begin{figure}
+	\centering		
+	 \includegraphics[scale=0.4]{images/AlgEL.jpg}
+	 	\caption{Algorithm comparison on EL-DB} The recognition rate of different algorithms. 
+	 	\label{fig:testEL}
+	%\caption{Experiments results :} a)    b) c) 
+\end{figure}
+
 \subsection{Shape Complexity Analysis}
@@
 		\includegraphics[scale=0.5]{images/featuresdifference.jpg}
-	\caption{Feature Comparison} The effect of different features on accuracy.  %The graph shows the recognition rate of each symbol using different algorithms. 
+	\caption{Feature Comparison on Hs-DB accuracy} The effect of different features on accuracy.  %The graph shows the recognition rate of each symbol using different algorithms. 
 	\label{fig:testFeaturesAll}
+\end{figure*}
+
+ \begin{figure*}
+	\centering
+		\includegraphics[scale=0.34]{images/featELc.jpg}
+	\caption{Feature Comparison on EL-DB} The effect of different features on accuracy.  %The graph shows the recognition rate of each symbol using different algorithms. 
+	\label{fig:ELtestFeaturesAll}
 \end{figure*}  
@@
 It is only based on what the user intends to draw while sketching the symbol.
- The main features affected by the segmentation stage are the geometric features computed in\textbf{FS1 }.
+ The main features affected by the segmentation stage are the geometric features computed in \textbf{FS1}.
 Hence those are the features we used to estimate the efficiency of the segmentation algorithm.
@@
 		\includegraphics[scale=0.4]{images/featureFS1only.jpg}
-	\caption{} %The effect of symbol complexity.  %The graph shows the recognition rate of each symbol using different algorithms. 
+	\caption{The effect of segmentation on Hs-DB accuracy} %The effect of symbol complexity.  %The graph shows the recognition rate of each symbol using different algorithms. 
 	\label{fig:testFeatonly}

### 2009-05-26 paper2/Paper2.tex
change after moving to the new computer and before chaing it to doctor magdah request.

@@
 \twoauthors
-  {Maha El Meseery, Mahmoud Fakhr El Din\sthanks{Associate Professor}, Samia Mashali\sthanks{Professor}}
+  {Maha El Meseery\sthanks{Corresponding Author}, Mahmoud Fakhr El Din\sthanks{Associate Professor}, Samia Mashali\sthanks{Professor}}
 	{Signals Processing Group \\ Computers and Systems Department \\
   Electronic Research Institute\\Cairo, Egypt\\
-	melmeseery@eri.sci.eg, samia@eri.sci.eg}
+	melmeseery@eri.sci.eg, mafakhr@mcit.gov.eg,\\ samia@eri.sci.eg}
   {Magda Fayek\sthanks{Associate Professor}, Nevin Darwish\sthanks{Professor}}
-	{ Computer Engineering Department\\
+	{Computer Engineering Department\\
 	Faculty of Engineering \\Cairo University\\Cairo, Egypt\\
@@
 \begin{abstract}
-Sketch recognition is defined as the process of identifying symbols that users draw using single or multiple strokes.
+%This study 
+Sketches and drawings are widely used as a simple method of expressing thought and designs.
+The goal of this paper is developing a sketch recognition system that facilitates using sketches in computer systems.
+The system uses Particle Swarm Optimization (PSO) algorithm to optimally segment the strokes the user draws into meaningful geometric primitives.
+These geometric primitives are grouped to formulate symbols which are further identified using a Support Vector Machines (SVM) classifier.
+Two different segmentation algorithms were evaluated and results show that both algorithms improve the final recognition results.
+Experiments were conducted on three different dataset, simple presentation (Hs-DB), Electrical Design (EL-DB) and Logic Design (LD-DB) symbols.
+Results show that the effectiveness of the proposed system on various dataset especially on datasets with low shape diversity as in Logic Design dataset (LD-DS).
+% sketch are divided  Users draw symbols and sketchs using a set of pen strokes.
+%Sketch recognition is defined as the process of identifying symbols that users draw using single or multiple strokes.
 Usually, users draw strokes using pens.
@@
 Similarly, direction information is used as it provides better distinctive maxima for the corners than the curvature information \cite{meanshift10}.% Figure \ref{fig:curvatures} shows the direction and curvature graphs for the stroke drawn in Figure \ref{fig:orignalStroke}
+ \begin{figure}[h]
+	\centering
+		\includegraphics[scale=0.5]{images/stroke3.jpg}
+	\caption{Input Stroke} Example of an input stroke to the system. 
+	\label{fig:orignalStroke}
+\end{figure}
  \begin{figure}
@@
 	\label{fig:speed2Distance}
-\end{figure}
-\begin{figure}[]
-	\centering
-		\includegraphics[scale=0.5]{images/stroke3.jpg}
-	\caption{Input Stroke} Example of an input stroke to the system. 
-	\label{fig:orignalStroke}
 \end{figure}
@@
 All the points labeled as possible dominant points $P_{pd}$ are saved into a single array.
-Figure \ref{fig:LabelsPPD} shows the particles labeled as Possible dominate point $P_{pd}$ as the preprocessing.
+Figure \ref{fig:ppd999} shows the particles labeled as Possible dominate point $P_{pd}$ as the preprocessing.
 It is clear that some of the $P_{pd}$ points are redundant as specified in the preprocessing stage.
@@
 	\centering
-		\subfigure[Possible dominate point]{ 	\label{fig:LabelsPPD}	\includegraphics[scale=0.7]{images/ppd.jpg}}
-		\hfill
-			\subfigure[DPSO algorithm encoding]{	\label{fig:LabelsPSO} \includegraphics[scale=0.7]{images/pso1.jpg}}			
-	\caption{Segmentation}% a) Possible dominate point b) Particle encoding  
-	\label{fig:pso1}
+\includegraphics[scale=0.7]{images/ppd.jpg}
+	\caption{Possible Dominante Points}
+	%\label{fig:ppd}
+	\label{fig:ppd999}	
 \end{figure}
@@
 \label{subsubsec:Discreteparticleswarmalgorithm}
+\begin{figure}
+	\centering
+		%\subfigure[Possible dominate point]{ 	\label{fig:LabelsPPD}	\includegraphics[scale=0.7]{images/ppd.jpg}}
+	%	\hfill
+	 \includegraphics[scale=0.7]{images/pso1.jpg}			
+	\caption{DPSO algorithm encoding}% a) Possible dominate point b) Particle encoding  
+	\label{fig:pso1}
+\end{figure}
 Two DPSO algorithms are used to generate different segmentations for each stroke the user draws.
@@
 Therefore, the system represents each particle $P_i$ by $P_i = \left\{ {p_{ij} \left| {j = 1,2 \cdots N} \right.} \right\}$ where $p_{ij}$ has only two values a)1 ($p_{ij}=1$); means that this point ($j$) is a dominate point, or b) 0($p_{ij}=0$) means this point ($j$) is not a dominate point.
-Figure\ref{fig:LabelsPSO} shows a particle encoding in the DPSO system.
+Figure\ref{fig:pso1} shows a particle encoding in the DPSO system.
 
@@
 %In the following section, the recognition accuracy means the average recognition accuracy in the four data splits.
-\begin{figure}
-\centering 
-\label{fig:symbolSet}
-\fbox{ \parbox{5cm}{% 
-		\includegraphics[scale=0.5]{images/symbolSet.PNG}	}}
-		\caption{The Symbol Set} Symbols used in the dataset.
-\end{figure}
+
 \subsection{Algorithms comparisons}
 \label{sec:AlgExp}
-A set of test were performed on the system to enhance both the swarm algorithms outputs.
-The effect of number of agents, the error thresholds and various other paramters was investigated.
+A set of tests were performed on the system to enhance non ellipse fitting algorithm.
+The effect of number of agents, the error thresholds and various other parameters were investigated.
 The result of these tests are shown in Figure \ref{fig:swarmtesting} and   \ref{fig:swarmtesting2}.
 Figure \ref{fig:swarmtesting} shows that as the number of agents increases the error value decrease (error of the segmentation algorithm).
-The final number of agents and maximum swarm iteration used in the system was based trade off between error achieved versus the computation time.
-Other algorithm parameters like $c_1$,$c_2$ that were mentioned in Section\ref{sec:ParticleSwarmAlgorithm} are tested using similar tests (Figure \ref{fig:swarmtesting} and \ref{fig:swarmtesting2}).
+The final number of agents and maximum swarm iteration used in the system was based on a trade off between error achieved versus the computation time.
+Other algorithm parameters such as $c_1$,$c_2$ that were mentioned in Section\ref{sec:ParticleSwarmAlgorithm} are tested using similar tests (Figure \ref{fig:swarmtesting} and \ref{fig:swarmtesting2}).
 
@@
 	\centering		
-	 \includegraphics[scale=0.3]{images/swarmTesting.jpg}
+	 \includegraphics[scale=0.3]{images/swarmAgentTest.jpg}
 	 	\caption{Swarm testing} Segmentation Error Vs. Number of agents
@@
 
-\begin{figure*}
+\begin{figure}
 	\centering		
-	 \includegraphics[width=18cm]{images/swarmTesting2.jpg}
+	 \includegraphics[scale=0.4]{images/swarmParamterC.jpg}
 	 	\caption{Swarm testing} Segmentation Error Vs. PSO parameters.
@@
 	 \includegraphics[scale=0.4]{images/testAlg.jpg}
-	 	\caption{Algorithm comparison} The recognition rate of different algorithms. 
+	 	\caption{Algorithm comparison on HS-DB} The recognition rate of different algorithms. 
 	 	\label{fig:test1}
+	%\caption{Experiments results :} a)    b) c) 
+\end{figure}
+
+\begin{figure}
+	\centering		
+	 \includegraphics[scale=0.4]{images/LDAlg.jpg}
+	 	\caption{Algorithm comparison on LD-DB} The recognition rate of different algorithms. 
+	 	\label{fig:testLD}
+	%\caption{Experiments results :} a)    b) c) 
+\end{figure}
+
+\begin{figure}
+	\centering		
+	 \includegraphics[scale=0.4]{images/AlgEL.jpg}
+	 	\caption{Algorithm comparison on EL-DB} The recognition rate of different algorithms. 
+	 	\label{fig:testEL}
 	%\caption{Experiments results :} a)    b) c) 
@@
 	\centering
-		\includegraphics[scale=0.5]{images/testFeat.jpg}
-	\caption{Feature Comparison} The effect of different features on accuracy.  %The graph shows the recognition rate of each symbol using different algorithms. 
+		\includegraphics[scale=0.5]{images/featuresdifference.jpg}
+	\caption{Feature Comparison on Hs-DB accuracy} The effect of different features on accuracy.  %The graph shows the recognition rate of each symbol using different algorithms. 
 	\label{fig:testFeaturesAll}
@@
 
+ \begin{figure*}
+	\centering
+		\includegraphics[scale=0.34]{images/featELc.jpg}
+	\caption{Feature Comparison on EL-DB} The effect of different features on accuracy.  %The graph shows the recognition rate of each symbol using different algorithms. 
+	\label{fig:ELtestFeaturesAll}
+\end{figure*}  
  We could not test the result of the segmentation algorithm directly due to the fact that the correct segmentation is highly ambiguous.
 It is only based on what the user intends to draw while sketching the symbol.
-Figure \ref{fig:SampleSeg} shows a sample of the output of the segmentation block.
+ The main features affected by the segmentation stage are the geometric features computed in \textbf{FS1}.
+Hence those are the features we used to estimate the efficiency of the segmentation algorithm.
+The recognition accuracy using only the geometrical features is reported in Figure \ref{fig:testFeatonly} .
+The figure shows that \textsl{AlgS2} achieve better results.
+%effect of the segmentation algorithm
+ %Figure \ref{fig:SampleSeg} shows a sample of the output of the segmentation block.
 The result of segmentation block is tested by their effect on the final recognition results using the features that will be highly affected using different algorithms like \textbf{FS1} and \textbf{FS4} (Section\ref{sec:Recognition}).
@@
 	\centering
-		\includegraphics[scale=0.4]{images/testF1.jpg}
-	\caption{} %The effect of symbol complexity.  %The graph shows the recognition rate of each symbol using different algorithms. 
+		\includegraphics[scale=0.4]{images/featureFS1only.jpg}
+	\caption{The effect of segmentation on Hs-DB accuracy} %The effect of symbol complexity.  %The graph shows the recognition rate of each symbol using different algorithms. 
 	\label{fig:testFeatonly}
@@
 The tradeoff between accuracy achieved and time complexity must be further investigated to achieve better results.
-The use of statistical moments and structural features improves the recognition rate.
-The system was tested on 13 different symbols and achieved an accuracy of 93.5\%.
+The final recognition is based on both global shape and hierarchical stroke properties.
+Adding stroke and segment geometrical features improved the final recognition.
+Global shape and moment properties was used to prevent the need to use graph matching which is computationally expensive.
+The system was tested on 13 different symbols and achieved an overall accuracy of 89.5\%.
 The system does not depend on low level recognizer but rather on a set of high level features.
@@
  A possible extension of this research is to complete the clustering algorithm for fully automated sketch recognition.
-Currently, the system process one symbol at a time, an enchantment would be to allow users to draw more than one symbol at a time and incorporate a method for separating symbols.
-Other area of enhancements is the features extraction methods.
+Currently, the system processes one symbol at a time, an enhancement would be to allow users to draw more than one symbol at a time and incorporate a method for separating symbols.
+Another area of enhancements is the features extraction methods.
 Introducing more spatial and geometrical features is believed to improve classifications.
+Features that represents the appearance of the symbols are likely to improve the recognition that have similar geometrical structure (i.e.
+Nand and And gate in logic symbols).
 \bibliographystyle{IEEEbib}

### 2009-05-26 trunk/paper2/Paper2.tex
the new pattern recognition format

@@
-\documentclass{article}
-\usepackage{spconf}
+%\documentclass{article}
+\documentclass[preprint,authoryear,10pt,3p]{elsarticle}
+%\usepackage{spconf}
 \usepackage{amsmath}
-\usepackage[dvips,pdftex]{graphicx}
+\usepackage{graphicx}
+%\usepackage[dvips,pdftex]{graphicx}
 \usepackage{subfigure}
-\usepackage{named}
-\usepackage[dvips,pdftex]{hyperref}
+%\usepackage{named}
+%\usepackage[dvips,pdftex]{hyperref}
+
+%\title{Sketch Recognition using Particle Swarm Optimizations}
+%\twoauthors
+%  {Maha El Meseery\sthanks{Corresponding Author}, Mahmoud Fakhr El Din\sthanks{Associate Professor}, Samia Mashali\sthanks{Professor}}
+%	{Signals Processing Group \\ Computers and Systems Department \\
+%  Electronic Research Institute\\Cairo, Egypt\\
+%	melmeseery@eri.sci.eg, mafakhr@mcit.gov.eg,\\ samia@eri.sci.eg}
+%  {Magda Fayek\sthanks{Associate Professor}, Nevin Darwish\sthanks{Professor}}
+%	{Computer Engineering Department\\
+%	Faculty of Engineering \\Cairo University\\Cairo, Egypt\\
+%	magdafayek@gmail.com,ndarwish@ieee.org}
+\begin{document}
+
+\begin{frontmatter}
+
+%% Title, authors and addresses
+
+%% use the tnoteref command within \title for footnotes;
+%% use the tnotetext command for the associated footnote;
+%% use the fnref command within \author or \address for footnotes;
+%% use the fntext command for the associated footnote;
+%% use the corref command within \author for corresponding author footnotes;
+%% use the cortext command for the associated footnote;
+%% use the ead command for the email address,
+%% and the form \ead[url] for the home page:
+%%
+%% \title{Title\tnoteref{label1}}
+%% \tnotetext[label1]{}
+%% \author{Name\corref{cor1}\fnref{label2}}
+%% \ead{email address}
+%% \ead[url]{home page}
+%% \fntext[label2]{}
+%% \cortext[cor1]{}
+%% \address{Address\fnref{label3}}
+%% \fntext[label3]{}
+
 \title{Sketch Recognition using Particle Swarm Optimizations}
-\twoauthors
-  {Maha El Meseery, Mahmoud Fakhr El Din\sthanks{Associate Professor}, Samia Mashali\sthanks{Professor}}
-	{Signals Processing Group \\ Computers and Systems Department \\
-  Electronic Research Institute\\Cairo, Egypt\\
-	melmeseery@eri.sci.eg, samia@eri.sci.eg}
-  {Magda Fayek\sthanks{Associate Professor}, Nevin Darwish\sthanks{Professor}}
-	{Computer Engineering Department\\
-	Faculty of Engineering \\Cairo University\\Cairo, Egypt\\
+
+%% use optional labels to link authors explicitly to addresses:
+%% \author[label1,label2]{<author name>}
+%% \address[label1]{<address>}
+%% \address[label2]{<address>}
+
+\author[ERI]{Maha El Meseery\footnote{Crossponding author}}
+\author[ERI]{Mahmoud Fakhr El Din}
+\author[ERI]{Samia Mashali}
+\author[FAC]{Magda Fayek}
+\author[FAC]{Nevin Darwish}
+\address[ERI]{Signals Processing Group, Computers and Systems Department \\
+  Electronic Research Institute\\Cairo, Egypt\\melmeseery@eri.sci.eg, mafakhr@mcit.gov.eg, samia@eri.sci.eg}
+\address[FAC]{Computer Engineering Department\\
+	Faculty of Engineering,Cairo University\\Cairo, Egypt\\
 	magdafayek@gmail.com,ndarwish@ieee.org}
-\begin{document}
-\maketitle
 \begin{abstract}
-Sketch recognition is defined as the process of identifying symbols that users draw using single or multiple strokes.
+%%This study 
+Sketches and drawings are widely used as a simple method of expressing thought and designs.
+The goal of this paper is developing a sketch recognition system that facilitates using sketches in computer systems.
+The system uses Particle Swarm Optimization (PSO) algorithm to optimally segment the strokes the user draws into meaningful geometric primitives.
+These geometric primitives are grouped to formulate symbols which are further identified using a Support Vector Machines (SVM) classifier.
+Two different segmentation algorithms were evaluated and results show that both algorithms improve the final recognition results.
+Experiments were conducted on three different dataset, simple presentation (Hs-DB), Electrical Design (EL-DB) and Logic Design (LD-DB) symbols.
+Results show that the effectiveness of the proposed system on various dataset especially on datasets with low shape diversity as in Logic Design dataset (LD-DS).
+% sketch are divided  Users draw symbols and sketchs using a set of pen strokes.
+%%Sketch recognition is defined as the process of identifying symbols that users draw using single or multiple strokes.
 Usually, users draw strokes using pens.
@@
 As for recognition we use Support Vector Machines (SVM) classifier which further improves the final recognition accuracy.
+
 \end{abstract}
+
+\begin{keyword}
+%% keywords here, in the form: keyword \sep keyword
+
+%% PACS codes here, in the form: \PACS code \sep code
+
+%% MSC codes here, in the form: \MSC code \sep code
+%% or \MSC[2008] code \sep code (2000 is the default)
+
+\end{keyword}
+
+\end{frontmatter}
+
+%\maketitle
+%\begin{abstract}
+%%This study 
+%Sketches and drawings are widely used as a simple method of expressing thought and designs.
+The goal of this paper is developing a sketch recognition system that facilitates using sketches in computer systems.
+The system uses Particle Swarm Optimization (PSO) algorithm to optimally segment the strokes the user draws into meaningful geometric primitives.
+These geometric primitives are grouped to formulate symbols which are further identified using a Support Vector Machines (SVM) classifier.
+Two different segmentation algorithms were evaluated and results show that both algorithms improve the final recognition results.
+Experiments were conducted on three different dataset, simple presentation (Hs-DB), Electrical Design (EL-DB) and Logic Design (LD-DB) symbols.
+Results show that the effectiveness of the proposed system on various dataset especially on datasets with low shape diversity as in Logic Design dataset (LD-DS).
+% sketch are divided  Users draw symbols and sketchs using a set of pen strokes.
+%%Sketch recognition is defined as the process of identifying symbols that users draw using single or multiple strokes.
+Usually, users draw strokes using pens.
+The sketch recognition system immediately interprets their strokes as objects that can be easily manipulated.
+This paper uses Particle Swarm Optimization (PSO) to divide the strokes the user draws into meaningful geometric primitives.
+These geometric primitives are grouped to formulate symbols which are further identified.
+The results show that using PSO improves segmentation results which guide the symbol recognition phase.
+As for recognition we use Support Vector Machines (SVM) classifier which further improves the final recognition accuracy.
+%\end{abstract}
 
@@
 The axes of the ellipse are estimated as $width/2$ and $height/2$ of the stroke bounding box.
-The least square fitting algorithm is used to minimize the fitting error of the ellipse equation \cite{chernov-2003}.
+The least square fitting algorithm is used to minimize the fitting error of the ellipse equation \cite{chernov}.
 This process tries to fit the stroke points into an ellipse arc; it starts with computing the center of the stroke bounding box.
@@
 The axes of the ellipse are estimated as the $width/2$ and $height/2$ of the stroke bounding box where $width$ is width and $height$ is height of the bonding box.
-The least square fitting algorithm \cite{chernov-2003} is used to minimize the fitting error of the ellipse Equation (\ref{eq:circleFit})
+The least square fitting algorithm \cite{chernov} is used to minimize the fitting error of the ellipse Equation (\ref{eq:circleFit})
 
@@
 Nand and And gate in logic symbols).
-\bibliographystyle{IEEEbib}
+ \bibliographystyle{elsarticle-harv} 
+%\bibliographystyle{IEEEbib}
 \bibliography{../../neededfiles/Bibliographies/Mybibliography}

### 2009-05-26 paper2/Paper2.tex
changing style

@@
-\documentclass{article}
-\usepackage{spconf}
+\documentclass[preprint,authoryear,10pt,3p]{elsarticle}
+
 \usepackage{amsmath}
-\usepackage[dvips,pdftex]{graphicx}
+\usepackage{graphicx}
 \usepackage{subfigure}
-\usepackage{named}
-\usepackage[dvips,pdftex]{hyperref}
+
+%% for the old format  ( remove this and the abstract and the bibgraphy style. )
+%\documentclass{article}
+%\usepackage{spconf}
+%\usepackage{amsmath}
+%\usepackage[dvips,pdftex]{graphicx}
+%\usepackage{named}
+%\usepackage[dvips,pdftex]{hyperref}
+
+%\title{Sketch Recognition using Particle Swarm Optimizations}
+%\twoauthors
+%  {Maha El Meseery\sthanks{Corresponding Author}, Mahmoud Fakhr El Din\sthanks{Associate Professor}, Samia Mashali\sthanks{Professor}}
+%	{Signals Processing Group \\ Computers and Systems Department \\
+%  Electronic Research Institute\\Cairo, Egypt\\
+%	melmeseery@eri.sci.eg, mafakhr@mcit.gov.eg,\\ samia@eri.sci.eg}
+%  {Magda Fayek\sthanks{Associate Professor}, Nevin Darwish\sthanks{Professor}}
+%	{Computer Engineering Department\\
+%	Faculty of Engineering \\Cairo University\\Cairo, Egypt\\
+%	magdafayek@gmail.com,ndarwish@ieee.org}
+%\maketitle
+
+\begin{document}
+\begin{frontmatter}
 \title{Sketch Recognition using Particle Swarm Optimizations}
-\twoauthors
-  {Maha El Meseery\sthanks{Corresponding Author}, Mahmoud Fakhr El Din\sthanks{Associate Professor}, Samia Mashali\sthanks{Professor}}
-	{Signals Processing Group \\ Computers and Systems Department \\
-  Electronic Research Institute\\Cairo, Egypt\\
-	melmeseery@eri.sci.eg, mafakhr@mcit.gov.eg,\\ samia@eri.sci.eg}
-  {Magda Fayek\sthanks{Associate Professor}, Nevin Darwish\sthanks{Professor}}
-	{Computer Engineering Department\\
-	Faculty of Engineering \\Cairo University\\Cairo, Egypt\\
+
+%% use optional labels to link authors explicitly to addresses:
+%% \author[label1,label2]{<author name>}
+%% \address[label1]{<address>}
+%% \address[label2]{<address>}
+
+\author[ERI]{Maha El Meseery\footnote{Crossponding author}}
+\author[ERI]{Mahmoud Fakhr El Din}
+\author[ERI]{Samia Mashali}
+\author[FAC]{Magda Fayek}
+\author[FAC]{Nevin Darwish}
+\address[ERI]{Signals Processing Group, Computers and Systems Department \\
+  Electronic Research Institute\\Cairo, Egypt\\melmeseery@eri.sci.eg, mafakhr@mcit.gov.eg, samia@eri.sci.eg}
+\address[FAC]{Computer Engineering Department\\
+	Faculty of Engineering,Cairo University\\Cairo, Egypt\\
 	magdafayek@gmail.com,ndarwish@ieee.org}
-\begin{document}
-\maketitle
 \begin{abstract}
-%This study 
+%%This study 
 Sketches and drawings are widely used as a simple method of expressing thought and designs.
@@
 % sketch are divided  Users draw symbols and sketchs using a set of pen strokes.
-%Sketch recognition is defined as the process of identifying symbols that users draw using single or multiple strokes.
+%%Sketch recognition is defined as the process of identifying symbols that users draw using single or multiple strokes.
 Usually, users draw strokes using pens.
@@
 As for recognition we use Support Vector Machines (SVM) classifier which further improves the final recognition accuracy.
+
 \end{abstract}
+
+\begin{keyword}
+%% keywords here, in the form: keyword \sep keyword
+
+%% PACS codes here, in the form: \PACS code \sep code
+
+%% MSC codes here, in the form: \MSC code \sep code
+%% or \MSC[2008] code \sep code (2000 is the default)
+
+\end{keyword}
+
+\end{frontmatter}
+
+%\begin{abstract}
+%%This study 
+%Sketches and drawings are widely used as a simple method of expressing thought and designs.
+The goal of this paper is developing a sketch recognition system that facilitates using sketches in computer systems.
+The system uses Particle Swarm Optimization (PSO) algorithm to optimally segment the strokes the user draws into meaningful geometric primitives.
+These geometric primitives are grouped to formulate symbols which are further identified using a Support Vector Machines (SVM) classifier.
+Two different segmentation algorithms were evaluated and results show that both algorithms improve the final recognition results.
+Experiments were conducted on three different dataset, simple presentation (Hs-DB), Electrical Design (EL-DB) and Logic Design (LD-DB) symbols.
+Results show that the effectiveness of the proposed system on various dataset especially on datasets with low shape diversity as in Logic Design dataset (LD-DS).
+% sketch are divided  Users draw symbols and sketchs using a set of pen strokes.
+%%Sketch recognition is defined as the process of identifying symbols that users draw using single or multiple strokes.
+Usually, users draw strokes using pens.
+The sketch recognition system immediately interprets their strokes as objects that can be easily manipulated.
+This paper uses Particle Swarm Optimization (PSO) to divide the strokes the user draws into meaningful geometric primitives.
+These geometric primitives are grouped to formulate symbols which are further identified.
+The results show that using PSO improves segmentation results which guide the symbol recognition phase.
+As for recognition we use Support Vector Machines (SVM) classifier which further improves the final recognition accuracy.
+%\end{abstract}
 

### 2009-05-27 paper2/Paper2.tex
final template of size
adding to commit at the begining of the day

@@
-\documentclass[preprint,authoryear,10pt,3p]{elsarticle}
+\documentclass[preprint,10pt,5p,twocolumn]{elsarticle}
+%\documentclass[preprint,authoryear,10pt,5p,twocolumn]{elsarticle}
 
@@
 %%This study 
-Sketches and drawings are widely used as a simple method of expressing thought and designs.
+Sketches and drawings are widely used as simple methods of expressing thought and designs.
 The goal of this paper is developing a sketch recognition system that facilitates using sketches in computer systems.
 The system uses Particle Swarm Optimization (PSO) algorithm to optimally segment the strokes the user draws into meaningful geometric primitives.
-These geometric primitives are grouped to formulate symbols which are further identified using a Support Vector Machines (SVM) classifier.
-Two different segmentation algorithms were evaluated and results show that both algorithms improve the final recognition results.
+These geometric primitives are grouped to form symbols which are further identified using a Support Vector Machines (SVM) classifier.
+Two versions of PSO segmentation algorithms were implemented.
 Experiments were conducted on three different dataset, simple presentation (Hs-DB), Electrical Design (EL-DB) and Logic Design (LD-DB) symbols.
-Results show that the effectiveness of the proposed system on various dataset especially on datasets with low shape diversity as in Logic Design dataset (LD-DS).
-% sketch are divided  Users draw symbols and sketchs using a set of pen strokes.
+Results show that both PSO algorithms produce better segmentation than most current systems.
+This segmentation improvement contributed in getting more accurate recognition results on all the datasets tested.
+The results prove that the system is effective specially in datasets with low diversity in shapes (i.e.
+shapes are very simillar) as in Logic Design dataset (LD-DS).
+%The results prove the effectiveness of the proposed system on various dataset, especially datasets with low shape diversity as in Logic Design dataset (LD-DS).% effectiveness of the proposed system on various dataset, especially datasets with low shape diversity as in Logic Design dataset (LD-DS).
+%Result show that the system is especially effictive on dataset with % sketch are divided  Users draw symbols and sketchs using a set of pen strokes.
 %%Sketch recognition is defined as the process of identifying symbols that users draw using single or multiple strokes.
@@
 % sketch are divided  Users draw symbols and sketchs using a set of pen strokes.
-%%Sketch recognition is defined as the process of identifying symbols that users draw using single or multiple strokes.
+%%Sketch recognition is defined as the process of identifying symbols that users draw using single or multiple %strokes.
 Usually, users draw strokes using pens.
-The sketch recognition system immediately interprets their strokes as objects that can be easily manipulated.
-This paper uses Particle Swarm Optimization (PSO) to divide the strokes the user draws into meaningful geometric primitives.
-These geometric primitives are grouped to formulate symbols which are further identified.
-The results show that using PSO improves segmentation results which guide the symbol recognition phase.
-As for recognition we use Support Vector Machines (SVM) classifier which further improves the final recognition accuracy.
+The sketch recognition system immediately interprets their %strokes as objects that can be easily manipulated.
+This paper uses Particle Swarm Optimization (PSO) to divide %the strokes the user draws into meaningful geometric primitives.
+These geometric primitives are grouped to %formulate symbols which are further identified.
+The results show that using PSO improves segmentation results %which guide the symbol recognition phase.
+As for recognition we use Support Vector Machines (SVM) classifier %which further improves the final recognition accuracy.
 %\end{abstract}
@@
 Pen-based interfaces give the user a pencil-paper like feeling that enhances interaction more than the current used keyboard-mouse computer interfaces.
-However, Using sketches as input to a computer system requires the system to understand the drawn sketches.
-In a truly freely sketch system this is not a trivial problem due to the sloppy and messy way users draw sketches.
+However, using sketches as input to a computer system requires the system to understand the drawn sketches.
+In a truly freely sketch system this is not a trivial problem due to the sloppy and messy  way in which users draw sketches.
 Sketches here stand for any hand drawing the user draws using digital stylus or similar device.
 Normally, sketches are composed of a set of strokes.
-The time between pen-up and pen-down events identify each stroke.
-The strokes are mainly the path of points, which the user path through between the pen-down and pen-up events while using the stylus or similar device.
+The time between pen-up and pen-down events identifies each stroke.
+The strokes are mainly the path of points, which the user draws between the pen-down and pen-up events while using the stylus or similar device.
 
-Sketch recognition can be based on: strokes, global shape properties, and appearance \cite{Oltmans07}.
+Sketch recognition can be based on either Pattern Recognition (PR) method or Artifical Intelligance (AI) based method.
+In AI based methods, Symbols are represented using hierarchical levels of simpler geometric shapes or primitives and spatial relations between them.
+Systems that employ such representation must first pre-process.
+strokes to break them into geometric primitives.
+Recognition can then is treated as an AI problem which can be solved using methods like sub graph matching, bayes network, and constrain statisfaction.
+A Hierarchical shape descriptions was widely used in sketch systems \cite{HierarchicalParsing7,Ladder30,AlvaradoFreedom42,SketchRead2007}.
+
+Recognition based on AI methods is applied in \cite{SketchRead2007}.
+The principle problem in these systems is the computational complexity of the matching between shape description and geometric primitives produced in the segmentation process.
+Natural sketch are difficult to segment and group to create reliable fragment of geometric primitives.
+This is partially due to the increase of noise and other phenomena such as overlapped, over traced strokes and sharing of strokes between different symbols.
+Since these systems are dependable on the segmentation results, they either perform recognition using undependable segmentation results or consider different ways to segment the input sketch.
+%The role of which each stroke plays in the structure of the sketched symbol is used to determine shapes in stroke based methods.  
+%Gestures are pen strokes drawn by the user that can immediately recognized by the system.
+%role of which each stroke plays in the structure of the sketched symbol is used to determine shapes 
+Systems that uses Pattern Recognition methods compute properties that represent different symbols then use a classification method to recognize symbols.
+The properties used can either be based on gesture properties, global shape properties or image based properties.
+Gestures properties are set of features that represent how the user draws a stroke not how it look like.
+In gestures properties approcah, the systems uses the role of each stroke in the structure of the symbol determine the shapes.
+System that uses gestures properties assume that users draw symbols in specific order and style \cite{gestureexample12}.
+In global shape properties approach, systems investigate general shape properties and its underlying strokes.
+Using shape properties (i.e.
+ratio of area of convex hull to area of shape) relaxes the assumption and role of each stroke in the symbol but it does not truly represent the appearance of the symbol\cite{DiagramOfflineConvexHull}.
+Unlike shape properties based methods, image based methods disregard individual strokes and focus on the overall appearance of shapes.
+Image based methods are used in \cite{Oltmans07} where they computed a code book of 1000 features to recognize symbols.
+
+Methods based on gesture and global shape properties produce systems that are dependent on assumptions about how strokes can be grouped.
+These systems assume that users will finish drawing a shape before moving to the next one \cite{Cali63,geometrydomain49}.
+This let them to temporally group strokes and then recognizes them based on their strokes or global shape properties.
+\cite{AlvaradoDigital} argued that this assumption does not hold in natural sketches.
+
+In this paper we choose to implement a hybrid method between shape property, guesture and hierarchical (AI method) based approaches.
+Our method tries to avoid many of the problems faced by approaches based on individual guestures or global shape properties while trying to gain the additional information that the stroke based system use.
+Firstly, the system uses particle swarm optimization (PSO) to optimally segment user's strokes.
+Users can draw symbols using any number of strokes in any order they like.
+The system segments each stroke using two different PSO algorithms.
+The first segmentation algorithm is based on dividing the digital curve into polygons\cite{PolygonApproximationPSO}.
+Our algorithm enhances its performance by adding curvature and other local information while segmenting strokes to achieve better results.
+To handle complex strokes, our system uses a second segmentation PSO algorithm to segment strokes into a set of lines and circular arcs.
+After segmentation, a feature vector composed of different statistical, geometrical and composite features is deduced.
+The final classification step uses a SVM classifier to classify symbols into one of the previously trained classes based on the features vector computed.
+
+ %In general, the  appearance based methods use image processing techniques to recognize sketches that the users draw. 
+%Sketch recognition can be based on: strokes, global shape properties, and appearance \cite{Oltmans07}.
 The role of which each stroke plays in the structure of the sketched symbol is used to determine shapes in stroke based methods.
@@
 Systems that employ such representation must first pre-process strokes to break them into geometric primitives.
-Recognition can then be treated as sub graph matching problem.
+Recognition can hen be treated as sub graph matching problem.
 
- In global shape properties approach, systems investigate general shape properties and its underlying strokes.
+ %In global shape properties approach, systems investigate general shape properties and its underlying strokes.
 Using shape properties (i.e.
@@
 
- Methods based on gesture and global shape properties produce systems that are dependent on assumptions about how strokes can be grouped.
+ %Methods based on gesture and global shape properties produce systems that are dependent on assumptions about how strokes can be grouped.
 These systems assume that users will finish drawing a shape before moving to the next one \cite{Cali63,geometrydomain49}.
@@
 
- Recognition based on strokes using hierarchical shape description is applied in \cite{SketchRead2007}.
+ %Recognition based on strokes using hierarchical shape description is applied in \cite{SketchRead2007}.
 The principle problem in these systems is the computational complexity of the matching between shape description and geometric primitives produced in the segmentation process.
@@
 
-In this paper we choose to implement a hybrid method between shape property and stroke based approaches.
+%In this paper we choose to implement a hybrid method between shape property and stroke based approaches.
 Our method tries to avoid many of the problems faced by approaches based on individual strokes or global shape properties while trying to gain the additional information that the stroke based system use.

### 2009-05-27 trunk/paper2/Paper2.tex
fininshign the introduction

@@
-%\documentclass{article}
-\documentclass[preprint,authoryear,10pt,3p]{elsarticle}
-%\usepackage{spconf}
+\documentclass[preprint,10pt,5p,twocolumn]{elsarticle}
+%\documentclass[preprint,authoryear,10pt,5p,twocolumn]{elsarticle}
+
 \usepackage{amsmath}
 \usepackage{graphicx}
+\usepackage{subfigure}
+
+%% for the old format  ( remove this and the abstract and the bibgraphy style. )
+%\documentclass{article}
+%\usepackage{spconf}
+%\usepackage{amsmath}
 %\usepackage[dvips,pdftex]{graphicx}
-\usepackage{subfigure}
 %\usepackage{named}
@@
 %	magdafayek@gmail.com,ndarwish@ieee.org}
+%\maketitle
+
 \begin{document}
-
 \begin{frontmatter}
-
-%% Title, authors and addresses
-
-%% use the tnoteref command within \title for footnotes;
-%% use the tnotetext command for the associated footnote;
-%% use the fnref command within \author or \address for footnotes;
-%% use the fntext command for the associated footnote;
-%% use the corref command within \author for corresponding author footnotes;
-%% use the cortext command for the associated footnote;
-%% use the ead command for the email address,
-%% and the form \ead[url] for the home page:
-%%
-%% \title{Title\tnoteref{label1}}
-%% \tnotetext[label1]{}
-%% \author{Name\corref{cor1}\fnref{label2}}
-%% \ead{email address}
-%% \ead[url]{home page}
-%% \fntext[label2]{}
-%% \cortext[cor1]{}
-%% \address{Address\fnref{label3}}
-%% \fntext[label3]{}
-
 \title{Sketch Recognition using Particle Swarm Optimizations}
@@
 \author[FAC]{Nevin Darwish}
-\address[ERI]{Signals Processing Group, Computers and Systems Department \\
-  Electronic Research Institute\\Cairo, Egypt\\melmeseery@eri.sci.eg, mafakhr@mcit.gov.eg, samia@eri.sci.eg}
-\address[FAC]{Computer Engineering Department\\
-	Faculty of Engineering,Cairo University\\Cairo, Egypt\\
+\address[ERI]{Signals Processing Group, Computers and Systems Department, 
+  Electronic Research Institute, Cairo, Egypt\\melmeseery@eri.sci.eg, mafakhr@mcit.gov.eg, samia@eri.sci.eg}
+\address[FAC]{Computer Engineering Department,
+	Faculty of Engineering,Cairo University, Cairo, Egypt\\
 	magdafayek@gmail.com,ndarwish@ieee.org}
@@
 %%This study 
-Sketches and drawings are widely used as a simple method of expressing thought and designs.
+Sketches and drawings are widely used as simple methods of expressing thought and designs.
 The goal of this paper is developing a sketch recognition system that facilitates using sketches in computer systems.
 The system uses Particle Swarm Optimization (PSO) algorithm to optimally segment the strokes the user draws into meaningful geometric primitives.
-These geometric primitives are grouped to formulate symbols which are further identified using a Support Vector Machines (SVM) classifier.
-Two different segmentation algorithms were evaluated and results show that both algorithms improve the final recognition results.
-Experiments were conducted on three different dataset, simple presentation (Hs-DB), Electrical Design (EL-DB) and Logic Design (LD-DB) symbols.
-Results show that the effectiveness of the proposed system on various dataset especially on datasets with low shape diversity as in Logic Design dataset (LD-DS).
-% sketch are divided  Users draw symbols and sketchs using a set of pen strokes.
+These geometric primitives are grouped to form symbols which are further identified using a Support Vector Machines (SVM) classifier.
+Two versions of PSO segmentation algorithms were implemented.
+Experiments were conducted on three different dataset, simple presentation (Hs-DS), Electrical Design (EL-DS) and Logic Design (LD-DS) symbols.
+Results show that both PSO algorithms produce better segmentation than most current systems.
+This segmentation improvement contributed in getting more accurate recognition results on all the datasets tested.
+The results prove that the system is effective specially in datasets with low diversity in shapes, i.e., shapes are very simillar, as in Logic Design dataset (LD-DS).
+%The results prove the effectiveness of the proposed system on various dataset, especially datasets with low shape diversity as in Logic Design dataset (LD-DS).% effectiveness of the proposed system on various dataset, especially datasets with low shape diversity as in Logic Design dataset (LD-DS).
+%Result show that the system is especially effictive on dataset with % sketch are divided  Users draw symbols and sketchs using a set of pen strokes.
 %%Sketch recognition is defined as the process of identifying symbols that users draw using single or multiple strokes.
@@
 %% keywords here, in the form: keyword \sep keyword
-
+Symbol Recognition, Particle Swarm Optimization, Sketch Recognition, Geometerical Features, Polygon approximation
 %% PACS codes here, in the form: \PACS code \sep code
@@
 
-%\maketitle
 %\begin{abstract}
@@
 Pen-based interfaces give the user a pencil-paper like feeling that enhances interaction more than the current used keyboard-mouse computer interfaces.
-However, Using sketches as input to a computer system requires the system to understand the drawn sketches.
-In a truly freely sketch system this is not a trivial problem due to the sloppy and messy way users draw sketches.
+However, using sketches as input to a computer system requires the system to understand the drawn sketches.
+In a truly freely sketch system this is not a trivial problem due to the sloppy and messy  way in which users draw sketches.
 Sketches here stand for any hand drawing the user draws using digital stylus or similar device.
 Normally, sketches are composed of a set of strokes.
-The time between pen-up and pen-down events identify each stroke.
-The strokes are mainly the path of points, which the user path through between the pen-down and pen-up events while using the stylus or similar device.
+The time between pen-up and pen-down events identifies each stroke.
+The strokes are mainly the path of points, which the user draws between the pen-down and pen-up events while using the stylus or similar device.
 
-Sketch recognition can be based on: strokes, global shape properties, and appearance \cite{Oltmans07}.
+Sketch recognition can be applied either using  Pattern Recognition (PR) based method or Artifical Intelligance (AI) based method.
+In AI based methods, Symbols are represented using hierarchical levels of simpler geometric shapes or primitives and spatial relations between them.
+Systems that employ such representation must first pre-process.
+strokes to break them into geometric primitives.
+Recognition can then is treated as an AI problem which can be solved using methods like sub graph matching, bayes belief network, and constrain statisfaction.
+A Hierarchical shape descriptions was widely used in sketch systems \cite{HierarchicalParsing7,SketchRead2007}.
+%Ladder30,AlvaradoFreedom42
+
+Recognition based on AI methods as applied in \cite{SketchRead2007} has the principle problem of computational complexity of the matching between shape description and geometric primitives produced in the segmentation process.
+Natural sketch are difficult to be segmented and grouped to create reliable fragment of geometric primitives.
+This is partially due to the increase of noise and other phenomena such as overlapped, over traced strokes and sharing of strokes between different symbols.
+Since these systems are dependable on the segmentation results, they either perform recognition using undependable segmentation results or consider different ways to segment the input sketch.
+This is because they assume that each stroke plays a role in the structure of the shape.
+%The role of which each stroke plays in the structure of the sketched symbol is used to determine shapes in stroke based methods.  
+%Gestures are pen strokes drawn by the user that can immediately recognized by the system.
+%role of which each stroke plays in the structure of the sketched symbol is used to determine shapes 
+%.
+discribing the role  how user draws individal storkes  properties that discribe how users draw individual strokes in drawing the symbol.
+to .
+The identification of individual stroke is used to determine  The symbols are recognized by identifing the individual strokes.
+In PR based methods, systems compute properties that discribe symbols then use a classification method to recognize symbols.
+The properties used can either be based on gesture properties, global shape properties or image based properties.
+Gesture properties are set of features that represent how the user draws a symbol not how it looks like (e.g., identify a regtangle when a user draws a L corner).
+In gesture properties approcah, systems identify symbols by using gesutre properties of some of its  underlying strokes.
+System that uses gestures properties assume that users draw symbols in specific order and style \cite{gestureexample12,aideddesgin22}.
+In global shape properties approach, systems investigate general shape properties and its underlying strokes.
+Using shape properties (e.g., ratio of the area of convex hull to the area of shape) relaxes the assumption and the role of each stroke in the symbol but it does not truly represent the appearance or fine details of the symbol \cite{DiagramOfflineConvexHull,Cali63}.
+Unlike shape properties based methods, image based methods disregard individual strokes and focus on the overall appearance of shapes.
+These system convertes the shapes into image and extracte image based features (e.g.
+local gradient feature)\cite{Oltmans07,imagetrainable48}.
+In \cite{imagetrainable48} used four different distance measures to to match user symbol with the prototype database.
+The prototype of a shape is represented as an image of the symbol that is normalized and scaled.
+However, the problem is that the shape prototype must represent all transformation and variation the input shape can have.
+%Image based methods are used in \cite{Oltmans07,imagetrainable48} where they computed a code book of 1000 features to recognize symbols.
+
+Methods based on gesture and global shape properties produce systems that are dependent on assumptions about how strokes can be grouped.
+These systems assume that users will finish drawing a shape before moving to the next one \cite{Cali63,geometrydomain49}.
+This assumption allows the system to temporally group strokes and then recognizes them based on their strokes or global shape properties.
+Alvarado \cite{AlvaradoDigital} argued that this assumption does not hold in natural sketches.
+
+In this paper, we choose to implement a hybrid method between shape property, guesture and hierarchical (AI method) based approaches.
+Our method attempts to avoid many of the problems faced by approaches based on individual guestures or global shape properties while trying to gain the additional information that the stroke based system uses.
+Firstly, the system uses particle swarm optimization (PSO) to optimally segment user's strokes.
+Users can draw symbols using any number of strokes in any order they like.
+The system segments each stroke using two different PSO algorithms.
+The first segmentation algorithm is based on dividing the digital curve into polygons\cite{PolygonApproximationPSO}.
+Our algorithm enhances its performance by adding curvature and other local information while segmenting strokes to achieve better results.
+To handle complex strokes, our system uses a second segmentation PSO algorithm to segment strokes into a set of lines and circular arcs.
+After segmentation, a feature vector composed of different statistical, geometrical and composite features is deduced.
+The final classification step uses a SVM classifier to classify symbols into one of the previously trained classes based on the features vector computed.
+
+ %In general, the  appearance based methods use image processing techniques to recognize sketches that the users draw. 
+%Sketch recognition can be based on: strokes, global shape properties, and appearance \cite{Oltmans07}.
 The role of which each stroke plays in the structure of the sketched symbol is used to determine shapes in stroke based methods.
@@
 %This paper introduces a new sketch recognition system, which uses  % uses the computed feature vector to classify the segments into one of the previously trained classes.
-
- The remaining of the paper is as follows; the literature review on stroke segmentation is in Section \ref{sec:review}.
+%Sections \ref{Prepross}, \ref{seg} and \ref{sec:Recognition} provide details on preprocessing, segmentation and recognition process respectively.
+ The remaining of the paper is as follows; the literature review on stroke segmentation is presented in Section \ref{sec:review}.
 Section \ref{sec:ParticleSwarmAlgorithm} introduces Particle Swarm Optimization.
-Section \ref{Sysdisc} describes the system in details.
-Sections \ref{Prepross}, \ref{seg} and \ref{sec:Recognition} provide details on preprocessing, segmentation and recognition process respectively.
+Section \ref{Sysdisc} describes the proposed system in details.
 Section \ref{sec:Experiments} describes the experiments that were preformed.
-We conclude in Section \ref{ConclusionandFutureWork}.
+Finally, the conclusion and future work is presented in Section \ref{ConclusionandFutureWork}.
 
@@
 Similarly, direction information is used as it provides better distinctive maxima for the corners than the curvature information \cite{meanshift10}.% Figure \ref{fig:curvatures} shows the direction and curvature graphs for the stroke drawn in Figure \ref{fig:orignalStroke}
+
  \begin{figure}[h]

### 2009-05-27 trunk/paper2/Paper2.tex
adding lot of changes till segmentation alg1

@@
 local gradient feature)\cite{Oltmans07,imagetrainable48}.
-In \cite{imagetrainable48} used four different distance measures to to match user symbol with the prototype database.
+In \cite{imagetrainable48}, they used four different distance measures to to match user symbol with the prototype database.
 The prototype of a shape is represented as an image of the symbol that is normalized and scaled.
@@
 \cite{CruveDivisionSwarm} uses digital curves scanned from paper as input to the system.
-Also, they did not take advantage of the curvature or local geometric properties of the digital curve.
-Yin \cite{PolygonApproximationPSO} used PSO to convert digital curves into polygons, our system adopts  Yin \cite{PolygonApproximationPSO} method and tries to improve it by adding curvature and other local information while segmenting strokes to achieve better segmentations.
-The next section presents the general particle swarm algorithm which is used in both our segmentation algorithms.
+However, they did not take the advantage of the curvature or local geometric properties of the digital curve.
+Yin \cite{PolygonApproximationPSO} used PSO to convert digital curves into polygons.
+Our system adopts  Yin \cite{PolygonApproximationPSO} method and tries to improve it by adding curvature and other local information while segmenting strokes to achieve better segmentations.
+The next section presents the general particle swarm algorithm which is used in this system.
 
@@
 %What is particle swarm algorithm and how it was used in related researches. 
-The main idea of \textit{Particle Swarm Algorithm (PSO)} is to represent each agent with a particle from the solution space \cite{PSOFirst}.
-Each agent moves the particle with a direction and velocity $v_{ij}$ based on equations \ref{eq:Swarm} \& \ref{eq:Swarm1}.
+%The main idea of \textit{Particle Swarm Algorithm (PSO)} is to represent each solution in the solution space with a particle\cite{PSOFirst}.
+Each particle moves the n with a direction and velocity $v_{ij}$ based on equations \ref{eq:Swarm1} \& \ref{eq:Swarm}.
+The main idea of \textit{Particle Swarm Algorithm (PSO)} is to represent each solution with a $N$ dimension particle from the solution space \cite{PSOFirst}.
+Each particle moves with a direction and velocity $v_{ij}$ based on equations \ref{eq:Swarm1} \& \ref{eq:Swarm}.
+
 \begin{equation}
@@
 \end{equation}
-where $p_{ij}$ represent the $jth$ particle in the $ith$ agent and $v_{ij}$ is the velocity of the $jth$ particle in the $ith$ agent.
+%where $p_{ij}$ represent the $jth$ particle in the $ith$ agent and $v_{ij}$ is the velocity of the $jth$ particle in the $ith$ agent.
+where $p_{ij}$ represent the $j$th dimension in the $i$th particle and $v_{ij}$ is the velocity of the $j$th dimension in the $i$th particle.
  %Equation [\ref{eq:Swarm}] shows how velocity and direction of each particle are computed
  \begin{equation}
-v_{ij}  = v_{ij}  + c_1 r_1 (lbest_{ij}  - p_{ij} ) + c_2 r_2 (gbest_{ij}  - p_{ij} )
+v_{ij}  = v_{ij} \omega + c_1 r_1 (lbest_{ij}  - p_{ij} ) + c_2 r_2 (gbest_{ij}  - p_{ij} ),
 \label{eq:Swarm}
 \end{equation}
- where $lbest_{ij}$ is the local best particle, $gbest_{ij}$ is the global best particle, $r_1$ \& $r_2$ are random variables and $c_1$ \& $c_2$ are the swarm system variables.
- After each iteration the global best $g_{best}$ particle and the agent local best $l_{best}$ particle are evaluated based on the maximum fitness functions of all particles in the solution space.
+ where  $\omega$ is the inertia weight parameter which controls the tradeoff between exploration and exploitation,  $lbest_{ij}$ is the local best particle, $gbest_{ij}$ is the global best particle, $r_1$ \& $r_2$ are random variables and $c_1$ \& $c_2$ are the swarm acceleration parameters.
+
+ After each iteration the global best $g_{best}$ particle and the agent local best particle $l_{best}$are evaluated based on the maximum fitness functions of all particles in the solution space.
 The solution is found after achieving a specific number of iteration or after an error threshold is achieved.
-Equation \ref{eq:descrite} is used to change the general swarm algorithm into binary particle (\textit{Discrete Particle Swarm Algorithm DPSO}) which handles particle values of either $0$ or $1$ \cite{PSODisceret}.
+Equation \ref{eq:descrite} is used to change the general swarm algorithm into binary particle (\textit{Discrete Particle Swarm Algorithm DPSO}).
+The \textit{DPSO} handles particle values of either $0$ or $1$ \cite{PSODisceret}.
 \begin{equation}
@@
 
-0 \quad \quad if\quad r_{3}<p_{i} 
+0 \quad \quad if\quad r_{3}<p_{i}, 
 \label{eq:descrite}
@@
 \label{Sysdisc}
- The suggested sketch recognition system is divided into three main steps 1) preprocessing, 2) segmentation and 3) Recognition. The following section provides a detailed description of each step.
+ The suggested sketch recognition system is divided into three main steps; 1) preprocessing, 2) segmentation and 3) Recognition. The following section provides a detailed description of each step.
 
@@
  Speed and time difference information were widely used in sketch recognition systems \cite{earlyprocess}.
-Figure \ref{fig:speed2Distance} shows the time difference and the speed graphs for the stroke drawn in Figure \ref{fig:orignalStroke}.
+Figure \ref{fig:orignalStroke} shows an example of an input stroke drawn by users.
+Figure \ref{fig:speed2Distance} shows the graph of time difference and the speed for the stroke in Figure \ref{fig:orignalStroke}.
 It was noted that time difference provides more distinct maxima than the speed information.
 Agar et al.
-\cite{polygonfeedback31} mention that the pointing device (for example the mouse) sampling rate is the reason for this phenomena.
+\cite{polygonfeedback31} mention that the pointing device (e.g., mouse) sampling rate is the reason for this phenomena.
 The points are sampled at regular time intervals as long as the pointing device is moving.
@@
 This leads to a nearly constant time difference between samples while the pen is moving and large difference while the pen is stationary (Figure \ref{fig:timediff}).
-Contrary to speed information, where the users draw with variant speeds, the speed information has a lot of noise in the data.
-Similarly, direction information is used as it provides better distinctive maxima for the corners than the curvature information \cite{meanshift10}.% Figure \ref{fig:curvatures} shows the direction and curvature graphs for the stroke drawn in Figure \ref{fig:orignalStroke}
+Contrary to time difference information, the speed information, where the users draw with variant speeds, the speed information has a lot of noise in the data (Figure \ref{fig:speed2Distance}).
+%Similarly, direction information is used as it provides better distinctive maxima for the corners than the curvature information \cite{meanshift10}.% Figure \ref{fig:curvatures} shows the direction and curvature graphs for the stroke drawn in Figure \ref{fig:orignalStroke}
 
@@
 \end{figure}
+
  \begin{figure}
 	\centering
-			\subfigure[ Speed Graph]{	\includegraphics[scale=0.4]{images/vel3.jpg}}
+			\subfigure[Speed versus Displacement Graph]{	\includegraphics[scale=0.4]{images/vel3.jpg}}
 			\hfill
-			\subfigure[ Time Difference Graph] {\label{fig:timediff}\includegraphics[scale=0.4]{images/td3.jpg}}
-	\caption{Speed and Time Difference Graphs}
+			\subfigure[ Time Difference  versus Displacement Graph] {\label{fig:timediff}\includegraphics[scale=0.4]{images/td3.jpg}}
+	\caption{Speed and Time Difference Graphs}  a) speed graph has lot of noise in the data with respect to b) time difference.
+The black points denote the location of the possible dominant points $P_{pd}$.
+Black rectangles represent regions $R_i$ that in a) are lower than the thershold.
+b) are above the thershold.
 	\label{fig:speed2Distance}
 \end{figure}
-Dominant points are characterized by lower speed values, higher curvature and direction values.
+%is calculated as the angle between two vectors and  where $Q_i=\overrightarrow {P_{i - 1} P_i }$ is the vector from the point $P_{i - 1}$ to point $P_i$ 
+Dominant points are characterized by low speed values, high curvature and direction values.
 However, the dominant points from the curvature information may not be the same points as the dominant points extracted from the speed information.
@@
 The speed is calculated as $v=\Delta s/\Delta t$ where $t$ is the time difference between two points and $s$ is the length between them.
-The direction $d$ is calculated as the angle between two vectors and curvature is considered as the change in direction with respect to length i.e.
+The direction $d$ calculated as the angle between vectors $\overrightarrow {P_{i - 1} P_i }$ and the $x-axis$ and the curvature is considered as the change in direction $d$ with respect to length $s$ i.e.
 $c= \Delta d/\Delta s$.
 
-After the system computes the speed, time difference and curvature information it proceeds to detect the points with low velocity and high curvature.
-However, using simple differentiation to detect local extreme points resulted in false points due to the non smooth curves.
-Hence, the system adopted a process presented by \cite{earlyprocess}, where the mean of each of the four calculated information  is calculated.
+After the system computes the speed, time difference and curvature information, it proceeds to detect the points that have low velocity and high curvature.
+However, using simple differentiation to detect local extreme points results in false points due to the non smooth curves.
+Hence, our system adopted a process presented by \cite{earlyprocess}, where the mean of each of the four calculated information  is calculated.
 This mean is taken as the threshold \textit{th} which is used to separate the curve into regions ($R_i$) (Figure\ref{fig:timediff}); each region $R_i$ is defined as a range of points, where the curve values are either above or below the threshold \textit{th}.
@@
 All the points labeled as possible dominant points $P_{pd}$ are saved into a single array.
-Figure \ref{fig:ppd999} shows the particles labeled as Possible dominate point $P_{pd}$ as the preprocessing.
+Figure \ref{fig:ppd999} shows the particles labeled as Possible dominate point $P_{pd}$ in the preprocessing phase.
 It is clear that some of the $P_{pd}$ points are redundant as specified in the preprocessing stage.
@@
 \includegraphics[scale=0.7]{images/ppd.jpg}
-	\caption{Possible Dominante Points}
+	\caption{Possible Dominante Points} An illustration of stroke points and the expected \textit{possible dominant points} $P_{pd}$.  
 	%\label{fig:ppd}
@@
 \label{seg}
-In the segmentation stage the input stroke is divided into a set of primitives.
+In the segmentation stage, the input stroke is divided into a set of primitives.
 As shown in Figure \ref{fig:segblock} first an attempt is made to fit the stroke points into a curve or an ellipse using a minimum square error fitting algorithm \cite{ellipsefit}.
-The ellipse fitting step helps to prevent the system from over segmenting the stroke into multiple lines or curves if the input stroke is an ellipse.
+The ellipse fitting step mainly helps to prevent the system from over segmenting the stroke into multiple lines or curves if the input stroke is an ellipse.
 If the stroke proved to be an ellipse arc then the segmentation process ends and the system proceeds to the next step.
 Otherwise, the stroke is passed to two further segmentation algorithms that divide the stroke to either lines or lines and curves.
-After two different segmentations are generated, the system chooses the segmentation with the minimum error.
+Hence we generate two different segmentations, the system chooses the segmentation with the minimum error.
 The following sections describe in details the ellipse detection algorithm and the two segmentation algorithms used to divide the stroke.
@@
 The output is a set of dominant points which are connected by either lines or curves.
-
+%After two different segmentations are generated, 
 %///AllBlockDiagram.jpg  you can change the block to the main block. 
@@
 The least square fitting algorithm is used to minimize the fitting error of the ellipse equation \cite{chernov}.
-This process tries to fit the stroke points into an ellipse arc; it starts with computing the center of the stroke bounding box.
+This process attempts to fit the stroke points into an ellipse arc; it starts with computing the center of the stroke bounding box.
 The bounding box center point is used as the first estimation of the center of the ellipse.
-The axes of the ellipse are estimated as the $width/2$ and $height/2$ of the stroke bounding box where $width$ is width and $height$ is height of the bonding box.
+The axes of the ellipse are estimated as the $width/2$ and $height/2$ of the stroke bounding box where $width$ and $height$ are the width and height of the bonding box.
 The least square fitting algorithm \cite{chernov} is used to minimize the fitting error of the ellipse Equation (\ref{eq:circleFit})
-
 \begin{equation}
@@
 \end{equation}
-
- where $N$ is number of points in the stroke, $a,b$ are the length of ellipse axes, $x_0$ \& $y_0$ are the coordinates of the center point, $x_i$ \& $y_i$ are the coordinates of point $i$ in the stroke.
+where $N$ is number of points in the stroke, $a$ \& $b$ are the length of ellipse axes, $x_0$ \& $y_0$ are the coordinates of the center point and $x_i$ \& $y_i$ are the coordinates of point $i$ in the stroke.
 A list of new values for $x_0$ , $y_0$ ,$a$ and $b$ are generated randomly from the older values with small increments after each loop.
-After few iteration, the final fit error of the estimated ellipse is reported.
+After few iterations, the final fit error of the estimated ellipse is reported.
 Another measure is used to compute the efficiency of the final estimated ellipse.
 Equation \ref{eq:circleError} ensures that the drawn percentage of ellipse is considered.
-This eliminates fitting a  line into very large ellipse but leaves small ellipses to be fitted as a partial or full ellipse.
+This eliminates fitting a  line into very large ellipse but leaves small ellipses to be fitted as a partial or a full ellipse.
 
  \begin{equation}
-eff= (P_{percent}/E)
+%eff= (P_{percent}/E)
+%eff = \left\{ {\begin{array}{*{20}c}
+%   {\frac{{P_{percent} }}{E}} & {ifP_{percent}  < 0.5}  \\
+%   {\frac{{P_{percent} }}{{E \times 2}}} & {ifP_{percent}  \ge 0.5}  \\
+%\end{array}} \right,
+eff= \begin{cases} 
+\frac{P_{percent}}{E},&if P_{percent}< 0.5 \\
+\frac{P_{percent}}{2\times E},&if P_{percent}\geq 0.5 \\
+% \frac{P_{percent}}{2\times E},&\mbox{if} P_{percent}\mbox{\geq0.5} 
+\end{cases},
 \label{eq:circleError}
 \end{equation}
- \begin{equation}
-P_{percent}  = L_{stroke} /P_{ellipse} 
-\label{eq:ErrorArea}
-\end{equation}
-where $E$ is the error computed by Equation(\ref{eq:circleFit}), $L_{stroke}$ total length of stroke and $P_{ellipse} $ is the perimeter of the estimated ellipse.
-If $eff$ is more than threshold $th_{Ellipse}$\footnote{By trial and error best threshold found was $th_{Ellipse}=0.2$} then the stroke is segmented as an ellipse otherwise the system proceeds to the next step.
+ where $E$ is the error computed by Equation (\ref{eq:circleFit}), and 
+\[
+P_{percent}  = L_{stroke} /P_{ellipse}, 
+%\label{eq:ErrorArea}
+\]
+ where $L_{stroke}$ total length of stroke and $P_{ellipse} $ is the perimeter of the estimated ellipse.
+If $eff$ has a value more than threshold $th_{Ellipse}$\footnote{By trial and error best threshold found was $th_{Ellipse}=0.2$} then the stroke is segmented as an ellipse otherwise the system proceeds to the next step.
 
@@
 	 \includegraphics[scale=0.7]{images/pso1.jpg}			
-	\caption{DPSO algorithm encoding}% a) Possible dominate point b) Particle encoding  
+	\caption{DPSO algorithm encoding} The swarm particle encoding where 1 represent the dominant points or segment points (i.e points are drawn as a big plus).
+Length L is the distance between a stroke point and the line connecting two successive segment points.
+%represent the length   % a) Possible dominate point b) Particle encoding
 	\label{fig:pso1}
@@
 Two DPSO algorithms are used to generate different segmentations for each stroke the user draws.
-For each segmentation generated an error is evaluated.
+For each segmentation generated, an error is evaluated.
 The segmentation with the minimum error value is chosen as the best stroke segmentation (Figure \ref{fig:segblock}).
@@
 
-Therefore, the system represents each particle $P_i$ by $P_i = \left\{ {p_{ij} \left| {j = 1,2 \cdots N} \right.} \right\}$ where $p_{ij}$ has only two values a)1 ($p_{ij}=1$); means that this point ($j$) is a dominate point, or b) 0($p_{ij}=0$) means this point ($j$) is not a dominate point.
+Therefore, the system represents each particle $P_i$ by $P_i = \left\{ {p_{ij} \left| {j = 1,2 \cdots N} \right.} \right\}$ where $p_{ij}$ has only two values a)1 ($p_{ij}=1$); means that this point ($j$) is a dominate point, or b) 0 ($p_{ij}=0$) means this point ($j$) is not a dominate point.
 Figure\ref{fig:pso1} shows a particle encoding in the DPSO system.
@@
 	\end{description}
-\subsubsection{\textit{Polygon approximation\textsl{AlgS1}}}
+\subsubsection{\textit{Polygon Approximation Algorithm \textsl{(AlgS1)}}}
 %\\
@@
 \end{equation} where the arc $\widehat{x_ix_j}$ is defined as the consecutive set of points from point $x_i$ to point $x_{j}$ as in $x_i,x_{i+1} \cdots,x_j$.
-The line $\overline{x_i x_j} $ as the straight line connecting point $x_i$ to point $x_j$, $M$ is the number of dominant points in this solution as generated by the swarm algorithm.
+The line $\overline{x_i x_j}$ is defined as the straight line connecting point $x_i$ to point $x_j$ and $M$ is the number of dominant points in this solution as generated by the swarm algorithm.
 The error $e ( \widehat{x_ix_j},\overline{x_i x_j})$ is computed as the sum of squared perpendicular distance from every point along the arc $\widehat{x_ix_j}$ to the line $\overline{x_i x_j}$.
-The fitness is computed using equation %\ref{eq:fitnessSwarm1}
+The fitness is computed using the following equation %\ref{eq:fitnessSwarm1}
 \begin{equation}
 \max fitness(p_i ) = \left\{ {\begin{array}{*{20}c}
-   { - E/\varepsilon N} & {ifE > \varepsilon ,}  \\
+   {\frac{-E}{\varepsilon N}} & {ifE > \varepsilon ,}  \\
+
    {D/\sum\limits_{j = 1}^N {p_{ij} } } & {otherwise}  \\

### 2009-05-27 trunk/paper2/Paper2.tex
change to the segmentation and modification

@@
 \usepackage{subfigure}
+\usepackage{array}
+\setlength{\extrarowheight}{0.3cm}
 
@@
 %\end{abstract}
-
+\pagenumbering{arabic}
 \section{Introduction}
@@
    P(i)\Leftarrow 
-\{
+\left \{
 \begin{array}{c} 
@@
 \label{eq:descrite}
-\end{array}\}
+\end{array}\right.
 \end{equation}
@@
 eff= \begin{cases} 
-\frac{P_{percent}}{E},&if P_{percent}< 0.5 \\
-\frac{P_{percent}}{2\times E},&if P_{percent}\geq 0.5 \\
+\displaystyle \frac{P_{percent}}{E},&if P_{percent}< 0.5 \\ \medskip
+\displaystyle  \frac{P_{percent}}{2\times E},&if P_{percent}\geq 0.5 \\
 % \frac{P_{percent}}{2\times E},&\mbox{if} P_{percent}\mbox{\geq0.5} 
@@
 %\\
-The approximation error is computed by the equation \ref{eq:ErrorSwarm1}. The graphical meaning of the error is shown in Figure\ref{fig:LabelsPSO}.
+The approximation error is computed by the equation \ref{eq:ErrorSwarm1}. The graphical meaning of the error is shown in Figure\ref{fig:pso1}.
 \begin{equation}
@@
 The fitness is computed using the following equation %\ref{eq:fitnessSwarm1}
+
 \begin{equation}
-\max fitness(p_i ) = \left\{ {\begin{array}{*{20}c}
-   {\frac{-E}{\varepsilon N}} & {ifE > \varepsilon ,}  \\
+\max fitness(p_i ) = \left\{ {\begin{array}{*{30}l}
+   {\displaystyle \frac{-E}{\varepsilon N}} & {ifE > \varepsilon }  \\
 
-   {D/\sum\limits_{j = 1}^N {p_{ij} } } & {otherwise}  \\
-\end{array}} \right.
+   { \frac {D}{\sum\limits_{j = 1}^N {p_{ij} } }} & {otherwise}  \\
+\end{array}} \right.,
 \label{eq:fitnessSwarm1}
-\end{equation} where $N$ is the number of points in the stroke, $D$ is the number of points in the solution that was previously labeled as a possible dominant point ($P_{pd}$), $E$ is the computed error and $\varepsilon$ is the error threshold.
-It should be noticed that when the error is larger than the threshold $\varepsilon$ the fitness is given a -ve value to lower the fitness value of the solution.
-Otherwise the system favors the lower number of vertices.
+\end{equation} where $N$ is the number of points in the stroke, $D$ is the number of points in the solution that were previously labeled as possible dominant points ($P_{pd}$), $E$ is the computed error and $\varepsilon$ is the error threshold.
+It should be noticed that when the error is larger than the threshold $\varepsilon$ the fitness is given a $-ve$ value to lower the fitness value of the solution.
+Otherwise, the system favors the least number of vertices.
 
- \subsubsection{\textit{Hybrid Fitting \textsl{AlgS2}}}
-The algorithm has the same problem formulation but different fitness and error functions are used.
+ \subsubsection{\textit{Hybrid Fitting Algorithm \textsl{(AlgS2)}}}
+The \textsl{(AlgS2)} algorithm has the same problem formulation as \textsl{(AlgS1)}.
+However, it has different fitness and error functions.
 An attempt is made to fit each segment into a line or circular arc.
 The errors of both circle and line fit estimations are computed for each segment $S_i=\widehat{x_ix_j}$.
-The approximation with the lower error value is chosen as the final approximation of this segment $S_i$\cite{CruveDivisionSwarm}.
-The sum of the approximation errors of all segments is the total error of the particle.
+The approximation with the lowest error value is chosen as the final approximation of this segment $S_i$\cite{CruveDivisionSwarm}.
+The sum of the approximation errors of all segments is considered the total error of the particle.
 The total error of the particle is computed by equation %\ref{eq:errorSwarm2}
  \begin{equation}
-E=\sum\nolimits_{i = 0}^M e(D_i) 
+E=\sum\nolimits_{i = 0}^{N_s} e(D_i) 
 \label{eq:errorSwarm2}
-\end{equation}where $M$ is the number of segments in the solution as generated by the swarm algorithm, $D_i$ is the minimum approximation error of curve and line approximations $min(d_c,d_l)$ where $d_c$ is the circle approximation error and $d_l$ is the line approximation error as computed by \cite{CruveDivisionSwarm}.
+\end{equation}where $N_s$ is the number of segments in the solution as generated by the swarm algorithm, $D_i$ is the minimum approximation error of curve and line approximations $min(d_c,d_l)$ where $d_c$ is the circle approximation error and $d_l$ is the line approximation error as computed by \cite{CruveDivisionSwarm}.
 The fitness is computed by the equation %\ref{eq:fitnessSwarm2}
 \begin{equation}
-\max fitness(P_i ) = \frac{1}{{E \times M^k }}
-\label{eq:fitnessSwarm2}
-\end{equation} where $E$ is the error and $M$ is the number of segments and $k$ is a parameter tweaked to get minimum number of segments.
-The larger $k$ is, the more effect the number of segments will have.
-For our system, $k$ is selected to be 0.5\cite{CruveDivisionSwarm}.
+\max fitness(P_i ) = \frac{1}{{E \times {N_s}^k }}
+\label{eq:fitnessSwarm2}%$N_s$ is the number of segments and
+\end{equation} where $E$ is the error and $k$ is a parameter tweaked to get minimum number of segments.
+The larger $k$ is, the higher the effect the number of segments will have.
+For our system, $k$ is selected to be 0.5 \cite{CruveDivisionSwarm}.
 
@@
 	\item For each particle $P_i$ each dominant point $P_{ij}$ is checked to find if it was labeled before as a \textit{possible dominant point} $P_{pd}$ (computed as in section \ref{Prepross}).
-If it was not labeled the point $P_{ij}$ is moved to the nearest labeled point.
+If it was not labeled, the point $P_{ij}$ is moved to the nearest labeled point.
 This ensures that all of the points generated by the DPSO are possible dominant points $P_{pd}$.
 	\item The particles are tested to make sure that the distance between every two successive dominate point is larger than $min_D$, where $min_D$ is 5\% of the total length of the stroke.
-Otherwise, one of these points are removed based on the error cased by its removal.
+Otherwise, one of these points is removed based on the error cased by its removal.
 	\item Each two successive segments are test to determine if they can be merged into a single segment.

### 2009-05-30 paper2/Paper2.tex
The beginig of satureday changes i did on thursday in ERI

@@
 Two versions of PSO segmentation algorithms were implemented.
-Experiments were conducted on three different dataset, simple presentation (Hs-DB), Electrical Design (EL-DB) and Logic Design (LD-DB) symbols.
+Experiments were conducted on three different dataset, simple presentation (Hs-DS), Electrical Design (EL-DS) and Logic Design (LD-DS) symbols.
 Results show that both PSO algorithms produce better segmentation than most current systems.
 This segmentation improvement contributed in getting more accurate recognition results on all the datasets tested.
-The results prove that the system is effective specially in datasets with low diversity in shapes (i.e.
-shapes are very simillar) as in Logic Design dataset (LD-DS).
+The results prove that the system is effective specially in datasets with low diversity in shapes, i.e., shapes are very simillar, as in Logic Design dataset (LD-DS).
 %The results prove the effectiveness of the proposed system on various dataset, especially datasets with low shape diversity as in Logic Design dataset (LD-DS).% effectiveness of the proposed system on various dataset, especially datasets with low shape diversity as in Logic Design dataset (LD-DS).
@@
 Two different segmentation algorithms were evaluated and results show that both algorithms improve the final recognition results.
-Experiments were conducted on three different dataset, simple presentation (Hs-DB), Electrical Design (EL-DB) and Logic Design (LD-DB) symbols.
+Experiments were conducted on three different dataset, simple presentation (Hs-DS), Electrical Design (EL-DS) and Logic Design (LD-DS) symbols.
 Results show that the effectiveness of the proposed system on various dataset especially on datasets with low shape diversity as in Logic Design dataset (LD-DS).
@@
 
-Sketch recognition can be based on either Pattern Recognition (PR) method or Artifical Intelligance (AI) based method.
+Sketch recognition can be applied either using  Pattern Recognition (PR) based method or Artifical Intelligance (AI) based method.
 In AI based methods, Symbols are represented using hierarchical levels of simpler geometric shapes or primitives and spatial relations between them.
@@
 strokes to break them into geometric primitives.
-Recognition can then is treated as an AI problem which can be solved using methods like sub graph matching, bayes network, and constrain statisfaction.
-A Hierarchical shape descriptions was widely used in sketch systems \cite{HierarchicalParsing7,Ladder30,AlvaradoFreedom42,SketchRead2007}.
+Recognition can then is treated as an AI problem which can be solved using methods like sub graph matching, bayes belief network, and constrain statisfaction.
+A Hierarchical shape descriptions was widely used in sketch systems \cite{HierarchicalParsing7,SketchRead2007}.
+%Ladder30,AlvaradoFreedom42
 
-Recognition based on AI methods is applied in \cite{SketchRead2007}.
-The principle problem in these systems is the computational complexity of the matching between shape description and geometric primitives produced in the segmentation process.
-Natural sketch are difficult to segment and group to create reliable fragment of geometric primitives.
+Recognition based on AI methods as applied in \cite{SketchRead2007} has the principle problem of computational complexity of the matching between shape description and geometric primitives produced in the segmentation process.
+Natural sketch are difficult to be segmented and grouped to create reliable fragment of geometric primitives.
 This is partially due to the increase of noise and other phenomena such as overlapped, over traced strokes and sharing of strokes between different symbols.
 Since these systems are dependable on the segmentation results, they either perform recognition using undependable segmentation results or consider different ways to segment the input sketch.
+This is because they assume that each stroke plays a role in the structure of the shape.
 %The role of which each stroke plays in the structure of the sketched symbol is used to determine shapes in stroke based methods.  
@@
 %role of which each stroke plays in the structure of the sketched symbol is used to determine shapes 
-Systems that uses Pattern Recognition methods compute properties that represent different symbols then use a classification method to recognize symbols.
+%.
+discribing the role  how user draws individal storkes  properties that discribe how users draw individual strokes in drawing the symbol.
+to .
+The identification of individual stroke is used to determine  The symbols are recognized by identifing the individual strokes.
+In PR based methods, systems compute properties that discribe symbols then use a classification method to recognize symbols.
 The properties used can either be based on gesture properties, global shape properties or image based properties.
-Gestures properties are set of features that represent how the user draws a stroke not how it look like.
-In gestures properties approcah, the systems uses the role of each stroke in the structure of the symbol determine the shapes.
-System that uses gestures properties assume that users draw symbols in specific order and style \cite{gestureexample12}.
+Gesture properties are set of features that represent how the user draws a symbol not how it looks like (e.g., identify a regtangle when a user draws a L corner).
+In gesture properties approcah, systems identify symbols by using gesutre properties of some of its  underlying strokes.
+System that uses gestures properties assume that users draw symbols in specific order and style \cite{gestureexample12,aideddesgin22}.
 In global shape properties approach, systems investigate general shape properties and its underlying strokes.
-Using shape properties (i.e.
-ratio of area of convex hull to area of shape) relaxes the assumption and role of each stroke in the symbol but it does not truly represent the appearance of the symbol\cite{DiagramOfflineConvexHull}.
+Using shape properties (e.g., ratio of the area of convex hull to the area of shape) relaxes the assumption and the role of each stroke in the symbol but it does not truly represent the appearance or fine details of the symbol \cite{DiagramOfflineConvexHull,Cali63}.
 Unlike shape properties based methods, image based methods disregard individual strokes and focus on the overall appearance of shapes.
-Image based methods are used in \cite{Oltmans07} where they computed a code book of 1000 features to recognize symbols.
+These system convertes the shapes into image and extracte image based features (e.g.
+local gradient feature)\cite{Oltmans07,imagetrainable48}.
+In \cite{imagetrainable48}, they used four different distance measures to to match user symbol with the prototype database.
+The prototype of a shape is represented as an image of the symbol that is normalized and scaled.
+However, the problem is that the shape prototype must represent all transformation and variation the input shape can have.
+%Image based methods are used in \cite{Oltmans07,imagetrainable48} where they computed a code book of 1000 features to recognize symbols.
 
@@
 These systems assume that users will finish drawing a shape before moving to the next one \cite{Cali63,geometrydomain49}.
-This let them to temporally group strokes and then recognizes them based on their strokes or global shape properties.
-\cite{AlvaradoDigital} argued that this assumption does not hold in natural sketches.
+This assumption allows the system to temporally group strokes and then recognizes them based on their strokes or global shape properties.
+Alvarado \cite{AlvaradoDigital} argued that this assumption does not hold in natural sketches.
 
-In this paper we choose to implement a hybrid method between shape property, guesture and hierarchical (AI method) based approaches.
-Our method tries to avoid many of the problems faced by approaches based on individual guestures or global shape properties while trying to gain the additional information that the stroke based system use.
+In this paper, we choose to implement a hybrid method between shape property, guesture and hierarchical (AI method) based approaches.
+Our method attempts to avoid many of the problems faced by approaches based on individual guestures or global shape properties while trying to gain the additional information that the stroke based system uses.
 Firstly, the system uses particle swarm optimization (PSO) to optimally segment user's strokes.
@@
 \end{equation}
-where $p_{ij}$ represent the $jth$ particle in the $ith$ agent and $v_{ij}$ is the velocity of the $jth$ particle in the $ith$ agent.
+%where $p_{ij}$ represent the $jth$ particle in the $ith$ agent and $v_{ij}$ is the velocity of the $jth$ particle in the $ith$ agent.
+where $p_{ij}$ represent the $j$th dimension in the $i$th particle and $v_{ij}$ is the velocity of the $j$th dimension in the $i$th particle.
  %Equation [\ref{eq:Swarm}] shows how velocity and direction of each particle are computed
  \begin{equation}
-v_{ij}  = v_{ij}  + c_1 r_1 (lbest_{ij}  - p_{ij} ) + c_2 r_2 (gbest_{ij}  - p_{ij} )
+v_{ij}  = v_{ij} \omega + c_1 r_1 (lbest_{ij}  - p_{ij} ) + c_2 r_2 (gbest_{ij}  - p_{ij} ),
 \label{eq:Swarm}
 \end{equation}
- where $lbest_{ij}$ is the local best particle, $gbest_{ij}$ is the global best particle, $r_1$ \& $r_2$ are random variables and $c_1$ \& $c_2$ are the swarm system variables.
- After each iteration the global best $g_{best}$ particle and the agent local best $l_{best}$ particle are evaluated based on the maximum fitness functions of all particles in the solution space.
+ where  $\omega$ is the inertia weight parameter which controls the tradeoff between exploration and exploitation,  $lbest_{ij}$ is the local best particle, $gbest_{ij}$ is the global best particle, $r_1$ \& $r_2$ are random variables and $c_1$ \& $c_2$ are the swarm acceleration parameters.
+
+ After each iteration the global best $g_{best}$ particle and the agent local best particle $l_{best}$are evaluated based on the maximum fitness functions of all particles in the solution space.
 The solution is found after achieving a specific number of iteration or after an error threshold is achieved.
-Equation \ref{eq:descrite} is used to change the general swarm algorithm into binary particle (\textit{Discrete Particle Swarm Algorithm DPSO}) which handles particle values of either $0$ or $1$ \cite{PSODisceret}.
+Equation \ref{eq:descrite} is used to change the general swarm algorithm into binary particle (\textit{Discrete Particle Swarm Algorithm DPSO}).
+The \textit{DPSO} handles particle values of either $0$ or $1$ \cite{PSODisceret}.
 \begin{equation}
    P(i)\Leftarrow 
-\{
+\left \{
 \begin{array}{c} 
@@
 
-0 \quad \quad if\quad r_{3}<p_{i} 
+0 \quad \quad if\quad r_{3}<p_{i}, 
 \label{eq:descrite}
-\end{array}\}
+\end{array}\right.
 \end{equation}
@@
 This leads to a nearly constant time difference between samples while the pen is moving and large difference while the pen is stationary (Figure \ref{fig:timediff}).
-Contrary to speed information, where the users draw with variant speeds, the speed information has a lot of noise in the data.
-Similarly, direction information is used as it provides better distinctive maxima for the corners than the curvature information \cite{meanshift10}.% Figure \ref{fig:curvatures} shows the direction and curvature graphs for the stroke drawn in Figure \ref{fig:orignalStroke}
+Contrary to time difference information, the speed information, where the users draw with variant speeds, the speed information has a lot of noise in the data (Figure \ref{fig:speed2Distance}).
+%Similarly, direction information is used as it provides better distinctive maxima for the corners than the curvature information \cite{meanshift10}.% Figure \ref{fig:curvatures} shows the direction and curvature graphs for the stroke drawn in Figure \ref{fig:orignalStroke}
+
  \begin{figure}[h]
@@
 \end{equation}
-
- where $N$ is number of points in the stroke, $a,b$ are the length of ellipse axes, $x_0$ \& $y_0$ are the coordinates of the center point, $x_i$ \& $y_i$ are the coordinates of point $i$ in the stroke.
+where $N$ is number of points in the stroke, $a$ \& $b$ are the length of ellipse axes, $x_0$ \& $y_0$ are the coordinates of the center point and $x_i$ \& $y_i$ are the coordinates of point $i$ in the stroke.
 A list of new values for $x_0$ , $y_0$ ,$a$ and $b$ are generated randomly from the older values with small increments after each loop.
-After few iteration, the final fit error of the estimated ellipse is reported.
+After few iterations, the final fit error of the estimated ellipse is reported.
 Another measure is used to compute the efficiency of the final estimated ellipse.
 Equation \ref{eq:circleError} ensures that the drawn percentage of ellipse is considered.
-This eliminates fitting a  line into very large ellipse but leaves small ellipses to be fitted as a partial or full ellipse.
+This eliminates fitting a  line into very large ellipse but leaves small ellipses to be fitted as a partial or a full ellipse.
 
  \begin{equation}
-eff= (P_{percent}/E)
+%eff= (P_{percent}/E)
+%eff = \left\{ {\begin{array}{*{20}c}
+%   {\frac{{P_{percent} }}{E}} & {ifP_{percent}  < 0.5}  \\
+%   {\frac{{P_{percent} }}{{E \times 2}}} & {ifP_{percent}  \ge 0.5}  \\
+%\end{array}} \right,
+eff= \begin{cases} 
+\displaystyle \frac{P_{percent}}{E},&if P_{percent}< 0.5 \\ \medskip
+\displaystyle  \frac{P_{percent}}{2\times E},&if P_{percent}\geq 0.5 \\
+% \frac{P_{percent}}{2\times E},&\mbox{if} P_{percent}\mbox{\geq0.5} 
+\end{cases},
 \label{eq:circleError}
 \end{equation}
- \begin{equation}
-P_{percent}  = L_{stroke} /P_{ellipse} 
-\label{eq:ErrorArea}
-\end{equation}
-where $E$ is the error computed by Equation(\ref{eq:circleFit}), $L_{stroke}$ total length of stroke and $P_{ellipse} $ is the perimeter of the estimated ellipse.
-If $eff$ is more than threshold $th_{Ellipse}$\footnote{By trial and error best threshold found was $th_{Ellipse}=0.2$} then the stroke is segmented as an ellipse otherwise the system proceeds to the next step.
+ where $E$ is the error computed by Equation (\ref{eq:circleFit}), and 
+\[
+P_{percent}  = L_{stroke} /P_{ellipse}, 
+%\label{eq:ErrorArea}
+\]
+ where $L_{stroke}$ total length of stroke and $P_{ellipse} $ is the perimeter of the estimated ellipse.
+If $eff$ has a value more than threshold $th_{Ellipse}$\footnote{By trial and error best threshold found was $th_{Ellipse}=0.2$} then the stroke is segmented as an ellipse otherwise the system proceeds to the next step.
 
@@
 	\end{description}
-\subsubsection{\textit{Polygon approximation\textsl{AlgS1}}}
+\subsubsection{\textit{Polygon Approximation Algorithm \textsl{(AlgS1)}}}
 %\\
-The approximation error is computed by the equation \ref{eq:ErrorSwarm1}. The graphical meaning of the error is shown in Figure\ref{fig:LabelsPSO}.
+The approximation error is computed by the equation \ref{eq:ErrorSwarm1}. The graphical meaning of the error is shown in Figure\ref{fig:pso1}.
 \begin{equation}
@@
 \end{equation} where the arc $\widehat{x_ix_j}$ is defined as the consecutive set of points from point $x_i$ to point $x_{j}$ as in $x_i,x_{i+1} \cdots,x_j$.
-The line $\overline{x_i x_j} $ as the straight line connecting point $x_i$ to point $x_j$, $M$ is the number of dominant points in this solution as generated by the swarm algorithm.
+The line $\overline{x_i x_j}$ is defined as the straight line connecting point $x_i$ to point $x_j$ and $M$ is the number of dominant points in this solution as generated by the swarm algorithm.
 The error $e ( \widehat{x_ix_j},\overline{x_i x_j})$ is computed as the sum of squared perpendicular distance from every point along the arc $\widehat{x_ix_j}$ to the line $\overline{x_i x_j}$.
-The fitness is computed using equation %\ref{eq:fitnessSwarm1}
+The fitness is computed using the following equation %\ref{eq:fitnessSwarm1}
+
 \begin{equation}
-\max fitness(p_i ) = \left\{ {\begin{array}{*{20}c}
-   { - E/\varepsilon N} & {ifE > \varepsilon ,}  \\
-   {D/\sum\limits_{j = 1}^N {p_{ij} } } & {otherwise}  \\
-\end{array}} \right.
+\max fitness(p_i ) = \left\{ {\begin{array}{*{30}l}
+   {\displaystyle \frac{-E}{\varepsilon N}} & {ifE > \varepsilon }  \\
+
+   { \frac {D}{\sum\limits_{j = 1}^N {p_{ij} } }} & {otherwise}  \\
+\end{array}} \right.,
 \label{eq:fitnessSwarm1}
-\end{equation} where $N$ is the number of points in the stroke, $D$ is the number of points in the solution that was previously labeled as a possible dominant point ($P_{pd}$), $E$ is the computed error and $\varepsilon$ is the error threshold.
-It should be noticed that when the error is larger than the threshold $\varepsilon$ the fitness is given a -ve value to lower the fitness value of the solution.
-Otherwise the system favors the lower number of vertices.
+\end{equation} where $N$ is the number of points in the stroke, $D$ is the number of points in the solution that were previously labeled as possible dominant points ($P_{pd}$), $E$ is the computed error and $\varepsilon$ is the error threshold.
+It should be noticed that when the error is larger than the threshold $\varepsilon$ the fitness is given a $-ve$ value to lower the fitness value of the solution.
+Otherwise, the system favors the least number of vertices.
 
- \subsubsection{\textit{Hybrid Fitting \textsl{AlgS2}}}
-The algorithm has the same problem formulation but different fitness and error functions are used.
+ \subsubsection{\textit{Hybrid Fitting Algorithm \textsl{(AlgS2)}}}
+The \textsl{(AlgS2)} algorithm has the same problem formulation as \textsl{(AlgS1)}.
+However, it has different fitness and error functions.
 An attempt is made to fit each segment into a line or circular arc.
 The errors of both circle and line fit estimations are computed for each segment $S_i=\widehat{x_ix_j}$.
-The approximation with the lower error value is chosen as the final approximation of this segment $S_i$\cite{CruveDivisionSwarm}.
-The sum of the approximation errors of all segments is the total error of the particle.
+The approximation with the lowest error value is chosen as the final approximation of this segment $S_i$\cite{CruveDivisionSwarm}.
+The sum of the approximation errors of all segments is considered the total error of the particle.
 The total error of the particle is computed by equation %\ref{eq:errorSwarm2}
  \begin{equation}
-E=\sum\nolimits_{i = 0}^M e(D_i) 
+E=\sum\nolimits_{i = 0}^{N_s} e(D_i) 
 \label{eq:errorSwarm2}
-\end{equation}where $M$ is the number of segments in the solution as generated by the swarm algorithm, $D_i$ is the minimum approximation error of curve and line approximations $min(d_c,d_l)$ where $d_c$ is the circle approximation error and $d_l$ is the line approximation error as computed by \cite{CruveDivisionSwarm}.
+\end{equation}where $N_s$ is the number of segments in the solution as generated by the swarm algorithm, $D_i$ is the minimum approximation error of curve and line approximations $min(d_c,d_l)$ where $d_c$ is the circle approximation error and $d_l$ is the line approximation error as computed by \cite{CruveDivisionSwarm}.
 The fitness is computed by the equation %\ref{eq:fitnessSwarm2}
 \begin{equation}
-\max fitness(P_i ) = \frac{1}{{E \times M^k }}
-\label{eq:fitnessSwarm2}
-\end{equation} where $E$ is the error and $M$ is the number of segments and $k$ is a parameter tweaked to get minimum number of segments.
-The larger $k$ is, the more effect the number of segments will have.
-For our system, $k$ is selected to be 0.5\cite{CruveDivisionSwarm}.
+\max fitness(P_i ) = \frac{1}{{E \times {N_s}^k }}
+\label{eq:fitnessSwarm2}%$N_s$ is the number of segments and
+\end{equation} where $E$ is the error and $k$ is a parameter tweaked to get minimum number of segments.
+The larger $k$ is, the higher the effect the number of segments will have.
+For our system, $k$ is selected to be 0.5 \cite{CruveDivisionSwarm}.
 
@@
 	\item For each particle $P_i$ each dominant point $P_{ij}$ is checked to find if it was labeled before as a \textit{possible dominant point} $P_{pd}$ (computed as in section \ref{Prepross}).
-If it was not labeled the point $P_{ij}$ is moved to the nearest labeled point.
+If it was not labeled, the point $P_{ij}$ is moved to the nearest labeled point.
 This ensures that all of the points generated by the DPSO are possible dominant points $P_{pd}$.
 	\item The particles are tested to make sure that the distance between every two successive dominate point is larger than $min_D$, where $min_D$ is 5\% of the total length of the stroke.
-Otherwise, one of these points are removed based on the error cased by its removal.
-	\item Each two successive segments are test to determine if they can be merged into a single segment.
-If merging the two segments have the same or smaller the error, the two segments are merged.
+Otherwise, two temperary particles are evaluated by removing either point from the current particle.
+After that, one of these points is removed based on the error cased by its removal.
+	\item Each two successive segments are tested to determine if they can be merged into a single segment.
+If merging the two segments results in segmentation that has an error value less than or equal to the error before merging, the two segments are merged.
 This step ensures that the final segmentation of the system has minimum number of segments.
+	% the same or smaller the error
 \end{enumerate}
@@
 %\begin{description}
-\subsubsection{Structural and geometrical Features(FS1)}
-  Features defines the structure of the geometrical symbol.  
+\subsubsection{Structural and geometrical Features (FS1)}
+  They are features defines the structure of the geometrical symbol such as;   
 %\begin{description}
@@
  \begin{itemize}
-	 \item \emph{Segments:} Number of segments in the symbol.
-	 \item \emph{Strokes:} Number of strokes or partial strokes that created the symbols.  
-		\item  \emph{Primitives:} Number of primitives in the symbol. The feature helps when identifying             symbols with mixed geometric primitives like cylinders and callouts.  
-		\item \emph{Curves:} Number of curves or ellipses in the symbol. 
-		\item \emph{Lines:} Number of lines in the symbol. 
-		\item \emph{Perpendicular} lines Number of perpendicular lines. 
-		\item \emph{Parallel} lines Number of parallel lines. 
-		\item \emph{Intersections} Number of intersection between lines and curves. 
-		\item \emph{T intersections} Number of T intersections. 
-		\item \emph{L intersections} Number of L intersections. 
-		\item \emph{X intersections} Number of X intersections.
-
+	 \item \emph{Segments:} The number of segments in the symbol.
+	 \item \emph{Strokes:} The number of strokes or partial strokes that created the symbols.  
+		\item  \emph{Primitives:} The umber of primitives in the symbol. This feature helps when identifying  symbols with mixed geometric primitives such as cylinders and callouts.  
+		\item \emph{Curves:} The number of curves or ellipses in the symbol. 
+		\item \emph{Lines:} The number of lines in the symbol. 
+		\item \emph{Perpendicular}\emph{lines:} The number of perpendicular lines. 
+		\item \emph{Parallel}\emph{lines:} The number of parallel lines. 
+		\item \emph{Intersections:} The number of intersection between lines and curves. 
+		\item \emph{T intersections:} The number of T intersections. 
+		\item \emph{L intersections:} The number of L intersections. 
+		\item \emph{X intersections:} The number of X intersections.	
 		%\item[No. of holes] Number of holes in the symbols. 
 \end{itemize}
-\subsubsection{Rubine Feature Set (FS2)}
-  Features introduced by Rubine\cite{gestureexample12} for single stroke gestures. Feature such as length of stroke, smoothness of symbol and properties of the bounding box of the symbol.
+\subsubsection{Rubine Features (FS2)}
+  These features introduced by Rubine \cite{gestureexample12} for single stroke gestures.
+Such as length of stroke, cosine of starting angle, smoothness of symbol and the properties of the  bounding box of the symbol.
 
 \subsubsection{Statistical Features (FS3)} 
+These features uses moments and statistical properties such as;
 \begin{itemize}
-	\item \emph{Zernike moments } Zernike moments of order n (10 or 20).  \cite{HeloiseBeautification}. 
+	\item \emph{Zernike moments:} The zernike moments are orthogonal moments which define the distribution of point in the input space.
+They are are invariant to both reflection and rotation of the input shape \cite{ZerMomentOrthogonal}.
+%The zernike moment were used in \cite{HeloiseBeautification} t.
 \end{itemize}
 \subsubsection{Composite Features (FS4)}
- Features that are composites of different geometrical and statistical symbol. 
+ These features are composites of different geometrical and statistical properties of the symbol. 
 	\begin{itemize}
-\item \emph{Size Ratio} Ratio between width to height of the symbol.
-	\item \emph{Ink density} Compute the density of points inside its bounding box\cite{GeometryAndDomain102}.   
- 	\item \emph{Convex Hull Area} Area of convex hull with respect to area of bounding box of symbol.
-	\item \emph{Convex Hull Perimeter} Perimeter of convex hull with respect to total length of symbol.
-		\item \emph{Mean Centroidal radius} The Mean of the centroidal radius which is the distance from each point in the symbol to the center of gravity.
-
-	\item \emph{Mean Time difference} The mean of the time difference between each two successive points in the symbol. %Different strokes are appended to construct a single path.
+\item \emph{Size Ratio:} The ratio between width and height of the symbol.
+	\item \emph{Ink density:} The density of points inside its bounding box\cite{GeometryAndDomain102}.   
+ 	\item \emph{Convex Hull Area:} The area of convex hull with respect to area of bounding box of symbol.
+	\item \emph{Convex Hull Perimeter:} The perimeter of convex hull with respect to total length of symbol.
+		\item \emph{Mean Centroidal radius:} The Mean of the centroidal radius is the distance from each point in the symbol to the center of gravity.
+	\item \emph{Mean Time difference:} The mean of the time difference between each two successive points in the symbol. %Different strokes are appended to construct a single path.
 	%\item [Ra]
   \end{itemize}  
-
 %\end{description}
 %The features used consist of Rubine feature set,  of order 21 , ink density as well as some structural and spatial information like number of perpendiculars lines ,number of parallel lines and number of different types of primitives in each symbol.
-After computing the features the symbol is introduced to the classifier.
+After computing the features, the symbol is introduced to the classifier as a feature vector.
 The system uses Support Vector Machine (SVM) classifier with Linear kernel \cite{libsvm}.
-An OVO classifier structure (object versus object) is used to handle multi-class.% A validation algorithm is used to choose the Kernel variables parameters $c, \gamma$.
+An OVO classifier structure (object versus object) is used to handle multi-class problem.% A validation algorithm is used to choose the Kernel variables parameters $c, \gamma$.
 
@@
 \label{sec:Experiments} 
+%Figure \ref{fig:symbolSet} shows a set of the shapes used in this data set. 
 In our experiment we used three different datasets.
-First dataset was collected by Hse and Newton\cite{HeloiseBeautification} (Hs-DB).
+The first dataset (Figure \ref{fig:HsSet}) was collected by Hse and Newton\cite{HeloiseBeautification} (Hs-DS).
 The data are drawn by 16 users each of them was asked to draw 13 shapes from 30 to 50 times.
-Figure \ref{fig:symbolSet} shows a set of the shapes used in the data set.
 We collected the second and third datasets.
-Figure \ref{fig:ELsymbolSet} and \ref{fig:LogicsymbolSet} shows the symbols gathered from 7 different users in electrical (EL-DB) and logic design (LD-DB) domain respectively.
+Figure \ref{fig:ELsymbolSet} and \ref{fig:LogicsymbolSet} shows the symbols gathered from 7 different users in electrical (EL-DS) and logic design (LD-DS) domain respectively.
 Table \ref{tab:datasets} shows a  comparison between the three datasets used in our experiments.
@@
 The accuracy is computed as the number of correctly recognized samples divided by the total number of samples in the test.
+\begin{figure}
+\centering
+
+\fbox{ \parbox{5cm}{% 
+		\includegraphics[scale=0.5]{images/symbolSet.PNG}	}}
+		\caption{The Hs-DS Symbol Set} The dataset consist of simple presentation symbols.
+		\label{fig:HsSet} 
+\end{figure}
 
  \begin{figure}
-\centering 
-\label{fig:symbolSet}
-\fbox{ \parbox{5cm}{% 
-		\includegraphics[scale=0.5]{images/symbolSet.PNG}	}}
-		\caption{The Hs-DB Symbol Set}.
+			\subfigure[The Digital Design Symbols Dataset(LD-DS)]{\label{fig:LogicsymbolSet} \includegraphics[scale=0.5]{images/logicSet.jpg}}
+			\hfill
+			\subfigure[The Electrical Symbols Dataset(EL-DS)]
+			 {\label{fig:ELsymbolSet}\includegraphics[scale=0.45]{images/EelectImage.jpg}}
+%			 \label{fig:ELsymbolSet} 
+%			 \fbox{ 
+%			 \parbox{4cm}{\includegraphics[scale=0.5]{images/EelectImage.jpg}	}}}
+
+	\caption{The datasets that were manually collected}  a)  The (LD-DS) dataset consist of symbols that are visually simillar (i.e.
+has low shape diversity).
+b) The (EL-DS) dataset consist of complex toplogical symbols.
 \end{figure}
-\begin{figure}
-\centering 
-\label{fig:LogicsymbolSet}
 
-		\includegraphics[scale=0.5]{images/logicSet.jpg}	 
-		\caption{The Digital Design Symbols Dataset(LD-DB)} Symbols used in the dataset.
-\end{figure}
-\begin{figure}
-\centering 
-\label{fig:ELsymbolSet}
-\fbox{ \parbox{4.5cm}{% 
-		\includegraphics[scale=0.5]{images/EelectImage.jpg}	}}
-		\caption{The Electrical Symbols Dataset(EL-DB)} .
-\end{figure}
+%\begin{figure}
+%\centering 
+%\label{fig:LogicsymbolSet}
+% 
+%		\includegraphics[scale=0.5]{images/logicSet.jpg}	 
+%		\caption{The Digital Design Symbols Dataset(LD-DS)} The dataset consist of symbols that are visually simillar (i.e. has low shape diversity). 
+%\end{figure}
+%\begin{figure}
+%\centering 
+%\label{fig:ELsymbolSet}
+%\fbox{ \parbox{4.5cm}{% 
+%		\includegraphics[scale=0.5]{images/EelectImage.jpg}	}}
+%		\caption{The Electrical Symbols Dataset(EL-DS)} . The dataset consist of complex toplogical symbols.
+%\end{figure}
 
@@
  \hline
- & Hs-DB &  EL-DB & LD-DB \\ \hline
+ & Hs-DS &  EL-DS & LD-DS \\ \hline
 No of samples  &  7791 & 2764 & 1859 \\ \hline
@@
 Splits  & 5 random splits  & 5 random splits  & 5 random splits  \\ \hline
-Examples & Figure \ref{fig:symbolSet}  & Figure \ref{fig:ELsymbolSet}  & Figure \ref{fig:LogicsymbolSet}   \\ \hline
+Examples & Figure \ref{fig:HsSet}  & Figure \ref{fig:ELsymbolSet}  & Figure \ref{fig:LogicsymbolSet}   \\ \hline
 \end{tabular}
@@
 	 \includegraphics[scale=0.3]{images/swarmAgentTest.jpg}
-	 	\caption{Swarm testing} Segmentation Error Vs. Number of agents
+	 	\caption{Swarm testing} Segmentation Error Vs. Number of particles
 	 	\label{fig:swarmtesting}
@@
 	 \includegraphics[scale=0.4]{images/testAlg.jpg}
-	 	\caption{Algorithm comparison on HS-DB} The recognition rate of different algorithms. 
+	 	\caption{Algorithm comparison on HS-DS} The recognition rate of different algorithms. 
 	 	\label{fig:test1}
@@
 	 \includegraphics[scale=0.4]{images/LDAlg.jpg}
-	 	\caption{Algorithm comparison on LD-DB} The recognition rate of different algorithms. 
+	 	\caption{Algorithm comparison on LD-DS} The recognition rate of different algorithms. 
 	 	\label{fig:testLD}
@@
 	 \includegraphics[scale=0.4]{images/AlgEL.jpg}
-	 	\caption{Algorithm comparison on EL-DB} The recognition rate of different algorithms. 
+	 	\caption{Algorithm comparison on EL-DS} The recognition rate of different algorithms. 
 	 	\label{fig:testEL}
@@
 		\includegraphics[scale=0.5]{images/testsym.jpg}
-	\caption{Symbols Comparison on HS-DB} The effect of symbol complexity.  %The graph shows the recognition rate of each symbol using different algorithms. 
+	\caption{Symbols Comparison on HS-DS} The effect of symbol complexity.  %The graph shows the recognition rate of each symbol using different algorithms. 
 	\label{fig:test2}
@@
 		\includegraphics[scale=0.5]{images/LDsymbols.jpg}
-	\caption{Symbols Comparison on LD-DB symbols}.  %The graph shows the recognition rate of each symbol using different algorithms. 
+	\caption{Symbols Comparison on LD-DS symbols}.  %The graph shows the recognition rate of each symbol using different algorithms. 
 	\label{fig:LDtest2}
@@
  	%}
-	\caption{Confusion Matrix of Hs-DB}
+	\caption{Confusion Matrix of Hs-DS}
 	\label{tab:ConfusionMatrix}
@@
 %  	%}
-% 	\caption{Confusion Matrix of DL-DB}
+% 	\caption{Confusion Matrix of DL-DS}
 % 	\label{tab:ConfusionMatrix}
@@
 		\includegraphics[scale=0.5]{images/featuresdifference.jpg}
-	\caption{Feature Comparison on Hs-DB accuracy} The effect of different features on accuracy.  %The graph shows the recognition rate of each symbol using different algorithms. 
+	\caption{Feature Comparison on Hs-DS accuracy} The effect of different features on accuracy.  %The graph shows the recognition rate of each symbol using different algorithms. 
 	\label{fig:testFeaturesAll}
@@
 		\includegraphics[scale=0.34]{images/featELc.jpg}
-	\caption{Feature Comparison on EL-DB} The effect of different features on accuracy.  %The graph shows the recognition rate of each symbol using different algorithms. 
+	\caption{Feature Comparison on EL-DS} The effect of different features on accuracy.  %The graph shows the recognition rate of each symbol using different algorithms. 
 	\label{fig:ELtestFeaturesAll}
@@
 		\includegraphics[scale=0.4]{images/featureFS1only.jpg}
-	\caption{The effect of segmentation on Hs-DB accuracy} %The effect of symbol complexity.  %The graph shows the recognition rate of each symbol using different algorithms. 
+	\caption{The effect of segmentation on Hs-DS accuracy} %The effect of symbol complexity.  %The graph shows the recognition rate of each symbol using different algorithms. 
 	\label{fig:testFeatonly}

### 2009-05-30 paper2/Paper2.tex
the same

@@
 The data are drawn by 16 users each of them was asked to draw 13 shapes from 30 to 50 times.
-We collected the second and third datasets.
-Figure \ref{fig:ELsymbolSet} and \ref{fig:LogicsymbolSet} shows the symbols gathered from 7 different users in electrical (EL-DS) and logic design (LD-DS) domain respectively.
+Figure \ref{fig:ELsymbolSet} and \ref{fig:LogicsymbolSet} shows the symbols we manually gathered from 7 different users in electrical (EL-DS) and logic design (LD-DS) domain respectively.
 Table \ref{tab:datasets} shows a  comparison between the three datasets used in our experiments.
-The datasets were divided into training set and test set.
+The datasets were divided into training and test sets.
 Five different splits  for the training and test data are generated from each dataset.
@@
 \label{sec:AlgExp}
-A set of tests were performed on the system to enhance non ellipse fitting algorithm.
-The effect of number of agents, the error thresholds and various other parameters were investigated.
-The result of these tests are shown in Figure \ref{fig:swarmtesting} and   \ref{fig:swarmtesting2}.
+A set of tests were performed on the system to enhance non-ellipse fitting algorithms.
+The effect of the number of particles, the error thresholds and various other parameters were investigated.
+The results of these tests are shown in Figures \ref{fig:swarmtesting} and   \ref{fig:swarmtesting2}.
 Figure \ref{fig:swarmtesting} shows that as the number of agents increases the error value decrease (error of the segmentation algorithm).
-The final number of agents and maximum swarm iteration used in the system was based on a trade off between error achieved versus the computation time.
-Other algorithm parameters such as $c_1$,$c_2$ that were mentioned in Section\ref{sec:ParticleSwarmAlgorithm} are tested using similar tests (Figure \ref{fig:swarmtesting} and \ref{fig:swarmtesting2}).
+The final number of particles and maximum swarm iterations used in the system was based on a trade off between error achieved versus the computation time.
+The effect of other algorithm parameters such as $c_1$, $c_2$ (mentioned in Section\ref{sec:ParticleSwarmAlgorithm}) is tested using similar tests (Figure \ref{fig:swarmtesting}, \ref{fig:swarmtesting2}).
 
@@
 We performed several experiments to evaluate the presented recognition system.
-Firstly we tested recognition accuracy of shapes in the data set with both algorithms (\textsl{AlgS1} and \textsl{AlgS2}).
+Firstly, we tested recognition accuracy of shapes in the data set with both algorithms (\textsl{AlgS1} and \textsl{AlgS2}).
 We also implemented the segmentation algorithm described in \cite{earlyprocess} (\textsl{Alg3}) to use as reference to compare it with our swarm algorithms.
@@
 The two swarm algorithms were tested with and without the ellipse fitting module.
-The ellipse detection module appears to be superior.
+Combining the ellipse detection module improves the performance.
 The result shows that combining both (\textsl{AlgS1} and \textsl{AlgS2}) out perform any single algorithm.
@@
 \end{figure*}  
-The second experiment we implemented was to investigate the effect of symbol complexity and type on the recognition rate.
-Figure \ref{fig:test2} and \ref{fig:LDtest2} shows the achieved accuracy of each symbol by our algorithms compared to \cite{earlyprocess} (\textsl{Alg3}).
-It is clearly noted that symbols that have only line segments achieve higher accuracy rate than other symbols.
-The results indicate that algorithm \textsl{AlgS1} achieve better performance than algorithm \textsl{AlgS2} in the symbols that consist only of lines.
+The second experiment we implemented was to investigate the effect of symbol complexity and its type on the recognition rate.
+Figure \ref{fig:test2} and \ref{fig:LDtest2} show the achieved accuracy of each symbol by our algorithms compared to \cite{earlyprocess} (\textsl{Alg3}).
+It is clearly noted that symbols that have only line segments achieve higher accuracy rate than other symbols (such as square, parallegram).
+The results also indicate that algorithm \textsl{AlgS1} achieves better performance than algorithm \textsl{AlgS2} in the symbols that consist only of lines.
 This is understandable because algorithm \textsl{AlgS1} divides strokes into line segments only but \textsl{AlgS2} is able to divide strokes into lines and curves based on the minimum error of the segment itself.
-Figure \ref{fig:LDtest2} shows that in logic design symbols \textsl{AlgS2} achieved better performance than \textsl{Alg3} \textsl{AlgS1} this emphases that complex shapes that are combination of lines and arc are better segmented using \textsl{AlgS2}.
-Algorithm \textsl{Alg3} gives good performance as long as the symbols consist of combination of lines and curves, if the stroke consists of only curve or lines the algorithm may lead to wrong segmentation result.
-This is because the system divides the stroke first to line segments then tries to decide if each segment can be represent better as a curve unlike algorithm \textsl{AlgS2} where the curve segments are tested while choosing the best segmentation.
+Figure \ref{fig:LDtest2} shows that in logic design symbols \textsl{AlgS2} achieved better performance than \textsl{Alg3} and  \textsl{AlgS1}.
+This proves that for complex shapes that are combination of lines and arcs are \textsl{AlgS2} acheives better segmentation.
+Algorithm \textsl{Alg3} gives good performance as long as the symbols consist of combination of lines and curves, if the stroke consists of curve or lines only the algorithm may lead to wrong segmentation result.
+This is because the system divides the stroke first into line segments then tries to decide whether each segment can be represented better as a curve.
+However, in \textsl{AlgS2} the curve segments are tested while choosing the best segmentation.
 Combining both algorithm \textsl{AlgS1} and \textsl{AlgS2} improved the recognition rate of all symbols.
 The penalty for this improved performance is the computational time required to run both swarm algorithms.
-Table \ref{tab:ConfusionMatrix} shows the confusion matrix of symbols.
-The table shows that errors are only between two sets of symbols moon with triangles and trapezoid with triangles.
+Table \ref{tab:ConfusionMatrix} shows the confusion matrix for the symbols.
+The table shows that errors are only between two sets of symbols: moon with triangles and trapezoid with triangles.
 This observation is understandable as the symbols are visually similar but the system must be able to differentiate between them.
-This lead us to the next experiment to choose best set of feature to recognize symbols.
+This leads us to the next experiment to choose best set of feature to recognize symbols.
 \begin{table*}
@@
 \label{sec:featexp}
-Different feature sets are tested to determine the best features that can used in sketch and symbol recognition.
-Figure\ref{fig:testFeaturesAll} shows the result of different feature sets from the basic four sets (\textbf{FS1,FS2,FS3,FS4}).
-Result shows that (\textbf{FS2}) Rubine features \cite{gestureexample12} gives the worst results when used alone without any other features.
+%Results show that (\textbf{FS2}) Rubine features \cite{gestureexample12} gives the worst results when used alone without any other features.
+Different feature sets are tested to determine the best features that can be used in sketch and symbol recognition.
+Figure\ref{fig:testFeaturesAll} shows the result of using different feature sets (from the basic four sets \textbf{FS1,FS2,FS3,FS4}).
+Results show that  geometerical features (\textbf{FS1}) and  Rubine features (\textbf{FS2}) \cite{gestureexample12} gives the low results when used alone without any other features.
 This is because they are mainly computed for single stroke gestures and fare bad in multi-stroke symbols \cite{compareFeaturSVM}.
-Features \textbf{(FS4)} gives good results but it is improved by adding structural features \textbf{(FS1)}.
+Features \textbf{(FS4)} gives good results but it is improved by adding geometerical structural features \textbf{(FS1)}.
  \begin{figure*}
@@
 \end{figure*}  
- We could not test the result of the segmentation algorithm directly due to the fact that the correct segmentation is highly ambiguous.
+ We could not test the result of the segmentation algorithm alone due to the fact that the correct segmentation is hightly qualitative.
 It is only based on what the user intends to draw while sketching the symbol.
@@
 \label{ConclusionandFutureWork}
+
 This paper presented a new approach to sketch recognition using PSO.
-The system uses both speed and curvature information which help improving the \textit{DPSO}.
+The system uses both speed and curvature information the \textit{DPSO}.
 It is noted that the \textit{DPSO} in general generates an optimized stroke segmentation which improves the final recognition rate.
-The tradeoff between accuracy achieved and time complexity must be further investigated to achieve better results.
+The tradeoff between accuracy achieved and time complexity can be further investigated to achieve better results.
 The final recognition is based on both global shape and hierarchical stroke properties.
 Adding stroke and segment geometrical features improved the final recognition.
-Global shape and moment properties was used to prevent the need to use graph matching which is computationally expensive.
+Global shape and moment properties were used to prevent the need to use graph matching which is computationally expensive \cite{SketchRead2007}.
 The system was tested on 13 different symbols and achieved an overall accuracy of 89.5\%.

### 2009-05-31 paper2/Paper2.tex
finish the paper

@@
 The second experiment we implemented was to investigate the effect of symbol complexity and its type on the recognition rate.
-Figure \ref{fig:test2} and \ref{fig:LDtest2} show the achieved accuracy of each symbol by our algorithms compared to \cite{earlyprocess} (\textsl{Alg3}).
+Figure \ref{fig:test2} and \ref{fig:LDtest2} show the achieved accuracy of each symbol by our algorithms compared to (\textsl{Alg3}) \cite{earlyprocess}.
 It is clearly noted that symbols that have only line segments achieve higher accuracy rate than other symbols (such as square, parallegram).
-The results also indicate that algorithm \textsl{AlgS1} achieves better performance than algorithm \textsl{AlgS2} in the symbols that consist only of lines.
+The results also indicate that algorithm \textsl{AlgS1} achieves better performance than algorithm \textsl{AlgS2} in the symbols that consists only of lines.
 This is understandable because algorithm \textsl{AlgS1} divides strokes into line segments only but \textsl{AlgS2} is able to divide strokes into lines and curves based on the minimum error of the segment itself.
-Figure \ref{fig:LDtest2} shows that in logic design symbols \textsl{AlgS2} achieved better performance than \textsl{Alg3} and  \textsl{AlgS1}.
+Figure \ref{fig:LDtest2} shows that in logic design (LD-DS) \textsl{AlgS2} achieved better performance than \textsl{Alg3} and  \textsl{AlgS1}.
 This proves that for complex shapes that are combination of lines and arcs are \textsl{AlgS2} acheives better segmentation.
@@
 The penalty for this improved performance is the computational time required to run both swarm algorithms.
-Table \ref{tab:ConfusionMatrix} shows the confusion matrix for the symbols.
-The table shows that errors are only between two sets of symbols: moon with triangles and trapezoid with triangles.
+Table \ref{tab:ConfusionMatrix} shows the confusion matrix for the (HS-DS) symbols.
+The table shows that errors occurs mostly between two sets of symbols: moon with triangles and trapezoid with triangles.
 This observation is understandable as the symbols are visually similar but the system must be able to differentiate between them.
-This leads us to the next experiment to choose best set of feature to recognize symbols.
+This leads us to the next experiment to choose the best set of feature to recognize symbols.
 \begin{table*}
@@
 Different feature sets are tested to determine the best features that can be used in sketch and symbol recognition.
-Figure\ref{fig:testFeaturesAll} shows the result of using different feature sets (from the basic four sets \textbf{FS1,FS2,FS3,FS4}).
+Figure\ref{fig:testFeaturesAll} shows the result of using different feature sets (from the basic four sets \textbf{FS1,FS2,FS3,FS4} in Section \ref{sec:Recognition}).
 Results show that  geometerical features (\textbf{FS1}) and  Rubine features (\textbf{FS2}) \cite{gestureexample12} gives the low results when used alone without any other features.
 This is because they are mainly computed for single stroke gestures and fare bad in multi-stroke symbols \cite{compareFeaturSVM}.
-Features \textbf{(FS4)} gives good results but it is improved by adding geometerical structural features \textbf{(FS1)}.
+Features \textbf{(FS4)} gives good results and it is furthur improved by adding geometerical structural features \textbf{(FS1)}.
  \begin{figure*}
@@
 This paper presented a new approach to sketch recognition using PSO.
-The system uses both speed and curvature information the \textit{DPSO}.
+The system uses both sthe peed and curvature information as input to the \textit{DPSO}.
 It is noted that the \textit{DPSO} in general generates an optimized stroke segmentation which improves the final recognition rate.

### 2009-05-31 paper2/Paper2.tex

@@
 }
-\caption[Datasets Comparisons]{Datasets Comparisons} 
+\caption[Datasets Comparisons]{Table compares between datasets used in experiments in terms of number of samples, number of samples per category, etc...} 
 \label{tab:datasets}
@@
 
-\subsection{Algorithms comparisons}
-\label{sec:AlgExp}
 A set of tests were performed on the system to enhance non-ellipse fitting algorithms.
@@
 	 \includegraphics[scale=0.3]{images/swarmAgentTest.jpg}
-	 	\caption{Swarm testing} Segmentation Error Vs. Number of particles
+	 	\caption{Segmentation Error Vs. Number of particles} The error decrease as the number of particle increase.  
 	 	\label{fig:swarmtesting}
@@
 	 \includegraphics[scale=0.4]{images/swarmParamterC.jpg}
-	 	\caption{Swarm testing} Segmentation Error Vs. PSO parameters.
+	 	\caption{Segmentation Error Vs. PSO parameters.} The behaviour of segmentation error versus c1. 
 	 	\label{fig:swarmtesting2}
@@
 \end{figure} 
-
+\subsection{Algorithms comparisons}
+\label{sec:AlgExp}
 We performed several experiments to evaluate the presented recognition system.

### 2009-05-31 trunk/bib/old bib/bib.tex
initall add

@article{ SketchRec1, 
  title ="  Sketch Recongition for software engineering  ", 
 YEAR = " 2003  ",
Month= "1",
 author =" Levet Burak  Kara   and  ",
}
@article{ sketchint2, 
  title ="  sketch interfaces: Towards more human interface design  ", 
 YEAR = " 2001  ",
Month= "5",
 author =" James Landay   and  Brad A. Myers   and  ",
}
@article{ Improving 3, 
  title ="  Improving sketching by utilizing Haptic feedback  ", 
 YEAR = " 2002  ",
Month= "5",
 author =" Chris Raymaekers   and  Gert Vansichem   and  Frank van Reeth   and  ",
}
@article{ early proc4, 
  title ="  early processing for sketch understanding  ", 
 YEAR = " 2001  ",
Month= "12",
 author =" Tevfik Metin sezgin   and  Thomas Stahovich   and  Randall Davis   and  ",
}
@article{ recognizin5, 
  title ="  recognizing multi stroke symbols  ", 
 YEAR = " 2002  ",
Month= "12",
 author =" Levet Burak  Kara   and  Thomas Stahovich   and  Tolga kurtoglu    and  Chris calhoun   and  ",
}
@article{ interpreti6, 
  title ="  interpreting schematic sketches using physical reasoning  ", 
 YEAR = " 2002  ",
Month= "12",
 author =" Thomas Stahovich   and  Tolga kurtoglu    and  ",
}
@article{ a sketch b8, 
  title ="  a sketch based interface for the design and analysis of a simple vibratory mechanical system  ", 
 YEAR = " 2004  ",
Month= "4",
 author =" Levet Burak  Kara   and  Thomas Stahovich   and  Leslie Gennari   and  ",
}
@article{ Experience9, 
  title ="  Experiences developing and E-whiteboard based circuit designer  ", 
 YEAR = " 2003  ",
Month= "1",
 author =" Ray Liu   and  Lisa  Wong    and  John Grundy   and  ",
}
@article{ recognitio10, 
  title ="  recognition of freehand sketch using mean shift  ", 
 YEAR = " 2003  ",
Month= "4",
 author =" bo yu   and  ",
}
@article{ a frame wo11, 
  title ="  a frame work for multi domain sketch recognition  ", 
 YEAR = " 2002  ",
Month= "5",
 author =" Christine Alverado   and  Randall Davis   and  Michael oltmans   and  ",
}
@article{ Specifying12, 
  title ="  Specifying Gesture by example  ", 
 YEAR = " 1991  ",
Month= "4",
 author =" Dean Rubine    and  ",
}
@article{ Hand drawn13, 
  title ="  Hand drawn  maps for robot navigation  ", 
 YEAR = " 2002  ",
Month= "4",
 author =" Majorie Skubic    and   Pascal matsakis   and  Sam bilsard   and  ",
}
@article{ Tahuti : a14, 
  title ="  Tahuti : a geometrical Sketch Recognition  ", 
 YEAR = " 2002  ",
Month= "4",
 author =" Tracy hammond   and  Randall Davis   and  ",
}
@article{ Functions 15, 
  title ="  Functions in design idea generation meetings  ", 
 YEAR = " 2002  ",
Month= "4",
 author =" Remko van der Lugt   and  ",
}
@article{ Toward com16, 
  title ="  Toward computational model of sketching   ", 
 YEAR = " 2002  ",
Month= "4",
 author =" Kenneth D. Forbus   and  Jeffery M.  Usher   and  Ronald W. Ferguson   and  ",
}
@article{ A domain  17, 
  title ="  A domain  independent system for sketch recognition   ", 
 YEAR = " 2001  ",
Month= "4",
 author =" bo yu   and  Shijie Cai   and  ",
}
@article{ A cognitiv18, 
  title ="  A cognitive approach to sketch understanding  ", 
 author =" Kenneth D. Forbus   and  Ronald W. Ferguson   and  ",
}
@article{ Sketching 19, 
  title ="  Sketching for knowledge capture a progress report  ", 
 author =" Kenneth D. Forbus   and  Jeffery M.  Usher   and  ",
}
@article{ A Sketch i20, 
  title ="  A Sketch interface for mobile robots  ", 
 YEAR = " 2003  ",
Month= "4",
 author =" Majorie Skubic    and  Craig Bailey   and  George Chornis   and  ",
}
@article{ Robust Ske21, 
  title ="  Robust Sketched symbol fragmentation using templates  ", 
 YEAR = " 2004  ",
Month= "12",
 author =" Heloise Hse   and  Michael Shilman   and  A. Richard Newton   and  ",
}
@article{ Computer A22, 
  title ="  Computer Aided sketching to capture preliminary design  ", 
 YEAR = " 2002  ",
Month= "4",
 author =" Beryl Plimmer   and  Mark Apperley   and  ",
}
@article{ handling o24, 
  title ="  handling over traced strokes in hand drawn sketches  ", 
 YEAR = " 2004  ",
Month= "4",
 author =" Tevfik Metin sezgin   and  Randall Davis   and  ",
}
@article{ Fluid Sket25, 
  title ="  Fluid Sketches : continuous Recognition and morphing of simple hand drawn shapes  ", 
 YEAR = " 2000  ",
Month= "4",
 author =" James Avro    and  Kevin Novins   and  ",
}
@article{ statical v26, 
  title ="  statical visual language models for ink parsing.  ", 
 YEAR = " 2000  ",
Month= "4",
 author =" Michael Shilman   and  A. Richard Newton   and  Hannna Psula    and  Stuart Russell   and  ",
}
@article{ speed and 27, 
  title ="  speed and sketching for multimodal design  ", 
 YEAR = " 2004  ",
Month= "4",
 author =" Randall Davis   and  Aron Adler   and  ",
}
@article{ Generic an28, 
  title ="  Generic and HMM based approaches to freehand sketch recognition  ", 
 YEAR = " 2003  ",
Month= "4",
 author =" Tevfik Metin sezgin   and  ",
}
@article{ SketchRead29, 
  title ="  SketchRead : A Multi domain sketch recognition engine  ", 
 YEAR = " 2004  ",
Month= "4",
 author =" Christine Alverado   and  Randall Davis   and  ",
}
@article{ LADDER: a 30, 
  title ="  LADDER: a language to describe drawing display and editing in sketch recognition  ", 
 YEAR = " 2003  ",
Month= "4",
 author =" Tracy hammond   and  Randall Davis   and  ",
}
@article{ hierarchic7, 
  title ="  hierarchical parsing and recognition of hand sketch diagrams  ", 
 YEAR = " 2003  ",
Month= "4",
 author =" Levet Burak  Kara   and  Thomas Stahovich   and  ",
}
@article{ A domain d31, 
  title ="  A domain description language for sketch recognition  ", 
 YEAR = " 2003  ",
Month= "4",
 author =" ",
}
@article{ Three main23, 
  title ="  Three main concerns is sketch recognition an and approach to address them  ", 
 YEAR = " 2002  ",
Month= "4",
 author =" James V. Mahoney    and  Markus p. j. formhez   and  ",
}
@article{ learning s1, 
  title ="  learning structural descriptions form examples   ", 
 YEAR = " 2005  ",
Month= "8",
 author =" ",
}
@article{ Interactio45, 
  title ="  Interactional Context in Sketch Understanding  ", 
 YEAR = " 2005  ",
Month= "8",
 author =" Patrick G.T. Healey   and  Simon Garrod   and  Nick Fay   and  John Lee   and  Jon Oberlander   and  ",
}
@article{ Polygon Re31, 
  title ="  Polygon Recognition in Sketch- Based interfaces with immediate and continous feedback  ", 
 YEAR = " 2003  ",
Month= "11",
 author =" Peter Agar   and  Kevin Novins   and  ",
}
@article{ A Multi-Le32, 
  title ="  A Multi-Level Sketching Tool for pencil_and_paper animation   ", 
 YEAR = " 2002  ",
Month= "11",
 author =" Frank van Reeth   and  Fabian Di Fiore   and  ",
}
@article{ Mixing Ima33, 
  title ="  Mixing Images and Sketches for Retrieving Vector Drawings  ", 
 YEAR = " 2004  ",
Month= "11",
 author =" Alfredo Ferreira   and  Manuel J. Fonseca   and  Joaquim A. Jorge   and  Mario Ramalho   and  ",
}
@article{ Analysis o34, 
  title ="  Analysis of Engineering Drawing : State of Art and Challenges  ", 
 YEAR = " 2005  ",
Month= "8",
 author =" Karl Tombre   and  ",
}
@article{ sketch und36, 
  title ="  sketch understanding in design: overview of work at the mit ai lab  ", 
 YEAR = " 2002  ",
Month= "12",
 author =" Randall Davis   and  ",
}
@article{ Sketching 37, 
  title ="  Sketching for military courses of action diagrams  ", 
 YEAR = " 2003  ",
Month= "1",
 author =" Kenneth D. Forbus   and  Jeffery M.  Usher   and  Vernell Chapman   and  ",
}
@article{ functional38, 
  title ="  functional and formal reasonoing in architectural sketches  ", 
 YEAR = " 2005  ",
Month= "7",
 author =" Ellen Yi- Luen Do   and  ",
}
@article{ efficient 39, 
  title ="  efficient search space exploration for sketch recognition  ", 
 YEAR = " 2004  ",
Month= "9",
 author =" Tevfik Metin sezgin   and  Randall Davis   and  ",
}
@article{ Drawing ma40, 
  title ="  Drawing marks, acts, and reacts: toward a computational sketching interface for architectural design  ", 
 YEAR = " 2005  ",
Month= "7",
 author =" Ellen Yi- Luen Do   and  ",
}
@article{ incrementa41, 
  title ="  incremental sketch understanding for intention extraction in sketch-based user interfaces  ", 
 YEAR = " 2003  ",
Month= "10",
 author =" Yang Li    and  ",
}
@article{ preserving42, 
  title ="  preserving the freedom of paper in a computer based sketch tool  ", 
 YEAR = " 2001  ",
Month= "12",
 author =" Christine Alverado   and  Randall Davis   and  ",
}
@article{ sketching 43, 
  title ="  sketching user interfaces with visual patterns  ", 
 YEAR = " 2005  ",
Month= "8",
 author =" Manuel J. Fonseca   and  Joaquim A. Jorge   and  Anabela Caetano   and  Neri Goulart   and  ",
}
@article{ stretch-a-44, 
  title ="  stretch-a-sketch: a dynamic diagrammer  ", 
 YEAR = " 2005  ",
Month= "8",
 author =" Mark D Gross   and  ",
}
@article{ mathpad2 :46, 
  title ="  mathpad2 : a system for the creation and exploration of mathematical sketchs  ", 
 YEAR = " 2005  ",
Month= "8",
 author =" Joseph J. La Viola Jr.   and  Robert C. Zelezink   and  ",
}
@article{ agent base47, 
  title ="  agent based sketch recognition  ", 
 YEAR = " 2003  ",
Month= "8",
 author =" Graham Mackenzie   and  ",
}
@article{ an image-b48, 
  title ="  an image-based trainable symbol recognizer for sketch-base interfaces  ", 
 YEAR = " 2004  ",
Month= "9",
 author =" Levet Burak  Kara   and  Thomas Stahovich   and  ",
}
@article{ combining 49, 
  title ="  combining geometry and domain knolwdge to interpret hand drawn diagrams  ", 
 YEAR = " 2004  ",
Month= "12",
 author =" Levet Burak  Kara   and  Thomas Stahovich   and  Leslie Gennari   and  ",
}
@article{ robust rec50, 
  title ="  robust recognigion and interpretation of mechanical design sketches  ", 
 YEAR = " 1999  ",
Month= "10",
 author =" Christine Alverado   and  ",
}
@article{ sim-u-sket51, 
  title ="  sim-u-sketch: a sketch-based interface for simulink  ", 
 YEAR = " 2005  ",
Month= "7",
 author =" Levet Burak  Kara   and  Thomas Stahovich   and  ",
}
@article{ smart sket56, 
  title ="  smart sketchpad-an on line  graphics recognition system  ", 
 YEAR = " 2001  ",
Month= "11",
 author =" Liu Wenyin   and  Wenjie Qian   and  Rong Xiao   and  Xiangyu jin   and  ",
}
@article{ on line gr35, 
  title ="  on line graphics recognition state of the art   ", 
 YEAR = " 2003  ",
Month= "12",
 author =" Liu Wenyin   and  ",
}
@article{ etcha sket52, 
  title ="  etcha sketches : lessons learned from collecting sketch data  ", 
 YEAR = " 2004  ",
Month= "8",
 author =" Christine Alverado   and  Randall Davis   and  Michael oltmans   and  ",
}
@article{ sketch rec54, 
  title ="  sketch recognition with continuous feedback based on incremental intention extraction  ", 
 YEAR = " 2005  ",
Month= "1",
 author =" Junfeng Li   and  Xiwen zhang   and  Xiang Ao    and  Guozhong Dai   and  ",
}
@article{ HMM based 53, 
  title ="  HMM based efficient sketch recognition  ", 
 YEAR = " 2005  ",
Month= "1",
 author =" Tevfik Metin sezgin   and  Randall Davis   and  ",
}
@article{ a srg base57, 
  title ="  a srg based online composite graphic recognition strategy for sketch based user interface  ", 
 YEAR = " 2002  ",
Month= "1",
 author =" Rong Xiao   and  Xiangyu jin   and  zheng xing sun   and  bin bin peng   and  ",
}
@article{ Handling A55, 
  title ="  Handling Ambiguity and Errors: visual Languages for calligraphic interation   ", 
 YEAR = " 2001  ",
Month= "10",
 author =" Manuel J. Fonseca   and  Joaquim A. Jorge   and  joao  paulo pereira   and  ",
}
@article{ A Retarget58, 
  title ="  A Retargetable Framework for Interactive Diagram Recognition  ", 
 YEAR = " 2003  ",
Month= "1",
 author =" Edward h. lank   and  ",
}
@article{ A parsing 59, 
  title ="  A parsing technique for sketch recongition  ", 
 YEAR = " 2004  ",
Month= "3",
 author =" gennaro costagliola   and  vincenzo deufemia   and  giuseppe polese   and  michele risi   and  ",
}

### 2009-06-06 paper2/Paper2.tex
after all editin gis done

@@
 His system achieved good accuracy in simple shapes (square, ellipse,...etc) but did not perform well in complex shapes.
-A hybrid algorithm was introduced in \cite{earlyprocess} where different sets of segments are generated based on both curvature and speed dominant points, followed by choosing a segmentation with the least error from a generated hybrid set.
+A hybrid algorithm was introduced in \cite{earlyprocess}\footnote{Alg3 used in Section \ref{sec:Experiments} } where different sets of segments are generated based on both curvature and speed dominant points, followed by choosing a segmentation with the least error from a generated hybrid set.
 However, this system is limited to recognizing only specific simple geometric shapes with a set of low level recognizers.
@@
 \begin{center}
-\scalebox{0.9}{
- \begin{tabular}{|p{2cm}|p{1.5cm}|p{1.5cm}|p{1.5cm}|}
+\scalebox{0.8}{
+ \begin{tabular}{|p{3cm}|p{1.5cm}|p{1.5cm}|p{1.5cm}|}
  \hline
@@
 	\centering		
-	 \includegraphics[scale=0.3]{images/swarmAgentTest.jpg}
-	 	\caption{Segmentation Error Vs. Number of particles} The error decrease as the number of particle increase.  
-	 	\label{fig:swarmtesting}
+	 	\subfigure[Segmentation Error Vs. Number of particles]{	\label{fig:swarmtesting} \includegraphics[scale=0.3]{images/swarmAgentTest.jpg}}
+	 %	\caption{}
+
 	%\caption{Experiments results :} a)    b) c) 
-\end{figure}
+	%\end{figure}
 
-\begin{figure}
-	\centering		
-	 \includegraphics[scale=0.4]{images/swarmParamterC.jpg}
-	 	\caption{Segmentation Error Vs. PSO parameters.} The behaviour of segmentation error versus c1. 
-	 	\label{fig:swarmtesting2}
-	%\caption{Experiments results :} a)    b) c)
+%\begin{figure}
+	%\centering		
+		\subfigure[Segmentation Error Vs. PSO parameters.]{	\label{fig:swarmtesting2}  \includegraphics[scale=0.4]{images/swarmParamterC.jpg}}
+	 	\caption{}a) The error decrease as the number of particle increase.   b) The behaviour of segmentation error versus c1.
+
+	%\Segmentation Error Vs. PSO parameters.caption{Experiments results :} a)    b) c)
 
@@
 Combining the ellipse detection module improves the performance.
-The result shows that combining both (\textsl{AlgS1} and \textsl{AlgS2}) out perform any single algorithm.
+The result shows that (\textsl{AlgS2}) gives better performance than Alg3 and AlgS1 in all datasets.
+The results also shows that combineing  (\textsl{AlgS1} and \textsl{AlgS2} using minimum error) outperforme AlgS2 in Hs-Ds only.
+This behaviour is explained by noticing that features are computed based on the choosen segmentation.
+This means that different segmentation for a symbol generates different set of features used for recognition.
+This behaviour affect Ls-Ds and EL-Ds more than Hs-Ds due to the higher topological and structure symbols in both datasets than than of Hs-Ds.
+%   can be explained by the simplicity of the symbols in Hs-DS % and LD-Ds unlike EL-DS where AlgS2 is the best algorithm.
+This behaviour is explained by the structure of the symbols in each dataset.
+El-DS consist of highly topological and structure symbols that are usaually drawn using multiple small strokes.
+This leads to    %   out perform any single algorithm in HS-DS and LD-DS.
 %show that combining algoirthmsachieve higher accuracy than any single algoirthm.
@@
 
- \begin{figure}
+ \begin{figure}[h]
 	\centering		
-	 \includegraphics[scale=0.4]{images/testAlg.jpg}
-	 	\caption{Algorithm comparison on HS-DS} The recognition rate of different algorithms. 
-	 	\label{fig:test1}
+	\subfigure[Algorithm comparison on HS-DS]{\label{fig:test1} \includegraphics[scale=0.4]{images/testAlg.jpg}}
+	 %	\caption{} The recognition rate of different algorithms.
+
 	%\caption{Experiments results :} a)    b) c) 
-\end{figure}
+%\end{figure}
 
-\begin{figure}
-	\centering		
-	 \includegraphics[scale=0.4]{images/LDAlg.jpg}
-	 	\caption{Algorithm comparison on LD-DS} The recognition rate of different algorithms. 
-	 	\label{fig:testLD}
+%\begin{figure}[h]
+%	\centering		
+		\subfigure[Algorithm comparison on LD-DS]{ 	 	\label{fig:testLD} \includegraphics[scale=0.4]{images/LDAlg.jpg}}
+	% 	\caption{} The recognition rate of different algorithms.
+
 	%\caption{Experiments results :} a)    b) c) 
-\end{figure}
+%\end{figure}
 
-\begin{figure}
-	\centering		
-	 \includegraphics[scale=0.4]{images/AlgEL.jpg}
-	 	\caption{Algorithm comparison on EL-DS} The recognition rate of different algorithms. 
-	 	\label{fig:testEL}
+%\begin{figure}[h]
+%	\centering		
+		\subfigure[Algorithm comparison on EL-DS]{	 	\label{fig:testEL}  \includegraphics[scale=0.4]{images/AlgE.jpg}}
+	 	\caption{ The recognition rate of different algorithms. }
+
 	%\caption{Experiments results :} a)    b) c) 
@@
 \label{sec:ShapeComplexityExperiments}
-
 \begin{figure*}
 	\centering 
-		\includegraphics[scale=0.5]{images/testsym.jpg}
-	\caption{Symbols Comparison on HS-DS} The effect of symbol complexity.  %The graph shows the recognition rate of each symbol using different algorithms. 
-	\label{fig:test2}
-\end{figure*}
-
-\begin{figure*}
-	\centering
-		\includegraphics[scale=0.5]{images/LDsymbols.jpg}
-	\caption{Symbols Comparison on LD-DS symbols}.  %The graph shows the recognition rate of each symbol using different algorithms. 
-	\label{fig:LDtest2}
+		\subfigure[Symbols Comparison on HS-DS]{\label{fig:test2}
+		\includegraphics[scale=0.35]{images/testsym.jpg}}
+	%\caption{} The effect of symbol complexity.  %The graph shows the recognition rate of each symbol using different algorithms. 
+%\end{figure*}  
+%\begin{figure*}
+%	\centering
+\subfigure[Symbols Comparison on LD-DS symbols]{\label{fig:LDtest2}
+		\includegraphics[scale=0.5]{images/DLSymbols.jpg}}
+	%\caption{}.  %The graph shows the recognition rate of each symbol using different algorithms. 
+	\caption {Symbol Comparison} The accuarcy of achieved by each algorithm in each symbol.
+a) Low accuracy level appears in triangle, heart, and moon symbols.
+b) Low accuracy level appears only in symbols that are visually confusing (i.e Nor, OR, Nand gates).
 \end{figure*}  
@@
 	\centering
-		\includegraphics[scale=0.5]{images/featuresdifference.jpg}
-	\caption{Feature Comparison on Hs-DS accuracy} The effect of different features on accuracy.  %The graph shows the recognition rate of each symbol using different algorithms. 
-	\label{fig:testFeaturesAll}
-\end{figure*}
+	\subfigure[Feature Comparison on Hs-DS]{	\label{fig:testFeaturesAll} \includegraphics[scale=0.5]{images/featuresdifference.jpg}}
+    %The graph shows the recognition rate of each symbol using different algorithms.
 
- \begin{figure*}
-	\centering
-		\includegraphics[scale=0.34]{images/featELc.jpg}
-	\caption{Feature Comparison on EL-DS} The effect of different features on accuracy.  %The graph shows the recognition rate of each symbol using different algorithms. 
-	\label{fig:ELtestFeaturesAll}
+%\end{figure*}
+
+ %\begin{figure*}
+	%\centering
+
+	\subfigure[Feature Comparison on EL-DS]{	\label{fig:ELtestFeaturesAll}	\includegraphics[scale=0.34]{images/FeatEl.jpg}}
+	\caption{Different Features } The effect of different feature set on accuracy.
+FS4 and FS3 give good results in a) Hs-DS and b) EL-Ds.
+The results is improved by combining FS1 with FS2 and FS1 with FS4.
+%The graph shows the recognition rate of each symbol using different algorithms.
+
 \end{figure*}  
@@
 Hence those are the features we used to estimate the efficiency of the segmentation algorithm.
-The recognition accuracy using only the geometrical features is reported in Figure \ref{fig:testFeatonly} .
+The recognition accuracy using only the geometrical features is reported in Figure \ref{fig:testFeatonly}.
 The figure shows that \textsl{AlgS2} achieve better results.
@@
 	\centering
-		\includegraphics[scale=0.4]{images/featureFS1only.jpg}
+		\includegraphics[scale=0.3]{images/OnlyF1EL.jpg}
 	\caption{The effect of segmentation on Hs-DS accuracy} %The effect of symbol complexity.  %The graph shows the recognition rate of each symbol using different algorithms. 
@@
 	\caption{Segmentation outputs}  The output segmentation of different input strokes.
+a) Sample of single input stroke b) The single stroke after segmentation.
+c) An example of multiple stroke and their segmentation.
 %The figure shows sample of the output of the segmentation phase.
@@
 
-This paper presented a new approach to sketch recognition using PSO.
-The system uses both sthe peed and curvature information as input to the \textit{DPSO}.
+This paper presented a new approach to sketch recognition using DPSO.
+The system uses multiple premilary information i.e., speed and curvature information, as input to the \textit{DPSO}.
 It is noted that the \textit{DPSO} in general generates an optimized stroke segmentation which improves the final recognition rate.
@@
 Global shape and moment properties were used to prevent the need to use graph matching which is computationally expensive \cite{SketchRead2007}.
-The system was tested on 13 different symbols and achieved an overall accuracy of 89.5\%.
+The experiments evaluated two\textit{DPSO} different segmenatation algorithms (AlgS1 and AlgS2) on three different datasets (Hs-Ds, EL-Ds and LD-Ds).
+Results show that AlgS2 achieves better segmentation  and final accuracy than Alg3 \cite{earlyprocess} on the datasets used.
+The results proved that the system is easily expandable to more and more symbols as it dose not depend on low level primitive recogniziers but on high level symbol and geometrical featuers.
+The system achieved an overall improvement of more than 2\% over Alg3 in all datasets.
+%The system was tested on 13 different symbols and achieved an overall accuracy of 89.5\%.
 The system does not depend on low level recognizer but rather on a set of high level features.

### 2009-06-06 paper2/Paper2.tex

@@
 
-\author[ERI]{Maha El Meseery\footnote{Crossponding author}}
+\author[ERI]{Maha El Meseery\footnote{Corresponding author}}
 \author[ERI]{Mahmoud Fakhr El Din}
@@
 \address[FAC]{Computer Engineering Department,
-	Faculty of Engineering,Cairo University, Cairo, Egypt\\
-	magdafayek@gmail.com,ndarwish@ieee.org}
+	Faculty of Engineering, Cairo University, Cairo, Egypt\\
+	magdafayek@gmail.com, ndarwish@ieee.org}
 \begin{abstract}
@@
 This segmentation improvement contributed in getting more accurate recognition results on all the datasets tested.
-The results prove that the system is effective specially in datasets with low diversity in shapes, i.e., shapes are very simillar, as in Logic Design dataset (LD-DS).
+The results prove that the system is effective specially in datasets with low diversity in shapes, i.e., shapes are very similar, as in Logic Design dataset (LD-DS).
 %The results prove the effectiveness of the proposed system on various dataset, especially datasets with low shape diversity as in Logic Design dataset (LD-DS).% effectiveness of the proposed system on various dataset, especially datasets with low shape diversity as in Logic Design dataset (LD-DS).
-%Result show that the system is especially effictive on dataset with % sketch are divided  Users draw symbols and sketchs using a set of pen strokes.
+%Result show that the system is especially effective on dataset with % sketch are divided Users draw symbols and sketches using a set of pen strokes.
 %%Sketch recognition is defined as the process of identifying symbols that users draw using single or multiple strokes.
@@
 %% keywords here, in the form: keyword \sep keyword
-Symbol Recognition, Particle Swarm Optimization, Sketch Recognition, Geometerical Features, Polygon approximation
+Symbol Recognition, Particle Swarm Optimization, Sketch Recognition, Geometrical Features, Polygon approximation
 %% PACS codes here, in the form: \PACS code \sep code
@@
 Results show that the effectiveness of the proposed system on various dataset especially on datasets with low shape diversity as in Logic Design dataset (LD-DS).
-% sketch are divided  Users draw symbols and sketchs using a set of pen strokes.
+% sketch are divided Users draw symbols and sketches using a set of pen strokes.
 %%Sketch recognition is defined as the process of identifying symbols that users draw using single or multiple %strokes.
@@
 
-Sketch recognition can be applied either using  Pattern Recognition (PR) based method or Artifical Intelligance (AI) based method.
+Sketch recognition can be applied either using Pattern Recognition (PR) based method or Artificial Intelligence (AI) based method.
 In AI based methods, Symbols are represented using hierarchical levels of simpler geometric shapes or primitives and spatial relations between them.
-Systems that employ such representation must first pre-process.
-strokes to break them into geometric primitives.
-Recognition can then is treated as an AI problem which can be solved using methods like sub graph matching, bayes belief network, and constrain statisfaction.
+Systems that employ such representation must first pre-process strokes to break them into geometric primitives.
+Recognition can then is treated as an AI problem which can be solved using methods like sub graph matching, Bayes belief network, and constrain satisfaction.
 A Hierarchical shape descriptions was widely used in sketch systems \cite{HierarchicalParsing7,SketchRead2007}.
@@
 %.
-discribing the role  how user draws individal storkes  properties that discribe how users draw individual strokes in drawing the symbol.
-to .
-The identification of individual stroke is used to determine  The symbols are recognized by identifing the individual strokes.
-In PR based methods, systems compute properties that discribe symbols then use a classification method to recognize symbols.
+Describing the role  how user draws individual strokes  properties that describe how users draw individual strokes in drawing the symbol.
+to.
+The identification of individual stroke is used to determine The symbols are recognized by identifying the individual strokes.
+In PR based methods, systems compute properties that describe symbols then use a classification method to recognize symbols.
 The properties used can either be based on gesture properties, global shape properties or image based properties.
-Gesture properties are set of features that represent how the user draws a symbol not how it looks like (e.g., identify a regtangle when a user draws a L corner).
-In gesture properties approcah, systems identify symbols by using gesutre properties of some of its  underlying strokes.
+Gesture properties are set of features that represent how the user draws a symbol not how it looks like (e.g., identify a rectangle when a user draws a L corner).
+In gesture properties approach, systems identify symbols by using gesture properties of some of its underlying strokes.
 System that uses gestures properties assume that users draw symbols in specific order and style \cite{gestureexample12,aideddesgin22}.
@@
 Unlike shape properties based methods, image based methods disregard individual strokes and focus on the overall appearance of shapes.
-These system convertes the shapes into image and extracte image based features (e.g.
+These system convert the shapes into image and extracted image based features (e.g.
 local gradient feature)\cite{Oltmans07,imagetrainable48}.
-In \cite{imagetrainable48}, they used four different distance measures to to match user symbol with the prototype database.
+In \cite{imagetrainable48}, they used four different distance measures to match user symbol with the prototype database.
 The prototype of a shape is represented as an image of the symbol that is normalized and scaled.
@@
 
-In this paper, we choose to implement a hybrid method between shape property, guesture and hierarchical (AI method) based approaches.
-Our method attempts to avoid many of the problems faced by approaches based on individual guestures or global shape properties while trying to gain the additional information that the stroke based system uses.
+In this paper, we choose to implement a hybrid method between shape property, gesture and hierarchical (AI method) based approaches.
+Our method attempts to avoid many of the problems faced by approaches based on individual gestures or global shape properties while trying to gain the additional information that the stroke based system uses.
 Firstly, the system uses particle swarm optimization (PSO) to optimally segment user's strokes.
@@
 Systems that employ such representation must first pre-process strokes to break them into geometric primitives.
-Recognition can hen be treated as sub graph matching problem.
+Recognition can then be treated as sub graph matching problem.
 
@@
 Agar et al.
-\cite{polygonfeedback31} mention that the pointing device (e.g., mouse) sampling rate is the reason for this phenomena.
+\cite{polygonfeedback31} mentions that the pointing device (e.g., mouse) sampling rate is the reason for this phenomenon.
 The points are sampled at regular time intervals as long as the pointing device is moving.
@@
 The black points denote the location of the possible dominant points $P_{pd}$.
-Black rectangles represent regions $R_i$ that in a) are lower than the thershold.
-b) are above the thershold.
+Black rectangles represent regions $R_i$ that in a) Are lower than the threshold.
+b) Are above the threshold.
 	\label{fig:speed2Distance}
@@
 Otherwise, the stroke is passed to two further segmentation algorithms that divide the stroke to either lines or lines and curves.
-Hence we generate two different segmentations, the system chooses the segmentation with the minimum error.
+Hence we generate two different segmentations; the system chooses the segmentation with the minimum error.
 The following sections describe in details the ellipse detection algorithm and the two segmentation algorithms used to divide the stroke.
@@
 Otherwise, two temperary particles are evaluated by removing either point from the current particle.
-After that, one of these points is removed based on the error cased by its removal.
+After that, one of these points is removed based on the error caused by its removal.
 	\item Each two successive segments are tested to determine if they can be merged into a single segment.
@@
 	 \item \emph{Strokes:} The number of strokes or partial strokes that created the symbols.  
-		\item  \emph{Primitives:} The umber of primitives in the symbol. This feature helps when identifying  symbols with mixed geometric primitives such as cylinders and callouts.  
+		\item  \emph{Primitives:} The number of primitives in the symbol. This feature helps when identifying  symbols with mixed geometric primitives such as cylinders and callouts.  
 		\item \emph{Curves:} The number of curves or ellipses in the symbol. 
@@
 \subsubsection{Statistical Features (FS3)} 
-These features uses moments and statistical properties such as;
+These features use moments and statistical properties such as;
 \begin{itemize}
 	\item \emph{Zernike moments:} The zernike moments are orthogonal moments which define the distribution of point in the input space.
-They are are invariant to both reflection and rotation of the input shape \cite{ZerMomentOrthogonal}.
+They are invariant to both reflection and rotation of the input shape \cite{ZerMomentOrthogonal}.
 %The zernike moment were used in \cite{HeloiseBeautification} t.
@@
 
-	\caption{The datasets that were manually collected}  a)  The (LD-DS) dataset consist of symbols that are visually simillar (i.e.
+	\caption{The datasets that were manually collected}  a)  The (LD-DS) dataset consist of symbols that are visually similar (i.e.
 has low shape diversity).
-b) The (EL-DS) dataset consist of complex toplogical symbols.
+b) The (EL-DS) dataset consist of complex topological symbols.
 \end{figure}
@@
 Figure \ref{fig:swarmtesting} shows that as the number of agents increases the error value decrease (error of the segmentation algorithm).
-The final number of particles and maximum swarm iterations used in the system was based on a trade off between error achieved versus the computation time.
+The final number of particles and maximum swarm iterations used in the system was based on a tradeoff between error achieved versus the computation time.
 The effect of other algorithm parameters such as $c_1$, $c_2$ (mentioned in Section\ref{sec:ParticleSwarmAlgorithm}) is tested using similar tests (Figure \ref{fig:swarmtesting}, \ref{fig:swarmtesting2}).
@@
 		\subfigure[Segmentation Error Vs. PSO parameters.]{	\label{fig:swarmtesting2}  \includegraphics[scale=0.4]{images/swarmParamterC.jpg}}
-	 	\caption{}a) The error decrease as the number of particle increase.   b) The behaviour of segmentation error versus c1.
+	 	\caption{}a) The error decrease as the number of particle increase.   b) The behavior of segmentation error versus c1.
 
-	%\Segmentation Error Vs. PSO parameters.caption{Experiments results :} a)    b) c)
+	%\Segmentation Error Vs. PSO parameters. caption{Experiments results :} a)    b) c)
 
@@
 The result shows that (\textsl{AlgS2}) gives better performance than Alg3 and AlgS1 in all datasets.
-The results also shows that combineing  (\textsl{AlgS1} and \textsl{AlgS2} using minimum error) outperforme AlgS2 in Hs-Ds only.
-This behaviour is explained by noticing that features are computed based on the choosen segmentation.
+The results also shows that combining (\textsl{AlgS1} and \textsl{AlgS2} using minimum error) outperforms AlgS2 in Hs-Ds only.
+This behavior is explained by noticing that features are computed based on the chosen segmentation.
 This means that different segmentation for a symbol generates different set of features used for recognition.
-This behaviour affect Ls-Ds and EL-Ds more than Hs-Ds due to the higher topological and structure symbols in both datasets than than of Hs-Ds.
+This behavior affect Ls-Ds and EL-Ds more than Hs-Ds due to the higher topological and structure symbols in both datasets than of Hs-Ds.
 %   can be explained by the simplicity of the symbols in Hs-DS % and LD-Ds unlike EL-DS where AlgS2 is the best algorithm.
-This behaviour is explained by the structure of the symbols in each dataset.
-El-DS consist of highly topological and structure symbols that are usaually drawn using multiple small strokes.
+This behavior is explained by the structure of the symbols in each dataset.
+El-DS consist of highly topological and structure symbols that are usually drawn using multiple small strokes.
 This leads to    %   out perform any single algorithm in HS-DS and LD-DS.
-%show that combining algoirthmsachieve higher accuracy than any single algoirthm.
+%show that combining algorithms achieve higher accuracy than any single algorithm.
 %improves the results.% with both the \textit{DPSO} algorithms.% The results show that both \textit{DPSO} algorithms achieve better result than other algorithms.
@@
 	%\caption{}.  %The graph shows the recognition rate of each symbol using different algorithms. 
-	\caption {Symbol Comparison} The accuarcy of achieved by each algorithm in each symbol.
+	\caption {Symbol Comparison} The accuracy of achieved by each algorithm in each symbol.
 a) Low accuracy level appears in triangle, heart, and moon symbols.
@@
 Figure \ref{fig:LDtest2} shows that in logic design (LD-DS) \textsl{AlgS2} achieved better performance than \textsl{Alg3} and  \textsl{AlgS1}.
-This proves that for complex shapes that are combination of lines and arcs are \textsl{AlgS2} acheives better segmentation.
+This proves that for complex shapes that are combination of lines and arcs are \textsl{AlgS2} achieves better segmentation.
 Algorithm \textsl{Alg3} gives good performance as long as the symbols consist of combination of lines and curves, if the stroke consists of curve or lines only the algorithm may lead to wrong segmentation result.
@@
 Table \ref{tab:ConfusionMatrix} shows the confusion matrix for the (HS-DS) symbols.
-The table shows that errors occurs mostly between two sets of symbols: moon with triangles and trapezoid with triangles.
+The table shows that an error occurs mostly between two sets of symbols: moon with triangles and trapezoid with triangles.
 This observation is understandable as the symbols are visually similar but the system must be able to differentiate between them.
@@
 This is because they are mainly computed for single stroke gestures and fare bad in multi-stroke symbols \cite{compareFeaturSVM}.
-Features \textbf{(FS4)} gives good results and it is furthur improved by adding geometerical structural features \textbf{(FS1)}.
+Features \textbf{(FS4)} gives good results and it is further improved by adding geometerical structural features \textbf{(FS1)}.
  \begin{figure*}
@@
 	\subfigure[Feature Comparison on EL-DS]{	\label{fig:ELtestFeaturesAll}	\includegraphics[scale=0.34]{images/FeatEl.jpg}}
-	\caption{Different Features } The effect of different feature set on accuracy.
+	\caption{Different Features} The effect of different feature set on accuracy.
 FS4 and FS3 give good results in a) Hs-DS and b) EL-Ds.
-The results is improved by combining FS1 with FS2 and FS1 with FS4.
+The results are improved by combining FS1 with FS2 and FS1 with FS4.
 %The graph shows the recognition rate of each symbol using different algorithms.
@@
 \end{figure*}  
- We could not test the result of the segmentation algorithm alone due to the fact that the correct segmentation is hightly qualitative.
+ We could not test the result of the segmentation algorithm alone due to the fact that the correct segmentation is highly qualitative.
 It is only based on what the user intends to draw while sketching the symbol.
@@
 The result of segmentation block is tested by their effect on the final recognition results using the features that will be highly affected using different algorithms like \textbf{FS1} and \textbf{FS4} (Section\ref{sec:Recognition}).
-Figure \ref{fig:testFeatonly} shows the result of the different algorithms using only \textbf{FS1 }features.
+Figure \ref{fig:testFeatonly} shows the result of the different algorithms using only \textbf{FS1} features.
 \begin{figure}
@@
 a) Sample of single input stroke b) The single stroke after segmentation.
-c) An example of multiple stroke and their segmentation.
+c) An example of multiple strokes and their segmentation.
 %The figure shows sample of the output of the segmentation phase.
@@
 This paper presented a new approach to sketch recognition using DPSO.
-The system uses multiple premilary information i.e., speed and curvature information, as input to the \textit{DPSO}.
+The system uses multiple preliminary information, i.e., speed and curvature information, as input to the \textit{DPSO}.
 It is noted that the \textit{DPSO} in general generates an optimized stroke segmentation which improves the final recognition rate.
@@
 Global shape and moment properties were used to prevent the need to use graph matching which is computationally expensive \cite{SketchRead2007}.
-The experiments evaluated two\textit{DPSO} different segmenatation algorithms (AlgS1 and AlgS2) on three different datasets (Hs-Ds, EL-Ds and LD-Ds).
+The experiments evaluated two\textit{DPSO} different segmentation algorithms (AlgS1 and AlgS2) on three different datasets (Hs-Ds, EL-Ds and LD-Ds).
 Results show that AlgS2 achieves better segmentation and final accuracy than Alg3 \cite{earlyprocess} on the datasets used.
-The results proved that the system is easily expandable to more and more symbols as it dose not depend on low level primitive recogniziers but on high level symbol and geometrical featuers.
+The results proved that the system is easily expandable to more and more symbols as it does not depend on low level primitive recognizers but on high level symbol and geometrical features.
 The system achieved an overall improvement of more than 2\% over Alg3 in all datasets.
@@
 Features that represents the appearance of the symbols are likely to improve the recognition that have similar geometrical structure (i.e.
-Nand and And gate in logic symbols).
+Nor and OR gate in logic symbols).
  \bibliographystyle{elsarticle-harv} 

### 2009-06-09 trunk/paper2/Paper2.tex
aFter all changes requiredc by doctor

@@
 Two different segmentation algorithms were evaluated and results show that both algorithms improve the final recognition results.
-Experiments were conducted on three different dataset, simple presentation (Hs-DB), Electrical Design (EL-DB) and Logic Design (LD-DB) symbols.
+Experiments were conducted on three different dataset, simple presentation (Hs-DS), Electrical Design (EL-DS) and Logic Design (LD-DS) symbols.
 Results show that the effectiveness of the proposed system on various dataset especially on datasets with low shape diversity as in Logic Design dataset (LD-DS).
-% sketch are divided  Users draw symbols and sketchs using a set of pen strokes.
+% sketch are divided Users draw symbols and sketches using a set of pen strokes.
 %%Sketch recognition is defined as the process of identifying symbols that users draw using single or multiple %strokes.
@@
 %After two different segmentations are generated, 
-%///AllBlockDiagram.jpg  you can change the block to the main block. 
+%///AllBlockDiagram.jpg  you can change the block to the main block. flowchart
  \begin{figure}
 	\centering
-		\includegraphics[scale=0.48]{images/blockSmall.jpg}
-	\caption{Segmentation Block} 
+		\includegraphics[scale=0.48]{images/flowchart.jpg}
+	\caption{Segmentation process flow chart} Inputs to the segmentation proces are the input stroke and \textit{possible dominant points} $P_{pd}$.
+The output is either the stroke fitted as an ellipse or the best segmentation generated from the two segmentation algorithms.
+% First, attempts to fit stroke as an ellipse.
+Then two different segmentation are generated using (AlgS1 and AlgS2).
+Finaly, the Minimum error of segmentation as the best segmentation.
 	\label{fig:segblock}
 \end{figure}
+
+% \begin{figure}
+%	\centering
+%		\includegraphics[scale=0.48]{images/blockSmall.jpg}
+%	\caption{Segmentation Block} 
+%	\label{fig:segblock}
+%\end{figure}
+
 \subsection{Ellipse fitting}
@@
 	\item The particles are tested to make sure that the distance between every two successive dominate point is larger than $min_D$, where $min_D$ is 5\% of the total length of the stroke.
-Otherwise, one of these points is removed based on the error cased by its removal.
-	\item Each two successive segments are test to determine if they can be merged into a single segment.
-If merging the two segments have the same or smaller the error, the two segments are merged.
+Otherwise, two temperary particles are evaluated by removing either point from the current particle.
+After that, one of these points is removed based on the error caused by its removal.
+	\item Each two successive segments are tested to determine if they can be merged into a single segment.
+If merging the two segments results in segmentation that has an error value less than or equal to the error before merging, the two segments are merged.
 This step ensures that the final segmentation of the system has minimum number of segments.
+	% the same or smaller the error
 \end{enumerate}
@@
  \begin{itemize}
-	 \item \emph{Segments:} Number of segments in the symbol.
-	 \item \emph{Strokes:} Number of strokes or partial strokes that created the symbols.  
-		\item  \emph{Primitives:} Number of primitives in the symbol. The feature helps when identifying             symbols with mixed geometric primitives like cylinders and callouts.  
-		\item \emph{Curves:} Number of curves or ellipses in the symbol. 
-		\item \emph{Lines:} Number of lines in the symbol. 
-		\item \emph{Perpendicular} lines Number of perpendicular lines. 
-		\item \emph{Parallel} lines Number of parallel lines. 
-		\item \emph{Intersections} Number of intersection between lines and curves. 
-		\item \emph{T intersections} Number of T intersections. 
-		\item \emph{L intersections} Number of L intersections. 
-		\item \emph{X intersections} Number of X intersections.
-
+	 \item \emph{Segments:} The number of segments in the symbol.
+	 \item \emph{Strokes:} The number of strokes or partial strokes that created the symbols.  
+		\item  \emph{Primitives:} The number of primitives in the symbol. This feature helps when identifying  symbols with mixed geometric primitives such as cylinders and callouts.  
+		\item \emph{Curves:} The number of curves or ellipses in the symbol. 
+		\item \emph{Lines:} The number of lines in the symbol. 
+		\item \emph{Perpendicular}\emph{lines:} The number of perpendicular lines. 
+		\item \emph{Parallel}\emph{lines:} The number of parallel lines. 
+		\item \emph{Intersections:} The number of intersection between lines and curves. 
+		\item \emph{T intersections:} The number of T intersections. 
+		\item \emph{L intersections:} The number of L intersections. 
+		\item \emph{X intersections:} The number of X intersections.	
 		%\item[No. of holes] Number of holes in the symbols. 
 \end{itemize}
-\subsubsection{Rubine Feature Set (FS2)}
-  Features introduced by Rubine\cite{gestureexample12} for single stroke gestures. Feature such as length of stroke, smoothness of symbol and properties of the bounding box of the symbol.
+\subsubsection{Rubine Features (FS2)}
+  These features introduced by Rubine \cite{gestureexample12} for single stroke gestures.
+Such as length of stroke, cosine of starting angle, smoothness of symbol and the properties of the bounding box of the symbol.
 
 \subsubsection{Statistical Features (FS3)} 
+These features use moments and statistical properties such as;
 \begin{itemize}
-	\item \emph{Zernike moments } Zernike moments of order n (10 or 20).  \cite{HeloiseBeautification}. 
+	\item \emph{Zernike moments:} The zernike moments are orthogonal moments which define the distribution of point in the input space.
+They are invariant to both reflection and rotation of the input shape \cite{ZerMomentOrthogonal}.
+%The zernike moment were used in \cite{HeloiseBeautification} t.
 \end{itemize}
 \subsubsection{Composite Features (FS4)}
- Features that are composites of different geometrical and statistical symbol. 
+ These features are composites of different geometrical and statistical properties of the symbol. 
 	\begin{itemize}
-\item \emph{Size Ratio} Ratio between width to height of the symbol.
-	\item \emph{Ink density} Compute the density of points inside its bounding box\cite{GeometryAndDomain102}.   
- 	\item \emph{Convex Hull Area} Area of convex hull with respect to area of bounding box of symbol.
-	\item \emph{Convex Hull Perimeter} Perimeter of convex hull with respect to total length of symbol.
-		\item \emph{Mean Centroidal radius} The Mean of the centroidal radius which is the distance from each point in the symbol to the center of gravity.
-
-	\item \emph{Mean Time difference} The mean of the time difference between each two successive points in the symbol. %Different strokes are appended to construct a single path.
+\item \emph{Size Ratio:} The ratio between width and height of the symbol.
+	\item \emph{Ink density:} The density of points inside its bounding box\cite{GeometryAndDomain102}.   
+ 	\item \emph{Convex Hull Area:} The area of convex hull with respect to area of bounding box of symbol.
+	\item \emph{Convex Hull Perimeter:} The perimeter of convex hull with respect to total length of symbol.
+		\item \emph{Mean Centroidal radius:} The Mean of the centroidal radius is the distance from each point in the symbol to the center of gravity.
+	\item \emph{Mean Time difference:} The mean of the time difference between each two successive points in the symbol. %Different strokes are appended to construct a single path.
 	%\item [Ra]
   \end{itemize}  
-
 %\end{description}
 %The features used consist of Rubine feature set,  of order 21 , ink density as well as some structural and spatial information like number of perpendiculars lines ,number of parallel lines and number of different types of primitives in each symbol.
-After computing the features the symbol is introduced to the classifier.
+After computing the features, the symbol is introduced to the classifier as a feature vector.
 The system uses Support Vector Machine (SVM) classifier with Linear kernel \cite{libsvm}.
-An OVO classifier structure (object versus object) is used to handle multi-class.% A validation algorithm is used to choose the Kernel variables parameters $c, \gamma$.
+An OVO classifier structure (object versus object) is used to handle multi-class problem.% A validation algorithm is used to choose the Kernel variables parameters $c, \gamma$.
 
@@
 \label{sec:Experiments} 
+%Figure \ref{fig:symbolSet} shows a set of the shapes used in this data set. 
 In our experiment we used three different datasets.
-First dataset was collected by Hse and Newton\cite{HeloiseBeautification} (Hs-DB).
+The first dataset (Figure \ref{fig:HsSet}) was collected by Hse and Newton\cite{HeloiseBeautification} (Hs-DS).
 The data are drawn by 16 users each of them was asked to draw 13 shapes from 30 to 50 times.
-Figure \ref{fig:symbolSet} shows a set of the shapes used in the data set.
-We collected the second and third datasets.
-Figure \ref{fig:ELsymbolSet} and \ref{fig:LogicsymbolSet} shows the symbols gathered from 7 different users in electrical (EL-DB) and logic design (LD-DB) domain respectively.
+Figure \ref{fig:ELsymbolSet} and \ref{fig:LogicsymbolSet} shows the symbols we manually gathered from 7 different users in electrical (EL-DS) and logic design (LD-DS) domain respectively.
 Table \ref{tab:datasets} shows a comparison between the three datasets used in our experiments.
-The datasets were divided into training set and test set.
+The datasets were divided into training and test sets.
 Five different splits for the training and test data are generated from each dataset.
@@
 The accuracy is computed as the number of correctly recognized samples divided by the total number of samples in the test.
+\begin{figure}
+\centering
+
+\fbox{ \parbox{5cm}{% 
+		\includegraphics[scale=0.5]{images/symbolSet.PNG}	}}
+		\caption{The Hs-DS Symbol Set} The dataset consist of simple presentation symbols.
+		\label{fig:HsSet} 
+\end{figure}
 
  \begin{figure}
-\centering 
-\label{fig:symbolSet}
-\fbox{ \parbox{5cm}{% 
-		\includegraphics[scale=0.5]{images/symbolSet.PNG}	}}
-		\caption{The Hs-DB Symbol Set}.
+			\subfigure[The Digital Design Symbols Dataset(LD-DS)]{\label{fig:LogicsymbolSet} \includegraphics[scale=0.5]{images/logicSet.jpg}}
+			\hfill
+			\subfigure[The Electrical Symbols Dataset(EL-DS)]
+			 {\label{fig:ELsymbolSet}\includegraphics[scale=0.45]{images/EelectImage.jpg}}
+%			 \label{fig:ELsymbolSet} 
+%			 \fbox{ 
+%			 \parbox{4cm}{\includegraphics[scale=0.5]{images/EelectImage.jpg}	}}}
+
+	\caption{The datasets that were manually collected}  a)  The (LD-DS) dataset consist of symbols that are visually similar (i.e.
+has low shape diversity).
+b) The (EL-DS) dataset consist of complex topological symbols.
 \end{figure}
-\begin{figure}
-\centering 
-\label{fig:LogicsymbolSet}
 
-		\includegraphics[scale=0.5]{images/logicSet.jpg}	 
-		\caption{The Digital Design Symbols Dataset(LD-DB)} Symbols used in the dataset.
-\end{figure}
-\begin{figure}
-\centering 
-\label{fig:ELsymbolSet}
-\fbox{ \parbox{4.5cm}{% 
-		\includegraphics[scale=0.5]{images/EelectImage.jpg}	}}
-		\caption{The Electrical Symbols Dataset(EL-DB)} .
-\end{figure}
+%\begin{figure}
+%\centering 
+%\label{fig:LogicsymbolSet}
+% 
+%		\includegraphics[scale=0.5]{images/logicSet.jpg}	 
+%		\caption{The Digital Design Symbols Dataset(LD-DS)} The dataset consist of symbols that are visually simillar (i.e. has low shape diversity). 
+%\end{figure}
+%\begin{figure}
+%\centering 
+%\label{fig:ELsymbolSet}
+%\fbox{ \parbox{4.5cm}{% 
+%		\includegraphics[scale=0.5]{images/EelectImage.jpg}	}}
+%		\caption{The Electrical Symbols Dataset(EL-DS)} . The dataset consist of complex toplogical symbols.
+%\end{figure}
 
@@
 \begin{center}
-\scalebox{0.9}{
- \begin{tabular}{|p{2cm}|p{1.5cm}|p{1.5cm}|p{1.5cm}|}
+\scalebox{0.8}{
+ \begin{tabular}{|p{3cm}|p{1.5cm}|p{1.5cm}|p{1.5cm}|}
  \hline
- & Hs-DB &  EL-DB & LD-DB \\ \hline
+ & Hs-DS &  EL-DS & LD-DS \\ \hline
 No of samples  &  7791 & 2764 & 1859 \\ \hline
@@
 Splits  & 5 random splits  & 5 random splits  & 5 random splits  \\ \hline
-Examples & Figure \ref{fig:symbolSet}  & Figure \ref{fig:ELsymbolSet}  & Figure \ref{fig:LogicsymbolSet}   \\ \hline
+Examples & Figure \ref{fig:HsSet}  & Figure \ref{fig:ELsymbolSet}  & Figure \ref{fig:LogicsymbolSet}   \\ \hline
 \end{tabular}
 }
-\caption[Datasets Comparisons]{Datasets Comparisons} 
+\caption[Datasets Comparisons]{Table compares between datasets used in experiments in terms of number of samples, number of samples per category, etc...} 
 \label{tab:datasets}
@@
 
-\subsection{Algorithms comparisons}
-\label{sec:AlgExp}
-A set of tests were performed on the system to enhance non ellipse fitting algorithm.
-The effect of number of agents, the error thresholds and various other parameters were investigated.
-The result of these tests are shown in Figure \ref{fig:swarmtesting} and   \ref{fig:swarmtesting2}.
+A set of tests were performed on the system to enhance non-ellipse fitting algorithms.
+The effect of the number of particles, the error thresholds and various other parameters were investigated.
+The results of these tests are shown in Figures \ref{fig:swarmtesting} and   \ref{fig:swarmtesting2}.
 Figure \ref{fig:swarmtesting} shows that as the number of agents increases the error value decrease (error of the segmentation algorithm).
-The final number of agents and maximum swarm iteration used in the system was based on a trade off between error achieved versus the computation time.
-Other algorithm parameters such as $c_1$,$c_2$ that were mentioned in Section\ref{sec:ParticleSwarmAlgorithm} are tested using similar tests (Figure \ref{fig:swarmtesting} and \ref{fig:swarmtesting2}).
+The final number of particles and maximum swarm iterations used in the system was based on a tradeoff between error achieved versus the computation time.
+The effect of other algorithm parameters such as $c_1$, $c_2$ (mentioned in Section\ref{sec:ParticleSwarmAlgorithm}) is tested using similar tests (Figure \ref{fig:swarmtesting}, \ref{fig:swarmtesting2}).
 
@@
 	\centering		
-	 \includegraphics[scale=0.3]{images/swarmAgentTest.jpg}
-	 	\caption{Swarm testing} Segmentation Error Vs. Number of agents
-	 	\label{fig:swarmtesting}
+	 	\subfigure[Segmentation Error Vs. Number of particles]{	\label{fig:swarmtesting} \includegraphics[scale=0.3]{images/swarmIterationVsParticles.jpg}}
+	 %\includegraphics[scale=0.3]{images/swarmAgentTest.jpg}} 	\caption{}
+
 	%\caption{Experiments results :} a)    b) c) 
-\end{figure}
+	%\end{figure}
 
-\begin{figure}
-	\centering		
-	 \includegraphics[scale=0.4]{images/swarmParamterC.jpg}
-	 	\caption{Swarm testing} Segmentation Error Vs. PSO parameters.
-	 	\label{fig:swarmtesting2}
-	%\caption{Experiments results :} a)    b) c)
+%\begin{figure}
+	%\centering
+
+		\subfigure[Segmentation Error Vs. PSO parameters.]{	\label{fig:swarmtesting2}  \includegraphics[scale=0.4]{images/swarmParamterC.jpg}}
+
+	 	\caption{}a) The error decrease as the number of swarm iteration increase until it reach a saturation.   b) The behavior of segmentation error versus c1.
+
+	%\Segmentation Error Vs. PSO parameters. caption{Experiments results :} a)    b) c)
 
 \end{figure} 
-
+\subsection{Algorithms comparisons}
+\label{sec:AlgExp}
 We performed several experiments to evaluate the presented recognition system.
-Firstly we tested recognition accuracy of shapes in the data set with both algorithms (\textsl{AlgS1} and \textsl{AlgS2}).
+Firstly, we tested recognition accuracy of shapes in the data set with both algorithms (\textsl{AlgS1} and \textsl{AlgS2}).
 We also implemented the segmentation algorithm described in \cite{earlyprocess} (\textsl{Alg3}) to use as reference to compare it with our swarm algorithms.
@@
 The two swarm algorithms were tested with and without the ellipse fitting module.
-The ellipse detection module appears to be superior.
-The result shows that combining both  (\textsl{AlgS1} and \textsl{AlgS2}) out perform any single algorithm.
-%show that combining algoirthmsachieve higher accuracy than any single algoirthm.
+Combining the ellipse detection module improves the performance.
+The result shows that (\textsl{AlgS2}) gives better performance than Alg3 and AlgS1 in all datasets.
+The results also shows that combining (\textsl{AlgS1} and \textsl{AlgS2} using minimum error) outperforms AlgS2 in Hs-Ds only.
+This behavior is explained by noticing that features are computed based on the chosen segmentation.
+This means that different segmentation for a symbol generates different set of features used for recognition.
+This behavior affect Ls-Ds and EL-Ds more than Hs-Ds due to the higher topological and structure symbols in both datasets than of Hs-Ds.
+%   can be explained by the simplicity of the symbols in Hs-DS % and LD-Ds unlike EL-DS where AlgS2 is the best algorithm.
+This behavior is explained by the structure of the symbols in each dataset.
+El-DS consist of highly topological and structure symbols that are usually drawn using multiple small strokes.
+This leads to    %   out perform any single algorithm in HS-DS and LD-DS.
+%show that combining algorithms achieve higher accuracy than any single algorithm.
 %improves the results.% with both the \textit{DPSO} algorithms.% The results show that both \textit{DPSO} algorithms achieve better result than other algorithms.
 
- \begin{figure}
+ \begin{figure}[h]
 	\centering		
-	 \includegraphics[scale=0.4]{images/testAlg.jpg}
-	 	\caption{Algorithm comparison on HS-DB} The recognition rate of different algorithms. 
-	 	\label{fig:test1}
+	\subfigure[Algorithm comparison on HS-DS]{\label{fig:test1} \includegraphics[scale=0.4]{images/testAlg.jpg}}
+	 %	\caption{} The recognition rate of different algorithms.
+
 	%\caption{Experiments results :} a)    b) c) 
-\end{figure}
+%\end{figure}
 
-\begin{figure}
-	\centering		
-	 \includegraphics[scale=0.4]{images/LDAlg.jpg}
-	 	\caption{Algorithm comparison on LD-DB} The recognition rate of different algorithms. 
-	 	\label{fig:testLD}
+%\begin{figure}[h]
+%	\centering		
+		\subfigure[Algorithm comparison on LD-DS]{ 	 	\label{fig:testLD} \includegraphics[scale=0.4]{images/LDAlg.jpg}}
+	% 	\caption{} The recognition rate of different algorithms.
+
 	%\caption{Experiments results :} a)    b) c) 
-\end{figure}
+%\end{figure}
 
-\begin{figure}
-	\centering		
-	 \includegraphics[scale=0.4]{images/AlgEL.jpg}
-	 	\caption{Algorithm comparison on EL-DB} The recognition rate of different algorithms. 
-	 	\label{fig:testEL}
+%\begin{figure}[h]
+%	\centering		
+		\subfigure[Algorithm comparison on EL-DS]{	 	\label{fig:testEL}  \includegraphics[scale=0.4]{images/AlgE.jpg}}
+	 	\caption{ The recognition rate of different algorithms. }
+
 	%\caption{Experiments results :} a)    b) c) 
@@
 \label{sec:ShapeComplexityExperiments}
-
 \begin{figure*}
 	\centering 
-		\includegraphics[scale=0.5]{images/testsym.jpg}
-	\caption{Symbols Comparison on HS-DB} The effect of symbol complexity.  %The graph shows the recognition rate of each symbol using different algorithms. 
-	\label{fig:test2}
+		\subfigure[Symbols Comparison on HS-DS]{\label{fig:test2}
+		\includegraphics[scale=0.35]{images/testsym.jpg}}
+	%\caption{} The effect of symbol complexity.  %The graph shows the recognition rate of each symbol using different algorithms. 
+%\end{figure*}  
+%\begin{figure*}
+%	\centering
+\subfigure[Symbols Comparison on LD-DS symbols]{\label{fig:LDtest2}
+		\includegraphics[scale=0.5]{images/DLSymbols.jpg}}
+	%\caption{}.  %The graph shows the recognition rate of each symbol using different algorithms. 
+	\caption {Symbol Comparison} The accuracy of achieved by each algorithm in each symbol.
+a) Low accuracy level appears in triangle, heart, and moon symbols.
+b) Low accuracy level appears only in symbols that are visually confusing (i.e Nor, OR, Nand gates).
 \end{figure*}  
-
-\begin{figure*}
-	\centering
-		\includegraphics[scale=0.5]{images/LDsymbols.jpg}
-	\caption{Symbols Comparison on LD-DB symbols}.  %The graph shows the recognition rate of each symbol using different algorithms. 
-	\label{fig:LDtest2}
-\end{figure*}  
-The second experiment we implemented was to investigate the effect of symbol complexity and type on the recognition rate.
-Figure \ref{fig:test2} and \ref{fig:LDtest2} shows the achieved accuracy of each symbol by our algorithms compared to \cite{earlyprocess} (\textsl{Alg3}).
-It is clearly noted that symbols that have only line segments achieve higher accuracy rate than other symbols.
-The results indicate that algorithm \textsl{AlgS1} achieve better performance than algorithm \textsl{AlgS2} in the symbols that consist only of lines.
+The second experiment we implemented was to investigate the effect of symbol complexity and its type on the recognition rate.
+Figure \ref{fig:test2} and \ref{fig:LDtest2} show the achieved accuracy of each symbol by our algorithms compared to (\textsl{Alg3}) \cite{earlyprocess}.
+It is clearly noted that symbols that have only line segments achieve higher accuracy rate than other symbols (such as square, parallegram).
+The results also indicate that algorithm \textsl{AlgS1} achieves better performance than algorithm \textsl{AlgS2} in the symbols that consists only of lines.
 This is understandable because algorithm \textsl{AlgS1} divides strokes into line segments only but \textsl{AlgS2} is able to divide strokes into lines and curves based on the minimum error of the segment itself.
-Figure \ref{fig:LDtest2} shows that in logic design symbols \textsl{AlgS2} achieved better performance than \textsl{Alg3} \textsl{AlgS1} this emphases that complex shapes that are combination of lines and arc are better segmented using \textsl{AlgS2}.
-Algorithm \textsl{Alg3} gives good performance as long as the symbols consist of combination of lines and curves, if the stroke consists of only curve or lines the algorithm may lead to wrong segmentation result.
-This is because the system divides the stroke first to line segments then tries to decide if each segment can be represent better as a curve unlike algorithm \textsl{AlgS2} where the curve segments are tested while choosing the best segmentation.
+Figure \ref{fig:LDtest2} shows that in logic design (LD-DS) \textsl{AlgS2} achieved better performance than \textsl{Alg3} and  \textsl{AlgS1}.
+This proves that for complex shapes that are combination of lines and arcs are \textsl{AlgS2} achieves better segmentation.
+Algorithm \textsl{Alg3} gives good performance as long as the symbols consist of combination of lines and curves, if the stroke consists of curve or lines only the algorithm may lead to wrong segmentation result.
+This is because the system divides the stroke first into line segments then tries to decide whether each segment can be represented better as a curve.
+However, in \textsl{AlgS2} the curve segments are tested while choosing the best segmentation.
 Combining both algorithm \textsl{AlgS1} and \textsl{AlgS2} improved the recognition rate of all symbols.
 The penalty for this improved performance is the computational time required to run both swarm algorithms.
-Table \ref{tab:ConfusionMatrix} shows the confusion matrix of symbols.
-The table shows that errors are only between two sets of symbols moon with triangles and trapezoid with triangles.
+Table \ref{tab:ConfusionMatrix} shows the confusion matrix for the (HS-DS) symbols.
+The table shows that an error occurs mostly between two sets of symbols: moon with triangles and trapezoid with triangles.
 This observation is understandable as the symbols are visually similar but the system must be able to differentiate between them.
-This lead us to the next experiment to choose best set of feature to recognize symbols.
+This leads us to the next experiment to choose the best set of feature to recognize symbols.
 \begin{table*}
@@
 \label{sec:featexp}
-Different feature sets are tested to determine the best features that can used in sketch and symbol recognition.
-Figure\ref{fig:testFeaturesAll} shows the result of different feature sets from the basic four sets (\textbf{FS1,FS2,FS3,FS4}).
-Result shows that (\textbf{FS2}) Rubine features \cite{gestureexample12} gives the worst results when used alone without any other features.
+%Results show that (\textbf{FS2}) Rubine features \cite{gestureexample12} gives the worst results when used alone without any other features.
+Different feature sets are tested to determine the best features that can be used in sketch and symbol recognition.
+Figure\ref{fig:testFeaturesAll} shows the result of using different feature sets (from the basic four sets \textbf{FS1,FS2,FS3,FS4} in Section \ref{sec:Recognition}).
+Results show that geometerical features (\textbf{FS1}) and Rubine features (\textbf{FS2}) \cite{gestureexample12} gives the low results when used alone without any other features.
 This is because they are mainly computed for single stroke gestures and fare bad in multi-stroke symbols \cite{compareFeaturSVM}.
-Features \textbf{(FS4)} gives good results but it is improved by adding structural features \textbf{(FS1)}.
+Features \textbf{(FS4)} gives good results and it is further improved by adding geometerical structural features \textbf{(FS1)}.
  \begin{figure*}
 	\centering
-		\includegraphics[scale=0.5]{images/featuresdifference.jpg}
-	\caption{Feature Comparison on Hs-DB accuracy} The effect of different features on accuracy.  %The graph shows the recognition rate of each symbol using different algorithms. 
-	\label{fig:testFeaturesAll}
+	\subfigure[Feature Comparison on Hs-DS]{	\label{fig:testFeaturesAll} \includegraphics[scale=0.5]{images/featuresdifference.jpg}}
+    %The graph shows the recognition rate of each symbol using different algorithms.
+
+%\end{figure*}
+
+ %\begin{figure*}
+	%\centering
+
+	\subfigure[Feature Comparison on EL-DS]{	\label{fig:ELtestFeaturesAll}	\includegraphics[scale=0.34]{images/FeatEl.jpg}}
+	\caption{Different Features} The effect of different feature set on accuracy.
+FS4 and FS3 give good results in a) Hs-DS and b) EL-Ds.
+The results are improved by combining FS1 with FS2 and FS1 with FS4.
+%The graph shows the recognition rate of each symbol using different algorithms.
+
 \end{figure*}  
-
- \begin{figure*}
-	\centering
-		\includegraphics[scale=0.34]{images/featELc.jpg}
-	\caption{Feature Comparison on EL-DB} The effect of different features on accuracy.  %The graph shows the recognition rate of each symbol using different algorithms. 
-	\label{fig:ELtestFeaturesAll}
-\end{figure*}  
- We could not test the result of the segmentation algorithm directly due to the fact that the correct segmentation is highly ambiguous.
+ We could not test the result of the segmentation algorithm alone due to the fact that the correct segmentation is highly qualitative.
 It is only based on what the user intends to draw while sketching the symbol.
@@
 The result of segmentation block is tested by their effect on the final recognition results using the features that will be highly affected using different algorithms like \textbf{FS1} and \textbf{FS4} (Section\ref{sec:Recognition}).
-Figure \ref{fig:testFeatonly} shows the result of the different algorithms using only \textbf{FS1 }features.
+Figure \ref{fig:testFeatonly} shows the result of the different algorithms using only \textbf{FS1} features.
 \begin{figure}
 	\centering
-		\includegraphics[scale=0.4]{images/featureFS1only.jpg}
-	\caption{The effect of segmentation on Hs-DB accuracy} %The effect of symbol complexity.  %The graph shows the recognition rate of each symbol using different algorithms. 
+		\includegraphics[scale=0.3]{images/OnlyF1EL.jpg}
+	\caption{The effect of segmentation on Hs-DS accuracy} %The effect of symbol complexity.  %The graph shows the recognition rate of each symbol using different algorithms. 
 	\label{fig:testFeatonly}
@@
 	\caption{Segmentation outputs} The output segmentation of different input strokes.
+a) Sample of single input stroke b) The single stroke after segmentation.
+c) An example of multiple strokes and their segmentation.
 %The figure shows sample of the output of the segmentation phase.
@@
 \label{ConclusionandFutureWork}
-This paper presented a new approach to sketch recognition using PSO.
-The system uses both speed and curvature information which help improving the \textit{DPSO}.
+
+This paper presented a new approach to sketch recognition using DPSO.
+The system uses multiple preliminary information, i.e., speed and curvature information, as input to the \textit{DPSO}.
 It is noted that the \textit{DPSO} in general generates an optimized stroke segmentation which improves the final recognition rate.
-The tradeoff between accuracy achieved and time complexity must be further investigated to achieve better results.
+The tradeoff between accuracy achieved and time complexity can be further investigated to achieve better results.
 The final recognition is based on both global shape and hierarchical stroke properties.
 Adding stroke and segment geometrical features improved the final recognition.
-Global shape and moment properties was used to prevent the need to use graph matching which is computationally expensive.
-The system was tested on 13 different symbols and achieved an overall accuracy of 89.5\%.
+Global shape and moment properties were used to prevent the need to use graph matching which is computationally expensive \cite{SketchRead2007}.
+The experiments evaluated two\textit{DPSO} different segmentation algorithms (AlgS1 and AlgS2) on three different datasets (Hs-Ds, EL-Ds and LD-Ds).
+Results show that AlgS2 achieves better segmentation and final accuracy than Alg3 \cite{earlyprocess} on the datasets used.
+The results proved that the system is easily expandable to more and more symbols as it does not depend on low level primitive recognizers but on high level symbol and geometrical features.
+The system achieved an overall improvement of more than 2\% over Alg3 in all datasets.
+%The system was tested on 13 different symbols and achieved an overall accuracy of 89.5\%.
 The system does not depend on low level recognizer but rather on a set of high level features.

### 2009-06-10 paper2/Paper2.tex
some editing after meeting with dr. magdah today 10 6

@@
 The goal of this paper is developing a sketch recognition system that facilitates using sketches in computer systems.
-The system uses Particle Swarm Optimization (PSO) algorithm to optimally segment the strokes the user draws into meaningful geometric primitives.
+The system uses the Particle Swarm Optimization (PSO) algorithm to optimally segment the strokes the user draws into meaningful geometric primitives.
 These geometric primitives are grouped to form symbols which are further identified using a Support Vector Machines (SVM) classifier.
 Two versions of PSO segmentation algorithms were implemented.
-Experiments were conducted on three different dataset, simple presentation (Hs-DS), Electrical Design (EL-DS) and Logic Design (LD-DS) symbols.
+Experiments were conducted on three different datasets, simple presentation (Hs-DS), Electrical Design (EL-DS) and Logic Design (LD-DS) symbols.
 Results show that both PSO algorithms produce better segmentation than most current systems.
 This segmentation improvement contributed in getting more accurate recognition results on all the datasets tested.
-The results prove that the system is effective specially in datasets with low diversity in shapes, i.e., shapes are very similar, as in Logic Design dataset (LD-DS).
+The results prove that the system is effective especially in datasets with low diversity in shapes, i.e., similar shapes, as in Logic Design dataset (LD-DS).
 %The results prove the effectiveness of the proposed system on various dataset, especially datasets with low shape diversity as in Logic Design dataset (LD-DS).% effectiveness of the proposed system on various dataset, especially datasets with low shape diversity as in Logic Design dataset (LD-DS).
@@
 %% keywords here, in the form: keyword \sep keyword
-Symbol Recognition, Particle Swarm Optimization, Sketch Recognition, Geometrical Features, Polygon approximation
+Symbol Recognition, Discrite Particle Swarm Optimization (DPSO), Sketch Recognition, Geometrical Features, Polygon approximation
 %% PACS codes here, in the form: \PACS code \sep code
@@
 
-Sketch recognition can be applied either using Pattern Recognition (PR) based method or Artificial Intelligence (AI) based method.
+Sketch recognition can be applied using either Pattern Recognition (PR) based method or Artificial Intelligence (AI) based method.
 In AI based methods, Symbols are represented using hierarchical levels of simpler geometric shapes or primitives and spatial relations between them.
@@
 Recognition can then is treated as an AI problem which can be solved using methods like sub graph matching, Bayes belief network, and constrain satisfaction.
-A Hierarchical shape descriptions was widely used in sketch systems \cite{HierarchicalParsing7,SketchRead2007}.
+A Hierarchical shape description was widely used in sketch systems \cite{HierarchicalParsing7,SketchRead2007}.
 %Ladder30,AlvaradoFreedom42
@@
 The identification of individual stroke is used to determine The symbols are recognized by identifying the individual strokes.
+
 In PR based methods, systems compute properties that describe symbols then use a classification method to recognize symbols.
@@
 In gesture properties approach, systems identify symbols by using gesture properties of some of its underlying strokes.
-System that uses gestures properties assume that users draw symbols in specific order and style \cite{gestureexample12,aideddesgin22}.
+A system that uses gestures properties assumes that users draws symbols in specific order and style \cite{gestureexample12,aideddesgin22}.
 In global shape properties approach, systems investigate general shape properties and its underlying strokes.
@@
 
-In this paper, we choose to implement a hybrid method between shape property, gesture and hierarchical (AI method) based approaches.
+In this paper, we choose to implement a hybrid method that includes shape property, gesture and hierarchical information.
 Our method attempts to avoid many of the problems faced by approaches based on individual gestures or global shape properties while trying to gain the additional information that the stroke based system uses.
@@
 The first segmentation algorithm is based on dividing the digital curve into polygons\cite{PolygonApproximationPSO}.
-Our algorithm enhances its performance by adding curvature and other local information while segmenting strokes to achieve better results.
-To handle complex strokes, our system uses a second segmentation PSO algorithm to segment strokes into a set of lines and circular arcs.
+Our algorithm enhances its performance by adding curvature and other local information during segmenting strokes to achieve better results.
+To handle complex strokes, our system uses a second segmentation PSO algorithm that segment strokes into a set of lines and circular arcs.
+The two segmentation algorithms can either by used  individually or in parallel then chosing the segmentation that gives the minimum error.
 After segmentation, a feature vector composed of different statistical, geometrical and composite features is deduced.
@@
 Section \ref{Sysdisc} describes the proposed system in details.
-Section \ref{sec:Experiments} describes the experiments that were preformed.
+Section \ref{sec:Experiments} describes the results obtained from the experiments that were preformed.
 Finally, the conclusion and future work is presented in Section \ref{ConclusionandFutureWork}.
@@
 Our system adopts Yin \cite{PolygonApproximationPSO} method and tries to improve it by adding curvature and other local information while segmenting strokes to achieve better segmentations.
-The next section presents the general particle swarm algorithm which is used in this system.
+The next section presents the general particle swarm algorithm used in the proposed system.
 
@@
 This leads to a nearly constant time difference between samples while the pen is moving and large difference while the pen is stationary (Figure \ref{fig:timediff}).
-Contrary to time difference information, the speed information, where the users draw with variant speeds, the speed information has a lot of noise in the data (Figure \ref{fig:speed2Distance}).
+Contrary to time difference information, the speed information has a lot of noise in the data because usually  users draw with variant speeds (Figure \ref{fig:speed2Distance}).
 %Similarly, direction information is used as it provides better distinctive maxima for the corners than the curvature information \cite{meanshift10}.% Figure \ref{fig:curvatures} shows the direction and curvature graphs for the stroke drawn in Figure \ref{fig:orignalStroke}
@@
 
-After the system computes the speed, time difference and curvature information, it proceeds to detect the points that have low velocity and high curvature.
+After the system computes the speed, time difference and curvature information, it proceeds to extracts the points that have low velocity and high curvature.
 However, using simple differentiation to detect local extreme points results in false points due to the non smooth curves.
-Hence, our system adopted a process presented by \cite{earlyprocess}, where the mean of each of the four calculated information is calculated.
+Hence, our system adopts a process presented by \cite{earlyprocess}, where the mean of each of the four calculated information is calculated.
 This mean is taken as the threshold \textit{th} which is used to separate the curve into regions ($R_i$) (Figure\ref{fig:timediff}); each region $R_i$ is defined as a range of points, where the curve values are either above or below the threshold \textit{th}.
@@
 The least square fitting algorithm is used to minimize the fitting error of the ellipse equation \cite{chernov}.
-This process attempts to fit the stroke points into an ellipse arc; it starts with computing the center of the stroke bounding box.
+This process attempts to fit the stroke points into an ellipse arc; it starts by computing the center of the stroke bounding box.
 The bounding box center point is used as the first estimation of the center of the ellipse.
-The axes of the ellipse are estimated as the $width/2$ and $height/2$ of the stroke bounding box where $width$ and $height$ are the width and height of the bonding box.
+The axis of the ellipse are estimated as the $width/2$ and $height/2$ of the stroke bounding box where $width$ and $height$ are the width and height of the bonding box.
 The least square fitting algorithm \cite{chernov} is used to minimize the fitting error of the ellipse Equation (\ref{eq:circleFit})
@@
 Another measure is used to compute the efficiency of the final estimated ellipse.
-Equation \ref{eq:circleError} ensures that the drawn percentage of ellipse is considered.
+The value of $eff$ in Equation \ref{eq:circleError} ensures that the drawn percentage of ellipse is considered.
 This eliminates fitting a line into very large ellipse but leaves small ellipses to be fitted as a partial or a full ellipse.
@@
 \subsubsection{Structural and geometrical Features (FS1)}
-  They are features defines the structure of the geometrical symbol such as;   
+  They are features that define the structure of the geometrical symbol such as;   
 %\begin{description}
@@
 \subsubsection{Rubine Features (FS2)}
-  These features introduced by Rubine \cite{gestureexample12} for single stroke gestures.
-Such as length of stroke, cosine of starting angle, smoothness of symbol and the properties of the bounding box of the symbol.
+  These features were introduced by Rubine \cite{gestureexample12} for single stroke gestures.
+Such as length of stroke, cosine of starting angle, smoothness of symbol and some features extracted from the bounding box of the symbol.
 
@@
 The results of these tests are shown in Figures \ref{fig:swarmtesting} and   \ref{fig:swarmtesting2}.
-Figure \ref{fig:swarmtesting} shows that as the number of agents increases the error value decrease (error of the segmentation algorithm).
+Figure \ref{fig:swarmtesting} shows that as the number of swarm iterations increase the error value decrease (error of the segmentation algorithm).
 The final number of particles and maximum swarm iterations used in the system was based on a tradeoff between error achieved versus the computation time.
-The effect of other algorithm parameters such as $c_1$, $c_2$ (mentioned in Section\ref{sec:ParticleSwarmAlgorithm}) is tested using similar tests (Figure \ref{fig:swarmtesting}, \ref{fig:swarmtesting2}).
+The effect of other algorithm parameters such as $c_1$ (mentioned in Section\ref{sec:ParticleSwarmAlgorithm}) is tested using similar tests (Figure \ref{fig:swarmtesting}, \ref{fig:swarmtesting2}).%, $c_2$
 
@@
 	\centering		
-	 	\subfigure[Segmentation Error Vs. Number of particles]{	\label{fig:swarmtesting} \includegraphics[scale=0.3]{images/swarmAgentTest.jpg}}
-	 %	\caption{}
+	 	\subfigure[Segmentation Error Vs. Swarm Iterations]{	\label{fig:swarmtesting} \includegraphics[scale=0.3]{images/swarmIterationVsParticles.jpg}}
+	 %\includegraphics[scale=0.3]{images/swarmAgentTest.jpg}} 	\caption{}
 
@@
 	%\centering
-		\subfigure[Segmentation Error Vs. PSO parameters.]{	\label{fig:swarmtesting2}  \includegraphics[scale=0.4]{images/swarmParamterC.jpg}}
-	 	\caption{}a) The error decrease as the number of particle increase.   b) The behavior of segmentation error versus c1.
+
+		\subfigure[Segmentation Error Vs. c1]{	\label{fig:swarmtesting2}  \includegraphics[scale=0.4]{images/swarmParamterC.jpg}}
+
+	 	\caption{Segmentation Error Vs. PSO parameters.}a) The error decrease as the number of swarm iteration increase until it reach a saturation.   b) The behavior of segmentation error versus c1.
 
@@
 Firstly, we tested recognition accuracy of shapes in the data set with both algorithms (\textsl{AlgS1} and \textsl{AlgS2}).
-We also implemented the segmentation algorithm described in \cite{earlyprocess} (\textsl{Alg3}) to use as reference to compare it with our swarm algorithms.
-Figure \ref{fig:test1} and \ref{fig:testLD} shows the accuracy achieved by each algorithm.
+We also implemented the segmentation algorithm described in \cite{earlyprocess} (\textsl{Alg3}) to compare it with our swarm algorithms.
+Figure \ref{fig:test1} and \ref{fig:testLD} show the accuracy achieved by each algorithm.
 The two swarm algorithms were tested with and without the ellipse fitting module.
-Combining the ellipse detection module improves the performance.
+Combining the ellipse detection module improves the performance in certain datasets.
 The result shows that (\textsl{AlgS2}) gives better performance than Alg3 and AlgS1 in all datasets.
@@
 This behavior is explained by noticing that features are computed based on the chosen segmentation.
-This means that different segmentation for a symbol generates different set of features used for recognition.
-This behavior affect Ls-Ds and EL-Ds more than Hs-Ds due to the higher topological and structure symbols in both datasets than of Hs-Ds.
+This means that different segmentation for a symbol generates a different set of features used for recognition.
+This behavior affects Ls-Ds and EL-Ds more than Hs-Ds due to the higher topological and structure nature of the symbols in both datasets than of Hs-Ds.
 %   can be explained by the simplicity of the symbols in Hs-DS % and LD-Ds unlike EL-DS where AlgS2 is the best algorithm.
@@
 %	\centering		
-		\subfigure[Algorithm comparison on EL-DS]{	 	\label{fig:testEL}  \includegraphics[scale=0.4]{images/AlgE.jpg}}
+		\subfigure[Algorithm comparison on EL-DS]{	 	\label{fig:testEL}  \includegraphics[scale=0.4]{images/AlgEL.jpg}}
 	 	\caption{ The recognition rate of different algorithms. }
@@
 \label{sec:ShapeComplexityExperiments}
-\begin{figure*}
-	\centering 
-		\subfigure[Symbols Comparison on HS-DS]{\label{fig:test2}
-		\includegraphics[scale=0.35]{images/testsym.jpg}}
-	%\caption{} The effect of symbol complexity.  %The graph shows the recognition rate of each symbol using different algorithms. 
-%\end{figure*}  
-%\begin{figure*}
-%	\centering
-\subfigure[Symbols Comparison on LD-DS symbols]{\label{fig:LDtest2}
-		\includegraphics[scale=0.5]{images/DLSymbols.jpg}}
-	%\caption{}.  %The graph shows the recognition rate of each symbol using different algorithms. 
-	\caption {Symbol Comparison} The accuracy of achieved by each algorithm in each symbol.
-a) Low accuracy level appears in triangle, heart, and moon symbols.
-b) Low accuracy level appears only in symbols that are visually confusing (i.e Nor, OR, Nand gates).
-\end{figure*}  
 The second experiment we implemented was to investigate the effect of symbol complexity and its type on the recognition rate.
@@
 This leads us to the next experiment to choose the best set of feature to recognize symbols.
+\begin{figure*}[h]
+	\centering 
+		\subfigure[Symbols Comparison on HS-DS]{\label{fig:test2}
+		\includegraphics[scale=0.35]{images/testsym.jpg}}
+	%\caption{} The effect of symbol complexity.  %The graph shows the recognition rate of each symbol using different algorithms. 
+%\end{figure*}  
+%\begin{figure*}
+%	\centering
+\subfigure[Symbols Comparison on LD-DS symbols]{\label{fig:LDtest2}
+		\includegraphics[scale=0.5]{images/DLSymbols.jpg}}
+	%\caption{}.  %The graph shows the recognition rate of each symbol using different algorithms. 
+	\caption {Symbol Comparison} The accuracy of achieved by each algorithm in each symbol.
+a) Low accuracy level appears in triangle, heart, and moon symbols.
+b) Low accuracy level appears only in symbols that are visually confusing (i.e Nor, OR, Nand gates).
+\end{figure*}  
 \begin{table*}
@@
 Figure\ref{fig:testFeaturesAll} shows the result of using different feature sets (from the basic four sets \textbf{FS1,FS2,FS3,FS4} in Section \ref{sec:Recognition}).
-Results show that geometerical features (\textbf{FS1}) and Rubine features (\textbf{FS2}) \cite{gestureexample12} gives the low results when used alone without any other features.
-This is because they are mainly computed for single stroke gestures and fare bad in multi-stroke symbols \cite{compareFeaturSVM}.
-Features \textbf{(FS4)} gives good results and it is further improved by adding geometerical structural features \textbf{(FS1)}.
- \begin{figure*}
+Results show that geometerical features (\textbf{FS1}) and Rubine features (\textbf{FS2}) \cite{gestureexample12} give the low results when used alone without any other features.
+This is because they are mainly computed for single stroke gestures and fare badly in multi-stroke symbols \cite{compareFeaturSVM}.
+Features \textbf{(FS4)} give good results and it is further improved by adding geometerical structural features \textbf{(FS1)}.
+
+ \begin{figure*}[h]
 	\centering
@@
 %The graph shows the recognition rate of each symbol using different algorithms.
+\end{figure*}
 
-\end{figure*}  
  We could not test the result of the segmentation algorithm alone due to the fact that the correct segmentation is highly qualitative.
 It is only based on what the user intends to draw while sketching the symbol.
+
  The main features affected by the segmentation stage are the geometric features computed in \textbf{FS1}.
@@
 Figure \ref{fig:testFeatonly} shows the result of the different algorithms using only \textbf{FS1} features.
-\begin{figure}
+\begin{figure}[h] 
 	\centering
 		\includegraphics[scale=0.3]{images/OnlyF1EL.jpg}
-	\caption{The effect of segmentation on Hs-DS accuracy} %The effect of symbol complexity.  %The graph shows the recognition rate of each symbol using different algorithms. 
+	\caption{The effect of segmentation on accuracy} The experiment is implemented on Hs-Ds using only FS1.
+%The effect of symbol complexity.
+%The graph shows the recognition rate of each symbol using different algorithms.
 	\label{fig:testFeatonly}

### 2009-06-10 trunk/paper2/Paper2.tex
some commment removal

@@
 \label{sec:ParticleSwarmAlgorithm}
-%\section{Particle Swarm Algorithm}
-%\label{PSO}
-%What is particle swarm algorithm and how it was used in related researches. 
-%The main idea of \textit{Particle Swarm Algorithm (PSO)} is to represent each solution in the solution space with a particle\cite{PSOFirst}.
-Each particle moves the n with a direction and velocity $v_{ij}$ based on equations \ref{eq:Swarm1} \& \ref{eq:Swarm}.
  The main idea of \textit{Particle Swarm Algorithm (PSO)} is to represent each solution with a $N$ dimension particle from the solution space \cite{PSOFirst}.
@@
 \subsubsection{Structural and geometrical Features (FS1)}
-  They are features defines the structure of the geometrical symbol such as;   
-%\begin{description}
-%	 \item[No. Of segments] Number of segments in the symbol. 
-%		\item[No. Of primitives] Number of primitives in the symbol. The feature helps when identifying             symbols with mixed geometric primitives like cylinders and callouts.  
-%		\item [No. of curves] Number of curves or ellipses in the symbol. 
-%		\item [No. of lines] Number of lines in the symbol. 
-%		\item [No. of perpendicular lines] Number of perpendicular lines. 
-%		\item [No. of parallel lines] Number of parallel lines. 
-%		\item [No. of intersections] Number of intersection between lines and curves. 
-%		\item [Size Ratio] Ratio between width to height of the symbol.
-%		%\item[No. of holes] Number of holes in the symbols. 
-%
-%\end{description}
+  They are features that define the structure of the geometrical symbol such as;   
  \begin{itemize}
@@
 		\item \emph{Mean Centroidal radius:} The Mean of the centroidal radius is the distance from each point in the symbol to the center of gravity.
-	\item \emph{Mean Time difference:} The mean of the time difference between each two successive points in the symbol. %Different strokes are appended to construct a single path.
-	%\item [Ra]
+	\item \emph{Mean Time difference:} The mean of the time difference between each two successive points in the symbol. 
   \end{itemize}
-%\end{description}
-%The features used consist of Rubine feature set,  of order 21 , ink density as well as some structural and spatial information like number of perpendiculars lines ,number of parallel lines and number of different types of primitives in each symbol.
+
  After computing the features, the symbol is introduced to the classifier as a feature vector.
@@
 
-%\begin{figure}
-%\centering 
-%\label{fig:LogicsymbolSet}
-% 
-%		\includegraphics[scale=0.5]{images/logicSet.jpg}	 
-%		\caption{The Digital Design Symbols Dataset(LD-DS)} The dataset consist of symbols that are visually simillar (i.e. has low shape diversity). 
-%\end{figure}
-%\begin{figure}
-%\centering 
-%\label{fig:ELsymbolSet}
-%\fbox{ \parbox{4.5cm}{% 
-%		\includegraphics[scale=0.5]{images/EelectImage.jpg}	}}
-%		\caption{The Electrical Symbols Dataset(EL-DS)} . The dataset consist of complex toplogical symbols.
-%\end{figure}
-
 \begin{table}
@@
 
-%The dataset collected by Hse and Newton\cite{HeloiseBeautification} is used to test the presented system.
-The data are drawn by 16 users each of them was asked to draw 13 shapes from 30 to 50 times.
-Figure \ref{fig:symbolSet} shows a set of the shapes used in the data set.
-The dataset was divided into training set and test set.
-Four different splits for the training and test data are generated from the dataset.
-The results displayed are the average recognition accuracy of the four splits.
-The accuracy is computed as the number of correctly recognized samples divided by the total number of samples in the test.
-%In the following section, the recognition accuracy means the average recognition accuracy in the four data splits.
-
 A set of tests were performed on the system to enhance non-ellipse fitting algorithms.
@@
 	\centering		
-	 	\subfigure[Segmentation Error Vs. Number of particles]{	\label{fig:swarmtesting} \includegraphics[scale=0.3]{images/swarmIterationVsParticles.jpg}}
+	 	\subfigure[Segmentation Error Vs. Swarm Iterations]{	\label{fig:swarmtesting} \includegraphics[scale=0.3]{images/swarmIterationVsParticles.jpg}}
 	 %\includegraphics[scale=0.3]{images/swarmAgentTest.jpg}} 	\caption{}
@@
 
-		\subfigure[Segmentation Error Vs. PSO parameters.]{	\label{fig:swarmtesting2}  \includegraphics[scale=0.4]{images/swarmParamterC.jpg}}
+		\subfigure[Segmentation Error Vs. c1]{	\label{fig:swarmtesting2}  \includegraphics[scale=0.4]{images/swarmParamterC.jpg}}
 
-	 	\caption{}a) The error decrease as the number of swarm iteration increase until it reach a saturation.   b) The behavior of segmentation error versus c1.
+	 	\caption{Segmentation Error Vs. PSO parameters.}a) The error decrease as the number of swarm iteration increase until it reach a saturation.   b) The behavior of segmentation error versus c1.
 
@@
 This behavior is explained by noticing that features are computed based on the chosen segmentation.
-This means that different segmentation for a symbol generates different set of features used for recognition.
-This behavior affect Ls-Ds and EL-Ds more than Hs-Ds due to the higher topological and structure symbols in both datasets than of Hs-Ds.
-%   can be explained by the simplicity of the symbols in Hs-DS % and LD-Ds unlike EL-DS where AlgS2 is the best algorithm.
-This behavior is explained by the structure of the symbols in each dataset.
-El-DS consist of highly topological and structure symbols that are usually drawn using multiple small strokes.
-This leads to    %   out perform any single algorithm in HS-DS and LD-DS.
-%show that combining algorithms achieve higher accuracy than any single algorithm.
-%improves the results.% with both the \textit{DPSO} algorithms.% The results show that both \textit{DPSO} algorithms achieve better result than other algorithms.
+This means that different segmentation for a symbol generates a different set of features used for recognition.
+This behavior affects Ls-Ds and EL-Ds more than Hs-Ds due to the higher topological and structure nature of the symbols in both datasets than of Hs-Ds.
 
- \begin{figure}[h]
+ \begin{figure} 
 	\centering		
 	\subfigure[Algorithm comparison on HS-DS]{\label{fig:test1} \includegraphics[scale=0.4]{images/testAlg.jpg}}
-	 %	\caption{} The recognition rate of different algorithms.
-
-	%\caption{Experiments results :} a)    b) c) 
-%\end{figure}
-
-%\begin{figure}[h]
-%	\centering		
  		\subfigure[Algorithm comparison on LD-DS]{ 	 	\label{fig:testLD} \includegraphics[scale=0.4]{images/LDAlg.jpg}}
-	% 	\caption{} The recognition rate of different algorithms.
-
-	%\caption{Experiments results :} a)    b) c) 
-%\end{figure}
-
-%\begin{figure}[h]
-%	\centering		
-		\subfigure[Algorithm comparison on EL-DS]{	 	\label{fig:testEL}  \includegraphics[scale=0.4]{images/AlgE.jpg}}
+ 		\subfigure[Algorithm comparison on EL-DS]{	 	\label{fig:testEL}  \includegraphics[scale=0.4]{images/AlgEL.jpg}}
 	 	\caption{ The recognition rate of different algorithms. }
-
-	%\caption{Experiments results :} a)    b) c) 
  \end{figure}
@@
 This leads us to the next experiment to choose the best set of feature to recognize symbols.
+\begin{figure*} 
+	\centering 
+		\subfigure[Symbols Comparison on HS-DS]{\label{fig:test2}
+		\includegraphics[scale=0.35]{images/testsym.jpg}}
+\subfigure[Symbols Comparison on LD-DS symbols]{\label{fig:LDtest2}
+		\includegraphics[scale=0.5]{images/DLSymbols.jpg}}
+	%\caption{}.  %The graph shows the recognition rate of each symbol using different algorithms. 
+	\caption {Symbol Comparison} The accuracy of achieved by each algorithm in each symbol.
+a) Low accuracy level appears in triangle, heart, and moon symbols.
+b) Low accuracy level appears only in symbols that are visually confusing (i.e Nor, OR, Nand gates).
+\end{figure*}  
 \begin{table*}
@@
 Figure\ref{fig:testFeaturesAll} shows the result of using different feature sets (from the basic four sets \textbf{FS1,FS2,FS3,FS4} in Section \ref{sec:Recognition}).
-Results show that geometerical features (\textbf{FS1}) and Rubine features (\textbf{FS2}) \cite{gestureexample12} gives the low results when used alone without any other features.
-This is because they are mainly computed for single stroke gestures and fare bad in multi-stroke symbols \cite{compareFeaturSVM}.
-Features \textbf{(FS4)} gives good results and it is further improved by adding geometerical structural features \textbf{(FS1)}.
- \begin{figure*}
+Results show that geometerical features (\textbf{FS1}) and Rubine features (\textbf{FS2}) \cite{gestureexample12} give the low results when used alone without any other features.
+This is because they are mainly computed for single stroke gestures and fare badly in multi-stroke symbols \cite{compareFeaturSVM}.
+Features \textbf{(FS4)} give good results and it is further improved by adding geometerical structural features \textbf{(FS1)}.
+
+ \begin{figure*}[h]
 	\centering
 	\subfigure[Feature Comparison on Hs-DS]{	\label{fig:testFeaturesAll} \includegraphics[scale=0.5]{images/featuresdifference.jpg}}
-    %The graph shows the recognition rate of each symbol using different algorithms.
-
-%\end{figure*}
-
- %\begin{figure*}
-	%\centering
-
 	\subfigure[Feature Comparison on EL-DS]{	\label{fig:ELtestFeaturesAll}	\includegraphics[scale=0.34]{images/FeatEl.jpg}}
@@
 Figure \ref{fig:testFeatonly} shows the result of the different algorithms using only \textbf{FS1} features.
-\begin{figure}
+\begin{figure*} 
 	\centering
 		\includegraphics[scale=0.3]{images/OnlyF1EL.jpg}
-	\caption{The effect of segmentation on Hs-DS accuracy} %The effect of symbol complexity.  %The graph shows the recognition rate of each symbol using different algorithms. 
+	\caption{The effect of segmentation on accuracy} The experiment is implemented on Hs-Ds using only FS1.
+%The effect of symbol complexity.
+%The graph shows the recognition rate of each symbol using different algorithms.
 	\label{fig:testFeatonly}
-\end{figure}
+\end{figure*}
 

### 2009-06-10 trunk/paper2/Paper2.tex
after add

@@
 This is because they assume that each stroke plays a role in the structure of the shape.
-%The role of which each stroke plays in the structure of the sketched symbol is used to determine shapes in stroke based methods.  
-%Gestures are pen strokes drawn by the user that can immediately recognized by the system.
-%role of which each stroke plays in the structure of the sketched symbol is used to determine shapes 
-%.
-Describing the role  how user draws individual strokes  properties that describe how users draw individual strokes in drawing the symbol.
-to.
-The identification of individual stroke is used to determine The symbols are recognized by identifying the individual strokes.
 
@@
 
- %In general, the appearance based methods use image processing techniques to recognize sketches that the users draw. 
-%Sketch recognition can be based on: strokes, global shape properties, and appearance \cite{Oltmans07}.
-The role of which each stroke plays in the structure of the sketched symbol is used to determine shapes in stroke based methods.
-Most stroke based systems either represent symbols as gestures or as a hierarchical shape.
-Gestures are pen strokes drawn by the user that can immediately recognized by the system.
-A Hierarchical shape descriptions was widely used in sketch systems \cite{HierarchicalParsing7,Ladder30,AlvaradoFreedom42,SketchRead2007}.
-Symbols are represented using hierarchical levels of simpler geometric shapes or primitives and spatial relations between them.
-Systems that employ such representation must first pre-process strokes to break them into geometric primitives.
-Recognition can then be treated as sub graph matching problem.
-
- %In global shape properties approach, systems investigate general shape properties and its underlying strokes.
-Using shape properties (i.e.
-ratio of area of convex hull to area of shape) relaxes the assumption and role of each stroke in the symbol but it does not truly represent the appearance of the symbol\cite{DiagramOfflineConvexHull}.
-Unlike appearance based methods where individual strokes are disregarded and they focus on the overall appearance of shapes.
-In general, the appearance based methods use image processing techniques to recognize sketches that the users draw.
-
- %Methods based on gesture and global shape properties produce systems that are dependent on assumptions about how strokes can be grouped.
-These systems assume that users will finish drawing a shape before moving to the next one \cite{Cali63,geometrydomain49}.
-This let them to temporally group strokes and then recognizes them based on their strokes or global shape properties.
-\cite{AlvaradoDigital} argued that this assumption does not hold in natural sketches.
-
- %Recognition based on strokes using hierarchical shape description is applied in \cite{SketchRead2007}.
-The principle problem in these systems is the computational complexity of the matching between shape description and geometric primitives produced in the segmentation process.
-Natural sketch are difficult to segment and group to create reliable fragment of geometric primitives.
-This is partially due to the increase of noise and other phenomena such as overlapped, over traced strokes and sharing of strokes between different symbols.
-Since these systems are dependable on the segmentation results, they either perform recognition using undependable segmentation results or consider different ways to segment the input sketch.
-
-%In this paper we choose to implement a hybrid method between shape property and stroke based approaches.
-Our method tries to avoid many of the problems faced by approaches based on individual strokes or global shape properties while trying to gain the additional information that the stroke based system use.
-Firstly, the system uses particle swarm optimization (PSO) to optimally segment user's strokes.
-Users can draw symbols using any number of strokes in any order they like.
-The system segments each stroke using two different PSO algorithms.
-The first segmentation algorithm is based on dividing the digital curve into polygons\cite{PolygonApproximationPSO}.
-Our algorithm enhances its performance by adding curvature and other local information while segmenting strokes to achieve better results.
-To handle complex strokes, our system uses a second segmentation PSO algorithm to segment strokes into a set of lines and circular arcs.
-After segmentation, a feature vector composed of different statistical, geometrical and composite features is deduced.
-The final classification step uses a SVM classifier to classify symbols into one of the previously trained classes based on the features vector computed.
-
- %The preprocessing phase captures user input stroke points and collects basic information about the stroke.
-In the segmentation phase, the strokes are divided into a set of simple geometrical primitives or segments.
-In the third phase, sketch recognition, strokes and segments are clustered to formulate symbols that can be recognized by a classifier.
-
-%This paper introduces a new sketch recognition system, which uses  % uses the computed feature vector to classify the segments into one of the previously trained classes.
-%Sections \ref{Prepross}, \ref{seg} and \ref{sec:Recognition} provide details on preprocessing, segmentation and recognition process respectively.
  The remaining of the paper is as follows; the literature review on stroke segmentation is presented in Section \ref{sec:review}.
@@
 \end{equation}
-%where $p_{ij}$ represent the $jth$ particle in the $ith$ agent and $v_{ij}$ is the velocity of the $jth$ particle in the $ith$ agent.
+
 where $p_{ij}$ represent the $j$th dimension in the $i$th particle and $v_{ij}$ is the velocity of the $j$th dimension in the $i$th particle.
@@
 \subsection{Preprocessing}
-\label{Prepross}% Agar \cite{polygonfeedback31} mentions that the pointing device (for example the mouse) sampling rate is the reason for this phenomena.
+\label{Prepross} 
  Speed and time difference information were widely used in sketch recognition systems \cite{earlyprocess}.
@@
 \end{figure}
-%is calculated as the angle between two vectors and  where $Q_i=\overrightarrow {P_{i - 1} P_i }$ is the vector from the point $P_{i - 1}$ to point $P_i$ 
+
 Dominant points are characterized by low speed values, high curvature and direction values.
@@
 The following sections describe in details the ellipse detection algorithm and the two segmentation algorithms used to divide the stroke.
-% The input to these algorithms are the stroke points along with the possible dominant points $P_{pd}$ computed during preprocessing.
-The output is a set of dominant points which are connected by either lines or curves.
-%After two different segmentations are generated, 
-%///AllBlockDiagram.jpg  you can change the block to the main block. flowchart
  \begin{figure}
@@
 The output is either the stroke fitted as an ellipse or the best segmentation generated from the two segmentation algorithms.
-% First, attempts to fit stroke as an ellipse.
-Then two different segmentation are generated using (AlgS1 and AlgS2).
-Finaly, the Minimum error of segmentation as the best segmentation.
 	\label{fig:segblock}
@@
 
-% \begin{figure}
-%	\centering
-%		\includegraphics[scale=0.48]{images/blockSmall.jpg}
-%	\caption{Segmentation Block} 
-%	\label{fig:segblock}
-%\end{figure}
+\subsection{Ellipse fitting}
 
-\subsection{Ellipse fitting}
-%The process starts by computing the center of the stroke bounding box.
-The bounding box center point is used as the first estimation of the ellipse center.
-The axes of the ellipse are estimated as $width/2$ and $height/2$ of the stroke bounding box.
-The least square fitting algorithm is used to minimize the fitting error of the ellipse equation \cite{chernov}.
 This process attempts to fit the stroke points into an ellipse arc; it starts by computing the center of the stroke bounding box.
@@
  \begin{equation}
-%eff= (P_{percent}/E)
-%eff = \left\{ {\begin{array}{*{20}c}
-%   {\frac{{P_{percent} }}{E}} & {ifP_{percent}  < 0.5}  \\
-%   {\frac{{P_{percent} }}{{E \times 2}}} & {ifP_{percent}  \ge 0.5}  \\
-%\end{array}} \right,
  eff= \begin{cases} 
@@
 \displaystyle  \frac{P_{percent}}{2\times E},&if P_{percent}\geq 0.5 \\
-% \frac{P_{percent}}{2\times E},&\mbox{if} P_{percent}\mbox{\geq0.5} 
+
 \end{cases},
@@
 	\centering
-		%\subfigure[Possible dominate point]{ 	\label{fig:LabelsPPD}	\includegraphics[scale=0.7]{images/ppd.jpg}}
-	%	\hfill
  	 \includegraphics[scale=0.7]{images/pso1.jpg}			
@@
 Length L is the distance between a stroke point and the line connecting two successive segment points.
-%represent the length   % a) Possible dominate point b) Particle encoding
 	\label{fig:pso1}
@@
 		\item \emph{X intersections:} The number of X intersections.
-		%\item[No. of holes] Number of holes in the symbols. 
+
 \end{itemize}
@@
 			 {\label{fig:ELsymbolSet}\includegraphics[scale=0.45]{images/EelectImage.jpg}}
-%			 \label{fig:ELsymbolSet} 
-%			 \fbox{ 
-%			 \parbox{4cm}{\includegraphics[scale=0.5]{images/EelectImage.jpg}	}}}
 
@@
 	 	\subfigure[Segmentation Error Vs. Swarm Iterations]{	\label{fig:swarmtesting} \includegraphics[scale=0.3]{images/swarmIterationVsParticles.jpg}}
-	 %\includegraphics[scale=0.3]{images/swarmAgentTest.jpg}} 	\caption{}
-
-	%\caption{Experiments results :} a)    b) c) 
-	%\end{figure}
-
-%\begin{figure}
-	%\centering
 
@@
 	 	\caption{Segmentation Error Vs. PSO parameters.}a) The error decrease as the number of swarm iteration increase until it reach a saturation.   b) The behavior of segmentation error versus c1.
-
-	%\Segmentation Error Vs. PSO parameters. caption{Experiments results :} a)    b) c)
 
@@
 \end{table*}
-% \begin{table*}
-% 	\centering
-% 	%\small
-%	\scalebox{0.7}{
-% 		%	\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|c|c|c|c|}\hline 
-%  \include{conf2}
-%  		%\end{tabular}
-%  	%}
-% 	\caption{Confusion Matrix of DL-DS}
-% 	\label{tab:ConfusionMatrix}
-% 	%\end{minipage}
-% \end{table*}
  \subsection{Features Analysis}
@@
 
- \begin{figure*}[h]
+ \begin{figure*}
 	\centering
@@
 The figure shows that \textsl{AlgS2} achieve better results.
-%effect of the segmentation algorithm
- %Figure \ref{fig:SampleSeg} shows a sample of the output of the segmentation block.
-The result of segmentation block is tested by their effect on the final recognition results using the features that will be highly affected using different algorithms like \textbf{FS1} and \textbf{FS4} (Section\ref{sec:Recognition}).
-Figure \ref{fig:testFeatonly} shows the result of the different algorithms using only \textbf{FS1} features.
 \begin{figure*} 
@@
 c) An example of multiple strokes and their segmentation.
-%The figure shows sample of the output of the segmentation phase.
-Figure a) shows the stroke after the user draws it.
-The red points are points labeled as \textit{possible dominate points}.
-Figure b) shows the symbol after being segmented.
-Figure c) shows the segmentation of different and overlapped strokes.
-Notice the system mark the center of the ellipse that was identified.
+identified.
 	\label{fig:SampleSeg}
@@
 The system uses multiple preliminary information, i.e., speed and curvature information, as input to the \textit{DPSO}.
-It is noted that the \textit{DPSO} in general generates an optimized stroke segmentation which improves the final recognition rate.
-The tradeoff between accuracy achieved and time complexity can be further investigated to achieve better results.
+Results show that the \textit{DPSO} in general generates an optimized stroke segmentation which improves the final recognition rate.
 The final recognition is based on both global shape and hierarchical stroke properties.
@@
 Global shape and moment properties were used to prevent the need to use graph matching which is computationally expensive \cite{SketchRead2007}.
-The experiments evaluated two\textit{DPSO} different segmentation algorithms (AlgS1 and AlgS2) on three different datasets (Hs-Ds, EL-Ds and LD-Ds).
+The experiments evaluated two \textit{DPSO} different segmentation algorithms (AlgS1 and AlgS2) on three different datasets (Hs-Ds, EL-Ds and LD-Ds).
 Results show that AlgS2 achieves better segmentation and final accuracy than Alg3 \cite{earlyprocess} on the datasets used.
 The results proved that the system is easily expandable to more and more symbols as it does not depend on low level primitive recognizers but on high level symbol and geometrical features.
-The system achieved an overall improvement of more than 2\% over Alg3 in all datasets.
-%The system was tested on 13 different symbols and achieved an overall accuracy of 89.5\%.
-The system does not depend on low level recognizer but rather on a set of high level features.
-This makes the system easily expandable to more and more symbols.
+The system achieved an average overall improvement of more than 2\% over Alg3 in all datasets.
 
- A possible extension of this research is to complete the clustering algorithm for fully automated sketch recognition.
+  The tradeoff between accuracy achieved and time complexity can be further investigated to achieve better results.
 Currently, the system processes one symbol at a time, an enhancement would be to allow users to draw more than one symbol at a time and incorporate a method for separating symbols.
+Thus, possible extension of this research is to complete the clustering algorithm for fully automated sketch recognition.
 Another area of enhancements is the features extraction methods.
 Introducing more spatial and geometrical features is believed to improve classifications.
-Features that represents the appearance of the symbols are likely to improve the recognition that have similar geometrical structure (i.e.
+Features that represent the appearance of the symbols are likely to improve the recognition of symbols that have similar geometrical structure (i.e.
 Nor and OR gate in logic symbols).

### 2009-06-10 paper2/Paper2.tex
figure comment removal
after chanign and addin omega image

@@
   They are features that define the structure of the geometrical symbol such as;   
-%\begin{description}
-%	 \item[No. Of segments] Number of segments in the symbol. 
-%		\item[No. Of primitives] Number of primitives in the symbol. The feature helps when identifying             symbols with mixed geometric primitives like cylinders and callouts.  
-%		\item [No. of curves] Number of curves or ellipses in the symbol. 
-%		\item [No. of lines] Number of lines in the symbol. 
-%		\item [No. of perpendicular lines] Number of perpendicular lines. 
-%		\item [No. of parallel lines] Number of parallel lines. 
-%		\item [No. of intersections] Number of intersection between lines and curves. 
-%		\item [Size Ratio] Ratio between width to height of the symbol.
-%		%\item[No. of holes] Number of holes in the symbols. 
-%
-%\end{description}
  \begin{itemize}
@@
 \end{figure}
-
-%\begin{figure}
-%\centering 
-%\label{fig:LogicsymbolSet}
-% 
-%		\includegraphics[scale=0.5]{images/logicSet.jpg}	 
-%		\caption{The Digital Design Symbols Dataset(LD-DS)} The dataset consist of symbols that are visually simillar (i.e. has low shape diversity). 
-%\end{figure}
-%\begin{figure}
-%\centering 
-%\label{fig:ELsymbolSet}
-%\fbox{ \parbox{4.5cm}{% 
-%		\includegraphics[scale=0.5]{images/EelectImage.jpg}	}}
-%		\caption{The Electrical Symbols Dataset(EL-DS)} . The dataset consist of complex toplogical symbols.
-%\end{figure}
 
@@
 The final number of particles and maximum swarm iterations used in the system was based on a tradeoff between error achieved versus the computation time.
-The effect of other algorithm parameters such as $c_1$ (mentioned in Section\ref{sec:ParticleSwarmAlgorithm}) is tested using similar tests (Figure \ref{fig:swarmtesting}, \ref{fig:swarmtesting2}).%, $c_2$
+The effect of other algorithm parameters such as $c_1$, $\omega$ (mentioned in Section\ref{sec:ParticleSwarmAlgorithm}) is tested using similar tests (Figure  \ref{fig:swarmtesting2},\ref{fig:swarmwtest}).%, $c_2$
 
@@
 	 	\subfigure[Segmentation Error Vs. Swarm Iterations]{	\label{fig:swarmtesting} \includegraphics[scale=0.3]{images/swarmIterationVsParticles.jpg}}
-	 %\includegraphics[scale=0.3]{images/swarmAgentTest.jpg}} 	\caption{}
 
-	%\caption{Experiments results :} a)    b) c) 
-	%\end{figure}
+		\subfigure[Segmentation Error Vs. $c_1$]{	\label{fig:swarmtesting2}  \includegraphics[scale=0.3]{images/C1Swarm.jpg}} %swarmParamterC.jpg
 
-%\begin{figure}
-	%\centering
+				\subfigure[Segmentation Error Vs. $\omega$]{\label{fig:swarmwtest}  \includegraphics[scale=0.3]{images/wSwarm.jpg}} %swarmParamterC.jpg
 
-		\subfigure[Segmentation Error Vs. c1]{	\label{fig:swarmtesting2}  \includegraphics[scale=0.4]{images/swarmParamterC.jpg}}
-
-	 	\caption{Segmentation Error Vs. PSO parameters.}a) The error decrease as the number of swarm iteration increase until it reach a saturation.   b) The behavior of segmentation error versus c1.
-
-	%\Segmentation Error Vs. PSO parameters. caption{Experiments results :} a)    b) c)
+	 	\caption{Segmentation Error Vs.
+PSO parameters.}a) The error decrease as the number of swarm iteration increase until it reach a saturation.
+b) The behavior of segmentation error versus $c_1$.
+c) The behavior of segmentation error versus  $\omega$
 
@@
 This behavior affects Ls-Ds and EL-Ds more than Hs-Ds due to the higher topological and structure nature of the symbols in both datasets than of Hs-Ds.
-%   can be explained by the simplicity of the symbols in Hs-DS % and LD-Ds unlike EL-DS where AlgS2 is the best algorithm.
-This behavior is explained by the structure of the symbols in each dataset.
-El-DS consist of highly topological and structure symbols that are usually drawn using multiple small strokes.
-This leads to    %   out perform any single algorithm in HS-DS and LD-DS.
-%show that combining algorithms achieve higher accuracy than any single algorithm.
-%improves the results.% with both the \textit{DPSO} algorithms.% The results show that both \textit{DPSO} algorithms achieve better result than other algorithms.
 
- \begin{figure}[h]
+ \begin{figure} 
 	\centering		
 	\subfigure[Algorithm comparison on HS-DS]{\label{fig:test1} \includegraphics[scale=0.4]{images/testAlg.jpg}}
-	 %	\caption{} The recognition rate of different algorithms.
-
-	%\caption{Experiments results :} a)    b) c) 
-%\end{figure}
-
-%\begin{figure}[h]
-%	\centering		
  		\subfigure[Algorithm comparison on LD-DS]{ 	 	\label{fig:testLD} \includegraphics[scale=0.4]{images/LDAlg.jpg}}
-	% 	\caption{} The recognition rate of different algorithms.
-
-	%\caption{Experiments results :} a)    b) c) 
-%\end{figure}
-
-%\begin{figure}[h]
-%	\centering		
  		\subfigure[Algorithm comparison on EL-DS]{	 	\label{fig:testEL}  \includegraphics[scale=0.4]{images/AlgEL.jpg}}
 	 	\caption{ The recognition rate of different algorithms. }
-
-	%\caption{Experiments results :} a)    b) c) 
  \end{figure}
@@
 This leads us to the next experiment to choose the best set of feature to recognize symbols.
-\begin{figure*}[h]
+\begin{figure*} 
 	\centering 
@@
 		\includegraphics[scale=0.35]{images/testsym.jpg}}
-	%\caption{} The effect of symbol complexity.  %The graph shows the recognition rate of each symbol using different algorithms. 
-%\end{figure*}  
-%\begin{figure*}
-%	\centering
 \subfigure[Symbols Comparison on LD-DS symbols]{\label{fig:LDtest2}
@@
 
- \begin{figure*}[h]
+ \begin{figure*}
 	\centering
 	\subfigure[Feature Comparison on Hs-DS]{	\label{fig:testFeaturesAll} \includegraphics[scale=0.5]{images/featuresdifference.jpg}}
-    %The graph shows the recognition rate of each symbol using different algorithms.
-
-%\end{figure*}
-
- %\begin{figure*}
-	%\centering
-
 	\subfigure[Feature Comparison on EL-DS]{	\label{fig:ELtestFeaturesAll}	\includegraphics[scale=0.34]{images/FeatEl.jpg}}
@@
 The figure shows that \textsl{AlgS2} achieve better results.
-%effect of the segmentation algorithm
- %Figure \ref{fig:SampleSeg} shows a sample of the output of the segmentation block.
-The result of segmentation block is tested by their effect on the final recognition results using the features that will be highly affected using different algorithms like \textbf{FS1} and \textbf{FS4} (Section\ref{sec:Recognition}).
-Figure \ref{fig:testFeatonly} shows the result of the different algorithms using only \textbf{FS1} features.
-\begin{figure}[h] 
+\begin{figure*} 
 	\centering
-		\includegraphics[scale=0.3]{images/OnlyF1EL.jpg}
+		\includegraphics[scale=0.45]{images/OnlyF1EL.jpg}
 	\caption{The effect of segmentation on accuracy} The experiment is implemented on Hs-Ds using only FS1.
@@
 	\label{fig:testFeatonly}
-\end{figure}
+\end{figure*}
 

### 2009-06-10 trunk/paper2/Paper2.tex
after finihsing

@@
 
-		\subfigure[Segmentation Error Vs. c1]{	\label{fig:swarmtesting2}  \includegraphics[scale=0.4]{images/swarmParamterC.jpg}}
+		\subfigure[Segmentation Error Vs. $c_1$]{	\label{fig:swarmtesting2}  \includegraphics[scale=0.3]{images/C1Swarm.jpg}} %swarmParamterC.jpg
 
-	 	\caption{Segmentation Error Vs. PSO parameters.}a) The error decrease as the number of swarm iteration increase until it reach a saturation.   b) The behavior of segmentation error versus c1.
+				\subfigure[Segmentation Error Vs. $\omega$]{\label{fig:swarmwtest}  \includegraphics[scale=0.3]{images/wSwarm.jpg}} %swarmParamterC.jpg
+
+	 	\caption{Segmentation Error Vs.
+PSO parameters.}a) The error decrease as the number of swarm iteration increase until it reach a saturation.
+b) The behavior of segmentation error versus $c_1$.
+c) The behavior of segmentation error versus  $\omega$
 
@@
 		\includegraphics[scale=0.5]{images/DLSymbols.jpg}}
+		\subfigure{ 
+		\includegraphics[scale=0.5]{images/captionSymbols.jpg}}
 	%\caption{}.  %The graph shows the recognition rate of each symbol using different algorithms. 
@@
 	\centering
-		\includegraphics[scale=0.3]{images/OnlyF1EL.jpg}
+		\includegraphics[scale=0.45]{images/OnlyF1EL.jpg}
 	\caption{The effect of segmentation on accuracy} The experiment is implemented on Hs-Ds using only FS1.

### 2009-07-17 trunk/draft/SketchMaster.tex
save to version before defense

@@
                                                     % APPROVAL                                                         %======================
-%\thispagestyle{empty}               % Don't do headers or footers.
-%\newpage                                                             
-%\begin{center}     
-%\LARGE  
-%\textbf{Sketch Recognition using Particle Swarm Optimization Algorithm}\\
-%\normalsize by      \\ 
-%\Large  \textbf{ Maha mohamed Nabeel El Meseery}\\
-%\vspace*{0.4in}        
-%\normalsize
-%A Thesis Submitted To the\\
-%Faculty of Engineering, Cairo University\\
-%in Partial Fulfillment of the\\
-%Requirements for the Degree of\\
-%MASTER OF SCIENCE
-%in\\
-%COMPUTER ENGINEERING\\
-%
-%\end{center}
-%          
-%    Approved by\\                                           
-%Examining Committee:\\
-% \makebox[6in]{\hrulefill}\\ 
-%Professor Nevin Darwish, Thesis Main Advisor\\
-% \makebox[6in]{\hrulefill}\\ 
-%Professor Samia Mashali,Thesis Advisor \\
-% \makebox[6in]{\hrulefill}\\  
-%Assoc. Professor Magda Fayek,Thesis Advisor\\ 
-% \makebox[6in]{\hrulefill}\\ 
-%Professor  Ashraf Abd El Wahab, Member\\
-% \makebox[6in]{\hrulefill}\\ 
-%Professor Mohamed Zaki Abd ElMegeed, Member\\
-% \makebox[6in]{\hrulefill}\\ 
+\thispagestyle{empty}               % Don't do headers or footers.
+\newpage                                                             
+\begin{center}     
+\LARGE  
+\textbf{Sketch Recognition and Understanding}\\
+\normalsize by      \\ 
+\Large  \textbf{ Maha Mohamed Nabeel El Meseery}\\
+\vspace*{0.4in}        
+\normalsize
+A Thesis Submitted To the\\
+Faculty of Engineering, Cairo University\\
+in Partial Fulfillment of the\\
+Requirements for the Degree of\\
+MASTER OF SCIENCE
+in\\
+COMPUTER ENGINEERING\\
+
+\end{center}
+
+    Approved by\\                                           
+Examining Committee:\\
+ \makebox[6in]{\hrulefill}\\ 
+Professor Nevin Darwish, Thesis Main Advisor\\
+ \makebox[6in]{\hrulefill}\\ 
+Professor Samia Mashali,Thesis Advisor \\
+ \makebox[6in]{\hrulefill}\\  
+Assoc. Professor Magda Fayek,Thesis Advisor\\ 
+ \makebox[6in]{\hrulefill}\\ 
+Professor  Ashraf Abd El Wahab, Member\\
+ \makebox[6in]{\hrulefill}\\ 
+Professor Mohamed Zaki Abd ElMegeed, Member\\
+ \makebox[6in]{\hrulefill}\\ 
   % \\   you need...\\

### 2009-07-17 trunk/smallpaper/SmallPaper.tex
changing the text for the  datasets

--- 2009-01-06 trunk/writing/outline.tex
@@
-\documentclass[a4paper,10pt]{IEEEconf}
-\usepackage[latin1]{inputenc}
-\usepackage[american]{babel}
-%\usepackage[T1]{fontenc}
+\documentclass{article}
+\usepackage{spconf}
+\usepackage{amsmath}
 \usepackage[dvips,pdftex]{graphicx}
-%\usepackage{multicol}
 \usepackage{subfigure}
 \usepackage{named}
-%\usepackage{fancybox}
-%\usepackage{natbib}
-\bibliographystyle{named}
 \usepackage[dvips,pdftex]{hyperref}
-%opening
-\title{Symbol Recognition using Particle Swarm Algorithms}
-%\title{A Study on Using Particle Swarm Algorithms in Symbol Recognition }
-\author{Maha El Meseery,  Mahmoud Fakhr El Din,  Samyah Mashaly,
-\\  Magdah Fayek,  Nevin Darwish}
+\title{Sketch Recognition using Particle Swarm Algorithms}
+\twoauthors
+  {Maha El Meseery, Mahmoud Fakhr El Din\sthanks{Associate Professor}, Samia Mashali\sthanks{Professor}}
+	{Signals Processing Group \\ Computers and Systems Department \\
+  Electronic Research Institute\\Cairo, Egypt\\
+	melmeseery@eri.sci.eg, samia@eri.sci.eg}
+  {Magda Fayek\sthanks{Associate Professor}, Nevin Darwish\sthanks{Professor}}
+	{ Computer Engineering Department\\
+	Faculty of Engineering \\Cairo University\\Cairo, Egypt\\
+	magdafayek@gmail.com,ndarwish@ieee.org}
 \begin{document}
@@
 Sketch recognition is defined as the process of identifying symbols that users draw using single or multiple strokes.
-Users draw strokes using a pen and the system should immediately interprets their strokes into objects that can be easily manipulated.
+Users draw strokes using a pen and the system immediately interprets their strokes as objects that can be easily manipulated.
 This paper uses Particle Swarm Algorithm (PSO) to divide the strokes the user draws into meaningful geometric primitives.
@@
 The results show that using PSO improves segmentation results which guide the symbol recognition phase.
-This paper uses SVM classifier which further improves the final recognition accuracy.
-%u Final recognition has improved accuracy after the PSO algorithm using a SVM classifier.
- %Hand drawn sketches are the natural method of exchanging thoughts between engineers and humans in general.
-Nevertheless, researchers barely give attention to improve current gesture and sketch recognition systems.
-In a sketch understanding system, users need to feel the freedom of paper and pencil along with the flexibility of computer manipulation.
+This paper uses Support Vector Machines (SVM) classifier which further improves the final recognition accuracy.
 \end{abstract}
 \section{Introduction}
-%A picture speak a thousand words is also true for sketches. 
+The field of symbol recognition has gained interest in the last few years.
 Scientists generally and engineers specifically express thoughts and designs using sketches.
 Engineers use sketches to exchange designs as a natural method of communication rather than writing or speaking.
-Engineers use paper and pencil design in early stages of design.
 Design engineers need a powerful computer based system with symbol recognition to help them design, manipulate and store sketches more effectively than using only papers.
-Increasing the interaction between computers and users in sketch and CAD systems has been the reason for the emerging of few advanced sketch recognition systems \cite{CAD}.
-Sketch recognition system converts the drawings and informal graphs user draws into objects and symbols that can be used in further interaction with the system.
+Increasing the interaction between computers and users in sketch and CAD systems has been the reason for the emerging of a few advanced sketch recognition systems.
 
-Sketch recognition is divided into four main steps preprocessing, segmentation, symbol recognition and finally sketch understanding.
-The preprocessing process saves user input stroke points from the hardware and collects basic information about the stroke then proceeds to remove noise and primary data calculation.
-In segmentation phase, the system divides the stroke into a set of simple geometrical primitives or segments.
-The third phase, sketch recognition, is clustering strokes and segments to formulate symbols that can be recognized by a classifier system.
-Finally, the sketch understanding phase uses priori information about the sketch to support identification of symbols by excluding any possible invalid symbols.
-%the basic symbols are identified in consistent with the whole sketch. \\
-%from the user information from the users converts the strokes the user draw into segments which is recognized into basic symbols that can be easily manipulated.
-Finally, the basic symbols are interpreted in consistent with the whole sketch.
-\\
-This paper introduces a sketch recognition system.
-The presented system uses' particle swarm algorithm (PSO) to optimally segment users strokes.
+Sketch recognition is divided into three main steps preprocessing, segmentation and symbol recognition.
+The preprocessing phase captures user input stroke points and collects basic information about the stroke then proceeds to remove noise and compute basic statistical and geometrical information.
+In segmentation phase, the strokes are divided into a set of simple geometrical primitives or segments.
+In the third phase, sketch recognition, strokes and segments are clustered to formulate symbols that can be recognized by a classifier system.
+
+A wide variety of techniques were suggested for sketch recognition.
+A hybrid algorithm was introduced in \cite{earlyprocess} where they generated different sets of segments based on both curvature and speed dominant points, followed by choosing a segmentation with the least error from a generated hybrid set.
+Their system is however limited to recognizing only specific simple geometric shapes with a set of low level recognizers.
+Each low level recognizer is designed to recognize only one geometric shape using spatial and geometric information extracted from the stroke.
+Yu \cite{meanshift10}introduced a feature area for each primitive and then computed the segmentation error for different types of primitives based on the computed feature area.
+His system achieved good accuracy in simple shapes (square, ellipse,...) but did not perform well in more complex shapes.
+Paulson and Hammond\cite{Paleosketch08} introduced a set of low level recognizers that were reported to achieve 98\%  accuracy but their system, similar to all low level recognizers, only identify a small set of simple shapes.
+A genetic algorithm was used by \cite{CruveDivisionSwarm} to optimally divide digital curves into lines and curves.
+Chen et al.\cite{CruveDivisionSwarm} uses digital curves scanned from paper as input to the system and did not take advantage of the curvature or local geometric properties of the digital curve.
+
+This paper introduces a new sketch recognition system, which uses particle swarm algorithm (PSO) to optimally segment user's strokes.
 Users can draw symbols using any number of strokes in any order they like.
-The system segments stroke user draws then cluster symbols together then passes them to the SVM symbol recognizer.
-The experiments in section \ref{sec:Experiments} shows that the proposed system improves the final symbol recognition accuracy.
+The system segments each stroke using two different PSO algorithms.
+Segmentation is based on dividing the digital curve into polygon\cite{PolygonApproximationPSO}.
+The algorithm enhances its performance by adding curvature and other local information while segmenting strokes to achieve better results.
+To handle complex strokes, our system uses a second segmentation PSO algorithm to segment strokes into a set of lines and circular arcs.
+After segmentation, a symbol features vector composed of different statistical, geometrical and spatial features is deduced.
+The final classification step, SVM classifier uses the computed feature vector to classify the segments into one of the previously trained classes.
 
-The remaining of this papers is organized as following; section \ref{Sec:RelatedWork} provides a brief literature review.
-Section \ref{PSO} explains the general particle swarm algorithm which is used in the presented system.
-The proposed system block diagram along with a detailed system description is presented in Section \ref{Sysdisc}.
-The experiments and results achieved are explained in Section \ref{sec:Experiments}.
-Finally, section \ref{ConclusionandFutureWork} states the conclusion and future work.
-%The next section will relate previous work in the field.
-The remaining of the paper will be as following; Section \ref{PSO} will explain the general particle swarm algorithm.
-Section \ref{Sysdisc} will describe the system block diagram and then section \ref{Prepross} will explain the preprocessing step, section \ref{seg} will describe the segmentation algorithm in details.
-Finally, section \ref{sec:Experiments} will describe the experiments that was preformed.\\
-\section{Literature Review}
-\label{Sec:RelatedWork}
-%%%What is sketch recognitions. Why we need sketch recognition systems. 
-The field of symbol recognition has gained interest in the last few years. A wide variety of techniques were used either on segmentation process or symbol recognition.
-%To segment strokes
-%In order to segment strokes, \cite{meanshift10} introduced a feature area for each primitive and then computed the segmentation error for different types of primitives based on this area.
-%\citeauthor{HeloiseBeautification} used a dynamic programming algorithm to solve the segmentation problem \cite{HeloiseBeautification}.
-A hybrid algorithm was introduced in \cite{earlyprocess} where they generated different sets of segmentation based on both curvature and speed dominant points, this is followed by choosing a segmentation with the least error from generated hybrid set.
-Their system has a draw back as it only recognizes simple specific shapes with a set of low level recognizers.
-\cite{meanshift10} introduced a feature area for each primitive and then computed the segmentation error for different types of primitives based on the computed feature area.
-Their system achieved good accuracy in simple shapes (square, ellipse,...) but did not perform well in more complex shapes\cite{meanshift10,Paleosketch08}.
-\citeauthor{Paleosketch08} \cite{Paleosketch08} introduced a set of low level recognizers that were reported to achieve 98\%  accuracy but their system similar to all low level recognizers only identify a small set of simple shapes.
-
-Many different approaches have been investigated to achieve final recognition of symbols.
-Graph searching and template matching where used in \cite{GraphBased07} where a graph of the drawn symbol is generated and matched with a stored set of graphs templates to reach a classification decision.
-These systems may generate good performance but they are very computationally expensive.
-Geometrical and spatial descriptive language was used in \cite{SketchRead2007,Ladder30} where an architecture similar to compiler was used to recognize symbols.
-The drawback of this architecture is that the recognition system can only parse symbols that were hard coded using the descriptive language.
-New symbols are not recognized unless the description of the symbol is provided or generated using another system.
-%%SVM and HMM classification was implemented in\cite{SVM300,HMM53} to correctly classify symbols.
-
-Genetic algorithm was used by \cite{CruveDivisionSwarm} to optimally divide digital curves into lines and curves.
-\citeauthor{CruveDivisionSwarm} uses digital curves scanned from paper as input to the system and did not take advantage of the curvature or local geometric properties of the digital curve.
-\citeauthor{PolygonApproximationPSO} \cite{PolygonApproximationPSO} used PSO to convert digital curves into polygons, our system adopts \citeauthor{PolygonApproximationPSO} method but improves it by adding curvature and other local information while segmenting strokes to achieve better segmentations.
-The next section presents the general particle swarm algorithm which is used in our proposed segmentation algorithms.
-% are presented in section \ref{subsubsec:Discreteparticleswarmalgorithm}
-% Other techniques as  [ ] was used  in []. \\  
-%In the next section the particle swarm algorithm will be explained.
-After that the system block description and detailed algorithms will be described.
-We will then proceed to describe the experiments preformed and the achieved results.
-%\subsection{Sketch Recognitions systems}
-%Details of Sketch systems. 
-%Related work and research in this area. 
-%Writing on the pros and cons of current systems.  
-%Write on the difficulties of the sketch recognition systems. 
-  %Some tried to detect dominant or corner points based on data generated from speed   \cite{earlyprocess} and mostly used curvature of the stroke drawn \cite{earlyprocess,meanshift10}. 
-%Particle swarm algorithms was used in \cite{} to fit a set of points into ellipse or higher order curves.
-%The first set of features introduced to be used on gesture recognition was introduced by Rubine in \cite{gestureexample12}.
-later,  Spatial relations and features where used in  \cite{cognitivesketch18,SRGraph57}  .
-\citeauthor{GeometryAndDomain102} introduced ink density feature which they used.
-The use of Zernike moments  was used  as the only features to the SVM classifiers in \cite{HeloiseBeautification}.
-\cite{SketchRead2007} used a set of features varied from relative spatial features and some compositional features.
-\\
-%After segmentation researchers varies on the methods in which to continue.
-\section{Particle Swarm Algorithm}
-\label{PSO}
-%What is particle swarm algorithm and how it was used in related researches. 
-The main idea of Particle Swarm Algorithm (PSO) is to represent each agent with a particle from the solution space.
-Each agent moves the particle with a direction and velocity based on equation \ref{eq:Swarm}.
- %Equation [\ref{eq:Swarm}] shows how velocity and direction of each particle are computed
- \begin{equation}
-v_{ij}  = v_{ij}  + c_1 r_1 (lbest_{ij}  - p_{ij} ) + c_2 r_2 (gbest_{ij}  - p_{ij} )
-\label{eq:Swarm}
-\end{equation}
- where $r_1$ \& $r_2$ are random variables and $c_1$ \& $c_2$ are the swarm system variables.
-\begin{equation}
-%\[
-p_{ij}=p_{ij}+v_{ij},
-%\
-\end{equation}
-where $p_{ij}$ represent the $jth$ particle in the $ith$ agent and $v_{ij}$ is the velocity of the $jth$ particle in the $ith$ agent.
-After each iteration the global best $g_{best}$ particle and the agent local best $l_{best}$ particle are evaluated based on the maximum fitness functions of all particles.
-The solution is found after achieving a specific number of iteration or after an error threshold is achieved.
-The following equation \ref{eq:descrite}  
-\begin{equation}
-   P(i)\Leftarrow 
-\{
-\begin{array}{c} 
-1 \quad \quad if\quad r_{3}>t  \\
-
-0 \quad \quad if\quad r_{3}<t 
-\label{eq:descrite}
-\end{array}\}
-\end{equation}
- (where $t$ is a threshold and $r_{3}$ is a random variable), is used in order to change the general swarm algorithm into binary particle which handles particle values of either $0$ or $1$.
-
-\section{System Details}
+ The remaining of the paper is as following; Section \ref{Sysdisc} describes the system in details.
+Sections \ref{Prepross}, \ref{seg} and \ref{sec:Recognition} provide details on preprocessing, segmentation and recognition process respectively.
+Section \ref{sec:Experiments} describes the experiments that were preformed.
+We conclude in Section \ref{ConclusionandFutureWork}.
+\section{The proposed system}
 \label{Sysdisc}
-%The stroke is the path of points from the instant the pen is down till it is up.
-Figure \ref{fig:Blockdiagram} shows that the main system blocks are 1) preprocessing, 2) segmentation, 3) clustering, 4) feature extraction, and finally 5) classification.
-
- %Due to sloppiness of users and hardware glitches the system preprocess the data before further processing. 
-The preprocessing phase is responsible of capturing the input data and removing the noise from it.
-The system then proceeds to estimate a set of possible dominant points to help in the segmentation process.
-% The time differences, direction, speed and curvature of all points of the stroke is computed along with some other geometrical data.
-The system then proceed to estimate a set of possible dominant points to help in segmentation process.
-\\
-  %then passes through the segmentation process which will try to
-
- %The goal of the segmentation stage is to divide stroke into segments of either curves or lines.
-The segmentation is divided into two steps ellipse fitting and curve segmentation.
-In the first step, the ellipse fitting process tries to fit the stroke into an ellipse.
-If the system fails, it passes the stroke to the second step; curve segmentation which consists of two PSO segmentation algorithms.
-The two algorithms generate two segmentations.
-The system then chooses the segmentation that has the minimum segmentation error.%generate the segmentations,  the minimum error will be the chosen segmentation.
-% is the ellipse fitting
-
- %the stroke first If the ellipse detection fails the stroke is passed to the segmentation algorithms which will pass it to the two PSO algorithms described below the segmentation with the minimum error will be the chosen segmentation.
-The segmentation is then added to the set of un-recognized segments in the system.
-\\% This part is repeated for each stroke.
- The clustering algorithm starts to group segments together after the segmentation step.
-The system allows the user draws the symbol by using any number of strokes; a set of unrecognized segments is passed to the clustering algorithm to generate a symbol and compute a feature vector for it.
-A set of hybrid features of statistical, geometric and spatial features is used to generate the symbol feature vector.
-The final classification step, SVM classifier uses the feature vector computed to classify the segments into one of the previously trained classes.
-The following sections provides a detail description of each step.%  Or to determine the symbol of the given segments from the set of preciously trained symbols.
- %with a symbol from the training set.% The system compute composite set of features some are statistical other are spatial features based on the type primitives.
-Finally, the strokes are classified into the corresponding classes.% the classifier identifies the strokes into a symbol from the set of known symbols.\\
-
-\begin{figure}[]
-%\begin{minipage}[b]{0.5\linewidth}
-	\centering
-		\includegraphics[scale=0.75]{images/block2.pdf}%Blockdiagram.pdf}
-	\caption{The system Block Diagram.} The figure shows the main system blocks. Preprocessing, segmentation, clustering, features extraction and classification.
-
-	\label{fig:Blockdiagram}
-%	\end{minipage}
-\end{figure}
-
+ The suggested sketch recognitions system is divided into three main steps 1) preprocessing, 2) segmentation and 3) Recognition. The following section provides a detailed description of each step.
 \subsection{Preprocessing}
 \label{Prepross}
-%Some data is extracted from the points in the stroke.
- It is noted that as the user draws a shape, the pen slows down near corners and picks up speed when drawing a straight lines.
-Therefore, the speed and time difference data were widely used in sketch understanding systems \cite{earlyprocess}.
-The direction and curvature data are used to determine the points with high angular changes along the path of points \cite{meanshift10}.
- %speed is calculated as 
- In the presented system, we computed time difference, direction, speed and curvature of each point along the stroke.
-The speed is calculated as  $v=\Delta t/\Delta s$ where $t$ is the time difference between two points and $s$ is the length between them.
-The direction is calculated as the angle between two vectors and curvature is considered as the change in direction with respect to length i.e.
+ Speed and time difference information were widely used in sketch understanding systems \cite{earlyprocess}.
+It was noted that time difference provides more distinct maxima than the speed information.
+Agar \cite{polygonfeedback31} mentions that the pointing device (for example the mouse) sampling rate is the reason for this phenomena.
+ In the presented system, we computed time difference, direction, speed and curvature of each point along a stroke.
+The speed is calculated as $v=\Delta s/\Delta t$ where $t$ is the time difference between two points and $s$ is the length between them.
+The direction $d$ is calculated as the angle between two vectors and curvature is considered as the change in direction with respect to length i.e.
 $c= \Delta d/\Delta s$.
+Dominant points are characterized by lower speed values and higher curvature and direction values.
 
- %system compute speed, distance data for all points in the stroke. Curvature is computed using estimation used in [] where direction computed is angle between two lines. Curvature is 
- %$\Delta d/  \Delta S $ where d is the difference direction of point and s is difference in distance between points. 
- It should be noted that, all these calculations are performed in real time while the user draws the strokes. The complexity of computation is $O(n)$ where $n$ is number of points. \\
- Figure \ref{fig:speed2} shows the computed data for the stroke drawn in fig. \ref{fig:orignalStroke}. %If you look at the data curves you will see by looking at the data curves s
- It is noted that the dominant points are characterized by lower speed values and higher curvature and direction values (fig.
-\ref{fig:speed2}).
-\\% the lower speed points correspond to vertex and dominant point's location.
-The higher the direction and curvature data correspond to location with higher change in curvature witch promote the location for vertices.
-\\
-After the system computes all the speed, time difference and curvature data it proceeds to detect the points with low velocity and high curvatures.
-Using simple differentiations to detect local extreme points resulted in false points due to the non smooth curves.
-Hence, the system adopted a process presented by \cite{earlyprocess}, where the mean
-%So instead the following procedure was used, firstly the mean of the curve is used as threshold for lower values then the high values are further investigated to find the extreme points.
-Hence, the mean
-of the curve is calculated and used as threshold \textit{th}.
-The threshold \textit{th} is used to separate the curve into regions  $Region_i$; each region $Region_i$ is defined as the part between two intersection points of the threshold line and the curve.
-Regions higher than the threshold \textit{th} are processed to find the maximum point $Max(Region_i)$ of each region $Region_i$.
+After the system computes all the speed, time difference and curvature information it proceeds to detect the points with low velocity and high curvatures.
+Using simple differentiation to detect local extreme points resulted in false points due to the non smooth curves.
+Hence, the system adopted a process presented by \cite{earlyprocess}, where the mean of the curve is calculated.
+Then a threshold \textit{th} is used to separate the curve into regions $Region_i$; each region $Region_i$ is defined as a range of points, where the curve values are either above or below the threshold \textit{th}.
+Those regions are further processed to find the maximum point $Max(Region_i)$ of each region $Region_i$.
 The stroke points $p_i(x,y)$ that correspond to those maximum values are labeled as \textit{possible dominant points} $P_{pd}$.
-\\
-   %If we tried to differentiate the curve the result will be false threshold values we divide the data into regions of data higher or lower that the threshold.
-This will let us only look for high data only.
-The maximum of each region is then computed and reported as a possible vertex point.
-\begin{figure}[]
+The system repeats this process for curvature, time difference and speed information.
+All the points labeled as possible dominant points $P_{pd}$ are saved into a single array.
+Figure \ref{fig:LabelsPPD} shows the particles labeled as Possible dominate point $P_{pd}$ by the preprocessing, it is noticed the redundancy of some $P_{pd}$ points.
+% (as shown are redundant)
+\begin{figure}
 	\centering
-		\includegraphics[scale=0.5]{images/orignalStroke.pdf}
-	\caption{Stroke Sample} An example of the stroke drawn in the system.
-	\label{fig:orignalStroke}
+		\subfigure[Posible domninate point]{ 	\label{fig:LabelsPPD}	\includegraphics[scale=0.5]{images/ppd.jpg}}
+		\hfill
+			\subfigure[DPSO algorithm encoding]{	\label{fig:LabelsPSO} \includegraphics[scale=0.5]{images/pso1.jpg}}			
+	\caption{Segmentation}% a) Possible dominate point b) Particle encoding  
+	\label{fig:pso1}
 \end{figure}
-The system repeats this process to curvature, time difference and speed curves.
-All the points labeled as possible dominant points $P_{pd}$ are saved into a single array.
-They are then used as input to guide the next segmentation phase.
-\begin{figure*}[]
-%\begin{minipage}[b]{0.9\linewidth}
-	\centering
-
-		\subfigure[ The Speed of data] {\includegraphics[scale=0.25]{images/speed2.pdf}}	
-	%	\subfigure[ Direction ] {\includegraphics[scale=0.15]{images/direction2.pdf}}
-					%\end{minipage}
-					\hspace{20 mm}
-					%\begin{minipage}[b]{0.5\linewidth}
-			\subfigure[The Curvature] {\includegraphics[scale=0.29]{images/curvature2.pdf}}
-	%		\subfigure[ Time Difference] {\includegraphics[scale=0.15]{images/time2.pdf}}
-%			\end{minipage}
-	\caption{Data Curves} The curve in a) shows the speed of the points in stroke in fig. \ref{fig:orignalStroke}. b) Shows the curvature curve for the same stroke.  %{The data of the stroke}
-
-	\label{fig:speed2}
-\end{figure*}
-%Description of input data. 
-%how to calculate speed, curvature, area, bounding box.
-%how to remove noise. 
-%Finally how to compute primarily dominant points. 
 \subsection{Segmentation}
 \label{seg}
-After computing the primary data the system tries to segment the stroke into a set of primitives.
-The segmentation algorithm first attempts to detect if the stroke is an ellipse.
-If the stroke proved to be an ellipse then the segmentation process ends and the system proceeds to the next step.
-Otherwise, the stroke is passed into two particle swarm algorithms that divides the stroke to either lines or lines and curves.
-The algorithms take the stroke points along with the possible dominant points $P_{pd}$ computed then produce a set of dominant points which are connected with either lines or curves (see fig.
-\ref{fig:Blockdiagram}).
-The next section describes the ellipse detection algorithm and the two particle swarm algorithms used to divide the stroke.\\
-%Paragraphs describe the segmentation algorithm.  
+In the segmentation stage the stroke is divided into a set of primitives.
+As shown in Figure \ref{fig:segblock} first an attempt is made to fit the stroke points into a curve or an ellipse using a minimum square error fitting algorithm \cite{chernov-2003}.
+If the stroke proved to be an ellipse arc then the segmentation process ends and the system proceeds to the next step.
+Otherwise, the stroke is passed to two particle swarm algorithms that divide the stroke to either lines or lines and curves.
+The algorithms takes the stroke points along with the possible dominant points $P_{pd}$ computed during preprocessing then produce a set of dominant points which are connected by either lines or curves.
+The next section describes the ellipse detection algorithm and the two particle swarm algorithms used to divide the stroke.
+ \begin{figure}
+	\centering
+		\includegraphics[scale=0.55]{images/segblock.jpg}
+	\caption{Segmentation Block}
+	\label{fig:segblock}
+\end{figure}
 \subsubsection{Ellipse Detection} 
-%Describe the ellipse detection algorithms 
 The process starts by computing the center of the stroke bounding box.
@@
 The axes of the ellipse are estimated as $width/2$ and $height/2$ of the stroke bounding box.
-The least square fitting algorithm is used to minimize the fitting error of the ellipse equation eq.
-\ref{eq:circleFit}.
-\begin{equation}
-E = \sum\limits_{i = 0}^N {\frac{{(x_i - x_0 )}}{{a^2 }}^2  + \frac{{(y_i - y_0 )}}{{b^2 }}^2  - 1} 
-\label{eq:circleFit}
-\end{equation} where $N$ is the number of points in the stroke, $a,b$ are the length of ellipse axes, $x_0$ , $y_0$ are the coordinates of the center point and $x_i$ \& $y_i$ are the coordinates of point $i$ in the stroke.
-After few loops, the error and a confidence values  are computed to check if the stroke can be labeled as ellipse or not.\\ %is computed to check if stroke is an ellipse.
-\\
-%We found that if the confidante is above threshold then the probability of ellipse is highest otherwise the stroke is passed to the next section to get the divisions of stroke and test its error.  
+The least square fitting algorithm is used to minimize the fitting error of the ellipse equation \cite{chernov-2003}.
 \subsubsection{Discrete particle swarm algorithm (\textit{DPSO})}
@@
 Two DPSO algorithms are used to generate segmentations for each stroke the user draws.
-The segmentation with the minimum error value is chosen as the stroke segmentation.
+For each generated segmentation evaluated an error is determined.
+The segmentation with the minimum error value is chosen as the best stroke segmentation (Figure \ref{fig:segblock}).
 The problem definition is the same in both algorithms but they differ in the method they use to compute fitness and error functions.
-\subsubsection{Problem definition}
-The input stroke with $N$ points can be represented by set $S = \left\{ {x_1 ,x_2  \ldots x_N } \right\}$ where $x_i$ is the location of the point $i$.
-The swarm algorithms consist of $M$ agents which are represented by the set
-$A  = \left\{ {P_i \left| {i = 1,2 \cdots M} \right.} \right\}$ where $P_i$ is a single solution particle from the solution space.
-Each particle decodes the problem with binary array with the same length $N$ as the input stroke.
-% The particles of the swarm represent a single solution of the solution space.
-For this problem the, each particle %will give a different segmentation for the input stroke.
-Firstly, we will define the stroke with N points by the set %S where.
-We define the arc %An array with the same length as the number of points of the strokes.
+
+\textsl{Problem definition:} The input stroke with $N$ points can be represented by set $S = \left\{ {x_1 ,x_2  \ldots x_N }\right\}$ where $x_i$ is the location of the point $i$.
+The swarm algorithms consist of $M$ agents which are represented by the set  $A = \left\{ {P_i \left| {i = 1,2 \cdots M} \right.} \right\}$ where $P_i$ is a single solution particle from the solution space.
+Each particle decodes the problem using a binary array with the same length $N$ as the input stroke.
+
 Therefore, the system represents each particle $P_i$ by $P_i = \left\{ {p_{ij} \left| {j = 1,2 \cdots N} \right.} \right\}$ where $p_{ij}$ has only two values 0 or 1.
-When $p_{ij}=1$ means that point $j$ is a dominant point.
-% a!402B!
-%$A = \left\{ {p_i \left| {i = 1,2 \cdots N} \right.} \right\}$
-% a!402A!
-%\begin{equation}
-%A = \left\{ {p_i \left| {i = 1,2 \cdots N} \right.} \right\}.
-%\end{equation}
-\subsubsection{The algorithm}
-The fitness function and error calculation are different in each of the two \textit{DPSO} algorithms. For the first algorithm \textsl{AlgS1}, 
+$p_{ij}=1$ means that point $j$ is a dominant point.
+Figure\ref{fig:LabelsPSO} shows a particle encoding in the DPSO system.
+
+\textsl{Fitness function:} The fitness function and error calculation are different in each of the two \textit{DPSO} algorithms.
+
+For the \textit{first algorithm \textsl{AlgS1}}, 
  the arc $\widehat{x_ix_j}$ is defined as the consecutive set of points from point$x_i$ to point $x_{j}$ as in $x_i,x_{i+1} \cdots,x_j$.
 The line $\overline{x_i x_j} $ as the straight line connecting point $x_i$ to point $x_j$.
-The approximation error is computed by the equation \ref{eq:ErrorSwarm1}
+The approximation error is computed by the equation \ref{eq:ErrorSwarm1}.
+The graphical meaning of the error is shown in Figure\ref{fig:LabelsPSO}.
 \begin{equation}
@@
 \label{eq:ErrorSwarm1}
-\end{equation} where $M$ is the number of dominant points in this solution.
+\end{equation} where $M$ is the number of dominant points in this solution as generated by the swarm algorithm.
 The error $e ( \widehat{x_ix_j},\overline{x_i x_j})$ is computed as the sum of squared perpendicular distance from every point along the arc $\widehat{x_ix_j}$ to the line $\overline{x_i x_j}$.
-\\
 The fitness is computed using equation \ref{eq:fitnessSwarm1}\begin{equation}
@@
 \end{equation} where $N$ is the number of points in the stroke, $D$ is the number of points in the solution that was previously labeled as a possible dominant point ($P_{pd}$), $E$ is the computed error and $\varepsilon$ is the error threshold.
-It should be noticed that when the error is larger than the threshold $E$ the fitness is given a -ve value to lower the value of solution.
-Otherwise the system favors the lower number of vertices.\\ %we will want to lower the number of vertices.
-\\
-%if we say that s
-%Alg1:
-%Alg2: o
-The second algorithm \textsl{AlgS2} has the same problem formulation but different fitness and error functions.
-It was previously introduced in \cite{CruveDivisionSwarm} but a genetic programming was used as the optimizing algorithm.
-The error of both circle and line estimation are computed for each segment $S_i=\widehat{x_ix_j}$, the approximation with the lower error value is the chosen approximation of this segment $S_i$.
+It should be noticed that when the error is larger than the threshold $\varepsilon$ the fitness is given a -ve value to lower the fitness value of the solution.
+Otherwise the system favors the lower number of vertices.
+
+The \textit{second algorithm \textsl{AlgS2}} has the same problem formulation but different fitness and error functions are used.
+An attempt is made to fit each segment into line or circular arc, the error of both circle and line fit estimations are computed for each segment $S_i=\widehat{x_ix_j}$, the approximation with the lower error value is the chosen approximation of this segment $S_i$\cite{CruveDivisionSwarm}.
 The sum of the approximation error of all segments is the total error of the particle.
-The error is computed by equation \ref{eq:errorSwarm2} \begin{equation}
+The total error of the particle is computed by equation \ref{eq:errorSwarm2}
+ \begin{equation}
 E=\sum\nolimits_{i = 0}^M e(D_i) 
 \label{eq:errorSwarm2}
-\end{equation}where $M$ is the number of segments in the solution, $D_i$ is the minimum approximation error of curve and line approximations $min(d_c,d_l)$ as computed by \cite{CruveDivisionSwarm}.
+\end{equation}where $M$ is the number of segments in the solution as generated by the swarm algorithm, $D_i$ is the minimum approximation error of curve and line approximations $min(d_c,d_l)$ as computed by \cite{CruveDivisionSwarm}.
 The fitness is computed by the equation \ref{eq:fitnessSwarm2} \begin{equation}
@@
 \label{eq:fitnessSwarm2}
-\end{equation} where $E$ is the error and $M$ is the number of segments and $k$ is a parameter tweaked to get minimum number of segments. \\
-%Describe how we used particle swarm in curve polygon approximation using first algorithm.  
-%Describe how we used PSO in to divide stroke into segments and curves. 
-%[These two algorithm was modified from the algorithm in paper \cite{PolygonApproximationPSO,CruveDivisionSwarm}].
-\citeauthor{PolygonApproximationPSO} \cite{PolygonApproximationPSO} used a merge and divide algorithm after each loop of the swarm system to refine the solution but this paper is introducing another enhancement method.
+\end{equation} where $E$ is the error and $M$ is the number of segments and $k$ is a parameter tweaked to get minimum number of segments.
+The larger k is, the more effect the number of segments will have.
+For our system, k is selected to be 0.5\cite{CruveDivisionSwarm}.
+
 After each loop of the swarm algorithm (\textsl{AlgS1} and \textsl{AlgS2}), each particle is refined using the following procedure.
-For each particle $P_{ij}$ %loops on the set of selected dominant points to enhance the solution.
-each dominant point is checked to find if it was labeled before as a \textit{possible dominant point} $P_{pd}$ (computed as in section \ref{Prepross}), if it is not labeled the point is moved to the nearest labeled point.
+For each particle $P_i$ each dominant point $P_{ij}$ is checked to find if it was labeled before as a \textit{possible dominant point} $P_{pd}$ (computed as in section \ref{Prepross}).
+If it was not labeled the point $P_{ij}$ is moved to the nearest labeled point.
+This ensures that all of the points generated by the DPSO are possible dominate points $P_{pd}$.
 After that the particles are tested to make sure that the distance between every two successive dominate point is larger than the constant $min_D$.
-If two points are found that they are nearer than $min_D$ then based on the decrease in the error $E$ in eq.
-\ref{eq:ErrorSwarm1} \& \ref{eq:errorSwarm2} one of the points are either removed or moved to the nearest possible dominate point.
-%check the new solution and check the new vertices if a vertex points is a possible dominant point it is left as a vertex.
-Otherwise, the system will search in the local neighborhood for the nearest possible dominant point to move these vertices to.
-After each change the error is computed and if new error is less that the old then the new set is approved.
-\\
+If two points are nearer than $min_D$ then one of the points is removed.
 \subsection{Recognition}
-After the user draws all strokes of the symbol, he/she has to wait 10 seconds or press finish button beside the drawing area.
-The set of unrecognized strokes is grouped together along with their segmentation as input to the feature extraction process.
+\label{sec:Recognition}
+After the user draws all strokes of the symbol, the set of un-recognized strokes is grouped together along with their segmentation as input to the feature extraction process.
 A composite set of features is used to generate a single feature vector.
-The features used consist of Rubine feature set \cite{gestureexample12},  Zernike moments of order 21 \cite{HeloiseBeautification}, ink density \cite{GeometryAndDomain102} as well as some structural and spatial information like number of perpendiculars lines ,number of parallel lines and types of primitives in each symbol.
-After computing the features the symbol is introduced to the classifier.
-The system uses a support vector machine (SVM) classifier with Gaussian kernel (RBF kernel)\cite{libsvm}.\\%Some features like number of parallel, perpendicular and intersecting lines are computed from the segmentation other like density, centroid and Zernike moments are computed from the raw stroke points.
- %trained with the 13 categories in the system.
-The same training set was used by \cite{HeloiseBeautification} in their symbol recognition system.
-\\ %  The table describe the features used in the system.
-%\begin{table}
-%\begin{minipage}[t][1\totalheight]{0.3\columnwidth}
-%
-%\begin{tabular}{|l|l|l|}
-%\hline 
-%Type & Features & Notes\tabularnewline
-%\hline 
-%\hline Structural & No. of parallel lines & \tabularnewline
-%\hline & No. of perpendicular lines & \tabularnewline
-%\hline & No. of intersections & \tabularnewline
-%\hline Statistical & Mean and standard deviation of the Curvature data   & \tabularnewline
-%\hline & Mean of the speed data & \tabularnewline
-%\hline 
-%Rubine & Set of features \\for gesture recognition & Introduced by Rubine in \cite{} \tabularnewline
-%\hline Zernike &  A Zernike\\ moments  &  Used by \cite{zernike61} \tabularnewline
-%\hline Composite & width to\\ height ratio  & \tabularnewline
-%\hline  & ink density  & Used by \cite{} \tabularnewline
-%\hline
-%\end{tabular}		
-%\end{minipage}
-%	\caption{Feature Table}
-%	\label{tab:FeatureTable}
-%\end{table}
-  % Rubine el al.\cite{gestureexample12}  used a set of statistical features,  There are a lot of features used in the sketch recognition literature
-%How to handle multi-stroked symbols. 
-%The clustering algorithm.  
-%The classifier used in recognition
-%Describe the features used in recognition [ref to \cite{zernike61}]
-\section{Experiments}\label{sec:Experiments}
-A dataset collected by \citeauthor{HeloiseBeautification} is used to test the presented system\cite{HeloiseBeautification}.
-The data are drawn by 16 users each of them was asked to draw the 13 shapes from 30 to 50 times.
+The features used consist of Rubine feature set,  Zernike moments of order 21 \cite{HeloiseBeautification}, ink density as well as some structural and spatial information like number of perpendiculars lines ,number of parallel lines and number of different types of primitives in each symbol.
+After computing the features the symbol is introduced to the (SVM) classifier.
+\section{Experimental}
+\label{sec:Experiments}
+
+The dataset collected by Hse and Newton\cite{HeloiseBeautification} is used to test the presented system.
 Figure \ref{fig:symbolSet} shows a set of the shapes used in the data set.
-We divide the dataset into four sets each set contains four different users.
-We divide the dataset equally into training set and test set.
-To generate trusted results, four different splits for the training and test data are generated from the dataset.
-The results displayed are the recognition accuracy resulted from the classifier after recognizing the symbols.
-In the following section, the recognition accuracy means the average recognition accuracy in the four data splits.
-%The result presented in this section are the average result of the four splits.
-%then result are stored The results displayed are the recognition accuracy resulted from the classifier after recognizing the symbols.
-%Furthermore, each set is divided into train and test set where the trained set where used to train the classifiers and the test set where used for testing.
-The algorithms are then tested using the four sets and the results are averaged.
-%We divide the dataset equally into training set and test set.
-The results displayed are the recognition accuracy resulted from the classifier after recognizing the symbols.
-\\
-%We divided the dataset into four sets each set contains four different users.
-Furthermore, each set is divided into train and test set where the trained set where used to train the classifiers and the test set where used for testing.
-The algorithms are then tested using the four sets and the results are averaged.
-\begin{figure}[]\centering
-\fbox{  \parbox{3cm}{% 
-		\includegraphics[scale=0.4]{images/symbolSet.pdf}	}}
-	\caption{The Symbol Set} Ths figure shows a set of the symbols used in the dataset. 
-	\label{fig:symbolSet}
-\end{figure}
-\begin{figure}[]
+The datasets were divided into training and test sets.
+Five different splits for the training and test data are generated from each dataset.
+The results displayed are the average recognition accuracy of the five splits.
+The accuracy is computed as the number of correctly recognized samples divided by the total number of samples in the test.
+
+ \begin{figure*}[]
 	\centering
-		\includegraphics[scale=0.5]{images/test1.jpg}
-	\caption{Algorithm comparison} The results of comparing the algorithms. The graphs displays the recognition rate of symbols. 
-	\label{fig:test1}
-\end{figure}
-%In the figure \ref{fig:test1}, the accuracy means that final hit rate of all symbols used in the test set.by the  four datasets . 
-We performed two experiments to test the system firstly we tested recognition accuracy of shapes in the data set with both algorithms (\textsl{AlgS1} and \textsl{AlgS2}).
-We also implemented the segmentation algorithm described in \cite{earlyprocess} (\textsl{Alg3}) to use as reference to our swarm algorithms.
+	\begin{minipage}[b]{5cm}	
+\begin{center}
+\subfigure [The Symbol Set] {\label{fig:symbolSet}
+\fbox{ \parbox{2cm}{% 
+		\includegraphics[scale=0.2]{images/symbolSet.PNG}	}}}
+
+		\subfigure[Algorithm comparison]{\includegraphics[width=5cm]{images/test1smallpaper.jpg} 	\label{fig:test1}}
+		\end{center}
+		\end{minipage}
+		\hfill
+  \subfigure [Symbols Comparison] {\includegraphics[width=9cm]{images/test2smallpaper.jpg} \label{fig:test2}}
+	\caption{Experiments results :} a) the figure shows a set of the symbols used in the dataset.
+b)The results of comparing the algorithms.
+The graph displays the recognition rate of symbols.
+c) The graph shows the recognition rate of each symbol using different algorithms.
+\end{figure*} 
+We performed two experiments to test the system.
+Firstly we tested recognition accuracy of shapes in the data set with both algorithms (\textsl{AlgS1} and \textsl{AlgS2}).
 Figure \ref{fig:test1} shows the accuracy achieved by each algorithm.
-The results shows that both \textit{DPSO} algorithms achieve better result than other algorithms.
-The swarm algorithms were tested with and without the ellipse detection module.
-The ellipse detection module appears to be superior to results with only the \textit{DPSO} algorithms.\\
+The two swarm algorithms were tested with and without the ellipse fitting module.
+Including the ellipse detection module improves the results with both the \textit{DPSO} algorithms.% The results show that both \textit{DPSO} algorithms achieve better result than other algorithms.
 
-The second experiment we implemented is testing the effect of symbol complexity and type on the recognition rate.
+The second experiment we implemented was to test the effect of symbol complexity and type on the recognition rate.
 Figure \ref{fig:test2} shows the accuracy of each symbol, it is clearly noted that symbols that have only line segments achieve higher accuracy rate than other symbols.
@@
 This is natural as algorithm \textsl{AlgS1} divides strokes into line segments only but \textsl{AlgS2} is able to divide strokes into lines and curves based on the minimum error of the segment itself.
-Algorithm \textsl{Alg3} gives good performance as long as the symbols consist of lines and curves, if the stroke consists of curve only  the algorithm may lead to wrong segmentation result.
-This is because the system divides the stroke first to line segments then tries to decide if each segment can be represent better as a curve unlike algorithm \textsl{AlgS2} where the curve segments are tested while choosing the best segmentation.
-Combining both algorithm \textsl{AlgS1} and \textsl{AlgS2} improved the recognition rate of all symbols because of combining the advantages of both algorithms.
-The penalty for this increased performance is in the computational time needed to run both the swarm algorithms.
-\\
-\begin{figure*}[]
-	\centering
-		\includegraphics[scale=0.8]{images/test2.jpg}
-	\caption{Symbols Comparison} The graph shows the recognition rate of each symbol using different algorithms. 
-	\label{fig:test2}
-\end{figure*}
-%We could not test the result of the segmentation algorithm directly due to the fact that the correct segmentation is highly ambiguous.
-It is only based on what the user intends to draw while sketching the symbol.
-Figure \ref{fig:SampleSeg} shows a sample of the output of the segmentation block.
-  %The final experiment we performed was to test the efficiency svm classifier.
-We implemented the linear discriminator classifier described in \cite{gestureexample12}(see figure \ref{fig:test3}).
-%The result shows that SVM is more scalable than the linear discriminator for this problem.
-%The test focus on the size of data needed to test each classifier and the achieved accuracy.
-%For this experiment we also shows the classification time for both classifiers
-%\begin{figure}[]
-%	\centering
-% 
-%				\subfigure[User strokes] {\includegraphics[scale=0.7]{images/strokebefore.jpg}}
-%		 	\subfigure[Segmented strokes] {\includegraphics[scale=0.7]{images/strokeafter.jpg}}
-%		 		\subfigure[Multiple symbols] {\includegraphics[scale=0.7]{images/results3.jpg}}
-%		
-%	\caption{Segmentation outputs} The figure shows sample of the output of the segmentation phase.
-Figure a) shows the stroke after the user draws it.
-The red points are points labeled as \textit{possible dominate points}.
-Figure b) shows the symbol after being segmented.
-Figure c) shows the segmentation of different and overlapped strokes.
-Notice the system mark the center of the ellipse that was identified.
-%	\label{fig:SampleSeg}
-%\end{figure}
-%\begin{figure}[]
-%	\centering
-%		\includegraphics[scale=0.4,clip=true, viewport=0 380 700 1280]{images/test1.pdf}
-%	\caption{The results of third experiment}
-%	\label{fig:test3}
-%\end{figure}
- %Details description of results got from comparing various algorithms  
- %Describe data set used for experiments ( No of data set, No. of categories, No of samples).
- %Describe other Algorithms [Alg 3 in paper \cite{earlyprocess}]used for comparing with the one described in the paper. 
+ Combining both algorithm \textsl{AlgS1} and \textsl{AlgS2} improved the recognition rate of all symbols.
+The penalty for this increased performance is in the computational time needed to run both swarm algorithms.
 \section{Conclusion and Future Work}
@@
 This paper presented a new approach to sketch recognition using PSO.
-The system uses both speed and curvature data which helps in improving the \textit{DPSO} algorithms over the original algorithms \cite{CruveDivisionSwarm,PolygonApproximationPSO}.
-It is noted that the \textit{DPSO}in general generates an optimized stroke segmentation which improves the final recognition rate.
+The system uses both speed and curvature information which help improving the \textit{DPSO}.
+It is noted that the \textit{DPSO} in general generates an optimized stroke segmentation which improves the final recognition rate.
 The tradeoff between accuracy achieved and time complexity must be further investigated to achieve better results.
 The use of statistical moments and structural features improves the recognition rate.
-The system was tested on 13 different symbols and achieved satisfying results.
-The system dose not depends on low level recognizer but rather on a set of high level features.
-This makes the system easily expandable with aspect to symbols recognized.
-\\
-%As you can see form the experiments PSO have proved a spuriously than other system. 
-%\section{Future Work}The next step in 
+The system was tested on 13 different symbols and achieved an accuracy of 91.5\%.
+The system does not depend on low level recognizer but rather on a set of high level features.
+This makes the system easily expandable to more and more symbols.
+
  A possible extension of this research is to complete the clustering algorithm for fully automated sketch recognition.
-The clustering must be performed without the user explicit involvement.
+Currently, the system process one symbol at a time, an enchantment would be to allow users to draw more than one symbol at a time and incorporate a method for separating symbols.
 Other area of enhancements is the features extraction methods.
 Introducing more spatial and geometrical features is believed to improve classifications.
-%Each new stroke is checked if it can be a symbol or is a part of already drawn un completed symbol.
-%\bibliography{Mybibliography}
-\bibliography{../../neededfiles/Bibliographies/Mybibliography2}
+\bibliographystyle{IEEEbib}
+\bibliography{../../neededfiles/Bibliographies/Mybibliography}
 \end{document}

### 2009-07-17 trunk/smallpaper/SmallPaper.tex
adding the alg3

@@
 \title{Sketch Recognition using Particle Swarm Algorithms}
+%\sthanks{Professor}\sthanks{Associate Professor}\sthanks{Professor}\sthanks{Associate Professor}
 \twoauthors
-  {Maha El Meseery, Mahmoud Fakhr El Din\sthanks{Associate Professor}, Samia Mashali\sthanks{Professor}}
+  {Maha El Meseery, Mahmoud Fakhr El Din, Samia Mashali}
 	{Signals Processing Group \\ Computers and Systems Department \\
@@
 	melmeseery@eri.sci.eg, samia@eri.sci.eg}
-  {Magda Fayek\sthanks{Associate Professor}, Nevin Darwish\sthanks{Professor}}
+  {Magda Fayek, Nevin Darwish}
 	{ Computer Engineering Department\\
@@
 
-		\subfigure[Algorithm comparison]{\includegraphics[width=5cm]{images/test1smallpaper.jpg} 	\label{fig:test1}}
+		\subfigure[Algorithm comparison]{\includegraphics[width=5cm]{images/testAlg.jpg} 	\label{fig:test1}}
 		\end{center}
@@
   \subfigure [Symbols Comparison] {\includegraphics[width=9cm]{images/test2smallpaper.jpg} \label{fig:test2}}
-	\caption{Experiments results :} a) the figure shows a set of the symbols used in the dataset.
-b)The results of comparing the algorithms.
-The graph displays the recognition rate of symbols.
-c) The graph shows the recognition rate of each symbol using different algorithms.
+	\caption{Experiments results :} a) The symbols set used in the dataset.   b) The recognition rate of different algorithms.  c) The recognition rate of each symbol using different algorithms. 
 \end{figure*} 
 We performed two experiments to test the system.
-Firstly we tested recognition accuracy of shapes in the data set with both algorithms (\textsl{AlgS1} and \textsl{AlgS2}).
+Firstly, we tested recognition accuracy of shapes in the data set with both algorithms (\textsl{AlgS1} and \textsl{AlgS2}).
+We also implemented the segmentation algorithm described in \cite{earlyprocess} (\textsl{Alg3}) to compare it with our swarm algorithms.
+Figure \ref{fig:test1} shows the accuracy achieved by each algorithm.
+The two swarm algorithms were tested with and without the ellipse fitting module.
+Combining the ellipse detection module improves the performance in the tested dataset.
+The result shows that (\textsl{AlgS2}) gives better performance than Alg3 and AlgS1.
+% Firstly we tested recognition accuracy of shapes in the data set with both algorithms (\textsl{AlgS1} and \textsl{AlgS2}).
 Figure \ref{fig:test1} shows the accuracy achieved by each algorithm.
@@
 \label{ConclusionandFutureWork}
-This paper presented a new approach to sketch recognition using PSO.
+This paper presented a new approach to sketch recognition using DPSO.
+The system uses multiple preliminary information, i.e., speed and curvature information, as input to the \textit{DPSO}.
+Results show that the \textit{DPSO} in general generates an optimized stroke segmentation which improves the final recognition rate.
+The final recognition is based on geomterical,statistical and structural features.
+The experiments evaluated two \textit{DPSO} different segmentation algorithms (AlgS1 and AlgS2) on simple presentation set datasets.
+Results show that AlgS2 achieves better segmentation and final accuracy than Alg3 \cite{earlyprocess} on the dataset used.
+The results proved that the system is easily expandable to more and more symbols as it does not depend on low level primitive recognizers but on high level symbol and geometrical features.
+The system achieved an average overall improvement of more than 2\% over Alg3.
+%This paper presented a new approach to sketch recognition using PSO.
 The system uses both speed and curvature information which help improving the \textit{DPSO}.

### 2009-07-20 smallpaper/SmallPaper.tex

--- 2009-01-06 trunk/writing/outline.tex
@@
-\documentclass[a4paper,10pt]{IEEEconf}
-\usepackage[latin1]{inputenc}
-\usepackage[american]{babel}
-%\usepackage[T1]{fontenc}
+\documentclass{article}
+\usepackage{spconf}
+\usepackage{amsmath}
 \usepackage[dvips,pdftex]{graphicx}
-%\usepackage{multicol}
 \usepackage{subfigure}
 \usepackage{named}
-%\usepackage{fancybox}
-%\usepackage{natbib}
-\bibliographystyle{named}
 \usepackage[dvips,pdftex]{hyperref}
-%opening
-\title{Symbol Recognition using Particle Swarm Algorithms}
-%\title{A Study on Using Particle Swarm Algorithms in Symbol Recognition }
-\author{Maha El Meseery,  Mahmoud Fakhr El Din,  Samyah Mashaly,
-\\  Magdah Fayek,  Nevin Darwish}
+\title{Sketch Recognition using Particle Swarm Algorithms}
+%\sthanks{Professor}\sthanks{Associate Professor}\sthanks{Professor}\sthanks{Associate Professor}
+\twoauthors
+  {Maha El Meseery, Mahmoud Fakhr El Din, Samia Mashali}
+	{Signals Processing Group \\ Computers and Systems Department \\
+  Electronic Research Institute\\Cairo, Egypt\\
+	melmeseery@eri.sci.eg, samia@eri.sci.eg}
+  {Magda Fayek, Nevin Darwish}
+	{ Computer Engineering Department\\
+	Faculty of Engineering \\Cairo University\\Cairo, Egypt\\
+	magdafayek@gmail.com,ndarwish@ieee.org}
 \begin{document}
@@
 Sketch recognition is defined as the process of identifying symbols that users draw using single or multiple strokes.
-Users draw strokes using a pen and the system should immediately interprets their strokes into objects that can be easily manipulated.
-This paper uses Particle Swarm Algorithm (PSO) to divide the strokes the user draws into meaningful geometric primitives.
+Users draw strokes using a pen and the system immediately interprets their strokes as objects that can be easily manipulated.
+This paper uses Particle Swarm Optimization Algorithm (PSO) to divide the strokes the user draws into meaningful geometric primitives.
 These geometric primitives are grouped to formulate symbols which are further identified.
 The results show that using PSO improves segmentation results which guide the symbol recognition phase.
-This paper uses SVM classifier which further improves the final recognition accuracy.
-%u Final recognition has improved accuracy after the PSO algorithm using a SVM classifier.
- %Hand drawn sketches are the natural method of exchanging thoughts between engineers and humans in general.
-Nevertheless, researchers barely give attention to improve current gesture and sketch recognition systems.
-In a sketch understanding system, users need to feel the freedom of paper and pencil along with the flexibility of computer manipulation.
+This paper uses Support Vector Machines (SVM) classifier which further improves the final recognition accuracy.
 \end{abstract}
 \section{Introduction}
-%A picture speak a thousand words is also true for sketches. 
+The field of symbol recognition has gained interest in the last few years.
 Scientists generally and engineers specifically express thoughts and designs using sketches.
 Engineers use sketches to exchange designs as a natural method of communication rather than writing or speaking.
-Engineers use paper and pencil design in early stages of design.
 Design engineers need a powerful computer based system with symbol recognition to help them design, manipulate and store sketches more effectively than using only papers.
-Increasing the interaction between computers and users in sketch and CAD systems has been the reason for the emerging of few advanced sketch recognition systems \cite{CAD}.
-Sketch recognition system converts the drawings and informal graphs user draws into objects and symbols that can be used in further interaction with the system.
+Increasing the interaction between computers and users in sketch and CAD systems has been the reason for the emerging of a few advanced sketch recognition systems.
 
-Sketch recognition is divided into four main steps preprocessing, segmentation, symbol recognition and finally sketch understanding.
-The preprocessing process saves user input stroke points from the hardware and collects basic information about the stroke then proceeds to remove noise and primary data calculation.
-In segmentation phase, the system divides the stroke into a set of simple geometrical primitives or segments.
-The third phase, sketch recognition, is clustering strokes and segments to formulate symbols that can be recognized by a classifier system.
-Finally, the sketch understanding phase uses priori information about the sketch to support identification of symbols by excluding any possible invalid symbols.
-%the basic symbols are identified in consistent with the whole sketch. \\
-%from the user information from the users converts the strokes the user draw into segments which is recognized into basic symbols that can be easily manipulated.
-Finally, the basic symbols are interpreted in consistent with the whole sketch.
-\\
-This paper introduces a sketch recognition system.
-The presented system uses' particle swarm algorithm (PSO) to optimally segment users strokes.
+Sketch recognition is divided into three main steps preprocessing, segmentation and symbol recognition.
+The preprocessing phase captures user input stroke points and collects basic information about the stroke then proceeds to remove noise and compute basic statistical and geometrical information.
+In segmentation phase, the strokes are divided into a set of simple geometrical primitives or segments.
+In the third phase, sketch recognition, strokes and segments are clustered to formulate symbols that can be recognized by a classifier system.
+
+A wide variety of techniques were suggested for sketch recognition.
+A hybrid algorithm was introduced in \cite{earlyprocess} where they generated different sets of segments based on both curvature and speed dominant points, followed by choosing a segmentation with the least error from a generated hybrid set.
+Their system is however limited to recognizing only specific simple geometric shapes with a set of low level recognizers.
+Each low level recognizer is designed to recognize only one geometric shape using spatial and geometric information extracted from the stroke.
+Yu \cite{meanshift10} introduced a feature area for each primitive and then computed the segmentation error for different types of primitives based on the computed feature area.
+His system achieved good accuracy in simple shapes (square, ellipse,...) but did not perform well in more complex shapes.
+Paulson and Hammond\cite{Paleosketch08} introduced a set of low level recognizers that were reported to achieve 98\%  accuracy but their system, similar to all low level recognizers, only identify a small set of simple shapes.
+One of the problems in these previous systems is that they only rely on curvature and speed data which are supjected to noise.
+They are also subjected to over and under segmentation problems due to use of thresholds.
+
+The segmentation problem is summarized as finding the best decomposition with least number of segments each represents a geometric primitive.
+This is an optimization problem, which evolutionary programming can solve efficiently.
+A genetic algorithm was used by \cite{CruveDivisionSwarm} to optimally divide digital curves into lines and curves.
+Chen et al.\cite{CruveDivisionSwarm} uses digital curves scanned from paper as input to the system and did not take advantage of the curvature or local geometric properties of the digital curve.
+
+This paper introduces a new sketch recognition system, which uses particle swarm algorithm (PSO) to optimally segment user's strokes.
 Users can draw symbols using any number of strokes in any order they like.
-The system segments stroke user draws then cluster symbols together then passes them to the SVM symbol recognizer.
-The experiments in section \ref{sec:Experiments} shows that the proposed system improves the final symbol recognition accuracy.
+The system segments each stroke using two different PSO algorithms.
+Segmentation is based on dividing the digital curve into polygon\cite{PolygonApproximationPSO}.
+The algorithm enhances its performance by adding curvature and other local information while segmenting strokes to achieve better results.
+To handle complex strokes, our system uses a second segmentation PSO algorithm to segment strokes into a set of lines and circular arcs.
+After segmentation, a symbol features vector composed of different statistical, geometrical and spatial features is deduced.
+The final classification step, SVM classifier uses the computed feature vector to classify the segments into one of the previously trained classes.
 
-The remaining of this papers is organized as following; section \ref{Sec:RelatedWork} provides a brief literature review.
-Section \ref{PSO} explains the general particle swarm algorithm which is used in the presented system.
-The proposed system block diagram along with a detailed system description is presented in Section \ref{Sysdisc}.
-The experiments and results achieved are explained in Section \ref{sec:Experiments}.
-Finally, section \ref{ConclusionandFutureWork} states the conclusion and future work.
-%The next section will relate previous work in the field.
-The remaining of the paper will be as following; Section \ref{PSO} will explain the general particle swarm algorithm.
-Section \ref{Sysdisc} will describe the system block diagram and then section \ref{Prepross} will explain the preprocessing step, section \ref{seg} will describe the segmentation algorithm in details.
-Finally, section \ref{sec:Experiments} will describe the experiments that was preformed.\\
-\section{Literature Review}
-\label{Sec:RelatedWork}
-%%%What is sketch recognitions. Why we need sketch recognition systems. 
-The field of symbol recognition has gained interest in the last few years. A wide variety of techniques were used either on segmentation process or symbol recognition.
-%To segment strokes
-%In order to segment strokes, \cite{meanshift10} introduced a feature area for each primitive and then computed the segmentation error for different types of primitives based on this area.
-%\citeauthor{HeloiseBeautification} used a dynamic programming algorithm to solve the segmentation problem \cite{HeloiseBeautification}.
-A hybrid algorithm was introduced in \cite{earlyprocess} where they generated different sets of segmentation based on both curvature and speed dominant points, this is followed by choosing a segmentation with the least error from generated hybrid set.
-Their system has a draw back as it only recognizes simple specific shapes with a set of low level recognizers.
-\cite{meanshift10} introduced a feature area for each primitive and then computed the segmentation error for different types of primitives based on the computed feature area.
-Their system achieved good accuracy in simple shapes (square, ellipse,...) but did not perform well in more complex shapes\cite{meanshift10,Paleosketch08}.
-\citeauthor{Paleosketch08} \cite{Paleosketch08} introduced a set of low level recognizers that were reported to achieve 98\%  accuracy but their system similar to all low level recognizers only identify a small set of simple shapes.
+The contribution in the system is mainly in attempting to generate the optimal segmentation using PSO which exhibited superiority with respect to simillar algoirthms in various other applications \cite{PolygonApproximationPSO}.
+The system attempts to find the optimial decomposition of the input stroke using PSO with the help of curvature and speed information.
+The use of PSO algorithm helps in eliminating the effects of input noise, problems of over and under segmentation that were in pervious systems.
+The proposed symbol recognizier uses the generated segmenation to compute the feature vector which is used to identify more complex symbols.
+This ensures that the system is scalable for more  complex shapes.
 
-Many different approaches have been investigated to achieve final recognition of symbols.
-Graph searching and template matching where used in \cite{GraphBased07} where a graph of the drawn symbol is generated and matched with a stored set of graphs templates to reach a classification decision.
-These systems may generate good performance but they are very computationally expensive.
-Geometrical and spatial descriptive language was used in \cite{SketchRead2007,Ladder30} where an architecture similar to compiler was used to recognize symbols.
-The drawback of this architecture is that the recognition system can only parse symbols that were hard coded using the descriptive language.
-New symbols are not recognized unless the description of the symbol is provided or generated using another system.
-%%SVM and HMM classification was implemented in\cite{SVM300,HMM53} to correctly classify symbols.
+ %The remaining of the paper is as following; Section \ref{Sysdisc} describes the system in details.
+Sections \ref{Prepross}, \ref{seg} and \ref{sec:Recognition} provide details on preprocessing, segmentation and recognition process respectively.
+Section \ref{sec:Experiments} describes the experiments that were preformed.
+We conclude in Section \ref{ConclusionandFutureWork}.
 
-Genetic algorithm was used by \cite{CruveDivisionSwarm} to optimally divide digital curves into lines and curves.
-\citeauthor{CruveDivisionSwarm} uses digital curves scanned from paper as input to the system and did not take advantage of the curvature or local geometric properties of the digital curve.
-\citeauthor{PolygonApproximationPSO} \cite{PolygonApproximationPSO} used PSO to convert digital curves into polygons, our system adopts \citeauthor{PolygonApproximationPSO} method but improves it by adding curvature and other local information while segmenting strokes to achieve better segmentations.
-The next section presents the general particle swarm algorithm which is used in our proposed segmentation algorithms.
-% are presented in section \ref{subsubsec:Discreteparticleswarmalgorithm}
-% Other techniques as  [ ] was used  in []. \\  
-%In the next section the particle swarm algorithm will be explained.
-After that the system block description and detailed algorithms will be described.
-We will then proceed to describe the experiments preformed and the achieved results.
-%\subsection{Sketch Recognitions systems}
-%Details of Sketch systems. 
-%Related work and research in this area. 
-%Writing on the pros and cons of current systems.  
-%Write on the difficulties of the sketch recognition systems. 
-  %Some tried to detect dominant or corner points based on data generated from speed   \cite{earlyprocess} and mostly used curvature of the stroke drawn \cite{earlyprocess,meanshift10}. 
-%Particle swarm algorithms was used in \cite{} to fit a set of points into ellipse or higher order curves.
-%The first set of features introduced to be used on gesture recognition was introduced by Rubine in \cite{gestureexample12}.
-later,  Spatial relations and features where used in  \cite{cognitivesketch18,SRGraph57}  .
-\citeauthor{GeometryAndDomain102} introduced ink density feature which they used.
-The use of Zernike moments  was used  as the only features to the SVM classifiers in \cite{HeloiseBeautification}.
-\cite{SketchRead2007} used a set of features varied from relative spatial features and some compositional features.
-\\
-%After segmentation researchers varies on the methods in which to continue.
-\section{Particle Swarm Algorithm}
-\label{PSO}
-%What is particle swarm algorithm and how it was used in related researches. 
-The main idea of Particle Swarm Algorithm (PSO) is to represent each agent with a particle from the solution space.
-Each agent moves the particle with a direction and velocity based on equation \ref{eq:Swarm}.
- %Equation [\ref{eq:Swarm}] shows how velocity and direction of each particle are computed
- \begin{equation}
-v_{ij}  = v_{ij}  + c_1 r_1 (lbest_{ij}  - p_{ij} ) + c_2 r_2 (gbest_{ij}  - p_{ij} )
-\label{eq:Swarm}
-\end{equation}
- where $r_1$ \& $r_2$ are random variables and $c_1$ \& $c_2$ are the swarm system variables.
-\begin{equation}
-%\[
-p_{ij}=p_{ij}+v_{ij},
-%\
-\end{equation}
-where $p_{ij}$ represent the $jth$ particle in the $ith$ agent and $v_{ij}$ is the velocity of the $jth$ particle in the $ith$ agent.
-After each iteration the global best $g_{best}$ particle and the agent local best $l_{best}$ particle are evaluated based on the maximum fitness functions of all particles.
-The solution is found after achieving a specific number of iteration or after an error threshold is achieved.
-The following equation \ref{eq:descrite}  
-\begin{equation}
-   P(i)\Leftarrow 
-\{
-\begin{array}{c} 
-1 \quad \quad if\quad r_{3}>t  \\
-
-0 \quad \quad if\quad r_{3}<t 
-\label{eq:descrite}
-\end{array}\}
-\end{equation}
- (where $t$ is a threshold and $r_{3}$ is a random variable), is used in order to change the general swarm algorithm into binary particle which handles particle values of either $0$ or $1$.
-
-\section{System Details}
+\section{The proposed system}
 \label{Sysdisc}
-%The stroke is the path of points from the instant the pen is down till it is up.
-Figure \ref{fig:Blockdiagram} shows that the main system blocks are 1) preprocessing, 2) segmentation, 3) clustering, 4) feature extraction, and finally 5) classification.
-
- %Due to sloppiness of users and hardware glitches the system preprocess the data before further processing. 
-The preprocessing phase is responsible of capturing the input data and removing the noise from it.
-The system then proceeds to estimate a set of possible dominant points to help in the segmentation process.
-% The time differences, direction, speed and curvature of all points of the stroke is computed along with some other geometrical data.
-The system then proceed to estimate a set of possible dominant points to help in segmentation process.
-\\
-  %then passes through the segmentation process which will try to
-
- %The goal of the segmentation stage is to divide stroke into segments of either curves or lines.
-The segmentation is divided into two steps ellipse fitting and curve segmentation.
-In the first step, the ellipse fitting process tries to fit the stroke into an ellipse.
-If the system fails, it passes the stroke to the second step; curve segmentation which consists of two PSO segmentation algorithms.
-The two algorithms generate two segmentations.
-The system then chooses the segmentation that has the minimum segmentation error.%generate the segmentations,  the minimum error will be the chosen segmentation.
-% is the ellipse fitting
-
- %the stroke first If the ellipse detection fails the stroke is passed to the segmentation algorithms which will pass it to the two PSO algorithms described below the segmentation with the minimum error will be the chosen segmentation.
-The segmentation is then added to the set of un-recognized segments in the system.
-\\% This part is repeated for each stroke.
- The clustering algorithm starts to group segments together after the segmentation step.
-The system allows the user draws the symbol by using any number of strokes; a set of unrecognized segments is passed to the clustering algorithm to generate a symbol and compute a feature vector for it.
-A set of hybrid features of statistical, geometric and spatial features is used to generate the symbol feature vector.
-The final classification step, SVM classifier uses the feature vector computed to classify the segments into one of the previously trained classes.
-The following sections provides a detail description of each step.%  Or to determine the symbol of the given segments from the set of preciously trained symbols.
- %with a symbol from the training set.% The system compute composite set of features some are statistical other are spatial features based on the type primitives.
-Finally, the strokes are classified into the corresponding classes.% the classifier identifies the strokes into a symbol from the set of known symbols.\\
-
-\begin{figure}[]
-%\begin{minipage}[b]{0.5\linewidth}
-	\centering
-		\includegraphics[scale=0.75]{images/block2.pdf}%Blockdiagram.pdf}
-	\caption{The system Block Diagram.} The figure shows the main system blocks. Preprocessing, segmentation, clustering, features extraction and classification.
-
-	\label{fig:Blockdiagram}
-%	\end{minipage}
-\end{figure}
-
+ The suggested sketch recognitions system is divided into three main steps 1) preprocessing, 2) segmentation and 3) Recognition. The following section provides a detailed description of each step.
 \subsection{Preprocessing}
 \label{Prepross}
-%Some data is extracted from the points in the stroke.
- It is noted that as the user draws a shape, the pen slows down near corners and picks up speed when drawing a straight lines.
-Therefore, the speed and time difference data were widely used in sketch understanding systems \cite{earlyprocess}.
-The direction and curvature data are used to determine the points with high angular changes along the path of points \cite{meanshift10}.
- %speed is calculated as 
- In the presented system, we computed time difference, direction, speed and curvature of each point along the stroke.
-The speed is calculated as  $v=\Delta t/\Delta s$ where $t$ is the time difference between two points and $s$ is the length between them.
-The direction is calculated as the angle between two vectors and curvature is considered as the change in direction with respect to length i.e.
+ Speed and time difference information were widely used in sketch understanding systems \cite{earlyprocess}.
+It was noted that time difference provides more distinct maxima than the speed information.
+Agar \cite{polygonfeedback31} mentions that the pointing device (for example the mouse) sampling rate is the reason for this phenomena.
+ In the presented system, we computed time difference, direction, speed and curvature of each point along a stroke.
+The speed is calculated as $v=\Delta s/\Delta t$ where $t$ is the time difference between two points and $s$ is the length between them.
+The direction $d$ is calculated as the angle between two vectors and curvature is considered as the change in direction with respect to length i.e.
 $c= \Delta d/\Delta s$.
+Dominant points are characterized by lower speed values and higher curvature and direction values.
 
- %system compute speed, distance data for all points in the stroke. Curvature is computed using estimation used in [] where direction computed is angle between two lines. Curvature is 
- %$\Delta d/  \Delta S $ where d is the difference direction of point and s is difference in distance between points. 
- It should be noted that, all these calculations are performed in real time while the user draws the strokes. The complexity of computation is $O(n)$ where $n$ is number of points. \\
- Figure \ref{fig:speed2} shows the computed data for the stroke drawn in fig. \ref{fig:orignalStroke}. %If you look at the data curves you will see by looking at the data curves s
- It is noted that the dominant points are characterized by lower speed values and higher curvature and direction values (fig.
-\ref{fig:speed2}).
-\\% the lower speed points correspond to vertex and dominant point's location.
-The higher the direction and curvature data correspond to location with higher change in curvature witch promote the location for vertices.
-\\
-After the system computes all the speed, time difference and curvature data it proceeds to detect the points with low velocity and high curvatures.
-Using simple differentiations to detect local extreme points resulted in false points due to the non smooth curves.
-Hence, the system adopted a process presented by \cite{earlyprocess}, where the mean
-%So instead the following procedure was used, firstly the mean of the curve is used as threshold for lower values then the high values are further investigated to find the extreme points.
-Hence, the mean
-of the curve is calculated and used as threshold \textit{th}.
-The threshold \textit{th} is used to separate the curve into regions  $Region_i$; each region $Region_i$ is defined as the part between two intersection points of the threshold line and the curve.
-Regions higher than the threshold \textit{th} are processed to find the maximum point $Max(Region_i)$ of each region $Region_i$.
+After the system computes all the speed, time difference and curvature information it proceeds to detect the points with low velocity and high curvatures.
+Using simple differentiation to detect local extreme points resulted in false points due to the non smooth curves.
+Hence, the system adopted a process presented by \cite{earlyprocess}, where the mean of the curve is calculated.
+Then a threshold \textit{th} is used to separate the curve into regions $Region_i$; each region $Region_i$ is defined as a range of points, where the curve values are either above or below the threshold \textit{th}.
+Those regions are further processed to find the maximum point $Max(Region_i)$ of each region $Region_i$.
 The stroke points $p_i(x,y)$ that correspond to those maximum values are labeled as \textit{possible dominant points} $P_{pd}$.
-\\
-   %If we tried to differentiate the curve the result will be false threshold values we divide the data into regions of data higher or lower that the threshold.
-This will let us only look for high data only.
-The maximum of each region is then computed and reported as a possible vertex point.
-\begin{figure}[]
+The system repeats this process for curvature, time difference and speed information.
+All the points labeled as possible dominant points $P_{pd}$ are saved into a single array.
+Figure \ref{fig:LabelsPPD} shows the particles labeled as Possible dominate point $P_{pd}$ by the preprocessing, it is noticed the redundancy of some $P_{pd}$ points.
+% (as shown are redundant)
+\begin{figure}
 	\centering
-		\includegraphics[scale=0.5]{images/orignalStroke.pdf}
-	\caption{Stroke Sample} An example of the stroke drawn in the system.
-	\label{fig:orignalStroke}
+		\subfigure[Posible domninate point]{ 	\label{fig:LabelsPPD}	\includegraphics[scale=0.5]{images/ppd.jpg}}
+		\hfill
+			\subfigure[DPSO algorithm encoding]{	\label{fig:LabelsPSO} \includegraphics[scale=0.5]{images/pso1.jpg}}			
+	\caption{Segmentation}% a) Possible dominate point b) Particle encoding  
+	\label{fig:pso1}
 \end{figure}
-The system repeats this process to curvature, time difference and speed curves.
-All the points labeled as possible dominant points $P_{pd}$ are saved into a single array.
-They are then used as input to guide the next segmentation phase.
-\begin{figure*}[]
-%\begin{minipage}[b]{0.9\linewidth}
-	\centering
-
-		\subfigure[ The Speed of data] {\includegraphics[scale=0.25]{images/speed2.pdf}}	
-	%	\subfigure[ Direction ] {\includegraphics[scale=0.15]{images/direction2.pdf}}
-					%\end{minipage}
-					\hspace{20 mm}
-					%\begin{minipage}[b]{0.5\linewidth}
-			\subfigure[The Curvature] {\includegraphics[scale=0.29]{images/curvature2.pdf}}
-	%		\subfigure[ Time Difference] {\includegraphics[scale=0.15]{images/time2.pdf}}
-%			\end{minipage}
-	\caption{Data Curves} The curve in a) shows the speed of the points in stroke in fig. \ref{fig:orignalStroke}. b) Shows the curvature curve for the same stroke.  %{The data of the stroke}
-
-	\label{fig:speed2}
-\end{figure*}
-%Description of input data. 
-%how to calculate speed, curvature, area, bounding box.
-%how to remove noise. 
-%Finally how to compute primarily dominant points. 
 \subsection{Segmentation}
 \label{seg}
-After computing the primary data the system tries to segment the stroke into a set of primitives.
-The segmentation algorithm first attempts to detect if the stroke is an ellipse.
-If the stroke proved to be an ellipse then the segmentation process ends and the system proceeds to the next step.
-Otherwise, the stroke is passed into two particle swarm algorithms that divides the stroke to either lines or lines and curves.
-The algorithms take the stroke points along with the possible dominant points $P_{pd}$ computed then produce a set of dominant points which are connected with either lines or curves (see fig.
-\ref{fig:Blockdiagram}).
-The next section describes the ellipse detection algorithm and the two particle swarm algorithms used to divide the stroke.\\
-%Paragraphs describe the segmentation algorithm.  
+In the segmentation stage the stroke is divided into a set of primitives.
+As shown in Figure \ref{fig:segblock} first an attempt is made to fit the stroke points into a curve or an ellipse using a minimum square error fitting algorithm \cite{chernov}.
+If the stroke proved to be an ellipse arc then the segmentation process ends and the system proceeds to the next step.
+Otherwise, the stroke is passed to two particle swarm algorithms that divide the stroke to either lines or lines and curves.
+The algorithms takes the stroke points along with the possible dominant points $P_{pd}$ computed during preprocessing then produce a set of dominant points which are connected by either lines or curves.
+The next section describes the ellipse detection algorithm and the two particle swarm algorithms used to divide the stroke.
+ \begin{figure}
+	\centering
+		\includegraphics[scale=0.55]{images/segblock.jpg}
+	\caption{Segmentation Block}
+	\label{fig:segblock}
+\end{figure}
 \subsubsection{Ellipse Detection} 
-%Describe the ellipse detection algorithms 
 The process starts by computing the center of the stroke bounding box.
@@
 The axes of the ellipse are estimated as $width/2$ and $height/2$ of the stroke bounding box.
-The least square fitting algorithm is used to minimize the fitting error of the ellipse equation eq.
-\ref{eq:circleFit}.
-\begin{equation}
-E = \sum\limits_{i = 0}^N {\frac{{(x_i - x_0 )}}{{a^2 }}^2  + \frac{{(y_i - y_0 )}}{{b^2 }}^2  - 1} 
-\label{eq:circleFit}
-\end{equation} where $N$ is the number of points in the stroke, $a,b$ are the length of ellipse axes, $x_0$ , $y_0$ are the coordinates of the center point and $x_i$ \& $y_i$ are the coordinates of point $i$ in the stroke.
-After few loops, the error and a confidence values  are computed to check if the stroke can be labeled as ellipse or not.\\ %is computed to check if stroke is an ellipse.
-\\
-%We found that if the confidante is above threshold then the probability of ellipse is highest otherwise the stroke is passed to the next section to get the divisions of stroke and test its error.  
+The least square fitting algorithm is used to minimize the fitting error of the ellipse equation \cite{chernov}.
 \subsubsection{Discrete particle swarm algorithm (\textit{DPSO})}
@@
 Two DPSO algorithms are used to generate segmentations for each stroke the user draws.
-The segmentation with the minimum error value is chosen as the stroke segmentation.
+For each generated segmentation evaluated an error is determined.
+The segmentation with the minimum error value is chosen as the best stroke segmentation (Figure \ref{fig:segblock}).
 The problem definition is the same in both algorithms but they differ in the method they use to compute fitness and error functions.
-\subsubsection{Problem definition}
-The input stroke with $N$ points can be represented by set $S = \left\{ {x_1 ,x_2  \ldots x_N } \right\}$ where $x_i$ is the location of the point $i$.
-The swarm algorithms consist of $M$ agents which are represented by the set
-$A  = \left\{ {P_i \left| {i = 1,2 \cdots M} \right.} \right\}$ where $P_i$ is a single solution particle from the solution space.
-Each particle decodes the problem with binary array with the same length $N$ as the input stroke.
-% The particles of the swarm represent a single solution of the solution space.
-For this problem the, each particle %will give a different segmentation for the input stroke.
-Firstly, we will define the stroke with N points by the set %S where.
-We define the arc %An array with the same length as the number of points of the strokes.
+
+\textsl{Problem definition:} The input stroke with $N$ points can be represented by set $S = \left\{ {x_1 ,x_2  \ldots x_N }\right\}$ where $x_i$ is the location of the point $i$.
+The swarm algorithms consist of $M$ agents which are represented by the set  $A = \left\{ {P_i \left| {i = 1,2 \cdots M} \right.} \right\}$ where $P_i$ is a single solution particle from the solution space.
+Each particle decodes the problem using a binary array with the same length $N$ as the input stroke.
+
 Therefore, the system represents each particle $P_i$ by $P_i = \left\{ {p_{ij} \left| {j = 1,2 \cdots N} \right.} \right\}$ where $p_{ij}$ has only two values 0 or 1.
-When $p_{ij}=1$ means that point $j$ is a dominant point.
-% a!402B!
-%$A = \left\{ {p_i \left| {i = 1,2 \cdots N} \right.} \right\}$
-% a!402A!
-%\begin{equation}
-%A = \left\{ {p_i \left| {i = 1,2 \cdots N} \right.} \right\}.
-%\end{equation}
-\subsubsection{The algorithm}
-The fitness function and error calculation are different in each of the two \textit{DPSO} algorithms. For the first algorithm \textsl{AlgS1}, 
+$p_{ij}=1$ means that point $j$ is a dominant point.
+Figure\ref{fig:LabelsPSO} shows a particle encoding in the DPSO system.
+
+\textsl{Fitness function:} The fitness function and error calculation are different in each of the two \textit{DPSO} algorithms.
+
+For the \textit{first algorithm \textsl{AlgS1}}, 
  the arc $\widehat{x_ix_j}$ is defined as the consecutive set of points from point$x_i$ to point $x_{j}$ as in $x_i,x_{i+1} \cdots,x_j$.
 The line $\overline{x_i x_j} $ as the straight line connecting point $x_i$ to point $x_j$.
-The approximation error is computed by the equation \ref{eq:ErrorSwarm1}
+The approximation error is computed by the equation \ref{eq:ErrorSwarm1}.
+The graphical meaning of the error is shown in Figure\ref{fig:LabelsPSO}.
 \begin{equation}
-E=\sum\nolimits_{i = 0}^M e ( \widehat{x_ix_{i+1}},\overline{x_i x_{i+1}})
+E=\sum\nolimits_{i = 1}^M e ( \widehat{x_ix_{i+1}},\overline{x_i x_{i+1}})
 \label{eq:ErrorSwarm1}
-\end{equation} where $M$ is the number of dominant points in this solution.
+\end{equation} where $M$ is the number of dominant points in this solution as generated by the swarm algorithm.
 The error $e ( \widehat{x_ix_j},\overline{x_i x_j})$ is computed as the sum of squared perpendicular distance from every point along the arc $\widehat{x_ix_j}$ to the line $\overline{x_i x_j}$.
-\\
 The fitness is computed using equation \ref{eq:fitnessSwarm1}\begin{equation}
@@
 \end{equation} where $N$ is the number of points in the stroke, $D$ is the number of points in the solution that was previously labeled as a possible dominant point ($P_{pd}$), $E$ is the computed error and $\varepsilon$ is the error threshold.
-It should be noticed that when the error is larger than the threshold $E$ the fitness is given a -ve value to lower the value of solution.
-Otherwise the system favors the lower number of vertices.\\ %we will want to lower the number of vertices.
-\\
-%if we say that s
-%Alg1:
-%Alg2: o
-The second algorithm \textsl{AlgS2} has the same problem formulation but different fitness and error functions.
-It was previously introduced in \cite{CruveDivisionSwarm} but a genetic programming was used as the optimizing algorithm.
-The error of both circle and line estimation are computed for each segment $S_i=\widehat{x_ix_j}$, the approximation with the lower error value is the chosen approximation of this segment $S_i$.
+It should be noticed that when the error is larger than the threshold $\varepsilon$ the fitness is given a -ve value to lower the fitness value of the solution.
+Otherwise the system favors the lower number of vertices.
+
+The \textit{second algorithm \textsl{AlgS2}} has the same problem formulation but different fitness and error functions are used.
+An attempt is made to fit each segment into line or circular arc, the error of both circle and line fit estimations are computed for each segment $S_i=\widehat{x_ix_j}$, the approximation with the lower error value is the chosen approximation of this segment $S_i$\cite{CruveDivisionSwarm}.
 The sum of the approximation error of all segments is the total error of the particle.
-The error is computed by equation \ref{eq:errorSwarm2} \begin{equation}
-E=\sum\nolimits_{i = 0}^M e(D_i) 
+The total error of the particle is computed by equation \ref{eq:errorSwarm2}
+ \begin{equation}
+E=\sum\nolimits_{i = 1}^M e(D_i) 
 \label{eq:errorSwarm2}
-\end{equation}where $M$ is the number of segments in the solution, $D_i$ is the minimum approximation error of curve and line approximations $min(d_c,d_l)$ as computed by \cite{CruveDivisionSwarm}.
+\end{equation}where $M$ is the number of segments in the solution as generated by the swarm algorithm, $D_i$ is the minimum approximation error of curve and line approximations $min(d_c,d_l)$ as computed by \cite{CruveDivisionSwarm}.
 The fitness is computed by the equation \ref{eq:fitnessSwarm2} \begin{equation}
@@
 \label{eq:fitnessSwarm2}
-\end{equation} where $E$ is the error and $M$ is the number of segments and $k$ is a parameter tweaked to get minimum number of segments. \\
-%Describe how we used particle swarm in curve polygon approximation using first algorithm.  
-%Describe how we used PSO in to divide stroke into segments and curves. 
-%[These two algorithm was modified from the algorithm in paper \cite{PolygonApproximationPSO,CruveDivisionSwarm}].
-\citeauthor{PolygonApproximationPSO} \cite{PolygonApproximationPSO} used a merge and divide algorithm after each loop of the swarm system to refine the solution but this paper is introducing another enhancement method.
+\end{equation} where $E$ is the error and $M$ is the number of segments and $k$ is a parameter tweaked to get minimum number of segments.
+The larger $k$ is, the more effect the number of segments will have.
+For our system, $k$ is selected to be $0.5$\cite{CruveDivisionSwarm}.
+
 After each loop of the swarm algorithm (\textsl{AlgS1} and \textsl{AlgS2}), each particle is refined using the following procedure.
-For each particle $P_{ij}$ %loops on the set of selected dominant points to enhance the solution.
-each dominant point is checked to find if it was labeled before as a \textit{possible dominant point} $P_{pd}$ (computed as in section \ref{Prepross}), if it is not labeled the point is moved to the nearest labeled point.
+For each particle $P_i$ each dominant point $P_{ij}$ is checked to find if it was labeled before as a \textit{possible dominant point} $P_{pd}$ (computed as in section \ref{Prepross}).
+If it was not labeled the point $P_{ij}$ is moved to the nearest labeled point.
+This ensures that all of the points generated by the DPSO are possible dominate points $P_{pd}$.
 After that the particles are tested to make sure that the distance between every two successive dominate point is larger than the constant $min_D$.
-If two points are found that they are nearer than $min_D$ then based on the decrease in the error $E$ in eq.
-\ref{eq:ErrorSwarm1} \& \ref{eq:errorSwarm2} one of the points are either removed or moved to the nearest possible dominate point.
-%check the new solution and check the new vertices if a vertex points is a possible dominant point it is left as a vertex.
-Otherwise, the system will search in the local neighborhood for the nearest possible dominant point to move these vertices to.
-After each change the error is computed and if new error is less that the old then the new set is approved.
-\\
+If two points are nearer than $min_D$ then one of the points is removed.
 \subsection{Recognition}
-After the user draws all strokes of the symbol, he/she has to wait 10 seconds or press finish button beside the drawing area.
-The set of unrecognized strokes is grouped together along with their segmentation as input to the feature extraction process.
+\label{sec:Recognition}
+After the user draws all strokes of the symbol, the set of un-recognized strokes is grouped together along with their segmentation as input to the feature extraction process.
 A composite set of features is used to generate a single feature vector.
-The features used consist of Rubine feature set \cite{gestureexample12},  Zernike moments of order 21 \cite{HeloiseBeautification}, ink density \cite{GeometryAndDomain102} as well as some structural and spatial information like number of perpendiculars lines ,number of parallel lines and types of primitives in each symbol.
-After computing the features the symbol is introduced to the classifier.
-The system uses a support vector machine (SVM) classifier with Gaussian kernel (RBF kernel)\cite{libsvm}.\\%Some features like number of parallel, perpendicular and intersecting lines are computed from the segmentation other like density, centroid and Zernike moments are computed from the raw stroke points.
- %trained with the 13 categories in the system.
-The same training set was used by \cite{HeloiseBeautification} in their symbol recognition system.
-\\ %  The table describe the features used in the system.
-%\begin{table}
-%\begin{minipage}[t][1\totalheight]{0.3\columnwidth}
-%
-%\begin{tabular}{|l|l|l|}
-%\hline 
-%Type & Features & Notes\tabularnewline
-%\hline 
-%\hline Structural & No. of parallel lines & \tabularnewline
-%\hline & No. of perpendicular lines & \tabularnewline
-%\hline & No. of intersections & \tabularnewline
-%\hline Statistical & Mean and standard deviation of the Curvature data   & \tabularnewline
-%\hline & Mean of the speed data & \tabularnewline
-%\hline 
-%Rubine & Set of features \\for gesture recognition & Introduced by Rubine in \cite{} \tabularnewline
-%\hline Zernike &  A Zernike\\ moments  &  Used by \cite{zernike61} \tabularnewline
-%\hline Composite & width to\\ height ratio  & \tabularnewline
-%\hline  & ink density  & Used by \cite{} \tabularnewline
-%\hline
-%\end{tabular}		
-%\end{minipage}
-%	\caption{Feature Table}
-%	\label{tab:FeatureTable}
-%\end{table}
-  % Rubine el al.\cite{gestureexample12}  used a set of statistical features,  There are a lot of features used in the sketch recognition literature
-%How to handle multi-stroked symbols. 
-%The clustering algorithm.  
-%The classifier used in recognition
-%Describe the features used in recognition [ref to \cite{zernike61}]
-\section{Experiments}\label{sec:Experiments}
-A dataset collected by \citeauthor{HeloiseBeautification} is used to test the presented system\cite{HeloiseBeautification}.
-The data are drawn by 16 users each of them was asked to draw the 13 shapes from 30 to 50 times.
+The features used consist of Rubine feature set,  Zernike moments of order 10 \cite{HeloiseBeautification}, ink density as well as some structural and spatial information like number of perpendiculars lines ,number of parallel lines and number of different types of primitives in each symbol.
+After computing the features the symbol is introduced to the (SVM) classifier.
+\section{Experimental}
+\label{sec:Experiments}
+
+The dataset collected by Hse and Newton\cite{HeloiseBeautification} is used to test the presented system.
 Figure \ref{fig:symbolSet} shows a set of the shapes used in the data set.
-We divide the dataset into four sets each set contains four different users.
-We divide the dataset equally into training set and test set.
-To generate trusted results, four different splits for the training and test data are generated from the dataset.
-The results displayed are the recognition accuracy resulted from the classifier after recognizing the symbols.
-In the following section, the recognition accuracy means the average recognition accuracy in the four data splits.
-%The result presented in this section are the average result of the four splits.
-%then result are stored The results displayed are the recognition accuracy resulted from the classifier after recognizing the symbols.
-%Furthermore, each set is divided into train and test set where the trained set where used to train the classifiers and the test set where used for testing.
-The algorithms are then tested using the four sets and the results are averaged.
-%We divide the dataset equally into training set and test set.
-The results displayed are the recognition accuracy resulted from the classifier after recognizing the symbols.
-\\
-%We divided the dataset into four sets each set contains four different users.
-Furthermore, each set is divided into train and test set where the trained set where used to train the classifiers and the test set where used for testing.
-The algorithms are then tested using the four sets and the results are averaged.
-\begin{figure}[]\centering
-\fbox{  \parbox{3cm}{% 
-		\includegraphics[scale=0.4]{images/symbolSet.pdf}	}}
-	\caption{The Symbol Set} Ths figure shows a set of the symbols used in the dataset. 
-	\label{fig:symbolSet}
-\end{figure}
-\begin{figure}[]
+The datasets were divided into training and test sets.
+Five different splits for the training and test data are generated from each dataset.
+The results displayed are the average recognition accuracy of the five splits.
+The accuracy is computed as the number of correctly recognized samples divided by the total number of samples in the test.
+ \begin{figure*}[]
 	\centering
-		\includegraphics[scale=0.5]{images/test1.jpg}
-	\caption{Algorithm comparison} The results of comparing the algorithms. The graphs displays the recognition rate of symbols. 
-	\label{fig:test1}
-\end{figure}
-%In the figure \ref{fig:test1}, the accuracy means that final hit rate of all symbols used in the test set.by the  four datasets . 
-We performed two experiments to test the system firstly we tested recognition accuracy of shapes in the data set with both algorithms (\textsl{AlgS1} and \textsl{AlgS2}).
-We also implemented the segmentation algorithm described in \cite{earlyprocess} (\textsl{Alg3}) to use as reference to our swarm algorithms.
+	\begin{minipage}[b]{5cm}	
+\begin{center}
+\subfigure [The Symbol Set] {\label{fig:symbolSet}
+\fbox{ \parbox{2cm}{% 
+		\includegraphics[scale=0.2]{images/symbolSet.PNG}	}}}
+
+		\subfigure[Algorithm comparison]{\includegraphics[width=5cm]{images/testAlg.jpg} 	\label{fig:test1}}
+		\end{center}
+		\end{minipage}
+		\hfill
+  \subfigure [Symbols Comparison] {\includegraphics[width=9cm]{images/test2smallpaper.jpg} \label{fig:test2}}
+	\caption{Experiments results :} a) The symbols set used in the dataset.   b) The recognition rate of different algorithms.  c) The recognition rate of each symbol using different algorithms. 
+\end{figure*} 
+We performed two experiments to test the system.
+Firstly, we tested recognition accuracy of shapes in the data set with both algorithms (\textsl{AlgS1} and \textsl{AlgS2}).
+We also implemented the segmentation algorithm described in \cite{earlyprocess} (\textsl{Alg3}) to compare it with our swarm algorithms.
 Figure \ref{fig:test1} shows the accuracy achieved by each algorithm.
-The results shows that both \textit{DPSO} algorithms achieve better result than other algorithms.
-The swarm algorithms were tested with and without the ellipse detection module.
-The ellipse detection module appears to be superior to results with only the \textit{DPSO} algorithms.\\
+The two swarm algorithms were tested with and without the ellipse fitting module.
+Combining the ellipse detection module improves the performance in the tested dataset.
+The result shows that (\textsl{AlgS2}) gives better performance than Alg3 and AlgS1.
+% Firstly we tested recognition accuracy of shapes in the data set with both algorithms (\textsl{AlgS1} and \textsl{AlgS2}).
+Figure \ref{fig:test1} shows the accuracy achieved by each algorithm.
+The two swarm algorithms were tested with and without the ellipse fitting module.
+Including the ellipse detection module improves the results with both the \textit{DPSO} algorithms.% The results show that both \textit{DPSO} algorithms achieve better result than other algorithms.
 
-The second experiment we implemented is testing the effect of symbol complexity and type on the recognition rate.
+The second experiment we implemented was to test the effect of symbol complexity and type on the recognition rate.
 Figure \ref{fig:test2} shows the accuracy of each symbol, it is clearly noted that symbols that have only line segments achieve higher accuracy rate than other symbols.
@@
 This is natural as algorithm \textsl{AlgS1} divides strokes into line segments only but \textsl{AlgS2} is able to divide strokes into lines and curves based on the minimum error of the segment itself.
-Algorithm \textsl{Alg3} gives good performance as long as the symbols consist of lines and curves, if the stroke consists of curve only  the algorithm may lead to wrong segmentation result.
-This is because the system divides the stroke first to line segments then tries to decide if each segment can be represent better as a curve unlike algorithm \textsl{AlgS2} where the curve segments are tested while choosing the best segmentation.
-Combining both algorithm \textsl{AlgS1} and \textsl{AlgS2} improved the recognition rate of all symbols because of combining the advantages of both algorithms.
-The penalty for this increased performance is in the computational time needed to run both the swarm algorithms.
-\\
-\begin{figure*}[]
-	\centering
-		\includegraphics[scale=0.8]{images/test2.jpg}
-	\caption{Symbols Comparison} The graph shows the recognition rate of each symbol using different algorithms. 
-	\label{fig:test2}
-\end{figure*}
-%We could not test the result of the segmentation algorithm directly due to the fact that the correct segmentation is highly ambiguous.
-It is only based on what the user intends to draw while sketching the symbol.
-Figure \ref{fig:SampleSeg} shows a sample of the output of the segmentation block.
-  %The final experiment we performed was to test the efficiency svm classifier.
-We implemented the linear discriminator classifier described in \cite{gestureexample12}(see figure \ref{fig:test3}).
-%The result shows that SVM is more scalable than the linear discriminator for this problem.
-%The test focus on the size of data needed to test each classifier and the achieved accuracy.
-%For this experiment we also shows the classification time for both classifiers
-%\begin{figure}[]
-%	\centering
-% 
-%				\subfigure[User strokes] {\includegraphics[scale=0.7]{images/strokebefore.jpg}}
-%		 	\subfigure[Segmented strokes] {\includegraphics[scale=0.7]{images/strokeafter.jpg}}
-%		 		\subfigure[Multiple symbols] {\includegraphics[scale=0.7]{images/results3.jpg}}
-%		
-%	\caption{Segmentation outputs} The figure shows sample of the output of the segmentation phase.
-Figure a) shows the stroke after the user draws it.
-The red points are points labeled as \textit{possible dominate points}.
-Figure b) shows the symbol after being segmented.
-Figure c) shows the segmentation of different and overlapped strokes.
-Notice the system mark the center of the ellipse that was identified.
-%	\label{fig:SampleSeg}
-%\end{figure}
-%\begin{figure}[]
-%	\centering
-%		\includegraphics[scale=0.4,clip=true, viewport=0 380 700 1280]{images/test1.pdf}
-%	\caption{The results of third experiment}
-%	\label{fig:test3}
-%\end{figure}
- %Details description of results got from comparing various algorithms  
- %Describe data set used for experiments ( No of data set, No. of categories, No of samples).
- %Describe other Algorithms [Alg 3 in paper \cite{earlyprocess}]used for comparing with the one described in the paper. 
+ Combining both algorithm \textsl{AlgS1} and \textsl{AlgS2} improved the recognition rate of all symbols.
+The penalty for this increased performance is in the computational time needed to run both swarm algorithms.
 \section{Conclusion and Future Work}
 \label{ConclusionandFutureWork}
-This paper presented a new approach to sketch recognition using PSO.
-The system uses both speed and curvature data which helps in improving the \textit{DPSO} algorithms over the original algorithms \cite{CruveDivisionSwarm,PolygonApproximationPSO}.
-It is noted that the \textit{DPSO}in general generates an optimized stroke segmentation which improves the final recognition rate.
+This paper presented a new approach to sketch recognition using DPSO.
+The system uses multiple preliminary information, i.e., speed and curvature information, as input to the \textit{DPSO}.
+Results show that the \textit{DPSO} in general generates an optimized stroke segmentation which improves the final recognition rate.
+The final recognition is based on geomterical,statistical and structural features.
+The experiments evaluated two \textit{DPSO} different segmentation algorithms (AlgS1 and AlgS2) on simple presentation set dataset.
+Results show that AlgS2 achieves better segmentation and final accuracy than Alg3 \cite{earlyprocess} on the dataset used.
+The results proved that the system is easily expandable to more and more symbols as it does not depend on low level primitive recognizers but on high level symbol and geometrical features.
+The system achieved an average overall improvement of more than 2\% over Alg3.
+%This paper presented a new approach to sketch recognition using PSO.
+The system uses both speed and curvature information which help improving the \textit{DPSO}.
+It is noted that the \textit{DPSO} in general generates an optimized stroke segmentation which improves the final recognition rate.
 The tradeoff between accuracy achieved and time complexity must be further investigated to achieve better results.
 The use of statistical moments and structural features improves the recognition rate.
-The system was tested on 13 different symbols and achieved satisfying results.
-The system dose not depends on low level recognizer but rather on a set of high level features.
-This makes the system easily expandable with aspect to symbols recognized.
-\\
-%As you can see form the experiments PSO have proved a spuriously than other system. 
-%\section{Future Work}The next step in 
+The system was tested on 13 different symbols and achieved an accuracy of 91.5\%.
+The system does not depend on low level recognizer but rather on a set of high level features.
+This makes the system easily expandable to more and more symbols.
+
  A possible extension of this research is to complete the clustering algorithm for fully automated sketch recognition.
-The clustering must be performed without the user explicit involvement.
+Currently, the system process one symbol at a time, an enchantment would be to allow users to draw more than one symbol at a time and incorporate a method for separating symbols.
 Other area of enhancements is the features extraction methods.
 Introducing more spatial and geometrical features is believed to improve classifications.
-%Each new stroke is checked if it can be a symbol or is a part of already drawn un completed symbol.
-%\bibliography{Mybibliography}
-\bibliography{../../neededfiles/Bibliographies/Mybibliography2}
+\bibliographystyle{IEEEbib}
+\bibliography{../../neededfiles/Bibliographies/Mybibliography}
 \end{document}

### 2009-07-20 trunk/smallpaper/SmallPaper.tex

@@
 Users draw strokes using a pen and the system immediately interprets their strokes as objects that can be easily manipulated.
-This paper uses Particle Swarm Algorithm (PSO) to divide the strokes the user draws into meaningful geometric primitives.
+This paper uses Particle Swarm Optimization Algorithm (PSO) to divide the strokes the user draws into meaningful geometric primitives.
 These geometric primitives are grouped to formulate symbols which are further identified.
@@
 Each low level recognizer is designed to recognize only one geometric shape using spatial and geometric information extracted from the stroke.
-Yu \cite{meanshift10}introduced a feature area for each primitive and then computed the segmentation error for different types of primitives based on the computed feature area.
+Yu \cite{meanshift10} introduced a feature area for each primitive and then computed the segmentation error for different types of primitives based on the computed feature area.
 His system achieved good accuracy in simple shapes (square, ellipse,...) but did not perform well in more complex shapes.
 Paulson and Hammond\cite{Paleosketch08} introduced a set of low level recognizers that were reported to achieve 98\%  accuracy but their system, similar to all low level recognizers, only identify a small set of simple shapes.
+One of the problems in these previous systems is that they only rely on curvature and speed data which are supjected to noise.
+They are also subjected to over and under segmentation problems due to use of thresholds.
+
+The segmentation problem is summarized as finding the best decomposition with least number of segments each represents a geometric primitive.
+This is an optimization problem, which evolutionary programming can solve efficiently.
 A genetic algorithm was used by \cite{CruveDivisionSwarm} to optimally divide digital curves into lines and curves.
@@
 
- The remaining of the paper is as following; Section \ref{Sysdisc} describes the system in details.
+The contribution in the system is mainly in attempting to generate the optimal segmentation using PSO which exhibited superiority with respect to simillar algoirthms in various other applications \cite{PolygonApproximationPSO}.
+The system attempts to find the optimial decomposition of the input stroke using PSO with the help of curvature and speed information.
+The use of PSO algorithm helps in eliminating the effects of input noise, problems of over and under segmentation that were in pervious systems.
+The proposed symbol recognizier uses the generated segmenation to compute the feature vector which is used to identify more complex symbols.
+This enables the system to be scalable for more  complex shapes.
+
+ %The remaining of the paper is as following; Section \ref{Sysdisc} describes the system in details.
 Sections \ref{Prepross}, \ref{seg} and \ref{sec:Recognition} provide details on preprocessing, segmentation and recognition process respectively.
@@
 We conclude in Section \ref{ConclusionandFutureWork}.
+
 \section{The proposed system}
@@
 In the segmentation stage the stroke is divided into a set of primitives.
-As shown in Figure \ref{fig:segblock} first an attempt is made to fit the stroke points into a curve or an ellipse using a minimum square error fitting algorithm \cite{chernov-2003}.
+As shown in Figure \ref{fig:segblock} first an attempt is made to fit the stroke points into a curve or an ellipse using a minimum square error fitting algorithm \cite{chernov}.
 If the stroke proved to be an ellipse arc then the segmentation process ends and the system proceeds to the next step.
@@
 The axes of the ellipse are estimated as $width/2$ and $height/2$ of the stroke bounding box.
-The least square fitting algorithm is used to minimize the fitting error of the ellipse equation \cite{chernov-2003}.
+The least square fitting algorithm is used to minimize the fitting error of the ellipse equation \cite{chernov}.
 \subsubsection{Discrete particle swarm algorithm (\textit{DPSO})}
@@
 \begin{equation}
-E=\sum\nolimits_{i = 0}^M e ( \widehat{x_ix_{i+1}},\overline{x_i x_{i+1}})
+E=\sum\nolimits_{i = 1}^M e ( \widehat{x_ix_{i+1}},\overline{x_i x_{i+1}})
 \label{eq:ErrorSwarm1}
@@
  \begin{equation}
-E=\sum\nolimits_{i = 0}^M e(D_i) 
+E=\sum\nolimits_{i = 1}^M e(D_i) 
 \label{eq:errorSwarm2}
@@
 \end{equation} where $E$ is the error and $M$ is the number of segments and $k$ is a parameter tweaked to get minimum number of segments.
-The larger k is, the more effect the number of segments will have.
-For our system, k is selected to be 0.5\cite{CruveDivisionSwarm}.
+The larger $k$ is, the more effect the number of segments will have.
+For our system, $k$ is selected to be $0.5$\cite{CruveDivisionSwarm}.
 
@@
 A composite set of features is used to generate a single feature vector.
-The features used consist of Rubine feature set,  Zernike moments of order 21 \cite{HeloiseBeautification}, ink density as well as some structural and spatial information like number of perpendiculars lines ,number of parallel lines and number of different types of primitives in each symbol.
+The features used consist of Rubine feature set,  Zernike moments of order 10 \cite{HeloiseBeautification}, ink density as well as some structural and spatial information like number of perpendiculars lines ,number of parallel lines and number of different types of primitives in each symbol.
 After computing the features the symbol is introduced to the (SVM) classifier.
@@
 The accuracy is computed as the number of correctly recognized samples divided by the total number of samples in the test.
-
  \begin{figure*}[]
@@
 The penalty for this increased performance is in the computational time needed to run both swarm algorithms.
+
 \section{Conclusion and Future Work}
@@
 The final recognition is based on geomterical,statistical and structural features.
-The experiments evaluated two \textit{DPSO} different segmentation algorithms (AlgS1 and AlgS2) on simple presentation set datasets.
+The experiments evaluated two \textit{DPSO} different segmentation algorithms (AlgS1 and AlgS2) on simple presentation set dataset.
 Results show that AlgS2 achieves better segmentation and final accuracy than Alg3 \cite{earlyprocess} on the dataset used.

### 2009-07-20 smallpaper/SmallPaper.tex

@@
 The proposed symbol recognizier uses the generated segmenation to compute the feature vector which is used to identify more complex symbols.
-This ensures that the system is scalable for more  complex shapes.
+This enables the system to be scalable for more  complex shapes.
 
@@
 For the \textit{first algorithm \textsl{AlgS1}}, 
- the arc $\widehat{x_ix_j}$ is defined as the consecutive set of points from point$x_i$ to point $x_{j}$ as in $x_i,x_{i+1} \cdots,x_j$.
-The line $\overline{x_i x_j} $ as the straight line connecting point $x_i$ to point $x_j$.
+ the arc $\widehat{x_lx_k}$ is defined as the consecutive set of points from point$x_l$ to point $x_{k}$ as in $x_l,x_{l+1} \cdots,x_k$.
+The line $\overline{x_l x_k} $ as the straight line connecting point $x_l$ to point $x_k$.
 The approximation error is computed by the equation \ref{eq:ErrorSwarm1}.
@@
 \begin{equation}
-E=\sum\nolimits_{i = 1}^M e ( \widehat{x_ix_{i+1}},\overline{x_i x_{i+1}})
+E=\sum\nolimits_{l = 1}^M e ( \widehat{x_lx_{l+1}},\overline{x_l x_{l+1}})
 \label{eq:ErrorSwarm1}
 \end{equation} where $M$ is the number of dominant points in this solution as generated by the swarm algorithm.
-The error $e ( \widehat{x_ix_j},\overline{x_i x_j})$ is computed as the sum of squared perpendicular distance from every point along the arc $\widehat{x_ix_j}$ to the line $\overline{x_i x_j}$.
+The error $e ( \widehat{x_lx_k},\overline{x_l x_k})$ is computed as the sum of squared perpendicular distance from every point along the arc $\widehat{x_lx_k}$ to the line $\overline{x_l x_k}$.
 The fitness is computed using equation \ref{eq:fitnessSwarm1}\begin{equation}
@@
 \end{equation} where $N$ is the number of points in the stroke, $D$ is the number of points in the solution that was previously labeled as a possible dominant point ($P_{pd}$), $E$ is the computed error and $\varepsilon$ is the error threshold.
-It should be noticed that when the error is larger than the threshold $\varepsilon$ the fitness is given a -ve value to lower the fitness value of the solution.
+It should be noticed that when the error is larger than the threshold $\varepsilon$ the fitness is given a $-ve$ value to lower the fitness value of the solution.
 Otherwise the system favors the lower number of vertices.
 
-The \textit{second algorithm \textsl{AlgS2}} has the same problem formulation but different fitness and error functions are used.
-An attempt is made to fit each segment into line or circular arc, the error of both circle and line fit estimations are computed for each segment $S_i=\widehat{x_ix_j}$, the approximation with the lower error value is the chosen approximation of this segment $S_i$\cite{CruveDivisionSwarm}.
+The pervious algorithm segments stroke into set of line segments but storkes dose not always consist of line only.
+It is more meanfull to segment stroke to  Therefor, a \textit{second algorithm \textsl{AlgS2}} is used to decompose strokes.
+\textit{\textsl{AlgS2}} has the same problem formulation but different fitness and error functions are used.
+An attempt is made to fit each segment into line or circular arc, the error of both circle and line fit estimations are computed for each segment $S_s=\widehat{x_lx_k}$, the approximation with the lower error value is the chosen approximation of this segment $S_s$\cite{CruveDivisionSwarm}.
 The sum of the approximation error of all segments is the total error of the particle.
@@
  \begin{equation}
-E=\sum\nolimits_{i = 1}^M e(D_i) 
+E=\sum\nolimits_{s = 1}^M e(D_s) 
 \label{eq:errorSwarm2}
-\end{equation}where $M$ is the number of segments in the solution as generated by the swarm algorithm, $D_i$ is the minimum approximation error of curve and line approximations $min(d_c,d_l)$ as computed by \cite{CruveDivisionSwarm}.
+\end{equation}where $M$ is the number of segments in the solution as generated by the swarm algorithm, $D_s$ is the minimum approximation error of curve and line approximations $min(d_c,d_l)$ as computed by \cite{CruveDivisionSwarm}.
 The fitness is computed by the equation \ref{eq:fitnessSwarm2} \begin{equation}
@@
 Figure \ref{fig:test1} shows the accuracy achieved by each algorithm.
-The two swarm algorithms were tested with and without the ellipse fitting module.
-Combining the ellipse detection module improves the performance in the tested dataset.
 The result shows that (\textsl{AlgS2}) gives better performance than Alg3 and AlgS1.
@@
 Including the ellipse detection module improves the results with both the \textit{DPSO} algorithms.% The results show that both \textit{DPSO} algorithms achieve better result than other algorithms.
+The two swarm algorithms were tested with and without the ellipse fitting module.
+Combining the ellipse detection module improves the performance in the tested dataset.
 

### 2009-07-20 trunk/smallpaper/SmallPaper.tex

@@
 \end{equation} where $N$ is the number of points in the stroke, $D$ is the number of points in the solution that was previously labeled as a possible dominant point ($P_{pd}$), $E$ is the computed error and $\varepsilon$ is the error threshold.
-It should be noticed that when the error is larger than the threshold $\varepsilon$ the fitness is given a -ve value to lower the fitness value of the solution.
+It should be noticed that when the error is larger than the threshold $\varepsilon$ the fitness is given a $-ve$ value to lower the fitness value of the solution.
 Otherwise the system favors the lower number of vertices.
 
-The \textit{second algorithm \textsl{AlgS2}} has the same problem formulation but different fitness and error functions are used.
-An attempt is made to fit each segment into line or circular arc, the error of both circle and line fit estimations are computed for each segment $S_i=\widehat{x_ix_j}$, the approximation with the lower error value is the chosen approximation of this segment $S_i$\cite{CruveDivisionSwarm}.
+The pervious algorithm segments stroke into set of line segments but storkes dose not always consist of line only.
+It is more meanful to use curves when segmenting strokes.
+Therefor, a \textit{second algorithm \textsl{AlgS2}} is used to decompose strokes to set of lines and circular arcs.
+\textit{\textsl{AlgS2}} has the same problem formulation but different fitness and error functions are used.
+An attempt is made to fit each segment into line or circular arc, the error of both circle and line fit estimations are computed for each segment $S_s=\widehat{x_lx_k}$, the approximation with the lower error value is the chosen approximation of this segment $S_s$\cite{CruveDivisionSwarm}.
 The sum of the approximation error of all segments is the total error of the particle.

### 2009-07-26 trunk/smallpaper/SmallPaper.tex
making the paper 4 pages

@@
 Design engineers need a powerful computer based system with symbol recognition to help them design, manipulate and store sketches more effectively than using only papers.
-Increasing the interaction between computers and users in sketch and CAD systems has been the reason for the emerging of a few advanced sketch recognition systems.
-
-Sketch recognition is divided into three main steps preprocessing, segmentation and symbol recognition.
+Increasing the interaction between computers and users in sketch and CAD systems has been the reason for the emerging of a few advanced sketch recognition systems.Sketch recognition is the process of identifying user strokes into a meanful symbol that can be furthur used.
+%Sketch recognition is divided into three main steps preprocessing, segmentation and symbol recognition.
 The preprocessing phase captures user input stroke points and collects basic information about the stroke then proceeds to remove noise and compute basic statistical and geometrical information.
@@
 Segmentation is based on dividing the digital curve into polygon\cite{PolygonApproximationPSO}.
-The algorithm enhances its performance by adding curvature and other local information while segmenting strokes to achieve better results.
 To handle complex strokes, our system uses a second segmentation PSO algorithm to segment strokes into a set of lines and circular arcs.
@@
 It was noted that time difference provides more distinct maxima than the speed information.
-Agar \cite{polygonfeedback31} mentions that the pointing device (for example the mouse) sampling rate is the reason for this phenomena.
+Agar \cite{polygonfeedback31} mentions that the pointing device (i.e.
+the mouse) sampling rate is the reason for this phenomena.
  In the presented system, we computed time difference, direction, speed and curvature of each point along a stroke.
@@
 All the points labeled as possible dominant points $P_{pd}$ are saved into a single array.
-Figure \ref{fig:LabelsPPD} shows the particles labeled as Possible dominate point $P_{pd}$ by the preprocessing, it is noticed the redundancy of some $P_{pd}$ points.
+Figure \ref{fig:LabelsPPD} shows the particles labeled as Possible dominant point $P_{pd}$ by the preprocessing, it is noticed the redundancy of some $P_{pd}$ points.
 % (as shown are redundant)
@@
 In the segmentation stage the stroke is divided into a set of primitives.
-As shown in Figure \ref{fig:segblock} first an attempt is made to fit the stroke points into a curve or an ellipse using a minimum square error fitting algorithm \cite{chernov}.
+As shown in Figure \ref{fig:segblock} first an attempt is made to fit the stroke points into a curve or an ellipse using a minimum square error fitting algorithm.
 If the stroke proved to be an ellipse arc then the segmentation process ends and the system proceeds to the next step.
@@
 The axes of the ellipse are estimated as $width/2$ and $height/2$ of the stroke bounding box.
-The least square fitting algorithm is used to minimize the fitting error of the ellipse equation \cite{chernov}.
+The least square fitting algorithm is used to minimize the fitting error of the ellipse equation.
 \subsubsection{Discrete particle swarm algorithm (\textit{DPSO})}
@@
 The graphical meaning of the error is shown in Figure\ref{fig:LabelsPSO}.
-\begin{equation}
+\small\begin{equation}
 E=\sum\nolimits_{l = 1}^M e ( \widehat{x_lx_{l+1}},\overline{x_l x_{l+1}})
 \label{eq:ErrorSwarm1}
-\end{equation} where $M$ is the number of dominant points in this solution as generated by the swarm algorithm.
+\end{equation} 
+\normalsize
+where $M$ is the number of dominant points in this solution as generated by the swarm algorithm.
 The error $e ( \widehat{x_lx_k},\overline{x_l x_k})$ is computed as the sum of squared perpendicular distance from every point along the arc $\widehat{x_lx_k}$ to the line $\overline{x_l x_k}$.
-The fitness is computed using equation \ref{eq:fitnessSwarm1}\begin{equation}
+The fitness is computed using equation \ref{eq:fitnessSwarm1}
+\small \begin{equation}
 \max fitness(p_i ) = \left\{ {\begin{array}{*{20}c}
@@
 \label{eq:fitnessSwarm1}
-\end{equation} where $N$ is the number of points in the stroke, $D$ is the number of points in the solution that was previously labeled as a possible dominant point ($P_{pd}$), $E$ is the computed error and $\varepsilon$ is the error threshold.
+\end{equation}\normalsize where $N$ is the number of points in the stroke, $D$ is the number of points in the solution that was previously labeled as a possible dominant point ($P_{pd}$), $E$ is the computed error and $\varepsilon$ is the error threshold.
 It should be noticed that when the error is larger than the threshold $\varepsilon$ the fitness is given a $-ve$ value to lower the fitness value of the solution.
@@
 This paper presented a new approach to sketch recognition using DPSO.
-The system uses multiple preliminary information, i.e., speed and curvature information, as input to the \textit{DPSO}.
 Results show that the \textit{DPSO} in general generates an optimized stroke segmentation which improves the final recognition rate.
@@
 Introducing more spatial and geometrical features is believed to improve classifications.
+
 \bibliographystyle{IEEEbib}

### 2009-07-26 trunk/smallpaper/SmallPaper.tex
final submitted
final after corrections

@@
 Design engineers need a powerful computer based system with symbol recognition to help them design, manipulate and store sketches more effectively than using only papers.
-Increasing the interaction between computers and users in sketch and CAD systems has been the reason for the emerging of a few advanced sketch recognition systems.Sketch recognition is the process of identifying user strokes into a meanful symbol that can be furthur used.
+Increasing the interaction between computers and users in sketch and CAD systems has been the reason for the emerging of a few advanced sketch recognition systems.Sketch recognition is the process of identifying user strokes into a meaningful symbols that can be further used.
 %Sketch recognition is divided into three main steps preprocessing, segmentation and symbol recognition.
@@
 Paulson and Hammond\cite{Paleosketch08} introduced a set of low level recognizers that were reported to achieve 98\%  accuracy but their system, similar to all low level recognizers, only identify a small set of simple shapes.
-One of the problems in these previous systems is that they only rely on curvature and speed data which are supjected to noise.
+One of the problems in these previous systems is that they only rely on curvature and speed data which are subjected to noise.
 They are also subjected to over and under segmentation problems due to use of thresholds.
@@
 
-The contribution in the system is mainly in attempting to generate the optimal segmentation using PSO which exhibited superiority with respect to simillar algoirthms in various other applications \cite{PolygonApproximationPSO}.
-The system attempts to find the optimial decomposition of the input stroke using PSO with the help of curvature and speed information.
+The contribution in the system is mainly in attempting to generate the optimal segmentation using PSO which exhibited superiority with respect to similar algorithms in various other applications \cite{PolygonApproximationPSO}.
+The system attempts to find the optimal decomposition of the input stroke using PSO with the help of curvature and speed information.
 The use of PSO algorithm helps in eliminating the effects of input noise, problems of over and under segmentation that were in pervious systems.
-The proposed symbol recognizier uses the generated segmenation to compute the feature vector which is used to identify more complex symbols.
+The proposed symbol recognizer uses the generated segmentation to compute the feature vector which is used to identify more complex symbols.
 This enables the system to be scalable for more  complex shapes.
@@
 
-The pervious algorithm segments stroke into set of line segments but storkes dose not always consist of line only.
-It is more meanful to use curves when segmenting strokes.
+The previous algorithm segments stroke into set of line segments but strokes dose not always consist of line only.
+It is more meaningful to use curves when segmenting strokes.
 Therefor, a \textit{second algorithm \textsl{AlgS2}} is used to decompose strokes to set of lines and circular arcs.
@@
 		\hfill
-  \subfigure [Symbols Comparison] {\includegraphics[width=9cm]{images/test2smallpaper.jpg} \label{fig:test2}}
+  \subfigure [Symbols Comparison] {\includegraphics[width=11cm]{images/symbols.jpg} \label{fig:test2}} %symbols.jpg  test2smallpaper.jpg
 	\caption{Experiments results :} a) The symbols set used in the dataset.   b) The recognition rate of different algorithms.  c) The recognition rate of each symbol using different algorithms. 
@@
 Results show that the \textit{DPSO} in general generates an optimized stroke segmentation which improves the final recognition rate.
-The final recognition is based on geomterical,statistical and structural features.
+The final recognition is based on geometrical,statistical and structural features.
 The experiments evaluated two \textit{DPSO} different segmentation algorithms (AlgS1 and AlgS2) on simple presentation set dataset.

### 2009-07-28 draft/System.tex
change i mad for pso1

@@
 The current implemented systems either segment strokes based on simple curvature\cite{meanshift10,earlySketchbased4} or speed information\cite{earlySketchbased4}.
+
+The contribution in the system is mainly in attempting to generate the optimal segmentation using PSO which exhibited superiority with respect to similar algorithms in various other applications \cite{PolygonApproximationPSO}.
+The system attempts to find the optimal decomposition of the input stroke using PSO with the help of curvature and speed information.
+The use of PSO algorithm helps in eliminating the effects of input noise, problems of over and under segmentation that were in pervious systems.
+The proposed symbol recognizer uses the generated segmentation to compute the feature vector which is used to identify more complex symbols.
+This enables the system to be scalable for more  complex shapes.
+
  %the performance of most algorithms is based on user style.

### 2009-07-28 trunk/draft/chapterSurvey.tex
save to version before defense
change i mad for pso1

@@
 The algorithm simulate the moving Kennedy and Eberhart \cite{PSOFirst,PSO2} introduced particle swarm algorithm (PSO) as a general optimization algorithm.
- \subsection*{Particle Swarm Optimization}
+ \subsection{Particle Swarm Optimization}
  In optimization problems, the algorithm attempts to find a feasible solution which satisfies all constrain of this problem.
@@
 
+\subsection{PSO vs. GA}
  PSO can be viewed as an algorithm between stochastic search and Evolutionary algorithms\cite{PSOpattern}.
-In general the PSO is a continuous optimization problem but experiments showed it performed as GA is discrete and binary optimization problems.
-The comparison between PSO and Genetic algorithm was performed in \cite{PSOGAComp} and proved that PSO generate faster, better solutions than GA especially in problems with high dimensionality.
+In general the PSO is a continuous optimization problem but experiments showed it performed as well as GA which is in general a discrete and binary optimization problem.
+
+ The comparison between PSO and Genetic algorithm was performed in \cite{PSOGAComp} and proved that PSO generate faster, better optimal solutions than GA especially in problems with high dimensionality.
+In addition, Rohinsons et al.
+\cite{RohinsonPSGA} proved that PSO performed better than GA in difficult engineering problems \cite{EngineerProblemGA}.
+
+% Eberhart \cite{PSOGAComp}
 
@@
 
-\subsection*{PSO Neighborhood Topology}  
+ %From the procedure, we can learn that PSO shares many common points with GA.
+Both algorithms start with a group of a randomly generated population, both have fitness values to evaluate the population.
+Both update the population and search for the optimium with random techniques.
+Both systems do not guarantee success.
+
+%However, PSO does not have genetic operators like crossover and mutation. Particles update themselves with the internal velocity. They also have memory, which is important to the algorithm.
+
+%Compared with genetic algorithms (GAs), the information sharing mechanism in PSO is significantly different.
+In GAs, chromosomes share information with each other.
+So the whole population moves like a one group towards an optimal area.
+In PSO, only gBest (or lBest) gives out the information to others.
+It is a one -way information sharing mechanism.
+The evolution only looks for the best solution.
+Compared with GA, all the particles tend to converge to the best solution quickly even in the local version in most cases.
+
+\subsection{PSO Neighborhood Topology}  
 In general the PSO algorithm have fully connected particle which means that every particle know the global best and local best of each other particle.
@@
 
-\subsection*{Drawbacks}
+%\subsection*{PSO vs. Genetic Algorithms}  
+% compare   between genetic programming and algorihtms.
+
+\subsection{Drawbacks}
 
@@
 
-\subsection*{Improvements to PSO}
+\subsection{Improvements to PSO}
 
@@
 
-\subsection*{Particle Swarm Applications}
+\subsection{Particle Swarm Applications}
 \label{sec:SwarmApp}
@@
 
-\subsection*{Discrite Particle Swarm Equations(DPSO)}
+\subsection{Discrete Particle Swarm Equations(DPSO)}
 %\section{Particle Swarm Algorithm}
@@
 The main idea of \textit{Particle Swarm Algorithm (PSO)} is to represent each solution with a $N$ dimension particle from the solution space \cite{PSOFirst}.
-Each particle moves with a direction and velocity $v_{ij}$ based on equations \ref{eq:Swarm} \& \ref{eq:Swarm1}.
+Each particle moves with a direction and velocity $v_{ij}$ based on equations \ref{eq:Swarm1} \& \ref{eq:Swarm}.
 \begin{equation}

### 2009-07-28 trunk/draft/chapterSurvey.tex
change i mad for pso1

@@
 
- The comparison between PSO and Genetic algorithm was performed in \cite{PSOGAComp} and proved that PSO generate faster, better optimal solutions than GA especially in problems with high dimensionality.
+ The comparison between PSO and Genetic algorithm was performed in \cite{PSOGAComp,} and proved that PSO generate faster, better optimal solutions than GA especially in problems with high dimensionality.
 In addition, Rohinsons et al.
@@
 % Eberhart \cite{PSOGAComp}
+%%PSO has been shown to have the same ability to find a global optimum as genetic
+%algorithms, but is able to find optimums faster than genetic algorithms (Panda and Padhy
+%2007; Mouser and Dunn 2005; Hassan et al. 2005)
 

### 2009-07-28 smallpaper/SmallPaper.tex
paper send to reviewers

@@
 Increasing the interaction between computers and users in sketch and CAD systems has been the reason for the emerging of a few advanced sketch recognition systems.
-
-Sketch recognition is divided into three main steps preprocessing, segmentation and symbol recognition.
+Sketch recognition is the process of identifying user strokes into meaningful symbols that can be further used.
+%Sketch recognition is divided into three main steps preprocessing, segmentation and symbol recognition.
 The preprocessing phase captures user input stroke points and collects basic information about the stroke then proceeds to remove noise and compute basic statistical and geometrical information.
@@
 
-The contribution in the system is mainly in attempting to generate the optimal segmentation using PSO which exhibited superiority with respect to simillar algoirthms in various other applications \cite{PolygonApproximationPSO}.
-The system attempts to find the optimial decomposition of the input stroke using PSO with the help of curvature and speed information.
+The contribution in the system is mainly in attempting to generate the optimal segmentation using PSO which exhibited superiority with respect to similar algorithms in various other applications \cite{PolygonApproximationPSO}.
+The system attempts to find the optimal decomposition of the input stroke using PSO with the help of curvature and speed information.
 The use of PSO algorithm helps in eliminating the effects of input noise, problems of over and under segmentation that were in pervious systems.
-The proposed symbol recognizier uses the generated segmenation to compute the feature vector which is used to identify more complex symbols.
+The proposed symbol recognizer uses the generated segmentation to compute the feature vector which is used to identify more complex symbols.
 This enables the system to be scalable for more complex shapes.
@@
 It was noted that time difference provides more distinct maxima than the speed information.
-Agar \cite{polygonfeedback31} mentions that the pointing device (for example the mouse) sampling rate is the reason for this phenomena.
+Agar \cite{polygonfeedback31} mentions that the pointing device (i.e.
+the mouse) sampling rate is the reason for this phenomenon.
  In the presented system, we computed time difference, direction, speed and curvature of each point along a stroke.
@@
 	\centering
-		\subfigure[Posible domninate point]{ 	\label{fig:LabelsPPD}	\includegraphics[scale=0.5]{images/ppd.jpg}}
+		\subfigure[Posible domninate point]{ 	
+		\label{fig:LabelsPPD}	\includegraphics[scale=0.5]{images/ppd.jpg}}
 		\hfill
-			\subfigure[DPSO algorithm encoding]{	\label{fig:LabelsPSO} \includegraphics[scale=0.5]{images/pso1.jpg}}			
+			\subfigure[DPSO algorithm encoding]{	
+			\label{fig:LabelsPSO} 
+			\includegraphics[scale=0.5]{images/pso1.jpg}}			
 	\caption{Segmentation}% a) Possible dominate point b) Particle encoding  
@@
 The axes of the ellipse are estimated as $width/2$ and $height/2$ of the stroke bounding box.
-The least square fitting algorithm is used to minimize the fitting error of the ellipse equation \cite{chernov}.
+The least square fitting algorithm is used to minimize the fitting error of the ellipse equation.
+
 \subsubsection{Discrete particle swarm algorithm (\textit{DPSO})}
@@
 Two DPSO algorithms are used to generate segmentations for each stroke the user draws.
-For each generated segmentation evaluated an error is determined.
+For each generated segmentation an error is determined.
 The segmentation with the minimum error value is chosen as the best stroke segmentation (Figure \ref{fig:segblock}).
-The problem definition is the same in both algorithms but they differ in the method they use to compute fitness and error functions.
+The problem definition is the same in both algorithms but they differ in the method they use to compute fitness and error functions \footnote{The parameters of the DPSO algorithms are determined using empirical experiments and conform with similar applications \cite{PolygonApproximationPSO} (e.g.
+15 particles, 80 maximimum swarm iteration)}.
 
@@
 
-For the \textit{first algorithm \textsl{AlgS1}}, 
- the arc $\widehat{x_lx_k}$ is defined as the consecutive set of points from point$x_l$ to point $x_{k}$ as in $x_l,x_{l+1} \cdots,x_k$.
+For the \textit{first algorithm \textsl{AlgS1}}, the arc $\widehat{x_lx_k}$ is defined as the consecutive set of points from point$x_l$ to point $x_{k}$ as in $x_l,x_{l+1} \cdots,x_k$.
 The line $\overline{x_l x_k} $ as the straight line connecting point $x_l$ to point $x_k$.
@@
 The graphical meaning of the error is shown in Figure\ref{fig:LabelsPSO}.
-\begin{equation}
+\small \begin{equation}
 E=\sum\nolimits_{l = 1}^M e ( \widehat{x_lx_{l+1}},\overline{x_l x_{l+1}})
 \label{eq:ErrorSwarm1}
-\end{equation} where $M$ is the number of dominant points in this solution as generated by the swarm algorithm.
+\end{equation} 
+\normalsize
+where $M$ is the number of dominant points in this solution as generated by the swarm algorithm.
 The error $e ( \widehat{x_lx_k},\overline{x_l x_k})$ is computed as the sum of squared perpendicular distance from every point along the arc $\widehat{x_lx_k}$ to the line $\overline{x_l x_k}$.
-The fitness is computed using equation \ref{eq:fitnessSwarm1}\begin{equation}
+The fitness is computed using equation \ref{eq:fitnessSwarm1}
+\small \begin{equation}
 \max fitness(p_i ) = \left\{ {\begin{array}{*{20}c}
@@
 
-The pervious algorithm segments stroke into set of line segments but storkes dose not always consist of line only.
-It is more meanfull to segment stroke to  Therefor, a \textit{second algorithm \textsl{AlgS2}} is used to decompose strokes.
+The previous algorithm segments stroke into set of line segments but strokes does not always consist of line only.
+It is more meaningful to use curves when segmenting strokes.
+Therefor, a \textit{second algorithm \textsl{AlgS2}} is used to decompose strokes to set of lines and circular arcs.
 \textit{\textsl{AlgS2}} has the same problem formulation but different fitness and error functions are used.
@@
 The total error of the particle is computed by equation \ref{eq:errorSwarm2}
- \begin{equation}
+\small \begin{equation}
 E=\sum\nolimits_{s = 1}^M e(D_s) 
 \label{eq:errorSwarm2}
-\end{equation}where $M$ is the number of segments in the solution as generated by the swarm algorithm, $D_s$ is the minimum approximation error of curve and line approximations $min(d_c,d_l)$ as computed by \cite{CruveDivisionSwarm}.
-The fitness is computed by the equation \ref{eq:fitnessSwarm2} \begin{equation}
+\end{equation} \normalsize where $M$ is the number of segments in the solution as generated by the swarm algorithm, $D_s$ is the minimum approximation error of curve and line approximations $min(d_c,d_l)$ as computed by \cite{CruveDivisionSwarm}.
+The fitness is computed by the equation \ref{eq:fitnessSwarm2} \small \begin{equation}
 \max fitness(P_i ) = \frac{1}{{E \times M^k }}
 \label{eq:fitnessSwarm2}
-\end{equation} where $E$ is the error and $M$ is the number of segments and $k$ is a parameter tweaked to get minimum number of segments.
+\end{equation} \normalsize where $E$ is the error and $M$ is the number of segments and $k$ is a parameter tweaked to get minimum number of segments.
 The larger $k$ is, the more effect the number of segments will have.
@@
 This paper presented a new approach to sketch recognition using DPSO.
-The system uses multiple preliminary information, i.e., speed and curvature information, as input to the \textit{DPSO}.
 Results show that the \textit{DPSO} in general generates an optimized stroke segmentation which improves the final recognition rate.
-The final recognition is based on geomterical,statistical and structural features.
+The final recognition is based on geometrical, statistical and structural features.
 The experiments evaluated two \textit{DPSO} different segmentation algorithms (AlgS1 and AlgS2) on simple presentation set dataset.
 Results show that AlgS2 achieves better segmentation and final accuracy than Alg3 \cite{earlyprocess} on the dataset used.
-The results proved that the system is easily expandable to more and more symbols as it does not depend on low level primitive recognizers but on high level symbol and geometrical features.
+The results proved that the system is easily expandable to more and more symbol as it does not depend on low level primitive recognizers but on high level symbol and geometrical features.
 The system achieved an average overall improvement of more than 2\% over Alg3.

### 2009-07-28 draft/chapterSurvey.tex

--- 2009-04-27 draft/chapterSurvey.tex
@@
 
-\section{Swarm Optimization Review}
+\section{Swarm Intelligence Review}
 \label{sec:ParticleSwramOptimizationReview}
-%This seems to emerge of what is called 'collective intelligence' from often large simple agents.  
-% start with swarm in general  .. say about flock , ant and usin in image processing 
-% then speeak about PSO and  its firstly generated   (see article)
-%mention that is best than ga (see thesis)
 
@@
 
-Image processing is one of the disciplines that used Swarm Intelligence.
-Ant colony was widely used to segment objects in \cite{antSegmentation,antthreshold,SwarmMedical}.
-Since 1995 when Kennedy and Eberhart \cite{PSOFirst,PSO2} introduced particle swarm algorithm (PSO) the algorithm had been used in various image processing system as an optimization  algorithm \cite{SwarmImageProcessing,Swarmimageprocessing2,Colorquantization,SwarmMedical}.
+The underling research in swarm intelligence led to optimization algorithm that simulates the animal social behaviors.
+Ant Colony Optimization (ACO) algorithm simulate how ants search for optimize path for food \cite{ACOFirst}.
+Real ants leave evidence of their exploration while they are searching for food.
+This trail is used by other ants to locate shortest distance between the colony and food location.
+ACO use similar type of cooperation between agents to find optimal solution.
+The synchronous bird flock movement was simulated in \cite{PSOFirst,PSO2}.
+Kennedy and Eberhart \cite{PSOFirst,PSO2} simulated the rules of bird flock movement and introduced the Particle Swarm Optimization Algorithm (PSO).
+
+   The next section describes in details the Particle Swarm Optimization and its advantages and disadvantages. 
+% Kennedy and Eberhart \cite{PSOFirst,PSO2} introduced Particle Swarm Optimization Algorithm (PSO).
+The algorithm simulate the moving Kennedy and Eberhart \cite{PSOFirst,PSO2} introduced particle swarm algorithm (PSO) as a general optimization algorithm.
+ \subsection{Particle Swarm Optimization}
+ In optimization problems, the algorithm attempts to find a feasible solution which satisfies all constrain of this problem.
+An optimal solution is the best feasible solution in the solution space.
+Several methods were used to solve optimization problems.
+Stochastic search is based on the assumption that good solution are near each other.
+Algorithms like simulating annealing and hill climbing used stochastic to find best solution.
+The problem in these methods is they may be trapped into a local best and fail to find the global best solution.
+
+  Evolutionary algorithms (e.g.
+Genatic Algoirthms) explore solution using a population which is evaluated using fitness functions.
+The algorithms initially start with a randomly chosen population which is used to generate new population using various operators (e.g.
+cross over, mutation...
+).
+After each iteration, the population is evaluated using a fitness function and only the strongest are used to generate new population.
+The use of population helps to avoid stochastic search problem and explores the solution space to find better global optimal solutions, but unfortunately they may not always converge to a global best solution\cite{PSOpattern}.
+
+ Particle Swarm Algorithm\cite{PSOFirst}is a population based stochastic search where each solution is simulated as a  particle that flies around in a multidimensional search space.
+A fitness function is used to evaluate the goodness of each particle in solution space.
+But unlike evolutionally algorithm the new population is the PSO is influenced by the social cooperation of the particle rather than survival to the fittest.
+Each particle saves a history of the best location in the swarm and its own best location.
+At the end each iteration, the particles moves to a new location based on the global best location and it own history of best location.
+This use of pervious particle locations gives PSO advantage over Evolutionary algorithms where no similar information is used.
+
+\subsection{PSO vs. GA}
+ PSO can be viewed as an algorithm between stochastic search and Evolutionary algorithms\cite{PSOpattern}.
+In general the PSO is a continuous optimization problem but experiments showed it performed as well as GA which is in general a discrete and binary optimization problem.
+In general, both GA and PSO share many points, e.g, both algorithms generate the initial populuation randomaly, both use a fitness function or value to evaluate the goodness of solutions and the next populations.
+Nevertheless, PSO dose not have genetic operators like crossover or mutations.
+Partices are updated by internal comunication among themselfs based on the local and global best.
+In contrast to GA where chromosomes shares information with each others, PSO shares information using memory from previous iteration in form of global best solution.
+This leads the population to move as one group to towards an optimal solution (i.e find solution quickly).
+%which significally
+
+  %From the procedure, we can learn that PSO shares many common points with GA.
+Both algorithms start with a group of a randomly generated population, both have fitness values to evaluate the population.
+Both update the population and search for the optimium with random techniques.
+Both systems do not guarantee success.
+
+%However, PSO does not have genetic operators like crossover and mutation. Particles update themselves with the internal velocity. They also have memory, which is important to the algorithm.
+
+%Compared with genetic algorithms (GAs), the information sharing mechanism in PSO is significantly different.
+In GAs, chromosomes share information with each other.
+So the whole population moves like a one group towards an optimal area.
+In PSO, only gBest (or lBest) gives out the information to others.
+It is a one -way information sharing mechanism.
+The evolution only looks for the best solution.
+Compared with GA, all the particles tend to converge to the best solution quickly even in the local version in most cases.
+ The comparison between PSO and Genetic algorithm was performed in \cite{PSOGAComp,ComparePSOGA05,CoPSAanGA05,PAOvsGA07} and proved that PSO generate faster, better optimal solutions than GA especially in problems with high dimensionality.
+In addition, Rohinsons et al.
+\cite{RohinsonPSGA} proved that PSO performed better than GA in difficult engineering problems \cite{EngineerProblemGA}.
+
+% Eberhart \cite{PSOGAComp}
+%%PSO has been shown to have the same ability to find a global optimum as genetic
+%algorithms, but is able to find optimums faster than genetic algorithms (Panda and Padhy
+%2007; Mouser and Dunn 2005; Hassan et al. 2005)
+
+ A hybrid of PSO and other algorithms was investigated in various systems.
+Veeramachaneni \cite{kaylanGAPSO} used a hybrid of GA and PSO to breed solutions using GA then feed them to PSO.
+They demonstrated that using PSO-GA and PSO outperform using GA as an optimization algorithm.
+
+\subsection{PSO Neighborhood Topology}  
+In general the PSO algorithm have fully connected particle which means that every particle know the global best and local best of each other particle.
+Other forms of cooperation and organization were suggested in various systems.
+\cite{grid} suggested using a grid like topology where particle are connected to four other particles around it.
+A ring and star topology were widely used in \cite{neighbourrhood}.
+The advantages over using different topology than the global best is that they slow down the swarm evolving (they need more time to propagate the best solution) which enable more space exploration and thus helps in eliminating the premature convergence problem.
+In \cite{hiericalPSO}, Janson and Middendorf proposed to use a dynamic and hierarchical neighborhood structure.
+They showed that using such structure helps in maintaining diversity over the solution space in a dynamic problem.
+% a PSO using a dynamic and hierarchical neighborhood structure to handle dynamic optimization problems.
+They demonstrated that such a structure is useful for maintaining some particle diversity in a dynamic environment.
+
+ Other forms of cooperation were also investigated in \cite{facotrsSwarm} where the problem is divided among particles and each particle solves part of the problem.
+\cite{facotrsSwarm} discuss a different behavior that governs cooperation particles in the swarm system.
+%were used There is various topologies used in connecting a swarm in PSO.
+The most used is the totally connected, star and ring topology.
+
+%\subsection*{PSO vs. Genetic Algorithms}  
+% compare   between genetic programming and algorihtms.
+
+\subsection{Drawbacks}
+
+The main drawback of the particle swarm algorithm is the premature convergence problem.
+This happens when global best cannot be improved after a few swarm iterations.
+This is because the use of global best makes PSO converge to one place which is between local best and global best.
+Another reason is because of the high amount of information between particle makes the solution less diverse which limit the explored solution space.
+This problem can be solved by using various methods; one of these methods is choosing another topology rather than the global best topology.
+Another solution is using adaptive inertia $w$ (Section \ref{sec:ParticleSwarmAlgorithm}).
+Other methods of solving problem were introduced as improvement to the PSO algorithm.
+Another problem of PSO is that performance is dependent on the problem and the PSO parameter selected for this specific problem.
+This means that any change in the PSO parameter change the efficiency of PSO solution.
+The solution to this problem is proposed by using self adaptive parameters.
+
+\subsection{Improvements to PSO}
+
+There was various improvement used to improve the PSO algorithm.
+They mostly try to solve the problems premature converge problem and increase the diversity of the particle population.
+Early work on dynamic optimization was introduced by Eberhard and shi \cite{paramterSelection}, where they added the inertia weight version of PSO to find the optimum of 3 dimensional function which change every 100 iteration.
+Others used constrained PSO, where the general PSO Equation \ref{eq:Swarm} is constrained.
+Guaranteed Convergence PSO is introduced in \cite{GCPSO} where a system is guaranteed to converge to at least a local optimal solution.
+A repetition of this algorithm is used to insure finding of an optimal global solution.
+
+Other hybrid system that combines PSO with other techniques was introduced.
+Bayesian network and probability estimation were used to move particle to better solutions.
+Fuzzy rules were used to improve the general PSO algorithms in \cite{shapeFuzzy}.
+
+\subsection{Particle Swarm Applications}
+\label{sec:SwarmApp}
+
+ Since 1995, when Kennedy and Eberhart \cite{PSOFirst,PSO2} introduced Particle Swarm Optimization Algorithm (PSO) the algorithm had been used in every discipline.
+The algorithm was used to train Neural network in \cite{PSOandNeuarl} and proved to perform better than GA algorithm.
+In Image processing discipline PSO was used an optimization algorithm in \cite{SwarmImageProcessing,Swarmimageprocessing2,Colorquantization,SwarmMedical}.
 PSO algorithm was used in \cite{PSOIMageClassification} to classify pixel to simplify the object segmentation process.
+Other system used in shape representation like \cite{shapeFuzzy}.
+Yin \cite{PolygonApproximationPSO} used Discrete PSO to optimally convert digital curves into polygons.
 
@@
 \label{sec:ParticleSwarmAlgorithm}
+
+  %It simulate the  % which has the best or  select a set of parameters that satisfy all constrains
+
+\subsection{Discrete Particle Swarm Equations(DPSO)}
 %\section{Particle Swarm Algorithm}
@@
 %What is particle swarm algorithm and how it was used in related researches. 
-The main idea of \textit{Particle Swarm Algorithm (PSO)} is to represent each agent with a particle from the solution space \cite{PSOFirst}.
-Each agent moves the particle with a direction and velocity $v_{ij}$ based on equations \ref{eq:Swarm} \& \ref{eq:Swarm1}.
+The main idea of \textit{Particle Swarm Algorithm (PSO)} is to represent each solution with a $N$ dimension particle from the solution space \cite{PSOFirst}.
+Each particle moves with a direction and velocity $v_{ij}$ based on equations \ref{eq:Swarm1} \& \ref{eq:Swarm}.
 \begin{equation}
@@
 \end{equation}
-where $p_{ij}$ represent the $jth$ particle in the $ith$ agent and $v_{ij}$ is the velocity of the $jth$ particle in the $ith$ agent.
+where $p_{ij}$ represent the $jth$ dimension in the $ith$ particle and $v_{ij}$ is the velocity of the $jth$ dimension in the $ith$ particle.
  %Equation [\ref{eq:Swarm}] shows how velocity and direction of each particle are computed
  \begin{equation}
-v_{ij}  = v_{ij}  + c_1 r_1 (lbest_{ij}  - p_{ij} ) + c_2 r_2 (gbest_{ij}  - p_{ij} )
+v_{ij}  = v_{ij} W  + c_1 r_1 (lbest_{ij}  - p_{ij} ) + c_2 r_2 (gbest_{ij}  - p_{ij} )
 \label{eq:Swarm}
 \end{equation}
- where $lbest_{ij}$ is the local best particle, $gbest_{ij}$ is the global best particle, $r_1$ \& $r_2$ are random variables and $c_1$ \& $c_2$ are the swarm system variables.
+ where $W$ is the inertia weight parameter which controls the tradeoff between exploration and exploitation, $lbest_{ij}$ is the local best particle, $gbest_{ij}$ is the global best particle, $r_1$ \& $r_2$ are random variables and $c_1$ \& $c_2$ are the swarm acceleration parameters.
  After each iteration the global best $g_{best}$ particle and the agent local best $l_{best}$ particle are evaluated based on the maximum fitness functions of all particles in the solution space.
+
+ The velocity is bounded to maximum of $V_{max}$ using Equation\ref{eq:vmax} to prevent exploding which leads to premature convergence.
 The solution is found after achieving a specific number of iteration or after an error threshold is achieved.
@@
 \end{array}\}
-\end{equation}
- where $p_{ij}$ is the numerical values of the particle and $r_{3}$ is a random variable.
+\end{equation} where $p_{ij}$ is the numerical values of the particle and $r_{3}$ is a random variable. 
+  \begin{equation}
+V_{ij}  = \frac{{V_{ij} }}{{2V_{\max } }} + 0.5
+\label{eq:vmax}
+\end{equation} where $V_{\max }$ is the velocity threshold.
+
+%\subsection{The Algorithm}
+%%\label{sec:ParticleSwarmAlgorithm} 
+%\begin{verbatim}
+% Initialize 
+%    Generate N particles with M dimensions (P_i...P_N) 
+%    Generate velocities for each particle for M dimension randomly. 
+% Repeat till Maximum number of iteration OR Fitness 
+%         For each particle P_i do the following 
+%                Compute fitness  
+%                Determine Lbest  
+%                Check if need to change Gbest. 
+%                Compute the Velocity for each dimension using eq. 
+%                Check if maximum Velocity reach and limit it 
+%                update particle values 
+%          end . 
+%        Improve all particle positions 
+%  end 
+%  
+%  Final solution reaches in Gbest.       
+%        
+%        
+%\end{verbatim}

### 2009-07-28 draft/System.tex

@@
 The next step is \textbf{Segmentation}, the goal of the segmentation stage is to divide strokes into segments of either curves or lines.
+The segmentation problem is summarized as finding the best decomposition with least number of segments each represents a geometric primitive.
+This is an optimization problem, which evolutionary programming can solve efficiently.
+A genetic algorithm was used by \cite{CruveDivisionSwarm} to optimally divide digital curves into lines and curves.
+Chen et al.\cite{CruveDivisionSwarm} uses digital curves scanned from paper as input to the system and did not take advantage of the curvature or local geometric properties of the digital curve.
+
  As shown in Figure \ref{fig:Blockdiagram} first an attempt is made to fit the stroke points into a curve or an ellipse using a minimum square error fitting algorithm \cite{chernov-2003}.
@@
 We used the library implementation of LibSVM \cite{libsvm} and its implementation of the one-vs-one strategy.
+\subsection{Support Vector Machines}
+
+% whate is support vector machines 
+% how decisions is made
+%  how training is done 
+% how testing is donel. 
+% kernels 
+% how multi classifier are implemented 
+% how parameter for cost is measured. 
+% advantages 
+% disadavantages 
+% importance and applications used in it.
+
+%Support vector machines (SVMs) are a set of related supervised learning methods used for classification and regression.
+Viewing input data as two sets of vectors in an n-dimensional space, an SVM will construct a separating hyperplane in that space, one which maximizes the margin between the two data sets.
+To calculate the margin, two parallel hyperplanes are constructed, one on each side of the separating hyperplane, which are "pushed up against" the two data sets.
+Intuitively, a good separation is achieved by the hyperplane that has the largest distance to the neighboring datapoints of both classes, since in general the larger the margin the lower the generalization error of the classifier.
+
+%Support Vector Machines are based on the concept of decision planes that define decision boundaries.
+A decision plane is one that separates between a set of objects having different class memberships.
+A schematic example is shown in the illustration below.
+In this example, the objects belong either to class GREEN or RED.
+The separating line defines a boundary on the right side of which all objects are GREEN and to the left of which all objects are RED.
+Any new object (white circle) falling to the right is labeled, i.e., classified, as GREEN (or classified as RED should it fall to the left of the separating line).
+
+%The above is a classic example of a linear classifier, i.e., a classifier that separates a set of objects into their respective groups (GREEN and RED in this case) with a line.
+Most classification tasks, however, are not that simple, and often more complex structures are needed in order to make an optimal separation, i.e., correctly classify new objects (test cases) on the basis of the examples that are available (train cases).
+This situation is depicted in the illustration below.
+Compared to the previous schematic, it is clear that a full separation of the GREEN and RED objects would require a curve (which is more complex than a line).
+Classification tasks based on drawing separating lines to distinguish between objects of different class memberships are known as hyperplane classifiers.
+Support Vector Machines are particularly suited to handle such tasks.

### 2009-08-01 trunk/draft/System.tex

@@
 \label{eq:direction}
-	\vartheta  = \cos ^{ - 1} \left( {\frac{{\overrightarrow {Q_i}  \cdot \overrightarrow {O_i } }}{{\left\| {\overrightarrow {Q_i} } \right\| \times \left\| {\overrightarrow {O_i} } \right\|}}} \right)
+	\vartheta  = \cos ^{ - 1} \left( {\frac{{\overrightarrow {Q_i}  \cdot \overrightarrow {Px_i } }}{{\left\| {\overrightarrow {Q_i} } \right\| \times \left\| {\overrightarrow {Px_i} } \right\|}}} \right)
 \end{equation}
-where $Q_i=\overrightarrow {P_{i - 1} P_i }$ is the vector from the point  $P_{i - 1}$ to point $P_i$ and $O_i$ represent the $x-axis$ at this point (Figure \ref{fig:curvatureCal.jpg}).
+where $Q_i=\overrightarrow {P_{i - 1} P_i }$ is the vector from the point  $P_{i - 1}$ to point $P_i$ and $Px_i$ represent the $x-axis$ at this point (Figure \ref{fig:curvatureCal.jpg}).
 $\|{\overrightarrow {Q_i }}\|$ is the normal of the  vector $\overrightarrow{Q_i}$
@@
 The additional computation is necessary to  compare the two segmentation fairly.
-The system choose the minimum segmentaiton as the final segmentation of this stroke.
+The system choose the minimum segmentation as the final segmentation of this stroke.
 % The segmentation which crossponds to the minimum $E_{tot}$ is the best segmentation.  
@@
 The system uses support vector machine (SVM) classifier with Linear kernel (Linear kernel) \cite{libsvm}.
-SVM is a binary classifier that maximize the margin between the decision boundary and the support vectors.
-To enable SVM to learn multiclassifiers we used one versus one strategy for combiinig set of binary classifer.
-This stratgy trains $ \frac{n(n-1)}{{2}}$ binary classifier each distinguish between only two shapes.
+SVM is a binary classifier that maximizes the margin between the decision boundary and the support vectors.
+To enable SVM to learn multi classes we used one versus one strategy for combining set of binary classifier.
+This strategy trains $ \frac{n(n-1)}{{2}}$ binary classifier each distinguish between only two shapes.
 The final result is based on a number of votes each shape has from all classifiers.
-We used the a library implementation of LibSVM \cite{libsvm} and its implementation of the one-vs-one strategy.
+We used the library implementation of LibSVM \cite{libsvm} and its implementation of the one-vs-one strategy.
+
+\subsection{Support Vector Machines}
+
+The foundation of Support Vector Machines (SVM) were developed by Vapnik \cite{} in 1995.
+It is gainig increase popularity due to the high performance and efficient implementation.
+The goal of SVM algorithms is to generate a sperating plane between two different classes which are represented using a set of n-dimentional vectors.
+The generated plane is used as a good classifier to the unseen exmaples.
+However, there can be more than one plane that can separates the data therefore SVM attempts to find the plane that maximize the margin (distance between classifier and nearest sample) between the two classes \ref{}.
+To calcuate the margin, two different hyperplanes are constructed, one plane on each side of the data, these two planes are then moved to generenate the maximum margin possilbe.
+
+%The foundations of Support Vector Machines (SVM) have
+
+% whate is support vector machines 
+The classification problem can be restricted to consideration of the two-class problem
+without loss of generality. In this problem the goal is to separate the two classes by a
+function which is induced from available examples. The goal is to produce a classifier
+that will work well on unseen examples, i.e. it generalises well. Consider the example
+in Figure 2.1. Here there are many possible linear classifiers that can separate the data,
+but there is only one that maximises the margin (maximises the distance between it
+and the nearest data point of each class). This linear classifier is termed the optimal
+separating hyperplane. Intuitively, we would expect this boundary to generalise well as
+opposed to the other possible boundaries.
+
+%The above is a classic example of a linear classifier, i.e., a classifier that separates a set of objects into their respective groups (GREEN and RED in this case) with a line.
+Most classification tasks, however, are not that simple, and often more complex structures are needed in order to make an optimal separation, i.e., correctly classify new objects (test cases) on the basis of the examples that are available (train cases).
+This situation is depicted in the illustration below.
+Compared to the previous schematic, it is clear that a full separation of the GREEN and RED objects would require a curve (which is more complex than a line).
+Classification tasks based on drawing separating lines to distinguish between objects of different class memberships are known as hyperplane classifiers.
+Support Vector Machines are particularly suited to handle such tasks.
+
+% how decisions is made
+%  how training is done 
+% how testing is donel. 
+% kernels 
+% how multi classifier are implemented 
+% how parameter for cost is measured. 
+% advantages 
+% disadavantages 
+% importance and applications used in it.

### 2009-08-02 draft/System.tex

@@
 We used the library implementation of LibSVM \cite{libsvm} and its implementation of the one-vs-one strategy.
+
 \subsection{Support Vector Machines}
 
+The foundation of Support Vector Machines (SVM) were developed by Vapnik \cite{} in 1995.
+It is gainig increase popularity due to the high performance and efficient implementation.
+The goal of SVM algorithms is to generate a sperating plane between two different classes which are represented using a set of n-dimentional vectors.
+The generated plane is used as a good classifier to the unseen exmaples.
+However, there can be more than one plane that can separates the data therefore SVM attempts to find the plane that maximize the margin (distance between classifier and nearest sample) between the two classes \ref{}.
+To calcuate the margin, two different hyperplanes are constructed, one plane on each side of the data, these two planes are then moved to generenate the maximum margin possilbe.
+
+%The foundations of Support Vector Machines (SVM) have
+
 % whate is support vector machines 
+The classification problem can be restricted to consideration of the two-class problem
+without loss of generality. In this problem the goal is to separate the two classes by a
+function which is induced from available examples. The goal is to produce a classifier
+that will work well on unseen examples, i.e. it generalises well. Consider the example
+in Figure 2.1. Here there are many possible linear classifiers that can separate the data,
+but there is only one that maximises the margin (maximises the distance between it
+and the nearest data point of each class). This linear classifier is termed the optimal
+separating hyperplane. Intuitively, we would expect this boundary to generalise well as
+opposed to the other possible boundaries.
+
+Let $x_i$ , $i$ = $1.2 \dots,N$ , be the feature vectors of the training set, $X$ . These
+belong to either of two classes that are linearly spereable $\omega_1=1$ and  $\omega_2=-1$.
+The goal, is to design a hyperplane which solve the equation 
+\begin{equation}
+g(x) = \omega^{T} x + \omega g = 0
+\label{eq:svmhyper}
+\end{equation}
+
+%The above is a classic example of a linear classifier, i.e., a classifier that separates a set of objects into their respective groups (GREEN and RED in this case) with a line.
+Most classification tasks, however, are not that simple, and often more complex structures are needed in order to make an optimal separation, i.e., correctly classify new objects (test cases) on the basis of the examples that are available (train cases).
+This situation is depicted in the illustration below.
+Compared to the previous schematic, it is clear that a full separation of the GREEN and RED objects would require a curve (which is more complex than a line).
+Classification tasks based on drawing separating lines to distinguish between objects of different class memberships are known as hyperplane classifiers.
+Support Vector Machines are particularly suited to handle such tasks.
+
 % how decisions is made
@@
 % importance and applications used in it.
-
-%Support vector machines (SVMs) are a set of related supervised learning methods used for classification and regression.
-Viewing input data as two sets of vectors in an n-dimensional space, an SVM will construct a separating hyperplane in that space, one which maximizes the margin between the two data sets.
-To calculate the margin, two parallel hyperplanes are constructed, one on each side of the separating hyperplane, which are "pushed up against" the two data sets.
-Intuitively, a good separation is achieved by the hyperplane that has the largest distance to the neighboring datapoints of both classes, since in general the larger the margin the lower the generalization error of the classifier.
-
-%Support Vector Machines are based on the concept of decision planes that define decision boundaries.
-A decision plane is one that separates between a set of objects having different class memberships.
-A schematic example is shown in the illustration below.
-In this example, the objects belong either to class GREEN or RED.
-The separating line defines a boundary on the right side of which all objects are GREEN and to the left of which all objects are RED.
-Any new object (white circle) falling to the right is labeled, i.e., classified, as GREEN (or classified as RED should it fall to the left of the separating line).
-
-%The above is a classic example of a linear classifier, i.e., a classifier that separates a set of objects into their respective groups (GREEN and RED in this case) with a line.
-Most classification tasks, however, are not that simple, and often more complex structures are needed in order to make an optimal separation, i.e., correctly classify new objects (test cases) on the basis of the examples that are available (train cases).
-This situation is depicted in the illustration below.
-Compared to the previous schematic, it is clear that a full separation of the GREEN and RED objects would require a curve (which is more complex than a line).
-Classification tasks based on drawing separating lines to distinguish between objects of different class memberships are known as hyperplane classifiers.
-Support Vector Machines are particularly suited to handle such tasks.

### 2009-08-04 draft/System.tex

@@
 If the solution produced an error that exceed the error threshold $\varepsilon$ the fitness function assign a -ve value for the solution to express it's in feasibility, otherwise the inverse of the number of vertex produced is used to access the solution fitness in terms of minimum number of vertices'.
+
 This fitness function optimize two goals the first is to move the solution into the feasible solution space with acceptable error bounds and second is to fly the particle to a new position which may result in a polygon with a fewer number of vertices'\cite{PolygonApproximationPSO}.
+The two optimization goals are pursued simultaneously since the DPSO evolves with swarm particles and each of which may invoke different fitness evaluation depending on the incurred approximation error.
 \begin{figure}
@@
 %if we say that s
-%Alg1:
+%Alg1:sinc
 %Alg2: 
@@
 % importance and applications used in it.
+
+\section{Step by Step Example}
+\label{secstepExample}
+In this section a step by step example is presented to clearify the presented sytesm.
+Let us make an assumption that the user will draw a clock using two strokes.
+Figure \ref{} show the two stroke and how the user draw them.
+The next steps demonstrate how ouput of the system after each step
+
+\begin{landscape}
+\begin{scriptsize}
+
+%\begin{table}
+	%\centering
+
+%\scalebox{0.6}{		
+\begin{longtable}{|p{2cm}|p{2cm}|p{2cm}|p{2cm}|p{13cm}|}
+	\caption{StepsStroke1}
+	\label{tab:StepsStroke1} \\
+
+\hline 
+\multicolumn{1}{|p{2cm}|}{\textbf{Step}} & 
+\multicolumn{1}{p{2cm}|}{\textbf{Discription}} &
+ \multicolumn{1}{p{2cm}|}{\textbf{Inputs}} &
+ \multicolumn{1}{p{2cm}|}{\textbf{Details}} &
+ \multicolumn{1}{p{13cm}|}{\textbf{Output}} 
+\\ \hline 
+\endfirsthead
+
+\multicolumn{3}{c}%
+{{\bfseries \tablename\ \thetable{} -- continued from previous page}} \\
+\multicolumn{1}{|p{2cm}|}{\textbf{Step}} & 
+\multicolumn{1}{p{2cm}|}{\textbf{Discription}} &
+ \multicolumn{1}{p{2cm}|}{\textbf{Inputs}} &
+ \multicolumn{1}{p{2cm}|}{\textbf{Details}} &
+ \multicolumn{1}{p{13cm}|}{\textbf{Output}} 
+\\ \hline 
+\endhead
+
+%Step & Discription of Step & Input & Steps & Output \\ \hline
+Stroke point extraction & Extracting point and computing boundary data & -- &  see Section \ref{} & 
+\begin{scriptsize}
+Stroke has 62 points = [P(198.0,153.0), P(197.0,150.0), P(197.0,145.0), P(198.0,140.0), P(203.0,125.0), P(205.0,120.0) , P(205.0,117.0), P(210.0,111.0) , P(214.0,103.0) , P(219.0,97.0), P(225.0,89.0) , P(229.0,85.0) , P(242.0,77.0) , P(247.0,73.0) , P(275.0,64.0) , P(287.0,62.0) , P(325.0,62.0) , P(333.0,63.0) , P(362.0,68.0) , P(372.0,72.0) , P(391.0,77.0) , P(406.0,83.0) , P(424.0,91.0) , P(440.0,100.0), P(474.0,133.0) , P(485.0,149.0) , P(492.0,164.0) , P(499.0,182.0) , P(501.0,191.0) , P(501.0,207.0) , P(499.0,222.0) , P(496.0,239.0) , P(471.0,292.0) , P(467.0,302.0) , P(456.0,315.0) , P(441.0,329.0) , P(427.0,339.0) , P(418.0,344.0) , P(398.0,354.0) , P(336.0,371.0) , P(283.0,376.0) , P(263.0,378.0) , P(242.0,378.0) , P(232.0,376.0) , P(213.0,373.0) , P(194.0,367.0) , P(177.0,362.0) , P(138.0,343.0) , P(129.0,336.0) , P(124.0,331.0) , P(120.0,321.0) , P(115.0,310.0) , P(113.0,297.0) , P(112.0,284.0) , P(114.0,256.0) , P(121.0,219.0) , P(128.0,202.0) , P(137.0,184.0) , P(143.0,172.0) , P(154.0,161.0) , P(172.0,146.0) , P(172.0,146.0)]
+
+ Bonding box is  = Corner P(112.0, 62.0), w= 389.0, h=  316.0
+ number of point in this stroke is   62
+\end{scriptsize}
+ \\ \hline
+Possible Dominate point Extraction & Computing and for details see Section\ref{} &  Stroke points  & 1. compute   the pdp  &    Number of $P_{dp}$  = 11
+ Pdp = [ 2 (197.0 , 145.0 ), 4 (203.0 , 125.0 ), 3 (198.0 , 140.0 ), 22 (424.0 , 91.0 ), 19 (372.0 , 72.0 ), 26 (492.0 , 164.0 ), 28 (501.0 , 191.0 ), 9 (219.0 , 97.0 ), 31 (496.0 , 239.0 ), 0 (198.0 , 153.0 ), 41 (263.0 , 378.0 )]
+
+\\ \hline 
+Segmentation & Segmentation of Ellipse & $P_{dp}$   &  Ellipse fitting  & 
+The circle is   a = 194.5 ,b = 194.5 , Center( 306.5,  220.0 )   with Error =  2.276869953758031 , Percent =0.9093857875960022
+ certainty  =  0.1997008626019765 \\ \hline
+
+		\end{longtable}
+%}
+
+%\end{table}
+
+\end{scriptsize}
+
+\end{landscape}
+
+%\begin{enumerate}
+%\begin{itemize}
+
+%	\item Processing for the first Stroke 1.
+
+%	\item Possible Dominiate Point Extraction
+%	\item Segmentation 
+%	   Computing 
+%	\begin{enumerate}
+%	\item Ellipse detection  
+%	      Calcualting estimated ellipse with radius = $\frac{w}{2}$  and center as same as the bounding box of stroke.  
+%	      The ouput of the system $$
+% \end{enumerate}
+%	
+%	
+%\end{enumerate}
+%  
+%
+%\end{itemize}

### 2009-08-04 draft/SketchMaster.tex

@@
 %\documentstyle[11pt,unhthesis]{report}
-\documentclass[11pt,doublespace]{Sketchthesis}
+\documentclass[11pt,doublespace]{SketchThesis}
 
@@
 \usepackage{longtable}
+\usepackage{lscape}
 \usepackage{array}
@@
 \degreeyear{}                                      %
+\superad{Professor Nevin Darwish\\Professor Samia Mashali\\Assoc. Professor Magda Fayek\\}
 %th today or first day of writing  December 26 2005 
@@
 \documenttype{Thesis}                                   % or dissertation
+
 \maketitle
 
-\providecommand{\tabularnewline}{\\}
                                                         % COPYRIGHT PAGE

### 2009-08-04 trunk/draft/System.tex

@@
 
+Let $x_i$ , $i$ = $1.2 \dots,N$ , be the feature vectors of the training set, $X$ . These
+belong to either of two classes that are linearly spereable $\omega_1=1$ and  $\omega_2=-1$.
+The goal, is to design a hyperplane which solve the equation 
+\begin{equation}
+g(x) = \omega^{T} x + \omega g = 0
+\label{eq:svmhyper}
+\end{equation}
+ % Vladimir N. Vapnik. Statistical Learning Theory. John Wiley & Sons, Inc., New
+% York, NY, September 1998.
+
  %The above is a classic example of a linear classifier, i.e., a classifier that separates a set of objects into their respective groups (GREEN and RED in this case) with a line.
@@
 % importance and applications used in it.
+
+\section{Step by Step Example}
+\label{secstepExample}
+In this section a step by step example is presented to clearify the presented sytesm.
+Let us make an assumption that the user will draw a clock using two strokes.
+Figure \ref{} show the two stroke and how the user draw them.
+The next steps demonstrate how ouput of the system after each step
+
+\begin{landscape}
+\begin{scriptsize}
+
+%\begin{table}
+	%\centering
+
+%\scalebox{0.6}{		
+\begin{longtable}{|p{2cm}|p{2cm}|p{2cm}|p{2cm}|p{13cm}|}
+	\caption{StepsStroke1}
+	\label{tab:StepsStroke1} \\
+
+\hline 
+\multicolumn{1}{|p{2cm}|}{\textbf{Step}} & 
+\multicolumn{1}{p{2cm}|}{\textbf{Discription}} &
+ \multicolumn{1}{p{2cm}|}{\textbf{Inputs}} &
+ \multicolumn{1}{p{2cm}|}{\textbf{Details}} &
+ \multicolumn{1}{p{13cm}|}{\textbf{Output}} 
+\\ \hline 
+\endfirsthead
+
+\hline
+\multicolumn{3}{c}%
+{{\bfseries \tablename\ \thetable{} -- continued from previous page}} \\
+\multicolumn{1}{|p{2cm}|}{\textbf{Step}} & 
+\multicolumn{1}{p{2cm}|}{\textbf{Discription}} &
+ \multicolumn{1}{p{2cm}|}{\textbf{Inputs}} &
+ \multicolumn{1}{p{2cm}|}{\textbf{Details}} &
+ \multicolumn{1}{p{13cm}|}{\textbf{Output}} 
+\\ \hline 
+\endhead
+
+%Step & Discription of Step & Input & Steps & Output \\ \hline
+Stroke 1 & Extracting point and computing boundary data & -- &  see Section \ref{sec:Preprocessing} & 
+\begin{scriptsize}
+Stroke has 62 points = [P(198.0,153.0), P(197.0,150.0), P(197.0,145.0), P(198.0,140.0), P(203.0,125.0), P(205.0,120.0) , P(205.0,117.0), P(210.0,111.0) , P(214.0,103.0) , P(219.0,97.0), P(225.0,89.0) , P(229.0,85.0) , P(242.0,77.0) , P(247.0,73.0) , P(275.0,64.0) , P(287.0,62.0) , P(325.0,62.0) , P(333.0,63.0) , P(362.0,68.0) , P(372.0,72.0) , P(391.0,77.0) , P(406.0,83.0) , P(424.0,91.0) , P(440.0,100.0), P(474.0,133.0) , P(485.0,149.0) , P(492.0,164.0) , P(499.0,182.0) , P(501.0,191.0) , P(501.0,207.0) , P(499.0,222.0) , P(496.0,239.0) , P(471.0,292.0) , P(467.0,302.0) , P(456.0,315.0) , P(441.0,329.0) , P(427.0,339.0) , P(418.0,344.0) , P(398.0,354.0) , P(336.0,371.0) , P(283.0,376.0) , P(263.0,378.0) , P(242.0,378.0) , P(232.0,376.0) , P(213.0,373.0) , P(194.0,367.0) , P(177.0,362.0) , P(138.0,343.0) , P(129.0,336.0) , P(124.0,331.0) , P(120.0,321.0) , P(115.0,310.0) , P(113.0,297.0) , P(112.0,284.0) , P(114.0,256.0) , P(121.0,219.0) , P(128.0,202.0) , P(137.0,184.0) , P(143.0,172.0) , P(154.0,161.0) , P(172.0,146.0) , P(172.0,146.0)]
+
+ Bonding box is  = Corner P(112.0, 62.0), w= 389.0, h=  316.0
+ number of point in this stroke is   62
+\end{scriptsize}
+ \\ \hline
+Possible Dominate point Extraction & Computing and for details see Section\ref{sec:Preprocessing} &  Stroke points  & 1. compute   the pdp  &    Number of $P_{dp}$  = 11
+ Pdp = [ 2 (197.0 , 145.0 ), 4 (203.0 , 125.0 ), 3 (198.0 , 140.0 ), 22 (424.0 , 91.0 ), 19 (372.0 , 72.0 ), 26 (492.0 , 164.0 ), 28 (501.0 , 191.0 ), 9 (219.0 , 97.0 ), 31 (496.0 , 239.0 ), 0 (198.0 , 153.0 ), 41 (263.0 , 378.0 )]
+
+\\ \hline 
+Segmentation & Segmentation of Ellipse & $P_{dp}$   &  Ellipse fitting  & 
+ Segmented as Ellipse with dimensions  a = 194.5 ,b = 194.5 , Center( 306.5,  220.0 )   with Error =  2.276869953758031 , Percent =0.9093857875960022
+ certainty  =  0.1997008626019765 
+\\ \hline
+
+Stroke 2 & Extracting point and computing boundary data & -- &  see Section \ref{sec:Preprocessing} &
+\begin{scriptsize}
+Stroke has 205 points = [P(201.0,179.0) , P(202.0,180.0) , P(202.0,182.0) , P(202.0,185.0) , P(202.0,187.0) , P(202.0,190.0) , P(202.0,192.0) , P(202.0,196.0) , P(202.0,202.0) , P(201.0,206.0) , P(200.0,219.0) , P(198.0,231.0) , P(198.0,234.0) , P(198.0,238.0) , P(197.0,244.0) , P(197.0,248.0) , P(197.0,251.0) , P(196.0,258.0) , %P(196.0,260.0) , P(196.0,263.0) , P(195.0,265.0) , P(195.0,266.0) , P(195.0,269.0) , P(195.0,271.0) , P(195.0,273.0) , P(195.0,278.0) , P(195.0,279.0) , P(195.0,281.0) , P(195.0,282.0) , P(195.0,284.0) , P(195.0,285.0) , P(195.0,288.0) , P(195.0,290.0) , P(195.0,295.0) , P(195.0,297.0) , P(195.0,298.0) , P(196.0,298.0) , P(196.0,299.0) , P(196.0,300.0) , P(196.0,301.0) , P(197.0,301.0) , P(198.0,301.0) , P(200.0,301.0) , P(203.0,301.0) , %P(206.0,301.0) , P(208.0,301.0) , P(213.0,301.0) , P(217.0,302.0) , P(219.0,302.0) , P(223.0,303.0) , P(234.0,304.0) , P(237.0,304.0) , P(240.0,306.0) , P(246.0,306.0) , P(247.0,307.0) , P(248.0,307.0) , P(249.0,307.0) , P(250.0,307.0) , P(250.0,306.0) , P(250.0,305.0) , P(250.0,304.0) , P(250.0,302.0) , P(251.0,299.0) , P(251.0,296.0) , P(251.0,294.0) , P(251.0,289.0) , P(252.0,285.0) , P(253.0,279.0) , P(253.0,274.0) , P(253.0,271.0) , P(253.0,264.0) , P(253.0,257.0) , P(254.0,252.0) , P(254.0,241.0) , P(254.0,237.0) , P(255.0,232.0) , P(255.0,229.0) , P(255.0,225.0) , P(256.0,212.0) , P(256.0,209.0) , P(256.0,206.0) , P(257.0,205.0) , P(257.0,200.0) , P(257.0,197.0) , P(258.0,195.0) , P(258.0,194.0) , P(258.0,189.0) , P(258.0,187.0) , P(258.0,186.0) , P(259.0,184.0) , P(259.0,181.0) , P(259.0,179.0) , P(259.0,178.0) , P(260.0,176.0) , P(260.0,174.0) , P(260.0,173.0) , P(260.0,172.0) , P(260.0,171.0) , P(260.0,169.0) , P(260.0,168.0) , P(260.0,167.0) , P(261.0,166.0) , P(261.0,165.0) , P(261.0,164.0) , P(261.0,162.0) , P(262.0,162.0) , P(263.0,162.0) , P(264.0,162.0) , P(268.0,162.0) , P(275.0,162.0) , P(279.0,162.0) , P(289.0,162.0) , P(291.0,162.0) , P(301.0,163.0) , P(305.0,164.0) , P(312.0,164.0) , P(315.0,165.0) , P(318.0,165.0) , P(319.0,165.0) , P(319.0,166.0) , P(320.0,168.0) , P(320.0,171.0) , P(320.0,173.0) , P(320.0,176.0) , P(321.0,182.0) , P(321.0,191.0) , P(322.0,198.0) , P(322.0,211.0) , P(323.0,226.0) , P(324.0,229.0) , P(324.0,238.0) , P(324.0,263.0) , P(325.0,272.0) , P(325.0,281.0) , P(325.0,288.0) , P(325.0,291.0) , P(325.0,297.0) , P(325.0,299.0) , P(325.0,301.0) , P(325.0,305.0) , P(324.0,306.0) , P(324.0,307.0) , P(323.0,308.0) , P(324.0,307.0) , P(325.0,307.0) , P(328.0,306.0) , P(332.0,306.0) , P(339.0,306.0) , P(347.0,305.0) , P(354.0,304.0) , P(365.0,304.0) , P(376.0,304.0) , P(379.0,304.0) , P(385.0,304.0) , P(391.0,304.0) , P(392.0,304.0) , P(393.0,304.0) , P(393.0,302.0) , P(393.0,301.0) , P(394.0,297.0) , P(394.0,290.0) , P(393.0,284.0) , P(393.0,274.0) , P(392.0,258.0) , P(391.0,245.0) , P(391.0,233.0) , P(389.0,213.0) , P(388.0,198.0) , P(388.0,189.0) , P(388.0,184.0) , P(388.0,178.0) , P(388.0,177.0) , P(388.0,176.0) , P(388.0,174.0)
+..................................................................................................................
+.................................................................................
+, P(388.0,173.0) , P(388.0,171.0) , P(388.0,170.0) , P(388.0,168.0) , P(388.0,167.0) , P(388.0,166.0) , P(388.0,165.0) , P(389.0,165.0) , P(390.0,165.0) , P(395.0,165.0) , P(401.0,165.0) , P(406.0,165.0) , P(418.0,166.0) , P(424.0,167.0) , P(426.0,167.0) , P(429.0,167.0) , P(430.0,167.0) , P(431.0,167.0) , P(432.0,167.0) , P(433.0,169.0) , P(434.0,170.0) , P(434.0,174.0) , P(436.0,185.0) , P(440.0,200.0) , P(442.0,220.0) , P(442.0,224.0) , P(446.0,267.0) , P(447.0,279.0) , P(448.0,286.0) , P(448.0,287.0) , P(448.0,287.0)]
+
+ Bonding box is  = Corner P(195.0, 162.0), w= 253.0, h=  146.0
+ \end{scriptsize}
+  \\ \hline
+  & Possible Dominate point Extraction for details see Section\ref{sec:Preprocessing} &  Stroke points  & 1. compute the $P_{pd}$  & \begin{scriptsize}   Number of $P_{pd}$  =  36
+
+ Pdp = [ 25 (195.0 , 278.0 ), 30 (195.0 , 285.0 ), 33 (195.0 , 295.0 ), 27 (195.0 , 281.0 ), 34 (195.0 , 297.0 ), 46 (213.0 , 301.0 ), 47 (217.0 , 302.0 ), 50 (234.0 , 304.0 ), 28 (195.0 , 282.0 ), 94 (260.0 , 174.0 ), 93 (260.0 , 176.0 ), 95 (260.0 , 173.0 ), 98 (260.0 , 169.0 ), 48 (219.0 , 302.0 ), 99 (260.0 , 168.0 ), 109 (275.0 , 162.0 ), 113 (301.0 , 163.0 ), 94 (260.0 , 174.0 ), 132 (325.0 , 272.0 ), 136 (325.0 , 297.0 ), 112 (291.0 , 162.0 ), 137 (325.0 , 299.0 ), 110 (279.0 , 162.0 ), 144 (325.0 , 307.0 ), 147 (339.0 , 306.0 ), 150 (365.0 , 304.0 ), 132 (325.0 , 272.0 ), 168 (388.0 , 189.0 ), 171 (388.0 , 177.0 ), 174 (388.0 , 173.0 ), 146 (332.0 , 306.0 ), 175 (388.0 , 171.0 ), 183 (395.0 , 165.0 ), 186 (418.0 , 166.0 ), 169 (388.0 , 184.0 ), 191 (431.0 , 167.0 ) ]
+
+  \end{scriptsize}
+    \\ \hline
+ \\ \hline 
+Segmentation & Segmentation of Ellipse & $P_{dp}$   &  Ellipse fitting  & 
+  The Ellipse is   a = 125.5 ,b = 125.5 , Center( 320.5,  235.0 )   with Error =  7.074966523515405 , Percent =1.1481777694994926 certainty  =  0.08114368920921669  
+ \\ \hline
+ Segmentation & PSO Segmentation ALgS1 & $P_{dp}$  & ALgS1 &   Particles =  [  1 0 1 0 0 1 1 1 1 1 1 0 0 0 1 0 0 0 1 ] 
+ The particle  Particles =  [  1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 1 0 0 0 0 1 0 ]
+
+ First Polygonization PSO = [[[  S0 Line[P(201.0,179.0) to P(195.0,281.0)]  ,      S1 Line[P(195.0,281.0) to P(234.0,304.0)]  ,      S2 Line[P(234.0,304.0) to P(260.0,174.0)]  ,      S3 Line[P(260.0,174.0) to P(325.0,272.0)]  ,      S4 Line[P(325.0,272.0) to P(365.0,304.0)]  ,      S5 Line[P(365.0,304.0) to P(388.0,189.0)]  ,      S6 Line[P(388.0,189.0) to P(431.0,167.0)]  ,      S7 Line[P(431.0,167.0) to P(442.0,220.0)]  ]]]
+ \\ \hline
+
+		\end{longtable}
+%}
+
+%\end{table}
+
+\end{scriptsize}
+
+\end{landscape}
+
+%\begin{enumerate}
+%\begin{itemize}
+
+%	\item Processing for the first Stroke 1.
+
+%	\item Possible Dominiate Point Extraction
+%	\item Segmentation 
+%	   Computing 
+%	\begin{enumerate}
+%	\item Ellipse detection  
+%	      Calcualting estimated ellipse with radius = $\frac{w}{2}$  and center as same as the bounding box of stroke.  
+%	      The ouput of the system $$
+% \end{enumerate}
+%	
+%	
+%\end{enumerate}
+%  
+%
+%\end{itemize}

### 2009-08-06 trunk/draft/System.tex
after the chages in eri for svm

@@
 
+Let us consider  a two class classification task with datapoints $x_i$ where $i=1,\dotes,m$ having corresponding labels $y_i=\pm1$ where if  $y_i=1$ if $x_i$ belonged to first class and $y_i=-1$ if $x_i$ belonged to second class.
+Each data  point $x_i$ is represented in $N$ dimensional input or attribule space.
+Let the classification plane be represented as function .
+\begin{equation}
+ y=f(x)=sign(\omega \cdot x - b)
+\label{eq:planeEq}
+\end{equation}
+
+ The vector $\omega$ determines th orientation of a seperation plane.
+The scalar $b$ determines the offeset of the plane from the orgin.
+Lets assume that the datatpointa re linearly seperable, i.e.
+there are infinitely many possible seperating planes that correctly classify the training data.
+Figure\ref{} illustrates two different seperating planes.
+It is intuitively noted from the figure that the () line is more likely to generalize more data and thus introduce less classification error.
+This can be geometrically characterized as plane furthest from both classes.
+% Which one is preferable ?
+Intuitively one perfers the solid pane since small perturbations of any point would not introduce misclassifications errors.
+Without any information, the solid plane is more likely to generalize better on future data.
+Geometrically we can characterize the solid plane as being furthuret from both classes.
+
+The SVM problem is how to construct the plane that is furthuest from both classes.
+One approcah is to maximize the margin between two parallel planes supporting.
+A plane supports a class if all points in that class are on one side of that plane.
+For the points with the class label $+1$ we would like there exist $\omega$ and $b$ such that $\omega \cdot x_i \> b$ or $\omega \cdot x_i -b \> 0$  depending on the class label.
+Let us suppose that smallest value of $\mid \omega \cdot x_i -b\mid$ iss $\kappa$, then when $ \omega \cdot x_i -b \geq \kappa$.
+The argument insidee the decision function is invariant under positive rescalling so we will implicity fix a scale by requiring $\omega \cdot x_i -b \geq 1 $.
+For the points with class label $-1$ we similary require $\omega \cdot x_i -b \leq -1 $.
+To find the plane furthest from both setss, we can simply maximize the distance or margin between the support planes for each class as illustrated in Figure \ref{}.
+The support planes are pussshed apart untill they bump inot small number of data points ( the support vecotrs ) for each class.
+The support vectors in Figure \ref{} are outlined in (),
+
+The distance or margin between these supporting planes 
+\begin{equation}
+ \omega \cdot x=b+1
+  \omega \cdot x=b-1
+\end{equation}
+is 
+\begin{equation}
+ \gamma = \frac{2}{\parallel \omega \parallel^2}
+\end{equation}
+Thus maximizing the margin is equivalent to minimizing $\frac{{\parallel \omega \parallel}^2}{2}$ int the following quadratic program
+
+\begin{equation}
+ min  \frac{{\parallel \omega \parallel}^2}{2}
+\end{equation}
+\begin{equation}
+ s.t.   \omega \cdot x_i \geq b+1    y_i \in Class 1 
+ \omega \cdot x_i \leq b-1    y_i \in Class -1 
+\end{equation}
+
+The constrain can be simplified to $y_i(\omega \codt x_i-b)\geq 1$.
+
+In the maximumm marg
+
 %The foundations of Support Vector Machines (SVM) have
@@
 % whate is support vector machines 
-The classification problem can be restricted to consideration of the two-class problem
-without loss of generality. In this problem the goal is to separate the two classes by a
-function which is induced from available examples. The goal is to produce a classifier
-that will work well on unseen examples, i.e. it generalises well. Consider the example
-in Figure 2.1. Here there are many possible linear classifiers that can separate the data,
-but there is only one that maximises the margin (maximises the distance between it
-and the nearest data point of each class). This linear classifier is termed the optimal
-separating hyperplane. Intuitively, we would expect this boundary to generalise well as
-opposed to the other possible boundaries.
-
-Let $x_i$ , $i$ = $1.2 \dots,N$ , be the feature vectors of the training set, $X$ . These
-belong to either of two classes that are linearly spereable $\omega_1=1$ and  $\omega_2=-1$.
-The goal, is to design a hyperplane which solve the equation 
-\begin{equation}
-g(x) = \omega^{T} x + \omega g = 0
-\label{eq:svmhyper}
-\end{equation}
+%%The classifcation problem can be restricted to consideration of the two-class problem
+%without loss of generality. In this problem the goal is to separate the two classes by a
+%function which is induced from available examples. The goal is to produce a classifier
+%that will work well on unseen examples, i.e. it generalises well. Consider the example
+%in Figure 2.1. Here there are many possible linear classifiers that can separate the data,
+%but there is only one that maximises the margin (maximises the distance between it
+%and the nearest data point of each class). This linear classifier is termed the optimal
+%separating hyperplane. Intuitively, we would expect this boundary to generalise well as
+%opposed to the other possible boundaries.
+%
+%Let $x_i$ , $i$ = $1.2 \dots,N$ , be the feature vectors of the training set, $X$ . These
+%belong to either of two classes that are linearly spereable $\omega_1=1$ and  $\omega_2=-1$.
+%The goal, is to design a hyperplane which solve the equation 
+%\begin{equation}
+%g(x) = \omega^{T} x + \omega g = 0
+%\label{eq:svmhyper}
+%\end{equation}
  % Vladimir N. Vapnik. Statistical Learning Theory. John Wiley & Sons, Inc., New

### 2009-08-07 draft/System.tex
adding the svm before statical approcah

@@
 
+Let us consider  a two class classification task with datapoints $x_i$ where $i=1,\cdots,m$ having corresponding labels $y_i=\pm1$ where if  $y_i=1$ if $x_i$ belonged to first class and $y_i=-1$ if $x_i$ belonged to second class.
+Each data  point $x_i$ is represented in $N$ dimensional input or attribule space.
+Let the classification plane be represented as function .
+\begin{equation}
+ y=f(x)=sign(\omega \cdot x - b)
+\label{eq:planeEq}
+\end{equation}
+
+ The vector $\omega$ determines th orientation of a seperation plane.
+The scalar $b$ determines the offeset of the plane from the orgin.
+Lets assume that the datatpointa re linearly seperable, i.e.
+there are infinitely many possible seperating planes that correctly classify the training data.
+Figure\ref{} illustrates two different seperating planes.
+It is intuitively noted from the figure that the () line is more likely to generalize more data and thus introduce less classification error.
+This can be geometrically characterized as plane furthest from both classes.
+% Which one is preferable ?
+Intuitively one perfers the solid pane since small perturbations of any point would not introduce misclassifications errors.
+Without any information, the solid plane is more likely to generalize better on future data.
+Geometrically we can characterize the solid plane as being furthuret from both classes.
+
+The SVM problem is how to construct the plane that is furthuest from both classes.
+One approcah is to maximize the margin between two parallel planes supporting.
+A plane supports a class if all points in that class are on one side of that plane.
+For the points with the class label $+1$ we would like there exist $\omega$ and $b$ such that $\omega \cdot x_i\> b$ or $\omega \cdot x_i -b \> 0$  depending on the class label.
+Let us suppose that smallest value of $\mid \omega \cdot x_i -b\mid$ iss $\kappa$, then when $ \omega \cdot x_i -b \geq \kappa$.
+The argument insidee the decision function is invariant under positive rescalling so we will implicity fix a scale by requiring $\omega \cdot x_i -b \geq 1 $.
+For the points with class label $-1$ we similary require $\omega \cdot x_i -b \leq -1 $.
+To find the plane furthest from both setss, we can simply maximize the distance or margin between the support planes for each class as illustrated in Figure \ref{}.
+The support planes are pussshed apart untill they bump inot small number of data points ( the support vecotrs ) for each class.
+The support vectors in Figure \ref{} are outlined in (),
+
+The distance or margin between these supporting planes 
+\begin{equation}
+ \omega \cdot x=b+1 \\
+  \omega \cdot x=b-1
+\end{equation}
+is 
+\begin{equation}
+ \gamma = \frac{2}{\parallel \omega \parallel^2}
+\end{equation}
+Thus maximizing the margin is equivalent to minimizing $\frac{{\parallel \omega \parallel}^2}{2}$ int the following quadratic program
+
+\begin{equation}
+ \min_{\omega,b}  \frac{{\parallel \omega \parallel}^2}{2}
+\end{equation}
+\begin{equation}	
+ s.t. \ \ \   \omega \cdot x_i \geq b+1 \ \ \    y_i \in Class 1 \\ 
+ \ \ \ \ \ \   \omega \cdot x_i \leq b-1 \ \ \    y_i \in Class -1 
+\end{equation}
+
+The constrain can be simplified to $y_i(\omega \cdot x_i-b)\geq 1$.
+
+The margin depends on the Support vecotrs that are bumpt into the planes.
+This is a primal problem Quadratic Programming (QP) problem.
+The concept of duality can also upsed to wields the problem to a  Largrnigian dual QP problem in Equation \ref{eq:dual}.
+% maximumm margin between parallel 
+\begin{equation}
+\min_{\alpha} \ \ \   \frac{1}{2} \sum_{i=1}^{m}{\sum_{j=1}^{m}{y_iy_j\alpha_i\alpha_jx_ix_j}} - \sum_{i=1}^{m}{\alpha_i} \\
+s.t. \ \ \sum_{i=1}^{m}{y_i\alpha_i}=0   \\
+  \alpha_i \geq 0  \ \ \ i=1,\cdots , m 
+\label{eq:dual}
+\end{equation}
+Both the dual form and the primal form weilds to the same plane $\omega=\sum_{i=1}^{m}y_i\alpha_i x_i$ and threshold $b$ determined by the support vectors (for which $\alpha_i\geq 0$).
+
 %The foundations of Support Vector Machines (SVM) have
@@
 % whate is support vector machines 
-The classification problem can be restricted to consideration of the two-class problem
-without loss of generality. In this problem the goal is to separate the two classes by a
-function which is induced from available examples. The goal is to produce a classifier
-that will work well on unseen examples, i.e. it generalises well. Consider the example
-in Figure 2.1. Here there are many possible linear classifiers that can separate the data,
-but there is only one that maximises the margin (maximises the distance between it
-and the nearest data point of each class). This linear classifier is termed the optimal
-separating hyperplane. Intuitively, we would expect this boundary to generalise well as
-opposed to the other possible boundaries.
-
-Let $x_i$ , $i$ = $1.2 \dots,N$ , be the feature vectors of the training set, $X$ . These
-belong to either of two classes that are linearly spereable $\omega_1=1$ and  $\omega_2=-1$.
-The goal, is to design a hyperplane which solve the equation 
-\begin{equation}
-g(x) = \omega^{T} x + \omega g = 0
-\label{eq:svmhyper}
-\end{equation}
+%%The classifcation problem can be restricted to consideration of the two-class problem
+%without loss of generality. In this problem the goal is to separate the two classes by a
+%function which is induced from available examples. The goal is to produce a classifier
+%that will work well on unseen examples, i.e. it generalises well. Consider the example
+%in Figure 2.1. Here there are many possible linear classifiers that can separate the data,
+%but there is only one that maximises the margin (maximises the distance between it
+%and the nearest data point of each class). This linear classifier is termed the optimal
+%separating hyperplane. Intuitively, we would expect this boundary to generalise well as
+%opposed to the other possible boundaries.
+%
+%Let $x_i$ , $i$ = $1.2 \dots,N$ , be the feature vectors of the training set, $X$ . These
+%belong to either of two classes that are linearly spereable $\omega_1=1$ and  $\omega_2=-1$.
+%The goal, is to design a hyperplane which solve the equation 
+%\begin{equation}
+%g(x) = \omega^{T} x + \omega g = 0
+%\label{eq:svmhyper}
+%\end{equation}
+ % Vladimir N. Vapnik. Statistical Learning Theory. John Wiley & Sons, Inc., New
+% York, NY, September 1998.
 
@@
 
+\hline
 \multicolumn{3}{c}%
@@
 %Step & Discription of Step & Input & Steps & Output \\ \hline
-Stroke point extraction & Extracting point and computing boundary data & -- &  see Section \ref{} & 
+Stroke 1 & Extracting point and computing boundary data & -- &  see Section \ref{sec:Preprocessing} & 
 \begin{scriptsize}
@@
  \\ \hline
-Possible Dominate point Extraction & Computing and for details see Section\ref{} &  Stroke points  & 1. compute   the pdp  &    Number of $P_{dp}$  = 11
+Possible Dominate point Extraction & Computing and for details see Section\ref{sec:Preprocessing} &  Stroke points  & 1. compute   the pdp  &    Number of $P_{dp}$  = 11
  Pdp = [ 2 (197.0 , 145.0 ), 4 (203.0 , 125.0 ), 3 (198.0 , 140.0 ), 22 (424.0 , 91.0 ), 19 (372.0 , 72.0 ), 26 (492.0 , 164.0 ), 28 (501.0 , 191.0 ), 9 (219.0 , 97.0 ), 31 (496.0 , 239.0 ), 0 (198.0 , 153.0 ), 41 (263.0 , 378.0 )]
@@
 Segmentation & Segmentation of Ellipse & $P_{dp}$   &  Ellipse fitting  & 
-The circle is   a = 194.5 ,b = 194.5 , Center( 306.5,  220.0 )   with Error =  2.276869953758031 , Percent =0.9093857875960022
- certainty  =  0.1997008626019765 \\ \hline
+ Segmented as Ellipse with dimensions  a = 194.5 ,b = 194.5 , Center( 306.5,  220.0 )   with Error =  2.276869953758031 , Percent =0.9093857875960022
+ certainty  =  0.1997008626019765 
+\\ \hline
+
+Stroke 2 & Extracting point and computing boundary data & -- &  see Section \ref{sec:Preprocessing} &
+\begin{scriptsize}
+Stroke has 205 points = [P(201.0,179.0) , P(202.0,180.0) , P(202.0,182.0) , P(202.0,185.0) , P(202.0,187.0) , P(202.0,190.0) , P(202.0,192.0) , P(202.0,196.0) , P(202.0,202.0) , P(201.0,206.0) , P(200.0,219.0) , P(198.0,231.0) , P(198.0,234.0) , P(198.0,238.0) , P(197.0,244.0) , P(197.0,248.0) , P(197.0,251.0) , P(196.0,258.0) , %P(196.0,260.0) , P(196.0,263.0) , P(195.0,265.0) , P(195.0,266.0) , P(195.0,269.0) , P(195.0,271.0) , P(195.0,273.0) , P(195.0,278.0) , P(195.0,279.0) , P(195.0,281.0) , P(195.0,282.0) , P(195.0,284.0) , P(195.0,285.0) , P(195.0,288.0) , P(195.0,290.0) , P(195.0,295.0) , P(195.0,297.0) , P(195.0,298.0) , P(196.0,298.0) , P(196.0,299.0) , P(196.0,300.0) , P(196.0,301.0) , P(197.0,301.0) , P(198.0,301.0) , P(200.0,301.0) , P(203.0,301.0) , %P(206.0,301.0) , P(208.0,301.0) , P(213.0,301.0) , P(217.0,302.0) , P(219.0,302.0) , P(223.0,303.0) , P(234.0,304.0) , P(237.0,304.0) , P(240.0,306.0) , P(246.0,306.0) , P(247.0,307.0) , P(248.0,307.0) , P(249.0,307.0) , P(250.0,307.0) , P(250.0,306.0) , P(250.0,305.0) , P(250.0,304.0) , P(250.0,302.0) , P(251.0,299.0) , P(251.0,296.0) , P(251.0,294.0) , P(251.0,289.0) , P(252.0,285.0) , P(253.0,279.0) , P(253.0,274.0) , P(253.0,271.0) , P(253.0,264.0) , P(253.0,257.0) , P(254.0,252.0) , P(254.0,241.0) , P(254.0,237.0) , P(255.0,232.0) , P(255.0,229.0) , P(255.0,225.0) , P(256.0,212.0) , P(256.0,209.0) , P(256.0,206.0) , P(257.0,205.0) , P(257.0,200.0) , P(257.0,197.0) , P(258.0,195.0) , P(258.0,194.0) , P(258.0,189.0) , P(258.0,187.0) , P(258.0,186.0) , P(259.0,184.0) , P(259.0,181.0) , P(259.0,179.0) , P(259.0,178.0) , P(260.0,176.0) , P(260.0,174.0) , P(260.0,173.0) , P(260.0,172.0) , P(260.0,171.0) , P(260.0,169.0) , P(260.0,168.0) , P(260.0,167.0) , P(261.0,166.0) , P(261.0,165.0) , P(261.0,164.0) , P(261.0,162.0) , P(262.0,162.0) , P(263.0,162.0) , P(264.0,162.0) , P(268.0,162.0) , P(275.0,162.0) , P(279.0,162.0) , P(289.0,162.0) , P(291.0,162.0) , P(301.0,163.0) , P(305.0,164.0) , P(312.0,164.0) , P(315.0,165.0) , P(318.0,165.0) , P(319.0,165.0) , P(319.0,166.0) , P(320.0,168.0) , P(320.0,171.0) , P(320.0,173.0) , P(320.0,176.0) , P(321.0,182.0) , P(321.0,191.0) , P(322.0,198.0) , P(322.0,211.0) , P(323.0,226.0) , P(324.0,229.0) , P(324.0,238.0) , P(324.0,263.0) , P(325.0,272.0) , P(325.0,281.0) , P(325.0,288.0) , P(325.0,291.0) , P(325.0,297.0) , P(325.0,299.0) , P(325.0,301.0) , P(325.0,305.0) , P(324.0,306.0) , P(324.0,307.0) , P(323.0,308.0) , P(324.0,307.0) , P(325.0,307.0) , P(328.0,306.0) , P(332.0,306.0) , P(339.0,306.0) , P(347.0,305.0) , P(354.0,304.0) , P(365.0,304.0) , P(376.0,304.0) , P(379.0,304.0) , P(385.0,304.0) , P(391.0,304.0) , P(392.0,304.0) , P(393.0,304.0) , P(393.0,302.0) , P(393.0,301.0) , P(394.0,297.0) , P(394.0,290.0) , P(393.0,284.0) , P(393.0,274.0) , P(392.0,258.0) , P(391.0,245.0) , P(391.0,233.0) , P(389.0,213.0) , P(388.0,198.0) , P(388.0,189.0) , P(388.0,184.0) , P(388.0,178.0) , P(388.0,177.0) , P(388.0,176.0) , P(388.0,174.0)
+..................................................................................................................
+.................................................................................
+, P(388.0,173.0) , P(388.0,171.0) , P(388.0,170.0) , P(388.0,168.0) , P(388.0,167.0) , P(388.0,166.0) , P(388.0,165.0) , P(389.0,165.0) , P(390.0,165.0) , P(395.0,165.0) , P(401.0,165.0) , P(406.0,165.0) , P(418.0,166.0) , P(424.0,167.0) , P(426.0,167.0) , P(429.0,167.0) , P(430.0,167.0) , P(431.0,167.0) , P(432.0,167.0) , P(433.0,169.0) , P(434.0,170.0) , P(434.0,174.0) , P(436.0,185.0) , P(440.0,200.0) , P(442.0,220.0) , P(442.0,224.0) , P(446.0,267.0) , P(447.0,279.0) , P(448.0,286.0) , P(448.0,287.0) , P(448.0,287.0)]
+
+ Bonding box is  = Corner P(195.0, 162.0), w= 253.0, h=  146.0
+ \end{scriptsize}
+  \\ \hline
+  & Possible Dominate point Extraction for details see Section\ref{sec:Preprocessing} &  Stroke points  & 1. compute the $P_{pd}$  & \begin{scriptsize}   Number of $P_{pd}$  =  36
+
+ Pdp = [ 25 (195.0 , 278.0 ), 30 (195.0 , 285.0 ), 33 (195.0 , 295.0 ), 27 (195.0 , 281.0 ), 34 (195.0 , 297.0 ), 46 (213.0 , 301.0 ), 47 (217.0 , 302.0 ), 50 (234.0 , 304.0 ), 28 (195.0 , 282.0 ), 94 (260.0 , 174.0 ), 93 (260.0 , 176.0 ), 95 (260.0 , 173.0 ), 98 (260.0 , 169.0 ), 48 (219.0 , 302.0 ), 99 (260.0 , 168.0 ), 109 (275.0 , 162.0 ), 113 (301.0 , 163.0 ), 94 (260.0 , 174.0 ), 132 (325.0 , 272.0 ), 136 (325.0 , 297.0 ), 112 (291.0 , 162.0 ), 137 (325.0 , 299.0 ), 110 (279.0 , 162.0 ), 144 (325.0 , 307.0 ), 147 (339.0 , 306.0 ), 150 (365.0 , 304.0 ), 132 (325.0 , 272.0 ), 168 (388.0 , 189.0 ), 171 (388.0 , 177.0 ), 174 (388.0 , 173.0 ), 146 (332.0 , 306.0 ), 175 (388.0 , 171.0 ), 183 (395.0 , 165.0 ), 186 (418.0 , 166.0 ), 169 (388.0 , 184.0 ), 191 (431.0 , 167.0 ) ]
+
+  \end{scriptsize}
+    \\ \hline
+ \\ \hline 
+Segmentation & Segmentation of Ellipse & $P_{dp}$   &  Ellipse fitting  & 
+  The Ellipse is   a = 125.5 ,b = 125.5 , Center( 320.5,  235.0 )   with Error =  7.074966523515405 , Percent =1.1481777694994926 certainty  =  0.08114368920921669  
+ \\ \hline
+ Segmentation & PSO Segmentation ALgS1 & $P_{dp}$  & ALgS1 &   Particles =  [  1 0 1 0 0 1 1 1 1 1 1 0 0 0 1 0 0 0 1 ] 
+ The particle  Particles =  [  1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 1 0 0 0 0 1 0 ]
+
+ First Polygonization PSO = [[[  S0 Line[P(201.0,179.0) to P(195.0,281.0)]  ,      S1 Line[P(195.0,281.0) to P(234.0,304.0)]  ,      S2 Line[P(234.0,304.0) to P(260.0,174.0)]  ,      S3 Line[P(260.0,174.0) to P(325.0,272.0)]  ,      S4 Line[P(325.0,272.0) to P(365.0,304.0)]  ,      S5 Line[P(365.0,304.0) to P(388.0,189.0)]  ,      S6 Line[P(388.0,189.0) to P(431.0,167.0)]  ,      S7 Line[P(431.0,167.0) to P(442.0,220.0)]  ]]]
+ \\ \hline
 

### 2009-08-07 trunk/draft/System.tex
80% ffinidhed of the svm

@@
 
-Let us consider  a two class classification task with datapoints $x_i$ where $i=1,\dotes,m$ having corresponding labels $y_i=\pm1$ where if  $y_i=1$ if $x_i$ belonged to first class and $y_i=-1$ if $x_i$ belonged to second class.
+Let us consider  a two class classification task with datapoints $x_i$ where $i=1,\cdots,m$ having corresponding labels $y_i=\pm1$ where if  $y_i=1$ if $x_i$ belonged to first class and $y_i=-1$ if $x_i$ belonged to second class.
 Each data  point $x_i$ is represented in $N$ dimensional input or attribule space.
@@
 A plane supports a class if all points in that class are on one side of that plane.
-For the points with the class label $+1$ we would like there exist $\omega$ and $b$ such that $\omega \cdot x_i \> b$ or $\omega \cdot x_i -b \> 0$  depending on the class label.
+For the points with the class label $+1$ we would like there exist $\omega$ and $b$ such that $\omega \cdot x_i\> b$ or $\omega \cdot x_i -b \> 0$  depending on the class label.
 Let us suppose that smallest value of $\mid \omega \cdot x_i -b\mid$ iss $\kappa$, then when $ \omega \cdot x_i -b \geq \kappa$.
@@
 \begin{equation}
- \omega \cdot x=b+1
+ \omega \cdot x=b+1 \\
   \omega \cdot x=b-1
@@
 \begin{equation}
- min  \frac{{\parallel \omega \parallel}^2}{2}
+ \min_{\omega,b}  \frac{{\parallel \omega \parallel}^2}{2}
 \end{equation}
 \begin{equation}	
- s.t.   \omega \cdot x_i \geq b+1    y_i \in Class 1 
- \omega \cdot x_i \leq b-1    y_i \in Class -1 
+ s.t. \ \ \   \omega \cdot x_i \geq b+1 \ \ \    y_i \in Class 1 \\ 
+ \ \ \ \ \ \   \omega \cdot x_i \leq b-1 \ \ \    y_i \in Class -1 
 \end{equation}
 
-The constrain can be simplified to $y_i(\omega \codt x_i-b)\geq 1$.
+The constrain can be simplified to $y_i(\omega \cdot x_i-b)\geq 1$.
 
-In the maximumm marg
+The margin depends on the Support vecotrs that are bumpt into the planes.
+This is a primal problem Quadratic Programming (QP) problem.
+The concept of duality can also upsed to wields the problem to a  Largrnigian dual QP problem in Equation \ref{eq:dual}.
+% maximumm margin between parallel 
+\begin{equation}
+\min_{\alpha} \ \ \   \frac{1}{2} \sum_{i=1}^{m}{\sum_{j=1}^{m}{y_iy_j\alpha_i\alpha_jx_ix_j}} - \sum_{i=1}^{m}{\alpha_i} \\
+s.t. \ \ \sum_{i=1}^{m}{y_i\alpha_i}=0   \\
+  \alpha_i \geq 0  \ \ \ i=1,\cdots , m 
+\label{eq:dual}
+\end{equation}
+Both the dual form and the primal form weilds to the same plane $\omega=\sum_{i=1}^{m}y_i\alpha_i x_i$ and threshold $b$ determined by the support vectors (for which $\alpha_i\geq 0$).
 
+\subsubsection{Theoretcical foundation}
+From a statistical learning theory perspective these QP formulation are well founded Rougly, Statistical Learning proves that bounds on the generalization error on future points not in the training set can be obtained.
+Theses bounds are a function fo misclassiffied error on the training data and tersm sthat measure the complexity or capacity fo classifcation function .
+For linear functions, maximizing the margin of seperation as discussed abouve reduces th efunction capacilty or complexity.
+Thus, by explicity maximizing the margin we are minimizing bound on the generalition error and can expect better generalization with high probability.
+The size of the margin is not directly dependent on the dimensionaliuty of the data.
+Thus, we can expect good performance event for very high dimensional data (i.e, with a very large number of attribute).
+In a sense, problems caused by oer fitting of high - dimensional data are greatly reduced.
+The problem is refered to the large volume of literature in, e.g., \cite{} for more technical discussion of the taht statistical theroy.
+
+We can gain inishgit into thesse result susing geomertic arguments. Classification function hat have capcity to fit the training data are more likely to overfit  resulting in poor generalization.
+
+\subsubsection{Linearly in seperable case}
+ So far we have assumed that the two datasets are linearly separable.
+Th this is not true, the startegy of constructing the plane that bisect the two closest points of the convex hulls will fail.
+As illustrated in Figure \ref{}, if the points are not linearly seperable the primal supporitng plane will fails.
+Since the QP task is not feasible for ethe linearly inseperable case , then we need to relax the constrains.
+Consider the linear inseperable problem in Figure\ref{}, Ideally we would like no points to be missclassified and no points to fall in the margin.
+But we must relax the constrain that insure that each point in on the approiate side of tis supporitng plane.
+Any point fallihng on the wrong side of tis supprting plane is considered to be an error.
+We want to maximize the margin and minimize the error.
+
+ This could be accomplished through minor change in the supporging plane QP problem.
+A nonegative slack or error variable $z_i$ is added to each constraint and then added as weighted penalty term in the objective as follows:
+ \begin{equation}
+\min_{w,b,z}  \quad \frac{{\parallel \omega \parallel}^2}{2}+C \sum_{i=1}^{l}z_i  \\
+s.t.  \quad y_i(\omega \cdot x_i -b)+z_i \geq 1  \\
+    z_i \geq 0  i=1,\cdots,m
+\label{eq:inseperable}
+\end{equation}
+
+Once again the primal relaxed supprting plane method is equivilent to the dual problem. The Largangian dual of the QP task is :
+
+\begin{equation}
+\min_{\alpha} \ \ \   \frac{1}{2} \sum_{i=1}^{m}{\sum_{j=1}^{m}{y_iy_j\alpha_i\alpha_jx_ix_j}} - \sum_{i=1}^{m}{\alpha_i} \\
+s.t. \ \ \sum_{i=1}^{m}{y_i\alpha_i}=0   \\
+ C \geq \alpha_i \geq 0  \ \ \ i=1,\cdots , m 
+\label{eq:dual}
+\end{equation}
+
+\subsubsection {Non linear function Using Kernels}
 %The foundations of Support Vector Machines (SVM) have
+\subsubsection{Summary of SVM}
+The resulting SVM method can be summarized as follows 
+\begin{enumerate}
+	\item Select paramter C representign the tradeoff between minimizign the training set error and maximizing the margin.
+Select the kernel function and any kernel paramters.
+For example for the radial basis function kernel, one must select width of the gaussian $\gamma$.
+	\item Solve dual QP Equation \ref{} or an alterantive SVM formulation using an approriate quadratic programming or linear programming alogrithm. 
+	\item Recover th primal threshold variable $b$ using the support vectors. 
+	\item Classify new points using $f(x)=sign\left(\sum{i}y_i\alpha_i K\left(x,x_i\right)-b\right)$
+\end{enumerate}
+
+The paramters in step 1. are slected using cross valiation if sufficent data are avaliable. %However, recent model  selectin strategies can
+
+\subsubsection{ SVM in Machine Learning }
+
+\susubsection { SVM vs. Other Learning Methods}
 

### 2009-08-09 draft/System.tex
change day 2

@@
 \begin{equation}	
- s.t. \ \ \   \omega \cdot x_i \geq b+1 \ \ \    y_i \in Class 1 \\ 
- \ \ \ \ \ \   \omega \cdot x_i \leq b-1 \ \ \    y_i \in Class -1 
+\begin{array}{cc}
+ s.t. \ \ \   \omega \cdot x_i \geq b+1 &    y_i \in Class 1 \\ 
+ \ \ \ \ \ \   \omega \cdot x_i \leq b-1 &    y_i \in Class -1 
+  \end{array} 
 \end{equation}
@@
 \begin{equation}
+\begin{array}{c}
 \min_{\alpha} \ \ \   \frac{1}{2} \sum_{i=1}^{m}{\sum_{j=1}^{m}{y_iy_j\alpha_i\alpha_jx_ix_j}} - \sum_{i=1}^{m}{\alpha_i} \\
@@
   \alpha_i \geq 0  \ \ \ i=1,\cdots , m 
+  \end{array} 
 \label{eq:dual}
@@
 
+\subsubsection{Theoretcical foundation}
+From a statistical learning theory perspective these QP formulation are well founded Rougly, Statistical Learning proves that bounds on the generalization error on future points not in the training set can be obtained.
+Theses bounds are a function fo misclassiffied error on the training data and tersm sthat measure the complexity or capacity fo classifcation function .
+For linear functions, maximizing the margin of seperation as discussed abouve reduces th efunction capacilty or complexity.
+Thus, by explicity maximizing the margin we are minimizing bound on the generalition error and can expect better generalization with high probability.
+The size of the margin is not directly dependent on the dimensionaliuty of the data.
+Thus, we can expect good performance event for very high dimensional data (i.e, with a very large number of attribute).
+In a sense, problems caused by oer fitting of high - dimensional data are greatly reduced.
+The problem is refered to the large volume of literature in, e.g., \cite{} for more technical discussion of the taht statistical theroy.
+
+We can gain inishgit into thesse result susing geomertic arguments. Classification function hat have capcity to fit the training data are more likely to overfit  resulting in poor generalization.
+
+\subsubsection{Linearly in seperable case}
+ So far we have assumed that the two datasets are linearly separable.
+Th this is not true, the startegy of constructing the plane that bisect the two closest points of the convex hulls will fail.
+As illustrated in Figure \ref{}, if the points are not linearly seperable the primal supporitng plane will fails.
+Since the QP task is not feasible for ethe linearly inseperable case , then we need to relax the constrains.
+Consider the linear inseperable problem in Figure\ref{}, Ideally we would like no points to be missclassified and no points to fall in the margin.
+But we must relax the constrain that insure that each point in on the approiate side of tis supporitng plane.
+Any point fallihng on the wrong side of tis supprting plane is considered to be an error.
+We want to maximize the margin and minimize the error.
+
+ This could be accomplished through minor change in the supporging plane QP problem.
+A nonegative slack or error variable $z_i$ is added to each constraint and then added as weighted penalty term in the objective as follows:
+ \begin{equation}
+ \begin{array}{c}
+\min_{w,b,z}  \quad \frac{{\parallel \omega \parallel}^2}{2}+C \sum_{i=1}^{l}z_i  \\
+s.t.  \quad y_i(\omega \cdot x_i -b)+z_i \geq 1  \\
+    z_i \geq 0  i=1,\cdots,m
+\end{array} 
+\label{eq:inseperable}
+\end{equation}
+
+Once again the primal relaxed supprting plane method is equivilent to the dual problem. The Largangian dual of the QP task is :
+
+\begin{equation}
+\begin{array}{c}
+\min_{\alpha} \ \ \   \frac{1}{2} \sum_{i=1}^{m}{\sum_{j=1}^{m}{y_iy_j\alpha_i\alpha_jx_ix_j}} - \sum_{i=1}^{m}{\alpha_i} \\
+s.t. \ \ \sum_{i=1}^{m}{y_i\alpha_i}=0   \\
+ C \geq \alpha_i \geq 0  \ \ \ i=1,\cdots , m 
+ \end{array} 
+\label{eq:dual}
+\end{equation}
+
+\subsubsection {Non linear function Using Kernels}
+If we consider a classification problem that cannot be seperated using a linear space (See Figure \ref{} ).
+A quadratic or  cubic function will work better in some casses.
+A classific method of converting a linear classification algoirhtm inot a nonolinear classification algoorithm is to simply add addional attribute to the data that are nonlinear functions of the original data.
+Existing linar classification algorithms can be then applied to the expanded dataset feature space producing non linear functions in the original input space.
+To construct a quadratic disceriminant in a two dimensional vector space with attributes $r$ and $s$, simply map the original two dimensional input space $\left[r,s\right]$ to the five dimensional features space $\left[r,s,rs,r^2,s^2\right]$ and construct a linear discriminiant in that space.
+specifically, define : $\vartheta(x):R^2  \rightarrow R^5$ then
+\begin{equation}
+\begin{array}{l}
+	x=[r,s] \\
+	\omega \cdot x=\omega_1 r+ \omega_2 s \\
+	\downarrow \\
+	\vartheta(x)=\left[r,s,rs,r^2,s^2\right] \\
+
+	 \omega \cdot \vartheta(x)= \omega_1 r+ \omega_2 s + \omega_3 rs+ \omega_4 r^2+\omega_5 r^2 
+\end{array}
+\label{eq:kernel}
+\end{equation}  
+The resulting classification function, 
+\begin{equation}
+\begin{array}{l}
+f(x)=sign\left( \omega \cdot\vartheta(x)-b \right) \\
+= sign\left( \omega_1 r+ \omega_2 s + \omega_3 rs+ \omega_4 r^2+\omega_5 r^2 -b \right) 
+\end{array}
+\label{eq:kern2}
+\end{equation}  
+is linear in the mapped five dimensional feature space but is quadratic in the two dimensional input space.
+
+For higher dimentional datasets, this nonlinear mapping method has two potential problems stemming from the fact taht dimensionality of the feature space explones exponentially.
+The first problem is that overfitting becomes a problem.
+SVM are largely immune to this porlblem since they rely on margin maximization.
+Provided an appropriate value of paaramter $C$ is chosen.
+The second concern is that it is not practical to actually compute $\vartheta(x)$.
+SVMs get around this issue through teh use of kernels.
+
+Exampine what happens when athe nonlinear mapping is introduced to equation \ref{eq:dual}. Let us define : $\vartheta(x):R^2\rightarrow R^5 \quad n^\'>> n$  we need to optimize
+
+\begin{equation}
+\begin{array}{c}
+\min_{\alpha} \ \ \   \frac{1}{2} \sum_{i=1}^{m}{\sum_{j=1}^{m}{y_iy_j\alpha_i\alpha_j\vartheta(x_i)\vartheta(x_j)}} - \sum_{i=1}^{m}{\alpha_i} \\
+s.t. \ \ \sum_{i=1}^{m}{y_i\alpha_i}=0   \\
+ C \geq \alpha_i \geq 0  \ \ \ i=1,\cdots , m 
+ \end{array} 
+\label{eq:dualkernel}
+\end{equation}
+Notice that the mapped data only occurs as inner product in the objective.
+Now we apply a little mathematically rigorous magie known as hilbern schmidt kernels, first applied to SVMs in \cite{}.
+By Mercer's theorem, we knoew that for ccertain mapping $\theta$ and any two points $u$ and $v$, the inner product of the the mapped points can be evalutated using the kernel function without ever explicity knowing ht emapping , e.g., $\theta(u)\theta(v)\equiv K(u,v)$.
+Some of the more popular known kernels are given below.
+New kernel are being developed to fit domian specific requriments.
+
+\begin{equation}
+\begin{array}{cc}
+\underline{\vartheta(u) & K(u,v)} \\ 
+Degree d polynomial  & \left(u\cdot v +1\right)^d \\
+Radial Basis Function & \exp \left( - \frac{\left\|u-v\right\|^2}{2\sigma}\right) \\
+Two layer Neural Network & sigmoid \left(\eta\left(u\cdot v +1\right)+c\right) 
+  \end{array} 
+\label{eq:diffkernel}
+\end{equation}
+
 %The foundations of Support Vector Machines (SVM) have
+\subsubsection{Summary of SVM}
+The resulting SVM method can be summarized as follows 
+\begin{enumerate}
+	\item Select paramter C representign the tradeoff between minimizign the training set error and maximizing the margin.
+Select the kernel function and any kernel paramters.
+For example for the radial basis function kernel, one must select width of the gaussian $\gamma$.
+	\item Solve dual QP Equation \ref{} or an alterantive SVM formulation using an approriate quadratic programming or linear programming alogrithm. 
+	\item Recover th primal threshold variable $b$ using the support vectors. 
+	\item Classify new points using $f(x)=sign\left(\sum{i}y_i\alpha_i K\left(x,x_i\right)-b\right)$
+\end{enumerate}
+
+The paramters in step 1. are selected using cross valiation if sufficent data are avaliable. %However, recent model  selectin strategies can
+
+\subsubsection{ SVM in Machine Learning }
+
+\subsubsection { SVM vs. Other Learning Methods}
 

### 2009-08-09 draft/System.tex

@@
 We used the library implementation of LibSVM \cite{libsvm} and its implementation of the one-vs-one strategy.
+Section \ref{Sec:SVMdetail} discribe in details the classifier, how classes (classification) are determined and the reasons behind using this specific classification methods
 
 \subsection{Support Vector Machines}
-
+\label{Sec:SVMdetail}
 The foundation of Support Vector Machines (SVM) were developed by Vapnik \cite{} in 1995.
@@
 
-Exampine what happens when athe nonlinear mapping is introduced to equation \ref{eq:dual}. Let us define : $\vartheta(x):R^2\rightarrow R^5 \quad n^\'>> n$  we need to optimize
+Exampine what happens when athe nonlinear mapping is introduced to equation \ref{eq:dual}. Let us define : $\vartheta(x):R^2\rightarrow R^5 \quad n>> n$  we need to optimize
 
@@
 New kernel are being developed to fit domian specific requriments.
-
-\begin{equation}
-\begin{array}{cc}
-\underline{\vartheta(u) & K(u,v)} \\ 
-Degree d polynomial  & \left(u\cdot v +1\right)^d \\
-Radial Basis Function & \exp \left( - \frac{\left\|u-v\right\|^2}{2\sigma}\right) \\
-Two layer Neural Network & sigmoid \left(\eta\left(u\cdot v +1\right)+c\right) 
-  \end{array} 
-\label{eq:diffkernel}
-\end{equation}
+\begin{center}
+	\begin{tabular}{cc}
+	$\vartheta(u)$  & $K(u,v)$ \\ \hline 
+	Degree $d$ polynomial & $\left(u\cdot v +1\right)^d$ \\ 
+	Radial Basis Function & $\exp \left( - \frac{\left\|u-v\right\|^2}{2\sigma}\right) $ \\ 
+		Two layer Neural Network& $ sigmoid \left(\eta\left(u\cdot v +1\right)+c\right)$ \\
+
+	\end{tabular}
+\end{center}
+
+%\begin{equation}
+%\begin{array}{cc}
+%\underline{&}\\  
+% & \left(u\cdot v +1\right)^d \\
+%Radial Basis Function & \exp \left( - \frac{\left\|u-v\right\|^2}{2\sigma}\right) \\
+%Two layer Neural Network & sigmoid \left(\eta\left(u\cdot v +1\right)+c\right) 
+% \end{array} 
+%\label{eq:diffkernel}
+%\end{equation}
 

### 2009-08-09 draft/System.tex

@@
 We used the library implementation of LibSVM \cite{libsvm} and its implementation of the one-vs-one strategy.
-Section \ref{Sec:SVMdetail} discribe in details the classifier, how classes (classification) are determined and the reasons behind using this specific classification methods
+Section \ref{Sec:SVMdetail} discribe in details the classifier, how classes (classification) are determined and the reasons behind using this specific classification methods.
 
@@
 
-\subsubsection{Linearly in seperable case}
+\subsubsection{Linearly Inseperable Case}
  So far we have assumed that the two datasets are linearly separable.
@@
 %The foundations of Support Vector Machines (SVM) have
-\subsubsection{Summary of SVM}
+\subsubsection{Classification}
 The resulting SVM method can be summarized as follows 
@@
 
-\subsubsection { SVM vs. Other Learning Methods}
+\subsubsection {SVM vs. Other Learning Methods}
+
+\subsubsection {Multi Class Classifiers}
+\subsubsection {SVM Implementations}
+\subsubsection {Summary Of SVM}
 
@@
 Figure \ref{} show the two stroke and how the user draw them.
-The next steps demonstrate how ouput of the system after each step
+The next steps demonstrate how ouput of the system after each step.
+The first section \ref{sec:stepseg} discribe how the segmentation is done step by step.
+Section \ref{sec:steprec} shows the details of feature calculation and the classification process.
+
+\subsection{Step by Step Segmentation}
+\label{sec:stepseg}
+In this step the user draw one or more stroke to draw a symbol.
+Figure \ref{} illustrate the sequence and direction of user draw.
+In this section, each stroke user draw will be labeled as it illustrated in Figure \ref{}.
+Table \ref{} shows the detailed output of the program at each step.
+The algorithm that compute the segmentation is demonstrated as follows:
+\begin{enumerate}
+	\item Extracting Stroke Points 
+	\item Extracting $P_dp$
+\begin{verbatim*}
+
+\end{verbatim*}	
+\item Ellipse Fitting: which is an attempt to fit as an ellipse
+\begin{verbatim*}
+
+\end{verbatim*}	
+\item Segmentation ALS1
+\begin{verbatim*}
+
+\end{verbatim*}	
+\item Segmentation ALS2
+\begin{verbatim*}
+
+\end{verbatim*}
+
+\end{enumerate}
 
@@
 \begin{longtable}{|p{2cm}|p{2cm}|p{2cm}|p{2cm}|p{13cm}|}
-	\caption{StepsStroke1}
+	\caption{ Detailed Output of System in Each Step of Segmentation}
 	\label{tab:StepsStroke1} \\
@@
 \end{landscape}
-
+\subsection{Step by Step Recognition}
+\label{sec:steprec}
 %\begin{enumerate}

### 2009-08-10 trunk/draft/System.tex

@@
 We used the library implementation of LibSVM \cite{libsvm} and its implementation of the one-vs-one strategy.
+Section \ref{Sec:SVMdetail} discribe in details the classifier, how classes (classification) are determined and the reasons behind using this specific classification methods.
 
 \subsection{Support Vector Machines}
-
+\label{Sec:SVMdetail}
 The foundation of Support Vector Machines (SVM) were developed by Vapnik \cite{} in 1995.
@@
  \begin{equation}
+ \begin{array}{c}
 \min_{w,b,z}  \quad \frac{{\parallel \omega \parallel}^2}{2}+C \sum_{i=1}^{l}z_i  \\
@@
     z_i \geq 0  i=1,\cdots,m
+\end{array} 
 \label{eq:inseperable}
@@
  C \geq \alpha_i \geq 0  \ \ \ i=1,\cdots , m 
+ \end{array} 
 \label{eq:dual}
@@
 \subsubsection {Non linear function Using Kernels}
+If we consider a classification problem that cannot be seperated using a linear space (See Figure \ref{} ).
+A quadratic or  cubic function will work better in some casses.
+A classific method of converting a linear classification algoirhtm inot a nonolinear classification algoorithm is to simply add addional attribute to the data that are nonlinear functions of the original data.
+Existing linar classification algorithms can be then applied to the expanded dataset feature space producing non linear functions in the original input space.
+To construct a quadratic disceriminant in a two dimensional vector space with attributes $r$ and $s$, simply map the original two dimensional input space $\left[r,s\right]$ to the five dimensional features space $\left[r,s,rs,r^2,s^2\right]$ and construct a linear discriminiant in that space.
+specifically, define : $\vartheta(x):R^2  \rightarrow R^5$ then
+\begin{equation}
+\begin{array}{l}
+	x=[r,s] \\
+	\omega \cdot x=\omega_1 r+ \omega_2 s \\
+	\downarrow \\
+	\vartheta(x)=\left[r,s,rs,r^2,s^2\right] \\
+
+	 \omega \cdot \vartheta(x)= \omega_1 r+ \omega_2 s + \omega_3 rs+ \omega_4 r^2+\omega_5 r^2 
+\end{array}
+\label{eq:kernel}
+\end{equation}  
+The resulting classification function, 
+\begin{equation}
+\begin{array}{l}
+f(x)=sign\left( \omega \cdot\vartheta(x)-b \right) \\
+= sign\left( \omega_1 r+ \omega_2 s + \omega_3 rs+ \omega_4 r^2+\omega_5 r^2 -b \right) 
+\end{array}
+\label{eq:kern2}
+\end{equation}  
+is linear in the mapped five dimensional feature space but is quadratic in the two dimensional input space.
+
+For higher dimentional datasets, this nonlinear mapping method has two potential problems stemming from the fact taht dimensionality of the feature space explones exponentially.
+The first problem is that overfitting becomes a problem.
+SVM are largely immune to this porlblem since they rely on margin maximization.
+Provided an appropriate value of paaramter $C$ is chosen.
+The second concern is that it is not practical to actually compute $\vartheta(x)$.
+SVMs get around this issue through teh use of kernels.
+
+Exampine what happens when athe nonlinear mapping is introduced to equation \ref{eq:dual}. Let us define : $\vartheta(x):R^2\rightarrow R^5 \quad n>> n$  we need to optimize
+
+\begin{equation}
+\begin{array}{c}
+\min_{\alpha} \ \ \   \frac{1}{2} \sum_{i=1}^{m}{\sum_{j=1}^{m}{y_iy_j\alpha_i\alpha_j\vartheta(x_i)\vartheta(x_j)}} - \sum_{i=1}^{m}{\alpha_i} \\
+s.t. \ \ \sum_{i=1}^{m}{y_i\alpha_i}=0   \\
+ C \geq \alpha_i \geq 0  \ \ \ i=1,\cdots , m 
+ \end{array} 
+\label{eq:dualkernel}
+\end{equation}
+Notice that the mapped data only occurs as inner product in the objective.
+Now we apply a little mathematically rigorous magie known as hilbern schmidt kernels, first applied to SVMs in \cite{}.
+By Mercer's theorem, we knoew that for ccertain mapping $\theta$ and any two points $u$ and $v$, the inner product of the the mapped points can be evalutated using the kernel function without ever explicity knowing ht emapping , e.g., $\theta(u)\theta(v)\equiv K(u,v)$.
+Some of the more popular known kernels are given below.
+New kernel are being developed to fit domian specific requriments.
+\begin{center}
+	\begin{tabular}{cc}
+	$\vartheta(u)$  & $K(u,v)$ \\ \hline 
+	Degree $d$ polynomial & $\left(u\cdot v +1\right)^d$ \\ 
+	Radial Basis Function & $\exp \left( - \frac{\left\|u-v\right\|^2}{2\sigma}\right) $ \\ 
+		Two layer Neural Network& $ sigmoid \left(\eta\left(u\cdot v +1\right)+c\right)$ \\
+
+	\end{tabular}
+\end{center}
+To change from a linear to nonlinear classifier, one must only substitute a kernel evaluation in the objective insted of the original dot product.
+Thus by changing kernels we can get different hightly nonlinear classifiers.
+No algorithmic changes are required from linear clase other than substitue of a kernel evaluation for simple dot product.
+All the benefits fo the original linear SVM methods are maintained.
+We can train a highly nonlinear classification function such as a ploynomial or a raadial basis function machine, or a sigmoidal neural network using robust, efficient algorithsm that have no problem with local minima.
+By using kernel substitution a linear algorithm (only capable of handling seperable data) can be turned into a general nonlinear algorithm.
+%\begin{equation}
+%\begin{array}{cc}
+%\underline{&}\\  
+% & \left(u\cdot v +1\right)^d \\
+%Radial Basis Function & \exp \left( - \frac{\left\|u-v\right\|^2}{2\sigma}\right) \\
+%Two layer Neural Network & sigmoid \left(\eta\left(u\cdot v +1\right)+c\right) 
+% \end{array} 
+%\label{eq:diffkernel}
+%\end{equation}
+
 %The foundations of Support Vector Machines (SVM) have
-\subsubsection{Summary of SVM}
+\subsubsection{Classification}
 The resulting SVM method can be summarized as follows 
@@
 
-The paramters in step 1. are slected using cross valiation if sufficent data are avaliable. %However, recent model  selectin strategies can
+The paramters in step 1. are selected using cross valiation if sufficent data are avaliable. %However, recent model  selectin strategies can
 
 \subsubsection{ SVM in Machine Learning }
-
-\susubsection { SVM vs. Other Learning Methods}
+SVM has been successfully applied to a number of applications ranging from particle identification \cite{}. Face detection  \cite{} and text categorization \cite{} to engine knock detection 
+\subsubsection {SVM vs. Other Learning Methods}
+
+\subsubsection {Multi Class Classifiers}
+\subsubsection {SVM Implementations}
+Typically an SVM approach requires the solution of a QP or LP problem.
+LP and QP type problems have been extensively studied inthe field of mathematical programming.
+These can be divided into three categories techniques in which kernel compoinents are evaluated and discarded during learning.
+Decomposition methods in which an evolving subset of data is used and new optimization approaches that specifically exploit the structure of the SVM problem.
+
+For the first category the most obvioius  approach is to sequentially update the $\alpha_i$  and this is the approach used by the kernel  Adtron (KA) algorithm \cite{}.
+
+Chunking and decomposition methods optimize the SVM with respect to subsets.
+Rather than sequentially updateing the $\alpha_i$ the alternative is to upodate the $\alpha_i$ in parallel but using only a subset or wokring set of data at each stage.
+There are many successful code based on these decomposition strategies.
+SVM codes available online such as SVMTorch \cite{} and SVMLight \cite{} use these working set strategies.
+
+The third approach is to directly attack the SVM  problem from and optimization prespective and create algorithms that explicity exploits the structure of teh problem.
+Keerthi et al.
+\cite{} proposed a very effective algorithm based on the dual geometry of finding the two closest points in the convex hulls.
+
+\subsubsection {Summary Of SVM}
 

### 2009-08-10 trunk/draft/System.tex

@@
 \subsubsection{ SVM in Machine Learning }
-SVM has been successfully applied to a number of applications ranging from particle identification \cite{}. Face detection  \cite{} and text categorization \cite{} to engine knock detection 
+SVM has been successfully applied to a number of applications ranging from particle identification \cite{}.
+Face detection  \cite{} and text categorization \cite{} to engine knock detection.
+SVMs are very suited to the classification task that commonly arise in machine vision.
+As an example we consider and application involving face identification\cite{}.
+Three different SVM classiifer were used on training set of 200 images (5 image per person).
+The error rate was less than 5\%.
+An RBF network and SVM classifiers were compared on the UPUS handwritten digit recognition database\cite{}.
+SVM outperformed and RBF network on all digits.
+
 \subsubsection {SVM vs. Other Learning Methods}
 
+SVM have been applied ot the a largest MNIST a Handwritten digit recognition dataset of 60000 training set and 10000 test set\cite{}.
+Sherif and Ezzat has performed a comparision between SVM and several other techniques on the MNIST dataset.
+They also compared the same classfication methods on arabic handwritten digits (MADBASE) \cite{}.
+Their results shows that SVM classifier outperform all other techniques on both MNIST and MADBase datasets and achieved recognition rate of over 98\%.
+Similar results was reported by \cite{} on the MNIST dataset.
+Ther are many other successful application SVM.
+They have proven to be robust to noise and perform well on many tasks.
+
+This seperior performance because they may elimnates many of problem experienced with other inference methodoliges with other inference methodollogies like neural network and decision trees.
+The first problem that is eliminated was the problems with local minima.
+SVM can construct highly nonlinear classification and regression function without woring about getting stuck at local minima.
+Another problem SVM eliminated is the number of parameter that are picked ot generate the model.
+For example, SVM classifier with RBF  kernel will only need 2 paramters to set the model the $C$ cost paramter and $\sigma$ paramter of the RBF kernel.
+The last reason is the final result are sable, reproducible and largely independent of the specific algorithm used to  optimize the SVM model.
+If two users apply the same SVM model with the same parameters to the same data, they will et the same solution modeule numeric issues.
+Compare this with neural networks where the results are dependent on the particular algorithms and starting point used.
+
+SVM classifiers are also simple to use. One need not be a SVM expert to successfully apply existing SVM software to new problems.
+
+%different classifier on MNIST and a
+
 \subsubsection {Multi Class Classifiers}
+
 \subsubsection {SVM Implementations}

### 2009-08-10 trunk/draft/System.tex
after 11 august change

@@
 \label{Sec:SVMdetail}
-The foundation of Support Vector Machines (SVM) were developed by Vapnik \cite{} in 1995.
+The foundation of Support Vector Machines (SVM) were developed by Vapnik \cite{svmintroduce} in 1995.
 It is gainig increase popularity due to the high performance and efficient implementation.
@@
 The generated plane is used as a good classifier to the unseen exmaples.
-However, there can be more than one plane that can separates the data therefore SVM attempts to find the plane that maximize the margin (distance between classifier and nearest sample) between the two classes \ref{}.
+However, there can be more than one plane that can separates the data therefore SVM attempts to find the plane that maximize the margin (distance between classifier and nearest sample) between the two classes \ref{fig:svm1}.
 To calcuate the margin, two different hyperplanes are constructed, one plane on each side of the data, these two planes are then moved to generenate the maximum margin possilbe.
+\begin{figure}
+	\centering
+		\includegraphics{afterDefense/svm1.jpg.eps}
+	\caption{Different Classifier planes }
+	\label{fig:svm1}
+\end{figure}
 
@@
 there are infinitely many possible seperating planes that correctly classify the training data.
-Figure\ref{} illustrates two different seperating planes.
-It is intuitively noted from the figure that the () line is more likely to generalize more data and thus introduce less classification error.
+Figure\ref{fig:svm1} illustrates two different seperating planes.
+It is intuitively noted from the figure that the solid line is more likely to generalize more data and thus introduce less classification error.
 This can be geometrically characterized as plane furthest from both classes.
@@
 For the points with class label $-1$ we similary require $\omega \cdot x_i -b \leq -1 $.
-To find the plane furthest from both setss, we can simply maximize the distance or margin between the support planes for each class as illustrated in Figure \ref{}.
+To find the plane furthest from both setss, we can simply maximize the distance or margin between the support planes for each class as illustrated in Figure \ref{fig:svm2}.
 The support planes are pussshed apart untill they bump inot small number of data points ( the support vecotrs ) for each class.
-The support vectors in Figure \ref{} are outlined in (),
+The support vectors in Figure \ref{fig:svm2} are filled datapoints ,
+\begin{figure}
+	\centering
+		\includegraphics{afterDefense/svm2.jpg.eps}
+	\caption{Margin of Support planes }
+	\label{fig:svm2}
+\end{figure}
 
@@
 Th this is not true, the startegy of constructing the plane that bisect the two closest points of the convex hulls will fail.
-As illustrated in Figure \ref{}, if the points are not linearly seperable the primal supporitng plane will fails.
+As illustrated in Figure \ref{fig:svm5}, if the points are not linearly seperable the primal supporitng plane will fails.
 Since the QP task is not feasible for ethe linearly inseperable case , then we need to relax the constrains.
-Consider the linear inseperable problem in Figure\ref{}, Ideally we would like no points to be missclassified and no points to fall in the margin.
+Consider the linear inseperable problem in Figure\ref{fig:svm5}, Ideally we would like no points to be missclassified and no points to fall in the margin.
 But we must relax the constrain that insure that each point in on the approiate side of tis supporitng plane.
@@
 We want to maximize the margin and minimize the error.
+
+\begin{figure}
+	\centering
+		\includegraphics{afterDefense/svm5.jpg.eps}
+	\caption{Linearly Inseperable}
+	\label{fig:svm5}
+\end{figure}
 
@@
 \subsubsection {Non linear function Using Kernels}
-If we consider a classification problem that cannot be seperated using a linear space (See Figure \ref{} ).
+If we consider a classification problem that cannot be seperated using a linear space (See Figure \ref{fig:svm6} ).
 A quadratic or  cubic function will work better in some casses.
@@
 \end{equation}
+
 The resulting classification function, 
@@
 is linear in the mapped five dimensional feature space but is quadratic in the two dimensional input space.
+
+\begin{figure}
+	\centering
+		\includegraphics{afterDefense/svm6.jpg.eps}
+	\caption{non Lineary problem }
+	\label{fig:svm6}
+\end{figure}
 
@@
 
-Exampine what happens when athe nonlinear mapping is introduced to equation \ref{eq:dual}. Let us define : $\vartheta(x):R^2\rightarrow R^5 \quad n>> n$  we need to optimize
+Examine what happens when athe nonlinear mapping is introduced to equation \ref{eq:dual}. Let us define : $\vartheta(x):R^2\rightarrow R^5 \quad n>> n$  we need to optimize
 

### 2009-08-15 draft/System.tex
after adding most cites

@@
 
+\subsection {Trainig the SVM}
+\subsection {Testing}
+
 \subsection{Support Vector Machines}
 \label{Sec:SVMdetail}
-The foundation of Support Vector Machines (SVM) were developed by Vapnik \cite{} in 1995.
+The foundation of Support Vector Machines (SVM) were developed by Vapnik \cite{svmintroduce} in 1995.
 It is gainig increase popularity due to the high performance and efficient implementation.
@@
 Each data  point $x_i$ is represented in $N$ dimensional input or attribule space.
-Let the classification plane be represented as function .
+Let the classification plane be represented as function.
 \begin{equation}
@@
 Let us suppose that smallest value of $\mid \omega \cdot x_i -b\mid$ iss $\kappa$, then when $ \omega \cdot x_i -b \geq \kappa$.
-The argument insidee the decision function is invariant under positive rescalling so we will implicity fix a scale by requiring $\omega \cdot x_i -b \geq 1 $.
+The argument inside the decision function is invariant under positive rescalling so we will implicity fix a scale by requiring $\omega \cdot x_i -b \geq 1 $.
 For the points with class label $-1$ we similary require $\omega \cdot x_i -b \leq -1 $.
-To find the plane furthest from both setss, we can simply maximize the distance or margin between the support planes for each class as illustrated in Figure \ref{}.
+To find the plane furthest from both setss, we can simply maximize the distance or margin between the support planes for each class as illustrated in Figure \ref{fig:svm2}.
 The support planes are pussshed apart untill they bump inot small number of data points ( the support vecotrs ) for each class.
-The support vectors in Figure \ref{} are outlined in (),
+The support vectors in Figure \ref{fig:svm2} are filled datapoints ,
+\begin{figure}
+	\centering
+		\includegraphics{afterDefense/svm2.jpg.eps}
+	\caption{Margin of Support planes }
+	\label{fig:svm2}
+\end{figure}
 
@@
 In a sense, problems caused by oer fitting of high - dimensional data are greatly reduced.
-The problem is refered to the large volume of literature in, e.g., \cite{} for more technical discussion of the taht statistical theroy.
+The problem is refered to the large volume of literature in, e.g., \cite{bookSVMoverfit11,statisticalSVM1} for more technical discussion of the taht statistical theroy.
 
@@
  So far we have assumed that the two datasets are linearly separable.
-Th this is not true, the startegy of constructing the plane that bisect the two closest points of the convex hulls will fail.
-As illustrated in Figure \ref{}, if the points are not linearly seperable the primal supporitng plane will fails.
+This is not true, the startegy of constructing the plane that bisect the two closest points of the convex hulls will fail.
+As illustrated in Figure \ref{fig:svm5}, if the points are not linearly seperable the primal supporitng plane will fails.
 Since the QP task is not feasible for ethe linearly inseperable case , then we need to relax the constrains.
-Consider the linear inseperable problem in Figure\ref{}, Ideally we would like no points to be missclassified and no points to fall in the margin.
+Consider the linear inseperable problem in Figure\ref{fig:svm5}, Ideally we would like no points to be missclassified and no points to fall in the margin.
 But we must relax the constrain that insure that each point in on the approiate side of tis supporitng plane.
@@
 Notice that the mapped data only occurs as inner product in the objective.
-Now we apply a little mathematically rigorous magie known as hilbern schmidt kernels, first applied to SVMs in \cite{}.
+Now we apply a little mathematically rigorous magie known as hilbern schmidt kernels, first applied to SVMs in \cite{svmintroduce}.
 By Mercer's theorem, we knoew that for ccertain mapping $\theta$ and any two points $u$ and $v$, the inner product of the the mapped points can be evalutated using the kernel function without ever explicity knowing ht emapping , e.g., $\theta(u)\theta(v)\equiv K(u,v)$.
@@
 \end{center}
-
+To change from a linear to nonlinear classifier, one must only substitute a kernel evaluation in the objective insted of the original dot product.
+Thus by changing kernels we can get different hightly nonlinear classifiers.
+No algorithmic changes are required from linear clase other than substitue of a kernel evaluation for simple dot product.
+All the benefits fo the original linear SVM methods are maintained.
+We can train a highly nonlinear classification function such as a ploynomial or a raadial basis function machine, or a sigmoidal neural network using robust, efficient algorithsm that have no problem with local minima.
+By using kernel substitution a linear algorithm (only capable of handling seperable data) can be turned into a general nonlinear algorithm.
 %\begin{equation}
@@
 \subsubsection{ SVM in Machine Learning }
+SVM has been successfully applied to a number of applications ranging from particle identification \cite{ParticleSVM7}.
+Face detection  \cite{faceSVM8,faceSVM7} and text categorization \cite{TextSVM7} to engine knock detection.
+SVMs are very suited to the classification task that commonly arise in machine vision.
+As an example we consider and application involving face identification\cite{faceSVM8}.
+Three different SVM classiifer were used on training set of 200 images (5 image per person).
+The error rate was less than 5\%.
+An RBF network and SVM classifiers were compared on the UPUS handwritten digit recognition database\cite{ORLDataset}.
+SVM outperformed and RBF network on all digits.
 
@@
 
+SVM have been applied ot the a largest MNIST a Handwritten digit recognition dataset of 60000 training set and 10000 test set\cite{IjdarSherifPaper}.
+Sherif and Ezzat has performed a comparision between SVM and several other techniques on the MNIST dataset.
+They also compared the same classfication methods on arabic handwritten digits (MADBASE) \cite{ADBase9,IjdarSherifPaper}.
+Their results shows that SVM classifier outperform all other techniques on both MNIST and MADBase datasets and achieved recognition rate of over 98\%.
+Similar results was reported by \cite{empiricalcomp11,SVMInvariantComDecoste02} on the MNIST dataset.
+Ther are many other successful application SVM.
+They have proven to be robust to noise and perform well on many tasks \cite{empiricalcomp11,Schölkopf97comparingsupport}.
+
+This seperior performance because they may elimnates many of problem experienced with other inference methodoliges with other inference methodollogies like neural network and decision trees.
+The first problem that is eliminated was the problems with local minima.
+SVM can construct highly nonlinear classification and regression function without woring about getting stuck at local minima.
+Another problem SVM eliminated is the number of parameter that are picked ot generate the model.
+For example, SVM classifier with RBF  kernel will only need 2 paramters to set the model the $C$ cost paramter and $\sigma$ paramter of the RBF kernel.
+The last reason is the final result are sable, reproducible and largely independent of the specific algorithm used to  optimize the SVM model.
+If two users apply the same SVM model with the same parameters to the same data, they will et the same solution modeule numeric issues.
+Compare this with neural networks where the results are dependent on the particular algorithms and starting point used.
+
+SVM classifiers are also simple to use. One need not be a SVM expert to successfully apply existing SVM software to new problems.
+
+%different classifier on MNIST and a
+
 \subsubsection {Multi Class Classifiers}
+
 \subsubsection {SVM Implementations}
+Typically an SVM approach requires the solution of a QP or LP problem.
+LP and QP type problems have been extensively studied inthe field of mathematical programming.
+These can be divided into three categories techniques in which kernel compoinents are evaluated and discarded during learning.
+Decomposition methods in which an evolving subset of data is used and new optimization approaches that specifically exploit the structure of the SVM problem.
+
+For the first category the most obvioius  approach is to sequentially update the $\alpha_i$  and this is the approach used by the kernel  Adtron (KA) algorithm \cite{}.
+
+Chunking and decomposition methods optimize the SVM with respect to subsets.
+Rather than sequentially updateing the $\alpha_i$ the alternative is to upodate the $\alpha_i$ in parallel but using only a subset or wokring set of data at each stage.
+There are many successful code based on these decomposition strategies.
+SVM codes available online such as SVMTorch \cite{} and SVMLight \cite{} use these working set strategies.
+
+The third approach is to directly attack the SVM  problem from and optimization prespective and create algorithms that explicity exploits the structure of teh problem.
+Keerthi et al.
+\cite{} proposed a very effective algorithm based on the dual geometry of finding the two closest points in the convex hulls.
+
 \subsubsection {Summary Of SVM}

### 2009-08-15 trunk/draft/System.tex
after finished cite of svm

@@
 
+\subsection {Trainig the SVM}
+\subsection {Testing}
+
 \subsection{Support Vector Machines}
@@
 Let us suppose that smallest value of $\mid \omega \cdot x_i -b\mid$ iss $\kappa$, then when $ \omega \cdot x_i -b \geq \kappa$.
-The argument insidee the decision function is invariant under positive rescalling so we will implicity fix a scale by requiring $\omega \cdot x_i -b \geq 1 $.
+The argument inside the decision function is invariant under positive rescalling so we will implicity fix a scale by requiring $\omega \cdot x_i -b \geq 1 $.
 For the points with class label $-1$ we similary require $\omega \cdot x_i -b \leq -1 $.
@@
  So far we have assumed that the two datasets are linearly separable.
-Th this is not true, the startegy of constructing the plane that bisect the two closest points of the convex hulls will fail.
+This is not true, the startegy of constructing the plane that bisect the two closest points of the convex hulls will fail.
 As illustrated in Figure \ref{fig:svm5}, if the points are not linearly seperable the primal supporitng plane will fails.
@@
 \subsubsection{ SVM in Machine Learning }
-SVM has been successfully applied to a number of applications ranging from particle identification \cite{}.
-Face detection  \cite{} and text categorization \cite{} to engine knock detection.
+SVM has been successfully applied to a number of applications ranging from particle identification \cite{ParticleSVM7}.
+Face detection  \cite{faceSVM8,faceSVM7} and text categorization \cite{TextSVM7} to engine knock detection.
 SVMs are very suited to the classification task that commonly arise in machine vision.
-As an example we consider and application involving face identification\cite{}.
+As an example we consider and application involving face identification\cite{faceSVM8}.
 Three different SVM classiifer were used on training set of 200 images (5 image per person).
 The error rate was less than 5\%.
-An RBF network and SVM classifiers were compared on the UPUS handwritten digit recognition database\cite{}.
+An RBF network and SVM classifiers were compared on the UPUS handwritten digit recognition database\cite{ORLDataset}.
 SVM outperformed and RBF network on all digits.
@@
 
-SVM have been applied ot the a largest MNIST a Handwritten digit recognition dataset of 60000 training set and 10000 test set\cite{}.
+SVM have been applied ot the a largest MNIST a Handwritten digit recognition dataset of 60000 training set and 10000 test set\cite{IjdarSherifPaper}.
 Sherif and Ezzat has performed a comparision between SVM and several other techniques on the MNIST dataset.
-They also compared the same classfication methods on arabic handwritten digits (MADBASE) \cite{}.
+They also compared the same classfication methods on arabic handwritten digits (MADBASE) \cite{ADBase9,IjdarSherifPaper}.
 Their results shows that SVM classifier outperform all other techniques on both MNIST and MADBase datasets and achieved recognition rate of over 98\%.
-Similar results was reported by \cite{} on the MNIST dataset.
+Similar results was reported by \cite{empiricalcomp11,SVMInvariantComDecoste02} on the MNIST dataset.
 Ther are many other successful application SVM.
-They have proven to be robust to noise and perform well on many tasks.
+They have proven to be robust to noise and perform well on many tasks \cite{empiricalcomp11,Scholkopf97comparingsupport}.
 
@@
 
-For the first category the most obvioius  approach is to sequentially update the $\alpha_i$  and this is the approach used by the kernel  Adtron (KA) algorithm \cite{}.
+For the first category the most obvioius  approach is to sequentially update the $\alpha_i$  and this is the approach used by the kernel  Adtron (KA) algorithm \cite{kernelSVM4}.
 
@@
 There are many successful code based on these decomposition strategies.
-SVM codes available online such as SVMTorch \cite{} and SVMLight \cite{} use these working set strategies.
+SVM codes available online such as SVMTorch \cite{svmTorch} and SVMLight \cite{SVMlight} use these working set strategies.
 
@@
 Keerthi et al.
-\cite{} proposed a very effective algorithm based on the dual geometry of finding the two closest points in the convex hulls.
+\cite{fastSVM6} proposed a very effective algorithm based on the dual geometry of finding the two closest points in the convex hulls.
 

### 2009-08-15 draft/System.tex
the commit of some svm and cite

@@
 
- The vector $\omega$ determines th orientation of a seperation plane.
-The scalar $b$ determines the offeset of the plane from the orgin.
-Lets assume that the datatpointa re linearly seperable, i.e.
+ The orientation of the speration plane is deteremined using the vector $\omega$ and the offest of the plane from the orgin is determined using the scalar $b$.
+Lets assume that the datatpoint are linearly seperable, i.e.
 there are infinitely many possible seperating planes that correctly classify the training data.
@@
 Ther are many other successful application SVM.
-They have proven to be robust to noise and perform well on many tasks \cite{empiricalcomp11,Schölkopf97comparingsupport}.
+They have proven to be robust to noise and perform well on many tasks \cite{empiricalcomp11,Scholkopf97comparingsupport}.
 

### 2009-08-15 draft/System.tex
after some rephrashing

@@
  The orientation of the speration plane is deteremined using the vector $\omega$ and the offest of the plane from the orgin is determined using the scalar $b$.
-Lets assume that the datatpoint are linearly seperable, i.e.
-there are infinitely many possible seperating planes that correctly classify the training data.
-Figure\ref{fig:svm1} illustrates two different seperating planes.
-It is intuitively noted from the figure that the solid line is more likely to generalize more data and thus introduce less classification error.
-This can be geometrically characterized as plane furthest from both classes.
+Lets make an assumption that the datatpoint are linearly seperable, i.e.
+there are infinite possibilities to find a plane that seperates and correctly classify the training data.
+This is idea is illustrated in Figure\ref{fig:svm1} where two different seperating planes can correctly classify the data.
+It is intuitively noted from the figure that the solid line is more likely to generalize more data.
+This means less classification error will be introduced.
+The geometrical characterization of this plane means the plane furthest from both classes.
 % Which one is preferable ?
@@
 Geometrically we can characterize the solid plane as being furthuret from both classes.
-
-The SVM problem is how to construct the plane that is furthuest from both classes.
-One approcah is to maximize the margin between two parallel planes supporting.
-A plane supports a class if all points in that class are on one side of that plane.
-For the points with the class label $+1$ we would like there exist $\omega$ and $b$ such that $\omega \cdot x_i\> b$ or $\omega \cdot x_i -b \> 0$  depending on the class label.
-Let us suppose that smallest value of $\mid \omega \cdot x_i -b\mid$ iss $\kappa$, then when $ \omega \cdot x_i -b \geq \kappa$.
+plane supports a class if all points in that class are on one side of that plane.
+
+The main problem of SVM is constructing the plane that is furthuest from both classes.
+The main approach is to use two parallel supporting planes and maximize the margin between them.
+All point in a class should be in one side of the plane in order to be a supporting plane.
+Therefore, for the points with the class label $+1$ we would like there exist $\omega$ and $b$ such that $\omega \cdot x_i\> b$ or $\omega \cdot x_i -b \> 0$%, and for the points with class label $-1$ we would like there exist $\omega$ and $b$ such that $\omega \cdot x_i \< b $.
+Suppose that smallest value of $\mid \omega \cdot x_i -b\mid$ is $\kappa$, then when $ \omega \cdot x_i -b \geq \kappa$.
 The argument inside the decision function is invariant under positive rescalling so we will implicity fix a scale by requiring $\omega \cdot x_i -b \geq 1 $.
-For the points with class label $-1$ we similary require $\omega \cdot x_i -b \leq -1 $.
-To find the plane furthest from both setss, we can simply maximize the distance or margin between the support planes for each class as illustrated in Figure \ref{fig:svm2}.
-The support planes are pussshed apart untill they bump inot small number of data points ( the support vecotrs ) for each class.
-The support vectors in Figure \ref{fig:svm2} are filled datapoints ,
+Similary the points with class label $-1$ requires $\omega \cdot x_i -b \leq -1 $.
+To find the plane furthest from both classes, we need to maximize the distance between the support planes for each class as illustrated in Figure \ref{fig:svm2}.
+To accomplish this, both planes are pushed apart untill they bump into small number of datapoints ( the support vecotrs ) for each class.
+Figure \ref{fig:svm2} shows the support vectors as filled datapoints,
 \begin{figure}
@@
 
-The distance or margin between these supporting planes 
+The distance between these supporting planes 
 \begin{equation}
@@
 \end{equation}
-Thus maximizing the margin is equivalent to minimizing $\frac{{\parallel \omega \parallel}^2}{2}$ int the following quadratic program
+
+Thus maximizing the margin is equivalent to minimizing $\frac{{\parallel \omega \parallel}^2}{2}$ in the next quadratic program
 
@@
 
-The margin depends on the Support vecotrs that are bumpt into the planes.
-This is a primal problem Quadratic Programming (QP) problem.
-The concept of duality can also upsed to wields the problem to a  Largrnigian dual QP problem in Equation \ref{eq:dual}.
+The Support vectors that are pumped into the supprting plane deteremine the margin.
+This is a primal form Quadratic Programming (QP) problem.
+The concept of duality can also used to wields the problem to a Largrnigian dual QP problem in Equation \ref{eq:dual}.
 % maximumm margin between parallel 
@@
 \end{equation}
+
 Both the dual form and the primal form weilds to the same plane $\omega=\sum_{i=1}^{m}y_i\alpha_i x_i$ and threshold $b$ determined by the support vectors (for which $\alpha_i\geq 0$).
@@
 \subsubsection{Theoretcical foundation}
-From a statistical learning theory perspective these QP formulation are well founded Rougly, Statistical Learning proves that bounds on the generalization error on future points not in the training set can be obtained.
-Theses bounds are a function fo misclassiffied error on the training data and tersm sthat measure the complexity or capacity fo classifcation function .
-For linear functions, maximizing the margin of seperation as discussed abouve reduces th efunction capacilty or complexity.
-Thus, by explicity maximizing the margin we are minimizing bound on the generalition error and can expect better generalization with high probability.
-The size of the margin is not directly dependent on the dimensionaliuty of the data.
-Thus, we can expect good performance event for very high dimensional data (i.e, with a very large number of attribute).
-In a sense, problems caused by oer fitting of high - dimensional data are greatly reduced.
-The problem is refered to the large volume of literature in, e.g., \cite{bookSVMoverfit11,statisticalSVM1} for more technical discussion of the taht statistical theroy.
+
+The generalization error on future points and not only in the training set can be obtained as proved by the statical learning theory \cite{BennettSVMP2000}.
+The generalized error bound is a function of the misclassified error on the training data and the terms that measure the complexiy or capacity of the classification.
+In linear functions, using planes to maximize the margin of seperation reduces the complexiy of the functions.
+Thus, by explicity maximizing the margin we are in the same time  minimizing bound on the generalization error and therefore expect better generalization with high probability.
+The dimensionality of the data dose not affect the size of the margin which   results in good performance  for very high dimensional data (i.e, with a very large number of attribute).
+In a sense, this also reduces the  problems caused by over fitting of high - dimensional data.
+A large volume of literature  , e.g., \cite{bookSVMoverfit11,statisticalSVM1} refers to this problem,  for more technical discussion of the   statistical theroy.
 

### 2009-08-16 trunk/draft/System.tex
after most changes

@@
 
-\section{Preprocessing}
+\section{Possible Dominant Point Extraction}
 \label{sec:Preprocessing}
@@
 The system choose the minimum segmentation as the final segmentation of this stroke.
+
+After the best fit segmentaion is choosen the paramter of each primitive is computed.
+For example, if the stroke is represented as an ellispe, the center and axii of the ellipse is computed and confirmed.
+For lines the slope, equation, start point and end point are computed.
+These values and paramter will help determine the spatial relations between different segments in the feature extraction step.
 % The segmentation which crossponds to the minimum $E_{tot}$ is the best segmentation.  
@@
 
+After the segmentation, each stroke is represented as a list of segments $L_s$.
+A segment can be either an ellipse, a line or a circular arc.
+The list of segments $L_s$ represents the segmentaion of the storke.
+When a user draw another stroke the generated segmentaion of the storke is appended to the list of segments.
+When the user finish drawing the symbol the segment list $L_s$ will contain all segments of the drawn symbol.
+The relation and properties between different segments are extracted from the segment list $L_s$.
+This step is considered as feature extration which is dircribed in details in the next sections.
+
 The system uses a composite set of features which include global shape properties like Rubine feature set \cite{gestureexample12},  ink density \cite{GeometryAndDomain102}, some appearance based properties like Zernike moments \cite{HeloiseBeautification}, and some stroke based structural information like number of perpendiculars lines , number of parallel lines and types of primitives in each symbol.
 The following list gives details on all the features used in the system.
+
 \subsection{Structural and geometrical Features(FS1)}
@@
 
- The vector $\omega$ determines th orientation of a seperation plane.
-The scalar $b$ determines the offeset of the plane from the orgin.
-Lets assume that the datatpointa re linearly seperable, i.e.
-there are infinitely many possible seperating planes that correctly classify the training data.
-Figure\ref{fig:svm1} illustrates two different seperating planes.
-It is intuitively noted from the figure that the solid line is more likely to generalize more data and thus introduce less classification error.
-This can be geometrically characterized as plane furthest from both classes.
+ The orientation of the speration plane is deteremined using the vector $\omega$ and the offest of the plane from the orgin is determined using the scalar $b$.
+Lets make an assumption that the datatpoint are linearly seperable, i.e.
+there are infinite possibilities to find a plane that seperates and correctly classify the training data.
+This is idea is illustrated in Figure\ref{fig:svm1} where two different seperating planes can correctly classify the data.
+It is intuitively noted from the figure that the solid line is more likely to generalize more data.
+This means less classification error will be introduced.
+The geometrical characterization of this plane means the plane furthest from both classes.
 % Which one is preferable ?
@@
 \subsubsection{Theoretcical foundation}
-From a statistical learning theory perspective these QP formulation are well founded Rougly, Statistical Learning proves that bounds on the generalization error on future points not in the training set can be obtained.
-Theses bounds are a function fo misclassiffied error on the training data and tersm sthat measure the complexity or capacity fo classifcation function .
-For linear functions, maximizing the margin of seperation as discussed abouve reduces th efunction capacilty or complexity.
-Thus, by explicity maximizing the margin we are minimizing bound on the generalition error and can expect better generalization with high probability.
-The size of the margin is not directly dependent on the dimensionaliuty of the data.
-Thus, we can expect good performance event for very high dimensional data (i.e, with a very large number of attribute).
-In a sense, problems caused by oer fitting of high - dimensional data are greatly reduced.
-The problem is refered to the large volume of literature in, e.g., \cite{bookSVMoverfit11,statisticalSVM1} for more technical discussion of the taht statistical theroy.
-
-We can gain inishgit into thesse result susing geomertic arguments. Classification function hat have capcity to fit the training data are more likely to overfit  resulting in poor generalization.
+
+The generalization error on future points and not only in the training set can be obtained as proved by the statical learning theory \cite{BennettSVMP2000}.
+The generalized error bound is a function of the misclassified error on the training data and the terms that measure the complexiy or capacity of the classification.
+In linear functions, using planes to maximize the margin of seperation reduces the complexiy of the functions.
+Thus, by explicity maximizing the margin we are in the same time  minimizing bound on the generalization error and therefore expect better generalization with high probability.
+The dimensionality of the data dose not affect the size of the margin which   results in good performance  for very high dimensional data (i.e, with a very large number of attribute).
+In a sense, this also reduces the  problems caused by over fitting of high - dimensional data.
+A large volume of literature  , e.g., \cite{bookSVMoverfit11,statisticalSVM1} refers to this problem,  for more technical discussion of the   statistical theroy.
+
+%Using geomteric arguments we can gain in sights to some of these results.
+Classification function that have capcity to fit the training data are more likely to overfit  resulting in poor generalization.
 
 \subsubsection{Linearly Inseperable Case}
- So far we have assumed that the two datasets are linearly separable.
-This is not true, the startegy of constructing the plane that bisect the two closest points of the convex hulls will fail.
-As illustrated in Figure \ref{fig:svm5}, if the points are not linearly seperable the primal supporitng plane will fails.
-Since the QP task is not feasible for ethe linearly inseperable case , then we need to relax the constrains.
-Consider the linear inseperable problem in Figure\ref{fig:svm5}, Ideally we would like no points to be missclassified and no points to fall in the margin.
+In all previous sections we assumed that the problem of classifiying two set is linearly separable.
+The assumption is not true in most casses, thus, trying to construct a plane to seperate the two dataset will fail.
+Figure \ref{fig:svm5} illustrates that we cannot construct a plane that can seperate the two sets correctly.
+The constrains for constructing the planes must be relaxed since the QP task is not feasible for the linearly inseperable case.
+Consider the linear inseperable problem in Figure\ref{fig:svm5}, in the ideal case there will be no point missclassified and no point falls on the margin.
 But we must relax the constrain that insure that each point in on the approiate side of tis supporitng plane.
-Any point fallihng on the wrong side of tis supprting plane is considered to be an error.
-We want to maximize the margin and minimize the error.
+Every point that falls on the wrong side of the supporting plane is considered an error.
+The problem is now to maximize the margin and minimize the error.
 
@@
 \end{figure}
-
- This could be accomplished through minor change in the supporging plane QP problem.
-A nonegative slack or error variable $z_i$ is added to each constraint and then added as weighted penalty term in the objective as follows:
+A minor change in the supporting plane QP problem can accomplish this.
+A nonegative slack or error variable $z_i$ is added to each constraint.
+A weighted penalty term is also added to the objective as follows:
  \begin{equation}
@@
 
-Once again the primal relaxed supprting plane method is equivilent to the dual problem. The Largangian dual of the QP task is :
+The primal relaxed supporting plane method is equivalent to the dual problem. The Largangian dual of the QP task is :
 
@@
 \subsubsection {Non linear function Using Kernels}
-If we consider a classification problem that cannot be seperated using a linear space (See Figure \ref{fig:svm6} ).
-A quadratic or  cubic function will work better in some casses.
-A classific method of converting a linear classification algoirhtm inot a nonolinear classification algoorithm is to simply add addional attribute to the data that are nonlinear functions of the original data.
-Existing linar classification algorithms can be then applied to the expanded dataset feature space producing non linear functions in the original input space.
-To construct a quadratic disceriminant in a two dimensional vector space with attributes $r$ and $s$, simply map the original two dimensional input space $\left[r,s\right]$ to the five dimensional features space $\left[r,s,rs,r^2,s^2\right]$ and construct a linear discriminiant in that space.
+Figure \ref{fig:svm6} shows a classification problem that cannot be seperated using linear space.
+A higher order plane can work better in similar casses.
+Converting a linear classification algorithm into non linear classification algoirhtm  can be achieved by adding addional attribute to the data that are nonlinear functions of the original data.
+The expanded dataset features space can be seperated using an  existing linear classification algorithm.
+The  linear classication that is applied on the feautre space produce a non linear classification on the orginal inputspace.
+In two dimensional vector space with attributes $r$ and $s$, to construct a quadratic  discriminant  simply map the original two dimensional input space $\left[r,s\right]$ to  five dimensional features space $\left[r,s,rs,r^2,s^2\right]$ and construct a linear discriminiant in that space.
 specifically, define : $\vartheta(x):R^2  \rightarrow R^5$ then
@@
 
-For higher dimentional datasets, this nonlinear mapping method has two potential problems stemming from the fact taht dimensionality of the feature space explones exponentially.
+For higher dimentional datasets, this method of nonlinear mapping  has two potential problems stemming from the fact that dimensionality of the feature space explones exponentially.
 The first problem is that overfitting becomes a problem.
-SVM are largely immune to this porlblem since they rely on margin maximization.
-Provided an appropriate value of paaramter $C$ is chosen.
-The second concern is that it is not practical to actually compute $\vartheta(x)$.
-SVMs get around this issue through teh use of kernels.
+However, since SVM rely on margin maximization they are largely immune to this porblem.
+Provided an appropriate value of paramter $C$ is chosen.
+The second concern is that it is the computaional complexity of   computing $\vartheta(x)$ but SVM use kernels to get around this problem.
 

### 2009-08-17 trunk/draft/Results.tex

@@
 Each part is evaluated separately.
-The preprocessing procedures and computation is evaluated using their effect on the segmentation algorithm.
+The Possible Dominant Point Extraction procedures and computation is evaluated using their effect on the segmentation algorithm.
 These results are reported in section \ref{sec:ResultsDetails}.
@@
 	\centering		
-	 \includegraphics[scale=0.5]{paperImages/swarmAgentTest.jpg.eps}
+	 \includegraphics[scale=0.5]{paperImages/swarmAgentTest.eps}
 	 	\caption{Segmentation Error Vs. Number of particles}
@@
 	%\caption{Experiments results :} a)    b) c) 
+\end{figure} 
+\begin{figure}
+	\centering
+		\includegraphics{paperImages/swarmIterationTest.eps}
+	\caption{Segmentation Error Vs. Number of Swarm iteration}
+	\label{fig:swarmIterationTest}
 \end{figure}

### 2009-08-17 draft/System.tex

@@
 
-The \textbf{preprocessing} step is responsible for capturing the input data and removing the noise from it.
-As shown in Figure \ref{fig:Blockdiagram} preprocessing consists of computing the time differences, direction, speed and curvature of every point in the stroke.
+The \textbf{Possible Dominant Point Extraction} step is responsible for capturing the input data and removing the noise from it.
+As shown in Figure \ref{fig:Blockdiagram} Possible Dominant Point Extraction consists of computing the time differences, direction, speed and curvature of every point in the stroke.
 After computing this information the local minimum and local maximum of each curve is extracted.
@@
 Otherwise, the stroke is passed to two particle swarm algorithms that divide the stroke to either a polygon or a set of lines and curves.
-The algorithms takes the stroke points along with the possible dominant points $P_{pd}$ computed during preprocessing then produce a set of dominant points which are connected by either lines or curves.
+The algorithms takes the stroke points along with the possible dominant points $P_{pd}$ computed during Possible Dominant Point Extraction then produce a set of dominant points which are connected by either lines or curves.
 The two algorithms will generate two segmentations, the system will choose the segmentation that has the minimum segmentation error.%
@@
 
-\section{Preprocessing}
+\section{Possible Dominant Point Extraction}
 \label{sec:Preprocessing}
 %Some data is extracted from the points in the stroke.
-The preprocessing step captures the points from the pointing devices then compute the information needed to determine a set of  \textit{Possible Dominant Points $P_{pd}$} which guides the segmentation step.
+The Possible Dominant Point Extraction step captures the points from the pointing devices then compute the information needed to determine a set of  \textit{Possible Dominant Points $P_{pd}$} which guides the segmentation step.
 It is noted as a person draws a shape the pen slows down near corners and picks up speed when drawing straight lines.
@@
 The array that contains possible points $P_{pd}$ is used as input for the next segmentation step.
-Figure \ref{fig:LabelsPPD} shows the particles labeled as Possible dominant points $P_{pd}$ by the preprocessing, it is noticed the redundancy of some $P_{pd}$ points.
+Figure \ref{fig:LabelsPPD} shows the particles labeled as Possible dominant points $P_{pd}$ by the Possible Dominant Point Extraction step, it is noticed the redundancy of some $P_{pd}$ points.
 
@@
 Otherwise, the stroke is passed to two particle swarm algorithms that divide the stroke to either lines or lines and curves.
-The algorithms take the stroke points along with the possible dominant points $P_{pd}$ computed during preprocessing then produce a set of dominant points which are connected by either lines or curves.
+The algorithms take the stroke points along with the possible dominant points $P_{pd}$ computed during Possible Dominant Point Extraction then produce a set of dominant points which are connected by either lines or curves.
 The next section describes the ellipse detection algorithm and the two particle swarm algorithms used to divide the stroke.
@@
 The system choose the minimum segmentation as the final segmentation of this stroke.
+
+After the best fit segmentation is chosen the parameter of each primitive is computed.
+For example, if the stroke is represented as an ellipse, the center and axii of the ellipse is computed and confirmed.
+For lines the slope, equation, start point and end point are computed.
+These values and parameter will help determine the spatial relations between different segments in the feature extraction step.
 % The segmentation which crossponds to the minimum $E_{tot}$ is the best segmentation.  
@@
 
+After the segmentation, each stroke is represented as a list of segments $L_s$.
+A segment can be either an ellipse, a line or a circular arc.
+The list of segments $L_s$ represents the segmentation of the stroke.
+When a user draw another stroke the generated segmentation of the stroke is appended to the list of segments.
+When the user finish drawing the symbol the segment list $L_s$ will contain all segments of the drawn symbol.
+The relation and properties between different segments are extracted from the segment list $L_s$.
+This step is considered as feature extraction which is described in details in the next sections.
+
 The system uses a composite set of features which include global shape properties like Rubine feature set \cite{gestureexample12},  ink density \cite{GeometryAndDomain102}, some appearance based properties like Zernike moments \cite{HeloiseBeautification}, and some stroke based structural information like number of perpendiculars lines , number of parallel lines and types of primitives in each symbol.
 The following list gives details on all the features used in the system.
+
 \subsection{Structural and geometrical Features(FS1)}
@@
 We used the library implementation of LibSVM \cite{libsvm} and its implementation of the one-vs-one strategy.
-Section \ref{Sec:SVMdetail} discribe in details the classifier, how classes (classification) are determined and the reasons behind using this specific classification methods.
-
-\subsection {Trainig the SVM}
-\subsection {Testing}
+Section \ref{Sec:SVMdetail} describe in details the classifier, how classes (classification) are determined and the reasons behind using this specific classification methods.
+
+\subsection {Training and Testing}
+A training set that contains positive samples of each category in the data set is used to train the SVM classifier.
+The training starts after computing the feature vector of each sample.
+The system trains $ \frac{n(n-1)}{{2}}$ binary classifiers each distinguish between only two categories.
+The goal of the training process is to generate a linear SVM plane that maximize margin between the two categories.
+A cross validation procedure is implemented to choose the best value for the $C$ paramter in SVM classifier.( $C=100$ was the best value)
+=
+For classification the decision is made from  producing the feature vector of the unkown symbol to the trained $\frac{n(n-1)}{{2}}$ binary classifiers.
+The final decision is given using simple voting strategy.
 
@@
 The foundation of Support Vector Machines (SVM) were developed by Vapnik \cite{svmintroduce} in 1995.
-It is gainig increase popularity due to the high performance and efficient implementation.
-The goal of SVM algorithms is to generate a sperating plane between two different classes which are represented using a set of n-dimentional vectors.
-The generated plane is used as a good classifier to the unseen exmaples.
+It is gaining increase popularity due to the high performance and efficient implementation.
+The goal of SVM algorithms is to generate a separating plane between two different classes which are represented using a set of n-dimensional vectors.
+The generated plane is used as a good classifier to the unseen examples.
 However, there can be more than one plane that can separates the data therefore SVM attempts to find the plane that maximize the margin (distance between classifier and nearest sample) between the two classes \ref{fig:svm1}.
-To calcuate the margin, two different hyperplanes are constructed, one plane on each side of the data, these two planes are then moved to generenate the maximum margin possilbe.
+To calculate the margin, two different hyperplanes are constructed, one plane on each side of the data, these two planes are then moved to generate the maximum margin possible.
 \begin{figure}
@@
 
-Let us consider  a two class classification task with datapoints $x_i$ where $i=1,\cdots,m$ having corresponding labels $y_i=\pm1$ where if  $y_i=1$ if $x_i$ belonged to first class and $y_i=-1$ if $x_i$ belonged to second class.
-Each data  point $x_i$ is represented in $N$ dimensional input or attribule space.
+Let us consider  a two class classification task with data points $x_i$ where $i=1,\cdots,m$ having corresponding labels $y_i=\pm1$ where if  $y_i=1$ if $x_i$ belonged to first class and $y_i=-1$ if $x_i$ belonged to second class.
+Each data  point $x_i$ is represented in $N$ dimensional input or attribute space.
 Let the classification plane be represented as function.
@@
 
- The orientation of the speration plane is deteremined using the vector $\omega$ and the offest of the plane from the orgin is determined using the scalar $b$.
-Lets make an assumption that the datatpoint are linearly seperable, i.e.
-there are infinite possibilities to find a plane that seperates and correctly classify the training data.
-This is idea is illustrated in Figure\ref{fig:svm1} where two different seperating planes can correctly classify the data.
+ The orientation of the separation plane is determined using the vector $\omega$ and the offset of the plane from the origin is determined using the scalar $b$.
+Lets make an assumption that the datat point are linearly separable, i.e.
+there are infinite possibilities to find a plane that separates and correctly classify the training data.
+This is idea is illustrated in Figure\ref{fig:svm1} where two different separating planes can correctly classify the data.
 It is intuitively noted from the figure that the solid line is more likely to generalize more data.
@@
 
-The main problem of SVM is constructing the plane that is furthuest from both classes.
+The main problem of SVM is constructing the plane that is furthest from both classes.
 The main approach is to use two parallel supporting planes and maximize the margin between them.
@@
 Suppose that smallest value of $\mid \omega \cdot x_i -b\mid$ is $\kappa$, then when $ \omega \cdot x_i -b \geq \kappa$.
-The argument inside the decision function is invariant under positive rescalling so we will implicity fix a scale by requiring $\omega \cdot x_i -b \geq 1 $.
-Similary the points with class label $-1$ requires $\omega \cdot x_i -b \leq -1 $.
+The argument inside the decision function is invariant under positive rescaling so we will implicitly fix a scale by requiring $\omega \cdot x_i -b \geq 1 $.
+Similarly the points with class label $-1$ requires $\omega \cdot x_i -b \leq -1 $.
 To find the plane furthest from both classes, we need to maximize the distance between the support planes for each class as illustrated in Figure \ref{fig:svm2}.
-To accomplish this, both planes are pushed apart untill they bump into small number of datapoints ( the support vecotrs ) for each class.
-Figure \ref{fig:svm2} shows the support vectors as filled datapoints,
+To accomplish this, both planes are pushed apart until they bump into small number of data points ( the support vectors ) for each class.
+Figure \ref{fig:svm2} shows the support vectors as filled data points,
 \begin{figure}
@@
 
-The Support vectors that are pumped into the supprting plane deteremine the margin.
+The Support vectors that are pumped into the supporting plane determine the margin.
 This is a primal form Quadratic Programming (QP) problem.
@@
 
-Both the dual form and the primal form weilds to the same plane $\omega=\sum_{i=1}^{m}y_i\alpha_i x_i$ and threshold $b$ determined by the support vectors (for which $\alpha_i\geq 0$).
-
-\subsubsection{Theoretcical foundation}
+Both the dual form and the primal form wields to the same plane $\omega=\sum_{i=1}^{m}y_i\alpha_i x_i$ and threshold $b$ determined by the support vectors (for which $\alpha_i\geq 0$).
+
+\subsubsection{Theoretical foundation}
 
 The generalization error on future points and not only in the training set can be obtained as proved by the statical learning theory \cite{BennettSVMP2000}.
-The generalized error bound is a function of the misclassified error on the training data and the terms that measure the complexiy or capacity of the classification.
-In linear functions, using planes to maximize the margin of seperation reduces the complexiy of the functions.
-Thus, by explicity maximizing the margin we are in the same time  minimizing bound on the generalization error and therefore expect better generalization with high probability.
+The generalized error bound is a function of the misclassified error on the training data and the terms that measure the complexity or capacity of the classification.
+In linear functions, using planes to maximize the margin of separation reduces the complexity of the functions.
+Thus, by explicitly maximizing the margin we are in the same time  minimizing bound on the generalization error and therefore expect better generalization with high probability.
 The dimensionality of the data dose not affect the size of the margin which   results in good performance  for very high dimensional data (i.e, with a very large number of attribute).
 In a sense, this also reduces the  problems caused by over fitting of high - dimensional data.
-A large volume of literature  , e.g., \cite{bookSVMoverfit11,statisticalSVM1} refers to this problem,  for more technical discussion of the   statistical theroy.
-
-We can gain inishgit into thesse result susing geomertic arguments. Classification function hat have capcity to fit the training data are more likely to overfit  resulting in poor generalization.
-
-\subsubsection{Linearly Inseperable Case}
- So far we have assumed that the two datasets are linearly separable.
-This is not true, the startegy of constructing the plane that bisect the two closest points of the convex hulls will fail.
-As illustrated in Figure \ref{fig:svm5}, if the points are not linearly seperable the primal supporitng plane will fails.
-Since the QP task is not feasible for ethe linearly inseperable case , then we need to relax the constrains.
-Consider the linear inseperable problem in Figure\ref{fig:svm5}, Ideally we would like no points to be missclassified and no points to fall in the margin.
-But we must relax the constrain that insure that each point in on the approiate side of tis supporitng plane.
-Any point fallihng on the wrong side of tis supprting plane is considered to be an error.
-We want to maximize the margin and minimize the error.
+A large volume of literature  , e.g., \cite{bookSVMoverfit11,statisticalSVM1} refers to this problem,  for more technical discussion of the   statistical theory.
+
+%Using geomteric arguments we can gain in sights to some of these results.
+Classification function that have capcity to fit the training data are more likely to overfit  resulting in poor generalization.
+
+\subsubsection{Linearly Inseparable Case}
+In all previous sections we assumed that the problem of classifying two set is linearly separable.
+The assumption is not true in most cases, thus, trying to construct a plane to separate the two dataset will fail.
+Figure \ref{fig:svm5} illustrates that we cannot construct a plane that can separate the two sets correctly.
+The constrains for constructing the planes must be relaxed since the QP task is not feasible for the linearly inseparable case.
+Consider the linear inseparable problem in Figure\ref{fig:svm5}, in the ideal case there will be no point misclassified and no point falls on the margin.
+But we must relax the constrain that insure that each point in on the appropriate side of tis supporting plane.
+Every point that falls on the wrong side of the supporting plane is considered an error.
+The problem is now to maximize the margin and minimize the error.
 
@@
 		\includegraphics{afterDefense/svm5.jpg.eps}
-	\caption{Linearly Inseperable}
+	\caption{Linearly Inseparable}
 	\label{fig:svm5}
 \end{figure}
-
- This could be accomplished through minor change in the supporging plane QP problem.
-A nonegative slack or error variable $z_i$ is added to each constraint and then added as weighted penalty term in the objective as follows:
+A minor change in the supporting plane QP problem can accomplish this.
+A non negative slack or error variable $z_i$ is added to each constraint.
+A weighted penalty term is also added to the objective as follows:
  \begin{equation}
@@
 \subsubsection {Non linear function Using Kernels}
-If we consider a classification problem that cannot be seperated using a linear space (See Figure \ref{fig:svm6} ).
-A quadratic or  cubic function will work better in some casses.
-A classific method of converting a linear classification algoirhtm inot a nonolinear classification algoorithm is to simply add addional attribute to the data that are nonlinear functions of the original data.
-Existing linar classification algorithms can be then applied to the expanded dataset feature space producing non linear functions in the original input space.
-To construct a quadratic disceriminant in a two dimensional vector space with attributes $r$ and $s$, simply map the original two dimensional input space $\left[r,s\right]$ to the five dimensional features space $\left[r,s,rs,r^2,s^2\right]$ and construct a linear discriminiant in that space.
+Figure \ref{fig:svm6} shows a classification problem that cannot be separated using linear space.
+A higher order plane can work better in similar cases.
+Converting a linear classification algorithm into non linear classification algorithm  can be achieved by adding additional attribute to the data that are nonlinear functions of the original data.
+The expanded dataset features space can be separated using an  existing linear classification algorithm.
+The  linear classification that is applied on the feature space produce a non linear classification on the original input space.
+In two dimensional vector space with attributes $r$ and $s$, to construct a quadratic  discriminant  simply map the original two dimensional input space $\left[r,s\right]$ to  five dimensional features space $\left[r,s,rs,r^2,s^2\right]$ and construct a linear discriminant in that space.
 specifically, define : $\vartheta(x):R^2  \rightarrow R^5$ then
@@
 
-For higher dimentional datasets, this nonlinear mapping method has two potential problems stemming from the fact taht dimensionality of the feature space explones exponentially.
-The first problem is that overfitting becomes a problem.
-SVM are largely immune to this porlblem since they rely on margin maximization.
-Provided an appropriate value of paaramter $C$ is chosen.
-The second concern is that it is not practical to actually compute $\vartheta(x)$.
-SVMs get around this issue through teh use of kernels.
+For higher dimensional datasets, this method of nonlinear mapping  has two potential problems stemming from the fact that dimensionality of the feature space explones exponentially.
+The first problem is that over fitting becomes a problem.
+However, since SVM rely on margin maximization they are largely immune to this problem.
+Provided an appropriate value of parameter $C$ is chosen.
+The second concern is that it is the computational complexity of   computing $\vartheta(x)$ but SVM use kernels to get around this problem.
 
@@
 Notice that the mapped data only occurs as inner product in the objective.
-Now we apply a little mathematically rigorous magie known as hilbern schmidt kernels, first applied to SVMs in \cite{svmintroduce}.
-By Mercer's theorem, we knoew that for ccertain mapping $\theta$ and any two points $u$ and $v$, the inner product of the the mapped points can be evalutated using the kernel function without ever explicity knowing ht emapping , e.g., $\theta(u)\theta(v)\equiv K(u,v)$.
+Now we apply a little mathematically magic known as hilbern schmidt kernels, first applied to SVMs in \cite{svmintroduce}.
+By Mercer's theorem, we know that for certain mapping $\theta$ and any two points $u$ and $v$, the inner product of the the mapped points can be evaluated using the kernel function without ever explicitly knowing the mapping , e.g., $\theta(u)\theta(v)\equiv K(u,v)$.
 Some of the more popular known kernels are given below.
-New kernel are being developed to fit domian specific requriments.
+New kernel are being developed to fit domain specific requirements.
 \begin{center}
@@
 \end{center}
-To change from a linear to nonlinear classifier, one must only substitute a kernel evaluation in the objective insted of the original dot product.
-Thus by changing kernels we can get different hightly nonlinear classifiers.
-No algorithmic changes are required from linear clase other than substitue of a kernel evaluation for simple dot product.
-All the benefits fo the original linear SVM methods are maintained.
-We can train a highly nonlinear classification function such as a ploynomial or a raadial basis function machine, or a sigmoidal neural network using robust, efficient algorithsm that have no problem with local minima.
-By using kernel substitution a linear algorithm (only capable of handling seperable data) can be turned into a general nonlinear algorithm.
+To change from a linear to nonlinear classifier, one must only substitute a kernel evaluation in the objective instead of the original dot product.
+Thus by changing kernels we can get different highly nonlinear classifiers.
+No algorithmic changes are required from linear case other than substitute of a kernel evaluation for simple dot product.
+All the benefits of the original linear SVM methods are maintained.
+We can train a highly nonlinear classification function such as a polynomial or a radial basis function machine, or a sigmoid neural network using robust, efficient algorithms that have no problem with local minima.
+By using kernel substitution a linear algorithm (only capable of handling separable data) can be turned into a general nonlinear algorithm.
 %\begin{equation}
@@
 \begin{enumerate}
-	\item Select paramter C representign the tradeoff between minimizign the training set error and maximizing the margin.
-Select the kernel function and any kernel paramters.
-For example for the radial basis function kernel, one must select width of the gaussian $\gamma$.
-	\item Solve dual QP Equation \ref{} or an alterantive SVM formulation using an approriate quadratic programming or linear programming alogrithm. 
+	\item Select parameter C representing the tradeoff between minimizing the training set error and maximizing the margin.
+Select the kernel function and any kernel parameters.
+For example for the radial basis function kernel, one must select width of the Gaussian $\gamma$.
+	\item Solve dual QP Equation \ref{} or an alternative SVM formulation using an appropriate quadratic programming or linear programming algorithm. 
 	\item Recover th primal threshold variable $b$ using the support vectors. 
@@
 
-The paramters in step 1. are selected using cross valiation if sufficent data are avaliable. %However, recent model  selectin strategies can
+The parameters in step 1. are selected using cross validation if sufficient data are available. %However, recent model  selectin strategies can 
+\subsubsection {Multi Class Classifiers}
 
 \subsubsection{ SVM in Machine Learning }
-SVM has been successfully applied to a number of applications ranging from particle identification \cite{ParticleSVM7}.
-Face detection  \cite{faceSVM8,faceSVM7} and text categorization \cite{TextSVM7} to engine knock detection.
-SVMs are very suited to the classification task that commonly arise in machine vision.
-As an example we consider and application involving face identification\cite{faceSVM8}.
-Three different SVM classiifer were used on training set of 200 images (5 image per person).
+SVM has been applied sucessfully to various application ranging from particle identification \cite{ParticleSVM7}.
+Other application like Face detection \cite{faceSVM8,faceSVM7} and text categorization \cite{TextSVM7}.
+Also SVM suits classification task that is commonly rise in Machine vision.
+For example we will consider face identification\cite{faceSVM8} problem.
+In \cite{faceSVM8} three different SVM classifier were used on training set of 200 images (5 image per person).
 The error rate was less than 5\%.
-An RBF network and SVM classifiers were compared on the UPUS handwritten digit recognition database\cite{ORLDataset}.
+Another example is digit recognition problem wehre an RBF network and SVM classifiers were compared on the UPUS handwritten digit recognition database\cite{ORLDataset}.
 SVM outperformed and RBF network on all digits.
 
-\subsubsection {SVM vs. Other Learning Methods}
+%\subsubsection {SVM vs. Other Learning Methods}
 
 SVM have been applied ot the a largest MNIST a Handwritten digit recognition dataset of 60000 training set and 10000 test set\cite{IjdarSherifPaper}.
-Sherif and Ezzat has performed a comparision between SVM and several other techniques on the MNIST dataset.
-They also compared the same classfication methods on arabic handwritten digits (MADBASE) \cite{ADBase9,IjdarSherifPaper}.
+Sherif and Ezzat has performed a comparison between SVM and several other techniques on the MNIST dataset.
+They also compared the same classification methods on arabic handwritten digits (MADBASE) \cite{ADBase9,IjdarSherifPaper}.
 Their results shows that SVM classifier outperform all other techniques on both MNIST and MADBase datasets and achieved recognition rate of over 98\%.
@@
 
-This seperior performance because they may elimnates many of problem experienced with other inference methodoliges with other inference methodollogies like neural network and decision trees.
+This superior performance because they may eliminates many of problem experienced with other inference methodoliges with other inference methodologies like neural network and decision trees.
 The first problem that is eliminated was the problems with local minima.
-SVM can construct highly nonlinear classification and regression function without woring about getting stuck at local minima.
+SVM can construct highly nonlinear classification and regression function without working about getting stuck at local minima.
 Another problem SVM eliminated is the number of parameter that are picked ot generate the model.
-For example, SVM classifier with RBF  kernel will only need 2 paramters to set the model the $C$ cost paramter and $\sigma$ paramter of the RBF kernel.
+For example, SVM classifier with RBF  kernel will only need 2 parameters to set the model the $C$ cost parameter and $\sigma$ parameter of the RBF kernel.
 The last reason is the final result are sable, reproducible and largely independent of the specific algorithm used to  optimize the SVM model.
-If two users apply the same SVM model with the same parameters to the same data, they will et the same solution modeule numeric issues.
+If two users apply the same SVM model with the same parameters to the same data, they will et the same solution module numeric issues.
 Compare this with neural networks where the results are dependent on the particular algorithms and starting point used.
@@
 
-\subsubsection {Multi Class Classifiers}
-
 \subsubsection {SVM Implementations}
-Typically an SVM approach requires the solution of a QP or LP problem.
-LP and QP type problems have been extensively studied inthe field of mathematical programming.
-These can be divided into three categories techniques in which kernel compoinents are evaluated and discarded during learning.
-Decomposition methods in which an evolving subset of data is used and new optimization approaches that specifically exploit the structure of the SVM problem.
-
-For the first category the most obvioius  approach is to sequentially update the $\alpha_i$  and this is the approach used by the kernel  Adtron (KA) algorithm \cite{kernelSVM4}.
+To solve SVM approach require solving a QP or LP problem.
+Mathematical programming have extensively studied the LP and QP type problems.
+These techniques can be divided into three categories, methods in which evaluate and discard kernel components during learning.
+Decomposition methods in which an evolving subset of data are used and new optimization approaches that specifically utilize the structure of the SVM problem.
+
+In the first category the most obvious approach is to sequentially update the $\alpha_i$,  kernel Adtron  used this approach in (KA) algorithm \cite{kernelSVM4}.
 
 Chunking and decomposition methods optimize the SVM with respect to subsets.
-Rather than sequentially updateing the $\alpha_i$ the alternative is to upodate the $\alpha_i$ in parallel but using only a subset or wokring set of data at each stage.
-There are many successful code based on these decomposition strategies.
-SVM codes available online such as SVMTorch \cite{svmTorch} and SVMLight \cite{SVMlight} use these working set strategies.
-
-The third approach is to directly attack the SVM  problem from and optimization prespective and create algorithms that explicity exploits the structure of teh problem.
+They update $\alpha_i$ in parallel but using only a subset or working set of data at each stage rather than sequentially updating the $\alpha_i$.
+Decomposition strategies have been successfully coded in many libraries.
+SVM libraries are available on line such as SVMTorch \cite{svmTorch} and SVMLight \cite{SVMlight}.
+
+The third approach is to directly attack the SVM  problem from and optimization perspective and create algorithms that explicitly exploits the structure of teh problem.
 Keerthi et al.
@@
 
-\subsubsection {Summary Of SVM}
+%\subsubsection {Summary Of SVM}
 
@@
 \label{secstepExample}
-In this section a step by step example is presented to clearify the presented sytesm.
+In this section a step by step example is presented to clarify the presented system.
 Let us make an assumption that the user will draw a clock using two strokes.
-Figure \ref{} show the two stroke and how the user draw them.
-The next steps demonstrate how ouput of the system after each step.
-The first section \ref{sec:stepseg} discribe how the segmentation is done step by step.
+Figure \ref{fig:clock} show the two stroke and how the user draw them.
+The next steps demonstrate how output of the system after each step.
+The first section \ref{sec:stepseg} describe how the segmentation is done step by step.
 Section \ref{sec:steprec} shows the details of feature calculation and the classification process.
@@
 \label{sec:stepseg}
+
+\begin{figure}
+	\centering
+
+	\subfigure[First Stroke]{\label{fig:clockstroke1}
+		\includegraphics{afterDefense/clockstroke1.jpg.eps} }
+
+		\subfigure[Second Stroke]{	\label{fig:clockstroke2}
+		\includegraphics{afterDefense/ClockComplete.jpg.eps}}
+
+		\subfigure[Completed symbol]{
+		\label{fig:Clock1}\includegraphics{afterDefense/Clock1.jpg.eps}	}
+	\caption{Stroke Drawing Sequence} a) The user first draws the ellipse b) The square wave is drawn inside the ellipse c) The arrows in the symbols illustrate how the user has drawn the symbo
+	\label{fig:clock}
+\end{figure}
+
 In this step the user draw one or more stroke to draw a symbol.
-Figure \ref{} illustrate the sequence and direction of user draw.
-In this section, each stroke user draw will be labeled as it illustrated in Figure \ref{}.
-Table \ref{} shows the detailed output of the program at each step.
+Figure \ref{fig:clock} illustrate the sequence and direction of user draw.
+In this section, each stroke user draw will be labeled as it illustrated in Figure \ref{fig:clockstroke1} and Figure \ref{fig:clockstroke2} .
+Table \ref{tab:StepsStroke1} shows the detailed output of the program at each step.
 The algorithm that compute the segmentation is demonstrated as follows:
 \begin{enumerate}
-	\item Extracting Stroke Points 
-	\item Extracting $P_dp$
+
+	\item Extracting $P_dp$ as in Section \ref{sec:Preprocessing}
 \begin{verbatim*}
-
+  1. For each point in Stroke do the follwoing:
+  		1.a Compute distance from previous point.
+  		1.b Compute Time difference from previous point and add it to Time_list.  
+  		1.c Compute Velocity and add it to Vel_list. (Eq.  ) 
+  		1.d Compute Direction and add it to Dir_list.(Eq.  )
+  	 	1.e Compute Curvature and add it to Cur_list.(Eq.  ) 
+  	end (for loop)
+  2. For each  list computed in previous step do the follwing:   (Section )
+  	 2.a Compute average of all points. 
+  	 2.b set average as threshold.  
+  	 2.c Create region from continous points larger than threshold. 
+  	 2.d for each region do the following. 
+  	 	   2.d.1 find the maximum 
+  	 	   2.d.1 Add point to P_dp 
+  	  end
+     end 
 \end{verbatim*}	
-\item Ellipse Fitting: which is an attempt to fit as an ellipse
+\item Ellipse Fitting: which is an attempt to fit as an ellipse as in  Section \ref{sec:EllipseDetection} 
 \begin{verbatim*}
+1. Initialize the
 
 \end{verbatim*}	
-\item Segmentation ALS1
+\item Segmentation ALS1 as explained in Section \ref{sec:SwarmSegmentation}
 \begin{verbatim*}
@@
 \end{verbatim*}	
-\item Segmentation ALS2
+\item Segmentation ALS2  as explained in Section \ref{sec:PolygonDivisionAlgorithm}
 \begin{verbatim*}
@@
 \begin{longtable}{|p{2cm}|p{2cm}|p{2cm}|p{2cm}|p{13cm}|}
-	\caption{ Detailed Output of System in Each Step of Segmentation}
+\caption{Detailed Output of System in Each Step of Segmentation}
 \label{tab:StepsStroke1} \\
@@
 \multicolumn{1}{|p{2cm}|}{\textbf{Step}} & 
-\multicolumn{1}{p{2cm}|}{\textbf{Discription}} &
+\multicolumn{1}{p{2cm}|}{\textbf{Description}} &
 \multicolumn{1}{p{2cm}|}{\textbf{Inputs}} &
@@
 \multicolumn{1}{|p{2cm}|}{\textbf{Step}} & 
-\multicolumn{1}{p{2cm}|}{\textbf{Discription}} &
+\multicolumn{1}{p{2cm}|}{\textbf{Description}} &
  \multicolumn{1}{p{2cm}|}{\textbf{Inputs}} &

### 2009-08-17 trunk/draft/Results.tex
adding the matrix

@@
 	\scalebox{0.7}{
-		\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|c|c|c|c|}\hline 
- Categories 	&ellipse	&heart	&trapezoid	&pentagon	&arch	&hexagon	&square	&triangle	&cube	&cylinder	&parallelogram	&moon	&callout\\ \hline 
-ellipse	&150	&20	&0	&0	&0	&5	&1	&0	&0	&0	&0	&1	&0\\  \hline 
-heart	&2	&130	&0	&1	&10	&1	&1	&0	&0	&0	&1	&4	&0\\  \hline 
-trapezoid	&0	&1	&147	&5	&1	&0	&1	&3	&2	&0	&37	&0	&0\\  \hline 
-pentagon	&0	&0	&1	&139	&0	&7	&0	&0	&0	&0	&1	&0	&0\\   \hline 
-arch	&0	&0	&1	&1	&125	&0	&4	&0	&3	&0	&0	&15	&1\\   \hline 
-hexagon	&0	&0	&0	&2	&0	&135	&0	&0	&0	&0	&0	&0	&0\\  \hline 
-square	&0	&0	&0	&0	&0	&0	&138	&0	&0	&0	&0	&1	&0\\   \hline 
-triangle	&0	&0	&1	&2	&5	&0	&1	&148	&0	&0	&4	&7	&0\\  \hline 
-cube	&0	&0	&0	&1	&0	&0	&0	&0	&146	&5	&0	&0	&0\\   \hline 
-cylinder	&0	&0	&0	&1	&0	&1	&0	&0	&0	&147	&0	&0	&0\\  \hline 
-parallelogram	&0	&0	&0	&0	&0	&0	&2	&0	&2	&0	&107	&0	&0\\  \hline 
-moon	&0	&1	&0	&0	&10	&2	&3	&0	&0	&0	&0	&125	&5\\   \hline 
-callout	&0	&0	&0	&0	&0	&0	&0	&0	&0	&0	&0	&0	&145\\  \hline 
+		\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|}\hline 
+		Categorie / classified as ==$>$	&ellipse	&heart	&trapezoid	&pentagon	&arch	&hexagon	&square	&triangle	&cube	&cylinder	&parallelogram	&moon	&callout	&Number of sample per class	&TP count	&FP count	&Fn	&Precision	&Recall ( TP rate )\\
+ellipse	&150	&20	&0	&0	&0	&5	&1	&0	&0	&0	&0	&1	&0	&152	&150	&27	&2	&0.84745762711864	&0.98684210526316\\
+heart	&2	&130	&0	&1	&10	&1	&1	&0	&0	&0	&1	&4	&0	&152	&130	&20	&22	&0.86666666666667	&0.85526315789474\\
+trapezoid	&0	&1	&147	&5	&1	&0	&1	&3	&2	&0	&37	&0	&0	&150	&147	&50	&3	&0.74619289340102	&0.98\\
+pentagon	&0	&0	&1	&139	&0	&7	&0	&0	&0	&0	&1	&0	&0	&152	&139	&9	&13	&0.93918918918919	&0.91447368421053\\
+arch	&0	&0	&1	&1	&125	&0	&4	&0	&3	&0	&0	&15	&1	&151	&125	&25	&26	&0.83333333333333	&0.82781456953642\\
+hexagon	&0	&0	&0	&2	&0	&135	&0	&0	&0	&0	&0	&0	&0	&151	&135	&2	&16	&0.98540145985401	&0.89403973509934\\
+square	&0	&0	&0	&0	&0	&0	&138	&0	&0	&0	&0	&1	&0	&151	&138	&1	&13	&0.99280575539568	&0.91390728476821\\
+triangle	&0	&0	&1	&2	&5	&0	&1	&148	&0	&0	&4	&7	&0	&151	&148	&20	&3	&0.88095238095238	&0.98013245033113\\
+cube	&0	&0	&0	&1	&0	&0	&0	&0	&146	&5	&0	&0	&0	&153	&146	&6	&7	&0.96052631578947	&0.95424836601307\\
+cylinder	&0	&0	&0	&1	&0	&1	&0	&0	&0	&147	&0	&0	&0	&152	&147	&2	&5	&0.98657718120805	&0.96710526315789\\
+parallelogram	&0	&0	&0	&0	&0	&0	&2	&0	&2	&0	&107	&0	&0	&150	&107	&4	&43	&0.96396396396396	&0.71333333333333\\
+moon	&0	&1	&0	&0	&10	&2	&3	&0	&0	&0	&0	&125	&5	&153	&125	&21	&28	&0.85616438356164	&0.81699346405229\\
+callout	&0	&0	&0	&0	&0	&0	&0	&0	&0	&0	&0	&0	&145	&151	&145	&0	&6	&1	&0.96026490066225\\
+Total 	&152	&152	&150	&152	&151	&151	&151	&151	&153	&152	&150	&153	&151	&	&	&	&	&	&\\ 
+% Categories 	&ellipse	&heart	&trapezoid	&pentagon	&arch	&hexagon	&square	&triangle	&cube	&cylinder	&parallelogram	&moon	&callout\\ \hline 
+%ellipse	&150	&20	&0	&0	&0	&5	&1	&0	&0	&0	&0	&1	&0\\  \hline 
+%heart	&2	&130	&0	&1	&10	&1	&1	&0	&0	&0	&1	&4	&0\\  \hline 
+%trapezoid	&0	&1	&147	&5	&1	&0	&1	&3	&2	&0	&37	&0	&0\\  \hline 
+%pentagon	&0	&0	&1	&139	&0	&7	&0	&0	&0	&0	&1	&0	&0\\   \hline 
+%arch	&0	&0	&1	&1	&125	&0	&4	&0	&3	&0	&0	&15	&1\\   \hline 
+%hexagon	&0	&0	&0	&2	&0	&135	&0	&0	&0	&0	&0	&0	&0\\  \hline 
+%square	&0	&0	&0	&0	&0	&0	&138	&0	&0	&0	&0	&1	&0\\   \hline 
+%triangle	&0	&0	&1	&2	&5	&0	&1	&148	&0	&0	&4	&7	&0\\  \hline 
+%cube	&0	&0	&0	&1	&0	&0	&0	&0	&146	&5	&0	&0	&0\\   \hline 
+%cylinder	&0	&0	&0	&1	&0	&1	&0	&0	&0	&147	&0	&0	&0\\  \hline 
+%parallelogram	&0	&0	&0	&0	&0	&0	&2	&0	&2	&0	&107	&0	&0\\  \hline 
+%moon	&0	&1	&0	&0	&10	&2	&3	&0	&0	&0	&0	&125	&5\\   \hline 
+%callout	&0	&0	&0	&0	&0	&0	&0	&0	&0	&0	&0	&0	&145\\  \hline 
  		\end{tabular}

### 2009-08-17 draft/System.tex
adding the matrix

@@
 
+ There is mainly two popular methods to change the binary SVM classification problem into Muliclass classification problem.
+The first methods is One Versus All (OVA) Strategy where M binary classsifiers is constructed one classifier for each class.
+The $ith$ classifier is trained using positive example form $ith$ class and negtive example from all other classes.
+The output function $p_i$ represents the distance and classification of the example from the $ith$ classifier.
+For a new example $x$, OVA SVM strategy assigns it to the class with the largest value of $p_i$.
+
+ The second method is One Versus One Strategy where a binary classifier for every pair of distinct classes ($M$ classes) and so, all together $\frac{M(M \cdot 1)}{2}$ binary classifiers are constructed.
+Each binary classifier is trianed using exapmles from both $p_i$ and $p_j$.
+For example, classifier $C_{ij}$ is trained taking the examples from $p_i$ as positive and the examples from $p_j$ as negative.
+The classification is based on the number of binary classifer that classifiy the new example as $p_i$.
+For a new example $x$, if classifier $C_{ij}$ says $x$ is in class $p_i$, then the vote for class $p_i$ is added by one.
+Otherwise, the vote for class $p_j$  is increased by one.
+After each of the $\frac{M(M \cdot 1)}{2}$  binary classifiers makes its vote,  OVO strategy assigns $x$ to the class with the largest number of votes.
+
+ Both methods proved to be competitive with each other and there is no clear superiority of one method
+over another \cite{Duan05whichis} .
+Some claims that OVO strategy trains in less time than OVA strategy because they train with only portion of the dataset \cite{libsvm} but this claim have not yet been proofed by expeirments.
+
 \subsubsection{ SVM in Machine Learning }
@@
 	\centering
-
 	\subfigure[First Stroke]{\label{fig:clockstroke1}
-		\includegraphics{afterDefense/clockstroke1.jpg.eps} }
+		\includegraphics[scale=0.7]{afterDefense/clockstroke1.jpg.eps} }
 
 		\subfigure[Second Stroke]{	\label{fig:clockstroke2}
-		\includegraphics{afterDefense/ClockComplete.jpg.eps}}
+		\includegraphics[scale=0.7]{afterDefense/ClockComplete.jpg.eps}}
 
 		\subfigure[Completed symbol]{
-		\label{fig:Clock1}\includegraphics{afterDefense/Clock1.jpg.eps}	}
-	\caption{Stroke Drawing Sequence} a) The user first draws the ellipse b) The square wave is drawn inside the ellipse c) The arrows in the symbols illustrate how the user has drawn the symbo
+		\label{fig:Clock1}\includegraphics[scale=0.7]{afterDefense/Clock1.jpg.eps}	}
+	\caption{Stroke Drawing Sequence} a) The user first draws the ellipse b) The square wave is drawn inside the ellipse c) The arrows in the symbols illustrate how the user has drawn the symbol
 	\label{fig:clock}
@@
 Table \ref{tab:StepsStroke1} shows the detailed output of the program at each step.
+The input of this step is the stroke points and the ouptut is a list of segments that wil be used in the recognition process.
 The algorithm that compute the segmentation is demonstrated as follows:
@@
 	\item Extracting $P_dp$ as in Section \ref{sec:Preprocessing}
-\begin{verbatim*}
+\begin{verbatim}
   1. For each point in Stroke do the follwoing:
@@
   		1.b Compute Time difference from previous point and add it to Time_list.  
-  		1.c Compute Velocity and add it to Vel_list. (Eq.  ) 
-  		1.d Compute Direction and add it to Dir_list.(Eq.  )
-  	 	1.e Compute Curvature and add it to Cur_list.(Eq.  ) 
+  		1.c Compute Velocity and add it to Vel_list. 
+  		1.d Compute Direction and add it to Dir_list.
+  	 	1.e Compute Curvature and add it to Cur_list.(Eq.4.1) 
   	end (for loop)
-  2. For each  list computed in previous step do the follwing:   (Section )
+  2. For each  list computed in previous step do the follwing:   
   	 2.a Compute average of all points. 
@@
      end 
-\end{verbatim*}	
+\end{verbatim}	
 \item Ellipse Fitting: which is an attempt to fit as an ellipse as in  Section \ref{sec:EllipseDetection} 
-\begin{verbatim*}
-1. Initialize the
-
-\end{verbatim*}	
-\item Segmentation ALS1 as explained in Section \ref{sec:SwarmSegmentation}
-\begin{verbatim*}
-
-\end{verbatim*}	
+%he axes of the ellipse are estimated as the $width/2$ and $height/2$ of the stroke bounding box where $width$ is width and $height$ is height of the bonding box.
+\begin{verbatim}
+1. Initialize the axes of ellipse  and center by bounding box information
+2. Repeat the follwoing for few loops: 
+	2.a Compute error in Equation  4.2 
+	2.b Update center of ellipse and axis from error. 
+3. Compute percent in Equation 4.4 
+4. Computer eff in Equation 4.3 
+5. if eff < 0.2  then fit as ellipse otherwise move to segment. 
+\end{verbatim}
+
+\item Segmentation ALS1 as explained in Section \ref{sec:SwarmSegmentation}. The alogrithm will be similar to the PSO algoirhtm  in section \ref{sec:ParticleSwarmAlgorithm} 
+\begin{verbatim}
+Initialize 
+  1. Generate N particles with M dimensions (P_i...P_N) 
+  2. Generate velocities for each particle for M dimension randomly. 
+  3. Repeat till Maximum number of iteration OR ErrorThreshold  
+         For each particle P_i do the following 
+         3.a   Compute fitness  (fitness) 
+         				3.a.1 Compute the Error of approximation as in Eq.4.5
+         				3.a.2. Computer the fitness function as in Eq. 4.4 
+         3.b   Determine Lbest  
+               if (lbest_fitness<fitness)
+                  3.b.1   lbest=P_i
+         3.c      Determine Gbest.
+               if (lGest_fitness<fitness)
+                  3.c.1   lGest=P_i
+         3.d   Compute the Velocity for each dimension using Eq.3.2 
+         3.e   Bound Velocity using Eq.3.4
+         3.f   Update particle values using Eq. 3.3
+         3.g   Refine the location of the particle 
+         end .
+     end 
+4. return Gbest
+\end{verbatim}	
 \item Segmentation ALS2  as explained in Section \ref{sec:PolygonDivisionAlgorithm}
-\begin{verbatim*}
-
-\end{verbatim*}
-
+\begin{verbatim}
+Initialize 
+  1. Generate N particles with M dimensions (P_i...P_N) 
+  2. Generate velocities for each particle for M dimension randomly. 
+  3. Repeat till Maximum number of iteration OR ErrorThreshold  
+         For each particle P_i do the following 
+         3.a   Compute fitness  (fitness) 
+              3.a.1 For each Segment in the particle do the following
+         				    3.a.1.I   Error of approximation as Line (Eq. 4.7 - 4.9 )
+         			      3.a.1.II  Error of approximation as circular arc (Eq. 4.10 - 4.19 )
+         				    3.a.1.III Computer minimum of both approximation 
+         			3.a.2 Compute Error of the particle as in Eq.  4.20
+         			3.a.3 Compute fitness of the particle as in Eq. 4.21
+         3.b   Determine Lbest  
+               if (lbest_fitness<fitness)
+                  3.b.1   lbest=P_i
+         3.c      Determine Gbest.
+               if (lGest_fitness<fitness)
+                  3.c.1   lGest=P_i
+         3.d   Compute the Velocity for each dimension using Eq.3.2 
+         3.e   Bound Velocity using Eq.3.4
+         3.f   Update particle values using Eq. 3.3
+         3.g   Refine the location of the particle 
+         end .
+     end 
+4. return Gbest
+\end{verbatim}	
+\item Append best segmentation to List of segmentation ($L_s$) of this symbols. 
 \end{enumerate}
@@
 \begin{scriptsize}
-
 %\begin{table}
 	%\centering 
-
 %\scalebox{0.6}{		
@@
  \\ \hline
-Possible Dominate point Extraction & Computing and for details see Section\ref{sec:Preprocessing} &  Stroke points  & 1. compute   the pdp  &    Number of $P_{dp}$  = 11
+Possible Dominate point Extraction & Computing and for details see Section\ref{sec:Preprocessing} &  Stroke points  &   &    Number of $P_{dp}$  = 11
  Pdp = [ 2 (197.0 , 145.0 ), 4 (203.0 , 125.0 ), 3 (198.0 , 140.0 ), 22 (424.0 , 91.0 ), 19 (372.0 , 72.0 ), 26 (492.0 , 164.0 ), 28 (501.0 , 191.0 ), 9 (219.0 , 97.0 ), 31 (496.0 , 239.0 ), 0 (198.0 , 153.0 ), 41 (263.0 , 378.0 )]
@@
 
+  Segmentation & PSO Segmentation ALgS2 & $P_{dp}$  & ALgS2 &   Particles =  [  1 0 1 0 0 1 1 1 1 1 1 0 0 0 1 0 0 0 1 ] 
+ The particle  Particles =  [  1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 1 0 0 0 0 1 0 ]
+
+ Second segmentation PSO = [[[  S0 Line[P(201.0,179.0) to P(195.0,281.0)]  ,      S1 Line[P(195.0,281.0) to P(234.0,304.0)]  ,      S2 Line[P(234.0,304.0) to P(260.0,174.0)]  ,      S3 Line[P(260.0,174.0) to P(325.0,272.0)]  ,      S4 Line[P(325.0,272.0) to P(365.0,304.0)]  ,      S5 Line[P(365.0,304.0) to P(388.0,189.0)]  ,      S6 Line[P(388.0,189.0) to P(431.0,167.0)]  ,      S7 Line[P(431.0,167.0) to P(442.0,220.0)]  ]]]
+ \\ \hline
+
 		\end{longtable}
@@
 \label{sec:steprec}
+
+\begin{figure}
+	\centering
+		\includegraphics{afterDefense/segmentatiofinal.jpg.eps}
+	\caption{Best Segmantation }
+	\label{fig:segmentatiofinal.jpg}
+\end{figure}
+
 %\begin{enumerate}
@@
 
-%	\item Processing for the first Stroke 1.
-
-%	\item Possible Dominiate Point Extraction
-%	\item Segmentation 
-%	   Computing 
-%	\begin{enumerate}
-%	\item Ellipse detection  
-%	      Calcualting estimated ellipse with radius = $\frac{w}{2}$  and center as same as the bounding box of stroke.  
-%	      The ouput of the system $$
-% \end{enumerate}
-%	
-%	
-%\end{enumerate}
-%  
-%
-%\end{itemize}
+	The input of the step is the list of segmentation $L_s$ that is used to computer the feature vector.
+Figure \ref{fig:segmentatiofinal.jpg} shows the best segmentation as given by the previous step.
+This step can be summerized as firstly using $L_s$ to computer the feature vector and then introduce this vector the SVM classifier to classify it to one of the trained categories.

### 2009-08-17 draft/Results.tex

@@
 	\scalebox{0.7}{
-		\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|c|c|c|c|}\hline 
- Categories 	&ellipse	&heart	&trapezoid	&pentagon	&arch	&hexagon	&square	&triangle	&cube	&cylinder	&parallelogram	&moon	&callout\\ \hline 
-ellipse	&150	&20	&0	&0	&0	&5	&1	&0	&0	&0	&0	&1	&0\\  \hline 
-heart	&2	&130	&0	&1	&10	&1	&1	&0	&0	&0	&1	&4	&0\\  \hline 
-trapezoid	&0	&1	&147	&5	&1	&0	&1	&3	&2	&0	&37	&0	&0\\  \hline 
-pentagon	&0	&0	&1	&139	&0	&7	&0	&0	&0	&0	&1	&0	&0\\   \hline 
-arch	&0	&0	&1	&1	&125	&0	&4	&0	&3	&0	&0	&15	&1\\   \hline 
-hexagon	&0	&0	&0	&2	&0	&135	&0	&0	&0	&0	&0	&0	&0\\  \hline 
-square	&0	&0	&0	&0	&0	&0	&138	&0	&0	&0	&0	&1	&0\\   \hline 
-triangle	&0	&0	&1	&2	&5	&0	&1	&148	&0	&0	&4	&7	&0\\  \hline 
-cube	&0	&0	&0	&1	&0	&0	&0	&0	&146	&5	&0	&0	&0\\   \hline 
-cylinder	&0	&0	&0	&1	&0	&1	&0	&0	&0	&147	&0	&0	&0\\  \hline 
-parallelogram	&0	&0	&0	&0	&0	&0	&2	&0	&2	&0	&107	&0	&0\\  \hline 
-moon	&0	&1	&0	&0	&10	&2	&3	&0	&0	&0	&0	&125	&5\\   \hline 
-callout	&0	&0	&0	&0	&0	&0	&0	&0	&0	&0	&0	&0	&145\\  \hline 
+		\begin{tabular}{|p{1cm}|c|c|c|c|c|c|c|c|c|c|c|c||c|c|c|c|c|c|c|}\hline 
+		Categorie / classified as ==$>$	&ellipse	&heart	&trapezoid	&pentagon	&arch	&hexagon	&square	&triangle	&cube	&cylinder	&parallelogram	&moon	&callout	&Number of sample per class	&TP count	&FP count	&Fn	&Precision	&Recall ( TP rate )\\ \hline
+ellipse	&150	&20	&0	&0	&0	&5	&1	&0	&0	&0	&0	&1	&0	&152	&150	&27	&2	&0.84745762711864	&0.98684210526316\\ \hline 
+heart	&2	&130	&0	&1	&10	&1	&1	&0	&0	&0	&1	&4	&0	&152	&130	&20	&22	&0.86666666666667	&0.85526315789474\\ \hline 
+trapezoid	&0	&1	&147	&5	&1	&0	&1	&3	&2	&0	&37	&0	&0	&150	&147	&50	&3	&0.74619289340102	&0.98\\ \hline 
+pentagon	&0	&0	&1	&139	&0	&7	&0	&0	&0	&0	&1	&0	&0	&152	&139	&9	&13	&0.93918918918919	&0.91447368421053\\ \hline 
+arch	&0	&0	&1	&1	&125	&0	&4	&0	&3	&0	&0	&15	&1	&151	&125	&25	&26	&0.83333333333333	&0.82781456953642\\ \hline 
+hexagon	&0	&0	&0	&2	&0	&135	&0	&0	&0	&0	&0	&0	&0	&151	&135	&2	&16	&0.98540145985401	&0.89403973509934\\ \hline 
+square	&0	&0	&0	&0	&0	&0	&138	&0	&0	&0	&0	&1	&0	&151	&138	&1	&13	&0.99280575539568	&0.91390728476821\\ \hline 
+triangle	&0	&0	&1	&2	&5	&0	&1	&148	&0	&0	&4	&7	&0	&151	&148	&20	&3	&0.88095238095238	&0.98013245033113\\ \hline 
+cube	&0	&0	&0	&1	&0	&0	&0	&0	&146	&5	&0	&0	&0	&153	&146	&6	&7	&0.96052631578947	&0.95424836601307\\ \hline  
+cylinder	&0	&0	&0	&1	&0	&1	&0	&0	&0	&147	&0	&0	&0	&152	&147	&2	&5	&0.98657718120805	&0.96710526315789\\ \hline 
+parallelogram	&0	&0	&0	&0	&0	&0	&2	&0	&2	&0	&107	&0	&0	&150	&107	&4	&43	&0.96396396396396	&0.71333333333333\\  \hline 
+moon	&0	&1	&0	&0	&10	&2	&3	&0	&0	&0	&0	&125	&5	&153	&125	&21	&28	&0.85616438356164	&0.81699346405229\\ \hline  
+callout	&0	&0	&0	&0	&0	&0	&0	&0	&0	&0	&0	&0	&145	&151	&145	&0	&6	&1	&0.96026490066225\\  \hline  
+Total 	&152	&152	&150	&152	&151	&151	&151	&151	&153	&152	&150	&153	&151	&	&	&	&	&	&\\ \hline  
+% Categories 	&ellipse	&heart	&trapezoid	&pentagon	&arch	&hexagon	&square	&triangle	&cube	&cylinder	&parallelogram	&moon	&callout\\ \hline 
+%ellipse	&150	&20	&0	&0	&0	&5	&1	&0	&0	&0	&0	&1	&0\\  \hline 
+%heart	&2	&130	&0	&1	&10	&1	&1	&0	&0	&0	&1	&4	&0\\  \hline 
+%trapezoid	&0	&1	&147	&5	&1	&0	&1	&3	&2	&0	&37	&0	&0\\  \hline 
+%pentagon	&0	&0	&1	&139	&0	&7	&0	&0	&0	&0	&1	&0	&0\\   \hline 
+%arch	&0	&0	&1	&1	&125	&0	&4	&0	&3	&0	&0	&15	&1\\   \hline 
+%hexagon	&0	&0	&0	&2	&0	&135	&0	&0	&0	&0	&0	&0	&0\\  \hline 
+%square	&0	&0	&0	&0	&0	&0	&138	&0	&0	&0	&0	&1	&0\\   \hline 
+%triangle	&0	&0	&1	&2	&5	&0	&1	&148	&0	&0	&4	&7	&0\\  \hline 
+%cube	&0	&0	&0	&1	&0	&0	&0	&0	&146	&5	&0	&0	&0\\   \hline 
+%cylinder	&0	&0	&0	&1	&0	&1	&0	&0	&0	&147	&0	&0	&0\\  \hline 
+%parallelogram	&0	&0	&0	&0	&0	&0	&2	&0	&2	&0	&107	&0	&0\\  \hline 
+%moon	&0	&1	&0	&0	&10	&2	&3	&0	&0	&0	&0	&125	&5\\   \hline 
+%callout	&0	&0	&0	&0	&0	&0	&0	&0	&0	&0	&0	&0	&145\\  \hline 
  		\end{tabular}

### 2009-08-17 draft/Results.tex
after confusion matrix

@@
 		%	\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|c|c|c|c|}\hline 
-	\scalebox{0.7}{
-		\begin{tabular}{|p{1cm}|c|c|c|c|c|c|c|c|c|c|c|c||c|c|c|c|c|c|c|}\hline 
-		Categorie / classified as ==$>$	&ellipse	&heart	&trapezoid	&pentagon	&arch	&hexagon	&square	&triangle	&cube	&cylinder	&parallelogram	&moon	&callout	&Number of sample per class	&TP count	&FP count	&Fn	&Precision	&Recall ( TP rate )\\ \hline
-ellipse	&150	&20	&0	&0	&0	&5	&1	&0	&0	&0	&0	&1	&0	&152	&150	&27	&2	&0.84745762711864	&0.98684210526316\\ \hline 
-heart	&2	&130	&0	&1	&10	&1	&1	&0	&0	&0	&1	&4	&0	&152	&130	&20	&22	&0.86666666666667	&0.85526315789474\\ \hline 
-trapezoid	&0	&1	&147	&5	&1	&0	&1	&3	&2	&0	&37	&0	&0	&150	&147	&50	&3	&0.74619289340102	&0.98\\ \hline 
-pentagon	&0	&0	&1	&139	&0	&7	&0	&0	&0	&0	&1	&0	&0	&152	&139	&9	&13	&0.93918918918919	&0.91447368421053\\ \hline 
-arch	&0	&0	&1	&1	&125	&0	&4	&0	&3	&0	&0	&15	&1	&151	&125	&25	&26	&0.83333333333333	&0.82781456953642\\ \hline 
-hexagon	&0	&0	&0	&2	&0	&135	&0	&0	&0	&0	&0	&0	&0	&151	&135	&2	&16	&0.98540145985401	&0.89403973509934\\ \hline 
-square	&0	&0	&0	&0	&0	&0	&138	&0	&0	&0	&0	&1	&0	&151	&138	&1	&13	&0.99280575539568	&0.91390728476821\\ \hline 
-triangle	&0	&0	&1	&2	&5	&0	&1	&148	&0	&0	&4	&7	&0	&151	&148	&20	&3	&0.88095238095238	&0.98013245033113\\ \hline 
-cube	&0	&0	&0	&1	&0	&0	&0	&0	&146	&5	&0	&0	&0	&153	&146	&6	&7	&0.96052631578947	&0.95424836601307\\ \hline  
-cylinder	&0	&0	&0	&1	&0	&1	&0	&0	&0	&147	&0	&0	&0	&152	&147	&2	&5	&0.98657718120805	&0.96710526315789\\ \hline 
-parallelogram	&0	&0	&0	&0	&0	&0	&2	&0	&2	&0	&107	&0	&0	&150	&107	&4	&43	&0.96396396396396	&0.71333333333333\\  \hline 
-moon	&0	&1	&0	&0	&10	&2	&3	&0	&0	&0	&0	&125	&5	&153	&125	&21	&28	&0.85616438356164	&0.81699346405229\\ \hline  
-callout	&0	&0	&0	&0	&0	&0	&0	&0	&0	&0	&0	&0	&145	&151	&145	&0	&6	&1	&0.96026490066225\\  \hline  
-Total 	&152	&152	&150	&152	&151	&151	&151	&151	&153	&152	&150	&153	&151	&	&	&	&	&	&\\ \hline  
+	\scalebox{0.5}{
+		\begin{tabular}{|p{2cm}|c|c|c|c|c|c|c|c|c|c|c|c|c||p{2cm}|c|c|c|c|p{2cm}|}\hline 
+		Categorie / classified as ==$>$	&ellipse	&heart	&trapezoid	&pentagon	&arch	&hexagon	&square	&triangle	&cube	&cylinder	&parallelogram	&moon	&callout	&Number of samples per class	&TP count	&FP count	&Fn	&Precision	&Recall ( TP rate )\\ \hline
+ellipse	&150	&20	&0	&0	&0	&5	&1	&0	&0	&0	&0	&1	&0	&152	&150	&27	&2	&0.8474	&0.9868\\ \hline 
+heart	&2	&130	&0	&1	&10	&1	&1	&0	&0	&0	&1	&4	&0	&152	&130	&20	&22	&0.86666	&0.8552\\ \hline 
+trapezoid	&0	&1	&147	&5	&1	&0	&1	&3	&2	&0	&37	&0	&0	&150	&147	&50	&3	&0.7461	&0.98\\ \hline 
+pentagon	&0	&0	&1	&139	&0	&7	&0	&0	&0	&0	&1	&0	&0	&152	&139	&9	&13	&0.9391	&0.91447\\ \hline 
+arch	&0	&0	&1	&1	&125	&0	&4	&0	&3	&0	&0	&15	&1	&151	&125	&25	&26	&0.83333	&0.82781\\ \hline 
+hexagon	&0	&0	&0	&2	&0	&135	&0	&0	&0	&0	&0	&0	&0	&151	&135	&2	&16	&0.9854	&0.89403\\ \hline 
+square	&0	&0	&0	&0	&0	&0	&138	&0	&0	&0	&0	&1	&0	&151	&138	&1	&13	&0.9928	&0.91390\\ \hline 
+triangle	&0	&0	&1	&2	&5	&0	&1	&148	&0	&0	&4	&7	&0	&151	&148	&20	&3	&0.88095	&0.98013\\ \hline 
+cube	&0	&0	&0	&1	&0	&0	&0	&0	&146	&5	&0	&0	&0	&153	&146	&6	&7	&0.96052	&0.95424\\ \hline  
+cylinder	&0	&0	&0	&1	&0	&1	&0	&0	&0	&147	&0	&0	&0	&152	&147	&2	&5	&0.98657	&0.967105\\ \hline 
+parallelogram	&0	&0	&0	&0	&0	&0	&2	&0	&2	&0	&107	&0	&0	&150	&107	&4	&43	&0.963	&0.71333\\  \hline 
+moon	&0	&1	&0	&0	&10	&2	&3	&0	&0	&0	&0	&125	&5	&153	&125	&21	&28	&0.8561	&0.8169\\ \hline  
+callout	&0	&0	&0	&0	&0	&0	&0	&0	&0	&0	&0	&0	&145	&151	&145	&0	&6	&1	&0.9602\\  \hline  
 % Categories 	&ellipse	&heart	&trapezoid	&pentagon	&arch	&hexagon	&square	&triangle	&cube	&cylinder	&parallelogram	&moon	&callout\\ \hline 

### 2009-08-17 draft/System.tex
after most changes

@@
 \subsection{Structural and geometrical Features(FS1)}
-Features defines the structure of the geometrical symbol.  
+Features defines the structure of the geometrical symbol. Section \ref{sec:steprec} descripe in details how these feature are computed. 
 %\begin{description}
@@
 This step can be summerized as firstly using $L_s$ to computer the feature vector and then introduce this vector the SVM classifier to classify it to one of the trained categories.
+The details of each step of feature extraction is detailed in the next algorithm.
+Feature extraction step is divided into four steps where each feature set is computed seperatally then all sets are appended to construct a final feature vector.
+
+\begin{enumerate}
+	\item Computing Structural and geometrical Features(FS1).
+This step compute number of parallel, perpendicular line and similar structural features.
+The follwoing algorithm is used to determine these features.
+	\begin{verbatim}
+Initialize 
+  1. For each segment Si in list Ls  do the follwoing. 
+      1.a If segment si is a line do the follwing 
+            1.a.I. Compute the slope and intercept of line si. 
+            1.a.II Increment Parallel count  if 
+                  parallel with each other line in Ls
+            1.a.III Increment PerpendicularCount if 
+                    perpendicular with each other line in Ls 
+            1.a.IV Increment Intersection Count if 
+                    intersected with any other line. 
+                    1.a.II.A  if Intersection found determine 
+                              type of intersection or (X,T,L)
+      1.b If segment si is curve do the follwing. 
+          1.b.I Increment Intersect if
+               Intersect other line or curve
+          1.b.II Increment number 
+                 of curves and primitives. 
+4. return Feature_Vector_FS1
+\end{verbatim}	
+	\item Rubine Feature Set (FS2).  Compute Rubine features as in \cite{gestureexample12} 
+	\item Statistical Features (FS3) compute the moments as in \cite{zernike61}
+	\item Global shape properties set (FS4). Firslty compute the convex hull of the stroke points then compute the rest of the features.
+
+	\item Introduce the computed feature vector to the classifier and check the label of the detected symbol. 
+\end{enumerate}
+
+ \begin{longtable}{|p{2cm}|p{2cm}|p{2cm}|p{2cm}|p{13cm}|}
+\caption{Detailed Output of System in Each Step of Recognition}
+\label{tab:StepsStroke1} \\
+
+\hline 
+\multicolumn{1}{|p{2cm}|}{\textbf{Step}} & 
+\multicolumn{1}{p{2cm}|}{\textbf{Description}} &
+\multicolumn{1}{p{2cm}|}{\textbf{Inputs}} &
+\multicolumn{1}{p{2cm}|}{\textbf{Details}} &
+\multicolumn{1}{p{13cm}|}{\textbf{Output}} 
+\\ \hline 
+\endfirsthead
+
+\hline
+\multicolumn{3}{c}%
+{{\bfseries \tablename\ \thetable{} -- continued from previous page}} \\
+\multicolumn{1}{|p{2cm}|}{\textbf{Step}} & 
+\multicolumn{1}{p{2cm}|}{\textbf{Description}} &
+ \multicolumn{1}{p{2cm}|}{\textbf{Inputs}} &
+ \multicolumn{1}{p{2cm}|}{\textbf{Details}} &
+ \multicolumn{1}{p{13cm}|}{\textbf{Output}} 
+\\ \hline 
+\endhead
+
+%Step & Discription of Step & Input & Steps & Output \\ \hline
+Stroke 1 & Extracting point and computing boundary data & -- &  see Section \ref{sec:Preprocessing} & 
+\begin{scriptsize}
+
+\end{scriptsize}
+ \\ \hline
+Possible Dominate point Extraction & Computing and for details see Section\ref{sec:Preprocessing} &  Stroke points  &   &
+
+\\ \hline 
+Segmentation & Segmentation of Ellipse & $P_{dp}$   &  Ellipse fitting  & 
+ \\ \hline
+
+Stroke 2 & Extracting point and computing boundary data & -- &  see Section \ref{sec:Preprocessing} &
+\begin{scriptsize}
+  \end{scriptsize}
+  \\ \hline
+  & Possible Dominate point Extraction for details see Section\ref{sec:Preprocessing} &  Stroke points  & 1. compute the $P_{pd}$  & \begin{scriptsize}   Number of $P_{pd}$  =  36
+
+  \end{scriptsize}
+    \\ \hline
+ \\ \hline 
+Segmentation & Segmentation of Ellipse & $P_{dp}$   &  Ellipse fitting  &
+
+ \\ \hline
+ Segmentation & PSO Segmentation ALgS1 & $P_{dp}$  & ALgS1 &   
+ \\ \hline
+
+  Segmentation & PSO Segmentation ALgS2 & $P_{dp}$  & ALgS2 &    
+ \\ \hline
+
+		\end{longtable}
+%}
+
+%\end{table}
+
+\end{scriptsize}
+
+\end{landscape}

### 2009-08-17 trunk/draft/Results.tex
after most changes

@@
 		%	\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|c|c|c|c|}\hline 
-	\scalebox{0.7}{
-		\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|}\hline 
-		Categorie / classified as ==$>$	&ellipse	&heart	&trapezoid	&pentagon	&arch	&hexagon	&square	&triangle	&cube	&cylinder	&parallelogram	&moon	&callout	&Number of sample per class	&TP count	&FP count	&Fn	&Precision	&Recall ( TP rate )\\
-ellipse	&150	&20	&0	&0	&0	&5	&1	&0	&0	&0	&0	&1	&0	&152	&150	&27	&2	&0.84745762711864	&0.98684210526316\\
-heart	&2	&130	&0	&1	&10	&1	&1	&0	&0	&0	&1	&4	&0	&152	&130	&20	&22	&0.86666666666667	&0.85526315789474\\
-trapezoid	&0	&1	&147	&5	&1	&0	&1	&3	&2	&0	&37	&0	&0	&150	&147	&50	&3	&0.74619289340102	&0.98\\
-pentagon	&0	&0	&1	&139	&0	&7	&0	&0	&0	&0	&1	&0	&0	&152	&139	&9	&13	&0.93918918918919	&0.91447368421053\\
-arch	&0	&0	&1	&1	&125	&0	&4	&0	&3	&0	&0	&15	&1	&151	&125	&25	&26	&0.83333333333333	&0.82781456953642\\
-hexagon	&0	&0	&0	&2	&0	&135	&0	&0	&0	&0	&0	&0	&0	&151	&135	&2	&16	&0.98540145985401	&0.89403973509934\\
-square	&0	&0	&0	&0	&0	&0	&138	&0	&0	&0	&0	&1	&0	&151	&138	&1	&13	&0.99280575539568	&0.91390728476821\\
-triangle	&0	&0	&1	&2	&5	&0	&1	&148	&0	&0	&4	&7	&0	&151	&148	&20	&3	&0.88095238095238	&0.98013245033113\\
-cube	&0	&0	&0	&1	&0	&0	&0	&0	&146	&5	&0	&0	&0	&153	&146	&6	&7	&0.96052631578947	&0.95424836601307\\
-cylinder	&0	&0	&0	&1	&0	&1	&0	&0	&0	&147	&0	&0	&0	&152	&147	&2	&5	&0.98657718120805	&0.96710526315789\\
-parallelogram	&0	&0	&0	&0	&0	&0	&2	&0	&2	&0	&107	&0	&0	&150	&107	&4	&43	&0.96396396396396	&0.71333333333333\\
-moon	&0	&1	&0	&0	&10	&2	&3	&0	&0	&0	&0	&125	&5	&153	&125	&21	&28	&0.85616438356164	&0.81699346405229\\
-callout	&0	&0	&0	&0	&0	&0	&0	&0	&0	&0	&0	&0	&145	&151	&145	&0	&6	&1	&0.96026490066225\\
-Total 	&152	&152	&150	&152	&151	&151	&151	&151	&153	&152	&150	&153	&151	&	&	&	&	&	&\\ 
+	\scalebox{0.5}{
+		\begin{tabular}{|p{2cm}|c|c|c|c|c|c|c|c|c|c|c|c|c||p{2cm}|c|c|c|c|p{2cm}|}\hline 
+		Categorie / classified as ==$>$	&ellipse	&heart	&trapezoid	&pentagon	&arch	&hexagon	&square	&triangle	&cube	&cylinder	&parallelogram	&moon	&callout	&Number of samples per class	&TP count	&FP count	&Fn	&Precision	&Recall ( TP rate )\\ \hline
+ellipse	&150	&20	&0	&0	&0	&5	&1	&0	&0	&0	&0	&1	&0	&152	&150	&27	&2	&0.8474	&0.9868\\ \hline 
+heart	&2	&130	&0	&1	&10	&1	&1	&0	&0	&0	&1	&4	&0	&152	&130	&20	&22	&0.86666	&0.8552\\ \hline 
+trapezoid	&0	&1	&147	&5	&1	&0	&1	&3	&2	&0	&37	&0	&0	&150	&147	&50	&3	&0.7461	&0.98\\ \hline 
+pentagon	&0	&0	&1	&139	&0	&7	&0	&0	&0	&0	&1	&0	&0	&152	&139	&9	&13	&0.9391	&0.91447\\ \hline 
+arch	&0	&0	&1	&1	&125	&0	&4	&0	&3	&0	&0	&15	&1	&151	&125	&25	&26	&0.83333	&0.82781\\ \hline 
+hexagon	&0	&0	&0	&2	&0	&135	&0	&0	&0	&0	&0	&0	&0	&151	&135	&2	&16	&0.9854	&0.89403\\ \hline 
+square	&0	&0	&0	&0	&0	&0	&138	&0	&0	&0	&0	&1	&0	&151	&138	&1	&13	&0.9928	&0.91390\\ \hline 
+triangle	&0	&0	&1	&2	&5	&0	&1	&148	&0	&0	&4	&7	&0	&151	&148	&20	&3	&0.88095	&0.98013\\ \hline 
+cube	&0	&0	&0	&1	&0	&0	&0	&0	&146	&5	&0	&0	&0	&153	&146	&6	&7	&0.96052	&0.95424\\ \hline  
+cylinder	&0	&0	&0	&1	&0	&1	&0	&0	&0	&147	&0	&0	&0	&152	&147	&2	&5	&0.98657	&0.967105\\ \hline 
+parallelogram	&0	&0	&0	&0	&0	&0	&2	&0	&2	&0	&107	&0	&0	&150	&107	&4	&43	&0.963	&0.71333\\  \hline 
+moon	&0	&1	&0	&0	&10	&2	&3	&0	&0	&0	&0	&125	&5	&153	&125	&21	&28	&0.8561	&0.8169\\ \hline  
+callout	&0	&0	&0	&0	&0	&0	&0	&0	&0	&0	&0	&0	&145	&151	&145	&0	&6	&1	&0.9602\\  \hline  
 % Categories 	&ellipse	&heart	&trapezoid	&pentagon	&arch	&hexagon	&square	&triangle	&cube	&cylinder	&parallelogram	&moon	&callout\\ \hline 

### 2009-08-18 trunk/draft/ChapterIntroduction.tex
After all edit

@@
 
-Sketch recognition is divided into three steps: preprocessing, segmentation and symbol recognition.
+Similar to most recognition system, sketch recognition is divided into three steps: preprocessing, segmentation and symbol recognition.
 The preprocessing phase captures user input stroke points and collects basic information about the stroke then proceeds to remove noise and compute basic statistical and geometrical information.
@@
 The problem could be solved by providing context or domain information which could not be provided in most sketches.
+
 \begin{figure}	
 	\centering
-		\includegraphics[scale=0.65]{../../neededfiles/Figures/VorU.eps}
-	\caption[Handwriting Ambiguities] { The same letter can be interpreted differently in various contexts; first as V in drive or as U in Cut. This example is adapted from \cite{Ambiguity}.}
+		\includegraphics[scale=0.65]{afterDefense/Amb.jpg.eps}
+	\caption[Handwriting Ambiguities] {The same handwriting can be interpreted differently in various contexts; In the first line the b can be viewed as a B and in the second one could see it as 8.}
 \label{fig:VorU}

### 2009-08-18 trunk/draft/System.tex
After all edit

@@
 
- As shown in Figure \ref{fig:Blockdiagram} first an attempt is made to fit the stroke points into a curve or an ellipse using a minimum square error fitting algorithm \cite{chernov-2003}.
+ As shown in Figure \ref{fig:Blockdiagram} first an attempt is made to fit the stroke points into a curve or an ellipse using a minimum square error fitting algorithm \cite{chernov}.
 If the stroke proved to be an elliptical arc then the segmentation process ends and the system proceeds to the next step.
 Otherwise, the stroke is passed to two particle swarm algorithms that divide the stroke to either a polygon or a set of lines and curves.
-The algorithms takes the stroke points along with the possible dominant points $P_{pd}$ computed during preprocessing then produce a set of dominant points which are connected by either lines or curves.
+The algorithms takes the stroke points along with the possible dominant points $P_{pd}$ computed during Possible Dominant Point Extraction then produce a set of dominant points which are connected by either lines or curves.
 The two algorithms will generate two segmentations, the system will choose the segmentation that has the minimum segmentation error.%
+
 % First an attempts the segmentation process is divided into two steps ellipse fitting and curve segmentation.
@@
 %Some data is extracted from the points in the stroke.
-The preprocessing step captures the points from the pointing devices then compute the information needed to determine a set of  \textit{Possible Dominant Points $P_{pd}$} which guides the segmentation step.
+The Possible Dominant Point Extraction step captures the points from the pointing devices then compute the information needed to determine a set of  \textit{Possible Dominant Points $P_{pd}$} which guides the segmentation step.
 It is noted as a person draws a shape the pen slows down near corners and picks up speed when drawing straight lines.
@@
 The segmentation tries to divide the stroke into a set of primitives.
-As shown in Figure \ref{fig:Blockdiagram} first an attempt is made to fit the stroke points into a curve or an ellipse using a minimum square error fitting algorithm \cite{chernov-2003}.
+As shown in Figure \ref{fig:Blockdiagram} first an attempt is made to fit the stroke points into a curve or an ellipse using a minimum square error fitting algorithm \cite{chernov}.
 If the stroke proved to be an ellipse arc then the segmentation process ends and the system proceeds to the next step.
 Otherwise, the stroke is passed to two particle swarm algorithms that divide the stroke to either lines or lines and curves.
-The algorithms take the stroke points along with the possible dominant points $P_{pd}$ computed during preprocessing then produce a set of dominant points which are connected by either lines or curves.
+The algorithms take the stroke points along with the possible dominant points $P_{pd}$ computed during Possible Dominant Point Extraction then produce a set of dominant points which are connected by either lines or curves.
 The next section describes the ellipse detection algorithm and the two particle swarm algorithms used to divide the stroke.
@@
 
-After the best fit segmentaion is choosen the paramter of each primitive is computed.
-For example, if the stroke is represented as an ellispe, the center and axii of the ellipse is computed and confirmed.
+After the best fit segmentation is chosen the parameter of each primitive is computed.
+For example, if the stroke is represented as an ellipse, the center and axii of the ellipse is computed and confirmed.
 For lines the slope, equation, start point and end point are computed.
-These values and paramter will help determine the spatial relations between different segments in the feature extraction step.
+These values and parameter will help determine the spatial relations between different segments in the feature extraction step.
 % The segmentation which crossponds to the minimum $E_{tot}$ is the best segmentation.  
@@
 A segment can be either an ellipse, a line or a circular arc.
-The list of segments $L_s$ represents the segmentaion of the storke.
-When a user draw another stroke the generated segmentaion of the storke is appended to the list of segments.
+The list of segments $L_s$ represents the segmentation of the stroke.
+When a user draw another stroke the generated segmentation of the stroke is appended to the list of segments.
 When the user finish drawing the symbol the segment list $L_s$ will contain all segments of the drawn symbol.
 The relation and properties between different segments are extracted from the segment list $L_s$.
-This step is considered as feature extration which is dircribed in details in the next sections.
+This step is considered as feature extraction which is described in details in the next sections.
 
@@
 We used the library implementation of LibSVM \cite{libsvm} and its implementation of the one-vs-one strategy.
-Section \ref{Sec:SVMdetail} discribe in details the classifier, how classes (classification) are determined and the reasons behind using this specific classification methods.
-
-\subsection {Trainig the SVM}
-\subsection {Testing}
+Section \ref{Sec:SVMdetail} describe in details the classifier, how classes (classification) are determined and the reasons behind using this specific classification methods.
+
+\subsection {Training and Testing}
+A training set that contains positive samples of each category in the data set is used to train the SVM classifier.
+The training starts after computing the feature vector of each sample.
+The system trains $ \frac{n(n-1)}{{2}}$ binary classifiers each distinguish between only two categories.
+The goal of the training process is to generate a linear SVM plane that maximize margin between the two categories.
+A cross validation procedure is implemented to choose the best value for the $C$ parameter in SVM classifier.( $C=100$ was the best value)
+=
+For classification the decision is made from producing the feature vector of the unknown symbol to the trained $\frac{n(n-1)}{{2}}$ binary classifiers.
+The final decision is given using simple voting strategy.
 
@@
 The foundation of Support Vector Machines (SVM) were developed by Vapnik \cite{svmintroduce} in 1995.
-It is gainig increase popularity due to the high performance and efficient implementation.
-The goal of SVM algorithms is to generate a sperating plane between two different classes which are represented using a set of n-dimentional vectors.
-The generated plane is used as a good classifier to the unseen exmaples.
+It is gaining increase popularity due to the high performance and efficient implementation.
+The goal of SVM algorithms is to generate a separating plane between two different classes which are represented using a set of n-dimensional vectors.
+The generated plane is used as a good classifier to the unseen examples.
 However, there can be more than one plane that can separates the data therefore SVM attempts to find the plane that maximize the margin (distance between classifier and nearest sample) between the two classes \ref{fig:svm1}.
-To calcuate the margin, two different hyperplanes are constructed, one plane on each side of the data, these two planes are then moved to generenate the maximum margin possilbe.
+To calculate the margin, two different hyper planes are constructed, one plane on each side of the data, these two planes are then moved to generate the maximum margin possible.
 \begin{figure}
@@
 
-Let us consider  a two class classification task with datapoints $x_i$ where $i=1,\cdots,m$ having corresponding labels $y_i=\pm1$ where if  $y_i=1$ if $x_i$ belonged to first class and $y_i=-1$ if $x_i$ belonged to second class.
-Each data  point $x_i$ is represented in $N$ dimensional input or attribule space.
+Let us consider a two class classification task with data points $x_i$ where $i=1,\cdots,m$ having corresponding labels $y_i=\pm1$ where if  $y_i=1$ if $x_i$ belonged to first class and $y_i=-1$ if $x_i$ belonged to second class.
+Each data point $x_i$ is represented in $N$ dimensional input or attribute space.
 Let the classification plane be represented as function.
@@
 
- The orientation of the speration plane is deteremined using the vector $\omega$ and the offest of the plane from the orgin is determined using the scalar $b$.
-Lets make an assumption that the datatpoint are linearly seperable, i.e.
-there are infinite possibilities to find a plane that seperates and correctly classify the training data.
-This is idea is illustrated in Figure\ref{fig:svm1} where two different seperating planes can correctly classify the data.
+ The orientation of the separation plane is determined using the vector $\omega$ and the offset of the plane from the origin is determined using the scalar $b$.
+Let's make an assumption that the data point are linearly separable, i.e.
+there are infinite possibilities to find a plane that separates and correctly classify the training data.
+This is idea is illustrated in Figure\ref{fig:svm1} where two different separating planes can correctly classify the data.
 It is intuitively noted from the figure that the solid line is more likely to generalize more data.
@@
 % Which one is preferable ?
-Intuitively one perfers the solid pane since small perturbations of any point would not introduce misclassifications errors.
+Intuitively one prefers the solid pane since small perturbations of any point would not introduce misclassifications errors.
 Without any information, the solid plane is more likely to generalize better on future data.
-Geometrically we can characterize the solid plane as being furthuret from both classes.
-plane supports a class if all points in that class are on one side of that plane.
-
-The main problem of SVM is constructing the plane that is furthuest from both classes.
+Geometrically we can characterize the solid plane as being furtherer from both classes.
+Plane supports a class if all points in that class are on one side of that plane.
+
+The main problem of SVM is constructing the plane that is furthest from both classes.
 The main approach is to use two parallel supporting planes and maximize the margin between them.
@@
 
-The Support vectors that are pumped into the supprting plane deteremine the margin.
+The Support vectors that are pumped into the supporting plane determine the margin.
 This is a primal form Quadratic Programming (QP) problem.
 The concept of duality can also used to wields the problem to a Largrnigian dual QP problem in Equation \ref{eq:dual}.
-% maximumm margin between parallel 
+% maximum margin between parallel 
 \begin{equation}
@@
 
-Both the dual form and the primal form weilds to the same plane $\omega=\sum_{i=1}^{m}y_i\alpha_i x_i$ and threshold $b$ determined by the support vectors (for which $\alpha_i\geq 0$).
-
-\subsubsection{Theoretcical foundation}
-
-The generalization error on future points and not only in the training set can be obtained as proved by the statical learning theory \cite{BennettSVMP2000}.
-The generalized error bound is a function of the misclassified error on the training data and the terms that measure the complexiy or capacity of the classification.
-In linear functions, using planes to maximize the margin of seperation reduces the complexiy of the functions.
-Thus, by explicity maximizing the margin we are in the same time  minimizing bound on the generalization error and therefore expect better generalization with high probability.
-The dimensionality of the data dose not affect the size of the margin which   results in good performance  for very high dimensional data (i.e, with a very large number of attribute).
+Both the dual form and the primal form wields to the same plane $\omega=\sum_{i=1}^{m}y_i\alpha_i x_i$ and threshold $b$ determined by the support vectors (for which $\alpha_i\geq 0$).
+
+\subsubsection{Theoretical foundation}
+
+The generalization error on future points and not only in the training set can be obtained as proved by the statistical learning theory \cite{BennettSVMP2000}.
+The generalized error bound is a function of the misclassified error on the training data and the terms that measure the complexity or capacity of the classification.
+In linear functions, using planes to maximize the margin of separation reduces the complexity of the functions.
+Thus, by explicitly maximizing the margin we are in the same time minimizing bound on the generalization error and therefore expect better generalization with high probability.
+The dimensionality of the data does not affect the size of the margin which   results in good performance for very high dimensional data (i.e, with a very large number of attribute).
 In a sense, this also reduces the problems caused by over fitting of high - dimensional data.
-A large volume of literature  , e.g., \cite{bookSVMoverfit11,statisticalSVM1} refers to this problem,  for more technical discussion of the   statistical theroy.
-
-%Using geomteric arguments we can gain in sights to some of these results.
-Classification function that have capcity to fit the training data are more likely to overfit  resulting in poor generalization.
-
-\subsubsection{Linearly Inseperable Case}
-In all previous sections we assumed that the problem of classifiying two set is linearly separable.
-The assumption is not true in most casses, thus, trying to construct a plane to seperate the two dataset will fail.
-Figure \ref{fig:svm5} illustrates that we cannot construct a plane that can seperate the two sets correctly.
-The constrains for constructing the planes must be relaxed since the QP task is not feasible for the linearly inseperable case.
-Consider the linear inseperable problem in Figure\ref{fig:svm5}, in the ideal case there will be no point missclassified and no point falls on the margin.
-But we must relax the constrain that insure that each point in on the approiate side of tis supporitng plane.
+A large volume of literature, e.g., \cite{bookSVMoverfit11,statisticalSVM1} refers to this problem,  for more technical discussion of the   statistical theory.
+
+%Using geometric arguments we can gain insights to some of these results.
+Classification function that have capacity to fit the training data are more likely to over fit resulting in poor generalization.
+
+\subsubsection{Linearly Inseparable Case}
+In all previous sections we assumed that the problem of classifying two set is linearly separable.
+The assumption is not true in most cases, thus, trying to construct a plane to separate the two dataset will fail.
+Figure \ref{fig:svm5} illustrates that we cannot construct a plane that can separate the two sets correctly.
+The constrains for constructing the planes must be relaxed since the QP task is not feasible for the linearly inseparable case.
+Consider the linear inseparable problem in Figure\ref{fig:svm5}, in the ideal case there will be no point misclassified and no point falls on the margin.
+But we must relax the constrain that insure that each point in on the appropriate side of it is supporting plane.
 Every point that falls on the wrong side of the supporting plane is considered an error.
@@
 		\includegraphics{afterDefense/svm5.jpg.eps}
-	\caption{Linearly Inseperable}
+	\caption{Linearly Inseparable}
 	\label{fig:svm5}
@@
 A minor change in the supporting plane QP problem can accomplish this.
-A nonegative slack or error variable $z_i$ is added to each constraint.
+A non negative slack or error variable $z_i$ is added to each constraint.
 A weighted penalty term is also added to the objective as follows:
@@
 \subsubsection {Non linear function Using Kernels}
-Figure \ref{fig:svm6} shows a classification problem that cannot be seperated using linear space.
-A higher order plane can work better in similar casses.
-Converting a linear classification algorithm into non linear classification algoirhtm  can be achieved by adding addional attribute to the data that are nonlinear functions of the original data.
-The expanded dataset features space can be seperated using an  existing linear classification algorithm.
-The  linear classication that is applied on the feautre space produce a non linear classification on the orginal inputspace.
-In two dimensional vector space with attributes $r$ and $s$, to construct a quadratic  discriminant  simply map the original two dimensional input space $\left[r,s\right]$ to  five dimensional features space $\left[r,s,rs,r^2,s^2\right]$ and construct a linear discriminiant in that space.
+Figure \ref{fig:svm6} shows a classification problem that cannot be separated using linear space.
+A higher order plane can work better in similar cases.
+Converting a linear classification algorithm into non linear classification algorithm can be achieved by adding additional attribute to the data that are nonlinear functions of the original data.
+The expanded dataset features space can be separated using an existing linear classification algorithm.
+The linear classification that is applied on the feature space produces a non linear classification on the original input space.
+In two dimensional vector space with attributes $r$ and $s$, to construct a quadratic  discriminate  simply map the original two dimensional input space $\left[r,s\right]$ to  five dimensional features space $\left[r,s,rs,r^2,s^2\right]$ and construct a linear discriminate in that space.
 specifically, define : $\vartheta(x):R^2  \rightarrow R^5$ then
@@
 
-For higher dimentional datasets, this method of nonlinear mapping  has two potential problems stemming from the fact that dimensionality of the feature space explones exponentially.
-The first problem is that overfitting becomes a problem.
-However, since SVM rely on margin maximization they are largely immune to this porblem.
-Provided an appropriate value of paramter $C$ is chosen.
-The second concern is that it is the computaional complexity of   computing $\vartheta(x)$ but SVM use kernels to get around this problem.
-
-Examine what happens when athe nonlinear mapping is introduced to equation \ref{eq:dual}. Let us define : $\vartheta(x):R^2\rightarrow R^5 \quad n>> n$  we need to optimize
+For higher dimensional datasets, this method of nonlinear mapping has two potential problems stemming from the fact that dimensionality of the feature space explodes exponentially.
+The first problem is that over fitting becomes a problem.
+However, since SVM rely on margin maximization they are largely immune to this problem.
+Provided an appropriate value of parameter $C$ is chosen.
+The second concern is that it is the computational complexity of   computing $\vartheta(x)$ but SVM use kernels to get around this problem.
+
+Examine what happens when  the nonlinear mapping is introduced to equation \ref{eq:dual}. Let us define: $\vartheta(x):R^2\rightarrow R^5 \quad n>> n$  we need to optimize
 
@@
 Notice that the mapped data only occurs as inner product in the objective.
-Now we apply a little mathematically rigorous magie known as hilbern schmidt kernels, first applied to SVMs in \cite{svmintroduce}.
-By Mercer's theorem, we knoew that for ccertain mapping $\theta$ and any two points $u$ and $v$, the inner product of the the mapped points can be evalutated using the kernel function without ever explicity knowing ht emapping , e.g., $\theta(u)\theta(v)\equiv K(u,v)$.
+Now we apply a little mathematically magic known as hilbern schmidt kernels, first applied to SVMs in \cite{svmintroduce}.
+By Mercer's theorem, we know that for certain mapping $\theta$ and any two points $u$ and $v$, the inner product of the mapped points can be evaluated using the kernel function without ever explicitly knowing the mapping, e.g., $\theta(u)\theta(v)\equiv K(u,v)$.
 Some of the more popular known kernels are given below.
-New kernel are being developed to fit domian specific requriments.
+New kernels are being developed to fit domain specific requirements.
 \begin{center}
@@
 	Two layer Neural Network& $ sigmoid \left(\eta\left(u\cdot v +1\right)+c\right)$ \\ 
-
 	\end{tabular}
 \end{center}
-To change from a linear to nonlinear classifier, one must only substitute a kernel evaluation in the objective insted of the original dot product.
-Thus by changing kernels we can get different hightly nonlinear classifiers.
-No algorithmic changes are required from linear clase other than substitue of a kernel evaluation for simple dot product.
-All the benefits fo the original linear SVM methods are maintained.
-We can train a highly nonlinear classification function such as a ploynomial or a raadial basis function machine, or a sigmoidal neural network using robust, efficient algorithsm that have no problem with local minima.
-By using kernel substitution a linear algorithm (only capable of handling seperable data) can be turned into a general nonlinear algorithm.
+To change from a linear to nonlinear classifier, one must only substitute a kernel evaluation in the objective instead of the original dot product.
+Thus by changing kernels we can get different highly nonlinear classifiers.
+No algorithmic changes are required from linear case other than substitute of a kernel evaluation for simple dot product.
+All the benefits of the original linear SVM methods are maintained.
+We can train a highly nonlinear classification function such as a polynomial or a radial basis function machine, or a sigmoid neural network using robust, efficient algorithms that have no problem with local minima.
+By using kernel substitution a linear algorithm (only capable of handling separable data) can be turned into a general nonlinear algorithm.
 %\begin{equation}
@@
 %The foundations of Support Vector Machines (SVM) have
-\subsubsection{Classification}
+\subsubsection{SVM  Classification}
 The resulting SVM method can be summarized as follows 
 \begin{enumerate}
-	\item Select paramter C representign the tradeoff between minimizign the training set error and maximizing the margin.
-Select the kernel function and any kernel paramters.
-For example for the radial basis function kernel, one must select width of the gaussian $\gamma$.
-	\item Solve dual QP Equation \ref{} or an alterantive SVM formulation using an approriate quadratic programming or linear programming alogrithm. 
-	\item Recover th primal threshold variable $b$ using the support vectors. 
+	\item Selecting the SVM classifier Paramters. Parameter $C$ representing the tradeoff between minimizing the training set error and maximizing the margin. 
+	\item Select A kernel function and their kernel parameters. For example for the radial basis function kernel, one must select width of the Gaussian $\gamma$. 
+	\item Solve dual QP Equation \ref{eq:dualkernel} or an alternative SVM formulation using an appropriate quadratic programming or linear programming algorithm. 
+	\item Get the support vector and recover the emprial threshold  variable $b$. 
 	\item Classify new points using $f(x)=sign\left(\sum{i}y_i\alpha_i K\left(x,x_i\right)-b\right)$
 \end{enumerate}
-
-The paramters in step 1. are selected using cross valiation if sufficent data are avaliable. %However, recent model  selectin strategies can
+The parameters in step 1. are selected using cross validation if sufficient data are available. %However, recent model  selectin strategies can 
+\subsubsection {Multi Class Classifiers}
+
+ There is mainly two popular methods to change the binary SVM classification problem into Multiclass classification problem.
+The first methods is One Versus All (OVA) Strategy where M binary classifiers is constructed one classifier for each class.
+The $ith$ classifier is trained using positive example form $ith$ class and negative example from all other classes.
+The output function $p_i$ represents the distance and classification of the example from the $ith$ classifier.
+For a new example $x$, OVA SVM strategy assigns it to the class with the largest value of $p_i$.
+
+ The second method is One Versus One Strategy where a binary classifier for every pair of distinct classes ($M$ classes) and so, all together $\frac{M(M \cdot 1)}{2}$ binary classifiers are constructed.
+Each binary classifier is trained using examples from both $p_i$ and $p_j$.
+For example, classifier $C_{ij}$ is trained taking the examples from $p_i$ as positive and the examples from $p_j$ as negative.
+The classification is based on the number of binary classifier that classify the new example as $p_i$.
+For a new example $x$, if classifier $C_{ij}$ says $x$ is in class $p_i$, then the vote for class $p_i$ is added by one.
+Otherwise, the vote for class $p_j$ is increased by one.
+After each of the $\frac{M(M \cdot 1)}{2}$  binary classifiers makes its vote,  OVO strategy assigns $x$ to the class with the largest number of votes.
+
+ Both methods proved to be competitive with each other and there is no clear superiority of one method
+over another \cite{Duan05whichis}.
+Some claims that OVO strategy trains in less time than OVA strategy because they train with only portion of the dataset \cite{libsvm} but this claim have not yet been proofed by expeirments.
 
 \subsubsection{ SVM in Machine Learning }
-SVM has been successfully applied to a number of applications ranging from particle identification \cite{ParticleSVM7}.
-Face detection  \cite{faceSVM8,faceSVM7} and text categorization \cite{TextSVM7} to engine knock detection.
-SVMs are very suited to the classification task that commonly arise in machine vision.
-As an example we consider and application involving face identification\cite{faceSVM8}.
-Three different SVM classiifer were used on training set of 200 images (5 image per person).
+SVM has been applied successfully to various application ranging from particle identification \cite{ParticleSVM7}.
+Other application like Face detection \cite{faceSVM8,faceSVM7} and text categorization \cite{TextSVM7}.
+Also SVM suits classification task that is commonly rise in Machine vision.
+For example we will consider face identification\cite{faceSVM8} problem.
+In \cite{faceSVM8} three different SVM classifier were used on training set of 200 images (5 image per person).
 The error rate was less than 5\%.
-An RBF network and SVM classifiers were compared on the UPUS handwritten digit recognition database\cite{ORLDataset}.
+Another example is digit recognition problem where an RBF network and SVM classifiers were compared on the UPUS handwritten digit recognition database\cite{ORLDataset}.
 SVM outperformed and RBF network on all digits.
 
-\subsubsection {SVM vs. Other Learning Methods}
+%\subsubsection {SVM vs. Other Learning Methods}
 
 SVM have been applied ot the a largest MNIST a Handwritten digit recognition dataset of 60000 training set and 10000 test set\cite{IjdarSherifPaper}.
-Sherif and Ezzat has performed a comparision between SVM and several other techniques on the MNIST dataset.
-They also compared the same classfication methods on arabic handwritten digits (MADBASE) \cite{ADBase9,IjdarSherifPaper}.
+Sherif and Ezzat has performed a comparison between SVM and several other techniques on the MNIST dataset.
+They also compared the same classification methods on Arabic handwritten digits (MADBASE) \cite{ADBase9,IjdarSherifPaper}.
 Their results shows that SVM classifier outperform all other techniques on both MNIST and MADBase datasets and achieved recognition rate of over 98\%.
 Similar results was reported by \cite{empiricalcomp11,SVMInvariantComDecoste02} on the MNIST dataset.
-Ther are many other successful application SVM.
+There are many other successful application SVM.
 They have proven to be robust to noise and perform well on many tasks \cite{empiricalcomp11,Scholkopf97comparingsupport}.
 
-This seperior performance because they may elimnates many of problem experienced with other inference methodoliges with other inference methodollogies like neural network and decision trees.
+This superior performance because they may eliminates many of problem experienced with other inference methodologies with other inference methodologies like neural network and decision trees.
 The first problem that is eliminated was the problems with local minima.
-SVM can construct highly nonlinear classification and regression function without woring about getting stuck at local minima.
-Another problem SVM eliminated is the number of parameter that are picked ot generate the model.
-For example, SVM classifier with RBF  kernel will only need 2 paramters to set the model the $C$ cost paramter and $\sigma$ paramter of the RBF kernel.
+SVM can construct highly nonlinear classification and regression function without working about getting stuck at local minima.
+Another problem SVM eliminated is the numbers of parameter that are picked to generate the model.
+For example, SVM classifier with RBF kernel will only need 2 parameters to set the model the $C$ cost parameter and $\sigma$ parameter of the RBF kernel.
 The last reason is the final result are sable, reproducible and largely independent of the specific algorithm used to  optimize the SVM model.
-If two users apply the same SVM model with the same parameters to the same data, they will et the same solution modeule numeric issues.
+If two users apply the same SVM model with the same parameters to the same data, they will et the same solution module numeric issues.
 Compare this with neural networks where the results are dependent on the particular algorithms and starting point used.
@@
 
-\subsubsection {Multi Class Classifiers}
-
 \subsubsection {SVM Implementations}
-Typically an SVM approach requires the solution of a QP or LP problem.
-LP and QP type problems have been extensively studied inthe field of mathematical programming.
-These can be divided into three categories techniques in which kernel compoinents are evaluated and discarded during learning.
-Decomposition methods in which an evolving subset of data is used and new optimization approaches that specifically exploit the structure of the SVM problem.
-
-For the first category the most obvioius  approach is to sequentially update the $\alpha_i$  and this is the approach used by the kernel  Adtron (KA) algorithm \cite{kernelSVM4}.
+To solve SVM approach requires solving a QP or LP problem.
+Mathematical programming have extensively studied the LP and QP type problems.
+These techniques can be divided into three categories, methods in which evaluate and discard kernel components during learning.
+Decomposition methods in which evolving subsets of data are used and new optimization approaches that specifically utilize the structure of the SVM problem.
+
+In the first category the most obvious approach is to sequentially update the $\alpha_i$,  kernel Adtron  used this approach in (KA) algorithm \cite{kernelSVM4}.
 
 Chunking and decomposition methods optimize the SVM with respect to subsets.
-Rather than sequentially updateing the $\alpha_i$ the alternative is to upodate the $\alpha_i$ in parallel but using only a subset or wokring set of data at each stage.
-There are many successful code based on these decomposition strategies.
-SVM codes available online such as SVMTorch \cite{svmTorch} and SVMLight \cite{SVMlight} use these working set strategies.
-
-The third approach is to directly attack the SVM  problem from and optimization prespective and create algorithms that explicity exploits the structure of teh problem.
+They update $\alpha_i$ in parallel but using only a subset or working set of data at each stage rather than sequentially updating the $\alpha_i$.
+Decomposition strategies have been successfully coded in many libraries.
+SVM libraries are available on line such as SVMTorch \cite{svmTorch} and SVMLight \cite{SVMlight}.
+
+The third approach is to directly attack the SVM problem from and optimization perspective and create algorithms that explicitly exploits the structure of ten problem.
 Keerthi et al.
@@
 %function which is induced from available examples. The goal is to produce a classifier
-%that will work well on unseen examples, i.e. it generalises well. Consider the example
+%that will work well on unseen examples, i.e. it generalizes well. Consider the example
 %in Figure 2.1. Here there are many possible linear classifiers that can separate the data,
-%but there is only one that maximises the margin (maximises the distance between it
+%but there is only one that maximizes the margin (maximizes the distance between it
 %and the nearest data point of each class). This linear classifier is termed the optimal
-%separating hyperplane. Intuitively, we would expect this boundary to generalise well as
+%separating hyper plane. Intuitively, we would expect this boundary to generalize well as
 %opposed to the other possible boundaries.
@@
 \label{secstepExample}
-In this section a step by step example is presented to clearify the presented sytesm.
+In this section a step by step example is presented to clarify the presented system.
 Let us make an assumption that the user will draw a clock using two strokes.
-Figure \ref{} show the two stroke and how the user draw them.
-The next steps demonstrate how ouput of the system after each step.
-The first section \ref{sec:stepseg} discribe how the segmentation is done step by step.
+Figure \ref{fig:clock} show the two strokes and how the user draws them.
+The next steps demonstrate how output of the system after each step.
+The first section \ref{sec:stepseg} describe how the segmentation is done step by step.
 Section \ref{sec:steprec} shows the details of feature calculation and the classification process.
@@
 \label{sec:stepseg}
+
+\begin{figure}
+	\centering
+	\subfigure[First Stroke]{\label{fig:clockstroke1}
+		\includegraphics[scale=0.7]{afterDefense/clockstroke1.jpg.eps} }
+
+		\subfigure[Second Stroke]{	\label{fig:clockstroke2}
+		\includegraphics[scale=0.7]{afterDefense/ClockComplete.jpg.eps}}
+
+		\subfigure[Completed symbol]{
+		\label{fig:Clock1}\includegraphics[scale=0.7]{afterDefense/Clock1.jpg.eps}	}
+	\caption{Stroke Drawing Sequence} a) The user first draws the ellipse b) The square wave is drawn inside the ellipse c) The arrows in the symbols illustrate how the user has drawn the symbol
+	\label{fig:clock}
+\end{figure}
+
 In this step the user draw one or more stroke to draw a symbol.
-Figure \ref{} illustrate the sequence and direction of user draw.
-In this section, each stroke user draw will be labeled as it illustrated in Figure \ref{}.
-Table \ref{} shows the detailed output of the program at each step.
-The algorithm that compute the segmentation is demonstrated as follows:
+Figure \ref{fig:clock} illustrate the sequence and direction of user draw.
+In this section, each stroke user draw will be labeled as it illustrated in Figure \ref{fig:clockstroke1} and Figure \ref{fig:clockstroke2} .
+Table \ref{tab:StepsStroke1} shows the detailed output of the program at each step.
+The input of this step is the stroke points and the output is a list of segments that will be used in the recognition process.
+The algorithm that computes the segmentation is demonstrated as follows:
 \begin{enumerate}
-	\item Extracting Stroke Points 
-	\item Extracting $P_dp$
-\begin{verbatim*}
-
-\end{verbatim*}	
-\item Ellipse Fitting: which is an attempt to fit as an ellipse
-\begin{verbatim*}
-
-\end{verbatim*}	
-\item Segmentation ALS1
-\begin{verbatim*}
-
-\end{verbatim*}	
-\item Segmentation ALS2
-\begin{verbatim*}
-
-\end{verbatim*}
-
+
+	\item Extracting $P_dp$ as in Section \ref{sec:Preprocessing}
+\begin{verbatim}
+  1. For each point in Stroke do the following:
+  		1.a Compute distance from previous point.
+  		1.b Compute Time difference from previous point and add it to Time_list.  
+  		1.c Compute Velocity and add it to Vel_list. 
+  		1.d Compute Direction and add it to Dir_list.
+  	 	1.e Compute Curvature and add it to Cur_list.(Eq.4.1) 
+  	end (for loop)
+  2. For each  list computed in previous step do the follwing:   
+  	 2.a Compute average of all points. 
+  	 2.b set average as threshold.  
+  	 2.c Create region from continuous points larger than threshold. 
+  	 2.d for each region do the following. 
+  	 	   2.d.1 find the maximum 
+  	 	   2.d.1 Add point to P_dp 
+  	  end
+     end 
+\end{verbatim}	
+\item Ellipse Fitting: which is an attempt to fit as an ellipse as in Section \ref{sec:EllipseDetection} 
+%he axes of the ellipse are estimated as the $width/2$ and $height/2$ of the stroke bounding box where $width$ is width and $height$ is height of the bonding box.
+\begin{verbatim}
+1. Initialize the axes of ellipse  and center by bounding box information
+2. Repeat the following for few loops: 
+	2.a Compute error in Equation  4.2 
+	2.b Update center of ellipse and axis from error. 
+3. Compute percent in Equation 4.4 
+4. Computer eff in Equation 4.3 
+5. if eff < 0.2  then fit as ellipse otherwise move to segment. 
+\end{verbatim}
+
+\item Segmentation ALS1 as explained in Section \ref{sec:SwarmSegmentation}. The alogrithm will be similar to the PSO algorithm  in section \ref{sec:ParticleSwarmAlgorithm} 
+\begin{verbatim}
+Initialize 
+  1. Generate N particles with M dimensions (P_i...P_N) 
+  2. Generate velocities for each particle for M dimension randomly. 
+  3. Repeat till Maximum number of iteration OR Error Threshold  
+         For each particle P_i do the following 
+         3.a   Compute fitness  (fitness) 
+         				3.a.1 Compute the Error of approximation as in Eq.4.5
+         				3.a.2. Computer the fitness function as in Eq. 4.4 
+         3.b   Determine Lbest  
+               if (lbest_fitness<fitness)
+                  3.b.1   lbest=P_i
+         3.c      Determine Gbest.
+               if (lGest_fitness<fitness)
+                  3.c.1   lGest=P_i
+         3.d   Compute the Velocity for each dimension using Eq.3.2 
+         3.e   Bound Velocity using Eq.3.4
+         3.f   Update particle values using Eq. 3.3
+         3.g   Refine the location of the particle 
+         end .
+     end 
+4. return Gbest
+\end{verbatim}	
+\item Segmentation ALS2  as explained in Section \ref{sec:PolygonDivisionAlgorithm}
+\begin{verbatim}
+Initialize 
+  1. Generate N particles with M dimensions (P_i...P_N) 
+  2. Generate velocities for each particle for M dimension randomly. 
+  3. Repeat till Maximum number of iteration OR Error Threshold  
+         For each particle P_i do the following 
+         3.a   Compute fitness  (fitness) 
+              3.a.1 For each Segment in the particle do the following
+         				    3.a.1.I   Error of approximation as Line (Eq. 4.7 - 4.9 )
+         			      3.a.1.II  Error of approximation as circular arc (Eq. 4.10 - 4.19 )
+         				    3.a.1.III Computer minimum of both approximation 
+         			3.a.2 Compute Error of the particle as in Eq.  4.20
+         			3.a.3 Compute fitness of the particle as in Eq. 4.21
+         3.b   Determine Lbest  
+               if (lbest_fitness<fitness)
+                  3.b.1   lbest=P_i
+         3.c      Determine Gbest.
+               if (lGest_fitness<fitness)
+                  3.c.1   lGest=P_i
+         3.d   Compute the Velocity for each dimension using Eq.3.2 
+         3.e   Bound Velocity using Eq.3.4
+         3.f   Update particle values using Eq. 3.3
+         3.g   Refine the location of the particle 
+         end .
+     end 
+4. return Gbest
+\end{verbatim}	
+\item Append best segmentation to List of segmentation ($L_s$) of this symbols. 
 \end{enumerate}
@@
 \begin{scriptsize}
-
 %\begin{table}
 	%\centering 
-
 %\scalebox{0.6}{		
 \begin{longtable}{|p{2cm}|p{2cm}|p{2cm}|p{2cm}|p{13cm}|}
-	\caption{ Detailed Output of System in Each Step of Segmentation}
+\caption{Detailed Output of System in Each Step of Segmentation}
 \label{tab:StepsStroke1} \\
@@
  \\ \hline
-Possible Dominate point Extraction & Computing and for details see Section\ref{sec:Preprocessing} &  Stroke points  & 1. compute   the pdp  &    Number of $P_{dp}$  = 11
+Possible Dominate point Extraction & for Computing details see Section\ref{sec:Preprocessing} &  Stroke points  &   &    Number of $P_{dp}$  = 11
  Pdp = [ 2 (197.0 , 145.0 ), 4 (203.0 , 125.0 ), 3 (198.0 , 140.0 ), 22 (424.0 , 91.0 ), 19 (372.0 , 72.0 ), 26 (492.0 , 164.0 ), 28 (501.0 , 191.0 ), 9 (219.0 , 97.0 ), 31 (496.0 , 239.0 ), 0 (198.0 , 153.0 ), 41 (263.0 , 378.0 )]
@@
 Segmentation & Segmentation of Ellipse & $P_{dp}$   &  Ellipse fitting  &
- Segmented as Ellipse with dimensions  a = 194.5 ,b = 194.5 , Center( 306.5,  220.0 )   with Error =  2.276869953758031 , Percent =0.9093857875960022
- certainty  =  0.1997008626019765 
+
+  The circle is   a = 103.05715108084719 ,b = 103.06851874747797 , Center( 283.03365107691775,  181.05035287667172 )   with Error =  1.9471676673301757 , Percent =1.0695193086531243
+eff  = 0.27463462099275116  
+ Ellipse detected...
+
  \\ \hline
@@
  \\ \hline
- Segmentation & PSO Segmentation ALgS1 & $P_{dp}$  & ALgS1 &   Particles =  [  1 0 1 0 0 1 1 1 1 1 1 0 0 0 1 0 0 0 1 ] 
- The particle  Particles =  [  1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 1 0 0 0 0 1 0 ]
-
- First Polygonization PSO = [[[  S0 Line[P(201.0,179.0) to P(195.0,281.0)]  ,      S1 Line[P(195.0,281.0) to P(234.0,304.0)]  ,      S2 Line[P(234.0,304.0) to P(260.0,174.0)]  ,      S3 Line[P(260.0,174.0) to P(325.0,272.0)]  ,      S4 Line[P(325.0,272.0) to P(365.0,304.0)]  ,      S5 Line[P(365.0,304.0) to P(388.0,189.0)]  ,      S6 Line[P(388.0,189.0) to P(431.0,167.0)]  ,      S7 Line[P(431.0,167.0) to P(442.0,220.0)]  ]]]
- \\ \hline
+ Segmentation & PSO Segmentation ALgS1 & $P_{dp}$  & ALgS1 (result if Figure \ref{fig:Stroke2Seg1labeled} ) & 
+Iteration 0 the global best fitness is 0.7 and error value is 675.0457637480733  and the particle is  Particles =  [  1 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 ]
+Iteration 10 the global best fitness is 0.7777777777777778 and error value is 719.5716152204503  and the particle is  Particles =  [  1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 ]
+Iteration 20 the global best fitness is 0.7777777777777778 and error value is 719.5716152204503  and the particle is  Particles =  [  1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 ]
+ After swarm  the global best fitness is 0.875 and error value is 729.248312158671  and the particle is  Particles =  [  1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 ]
+
+ F First Polygonization PSO = [ S0 Line[P(226.0,139.0) to P(241.0,209.0)]  ,      S1 Line[P(241.0,209.0) to P(266.0,226.0)]  ,      S2 Line[P(266.0,226.0) to P(296.0,128.0)]  ,      S3 Line[P(296.0,128.0) to P(312.0,216.0)]  ,    ]
+ \\ \hline
+
+  Segmentation & PSO Segmentation ALgS2 & $P_{dp}$  & ALgS2  (result if Figure \ref{fig:Stroke2Seg2labeled} )&    
+ Iteration 0 the global best fitness is 8.725462920280576E-11 and error value is 9.151490748550003  and the particle is  Particles =  [  1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 ]
+
+Iteration 10 the global best fitness is 3.126281513587432E-7 and error value is 8.61499750333895  and the particle is  Particles =  [  1 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 ]
+
+Iteration 20 the global best fitness is 3.126281513587432E-7 and error value is 8.61499750333895  and the particle is  Particles =  [  1 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 ]
+
+ After swarm  the global best fitness is 3.126281513587432E-7 and error value is 8.61499750333895  and the particle is  Particles =  [  1 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 ]
+
+  second PSO =   Segments = [   S0 Line{P(226.0,139.0) to P(238.0,183.0)},      S1 Line{P(238.0,183.0) to P(241.0,213.0)},      S2 Line{P(241.0,213.0) to P(276.0,227.0)},      S3 Line{P(276.0,227.0) to P(287.0,144.0)},      S4 Curve{P(287.0,144.0) to P(310.0,131.0),  Center ( 299.02634312229503,137.80386092517725) Radius =11.907214051459519 },  S5 Curve{P(310.0,131.0) to P(312.0,216.0),  Center ( 82.44520390380922,172.2089170470769) Radius =231.70851748053468 },  S6 Line{P(312.0,216.0) to P(336.0,224.0)}, ]
+
+ \\ \hline 
+Choosing best fit & Compare between ALgS1 and AlgS2 &  First PSO and Second PSO  &  (result if Figure \ref{fig:FinalSegmentation} ) & 
+   The error ALgS1 780.6652975044847  Algs2 161.60692855994625    Best = ALgS2  \\ \hline
 
@@
 \end{landscape}
+
+\begin{figure}
+	\centering
+	\subfigure [AlgS1 Output]{
+		\includegraphics{afterDefense/Stroke2Seg1labeled.jpg.eps}
+	\label{fig:Stroke2Seg1labeled}}
+
+	\subfigure[AlgS2 Output]{	\includegraphics{afterDefense/Stroke2Seg2labeled.jpg.eps}	\label{fig:Stroke2Seg2labeled}
+	}
+	\subfigure[Final Segmentation of Symbol]{		\label{fig:FinalSegmentation}	\includegraphics[scale=0.65]{afterDefense/FinalSegmentation.jpg.eps}
+	}
+	\caption{Segmentation Ouputs For Second Stroke}
+\end{figure}
+
+The final output of this step is list $L_s$ of segments There are  9  segments in this symbol
+
+\begin{itemize}
+	\item    Segment S0  = Ellipse ( Stroke 1)
+   \item Segment S1  =   Line ( Stroke 2)
+     \item  Segment S2 = Line ( Stroke 2)
+     \item  Segment S3  =  Line ( Stroke 2)
+     \item Segment S4  =   Line ( Stroke 2)
+  \item    Segment S5  =  Curve  ( Stroke 2)
+     \item  Segment S6  =  Curve  ( Stroke 2)
+     \item  Segment S7  =  Line   ( Stroke 2)
+     \item  Segment S8  =  Line   ( Stroke 2)
+\end{itemize}
+
 \subsection{Step by Step Recognition}
 \label{sec:steprec}
+
+\begin{figure}
+	\centering
+		\includegraphics[scale=0.85]{afterDefense/FinalSegmentation.jpg.eps}
+	\caption{Best Segmantation }
+	\label{fig:segmentatiofinal.jpg}
+\end{figure}
+
 %\begin{enumerate}
@@
 
-%	\item Processing for the first Stroke 1.
-
-%	\item Possible Dominiate Point Extraction
-%	\item Segmentation 
-%	   Computing 
-%	\begin{enumerate}
-%	\item Ellipse detection  
-%	      Calcualting estimated ellipse with radius = $\frac{w}{2}$  and center as same as the bounding box of stroke.  
-%	      The ouput of the system $$
-% \end{enumerate}
-%	
-%	
-%\end{enumerate}
-%  
-%
-%\end{itemize}
+	The input of the step is the list of segmentation $L_s$ that is used to compute the feature vector.
+Figure \ref{fig:segmentatiofinal.jpg} shows the best segmentation as given by the previous step.
+This step can be summarized as firstly using $L_s$ to computer the feature vector and then introduce this vector the SVM classifier to classify it to one of the trained categories.
+The details of each step of feature extraction are detailed in the next algorithm.
+Feature extraction step is divided into four steps where each feature set is computed separately then all sets are appended to construct a final feature vector.
+
+\begin{enumerate}
+	\item Computing Structural and geometrical Features (FS1).
+This step computes number of parallel, perpendicular line and similar structural features.
+The following algorithm is used to determine these features.
+
+	\begin{verbatim}
+Initialize 
+  1. For each segment Si in list Ls  do the following. 
+      1.a If segment si is a line do the following 
+            1.a.I. Compute the slope and intercept of line si. 
+            1.a.II Increment Parallel count  if 
+                  parallel with each other line in Ls
+            1.a.III Increment PerpendicularCount if 
+                    perpendicular with each other line in Ls 
+            1.a.IV Increment Intersection Count if 
+                    intersected with any other line. 
+                    1.a.II.A  if Intersection found determine 
+                              type of intersection or (X,T,L)
+      1.b If segment si is curve do the following. 
+          1.b.I Increment Intersect if
+               Intersect other line or curve
+    1.b.II Increment number 
+                 of curves and primitives. 
+4. return Feature_Vector_FS1
+\end{verbatim}
+
+	\item Rubine Feature Set (FS2).  Compute Rubine features as in \cite{gestureexample12} 
+	\item Statistical Features (FS3) compute the moments as in \cite{zernike61}
+	\item Global shape properties set (FS4). Firstly compute the convex hull of the stroke points then compute the rest of the features.
+
+	\item Introduce the computed feature vector to the classifier and check the label of the detected symbol. 
+\end{enumerate}
+
+\begin{landscape}
+\begin{scriptsize}
+
+ \begin{longtable}{|p{2cm}|p{2cm}|p{2cm}|p{2cm}|p{13cm}|}
+\caption{Detailed Output of System in Each Step of Recognition}
+\label{tab:StepsReg} \\
+
+\hline 
+\multicolumn{1}{|p{2cm}|}{\textbf{Step}} & 
+\multicolumn{1}{p{2cm}|}{\textbf{Description}} &
+\multicolumn{1}{p{2cm}|}{\textbf{Inputs}} &
+\multicolumn{1}{p{2cm}|}{\textbf{Details}} &
+\multicolumn{1}{p{13cm}|}{\textbf{Output}} 
+\\ \hline 
+\endfirsthead
+
+\hline
+\multicolumn{3}{c}%
+{{\bfseries \tablename\ \thetable{} -- continued from previous page}} \\
+\multicolumn{1}{|p{2cm}|}{\textbf{Step}} & 
+\multicolumn{1}{p{2cm}|}{\textbf{Description}} &
+ \multicolumn{1}{p{2cm}|}{\textbf{Inputs}} &
+ \multicolumn{1}{p{2cm}|}{\textbf{Details}} &
+ \multicolumn{1}{p{13cm}|}{\textbf{Output}} 
+\\ \hline 
+\endhead
+
+ Feature Extraction of FS1 & Preprocessing of Structural Featues  &   There are  9  segments in this symbol $L_s$ = 
+   Segment S0  = Ellipse
+   Segment S1  = Line
+   Segment S2  = Line
+   Segment S3  = Line
+   Segment S4  = Line
+   Segment S5  = Curve
+   Segment S6  = Curve
+   Segment S7  = Line
+   Segment S8  = Line  &  (see First Algorithm in Section \ref{sec:steprec} )  &    Line from  24 x= 241.0y=213.0  to 49 x2= 276.0y2=227.0  is\textbf{ parallel} to Line from  163 x= 312.0y=216.0  to 185 x2= 336.0y2=224.0
+
+ Line from  0 x= 226.0y=139.0  to 14 x2= 238.0y2=183.0 \textbf{ intersects }Line from  14 x= 238.0y=183.0  to 24 x2= 241.0y2=213.0
+
+ Line from  0 x= 226.0y=139.0  to 14 x2= 238.0y2=183.0  \textbf{intersects} Line from  24 x= 241.0y=213.0  to 49 x2= 276.0y2=227.0
+
+ Line from  0 x= 226.0y=139.0  to 14 x2= 238.0y2=183.0 \textbf{ intersects} Line from  49 x= 276.0y=227.0  to 98 x2= 287.0y2=144.0
+
+ Line from  0 x= 226.0y=139.0  to 14 x2= 238.0y2=183.0\textbf{  intersects }Line from  163 x= 312.0y=216.0  to 185 x2= 336.0y2=224.0
+
+ Line from  0 x= 226.0y=139.0  to 14 x2= 238.0y2=183.0 \textbf{ intersects} Line from  185 x= 336.0y=224.0  to 223 x2= 344.0y2=137.0
+ %Line from  14 x= 238.0y=183.0  to 24 x2= 241.0y2=213.0  intersects Line from  24 x= 241.0y=213.0  to 49 x2= 276.0y2=227.0
+ %Line from  14 x= 238.0y=183.0  to 24 x2= 241.0y2=213.0  intersects Line from  49 x= 276.0y=227.0  to 98 x2= 287.0y2=144.0
+ %Line from  14 x= 238.0y=183.0  to 24 x2= 241.0y2=213.0  intersects Line from  163 x= 312.0y=216.0  to 185 x2= 336.0y2=224.0
+ %Line from  14 x= 238.0y=183.0  to 24 x2= 241.0y2=213.0  intersects Line from  185 x= 336.0y=224.0  to 223 x2= 344.0y2=137.0
+ %Line from  24 x= 241.0y=213.0  to 49 x2= 276.0y2=227.0  intersects Line from  49 x= 276.0y=227.0  to 98 x2= 287.0y2=144.0
+ %Line from  24 x= 241.0y=213.0  to 49 x2= 276.0y2=227.0  intersects Line from  163 x= 312.0y=216.0  to 185 x2= 336.0y2=224.0
+ %Line from  24 x= 241.0y=213.0  to 49 x2= 276.0y2=227.0  intersects Line from  185 x= 336.0y=224.0  to 223 x2= 344.0y2=137.0
+ %Line from  49 x= 276.0y=227.0  to 98 x2= 287.0y2=144.0  intersects Line from  163 x= 312.0y=216.0  to 185 x2= 336.0y2=224.0
+
+ Line from  49 x= 276.0y=227.0  to 98 x2= 287.0y2=144.0  intersects Line from  185 x= 336.0y=224.0  to 223 x2= 344.0y2=137.0
+
+ Line from  163 x= 312.0y=216.0  to 185 x2= 336.0y2=224.0  intersects Line from  185 x= 336.0y=224.0  to 223 x2= 344.0y2=137.0
+
+\\ \hline
+
+ FS1 & Structural and geometrical Features & Segment List $L_s$  &  & 
+ The final Feature vector of the FS1==>
+ature Number of primitives 3.0   ,    Feature Number of segments 9.0   ,    Feature Line count 6.0   ,    Feature Curves Count 3.0   ,    Feature Ellipse Count 1.0   ,    Feature Intersection T type -1.0   ,    Feature Intersection L type 4.0   ,    Feature Intersection X type -1.0   ,    Feature Parallel Count 1.0   ,    Feature PerPendicular Count -1.0   ,    Feature Intersection Lines Count 4.0   ,    Feature Min Radius 1.7976931348623157E308   ,    Feature Max Radius 231.70851748053468
+
+ \\ \hline
+FS2  & Rubine Features & Segment List $L_s$  &   see \cite{gestureexample12}   & 
+\begin{scriptsize}
+ Feature Rubine 0 -1.0   ,    Feature Rubine 1 -1.0   ,    Feature Rubine 2 477.6107201476952   ,    Feature Rubine 3 0.6368010415482491   ,    Feature Rubine 4 169.49926253526885   ,    Feature Rubine 5 0.8908593331996381   ,    Feature Rubine 6 0.454279262624981   ,    Feature Rubine 7 1179.0172885029808   ,    Feature Rubine 8 4.712388980384688   ,    Feature Rubine 9 70.80157037633609   ,    Feature Rubine 10 55.4044569978846   ,    Feature Rubine 11 9.140625   ,    Feature Rubine 12 41783.0   ,
+\end{scriptsize}
+ \\ \hline
+
+FS3 & Momments Computation   &  Segment List $L_s$  &  see \cite{zernike61}  &
+
+\begin{scriptsize}
+   Feature Zernike moments 0 158.87520185460517   ,    Feature Zernike moments 0 7.007557168632727   ,    Feature Zernike moments 1 34.26844061831789   ,    Feature Zernike moments 1 6.658086415801404   ,    Feature Zernike moments 2 85.70032052974454   ,    Feature Zernike moments 2 9.441591321498212   ,    Feature Zernike moments 3 11.73170471301825   ,    Feature Zernike moments 3 66.52997844880663   ,    Feature Zernike moments 4 20.42774127507497   ,    Feature Zernike moments 4 10.196341027811322   ,    Feature Zernike moments 5 52.821799077785066   ,    Feature Zernike moments 5 12.078236804336935   ,    Feature Zernike moments 6 27.748912050714   ,    Feature Zernike moments 6 9.870152472028517   ,    Feature Zernike moments 7 26.079803494397133   ,    Feature Zernike moments 7 28.109003184347657   ,    Feature Zernike moments 8 23.32097186060531   ,    Feature Zernike moments 8 7.632681250023639   ,    Feature Zernike moments 9 6.5841386188472395   ,    Feature Zernike moments 9 51.34195037191517   ,    Feature Zernike moments 10 30.122054353503337   ,    Feature Zernike moments 10 26.902895199361584   ,    Feature Zernike moments 11 1.8295982451659387   ,    Feature Zernike moments 11 29.93196640188875   ,    Feature Zernike moments 12 64.0879253626976   ,    Feature Zernike moments 12 31.811780133580722   ,    Feature Zernike moments 13 23.614051070558084   ,    Feature Zernike moments 13 1.8043229719502216   ,    Feature Zernike moments 14 113.92181136864328   ,    Feature Zernike moments 14 67.26888519955787   ,    Feature Zernike moments 15 47.7397132180995   ,    Feature Zernike moments 15 31.490125533238682   ,    Feature Zernike moments 16 10.320112066817984   ,    Feature Zernike moments 16
+\end{scriptsize}
+
+  \\ \hline
+FS4   &  &Segment List $L_s$   &  &  \begin{scriptsize} 
+ The convex hull    [ X = 383.0   Y = 146.0  Time = 1250452568824,  X = 380.0   Y = 139.0  Time = 1250452568816,  X = 368.0   Y = 119.0  Time = 1250452568800,  X = 363.0   Y = 113.0  Time = 1250452568792,  X = 352.0   Y = 101.0  Time = 1250452568776,  X = 348.0   Y = 97.0  Time = 1250452568768,  X = 342.0   Y = 92.0  Time = 1250452568760,  X = 335.0   Y = 88.0  Time = 1250452568752,  X = 329.0   Y = 85.0  Time = 1250452568744, % X = 322.0   Y = 83.0  Time = 1250452568736,  X = 302.0   Y = 78.0  Time = 1250452568728,  X = 293.0   Y = 77.0  Time = 1250452568720,  X = 282.0   Y = 76.0  Time = 1250452568712,  X = 250.0   Y = 76.0  Time = 1250452568688,  X = 239.0   Y = 77.0  Time = 1250452568680,  X = 229.0   Y = 79.0  Time = 1250452568672,  X = 220.0   Y = 81.0  Time = 1250452568664,  X = 217.0   Y = 83.0  Time = 1250452568656,  X = 201.0   Y = 96.0  Time = 1250452568648,  X = 196.0   Y = 103.0  Time = 1250452568640,  X = 186.0   Y = 118.0  Time = 1250452568624,  X = 182.0   Y = 127.0  Time = 1250452568608,  X = 180.0   Y = 209.0  Time = 1250452569139,  X = 180.0   Y = 247.0  Time = 1250452569092,  X = 181.0   Y = 253.0  Time = 1250452569080,  X = 185.0   Y = 260.0  Time = 1250452569073,  X = 190.0   Y = 265.0  Time = 1250452569056,  X = 194.0   Y = 267.0  Time = 1250452569047,  X = 213.0   Y = 274.0  Time = 1250452569031,  X = 220.0   Y = 276.0  Time = 1250452569023,  X = 256.0   Y = 284.0  Time = 1250452568995,  X = 266.0   Y = 284.0  Time = 1250452568976,
+ ...................................................X = 274.0   Y = 283.0  Time = 1250452568969,  X = 282.0   Y = 281.0  Time = 1250452568960,  X = 290.0   Y = 278.0  Time = 1250452568952,  X = 298.0   Y = 274.0  Time = 1250452568944,  X = 319.0   Y = 262.0  Time = 1250452568936,  X = 339.0   Y = 243.0  Time = 1250452568912,  X = 357.0   Y = 222.0  Time = 1250452568889,  X = 368.0   Y = 202.0  Time = 1250452568872,  X = 378.0   Y = 183.0  Time = 1250452568864,  X = 384.0   Y = 167.0  Time = 1250452568848,  X = 384.0   Y = 160.0  Time = 1250452568840]
+ Area of convex hull =  -34160.5   preimiter of hull 656.1576413855265
+
+ --------------------------------------------------------------------------------
+
+Final Feature vector for the FS4: 
+  Feature Centroid time 1.2504525968640884E12  , Feature Centroid time difference 28319.088339222613   ,    Feature area of convexhull 34160.5   ,    Feature area of convexhull/area of symbol 1.2365344240932454   ,    Feature N.Points ConvexHull/ N.
+points symbol 0.15302491103202848   ,    Feature Convext Perimeter/symbol perimeter 0.556529278903672   ,    Feature  percentage of convex area to bounding box area 19.493542609351433   ,    Feature area  35519.5   ,    Feature diff area of symbol to area of bounding box 16.290771116138764   ,    Feature  Ration between w/h  0.9807692307692307   ,    Feature  Log of ration between hight and width -0.008433167536862813   ,
+ Feature  Log of computed area 4.550466843684004   ,    Feature  abs of area 27626.0   ,    Feature log of  abs of area 4.441318007475848   ,    Feature  Density(length/size)  4.046853548165911   ,    Feature Mean CentroidalRadius x 272.48042704626334   ,    Feature Mean CentroidalRadius y 177.19928825622776   ,    Feature Mean t 1.2504525970626514E12   ,    Feature Mean time difference 148.69395017793593   ,    Feature Mean Radius 62.000799881019084
+  \end{scriptsize}
+    \\ \hline
+ Feature Vector & Appends the selected Feature FS1, FS2, FS3, FS4 & Previous Steps  & &
+
+   \begin{scriptsize}
+   Feature Number of primitives 3.0   ,    Feature Number of segments 9.0   ,    Feature Line count 6.0   ,    Feature Curves Count 3.0   ,    Feature Ellipse Count 1.0   ,    Feature Intersection T type -1.0   ,    Feature Intersection L type 4.0   ,    Feature Intersection X type -1.0   ,    Feature Parallel Count 1.0   ,    Feature PerPendicular Count -1.0   ,    Feature Intersection Lines Count 4.0   ,    Feature Min Radius 1.7976931348623157E308   ,    Feature Max Radius 231.70851748053468   ,    Feature Centroid x 92.54770318021201   ,    Feature Centroid y 101.3851590106007   ,    Feature Centroid time 1.2504525968640884E12   ,    Feature Centroid time difference 28319.088339222613   ,    Feature area of convexhull 34160.5   ,    Feature area of convexhull/area of symbol 1.2365344240932454   ,    Feature N.Points ConvexHull/ N.
+points symbol 0.15302491103202848   ,    Feature Convext Perimeter/symbol perimeter 0.556529278903672   ,    Feature  percentage of convex area to bounding box area 19.493542609351433   ,    Feature area  35519.5   ,    Feature diff area of symbol to area of bounding box 16.290771116138764   ,    Feature  Ration between w/h  0.9807692307692307   ,    Feature  Log of ration between hight and width -0.008433167536862813   ,    Feature Zernike moments 0 158.87520185460517   ,    Feature Zernike moments 0 7.007557168632727   ,    Feature Zernike moments 1 34.26844061831789   ,    Feature Zernike moments 1 6.658086415801404   ,    Feature Zernike moments 2 85.70032052974454   ,    Feature Zernike moments 2 9.441591321498212   ,    Feature Zernike moments 3 11.73170471301825   ,    Feature Zernike moments 3 66.52997844880663   ,    Feature Zernike moments 4 20.42774127507497   ,    %Feature Zernike moments 4 10.196341027811322   ,    Feature Zernike moments 5 52.821799077785066   ,    Feature Zernike moments 5 12.078236804336935   ,    Feature Zernike moments 6 27.748912050714   ,    Feature Zernike moments 6 % 9.870152472028517   ,    Feature Zernike moments 7 26.079803494397133   ,    Feature Zernike moments 7 28.109003184347657   ,    Feature Zernike moments 8 23.32097186060531   ,    Feature Zernike moments 8 7.632681250023639   ,    Feature Zernike moments 9 6.5841386188472395   ,    Feature Zernike moments 9 51.34195037191517   ,    Feature Zernike moments 10 30.122054353503337   ,    Feature Zernike moments 10 26.902895199361584   ,    Feature Zernike moments 11 1.8295982451659387   ,    Feature Zernike moments 11 29.93196640188875   ,    Feature Zernike moments 12 64.0879253626976   ,    Feature Zernike moments 12 31.811780133580722   ,    Feature Zernike moments 13 23.614051070558084   ,    Feature Zernike moments 1    1.8043229719502216   ,
+%Feature Zernike moments 14 113.92181136864328   ,    Feature Zernike moments 14 67.26888519955787   ,    Feature Zernike moments 15 47.7397132180995   ,    Feature Zernike moments 15 31.490125533238682   ,
+Feature Zernike moments 16 10.320112066817984   ,    Feature Zernike moments 16 4.007106145033957   ,    Feature Rubine 0 -1.0   ,    Feature Rubine 1 -1.0   ,    Feature Rubine 2 477.6107201476952   ,    Feature Rubine 3 0.6368010415482491   ,    Feature Rubine 4 169.49926253526885   ,    Feature Rubine 5 0.8908593331996381   ,    Feature Rubine 6 0.454279262624981   ,    Feature Rubine 7 1179.0172885029808   ,    Feature Rubine 8 4.712388980384688   ,    Feature Rubine 9 70.80157037633609   ,    Feature Rubine 10 55.4044569978846   ,    Feature Rubine 11 9.140625   ,    Feature Rubine 12 41783.0   ,    Feature  Log of computed area 4.550466843684004   ,    Feature  abs of area 27626.0   ,    Feature log of  abs of area 4.441318007475848   ,    Feature  Density(length/size)  4.046853548165911   ,    Feature Mean CentroidalRadius x 272.48042704626334   ,    Feature Mean CentroidalRadius y 177.19928825622776   ,    Feature Mean t 1.2504525970626514E12   ,    Feature Mean time difference 148.69395017793593   ,    Feature Mean Radius 62.000799881019084   ,
+   \end{scriptsize} \\ \hline
+   Classification  & SVM classifier decision & Feature Vector  &  & 
+   Symbol = Clock.
+
+ \\ \hline
+
+ 		\end{longtable}
+%}
+
+%\end{table}
+
+\end{scriptsize}
+
+\end{landscape}

### 2009-09-01 draft/ChapterIntroduction.tex
after dr. mahmoud meeting

@@
 \begin{center}
-		\includegraphics[scale=0.8]{../../neededfiles/Figures/strokeinterpolation.eps}
+		\includegraphics[scale=0.6]{../../neededfiles/Figures/strokeinterpolation.eps}
 	\caption[Segmentation Error]{Different ways for segmenting a stroke drawn by user.
@@
 The problem could be solved by providing context or domain information which could not be provided in most sketches.
+
 \begin{figure}	
 	\centering
-		\includegraphics[scale=0.85]{../../neededfiles/Figures/VorU.eps}
-	\caption[Handwriting Ambiguities] { The same letter can be interpreted differently in various contexts; first as V in drive or as U in Cut. This example is adapted from \cite{Ambiguity}.}
+		\includegraphics[scale=0.65]{afterDefense/Amb.jpg.eps}
+	\caption[Handwriting Ambiguities] {The same handwriting can be interpreted differently in various contexts; In the first line the b can be viewed as a B and in the second one could see it as 8.
+This example is adapted from \cite{ambiguityWebsite}.
+}
 \label{fig:VorU}

### 2009-09-01 draft/System.tex
after dr. mahmoud meeting

--- 2009-08-17 draft/System.tex
@@
 %the performance of most algorithms is based on user style. 
-
 % Stroke segmentation is one of the most complex problems in the sketch recognition system. It is dividing the   
 %In stroke segmentation problem is Particle swarm optimization is used as the 
-
 \section{System Overview}
@@
 \begin{center}
-	\includegraphics[scale=0.75]{images/AllBlockDiagram.eps}
+	\includegraphics[scale=0.9]{images/AllBlockDiagram.eps}
 	\caption[The System Block Diagram]{Block diagram of the system}
@@
 Chen et al.\cite{CruveDivisionSwarm} uses digital curves scanned from paper as input to the system and did not take advantage of the curvature or local geometric properties of the digital curve.
-
- As shown in Figure \ref{fig:Blockdiagram} first an attempt is made to fit the stroke points into a curve or an ellipse using a minimum square error fitting algorithm \cite{chernov-2003}.
+However, comparisons between PSO and GA \cite{ComparePSOGA05} shows that PSO converges to optimal solution faster than GA.
+
+ As shown in Figure \ref{fig:Blockdiagram} first an attempt is made to fit the stroke points into a curve or an ellipse using a minimum square error fitting algorithm \cite{chernov}.
 If the stroke proved to be an elliptical arc then the segmentation process ends and the system proceeds to the next step.
@@
 The two algorithms will generate two segmentations, the system will choose the segmentation that has the minimum segmentation error.%
+
 % First an attempts the segmentation process is divided into two steps ellipse fitting and curve segmentation.
@@
 
- In \textbf{Feature Extraction} step, a feature vector consists of a set of statistical, geometrical and spatial features are extracted from the segmented strokes.
+ In \textbf{Feature Extraction} step, the approximated segments that were computed in the previous step are appended to a list of segments.
+A feature vector consists of a set of statistical, geometrical and spatial features are extracted from the segmented strokes.
 The feature vector is used as input for the next step.
@@
 
-   The final step is \textbf{Classification} using a SVM classifier.
-The classifier attempts to classify the symbol into one the previously trained classes.% c  that will use the features computed to classify the segments into one of the previously trained classes.
+   The final step is \textbf{Classification} using a SVM classifier. The classifier attempts to classify the symbol into one the previously trained classes.
+   % c  that will use the features computed to classify the segments into one of the previously trained classes.
 % Or to determine the symbol of the given segments from the set of preciously trained symbols.
-
  %with a symbol from the training set.% The system compute composite set of features some are statistical other are spatial features based on the type primitives.
 Finally, the strokes are classified into the corresponding classes.% the classifier identifies the strokes into a symbol from the set of known symbols.\\
-
 %\chapter{System Details}
 %\label{sec:SystemDetails}
-
 \section{Possible Dominant Point Extraction}
@@
 The segmentation tries to divide the stroke into a set of primitives.
-As shown in Figure \ref{fig:Blockdiagram} first an attempt is made to fit the stroke points into a curve or an ellipse using a minimum square error fitting algorithm \cite{chernov-2003}.
+As shown in Figure \ref{fig:Blockdiagram} first an attempt is made to fit the stroke points into a curve or an ellipse using a minimum square error fitting algorithm \cite{chernov}.
 If the stroke proved to be an ellipse arc then the segmentation process ends and the system proceeds to the next step.
@@
 \end{equation}%\]
-where $N$ is the number of points in the stroke, $D$ is the number of point in the solution that was previously labeled as $P_{pd}$  and $E$ is the computed error and $\varepsilon$ is the error threshold.
+where $N$ is the number of points in the stroke, $D$ is the number of point in the solution that was previously labeled as $P_{pd}$  and $E$ is the computed error and $\varepsilon$ is the error threshold which is computed as 20 \% of the input stroke area.
 If the solution produced an error that exceed the error threshold $\varepsilon$ the fitness function assign a -ve value for the solution to express it's in feasibility, otherwise the inverse of the number of vertex produced is used to access the solution fitness in terms of minimum number of vertices'.
@@
 The goal of the training process is to generate a linear SVM plane that maximize margin between the two categories.
-A cross validation procedure is implemented to choose the best value for the $C$ paramter in SVM classifier.( $C=100$ was the best value)
+A cross validation procedure is implemented to choose the best value for the $C$ parameter in SVM classifier.( $C=100$ was the best value)
 =
-For classification the decision is made from  producing the feature vector of the unkown symbol to the trained $\frac{n(n-1)}{{2}}$ binary classifiers.
+For classification the decision is made from producing the feature vector of the unknown symbol to the trained $\frac{n(n-1)}{{2}}$ binary classifiers.
 The final decision is given using simple voting strategy.
@@
 However, there can be more than one plane that can separates the data therefore SVM attempts to find the plane that maximize the margin (distance between classifier and nearest sample) between the two classes \ref{fig:svm1}.
-To calculate the margin, two different hyperplanes are constructed, one plane on each side of the data, these two planes are then moved to generate the maximum margin possible.
+To calculate the margin, two different hyper planes are constructed, one plane on each side of the data, these two planes are then moved to generate the maximum margin possible.
 \begin{figure}
@@
  The orientation of the separation plane is determined using the vector $\omega$ and the offset of the plane from the origin is determined using the scalar $b$.
-Lets make an assumption that the datat point are linearly separable, i.e.
+Let's make an assumption that the data point are linearly separable, i.e.
 there are infinite possibilities to find a plane that separates and correctly classify the training data.
@@
 % Which one is preferable ?
-Intuitively one perfers the solid pane since small perturbations of any point would not introduce misclassifications errors.
+Intuitively one prefers the solid pane since small perturbations of any point would not introduce misclassifications errors.
 Without any information, the solid plane is more likely to generalize better on future data.
-Geometrically we can characterize the solid plane as being furthuret from both classes.
-plane supports a class if all points in that class are on one side of that plane.
+Geometrically we can characterize the solid plane as being furtherer from both classes.
+Plane supports a class if all points in that class are on one side of that plane.
 
@@
 The concept of duality can also used to wields the problem to a Largrnigian dual QP problem in Equation \ref{eq:dual}.
-% maximumm margin between parallel 
+% maximum margin between parallel 
 \begin{equation}
@@
 
-The generalization error on future points and not only in the training set can be obtained as proved by the statical learning theory \cite{BennettSVMP2000}.
+The generalization error on future points and not only in the training set can be obtained as proved by the statistical learning theory \cite{BennettSVMP2000}.
 The generalized error bound is a function of the misclassified error on the training data and the terms that measure the complexity or capacity of the classification.
@@
 Thus, by explicitly maximizing the margin we are in the same time minimizing bound on the generalization error and therefore expect better generalization with high probability.
-The dimensionality of the data dose not affect the size of the margin which   results in good performance  for very high dimensional data (i.e, with a very large number of attribute).
+The dimensionality of the data does not affect the size of the margin which   results in good performance for very high dimensional data (i.e, with a very large number of attribute).
 In a sense, this also reduces the problems caused by over fitting of high - dimensional data.
-A large volume of literature  , e.g., \cite{bookSVMoverfit11,statisticalSVM1} refers to this problem,  for more technical discussion of the   statistical theory.
-
-%Using geomteric arguments we can gain in sights to some of these results.
-Classification function that have capcity to fit the training data are more likely to overfit  resulting in poor generalization.
+A large volume of literature, e.g., \cite{bookSVMoverfit11,statisticalSVM1} refers to this problem,  for more technical discussion of the   statistical theory.
+
+%Using geometric arguments we can gain insights to some of these results.
+Classification function that have capacity to fit the training data are more likely to over fit resulting in poor generalization.
 
@@
 Consider the linear inseparable problem in Figure\ref{fig:svm5}, in the ideal case there will be no point misclassified and no point falls on the margin.
-But we must relax the constrain that insure that each point in on the appropriate side of tis supporting plane.
+But we must relax the constrain that insure that each point in on the appropriate side of it is supporting plane.
 Every point that falls on the wrong side of the supporting plane is considered an error.
@@
 The expanded dataset features space can be separated using an existing linear classification algorithm.
-The  linear classification that is applied on the feature space produce a non linear classification on the original input space.
-In two dimensional vector space with attributes $r$ and $s$, to construct a quadratic  discriminant  simply map the original two dimensional input space $\left[r,s\right]$ to  five dimensional features space $\left[r,s,rs,r^2,s^2\right]$ and construct a linear discriminant in that space.
+The linear classification that is applied on the feature space produces a non linear classification on the original input space.
+In two dimensional vector space with attributes $r$ and $s$, to construct a quadratic  discriminate  simply map the original two dimensional input space $\left[r,s\right]$ to  five dimensional features space $\left[r,s,rs,r^2,s^2\right]$ and construct a linear discriminate in that space.
 specifically, define : $\vartheta(x):R^2  \rightarrow R^5$ then
@@
 
-For higher dimensional datasets, this method of nonlinear mapping  has two potential problems stemming from the fact that dimensionality of the feature space explones exponentially.
+For higher dimensional datasets, this method of nonlinear mapping has two potential problems stemming from the fact that dimensionality of the feature space explodes exponentially.
 The first problem is that over fitting becomes a problem.
@@
 
-Examine what happens when athe nonlinear mapping is introduced to equation \ref{eq:dual}. Let us define : $\vartheta(x):R^2\rightarrow R^5 \quad n>> n$  we need to optimize
+Examine what happens when  the nonlinear mapping is introduced to equation \ref{eq:dual}. Let us define: $\vartheta(x):R^2\rightarrow R^5 \quad n>> n$  we need to optimize
 
@@
 Now we apply a little mathematically magic known as hilbern schmidt kernels, first applied to SVMs in \cite{svmintroduce}.
-By Mercer's theorem, we know that for certain mapping $\theta$ and any two points $u$ and $v$, the inner product of the the mapped points can be evaluated using the kernel function without ever explicitly knowing the mapping , e.g., $\theta(u)\theta(v)\equiv K(u,v)$.
+By Mercer's theorem, we know that for certain mapping $\theta$ and any two points $u$ and $v$, the inner product of the mapped points can be evaluated using the kernel function without ever explicitly knowing the mapping, e.g., $\theta(u)\theta(v)\equiv K(u,v)$.
 Some of the more popular known kernels are given below.
-New kernel are being developed to fit domain specific requirements.
+New kernels are being developed to fit domain specific requirements.
 \begin{center}
@@
 	Two layer Neural Network& $ sigmoid \left(\eta\left(u\cdot v +1\right)+c\right)$ \\ 
-
 	\end{tabular}
@@
 %The foundations of Support Vector Machines (SVM) have
-\subsubsection{Classification}
+\subsubsection{SVM  Classification}
 The resulting SVM method can be summarized as follows 
 \begin{enumerate}
-	\item Select parameter C representing the tradeoff between minimizing the training set error and maximizing the margin.
-Select the kernel function and any kernel parameters.
-For example for the radial basis function kernel, one must select width of the Gaussian $\gamma$.
-	\item Solve dual QP Equation \ref{} or an alternative SVM formulation using an appropriate quadratic programming or linear programming algorithm. 
-	\item Recover th primal threshold variable $b$ using the support vectors. 
+	\item Selecting the SVM classifier Paramters. Parameter $C$ representing the tradeoff between minimizing the training set error and maximizing the margin. 
+	\item Select A kernel function and their kernel parameters. For example for the radial basis function kernel, one must select width of the Gaussian $\gamma$. 
+	\item Solve dual QP Equation \ref{eq:dualkernel} or an alternative SVM formulation using an appropriate quadratic programming or linear programming algorithm. 
+	\item Get the support vector and recover the emprial threshold  variable $b$. 
 	\item Classify new points using $f(x)=sign\left(\sum{i}y_i\alpha_i K\left(x,x_i\right)-b\right)$
 \end{enumerate}
-
 The parameters in step 1. are selected using cross validation if sufficient data are available. %However, recent model  selectin strategies can 
@@
 
- There is mainly two popular methods to change the binary SVM classification problem into Muliclass classification problem.
-The first methods is One Versus All (OVA) Strategy where M binary classsifiers is constructed one classifier for each class.
-The $ith$ classifier is trained using positive example form $ith$ class and negtive example from all other classes.
+ There is mainly two popular methods to change the binary SVM classification problem into Multiclass classification problem.
+The first methods is One Versus All (OVA) Strategy where M binary classifiers is constructed one classifier for each class.
+The $ith$ classifier is trained using positive example form $ith$ class and negative example from all other classes.
 The output function $p_i$ represents the distance and classification of the example from the $ith$ classifier.
@@
  The second method is One Versus One Strategy where a binary classifier for every pair of distinct classes ($M$ classes) and so, all together $\frac{M(M \cdot 1)}{2}$ binary classifiers are constructed.
-Each binary classifier is trianed using exapmles from both $p_i$ and $p_j$.
+Each binary classifier is trained using examples from both $p_i$ and $p_j$.
 For example, classifier $C_{ij}$ is trained taking the examples from $p_i$ as positive and the examples from $p_j$ as negative.
-The classification is based on the number of binary classifer that classifiy the new example as $p_i$.
+The classification is based on the number of binary classifier that classify the new example as $p_i$.
 For a new example $x$, if classifier $C_{ij}$ says $x$ is in class $p_i$, then the vote for class $p_i$ is added by one.
@@
  Both methods proved to be competitive with each other and there is no clear superiority of one method
-over another \cite{Duan05whichis} .
+over another \cite{Duan05whichis}.
 Some claims that OVO strategy trains in less time than OVA strategy because they train with only portion of the dataset \cite{libsvm} but this claim have not yet been proofed by expeirments.
@@
 \subsubsection{ SVM in Machine Learning }
-SVM has been applied sucessfully to various application ranging from particle identification \cite{ParticleSVM7}.
+SVM has been applied successfully to various application ranging from particle identification \cite{ParticleSVM7}.
 Other application like Face detection \cite{faceSVM8,faceSVM7} and text categorization \cite{TextSVM7}.
@@
 The error rate was less than 5\%.
-Another example is digit recognition problem wehre an RBF network and SVM classifiers were compared on the UPUS handwritten digit recognition database\cite{ORLDataset}.
+Another example is digit recognition problem where an RBF network and SVM classifiers were compared on the UPUS handwritten digit recognition database\cite{ORLDataset}.
 SVM outperformed and RBF network on all digits.
@@
 Sherif and Ezzat has performed a comparison between SVM and several other techniques on the MNIST dataset.
-They also compared the same classification methods on arabic handwritten digits (MADBASE) \cite{ADBase9,IjdarSherifPaper}.
+They also compared the same classification methods on Arabic handwritten digits (MADBASE) \cite{ADBase9,IjdarSherifPaper}.
 Their results shows that SVM classifier outperform all other techniques on both MNIST and MADBase datasets and achieved recognition rate of over 98\%.
 Similar results was reported by \cite{empiricalcomp11,SVMInvariantComDecoste02} on the MNIST dataset.
-Ther are many other successful application SVM.
+There are many other successful application SVM.
 They have proven to be robust to noise and perform well on many tasks \cite{empiricalcomp11,Scholkopf97comparingsupport}.
 
-This superior performance because they may eliminates many of problem experienced with other inference methodoliges with other inference methodologies like neural network and decision trees.
+This superior performance because they may eliminates many of problem experienced with other inference methodologies with other inference methodologies like neural network and decision trees.
 The first problem that is eliminated was the problems with local minima.
 SVM can construct highly nonlinear classification and regression function without working about getting stuck at local minima.
-Another problem SVM eliminated is the number of parameter that are picked ot generate the model.
+Another problem SVM eliminated is the numbers of parameter that are picked to generate the model.
 For example, SVM classifier with RBF kernel will only need 2 parameters to set the model the $C$ cost parameter and $\sigma$ parameter of the RBF kernel.
@@
 \subsubsection {SVM Implementations}
-To solve SVM approach require solving a QP or LP problem.
+To solve SVM approach requires solving a QP or LP problem.
 Mathematical programming have extensively studied the LP and QP type problems.
 These techniques can be divided into three categories, methods in which evaluate and discard kernel components during learning.
-Decomposition methods in which an evolving subset of data are used and new optimization approaches that specifically utilize the structure of the SVM problem.
+Decomposition methods in which evolving subsets of data are used and new optimization approaches that specifically utilize the structure of the SVM problem.
 
@@
 
-The third approach is to directly attack the SVM  problem from and optimization perspective and create algorithms that explicitly exploits the structure of teh problem.
+The third approach is to directly attack the SVM problem from and optimization perspective and create algorithms that explicitly exploits the structure of ten problem.
 Keerthi et al.
@@
 Let us make an assumption that the user will draw a clock using two strokes.
-Figure \ref{fig:clock} show the two stroke and how the user draw them.
+Figure \ref{fig:clock} show the two strokes and how the user draws them.
 The next steps demonstrate how output of the system after each step.
@@
 Table \ref{tab:StepsStroke1} shows the detailed output of the program at each step.
-The input of this step is the stroke points and the ouptut is a list of segments that wil be used in the recognition process.
-The algorithm that compute the segmentation is demonstrated as follows:
+The input of this step is the stroke points and the output is a list of segments that will be used in the recognition process.
+The algorithm that computes the segmentation is demonstrated as follows:
 \begin{enumerate}
@@
 \begin{verbatim}
-  1. For each point in Stroke do the follwoing:
+  1. For each point in Stroke do the following:
   		1.a Compute distance from previous point.
@@
   	 2.b set average as threshold.  
-  	 2.c Create region from continous points larger than threshold. 
+  	 2.c Create region from continuous points larger than threshold. 
   	 2.d for each region do the following. 
@@
 1. Initialize the axes of ellipse  and center by bounding box information
-2. Repeat the follwoing for few loops: 
+2. Repeat the following for few loops: 
 	2.a Compute error in Equation  4.2 
@@
 
-\item Segmentation ALS1 as explained in Section \ref{sec:SwarmSegmentation}. The alogrithm will be similar to the PSO algoirhtm  in section \ref{sec:ParticleSwarmAlgorithm} 
+\item Segmentation ALS1 as explained in Section \ref{sec:SwarmSegmentation}. The alogrithm will be similar to the PSO algorithm  in section \ref{sec:ParticleSwarmAlgorithm} 
 \begin{verbatim}
@@
   2. Generate velocities for each particle for M dimension randomly. 
-  3. Repeat till Maximum number of iteration OR ErrorThreshold  
+  3. Repeat till Maximum number of iteration OR Error Threshold  
          For each particle P_i do the following 
@@
  \\ \hline
-Possible Dominate point Extraction & Computing and for details see Section\ref{sec:Preprocessing} &  Stroke points  &   &    Number of $P_{dp}$  = 11
+Possible Dominate point Extraction & for Computing details see Section\ref{sec:Preprocessing} &  Stroke points  &   &    Number of $P_{dp}$  = 11
  Pdp = [ 2 (197.0 , 145.0 ), 4 (203.0 , 125.0 ), 3 (198.0 , 140.0 ), 22 (424.0 , 91.0 ), 19 (372.0 , 72.0 ), 26 (492.0 , 164.0 ), 28 (501.0 , 191.0 ), 9 (219.0 , 97.0 ), 31 (496.0 , 239.0 ), 0 (198.0 , 153.0 ), 41 (263.0 , 378.0 )]
@@
  \\ \hline
- Segmentation & PSO Segmentation ALgS1 & $P_{dp}$  & ALgS1 &   Particles =  [  1 0 1 0 0 1 1 1 1 1 1 0 0 0 1 0 0 0 1 ] 
- The particle  Particles =  [  1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 1 0 0 0 0 1 0 ]
-
- First Polygonization PSO = [[[  S0 Line[P(201.0,179.0) to P(195.0,281.0)]  ,      S1 Line[P(195.0,281.0) to P(234.0,304.0)]  ,      S2 Line[P(234.0,304.0) to P(260.0,174.0)]  ,      S3 Line[P(260.0,174.0) to P(325.0,272.0)]  ,      S4 Line[P(325.0,272.0) to P(365.0,304.0)]  ,      S5 Line[P(365.0,304.0) to P(388.0,189.0)]  ,      S6 Line[P(388.0,189.0) to P(431.0,167.0)]  ,      S7 Line[P(431.0,167.0) to P(442.0,220.0)]  ]]]
- \\ \hline
-
-  Segmentation & PSO Segmentation ALgS2 & $P_{dp}$  & ALgS2 &   Particles =  [  1 0 1 0 0 1 1 1 1 1 1 0 0 0 1 0 0 0 1 ] 
- The particle  Particles =  [  1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 1 0 0 0 0 1 0 ]
-
- Second segmentation PSO = [[[  S0 Line[P(201.0,179.0) to P(195.0,281.0)]  ,      S1 Line[P(195.0,281.0) to P(234.0,304.0)]  ,      S2 Line[P(234.0,304.0) to P(260.0,174.0)]  ,      S3 Line[P(260.0,174.0) to P(325.0,272.0)]  ,      S4 Line[P(325.0,272.0) to P(365.0,304.0)]  ,      S5 Line[P(365.0,304.0) to P(388.0,189.0)]  ,      S6 Line[P(388.0,189.0) to P(431.0,167.0)]  ,      S7 Line[P(431.0,167.0) to P(442.0,220.0)]  ]]]
- \\ \hline
+ Segmentation & PSO Segmentation ALgS1 & $P_{dp}$  & ALgS1 (result if Figure \ref{fig:Stroke2Seg1labeled} ) & 
+Iteration 0 the global best fitness is 0.7 and error value is 675.0457637480733  and the particle is  Particles =  [  1 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 ]
+Iteration 10 the global best fitness is 0.7777777777777778 and error value is 719.5716152204503  and the particle is  Particles =  [  1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 ]
+Iteration 20 the global best fitness is 0.7777777777777778 and error value is 719.5716152204503  and the particle is  Particles =  [  1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 ]
+ After swarm  the global best fitness is 0.875 and error value is 729.248312158671  and the particle is  Particles =  [  1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 ]
+
+ F First Polygonization PSO = [ S0 Line[P(226.0,139.0) to P(241.0,209.0)]  ,      S1 Line[P(241.0,209.0) to P(266.0,226.0)]  ,      S2 Line[P(266.0,226.0) to P(296.0,128.0)]  ,      S3 Line[P(296.0,128.0) to P(312.0,216.0)]  ,    ]
+ \\ \hline
+
+  Segmentation & PSO Segmentation ALgS2 & $P_{dp}$  & ALgS2  (result if Figure \ref{fig:Stroke2Seg2labeled} )&    
+ Iteration 0 the global best fitness is 8.725462920280576E-11 and error value is 9.151490748550003  and the particle is  Particles =  [  1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 ]
+
+Iteration 10 the global best fitness is 3.126281513587432E-7 and error value is 8.61499750333895  and the particle is  Particles =  [  1 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 ]
+
+Iteration 20 the global best fitness is 3.126281513587432E-7 and error value is 8.61499750333895  and the particle is  Particles =  [  1 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 ]
+
+ After swarm  the global best fitness is 3.126281513587432E-7 and error value is 8.61499750333895  and the particle is  Particles =  [  1 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 ]
+
+  second PSO =   Segments = [   S0 Line{P(226.0,139.0) to P(238.0,183.0)},      S1 Line{P(238.0,183.0) to P(241.0,213.0)},      S2 Line{P(241.0,213.0) to P(276.0,227.0)},      S3 Line{P(276.0,227.0) to P(287.0,144.0)},      S4 Curve{P(287.0,144.0) to P(310.0,131.0),  Center ( 299.02634312229503,137.80386092517725) Radius =11.907214051459519 },  S5 Curve{P(310.0,131.0) to P(312.0,216.0),  Center ( 82.44520390380922,172.2089170470769) Radius =231.70851748053468 },  S6 Line{P(312.0,216.0) to P(336.0,224.0)}, ]
+
+ \\ \hline 
+Choosing best fit & Compare between ALgS1 and AlgS2 &  First PSO and Second PSO  &  (result if Figure \ref{fig:FinalSegmentation} ) & 
+   The error ALgS1 780.6652975044847  Algs2 161.60692855994625    Best = ALgS2  \\ \hline
 
@@
 \end{landscape}
+
+\begin{figure}
+	\centering
+	\subfigure [AlgS1 Output]{
+		\includegraphics{afterDefense/Stroke2Seg1labeled.jpg.eps}
+	\label{fig:Stroke2Seg1labeled}}
+
+	\subfigure[AlgS2 Output]{	\includegraphics{afterDefense/Stroke2Seg2labeled.jpg.eps}	\label{fig:Stroke2Seg2labeled}
+	}
+	\subfigure[Final Segmentation of Symbol]{		\label{fig:FinalSegmentation}	\includegraphics[scale=0.65]{afterDefense/FinalSegmentation.jpg.eps}
+	}
+	\caption{Segmentation Ouputs For Second Stroke}
+\end{figure}
+
+The final output of this step is list $L_s$ of segments There are  9  segments in this symbol
+
+\begin{itemize}
+	\item    Segment S0  = Ellipse ( Stroke 1)
+   \item Segment S1  =   Line ( Stroke 2)
+     \item  Segment S2 = Line ( Stroke 2)
+     \item  Segment S3  =  Line ( Stroke 2)
+     \item Segment S4  =   Line ( Stroke 2)
+  \item    Segment S5  =  Curve  ( Stroke 2)
+     \item  Segment S6  =  Curve  ( Stroke 2)
+     \item  Segment S7  =  Line   ( Stroke 2)
+     \item  Segment S8  =  Line   ( Stroke 2)
+\end{itemize}
+
 \subsection{Step by Step Recognition}
@@
 	\centering
-		\includegraphics{afterDefense/segmentatiofinal.jpg.eps}
+		\includegraphics[scale=0.85]{afterDefense/FinalSegmentation.jpg.eps}
 	\caption{Best Segmantation }
@@
 
-	The input of the step is the list of segmentation $L_s$ that is used to computer the feature vector.
+	The input of the step is the list of segmentation $L_s$ that is used to compute the feature vector.
 Figure \ref{fig:segmentatiofinal.jpg} shows the best segmentation as given by the previous step.
-This step can be summerized as firstly using $L_s$ to computer the feature vector and then introduce this vector the SVM classifier to classify it to one of the trained categories.
+This step can be summarized as firstly using $L_s$ to computer the feature vector and then introduce this vector the SVM classifier to classify it to one of the trained categories.
+The details of each step of feature extraction are detailed in the next algorithm.
+Feature extraction step is divided into four steps where each feature set is computed separately then all sets are appended to construct a final feature vector.
+
+\begin{enumerate}
+	\item Computing Structural and geometrical Features (FS1).
+This step computes number of parallel, perpendicular line and similar structural features.
+The following algorithm is used to determine these features.
+
+	\begin{verbatim}
+Initialize 
+  1. For each segment Si in list Ls  do the following. 
+      1.a If segment si is a line do the following 
+            1.a.I. Compute the slope and intercept of line si. 
+            1.a.II Increment Parallel count  if 
+                  parallel with each other line in Ls
+            1.a.III Increment PerpendicularCount if 
+                    perpendicular with each other line in Ls 
+            1.a.IV Increment Intersection Count if 
+                    intersected with any other line. 
+                    1.a.II.A  if Intersection found determine 
+                              type of intersection or (X,T,L)
+      1.b If segment si is curve do the following. 
+          1.b.I Increment Intersect if
+               Intersect other line or curve
+    1.b.II Increment number 
+                 of curves and primitives. 
+4. return Feature_Vector_FS1
+\end{verbatim}
+
+	\item Rubine Feature Set (FS2).  Compute Rubine features as in \cite{gestureexample12} 
+	\item Statistical Features (FS3) compute the moments as in \cite{zernike61}
+	\item Global shape properties set (FS4). Firstly compute the convex hull of the stroke points then compute the rest of the features.
+
+	\item Introduce the computed feature vector to the classifier and check the label of the detected symbol. 
+\end{enumerate}
+
+\begin{landscape}
+\begin{scriptsize}
+
+ \begin{longtable}{|p{2cm}|p{2cm}|p{2cm}|p{2cm}|p{13cm}|}
+\caption{Detailed Output of System in Each Step of Recognition}
+\label{tab:StepsReg} \\
+
+\hline 
+\multicolumn{1}{|p{2cm}|}{\textbf{Step}} & 
+\multicolumn{1}{p{2cm}|}{\textbf{Description}} &
+\multicolumn{1}{p{2cm}|}{\textbf{Inputs}} &
+\multicolumn{1}{p{2cm}|}{\textbf{Details}} &
+\multicolumn{1}{p{13cm}|}{\textbf{Output}} 
+\\ \hline 
+\endfirsthead
+
+\hline
+\multicolumn{3}{c}%
+{{\bfseries \tablename\ \thetable{} -- continued from previous page}} \\
+\multicolumn{1}{|p{2cm}|}{\textbf{Step}} & 
+\multicolumn{1}{p{2cm}|}{\textbf{Description}} &
+ \multicolumn{1}{p{2cm}|}{\textbf{Inputs}} &
+ \multicolumn{1}{p{2cm}|}{\textbf{Details}} &
+ \multicolumn{1}{p{13cm}|}{\textbf{Output}} 
+\\ \hline 
+\endhead
+
+ Feature Extraction of FS1 & Preprocessing of Structural Featues  &   There are  9  segments in this symbol $L_s$ = 
+   Segment S0  = Ellipse
+   Segment S1  = Line
+   Segment S2  = Line
+   Segment S3  = Line
+   Segment S4  = Line
+   Segment S5  = Curve
+   Segment S6  = Curve
+   Segment S7  = Line
+   Segment S8  = Line  &  (see First Algorithm in Section \ref{sec:steprec} )  &    Line from  24 x= 241.0y=213.0  to 49 x2= 276.0y2=227.0  is\textbf{ parallel} to Line from  163 x= 312.0y=216.0  to 185 x2= 336.0y2=224.0
+
+ Line from  0 x= 226.0y=139.0  to 14 x2= 238.0y2=183.0 \textbf{ intersects }Line from  14 x= 238.0y=183.0  to 24 x2= 241.0y2=213.0
+
+ Line from  0 x= 226.0y=139.0  to 14 x2= 238.0y2=183.0  \textbf{intersects} Line from  24 x= 241.0y=213.0  to 49 x2= 276.0y2=227.0
+
+ Line from  0 x= 226.0y=139.0  to 14 x2= 238.0y2=183.0 \textbf{ intersects} Line from  49 x= 276.0y=227.0  to 98 x2= 287.0y2=144.0
+
+ Line from  0 x= 226.0y=139.0  to 14 x2= 238.0y2=183.0\textbf{  intersects }Line from  163 x= 312.0y=216.0  to 185 x2= 336.0y2=224.0
+
+ Line from  0 x= 226.0y=139.0  to 14 x2= 238.0y2=183.0 \textbf{ intersects} Line from  185 x= 336.0y=224.0  to 223 x2= 344.0y2=137.0
+ %Line from  14 x= 238.0y=183.0  to 24 x2= 241.0y2=213.0  intersects Line from  24 x= 241.0y=213.0  to 49 x2= 276.0y2=227.0
+ %Line from  14 x= 238.0y=183.0  to 24 x2= 241.0y2=213.0  intersects Line from  49 x= 276.0y=227.0  to 98 x2= 287.0y2=144.0
+ %Line from  14 x= 238.0y=183.0  to 24 x2= 241.0y2=213.0  intersects Line from  163 x= 312.0y=216.0  to 185 x2= 336.0y2=224.0
+ %Line from  14 x= 238.0y=183.0  to 24 x2= 241.0y2=213.0  intersects Line from  185 x= 336.0y=224.0  to 223 x2= 344.0y2=137.0
+ %Line from  24 x= 241.0y=213.0  to 49 x2= 276.0y2=227.0  intersects Line from  49 x= 276.0y=227.0  to 98 x2= 287.0y2=144.0
+ %Line from  24 x= 241.0y=213.0  to 49 x2= 276.0y2=227.0  intersects Line from  163 x= 312.0y=216.0  to 185 x2= 336.0y2=224.0
+ %Line from  24 x= 241.0y=213.0  to 49 x2= 276.0y2=227.0  intersects Line from  185 x= 336.0y=224.0  to 223 x2= 344.0y2=137.0
+ %Line from  49 x= 276.0y=227.0  to 98 x2= 287.0y2=144.0  intersects Line from  163 x= 312.0y=216.0  to 185 x2= 336.0y2=224.0
+
+ Line from  49 x= 276.0y=227.0  to 98 x2= 287.0y2=144.0  intersects Line from  185 x= 336.0y=224.0  to 223 x2= 344.0y2=137.0
+
+ Line from  163 x= 312.0y=216.0  to 185 x2= 336.0y2=224.0  intersects Line from  185 x= 336.0y=224.0  to 223 x2= 344.0y2=137.0
+
+\\ \hline
+
+ FS1 & Structural and geometrical Features & Segment List $L_s$  &  & 
+ The final Feature vector of the FS1==>
+ature Number of primitives 3.0   ,    Feature Number of segments 9.0   ,    Feature Line count 6.0   ,    Feature Curves Count 3.0   ,    Feature Ellipse Count 1.0   ,    Feature Intersection T type -1.0   ,    Feature Intersection L type 4.0   ,    Feature Intersection X type -1.0   ,    Feature Parallel Count 1.0   ,    Feature PerPendicular Count -1.0   ,    Feature Intersection Lines Count 4.0   ,    Feature Min Radius 1.7976931348623157E308   ,    Feature Max Radius 231.70851748053468
+
+ \\ \hline
+FS2  & Rubine Features & Segment List $L_s$  &   see \cite{gestureexample12}   & 
+\begin{scriptsize}
+ Feature Rubine 0 -1.0   ,    Feature Rubine 1 -1.0   ,    Feature Rubine 2 477.6107201476952   ,    Feature Rubine 3 0.6368010415482491   ,    Feature Rubine 4 169.49926253526885   ,    Feature Rubine 5 0.8908593331996381   ,    Feature Rubine 6 0.454279262624981   ,    Feature Rubine 7 1179.0172885029808   ,    Feature Rubine 8 4.712388980384688   ,    Feature Rubine 9 70.80157037633609   ,    Feature Rubine 10 55.4044569978846   ,    Feature Rubine 11 9.140625   ,    Feature Rubine 12 41783.0   ,
+\end{scriptsize}
+ \\ \hline
+
+FS3 & Momments Computation   &  Segment List $L_s$  &  see \cite{zernike61}  &
+
+\begin{scriptsize}
+   Feature Zernike moments 0 158.87520185460517   ,    Feature Zernike moments 0 7.007557168632727   ,    Feature Zernike moments 1 34.26844061831789   ,    Feature Zernike moments 1 6.658086415801404   ,    Feature Zernike moments 2 85.70032052974454   ,    Feature Zernike moments 2 9.441591321498212   ,    Feature Zernike moments 3 11.73170471301825   ,    Feature Zernike moments 3 66.52997844880663   ,    Feature Zernike moments 4 20.42774127507497   ,    Feature Zernike moments 4 10.196341027811322   ,    Feature Zernike moments 5 52.821799077785066   ,    Feature Zernike moments 5 12.078236804336935   ,    Feature Zernike moments 6 27.748912050714   ,    Feature Zernike moments 6 9.870152472028517   ,    Feature Zernike moments 7 26.079803494397133   ,    Feature Zernike moments 7 28.109003184347657   ,    Feature Zernike moments 8 23.32097186060531   ,    Feature Zernike moments 8 7.632681250023639   ,    Feature Zernike moments 9 6.5841386188472395   ,    Feature Zernike moments 9 51.34195037191517   ,    Feature Zernike moments 10 30.122054353503337   ,    Feature Zernike moments 10 26.902895199361584   ,    Feature Zernike moments 11 1.8295982451659387   ,    Feature Zernike moments 11 29.93196640188875   ,    Feature Zernike moments 12 64.0879253626976   ,    Feature Zernike moments 12 31.811780133580722   ,    Feature Zernike moments 13 23.614051070558084   ,    Feature Zernike moments 13 1.8043229719502216   ,    Feature Zernike moments 14 113.92181136864328   ,    Feature Zernike moments 14 67.26888519955787   ,    Feature Zernike moments 15 47.7397132180995   ,    Feature Zernike moments 15 31.490125533238682   ,    Feature Zernike moments 16 10.320112066817984   ,    Feature Zernike moments 16
+\end{scriptsize}
+
+  \\ \hline
+FS4   &  &Segment List $L_s$   &  &  \begin{scriptsize} 
+ The convex hull    [ X = 383.0   Y = 146.0  Time = 1250452568824,  X = 380.0   Y = 139.0  Time = 1250452568816,  X = 368.0   Y = 119.0  Time = 1250452568800,  X = 363.0   Y = 113.0  Time = 1250452568792,  X = 352.0   Y = 101.0  Time = 1250452568776,  X = 348.0   Y = 97.0  Time = 1250452568768,  X = 342.0   Y = 92.0  Time = 1250452568760,  X = 335.0   Y = 88.0  Time = 1250452568752,  X = 329.0   Y = 85.0  Time = 1250452568744, % X = 322.0   Y = 83.0  Time = 1250452568736,  X = 302.0   Y = 78.0  Time = 1250452568728,  X = 293.0   Y = 77.0  Time = 1250452568720,  X = 282.0   Y = 76.0  Time = 1250452568712,  X = 250.0   Y = 76.0  Time = 1250452568688,  X = 239.0   Y = 77.0  Time = 1250452568680,  X = 229.0   Y = 79.0  Time = 1250452568672,  X = 220.0   Y = 81.0  Time = 1250452568664,  X = 217.0   Y = 83.0  Time = 1250452568656,  X = 201.0   Y = 96.0  Time = 1250452568648,  X = 196.0   Y = 103.0  Time = 1250452568640,  X = 186.0   Y = 118.0  Time = 1250452568624,  X = 182.0   Y = 127.0  Time = 1250452568608,  X = 180.0   Y = 209.0  Time = 1250452569139,  X = 180.0   Y = 247.0  Time = 1250452569092,  X = 181.0   Y = 253.0  Time = 1250452569080,  X = 185.0   Y = 260.0  Time = 1250452569073,  X = 190.0   Y = 265.0  Time = 1250452569056,  X = 194.0   Y = 267.0  Time = 1250452569047,  X = 213.0   Y = 274.0  Time = 1250452569031,  X = 220.0   Y = 276.0  Time = 1250452569023,  X = 256.0   Y = 284.0  Time = 1250452568995,  X = 266.0   Y = 284.0  Time = 1250452568976,
+ ...................................................X = 274.0   Y = 283.0  Time = 1250452568969,  X = 282.0   Y = 281.0  Time = 1250452568960,  X = 290.0   Y = 278.0  Time = 1250452568952,  X = 298.0   Y = 274.0  Time = 1250452568944,  X = 319.0   Y = 262.0  Time = 1250452568936,  X = 339.0   Y = 243.0  Time = 1250452568912,  X = 357.0   Y = 222.0  Time = 1250452568889,  X = 368.0   Y = 202.0  Time = 1250452568872,  X = 378.0   Y = 183.0  Time = 1250452568864,  X = 384.0   Y = 167.0  Time = 1250452568848,  X = 384.0   Y = 160.0  Time = 1250452568840]
+ Area of convex hull =  -34160.5   preimiter of hull 656.1576413855265
+
+ --------------------------------------------------------------------------------
+
+Final Feature vector for the FS4: 
+  Feature Centroid time 1.2504525968640884E12  , Feature Centroid time difference 28319.088339222613   ,    Feature area of convexhull 34160.5   ,    Feature area of convexhull/area of symbol 1.2365344240932454   ,    Feature N.Points ConvexHull/ N.
+points symbol 0.15302491103202848   ,    Feature Convext Perimeter/symbol perimeter 0.556529278903672   ,    Feature  percentage of convex area to bounding box area 19.493542609351433   ,    Feature area  35519.5   ,    Feature diff area of symbol to area of bounding box 16.290771116138764   ,    Feature  Ration between w/h  0.9807692307692307   ,    Feature  Log of ration between hight and width -0.008433167536862813   ,
+ Feature  Log of computed area 4.550466843684004   ,    Feature  abs of area 27626.0   ,    Feature log of  abs of area 4.441318007475848   ,    Feature  Density(length/size)  4.046853548165911   ,    Feature Mean CentroidalRadius x 272.48042704626334   ,    Feature Mean CentroidalRadius y 177.19928825622776   ,    Feature Mean t 1.2504525970626514E12   ,    Feature Mean time difference 148.69395017793593   ,    Feature Mean Radius 62.000799881019084
+  \end{scriptsize}
+    \\ \hline
+ Feature Vector & Appends the selected Feature FS1, FS2, FS3, FS4 & Previous Steps  & &
+
+   \begin{scriptsize}
+   Feature Number of primitives 3.0   ,    Feature Number of segments 9.0   ,    Feature Line count 6.0   ,    Feature Curves Count 3.0   ,    Feature Ellipse Count 1.0   ,    Feature Intersection T type -1.0   ,    Feature Intersection L type 4.0   ,    Feature Intersection X type -1.0   ,    Feature Parallel Count 1.0   ,    Feature PerPendicular Count -1.0   ,    Feature Intersection Lines Count 4.0   ,    Feature Min Radius 1.7976931348623157E308   ,    Feature Max Radius 231.70851748053468   ,    Feature Centroid x 92.54770318021201   ,    Feature Centroid y 101.3851590106007   ,    Feature Centroid time 1.2504525968640884E12   ,    Feature Centroid time difference 28319.088339222613   ,    Feature area of convexhull 34160.5   ,    Feature area of convexhull/area of symbol 1.2365344240932454   ,    Feature N.Points ConvexHull/ N.
+points symbol 0.15302491103202848   ,    Feature Convext Perimeter/symbol perimeter 0.556529278903672   ,    Feature  percentage of convex area to bounding box area 19.493542609351433   ,    Feature area  35519.5   ,    Feature diff area of symbol to area of bounding box 16.290771116138764   ,    Feature  Ration between w/h  0.9807692307692307   ,    Feature  Log of ration between hight and width -0.008433167536862813   ,    Feature Zernike moments 0 158.87520185460517   ,    Feature Zernike moments 0 7.007557168632727   ,    Feature Zernike moments 1 34.26844061831789   ,    Feature Zernike moments 1 6.658086415801404   ,    Feature Zernike moments 2 85.70032052974454   ,    Feature Zernike moments 2 9.441591321498212   ,    Feature Zernike moments 3 11.73170471301825   ,    Feature Zernike moments 3 66.52997844880663   ,    Feature Zernike moments 4 20.42774127507497   ,    %Feature Zernike moments 4 10.196341027811322   ,    Feature Zernike moments 5 52.821799077785066   ,    Feature Zernike moments 5 12.078236804336935   ,    Feature Zernike moments 6 27.748912050714   ,    Feature Zernike moments 6 % 9.870152472028517   ,    Feature Zernike moments 7 26.079803494397133   ,    Feature Zernike moments 7 28.109003184347657   ,    Feature Zernike moments 8 23.32097186060531   ,    Feature Zernike moments 8 7.632681250023639   ,    Feature Zernike moments 9 6.5841386188472395   ,    Feature Zernike moments 9 51.34195037191517   ,    Feature Zernike moments 10 30.122054353503337   ,    Feature Zernike moments 10 26.902895199361584   ,    Feature Zernike moments 11 1.8295982451659387   ,    Feature Zernike moments 11 29.93196640188875   ,    Feature Zernike moments 12 64.0879253626976   ,    Feature Zernike moments 12 31.811780133580722   ,    Feature Zernike moments 13 23.614051070558084   ,    Feature Zernike moments 1    1.8043229719502216   ,
+%Feature Zernike moments 14 113.92181136864328   ,    Feature Zernike moments 14 67.26888519955787   ,    Feature Zernike moments 15 47.7397132180995   ,    Feature Zernike moments 15 31.490125533238682   ,
+Feature Zernike moments 16 10.320112066817984   ,    Feature Zernike moments 16 4.007106145033957   ,    Feature Rubine 0 -1.0   ,    Feature Rubine 1 -1.0   ,    Feature Rubine 2 477.6107201476952   ,    Feature Rubine 3 0.6368010415482491   ,    Feature Rubine 4 169.49926253526885   ,    Feature Rubine 5 0.8908593331996381   ,    Feature Rubine 6 0.454279262624981   ,    Feature Rubine 7 1179.0172885029808   ,    Feature Rubine 8 4.712388980384688   ,    Feature Rubine 9 70.80157037633609   ,    Feature Rubine 10 55.4044569978846   ,    Feature Rubine 11 9.140625   ,    Feature Rubine 12 41783.0   ,    Feature  Log of computed area 4.550466843684004   ,    Feature  abs of area 27626.0   ,    Feature log of  abs of area 4.441318007475848   ,    Feature  Density(length/size)  4.046853548165911   ,    Feature Mean CentroidalRadius x 272.48042704626334   ,    Feature Mean CentroidalRadius y 177.19928825622776   ,    Feature Mean t 1.2504525970626514E12   ,    Feature Mean time difference 148.69395017793593   ,    Feature Mean Radius 62.000799881019084   ,
+   \end{scriptsize} \\ \hline
+   Classification  & SVM classifier decision & Feature Vector  &  & 
+   Symbol = Clock.
+
+ \\ \hline
+
+ 		\end{longtable}
+%}
+
+%\end{table}
+
+\end{scriptsize}
+
+\end{landscape}

### 2009-09-06 trunk/draft/ConFut.tex
after the changes in examples

@@
 
-The system was tested on three different datasets, a simple presentation symbols, electrical symbols and logic design dataset.
+The system was tested on three different datasets, simple presentation symbols, electrical symbols and logic design dataset.
 The number and the complexity of symbols recognized by the system varied from simple to complex but did not affect the final recognition of the system.
@@
 The result is justified because logic design dataset contains symbols that are differentiated by number of curves in OR-gate and XOr gate, therefore representing the symbol as lines will remove important information.
+
+The system uses different set of feature that represents shape geometrical and global properties.
+Literature review showed that system that relies only on global shape properties fails to identify the shapes that have small geometrical difference (ie.
+Xor gate and Nor gate).
+Using geometrical and stroke based features with the global shape properties helps to improve the final recognition rate.
 
@@
 %Each new stroke is checked if it can be a symbol or is a part of already drawn un completed symbol.
+\appendix
+\chapter{Appendix: Step by Step Example Details}
+%\addcontentsline{toc}{chapter}{Appendix: Step by Step Example Details}
+\chaptermark{Appendix}
+\markboth{Appendix}{Appendix}
+\label{ChapterstepExample}%\chapter{Step by Step Example}
+%In this chapter a step by step example is presented to clarify the presented system.
+Let us make an assumption that the user will draw a clock using two strokes.
+Figure \ref{fig:clock} show the two strokes and how the user draws them.
+The next steps demonstrate how output of the system after each step.
+The first section \ref{sec:stepseg} describe how the segmentation is done step by step.
+Section \ref{sec:steprec} shows the details of feature calculation and the classification process.
+
+\section{Step by Step Segmentation}
+\label{apsec:stepseg}
+
+In this step the user draw one or more stroke to draw a symbol.
+Figure \ref{fig:clock} illustrate the sequence and direction of user draw.
+In this section, each stroke user draw will be labeled as it illustrated in Figure \ref{fig:clockstroke1} and Figure \ref{fig:clockstroke2} .
+Table \ref{tab:StepsStroke1Details} shows the detailed output of the program at each step.
+The input of this step is the stroke points and the output is a list of segments that will be used in the recognition process.
+The algorithm that computes the segmentation is demonstrated as follows:
+
+%\begin{enumerate}
+
+%\item Extracting $P_dp$ as in Section \ref{sec:Preprocessing} and Algorithm \ref{extractpdp}
+\begin{algorithm}
+\caption{Extracting $P_dp$}
+\label{extractpdp}
+\begin{verbatim}
+  1. For each point in Stroke do the following:
+  		1.a Compute distance from previous point.
+  		1.b Compute Time difference from previous point and add it to Time_list.  
+  		1.c Compute Velocity and add it to Vel_list. 
+  		1.d Compute Direction and add it to Dir_list.
+  	 	1.e Compute Curvature and add it to Cur_list.(Eq.4.1) 
+  	end (for loop)
+  2. For each  list computed in previous step do the follwing:   
+  	 2.a Compute average of all points. 
+  	 2.b set average as threshold.  
+  	 2.c Create region from continuous points larger than threshold. 
+  	 2.d for each region do the following. 
+  	 	   2.d.1 find the maximum 
+  	 	   2.d.1 Add point to P_dp 
+  	  end
+     end 
+\end{verbatim}	
+\end{algorithm}
+%\item Ellipse Fitting: which is an attempt to fit as an ellipse as in Section \ref{sec:EllipseDetection} 
+%he axes of the ellipse are estimated as the $width/2$ and $height/2$ of the stroke bounding box where $width$ is width and $height$ is height of the bonding box.
+\begin{algorithm}
+\caption{Ellipse Fitting Algorithm}
+\label{ellipseFitAlg}
+\begin{verbatim}
+1. Initialize the axes of ellipse  and center by bounding box information
+2. Repeat the following for few loops: 
+	2.a Compute error in Equation  4.2 
+	2.b Update center of ellipse and axis from error. 
+3. Compute percent in Equation 4.4 
+4. Computer eff in Equation 4.3 
+5. if eff < 0.2  then fit as ellipse otherwise move to segment. 
+\end{verbatim}	
+\end{algorithm}
+%\item Segmentation ALS1 as explained in Section \ref{sec:SwarmSegmentation}. The alogrithm will be similar to the PSO algorithm  in section \ref{sec:ParticleSwarmAlgorithm}.
+
+\begin{algorithm}
+\caption{ First Swarm Algorithm}
+\label{AlgS1alg}
+\begin{verbatim}
+Initialize 
+  1. Generate N particles with M dimensions (P_i...P_N) 
+  2. Generate velocities for each particle for M dimension randomly. 
+  3. Repeat till Maximum number of iteration OR Error Threshold  
+         For each particle P_i do the following 
+         3.a   Compute fitness  (fitness) 
+         				3.a.1 Compute the Error of approximation as in Eq.4.5
+         				3.a.2. Computer the fitness function as in Eq. 4.4 
+         3.b   Determine Lbest  
+               if (lbest_fitness<fitness)
+                  3.b.1   lbest=P_i
+         3.c      Determine Gbest.
+               if (lGest_fitness<fitness)
+                  3.c.1   lGest=P_i
+         3.d   Compute the Velocity for each dimension using Eq.3.2 
+         3.e   Bound Velocity using Eq.3.4
+         3.f   Update particle values using Eq. 3.3
+         3.g   Refine the location of the particle 
+         end .
+     end 
+4. return Gbest
+\end{verbatim}
+\end{algorithm}	
+%\item Segmentation ALS2  as explained in Section \ref{sec:PolygonDivisionAlgorithm}
+\begin{algorithm}
+\caption{ Second Swarm Algorithm}
+\label{AlgS2alg}
+\begin{verbatim}
+Initialize 
+  1. Generate N particles with M dimensions (P_i...P_N) 
+  2. Generate velocities for each particle for M dimension randomly. 
+  3. Repeat till Maximum number of iteration OR Error Threshold  
+         For each particle P_i do the following 
+         3.a   Compute fitness  (fitness) 
+              3.a.1 For each Segment in the particle do the following
+         				    3.a.1.I   Error of approximation as Line (Eq. 4.7 - 4.9 )
+         			      3.a.1.II  Error of approximation as circular arc (Eq. 4.10 - 4.19 )
+         				    3.a.1.III Computer minimum of both approximation 
+         			3.a.2 Compute Error of the particle as in Eq.  4.20
+         			3.a.3 Compute fitness of the particle as in Eq. 4.21
+         3.b   Determine Lbest  
+               if (lbest_fitness<fitness)
+                  3.b.1   lbest=P_i
+         3.c      Determine Gbest.
+               if (lGest_fitness<fitness)
+                  3.c.1   lGest=P_i
+         3.d   Compute the Velocity for each dimension using Eq.3.2 
+         3.e   Bound Velocity using Eq.3.4
+         3.f   Update particle values using Eq. 3.3
+         3.g   Refine the location of the particle 
+         end .
+     end 
+4. return Gbest
+\end{verbatim}	
+\end{algorithm}
+
+\begin{landscape}
+\begin{scriptsize}
+%\begin{table}
+	%\centering 
+%\scalebox{0.6}{		
+\begin{longtable}{|p{2cm}|p{2cm}|p{2cm}|p{2cm}|p{13cm}|}
+\caption{Detailed Output of System in Each Step of Segmentation}
+\label{tab:StepsStroke1Details} \\
+
+\hline 
+\multicolumn{1}{|p{2cm}|}{\textbf{Step}} & 
+\multicolumn{1}{p{2cm}|}{\textbf{Description}} &
+\multicolumn{1}{p{2cm}|}{\textbf{Inputs}} &
+\multicolumn{1}{p{2cm}|}{\textbf{Details}} &
+\multicolumn{1}{p{13cm}|}{\textbf{Output}} 
+\\ \hline 
+\endfirsthead
+
+\multicolumn{3}{c}%
+{{\bfseries \tablename\ \thetable{} -- continued from previous page}} \\
+\hline 
+\multicolumn{1}{|p{2cm}|}{\textbf{Step}} & 
+\multicolumn{1}{p{2cm}|}{\textbf{Description}} &
+ \multicolumn{1}{p{2cm}|}{\textbf{Inputs}} &
+ \multicolumn{1}{p{2cm}|}{\textbf{Details}} &
+ \multicolumn{1}{p{13cm}|}{\textbf{Output}} 
+\\ \hline 
+\endhead
+
+%Step & Discription of Step & Input & Steps & Output \\ \hline
+Stroke 1 & Extracting point and computing boundary data & -- &  see Section \ref{sec:Preprocessing} & 
+\begin{scriptsize}
+Stroke has 62 points = [P(198.0,153.0), P(197.0,150.0), P(197.0,145.0), P(198.0,140.0), P(203.0,125.0), P(205.0,120.0) , P(205.0,117.0), P(210.0,111.0) , P(214.0,103.0) , P(219.0,97.0), P(225.0,89.0) , P(229.0,85.0) , P(242.0,77.0) , P(247.0,73.0) , P(275.0,64.0) , P(287.0,62.0) , P(325.0,62.0) , P(333.0,63.0) , P(362.0,68.0) , P(372.0,72.0) , P(391.0,77.0) , P(406.0,83.0) , P(424.0,91.0) , P(440.0,100.0), P(474.0,133.0) , P(485.0,149.0) , P(492.0,164.0) , P(499.0,182.0) , P(501.0,191.0) , P(501.0,207.0) , P(499.0,222.0) , P(496.0,239.0) , P(471.0,292.0) , P(467.0,302.0) , P(456.0,315.0) , P(441.0,329.0) , P(427.0,339.0) , P(418.0,344.0) , P(398.0,354.0) , P(336.0,371.0) , P(283.0,376.0) , P(263.0,378.0) , P(242.0,378.0) , P(232.0,376.0) , P(213.0,373.0) , P(194.0,367.0) , P(177.0,362.0) , P(138.0,343.0) , P(129.0,336.0) , P(124.0,331.0) , P(120.0,321.0) , P(115.0,310.0) , P(113.0,297.0) , P(112.0,284.0) , P(114.0,256.0) , P(121.0,219.0) , P(128.0,202.0) , P(137.0,184.0) , P(143.0,172.0) , P(154.0,161.0) , P(172.0,146.0) , P(172.0,146.0)]
+
+ Bonding box is  = Corner P(112.0, 62.0), w= 389.0, h=  316.0
+ number of point in this stroke is   62
+\end{scriptsize}
+ \\ \hline
+Possible Dominate point Extraction & for Computing details see Section\ref{sec:Preprocessing} &  Stroke points  &   &    Number of $P_{dp}$  = 11
+ Pdp = [ 2 (197.0 , 145.0 ), 4 (203.0 , 125.0 ), 3 (198.0 , 140.0 ), 22 (424.0 , 91.0 ), 19 (372.0 , 72.0 ), 26 (492.0 , 164.0 ), 28 (501.0 , 191.0 ), 9 (219.0 , 97.0 ), 31 (496.0 , 239.0 ), 0 (198.0 , 153.0 ), 41 (263.0 , 378.0 )]
+
+\\ \hline 
+Segmentation & Segmentation of Ellipse & $P_{dp}$   &  Ellipse fitting  &
+
+  The circle is   a = 103.05715108084719 ,b = 103.06851874747797 , Center( 283.03365107691775,  181.05035287667172 )   with Error =  1.9471676673301757 , Percent =1.0695193086531243
+eff  = 0.27463462099275116  
+ Ellipse detected...
+
+ \\ \hline
+
+Stroke 2 & Extracting point and computing boundary data & -- &  see Section \ref{sec:Preprocessing} &
+\begin{scriptsize}
+Stroke has 205 points = [P(201.0,179.0) , P(202.0,180.0) , P(202.0,182.0) , P(202.0,185.0) , P(202.0,187.0) , P(202.0,190.0) , P(202.0,192.0) , P(202.0,196.0) , P(202.0,202.0) , P(201.0,206.0) , P(200.0,219.0) , P(198.0,231.0) , P(198.0,234.0) , P(198.0,238.0) , P(197.0,244.0) , P(197.0,248.0) , P(197.0,251.0) , P(196.0,258.0) , %P(196.0,260.0) , P(196.0,263.0) , P(195.0,265.0) , P(195.0,266.0) , P(195.0,269.0) , P(195.0,271.0) , P(195.0,273.0) , P(195.0,278.0) , P(195.0,279.0) , P(195.0,281.0) , P(195.0,282.0) , P(195.0,284.0) , P(195.0,285.0) , P(195.0,288.0) , P(195.0,290.0) , P(195.0,295.0) , P(195.0,297.0) , P(195.0,298.0) , P(196.0,298.0) , P(196.0,299.0) , P(196.0,300.0) , P(196.0,301.0) , P(197.0,301.0) , P(198.0,301.0) , P(200.0,301.0) , P(203.0,301.0) , %P(206.0,301.0) , P(208.0,301.0) , P(213.0,301.0) , P(217.0,302.0) , P(219.0,302.0) , P(223.0,303.0) , P(234.0,304.0) , P(237.0,304.0) , P(240.0,306.0) , P(246.0,306.0) , P(247.0,307.0) , P(248.0,307.0) , P(249.0,307.0) , P(250.0,307.0) , P(250.0,306.0) , P(250.0,305.0) , P(250.0,304.0) , P(250.0,302.0) , P(251.0,299.0) , P(251.0,296.0) , P(251.0,294.0) , P(251.0,289.0) , P(252.0,285.0) , P(253.0,279.0) , P(253.0,274.0) , P(253.0,271.0) , P(253.0,264.0) , P(253.0,257.0) , P(254.0,252.0) , P(254.0,241.0) , P(254.0,237.0) , P(255.0,232.0) , P(255.0,229.0) , P(255.0,225.0) , P(256.0,212.0) , P(256.0,209.0) , P(256.0,206.0) , P(257.0,205.0) , P(257.0,200.0) , P(257.0,197.0) , P(258.0,195.0) , P(258.0,194.0) , P(258.0,189.0) , P(258.0,187.0) , P(258.0,186.0) , P(259.0,184.0) , P(259.0,181.0) , P(259.0,179.0) , P(259.0,178.0) , P(260.0,176.0) , P(260.0,174.0) , P(260.0,173.0) , P(260.0,172.0) , P(260.0,171.0) , P(260.0,169.0) , P(260.0,168.0) , P(260.0,167.0) , P(261.0,166.0) , P(261.0,165.0) , P(261.0,164.0) , P(261.0,162.0) , P(262.0,162.0) , P(263.0,162.0) , P(264.0,162.0) , P(268.0,162.0) , P(275.0,162.0) , P(279.0,162.0) , P(289.0,162.0) , P(291.0,162.0) , P(301.0,163.0) , P(305.0,164.0) , P(312.0,164.0) , P(315.0,165.0) , P(318.0,165.0) , P(319.0,165.0) , P(319.0,166.0) , P(320.0,168.0) , P(320.0,171.0) , P(320.0,173.0) , P(320.0,176.0) , P(321.0,182.0) , P(321.0,191.0) , P(322.0,198.0) , P(322.0,211.0) , P(323.0,226.0) , P(324.0,229.0) , P(324.0,238.0) , P(324.0,263.0) , P(325.0,272.0) , P(325.0,281.0) , P(325.0,288.0) , P(325.0,291.0) , P(325.0,297.0) , P(325.0,299.0) , P(325.0,301.0) , P(325.0,305.0) , P(324.0,306.0) , P(324.0,307.0) , P(323.0,308.0) , P(324.0,307.0) , P(325.0,307.0) , P(328.0,306.0) , P(332.0,306.0) , P(339.0,306.0) , P(347.0,305.0) , P(354.0,304.0) , P(365.0,304.0) , P(376.0,304.0) , P(379.0,304.0) , P(385.0,304.0) , P(391.0,304.0) , P(392.0,304.0) , P(393.0,304.0) , P(393.0,302.0) , P(393.0,301.0) , P(394.0,297.0) , P(394.0,290.0) , P(393.0,284.0) , P(393.0,274.0) , P(392.0,258.0) , P(391.0,245.0) , P(391.0,233.0) , P(389.0,213.0) , P(388.0,198.0) , P(388.0,189.0) , P(388.0,184.0) , P(388.0,178.0) , P(388.0,177.0) , P(388.0,176.0) , P(388.0,174.0)
+..................................................................................................................
+.................................................................................
+, P(388.0,173.0) , P(388.0,171.0) , P(388.0,170.0) , P(388.0,168.0) , P(388.0,167.0) , P(388.0,166.0) , P(388.0,165.0) , P(389.0,165.0) , P(390.0,165.0) , P(395.0,165.0) , P(401.0,165.0) , P(406.0,165.0) , P(418.0,166.0) , P(424.0,167.0) , P(426.0,167.0) , P(429.0,167.0) , P(430.0,167.0) , P(431.0,167.0) , P(432.0,167.0) , P(433.0,169.0) , P(434.0,170.0) , P(434.0,174.0) , P(436.0,185.0) , P(440.0,200.0) , P(442.0,220.0) , P(442.0,224.0) , P(446.0,267.0) , P(447.0,279.0) , P(448.0,286.0) , P(448.0,287.0) , P(448.0,287.0)]
+
+ Bonding box is  = Corner P(195.0, 162.0), w= 253.0, h=  146.0
+ \end{scriptsize}
+  \\ \hline
+  & Possible Dominate point Extraction for details see Section\ref{sec:Preprocessing} &  Stroke points  & 1. compute the $P_{pd}$  & \begin{scriptsize}   Number of $P_{pd}$  =  36
+
+ Pdp = [ 25 (195.0 , 278.0 ), 30 (195.0 , 285.0 ), 33 (195.0 , 295.0 ), 27 (195.0 , 281.0 ), 34 (195.0 , 297.0 ), 46 (213.0 , 301.0 ), 47 (217.0 , 302.0 ), 50 (234.0 , 304.0 ), 28 (195.0 , 282.0 ), 94 (260.0 , 174.0 ), 93 (260.0 , 176.0 ), 95 (260.0 , 173.0 ), 98 (260.0 , 169.0 ), 48 (219.0 , 302.0 ), 99 (260.0 , 168.0 ), 109 (275.0 , 162.0 ), 113 (301.0 , 163.0 ), 94 (260.0 , 174.0 ), 132 (325.0 , 272.0 ), 136 (325.0 , 297.0 ), 112 (291.0 , 162.0 ), 137 (325.0 , 299.0 ), 110 (279.0 , 162.0 ), 144 (325.0 , 307.0 ), 147 (339.0 , 306.0 ), 150 (365.0 , 304.0 ), 132 (325.0 , 272.0 ), 168 (388.0 , 189.0 ), 171 (388.0 , 177.0 ), 174 (388.0 , 173.0 ), 146 (332.0 , 306.0 ), 175 (388.0 , 171.0 ), 183 (395.0 , 165.0 ), 186 (418.0 , 166.0 ), 169 (388.0 , 184.0 ), 191 (431.0 , 167.0 ) ]
+
+  \end{scriptsize}
+    \\ \hline
+ \\ \hline 
+Segmentation & Segmentation of Ellipse & $P_{dp}$   &  Ellipse fitting  & 
+  The Ellipse is   a = 125.5 ,b = 125.5 , Center( 320.5,  235.0 )   with Error =  7.074966523515405 , Percent =1.1481777694994926 certainty  =  0.08114368920921669  
+ \\ \hline
+ Segmentation & PSO Segmentation ALgS1 & $P_{dp}$  & ALgS1 (result if Figure \ref{fig:Stroke2Seg1labeled} ) & 
+Iteration 0 the global best fitness is 0.7 and error value is 675.0457637480733  and the particle is  Particles =  [  1 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 ]
+Iteration 10 the global best fitness is 0.7777777777777778 and error value is 719.5716152204503  and the particle is  Particles =  [  1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 ]
+Iteration 20 the global best fitness is 0.7777777777777778 and error value is 719.5716152204503  and the particle is  Particles =  [  1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 ]
+ After swarm  the global best fitness is 0.875 and error value is 729.248312158671  and the particle is  Particles =  [  1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 ]
+
+ F First Polygonization PSO = [ S0 Line[P(226.0,139.0) to P(241.0,209.0)]  ,      S1 Line[P(241.0,209.0) to P(266.0,226.0)]  ,      S2 Line[P(266.0,226.0) to P(296.0,128.0)]  ,      S3 Line[P(296.0,128.0) to P(312.0,216.0)]  ,    ]
+ \\ \hline
+
+  Segmentation & PSO Segmentation ALgS2 & $P_{dp}$  & ALgS2  (result if Figure \ref{fig:Stroke2Seg2labeled} )&    
+ Iteration 0 the global best fitness is 8.725462920280576E-11 and error value is 9.151490748550003  and the particle is  Particles =  [  1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 ]
+
+Iteration 10 the global best fitness is 3.126281513587432E-7 and error value is 8.61499750333895  and the particle is  Particles =  [  1 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 ]
+
+Iteration 20 the global best fitness is 3.126281513587432E-7 and error value is 8.61499750333895  and the particle is  Particles =  [  1 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 ]
+
+ After swarm  the global best fitness is 3.126281513587432E-7 and error value is 8.61499750333895  and the particle is  Particles =  [  1 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 ]
+
+  second PSO =   Segments = [   S0 Line{P(226.0,139.0) to P(238.0,183.0)},      S1 Line{P(238.0,183.0) to P(241.0,213.0)},      S2 Line{P(241.0,213.0) to P(276.0,227.0)},      S3 Line{P(276.0,227.0) to P(287.0,144.0)},      S4 Curve{P(287.0,144.0) to P(310.0,131.0),  Center ( 299.02634312229503,137.80386092517725) Radius =11.907214051459519 },  S5 Curve{P(310.0,131.0) to P(312.0,216.0),  Center ( 82.44520390380922,172.2089170470769) Radius =231.70851748053468 },  S6 Line{P(312.0,216.0) to P(336.0,224.0)}, ]
+
+ \\ \hline 
+Choosing best fit & Compare between ALgS1 and AlgS2 &  First PSO and Second PSO  &  (result if Figure \ref{fig:FinalSegmentation} ) & 
+   The error ALgS1 780.6652975044847  Algs2 161.60692855994625    Best = ALgS2  \\ \hline
+
+		\end{longtable}
+%}
+
+%\end{table}
+
+\end{scriptsize}
+
+\end{landscape}
+
+\begin{figure}
+	\centering
+	\subfigure [AlgS1 Output]{
+		\includegraphics{afterDefense/Stroke2Seg1labeled.jpg.eps}
+	\label{fig:Stroke2Seg1labeled}}
+
+	\subfigure[AlgS2 Output]{	\includegraphics{afterDefense/Stroke2Seg2labeled.jpg.eps}	\label{fig:Stroke2Seg2labeled}
+	}
+	\subfigure[Final Segmentation of Symbol]{		\label{fig:FinalSegmentation}	\includegraphics[scale=0.65]{afterDefense/FinalSegmentation.jpg.eps}
+	}
+	\caption{Segmentation Ouputs For Second Stroke}
+\end{figure}
+
+The final output of this step is list $L_s$ of segments There are  9  segments in this symbol
+
+\begin{itemize}
+	\item    Segment S0  = Ellipse ( Stroke 1)
+   \item Segment S1  =   Line ( Stroke 2)
+     \item  Segment S2 = Line ( Stroke 2)
+     \item  Segment S3  =  Line ( Stroke 2)
+     \item Segment S4  =   Line ( Stroke 2)
+  \item    Segment S5  =  Curve  ( Stroke 2)
+     \item  Segment S6  =  Curve  ( Stroke 2)
+     \item  Segment S7  =  Line   ( Stroke 2)
+     \item  Segment S8  =  Line   ( Stroke 2)
+\end{itemize}
+
+\section{Step by Step Recognition}
+\label{apsec:steprec}
+
+\begin{figure}
+	\centering
+		\includegraphics[scale=0.85]{afterDefense/FinalSegmentation.jpg.eps}
+	\caption{Best Segmantation }
+	\label{fig:segmentatiofinal.jpg}
+\end{figure}
+
+%\begin{enumerate}
+%\begin{itemize}
+
+	The input of the step is the list of segmentation $L_s$ that is used to compute the feature vector.
+Figure \ref{fig:segmentatiofinal.jpg} shows the best segmentation as given by the previous step.
+This step can be summarized as firstly using $L_s$ to computer the feature vector and then introduce this vector the SVM classifier to classify it to one of the trained categories.
+The details of each step of feature extraction are detailed in the next algorithm.
+Feature extraction step is divided into four steps where each feature set is computed separately then all sets are appended to construct a final feature vector.
+
+\begin{enumerate}
+	\item Computing Structural and geometrical Features (FS1).
+This step computes number of parallel, perpendicular line and similar structural features.
+Algorithm \ref{FeatureAlg} is used to determine these features.
+	\item Rubine Feature Set (FS2).  Compute Rubine features as in \cite{gestureexample12} 
+	\item Statistical Features (FS3) compute the moments as in \cite{zernike61}
+	\item Global shape properties set (FS4). Firstly compute the convex hull of the stroke points then compute the rest of the features.
+
+	\item Introduce the computed feature vector to the classifier and check the label of the detected symbol. 
+\end{enumerate}
+
+	 \begin{algorithm}
+\caption{Computing Structural Features}
+\label{FeatureAlg}
+	\begin{verbatim}
+Initialize 
+  1. For each segment Si in list Ls  do the following. 
+      1.a If segment si is a line do the following 
+            1.a.I. Compute the slope and intercept of line si. 
+            1.a.II Increment Parallel count  if 
+                  parallel with each other line in Ls
+            1.a.III Increment PerpendicularCount if 
+                    perpendicular with each other line in Ls 
+            1.a.IV Increment Intersection Count if 
+                    intersected with any other line. 
+                    1.a.II.A  if Intersection found determine 
+                              type of intersection or (X,T,L)
+      1.b If segment si is curve do the following. 
+          1.b.I Increment Intersect if
+               Intersect other line or curve
+    1.b.II Increment number 
+                 of curves and primitives. 
+4. return Feature_Vector_FS1
+\end{verbatim}	
+ \end{algorithm}
+
+\begin{landscape}
+\begin{scriptsize}
+
+ \begin{longtable}{|p{2cm}|p{2cm}|p{2cm}|p{2cm}|p{13cm}|}
+\caption{Detailed Output of System in Each Step of Recognition}
+\label{tab:StepsRegDetails} \\
+
+\hline 
+\multicolumn{1}{|p{2cm}|}{\textbf{Step}} & 
+\multicolumn{1}{p{2cm}|}{\textbf{Description}} &
+\multicolumn{1}{p{2cm}|}{\textbf{Inputs}} &
+\multicolumn{1}{p{2cm}|}{\textbf{Details}} &
+\multicolumn{1}{p{13cm}|}{\textbf{Output}} 
+\\ \hline 
+\endfirsthead
+
+\multicolumn{3}{c}%
+{{\bfseries \tablename\ \thetable{} -- continued from previous page}} \\
+\hline 
+\multicolumn{1}{|p{2cm}|}{\textbf{Step}} & 
+\multicolumn{1}{p{2cm}|}{\textbf{Description}} &
+ \multicolumn{1}{p{2cm}|}{\textbf{Inputs}} &
+ \multicolumn{1}{p{2cm}|}{\textbf{Details}} &
+ \multicolumn{1}{p{13cm}|}{\textbf{Output}} 
+\\ \hline 
+\endhead
+
+ Feature Extraction of FS1 & Preprocessing of Structural Featues  &   There are  9  segments in this symbol $L_s$ = 
+   Segment S0  = Ellipse
+   Segment S1  = Line
+   Segment S2  = Line
+   Segment S3  = Line
+   Segment S4  = Line
+   Segment S5  = Curve
+   Segment S6  = Curve
+   Segment S7  = Line
+   Segment S8  = Line  &  (see First Algorithm in Section \ref{sec:steprec} )  &    Line from  24 x= 241.0y=213.0  to 49 x2= 276.0y2=227.0  is\textbf{ parallel} to Line from  163 x= 312.0y=216.0  to 185 x2= 336.0y2=224.0
+
+ Line from  0 x= 226.0y=139.0  to 14 x2= 238.0y2=183.0 \textbf{ intersects }Line from  14 x= 238.0y=183.0  to 24 x2= 241.0y2=213.0
+
+ Line from  0 x= 226.0y=139.0  to 14 x2= 238.0y2=183.0  \textbf{intersects} Line from  24 x= 241.0y=213.0  to 49 x2= 276.0y2=227.0
+
+ Line from  0 x= 226.0y=139.0  to 14 x2= 238.0y2=183.0 \textbf{ intersects} Line from  49 x= 276.0y=227.0  to 98 x2= 287.0y2=144.0
+
+ Line from  0 x= 226.0y=139.0  to 14 x2= 238.0y2=183.0\textbf{  intersects }Line from  163 x= 312.0y=216.0  to 185 x2= 336.0y2=224.0
+
+ Line from  0 x= 226.0y=139.0  to 14 x2= 238.0y2=183.0 \textbf{ intersects} Line from  185 x= 336.0y=224.0  to 223 x2= 344.0y2=137.0
+ %Line from  14 x= 238.0y=183.0  to 24 x2= 241.0y2=213.0  intersects Line from  24 x= 241.0y=213.0  to 49 x2= 276.0y2=227.0
+ %Line from  14 x= 238.0y=183.0  to 24 x2= 241.0y2=213.0  intersects Line from  49 x= 276.0y=227.0  to 98 x2= 287.0y2=144.0
+ %Line from  14 x= 238.0y=183.0  to 24 x2= 241.0y2=213.0  intersects Line from  163 x= 312.0y=216.0  to 185 x2= 336.0y2=224.0
+ %Line from  14 x= 238.0y=183.0  to 24 x2= 241.0y2=213.0  intersects Line from  185 x= 336.0y=224.0  to 223 x2= 344.0y2=137.0
+ %Line from  24 x= 241.0y=213.0  to 49 x2= 276.0y2=227.0  intersects Line from  49 x= 276.0y=227.0  to 98 x2= 287.0y2=144.0
+ %Line from  24 x= 241.0y=213.0  to 49 x2= 276.0y2=227.0  intersects Line from  163 x= 312.0y=216.0  to 185 x2= 336.0y2=224.0
+ %Line from  24 x= 241.0y=213.0  to 49 x2= 276.0y2=227.0  intersects Line from  185 x= 336.0y=224.0  to 223 x2= 344.0y2=137.0
+ %Line from  49 x= 276.0y=227.0  to 98 x2= 287.0y2=144.0  intersects Line from  163 x= 312.0y=216.0  to 185 x2= 336.0y2=224.0
+
+ Line from  49 x= 276.0y=227.0  to 98 x2= 287.0y2=144.0  intersects Line from  185 x= 336.0y=224.0  to 223 x2= 344.0y2=137.0
+
+ Line from  163 x= 312.0y=216.0  to 185 x2= 336.0y2=224.0  intersects Line from  185 x= 336.0y=224.0  to 223 x2= 344.0y2=137.0
+
+\\ \hline
+
+ FS1 & Structural and geometrical Features & Segment List $L_s$  &  & 
+ The final Feature vector of the FS1==>
+ature Number of primitives 3.0   ,    Feature Number of segments 9.0   ,    Feature Line count 6.0   ,    Feature Curves Count 3.0   ,    Feature Ellipse Count 1.0   ,    Feature Intersection T type -1.0   ,    Feature Intersection L type 4.0   ,    Feature Intersection X type -1.0   ,    Feature Parallel Count 1.0   ,    Feature PerPendicular Count -1.0   ,    Feature Intersection Lines Count 4.0   ,    Feature Min Radius 1.7976931348623157E308   ,    Feature Max Radius 231.70851748053468
+
+ \\ \hline
+FS2  & Rubine Features & Segment List $L_s$  &   see \cite{gestureexample12}   & 
+\begin{scriptsize}
+ Feature Rubine 0 -1.0   ,    Feature Rubine 1 -1.0   ,    Feature Rubine 2 477.6107201476952   ,    Feature Rubine 3 0.6368010415482491   ,    Feature Rubine 4 169.49926253526885   ,    Feature Rubine 5 0.8908593331996381   ,    Feature Rubine 6 0.454279262624981   ,    Feature Rubine 7 1179.0172885029808   ,    Feature Rubine 8 4.712388980384688   ,    Feature Rubine 9 70.80157037633609   ,    Feature Rubine 10 55.4044569978846   ,    Feature Rubine 11 9.140625   ,    Feature Rubine 12 41783.0   ,
+\end{scriptsize}
+ \\ \hline
+
+FS3 & Momments Computation   &  Segment List $L_s$  &  see \cite{zernike61}  &
+
+\begin{scriptsize}
+   Feature Zernike moments 0 158.87520185460517   ,    Feature Zernike moments 0 7.007557168632727   ,    Feature Zernike moments 1 34.26844061831789   ,    Feature Zernike moments 1 6.658086415801404   ,    Feature Zernike moments 2 85.70032052974454   ,    Feature Zernike moments 2 9.441591321498212   ,    Feature Zernike moments 3 11.73170471301825   ,    Feature Zernike moments 3 66.52997844880663   ,    Feature Zernike moments 4 20.42774127507497   ,    Feature Zernike moments 4 10.196341027811322   ,    Feature Zernike moments 5 52.821799077785066   ,    Feature Zernike moments 5 12.078236804336935   ,    Feature Zernike moments 6 27.748912050714   ,    Feature Zernike moments 6 9.870152472028517   ,    Feature Zernike moments 7 26.079803494397133   ,    Feature Zernike moments 7 28.109003184347657   ,    Feature Zernike moments 8 23.32097186060531   ,    Feature Zernike moments 8 7.632681250023639   ,    Feature Zernike moments 9 6.5841386188472395   ,    Feature Zernike moments 9 51.34195037191517   ,    Feature Zernike moments 10 30.122054353503337   ,    Feature Zernike moments 10 26.902895199361584   ,    Feature Zernike moments 11 1.8295982451659387   ,    Feature Zernike moments 11 29.93196640188875   ,    Feature Zernike moments 12 64.0879253626976   ,    Feature Zernike moments 12 31.811780133580722   ,    Feature Zernike moments 13 23.614051070558084   ,    Feature Zernike moments 13 1.8043229719502216   ,    Feature Zernike moments 14 113.92181136864328   ,    Feature Zernike moments 14 67.26888519955787   ,    Feature Zernike moments 15 47.7397132180995   ,    Feature Zernike moments 15 31.490125533238682   ,    Feature Zernike moments 16 10.320112066817984   ,    Feature Zernike moments 16
+\end{scriptsize}
+
+  \\ \hline
+FS4   &  &Segment List $L_s$   &  &  \begin{scriptsize} 
+ The convex hull    [ X = 383.0   Y = 146.0  Time = 1250452568824,  X = 380.0   Y = 139.0  Time = 1250452568816,  X = 368.0   Y = 119.0  Time = 1250452568800,  X = 363.0   Y = 113.0  Time = 1250452568792,  X = 352.0   Y = 101.0  Time = 1250452568776,  X = 348.0   Y = 97.0  Time = 1250452568768,  X = 342.0   Y = 92.0  Time = 1250452568760,  X = 335.0   Y = 88.0  Time = 1250452568752,  X = 329.0   Y = 85.0  Time = 1250452568744, % X = 322.0   Y = 83.0  Time = 1250452568736,  X = 302.0   Y = 78.0  Time = 1250452568728,  X = 293.0   Y = 77.0  Time = 1250452568720,  X = 282.0   Y = 76.0  Time = 1250452568712,  X = 250.0   Y = 76.0  Time = 1250452568688,  X = 239.0   Y = 77.0  Time = 1250452568680,  X = 229.0   Y = 79.0  Time = 1250452568672,  X = 220.0   Y = 81.0  Time = 1250452568664,  X = 217.0   Y = 83.0  Time = 1250452568656,  X = 201.0   Y = 96.0  Time = 1250452568648,  X = 196.0   Y = 103.0  Time = 1250452568640,  X = 186.0   Y = 118.0  Time = 1250452568624,  X = 182.0   Y = 127.0  Time = 1250452568608,  X = 180.0   Y = 209.0  Time = 1250452569139,  X = 180.0   Y = 247.0  Time = 1250452569092,  X = 181.0   Y = 253.0  Time = 1250452569080,  X = 185.0   Y = 260.0  Time = 1250452569073,  X = 190.0   Y = 265.0  Time = 1250452569056,  X = 194.0   Y = 267.0  Time = 1250452569047,  X = 213.0   Y = 274.0  Time = 1250452569031,  X = 220.0   Y = 276.0  Time = 1250452569023,  X = 256.0   Y = 284.0  Time = 1250452568995,  X = 266.0   Y = 284.0  Time = 1250452568976,
+ ...................................................X = 274.0   Y = 283.0  Time = 1250452568969,  X = 282.0   Y = 281.0  Time = 1250452568960,  X = 290.0   Y = 278.0  Time = 1250452568952,  X = 298.0   Y = 274.0  Time = 1250452568944,  X = 319.0   Y = 262.0  Time = 1250452568936,  X = 339.0   Y = 243.0  Time = 1250452568912,  X = 357.0   Y = 222.0  Time = 1250452568889,  X = 368.0   Y = 202.0  Time = 1250452568872,  X = 378.0   Y = 183.0  Time = 1250452568864,  X = 384.0   Y = 167.0  Time = 1250452568848,  X = 384.0   Y = 160.0  Time = 1250452568840]
+ Area of convex hull =  -34160.5   preimiter of hull 656.1576413855265
+
+ --------------------------------------------------------------------------------
+
+Final Feature vector for the FS4: 
+  Feature Centroid time 1.2504525968640884E12  , Feature Centroid time difference 28319.088339222613   ,    Feature area of convexhull 34160.5   ,    Feature area of convexhull/area of symbol 1.2365344240932454   ,    Feature N.Points ConvexHull/ N.
+points symbol 0.15302491103202848   ,    Feature Convext Perimeter/symbol perimeter 0.556529278903672   ,    Feature  percentage of convex area to bounding box area 19.493542609351433   ,    Feature area  35519.5   ,    Feature diff area of symbol to area of bounding box 16.290771116138764   ,    Feature  Ration between w/h  0.9807692307692307   ,    Feature  Log of ration between hight and width -0.008433167536862813   ,
+ Feature  Log of computed area 4.550466843684004   ,    Feature  abs of area 27626.0   ,    Feature log of  abs of area 4.441318007475848   ,    Feature  Density(length/size)  4.046853548165911   ,    Feature Mean CentroidalRadius x 272.48042704626334   ,    Feature Mean CentroidalRadius y 177.19928825622776   ,    Feature Mean t 1.2504525970626514E12   ,    Feature Mean time difference 148.69395017793593   ,    Feature Mean Radius 62.000799881019084
+  \end{scriptsize}
+    \\ \hline
+ Feature Vector & Appends the selected Feature FS1, FS2, FS3, FS4 & Previous Steps  & &
+
+   \begin{scriptsize}
+   Feature Number of primitives 3.0   ,    Feature Number of segments 9.0   ,    Feature Line count 6.0   ,    Feature Curves Count 3.0   ,    Feature Ellipse Count 1.0   ,    Feature Intersection T type -1.0   ,    Feature Intersection L type 4.0   ,    Feature Intersection X type -1.0   ,    Feature Parallel Count 1.0   ,    Feature PerPendicular Count -1.0   ,    Feature Intersection Lines Count 4.0   ,    Feature Min Radius 1.7976931348623157E308   ,    Feature Max Radius 231.70851748053468   ,    Feature Centroid x 92.54770318021201   ,    Feature Centroid y 101.3851590106007   ,    Feature Centroid time 1.2504525968640884E12   ,    Feature Centroid time difference 28319.088339222613   ,    Feature area of convexhull 34160.5   ,    Feature area of convexhull/area of symbol 1.2365344240932454   ,    Feature N.Points ConvexHull/ N.
+points symbol 0.15302491103202848   ,    Feature Convext Perimeter/symbol perimeter 0.556529278903672   ,    Feature  percentage of convex area to bounding box area 19.493542609351433   ,    Feature area  35519.5   ,    Feature diff area of symbol to area of bounding box 16.290771116138764   ,    Feature  Ration between w/h  0.9807692307692307   ,    Feature  Log of ration between hight and width -0.008433167536862813   ,    Feature Zernike moments 0 158.87520185460517   ,    Feature Zernike moments 0 7.007557168632727   ,    Feature Zernike moments 1 34.26844061831789   ,    Feature Zernike moments 1 6.658086415801404   ,    Feature Zernike moments 2 85.70032052974454   ,    Feature Zernike moments 2 9.441591321498212   ,    Feature Zernike moments 3 11.73170471301825   ,    Feature Zernike moments 3 66.52997844880663   ,    Feature Zernike moments 4 20.42774127507497   ,    %Feature Zernike moments 4 10.196341027811322   ,    Feature Zernike moments 5 52.821799077785066   ,    Feature Zernike moments 5 12.078236804336935   ,    Feature Zernike moments 6 27.748912050714   ,    Feature Zernike moments 6 % 9.870152472028517   ,    Feature Zernike moments 7 26.079803494397133   ,    Feature Zernike moments 7 28.109003184347657   ,    Feature Zernike moments 8 23.32097186060531   ,    Feature Zernike moments 8 7.632681250023639   ,    Feature Zernike moments 9 6.5841386188472395   ,    Feature Zernike moments 9 51.34195037191517   ,    Feature Zernike moments 10 30.122054353503337   ,    Feature Zernike moments 10 26.902895199361584   ,    Feature Zernike moments 11 1.8295982451659387   ,    Feature Zernike moments 11 29.93196640188875   ,    Feature Zernike moments 12 64.0879253626976   ,    Feature Zernike moments 12 31.811780133580722   ,    Feature Zernike moments 13 23.614051070558084   ,    Feature Zernike moments 1    1.8043229719502216   ,
+%Feature Zernike moments 14 113.92181136864328   ,    Feature Zernike moments 14 67.26888519955787   ,    Feature Zernike moments 15 47.7397132180995   ,    Feature Zernike moments 15 31.490125533238682   ,
+Feature Zernike moments 16 10.320112066817984   ,    Feature Zernike moments 16 4.007106145033957   ,    Feature Rubine 0 -1.0   ,    Feature Rubine 1 -1.0   ,    Feature Rubine 2 477.6107201476952   ,    Feature Rubine 3 0.6368010415482491   ,    Feature Rubine 4 169.49926253526885   ,    Feature Rubine 5 0.8908593331996381   ,    Feature Rubine 6 0.454279262624981   ,    Feature Rubine 7 1179.0172885029808   ,    Feature Rubine 8 4.712388980384688   ,    Feature Rubine 9 70.80157037633609   ,    Feature Rubine 10 55.4044569978846   ,    Feature Rubine 11 9.140625   ,    Feature Rubine 12 41783.0   ,    Feature  Log of computed area 4.550466843684004   ,    Feature  abs of area 27626.0   ,    Feature log of  abs of area 4.441318007475848   ,    Feature  Density(length/size)  4.046853548165911   ,    Feature Mean CentroidalRadius x 272.48042704626334   ,    Feature Mean CentroidalRadius y 177.19928825622776   ,    Feature Mean t 1.2504525970626514E12   ,    Feature Mean time difference 148.69395017793593   ,    Feature Mean Radius 62.000799881019084   ,
+   \end{scriptsize} \\ \hline
+   Classification  & SVM classifier decision & Feature Vector  &  & 
+   Symbol = Clock.
+
+ \\ \hline
+
+ 		\end{longtable}
+%}
+
+%\end{table}
+
+\end{scriptsize}
+
+\end{landscape}

### 2009-09-06 draft/System.tex
after the changes in examples

--- 2009-08-17 draft/System.tex
@@
 Section \ref{sec:steprec} shows the details of feature calculation and the classification process.
+\begin{figure}
+	\centering
+	\subfigure[First Stroke]{\label{fig:clockstroke1}
+		\includegraphics[scale=0.7]{afterDefense/clockstroke1.jpg.eps} }
+		\subfigure[Second Stroke] {	\label{fig:clockstroke2}
+		\includegraphics[scale=0.7]{afterDefense/clockstroke2.jpg.eps}}		
+		\subfigure[Completed symbol]{
+		\label{fig:Clock1}\includegraphics[scale=0.7]{afterDefense/ClockDirections.jpg.eps}	}
+	\caption{Stroke Drawing Sequence} a) The user first draws the ellipse b) The square wave is drawn inside the ellipse c) The final symbol the user draws along with arrows to illustrate how the directions the user used while drawing the symbol.
+	\label{fig:clock}
+\end{figure}
 
@@
 
-\begin{figure}
-	\centering
-	\subfigure[First Stroke]{\label{fig:clockstroke1}
-		\includegraphics[scale=0.7]{afterDefense/clockstroke1.jpg.eps} }
-
-		\subfigure[Second Stroke]{	\label{fig:clockstroke2}
-		\includegraphics[scale=0.7]{afterDefense/ClockComplete.jpg.eps}}
-
-		\subfigure[Completed symbol]{
-		\label{fig:Clock1}\includegraphics[scale=0.7]{afterDefense/Clock1.jpg.eps}	}
-	\caption{Stroke Drawing Sequence} a) The user first draws the ellipse b) The square wave is drawn inside the ellipse c) The arrows in the symbols illustrate how the user has drawn the symbol
-	\label{fig:clock}
-\end{figure}
-
 In this step the user draw one or more stroke to draw a symbol.
 Figure \ref{fig:clock} illustrate the sequence and direction of user draw.
-In this section, each stroke user draw will be labeled as it illustrated in Figure \ref{fig:clockstroke1} and Figure \ref{fig:clockstroke2} .
+In this section, each stroke user draws will be labeled as it illustrated in Figure \ref{fig:clockstroke1} and Figure \ref{fig:clockstroke2}.
 Table \ref{tab:StepsStroke1} shows the detailed output of the program at each step.
-The input of this step is the stroke points and the ouptut is a list of segments that wil be used in the recognition process.
-The algorithm that compute the segmentation is demonstrated as follows:
+The input of this step is the a set of strokes, each stroke has a set of  points and the output is a list of segments that will be used in the recognition process.
+More detailed example and algorithms are demonstrated in the Appendex \ref{ChapterstepExample} \\
+
+For each stroke the user draw the system segments the stroke using the following procedure
+
 \begin{enumerate}
-
-	\item Extracting $P_dp$ as in Section \ref{sec:Preprocessing}
-\begin{verbatim}
-  1. For each point in Stroke do the follwoing:
-  		1.a Compute distance from previous point.
-  		1.b Compute Time difference from previous point and add it to Time_list.  
-  		1.c Compute Velocity and add it to Vel_list. 
-  		1.d Compute Direction and add it to Dir_list.
-  	 	1.e Compute Curvature and add it to Cur_list.(Eq.4.1) 
-  	end (for loop)
-  2. For each  list computed in previous step do the follwing:   
-  	 2.a Compute average of all points. 
-  	 2.b set average as threshold.  
-  	 2.c Create region from continous points larger than threshold. 
-  	 2.d for each region do the following. 
-  	 	   2.d.1 find the maximum 
-  	 	   2.d.1 Add point to P_dp 
-  	  end
-     end 
-\end{verbatim}	
-\item Ellipse Fitting: which is an attempt to fit as an ellipse as in  Section \ref{sec:EllipseDetection} 
+\item Extracting $P_dp$ as in Section \ref{sec:Preprocessing} and Algorithm \ref{extractpdp}.
+\item Ellipse Fitting: which is an attempt to fit as an ellipse as in details in Section \ref{sec:EllipseDetection} and Algorithm \ref{ellipseFitAlg} in Appendex \ref{ChapterstepExample}.
+\item Segmentation ALS1 as explained in Section \ref{sec:SwarmSegmentation}.
+The alogrithm will be similar to the PSO algorithm  in section \ref{sec:ParticleSwarmAlgorithm}.
+(see Algorithm \ref{AlgS1alg} in Appendex \ref{ChapterstepExample})
 %he axes of the ellipse are estimated as the $width/2$ and $height/2$ of the stroke bounding box where $width$ is width and $height$ is height of the bonding box.
-\begin{verbatim}
-1. Initialize the axes of ellipse  and center by bounding box information
-2. Repeat the follwoing for few loops: 
-	2.a Compute error in Equation  4.2 
-	2.b Update center of ellipse and axis from error. 
-3. Compute percent in Equation 4.4 
-4. Computer eff in Equation 4.3 
-5. if eff < 0.2  then fit as ellipse otherwise move to segment. 
-\end{verbatim}
-
-\item Segmentation ALS1 as explained in Section \ref{sec:SwarmSegmentation}. The alogrithm will be similar to the PSO algoirhtm  in section \ref{sec:ParticleSwarmAlgorithm} 
-\begin{verbatim}
-Initialize 
-  1. Generate N particles with M dimensions (P_i...P_N) 
-  2. Generate velocities for each particle for M dimension randomly. 
-  3. Repeat till Maximum number of iteration OR ErrorThreshold  
-         For each particle P_i do the following 
-         3.a   Compute fitness  (fitness) 
-         				3.a.1 Compute the Error of approximation as in Eq.4.5
-         				3.a.2. Computer the fitness function as in Eq. 4.4 
-         3.b   Determine Lbest  
-               if (lbest_fitness<fitness)
-                  3.b.1   lbest=P_i
-         3.c      Determine Gbest.
-               if (lGest_fitness<fitness)
-                  3.c.1   lGest=P_i
-         3.d   Compute the Velocity for each dimension using Eq.3.2 
-         3.e   Bound Velocity using Eq.3.4
-         3.f   Update particle values using Eq. 3.3
-         3.g   Refine the location of the particle 
-         end .
-     end 
-4. return Gbest
-\end{verbatim}	
-\item Segmentation ALS2  as explained in Section \ref{sec:PolygonDivisionAlgorithm}
-\begin{verbatim}
-Initialize 
-  1. Generate N particles with M dimensions (P_i...P_N) 
-  2. Generate velocities for each particle for M dimension randomly. 
-  3. Repeat till Maximum number of iteration OR ErrorThreshold  
-         For each particle P_i do the following 
-         3.a   Compute fitness  (fitness) 
-              3.a.1 For each Segment in the particle do the following
-         				    3.a.1.I   Error of approximation as Line (Eq. 4.7 - 4.9 )
-         			      3.a.1.II  Error of approximation as circular arc (Eq. 4.10 - 4.19 )
-         				    3.a.1.III Computer minimum of both approximation 
-         			3.a.2 Compute Error of the particle as in Eq.  4.20
-         			3.a.3 Compute fitness of the particle as in Eq. 4.21
-         3.b   Determine Lbest  
-               if (lbest_fitness<fitness)
-                  3.b.1   lbest=P_i
-         3.c      Determine Gbest.
-               if (lGest_fitness<fitness)
-                  3.c.1   lGest=P_i
-         3.d   Compute the Velocity for each dimension using Eq.3.2 
-         3.e   Bound Velocity using Eq.3.4
-         3.f   Update particle values using Eq. 3.3
-         3.g   Refine the location of the particle 
-         end .
-     end 
-4. return Gbest
-\end{verbatim}	
+\item Segmentation ALS2 as explained in Section \ref{sec:PolygonDivisionAlgorithm}. ( see Algorithm \ref{AlgS2alg} in Appendex \ref{ChapterstepExample} ). 
 \item Append best segmentation to List of segmentation ($L_s$) of this symbols. 
@@
 \begin{landscape}
+The steps of the first stroke (details of the ouputs are in  Appendex \ref{ChapterstepExample} in Table \ref{tab:StepsStroke1Details}):
 \begin{scriptsize}
@@
 %\scalebox{0.6}{		
-\begin{longtable}{|p{2cm}|p{2cm}|p{2cm}|p{2cm}|p{13cm}|}
-\caption{Detailed Output of System in Each Step of Segmentation}
+\begin{longtable}{|p{2cm}|p{2cm}|p{5cm}|p{2cm}|p{5cm}|}
+\caption{ Output of System in Each Step of Segmentation of First Stroke}
 \label{tab:StepsStroke1} \\
@@
 \multicolumn{1}{p{2cm}|}{\textbf{Description}} &
-\multicolumn{1}{p{2cm}|}{\textbf{Inputs}} &
+\multicolumn{1}{p{5cm}|}{\textbf{Inputs}} &
 \multicolumn{1}{p{2cm}|}{\textbf{Details}} &
-\multicolumn{1}{p{13cm}|}{\textbf{Output}} 
+\multicolumn{1}{p{5cm}|}{\textbf{Output}} 
 \\ \hline 
@@
 \multicolumn{1}{p{2cm}|}{\textbf{Description}} &
- \multicolumn{1}{p{2cm}|}{\textbf{Inputs}} &
+ \multicolumn{1}{p{5cm}|}{\textbf{Inputs}} &
  \multicolumn{1}{p{2cm}|}{\textbf{Details}} &
- \multicolumn{1}{p{13cm}|}{\textbf{Output}} 
+ \multicolumn{1}{p{5cm}|}{\textbf{Output}} 
 \\ \hline 
@@
 %Step & Discription of Step & Input & Steps & Output \\ \hline
-Stroke 1 & Extracting point and computing boundary data & -- &  see Section \ref{sec:Preprocessing} & 
-\begin{scriptsize}
-Stroke has 62 points = [P(198.0,153.0), P(197.0,150.0), P(197.0,145.0), P(198.0,140.0), P(203.0,125.0), P(205.0,120.0) , P(205.0,117.0), P(210.0,111.0) , P(214.0,103.0) , P(219.0,97.0), P(225.0,89.0) , P(229.0,85.0) , P(242.0,77.0) , P(247.0,73.0) , P(275.0,64.0) , P(287.0,62.0) , P(325.0,62.0) , P(333.0,63.0) , P(362.0,68.0) , P(372.0,72.0) , P(391.0,77.0) , P(406.0,83.0) , P(424.0,91.0) , P(440.0,100.0), P(474.0,133.0) , P(485.0,149.0) , P(492.0,164.0) , P(499.0,182.0) , P(501.0,191.0) , P(501.0,207.0) , P(499.0,222.0) , P(496.0,239.0) , P(471.0,292.0) , P(467.0,302.0) , P(456.0,315.0) , P(441.0,329.0) , P(427.0,339.0) , P(418.0,344.0) , P(398.0,354.0) , P(336.0,371.0) , P(283.0,376.0) , P(263.0,378.0) , P(242.0,378.0) , P(232.0,376.0) , P(213.0,373.0) , P(194.0,367.0) , P(177.0,362.0) , P(138.0,343.0) , P(129.0,336.0) , P(124.0,331.0) , P(120.0,321.0) , P(115.0,310.0) , P(113.0,297.0) , P(112.0,284.0) , P(114.0,256.0) , P(121.0,219.0) , P(128.0,202.0) , P(137.0,184.0) , P(143.0,172.0) , P(154.0,161.0) , P(172.0,146.0) , P(172.0,146.0)]
-
- Bonding box is  = Corner P(112.0, 62.0), w= 389.0, h=  316.0
- number of point in this stroke is   62
-\end{scriptsize}
- \\ \hline
-Possible Dominate point Extraction & Computing and for details see Section\ref{sec:Preprocessing} &  Stroke points  &   &    Number of $P_{dp}$  = 11
- Pdp = [ 2 (197.0 , 145.0 ), 4 (203.0 , 125.0 ), 3 (198.0 , 140.0 ), 22 (424.0 , 91.0 ), 19 (372.0 , 72.0 ), 26 (492.0 , 164.0 ), 28 (501.0 , 191.0 ), 9 (219.0 , 97.0 ), 31 (496.0 , 239.0 ), 0 (198.0 , 153.0 ), 41 (263.0 , 378.0 )]
-
+Stroke 1 & Extracting point and computing boundary data & -- &   &
+
+	\includegraphics[scale=0.5]{afterDefense/clockstroke1.eps}
+
+  \\ \hline
+Possible Dominate point Extraction & for Computing details see Section\ref{sec:Preprocessing} &
+
+	\includegraphics[scale=0.5]{afterDefense/clockstroke1.eps}
+  &   &    
 \\ \hline 
 Segmentation & Segmentation of Ellipse & $P_{dp}$   &  Ellipse fitting  & 
- Segmented as Ellipse with dimensions  a = 194.5 ,b = 194.5 , Center( 306.5,  220.0 )   with Error =  2.276869953758031 , Percent =0.9093857875960022
- certainty  =  0.1997008626019765 
-\\ \hline
+The circle is   a = 103.05715108084719 ,b = 103.06851874747797 , Center( 283.03365107691775,  181.05035287667172 )   with Error =  1.9471676673301757 , Percent =1.0695193086531243
+eff  = 0.27463462099275116
+
+ \includegraphics[scale=0.5]{afterDefense/ellipse.jpg.eps}
+ \\ \hline
+ 	\end{longtable}
+ 	\newpage
+ 	The steps of the second stroke:(details of the ouputs are in  Appendex \ref{ChapterstepExample} in Table \ref{tab:StepsStroke1Details})
+
+\begin{longtable}{|p{2cm}|p{2cm}|p{5cm}|p{2cm}|p{5cm}|}
+\caption{ Output of System in Each Step of Segmentation of Second Stroke}
+\label{tab:StepsStroke2} \\
+
+\hline 
+\multicolumn{1}{|p{2cm}|}{\textbf{Step}} & 
+\multicolumn{1}{p{2cm}|}{\textbf{Description}} &
+\multicolumn{1}{p{5cm}|}{\textbf{Inputs}} &
+\multicolumn{1}{p{2cm}|}{\textbf{Details}} &
+\multicolumn{1}{p{5cm}|}{\textbf{Output}} 
+\\ \hline 
+\endfirsthead
+
+\multicolumn{3}{c}%
+{{\bfseries \tablename\ \thetable{} -- continued from previous page}} \\ \hline
+\multicolumn{1}{|p{2cm}|}{\textbf{Step}} & 
+\multicolumn{1}{p{2cm}|}{\textbf{Description}} &
+ \multicolumn{1}{p{5cm}|}{\textbf{Inputs}} &
+ \multicolumn{1}{p{2cm}|}{\textbf{Details}} &
+ \multicolumn{1}{p{5cm}|}{\textbf{Output}} 
+\\ \hline 
+\endhead
 
 Stroke 2 & Extracting point and computing boundary data & -- &  see Section \ref{sec:Preprocessing} &
-\begin{scriptsize}
-Stroke has 205 points = [P(201.0,179.0) , P(202.0,180.0) , P(202.0,182.0) , P(202.0,185.0) , P(202.0,187.0) , P(202.0,190.0) , P(202.0,192.0) , P(202.0,196.0) , P(202.0,202.0) , P(201.0,206.0) , P(200.0,219.0) , P(198.0,231.0) , P(198.0,234.0) , P(198.0,238.0) , P(197.0,244.0) , P(197.0,248.0) , P(197.0,251.0) , P(196.0,258.0) , %P(196.0,260.0) , P(196.0,263.0) , P(195.0,265.0) , P(195.0,266.0) , P(195.0,269.0) , P(195.0,271.0) , P(195.0,273.0) , P(195.0,278.0) , P(195.0,279.0) , P(195.0,281.0) , P(195.0,282.0) , P(195.0,284.0) , P(195.0,285.0) , P(195.0,288.0) , P(195.0,290.0) , P(195.0,295.0) , P(195.0,297.0) , P(195.0,298.0) , P(196.0,298.0) , P(196.0,299.0) , P(196.0,300.0) , P(196.0,301.0) , P(197.0,301.0) , P(198.0,301.0) , P(200.0,301.0) , P(203.0,301.0) , %P(206.0,301.0) , P(208.0,301.0) , P(213.0,301.0) , P(217.0,302.0) , P(219.0,302.0) , P(223.0,303.0) , P(234.0,304.0) , P(237.0,304.0) , P(240.0,306.0) , P(246.0,306.0) , P(247.0,307.0) , P(248.0,307.0) , P(249.0,307.0) , P(250.0,307.0) , P(250.0,306.0) , P(250.0,305.0) , P(250.0,304.0) , P(250.0,302.0) , P(251.0,299.0) , P(251.0,296.0) , P(251.0,294.0) , P(251.0,289.0) , P(252.0,285.0) , P(253.0,279.0) , P(253.0,274.0) , P(253.0,271.0) , P(253.0,264.0) , P(253.0,257.0) , P(254.0,252.0) , P(254.0,241.0) , P(254.0,237.0) , P(255.0,232.0) , P(255.0,229.0) , P(255.0,225.0) , P(256.0,212.0) , P(256.0,209.0) , P(256.0,206.0) , P(257.0,205.0) , P(257.0,200.0) , P(257.0,197.0) , P(258.0,195.0) , P(258.0,194.0) , P(258.0,189.0) , P(258.0,187.0) , P(258.0,186.0) , P(259.0,184.0) , P(259.0,181.0) , P(259.0,179.0) , P(259.0,178.0) , P(260.0,176.0) , P(260.0,174.0) , P(260.0,173.0) , P(260.0,172.0) , P(260.0,171.0) , P(260.0,169.0) , P(260.0,168.0) , P(260.0,167.0) , P(261.0,166.0) , P(261.0,165.0) , P(261.0,164.0) , P(261.0,162.0) , P(262.0,162.0) , P(263.0,162.0) , P(264.0,162.0) , P(268.0,162.0) , P(275.0,162.0) , P(279.0,162.0) , P(289.0,162.0) , P(291.0,162.0) , P(301.0,163.0) , P(305.0,164.0) , P(312.0,164.0) , P(315.0,165.0) , P(318.0,165.0) , P(319.0,165.0) , P(319.0,166.0) , P(320.0,168.0) , P(320.0,171.0) , P(320.0,173.0) , P(320.0,176.0) , P(321.0,182.0) , P(321.0,191.0) , P(322.0,198.0) , P(322.0,211.0) , P(323.0,226.0) , P(324.0,229.0) , P(324.0,238.0) , P(324.0,263.0) , P(325.0,272.0) , P(325.0,281.0) , P(325.0,288.0) , P(325.0,291.0) , P(325.0,297.0) , P(325.0,299.0) , P(325.0,301.0) , P(325.0,305.0) , P(324.0,306.0) , P(324.0,307.0) , P(323.0,308.0) , P(324.0,307.0) , P(325.0,307.0) , P(328.0,306.0) , P(332.0,306.0) , P(339.0,306.0) , P(347.0,305.0) , P(354.0,304.0) , P(365.0,304.0) , P(376.0,304.0) , P(379.0,304.0) , P(385.0,304.0) , P(391.0,304.0) , P(392.0,304.0) , P(393.0,304.0) , P(393.0,302.0) , P(393.0,301.0) , P(394.0,297.0) , P(394.0,290.0) , P(393.0,284.0) , P(393.0,274.0) , P(392.0,258.0) , P(391.0,245.0) , P(391.0,233.0) , P(389.0,213.0) , P(388.0,198.0) , P(388.0,189.0) , P(388.0,184.0) , P(388.0,178.0) , P(388.0,177.0) , P(388.0,176.0) , P(388.0,174.0)
-..................................................................................................................
-.................................................................................
-, P(388.0,173.0) , P(388.0,171.0) , P(388.0,170.0) , P(388.0,168.0) , P(388.0,167.0) , P(388.0,166.0) , P(388.0,165.0) , P(389.0,165.0) , P(390.0,165.0) , P(395.0,165.0) , P(401.0,165.0) , P(406.0,165.0) , P(418.0,166.0) , P(424.0,167.0) , P(426.0,167.0) , P(429.0,167.0) , P(430.0,167.0) , P(431.0,167.0) , P(432.0,167.0) , P(433.0,169.0) , P(434.0,170.0) , P(434.0,174.0) , P(436.0,185.0) , P(440.0,200.0) , P(442.0,220.0) , P(442.0,224.0) , P(446.0,267.0) , P(447.0,279.0) , P(448.0,286.0) , P(448.0,287.0) , P(448.0,287.0)]
-
- Bonding box is  = Corner P(195.0, 162.0), w= 253.0, h=  146.0
- \end{scriptsize}
-  \\ \hline
-  & Possible Dominate point Extraction for details see Section\ref{sec:Preprocessing} &  Stroke points  & 1. compute the $P_{pd}$  & \begin{scriptsize}   Number of $P_{pd}$  =  36
-
- Pdp = [ 25 (195.0 , 278.0 ), 30 (195.0 , 285.0 ), 33 (195.0 , 295.0 ), 27 (195.0 , 281.0 ), 34 (195.0 , 297.0 ), 46 (213.0 , 301.0 ), 47 (217.0 , 302.0 ), 50 (234.0 , 304.0 ), 28 (195.0 , 282.0 ), 94 (260.0 , 174.0 ), 93 (260.0 , 176.0 ), 95 (260.0 , 173.0 ), 98 (260.0 , 169.0 ), 48 (219.0 , 302.0 ), 99 (260.0 , 168.0 ), 109 (275.0 , 162.0 ), 113 (301.0 , 163.0 ), 94 (260.0 , 174.0 ), 132 (325.0 , 272.0 ), 136 (325.0 , 297.0 ), 112 (291.0 , 162.0 ), 137 (325.0 , 299.0 ), 110 (279.0 , 162.0 ), 144 (325.0 , 307.0 ), 147 (339.0 , 306.0 ), 150 (365.0 , 304.0 ), 132 (325.0 , 272.0 ), 168 (388.0 , 189.0 ), 171 (388.0 , 177.0 ), 174 (388.0 , 173.0 ), 146 (332.0 , 306.0 ), 175 (388.0 , 171.0 ), 183 (395.0 , 165.0 ), 186 (418.0 , 166.0 ), 169 (388.0 , 184.0 ), 191 (431.0 , 167.0 ) ]
-
-  \end{scriptsize}
-    \\ \hline
- \\ \hline 
+ 	\includegraphics[scale=0.5]{afterDefense/clockstroke2.jpg.eps} \\ \hline
+  & Possible Dominate point Extraction for details see Section\ref{sec:Preprocessing} &   	\includegraphics[scale=0.5]{afterDefense/clockstroke2.jpg.eps}  & 1. compute the $P_{pd}$  & 
+    \\ \hline
+
 Segmentation & Segmentation of Ellipse & $P_{dp}$   &  Ellipse fitting  & 
@@
  \\ \hline
- Segmentation & PSO Segmentation ALgS1 & $P_{dp}$  & ALgS1 &   Particles =  [  1 0 1 0 0 1 1 1 1 1 1 0 0 0 1 0 0 0 1 ] 
- The particle  Particles =  [  1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 1 0 0 0 0 1 0 ]
-
- First Polygonization PSO = [[[  S0 Line[P(201.0,179.0) to P(195.0,281.0)]  ,      S1 Line[P(195.0,281.0) to P(234.0,304.0)]  ,      S2 Line[P(234.0,304.0) to P(260.0,174.0)]  ,      S3 Line[P(260.0,174.0) to P(325.0,272.0)]  ,      S4 Line[P(325.0,272.0) to P(365.0,304.0)]  ,      S5 Line[P(365.0,304.0) to P(388.0,189.0)]  ,      S6 Line[P(388.0,189.0) to P(431.0,167.0)]  ,      S7 Line[P(431.0,167.0) to P(442.0,220.0)]  ]]]
- \\ \hline
-
-  Segmentation & PSO Segmentation ALgS2 & $P_{dp}$  & ALgS2 &   Particles =  [  1 0 1 0 0 1 1 1 1 1 1 0 0 0 1 0 0 0 1 ] 
- The particle  Particles =  [  1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 1 0 0 0 0 1 0 ]
-
- Second segmentation PSO = [[[  S0 Line[P(201.0,179.0) to P(195.0,281.0)]  ,      S1 Line[P(195.0,281.0) to P(234.0,304.0)]  ,      S2 Line[P(234.0,304.0) to P(260.0,174.0)]  ,      S3 Line[P(260.0,174.0) to P(325.0,272.0)]  ,      S4 Line[P(325.0,272.0) to P(365.0,304.0)]  ,      S5 Line[P(365.0,304.0) to P(388.0,189.0)]  ,      S6 Line[P(388.0,189.0) to P(431.0,167.0)]  ,      S7 Line[P(431.0,167.0) to P(442.0,220.0)]  ]]]
+ Segmentation & PSO Segmentation ALgS1 & $P_{dp}$  & ALgS1   & Global best fitness is 0.875 and error value is 729.248312158671 
+ \includegraphics[scale=0.7]{afterDefense/Stroke2Seg1labeled.jpg.eps}
+ \\ \hline 
+  Segmentation & PSO Segmentation ALgS2 & $P_{dp}$  & ALgS2   & Error value is 8.61499750333895
+ \includegraphics[scale=0.7]{afterDefense/Stroke2Segmentaion.jpg.eps}
+ \\ \hline 
+Choosing best fit & Compare between ALgS1 and AlgS2 &  First PSO and Second PSO  &   & The error ALgS1 780.6652975044847  Algs2 161.60692855994625    Best = ALgS2
+
+\includegraphics[scale=0.7]{afterDefense/Stroke2Seg2labeled.jpg.eps}	
+  \\ \hline 
+		\end{longtable}
+\end{scriptsize}
+\end{landscape}
+The final output of this step is list $L_s$ of segments There are  9  segments in this symbol (\ref{fig:FinalSegmentationLabeled})
+
+\begin{itemize}
+	\item    Segment S0  = Ellipse ( Stroke 1)
+   \item Segment S1  =   Line ( Stroke 2)
+     \item  Segment S2 = Line ( Stroke 2)
+     \item  Segment S3  =  Line ( Stroke 2)
+     \item Segment S4  =   Line ( Stroke 2)
+  \item    Segment S5  =  Curve  ( Stroke 2)
+     \item  Segment S6  =  Curve  ( Stroke 2)
+     \item  Segment S7  =  Line   ( Stroke 2)
+     \item  Segment S8  =  Line   ( Stroke 2)
+\end{itemize}
+\begin{figure}
+	\centering
+ 		\includegraphics[scale=0.65]{afterDefense/FinalSegmentation.jpg.eps}
+
+		\label{fig:FinalSegmentationLabeled}
+	\caption{Final Segmentation of Symbol} The segments are labeled on the symbol. 
+\end{figure}
+
+\subsection{Step by Step Recognition}
+\label{sec:steprec}
+
+\begin{enumerate}
+	\item Computing Structural and geometrical Features (FS1).
+This step computes number of parallel, perpendicular line and similar structural features [10 features].
+Algorithm \ref{FeatureAlg} in Appendex \ref{ChapterstepExample}  is used to determine these features.
+ 	\item Rubine Feature Set (FS2) [12 features] .  Compute Rubine features as in \cite{gestureexample12} 
+	\item Statistical Features (FS3) compute the moments as in \cite{zernike61} for N=10 there are 32 features.
+	\item Global shape properties set (FS4). Firstly compute the convex hull of the stroke points then compute the rest of the features [12 features]. 
+	\item Introduce the computed feature vector to the classifier and check the label of the detected symbol. 
+\end{enumerate}
+
+\begin{landscape}
+
+More details are in Table \ref{tab:StepsRegDetails} in Appendex \ref{ChapterstepExample}
+
+\begin{scriptsize}
+
+ \begin{longtable}{|p{2cm}|p{2cm}|p{5cm}|p{2cm}|p{10cm}|}
+\caption{Detailed Output of System in Each Step of Recognition}
+\label{tab:StepsReg} \\
+
+\hline 
+\multicolumn{1}{|p{2cm}|}{\textbf{Step}} & 
+\multicolumn{1}{p{2cm}|}{\textbf{Description}} &
+\multicolumn{1}{p{5cm}|}{\textbf{Inputs}} &
+\multicolumn{1}{p{2cm}|}{\textbf{Details}} &
+\multicolumn{1}{p{10cm}|}{\textbf{Output}} 
+\\ \hline 
+\endfirsthead
+
+\multicolumn{3}{c}%
+{{\bfseries \tablename\ \thetable{} -- continued from previous page}} \\  
+\hline
+\multicolumn{1}{|p{2cm}|}{\textbf{Step}} & 
+\multicolumn{1}{p{2cm}|}{\textbf{Description}} &
+ \multicolumn{1}{p{5cm}|}{\textbf{Inputs}} &
+ \multicolumn{1}{p{2cm}|}{\textbf{Details}} &
+ \multicolumn{1}{p{10cm}|}{\textbf{Output}} 
+\\ \hline 
+\endhead
+
+ Feature Extraction of FS1 & Preprocessing of Structural Featues  &  \includegraphics[scale=0.5]{afterDefense/FinalSegmentation.jpg.eps}
+ &  (see First Algorithm in Section \ref{sec:steprec} )  &    Line 1    is\textbf{ parallel} to Line from  L3
+
+ Line 1  \textbf{ intersects }Line 2
+  Line 2  \textbf{ intersects }Line 3
+   Line 3  \textbf{ intersects }Line 4
+    Line 4  \textbf{ intersects }Line 5
+
+\\ \hline
+
+ FS1 & Structural and geometrical Features & Segment List $L_s$  &  & 
+ The final Feature vector of the FS1==>
+ature Number of primitives 3.0   ,    Feature Number of segments 9.0   ,    Feature Line count 6.0   ,    Feature Curves Count 3.0   ,    Feature Ellipse Count 1.0   ,    Feature Intersection T type -1.0   ,    Feature Intersection L type 4.0   ,    Feature Intersection X type -1.0   ,    Feature Parallel Count 1.0   ,    Feature PerPendicular Count -1.0   ,    Feature Intersection Lines Count 4.0   ,    Feature Min Radius 1.7976931348623157E308   ,    Feature Max Radius 231.70851748053468
+
+ \\ \hline
+FS2  & Rubine Features & Segment List $L_s$  &   see \cite{gestureexample12}   & 
+\begin{scriptsize}
+ Feature Rubine 0 -1.0   ,    Feature Rubine 1 -1.0   ,    Feature Rubine 2 477.6107201476952   ,    Feature Rubine 3 0.6368010415482491   ,    Feature Rubine 4 169.49926253526885   ,    Feature Rubine 5 0.8908593331996381   ,    Feature Rubine 6 0.454279262624981   ,    Feature Rubine 7 1179.0172885029808   ,    Feature Rubine 8 4.712388980384688   ,    Feature Rubine 9 70.80157037633609   ,    Feature Rubine 10 55.4044569978846   ,    Feature Rubine 11 9.140625   ,    Feature Rubine 12 41783.0   ,
+\end{scriptsize}
+ \\ \hline
+
+FS3 & Momments Computation   &  Segment List $L_s$  &  see \cite{zernike61}  &
+
+\begin{scriptsize}
+   Feature Zernike moments 0 158.87520185460517   ,    Feature Zernike moments 0 7.007557168632727   ,    Feature Zernike moments 1 34.26844061831789   ,    Feature Zernike moments 1 6.658086415801404   ,.......
+......
+....
+.......
+.......
+.......
+.....
+.....
+.......
+.......
+.......
+.....
+.....
+\end{scriptsize}
+
+  \\ \hline
+FS4   &  &Segment List $L_s$   &  &  \begin{scriptsize} 
+ The convex hull    [ ]
+ Area of convex hull =  -34160.5   preimiter of hull 656.1576413855265
+
+ --------------------------------------------------------------------------------
+
+Final Feature vector for the FS4: 
+  Feature Centroid time 1.2504525968640884E12  , Feature Centroid time difference 28319.088339222613   ,    Feature area of convexhull 34160.5   ,    Feature area of convexhull/area of symbol 1.2365344240932454   ,    Feature N.Points ConvexHull/ N.
+points symbol 0.15302491103202848   ,    Feature Convext Perimeter/symbol perimeter 0.556529278903672   ,    .......
+.......
+.......
+.....
+.....
+.......
+.......
+.......
+.....
+.....
+  \end{scriptsize}
+    \\ \hline
+ Feature Vector & Appends the selected Feature FS1, FS2, FS3, FS4 & Previous Steps  & &
+
+  [FS1,FS2,FS3,FS4]  \\ \hline
+   Classification  & SVM classifier decision & Feature Vector  &  & 
+   Symbol = Clock.
+
  \\ \hline
@@
 \end{scriptsize}
-
 \end{landscape}
-\subsection{Step by Step Recognition}
-\label{sec:steprec}
-
-\begin{figure}
-	\centering
-		\includegraphics{afterDefense/segmentatiofinal.jpg.eps}
-	\caption{Best Segmantation }
-	\label{fig:segmentatiofinal.jpg}
-\end{figure}
-
-%\begin{enumerate}
-%\begin{itemize}
-
-	The input of the step is the list of segmentation $L_s$ that is used to computer the feature vector.
-Figure \ref{fig:segmentatiofinal.jpg} shows the best segmentation as given by the previous step.
-This step can be summerized as firstly using $L_s$ to computer the feature vector and then introduce this vector the SVM classifier to classify it to one of the trained categories.

### 2009-09-26 trunk/draft/Results.tex
after approval of dr. samia

@@
 \begin{center}
+\caption[Datasets Comparisons]{Datasets Comparisons} 
 \scalebox{0.9}{
@@
 }
-\caption[Datasets Comparisons]{Datasets Comparisons} 
+
 \label{tab:datasets}
@@
  \begin{table*}
+ \caption{Confusion Matrix of HS-DB}
 	\centering
@@
  	%}
-	\caption{Confusion Matrix of HS-DB}
+
 	\label{tab:ConfusionMatrix}
@@
 \label{sec:featuresComparisions}
-Different feature sets are tested to determine the best features that can used in sketch and symbol recognition.
+Different feature sets are tested to determine the best features that can use in sketch and symbol recognition.
+The system computed in total 70 features in four different features sets.
+The 13 features of \textbf{FS1},  12 Rubine features in \textbf{FS2} and 32 moments features in \textbf{FS3} and 13 features of  \textbf{FS4} where described in details  section \ref{sec:FeatureExtraction}.
 Figure\ref{fig:testFeaturesAll} shows the result of different feature sets from the basic four sets (\textbf{FS1,FS2,FS3,FS4}).
@@
 % achieved  HS-DB a
+
  \begin{figure}
@@
 %\end{figure}  
+%Another expirements where conducted to evaluates the worth of an attribute.
+The worth of each attribute is measured using two different attribure measures.
+The first is by measuring the gain ratio with respect to the class, where the gain ration is computed as $Gain_{ration}(class,attribute)=\frac{H(class)-H(class|Attribure}{H(Attribute)}$.
+The second method is using
 

### 2009-09-26 trunk/draft/System.tex
after approval of dr. samia

@@
 The system attempts to find the optimal decomposition of the input stroke using PSO with the help of curvature and speed information.
-The use of PSO algorithm helps in eliminating the effects of input noise, problems of over and under segmentation that were in pervious systems.
+The use of PSO algorithm helps in eliminating the effects of input noise, problems of over and under segmentation that were in previous systems.
 The proposed symbol recognizer uses the generated segmentation to compute the feature vector which is used to identify more complex symbols.
@@
 Chen et al.\cite{CruveDivisionSwarm} uses digital curves scanned from paper as input to the system and did not take advantage of the curvature or local geometric properties of the digital curve.
+However, comparisons between PSO and GA \cite{ComparePSOGA05} shows that PSO converges to optimal solution faster than GA.
 
@@
 
-Also the time difference between samples was used in \cite{polygonfeedback31} as it provides more distinct maxima than the speed information.
+Also the time difference between samples was used in \cite{polygonfeedback31} as it provides more distinct maximum than the speed information.
 Agar et al.
@@
 Figure \ref{fig:speed2Distance} shows the time difference and the speed graphs for the stroke drawn in Figure \ref{fig:orignalStroke}.
-Similarly, direction information is used as it provides better distinctive maxima for the corners than the curvature information \cite{meanshift10}.
+Similarly, direction information is used as it provides better distinctive maximum for the corners than the curvature information \cite{meanshift10}.
 Figure \ref{fig:curvatures} shows the direction and curvature graphs for the stroke drawn in Figure \ref{fig:orignalStroke}
@@
 
-The system uses a composite set of features which include global shape properties like Rubine feature set \cite{gestureexample12},  ink density \cite{GeometryAndDomain102}, some appearance based properties like Zernike moments \cite{HeloiseBeautification}, and some stroke based structural information like number of perpendiculars lines , number of parallel lines and types of primitives in each symbol.
+The system uses a composite set of features (total of 70 features) which include global shape properties like Rubine feature set \cite{gestureexample12},  ink density \cite{GeometryAndDomain102}, some appearance based properties like Zernike moments \cite{HeloiseBeautification}, and some stroke based structural information like number of perpendiculars lines , number of parallel lines and types of primitives in each symbol.
 The following list gives details on all the features used in the system.
@@
 \subsection{Structural and geometrical Features(FS1)}
-Features defines the structure of the geometrical symbol. Section \ref{sec:steprec} descripe in details how these feature are computed. 
+ Features defines the structure of the geometrical symbol. Section \ref{sec:steprec} descripe in details how these 13 feature are computed.  
 %\begin{description}
@@
 		\item \emph{X intersections} Number of X intersections.
-
+		\item \emph{Min Radius} Minimum Radius of all curves in the symbol.
+		\item \emph{Max Radius} Maximum Radius of all curves in the symbol.
 		%\item[No. of holes] Number of holes in the symbols. 
@@
 \subsection{Rubine Feature Set (FS2)} Features introduced by Rubine\cite{gestureexample12} for single stroke gestures.
-These features represent the global shape of the symbol drawn.
+These 12 features represent the global shape of the symbol drawn.
 To compute the features we append all segments points into single path $ink_{path}$.
@@
 \begin{enumerate}
-	\item \textsl{Zernike moments:} Zernike moments of order n\footnote{ We tried different values of n. Results show the analysis of each order.} \cite{HeloiseBeautification}. 
+	\item \textsl{Zernike moments:} Zernike moments of order $n$\footnote{ We tried different values of n.
+Results show the analysis of each order.} \cite{HeloiseBeautification}.
+For $n=10$ there are 32 moment features.
 \end{enumerate}
 \subsection{Global shape properties set (FS4)}
- Features that are composites of different geometrical and statistical symbol. These features represent the global shape of the symbol drawn.
+ Features that are composites of different geometrical and statistical symbol. These 13 features represent the global shape of the symbol drawn.
 
@@
 \item \emph{Size Ratio} Ratio between width to height of the symbol.
+		\item \emp{Log of size Ratio} Log of the ratio between width and height of symbol.
 	\item \emph{Ink density} Compute the density of points inside its bounding box\cite{GeometryAndDomain102}.   
  	\item \emph{Convex Hull Area} Area of convex hull with respect to area of bounding box of symbol.
+
 	\item \emph{Convex Hull Perimeter} Perimeter of convex hull with respect to total length of symbol.
+		\item \emph{Convex Hull Point} Number of points of convex hull with respect to Number of points of symbol.
 		\item \emph{Mean Centroidal radius} The Mean of the centroidal radius which is the distance from each point in the symbol to the center of gravity.
+				\item \emph{Center of Gravity} The center of gravity of the symbol (both in y and x direction).
+					\item \emph{Mean Time} The mean of the time values between each two successive points in the symbol. %Different strokes are appended to construct a single path.
+	%\item [Ra]
 
@@
 	%\item [Ra]		
+			\item \emp{Log of Area} Log of area of the symbol.
+	\item \emph{Diff Area} Absolute difference of the area of bounding box and area of symbol.
+
   \end{itemize}
@@
 \begin{enumerate}
-	\item Selecting the SVM classifier Paramters. Parameter $C$ representing the tradeoff between minimizing the training set error and maximizing the margin. 
+	\item Selecting the SVM classifier parameters. Parameter $C$ representing the tradeoff between minimizing the training set error and maximizing the margin. 
 	\item Select A kernel function and their kernel parameters. For example for the radial basis function kernel, one must select width of the Gaussian $\gamma$. 
 	\item Solve dual QP Equation \ref{eq:dualkernel} or an alternative SVM formulation using an appropriate quadratic programming or linear programming algorithm. 
-	\item Get the support vector and recover the emprial threshold  variable $b$. 
+	\item Get the support vector and recover the imperial threshold  variable $b$. 
 	\item Classify new points using $f(x)=sign\left(\sum{i}y_i\alpha_i K\left(x,x_i\right)-b\right)$
@@
 
- There is mainly two popular methods to change the binary SVM classification problem into Multiclass classification problem.
+ There is mainly two popular methods to change the binary SVM classification problem into Multi class classification problem.
 The first methods is One Versus All (OVA) Strategy where M binary classifiers is constructed one classifier for each class.
@@
 over another \cite{Duan05whichis}.
-Some claims that OVO strategy trains in less time than OVA strategy because they train with only portion of the dataset \cite{libsvm} but this claim have not yet been proofed by expeirments.
+Some claims that OVO strategy trains in less time than OVA strategy because they train with only portion of the dataset \cite{libsvm} but this claim have not yet been proofed by experiments.
 
@@
 
-\begin{figure}
-	\centering
-	\subfigure[First Stroke]{\label{fig:clockstroke1}
-		\includegraphics[scale=0.7]{afterDefense/clockstroke1.jpg.eps} }
-
-		\subfigure[Second Stroke]{	\label{fig:clockstroke2}
-		\includegraphics[scale=0.7]{afterDefense/ClockComplete.jpg.eps}}
-
-		\subfigure[Completed symbol]{
-		\label{fig:Clock1}\includegraphics[scale=0.7]{afterDefense/Clock1.jpg.eps}	}
-	\caption{Stroke Drawing Sequence} a) The user first draws the ellipse b) The square wave is drawn inside the ellipse c) The arrows in the symbols illustrate how the user has drawn the symbol
-	\label{fig:clock}
-\end{figure}
-
 In this step the user draw one or more stroke to draw a symbol.
 Figure \ref{fig:clock} illustrate the sequence and direction of user draw.
-In this section, each stroke user draw will be labeled as it illustrated in Figure \ref{fig:clockstroke1} and Figure \ref{fig:clockstroke2} .
+In this section, each stroke user draws will be labeled as it illustrated in Figure \ref{fig:clockstroke1} and Figure \ref{fig:clockstroke2}.
 Table \ref{tab:StepsStroke1} shows the detailed output of the program at each step.
-The input of this step is the stroke points and the output is a list of segments that will be used in the recognition process.
-The algorithm that computes the segmentation is demonstrated as follows:
+The input of this step is the a set of strokes, each stroke has a set of  points and the output is a list of segments that will be used in the recognition process.
+More detailed example and algorithms are demonstrated in the Appendix \ref{ChapterstepExample} \\
+
+For each stroke the user draw the system segments the stroke using the following procedure
+
 \begin{enumerate}
-
-	\item Extracting $P_dp$ as in Section \ref{sec:Preprocessing}
-\begin{verbatim}
-  1. For each point in Stroke do the following:
-  		1.a Compute distance from previous point.
-  		1.b Compute Time difference from previous point and add it to Time_list.  
-  		1.c Compute Velocity and add it to Vel_list. 
-  		1.d Compute Direction and add it to Dir_list.
-  	 	1.e Compute Curvature and add it to Cur_list.(Eq.4.1) 
-  	end (for loop)
-  2. For each  list computed in previous step do the follwing:   
-  	 2.a Compute average of all points. 
-  	 2.b set average as threshold.  
-  	 2.c Create region from continuous points larger than threshold. 
-  	 2.d for each region do the following. 
-  	 	   2.d.1 find the maximum 
-  	 	   2.d.1 Add point to P_dp 
-  	  end
-     end 
-\end{verbatim}	
-\item Ellipse Fitting: which is an attempt to fit as an ellipse as in Section \ref{sec:EllipseDetection} 
+\item Extracting $P_dp$ as in Section \ref{sec:Preprocessing} and Algorithm \ref{extractpdp}.
+\item Ellipse Fitting: which is an attempt to fit as an ellipse as in details in Section \ref{sec:EllipseDetection} and Algorithm \ref{ellipseFitAlg} in Appendix \ref{ChapterstepExample}.
+\item Segmentation ALS1 as explained in Section \ref{sec:SwarmSegmentation}.
+The algorithm will be similar to the PSO algorithm  in section \ref{sec:ParticleSwarmAlgorithm}.
+(see Algorithm \ref{AlgS1alg} in Appendix \ref{ChapterstepExample})
 %he axes of the ellipse are estimated as the $width/2$ and $height/2$ of the stroke bounding box where $width$ is width and $height$ is height of the bonding box.
-\begin{verbatim}
-1. Initialize the axes of ellipse  and center by bounding box information
-2. Repeat the following for few loops: 
-	2.a Compute error in Equation  4.2 
-	2.b Update center of ellipse and axis from error. 
-3. Compute percent in Equation 4.4 
-4. Computer eff in Equation 4.3 
-5. if eff < 0.2  then fit as ellipse otherwise move to segment. 
-\end{verbatim}
-
-\item Segmentation ALS1 as explained in Section \ref{sec:SwarmSegmentation}. The alogrithm will be similar to the PSO algorithm  in section \ref{sec:ParticleSwarmAlgorithm} 
-\begin{verbatim}
-Initialize 
-  1. Generate N particles with M dimensions (P_i...P_N) 
-  2. Generate velocities for each particle for M dimension randomly. 
-  3. Repeat till Maximum number of iteration OR Error Threshold  
-         For each particle P_i do the following 
-         3.a   Compute fitness  (fitness) 
-         				3.a.1 Compute the Error of approximation as in Eq.4.5
-         				3.a.2. Computer the fitness function as in Eq. 4.4 
-         3.b   Determine Lbest  
-               if (lbest_fitness<fitness)
-                  3.b.1   lbest=P_i
-         3.c      Determine Gbest.
-               if (lGest_fitness<fitness)
-                  3.c.1   lGest=P_i
-         3.d   Compute the Velocity for each dimension using Eq.3.2 
-         3.e   Bound Velocity using Eq.3.4
-         3.f   Update particle values using Eq. 3.3
-         3.g   Refine the location of the particle 
-         end .
-     end 
-4. return Gbest
-\end{verbatim}	
-\item Segmentation ALS2  as explained in Section \ref{sec:PolygonDivisionAlgorithm}
-\begin{verbatim}
-Initialize 
-  1. Generate N particles with M dimensions (P_i...P_N) 
-  2. Generate velocities for each particle for M dimension randomly. 
-  3. Repeat till Maximum number of iteration OR Error Threshold  
-         For each particle P_i do the following 
-         3.a   Compute fitness  (fitness) 
-              3.a.1 For each Segment in the particle do the following
-         				    3.a.1.I   Error of approximation as Line (Eq. 4.7 - 4.9 )
-         			      3.a.1.II  Error of approximation as circular arc (Eq. 4.10 - 4.19 )
-         				    3.a.1.III Computer minimum of both approximation 
-         			3.a.2 Compute Error of the particle as in Eq.  4.20
-         			3.a.3 Compute fitness of the particle as in Eq. 4.21
-         3.b   Determine Lbest  
-               if (lbest_fitness<fitness)
-                  3.b.1   lbest=P_i
-         3.c      Determine Gbest.
-               if (lGest_fitness<fitness)
-                  3.c.1   lGest=P_i
-         3.d   Compute the Velocity for each dimension using Eq.3.2 
-         3.e   Bound Velocity using Eq.3.4
-         3.f   Update particle values using Eq. 3.3
-         3.g   Refine the location of the particle 
-         end .
-     end 
-4. return Gbest
-\end{verbatim}	
+\item Segmentation ALS2 as explained in Section \ref{sec:PolygonDivisionAlgorithm}. ( see Algorithm \ref{AlgS2alg} in Appendix \ref{ChapterstepExample} ). 
 \item Append best segmentation to List of segmentation ($L_s$) of this symbols. 
@@
 \begin{landscape}
+The steps of the first stroke (details of the outputs are in  Appendix \ref{ChapterstepExample} in Table \ref{tab:StepsStroke1Details}):
 \begin{scriptsize}
@@
 %Step & Discription of Step & Input & Steps & Output \\ \hline
-Stroke 1 & Extracting point and computing boundary data & -- &  see Section \ref{sec:Preprocessing} & 
-\begin{scriptsize}
-Stroke has 62 points = [P(198.0,153.0), P(197.0,150.0), P(197.0,145.0), P(198.0,140.0), P(203.0,125.0), P(205.0,120.0) , P(205.0,117.0), P(210.0,111.0) , P(214.0,103.0) , P(219.0,97.0), P(225.0,89.0) , P(229.0,85.0) , P(242.0,77.0) , P(247.0,73.0) , P(275.0,64.0) , P(287.0,62.0) , P(325.0,62.0) , P(333.0,63.0) , P(362.0,68.0) , P(372.0,72.0) , P(391.0,77.0) , P(406.0,83.0) , P(424.0,91.0) , P(440.0,100.0), P(474.0,133.0) , P(485.0,149.0) , P(492.0,164.0) , P(499.0,182.0) , P(501.0,191.0) , P(501.0,207.0) , P(499.0,222.0) , P(496.0,239.0) , P(471.0,292.0) , P(467.0,302.0) , P(456.0,315.0) , P(441.0,329.0) , P(427.0,339.0) , P(418.0,344.0) , P(398.0,354.0) , P(336.0,371.0) , P(283.0,376.0) , P(263.0,378.0) , P(242.0,378.0) , P(232.0,376.0) , P(213.0,373.0) , P(194.0,367.0) , P(177.0,362.0) , P(138.0,343.0) , P(129.0,336.0) , P(124.0,331.0) , P(120.0,321.0) , P(115.0,310.0) , P(113.0,297.0) , P(112.0,284.0) , P(114.0,256.0) , P(121.0,219.0) , P(128.0,202.0) , P(137.0,184.0) , P(143.0,172.0) , P(154.0,161.0) , P(172.0,146.0) , P(172.0,146.0)]
-
- Bonding box is  = Corner P(112.0, 62.0), w= 389.0, h=  316.0
- number of point in this stroke is   62
-\end{scriptsize}
- \\ \hline
-Possible Dominate point Extraction & for Computing details see Section\ref{sec:Preprocessing} &  Stroke points  &   &    Number of $P_{dp}$  = 11
- Pdp = [ 2 (197.0 , 145.0 ), 4 (203.0 , 125.0 ), 3 (198.0 , 140.0 ), 22 (424.0 , 91.0 ), 19 (372.0 , 72.0 ), 26 (492.0 , 164.0 ), 28 (501.0 , 191.0 ), 9 (219.0 , 97.0 ), 31 (496.0 , 239.0 ), 0 (198.0 , 153.0 ), 41 (263.0 , 378.0 )]
-
+Stroke 1 & Extracting point and computing boundary data & -- &   &
+
+	\includegraphics[scale=0.5]{afterDefense/clockstroke1.eps}
+
+  \\ \hline
+Possible Dominate point Extraction & for Computing details see Section\ref{sec:Preprocessing} &
+
+	\includegraphics[scale=0.5]{afterDefense/clockstroke1.eps}
+  &   &    
 \\ \hline 
 Segmentation & Segmentation of Ellipse & $P_{dp}$   &  Ellipse fitting  & 
-
 The circle is   a = 103.05715108084719 ,b = 103.06851874747797 , Center( 283.03365107691775,  181.05035287667172 )   with Error =  1.9471676673301757 , Percent =1.0695193086531243
 eff  = 0.27463462099275116
- Ellipse detected...
-
- \\ \hline
+
+ \includegraphics[scale=0.5]{afterDefense/ellipse.jpg.eps}
+ \\ \hline
+ 	\end{longtable}
+ 	\newpage
+ 	The steps of the second stroke:(details of the outputs are in  Appendix \ref{ChapterstepExample} in Table \ref{tab:StepsStroke1Details})
+
+\begin{longtable}{|p{2cm}|p{2cm}|p{5cm}|p{2cm}|p{5cm}|}
+\caption{ Output of System in Each Step of Segmentation of Second Stroke}
+\label{tab:StepsStroke2} \\
+
+\hline 
+\multicolumn{1}{|p{2cm}|}{\textbf{Step}} & 
+\multicolumn{1}{p{2cm}|}{\textbf{Description}} &
+\multicolumn{1}{p{5cm}|}{\textbf{Inputs}} &
+\multicolumn{1}{p{2cm}|}{\textbf{Details}} &
+\multicolumn{1}{p{5cm}|}{\textbf{Output}} 
+\\ \hline 
+\endfirsthead
+
+\multicolumn{3}{c}%
+{{\bfseries \tablename\ \thetable{} -- continued from previous page}} \\ \hline
+\multicolumn{1}{|p{2cm}|}{\textbf{Step}} & 
+\multicolumn{1}{p{2cm}|}{\textbf{Description}} &
+ \multicolumn{1}{p{5cm}|}{\textbf{Inputs}} &
+ \multicolumn{1}{p{2cm}|}{\textbf{Details}} &
+ \multicolumn{1}{p{5cm}|}{\textbf{Output}} 
+\\ \hline 
+\endhead
 
 Stroke 2 & Extracting point and computing boundary data & -- &  see Section \ref{sec:Preprocessing} &
-\begin{scriptsize}
-Stroke has 205 points = [P(201.0,179.0) , P(202.0,180.0) , P(202.0,182.0) , P(202.0,185.0) , P(202.0,187.0) , P(202.0,190.0) , P(202.0,192.0) , P(202.0,196.0) , P(202.0,202.0) , P(201.0,206.0) , P(200.0,219.0) , P(198.0,231.0) , P(198.0,234.0) , P(198.0,238.0) , P(197.0,244.0) , P(197.0,248.0) , P(197.0,251.0) , P(196.0,258.0) , %P(196.0,260.0) , P(196.0,263.0) , P(195.0,265.0) , P(195.0,266.0) , P(195.0,269.0) , P(195.0,271.0) , P(195.0,273.0) , P(195.0,278.0) , P(195.0,279.0) , P(195.0,281.0) , P(195.0,282.0) , P(195.0,284.0) , P(195.0,285.0) , P(195.0,288.0) , P(195.0,290.0) , P(195.0,295.0) , P(195.0,297.0) , P(195.0,298.0) , P(196.0,298.0) , P(196.0,299.0) , P(196.0,300.0) , P(196.0,301.0) , P(197.0,301.0) , P(198.0,301.0) , P(200.0,301.0) , P(203.0,301.0) , %P(206.0,301.0) , P(208.0,301.0) , P(213.0,301.0) , P(217.0,302.0) , P(219.0,302.0) , P(223.0,303.0) , P(234.0,304.0) , P(237.0,304.0) , P(240.0,306.0) , P(246.0,306.0) , P(247.0,307.0) , P(248.0,307.0) , P(249.0,307.0) , P(250.0,307.0) , P(250.0,306.0) , P(250.0,305.0) , P(250.0,304.0) , P(250.0,302.0) , P(251.0,299.0) , P(251.0,296.0) , P(251.0,294.0) , P(251.0,289.0) , P(252.0,285.0) , P(253.0,279.0) , P(253.0,274.0) , P(253.0,271.0) , P(253.0,264.0) , P(253.0,257.0) , P(254.0,252.0) , P(254.0,241.0) , P(254.0,237.0) , P(255.0,232.0) , P(255.0,229.0) , P(255.0,225.0) , P(256.0,212.0) , P(256.0,209.0) , P(256.0,206.0) , P(257.0,205.0) , P(257.0,200.0) , P(257.0,197.0) , P(258.0,195.0) , P(258.0,194.0) , P(258.0,189.0) , P(258.0,187.0) , P(258.0,186.0) , P(259.0,184.0) , P(259.0,181.0) , P(259.0,179.0) , P(259.0,178.0) , P(260.0,176.0) , P(260.0,174.0) , P(260.0,173.0) , P(260.0,172.0) , P(260.0,171.0) , P(260.0,169.0) , P(260.0,168.0) , P(260.0,167.0) , P(261.0,166.0) , P(261.0,165.0) , P(261.0,164.0) , P(261.0,162.0) , P(262.0,162.0) , P(263.0,162.0) , P(264.0,162.0) , P(268.0,162.0) , P(275.0,162.0) , P(279.0,162.0) , P(289.0,162.0) , P(291.0,162.0) , P(301.0,163.0) , P(305.0,164.0) , P(312.0,164.0) , P(315.0,165.0) , P(318.0,165.0) , P(319.0,165.0) , P(319.0,166.0) , P(320.0,168.0) , P(320.0,171.0) , P(320.0,173.0) , P(320.0,176.0) , P(321.0,182.0) , P(321.0,191.0) , P(322.0,198.0) , P(322.0,211.0) , P(323.0,226.0) , P(324.0,229.0) , P(324.0,238.0) , P(324.0,263.0) , P(325.0,272.0) , P(325.0,281.0) , P(325.0,288.0) , P(325.0,291.0) , P(325.0,297.0) , P(325.0,299.0) , P(325.0,301.0) , P(325.0,305.0) , P(324.0,306.0) , P(324.0,307.0) , P(323.0,308.0) , P(324.0,307.0) , P(325.0,307.0) , P(328.0,306.0) , P(332.0,306.0) , P(339.0,306.0) , P(347.0,305.0) , P(354.0,304.0) , P(365.0,304.0) , P(376.0,304.0) , P(379.0,304.0) , P(385.0,304.0) , P(391.0,304.0) , P(392.0,304.0) , P(393.0,304.0) , P(393.0,302.0) , P(393.0,301.0) , P(394.0,297.0) , P(394.0,290.0) , P(393.0,284.0) , P(393.0,274.0) , P(392.0,258.0) , P(391.0,245.0) , P(391.0,233.0) , P(389.0,213.0) , P(388.0,198.0) , P(388.0,189.0) , P(388.0,184.0) , P(388.0,178.0) , P(388.0,177.0) , P(388.0,176.0) , P(388.0,174.0)
-..................................................................................................................
-.................................................................................
-, P(388.0,173.0) , P(388.0,171.0) , P(388.0,170.0) , P(388.0,168.0) , P(388.0,167.0) , P(388.0,166.0) , P(388.0,165.0) , P(389.0,165.0) , P(390.0,165.0) , P(395.0,165.0) , P(401.0,165.0) , P(406.0,165.0) , P(418.0,166.0) , P(424.0,167.0) , P(426.0,167.0) , P(429.0,167.0) , P(430.0,167.0) , P(431.0,167.0) , P(432.0,167.0) , P(433.0,169.0) , P(434.0,170.0) , P(434.0,174.0) , P(436.0,185.0) , P(440.0,200.0) , P(442.0,220.0) , P(442.0,224.0) , P(446.0,267.0) , P(447.0,279.0) , P(448.0,286.0) , P(448.0,287.0) , P(448.0,287.0)]
-
- Bonding box is  = Corner P(195.0, 162.0), w= 253.0, h=  146.0
- \end{scriptsize}
-  \\ \hline
-  & Possible Dominate point Extraction for details see Section\ref{sec:Preprocessing} &  Stroke points  & 1. compute the $P_{pd}$  & \begin{scriptsize}   Number of $P_{pd}$  =  36
-
- Pdp = [ 25 (195.0 , 278.0 ), 30 (195.0 , 285.0 ), 33 (195.0 , 295.0 ), 27 (195.0 , 281.0 ), 34 (195.0 , 297.0 ), 46 (213.0 , 301.0 ), 47 (217.0 , 302.0 ), 50 (234.0 , 304.0 ), 28 (195.0 , 282.0 ), 94 (260.0 , 174.0 ), 93 (260.0 , 176.0 ), 95 (260.0 , 173.0 ), 98 (260.0 , 169.0 ), 48 (219.0 , 302.0 ), 99 (260.0 , 168.0 ), 109 (275.0 , 162.0 ), 113 (301.0 , 163.0 ), 94 (260.0 , 174.0 ), 132 (325.0 , 272.0 ), 136 (325.0 , 297.0 ), 112 (291.0 , 162.0 ), 137 (325.0 , 299.0 ), 110 (279.0 , 162.0 ), 144 (325.0 , 307.0 ), 147 (339.0 , 306.0 ), 150 (365.0 , 304.0 ), 132 (325.0 , 272.0 ), 168 (388.0 , 189.0 ), 171 (388.0 , 177.0 ), 174 (388.0 , 173.0 ), 146 (332.0 , 306.0 ), 175 (388.0 , 171.0 ), 183 (395.0 , 165.0 ), 186 (418.0 , 166.0 ), 169 (388.0 , 184.0 ), 191 (431.0 , 167.0 ) ]
-
-  \end{scriptsize}
-    \\ \hline
- \\ \hline 
+ 	\includegraphics[scale=0.5]{afterDefense/clockstroke2.jpg.eps} \\ \hline
+  & Possible Dominate point Extraction for details see Section\ref{sec:Preprocessing} &   	\includegraphics[scale=0.5]{afterDefense/clockstroke2.jpg.eps}  & 1. compute the $P_{pd}$  & 
+    \\ \hline
+
 Segmentation & Segmentation of Ellipse & $P_{dp}$   &  Ellipse fitting  & 
@@
  \\ \hline
- Segmentation & PSO Segmentation ALgS1 & $P_{dp}$  & ALgS1 (result if Figure \ref{fig:Stroke2Seg1labeled} ) & 
-Iteration 0 the global best fitness is 0.7 and error value is 675.0457637480733  and the particle is  Particles =  [  1 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 ]
-Iteration 10 the global best fitness is 0.7777777777777778 and error value is 719.5716152204503  and the particle is  Particles =  [  1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 ]
-Iteration 20 the global best fitness is 0.7777777777777778 and error value is 719.5716152204503  and the particle is  Particles =  [  1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 ]
- After swarm  the global best fitness is 0.875 and error value is 729.248312158671  and the particle is  Particles =  [  1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 ]
-
- F First Polygonization PSO = [ S0 Line[P(226.0,139.0) to P(241.0,209.0)]  ,      S1 Line[P(241.0,209.0) to P(266.0,226.0)]  ,      S2 Line[P(266.0,226.0) to P(296.0,128.0)]  ,      S3 Line[P(296.0,128.0) to P(312.0,216.0)]  ,    ]
- \\ \hline
-
-  Segmentation & PSO Segmentation ALgS2 & $P_{dp}$  & ALgS2  (result if Figure \ref{fig:Stroke2Seg2labeled} )&    
- Iteration 0 the global best fitness is 8.725462920280576E-11 and error value is 9.151490748550003  and the particle is  Particles =  [  1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 ]
-
-Iteration 10 the global best fitness is 3.126281513587432E-7 and error value is 8.61499750333895  and the particle is  Particles =  [  1 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 ]
-
-Iteration 20 the global best fitness is 3.126281513587432E-7 and error value is 8.61499750333895  and the particle is  Particles =  [  1 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 ]
-
- After swarm  the global best fitness is 3.126281513587432E-7 and error value is 8.61499750333895  and the particle is  Particles =  [  1 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 ]
-
-  second PSO =   Segments = [   S0 Line{P(226.0,139.0) to P(238.0,183.0)},      S1 Line{P(238.0,183.0) to P(241.0,213.0)},      S2 Line{P(241.0,213.0) to P(276.0,227.0)},      S3 Line{P(276.0,227.0) to P(287.0,144.0)},      S4 Curve{P(287.0,144.0) to P(310.0,131.0),  Center ( 299.02634312229503,137.80386092517725) Radius =11.907214051459519 },  S5 Curve{P(310.0,131.0) to P(312.0,216.0),  Center ( 82.44520390380922,172.2089170470769) Radius =231.70851748053468 },  S6 Line{P(312.0,216.0) to P(336.0,224.0)}, ]
-
- \\ \hline 
-Choosing best fit & Compare between ALgS1 and AlgS2 &  First PSO and Second PSO  &  (result if Figure \ref{fig:FinalSegmentation} ) & 
-   The error ALgS1 780.6652975044847  Algs2 161.60692855994625    Best = ALgS2  \\ \hline
-
+ Segmentation & PSO Segmentation ALgS1 & $P_{dp}$  & ALgS1   & Global best fitness is 0.875 and error value is 729.248312158671 
+ \includegraphics[scale=0.7]{afterDefense/Stroke2Seg1labeled.jpg.eps}
+ \\ \hline 
+  Segmentation & PSO Segmentation ALgS2 & $P_{dp}$  & ALgS2   & Error value is 8.61499750333895
+ \includegraphics[scale=0.7]{afterDefense/Stroke2Segmentaion.jpg.eps}
+ \\ \hline 
+Choosing best fit & Compare between ALgS1 and AlgS2 &  First PSO and Second PSO  &   & The error ALgS1 780.6652975044847  Algs2 161.60692855994625    Best = ALgS2
+
+\includegraphics[scale=0.7]{afterDefense/Stroke2Seg2labeled.jpg.eps}	
+  \\ \hline 
 		\end{longtable}
-%}
-
-%\end{table}
-
 \end{scriptsize}
-
 \end{landscape}
-
-\begin{figure}
-	\centering
-	\subfigure [AlgS1 Output]{
-		\includegraphics{afterDefense/Stroke2Seg1labeled.jpg.eps}
-	\label{fig:Stroke2Seg1labeled}}
-
-	\subfigure[AlgS2 Output]{	\includegraphics{afterDefense/Stroke2Seg2labeled.jpg.eps}	\label{fig:Stroke2Seg2labeled}
-	}
-	\subfigure[Final Segmentation of Symbol]{		\label{fig:FinalSegmentation}	\includegraphics[scale=0.65]{afterDefense/FinalSegmentation.jpg.eps}
-	}
-	\caption{Segmentation Ouputs For Second Stroke}
-\end{figure}
-
-The final output of this step is list $L_s$ of segments There are  9  segments in this symbol
+The final output of this step is list $L_s$ of segments There are  9  segments in this symbol (\ref{fig:FinalSegmentationLabeled})
 
@@
 \end{itemize}
+\begin{figure}
+	\centering
+ 		\includegraphics[scale=0.65]{afterDefense/FinalSegmentation.jpg.eps}
+
+		\label{fig:FinalSegmentationLabeled}
+	\caption{Final Segmentation of Symbol} The segments are labeled on the symbol. 
+\end{figure}
 
@@
 
-\begin{figure}
-	\centering
-		\includegraphics[scale=0.85]{afterDefense/FinalSegmentation.jpg.eps}
-	\caption{Best Segmantation }
-	\label{fig:segmentatiofinal.jpg}
-\end{figure}
-
-%\begin{enumerate}
-%\begin{itemize}
-
 	The input of the step is the list of segmentation $L_s$ that is used to compute the feature vector.
 Figure \ref{fig:segmentatiofinal.jpg} shows the best segmentation as given by the previous step.
-This step can be summarized as firstly using $L_s$ to computer the feature vector and then introduce this vector the SVM classifier to classify it to one of the trained categories.
-The details of each step of feature extraction are detailed in the next algorithm.
+This step can be summarized as firstly using $L_s$ to computer the feature vector (Total of 70 features) and then introduce this vector the SVM classifier to classify it to one of the trained categories.
+The details of each step of feature extraction are detailed in the Algorithm \ref{FeatureAlg} .
 Feature extraction step is divided into four steps where each feature set is computed separately then all sets are appended to construct a final feature vector.
@@
 \begin{enumerate}
-	\item Computing Structural and geometrical Features (FS1).
-This step computes number of parallel, perpendicular line and similar structural features.
-The following algorithm is used to determine these features.
-
-	\begin{verbatim}
-Initialize 
-  1. For each segment Si in list Ls  do the following. 
-      1.a If segment si is a line do the following 
-            1.a.I. Compute the slope and intercept of line si. 
-            1.a.II Increment Parallel count  if 
-                  parallel with each other line in Ls
-            1.a.III Increment PerpendicularCount if 
-                    perpendicular with each other line in Ls 
-            1.a.IV Increment Intersection Count if 
-                    intersected with any other line. 
-                    1.a.II.A  if Intersection found determine 
-                              type of intersection or (X,T,L)
-      1.b If segment si is curve do the following. 
-          1.b.I Increment Intersect if
-               Intersect other line or curve
-    1.b.II Increment number 
-                 of curves and primitives. 
-4. return Feature_Vector_FS1
-\end{verbatim}
-
-	\item Rubine Feature Set (FS2).  Compute Rubine features as in \cite{gestureexample12} 
-	\item Statistical Features (FS3) compute the moments as in \cite{zernike61}
-	\item Global shape properties set (FS4). Firstly compute the convex hull of the stroke points then compute the rest of the features.
-
-	\item Introduce the computed feature vector to the classifier and check the label of the detected symbol. 
+	\item Computing Structural and geometrical Features(FS1).
+This step computes number of parallel, perpendicular line and similar structural features [13 features].
+Algorithm \ref{FeatureAlg} in Appendix \ref{ChapterstepExample}  is used to determine these features.
+ 	\item Rubine Feature Set (FS2) [12 features] .  Compute Rubine features as in \cite{gestureexample12} 
+	\item Statistical Features (FS3) compute the moments as in \cite{zernike61} for N=10 there are 32 features.
+	\item Global shape properties set (FS4). Firstly compute the convex hull of the stroke points then compute the rest of the features [13 features]. 
+	\item Introduce the computed feature vector (All 70 features) to the classifier and check the label of the detected symbol. 
 \end{enumerate}
@@
 \begin{landscape}
+
+More details are in Table \ref{tab:StepsRegDetails} in Appendix \ref{ChapterstepExample}
+
 \begin{scriptsize}
 
- \begin{longtable}{|p{2cm}|p{2cm}|p{2cm}|p{2cm}|p{13cm}|}
+ \begin{longtable}{|p{2cm}|p{2cm}|p{5cm}|p{2cm}|p{10cm}|}
 \caption{Detailed Output of System in Each Step of Recognition}
@@
 \multicolumn{1}{p{2cm}|}{\textbf{Description}} &
-\multicolumn{1}{p{2cm}|}{\textbf{Inputs}} &
+\multicolumn{1}{p{5cm}|}{\textbf{Inputs}} &
 \multicolumn{1}{p{2cm}|}{\textbf{Details}} &
-\multicolumn{1}{p{13cm}|}{\textbf{Output}} 
+\multicolumn{1}{p{10cm}|}{\textbf{Output}} 
 \\ \hline 
@@
 
-\hline
 \multicolumn{3}{c}%
 {{\bfseries \tablename\ \thetable{} -- continued from previous page}} \\  
+\hline
 \multicolumn{1}{|p{2cm}|}{\textbf{Step}} & 
 \multicolumn{1}{p{2cm}|}{\textbf{Description}} &
- \multicolumn{1}{p{2cm}|}{\textbf{Inputs}} &
+ \multicolumn{1}{p{5cm}|}{\textbf{Inputs}} &
  \multicolumn{1}{p{2cm}|}{\textbf{Details}} &
- \multicolumn{1}{p{13cm}|}{\textbf{Output}} 
+ \multicolumn{1}{p{10cm}|}{\textbf{Output}} 
 \\ \hline 
@@
 
- Feature Extraction of FS1 & Preprocessing of Structural Featues  &   There are  9  segments in this symbol $L_s$ = 
-   Segment S0  = Ellipse
-   Segment S1  = Line
-   Segment S2  = Line
-   Segment S3  = Line
-   Segment S4  = Line
-   Segment S5  = Curve
-   Segment S6  = Curve
-   Segment S7  = Line
-   Segment S8  = Line  &  (see First Algorithm in Section \ref{sec:steprec} )  &    Line from  24 x= 241.0y=213.0  to 49 x2= 276.0y2=227.0  is\textbf{ parallel} to Line from  163 x= 312.0y=216.0  to 185 x2= 336.0y2=224.0
-
- Line from  0 x= 226.0y=139.0  to 14 x2= 238.0y2=183.0 \textbf{ intersects }Line from  14 x= 238.0y=183.0  to 24 x2= 241.0y2=213.0
-
- Line from  0 x= 226.0y=139.0  to 14 x2= 238.0y2=183.0  \textbf{intersects} Line from  24 x= 241.0y=213.0  to 49 x2= 276.0y2=227.0
-
- Line from  0 x= 226.0y=139.0  to 14 x2= 238.0y2=183.0 \textbf{ intersects} Line from  49 x= 276.0y=227.0  to 98 x2= 287.0y2=144.0
-
- Line from  0 x= 226.0y=139.0  to 14 x2= 238.0y2=183.0\textbf{  intersects }Line from  163 x= 312.0y=216.0  to 185 x2= 336.0y2=224.0
-
- Line from  0 x= 226.0y=139.0  to 14 x2= 238.0y2=183.0 \textbf{ intersects} Line from  185 x= 336.0y=224.0  to 223 x2= 344.0y2=137.0
- %Line from  14 x= 238.0y=183.0  to 24 x2= 241.0y2=213.0  intersects Line from  24 x= 241.0y=213.0  to 49 x2= 276.0y2=227.0
- %Line from  14 x= 238.0y=183.0  to 24 x2= 241.0y2=213.0  intersects Line from  49 x= 276.0y=227.0  to 98 x2= 287.0y2=144.0
- %Line from  14 x= 238.0y=183.0  to 24 x2= 241.0y2=213.0  intersects Line from  163 x= 312.0y=216.0  to 185 x2= 336.0y2=224.0
- %Line from  14 x= 238.0y=183.0  to 24 x2= 241.0y2=213.0  intersects Line from  185 x= 336.0y=224.0  to 223 x2= 344.0y2=137.0
- %Line from  24 x= 241.0y=213.0  to 49 x2= 276.0y2=227.0  intersects Line from  49 x= 276.0y=227.0  to 98 x2= 287.0y2=144.0
- %Line from  24 x= 241.0y=213.0  to 49 x2= 276.0y2=227.0  intersects Line from  163 x= 312.0y=216.0  to 185 x2= 336.0y2=224.0
- %Line from  24 x= 241.0y=213.0  to 49 x2= 276.0y2=227.0  intersects Line from  185 x= 336.0y=224.0  to 223 x2= 344.0y2=137.0
- %Line from  49 x= 276.0y=227.0  to 98 x2= 287.0y2=144.0  intersects Line from  163 x= 312.0y=216.0  to 185 x2= 336.0y2=224.0
-
- Line from  49 x= 276.0y=227.0  to 98 x2= 287.0y2=144.0  intersects Line from  185 x= 336.0y=224.0  to 223 x2= 344.0y2=137.0
-
- Line from  163 x= 312.0y=216.0  to 185 x2= 336.0y2=224.0  intersects Line from  185 x= 336.0y=224.0  to 223 x2= 344.0y2=137.0
+ Feature Extraction of FS1 & Preprocessing of Structural Featues  &  \includegraphics[scale=0.5]{afterDefense/FinalSegmentation.jpg.eps}
+ &  (see First Algorithm in Section \ref{sec:steprec} )  &    Line 1    is\textbf{ parallel} to Line from  L3
+
+ Line 1  \textbf{ intersects }Line 2
+  Line 2  \textbf{ intersects }Line 3
+   Line 3  \textbf{ intersects }Line 4
+    Line 4  \textbf{ intersects }Line 5
 
@@
 \begin{scriptsize}
-   Feature Zernike moments 0 158.87520185460517   ,    Feature Zernike moments 0 7.007557168632727   ,    Feature Zernike moments 1 34.26844061831789   ,    Feature Zernike moments 1 6.658086415801404   ,    Feature Zernike moments 2 85.70032052974454   ,    Feature Zernike moments 2 9.441591321498212   ,    Feature Zernike moments 3 11.73170471301825   ,    Feature Zernike moments 3 66.52997844880663   ,    Feature Zernike moments 4 20.42774127507497   ,    Feature Zernike moments 4 10.196341027811322   ,    Feature Zernike moments 5 52.821799077785066   ,    Feature Zernike moments 5 12.078236804336935   ,    Feature Zernike moments 6 27.748912050714   ,    Feature Zernike moments 6 9.870152472028517   ,    Feature Zernike moments 7 26.079803494397133   ,    Feature Zernike moments 7 28.109003184347657   ,    Feature Zernike moments 8 23.32097186060531   ,    Feature Zernike moments 8 7.632681250023639   ,    Feature Zernike moments 9 6.5841386188472395   ,    Feature Zernike moments 9 51.34195037191517   ,    Feature Zernike moments 10 30.122054353503337   ,    Feature Zernike moments 10 26.902895199361584   ,    Feature Zernike moments 11 1.8295982451659387   ,    Feature Zernike moments 11 29.93196640188875   ,    Feature Zernike moments 12 64.0879253626976   ,    Feature Zernike moments 12 31.811780133580722   ,    Feature Zernike moments 13 23.614051070558084   ,    Feature Zernike moments 13 1.8043229719502216   ,    Feature Zernike moments 14 113.92181136864328   ,    Feature Zernike moments 14 67.26888519955787   ,    Feature Zernike moments 15 47.7397132180995   ,    Feature Zernike moments 15 31.490125533238682   ,    Feature Zernike moments 16 10.320112066817984   ,    Feature Zernike moments 16
+   Feature Zernike moments 0 158.87520185460517   ,    Feature Zernike moments 0 7.007557168632727   ,    Feature Zernike moments 1 34.26844061831789   ,    Feature Zernike moments 1 6.658086415801404   ,.......
+......
+....
+.......
+.......
+.......
+.....
+.....
+.......
+.......
+.......
+.....
+.....
 \end{scriptsize}
@@
 FS4   &  &Segment List $L_s$   &  &  \begin{scriptsize} 
- The convex hull    [ X = 383.0   Y = 146.0  Time = 1250452568824,  X = 380.0   Y = 139.0  Time = 1250452568816,  X = 368.0   Y = 119.0  Time = 1250452568800,  X = 363.0   Y = 113.0  Time = 1250452568792,  X = 352.0   Y = 101.0  Time = 1250452568776,  X = 348.0   Y = 97.0  Time = 1250452568768,  X = 342.0   Y = 92.0  Time = 1250452568760,  X = 335.0   Y = 88.0  Time = 1250452568752,  X = 329.0   Y = 85.0  Time = 1250452568744, % X = 322.0   Y = 83.0  Time = 1250452568736,  X = 302.0   Y = 78.0  Time = 1250452568728,  X = 293.0   Y = 77.0  Time = 1250452568720,  X = 282.0   Y = 76.0  Time = 1250452568712,  X = 250.0   Y = 76.0  Time = 1250452568688,  X = 239.0   Y = 77.0  Time = 1250452568680,  X = 229.0   Y = 79.0  Time = 1250452568672,  X = 220.0   Y = 81.0  Time = 1250452568664,  X = 217.0   Y = 83.0  Time = 1250452568656,  X = 201.0   Y = 96.0  Time = 1250452568648,  X = 196.0   Y = 103.0  Time = 1250452568640,  X = 186.0   Y = 118.0  Time = 1250452568624,  X = 182.0   Y = 127.0  Time = 1250452568608,  X = 180.0   Y = 209.0  Time = 1250452569139,  X = 180.0   Y = 247.0  Time = 1250452569092,  X = 181.0   Y = 253.0  Time = 1250452569080,  X = 185.0   Y = 260.0  Time = 1250452569073,  X = 190.0   Y = 265.0  Time = 1250452569056,  X = 194.0   Y = 267.0  Time = 1250452569047,  X = 213.0   Y = 274.0  Time = 1250452569031,  X = 220.0   Y = 276.0  Time = 1250452569023,  X = 256.0   Y = 284.0  Time = 1250452568995,  X = 266.0   Y = 284.0  Time = 1250452568976,
- ...................................................X = 274.0   Y = 283.0  Time = 1250452568969,  X = 282.0   Y = 281.0  Time = 1250452568960,  X = 290.0   Y = 278.0  Time = 1250452568952,  X = 298.0   Y = 274.0  Time = 1250452568944,  X = 319.0   Y = 262.0  Time = 1250452568936,  X = 339.0   Y = 243.0  Time = 1250452568912,  X = 357.0   Y = 222.0  Time = 1250452568889,  X = 368.0   Y = 202.0  Time = 1250452568872,  X = 378.0   Y = 183.0  Time = 1250452568864,  X = 384.0   Y = 167.0  Time = 1250452568848,  X = 384.0   Y = 160.0  Time = 1250452568840]
+ The convex hull    [ ]
  Area of convex hull =  -34160.5   preimiter of hull 656.1576413855265
@@
   Feature Centroid time 1.2504525968640884E12  , Feature Centroid time difference 28319.088339222613   ,    Feature area of convexhull 34160.5   ,    Feature area of convexhull/area of symbol 1.2365344240932454   ,    Feature N.Points ConvexHull/ N.
-points symbol 0.15302491103202848   ,    Feature Convext Perimeter/symbol perimeter 0.556529278903672   ,    Feature  percentage of convex area to bounding box area 19.493542609351433   ,    Feature area  35519.5   ,    Feature diff area of symbol to area of bounding box 16.290771116138764   ,    Feature  Ration between w/h  0.9807692307692307   ,    Feature  Log of ration between hight and width -0.008433167536862813   ,
- Feature  Log of computed area 4.550466843684004   ,    Feature  abs of area 27626.0   ,    Feature log of  abs of area 4.441318007475848   ,    Feature  Density(length/size)  4.046853548165911   ,    Feature Mean CentroidalRadius x 272.48042704626334   ,    Feature Mean CentroidalRadius y 177.19928825622776   ,    Feature Mean t 1.2504525970626514E12   ,    Feature Mean time difference 148.69395017793593   ,    Feature Mean Radius 62.000799881019084
+points symbol 0.15302491103202848   ,    Feature Convext Perimeter/symbol perimeter 0.556529278903672   ,    .......
+.......
+.......
+.....
+.....
+.......
+.......
+.......
+.....
+.....
   \end{scriptsize}
@@
 
-   \begin{scriptsize}
-   Feature Number of primitives 3.0   ,    Feature Number of segments 9.0   ,    Feature Line count 6.0   ,    Feature Curves Count 3.0   ,    Feature Ellipse Count 1.0   ,    Feature Intersection T type -1.0   ,    Feature Intersection L type 4.0   ,    Feature Intersection X type -1.0   ,    Feature Parallel Count 1.0   ,    Feature PerPendicular Count -1.0   ,    Feature Intersection Lines Count 4.0   ,    Feature Min Radius 1.7976931348623157E308   ,    Feature Max Radius 231.70851748053468   ,    Feature Centroid x 92.54770318021201   ,    Feature Centroid y 101.3851590106007   ,    Feature Centroid time 1.2504525968640884E12   ,    Feature Centroid time difference 28319.088339222613   ,    Feature area of convexhull 34160.5   ,    Feature area of convexhull/area of symbol 1.2365344240932454   ,    Feature N.Points ConvexHull/ N.
-points symbol 0.15302491103202848   ,    Feature Convext Perimeter/symbol perimeter 0.556529278903672   ,    Feature  percentage of convex area to bounding box area 19.493542609351433   ,    Feature area  35519.5   ,    Feature diff area of symbol to area of bounding box 16.290771116138764   ,    Feature  Ration between w/h  0.9807692307692307   ,    Feature  Log of ration between hight and width -0.008433167536862813   ,    Feature Zernike moments 0 158.87520185460517   ,    Feature Zernike moments 0 7.007557168632727   ,    Feature Zernike moments 1 34.26844061831789   ,    Feature Zernike moments 1 6.658086415801404   ,    Feature Zernike moments 2 85.70032052974454   ,    Feature Zernike moments 2 9.441591321498212   ,    Feature Zernike moments 3 11.73170471301825   ,    Feature Zernike moments 3 66.52997844880663   ,    Feature Zernike moments 4 20.42774127507497   ,    %Feature Zernike moments 4 10.196341027811322   ,    Feature Zernike moments 5 52.821799077785066   ,    Feature Zernike moments 5 12.078236804336935   ,    Feature Zernike moments 6 27.748912050714   ,    Feature Zernike moments 6 % 9.870152472028517   ,    Feature Zernike moments 7 26.079803494397133   ,    Feature Zernike moments 7 28.109003184347657   ,    Feature Zernike moments 8 23.32097186060531   ,    Feature Zernike moments 8 7.632681250023639   ,    Feature Zernike moments 9 6.5841386188472395   ,    Feature Zernike moments 9 51.34195037191517   ,    Feature Zernike moments 10 30.122054353503337   ,    Feature Zernike moments 10 26.902895199361584   ,    Feature Zernike moments 11 1.8295982451659387   ,    Feature Zernike moments 11 29.93196640188875   ,    Feature Zernike moments 12 64.0879253626976   ,    Feature Zernike moments 12 31.811780133580722   ,    Feature Zernike moments 13 23.614051070558084   ,    Feature Zernike moments 1    1.8043229719502216   ,
-%Feature Zernike moments 14 113.92181136864328   ,    Feature Zernike moments 14 67.26888519955787   ,    Feature Zernike moments 15 47.7397132180995   ,    Feature Zernike moments 15 31.490125533238682   ,
-Feature Zernike moments 16 10.320112066817984   ,    Feature Zernike moments 16 4.007106145033957   ,    Feature Rubine 0 -1.0   ,    Feature Rubine 1 -1.0   ,    Feature Rubine 2 477.6107201476952   ,    Feature Rubine 3 0.6368010415482491   ,    Feature Rubine 4 169.49926253526885   ,    Feature Rubine 5 0.8908593331996381   ,    Feature Rubine 6 0.454279262624981   ,    Feature Rubine 7 1179.0172885029808   ,    Feature Rubine 8 4.712388980384688   ,    Feature Rubine 9 70.80157037633609   ,    Feature Rubine 10 55.4044569978846   ,    Feature Rubine 11 9.140625   ,    Feature Rubine 12 41783.0   ,    Feature  Log of computed area 4.550466843684004   ,    Feature  abs of area 27626.0   ,    Feature log of  abs of area 4.441318007475848   ,    Feature  Density(length/size)  4.046853548165911   ,    Feature Mean CentroidalRadius x 272.48042704626334   ,    Feature Mean CentroidalRadius y 177.19928825622776   ,    Feature Mean t 1.2504525970626514E12   ,    Feature Mean time difference 148.69395017793593   ,    Feature Mean Radius 62.000799881019084   ,
-   \end{scriptsize} \\ \hline
+  [FS1,FS2,FS3,FS4]  \\ \hline
    Classification  & SVM classifier decision & Feature Vector  &  & 
@@
 \end{scriptsize}
-
 \end{landscape}

### 2009-09-26 draft/ConFut.tex
after approval of dr. samia

@@
 The next step in this research is to complete the clustering algorithm for fully automated sketch recognition.
-This will enable testing the sytsem for more than single symbol at a time.
+This will enable testing the system for more than single symbol at a time.
 The segmentation algorithm may be more powerful if it can segments the stroke into more types of primitives other than the line and curve.
@@
 %Each new stroke is checked if it can be a symbol or is a part of already drawn un completed symbol.
+\appendix
+\chapter{Appendix: Step by Step Example Details}
+%\addcontentsline{toc}{chapter}{Appendix: Step by Step Example Details}
+\chaptermark{Appendix}
+\markboth{Appendix}{Appendix}
+\label{ChapterstepExample}%\chapter{Step by Step Example}
+%In this chapter a step by step example is presented to clarify the presented system.
+Let us make an assumption that the user will draw a clock using two strokes.
+Figure \ref{fig:clock} show the two strokes and how the user draws them.
+The next steps demonstrate how output of the system after each step.
+The first section \ref{sec:stepseg} describe how the segmentation is done step by step.
+Section \ref{sec:steprec} shows the details of feature calculation and the classification process.
+
+\section{Step by Step Segmentation}
+\label{apsec:stepseg}
+
+In this step the user draw one or more stroke to draw a symbol.
+Figure \ref{fig:clock} illustrate the sequence and direction of user draw.
+In this section, each stroke user draw will be labeled as it illustrated in Figure \ref{fig:clockstroke1} and Figure \ref{fig:clockstroke2} .
+Table \ref{tab:StepsStroke1Details} shows the detailed output of the program at each step.
+The input of this step is the stroke points and the output is a list of segments that will be used in the recognition process.
+The algorithm that computes the segmentation is demonstrated as follows:
+
+%\begin{enumerate}
+
+%\item Extracting $P_dp$ as in Section \ref{sec:Preprocessing} and Algorithm \ref{extractpdp}
+\begin{algorithm}
+\caption{Extracting $P_dp$}
+\label{extractpdp}
+\begin{verbatim}
+  1. For each point in Stroke do the following:
+  		1.a Compute distance from previous point.
+  		1.b Compute Time difference from previous point and add it to Time_list.  
+  		1.c Compute Velocity and add it to Vel_list. 
+  		1.d Compute Direction and add it to Dir_list.
+  	 	1.e Compute Curvature and add it to Cur_list.(Eq.4.1) 
+  	end (for loop)
+  2. For each  list computed in previous step do the follwing:   
+  	 2.a Compute average of all points. 
+  	 2.b set average as threshold.  
+  	 2.c Create region from continuous points larger than threshold. 
+  	 2.d for each region do the following. 
+  	 	   2.d.1 find the maximum 
+  	 	   2.d.1 Add point to P_dp 
+  	  end
+     end 
+\end{verbatim}	
+\end{algorithm}
+%\item Ellipse Fitting: which is an attempt to fit as an ellipse as in Section \ref{sec:EllipseDetection} 
+%he axes of the ellipse are estimated as the $width/2$ and $height/2$ of the stroke bounding box where $width$ is width and $height$ is height of the bonding box.
+\begin{algorithm}
+\caption{Ellipse Fitting Algorithm}
+\label{ellipseFitAlg}
+\begin{verbatim}
+1. Initialize the axes of ellipse  and center by bounding box information
+2. Repeat the following for few loops: 
+	2.a Compute error in Equation  4.2 
+	2.b Update center of ellipse and axis from error. 
+3. Compute percent in Equation 4.4 
+4. Computer eff in Equation 4.3 
+5. if eff < 0.2  then fit as ellipse otherwise move to segment. 
+\end{verbatim}	
+\end{algorithm}
+%\item Segmentation ALS1 as explained in Section \ref{sec:SwarmSegmentation}. The alogrithm will be similar to the PSO algorithm  in section \ref{sec:ParticleSwarmAlgorithm}.
+
+\begin{algorithm}
+\caption{ First Swarm Algorithm}
+\label{AlgS1alg}
+\begin{verbatim}
+Initialize 
+  1. Generate N particles with M dimensions (P_i...P_N) 
+  2. Generate velocities for each particle for M dimension randomly. 
+  3. Repeat till Maximum number of iteration OR Error Threshold  
+         For each particle P_i do the following 
+         3.a   Compute fitness  (fitness) 
+         				3.a.1 Compute the Error of approximation as in Eq.4.5
+         				3.a.2. Computer the fitness function as in Eq. 4.4 
+         3.b   Determine Lbest  
+               if (lbest_fitness<fitness)
+                  3.b.1   lbest=P_i
+         3.c      Determine Gbest.
+               if (lGest_fitness<fitness)
+                  3.c.1   lGest=P_i
+         3.d   Compute the Velocity for each dimension using Eq.3.2 
+         3.e   Bound Velocity using Eq.3.4
+         3.f   Update particle values using Eq. 3.3
+         3.g   Refine the location of the particle 
+         end .
+     end 
+4. return Gbest
+\end{verbatim}
+\end{algorithm}	
+%\item Segmentation ALS2  as explained in Section \ref{sec:PolygonDivisionAlgorithm}
+\begin{algorithm}
+\caption{ Second Swarm Algorithm}
+\label{AlgS2alg}
+\begin{verbatim}
+Initialize 
+  1. Generate N particles with M dimensions (P_i...P_N) 
+  2. Generate velocities for each particle for M dimension randomly. 
+  3. Repeat till Maximum number of iteration OR Error Threshold  
+         For each particle P_i do the following 
+         3.a   Compute fitness  (fitness) 
+              3.a.1 For each Segment in the particle do the following
+         				    3.a.1.I   Error of approximation as Line (Eq. 4.7 - 4.9 )
+         			      3.a.1.II  Error of approximation as circular arc (Eq. 4.10 - 4.19 )
+         				    3.a.1.III Computer minimum of both approximation 
+         			3.a.2 Compute Error of the particle as in Eq.  4.20
+         			3.a.3 Compute fitness of the particle as in Eq. 4.21
+         3.b   Determine Lbest  
+               if (lbest_fitness<fitness)
+                  3.b.1   lbest=P_i
+         3.c      Determine Gbest.
+               if (lGest_fitness<fitness)
+                  3.c.1   lGest=P_i
+         3.d   Compute the Velocity for each dimension using Eq.3.2 
+         3.e   Bound Velocity using Eq.3.4
+         3.f   Update particle values using Eq. 3.3
+         3.g   Refine the location of the particle 
+         end .
+     end 
+4. return Gbest
+\end{verbatim}	
+\end{algorithm}
+
+\begin{landscape}
+\begin{scriptsize}
+%\begin{table}
+	%\centering 
+%\scalebox{0.6}{		
+\begin{longtable}{|p{2cm}|p{2cm}|p{2cm}|p{2cm}|p{13cm}|}
+\caption{Detailed Output of System in Each Step of Segmentation}
+\label{tab:StepsStroke1Details} \\
+
+\hline 
+\multicolumn{1}{|p{2cm}|}{\textbf{Step}} & 
+\multicolumn{1}{p{2cm}|}{\textbf{Description}} &
+\multicolumn{1}{p{2cm}|}{\textbf{Inputs}} &
+\multicolumn{1}{p{2cm}|}{\textbf{Details}} &
+\multicolumn{1}{p{13cm}|}{\textbf{Output}} 
+\\ \hline 
+\endfirsthead
+
+\multicolumn{3}{c}%
+{{\bfseries \tablename\ \thetable{} -- continued from previous page}} \\
+\hline 
+\multicolumn{1}{|p{2cm}|}{\textbf{Step}} & 
+\multicolumn{1}{p{2cm}|}{\textbf{Description}} &
+ \multicolumn{1}{p{2cm}|}{\textbf{Inputs}} &
+ \multicolumn{1}{p{2cm}|}{\textbf{Details}} &
+ \multicolumn{1}{p{13cm}|}{\textbf{Output}} 
+\\ \hline 
+\endhead
+
+%Step & Discription of Step & Input & Steps & Output \\ \hline
+Stroke 1 & Extracting point and computing boundary data & -- &  see Section \ref{sec:Preprocessing} & 
+\begin{scriptsize}
+Stroke has 62 points = [P(198.0,153.0), P(197.0,150.0), P(197.0,145.0), P(198.0,140.0), P(203.0,125.0), P(205.0,120.0) , P(205.0,117.0), P(210.0,111.0) , P(214.0,103.0) , P(219.0,97.0), P(225.0,89.0) , P(229.0,85.0) , P(242.0,77.0) , P(247.0,73.0) , P(275.0,64.0) , P(287.0,62.0) , P(325.0,62.0) , P(333.0,63.0) , P(362.0,68.0) , P(372.0,72.0) , P(391.0,77.0) , P(406.0,83.0) , P(424.0,91.0) , P(440.0,100.0), P(474.0,133.0) , P(485.0,149.0) , P(492.0,164.0) , P(499.0,182.0) , P(501.0,191.0) , P(501.0,207.0) , P(499.0,222.0) , P(496.0,239.0) , P(471.0,292.0) , P(467.0,302.0) , P(456.0,315.0) , P(441.0,329.0) , P(427.0,339.0) , P(418.0,344.0) , P(398.0,354.0) , P(336.0,371.0) , P(283.0,376.0) , P(263.0,378.0) , P(242.0,378.0) , P(232.0,376.0) , P(213.0,373.0) , P(194.0,367.0) , P(177.0,362.0) , P(138.0,343.0) , P(129.0,336.0) , P(124.0,331.0) , P(120.0,321.0) , P(115.0,310.0) , P(113.0,297.0) , P(112.0,284.0) , P(114.0,256.0) , P(121.0,219.0) , P(128.0,202.0) , P(137.0,184.0) , P(143.0,172.0) , P(154.0,161.0) , P(172.0,146.0) , P(172.0,146.0)]
+
+ Bonding box is  = Corner P(112.0, 62.0), w= 389.0, h=  316.0
+ number of point in this stroke is   62
+\end{scriptsize}
+ \\ \hline
+Possible Dominate point Extraction & for Computing details see Section\ref{sec:Preprocessing} &  Stroke points  &   &    Number of $P_{dp}$  = 11
+ Pdp = [ 2 (197.0 , 145.0 ), 4 (203.0 , 125.0 ), 3 (198.0 , 140.0 ), 22 (424.0 , 91.0 ), 19 (372.0 , 72.0 ), 26 (492.0 , 164.0 ), 28 (501.0 , 191.0 ), 9 (219.0 , 97.0 ), 31 (496.0 , 239.0 ), 0 (198.0 , 153.0 ), 41 (263.0 , 378.0 )]
+
+\\ \hline 
+Segmentation & Segmentation of Ellipse & $P_{dp}$   &  Ellipse fitting  &
+
+  The circle is   a = 103.05715108084719 ,b = 103.06851874747797 , Center( 283.03365107691775,  181.05035287667172 )   with Error =  1.9471676673301757 , Percent =1.0695193086531243
+eff  = 0.27463462099275116  
+ Ellipse detected...
+
+ \\ \hline
+
+Stroke 2 & Extracting point and computing boundary data & -- &  see Section \ref{sec:Preprocessing} &
+\begin{scriptsize}
+Stroke has 205 points = [P(201.0,179.0) , P(202.0,180.0) , P(202.0,182.0) , P(202.0,185.0) , P(202.0,187.0) , P(202.0,190.0) , P(202.0,192.0) , P(202.0,196.0) , P(202.0,202.0) , P(201.0,206.0) , P(200.0,219.0) , P(198.0,231.0) , P(198.0,234.0) , P(198.0,238.0) , P(197.0,244.0) , P(197.0,248.0) , P(197.0,251.0) , P(196.0,258.0) , %P(196.0,260.0) , P(196.0,263.0) , P(195.0,265.0) , P(195.0,266.0) , P(195.0,269.0) , P(195.0,271.0) , P(195.0,273.0) , P(195.0,278.0) , P(195.0,279.0) , P(195.0,281.0) , P(195.0,282.0) , P(195.0,284.0) , P(195.0,285.0) , P(195.0,288.0) , P(195.0,290.0) , P(195.0,295.0) , P(195.0,297.0) , P(195.0,298.0) , P(196.0,298.0) , P(196.0,299.0) , P(196.0,300.0) , P(196.0,301.0) , P(197.0,301.0) , P(198.0,301.0) , P(200.0,301.0) , P(203.0,301.0) , %P(206.0,301.0) , P(208.0,301.0) , P(213.0,301.0) , P(217.0,302.0) , P(219.0,302.0) , P(223.0,303.0) , P(234.0,304.0) , P(237.0,304.0) , P(240.0,306.0) , P(246.0,306.0) , P(247.0,307.0) , P(248.0,307.0) , P(249.0,307.0) , P(250.0,307.0) , P(250.0,306.0) , P(250.0,305.0) , P(250.0,304.0) , P(250.0,302.0) , P(251.0,299.0) , P(251.0,296.0) , P(251.0,294.0) , P(251.0,289.0) , P(252.0,285.0) , P(253.0,279.0) , P(253.0,274.0) , P(253.0,271.0) , P(253.0,264.0) , P(253.0,257.0) , P(254.0,252.0) , P(254.0,241.0) , P(254.0,237.0) , P(255.0,232.0) , P(255.0,229.0) , P(255.0,225.0) , P(256.0,212.0) , P(256.0,209.0) , P(256.0,206.0) , P(257.0,205.0) , P(257.0,200.0) , P(257.0,197.0) , P(258.0,195.0) , P(258.0,194.0) , P(258.0,189.0) , P(258.0,187.0) , P(258.0,186.0) , P(259.0,184.0) , P(259.0,181.0) , P(259.0,179.0) , P(259.0,178.0) , P(260.0,176.0) , P(260.0,174.0) , P(260.0,173.0) , P(260.0,172.0) , P(260.0,171.0) , P(260.0,169.0) , P(260.0,168.0) , P(260.0,167.0) , P(261.0,166.0) , P(261.0,165.0) , P(261.0,164.0) , P(261.0,162.0) , P(262.0,162.0) , P(263.0,162.0) , P(264.0,162.0) , P(268.0,162.0) , P(275.0,162.0) , P(279.0,162.0) , P(289.0,162.0) , P(291.0,162.0) , P(301.0,163.0) , P(305.0,164.0) , P(312.0,164.0) , P(315.0,165.0) , P(318.0,165.0) , P(319.0,165.0) , P(319.0,166.0) , P(320.0,168.0) , P(320.0,171.0) , P(320.0,173.0) , P(320.0,176.0) , P(321.0,182.0) , P(321.0,191.0) , P(322.0,198.0) , P(322.0,211.0) , P(323.0,226.0) , P(324.0,229.0) , P(324.0,238.0) , P(324.0,263.0) , P(325.0,272.0) , P(325.0,281.0) , P(325.0,288.0) , P(325.0,291.0) , P(325.0,297.0) , P(325.0,299.0) , P(325.0,301.0) , P(325.0,305.0) , P(324.0,306.0) , P(324.0,307.0) , P(323.0,308.0) , P(324.0,307.0) , P(325.0,307.0) , P(328.0,306.0) , P(332.0,306.0) , P(339.0,306.0) , P(347.0,305.0) , P(354.0,304.0) , P(365.0,304.0) , P(376.0,304.0) , P(379.0,304.0) , P(385.0,304.0) , P(391.0,304.0) , P(392.0,304.0) , P(393.0,304.0) , P(393.0,302.0) , P(393.0,301.0) , P(394.0,297.0) , P(394.0,290.0) , P(393.0,284.0) , P(393.0,274.0) , P(392.0,258.0) , P(391.0,245.0) , P(391.0,233.0) , P(389.0,213.0) , P(388.0,198.0) , P(388.0,189.0) , P(388.0,184.0) , P(388.0,178.0) , P(388.0,177.0) , P(388.0,176.0) , P(388.0,174.0)
+..................................................................................................................
+.................................................................................
+, P(388.0,173.0) , P(388.0,171.0) , P(388.0,170.0) , P(388.0,168.0) , P(388.0,167.0) , P(388.0,166.0) , P(388.0,165.0) , P(389.0,165.0) , P(390.0,165.0) , P(395.0,165.0) , P(401.0,165.0) , P(406.0,165.0) , P(418.0,166.0) , P(424.0,167.0) , P(426.0,167.0) , P(429.0,167.0) , P(430.0,167.0) , P(431.0,167.0) , P(432.0,167.0) , P(433.0,169.0) , P(434.0,170.0) , P(434.0,174.0) , P(436.0,185.0) , P(440.0,200.0) , P(442.0,220.0) , P(442.0,224.0) , P(446.0,267.0) , P(447.0,279.0) , P(448.0,286.0) , P(448.0,287.0) , P(448.0,287.0)]
+
+ Bonding box is  = Corner P(195.0, 162.0), w= 253.0, h=  146.0
+ \end{scriptsize}
+  \\ \hline
+  & Possible Dominate point Extraction for details see Section\ref{sec:Preprocessing} &  Stroke points  & 1. compute the $P_{pd}$  & \begin{scriptsize}   Number of $P_{pd}$  =  36
+
+ Pdp = [ 25 (195.0 , 278.0 ), 30 (195.0 , 285.0 ), 33 (195.0 , 295.0 ), 27 (195.0 , 281.0 ), 34 (195.0 , 297.0 ), 46 (213.0 , 301.0 ), 47 (217.0 , 302.0 ), 50 (234.0 , 304.0 ), 28 (195.0 , 282.0 ), 94 (260.0 , 174.0 ), 93 (260.0 , 176.0 ), 95 (260.0 , 173.0 ), 98 (260.0 , 169.0 ), 48 (219.0 , 302.0 ), 99 (260.0 , 168.0 ), 109 (275.0 , 162.0 ), 113 (301.0 , 163.0 ), 94 (260.0 , 174.0 ), 132 (325.0 , 272.0 ), 136 (325.0 , 297.0 ), 112 (291.0 , 162.0 ), 137 (325.0 , 299.0 ), 110 (279.0 , 162.0 ), 144 (325.0 , 307.0 ), 147 (339.0 , 306.0 ), 150 (365.0 , 304.0 ), 132 (325.0 , 272.0 ), 168 (388.0 , 189.0 ), 171 (388.0 , 177.0 ), 174 (388.0 , 173.0 ), 146 (332.0 , 306.0 ), 175 (388.0 , 171.0 ), 183 (395.0 , 165.0 ), 186 (418.0 , 166.0 ), 169 (388.0 , 184.0 ), 191 (431.0 , 167.0 ) ]
+
+  \end{scriptsize}
+    \\ \hline
+ \\ \hline 
+Segmentation & Segmentation of Ellipse & $P_{dp}$   &  Ellipse fitting  & 
+  The Ellipse is   a = 125.5 ,b = 125.5 , Center( 320.5,  235.0 )   with Error =  7.074966523515405 , Percent =1.1481777694994926 certainty  =  0.08114368920921669  
+ \\ \hline
+ Segmentation & PSO Segmentation ALgS1 & $P_{dp}$  & ALgS1 (result if Figure \ref{fig:Stroke2Seg1labeled} ) & 
+Iteration 0 the global best fitness is 0.7 and error value is 675.0457637480733  and the particle is  Particles =  [  1 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 ]
+Iteration 10 the global best fitness is 0.7777777777777778 and error value is 719.5716152204503  and the particle is  Particles =  [  1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 ]
+Iteration 20 the global best fitness is 0.7777777777777778 and error value is 719.5716152204503  and the particle is  Particles =  [  1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 ]
+ After swarm  the global best fitness is 0.875 and error value is 729.248312158671  and the particle is  Particles =  [  1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 ]
+
+ F First Polygonization PSO = [ S0 Line[P(226.0,139.0) to P(241.0,209.0)]  ,      S1 Line[P(241.0,209.0) to P(266.0,226.0)]  ,      S2 Line[P(266.0,226.0) to P(296.0,128.0)]  ,      S3 Line[P(296.0,128.0) to P(312.0,216.0)]  ,    ]
+ \\ \hline
+
+  Segmentation & PSO Segmentation ALgS2 & $P_{dp}$  & ALgS2  (result if Figure \ref{fig:Stroke2Seg2labeled} )&    
+ Iteration 0 the global best fitness is 8.725462920280576E-11 and error value is 9.151490748550003  and the particle is  Particles =  [  1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 ]
+
+Iteration 10 the global best fitness is 3.126281513587432E-7 and error value is 8.61499750333895  and the particle is  Particles =  [  1 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 ]
+
+Iteration 20 the global best fitness is 3.126281513587432E-7 and error value is 8.61499750333895  and the particle is  Particles =  [  1 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 ]
+
+ After swarm  the global best fitness is 3.126281513587432E-7 and error value is 8.61499750333895  and the particle is  Particles =  [  1 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 ]
+
+  second PSO =   Segments = [   S0 Line{P(226.0,139.0) to P(238.0,183.0)},      S1 Line{P(238.0,183.0) to P(241.0,213.0)},      S2 Line{P(241.0,213.0) to P(276.0,227.0)},      S3 Line{P(276.0,227.0) to P(287.0,144.0)},      S4 Curve{P(287.0,144.0) to P(310.0,131.0),  Center ( 299.02634312229503,137.80386092517725) Radius =11.907214051459519 },  S5 Curve{P(310.0,131.0) to P(312.0,216.0),  Center ( 82.44520390380922,172.2089170470769) Radius =231.70851748053468 },  S6 Line{P(312.0,216.0) to P(336.0,224.0)}, ]
+
+ \\ \hline 
+Choosing best fit & Compare between ALgS1 and AlgS2 &  First PSO and Second PSO  &  (result if Figure \ref{fig:FinalSegmentation} ) & 
+   The error ALgS1 780.6652975044847  Algs2 161.60692855994625    Best = ALgS2  \\ \hline
+
+		\end{longtable}
+%}
+
+%\end{table}
+
+\end{scriptsize}
+
+\end{landscape}
+
+\begin{figure}
+	\centering
+	\subfigure [AlgS1 Output]{
+		\includegraphics{afterDefense/Stroke2Seg1labeled.jpg.eps}
+	\label{fig:Stroke2Seg1labeled}}
+
+	\subfigure[AlgS2 Output]{	\includegraphics{afterDefense/Stroke2Seg2labeled.jpg.eps}	\label{fig:Stroke2Seg2labeled}
+	}
+	\subfigure[Final Segmentation of Symbol]{		\label{fig:FinalSegmentation}	\includegraphics[scale=0.65]{afterDefense/FinalSegmentation.jpg.eps}
+	}
+	\caption{Segmentation Outputs For Second Stroke}
+\end{figure}
+
+The final output of this step is list $L_s$ of segments There are  9  segments in this symbol
+
+\begin{itemize}
+	\item    Segment S0  = Ellipse ( Stroke 1)
+   \item Segment S1  =   Line ( Stroke 2)
+     \item  Segment S2 = Line ( Stroke 2)
+     \item  Segment S3  =  Line ( Stroke 2)
+     \item Segment S4  =   Line ( Stroke 2)
+  \item    Segment S5  =  Curve  ( Stroke 2)
+     \item  Segment S6  =  Curve  ( Stroke 2)
+     \item  Segment S7  =  Line   ( Stroke 2)
+     \item  Segment S8  =  Line   ( Stroke 2)
+\end{itemize}
+
+\section{Step by Step Recognition}
+\label{apsec:steprec}
+
+\begin{figure}
+	\centering
+		\includegraphics[scale=0.85]{afterDefense/FinalSegmentation.jpg.eps}
+	\caption{Best Segmentation }
+	\label{fig:segmentatiofinal.jpg}
+\end{figure}
+
+%\begin{enumerate}
+%\begin{itemize}
+
+	The input of the step is the list of segmentation $L_s$ that is used to compute the feature vector.
+Figure \ref{fig:segmentatiofinal.jpg} shows the best segmentation as given by the previous step.
+This step can be summarized as firstly using $L_s$ to computer the feature vector and then introduce this vector the SVM classifier to classify it to one of the trained categories.
+The details of each step of feature extraction are detailed in the next algorithm.
+Feature extraction step is divided into four steps where each feature set is computed separately then all sets are appended to construct a final feature vector.
+	\begin{enumerate}
+	\item Computing Structural and geometrical Features(FS1).
+This step computes number of parallel, perpendicular line and similar structural features [13 features].
+Algorithm \ref{FeatureAlg} in Appendix \ref{ChapterstepExample}  is used to determine these features.
+ 	\item Rubine Feature Set (FS2) [12 features] .  Compute Rubine features as in \cite{gestureexample12} 
+	\item Statistical Features (FS3) compute the moments as in \cite{zernike61} for N=10 there are 32 features.
+	\item Global shape properties set (FS4). Firstly compute the convex hull of the stroke points then compute the rest of the features [13 features]. 
+	\item Introduce the computed feature vector (All 70 features) to the classifier and check the label of the detected symbol. 
+\end{enumerate}
+
+	 \begin{algorithm}
+\caption{Computing Structural Features}
+\label{FeatureAlg}
+	\begin{verbatim}
+Initialize 
+  1. For each segment Si in list Ls  do the following. 
+      1.a If segment si is a line do the following 
+            1.a.I. Compute the slope and intercept of line si. 
+            1.a.II Increment Parallel count  if 
+                  parallel with each other line in Ls
+            1.a.III Increment PerpendicularCount if 
+                    perpendicular with each other line in Ls 
+            1.a.IV Increment Intersection Count if 
+                    intersected with any other line. 
+                    1.a.II.A  if Intersection found determine 
+                              type of intersection or (X,T,L)
+      1.b If segment si is curve do the following. 
+          1.b.I Increment Intersect if
+               Intersect other line or curve
+    1.b.II Increment number 
+                 of curves and primitives. 
+4. return Feature_Vector_FS1
+\end{verbatim}	
+ \end{algorithm}
+
+\begin{landscape}
+\begin{scriptsize}
+
+ \begin{longtable}{|p{2cm}|p{2cm}|p{2cm}|p{2cm}|p{13cm}|}
+\caption{Detailed Output of System in Each Step of Recognition}
+\label{tab:StepsRegDetails} \\
+
+\hline 
+\multicolumn{1}{|p{2cm}|}{\textbf{Step}} & 
+\multicolumn{1}{p{2cm}|}{\textbf{Description}} &
+\multicolumn{1}{p{2cm}|}{\textbf{Inputs}} &
+\multicolumn{1}{p{2cm}|}{\textbf{Details}} &
+\multicolumn{1}{p{13cm}|}{\textbf{Output}} 
+\\ \hline 
+\endfirsthead
+
+\multicolumn{3}{c}%
+{{\bfseries \tablename\ \thetable{} -- continued from previous page}} \\
+\hline 
+\multicolumn{1}{|p{2cm}|}{\textbf{Step}} & 
+\multicolumn{1}{p{2cm}|}{\textbf{Description}} &
+ \multicolumn{1}{p{2cm}|}{\textbf{Inputs}} &
+ \multicolumn{1}{p{2cm}|}{\textbf{Details}} &
+ \multicolumn{1}{p{13cm}|}{\textbf{Output}} 
+\\ \hline 
+\endhead
+
+ Feature Extraction of FS1 & Preprocessing of Structural Featues  &   There are  9  segments in this symbol $L_s$ = 
+   Segment S0  = Ellipse
+   Segment S1  = Line
+   Segment S2  = Line
+   Segment S3  = Line
+   Segment S4  = Line
+   Segment S5  = Curve
+   Segment S6  = Curve
+   Segment S7  = Line
+   Segment S8  = Line  &  (see First Algorithm in Section \ref{sec:steprec} )  &    Line from  24 x= 241.0y=213.0  to 49 x2= 276.0y2=227.0  is\textbf{ parallel} to Line from  163 x= 312.0y=216.0  to 185 x2= 336.0y2=224.0
+
+ Line from  0 x= 226.0y=139.0  to 14 x2= 238.0y2=183.0 \textbf{ intersects }Line from  14 x= 238.0y=183.0  to 24 x2= 241.0y2=213.0
+
+ Line from  0 x= 226.0y=139.0  to 14 x2= 238.0y2=183.0  \textbf{intersects} Line from  24 x= 241.0y=213.0  to 49 x2= 276.0y2=227.0
+
+ Line from  0 x= 226.0y=139.0  to 14 x2= 238.0y2=183.0 \textbf{ intersects} Line from  49 x= 276.0y=227.0  to 98 x2= 287.0y2=144.0
+
+ Line from  0 x= 226.0y=139.0  to 14 x2= 238.0y2=183.0\textbf{  intersects }Line from  163 x= 312.0y=216.0  to 185 x2= 336.0y2=224.0
+
+ Line from  0 x= 226.0y=139.0  to 14 x2= 238.0y2=183.0 \textbf{ intersects} Line from  185 x= 336.0y=224.0  to 223 x2= 344.0y2=137.0
+ %Line from  14 x= 238.0y=183.0  to 24 x2= 241.0y2=213.0  intersects Line from  24 x= 241.0y=213.0  to 49 x2= 276.0y2=227.0
+ %Line from  14 x= 238.0y=183.0  to 24 x2= 241.0y2=213.0  intersects Line from  49 x= 276.0y=227.0  to 98 x2= 287.0y2=144.0
+ %Line from  14 x= 238.0y=183.0  to 24 x2= 241.0y2=213.0  intersects Line from  163 x= 312.0y=216.0  to 185 x2= 336.0y2=224.0
+ %Line from  14 x= 238.0y=183.0  to 24 x2= 241.0y2=213.0  intersects Line from  185 x= 336.0y=224.0  to 223 x2= 344.0y2=137.0
+ %Line from  24 x= 241.0y=213.0  to 49 x2= 276.0y2=227.0  intersects Line from  49 x= 276.0y=227.0  to 98 x2= 287.0y2=144.0
+ %Line from  24 x= 241.0y=213.0  to 49 x2= 276.0y2=227.0  intersects Line from  163 x= 312.0y=216.0  to 185 x2= 336.0y2=224.0
+ %Line from  24 x= 241.0y=213.0  to 49 x2= 276.0y2=227.0  intersects Line from  185 x= 336.0y=224.0  to 223 x2= 344.0y2=137.0
+ %Line from  49 x= 276.0y=227.0  to 98 x2= 287.0y2=144.0  intersects Line from  163 x= 312.0y=216.0  to 185 x2= 336.0y2=224.0
+
+ Line from  49 x= 276.0y=227.0  to 98 x2= 287.0y2=144.0  intersects Line from  185 x= 336.0y=224.0  to 223 x2= 344.0y2=137.0
+
+ Line from  163 x= 312.0y=216.0  to 185 x2= 336.0y2=224.0  intersects Line from  185 x= 336.0y=224.0  to 223 x2= 344.0y2=137.0
+
+\\ \hline
+
+ FS1 & Structural and geometrical Features & Segment List $L_s$  &  & 
+ The final Feature vector of the FS1==>
+ature Number of primitives 3.0   ,    Feature Number of segments 9.0   ,    Feature Line count 6.0   ,    Feature Curves Count 3.0   ,    Feature Ellipse Count 1.0   ,    Feature Intersection T type -1.0   ,    Feature Intersection L type 4.0   ,    Feature Intersection X type -1.0   ,    Feature Parallel Count 1.0   ,    Feature PerPendicular Count -1.0   ,    Feature Intersection Lines Count 4.0   ,    Feature Min Radius 1.7976931348623157E308   ,    Feature Max Radius 231.70851748053468
+
+ \\ \hline
+FS2  & Rubine Features & Segment List $L_s$  &   see \cite{gestureexample12}   & 
+\begin{scriptsize}
+ Feature Rubine 0 -1.0   ,    Feature Rubine 1 -1.0   ,    Feature Rubine 2 477.6107201476952   ,    Feature Rubine 3 0.6368010415482491   ,    Feature Rubine 4 169.49926253526885   ,    Feature Rubine 5 0.8908593331996381   ,    Feature Rubine 6 0.454279262624981   ,    Feature Rubine 7 1179.0172885029808   ,    Feature Rubine 8 4.712388980384688   ,    Feature Rubine 9 70.80157037633609   ,    Feature Rubine 10 55.4044569978846   ,    Feature Rubine 11 9.140625   ,    Feature Rubine 12 41783.0   ,
+\end{scriptsize}
+ \\ \hline
+
+FS3 & Momments Computation   &  Segment List $L_s$  &  see \cite{zernike61}  &
+
+\begin{scriptsize}
+   Feature Zernike moments 0 158.87520185460517   ,    Feature Zernike moments 0 7.007557168632727   ,    Feature Zernike moments 1 34.26844061831789   ,    Feature Zernike moments 1 6.658086415801404   ,    Feature Zernike moments 2 85.70032052974454   ,    Feature Zernike moments 2 9.441591321498212   ,    Feature Zernike moments 3 11.73170471301825   ,    Feature Zernike moments 3 66.52997844880663   ,    Feature Zernike moments 4 20.42774127507497   ,    Feature Zernike moments 4 10.196341027811322   ,    Feature Zernike moments 5 52.821799077785066   ,    Feature Zernike moments 5 12.078236804336935   ,    Feature Zernike moments 6 27.748912050714   ,    Feature Zernike moments 6 9.870152472028517   ,    Feature Zernike moments 7 26.079803494397133   ,    Feature Zernike moments 7 28.109003184347657   ,    Feature Zernike moments 8 23.32097186060531   ,    Feature Zernike moments 8 7.632681250023639   ,    Feature Zernike moments 9 6.5841386188472395   ,    Feature Zernike moments 9 51.34195037191517   ,    Feature Zernike moments 10 30.122054353503337   ,    Feature Zernike moments 10 26.902895199361584   ,    Feature Zernike moments 11 1.8295982451659387   ,    Feature Zernike moments 11 29.93196640188875   ,    Feature Zernike moments 12 64.0879253626976   ,    Feature Zernike moments 12 31.811780133580722   ,    Feature Zernike moments 13 23.614051070558084   ,    Feature Zernike moments 13 1.8043229719502216   ,    Feature Zernike moments 14 113.92181136864328   ,    Feature Zernike moments 14 67.26888519955787   ,    Feature Zernike moments 15 47.7397132180995   ,    Feature Zernike moments 15 31.490125533238682   ,    Feature Zernike moments 16 10.320112066817984   ,    Feature Zernike moments 16
+\end{scriptsize}
+
+  \\ \hline
+FS4   &  &Segment List $L_s$   &  &  \begin{scriptsize} 
+ The convex hull    [ X = 383.0   Y = 146.0  Time = 1250452568824,  X = 380.0   Y = 139.0  Time = 1250452568816,  X = 368.0   Y = 119.0  Time = 1250452568800,  X = 363.0   Y = 113.0  Time = 1250452568792,  X = 352.0   Y = 101.0  Time = 1250452568776,  X = 348.0   Y = 97.0  Time = 1250452568768,  X = 342.0   Y = 92.0  Time = 1250452568760,  X = 335.0   Y = 88.0  Time = 1250452568752,  X = 329.0   Y = 85.0  Time = 1250452568744, % X = 322.0   Y = 83.0  Time = 1250452568736,  X = 302.0   Y = 78.0  Time = 1250452568728,  X = 293.0   Y = 77.0  Time = 1250452568720,  X = 282.0   Y = 76.0  Time = 1250452568712,  X = 250.0   Y = 76.0  Time = 1250452568688,  X = 239.0   Y = 77.0  Time = 1250452568680,  X = 229.0   Y = 79.0  Time = 1250452568672,  X = 220.0   Y = 81.0  Time = 1250452568664,  X = 217.0   Y = 83.0  Time = 1250452568656,  X = 201.0   Y = 96.0  Time = 1250452568648,  X = 196.0   Y = 103.0  Time = 1250452568640,  X = 186.0   Y = 118.0  Time = 1250452568624,  X = 182.0   Y = 127.0  Time = 1250452568608,  X = 180.0   Y = 209.0  Time = 1250452569139,  X = 180.0   Y = 247.0  Time = 1250452569092,  X = 181.0   Y = 253.0  Time = 1250452569080,  X = 185.0   Y = 260.0  Time = 1250452569073,  X = 190.0   Y = 265.0  Time = 1250452569056,  X = 194.0   Y = 267.0  Time = 1250452569047,  X = 213.0   Y = 274.0  Time = 1250452569031,  X = 220.0   Y = 276.0  Time = 1250452569023,  X = 256.0   Y = 284.0  Time = 1250452568995,  X = 266.0   Y = 284.0  Time = 1250452568976,
+ ...................................................X = 274.0   Y = 283.0  Time = 1250452568969,  X = 282.0   Y = 281.0  Time = 1250452568960,  X = 290.0   Y = 278.0  Time = 1250452568952,  X = 298.0   Y = 274.0  Time = 1250452568944,  X = 319.0   Y = 262.0  Time = 1250452568936,  X = 339.0   Y = 243.0  Time = 1250452568912,  X = 357.0   Y = 222.0  Time = 1250452568889,  X = 368.0   Y = 202.0  Time = 1250452568872,  X = 378.0   Y = 183.0  Time = 1250452568864,  X = 384.0   Y = 167.0  Time = 1250452568848,  X = 384.0   Y = 160.0  Time = 1250452568840]
+ Area of convex hull =  -34160.5   preimiter of hull 656.1576413855265
+
+ --------------------------------------------------------------------------------
+
+Final Feature vector for the FS4: 
+  Feature Centroid time 1.2504525968640884E12  , Feature Centroid time difference 28319.088339222613   ,    Feature area of convexhull 34160.5   ,    Feature area of convexhull/area of symbol 1.2365344240932454   ,    Feature N.Points ConvexHull/ N.
+points symbol 0.15302491103202848   ,    Feature Convext Perimeter/symbol perimeter 0.556529278903672   ,    Feature  percentage of convex area to bounding box area 19.493542609351433   ,    Feature area  35519.5   ,    Feature diff area of symbol to area of bounding box 16.290771116138764   ,    Feature  Ration between w/h  0.9807692307692307   ,    Feature  Log of ration between hight and width -0.008433167536862813   ,
+ Feature  Log of computed area 4.550466843684004   ,    Feature  abs of area 27626.0   ,    Feature log of  abs of area 4.441318007475848   ,    Feature  Density(length/size)  4.046853548165911   ,    Feature Mean CentroidalRadius x 272.48042704626334   ,    Feature Mean CentroidalRadius y 177.19928825622776   ,    Feature Mean t 1.2504525970626514E12   ,    Feature Mean time difference 148.69395017793593   ,    Feature Mean Radius 62.000799881019084
+  \end{scriptsize}
+    \\ \hline
+ Feature Vector & Appends the selected Feature FS1, FS2, FS3, FS4 & Previous Steps  & &
+
+   \begin{scriptsize}
+   Feature Number of primitives 3.0   ,    Feature Number of segments 9.0   ,    Feature Line count 6.0   ,    Feature Curves Count 3.0   ,    Feature Ellipse Count 1.0   ,    Feature Intersection T type -1.0   ,    Feature Intersection L type 4.0   ,    Feature Intersection X type -1.0   ,    Feature Parallel Count 1.0   ,    Feature PerPendicular Count -1.0   ,    Feature Intersection Lines Count 4.0   ,    Feature Min Radius 1.7976931348623157E308   ,    Feature Max Radius 231.70851748053468   ,    Feature Centroid x 92.54770318021201   ,    Feature Centroid y 101.3851590106007   ,    Feature Centroid time 1.2504525968640884E12   ,    Feature Centroid time difference 28319.088339222613   ,    Feature area of convexhull 34160.5   ,    Feature area of convexhull/area of symbol 1.2365344240932454   ,    Feature N.Points ConvexHull/ N.
+points symbol 0.15302491103202848   ,    Feature Convext Perimeter/symbol perimeter 0.556529278903672   ,    Feature  percentage of convex area to bounding box area 19.493542609351433   ,    Feature area  35519.5   ,    Feature diff area of symbol to area of bounding box 16.290771116138764   ,    Feature  Ration between w/h  0.9807692307692307   ,    Feature  Log of ration between hight and width -0.008433167536862813   ,    Feature Zernike moments 0 158.87520185460517   ,    Feature Zernike moments 0 7.007557168632727   ,    Feature Zernike moments 1 34.26844061831789   ,    Feature Zernike moments 1 6.658086415801404   ,    Feature Zernike moments 2 85.70032052974454   ,    Feature Zernike moments 2 9.441591321498212   ,    Feature Zernike moments 3 11.73170471301825   ,    Feature Zernike moments 3 66.52997844880663   ,    Feature Zernike moments 4 20.42774127507497   ,    %Feature Zernike moments 4 10.196341027811322   ,    Feature Zernike moments 5 52.821799077785066   ,    Feature Zernike moments 5 12.078236804336935   ,    Feature Zernike moments 6 27.748912050714   ,    Feature Zernike moments 6 % 9.870152472028517   ,    Feature Zernike moments 7 26.079803494397133   ,    Feature Zernike moments 7 28.109003184347657   ,    Feature Zernike moments 8 23.32097186060531   ,    Feature Zernike moments 8 7.632681250023639   ,    Feature Zernike moments 9 6.5841386188472395   ,    Feature Zernike moments 9 51.34195037191517   ,    Feature Zernike moments 10 30.122054353503337   ,    Feature Zernike moments 10 26.902895199361584   ,    Feature Zernike moments 11 1.8295982451659387   ,    Feature Zernike moments 11 29.93196640188875   ,    Feature Zernike moments 12 64.0879253626976   ,    Feature Zernike moments 12 31.811780133580722   ,    Feature Zernike moments 13 23.614051070558084   ,    Feature Zernike moments 1    1.8043229719502216   ,
+%Feature Zernike moments 14 113.92181136864328   ,    Feature Zernike moments 14 67.26888519955787   ,    Feature Zernike moments 15 47.7397132180995   ,    Feature Zernike moments 15 31.490125533238682   ,
+Feature Zernike moments 16 10.320112066817984   ,    Feature Zernike moments 16 4.007106145033957   ,    Feature Rubine 0 -1.0   ,    Feature Rubine 1 -1.0   ,    Feature Rubine 2 477.6107201476952   ,    Feature Rubine 3 0.6368010415482491   ,    Feature Rubine 4 169.49926253526885   ,    Feature Rubine 5 0.8908593331996381   ,    Feature Rubine 6 0.454279262624981   ,    Feature Rubine 7 1179.0172885029808   ,    Feature Rubine 8 4.712388980384688   ,    Feature Rubine 9 70.80157037633609   ,    Feature Rubine 10 55.4044569978846   ,    Feature Rubine 11 9.140625   ,    Feature Rubine 12 41783.0   ,    Feature  Log of computed area 4.550466843684004   ,    Feature  abs of area 27626.0   ,    Feature log of  abs of area 4.441318007475848   ,    Feature  Density(length/size)  4.046853548165911   ,    Feature Mean CentroidalRadius x 272.48042704626334   ,    Feature Mean CentroidalRadius y 177.19928825622776   ,    Feature Mean t 1.2504525970626514E12   ,    Feature Mean time difference 148.69395017793593   ,    Feature Mean Radius 62.000799881019084   ,
+   \end{scriptsize} \\ \hline
+   Classification  & SVM classifier decision & Feature Vector  &  & 
+   Symbol = Clock.
+
+ \\ \hline
+
+ 		\end{longtable}
+%}
+
+%\end{table}
+
+\end{scriptsize}
+
+\end{landscape}

### 2009-09-26 draft/Results.tex
after dr. mahmoud meeting
after the changes in examples
after spell check

@@
 		}
-	\caption[Hs-DB Symbol Set]{The symboles Set in \cite{HeloiseBeautification} } 
+	\caption[Hs-DB Symbol Set]{The symbols Set in \cite{HeloiseBeautification} } 
 	\label{fig:symbolSet}
@@
 		\begin{tabular}{|p{2cm}|c|c|c|c|c|c|c|c|c|c|c|c|c||p{2cm}|c|c|c|c|p{2cm}|}\hline 
-		Categorie / classified as ==$>$	&ellipse	&heart	&trapezoid	&pentagon	&arch	&hexagon	&square	&triangle	&cube	&cylinder	&parallelogram	&moon	&callout	&Number of samples per class	&TP count	&FP count	&Fn	&Precision	&Recall ( TP rate )\\ \hline
+		Category / classified as ==$>$	&ellipse	&heart	&trapezoid	&pentagon	&arch	&hexagon	&square	&triangle	&cube	&cylinder	&parallelogram	&moon	&callout	&Number of samples per class	&TP count	&FP count	&Fn	&Precision	&Recall ( TP rate )\\ \hline
 ellipse	&150	&20	&0	&0	&0	&5	&1	&0	&0	&0	&0	&1	&0	&152	&150	&27	&2	&0.8474	&0.9868\\ \hline 

### 2009-09-26 trunk/draft/System.tex
after spell check

@@
 After the best fit segmentation is chosen the parameter of each primitive is computed.
-For example, if the stroke is represented as an ellipse, the center and axii of the ellipse is computed and confirmed.
+For example, if the stroke is represented as an ellipse, the center and axis of the ellipse is computed and confirmed.
 For lines the slope, equation, start point and end point are computed.
@@
 \subsection{Structural and geometrical Features(FS1)}
- Features defines the structure of the geometrical symbol. Section \ref{sec:steprec} descripe in details how these 13 feature are computed.  
+ Features defines the structure of the geometrical symbol. Section \ref{sec:steprec} describe in details how these 13 feature are computed.  
 %\begin{description}
@@
 		\includegraphics{afterDefense/svm6.jpg.eps}
-	\caption{non Lineary problem }
+	\caption{non Linearly problem }
 	\label{fig:svm6}
@@
 Notice that the mapped data only occurs as inner product in the objective.
-Now we apply a little mathematically magic known as hilbern schmidt kernels, first applied to SVMs in \cite{svmintroduce}.
+Now we apply a little mathematically magic known as Hilbert Schmidt kernels, first applied to SVM in \cite{svmintroduce}.
 By Mercer's theorem, we know that for certain mapping $\theta$ and any two points $u$ and $v$, the inner product of the mapped points can be evaluated using the kernel function without ever explicitly knowing the mapping, e.g., $\theta(u)\theta(v)\equiv K(u,v)$.
@@
 
-SVM have been applied ot the a largest MNIST a Handwritten digit recognition dataset of 60000 training set and 10000 test set\cite{IjdarSherifPaper}.
+SVM have been applied to the a largest MNIST a Handwritten digit recognition dataset of 60000 training set and 10000 test set\cite{IjdarSherifPaper}.
 Sherif and Ezzat has performed a comparison between SVM and several other techniques on the MNIST dataset.
@@
  \\ \hline 
-  Segmentation & PSO Segmentation ALgS2 & $P_{dp}$  & ALgS2   & Error value is 8.61499750333895
+  Segmentation & PSO Segmentation AlgS2 & $P_{dp}$  & AlgS2   & Error value is 8.61499750333895
  \includegraphics[scale=0.7]{afterDefense/Stroke2Segmentaion.jpg.eps}
  \\ \hline 
-Choosing best fit & Compare between ALgS1 and AlgS2 &  First PSO and Second PSO  &   & The error ALgS1 780.6652975044847  Algs2 161.60692855994625    Best = ALgS2
+Choosing best fit & Compare between ALgS1 and AlgS2 &  First PSO and Second PSO  &   & The error ALgS1 780.6652975044847  AlgS2 161.60692855994625    Best = AlgS2
 
@@
  The final Feature vector of the FS1==>
-ature Number of primitives 3.0   ,    Feature Number of segments 9.0   ,    Feature Line count 6.0   ,    Feature Curves Count 3.0   ,    Feature Ellipse Count 1.0   ,    Feature Intersection T type -1.0   ,    Feature Intersection L type 4.0   ,    Feature Intersection X type -1.0   ,    Feature Parallel Count 1.0   ,    Feature PerPendicular Count -1.0   ,    Feature Intersection Lines Count 4.0   ,    Feature Min Radius 1.7976931348623157E308   ,    Feature Max Radius 231.70851748053468
+ature Number of primitives 3.0   ,    Feature Number of segments 9.0   ,    Feature Line count 6.0   ,    Feature Curves Count 3.0   ,    Feature Ellipse Count 1.0   ,    Feature Intersection T type -1.0   ,    Feature Intersection L type 4.0   ,    Feature Intersection X type -1.0   ,    Feature Parallel Count 1.0   ,    Feature Perpendicular Count -1.0   ,    Feature Intersection Lines Count 4.0   ,    Feature Min Radius 1.7976931348623157E308   ,    Feature Max Radius 231.70851748053468
 
@@
 
-FS3 & Momments Computation   &  Segment List $L_s$  &  see \cite{zernike61}  &
+FS3 & Moments Computation   &  Segment List $L_s$  &  see \cite{zernike61}  &
 
@@
  The convex hull    [ ]
- Area of convex hull =  -34160.5   preimiter of hull 656.1576413855265
+ Area of convex hull =  -34160.5   perimeter of hull 656.1576413855265
 

### 2009-09-26 draft/Results.tex

@@
 	 \includegraphics[scale=0.4]{paperImages/testAlg.jpg.eps}
-	 	\caption{Algorithm comparison on HS-DB} The recognition rate of different algorithms. 
+	 	\caption{Algorithm Comparison on HS-DB} The recognition rate of different algorithms. 
 	 	\label{fig:test1}
@@
 	 \includegraphics[scale=0.8]{results/AlgLD.eps}
-	 	\caption{Algorithm comparison on LD-DB} The recognition rate of different algorithms. 
+	 	\caption{Algorithm Comparison on LD-DB} The recognition rate of different algorithms. 
 	 	\label{fig:testLD}
@@
 	 \includegraphics[scale=0.8]{results/AlgEL.eps}
-	 	\caption{Algorithm comparison on EL-DB} The recognition rate of different algorithms. 
+	 	\caption{Algorithm Comparison on EL-DB} The recognition rate of different algorithms. 
 	 	\label{fig:testEL}
@@
 		\includegraphics{results/Alg1Ds.eps}
-	\caption{AlgS1 on different datasets}
+	\caption{AlgS1 on Different Datasets}
 	\label{fig:Alg1Ds}
@@
 		\includegraphics{results/Alg2Ds.eps}
-	\caption{AlgS2 on different Datasets }
+	\caption{AlgS2 on Different Datasets }
 	\label{fig:Alg2Ds}
@@
 	\includegraphics[scale=0.5]{results/samplesofHS-DB.jpg.eps}
-	\caption[Hs-DB shapes drawn by single user]{Hs-DB shapes drawn by single user \cite{HeloiseBeautification} }
+	\caption[Hs-DB Shapes Drawn by Single User]{Hs-DB shapes drawn by single user \cite{HeloiseBeautification} }
 	\label{fig:samplesofHS-DB.jpg}

### 2009-09-26 draft/System.tex

@@
 \subsection{Structural and geometrical Features(FS1)}
-Features defines the structure of the geometrical symbol. Section \ref{sec:steprec} descripe in details how these feature are computed. 
+ Features defines the structure of the geometrical symbol. Section \ref{sec:steprec} describe in details how these 13 feature are computed.  
 %\begin{description}
@@
 \item \emph{Size Ratio} Ratio between width to height of the symbol.
+		\item \emph{Log of size Ratio} Log of the ratio between width and height of symbol.
 	\item \emph{Ink density} Compute the density of points inside its bounding box\cite{GeometryAndDomain102}.   
  	\item \emph{Convex Hull Area} Area of convex hull with respect to area of bounding box of symbol.
+
 	\item \emph{Convex Hull Perimeter} Perimeter of convex hull with respect to total length of symbol.
+		\item \emph{Convex Hull Point} Number of points of convex hull with respect to Number of points of symbol.
 		\item \emph{Mean Centroidal radius} The Mean of the centroidal radius which is the distance from each point in the symbol to the center of gravity.
+				\item \emph{Center of Gravity} The center of gravity of the symbol (both in y and x direction).
+					\item \emph{Mean Time} The mean of the time values between each two successive points in the symbol. %Different strokes are appended to construct a single path.
+	%\item [Ra]
 
@@
 	%\item [Ra]		
+			\item \emph{Log of Area} Log of area of the symbol.
+	\item \emph{Diff Area} Absolute difference of the area of bounding box and area of symbol.
+
   \end{itemize}
@@
 		\includegraphics{afterDefense/svm1.jpg.eps}
-	\caption{Different Classifier planes }
+	\caption{Different Classifier Planes }
 	\label{fig:svm1}
@@
 		\includegraphics{afterDefense/svm2.jpg.eps}
-	\caption{Margin of Support planes }
+	\caption{Margin of SVM Support planes }
 	\label{fig:svm2}
@@
 		\includegraphics{afterDefense/svm5.jpg.eps}
-	\caption{Linearly Inseparable}
+	\caption{Example of SVM Linearly Inseparable Problem}
 	\label{fig:svm5}
@@
 		\includegraphics{afterDefense/svm6.jpg.eps}
-	\caption{non Lineary problem }
+	\caption{Example of an non Linearly SVM Problem }
 	\label{fig:svm6}
@@
 The input of this step is the a set of strokes, each stroke has a set of  points and the output is a list of segments that will be used in the recognition process.
-More detailed example and algorithms are demonstrated in the Appendex \ref{ChapterstepExample} \\
+More detailed example and algorithms are demonstrated in the Appendix \ref{ChapterstepExample} \\
 
@@
 \item Extracting $P_dp$ as in Section \ref{sec:Preprocessing} and Algorithm \ref{extractpdp}.
-\item Ellipse Fitting: which is an attempt to fit as an ellipse as in details in Section \ref{sec:EllipseDetection} and Algorithm \ref{ellipseFitAlg} in Appendex \ref{ChapterstepExample}.
+\item Ellipse Fitting: which is an attempt to fit as an ellipse as in details in Section \ref{sec:EllipseDetection} and Algorithm \ref{ellipseFitAlg} in Appendix \ref{ChapterstepExample}.
 \item Segmentation ALS1 as explained in Section \ref{sec:SwarmSegmentation}.
-The alogrithm will be similar to the PSO algorithm  in section \ref{sec:ParticleSwarmAlgorithm}.
-(see Algorithm \ref{AlgS1alg} in Appendex \ref{ChapterstepExample})
+The algorithm will be similar to the PSO algorithm  in section \ref{sec:ParticleSwarmAlgorithm}.
+(see Algorithm \ref{AlgS1alg} in Appendix \ref{ChapterstepExample})
 %he axes of the ellipse are estimated as the $width/2$ and $height/2$ of the stroke bounding box where $width$ is width and $height$ is height of the bonding box.
-\item Segmentation ALS2 as explained in Section \ref{sec:PolygonDivisionAlgorithm}. ( see Algorithm \ref{AlgS2alg} in Appendex \ref{ChapterstepExample} ). 
+\item Segmentation ALS2 as explained in Section \ref{sec:PolygonDivisionAlgorithm}. ( see Algorithm \ref{AlgS2alg} in Appendix \ref{ChapterstepExample} ). 
 \item Append best segmentation to List of segmentation ($L_s$) of this symbols. 
@@
 \begin{landscape}
-The steps of the first stroke (details of the ouputs are in  Appendex \ref{ChapterstepExample} in Table \ref{tab:StepsStroke1Details}):
+The steps of the first stroke (details of the outputs are in  Appendix \ref{ChapterstepExample} in Table \ref{tab:StepsStroke1Details}):
 \begin{scriptsize}
@@
  	\newpage
- 	The steps of the second stroke:(details of the ouputs are in  Appendex \ref{ChapterstepExample} in Table \ref{tab:StepsStroke1Details})
+ 	The steps of the second stroke:(details of the outputs are in  Appendix \ref{ChapterstepExample} in Table \ref{tab:StepsStroke1Details})
 
@@
 
+	The input of the step is the list of segmentation $L_s$ that is used to compute the feature vector.
+Figure \ref{fig:segmentatiofinal.jpg} shows the best segmentation as given by the previous step.
+This step can be summarized as firstly using $L_s$ to computer the feature vector (Total of 70 features) and then introduce this vector the SVM classifier to classify it to one of the trained categories.
+The details of each step of feature extraction are detailed in the Algorithm \ref{FeatureAlg} .
+Feature extraction step is divided into four steps where each feature set is computed separately then all sets are appended to construct a final feature vector.
+
 \begin{enumerate}
-	\item Computing Structural and geometrical Features (FS1).
-This step computes number of parallel, perpendicular line and similar structural features [10 features].
-Algorithm \ref{FeatureAlg} in Appendex \ref{ChapterstepExample}  is used to determine these features.
+	\item Computing Structural and geometrical Features(FS1).
+This step computes number of parallel, perpendicular line and similar structural features [13 features].
+Algorithm \ref{FeatureAlg} in Appendix \ref{ChapterstepExample}  is used to determine these features.
  	\item Rubine Feature Set (FS2) [12 features] .  Compute Rubine features as in \cite{gestureexample12} 
 	\item Statistical Features (FS3) compute the moments as in \cite{zernike61} for N=10 there are 32 features.
-	\item Global shape properties set (FS4). Firstly compute the convex hull of the stroke points then compute the rest of the features [12 features]. 
-	\item Introduce the computed feature vector to the classifier and check the label of the detected symbol. 
+	\item Global shape properties set (FS4). Firstly compute the convex hull of the stroke points then compute the rest of the features [13 features]. 
+	\item Introduce the computed feature vector (All 70 features) to the classifier and check the label of the detected symbol. 
 \end{enumerate}
@@
 
-More details are in Table \ref{tab:StepsRegDetails} in Appendex \ref{ChapterstepExample}
+More details are in Table \ref{tab:StepsRegDetails} in Appendix \ref{ChapterstepExample}
 

### 2009-09-26 trunk/draft/System.tex

@@
 \item \emph{Size Ratio} Ratio between width to height of the symbol.
-		\item \emp{Log of size Ratio} Log of the ratio between width and height of symbol.
+		\item \emph{Log of size Ratio} Log of the ratio between width and height of symbol.
 	\item \emph{Ink density} Compute the density of points inside its bounding box\cite{GeometryAndDomain102}.   
@@
 	%\item [Ra]		
-			\item \emp{Log of Area} Log of area of the symbol.
+			\item \emph{Log of Area} Log of area of the symbol.
 	\item \emph{Diff Area} Absolute difference of the area of bounding box and area of symbol.
@@
 		\includegraphics{afterDefense/svm6.jpg.eps}
-	\caption{non Linearly problem }
+	\caption{Example of an non Linearly SVM Problem }
 	\label{fig:svm6}
@@
 	\includegraphics[scale=0.5]{afterDefense/clockstroke1.eps}
-  &   &    
+  &   &    \includegraphics[scale=0.5]{afterDefense/stroke1withpdp.jpg.eps}
 \\ \hline 
@@
   & Possible Dominate point Extraction for details see Section\ref{sec:Preprocessing} &   	\includegraphics[scale=0.5]{afterDefense/clockstroke2.jpg.eps}  & 1. compute the $P_{pd}$  &
+
+  \includegraphics[scale=0.5]{afterDefense/stroke2WithPdp.jpg.eps}
     \\ \hline

### 2009-09-26 trunk/draft/System.tex

@@
 
-	\includegraphics[scale=0.5]{afterDefense/clockstroke1.eps}
+%	\includegraphics[scale=0.5]{afterDefense/clockstroke1.eps}
   &   &    \includegraphics[scale=0.5]{afterDefense/stroke1withpdp.jpg.eps}
@@
  	\includegraphics[scale=0.5]{afterDefense/clockstroke2.jpg.eps} \\ \hline
-  & Possible Dominate point Extraction for details see Section\ref{sec:Preprocessing} &   	\includegraphics[scale=0.5]{afterDefense/clockstroke2.jpg.eps}  & 1. compute the $P_{pd}$  &
-
+  & Possible Dominate point Extraction for details see Section\ref{sec:Preprocessing} &   	%\includegraphics[scale=0.5]{afterDefense/clockstroke2.jpg.eps}
+    & 1. compute the $P_{pd}$  &  
   \includegraphics[scale=0.5]{afterDefense/stroke2WithPdp.jpg.eps}

### 2009-10-10 draft/SketchMaster.tex
after the changes in examples
changes done to thesis after dr. nevin approval

@@
 % \usepackage command:
-
+%\usepackage{setspace}
 \usepackage{psfig}
@@
 \hyphenation{gno-mon-ly}
-
+%\usepackage{setspace}
 \usepackage {graphicx}
@@
 \usepackage{array}
+\usepackage{algorithmic}
+\usepackage{algorithm} 
+%\numberwithin{algorithm}{chapter}
 %\usepackage[T1]{fontenc}
@@
                                                         %======================
-\title{Sketch Recognition and Understanding}             % My theses title  using Particle Swarm Optimization Algorithm
-\author{Maha Mohammed Nabeel El Meseery}                               % Y name
+\title{SKETCH RECOGNITION AND UNDERSTANDING}            % My theses title  using Particle Swarm Optimization
+% Algorithm   Sketch Recognition and Understanding
+\author{Maha Mohamed Nabeel El Meseery Moustafa El Meseery}                               % Y name
 \prevdegrees{B.S.,Computer Engineering}       		% Your old degree
-\major{Computer Engineering}                            % My current major
-\degree{Master of Science}                              % My new degree
-\degreemonth{}                                      % When awarded. to be added when known
-\degreeyear{}                                      %
-\superad{Professor Nevin Darwish\\Professor Samia Mashali\\Assoc. Professor Magda Fayek\\}
+\major{COMPUTER ENGINEERING}                            % My current major
+\degree{MASTER OF SCIENCE}                              % My new degree
+\degreemonth{July}                                      % When awarded. to be added when known
+\degreeyear{2009}      
+\FacultyUniversity{Faculty of Engineering, Cairo University \\ Giza, Egypt \\ July 2009}
+\superad{}                                %
+%\superad{Professor Nevin Darwish\\Professor Samia Mashali\\Assoc. Professor Magda Fayek\\}
 %th today or first day of writing  December 26 2005 
 %\thesisdate[]                                     % Date of document.
-\DOCUMENTtype{Thesis}                                   % or DISSERTATION
-\Documenttype{Thesis}                                   % or Dissertation
-\documenttype{Thesis}                                   % or dissertation
+\DOCUMENTtype{ Thesis }                                   % or DISSERTATION
+\Documenttype{ Thesis }                                   % or Dissertation
+\documenttype{ Thesis }                                   % or dissertation
 
@@
 %\makecopyright                                          % if no copyright
-                                                    % APPROVAL                                                         %======================
+  %=======================================Supervisors page ===============================
+\newpage    
 \thispagestyle{empty}               % Don't do headers or footers.                                                  
-\newpage                                                             
  \begin{center}     
-\LARGE  
-\textbf{Sketch Recognition and Understanding}\\
+ SKETCH RECOGNITION AND UNDERSTANDING\\
 \normalsize by      \\ 
-\Large  \textbf{ Maha Mohamed Nabeel El Meseery}\\
+  Maha Mohamed Nabeel El Meseery Moustafa El Meseery\\
 \vspace*{0.4in}        
@@
 COMPUTER ENGINEERING\\
-
+\vspace*{0.4in} 
+Under the Supervision of 
+\end{center}
+  \vspace*{0.4in} 
+\begin{center}
+ 		\begin{tabular}{ccc}
+		Dr. Nevin Darwish  & Dr. Samia Mashali & Dr. Magda Fayek \\
+		 Professor   & Professor & Associate Professor  \\
+		 Faculty of Engineering& Electronic Research Institute & 		 Faculty of Engineering \\
+		  Cairo University && Cairo University\\
+		\end{tabular}
+\end{center}
+  \vspace*{0.4in}       
+  \vfill
+\begin{center}
+   Faculty of Engineering, Cairo University \\ Giza, Egypt \\ July 2009
 \end{center}
 
+  %=======================================Supervisors page ===============================
+                                                     % APPROVAL                                                         %======================
+\newpage       
+\thispagestyle{empty}               % Don't do headers or footers.
+
+\begin{center}     
+SKETCH RECOGNITION AND UNDERSTANDING\\
+\normalsize by      \\ 
+Maha Mohamed Nabeel El Meseery Moustafa El Meseery\\
+\vspace*{0.2cm}        
+A Thesis Submitted To the\\
+Faculty of Engineering, Cairo University\\
+in Partial Fulfillment of the\\
+Requirements for the Degree of\\
+MASTER OF SCIENCE
+in\\
+COMPUTER ENGINEERING\\
+\end{center}         
 Approved by\\                                           
-Examining Committee:\\
- \makebox[6in]{\hrulefill}\\ 
-Professor Nevin Darwish, Thesis Main Advisor\\
- \makebox[6in]{\hrulefill}\\ 
-Professor Samia Mashali,Thesis Advisor \\
- \makebox[6in]{\hrulefill}\\  
-Assoc. Professor Magda Fayek,Thesis Advisor\\ 
- \makebox[6in]{\hrulefill}\\ 
-Professor  Ashraf Abd El Wahab, Member\\
- \makebox[6in]{\hrulefill}\\ 
+Examining Committee:\\ \vspace*{0.2cm} 
+\begin{tabular*}{0.9\textwidth}{l   }
+ %\makebox[5in]{\hrulefill}\\ 
+\\ \hline 
+Professor Nevin Darwish, Thesis Main Advisor\\ \\  \hline  
+ %\makebox[5in]{\hrulefill}\\ 
+Professor Samia Mashali,Thesis Advisor \\  \\ \hline   
+ %\makebox[6in]{\hrulefill}\\  
+Assoc. Professor Magda Fayek,Thesis Advisor\\  \\  \hline  
+ %\makebox[5in]{\hrulefill}\\ 
+Professor  Ashraf Abd El Wahab, Member\\ \\ \hline  
+ %\makebox[6in]{\hrulefill}\\ 
 Professor Mohamed Zaki Abd ElMegeed, Member\\  
- \makebox[6in]{\hrulefill}\\ 
+ %\makebox[6in]{\hrulefill}\\ 
+ %\end{spacing}
+
+		\end{tabular*}
+
+%\begin{spacing}{2}
+% \makebox[5in]{\hrulefill}\\ 
+%Professor Nevin Darwish, Thesis Main Advisor\\
+% \makebox[5in]{\hrulefill}\\ 
+%Professor Samia Mashali,Thesis Advisor \\
+% \makebox[6in]{\hrulefill}\\  
+%Assoc. Professor Magda Fayek,Thesis Advisor\\ 
+% \makebox[5in]{\hrulefill}\\ 
+%Professor  Ashraf Abd El Wahab, Member\\
+% \makebox[6in]{\hrulefill}\\ 
+%Professor Mohamed Zaki Abd ElMegeed, Member\\
+% \makebox[6in]{\hrulefill}\\ 
+ %\end{spacing}
+ \begin{center}
+   Faculty of Engineering, Cairo University \\ Giza, Egypt \\ July 2009
+\end{center}
   % \\   you need...\\
@@
                                                         % FOREWORD PAGE
-%\begin{foreword}                                        %======================
-    											                             % Delete these if
-%\end{foreword}                                          % no foreword page.
-
-                                                        % OTHER PAGES
-                                                        %======================
-\tableofcontents                                        % Always needed...
-%\listoftables                                           % Delete if no tables.
-\listoffigures                                          % Delete if no figures.
 
@@
 % didn't finish it all Need some addint
-The need for new level of human computer interaction had lead the researchers to attempt improve the current interfaces.
+The need for new level of human computer interaction has lead researchers to attempt improve the current interfaces.
 Even though the area of research progressed quickly in speech and handwriting systems, gesture and sketch understanding systems are still starting to get noticed.
 
-Sketch recognition is defined as the process of identifying the symbols the user draws using a single or multiple strokes.
+Sketch recognition is defined as the process of identifying the symbols the user draws using single or multiple strokes.
 Users draw strokes using a pen and the system immediately interprets their strokes into objects that can be easily manipulated.
-This research uses Particle Swarm Algorithm (PSO) to divide the strokes the user draws into meaningful geometric primitives.
+This research uses Particle Swarm Optimization algorithm (PSO) to divide the strokes the user draws into meaningful geometric primitives.
 These geometric primitives are grouped to formulate symbols which are further identified.
-Two algorithm are used to divide strokes, the first algorithm \textit{ALgS1} divides stroke to polygons.
-The second one \textit{AlgS2}, divides stroke to segments of either lines or curves.
-The research focuses on the effect of PSO segmentations to the sketch recognition systems.
+Two algorithms are used to divide strokes, the first algorithm \textit{ALgS1} divides strokes to polygons.
+The second one \textit{AlgS2}, divides strokes to segments of either lines or curves.
+The research focuses on the effect of PSO segmentations on the sketch recognition process.
 The final recognition is achieved using a set of geometrical and global shape properties features.
@@
 Experiments were conducted on three different datasets; Hs-DB a benchmark dataset which contains simple presentation symbols, a LD-DB a set of logic design symbols and EL-DB Electrical symbols.
-Results show that using \textit{ALgS2} to segment stroke improves segmentation than known algorithms.
-\textit{ALgS1} achieved good performance on EL-DB and Hs-DB but performed poorly on LD-DB due to high number of curves in every symbol.
-%The segmentation improvement is tested by final symbol recognition accuracy.
+Results show that using \textit{ALgS2} to segment strokes improves segmentation as compared to other algorithms.
+To be specific \textit{AlgS2} achieves good performance with datasets like EL-DB and Hs-DB however is not as good in datasets characterized by a higher number of curves on every symbolic such as LD-DB.
+  %\textit{ALgS1} achieved good performance on EL-DB and Hs-DB but performed poorly on LD-DB due to high number of curves in every symbol
+. %The segmentation improvement is tested by final symbol recognition accuracy.
 
 \end{abstractpage}                              % This ends the page.
+%\begin{foreword}                                        %======================
+    											                             % Delete these if
+%\end{foreword}                                          % no foreword page.
+
+                                                        % OTHER PAGES
+                                                        %======================
+\tableofcontents                                        % Always needed...
+%\listoftables                                           % Delete if no tables.
+\listoffigures                                          % Delete if no figures.
 
@@
 \include{Results}
+
 \include{ConFut}

### 2009-10-10 trunk/draft/Results.tex
after changes of dr. nevin

@@
 \begin{description}
-	\item [Hse Dataset] (denoted as Hs-DB) Data set collected by \cite{HeloiseBeautification} and was previously used by \cite{Oltmans07} as benchmark for sketch recognition problem.
+	\item [Hse Dataset] (denoted as Hs-DS) Data set collected by \cite{HeloiseBeautification} and was previously used by \cite{Oltmans07} as benchmark for sketch recognition problem.
 The data are drawn by 16 users each of them have drawn from 50 to 60 samples for each shape.
@@
 %is comparable to the 96.7%  %The results displayed are the average recognition accuracy of the four splits resulted from the classifier after recognizing the symbols.
-		\item [Multi Domain Dataset] (denoted as MD-DB) A dataset we collected manually from different users.
+\item [Electrical Dataset] (denoted as EL-DS) A dataset we collected manually from different users.
+The dataset contains  Electrical Engineering symbols (EL-DS).
+Data are collected from 7 different users each of them have drawn from 20 to 40 samples of each shapes.
+Figure \ref{fig:ElectSet} shows some of the shapes used in the dataset.
+Symbols were gathered to evaluate the system throughout different types of shapes.
+Users were requested to draw symbols without any restriction on style, size or orientations.
+The dataset was divided into 75\% training set and 25\% test set.
+Five different splits for the training and test data are generated from the dataset.
+%In worth noting that \cite{Oltmans07} gathered similar dataset and achieved 89\% on it   %The results displayed are the average recognition accuracy of the four splits resulted from the classifier after recognizing the symbols
+\item [Digital Logic Dataset](denoted as LD-DS) Similar to EL-Ds the dataset was collected manually from different users.
+The dataset contains symbols Logic Design symbols (LD-DS).
+Data are collected from 7 different users each of them have drawn from 20 to 40 samples of each shapes.
+Figures \ref{fig:LogicSet} shows some of the shapes used in the dataset.
+The dataset was divided to train and test set similar to EL-DS.
+%Symbols from different domains were gathered to evaluate the system throughout different types of shapes.
+Users were requested to draw symbols without any restriction on style, size or orientations.
+The dataset was divided into 75\% training set and 25\% test set.
+Five different splits for the training and test data are generated from the dataset.
+%In worth noting that \cite{Oltmans07} gathered similar dataset and achieved 89\% on it   %The results displayed are the average recognition accuracy of the four splits resulted from the classifier after recognizing the symbols
+	%	\item [Multi Domain Dataset] (denoted as MD-DB) A dataset we collected manually from different users.
 The dataset contains symbols from two different domains Electrical Engineering symbols (EL-DB) and logic design symbols (LD-DB).
@@
 The result of the segmentation algorithm can be viewed in the Figure \ref{fig:results1}.
-This result shows the originals stroke and the segmentation that the system generated for the stroke.
+This result shows the original stroke and the segmentation that the system generated for the stroke.
 Figure \ref{fig:results2} shows the result of \textsl{AlgS1} and \textsl{AlgS2} for the same input strokes.
 Figure \ref{fig:results3} shows the output of \textsl{Alg3} to the same set of input strokes.
-It clearly shows that \textsl{AlgS2} have better segmentation results than both \textsl{AlgS1} and \textsl{Alg3} but since these segmentation are subjective to what the user intended to draw the next section will focus on the evaluating the segmentation algorithms using the effect on the recognition accuracy.
+It clearly shows that \textsl{AlgS2} has better segmentation results than both \textsl{AlgS1} and \textsl{Alg3} but since these segmentations are subjective to what the user intended to draw the next section will focus on evaluating the segmentation algorithms and the effect on the recognition accuracy.
 
@@
 
-several experiments is presented to test the different segmentation algorithms on the three data set used in our system.
+Several experiments are presented to test the different segmentation algorithms on the three data sets used in our system.
 Algorithms (\textsl{AlgS1}, \textsl{AlgS2} and \textsl{Alg3} (in section \ref{sec:BenchMarckAlgorithm})) were tested individually with and without the ellipse fitting module.
@@
 The results shows that \textsl{AlgS2} with the ellipse fitting module gives better recognition rate across all datasets (Figure \ref{fig:test1},\ref{fig:testEL} and \ref{fig:testLD} ).
-Even though, \textsl{AlgS1} gives better result than \textsl{Alg3} in both EL-DB and HS-DB it perform poorly in LD-DB due to the high number of curves in every symbol in this dataset.
-These result are proved using different number of samples per category as shown in Figures \ref{fig:LDRes}, \ref{fig:ElRes}.
+Even though, \textsl{AlgS1} gives better result than \textsl{Alg3} in both EL-DB and HS-DB it performs poorly in LD-DB due to the high number of curves in every symbol in this dataset.
+These result are obtained using different number of samples per category as shown in Figures \ref{fig:LDRes}, \ref{fig:ElRes}.
 % i need to organize this as following. 
@@
 
- The effects of symbol complexity in a dataset were testing in the following experiment.
+ The effect of symbol complexity in a dataset is tested in the following experiment.
 These experiments test how many samples will be required to finally obtain good performance from the recognizer system.
 In the following experiments, we tested the system using different number of sample per category.
- Figure \ref{fig:Alg1Ds} and \ref{fig:Alg2Ds} shows that Hs-DB requires more number of sample for training to achieve same result as in other data sets.
-This is due that it contains less signal and conceptual variation that the other two datasets.
-Figure \ref{fig:samplesofHS-DB.jpg} shows sample of data collected from one user in Hs-Db it is clear that it has reasonable amount of variation between users but it is limited.
+ Figure \ref{fig:Alg1Ds} and \ref{fig:Alg2Ds} shows that Hs-DB requires more number of samples for training to achieve same result as in other data sets.
+This is due to the fact that it contains less signal and conceptual variation that the other two datasets.
+Figure \ref{fig:samplesofHS-DB.jpg} shows sample of data collected from one user in Hs-Db.
+It is clear that it has reasonable amount of variation between users but it is limited.
 Recognizing symbols that contains high variation in drawing style using fewer symbols in training prove the efficiency of our system.
@@
 	\includegraphics[scale=0.5]{results/samplesofHS-DB.jpg.eps}
-	\caption[Hs-DB shapes drawn by single user]{Hs-DB shapes drawn by single user \cite{HeloiseBeautification} }
+	\caption[Hs-DB Shapes Drawn by Single User]{Hs-DB shapes drawn by single user \cite{HeloiseBeautification} }
 	\label{fig:samplesofHS-DB.jpg}
 \end{figure}
+
   Figure \ref{fig:correct2.jpg} shows examples of correctly recognized symbol using our system.
 The figure shows that system recognized some hard to recognize example in both EL-DB and DL-DB.
-An example of symbols that was not correctly recognized are in Figure \ref{fig:erros}.
-Most of the wrong classifications were due to how messy the user draw symbols and how similarly they are conceptual.
+Example of symbols that were not correctly recognized are in Figure \ref{fig:erros}.
+Most of the wrong classifications were due to how messy the user draws symbols and how similar they are conceptual.
 Table \ref{tab:ConfusionMatrix} (confusion matrix) shows the types of error in Hs-DB which shows that the system was confused between parallelogram and trapezoid, moon and arch.
@@
 It is clearly noted that symbols that have only line segments achieve higher accuracy rate than other symbols.
-The results indicate that algorithm \textsl{AlgS1} achieve better performance than algorithm \textsl{AlgS2} in the symbols that consist only of lines.
+The results indicate that algorithm \textsl{AlgS1} achieves better performance than algorithm \textsl{AlgS2} for the symbols that consist only of lines.
 This is understandable because algorithm \textsl{AlgS1} divides strokes into line segments only but \textsl{AlgS2} is able to divide strokes into lines and curves based on the minimum error of the segment itself.

### 2009-10-10 draft/ConFut.tex
after changes of dr. nevin

@@
 This research presented a new approach to sketch recognition using PSO.
-It was noted that the PSO in general improve the accuracy of the final symbol recognition in the system.
+It was noted that the PSO in general improves the accuracy of the final symbol recognition in the system.
 The use of different preliminary information (speed, time difference and curvature)  helps to improve of the PSO algorithm over the original algorithms \cite{CruveDivisionSwarm,PolygonApproximationPSO}.
@@
 The system introduced an efficient method to sketch recognition using the PSO.
-The features computed was a hybrid of a stroke based features which gives more information on the geometrical and structure of the shape tested.
+The features computed were a hybrid set of a stroke based features which gives more information on the geometrical and structure of the shape tested.
 
@@
 The number and the complexity of symbols recognized by the system varied from simple to complex but did not affect the final recognition of the system.
-Test performed confirmed that PSO achieve better performance and optimization than other algorithms.
+Tests performed confirmed that PSO achieve better performance and optimization than other algorithms.
 In general \textsl{ALS2} proved to generate the best result on different datasets.
-Algorithm \textsl{ALS2} achieved better segmentation result that any other algorithm in shapes that are combination of curves and lines as in digital design dataset.
-Algorithm \textsl{ALS1} achieve high performance in electrical symbols but gives poor result in logic design set.
-The result is justified because logic design dataset contains symbols that are differentiated by number of curves in OR-gate and XOr gate, therefore representing the symbol as lines will remove important information.
+Algorithm \textsl{ALS2} achieved better segmentation results than any other algorithm in shapes that are combination of curves and lines as in digital design dataset.
+Algorithm \textsl{ALS1} achieves high performance in electrical symbols but gives poor results in logic design set.
+The result is justifiable because logic design dataset contains symbols that are differentiated by number of curves in OR-gate and XOr gate, therefore representing the symbol as lines will remove important information.
 
-The system uses different set of feature that represents shape geometrical and global properties.
+The system uses different sets of features that represent shape geometrical and global properties.
 Literature review showed that system that relies only on global shape properties fails to identify the shapes that have small geometrical difference (ie.
@@
 
- On the other hand, the system still has drawback as it lack in handling over traced strokes.
+%The system still has drawback as it lack in handling over traced strokes.
 Another drawback is that until now we only applied the system to single symbols.
@@
 %\section{Future Work}
-The next step in this research is to complete the clustering algorithm for fully automated sketch recognition.
+The system still has some drawbacks, it needs to be improved to handle over traced strokes.
+Another improvement required is handling of multia le symbols since we only applied the system to single symbols.
+In addition, full free hand sketch test has not been applied on the system.
+%sketched
+
+Hence, the next step in this research is to complete the clustering algorithm for fully automated sketch recognition.
 This will enable testing the system for more than single symbol at a time.
-The segmentation algorithm may be more powerful if it can segments the stroke into more types of primitives other than the line and curve.
+The segmentation algorithm may be more powerful if it can segment the stroke into more types of primitives other than the line and curve catagorization used in this work.
 Other area of modification can be the features extraction and classifier.
-Adding more spatial and structure features will improve classifications.
+Adding few more spatial and structure features should improve classifications.
 After segmenting the stroke the system can generate a structural graph from segments and there spatial relations.
 This graph will be matched with the set of known symbol template to make the identification.
-Those results along with the SVM classifier results will enhance the system as a whole and achieve better recognition rate.
+Those results along with the SVM classifier results are expected to enhance the system as a whole and achieve even better recognition rates.
 %Each new stroke is checked if it can be a symbol or is a part of already drawn un completed symbol.
+
+%drawback of the system is that it takes more time to segment strokes in comparison with other similar algorithms.
+
 \appendix

### 2009-10-13 draft/System.tex
FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL

@@
 Each DPSO algorithms tries to find the best particle in the solution space.
-Both algorithms has the same problem formulation but different fitness functions.
+Both algorithms have the same problem formulation but different fitness functions.
 After generating two segmentations the system chooses the best segmentation from the outputs of the DPSO.
@@
 The system represents each particle $P_i$ by $P_i = \left\{ {p_{ij} \left| {j = 1,2 \cdots N} \right.} \right\}$ where $p_{ij}$ is either 0 or 1 where $p_{ij}=1$ means that point $j$ is a dominant point(see Figure \ref{fig:CodingSwarm}.
-Thus , the particle represents points that are chosen to be used as dominant points for this segmentation.
+Thus, the particle represents points that are chosen to be used as dominant points for this segmentation.
 The goal of the DPSO algorithm is to find the solution $P_i$ that generates the minimum set of dominant points that defines the stroke with minimum segmentation error.
@@
 Each segment is also fitted into a circular arc $(x - a)^2  + (y - b)^2  = R^2, $ where ($a,b$) are the coordinate of the center of the circle and $R$ is the radius of the circle.
-They is estimated using the following set of equations:
+They are estimated using the following set of equations:
 
@@
 This evaluation is based on the area of both the segmentation and the input stroke.
-The segmentation error($E_{alg}$ )represents the sum of Square Root Error of the length from the point on the stroke to the estimated curve or line.
+The segmentation error ($E_{alg}$ )represents the sum of Square Root Error of the length from the point on the stroke to the estimated curve or line.
 In \textsl{AlgS2}, this error is the same as the computed error in Equation \ref{eq:ErrorSwarm1} but For the second algorithm \textsl{AlgS2} it is different than the computed error in Equation \ref{eq:errorSwarm2}.
-The additional computation is necessary to  compare the two segmentation fairly.
-The system choose the minimum segmentation as the final segmentation of this stroke.
+The additional computation is necessary to compare the two segmentations fairly.
+The system chooses the minimum segmentation as the final segmentation of this stroke.
 
@@
 These values and parameter will help determine the spatial relations between different segments in the feature extraction step.
-% The segmentation which crossponds to the minimum $E_{tot}$ is the best segmentation.  
+% The segmentation which crosseponds to the minimum $E_{tot}$ is the best segmentation.  
 %\begin{equation}
@@
 After the segmentation, each stroke is represented as a list of segments $L_s$.
-A segment can be either an ellipse, a line or a circular arc.
+A segment can be an ellipse, a line or a circular arc.
 The list of segments $L_s$ represents the segmentation of the stroke.
@@
 \subsection{Structural and geometrical Features(FS1)}
- Features defines the structure of the geometrical symbol. Section \ref{sec:steprec} describe in details how these 13 feature are computed.  
+ Features define the structure of the geometrical symbol. Section \ref{sec:steprec} describe in details how these 13 feature are computed.  
 %\begin{description}
@@
 \subsection{Global shape properties set (FS4)}
- Features that are composites of different geometrical and statistical symbol. These 13 features represent the global shape of the symbol drawn.
+ Features composed of different geometrical and statistical properties. These 13 features represent the global shape of the symbol drawn.
 
@@
 We used the library implementation of LibSVM \cite{libsvm} and its implementation of the one-vs-one strategy.
-Section \ref{Sec:SVMdetail} describe in details the classifier, how classes (classification) are determined and the reasons behind using this specific classification methods.
+Section \ref{Sec:SVMdetail} describe in details the classifier, how classes (classification) are determined and the reasons behind using this specific classification method.
 
@@
 		\includegraphics{afterDefense/svm1.jpg.eps}
-	\caption{Different Classifier Planes }
+	\caption{Different Classifier Planes}
 	\label{fig:svm1}
@@
  The orientation of the separation plane is determined using the vector $\omega$ and the offset of the plane from the origin is determined using the scalar $b$.
-Let's make an assumption that the data point are linearly separable, i.e.
+Let's make an assumption that the data points are linearly separable, i.e.
 there are infinite possibilities to find a plane that separates and correctly classify the training data.
@@
 Figure \ref{fig:svm5} illustrates that we cannot construct a plane that can separate the two sets correctly.
-The constrains for constructing the planes must be relaxed since the QP task is not feasible for the linearly inseparable case.
+Constrains for constructing the planes must be relaxed since the QP task is not feasible for the linearly inseparable case.
 Consider the linear inseparable problem in Figure\ref{fig:svm5}, in the ideal case there will be no point misclassified and no point falls on the margin.
-But we must relax the constrain that insure that each point in on the appropriate side of it is supporting plane.
+But we must relax constrains that insure that each point in on the appropriate side of it is supporting plane.
 Every point that falls on the wrong side of the supporting plane is considered an error.
@@
 In two dimensional vector space with attributes $r$ and $s$, to construct a quadratic  discriminate  simply map the original two dimensional input space $\left[r,s\right]$ to  five dimensional features space $\left[r,s,rs,r^2,s^2\right]$ and construct a linear discriminate in that space.
-specifically, define : $\vartheta(x):R^2  \rightarrow R^5$ then
+Specifically, define : $\vartheta(x):R^2  \rightarrow R^5$ then
 \begin{equation}
@@
  There is mainly two popular methods to change the binary SVM classification problem into Multi class classification problem.
-The first methods is One Versus All (OVA) Strategy where M binary classifiers is constructed one classifier for each class.
+The first method is One Versus All (OVA) Strategy where M binary classifiers is constructed one classifier for each class.
 The $ith$ classifier is trained using positive example form $ith$ class and negative example from all other classes.
@@
 They also compared the same classification methods on Arabic handwritten digits (MADBASE) \cite{ADBase9,IjdarSherifPaper}.
-Their results shows that SVM classifier outperform all other techniques on both MNIST and MADBase datasets and achieved recognition rate of over 98\%.
+Their result shows that SVM classifier outperforms all other techniques on both MNIST and MADBase datasets and achieved recognition rate of over 98\%.
 Similar results was reported by \cite{empiricalcomp11,SVMInvariantComDecoste02} on the MNIST dataset.
-There are many other successful application SVM.
-They have proven to be robust to noise and perform well on many tasks \cite{empiricalcomp11,Scholkopf97comparingsupport}.
-
-This superior performance because they may eliminates many of problem experienced with other inference methodologies with other inference methodologies like neural network and decision trees.
+There are many other successful applications.
+SVM have proven to be robust to noise and perform well on many tasks \cite{empiricalcomp11,Scholkopf97comparingsupport}.
+
+This superior performance because they may eliminates many of problems experienced with other inference methodologies with other inference methodologies like neural network and decision trees.
 The first problem that is eliminated was the problems with local minima.
@@
 The last reason is the final result are sable, reproducible and largely independent of the specific algorithm used to optimize the SVM model.
-If two users apply the same SVM model with the same parameters to the same data, they will et the same solution module numeric issues.
+If two users apply the same SVM model with the same parameters to the same data, they will get the same solution module numeric issues.
 Compare this with neural networks where the results are dependent on the particular algorithms and starting point used.
@@
 To solve SVM approach requires solving a QP or LP problem.
-Mathematical programming have extensively studied the LP and QP type problems.
+Mathematical programming has extensively studied the LP and QP type problems.
 These techniques can be divided into three categories, methods in which evaluate and discard kernel components during learning.
@@
 
-The third approach is to directly attack the SVM problem from and optimization perspective and create algorithms that explicitly exploits the structure of ten problem.
+The third approach is to directly attack the SVM problem from and optimization perspective and create an algorithm that explicitly exploits the structure of ten problems.
 Keerthi et al.
@@
 Table \ref{tab:StepsStroke1} shows the detailed output of the program at each step.
-The input of this step is the a set of strokes, each stroke has a set of  points and the output is a list of segments that will be used in the recognition process.
+The input of this step is a set of strokes, each stroke has a set of points and the output is a list of segments that will be used in the recognition process.
 More detailed example and algorithms are demonstrated in the Appendix \ref{ChapterstepExample} \\
@@
 
-	\includegraphics[scale=0.5]{afterDefense/clockstroke1.eps}
-  &   &    
+%	\includegraphics[scale=0.5]{afterDefense/clockstroke1.eps}
+  &   &    \includegraphics[scale=0.5]{afterDefense/stroke1withpdp.jpg.eps}
 \\ \hline 
@@
  	\includegraphics[scale=0.5]{afterDefense/clockstroke2.jpg.eps} \\ \hline
-  & Possible Dominate point Extraction for details see Section\ref{sec:Preprocessing} &   	\includegraphics[scale=0.5]{afterDefense/clockstroke2.jpg.eps}  & 1. compute the $P_{pd}$  & 
+  & Possible Dominate point Extraction for details see Section\ref{sec:Preprocessing} &   	%\includegraphics[scale=0.5]{afterDefense/clockstroke2.jpg.eps}
+    & 1. compute the $P_{pd}$  &  
+  \includegraphics[scale=0.5]{afterDefense/stroke2WithPdp.jpg.eps}
     \\ \hline

### 2009-10-13 draft/chapterSurvey.tex
FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL

@@
 To demonstrate our system inspiration we will first present current research in the area of sketch recognition.
-In Section \ref{sec:preprocessing}, preprocessing and segmentation methods will be discussed in details.
+In Section \ref{sec:preprocessingSeg}, preprocessing and segmentation methods will be discussed in details.
 Later, symbol recognition methods will be presented and analyzed in Section \ref{sec:symbolrecogntion}.
@@
 \section{Preprocessing and Segmentation}
-\label{sec:preprocessing}
+\label{sec:preprocessingSeg}
   In a system that supports free hand sketching, sketching should be as natural as possible.
@@
 		\includegraphics[scale=0.7]{../../neededfiles/Figures/overstroked.eps}
-	\caption[Example of an over traced stroke]{Example of an over traced stroke \cite{overtraced24}}
+	\caption[Example of an Over-traced Stroke]{Example of an over traced stroke \cite{overtraced24}}
 	\label{fig:overstroked}
@@
  Agar et al.
-\cite{polygonfeedback31} used the timestamp for each point and compute time difference between every two successive point to help him find corners as he argued that it provides more distinct maximum than the speed information.
+\cite{polygonfeedback31} used the time-stamp for each point and compute time difference between every two successive point to help him find corners as he argued that it provides more distinct maximum than the speed information.
 He mentions that the pointing device (for example the mouse) sampling rate is the reason for this phenomena.
@@
 This leads to a nearly constant time difference between samples while the pen is moving and large difference while the pen is stationary.
-Contrary to speed information, where the user draws with variant speeds which tends to make the speed information include a lot of noise in the data.
+Contrary to speed information, where the user draws with variant speed, which tends to make the speed information include a lot of noise in the data.
 
@@
 Their system gives better results than using only speed or direction information.
-This is because the speed information tend to be noisy when drawing large symbols, on the other side when using only direction information the some corner are missed.
+This is because the speed information tends to be noisy when drawing large symbols, on the other side when using only direction information the some corner are missed.
 Our system uses speed, direction and time difference information in its preliminary calculation step.
@@
 
-Arvo \cite{Fluid25} Tried to combine fast morphing and basic recognition for the sketch drawn as the user draw it.
+Arvo \cite{Fluid25} tried to combine fast morphing and basic recognition for the sketch drawn as the user draws it.
 It tries to move the points of strokes drawn by the user draw into one of the known shapes (ex: box, lines and circle).
@@
 
- Systems that employ such representation must first pre-process strokes to break them into geometric primitives ( section \ref{sec:preprocessing} reviews on preprocessing methods).
+ Systems that employ such representation must first pre-process strokes to break them into geometric primitives ( section \ref{sec:preprocessingSeg} reviews on preprocessing methods).
 Recognition can then be treated as sub graph matching problem or constrain satisfaction problem.
@@
  Fonseca et al.\cite{Cali63} created a system using similar set of features, their features were used by a number of fuzzy logic rules to identify symbols.
-They also used Naïve Bayes model that was trained to classify shapes based on those features.
+They also used Naï?½ve Bayes model that was trained to classify shapes based on those features.
 
@@
 In all of these system only \cite{SketchRead2007} using stroke based recognition have been implemented to the fully create a system that free draw sketch and understand them.
-This is because the emphasis on the interactive side of the sketch system which result in large difficulties in processing free hand sketches.
+This is because the emphasis on the interactive side of the sketch system which results in large difficulties in processing free hand sketches.
 Methods based on gesture and global stroke properties produce systems that are dependent on assumptions about how strokes can be grouped.
@@
 To avoid most problems in such system, our system uses hybrid of both global shape and hierarchical description methods.
-Currently, our system recognize independent symbols.
+Currently, our system recognizes independent symbols.
 Each symbol is segmented and hierarchical features are computed.
@@
   Evolutionary algorithms (e.g.
-Genatic Algoirthms) explore solution using a population which is evaluated using fitness functions.
+Genetic Algorithms) explore solution using a population which is evaluated using fitness functions.
 The algorithms initially start with a randomly chosen population which is used to generate new population using various operators (e.g.
@@
 A fitness function is used to evaluate the goodness of each particle in solution space.
-But unlike evolutionally algorithm the new population is the PSO is influenced by the social cooperation of the particle rather than survival to the fittest.
+But unlike evolutionarily algorithm the new population is the PSO is influenced by the social cooperation of the particle rather than survival to the fittest.
 Each particle saves a history of the best location in the swarm and its own best location.
 At the end each iteration, the particles moves to a new location based on the global best location and it own history of best location.
-This use of pervious particle locations gives PSO advantage over Evolutionary algorithms where no similar information is used.
+This use of previous particle locations gives PSO advantage over Evolutionary algorithms where no similar information is used.
 
@@
 In general the PSO is a continuous optimization problem but experiments showed it performed as well as GA which is in general a discrete and binary optimization problem.
-In general, both GA and PSO share many points, e.g, both algorithms generate the initial populuation randomaly, both use a fitness function or value to evaluate the goodness of solutions and the next populations.
-Nevertheless, PSO dose not have genetic operators like crossover or mutations.
-Partices are updated by internal comunication among themselfs based on the local and global best.
+In general, both GA and PSO share many points, e.g, both algorithms generate the initial population randomly; both use a fitness function or value to evaluate the goodness of solutions and the next populations.
+Nevertheless, PSO does not have genetic operators like crossover or mutations.
+Particles are updated by internal communication among themselves based on the local and global best.
 In contrast to GA where chromosomes shares information with each others, PSO shares information using memory from previous iteration in form of global best solution.
@@
 Both algorithms start with a group of a randomly generated population, both have fitness values to evaluate the population.
-Both update the population and search for the optimium with random techniques.
+Both update the population and search for the optimum with random techniques.
 Both systems do not guarantee success.
@@
 
-%\subsection{The Algorithm}
-%%\label{sec:ParticleSwarmAlgorithm} 
-%\begin{verbatim}
-% Initialize 
-%    Generate N particles with M dimensions (P_i...P_N) 
-%    Generate velocities for each particle for M dimension randomly. 
-% Repeat till Maximum number of iteration OR Fitness 
-%         For each particle P_i do the following 
-%                Compute fitness  
-%                Determine Lbest  
-%                Check if need to change Gbest. 
-%                Compute the Velocity for each dimension using eq. 
-%                Check if maximum Velocity reach and limit it 
-%                update particle values 
-%          end . 
-%        Improve all particle positions 
-%  end 
-%  
-%  Final solution reaches in Gbest.       
-%        
-%        
-%\end{verbatim}
+\subsection{The DPSO Algorithm}
+\label{sec:ParticleSwarmAlgorithm}
+
+This section will describe each step of the PSO Algorithm.
+Figure\ref{fig:BlockPSO} illustrates the DPSO algorithm in Algorithm \ref{DPSOalg} steps in details.
+Initially, the swarm particles are created and each of which is a binary vector corresponding to candidate solution.
+Additionally, the velocities of each particle are initialized at random by drawing uniformly real values from the range [0.0 to 1.0].
+Then the evolutionary process of particle movement is iterated until maximum number of iteration or a fitness error achieved.
+At the beginning of each iteration, the fitness value of each particle is evaluated, and the particles individual local best and global best is updated.
+The particle position is adjusted as follow; the velocities is computed using Equation \ref{eq:Swarm}.
+Equation \ref{eq:vmax} is used to scale the velocity to values between 0 and 1 which will serve as a probability that the corresponding bit takes the value of 1.
+The particles bit vectors are updated by Equation \ref{eq:descrite} according to their velocity which measure the trajectory to the target solution based on local experience ($lbest_{i}$) and global intelligence ($gbest_{i}$).
+Before the end of the iteration the particle may move to a refined location based on the refine procedure used.
+\begin{algorithm}
+\caption{ Discrete Particle Swarm Algorithm }
+\label{DPSOalg}
+\begin{verbatim}
+Initialize 
+  1. Generate N particles with M dimensions (P_i...P_N) 
+  2. Generate velocities for each particle for M dimension randomly. 
+  3. Repeat till Maximum number of iteration OR ErrorThreshold  
+         For each particle P_i do the following 
+         3.a   Compute fitness  (fitness)
+         3.b   Determine Lbest  
+               if (lbest_fitness<fitness)
+                  3.b.1   lbest=P_i
+         3.c      Determine Gbest.
+               if (lGest_fitness<fitness)
+                  3.c.1   lGest=P_i
+         3.d   Compute the Velocity for each dimension using eq.3.2 
+         3.e   Bound Velocity using eq.3.4
+         3.f   Update particle values using eq. 3.3
+         3.g   Refine the location of the particle
+         end .
+     end 
+4. return Gbest
+\end{verbatim}
+  \end{algorithm}
+
+\begin{figure}
+	\centering
+		\includegraphics[scale=0.7]{afterDefense/BlockPSO.eps}
+	\caption{Flow Chart of PSO Algorithm }
+	\label{fig:BlockPSO}
+\end{figure}

### 2009-10-13 trunk/draft/ConFut.tex
FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL

@@
 This research presented a new approach to sketch recognition using PSO.
-It was noted that the PSO in general improve the accuracy of the final symbol recognition in the system.
+It was noted that the PSO in general improves the accuracy of the final symbol recognition in the system.
 The use of different preliminary information (speed, time difference and curvature) helps to improve of the PSO algorithm over the original algorithms \cite{CruveDivisionSwarm,PolygonApproximationPSO}.
@@
 %\section{Future Work}
-The next step in this research is to complete the clustering algorithm for fully automated sketch recognition.
-This will enable testing the sytsem for more than single symbol at a time.
-The segmentation algorithm may be more powerful if it can segments the stroke into more types of primitives other than the line and curve.
+The system still has some drawbacks; it needs to be improved to handle over traced strokes.
+Another improvement required is handling of multiple symbols since we only applied the system to single symbols.
+In addition, full free hand sketch test has not been applied on the system.
+%sketched
+
+Hence, the next step in this research is to complete the clustering algorithm for fully automated sketch recognition.
+This will enable testing the system for more than single symbol at a time.
+The segmentation algorithm may be more powerful if it can segment the stroke into more types of primitives other than the line and curve categorization used in this work.
 Other area of modification can be the features extraction and classifier.
-Adding more spatial and structure features will improve classifications.
+Adding few more spatial and structure features should improve classifications.
 After segmenting the stroke the system can generate a structural graph from segments and there spatial relations.
 This graph will be matched with the set of known symbol template to make the identification.
-Those results along with the SVM classifier results will enhance the system as a whole and achieve better recognition rate.
+Those results along with the SVM classifier results are expected to enhance the system as a whole and achieve even better recognition rates.
 %Each new stroke is checked if it can be a symbol or is a part of already drawn un completed symbol.
+
+%drawback of the system is that it takes more time to segment strokes in comparison with other similar algorithms.
+
 \appendix
@@
 \end{algorithm}
-%\item Segmentation ALS1 as explained in Section \ref{sec:SwarmSegmentation}. The alogrithm will be similar to the PSO algorithm  in section \ref{sec:ParticleSwarmAlgorithm}.
+%\item Segmentation ALS1 as explained in Section \ref{sec:SwarmSegmentation}. The algorithm will be similar to the PSO algorithm in section \ref{sec:ParticleSwarmAlgorithm}.
 
@@
 
-%Step & Discription of Step & Input & Steps & Output \\ \hline
+%Step & Description of Step & Input & Steps & Output \\ \hline
 Stroke 1 & Extracting point and computing boundary data & -- &  see Section \ref{sec:Preprocessing} & 
@@
 	}
-	\caption{Segmentation Ouputs For Second Stroke}
+	\caption{Segmentation Outputs For Second Stroke}
 \end{figure}
@@
 		\includegraphics[scale=0.85]{afterDefense/FinalSegmentation.jpg.eps}
-	\caption{Best Segmantation }
+	\caption{Best Segmentation }
 	\label{fig:segmentatiofinal.jpg}
@@
 Feature extraction step is divided into four steps where each feature set is computed separately then all sets are appended to construct a final feature vector.
-
 	\begin{enumerate}
-	\item Computing Structural and geometrical Features (FS1).
-This step computes number of parallel, perpendicular line and similar structural features.
-Algorithm \ref{FeatureAlg} is used to determine these features.
-	\item Rubine Feature Set (FS2).  Compute Rubine features as in \cite{gestureexample12} 
-	\item Statistical Features (FS3) compute the moments as in \cite{zernike61}
-	\item Global shape properties set (FS4). Firstly compute the convex hull of the stroke points then compute the rest of the features.
-
-	\item Introduce the computed feature vector to the classifier and check the label of the detected symbol. 
+	\item Computing Structural and geometrical Features(FS1).
+This step computes number of parallel, perpendicular line and similar structural features [13 features].
+Algorithm \ref{FeatureAlg} in Appendix \ref{ChapterstepExample}  is used to determine these features.
+ 	\item Rubine Feature Set (FS2) [12 features] .  Compute Rubine features as in \cite{gestureexample12} 
+	\item Statistical Features (FS3) compute the moments as in \cite{zernike61} for N=10 there are 32 features.
+	\item Global shape properties set (FS4). Firstly compute the convex hull of the stroke points then compute the rest of the features [13 features]. 
+	\item Introduce the computed feature vector (All 70 features) to the classifier and check the label of the detected symbol. 
 \end{enumerate}
@@
  The final Feature vector of the FS1==>
-ature Number of primitives 3.0   ,    Feature Number of segments 9.0   ,    Feature Line count 6.0   ,    Feature Curves Count 3.0   ,    Feature Ellipse Count 1.0   ,    Feature Intersection T type -1.0   ,    Feature Intersection L type 4.0   ,    Feature Intersection X type -1.0   ,    Feature Parallel Count 1.0   ,    Feature PerPendicular Count -1.0   ,    Feature Intersection Lines Count 4.0   ,    Feature Min Radius 1.7976931348623157E308   ,    Feature Max Radius 231.70851748053468
+  Number of primitives 3.0   ,    Feature Number of segments 9.0   ,    Feature Line count 6.0   ,    Feature Curves Count 3.0   ,    Feature Ellipse Count 1.0   ,    Feature Intersection T type -1.0   ,    Feature Intersection L type 4.0   ,    Feature Intersection X type -1.0   ,    Feature Parallel Count 1.0   ,    Feature PerPendicular Count -1.0   ,    Feature Intersection Lines Count 4.0   ,    Feature Min Radius 1.7976931348623157E308   ,    Feature Max Radius 231.70851748053468
 

### 2009-10-13 trunk/draft/Results.tex
FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL

@@
 Algorithms (\textsl{AlgS1}, \textsl{AlgS2} and \textsl{Alg3} (in section \ref{sec:BenchMarckAlgorithm})) were tested individually with and without the ellipse fitting module.
-The final system (Choosing the Minimum error of (\textsl{AlgS1}, \textsl{AlgS2}) was evaluated on the three dataset(HS, EL and LD datasets ).
+The final system (Choosing the Minimum error of (\textsl{AlgS1}, \textsl{AlgS2}) was evaluated on the three dataset (HS, EL and LD datasets).
 Figure \ref{fig:test1},\ref{fig:testEL} and \ref{fig:testLD} shows the accuracy achieved by each algorithm.
-Theses experiments were done using the whole dataset as 75\% of the data as test set and 25\% as training.
+These experiments were done using the whole dataset as 75\% of the data as test set and 25\% as training.
 The results shows that \textsl{AlgS2} with the ellipse fitting module gives better recognition rate across all datasets (Figure \ref{fig:test1},\ref{fig:testEL} and \ref{fig:testLD} ).
@@
 %\end{figure}  
-%Another expirements where conducted to evaluates the worth of an attribute.
-The worth of each attribute is measured using two different attribure measures.
+%Another experiments where conducted to evaluates the worth of an attribute.
+The worth of each attribute is measured using two different attribute measures.
 The first is by measuring the gain ratio with respect to the class, where the gain ration is computed as $Gain_{ration}(class,attribute)=\frac{H(class)-H(class|Attribure}{H(Attribute)}$.

### 2009-10-13 trunk/draft/SketchMaster.tex
FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL

--- 2009-04-27 trunk/draft/SketchMaster.tex
@@
 \usepackage{longtable}
+\usepackage{lscape}
 \usepackage{array}
+\usepackage{algorithmic}
+\usepackage{algorithm} 
+%\numberwithin{algorithm}{chapter}
 %\usepackage[T1]{fontenc}
@@
                                                         %======================
-\title{Sketch Recognition and Understanding}             % My theses title  using Particle Swarm Optimization Algorithm
-\author{Maha Mohammed Nabeel El Meseery}                               % Y name
+\title{SKETCH RECOGNITION AND UNDERSTANDING}            % My theses title  using Particle Swarm Optimization
+% Algorithm   Sketch Recognition and Understanding
+\author{Maha Mohamed Nabeel El Meseery Moustafa El Meseery}                               % Y name
 \prevdegrees{B.S.,Computer Engineering}       		% Your old degree
-\major{Computer Engineering}                            % My current major
-\degree{Master of Science}                              % My new degree
-\degreemonth{}                                      % When awarded. to be added when known
-\degreeyear{}                                      %
-\superad{Professor Nevin Darwish\\Professor Samia Mashali\\Assoc. Professor Magda Fayek\\}
+\major{COMPUTER ENGINEERING}                            % My current major
+\degree{MASTER OF SCIENCE}                              % My new degree
+\degreemonth{ }                                      % When awarded. to be added when known
+\degreeyear{2009}      
+\FacultyUniversity{Faculty of Engineering, Cairo University \\ Giza, Egypt \\  2009}
+\superad{}                                %
+%\superad{Professor Nevin Darwish\\Professor Samia Mashali\\Assoc. Professor Magda Fayek\\}
 %th today or first day of writing  December 26 2005 
 %\thesisdate[]                                     % Date of document.
-\DOCUMENTtype{Thesis}                                   % or DISSERTATION
-\Documenttype{Thesis}                                   % or Dissertation
-\documenttype{Thesis}                                   % or dissertation
+\DOCUMENTtype{ Thesis }                                   % or DISSERTATION
+\Documenttype{ Thesis }                                   % or Dissertation
+\documenttype{ Thesis }                                   % or dissertation
 
@@
 %\makecopyright                                          % if no copyright
+  %=======================================Supervisors page ===============================
+\newpage    
+\thispagestyle{empty}               % Don't do headers or footers.                                                  
+ \begin{center}     
+ SKETCH RECOGNITION AND UNDERSTANDING\\
+\normalsize by      \\ 
+  Maha Mohamed Nabeel El Meseery Moustafa El Meseery\\
+\vspace*{0.4in}        
+\normalsize
+A Thesis Submitted To the\\
+Faculty of Engineering, Cairo University\\
+in Partial Fulfillment of the\\
+Requirements for the Degree of\\
+MASTER OF SCIENCE
+in\\
+COMPUTER ENGINEERING\\
+\vspace*{0.4in} 
+Under the Supervision of 
+\end{center}
+  \vspace*{0.4in} 
+\begin{center}
+ 		\begin{tabular}{ccc}
+		Dr. Nevin Darwish  & Dr. Samia Mashali & Dr. Magda Fayek \\
+		 Professor   & Professor & Associate Professor  \\
+		 Faculty of Engineering& Electronic Research Institute & 		 Faculty of Engineering \\
+		  Cairo University && Cairo University\\
+		\end{tabular}
+\end{center}
+  \vspace*{0.4in}       
+  \vfill
+\begin{center}
+   Faculty of Engineering, Cairo University \\ Giza, Egypt \\ 2009
+\end{center}
+
+  %=======================================Supervisors page ===============================
                                                      % APPROVAL                                                         %======================
-%\thispagestyle{empty}               % Don't do headers or footers.
-%\newpage                                                             
-%\begin{center}     
-%\LARGE  
-%\textbf{Sketch Recognition using Particle Swarm Optimization Algorithm}\\
-%\normalsize by      \\ 
-%\Large  \textbf{ Maha mohamed Nabeel El Meseery}\\
-%\vspace*{0.4in}        
-%\normalsize
-%A Thesis Submitted To the\\
-%Faculty of Engineering, Cairo University\\
-%in Partial Fulfillment of the\\
-%Requirements for the Degree of\\
-%MASTER OF SCIENCE
-%in\\
-%COMPUTER ENGINEERING\\
-%
-%\end{center}
-%          
-%    Approved by\\                                           
-%Examining Committee:\\
-% \makebox[6in]{\hrulefill}\\ 
+\newpage       
+\thispagestyle{empty}               % Don't do headers or footers.
+
+\begin{center}     
+SKETCH RECOGNITION AND UNDERSTANDING\\
+\normalsize by      \\ 
+Maha Mohamed Nabeel El Meseery Moustafa El Meseery\\
+\vspace*{0.2cm}        
+A Thesis Submitted To the\\
+Faculty of Engineering, Cairo University\\
+in Partial Fulfillment of the\\
+Requirements for the Degree of\\
+MASTER OF SCIENCE
+in\\
+COMPUTER ENGINEERING\\
+\end{center}         
+Approved by\\                                           
+Examining Committee:\\ \vspace*{0.2cm} 
+\begin{tabular*}{0.9\textwidth}{l   }
+ %\makebox[5in]{\hrulefill}\\ 
+\\ \hline 
+Professor Nevin Darwish, Thesis Main Advisor\\ \\  \hline  
+ %\makebox[5in]{\hrulefill}\\ 
+Professor Samia Mashali,Thesis Advisor \\  \\ \hline   
+ %\makebox[6in]{\hrulefill}\\  
+Assoc. Professor Magda Fayek,Thesis Advisor\\  \\  \hline  
+ %\makebox[5in]{\hrulefill}\\ 
+Professor  Ashraf Abd El Wahab, Member\\ \\ \hline  
+ %\makebox[6in]{\hrulefill}\\ 
+Professor Mohamed Zaki Abd ElMegeed, Member\\  
+ %\makebox[6in]{\hrulefill}\\ 
+ %\end{spacing}
+
+		\end{tabular*}
+
+%\begin{spacing}{2}
+% \makebox[5in]{\hrulefill}\\ 
 %Professor Nevin Darwish, Thesis Main Advisor\\
-% \makebox[6in]{\hrulefill}\\ 
+% \makebox[5in]{\hrulefill}\\ 
 %Professor Samia Mashali,Thesis Advisor \\
@@
 %Assoc. Professor Magda Fayek,Thesis Advisor\\ 
-% \makebox[6in]{\hrulefill}\\ 
+% \makebox[5in]{\hrulefill}\\ 
 %Professor  Ashraf Abd El Wahab, Member\\
@@
 % \makebox[6in]{\hrulefill}\\ 
+ %\end{spacing}
+ \begin{center}
+   Faculty of Engineering, Cairo University \\ Giza, Egypt \\  2009
+\end{center}
   % \\   you need...\\
@@
  I thank my family and friends for their moral and support.
-I specifically thank, dina said, noha mansour and marwah shafee.
+I specifically thank, Dina Said, Noha Mansour and Marwah Shafee.
 Thanks, for being always there when I needed you the most.

### 2009-10-14 draft/SketchMaster.tex

@@
 %\documentstyle[11pt,unhthesis]{report}
-\documentclass[11pt,doublespace]{SketchThesis}
+\documentclass[12pt,onehalfspacing]{SketchThesis}
 
@@
 \degree{MASTER OF SCIENCE}                              % My new degree
-\degreemonth{July}                                      % When awarded. to be added when known
+\degreemonth{ }                                      % When awarded. to be added when known
 \degreeyear{2009}      
-\FacultyUniversity{Faculty of Engineering, Cairo University \\ Giza, Egypt \\ July 2009}
+\FacultyUniversity{Faculty of Engineering, Cairo University \\ Giza, Egypt \\  2009}
 \superad{}                                %
@@
 \begin{center}
-   Faculty of Engineering, Cairo University \\ Giza, Egypt \\ July 2009
+   Faculty of Engineering, Cairo University \\ Giza, Egypt \\ 2009
 \end{center}
@@
  \begin{center}
-   Faculty of Engineering, Cairo University \\ Giza, Egypt \\ July 2009
+   Faculty of Engineering, Cairo University \\ Giza, Egypt \\  2009
 \end{center}

### 2010-06-29 Swarm/swarmModified.tex
Initial import

\documentclass[10pt]{article}
\usepackage{spconf}
\usepackage{amsmath}
\usepackage[dvips,pdftex]{graphicx}
\usepackage{named}
\usepackage[dvips,pdftex]{hyperref}
\usepackage{subfigure}
\usepackage{array}

\title{Enhanced Particle Swarm Optimization}
\twoauthors
  {Tarek abouldahab\sthanks{Corresponding Author}} {Cairo Metro Company \\
  Ministry of transport \\Cairo, Egypt\\ heshoaboda@hotmail.com}
 {Maha El Meseery, Mahmoud Fakhr El Din}
	{Signals Processing Group \\ Computers and Systems Department \\
  Electronic Research Institute\\Cairo, Egypt\\
	melmeseery@eri.sci.eg, mafakhr@mcit.gov.eg}

\begin{document}
\maketitle
\begin{abstract}
%This study 
Sketches and drawings are widely used as a simple method of expressing thought and designs. The goal of this paper is developing a sketch recognition system that facilitates using sketches in computer systems. The system uses Particle Swarm Optimization (PSO) algorithm to optimally segment the strokes the user draws into meaningful geometric primitives.  These geometric primitives are grouped to formulate symbols which are further identified using a Support Vector Machines (SVM) classifier. Two different segmentation algorithms were evaluated and results show that both algorithms improve the final recognition results. Experiments were conducted on three different dataset, simple presentation (Hs-DS), Electrical Design (EL-DS) and Logic Design (LD-DS) symbols. Results show that the effectiveness of the proposed system on various dataset especially on datasets with low shape diversity as in Logic Design dataset (LD-DS).  % sketch are divided Users draw symbols and sketches using a set of pen strokes. 
%Sketch recognition is defined as the process of identifying symbols that users draw using single or multiple %strokes. Usually, users draw strokes using pens. The sketch recognition system immediately interprets their %strokes as objects that can be easily manipulated. This paper uses Particle Swarm Optimization (PSO) to divide %the strokes the user draws into meaningful geometric primitives. These geometric primitives are grouped to %formulate symbols which are further identified. The results show that using PSO improves segmentation results %which guide the symbol recognition phase. As for recognition we use Support Vector Machines (SVM) classifier %which further improves the final recognition accuracy.  
\end{abstract}
\pagenumbering{arabic}
\section{Introduction}

Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.
Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.
Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.
Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.
Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.

\section{Basic Particle Swarm Optimization}
\label{sec:ParticleSwarmAlgorithm}
 The main idea of \textit{Particle Swarm Algorithm (PSO)} is to represent each solution with a $N$ dimension particle from the solution space \cite{PSOFirst}. Each particle moves with a direction and velocity $v_{ij}$ based on equations \ref{eq:Swarm1} \& \ref{eq:Swarm}.

\begin{equation}
%\[
p_{ij}=p_{ij}+v_{ij},
%\
\label{eq:Swarm1}
\end{equation}

where $p_{ij}$ represent the $j$th dimension in the $i$th particle and $v_{ij}$ is the velocity of the $j$th dimension in the $i$th particle.
 %Equation [\ref{eq:Swarm}] shows how velocity and direction of each particle are computed
 \begin{equation}
v_{ij}  = v_{ij} \omega + c_1 r_1 (lbest_{ij}  - p_{ij} ) + c_2 r_2 (gbest_{ij}  - p_{ij} ),
\label{eq:Swarm}
\end{equation}
 where $\omega$ is the inertia weight parameter which controls the tradeoff between exploration and exploitation,  $lbest_{ij}$ is the local best particle, $gbest_{ij}$ is the global best particle, $r_1$ \& $r_2$ are random variables and $c_1$ \& $c_2$ are the swarm acceleration parameters.

 After each iteration the global best $g_{best}$ particle and the agent local best particle $l_{best}$are evaluated based on the maximum fitness functions of all particles in the solution space. The solution is found after achieving a specific number of iteration or after an error threshold is achieved.

\section{Enhanced Partilce Swarm}
This paper modifies the general PSO algorithm by changeing the local term into a random term using the following equation .  
  \begin{equation}
  v_{ij}  = v_{ij} \omega + c_1 r_1 (p_{iRandom_i}  - p_{iRandom_j} ) + c_2 r_2 (gbest_{ij}  - p_{ij} ),
  \label{eq:ModSwarm}
  \end{equation}

where $iRandom_i$ and $Random_j$ are two particles in the solution space. This modification replace the distance between the current particle and the local best with two random particle from the space. The effect of this change in more expolration and avoiding of local min problems. Equation \ref{eq:ModSwarm} is computed and the new particle locaiton is calculated. The new location is compared the location of particle if with equation \ref{eq:Swarm} if the modified location is better (mean better fitness function) then the new location is selected otherwise the old equation is used. 
\section{Applications }
The modified swarm algorithm is used to improve the performance of the sketch recoginition system in \cite{myPaper}. The

\cite{myPaper}  
\section{Experiments} 
 The goal of the system is to modify the segmentation and recognition of the sketch system. It made the same type of experiement done in \cite{myPaper}. The dataset is divided into 13 category of goemetric symbols 
\section{Conclusion and Future Work}
Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.
Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.
Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.
Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.
Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.

\bibliographystyle{IEEEbib}
\bibliography{../../../neededfiles/Bibliographies/Mybibliography}

\end{document}

### 2010-08-07 Swarm/swarmModified.tex

@@
 As for recognition we use Support Vector Machines (SVM) classifier %which further improves the final recognition accuracy.
+%%% old abstract.. 
+ Also another enhanced Particle Swarm Optimization (EPSO) based on the perturbation of the particle velocity is introduced to train the weights of this proposed neural network architecture.
+This proposed architecture with its relevant training algorithm is mentioned as (SDRNN-EPSO) and its performance is compare with the proposed network architecture trained be Basic-PSO algorithm.
+Simulation results will be carried out on stock market indices as an example of time series prediction which shows that our proposed training algorithm is more efficient and accurate than the Basic-PSO in stock market prediction.
 \end{abstract}
@@
 
-Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.
-Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.
-Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.
-Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.
-Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.
+However, in PSO algorithm as the number of iterations increases, the particles speed decrease leading to converge to the global best point found so far, which is not guaranteed even to be a local minimum..
+To enhance the performance and explore better solutions, many researchers introduced GAs operations like crossover, mutation, and selection operators or tuning the PSO parameters leading to increase the computational effort  \cite{Ouyang2009IJCAI}.
+
+As the need for higher effective neural network architecture that increases the neural network output accuracy is highly required.
+So, in this study we propose a new architecture based on introducing the hidden layer sigmoid weight which adapts the shape of the hidden layer output sigmoid function.
+Also, another Enhanced Particle Swarm Optimization training algorithm is applied in training network weights.
+
+The field of symbol recognition has gained interest in the last few years.
+Scientists generally and engineers specifically express thoughts and designs using sketches.
+Engineers use sketches to exchange designs as a natural method of communication rather than writing or speaking.
+Design engineers need a powerful computer based system with symbol recognition to help them design, manipulate and store sketches more effectively than using only papers.
+Increasing the interaction between computers and users in sketch and CAD systems has been the reason for the emerging of a few advanced sketch recognition systems.
+Sketch recognition is the process of identifying user strokes into meaningful symbols that can be further used.
+%Sketch recognition is divided into three main steps preprocessing, segmentation and symbol recognition.
+The preprocessing phase captures user input stroke points and collects basic information about the stroke then proceeds to remove noise and compute basic statistical and geometrical information.
+In segmentation phase, the strokes are divided into a set of simple geometrical primitives or segments.
+In the third phase, sketch recognition, strokes and segments are clustered to formulate symbols that can be recognized by a classifier system.
+
+The segmentation problem is summarized as finding the best decomposition with least number of segments each represents a geometric primitive.
+This is an optimization problem, which evolutionary programming can solve efficiently.
+A genetic algorithm was used by \cite{CruveDivisionSwarm} to optimally divide digital curves into lines and curves.
+Chen et al.\cite{CruveDivisionSwarm} uses digital curves scanned from paper as input to the system and did not take advantage of the curvature or local geometric properties of the digital curve.
+
+This paper introduces a new sketch recognition system, which uses particle swarm algorithm (PSO) to optimally segment user's strokes.
+Users can draw symbols using any number of strokes in any order they like.
+The system segments each stroke using two different PSO algorithms.
+Segmentation is based on dividing the digital curve into polygon\cite{PolygonApproximationPSO}.
+To handle complex strokes, our system uses a second segmentation PSO algorithm to segment strokes into a set of lines and circular arcs.
+After segmentation, a symbol features vector composed of different statistical, geometrical and spatial features is deduced.
+The final classification step, SVM classifier uses the computed feature vector to classify the segments into one of the previously trained classes.
+
+%The remaining of the paper is as following; Section \ref{Sysdisc} describes the system in details.
+Sections \ref{Prepross}, \ref{seg} and \ref{sec:Recognition} provide details on preprocessing, segmentation and recognition process respectively.
+Section \ref{sec:Experiments} describes the experiments that were preformed.
+We conclude in Section \ref{ConclusionandFutureWork}.
 
@@
 \section{Enhanced Partilce Swarm}
+
+In this above standard PSO, the particle cognition term is associated with the own thinking without consideration of other swarm particles thinking which leads to a lake of diversity and could result is suboptimum solutions.
+So, to add social thinking of the particle and keep diversity while searching for the global optimum, we add perturbation to modify this cognitive term by the weighted difference of the position vectors of any other two distinct particles randomly chosen from the swarm.
+The particle is shifted to this new location only if it yields a fitness value better than obtained using the standard PSO.
+This modification is done as follows:-
+\cite{stockPaper}.
+First: - For each particle P in the swarm two other distinct particles, say L and M where PLM, are selected randomly.
+The difference between their positional coordinates is taken as a difference vector:
+ \begin{equation}
+ \lable{eq:mod}
+ \end{equation}
+And the associated velocity vector is updated according to this difference vector as follows:-
+ In essence the cognitive part of the velocity update formula in (1) is replaced with the vector differential operator P?to produce some additional exploration capability.
+Second: - a new trial position ) is created by adding the updated velocity )1(?tVP to the previous position)( 
+ \begin{equation}
+ \lable{eq:mod1}
+ \end{equation}
+Third: - a new trial objective function value )(tP?
+is calculated based on this new trial position vector and the particle is shifted to this trial position if the trial objective function is better than the objective function )(tFPassociated with standard PSO.
+Thus the target particle is relocated as follows:-
+  \begin{equation}
+ \lable{eq:mod2}
+ \end{equation}
+Otherwise add perturbation to modify this cognitive term by the weighted difference of the position vectors of any other two distinct particles randomly chosen from the swarm.
+The particle is shifted to this new location only if it yields a fitness value better than obtained using the standard PSO.
+
 This paper modifies the general PSO algorithm by changeing the local term into a random term using the following equation .  
@@
 The modified swarm algorithm is used to improve the performance of the sketch recoginition system in \cite{myPaper}. The 
+\subsection{The proposed system}
+\label{Sysdisc}
+ The suggested sketch recognitions system is divided into three main steps 1) preprocessing, 2) segmentation and 3) Recognition. The following section provides a detailed description of each step.
+\subsubsection{Preprocessing}
+\label{Prepross}
+ Speed and time difference information were widely used in sketch understanding systems \cite{earlyprocess}.
+It was noted that time difference provides more distinct maxima than the speed information.
+Agar \cite{polygonfeedback31} mentions that the pointing device (i.e.
+the mouse) sampling rate is the reason for this phenomenon.
+ In the presented system, we computed time difference, direction, speed and curvature of each point along a stroke.
+The speed is calculated as $v=\Delta s/\Delta t$ where $t$ is the time difference between two points and $s$ is the length between them.
+The direction $d$ is calculated as the angle between two vectors and curvature is considered as the change in direction with respect to length i.e.
+$c= \Delta d/\Delta s$.
+Dominant points are characterized by lower speed values and higher curvature and direction values.
+
+After the system computes all the speed, time difference and curvature information it proceeds to detect the points with low velocity and high curvatures.
+Using simple differentiation to detect local extreme points resulted in false points due to the non smooth curves.
+Hence, the system adopted a process presented by \cite{earlyprocess}, where the mean of the curve is calculated.
+Then a threshold \textit{th} is used to separate the curve into regions $Region_i$; each region $Region_i$ is defined as a range of points, where the curve values are either above or below the threshold \textit{th}.
+Those regions are further processed to find the maximum point $Max(Region_i)$ of each region $Region_i$.
+The stroke points $p_i(x,y)$ that correspond to those maximum values are labeled as \textit{possible dominant points} $P_{pd}$.
+The system repeats this process for curvature, time difference and speed information.
+All the points labeled as possible dominant points $P_{pd}$ are saved into a single array.
+Figure \ref{fig:LabelsPPD} shows the particles labeled as Possible dominant point $P_{pd}$ by the preprocessing, it is noticed the redundancy of some $P_{pd}$ points.
+% (as shown are redundant)
+\subsubsection{Recognition}
+\label{sec:Recognition}
+After the user draws all strokes of the symbol, the set of un-recognized strokes is grouped together along with their segmentation as input to the feature extraction process.
+A composite set of features is used to generate a single feature vector.
+The features used consist of Rubine feature set,  Zernike moments of order 10 \cite{HeloiseBeautification}, ink density as well as some structural and spatial information like number of perpendiculars lines ,number of parallel lines and number of different types of primitives in each symbol.
+After computing the features the symbol is introduced to the (SVM) classifier.
 
@@
 \section{Experiments}
+
  The goal of the system is to modify the segmentation and recognition of the sketch system.
@@
 The dataset is divided into 13 category of goemetric symbols
+
+ The dataset collected by Hse and Newton\cite{HeloiseBeautification} is used to test the presented system.
+Figure \ref{fig:symbolSet} shows a set of the shapes used in the data set.
+The datasets were divided into training and test sets.
+Five different splits for the training and test data are generated from each dataset.
+The results displayed are the average recognition accuracy of the five splits.
+The accuracy is computed as the number of correctly recognized samples divided by the total number of samples in the test.
+
+ We performed two experiments to test the system.
+Firstly, we tested recognition accuracy of shapes in the data set with both algorithms (\textsl{AlgS1} and \textsl{AlgS2}).
+We also implemented the segmentation algorithm described in \cite{earlyprocess} (\textsl{Alg3}) to compare it with our swarm algorithms.
+Figure \ref{fig:test1} shows the accuracy achieved by each algorithm.
+The result shows that (\textsl{AlgS2}) gives better performance than Alg3 and AlgS1.
+
 \section{Conclusion and Future Work}
-Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.
-Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.
-Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.
-Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.
-Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.
+
+The training and testing performance measurements using both PSO and Enhanced-PSO in training and testing phases are shown in table (1)
+
+Table 1:- S&P measurement performance using both PSO and Enhanced-PSO
+We have studied the shape of the sigmoid weight of the hidden layer neurons as a parameter which was usually determined using trial and error.
+Thus, we add additional parameter called sigmoid weight to the hidden layer neurons to approximate the required non linear dynamical mapping more precisely.
+We prove that our proposed neural network architecture is a dynamical one because it could be represented by a dynamical nonlinear mapping function.
+Also, we introduce a new enhanced particle Swarm Optimization method for adjusting
+weights of this new proposed neural network architecture based on random perturbation of the particles instead of the standard cognitive term.
+We prove that our proposed (SDRNN-EPSO) architecture could efficiently give better prediction accuracy and enhanced performance compared to using the SDRNN trained by the Basic-PSO
+For the future work, it is suggested that this new enhanced architecture will be used in other different applications of time series prediction and also extended to be used in other applications rather than time series prediction such as pattern recognition and speech verification
+
+This paper presented a new approach to sketch recognition using DPSO.
+Results show that the \textit{DPSO} in general generates an optimized stroke segmentation which improves the final recognition rate.
+The final recognition is based on geometrical, statistical and structural features.
+The experiments evaluated two \textit{DPSO} different segmentation algorithms (AlgS1 and AlgS2) on simple presentation set dataset.
+Results show that AlgS2 achieves better segmentation and final accuracy than Alg3 \cite{earlyprocess} on the dataset used.
+The results proved that the system is easily expandable to more and more symbol as it does not depend on low level primitive recognizers but on high level symbol and geometrical features.
+The system achieved an average overall improvement of more than 2\% over Alg3.
+
+This paper presented a new approach to sketch recognition using PSO.
+The system uses both speed and curvature information which help improving the \textit{DPSO}.
+It is noted that the \textit{DPSO} in general generates an optimized stroke segmentation which improves the final recognition rate.
+The tradeoff between accuracy achieved and time complexity must be further investigated to achieve better results.
+The use of statistical moments and structural features improves the recognition rate.
+The system was tested on 13 different symbols and achieved an accuracy of 91.5\%.
+The system does not depend on low level recognizer but rather on a set of high level features.
+This makes the system easily expandable to more and more symbols.
+
+ A possible extension of this research is to complete the clustering algorithm for fully automated sketch recognition.
+Currently, the system process one symbol at a time, an enchantment would be to allow users to draw more than one symbol at a time and incorporate a method for separating symbols.
+Other area of enhancements is the features extraction methods.
+Introducing more spatial and geometrical features is believed to improve classifications.
 

### 2010-08-13 Swarm/tmp/swarmModified.tex
adding

--- 2010-08-07 Swarm/swarmModified.tex
@@
 \twoauthors
-  {Tarek abouldahab\sthanks{Corresponding Author}} {Cairo Metro Company \\
-  Ministry of transport \\Cairo, Egypt\\ heshoaboda@hotmail.com}
- {Maha El Meseery, Mahmoud Fakhr El Din}
+{Maha El Meseery\sthanks{Corresponding Author}, Mahmoud Fakhr El Din}
 	{Signals Processing Group \\ Computers and Systems Department \\
@@
 	melmeseery@eri.sci.eg, mafakhr@mcit.gov.eg}
+  {Tarek abouldahab} {Cairo Metro Company \\
+  Ministry of transport \\Cairo, Egypt\\ heshoaboda@hotmail.com}
 
@@
 %This study
+
 Sketches and drawings are widely used as a simple method of expressing thought and designs.
-The goal of this paper is developing a sketch recognition system that facilitates using sketches in computer systems.
-The system uses Particle Swarm Optimization (PSO) algorithm to optimally segment the strokes the user draws into meaningful geometric primitives.
+The goal of this paper is to improve our sketch recognition system by using an enhanced particle swarm optimization algorithm.
+The enhanced Particle Swarm Optimization (EPSO) is based on the perturbation of the particle velocity is used to optimally segment the strokes the user draws into meaningful geometric primitives.
 These geometric primitives are grouped to formulate symbols which are further identified using a Support Vector Machines (SVM) classifier.
-Two different segmentation algorithms were evaluated and results show that both algorithms improve the final recognition results.
-Experiments were conducted on three different dataset, simple presentation (Hs-DS), Electrical Design (EL-DS) and Logic Design (LD-DS) symbols.
-Results show that the effectiveness of the proposed system on various dataset especially on datasets with low shape diversity as in Logic Design dataset (LD-DS).
+Experiments were conducted on the benchmark dataset of simple presentation (Hs-DS) symbols.
+Results show that that the enhanced particle swarm optimization improves the final recognition accuracy.
  % sketch are divided Users draw symbols and sketches using a set of pen strokes. 
@@
 %%% old abstract.. 
- Also another enhanced Particle Swarm Optimization (EPSO) based on the perturbation of the particle velocity is introduced to train the weights of this proposed neural network architecture.
+ %Also another enhanced Particle Swarm Optimization (EPSO) based on the perturbation of the particle velocity is introduced to train the weights of this proposed neural network architecture.
 This proposed architecture with its relevant training algorithm is mentioned as (SDRNN-EPSO) and its performance is compare with the proposed network architecture trained be Basic-PSO algorithm.
@@
 
-In this above standard PSO, the particle cognition term is associated with the own thinking without consideration of other swarm particles thinking which leads to a lake of diversity and could result is suboptimum solutions.
+In this above standard PSO, the particle cognition term is associated with the own thinking without consideration of other swarm particles thinking which leads to a lake of diversity and could result is sub optimum solutions.
 So, to add social thinking of the particle and keep diversity while searching for the global optimum, we add perturbation to modify this cognitive term by the weighted difference of the position vectors of any other two distinct particles randomly chosen from the swarm.
@@
  \begin{equation}
- \lable{eq:mod}
+ \label{eqmod}
+ x=y
  \end{equation}
@@
  \begin{equation}
- \lable{eq:mod1}
+ x=x
+ \label{eq:mod1}
  \end{equation}
@@
   \begin{equation}
- \lable{eq:mod2}
+ \label{eq:mod2}
+ x=z
  \end{equation}
@@
 The accuracy is computed as the number of correctly recognized samples divided by the total number of samples in the test.
-
+ \begin{figure}
+  \centering
+  \includegraphics[width=4in]{Experiment68.jpg}
+  \caption[Close up of \textit{Hemidactylus} sp.]%
+  {Close up of \textit{Hemidactylus} sp., which is
+   part the genus of the gecko family. It is the
+   second most speciose genus in the family.}
+   \label{exp1}
+\end{figure}
+   \begin{figure}
+  \centering
+  \includegraphics[scale=0.5]{Experiment92.jpg}
+  \caption[Close up of \textit{Hemidactylus} sp.]%
+  {Close up of \textit{Hemidactylus} sp., which is
+   part the genus of the gecko family. It is the
+   second most speciose genus in the family.}
+   \label{exp2}
+\end{figure}
  We performed two experiments to test the system.
@@
 
-Table 1:- S&P measurement performance using both PSO and Enhanced-PSO
+Table 1:- S P measurement performance using both PSO and Enhanced-PSO
 We have studied the shape of the sigmoid weight of the hidden layer neurons as a parameter which was usually determined using trial and error.

### 2010-08-15 trunk/Swarm/tmp/swarmModified.tex
adding modification in  pso and enhanced sections

--- 2010-08-13 Swarm/tmp/swarmModified.tex
@@
 \label{sec:ParticleSwarmAlgorithm}
- The main idea of \textit{Particle Swarm Algorithm (PSO)} is to represent each solution with a $N$ dimension particle from the solution space \cite{PSOFirst}.
+
+ Particle Swarm Algorithm\cite{PSOFirst}is a population based stochastic search where each solution is simulated as a  particle that flies around in a multidimensional search space.
+A fitness function is used to evaluate the goodness of each particle in solution space.
+But unlike evolutionarily algorithm the new population is the PSO is influenced by the social cooperation of the particle rather than survival to the fittest.
+Each particle saves a history of the best location in the swarm and its own best location.
+At the end each iteration, the particles moves to a new location based on the global best location and it own history of best location.
+This use of previous particle locations gives PSO advantage over Evolutionary algorithms where no similar information is used.
+
+ The \textit{Basic Particle Swarm Algorithm (PSO)} represents each solution with a $N$ dimension particle from the solution space \cite{PSOFirst}.
 Each particle moves with a direction and velocity $v_{ij}$ based on equations \ref{eq:Swarm1} \& \ref{eq:Swarm}.
@@
 
-\section{Enhanced Partilce Swarm}
+\section{Enhanced Particle Swarm}
 
-In this above standard PSO, the particle cognition term is associated with the own thinking without consideration of other swarm particles thinking which leads to a lake of diversity and could result is sub optimum solutions.
+In the above basic PSO, the velocity update is based on three terms an interia term, local cognition term ( $l_{best}$ term) and global term.
+The particle cognitive term is associated with the particle own history without consideration of other swarm particles locations and histories which leads to a lake of diversity and could result in sub optimum solutions.
 So, to add social thinking of the particle and keep diversity while searching for the global optimum, we add perturbation to modify this cognitive term by the weighted difference of the position vectors of any other two distinct particles randomly chosen from the swarm.
@@
 
- The goal of the system is to modify the segmentation and recognition of the sketch system.
+In our experiment we used the same benchmark dataset used in \cite{myPaper}, the dataset was collected by Hse and Newton\cite{HeloiseBeautification}.
+The data are drawn by 16 users each of them was asked to draw 13 shapes from 30 to 50 times.
+Figure \ref{fig:symbolSet} shows a set of the shapes used in the data set.
+Five different splits were generated to divide the dataset into training and test sets.
+The results displayed are the average recognition accuracy of the five splits.
+The accuracy is computed as the number of correctly recognized samples divided by the total number of samples in the test.
+
+ \begin{figure}
+  \centering
+
+		\includegraphics[scale=0.5]{symbolSet.PNG} 
+		\caption[The Symbol Set] {The symbol set}   \label{fig:symbolSet}
+	 \end{figure}
+	  %The first experiment with the whole feature set and the second experiments with only half of the features.
+To demonstrate the change in accuracy is due to the enhanced swarm optimization algorithm we tested the recognition system using two sets of features.
+Firstly, we tested recognition accuracy of shapes in the data set with both algorithms (\textsl{AlgS1} and \textsl{AlgS2}) using the whole feature set.
+The second experiment was tested on both algorithm but only using half the feature set.
+
+Figures \ref{exp2} and \ref{exp1} show the recognition rate of each algorithm using the full feature set and half the features set respectivily.
+The results show that the enhanced swarm optimization algorithm improve the segmentation of the system which leds to an increase in the recognition rate of the whole system by an average of $1.5\%$.
+The increase in recognition rate is consistant in all swarm algorithms and in both experiments.
+%n the We performed two experiments to test the system.
+Firstly, we tested recognition accuracy of shapes in the data set using the orginal swarm segmentation algorithm and then after using the enhanched swarm optimization.
+To  both algorithms (\textsl{AlgS1} and \textsl{AlgS2}).
+We also implemented the segmentation algorithm described in \cite{earlyprocess} (\textsl{Alg3}) to compare it with our swarm algorithms.
+Figure \ref{fig:test1} shows the accuracy achieved by each algorithm.
+The result shows that (\textsl{AlgS2}) gives better performance than Alg3 and AlgS1.
+
+%In our experiment we used three different datasets.
+First dataset was collected by Hse and Newton\cite{HeloiseBeautification} (Hs-DB).
+The data are drawn by 16 users each of them was asked to draw 13 shapes from 30 to 50 times.
+Figure \ref{fig:symbolSet} shows a set of the shapes used in the data set.
+We collected the second and third datasets.
+Figure \ref{fig:ELsymbolSet} and \ref{fig:LogicsymbolSet} shows the symbols gathered from 7 different users in electrical (EL-DB) and logic design (LD-DB) domain respectively.
+Table \ref{tab:datasets} shows a  comparison between the three datasets used in our experiments.
+The datasets were divided into training set and test set.
+Five different splits  for the training and test data are generated from each dataset.
+The results displayed are the average recognition accuracy of the five splits.
+The accuracy is computed as the number of correctly recognized samples divided by the total number of samples in the test.
+
+ %The goal of the system is to modify the segmentation and recognition of the sketch system.
 It made the same type of experiement done in \cite{myPaper}.
-The dataset is divided into 13 category of goemetric symbols
+The dataset is divided into 13 category of goemetric %symbols
 
- The dataset collected by Hse and Newton\cite{HeloiseBeautification} is used to test the presented system.
+% The dataset collected by Hse and Newton\cite{HeloiseBeautification} is used to test the presented system.
 Figure \ref{fig:symbolSet} shows a set of the shapes used in the data set.
@@
   \centering
-  \includegraphics[width=4in]{Experiment68.jpg}
-  \caption[Close up of \textit{Hemidactylus} sp.]%
-  {Close up of \textit{Hemidactylus} sp., which is
-   part the genus of the gecko family. It is the
-   second most speciose genus in the family.}
-   \label{exp1}
+  \includegraphics[scale=0.5]{Experiment92.jpg}
+  \caption[Recognition accuracy]%
+  {The recognition rate of system using the feature set.}
+   \label{exp2}
 \end{figure}
@@
   \centering
-  \includegraphics[scale=0.5]{Experiment92.jpg}
-  \caption[Close up of \textit{Hemidactylus} sp.]%
-  {Close up of \textit{Hemidactylus} sp., which is
-   part the genus of the gecko family. It is the
-   second most speciose genus in the family.}
-   \label{exp2}
+  \includegraphics[width=4in]{Experiment68.jpg}
+  \caption[Recognition accuracy]%
+  {The recognition rate of system using half the feature set.}
+   \label{exp1}
 \end{figure}
- We performed two experiments to test the system.
+
+% We performed two experiments to test the system.
 Firstly, we tested recognition accuracy of shapes in the data set with both algorithms (\textsl{AlgS1} and \textsl{AlgS2}).

### 2010-08-16 trunk/Swarm/tmp/swarmModified.tex
adding the paper after i finished it before correcctions

@@
 
-\title{Enhanced Particle Swarm Optimization}
+\title{Enhanced Particle Swarm Optimization In Sketch Recognition}
 \twoauthors
@@
 
-However, in PSO algorithm as the number of iterations increases, the particles speed decrease leading to converge to the global best point found so far, which is not guaranteed even to be a local minimum..
+%However, in PSO algorithm as the number of iterations increases, the particles speed decrease leading to converge to the global best point found so far, which is not guaranteed even to be a local minimum..
 To enhance the performance and explore better solutions, many researchers introduced GAs operations like crossover, mutation, and selection operators or tuning the PSO parameters leading to increase the computational effort  \cite{Ouyang2009IJCAI}.
 
-As the need for higher effective neural network architecture that increases the neural network output accuracy is highly required.
+%As the need for higher effective neural network architecture that increases the neural network output accuracy is highly required.
 So, in this study we propose a new architecture based on introducing the hidden layer sigmoid weight which adapts the shape of the hidden layer output sigmoid function.
@@
 Sketch recognition is the process of identifying user strokes into meaningful symbols that can be further used.
-%Sketch recognition is divided into three main steps preprocessing, segmentation and symbol recognition.
+
+Sketch recognition is divided into three main steps preprocessing, segmentation and symbol recognition.
 The preprocessing phase captures user input stroke points and collects basic information about the stroke then proceeds to remove noise and compute basic statistical and geometrical information.
@@
 
-The segmentation problem is summarized as finding the best decomposition with least number of segments each represents a geometric primitive.
+%The segmentation problem is summarized as finding the best decomposition with least number of segments each represents a geometric primitive.
 This is an optimization problem, which evolutionary programming can solve efficiently.
@@
 
-This paper introduces a new sketch recognition system, which uses particle swarm algorithm (PSO) to optimally segment user's strokes.
-Users can draw symbols using any number of strokes in any order they like.
-The system segments each stroke using two different PSO algorithms.
+This paper modifies the sketch recognition system presented in \cite{mypaper}, to use Enhanced particle swarm algorithm (EPSO) to optimally segment user's strokes.
+As in the Orginal system the users can draw symbols using any number of strokes in any order they like.
+The system segments each stroke using two different EPSO algorithms.
 Segmentation is based on dividing the digital curve into polygon\cite{PolygonApproximationPSO}.
-To handle complex strokes, our system uses a second segmentation PSO algorithm to segment strokes into a set of lines and circular arcs.
+To handle complex strokes, our system uses a second segmentation EPSO algorithm to segment strokes into a set of lines and circular arcs.
 After segmentation, a symbol features vector composed of different statistical, geometrical and spatial features is deduced.
@@
 
-%The remaining of the paper is as following; Section \ref{Sysdisc} describes the system in details.
-Sections \ref{Prepross}, \ref{seg} and \ref{sec:Recognition} provide details on preprocessing, segmentation and recognition process respectively.
+The remaining of the paper is as following; Section \ref{sec:ParticleSwarmAlgorithm} describes the Basic particle swarm algorithm in details.
+Enhanced Particle Swarm Algorithm is discribed in Section \ref{sec:EnhancedSwarmAlgorithm}.
+Section  \ref{Sysdisc} provide details on preprocessing, segmentation and recognition process.
 Section \ref{sec:Experiments} describes the experiments that were preformed.
@@
 \section{Enhanced Particle Swarm}
+\label{sec:EnhancedSwarmAlgorithm}
+In the above basic PSO, the velocity update is based on three terms an interia term, local cognitive ($l_{best}$) term and global cognitive ($g_{best}$)term.
+The particle local cognitive term is associated with the particle own history without consideration of other swarm particles locations and histories which leads to a lake of diversity and sub optimum solutions.
+Therefore, we are tyring to add more social influence in the particle movement by using the weighted difference of the position vectors of any other two distinct particles randomly chosen from the swarm to modify the local cognitive term.
+The modified term adds social behvaiour of other particle and keep diversity while searching for the global optimum.
+The new social cognitive term of the particle is shifted to the new location only if it yields a fitness value better than obtained using the standard PSO\cite{stockPaper}.
 
-In the above basic PSO, the velocity update is based on three terms an interia term, local cognition term ( $l_{best}$ term) and global term.
-The particle cognitive term is associated with the particle own history without consideration of other swarm particles locations and histories which leads to a lake of diversity and could result in sub optimum solutions.
-So, to add social thinking of the particle and keep diversity while searching for the global optimum, we add perturbation to modify this cognitive term by the weighted difference of the position vectors of any other two distinct particles randomly chosen from the swarm.
-The particle is shifted to this new location only if it yields a fitness value better than obtained using the standard PSO.
 This modification is done as follows:-
-\cite{stockPaper}.
-First: - For each particle P in the swarm two other distinct particles, say L and M where PLM, are selected randomly.
-The difference between their positional coordinates is taken as a difference vector:
+  \begin{enumerate}
+\item For each particle $p_{i}$ in the swarm two other distinct particles, $p_L$ and $p_M$ are selected randomly where $i\neq L\neq M$.
+The difference between their positional coordinates $p_{d}$ is taken as a difference vector in equation \ref{eqmod}.
  \begin{equation}
  \label{eqmod}
- x=y
+ p_{dj}= p_{Lj}-p_{Mj} 
  \end{equation}
-And the associated velocity vector is updated according to this difference vector as follows:-
- In essence the cognitive part of the velocity update formula in (1) is replaced with the vector differential operator P?to produce some additional exploration capability.
-Second: - a new trial position ) is created by adding the updated velocity )1(?tVP to the previous position)( 
+where $p_{dj}$ represent the $j$th dimension in the difference vector and $p_{Lj}$ represent the $j$th dimension in the $L$th particle.
+
+Then the associated velocity vector of particle $p_{i}$ is updated according to this difference vector as in equation \ref{eq:mod1}.
+In essence the local cognitive term of the velocity update formula in \ref{eq:Swarm} is replaced with the vector differential operator $p_d$ to produce some additional exploration capability.
  \begin{equation}
- x=x
+ v^{T}_{ij}  = v_{ij} \omega + c_1 r_1 (p_{dj}) + c_2 r_2 (gbest_{ij}  - p_{ij} ),
  \label{eq:mod1}
  \end{equation}
-Third: - a new trial objective function value )(tP?
-is calculated based on this new trial position vector and the particle is shifted to this trial position if the trial objective function is better than the objective function )(tFPassociated with standard PSO.
-Thus the target particle is relocated as follows:-
+\item The new trial position $p^{T}_{ij}$ is created by adding the updated velocity $v^{T}_{ij}$ in Eq. \ref{eq:mod1}.
   \begin{equation}
  \label{eq:mod2}
- x=z
+p^{T}_{ij}=p_{ij}+ v^{T}_{ij} ,
  \end{equation}
-Otherwise add perturbation to modify this cognitive term by the weighted difference of the position vectors of any other two distinct particles randomly chosen from the swarm.
+\item The new position $p^{T}_{ij}$ is used to generate the new trial objective fitness function value $fitness(p^{T})$.
+The fitness value is calculated based on this new trial position vector and the particle is shifted to this trial position if the trial objective function is better than the original objective function associated with standard PSO.
+Thus the target particle is relocated using the following equation (Eq.
+\ref{eq:objective})
+  \begin{equation}
+ \label{eq:objective}
+p^{new}_{ij}\Leftarrow 
+\{
+\begin{array}{c} 
+ p^{T}_{ij} \quad \quad if\quad fitness(p^{T})> fitness(p^{old})  \\
+ p_{ij} \quad \quad Otherwise\quad  
+\end{array}\}
+ \end{equation}
+ where $fitness(p^{T})$ is the objective fitness value of the new trail position $ p^{T}$ and  $fitness(p^{old})$is the objective fitness value of the new basic position $ p{i}$ computed by equation \ref{eq:Swarm1}.
+   \end{enumerate}
+
+%Otherwise add perturbation to modify this cognitive term by the weighted difference of the position vectors of any other two distinct particles randomly chosen from the swarm.
 The particle is shifted to this new location only if it yields a fitness value better than obtained using the standard PSO.
 
-This paper modifies the general PSO algorithm by changeing the local term into a random term using the following equation .  
-  \begin{equation}
-  v_{ij}  = v_{ij} \omega + c_1 r_1 (p_{iRandom_i}  - p_{iRandom_j} ) + c_2 r_2 (gbest_{ij}  - p_{ij} ),
-  \label{eq:ModSwarm}
-  \end{equation}
+%This paper modifies the general PSO algorithm by changeing the local term into a random term using the following equation .  
+ % \begin{equation}
+ % v_{ij}  = v_{ij} \omega + c_1 r_1 (p_{iRandom_i}  - p_{iRandom_j} ) + c_2 r_2 (gbest_{ij}  - p_{ij} ),
+ % \label{eq:ModSwarm}
+ % \end{equation}
 
-where $iRandom_i$ and $Random_j$ are two particles in the solution space.
+%where $iRandom_i$ and $Random_j$ are two particles in the solution space.
 This modification replace the distance between the current particle and the local best with two random particle from the space.
@@
 The new location is compared the location of particle if with equation \ref{eq:Swarm} if the modified location is better (mean better fitness function) then the new location is selected otherwise the old equation is used.
-\section{Applications }
+\section{System Discribtion  }\label{Sysdisc}
+%\label{sec:application}
 The modified swarm algorithm is used to improve the performance of the sketch recoginition system in \cite{myPaper}. The 
-\subsection{The proposed system}
-\label{Sysdisc}
+%\subsection{The system}
+
  The suggested sketch recognitions system is divided into three main steps 1) preprocessing, 2) segmentation and 3) Recognition. The following section provides a detailed description of each step.
-\subsubsection{Preprocessing}
+\subsection{Preprocessing}
 \label{Prepross}
@@
 % (as shown are redundant)
-\subsubsection{Recognition}
+\subsection{Segmentation}
+\label{seg}
+In the segmentation stage the input stroke is divided into a set of primitives.
+As shown in Figure \ref{fig:segblock} first an attempt is made to fit the stroke points into a curve or an ellipse using a minimum square error fitting algorithm \cite{ellipsefit}.
+The ellipse fitting step helps to prevent the system from over segmenting the stroke into multiple lines or curves if the input stroke is an ellipse.
+If the stroke proved to be an ellipse arc then the segmentation process ends and the system proceeds to the next step.
+Otherwise, the stroke is passed to two further segmentation algorithms that divide the stroke to either lines or lines and curves.
+After two different segmentations are generated, the system chooses the segmentation with the minimum error.
+The following sections describe in details the ellipse detection algorithm and the two segmentation algorithms used to divide the stroke.
+% The input to these algorithms are the stroke points along with the possible dominant points $P_{pd}$ computed during preprocessing.
+The output is a set of dominant points which are connected by either lines or curves.
+
+%///AllBlockDiagram.jpg  you can change the block to the main block. 
+ \begin{figure}
+	\centering
+		\includegraphics[scale=0.48]{blockSmall.jpg}
+	\caption{Segmentation Block}
+	\label{fig:segblock}
+\end{figure}
+\subsection{Ellipse fitting}
+%The process starts by computing the center of the stroke bounding box.
+The bounding box center point is used as the first estimation of the ellipse center.
+The axes of the ellipse are estimated as $width/2$ and $height/2$ of the stroke bounding box.
+The least square fitting algorithm is used to minimize the fitting error of the ellipse equation \cite{chernov}.
+This process tries to fit the stroke points into an ellipse arc; it starts with computing the center of the stroke bounding box.
+The bounding box center point is used as the first estimation of the center of the ellipse.
+The axes of the ellipse are estimated as the $width/2$ and $height/2$ of the stroke bounding box where $width$ is width and $height$ is height of the bonding box.
+The least square fitting algorithm \cite{chernov} is used to minimize the fitting error of the ellipse Equation (\ref{eq:circleFit})
+
+\begin{equation}
+E = \sum\limits_{i = 0}^N {\frac{{(x_i - x_0 )}}{{a^2 }}^2  + \frac{{(y_i - y_0 )}}{{b^2 }}^2  - 1} 
+\label{eq:circleFit}
+\end{equation}
+
+ where $N$ is number of points in the stroke, $a,b$ are the length of ellipse axes, $x_0$ \& $y_0$ are the coordinates of the center point, $x_i$ \& $y_i$ are the coordinates of point $i$ in the stroke.
+A list of new values for $x_0$ , $y_0$ ,$a$ and $b$ are generated randomly from the older values with small increments after each loop.
+After few iteration, the final fit error of the estimated ellipse is reported.
+Another measure is used to compute the efficiency of the final estimated ellipse.
+Equation \ref{eq:circleError} ensures that the drawn percentage of ellipse is considered.
+This eliminates fitting a  line into very large ellipse but leaves small ellipses to be fitted as a partial or full ellipse.
+
+ \begin{equation}
+eff= (P_{percent}/E)
+\label{eq:circleError}
+\end{equation}
+ \begin{equation}
+P_{percent}  = L_{stroke} /P_{ellipse} 
+\label{eq:ErrorArea}
+\end{equation}
+where $E$ is the error computed by Equation(\ref{eq:circleFit}), $L_{stroke}$ total length of stroke and $P_{ellipse} $ is the perimeter of the estimated ellipse.
+If $eff$ is more than threshold $th_{Ellipse}$\footnote{By trial and error best threshold found was $th_{Ellipse}=0.2$} then the stroke is segmented as an ellipse otherwise the system proceeds to the next step.
+
+\subsection{Non ellipse fitting algorithm}
+\label{subsubsec:Discreteparticleswarmalgorithm}
+\begin{figure}
+	\centering
+		%\subfigure[Possible dominate point]{ 	\label{fig:LabelsPPD}	\includegraphics[scale=0.7]{ppd.jpg}}
+	%	\hfill
+	 \includegraphics[scale=0.7]{pso1.jpg}			
+	\caption{DPSO algorithm encoding}% a) Possible dominate point b) Particle encoding  
+	\label{fig:pso1}
+\end{figure}
+Two DPSO algorithms are used to generate different segmentations for each stroke the user draws.
+For each segmentation generated an error is evaluated.
+The segmentation with the minimum error value is chosen as the best stroke segmentation (Figure \ref{fig:segblock}).
+The problem definition is the same in both algorithms but they differ in the method they use to compute fitness and error functions.
+\begin{description}
+	\item[ Problem definition:] The input stroke $S$  with $N$ points can be represented by set $S = \left\{ {x_1 ,x_2  \ldots x_N }\right\}$ where $x_i$ is the location of the point $i$.
+The swarm algorithms consist of $M$ agents which are represented by the set $A = \left\{ {P_i \left| {i = 1,2 \cdots M} \right.} \right\}$ where $P_i$ is a single solution particle from the solution space.
+Each particle decodes the problem using a binary array with the same length $N$ as the input stroke.
+
+Therefore, the system represents each particle $P_i$ by $P_i = \left\{ {p_{ij} \left| {j = 1,2 \cdots N} \right.} \right\}$ where $p_{ij}$ has only two values a)1 ($p_{ij}=1$); means that this point ($j$) is a dominate point, or b) 0($p_{ij}=0$) means this point ($j$) is not a dominate point.
+Figure\ref{fig:pso1} shows a particle encoding in the DPSO system.
+
+	\item[Fitness function:] The fitness function and error calculation are different in each of the two \textit{DPSO} algorithms. 
+	\end{description}
+\subsubsection{\textit{Polygon approximation\textsl{AlgS1}}}
+%\\
+The approximation error is computed by the equation \ref{eq:ErrorSwarm1}. The graphical meaning of the error is shown in Figure\ref{fig:LabelsPSO}.
+\begin{equation}
+E=\sum\nolimits_{i = 0}^M e ( \widehat{x_ix_{i+1}},\overline{x_i x_{i+1}})
+\label{eq:ErrorSwarm1}
+\end{equation} where the arc $\widehat{x_ix_j}$ is defined as the consecutive set of points from point $x_i$ to point $x_{j}$ as in $x_i,x_{i+1} \cdots,x_j$.
+The line $\overline{x_i x_j} $ as the straight line connecting point $x_i$ to point $x_j$, $M$ is the number of dominant points in this solution as generated by the swarm algorithm.
+The error $e ( \widehat{x_ix_j},\overline{x_i x_j})$ is computed as the sum of squared perpendicular distance from every point along the arc $\widehat{x_ix_j}$ to the line $\overline{x_i x_j}$.
+The fitness is computed using equation %\ref{eq:fitnessSwarm1}
+\begin{equation}
+\max fitness(p_i ) = \left\{ {\begin{array}{*{20}c}
+   { - E/\varepsilon N} & {ifE > \varepsilon ,}  \\
+   {D/\sum\limits_{j = 1}^N {p_{ij} } } & {otherwise}  \\
+\end{array}} \right.
+\label{eq:fitnessSwarm1}
+\end{equation} where $N$ is the number of points in the stroke, $D$ is the number of points in the solution that was previously labeled as a possible dominant point ($P_{pd}$), $E$ is the computed error and $\varepsilon$ is the error threshold.
+It should be noticed that when the error is larger than the threshold $\varepsilon$ the fitness is given a -ve value to lower the fitness value of the solution.
+Otherwise the system favors the lower number of vertices.
+
+ \subsubsection{\textit{Hybrid Fitting \textsl{AlgS2}}}
+The algorithm has the same problem formulation but different fitness and error functions are used.
+An attempt is made to fit each segment into a line or circular arc.
+The errors of both circle and line fit estimations are computed for each segment $S_i=\widehat{x_ix_j}$.
+The approximation with the lower error value is chosen as the final approximation of this segment $S_i$\cite{CruveDivisionSwarm}.
+The sum of the approximation errors of all segments is the total error of the particle.
+The total error of the particle is computed by equation %\ref{eq:errorSwarm2}
+ \begin{equation}
+E=\sum\nolimits_{i = 0}^M e(D_i) 
+\label{eq:errorSwarm2}
+\end{equation}where $M$ is the number of segments in the solution as generated by the swarm algorithm, $D_i$ is the minimum approximation error of curve and line approximations $min(d_c,d_l)$ where $d_c$ is the circle approximation error and $d_l$ is the line approximation error as computed by \cite{CruveDivisionSwarm}.
+The fitness is computed by the equation %\ref{eq:fitnessSwarm2}
+\begin{equation}
+\max fitness(P_i ) = \frac{1}{{E \times M^k }}
+\label{eq:fitnessSwarm2}
+\end{equation} where $E$ is the error and $M$ is the number of segments and $k$ is a parameter tweaked to get minimum number of segments.
+The larger $k$ is, the more effect the number of segments will have.
+For our system, $k$ is selected to be 0.5\cite{CruveDivisionSwarm}.
+
+After each iteration of the swarm algorithm (\textsl{AlgS1} and \textsl{AlgS2}), each particle is refined using the following procedures: 
+\begin{enumerate}
+	\item For each particle $P_i$ each dominant point $P_{ij}$ is checked to find if it was labeled before as a \textit{possible dominant point} $P_{pd}$ (computed as in section \ref{Prepross}).
+If it was not labeled the point $P_{ij}$ is moved to the nearest labeled point.
+This ensures that all of the points generated by the DPSO are possible dominant points $P_{pd}$.
+	\item The particles are tested to make sure that the distance between every two successive dominate point is larger than $min_D$, where $min_D$ is 5\% of the total length of the stroke.
+Otherwise, one of these points are removed based on the error cased by its removal.
+	\item Each two successive segments are test to determine if they can be merged into a single segment.
+If merging the two segments have the same or smaller the error, the two segments are merged.
+This step ensures that the final segmentation of the system has minimum number of segments.
+\end{enumerate}
+%1) 2)  % If two points are nearer than $min_D$ then one of the points is removed. 
+\subsection{Recognition}
 \label{sec:Recognition}
@@
 \section{Experiments} 
-
+\label{sec:Experiments}
 In our experiment we used the same benchmark dataset used in \cite{myPaper}, the dataset was collected by Hse and Newton\cite{HeloiseBeautification}.
@@
   \centering
-  \includegraphics[scale=0.5]{Experiment92.jpg}
+  \includegraphics[width=3.5in]{Experiment92.jpg}
   \caption[Recognition accuracy]%
@@
   \centering
-  \includegraphics[width=4in]{Experiment68.jpg}
+  \includegraphics[width=3.5in]{Experiment68.jpg}
   \caption[Recognition accuracy]%
@@
 
-\section{Conclusion and Future Work}
+\section{Conclusion }
+ \label{ConclusionandFutureWork}
 
-The training and testing performance measurements using both PSO and Enhanced-PSO in training and testing phases are shown in table (1)
+%The training and testing performance measurements using both PSO and Enhanced-PSO in training and testing phases are shown in figure\ref{exp2} and figure \ref{exp1}.
 
-Table 1:- S P measurement performance using both PSO and Enhanced-PSO
-We have studied the shape of the sigmoid weight of the hidden layer neurons as a parameter which was usually determined using trial and error.
-Thus, we add additional parameter called sigmoid weight to the hidden layer neurons to approximate the required non linear dynamical mapping more precisely.
-We prove that our proposed neural network architecture is a dynamical one because it could be represented by a dynamical nonlinear mapping function.
-Also, we introduce a new enhanced particle Swarm Optimization method for adjusting
-weights of this new proposed neural network architecture based on random perturbation of the particles instead of the standard cognitive term.
-We prove that our proposed (SDRNN-EPSO) architecture could efficiently give better prediction accuracy and enhanced performance compared to using the SDRNN trained by the Basic-PSO
-For the future work, it is suggested that this new enhanced architecture will be used in other different applications of time series prediction and also extended to be used in other applications rather than time series prediction such as pattern recognition and speech verification
+This paper used an  Enhanced-PSO  to modify a new sketch recognition using DPSO represented in \cite{mypaper}.
+Results show that enhancing the solutions that the \textit{PSO} in general generates an optimized stroke segmentation which improves the final recognition rate.
+The final recognition improved by modfiying the particle optimization algorithm without depending on the higher level features.
+The experiments evaluated two \textit{DPSO} different segmentation algorithms (AlgS1 and AlgS2) on simple presentation set dataset.
+Results show that the enhanced PSO algorithm improves the segmenattion on both algorithms which achieves better segmentation and final accuracy.
+The results proved that the Enhancment made on the PSO improves the efficency of the Basic PSO algorithm without being dependable on special objective function or application.
+The system achieved an average overall improvement of more than 1\% over Basic PSO.
 
-This paper presented a new approach to sketch recognition using DPSO.
-Results show that the \textit{DPSO} in general generates an optimized stroke segmentation which improves the final recognition rate.
-The final recognition is based on geometrical, statistical and structural features.
-The experiments evaluated two \textit{DPSO} different segmentation algorithms (AlgS1 and AlgS2) on simple presentation set dataset.
-Results show that AlgS2 achieves better segmentation and final accuracy than Alg3 \cite{earlyprocess} on the dataset used.
-The results proved that the system is easily expandable to more and more symbol as it does not depend on low level primitive recognizers but on high level symbol and geometrical features.
-The system achieved an average overall improvement of more than 2\% over Alg3.
+ For the future work, it is suggested that we continue to improve both the enhanced PSO segmentation algorithm.
+Also, a possible extentsion of this research is to improve the recognition algorithm of the Sketch Recognition system.
+Other area of enhancements is the features extraction methods.
+Introducing more spatial and geometrical features is believed to improve classifications.
+%this new enhanced architecture will be used in other different applications of time series prediction and also extended to be used in other applications rather than time series prediction such as pattern recognition and speech verification
 
-This paper presented a new approach to sketch recognition using PSO.
-The system uses both speed and curvature information which help improving the \textit{DPSO}.
-It is noted that the \textit{DPSO} in general generates an optimized stroke segmentation which improves the final recognition rate.
-The tradeoff between accuracy achieved and time complexity must be further investigated to achieve better results.
-The use of statistical moments and structural features improves the recognition rate.
-The system was tested on 13 different symbols and achieved an accuracy of 91.5\%.
-The system does not depend on low level recognizer but rather on a set of high level features.
-This makes the system easily expandable to more and more symbols.
-
- A possible extension of this research is to complete the clustering algorithm for fully automated sketch recognition.
+ %A possible extension of this research is to complete the clustering algorithm for fully automated sketch recognition.
 Currently, the system process one symbol at a time, an enchantment would be to allow users to draw more than one symbol at a time and incorporate a method for separating symbols.

### 2010-08-16 trunk/Swarm/tmp/swarmModified.tex
correcting

@@
 Results show that that the enhanced particle swarm optimization improves the final recognition accuracy.
+
  % sketch are divided Users draw symbols and sketches using a set of pen strokes. 
@@
 
-This paper modifies the sketch recognition system presented in \cite{mypaper}, to use Enhanced particle swarm algorithm (EPSO) to optimally segment user's strokes.
-As in the Orginal system the users can draw symbols using any number of strokes in any order they like.
+This paper modifies the sketch recognition system presented in \cite{mypaper}, to use enhanced particle swarm algorithm (EPSO) to optimally segment user's strokes.
+As in the Original system the users can draw symbols using any number of strokes in any order they like.
 The system segments each stroke using two different EPSO algorithms.
@@
 The remaining of the paper is as following; Section \ref{sec:ParticleSwarmAlgorithm} describes the Basic particle swarm algorithm in details.
-Enhanced Particle Swarm Algorithm is discribed in Section \ref{sec:EnhancedSwarmAlgorithm}.
+Enhanced Particle Swarm Algorithm is described in Section \ref{sec:EnhancedSwarmAlgorithm}.
 Section \ref{Sysdisc} provide details on preprocessing, segmentation and recognition process.
@@
 \label{sec:EnhancedSwarmAlgorithm}
-In the above basic PSO, the velocity update is based on three terms an interia term, local cognitive ($l_{best}$) term and global cognitive ($g_{best}$)term.
+In the above basic PSO, the velocity update is based on three terms an inertia term, local cognitive ($l_{best}$) term and global cognitive ($g_{best}$)term.
 The particle local cognitive term is associated with the particle own history without consideration of other swarm particles locations and histories which leads to a lake of diversity and sub optimum solutions.
-Therefore, we are tyring to add more social influence in the particle movement by using the weighted difference of the position vectors of any other two distinct particles randomly chosen from the swarm to modify the local cognitive term.
-The modified term adds social behvaiour of other particle and keep diversity while searching for the global optimum.
+Therefore, we are trying to add more social influence in the particle movement by using the weighted difference of the position vectors of any other two distinct particles randomly chosen from the swarm to modify the local cognitive term.
+The modified term adds social behavior of other particle and keeps diversity while searching for the global optimum.
 The new social cognitive term of the particle is shifted to the new location only if it yields a fitness value better than obtained using the standard PSO\cite{stockPaper}.
@@
 %\label{sec:application}
-The modified swarm algorithm is used to improve the performance of the sketch recoginition system in \cite{myPaper}. The 
+The modified swarm algorithm is used to improve the performance of the sketch recognition system in \cite{myPaper}. 
 %\subsection{The system}
 
- The suggested sketch recognitions system is divided into three main steps 1) preprocessing, 2) segmentation and 3) Recognition. The following section provides a detailed description of each step.
+ The sketch recognitions system is divided into three main steps 1) preprocessing, 2) segmentation and 3) Recognition. The following section provides a detailed description of each step.
 \subsection{Preprocessing}
@@
 
-After each iteration of the swarm algorithm (\textsl{AlgS1} and \textsl{AlgS2}), each particle is refined using the following procedures: 
+After each iteration, of the swarm algorithm (\textsl{AlgS1} and \textsl{AlgS2}), each particle is refined using the following procedures: 
 \begin{enumerate}
@@
 
-Figures \ref{exp2} and \ref{exp1} show the recognition rate of each algorithm using the full feature set and half the features set respectivily.
-The results show that the enhanced swarm optimization algorithm improve the segmentation of the system which leds to an increase in the recognition rate of the whole system by an average of $1.5\%$.
-The increase in recognition rate is consistant in all swarm algorithms and in both experiments.
+Figures \ref{exp2} and \ref{exp1} show the recognition rate of each algorithm using the full feature set and half the features set respectively.
+The results show that the enhanced swarm optimization algorithm improves the segmentation of the system which leads to an increase in the recognition rate of the whole system by an average of $1.5\%$.
+The increase in recognition rate is consistent in all swarm algorithms and in both experiments.
 %n the We performed two experiments to test the system.
@@
 Results show that enhancing the solutions that the \textit{PSO} in general generates an optimized stroke segmentation which improves the final recognition rate.
-The final recognition improved by modfiying the particle optimization algorithm without depending on the higher level features.
+The final recognition improved by modifying the particle optimization algorithm without depending on the higher level features.
 The experiments evaluated two \textit{DPSO} different segmentation algorithms (AlgS1 and AlgS2) on simple presentation set dataset.
-Results show that the enhanced PSO algorithm improves the segmenattion on both algorithms which achieves better segmentation and final accuracy.
-The results proved that the Enhancment made on the PSO improves the efficency of the Basic PSO algorithm without being dependable on special objective function or application.
+Results show that the enhanced PSO algorithm improves the segmentation on both algorithms which achieves better segmentation and final accuracy.
+The results proved that the Enhancement made on the PSO improves the efficiency of the Basic PSO algorithm without being dependable on special objective function or application.
 The system achieved an average overall improvement of more than 1\% over Basic PSO.
@@
  For the future work, it is suggested that we continue to improve both the enhanced PSO segmentation algorithm.
-Also, a possible extentsion of this research is to improve the recognition algorithm of the Sketch Recognition system.
+Also, a possible extension of this research is to improve the recognition algorithm of the Sketch Recognition system.
 Other area of enhancements is the features extraction methods.

### 2010-08-18 Swarm/tmp/swarmModified.tex
after finishing the code

--- 2010-06-29 Swarm/swarmModified.tex
@@
 
-\title{Enhanced Particle Swarm Optimization}
+\title{Enhanced Particle Swarm Optimization In Sketch Recognition}
 \twoauthors
-  {Tarek abouldahab\sthanks{Corresponding Author}} {Cairo Metro Company \\
-  Ministry of transport \\Cairo, Egypt\\ heshoaboda@hotmail.com}
- {Maha El Meseery, Mahmoud Fakhr El Din}
+{Maha El Meseery\sthanks{Corresponding Author}, Mahmoud Fakhr El Din}
 	{Signals Processing Group \\ Computers and Systems Department \\
@@
 %This study
+
 Sketches and drawings are widely used as a simple method of expressing thought and designs.
-The goal of this paper is developing a sketch recognition system that facilitates using sketches in computer systems.
-The system uses Particle Swarm Optimization (PSO) algorithm to optimally segment the strokes the user draws into meaningful geometric primitives.
+The goal of this paper is to improve our sketch recognition system by using an enhanced particle swarm optimization algorithm.
+The enhanced Particle Swarm Optimization (EPSO) is based on the perturbation of the particle velocity is used to optimally segment the strokes the user draws into meaningful geometric primitives.
 These geometric primitives are grouped to formulate symbols which are further identified using a Support Vector Machines (SVM) classifier.
-Two different segmentation algorithms were evaluated and results show that both algorithms improve the final recognition results.
-Experiments were conducted on three different dataset, simple presentation (Hs-DS), Electrical Design (EL-DS) and Logic Design (LD-DS) symbols.
-Results show that the effectiveness of the proposed system on various dataset especially on datasets with low shape diversity as in Logic Design dataset (LD-DS).
+Experiments were conducted on the benchmark dataset of simple presentation (Hs-DS) symbols.
+Results show that that the enhanced particle swarm optimization improves the final recognition accuracy.
+
  % sketch are divided Users draw symbols and sketches using a set of pen strokes. 
@@
 As for recognition we use Support Vector Machines (SVM) classifier %which further improves the final recognition accuracy.
+%%% old abstract.. 
+ %Also another enhanced Particle Swarm Optimization (EPSO) based on the perturbation of the particle velocity is introduced to train the weights of this proposed neural network architecture.
+This proposed architecture with its relevant training algorithm is mentioned as (SDRNN-EPSO) and its performance is compare with the proposed network architecture trained be Basic-PSO algorithm.
+Simulation results will be carried out on stock market indices as an example of time series prediction which shows that our proposed training algorithm is more efficient and accurate than the Basic-PSO in stock market prediction.
 \end{abstract}
@@
 
-Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.
-Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.
-Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.
-Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.
-Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.
+%However, in PSO algorithm as the number of iterations increases, the particles speed decrease leading to converge to the global best point found so far, which is not guaranteed even to be a local minimum..
+To enhance the performance and explore better solutions, many researchers introduced GAs operations like crossover, mutation, and selection operators or tuning the PSO parameters leading to increase the computational effort  \cite{Ouyang2009IJCAI}.
+
+%As the need for higher effective neural network architecture that increases the neural network output accuracy is highly required.
+So, in this study we propose a new architecture based on introducing the hidden layer sigmoid weight which adapts the shape of the hidden layer output sigmoid function.
+Also, another Enhanced Particle Swarm Optimization training algorithm is applied in training network weights.
+
+The field of symbol recognition has gained interest in the last few years.
+Scientists generally and engineers specifically express thoughts and designs using sketches.
+Engineers use sketches to exchange designs as a natural method of communication rather than writing or speaking.
+Design engineers need a powerful computer based system with symbol recognition to help them design, manipulate and store sketches more effectively than using only papers.
+Increasing the interaction between computers and users in sketch and CAD systems has been the reason for the emerging of a few advanced sketch recognition systems.
+Sketch recognition is the process of identifying user strokes into meaningful symbols that can be further used.
+
+Sketch recognition is divided into three main steps preprocessing, segmentation and symbol recognition.
+The preprocessing phase captures user input stroke points and collects basic information about the stroke then proceeds to remove noise and compute basic statistical and geometrical information.
+In segmentation phase, the strokes are divided into a set of simple geometrical primitives or segments.
+In the third phase, sketch recognition, strokes and segments are clustered to formulate symbols that can be recognized by a classifier system.
+
+%The segmentation problem is summarized as finding the best decomposition with least number of segments each represents a geometric primitive.
+This is an optimization problem, which evolutionary programming can solve efficiently.
+A genetic algorithm was used by \cite{CruveDivisionSwarm} to optimally divide digital curves into lines and curves.
+Chen et al.\cite{CruveDivisionSwarm} uses digital curves scanned from paper as input to the system and did not take advantage of the curvature or local geometric properties of the digital curve.
+
+This paper modifies the sketch recognition system presented in \cite{mypaper}, to use enhanced particle swarm algorithm (EPSO) to optimally segment user's strokes.
+As in the Original system the users can draw symbols using any number of strokes in any order they like.
+The system segments each stroke using two different EPSO algorithms.
+Segmentation is based on dividing the digital curve into polygon\cite{PolygonApproximationPSO}.
+To handle complex strokes, our system uses a second segmentation EPSO algorithm to segment strokes into a set of lines and circular arcs.
+After segmentation, a symbol features vector composed of different statistical, geometrical and spatial features is deduced.
+The final classification step, SVM classifier uses the computed feature vector to classify the segments into one of the previously trained classes.
+
+The remaining of the paper is as following; Section \ref{sec:ParticleSwarmAlgorithm} describes the Basic particle swarm algorithm in details.
+Enhanced Particle Swarm Algorithm is described in Section \ref{sec:EnhancedSwarmAlgorithm}.
+Section \ref{Sysdisc} provide details on preprocessing, segmentation and recognition process.
+Section \ref{sec:Experiments} describes the experiments that were preformed.
+We conclude in Section \ref{ConclusionandFutureWork}.
 
@@
 
-\section{Enhanced Partilce Swarm}
-This paper modifies the general PSO algorithm by changeing the local term into a random term using the following equation .  
+\section{Enhanced Particle Swarm}
+\label{sec:EnhancedSwarmAlgorithm}
+In the above basic PSO, the velocity update is based on three terms an inertia term, local cognitive ($l_{best}$) term and global cognitive ($g_{best}$)term.
+The particle local cognitive term is associated with the particle own history without consideration of other swarm particles locations and histories which leads to a lake of diversity and sub optimum solutions.
+Therefore, we are trying to add more social influence in the particle movement by using the weighted difference of the position vectors of any other two distinct particles randomly chosen from the swarm to modify the local cognitive term.
+The modified term adds social behavior of other particle and keeps diversity while searching for the global optimum.
+The new social cognitive term of the particle is shifted to the new location only if it yields a fitness value better than obtained using the standard PSO\cite{stockPaper}.
+
+This modification is done as follows:-
+  \begin{enumerate}
+\item For each particle $p_{i}$ in the swarm two other distinct particles, $p_L$ and $p_M$ are selected randomly where $i\neq L\neq M$.
+The difference between their positional coordinates $p_{d}$ is taken as a difference vector in equation \ref{eqmod}.
  \begin{equation}
-  v_{ij}  = v_{ij} \omega + c_1 r_1 (p_{iRandom_i}  - p_{iRandom_j} ) + c_2 r_2 (gbest_{ij}  - p_{ij} ),
-  \label{eq:ModSwarm}
+ \label{eqmod}
+ p_{dj}= p_{Lj}-p_{Mj} 
  \end{equation}
+where $p_{dj}$ represent the $j$th dimension in the difference vector and $p_{Lj}$ represent the $j$th dimension in the $L$th particle.
 
-where $iRandom_i$ and $Random_j$ are two particles in the solution space.
+Then the associated velocity vector of particle $p_{i}$ is updated according to this difference vector as in equation \ref{eq:mod1}.
+In essence the local cognitive term of the velocity update formula in \ref{eq:Swarm} is replaced with the vector differential operator $p_d$ to produce some additional exploration capability.
+ \begin{equation}
+ v^{T}_{ij}  = v_{ij} \omega + c_1 r_1 (p_{dj}) + c_2 r_2 (gbest_{ij}  - p_{ij} ),
+ \label{eq:mod1}
+ \end{equation}
+\item The new trial position $p^{T}_{ij}$ is created by adding the updated velocity $v^{T}_{ij}$ in Eq. \ref{eq:mod1}.
+  \begin{equation}
+ \label{eq:mod2}
+p^{T}_{ij}=p_{ij}+ v^{T}_{ij} ,
+ \end{equation}
+\item The new position $p^{T}_{ij}$ is used to generate the new trial objective fitness function value $fitness(p^{T})$.
+The fitness value is calculated based on this new trial position vector and the particle is shifted to this trial position if the trial objective function is better than the original objective function associated with standard PSO.
+Thus the target particle is relocated using the following equation (Eq.
+\ref{eq:objective})
+  \begin{equation}
+ \label{eq:objective}
+p^{new}_{ij}\Leftarrow 
+\{
+\begin{array}{c} 
+ p^{T}_{ij} \quad \quad if\quad fitness(p^{T})> fitness(p^{old})  \\
+ p_{ij} \quad \quad Otherwise\quad  
+\end{array}\}
+ \end{equation}
+ where $fitness(p^{T})$ is the objective fitness value of the new trail position $ p^{T}$ and  $fitness(p^{old})$is the objective fitness value of the new basic position $ p{i}$ computed by equation \ref{eq:Swarm1}.
+   \end{enumerate}
+
+%Otherwise add perturbation to modify this cognitive term by the weighted difference of the position vectors of any other two distinct particles randomly chosen from the swarm.
+The particle is shifted to this new location only if it yields a fitness value better than obtained using the standard PSO.
+
+%This paper modifies the general PSO algorithm by changeing the local term into a random term using the following equation .  
+ % \begin{equation}
+ % v_{ij}  = v_{ij} \omega + c_1 r_1 (p_{iRandom_i}  - p_{iRandom_j} ) + c_2 r_2 (gbest_{ij}  - p_{ij} ),
+ % \label{eq:ModSwarm}
+ % \end{equation}
+
+%where $iRandom_i$ and $Random_j$ are two particles in the solution space.
 This modification replace the distance between the current particle and the local best with two random particle from the space.
@@
 The new location is compared the location of particle if with equation \ref{eq:Swarm} if the modified location is better (mean better fitness function) then the new location is selected otherwise the old equation is used.
-\section{Applications }
-The modified swarm algorithm is used to improve the performance of the sketch recoginition system in \cite{myPaper}. The
+\section{System Discribtion  }\label{Sysdisc}
+%\label{sec:application}
+The modified swarm algorithm is used to improve the performance of the sketch recognition system in \cite{mypaper}.
+The sketch recognitions system is divided into three main steps 1) preprocessing, 2) segmentation and 3) Recognition.
+The following section provides a detailed description of each step.
+The main block of the system remains the same but their is some modification in the segmentation algorithm to improve its efficiency.
+The main modification is using Enhanced Particle swarm algorithm instead of Basic Particle swarm algorithms.
+Other modifications are explained in the details in the next sections.
+\subsection{Preprocessing}
+\label{Prepross}
+ Speed and time difference information were widely used in sketch understanding systems \cite{earlyprocess}.
+It was noted that time difference provides more distinct maxima than the speed information.
+Agar \cite{polygonfeedback31} mentions that the pointing device (i.e.
+the mouse) sampling rate is the reason for this phenomenon.
+ In the presented system, we computed time difference, direction, speed and curvature of each point along a stroke.
+The speed is calculated as $v=\Delta s/\Delta t$ where $t$ is the time difference between two points and $s$ is the length between them.
+The direction $d$ is calculated as the angle between two vectors and curvature is considered as the change in direction with respect to length i.e.
+$c= \Delta d/\Delta s$.
+Dominant points are characterized by lower speed values and higher curvature and direction values.
 
-\cite{myPaper}  
+After the system computes all the speed, time difference and curvature information it proceeds to detect the points with low velocity and high curvatures.
+Using simple differentiation to detect local extreme points resulted in false points due to the non smooth curves.
+Hence, the system adopted a process presented by \cite{earlyprocess}, where the mean of the curve is calculated.
+Then a threshold \textit{th} is used to separate the curve into regions $Region_i$; each region $Region_i$ is defined as a range of points, where the curve values are either above or below the threshold \textit{th}.
+Those regions are further processed to find the maximum point $Max(Region_i)$ of each region $Region_i$.
+The stroke points $p_i(x,y)$ that correspond to those maximum values are labeled as \textit{possible dominant points} $P_{pd}$.
+The system repeats this process for curvature, time difference and speed information.
+All the points labeled as possible dominant points $P_{pd}$ are saved into a single array.
+Figure \ref{fig:LabelsPPD} shows the particles labeled as Possible dominant point $P_{pd}$ by the preprocessing, it is noticed the redundancy of some $P_{pd}$ points.
+% (as shown are redundant)
+\subsection{Segmentation}
+\label{seg}
+In the segmentation stage the input stroke is divided into a set of primitives.
+As shown in Figure \ref{fig:segblock} first an attempt is made to fit the stroke points into a curve or an ellipse using a minimum square error fitting algorithm \cite{ellipsefit}.
+The ellipse fitting step helps to prevent the system from over segmenting the stroke into multiple lines or curves if the input stroke is an ellipse.
+If the stroke proved to be an ellipse arc then the segmentation process ends and the system proceeds to the next step.
+Otherwise, the stroke is passed to two further segmentation algorithms that divide the stroke to either lines or lines and curves.
+After two different segmentations are generated, the system chooses the segmentation with the minimum error.
+The following sections describe in details the ellipse detection algorithm and the two segmentation algorithms used to divide the stroke.
+% The input to these algorithms are the stroke points along with the possible dominant points $P_{pd}$ computed during preprocessing.
+The output is a set of dominant points which are connected by either lines or curves.
+
+%///AllBlockDiagram.jpg  you can change the block to the main block. 
+ \begin{figure}
+	\centering
+		\includegraphics[scale=0.48]{blockSmall.jpg}
+	\caption{Segmentation Block}
+	\label{fig:segblock}
+\end{figure}
+\subsection{Ellipse fitting}
+%The process starts by computing the center of the stroke bounding box.
+The bounding box center point is used as the first estimation of the ellipse center.
+The axes of the ellipse are estimated as $width/2$ and $height/2$ of the stroke bounding box.
+The least square fitting algorithm is used to minimize the fitting error of the ellipse equation \cite{chernov}.
+This process tries to fit the stroke points into an ellipse arc; it starts with computing the center of the stroke bounding box.
+The bounding box center point is used as the first estimation of the center of the ellipse.
+The axes of the ellipse are estimated as the $width/2$ and $height/2$ of the stroke bounding box where $width$ is width and $height$ is height of the bonding box.
+The least square fitting algorithm \cite{chernov} is used to minimize the fitting error of the ellipse Equation (\ref{eq:circleFit})
+
+\begin{equation}
+E = \sum\limits_{i = 0}^N {\frac{{(x_i - x_0 )}}{{a^2 }}^2  + \frac{{(y_i - y_0 )}}{{b^2 }}^2  - 1} 
+\label{eq:circleFit}
+\end{equation}
+
+ where $N$ is number of points in the stroke, $a,b$ are the length of ellipse axes, $x_0$ \& $y_0$ are the coordinates of the center point, $x_i$ \& $y_i$ are the coordinates of point $i$ in the stroke.
+A list of new values for $x_0$ , $y_0$ ,$a$ and $b$ are generated randomly from the older values with small increments after each loop.
+After few iteration, the final fit error of the estimated ellipse is reported.
+A modification to the system in \cite{mypaper} is trying to add another measure to compute the efficiency of the final estimated ellipse.
+Equation \ref{eq:circleError} ensures that the drawn percentage of ellipse is considered.
+This eliminates fitting a line into very large ellipse but leaves small ellipses to be fitted as a partial or full ellipse.
+
+ \begin{equation}
+eff= (P_{percent}/E)
+\label{eq:circleError}
+\end{equation}
+ \begin{equation}
+P_{percent}  = L_{stroke} /P_{ellipse} 
+\label{eq:ErrorArea}
+\end{equation}
+where $E$ is the error computed by Equation(\ref{eq:circleFit}), $L_{stroke}$ total length of stroke and $P_{ellipse} $ is the perimeter of the estimated ellipse.
+If $eff$ is more than threshold $th_{Ellipse}$\footnote{By trial and error best threshold found was $th_{Ellipse}=0.2$} then the stroke is segmented as an ellipse otherwise the system proceeds to the next step.
+
+\subsection{Non ellipse fitting algorithm}
+\label{subsubsec:Discreteparticleswarmalgorithm}
+\begin{figure}
+	\centering
+		%\subfigure[Possible dominate point]{ 	\label{fig:LabelsPPD}	\includegraphics[scale=0.7]{ppd.jpg}}
+	%	\hfill
+	 \includegraphics[scale=0.7]{pso1.jpg}			
+	\caption{DPSO algorithm encoding}% a) Possible dominate point b) Particle encoding  
+	\label{fig:pso1}
+\end{figure}
+Two DPSO algorithms are used to generate different segmentations for each stroke the user draws.
+For each segmentation generated, an error is evaluated.
+The segmentation with the minimum error value is chosen as the best stroke segmentation (Figure \ref{fig:segblock}).
+The problem definition is the same in both algorithms but they differ in the method they use to compute fitness and error functions.
+\begin{description}
+	\item[ Problem definition:] The input stroke $S$  with $N$ points can be represented by set $S = \left\{ {x_1 ,x_2  \ldots x_N }\right\}$ where $x_i$ is the location of the point $i$.
+The swarm algorithms consist of $M$ agents which are represented by the set $A = \left\{ {P_i \left| {i = 1,2 \cdots M} \right.} \right\}$ where $P_i$ is a single solution particle from the solution space.
+Each particle decodes the problem using a binary array with the same length $N$ as the input stroke.
+
+Therefore, the system represents each particle $P_i$ by $P_i = \left\{ {p_{ij} \left| {j = 1,2 \cdots N} \right.} \right\}$ where $p_{ij}$ has only two values a)1 ($p_{ij}=1$); means that this point ($j$) is a dominate point, or b) 0($p_{ij}=0$) means this point ($j$) is not a dominate point.
+Figure\ref{fig:pso1} shows a particle encoding in the DPSO system.
+
+	\item[Fitness function:] The fitness function and error calculation are different in each of the two \textit{DPSO} algorithms. 
+	\end{description}
+\subsubsection{\textit{Polygon approximation\textsl{AlgS1}}}
+%\\
+The approximation error is computed by the equation \ref{eq:ErrorSwarm1}. The graphical meaning of the error is shown in Figure\ref{fig:LabelsPSO}.
+\begin{equation}
+E=\sum\nolimits_{i = 0}^M e ( \widehat{x_ix_{i+1}},\overline{x_i x_{i+1}})
+\label{eq:ErrorSwarm1}
+\end{equation} where the arc $\widehat{x_ix_j}$ is defined as the consecutive set of points from point $x_i$ to point $x_{j}$ as in $x_i,x_{i+1} \cdots,x_j$.
+The line $\overline{x_i x_j} $ as the straight line connecting point $x_i$ to point $x_j$, $M$ is the number of dominant points in this solution as generated by the swarm algorithm.
+The error $e ( \widehat{x_ix_j},\overline{x_i x_j})$ is computed as the sum of squared perpendicular distance from every point along the arc $\widehat{x_ix_j}$ to the line $\overline{x_i x_j}$.
+The fitness is computed using equation %\ref{eq:fitnessSwarm1}
+\begin{equation}
+\max fitness(p_i ) = \left\{ {\begin{array}{*{20}c}
+   { - E/\varepsilon N} & {ifE > \varepsilon ,}  \\
+   {D/\sum\limits_{j = 1}^N {p_{ij} } } & {otherwise}  \\
+\end{array}} \right.
+\label{eq:fitnessSwarm1}
+\end{equation} where $N$ is the number of points in the stroke, $D$ is the number of points in the solution that was previously labeled as a possible dominant point ($P_{pd}$), $E$ is the computed error and $\varepsilon$ is the error threshold.
+It should be noticed that when the error is larger than the threshold $\varepsilon$ the fitness is given a -ve value to lower the fitness value of the solution.
+Otherwise the system favors the lower number of vertices.
+
+ \subsubsection{\textit{Hybrid Fitting \textsl{AlgS2}}}
+The algorithm has the same problem formulation but different fitness and error functions are used.
+An attempt is made to fit each segment into a line or circular arc.
+The errors of both circle and line fit estimations are computed for each segment $S_i=\widehat{x_ix_j}$.
+The approximation with the lower error value is chosen as the final approximation of this segment $S_i$\cite{CruveDivisionSwarm}.
+The sum of the approximation errors of all segments is the total error of the particle.
+The total error of the particle is computed by equation %\ref{eq:errorSwarm2}
+ \begin{equation}
+E=\sum\nolimits_{i = 0}^M e(D_i) 
+\label{eq:errorSwarm2}
+\end{equation}where $M$ is the number of segments in the solution as generated by the swarm algorithm, $D_i$ is the minimum approximation error of curve and line approximations $min(d_c,d_l)$ where $d_c$ is the circle approximation error and $d_l$ is the line approximation error as computed by \cite{CruveDivisionSwarm}.
+The fitness is computed by the equation %\ref{eq:fitnessSwarm2}
+\begin{equation}
+\max fitness(P_i ) = \frac{1}{{E \times M^k }}
+\label{eq:fitnessSwarm2}
+\end{equation} where $E$ is the error and $M$ is the number of segments and $k$ is a parameter tweaked to get minimum number of segments.
+The larger $k$ is, the more effect the number of segments will have.
+For our system, $k$ is selected to be 0.5\cite{CruveDivisionSwarm}.
+
+As a modification to the system in \cite{mypaper}. After each iteration, of the swarm algorithm (\textsl{AlgS1} and \textsl{AlgS2}), each particle is refined using the following procedures: 
+\begin{enumerate}
+	\item For each particle $P_i$ each dominant point $P_{ij}$ is checked to find if it was labeled before as a \textit{possible dominant point} $P_{pd}$ (computed as in section \ref{Prepross}).
+If it was not labeled the point $P_{ij}$ is moved to the nearest labeled point.
+This ensures that all of the points generated by the DPSO are possible dominant points $P_{pd}$.
+	\item The particles are tested to make sure that the distance between every two successive dominate point is larger than $min_D$, where $min_D$ is 5\% of the total length of the stroke.
+Otherwise, one of these points are removed based on the error cased by its removal.
+	\item Each two successive segments are test to determine if they can be merged into a single segment.
+If merging the two segments have the same or smaller the error, the two segments are merged.
+This step ensures that the final segmentation of the system has minimum number of segments.
+\end{enumerate}
+%1) 2)  % If two points are nearer than $min_D$ then one of the points is removed. 
+\subsection{Recognition}
+\label{sec:Recognition}
+After the user draws all strokes of the symbol, the set of un-recognized strokes is grouped together along with their segmentation as input to the feature extraction process.
+A composite set of features is used to generate a single feature vector.
+The features used consist of Rubine feature set,  Zernike moments of order 10 \cite{HeloiseBeautification}, ink density as well as some structural and spatial information like number of perpendiculars lines ,number of parallel lines and number of different types of primitives in each symbol.
+After computing the features the symbol is introduced to the (SVM) classifier.
+
+\cite{mypaper}  
 \section{Experiments} 
- The goal of the system is to modify the segmentation and recognition of the sketch system.
-It made the same type of experiement done in \cite{myPaper}.
-The dataset is divided into 13 category of goemetric symbols
-\section{Conclusion and Future Work}
-Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.
-Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.
-Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.
-Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.
-Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.
+\label{sec:Experiments}
+In our experiment we used the same benchmark dataset used in \cite{mypaper}, the dataset was collected by Hse and Newton\cite{HeloiseBeautification}.
+The data are drawn by 16 users each of them was asked to draw 13 shapes from 30 to 50 times.
+Figure \ref{fig:symbolSet} shows a set of the shapes used in the data set.
+Five different splits were generated to divide the dataset into training and test sets.
+The results displayed are the average recognition accuracy of the five splits.
+The accuracy is computed as the number of correctly recognized samples divided by the total number of samples in the test.
+
+ \begin{figure}
+  \centering
+
+		\includegraphics[scale=0.5]{symbolSet.PNG} 
+		\caption[The Symbol Set] {The symbol set}   \label{fig:symbolSet}
+	 \end{figure}
+	  %The first experiment with the whole feature set and the second experiments with only half of the features.
+To demonstrate the change in accuracy is due to the enhanced swarm optimization algorithm we tested the recognition system using two sets of features.
+Firstly, we tested recognition accuracy of shapes in the data set with both algorithms (\textsl{AlgS1} and \textsl{AlgS2}) using the whole feature set.
+The second experiment was tested on both algorithm but only using half the feature set.
+Both experiments were compared with the orginal sketch recognition system prsenented in \cite{mypaper}.
+
+Figures \ref{exp2} and \ref{exp1} show the recognition rate of each algorithm using the full feature set and half the features set respectively.
+The figures compares between the result that can be obtained by system in \cite{mypaper} and modifying the sytem using Enhanced PSO.
+The results show that the enhanced swarm optimization algorithm improves the segmentation of the system which leads to an increase in the recognition rate of the whole system by an average of $1.5\%$.
+The increase in recognition rate is consistent in all swarm algorithms and in both experiments.
+%n the We performed two experiments to test the system.
+Firstly, we tested recognition accuracy of shapes in the data set using the orginal swarm segmentation algorithm and then after using the enhanched swarm optimization.
+To  both algorithms (\textsl{AlgS1} and \textsl{AlgS2}).
+We also implemented the segmentation algorithm described in \cite{earlyprocess} (\textsl{Alg3}) to compare it with our swarm algorithms.
+Figure \ref{fig:test1} shows the accuracy achieved by each algorithm.
+The result shows that (\textsl{AlgS2}) gives better performance than Alg3 and AlgS1.
+
+%In our experiment we used three different datasets.
+First dataset was collected by Hse and Newton\cite{HeloiseBeautification} (Hs-DB).
+The data are drawn by 16 users each of them was asked to draw 13 shapes from 30 to 50 times.
+Figure \ref{fig:symbolSet} shows a set of the shapes used in the data set.
+We collected the second and third datasets.
+Figure \ref{fig:ELsymbolSet} and \ref{fig:LogicsymbolSet} shows the symbols gathered from 7 different users in electrical (EL-DB) and logic design (LD-DB) domain respectively.
+Table \ref{tab:datasets} shows a  comparison between the three datasets used in our experiments.
+The datasets were divided into training set and test set.
+Five different splits  for the training and test data are generated from each dataset.
+The results displayed are the average recognition accuracy of the five splits.
+The accuracy is computed as the number of correctly recognized samples divided by the total number of samples in the test.
+
+ %The goal of the system is to modify the segmentation and recognition of the sketch system.
+It made the same type of experiement done in \cite{mypaper}.
+The dataset is divided into 13 category of goemetric %symbols
+
+% The dataset collected by Hse and Newton\cite{HeloiseBeautification} is used to test the presented system.
+Figure \ref{fig:symbolSet} shows a set of the shapes used in the data set.
+The datasets were divided into training and test sets.
+Five different splits for the training and test data are generated from each dataset.
+The results displayed are the average recognition accuracy of the five splits.
+The accuracy is computed as the number of correctly recognized samples divided by the total number of samples in the test.
+    \begin{figure}
+  \centering
+  \includegraphics[width=3.3in]{Experiment92.jpg}
+  \caption[Recognition accuracy]%
+  {The chart compares between recognition rate of the full feature set on both the orginal system presented in \cite{mypaper} and Enhanced PSO.
+The results shows that Enhanced PSO increases the performance over the orginal system.
+}
+   \label{exp2}
+\end{figure}
+\begin{figure}
+  \centering
+  \includegraphics[width=3.3in]{Experiment68.jpg}
+  \caption[Recognition accuracy]%
+  {The chart compares between recognition rate of half the feature set on both the orginal system presented in \cite{mypaper} and Enhanced PSO.
+The results shows that Enhanced PSO increase the recognition rate.
+}
+   \label{exp1}
+\end{figure}
+
+% We performed two experiments to test the system.
+Firstly, we tested recognition accuracy of shapes in the data set with both algorithms (\textsl{AlgS1} and \textsl{AlgS2}).
+We also implemented the segmentation algorithm described in \cite{earlyprocess} (\textsl{Alg3}) to compare it with our swarm algorithms.
+Figure \ref{fig:test1} shows the accuracy achieved by each algorithm.
+The result shows that (\textsl{AlgS2}) gives better performance than Alg3 and AlgS1.
+
+\section{Conclusion }
+ \label{ConclusionandFutureWork}
+
+%The training and testing performance measurements using both PSO and Enhanced-PSO in training and testing phases are shown in figure\ref{exp2} and figure \ref{exp1}.
+
+This paper used an Enhanced-PSO  to modify a new sketch recognition using DPSO represented in \cite{mypaper}.
+Results show that enhancing the solutions that the \textit{PSO} in general generates an optimized stroke segmentation which improves the final recognition rate.
+The final recognition improved by modifying the particle optimization algorithm without depending on the higher level features.
+The experiments evaluated two \textit{DPSO} different segmentation algorithms (AlgS1 and AlgS2) on simple presentation set dataset.
+Results show that the enhanced PSO algorithm improves the segmentation on both algorithms which achieves better segmentation and final accuracy.
+The results proved that the Enhancement made on the PSO improves the efficiency of the Basic PSO algorithm without being dependable on special objective function or application.
+The system achieved an average overall improvement of more than 1\% over Basic PSO.
+
+ For the future work, it is suggested that we continue to improve both the enhanced PSO segmentation algorithm.
+Also, a possible extension of this research is to improve the recognition algorithm of the Sketch Recognition system.
+Other area of enhancements is the features extraction methods.
+Introducing more spatial and geometrical features is believed to improve classifications.
+%this new enhanced architecture will be used in other different applications of time series prediction and also extended to be used in other applications rather than time series prediction such as pattern recognition and speech verification
+
+ %A possible extension of this research is to complete the clustering algorithm for fully automated sketch recognition.
+Currently, the system process one symbol at a time, an enchantment would be to allow users to draw more than one symbol at a time and incorporate a method for separating symbols.
+Other area of enhancements is the features extraction methods.
+Introducing more spatial and geometrical features is believed to improve classifications.
 

### 2010-08-18 trunk/Swarm/tmp/swarmModified.tex

@@
 %\label{sec:application}
-The modified swarm algorithm is used to improve the performance of the sketch recognition system in \cite{myPaper}. 
-%\subsection{The system}
-
- The sketch recognitions system is divided into three main steps 1) preprocessing, 2) segmentation and 3) Recognition. The following section provides a detailed description of each step.
+The modified swarm algorithm is used to improve the performance of the sketch recognition system in \cite{mypaper}.
+The sketch recognitions system is divided into three main steps 1) preprocessing, 2) segmentation and 3) Recognition.
+The following section provides a detailed description of each step.
+The main block of the system remains the same but their is some modification in the segmentation algorithm to improve its efficiency.
+The main modification is using Enhanced Particle swarm algorithm instead of Basic Particle swarm algorithms.
+Other modifications are explained in the details in the next sections.
 \subsection{Preprocessing}
@@
 After few iteration, the final fit error of the estimated ellipse is reported.
-Another measure is used to compute the efficiency of the final estimated ellipse.
+A modification to the system in \cite{mypaper} is trying to add another measure to compute the efficiency of the final estimated ellipse.
 Equation \ref{eq:circleError} ensures that the drawn percentage of ellipse is considered.
@@
 
-After each iteration, of the swarm algorithm (\textsl{AlgS1} and \textsl{AlgS2}), each particle is refined using the following procedures: 
+As a modification to the system in \cite{mypaper}. After each iteration, of the swarm algorithm (\textsl{AlgS1} and \textsl{AlgS2}), each particle is refined using the following procedures: 
 \begin{enumerate}
@@
 
-\cite{myPaper}  
+\cite{mypaper}  
 \section{Experiments} 
 \label{sec:Experiments}
-In our experiment we used the same benchmark dataset used in \cite{myPaper}, the dataset was collected by Hse and Newton\cite{HeloiseBeautification}.
+In our experiment we used the same benchmark dataset used in \cite{mypaper}, the dataset was collected by Hse and Newton\cite{HeloiseBeautification}.
 The data are drawn by 16 users each of them was asked to draw 13 shapes from 30 to 50 times.
@@
 The second experiment was tested on both algorithm but only using half the feature set.
+Both experiments were compared with the original sketch recognition system presented in \cite{mypaper}.
 
 Figures \ref{exp2} and \ref{exp1} show the recognition rate of each algorithm using the full feature set and half the features set respectively.
+The figures compares between the result that can be obtained by system in \cite{mypaper} and modifying the system using Enhanced PSO.
 The results show that the enhanced swarm optimization algorithm improves the segmentation of the system which leads to an increase in the recognition rate of the whole system by an average of $1.5\%$.
@@
  %The goal of the system is to modify the segmentation and recognition of the sketch system.
-It made the same type of experiement done in \cite{myPaper}.
+It made the same type of experiement done in \cite{mypaper}.
 The dataset is divided into 13 category of goemetric %symbols
@@
   \centering
-  \includegraphics[width=3.5in]{Experiment92.jpg}
+  \includegraphics[width=3.3in]{Experiment92.jpg}
   \caption[Recognition accuracy]%
-  {The recognition rate of system using the feature set.}
+  {The chart compares between recognition rate of the full feature set on both the original system presented in \cite{mypaper} and Enhanced PSO.
+The results shows that Enhanced PSO increases the performance over the original system.
+}
    \label{exp2}
@@
   \centering
-  \includegraphics[width=3.5in]{Experiment68.jpg}
+  \includegraphics[width=3.3in]{Experiment68.jpg}
   \caption[Recognition accuracy]%
-  {The recognition rate of system using half the feature set.}
+  {The chart compares between recognition rate of half the feature set on both the original system presented in \cite{mypaper} and Enhanced PSO.
+The results shows that Enhanced PSO increase the recognition rate.
+}
    \label{exp1}

### 2010-08-18 Swarm/tmp/swarmModified.tex
removing the comments

@@
 
- % sketch are divided Users draw symbols and sketches using a set of pen strokes. 
-%Sketch recognition is defined as the process of identifying symbols that users draw using single or multiple %strokes.
-Usually, users draw strokes using pens.
-The sketch recognition system immediately interprets their %strokes as objects that can be easily manipulated.
-This paper uses Particle Swarm Optimization (PSO) to divide %the strokes the user draws into meaningful geometric primitives.
-These geometric primitives are grouped to %formulate symbols which are further identified.
-The results show that using PSO improves segmentation results %which guide the symbol recognition phase.
-As for recognition we use Support Vector Machines (SVM) classifier %which further improves the final recognition accuracy.
-%%% old abstract.. 
- %Also another enhanced Particle Swarm Optimization (EPSO) based on the perturbation of the particle velocity is introduced to train the weights of this proposed neural network architecture.
-This proposed architecture with its relevant training algorithm is mentioned as (SDRNN-EPSO) and its performance is compare with the proposed network architecture trained be Basic-PSO algorithm.
-Simulation results will be carried out on stock market indices as an example of time series prediction which shows that our proposed training algorithm is more efficient and accurate than the Basic-PSO in stock market prediction.
 \end{abstract}
@@
  \section{Introduction}
-
-%However, in PSO algorithm as the number of iterations increases, the particles speed decrease leading to converge to the global best point found so far, which is not guaranteed even to be a local minimum..
-To enhance the performance and explore better solutions, many researchers introduced GAs operations like crossover, mutation, and selection operators or tuning the PSO parameters leading to increase the computational effort  \cite{Ouyang2009IJCAI}.
-
-%As the need for higher effective neural network architecture that increases the neural network output accuracy is highly required.
-So, in this study we propose a new architecture based on introducing the hidden layer sigmoid weight which adapts the shape of the hidden layer output sigmoid function.
-Also, another Enhanced Particle Swarm Optimization training algorithm is applied in training network weights.
 
@@
 In the third phase, sketch recognition, strokes and segments are clustered to formulate symbols that can be recognized by a classifier system.
-
-%The segmentation problem is summarized as finding the best decomposition with least number of segments each represents a geometric primitive.
-This is an optimization problem, which evolutionary programming can solve efficiently.
-A genetic algorithm was used by \cite{CruveDivisionSwarm} to optimally divide digital curves into lines and curves.
-Chen et al.\cite{CruveDivisionSwarm} uses digital curves scanned from paper as input to the system and did not take advantage of the curvature or local geometric properties of the digital curve.
 
@@
 where $p_{ij}$ represent the $j$th dimension in the $i$th particle and $v_{ij}$ is the velocity of the $j$th dimension in the $i$th particle.
- %Equation [\ref{eq:Swarm}] shows how velocity and direction of each particle are computed
+
  \begin{equation}
@@
 
-%Otherwise add perturbation to modify this cognitive term by the weighted difference of the position vectors of any other two distinct particles randomly chosen from the swarm.
-The particle is shifted to this new location only if it yields a fitness value better than obtained using the standard PSO.
-
-%This paper modifies the general PSO algorithm by changeing the local term into a random term using the following equation .  
- % \begin{equation}
- % v_{ij}  = v_{ij} \omega + c_1 r_1 (p_{iRandom_i}  - p_{iRandom_j} ) + c_2 r_2 (gbest_{ij}  - p_{ij} ),
- % \label{eq:ModSwarm}
- % \end{equation}
-
-%where $iRandom_i$ and $Random_j$ are two particles in the solution space.
-This modification replace the distance between the current particle and the local best with two random particle from the space.
-The effect of this change in more expolration and avoiding of local min problems.
-Equation \ref{eq:ModSwarm} is computed and the new particle locaiton is calculated.
-The new location is compared the location of particle if with equation \ref{eq:Swarm} if the modified location is better (mean better fitness function) then the new location is selected otherwise the old equation is used.
 \section{System Discribtion  }\label{Sysdisc}
@@
 The following sections describe in details the ellipse detection algorithm and the two segmentation algorithms used to divide the stroke.
-% The input to these algorithms are the stroke points along with the possible dominant points $P_{pd}$ computed during preprocessing.
-The output is a set of dominant points which are connected by either lines or curves.
-
-%///AllBlockDiagram.jpg  you can change the block to the main block. 
  \begin{figure}
@@
 \subsection{Ellipse fitting}
-%The process starts by computing the center of the stroke bounding box.
-The bounding box center point is used as the first estimation of the ellipse center.
-The axes of the ellipse are estimated as $width/2$ and $height/2$ of the stroke bounding box.
-The least square fitting algorithm is used to minimize the fitting error of the ellipse equation \cite{chernov}.
+
 This process tries to fit the stroke points into an ellipse arc; it starts with computing the center of the stroke bounding box.
@@
 	\centering
-		%\subfigure[Possible dominate point]{ 	\label{fig:LabelsPPD}	\includegraphics[scale=0.7]{ppd.jpg}}
-	%	\hfill
+
 	 \includegraphics[scale=0.7]{pso1.jpg}			
-	\caption{DPSO algorithm encoding}% a) Possible dominate point b) Particle encoding  
+	\caption{DPSO algorithm encoding}
 	\label{fig:pso1}
@@
 \end{enumerate}
-%1) 2)  % If two points are nearer than $min_D$ then one of the points is removed. 
+
 \subsection{Recognition}
@@
 	 \end{figure}
-	  %The first experiment with the whole feature set and the second experiments with only half of the features.
+
 To demonstrate the change in accuracy is due to the enhanced swarm optimization algorithm we tested the recognition system using two sets of features.
@@
 The second experiment was tested on both algorithm but only using half the feature set.
-Both experiments were compared with the orginal sketch recognition system prsenented in \cite{mypaper}.
+Both experiments were compared with the original sketch recognition system presented in \cite{mypaper}.
 
 Figures \ref{exp2} and \ref{exp1} show the recognition rate of each algorithm using the full feature set and half the features set respectively.
-The figures compares between the result that can be obtained by system in \cite{mypaper} and modifying the sytem using Enhanced PSO.
+The figures compares between the result that can be obtained by system in \cite{mypaper} and modifying the system using Enhanced PSO.
 The results show that the enhanced swarm optimization algorithm improves the segmentation of the system which leads to an increase in the recognition rate of the whole system by an average of $1.5\%$.
 The increase in recognition rate is consistent in all swarm algorithms and in both experiments.
-%n the We performed two experiments to test the system.
-Firstly, we tested recognition accuracy of shapes in the data set using the orginal swarm segmentation algorithm and then after using the enhanched swarm optimization.
-To  both algorithms (\textsl{AlgS1} and \textsl{AlgS2}).
-We also implemented the segmentation algorithm described in \cite{earlyprocess} (\textsl{Alg3}) to compare it with our swarm algorithms.
-Figure \ref{fig:test1} shows the accuracy achieved by each algorithm.
-The result shows that (\textsl{AlgS2}) gives better performance than Alg3 and AlgS1.
-
-%In our experiment we used three different datasets.
-First dataset was collected by Hse and Newton\cite{HeloiseBeautification} (Hs-DB).
-The data are drawn by 16 users each of them was asked to draw 13 shapes from 30 to 50 times.
-Figure \ref{fig:symbolSet} shows a set of the shapes used in the data set.
-We collected the second and third datasets.
-Figure \ref{fig:ELsymbolSet} and \ref{fig:LogicsymbolSet} shows the symbols gathered from 7 different users in electrical (EL-DB) and logic design (LD-DB) domain respectively.
-Table \ref{tab:datasets} shows a  comparison between the three datasets used in our experiments.
-The datasets were divided into training set and test set.
-Five different splits  for the training and test data are generated from each dataset.
-The results displayed are the average recognition accuracy of the five splits.
-The accuracy is computed as the number of correctly recognized samples divided by the total number of samples in the test.
-
- %The goal of the system is to modify the segmentation and recognition of the sketch system.
-It made the same type of experiement done in \cite{mypaper}.
-The dataset is divided into 13 category of goemetric %symbols
-
-% The dataset collected by Hse and Newton\cite{HeloiseBeautification} is used to test the presented system.
-Figure \ref{fig:symbolSet} shows a set of the shapes used in the data set.
-The datasets were divided into training and test sets.
-Five different splits for the training and test data are generated from each dataset.
-The results displayed are the average recognition accuracy of the five splits.
-The accuracy is computed as the number of correctly recognized samples divided by the total number of samples in the test.
     \begin{figure}
@@
   \caption[Recognition accuracy]%
-  {The chart compares between recognition rate of the full feature set on both the orginal system presented in \cite{mypaper} and Enhanced PSO.
-The results shows that Enhanced PSO increases the performance over the orginal system.
+  {The chart compares between recognition rate of the full feature set on both the original system presented in \cite{mypaper} and Enhanced PSO.
+The results shows that Enhanced PSO increases the performance over the original system.
 }
@@
   \caption[Recognition accuracy]%
-  {The chart compares between recognition rate of half the feature set on both the orginal system presented in \cite{mypaper} and Enhanced PSO.
+  {The chart compares between recognition rate of half the feature set on both the original system presented in \cite{mypaper} and Enhanced PSO.
 The results shows that Enhanced PSO increase the recognition rate.
@@
 
-% We performed two experiments to test the system.
-Firstly, we tested recognition accuracy of shapes in the data set with both algorithms (\textsl{AlgS1} and \textsl{AlgS2}).
-We also implemented the segmentation algorithm described in \cite{earlyprocess} (\textsl{Alg3}) to compare it with our swarm algorithms.
-Figure \ref{fig:test1} shows the accuracy achieved by each algorithm.
-The result shows that (\textsl{AlgS2}) gives better performance than Alg3 and AlgS1.
-
 \section{Conclusion }
  \label{ConclusionandFutureWork}
-
-%The training and testing performance measurements using both PSO and Enhanced-PSO in training and testing phases are shown in figure\ref{exp2} and figure \ref{exp1}.
 
@@
 Introducing more spatial and geometrical features is believed to improve classifications.
-%this new enhanced architecture will be used in other different applications of time series prediction and also extended to be used in other applications rather than time series prediction such as pattern recognition and speech verification
-
- %A possible extension of this research is to complete the clustering algorithm for fully automated sketch recognition.
-Currently, the system process one symbol at a time, an enchantment would be to allow users to draw more than one symbol at a time and incorporate a method for separating symbols.
-Other area of enhancements is the features extraction methods.
-Introducing more spatial and geometrical features is believed to improve classifications.
 

### 2010-09-24 MySimpleSketch/src/SketchMaster/Stroke/StrokeData/DominatePointStructure.java
This is saving this version into a new tag to be reterifed latter before radwas comments

@@ -1,121 +1,0 @@
-/**
- * 
- */
-package SketchMaster.Stroke.StrokeData;
-import java.util.ArrayList;
-
-/**
- * @author maha
- * 
- */
-public class DominatePointStructure {
-	PointData point = null;
-
-	int indexInInk = 0;
-
-	ArrayList<Integer> InFunctions = null;
-
-	int FunctionCount = 0;
-
-	double Certainty = 0.0;
-
-	public void addFunction(int f) {
-		if (InFunctions == null)
-			InFunctions = new ArrayList<Integer>();
-
-		InFunctions.add(Integer.valueOf(f));
-		FunctionCount++;
-	}
-
-	public double getCertainty() {
-		return Certainty;
-	}
-
-	public void setCertainty(double certainty) {
-		Certainty = certainty;
-	}
-
-	public int getIndexInInk() {
-		return indexInInk;
-	}
-
-	public void setIndexInInk(int indexInInk) {
-		this.indexInInk = indexInInk;
-	}
-
-	public PointData getPoint() {
-		return point;
-	}
-
-	public void setPoint(PointData point) {
-		this.point = point;
-	}
-
-	public int getFunctionCount() {
-		return FunctionCount;
-	}
-
-	public ArrayList<Integer> getInFunctions() {
-		return InFunctions;
-	}
-
-	/* (non-Javadoc)
-	 * @see java.lang.Object#hashCode()
-	 */
-	@Override
-	public int hashCode() {
-		final int prime = 31;
-		int result = 1;
-		long temp;
-		temp = Double.doubleToLongBits(Certainty);
-		result = prime * result + (int) (temp ^ (temp >>> 32));
-		result = prime * result + FunctionCount;
-		result = prime * result
-				+ ((InFunctions == null) ? 0 : InFunctions.hashCode());
-		result = prime * result + indexInInk;
-		result = prime * result + ((point == null) ? 0 : point.hashCode());
-		return result;
-	}
-
-	/* (non-Javadoc)
-	 * @see java.lang.Object#equals(java.lang.Object)
-	 */
-	@Override
-	public boolean equals(Object obj) {
-		if (this == obj)
-			return true;
-		if (obj == null)
-			return false;
-		if (getClass() != obj.getClass())
-			return false;
-		final DominatePointStructure other = (DominatePointStructure) obj;
-		if (Double.doubleToLongBits(Certainty) != Double
-				.doubleToLongBits(other.Certainty))
-			return false;
[…]

### 2010-09-24 MySimpleSketch/src/SketchMaster/Stroke/StrokeData/InkInterface.java

@@ -1,55 +1,0 @@
-/**
- * 
- */
-package SketchMaster.Stroke.StrokeData;
-
-import java.awt.Graphics2D;
-import java.awt.Rectangle;
-import java.io.Serializable;
-import java.lang.reflect.Array;
-import java.util.ArrayList;
-
-import SketchMaster.Stroke.graphics.shapes.GuiShape;
-import SketchMaster.Stroke.graphics.shapes.Line;
-
-/**
- * @author maha
- * 
- */
-public interface InkInterface extends GuiShape {
-
-	// private Rectangle box = null;
-
-	/**
-	 * @return Returns the points.
-	 * @uml.property name="points"
-	 */
-	public ArrayList<PointData> getPoints();
-   
-	public int getPointsCount();
-
-	public void addPoint(PointData point);
-
-	public void deletePoint(int index);
-
-	public void setPoints(ArrayList<PointData> po);
-
-	public PointData getPoint(int index);
-
-	public void setPoints(int index, PointData point);
-
-	public InkInterface createSubInkObject(int start, int end);
-
-	public boolean canIntersect(InkInterface end);
-	
-	public  ArrayList<PointData> IntersectionPoints(Line l);
-	
-	public 	 ArrayList<Line> toLines();
-	
-	public 	 ArrayList< InkInterface > divideDirection();
-	
-	
-
-	//public double getBox();
-
-}

### 2010-09-24 MySimpleSketch/src/SketchMaster/Stroke/StrokeData/InkPart.java

@@ -1,326 +1,0 @@
-/**
- * 
- */
-package SketchMaster.Stroke.StrokeData;
-
-import java.awt.Graphics2D;
-import java.awt.geom.Rectangle2D;
-import java.awt.geom.Rectangle2D.Double;
-import java.util.ArrayList;
-import java.util.Collections;
-import java.util.Comparator;
-
-import org.apache.log4j.Logger;
- 
-
-import SketchMaster.Stroke.graphics.shapes.FittedShape;
-import SketchMaster.Stroke.graphics.shapes.Line;
-import SketchMaster.lib.ComputationsGeometry;
-import SketchMaster.lib.CurveFitData;
-import SketchMaster.system.SystemSettings;
-
-/**
- * @author maha
- *
- */
-public class InkPart implements InkInterface {
-	
-	
-	private static final Logger logger = Logger.getLogger(InkPart.class);
-	private static final double DivideStrokePercent = SystemSettings.DivideStrokePercent;
-	int start, end;
-	boolean OnlyPart;
-	CurveFitData sums;
-	private boolean sumsComputed=false;
-	private Rectangle2D box;
-	private boolean box_valid=false; 
-	double length;
-	boolean lengthComputed=false;
-	double revolution;
-	protected ArrayList<PointData> points = null;
-	protected boolean closed=false,open=true;
-	private double rotation;
-	boolean rotationComputed=false;
-	
-	/* (non-Javadoc)
-	 * @see SketchMaster.Stroke.StrokeData.InkInterface#addPoint(SketchMaster.Stroke.StrokeData.PointData)
-	 */
-	public void addPoint(PointData point) {
-	 
-	}
-
-	/**
-	 * @return the start
-	 */
-	public int getStart() {
-		return start;
-	}
-
-	/**
-	 * @param start the start to set
-	 */
-	public void setStart(int start) {
-		this.start = start;
-	}
-
-	/**
-	 * @return the end
-	 */
-	public int getEnd() {
-		return end;
-	}
-
-	/**
-	 * @param end the end to set
-	 */
-	public void setEnd(int end) {
-		this.end = end;
-	}
-
-	/**
-	 * @return the sums
-	 */
-	public CurveFitData getSums() {
-		return sums;
-	}
-
-	/* (non-Javadoc)
-	 * @see SketchMaster.Stroke.StrokeData.InkInterface#canIntersect(SketchMaster.Stroke.StrokeData.InkInterface)
-	 */
[…]

### 2010-09-24 MySimpleSketch/src/SketchMaster/Stroke/StrokeData/PointData.java

@@ -1,613 +1,0 @@
-package SketchMaster.Stroke.StrokeData;
-
-import org.apache.log4j.Logger;
-
-import java.awt.Graphics;
-import java.awt.Graphics2D;
-import java.awt.Point;
-import java.awt.Rectangle;
-import java.awt.geom.Point2D;
-import java.awt.geom.Rectangle2D;
-import java.awt.geom.Point2D.Double;
-
-import java.io.IOException;
-import java.io.ObjectInputStream;
-import java.io.ObjectOutputStream;
-import java.io.Serializable;
-import java.util.ArrayList;
-
-
-
-import SketchMaster.Stroke.graphics.shapes.GeometricObject;
-import SketchMaster.Stroke.graphics.shapes.PolygonShape;
-//import SketchMaster.Stroke.graphics.shapes.Polygon;
-import SketchMaster.swarm.polygonApproximations.PolygonSolution;
-
-/**
- * @author Mahi
- */
-public class PointData extends Point2D.Double implements Serializable,GeometricObject, Comparable<PointData> {
-	/**
-	 * Logger for this class
-	 */
-	private static final Logger logger = Logger.getLogger(PointData.class);
-
-	public PointData() {
-		super();
-        magComputed=false;
-	}
-
-	public PointData(double arg0, double arg1) {
-		super(arg0, arg1);
-        magComputed=false;
-	}
-	
-	public PointData(double x,double y , long t, double p){
-		this(x,y);
-		this.setTime(t);
-		this.setPresureValue(p);
-		
-		
-		
-	}
-
-	public PointData(Point2D f) {
-		super(f.getX(),f.getY());
-        magComputed=false;
-	}
-
-	@Override
-	public Object clone() {
-		PointData temp = null;
-		temp = (PointData) super.clone();
-		temp.time = time;
-		temp.startPoint = this.startPoint;
-		temp.endPoint = this.endPoint;
-		temp.CompulativeLength = CompulativeLength;
-		temp.presureValue = this.presureValue;
-		temp.controlPoint=this.controlPoint;
-		temp.magComputed=false;
-
-		// tempSolution.polygonVertices=(ArrayList<Point2D>)this.polygonVertices.clone();
-		return temp;
-
-	}
-	
-	private void writeObject(ObjectOutputStream os) throws IOException {
-		try {
-			os.defaultWriteObject();
-		
-			
-//			os.writeBoolean(this.endPoint);
-//			os.writeBoolean( this.startPoint);
-//			os.writeBoolean(this.controlPoint);
[…]

### 2010-09-24 MySimpleSketch/src/SketchMaster/Stroke/StrokeData/Segment.java

@@ -1,95 +1,0 @@
-package SketchMaster.Stroke.StrokeData;
-import java.awt.Graphics;
-import java.awt.Graphics2D;
-import java.io.Serializable;
-import java.util.ArrayList;
-
-import SketchMaster.Stroke.graphics.shapes.GeometricPrimitive;
-import SketchMaster.Stroke.graphics.shapes.GuiShape;
-
-public class Segment extends SimpleInkObject implements Serializable {
-	public Segment() {
-		super();
-		
-	}
-
-	/**
-	 * 
-	 */
-	private static final long serialVersionUID = 8230908701712079078L;
-	GeometricPrimitive SegmentType;
-	private PointData StartPoint = null;
-
-	private PointData EndPoint = null;
-
-	// public Stroke getStroke(){
-	//	
-	// return stroke;
-	// }
-	//
-	// public void setStroke(Stroke stroke){
-	// this.stroke= stroke;
-	// }
-
-	public Segment(InkInterface object) {
-		this.setPoints(object.getPoints());
-
-	}
-
-	public void setParam(ArrayList Param) {
-		// 
-
-	}
-
-	// private Stroke stroke; //orginating stroke
-
-	public void paint(Graphics2D g) {
-		// 
-		this.getSegmentType().paint(g);
-	}
-
-	public void setEndPoint(PointData endPoint) {
-		EndPoint = endPoint;
-	}
-
-	public void setSegmentType(GeometricPrimitive segmentType) {
-		SegmentType = segmentType;
-	}
-
-	public void setStartPoint(PointData startPoint) {
-		StartPoint = startPoint;
-	}
-
-	@Override
-	public InkInterface createSubInkObject(int start, int end) {
-		Segment ink = new Segment();
-
-		ArrayList<PointData> po = new ArrayList<PointData>();
-		if (this.points != null) {
-			for (int i = start; (i < this.points.size()) && (i < end); i++) {
-				po.add(this.points.get(i));
-			}
-
-		}
-		ink.setSegmentType(this.SegmentType);
-		ink.setPoints(po);
-		// return a segment from this stroke that will contain the points of the
-		// stroke.
-		return ink;
-		// return null;
-	}
-
-	public GeometricPrimitive getSegmentType() {
-		return SegmentType;
-	}
-
-	/* (non-Javadoc)
-	 * @see java.lang.Object#toString()
-	 */
-	@Override
-	public String toString() {
-	
-		return "Segment  "+this.SegmentType.getClass().getSimpleName();
-	}
-
-}

### 2010-09-24 MySimpleSketch/src/SketchMaster/Stroke/StrokeData/SegmentCluster.java

@@ -1,481 +1,0 @@
-/**
- * 
- */
-package SketchMaster.Stroke.StrokeData;
-
-import org.apache.log4j.Level;
-import org.apache.log4j.Logger;
-
-import java.awt.BasicStroke;
-import java.awt.Color;
-import java.awt.Graphics2D;
-import java.awt.geom.Rectangle2D;
-import java.awt.geom.Rectangle2D.Double;
-import java.io.IOException;
-import java.io.ObjectInputStream;
-import java.io.ObjectOutputStream;
-import java.io.Serializable;
-import java.util.ArrayList;
-import java.util.Iterator;
-
-import SketchMaster.Stroke.features.SegmentClusterFeatureSet;
-import SketchMaster.Stroke.graphics.shapes.GuiShape;
-import SketchMaster.Stroke.graphics.shapes.Line;
-import SketchMaster.Stroke.graphics.shapes.SegmentedShape;
-import SketchMaster.gui.DrawingDebugUtils;
-import SketchMaster.system.SystemSettings;
-
-/**
- * @author maha a coollection of segment that will be used to get a sembol
- */
-public class SegmentCluster extends ArrayList<Segment> implements InkInterface,
-		Serializable {
-	/**
-	 * Logger for this class
-	 */
-	private static final Logger logger = Logger.getLogger(SegmentCluster.class);
-	private static final  Logger logE=Logger.getLogger("ExampleLogging");;
-	private Rectangle2D box;
-	private boolean box_valid=false; 
-	// it must compute features
-
-	// the full ink path of the colleection
-	transient InkInterface FullinkPath;
-
-	// strokes of a symbol
-	ArrayList<Stroke> strokeInSymbol = new ArrayList<Stroke>();
-
-	int strokeCount = 0; // number of strokes consist the segmets in the
-							// cluster
-	int segmentCount = 0;
-	double area=0.0;
-	transient SegmentClusterFeatureSet FeatureSet = new SegmentClusterFeatureSet();
-
-	private String curString;
-
-	
-
-	/**
-	 * 
-	 */
-	private static final long serialVersionUID = 7346329178510124243L;
-
-	public static final int DRAW_RAW_INK = 0;
-
-	public static final int DRAW_CLEAN_SEGMENTS = 1;
-	public static final int	DRAW_BOTH_RAW_SEGMENT=2;
[…]

### 2010-09-24 MySimpleSketch/src/SketchMaster/Stroke/StrokeData/SimpleInkObject.java

@@ -1,391 +1,0 @@
-/**
- * 
- */
-package SketchMaster.Stroke.StrokeData;
-
- 
-
-import org.apache.log4j.Logger;
-
-import java.awt.Graphics2D;
-import java.awt.geom.AffineTransform;
-import java.awt.geom.Point2D;
-import java.awt.geom.Rectangle2D;
-import java.awt.geom.Rectangle2D.Double;
-import java.util.ArrayList;
-import java.util.Collections;
-import java.io.Serializable;
-import java.lang.reflect.Array;
-
-import SketchMaster.Stroke.graphics.shapes.GuiShape;
-import SketchMaster.Stroke.graphics.shapes.Line;
-import SketchMaster.lib.ComputationsGeometry;
-import SketchMaster.lib.CurveFitData;
-import SketchMaster.system.SystemSettings;
-
-/**
- * @author melmeseery
- * 
- */
-public class SimpleInkObject implements Serializable, InkInterface {
-	/**
-	 * Logger for this class
-	 */
-	
-	private static final double DivideStrokePercent = SystemSettings.DivideStrokePercent;
-	private static final Logger logger = Logger.getLogger(SimpleInkObject.class);
-
-	/**
-	 * @uml.property name="points"
-	 */
-	protected ArrayList<PointData> points = null;
-	protected boolean closed=false,open=true;
-	private Rectangle2D box;
-	private boolean box_valid=false; 
-	double length;
-	boolean lengthComputed=false;
-	private boolean sumsComputed=false;
-	 CurveFitData  sums;
-		private double rotation;
-		boolean rotationComputed=false;
-		double revolution;
-	/**
-	 * @return the closed
-	 */
-	public boolean isClosed() {
-		return closed;
-	}
-
-	/**
-	 * @param closed the closed to set
-	 */
-	public void setClosed(boolean closed) {
-		this.closed = closed;
-	}
-
-	/**
-	 * @return the open
-	 */
-	public boolean isOpen() {
-		return open;
-	}
-
-	/**
-	 * @param open the open to set
-	 */
-	public void setOpen(boolean open) {
-		this.open = open;
-	}
-
-	/**
-	 * 
-	 */
-	public SimpleInkObject() {
-	
-	 points=new ArrayList<PointData>();
-	}
-
-	public ArrayList<PointData> getPoints() {
-		return points;
-	}
-
-	/**
-	 * getPointsCount
-	 */
-	public int getPointsCount() {
-		int returnint = points.size();
-		return returnint;
-	}
-
[…]

### 2010-09-24 MySimpleSketch/src/SketchMaster/Stroke/StrokeData/Stroke.java

@@ -1,1811 +1,0 @@
-package SketchMaster.Stroke.StrokeData;
-
-import org.apache.log4j.Logger;
-
-import java.awt.Color;
-import java.awt.Graphics;
-import java.awt.Graphics2D;
-import java.awt.Point;
-import java.awt.geom.AffineTransform;
-import java.awt.geom.Point2D;
-import java.awt.geom.Rectangle2D;
-import java.io.Serializable;
-import java.util.ArrayList;
-
-
-
-
-import SketchMaster.system.SystemSettings;
-import SketchMaster.Stroke.features.FeatureFunction;
-import SketchMaster.Stroke.graphics.shapes.GuiShape;
-import SketchMaster.Stroke.graphics.shapes.Line;
-import SketchMaster.Stroke.graphics.shapes.SegmentedShape;
-import SketchMaster.gui.DrawingDebugUtils;
-import SketchMaster.lib.ComputationsGeometry;
-import SketchMaster.system.SystemSettings;
-
-/**
- * @author Mahi
- */
-public class Stroke extends SimpleInkObject implements Serializable, GuiShape {
-	
-//	class pointChange{
-//		final static int  TYPE_INC=0;
-//		 final static  	int TYPE_DEC=1;
-//		 final static 	int TYPE_ULTERNATING=2;
-//		 final static int 	TURN_X=0;
-//		 final static int 	TURN_y=1;
-//		 final static int 	TURN_M=2;
-//		 
-//
-//		  final static int 	TURN_YM=3;
-//		 final static int 	TURN_XM=4;
-//	
-//	   final static int 	TURN_XY=5; 
-//	   
-//		 final static int 	TURN_ALL=6;
-//		 
-//		double inX;
-//		double inY;
-//		double inMag;
-//		
-//		int typeX;
-//		int typeY;
-//		int typeMag;
-//		int turnType=0;
-//		int index;
-//		boolean turn=false;
-//		public void add(double x, double y, double mag){
-//			inX=x;
-//			if (x>0){
-//				typeX=TYPE_INC;
-//			}
-//			else {
-//				typeX=TYPE_DEC;
-//			}
-//			
-//			
-//			
-//			
-//			
-//			inY=y;
-//			if (y>0){
-//				typeY=TYPE_INC;
-//			}
-//			else {
-//				typeY=TYPE_DEC;
-//			}
-//			
-//			inMag=mag;
-//			if (mag>0){
[…]

### 2010-09-24 MySimpleSketch/src/SketchMaster/Stroke/features/FeatureFunction.java

@@ -1,781 +1,0 @@
-/**
- *
- * 
- * 
- * 
- */
-package SketchMaster.Stroke.features;
-
-import org.apache.log4j.Logger;
-
-import java.awt.Point;
-import java.awt.geom.Point2D;
-import java.io.Serializable;
-import java.util.ArrayList;
-
-import SketchMaster.system.SystemSettings;
-import SketchMaster.Stroke.StrokeData.InkInterface;
-import SketchMaster.Stroke.StrokeData.PointData;
-import SketchMaster.Stroke.StrokeData.Stroke;
-
-/**
- * @author Mahi
- */
-public class FeatureFunction implements Serializable {
-	/**
-	 * 
-	 */
-	private static final long serialVersionUID = 8911153265345792426L;
-
-	/**
-	 * Logger for this class
-	 */
-	private static final Logger logger = Logger.getLogger(FeatureFunction.class);
-
-	private StrokeFeatures func = null;
-	private ArrayList<PointData> data = null;
-	private static double ZERO = 10e-3;
-	private double max = Double.MIN_VALUE;
-	private double min = Double.MAX_VALUE;
-	private String Name = "";
-	private String xName = " X - axis ";
-	private String yName = " Y - axis ";
-	private int maxi = 0;
-	private int mini = 0;
-	private double maxX = 0.0;
-	private double minX = 0.0;
-	private double average = 0.0;
-	private double absaverage = 0.0;
-	private ArrayList pointIndex = null;
-	private double sumAbs = 0.0;
-	private double sumUpNow = 0.0;
-	private double DataThreshold = 0.0;
-	private int functionType = ONLINE_TYPE;
-	public static final int BATCH_TYPE = 0;
-	public static final int ONLINE_TYPE = 1;
-
-	private   transient ArrayList<Point2D> ImportantList = new ArrayList<Point2D>();
-
-	/**
-	 * @return the data
-	 * @uml.property name="data"
-	 */
-	public ArrayList<PointData> getData() {
-
-		return data;
-	}
-
-	public StrokeFeatures getFunc() {
-		return func;
-	}
-
-	/**
-	 * @param func
-	 *            the func to set
-	 * @uml.property name="func"
-	 */
-	public void setFunc(StrokeFeatures func) {
-		this.func = func;
-	}
- 
-	@Deprecated
-	public void updateFunctionWithPoint(PointData point, InkInterface stroke) {
-		if (func != null) {
-			ArrayList points = stroke.getPoints();
-			if ((points.size() == 1) || data == null) {
[…]

### 2010-09-24 MySimpleSketch/src/SketchMaster/Stroke/features/SVMFeatureSet.java

@@ -1,119 +1,0 @@
-/**
- * 
- */
-package SketchMaster.Stroke.features;
-
-import org.apache.log4j.Logger;
-
-import SketchMaster.classifier.svmTrainable.SVMClassifier;
-
-/**
- * @author Maha
- *
- */
-public class SVMFeatureSet {
-	private static transient final Logger logger = Logger.getLogger(SVMFeatureSet.class);
-	private static  int numFeatures=0;
-	  /** The feature vector. */
-    private double _features[]=null;
-    private String featuresName[]=null;
-	/**
-	 * 
-	 */
-	
-	public SVMFeatureSet(int count){
-		initFeature(count);
-	}
-	
-public SVMFeatureSet(double[] vals) {
-	 _features=vals.clone();
-	}
-
-public SVMFeatureSet(double[] vals,String[] names) {
-
-	 _features=vals.clone();
-	 featuresName=new String[names.length];
-	 for (int i = 0; i < names.length; i++) {
-		 featuresName[i]=names[i];
-	}
-	 
-	//featuresName=names.clone();
-	 
-//		for (int i = 0; i < vals.length; i++) {
-//			logger.info("   features "+featuresName[i]+"  = "+ _features[i]);
-//		}
-//		logger.info("--------------------------------------Features------------------------------------");
-	}
-public SVMFeatureSet(SVMFeatureSet s) {
-	 _features=s._features.clone();
-	 featuresName=new String[s.featuresName.length];
-	 for (int i = 0; i < s.featuresName.length; i++) {
-		 featuresName[i]=s.featuresName[i];
-	}
-}
-
-public static int getFeatureMaxCount(){
-	return numFeatures;
-}
-public int getFeatureCount() {// this feature set with no zero values 
-
-		return  _features.length;
-	}
-
-public int getFeatureNoZeroCount(){
-	
-	int count=0;
-	for (int i = 0; i < _features.length; i++) {
-		if (_features[i]!=0)
-			count++;
-	}
-	return count;
-}
-
-public double getFeature(int i) { // get the value of feature of value i 
-	 return _features[i];
-	//return 0;
-}
-public String getFeatureName(int i) {
-	
-	if (featuresName!=null)
-		return featuresName[i];
-	return "";
-}
-
-public void setFeature(int i, double val, String name){
-	_features[i]=val;
-	featuresName[i]=name;
-}
-public void setFeature(int i, double val){
-	_features[i]=val;
-}
-
[…]

### 2010-09-24 MySimpleSketch/src/SketchMaster/Stroke/features/SegmentClusterFeature.java

@@ -1,142 +1,0 @@
-/**
- * 
- */
-package SketchMaster.Stroke.features;
-/**
- * @author maha
- * 
- */
-public abstract class SegmentClusterFeature {
-	static double ZERO_REPLACMENT=-1;
-	protected double Value = 0.0;
-	protected double[] Values;
-	protected boolean ValueOk = false;
-
-	protected boolean computed = false;
-
-	protected double minValue = 0.0;
-
-	protected double maxValue = Double.MAX_VALUE;
-
-	protected String Name="";
-	protected String[] Names=null;
-
-	public double[] getValues() {
-		if (!computed) {
-			computed = true;
-			double[] returndoubleArray = Values();
-			for (int i = 0; i < returndoubleArray.length; i++) {
-				if (returndoubleArray[i]==0||Double.isInfinite(returndoubleArray[i])||Double.isNaN(returndoubleArray[i]))
-				{
-					returndoubleArray[i]=ZERO_REPLACMENT;
-				}
-			}
-			return returndoubleArray;
-			
-			//return returndoubleArray;
-		} else {
-			
-			for (int i = 0; i < Values.length; i++) {
-				if (Values[i]==0||Double.isInfinite(Values[i])||Double.isNaN(Values[i]))
-				{
-					Values[i]=ZERO_REPLACMENT;
-				}
-			}
-			return Values;
-		}
-		// return this.MinValue();
-	}
-
-//	public double getValue() {
-//		if (!computed) {
-//			computed = true;
-//			double returndouble = Value();
-//			return returndouble;
-//		} else {
-//			if (Value==0||Double.isInfinite(Value)||Double.isNaN(Value))
-//				
-//				Value=ZERO_REPLACMENT;
-//			return Value;
-//		}
-//		// return this.MinValue();
-//	}
-
-//	abstract public double Value();
-
-	abstract public double[] Values();
-
-	abstract void setData(Object data);
-
-	public boolean isValueOk() {
-		return ValueOk;
-	}
-
-	public String toString() {
-	
-			
-		
-		StringBuilder string =new StringBuilder();
-		string.append( "  Function " );
-		string.append(Name ); 
-		string.append("  and Value = "); 
-		string.append(Value);
-		string.append( " with Status " );
-		string.append(ValueOk );
-		string.append( "  ");
-		
-		if (NoOfValues()>1){
-			//string =  "  Function " + Name + "  and Value = "   ;
-			string =new StringBuilder();
[…]

### 2010-09-24 MySimpleSketch/src/SketchMaster/Stroke/features/SegmentClusterFeatureSet.java

@@ -1,2370 +1,0 @@
-/**
- * 
- */
-package SketchMaster.Stroke.features;
-import java.awt.Point;
-import java.awt.Rectangle;
-import java.awt.geom.Point2D;
-import java.awt.geom.Rectangle2D;
-import java.io.Serializable;
-import java.util.ArrayList;
-
-import org.apache.log4j.Logger;
-
-import SketchMaster.Stroke.StrokeData.InkInterface;
-import SketchMaster.Stroke.StrokeData.PointData;
-import SketchMaster.Stroke.StrokeData.Segment;
-import SketchMaster.Stroke.StrokeData.SegmentCluster;
-import SketchMaster.Stroke.StrokeData.SimpleInkObject;
-import SketchMaster.Stroke.StrokeData.Stroke;
-import SketchMaster.Stroke.graphics.shapes.BezireCurve;
-import SketchMaster.Stroke.graphics.shapes.Circle;
-import SketchMaster.Stroke.graphics.shapes.Ellipse;
-import SketchMaster.Stroke.graphics.shapes.GeometricPrimitive;
-import SketchMaster.Stroke.graphics.shapes.Line;
-import SketchMaster.lib.ComputationsGeometry;
-import SketchMaster.lib.ZernikeFE;
-import SketchMaster.system.SketchSheet;
-import SketchMaster.system.SystemSettings;
-
-/**
- * @author maha
- * 
- */
-public class SegmentClusterFeatureSet implements InkFeatureSet, Serializable {
-	/**
-	 * 
-	 */
-	private static final long serialVersionUID = -5502467430484342506L;
-	transient ArrayList<SegmentClusterFeature> Features = null;
-	private static final  Logger logE=Logger.getLogger("ExampleLogging");;
-	private static final Logger logger = Logger.getLogger(SegmentClusterFeatureSet.class);
-	SegmentCluster inkCluster;
-	static double ZERO_REPLACMENT=-1;
-
-	/***************************************************************************
-	 * Features 1. number of primitives ---------- done 2. number of lines ,
-	 * ---------------done (number of strokes ) 3. number of curves - arcs ,
-	 * -----------done 4. number of ellipses ----------------done 5. number of
-	 * parralles lines. ----------------50 % done 6. number of perpendicular
-	 * lines ----------------50 % done 7. number of intersecitons at t , x , l
[…]

### 2010-09-24 MySimpleSketch/src/SketchMaster/Stroke/features/StrokeFeatures.java

@@ -1,58 +1,0 @@
-/**
- * 
- */
-package SketchMaster.Stroke.features;
-import java.awt.geom.Point2D;
-import java.io.Serializable;
-
-import SketchMaster.Stroke.StrokeData.PointData;
-import SketchMaster.Stroke.StrokeData.Segment;
-
-/**
- * @author Mahi
- */
-public abstract class StrokeFeatures implements Serializable {
-	public static final int THERSHOLD_AVERAGE = 0;
-	public static final int THERSHOLD_ABS_AVERAGE = 1;
-	public static final int THERSHOLD_STATIC = 2;
-	public static final int THERSHOLD_CUSTOM = 3;
-	public static final int THERSHOLD_NO = -1;
-	protected PointData prev;
-	protected PointData next;
-	protected PointData current;
-	protected String Name;
-
-	abstract public Point2D f(PointData p);
-
-	abstract public Point2D f();
-
-	abstract public boolean isInFunction(double y, double Threshold);
-
-	abstract public int getThresholdType();
-
-	protected double minValue = 0.0;
-
-	public double MinValue() {
-		return minValue;
-	}
-
-	protected double maxValue = Double.MAX_VALUE;
-
-	public double MaxValue() {
-		return maxValue;
-	}
-
-	public double ScaledFunction(double factor) {
-		double returndouble = f().getY() * factor;
-		return returndouble;
-	}
-
-	public void setName(String name) {
-		Name = name;
-	}
-
-	public String getName() {
-		return Name;
-	}
-
-}

### 2010-09-24 MySimpleSketch/src/SketchMaster/Stroke/features/StrokeRubineFeatureSet.java

@@ -1,970 +1,0 @@
-/**
- * 
- */
-package SketchMaster.Stroke.features;
-import java.awt.Point;
-import java.awt.Rectangle;
-import java.awt.geom.Point2D;
-import java.io.Serializable;
-import java.util.ArrayList;
-
-import org.apache.log4j.Logger;
-
-import SketchMaster.Stroke.StrokeData.InkInterface;
-import SketchMaster.Stroke.StrokeData.PointData;
-import SketchMaster.Stroke.StrokeData.SimpleInkObject;
-import SketchMaster.Stroke.StrokeData.Stroke;
-import SketchMaster.Stroke.StrokeData.StrokeStatisticalData;
-import SketchMaster.classifier.Symbol.SymbolCategory;
-import SketchMaster.lib.ComputationsGeometry;
-import SketchMaster.system.SystemSettings;
-
-/**
- * @author maha
- * 
- */
-public class StrokeRubineFeatureSet  implements
-		Serializable, InkFeatureSet {
-	
-	private static final Logger logger = Logger.getLogger(StrokeRubineFeatureSet.class);
-	/**
-	 * 
-	 */
-	private static final long serialVersionUID = 566823144662422035L;
-         SimpleInkObject stroke;
-	/**
-	 * @return the stroke
-	 */
-	public InkInterface getStroke() {
-		return stroke;
-	}
-
-	public InkInterface getInk() {
-		return stroke;
-	}
-
-	transient ArrayList<StrokeStatisticalFeatures> Features = null;
-
-	/**
-	 * 
-	 */
-	public StrokeRubineFeatureSet() {
-		super();
-	}
-
-	/**
-	 * 
-	 */
-	public StrokeRubineFeatureSet(InkInterface ink) {
-		 this.stroke=(SimpleInkObject) ink;
-		//super(ink);
-	}
-
-
-	public void initAll() {
-		Features = new ArrayList<StrokeStatisticalFeatures>();
-
-		// init the f1
-		StrokeStatisticalFeatures temp = new cosineStartingPoint();
-		temp.setSubStroke(stroke);
-		Features.add(temp);
-
-		// f2
-		temp = new sineStartingPoint();
-		temp.setSubStroke(stroke);
-		Features.add(temp);
-		// f3
-		temp = new lengthDiagonal();
-		temp.setSubStroke(stroke);
-		Features.add(temp);
-		// f4
-		temp = new AngleOfDiagonal();
[…]

### 2010-09-24 MySimpleSketch/src/SketchMaster/Stroke/features/StrokeStatisticalFeatures.java

@@ -1,126 +1,0 @@
-/**
- * 
- */
-package SketchMaster.Stroke.features;
-import java.awt.geom.Point2D;
-import java.io.Serializable;
-
-import SketchMaster.Stroke.StrokeData.InkInterface;
-import SketchMaster.Stroke.StrokeData.PointData;
-import SketchMaster.Stroke.StrokeData.Segment;
-
-/**
- * @author maha
- * 
- */
-public abstract class StrokeStatisticalFeatures extends StrokeFeatures {
-	@Override
-	public Point2D f() {
-
-		return null;
-	}
-
-	@Override
-	public int getThresholdType() {
-
-		return 0;
-	}
-
-	@Override
-	public boolean isInFunction(double y, double Threshold) {
-		return false;
-	}
-
-	@Override
-	public String toString() {
-		String string = "Function " + Name + "  and Value = " + Value
-				+ " with Status " + ValueOk;
-
-		return string;
-	}
-
-	protected InkInterface subStroke = null;
-
-	protected boolean ValueOk = false;
-	protected boolean computed = false;
-	protected Object data;
-	protected double Value = 0.0;
-
-	public double getValue() {
-		if (!computed) {
-			computed = true;
-			double returndouble = Value();
-			return returndouble;
-		} else {
-			return Value;
-		}
-		// return this.MinValue();
-	}
-
-	/*
-	 * (non-Javadoc)
-	 * 
-	 * @see SketchMaster.Stroke.features.StrokeFeatures#f(SketchMaster.Stroke.StrokeData.PointData)
-	 */
-	@Override
-	public Point2D f(PointData p) {
-		return null;
-	}
-
-	/*
-	 * (non-Javadoc)
-	 * 
-	 * @see SketchMaster.Stroke.features.StrokeFeatures#f()
-	 */
-	// @Override
-	// public Point2D f() {
-	//		
-	// return null;
-	// }
-	abstract public double Value();
-
-	// {
-	// return 0.0;
-	// }
-	/*
-	 * (non-Javadoc)
-	 * 
-	 * @see SketchMaster.Stroke.features.StrokeFeatures#getThresholdType()
-	 */
-	// @Override
-	// public int getThresholdType() {
-	//		
-	// return 0;
-	// }
-	/*
-	 * (non-Javadoc)
-	 * 
-	 * @see SketchMaster.Stroke.features.StrokeFeatures#isInFunction(double,
-	 *      double)
-	 */
[…]

### 2010-09-24 MySimpleSketch/src/SketchMaster/Stroke/graphics/layers/SketchLayer.java

@@ -1,81 +1,0 @@
-/**
- * 
- */
-package SketchMaster.Stroke.graphics.layers;
-import java.awt.Graphics2D;
-import java.util.ArrayList;
-import java.util.Observable;
-import java.util.Observer;
-
-import SketchMaster.Stroke.graphics.shapes.GuiShape;
-import SketchMaster.gui.GraphWatched;
-
-/**
- * @author Mahi
- * 
- */
-public class SketchLayer implements Observer {
-	String LayerName = "layer ";
-
-	public void addObject(Object object) {
-		if (object instanceof GuiShape) {
-			GuiShape NewGuiShape = (GuiShape) object;
-			GuiItems.add(NewGuiShape);
-		}
-	}
-
-	/**
-	 * 
-	 */
-
-	protected ArrayList<GuiShape> GuiItems = new ArrayList<GuiShape>();
-
-	// public SketchLayer() {
-	// super();
-
-	// }
-	public void paint(Graphics2D arg0) {
-		for (int i = 0; i < GuiItems.size(); i++) {
-			GuiItems.get(i).paint(arg0);
-		}
-	}
-
-	public void ClearAll() {
-		GuiItems.clear();
-	};
-
-	protected boolean Active = false; // if true then this layer will be
-										// displayed
-
-	public boolean isActive() {
-		return Active;
-	}
-
-	public void setActive(boolean active) {
-		Active = active;
-	}
-
-	public String getLayerName() {
-		return LayerName;
-	}
-
-	public void setLayerName(String layerName) {
-		LayerName = layerName;
-	}
-
-	public void update(Observable ob, Object arg1) {
-		if (ob instanceof GraphWatched) {
-			GraphWatched gr = (GraphWatched) ob;
-
-			if (this.Active)
-
-			{
-				// /logger.info("iiiiiiiiiiiiiiiiiiiiiiiiiiiiiiiii");
-				paint(gr.getG());
-			}
-			// else
-			// logger.info("in update but not active");
-		}
-	}
-
-}

### 2010-09-24 MySimpleSketch/src/SketchMaster/Stroke/graphics/layers/SketchSegmentionLayer.java

@@ -1,80 +1,0 @@
-/**
- * 
- */
-package SketchMaster.Stroke.graphics.layers;
-
-import java.awt.Graphics;
-import java.awt.Graphics2D;
-import java.util.ArrayList;
-
-import SketchMaster.system.SystemSettings;
-
-// import SketchMaster.Stroke.StrokeData.Symbol;
-// import SketchMaster.Stroke.graphics.shapes.StrokeFitObject;
-
-/**
- * @author Mahi
- */
-public class SketchSegmentionLayer extends SketchLayer {
-
-	/**
-	 * 
-	 */
-	public SketchSegmentionLayer() {
-		super();
-
-	}
-
-	// /**
-	// * @directed
-	// * @uml.property name="shapes"
-	// * @uml.associationEnd multiplicity="(0 -1)"
-	// */
-	// protected ArrayList<StrokeFitObject> shapes=new
-	// ArrayList<StrokeFitObject>();
-	// public StrokeFitObject getSymbols(int index){
-	// return shapes.get(index);
-	// }
-	//	
-	// public void AddSymbols( StrokeFitObject symbols){
-	// this.shapes.add(symbols);
-	// }
-	//	
-	// /**
-	// * @return the shapes
-	// * @uml.property name="shapes"
-	// */
-	// public ArrayList<StrokeFitObject> getShapes(){
-	// return shapes;
-	// }
-	//
-	// @Override
-	// public void paint(Graphics2D arg0) {
-	//			
-	// for (int i = 0; i < shapes.size(); i++) {
-	// //shapes.get(i).setFitType(SystemSettings.STROKE_POLYGON);
-	// shapes.get(i).paint(arg0);
-	// }
-	// }
-	//
-	// @Override
-	// public void addObject(Object object) {
-	// if (object instanceof StrokeFitObject) {
-	// StrokeFitObject temp = (StrokeFitObject) object;
-	// shapes.add(temp);
-	//				
-	// }
-	//			
-	//			
-	//			
-	// }
-	//	
-	//	
-	// @Override
-	// public void ClearAll() {
-	//
-	// shapes.clear();
-	//			
-	// }
-
-}

### 2010-09-24 MySimpleSketch/src/SketchMaster/Stroke/graphics/layers/SketchStrokesLayer.java

@@ -1,105 +1,0 @@
-/**
- * 
- */
-package SketchMaster.Stroke.graphics.layers;
-import java.awt.Color;
-import java.awt.Graphics;
-import java.awt.Graphics2D;
-import java.util.ArrayList;
-
-import SketchMaster.Stroke.StrokeData.Stroke;
-import SketchMaster.Stroke.graphics.shapes.GuiShape;
-
-// import SketchMaster.Stroke.graphics.shapes.StrokeFitObject;
-
-/**
- * @author Mahi
- */
-public class SketchStrokesLayer extends SketchLayer {
-	/**
-	 * 
-	 */
-	public SketchStrokesLayer() {
-		super();
-
-	}
-
-	/**
-	 * @return the strokes
-	 * @uml.property name="strokes"
-	 */
-	public ArrayList getStrokes() {
-		return Strokes;
-	}
-
-	public void addStroke(SketchMaster.Stroke.StrokeData.Stroke stroke) {
-		Strokes.add(stroke);
-	}
-
-	private void clearAllStroke() {
-		for (int i = 0; i < Strokes.size(); i++) {
-			((Stroke) (Strokes.get(i))).clearAllPoints();
-		}
-		Strokes.clear();
-	}
-
-	/**
-	 * @param strokes
-	 *            the strokes to set
-	 * @uml.property name="strokes"
-	 */
-	public void setStrokes(ArrayList Strokes) {
-		this.Strokes = Strokes;
-	}
-
-	public void drawStrokes(Graphics2D g) {
-		// logger.info("i am on the draw");
-
-		int max = Strokes.size();
-
-		for (int i = 0; i < Strokes.size(); i++) {
-			// logger.info("stroke "+ i);
-			// g.setColor();
-			Color c = Color.getHSBColor((float) i / (float) max, (float) 0.7,
-					(float) 0.5);
-			if (Math.abs(i) % 2 == 1)
-				((Stroke) (Strokes.get(i)))
-						.drawStroke(g, Color.BLUE, Color.RED);
-			else
-				((Stroke) (Strokes.get(i)))
-						.drawStroke(g, Color.red, Color.black);
-
-		}
-	}
-
-	/** @link dependency */
-	/* # Stroke lnkStroke; */
-
-	/**
-	 * @uml.property name="strokes"
-	 */
-	private ArrayList<GuiShape> Strokes = new ArrayList();
-
-	@Override
-	public void paint(Graphics2D g) {
[…]

### 2010-09-24 MySimpleSketch/src/SketchMaster/Stroke/graphics/shapes/BezireCurve.java

@@ -1,324 +1,0 @@
-/**
- * 
- */
-package SketchMaster.Stroke.graphics.shapes;
-import java.awt.Color;
-import java.awt.Graphics2D;
-import java.awt.Stroke;
-import java.awt.geom.Point2D;
-import java.util.ArrayList;
-
-import SketchMaster.Stroke.StrokeData.PointData;
-import SketchMaster.lib.ComputationsGeometry;
-
-/**
- * @author maha
- * 
- */
-public class BezireCurve extends Circle {
-	/**
-	 * 
-	 */
-	private static final long serialVersionUID = -3510501677272295969L;
-	// StartPoint, EndPoint;
-	transient Point2D[] Curve = new Point2D[100];
-	 PointData[] Cp;
-	double ax, bx, cx; // /coffecient of the curve
-	double ay, by, cy;//
-	static int NO_OF_SEGMENT = 5;
-
-	@Override
-	public void paint(Graphics2D g) {
-		// g.setColor(Color.BLACK);
-		// g.drawRect((int)StartPoint.x, (int)StartPoint.y,3, 3);
-		// g.drawRect((int)EndPoint.x, (int)EndPoint.y,3, 3);
-		//		
-		//		
-		// g.fillRect((int)StartPoint.x, (int)StartPoint.y,3, 3);
-		// g.fillRect((int)EndPoint.x, (int)EndPoint.y,3, 3);
-		//		
-		// // 
-		// logger.info("this curve is beizer // TODO: drawing of the
-		// paint bizire curve. ");
-		// super.paint(g);
-
-		drawCurve(g);
-		super.paint(g);
-	}
-
-	public void drawCurve(Graphics2D g) {
-		// logger.info("000000000000000000000000000000000000000000000000000000000000000000000000000");
-
-		Stroke s = g.getStroke();
-		g.setStroke(bs);
-
-		g.setColor(Color.ORANGE);
-		if (Curve != null)
-
-		{
-			for (int i = 0; i < Curve.length - 1; i++) {
-				if (Curve[i] != null) {
-					// logger.info(Curve[i]);
-					if (!Double.isNaN(Curve[i].getX())) {
-						if (!Double.isNaN(Curve[i].getY())) {
-							// logger.info("The curev e"+Curve[i]);
-							g.drawLine((int) Curve[i].getX(), (int) Curve[i]
-									.getY(), (int) Curve[i + 1].getX(),
-									(int) Curve[i + 1].getY());
-						}
[…]

### 2010-09-24 MySimpleSketch/src/SketchMaster/Stroke/graphics/shapes/Circle.java

@@ -1,433 +1,0 @@
-/**
- * 
- */
-package SketchMaster.Stroke.graphics.shapes;
-import java.awt.BasicStroke;
-import java.awt.Color;
-import java.awt.Graphics;
-import java.awt.Graphics2D;
-import java.awt.Point;
-import java.awt.Rectangle;
-import java.awt.Stroke;
-import java.awt.geom.Point2D;
-import java.awt.geom.Rectangle2D;
-import java.util.ArrayList;
-
-import org.apache.log4j.Logger;
-
-import SketchMaster.Stroke.StrokeData.PointData;
-import SketchMaster.lib.ComputationsGeometry;
-import SketchMaster.lib.MathLib;
-
-/**
- * @author maha
- * 
- */
-public class Circle extends GeometricPrimitive {
-
-
-	private static final Logger logger = Logger.getLogger(Circle .class);
-	
-	@Override
-	public String toString() {
-		
-		return " Center ( "+a+","+b+") Radius ="+r;
-	}
-
-	double r, a, b, x, y;
-	public int type;
-	PointData d1, d2,  centerpoint;
-	double start, end;
-	int percentDrawnWillDraw;
-	int startAngle=0,arcAngle=360;
-	
-	final static double toTheta = 180.0 / Math.PI;
-	Rectangle arcBox = null;
-	static protected Stroke bs = new BasicStroke(3);
-
-	public void paint(Graphics2D g) {
-		  super.paint(g);
-		// logger.info("draw oval ");
-		// g.setColor(Color.gray);
-		// g.draw(arcBox);
-
-		// g.setColor(Color.gray);
-		// g.drawRect((int)x,(int)y , (int)(2*r), (int)(2*r));
-		// g.setColor(Color.MAGENTA);
-		// g.draw(arcBox);
-		// g.drawRect((int)d1.getX(),(int)d1.getY() , (int)(4), (int)(4));
-		// g.drawRect((int)d2.getX(),(int)d2.getY() , (int)(4), (int)(4));
- if (arcBox!=null){
-		Stroke s = g.getStroke();
-		g.setStroke(bs);
-		// logger.info("x "+(int)x+" y "+(int)y +" 2r "+ (int)(2*r)+"
-		// start "+ (int)start+" end "+end+" angle "+(int)(end-start));
-
-		g.setColor(Color.RED);
-		// g.drawArc((int)x,(int)y , (int)(2*r),
[…]

### 2010-09-24 MySimpleSketch/src/SketchMaster/Stroke/graphics/shapes/ComplexPolygon.java

@@ -1,30 +1,0 @@
-/**
- * 
- */
-package SketchMaster.Stroke.graphics.shapes;
-import java.awt.Graphics2D;
-import java.io.Serializable;
-import java.util.ArrayList;
-
-import SketchMaster.Stroke.StrokeData.PointData;
-
-/**
- * @author maha
- * 
- */
-public class ComplexPolygon extends GeometricPrimitive {
-	public void paint(Graphics2D g) {
-           super.paint(g);
-	}
-
-	public void setParam(ArrayList Param) {
-
-		// this.param = param;
-	}
-
-	@Override
-	public double DifferanceFromPoint(PointData point) {
-		  
-		return 0;
-	}
-}

### 2010-09-24 MySimpleSketch/src/SketchMaster/Stroke/graphics/shapes/Direction.java

@@ -1,28 +1,0 @@
-// Decompiled by Jad v1.5.8f. Copyright 2001 Pavel Kouznetsov.
-// Jad home page: http://www.kpdus.com/jad.html
-// Decompiler options: packimports(3) 
-// Source File Name:   Direction.java
-
-package SketchMaster.Stroke.graphics.shapes;
-class Direction
-{
-    Direction()
-    {
-    }
-
-    public static String toString(int relation)
-    {
-        switch(relation)
-        {
-        case 0: // '\0'
-            return "left_right";
-
-        case 1: // '\001'
-            return "above_below";
-        }
-        return "unknown relation";
-    }
-
-    public static final int LEFT_RIGHT = 0;
-    public static final int ABOVE_BELOW = 1;
-}

### 2010-09-24 MySimpleSketch/src/SketchMaster/Stroke/graphics/shapes/FittedShape.java

@@ -1,144 +1,0 @@
-/**
- * 
- */
-package SketchMaster.Stroke.graphics.shapes;
-
-import java.awt.Graphics2D;
-import java.awt.geom.Point2D;
-import java.util.ArrayList;
-
-import org.apache.log4j.Logger;
-
-import SketchMaster.Stroke.StrokeData.DominatePointStructure;
-import SketchMaster.Stroke.StrokeData.InkInterface;
-import SketchMaster.Stroke.StrokeData.InkPart;
-import SketchMaster.Stroke.StrokeData.PointData;
-import SketchMaster.Stroke.StrokeData.Segment;
-import SketchMaster.Stroke.StrokeData.Stroke;
-import SketchMaster.lib.ComputationsGeometry;
-import SketchMaster.lib.CurveFitData;
-import SketchMaster.system.SystemSettings;
-import SketchMaster.system.segmentation.SketchSegmentors;
-
-/**
- * @author TOSHIBA
- *
- */
-public class FittedShape implements GuiShape,SegmentedShape{
-
-
-	private static final Logger logger = Logger.getLogger(FittedShape.class);
-	GeometricPrimitive  shape;
-	double Error;
-	boolean accepted;
-	int type;
-	
-	private InkInterface Originalink;
-
-	Segment seg=null;
-	
-	
-	public final static int TYPE_LINE=0;  
-	public final static int TYPE_CIRCLE=1; 
-	public final static int TYPE_ELLIPSE=2;  
-	public final static int TYPE_POLYLINE=3;
-	public final static int TYPE_HELIX=4;
-	public final static int TYPE_SPRIAL=5;
-	public final static int TYPE_ARC=6;
-	
-//	public	double slope;
-//	public double  intercept;
-	
-	
-	public void paint(Graphics2D g) {
-		shape.paint(g);
-		
-	}
-
-
-	public GeometricPrimitive getShape() {
-		return shape;
-	}
-
-
-	public void setShape(GeometricPrimitive shape) {
-		this.shape = shape;
-	}
-
-
-	public double getError() {
-		return Error;
-	}
-
-
-	public void setError(double error) {
-		Error = error;
-	}
-
-
-	public boolean isAccepted() {
-		return accepted;
-	}
-
-
-	public void setAccepted(boolean accepted) {
-		this.accepted = accepted;
-	}
-
-
-	public int getType() {
-		return type;
-	}
-
-
[…]

### 2010-09-24 MySimpleSketch/src/SketchMaster/Stroke/graphics/shapes/GeometricObject.java

@@ -1,43 +1,0 @@
-// Decompiled by Jad v1.5.8f. Copyright 2001 Pavel Kouznetsov.
-// Jad home page: http://www.kpdus.com/jad.html
-// Decompiler options: packimports(3) 
-// Source File Name:   GeometricObject.java
-
-package SketchMaster.Stroke.graphics.shapes;
-
-
-
-
-// Referenced classes of package edu.mit.sketch.geom:
-//            Translatable, Polygon, Rectangle, Vertex
-
-public interface GeometricObject
-    extends Terminal, Translatable, GuiShape
-{
-
-    public abstract int getIntType();
-
-    public abstract PolygonShape getPolygonalBounds();
-
-    public abstract boolean touches(GeometricObject geometricobject);
-
-    public abstract int spatialRelation(GeometricObject geometricobject);
-
-    public abstract  SketchMaster.Stroke.graphics.shapes.Rectangle getRectangularBounds();
-
-    public abstract java.awt.Rectangle getBounds();
-
-    public abstract boolean containsGeometricObject(GeometricObject geometricobject);
-
-    public abstract boolean containsGeometricObjects(GeometricObject ageometricobject[]);
-
-  //  public abstract void setDataPoints(Polygon polygon);
-
-   // public abstract Polygon getDataPoints();
-
- //   public abstract Vertex[] getOriginalVertices();
-
- //   public abstract void setOriginalVertices(Vertex avertex[]);
-
-    public abstract GeometricObject copy();
-}

### 2010-09-24 MySimpleSketch/src/SketchMaster/Stroke/graphics/shapes/GeometricPrimitive.java

@@ -1,116 +1,0 @@
-/**
- * 
- */
-package SketchMaster.Stroke.graphics.shapes;
-import java.awt.Color;
-import java.awt.Graphics2D;
-import java.awt.geom.Point2D;
-import java.io.Serializable;
-import java.util.ArrayList;
-
-import SketchMaster.Stroke.StrokeData.PointData;
-
-/**
- * @author maha
- * 
- */
-public abstract class GeometricPrimitive implements GuiShape, Serializable {
-	private double ErrorComputed=0.0;
-	
-	PointData StartPoint,EndPoint;
-	int iStart,iEnd;
-	public  void paint(Graphics2D g){
-		Point2D pointS = getStartPoint();
-		Point2D pointe=	getEndPoint();
-		if (	pointS!=null){
-		g.setColor(Color.BLUE);
-		g.drawRect((int)pointS.getX(),(int)pointS.getY(),4,4);
-		g.fillRect((int)pointS.getX(),(int)pointS.getY(),3,3);
-		}
-		if (pointe!=null)
-		{
-		g.setColor(Color.RED);
-		g.drawRect((int)pointe.getX(),(int)pointe.getY(),6,6);
-		g.fillRect((int)pointe.getX(),(int)pointe.getY(),5,5);
-		}
-	}
-
-	public abstract void setParam(ArrayList Param);
-//
-//	abstract public Point2D getStartPoint();
-//
-//	abstract public Point2D getEndPoint();
-
-	public abstract double DifferanceFromPoint(PointData point);
-	
-	public Point2D getEndPoint() {
-		
-		return EndPoint;
-	}
-
-	public Point2D getStartPoint() {
-		
-		return StartPoint;
-	}
-
-	
-	public void setStartPoint(PointData p) {
-		
-//		logger.info("setting the start point of the class "+" (" + this.getClass().getSimpleName()
-//				+ "    " + (new Throwable()).getStackTrace()[0].getLineNumber()
-//				+ "  )  ");
-			StartPoint=p;
-	}
-
-	public void setEndPoint(PointData p) {
-//	        logger.info(" setting the end point  "+" (" + this.getClass().getSimpleName()
-//					+ "    "
-//					+ (new Throwable()).getStackTrace()[0].getLineNumber()
-//					+ "  )  ");
-		EndPoint=p;
-		
-	}
-
-	/**
-	 * @return the errorComputed
-	 */
-	public double getErrorComputed() {
[…]

### 2010-09-24 MySimpleSketch/src/SketchMaster/Stroke/graphics/shapes/GeometryUtil.java

@@ -1,201 +1,0 @@
-// Decompiled by Jad v1.5.8f. Copyright 2001 Pavel Kouznetsov.
-// Jad home page: http://www.kpdus.com/jad.html
-// Decompiler options: packimports(3) 
-// Source File Name:   GeometryUtil.java
-
-package SketchMaster.Stroke.graphics.shapes;
-
-import org.apache.log4j.Logger;
-
-import java.io.PrintStream;
-
-import SketchMaster.Stroke.StrokeData.PointData;
-
-// Referenced classes of package edu.mit.sketch.geom:
-//            Line, Point, Vertex
-
-public class GeometryUtil
-{
-	/**
-	 * Logger for this class
-	 */
-	private static final Logger logger = Logger.getLogger(GeometryUtil.class);
-
-    public GeometryUtil()
-    {
-    }
-
-    public static double[] roundAngles(double angles[], double dx)
-    {
-        double result[] = new double[angles.length];
-        for(int i = 0; i < angles.length; i++)
-            if(Math.abs(angles[i] - (double)(int)(angles[i] / dx) * dx) < Math.abs(angles[i] - (double)((int)(angles[i] / dx) + 1) * dx))
-                result[i] = (double)(int)(angles[i] / dx) * dx;
-            else
-                result[i] = (double)((int)(angles[i] / dx) + 1) * dx;
-        return result;
-    }
-
-    public static boolean equalDoubles(double a, double b, double tolerance)
-    {
-		boolean returnboolean = Math.abs(a - b) < tolerance;
-        return returnboolean;
-    }
-
-    public static boolean linesParallel(Line a, Line b, double t)
-    {
-        double angle_a = Math.atan2(a.y2 - a.y1, a.x2 - a.x1);
-        double angle_b = Math.atan2(b.y2 - b.y1, b.x2 - b.x1);
-		boolean returnboolean = parallelAngles(angle_a, angle_b, t);
-        return returnboolean;
-    }
-
-    public static boolean parallelAngles(double angle_a, double angle_b, double t)
-    {
-		if (logger.isDebugEnabled()) {
-			//  logger.debug("parallelAngles(double, double, double) - GeometryUtil : parallelAngles()"); //$NON-NLS-1$
-		}
-		if (logger.isDebugEnabled()) {
[…]

### 2010-09-24 MySimpleSketch/src/SketchMaster/Stroke/graphics/shapes/GraphicsUtil.java

@@ -1,72 +1,0 @@
-// Decompiled by Jad v1.5.8f. Copyright 2001 Pavel Kouznetsov.
-// Jad home page: http://www.kpdus.com/jad.html
-// Decompiler options: packimports(3) 
-// Source File Name:   GraphicsUtil.java
-
-package SketchMaster.Stroke.graphics.shapes;
-import java.awt.*;
-
-public class GraphicsUtil
-{
-    public GraphicsUtil()
-    {
-    }
-
-    public static void drawThickLine(int r, Graphics g, int x1, int y1, int x2, int y2)
-    {
-        if(r == 1)
-        {
-            g.drawLine(x1, y1, x2, y2);
-            return;
-        }
-        int iterations = (int)Math.sqrt((x1 - x2) * (x1 - x2) + (y1 - y2) * (y1 - y2));
-        for(int i = 0; i < iterations; i++)
-            g.fillOval((x1 * i + x2 * (iterations - i)) / iterations, (y1 * i + y2 * (iterations - i)) / iterations, r, r);
-    }
-
-    public static void drawThickLine(double r, Graphics g, double x1, double y1, double x2, double y2)
-    {
-        if(r == 1.0D)
-        {
-            g.drawLine((int)x1, (int)y1, (int)x2, (int)y2);
-            return;
-        }
-        int iterations = (int)Math.sqrt((x1 - x2) * (x1 - x2) + (y1 - y2) * (y1 - y2));
-        for(int i = 0; i < iterations; i++)
-            g.fillOval((int)((x1 * (double)i + x2 * (double)(iterations - i)) / (double)iterations), (int)((y1 * (double)i + y2 * (double)(iterations - i)) / (double)iterations), (int)r, (int)r);
-    }
-
-    public static void drawThickOval(int r, Graphics g, int x, int y, int r1, int r2)
-    {
-        if(r == 1)
-        {
-            g.drawOval(x, y, r1, r2);
-            return;
-        }
-        for(int i = 0; i < r; i++)
-            g.drawOval(x + i, y + i, r1 - i * 2, r2 - i * 2);
-    }
-
-    public static void setConstraints(GridBagLayout layout, Component component, int x, int y, int column_span, int row_span, double weightx, 
-            double weighty, int fill, int anchor, Insets insets)
-    {
-        GridBagConstraints gbc = new GridBagConstraints();
-        gbc.gridx = x;
-        gbc.gridy = y;
-        gbc.gridwidth = column_span;
[…]

### 2010-09-24 MySimpleSketch/src/SketchMaster/Stroke/graphics/shapes/GuiShape.java

@@ -1,20 +1,0 @@
-package SketchMaster.Stroke.graphics.shapes;
-
-import java.awt.Graphics;
-import java.awt.Graphics2D;
-import java.util.ArrayList;
-
-import org.apache.log4j.Logger;
-
-import SketchMaster.Stroke.StrokeData.Stroke;
-
-public interface GuiShape {
-
-	public void paint(Graphics2D g);
-
-
-
-
-
-	
-}

### 2010-09-24 MySimpleSketch/src/SketchMaster/Stroke/graphics/shapes/PolygonShape.java

@@ -1,828 +1,0 @@
-// Decompiled by Jad v1.5.8f. Copyright 2001 Pavel Kouznetsov.
-// Jad home page: http://www.kpdus.com/jad.html
-// Decompiler options: packimports(3) 
-// Source File Name:   Polygon.java
-
-package SketchMaster.Stroke.graphics.shapes;
-
-import org.apache.log4j.Logger;
-
-import java.awt.Color;
-import java.awt.Graphics;
-import java.awt.Graphics2D;
-import java.awt.Rectangle;
-import java.io.PrintStream;
-import java.io.Serializable;
-
-
-
-
-
-import SketchMaster.Stroke.StrokeData.PointData;
-
-// Referenced classes of package edu.mit.sketch.geom:
-//            Range, Point, Rectangle, Line, 
-//            Vertex, GeometricObject, GeometryUtil
-
-public class PolygonShape extends java.awt.Polygon
-    implements GeometricObject, Serializable
-{
-	/**
-	 * Logger for this class
-	 */
-	private static final Logger logger = Logger.getLogger(PolygonShape.class);
-
-    public void setIndices(int indices[])
-    {
-//        this.indices = indices;
-//        ranges = new Range[indices.length - 1];
-//        for(int i = 0; i < ranges.length; i++)
-//            ranges[i] = getRange(i);
-
-    }
-
-//    public Range[] getRanges()
-//    {
-//        return ranges;
-//    }
-//
-//    public Range getRange(int i)
-//    {
-//        if(i + 1 < indices.length)
-//            return new Range(indices[i], indices[i + 1]);
-//        else
-//            return null;
-//    }
-
-    public PolygonShape()
-    {
-    }
-
-    public PolygonShape(java.awt.Polygon p)
-    {
-        npoints = p.npoints;
-        xpoints = new int[p.npoints];
-        ypoints = new int[p.npoints];
-        for(int i = 0; i < npoints; i++)
-        {
-            xpoints[i] = p.xpoints[i];
-            ypoints[i] = p.ypoints[i];
-        }
-
-    }
-
-    public PolygonShape(PointData points[])
-    {
-        npoints = points.length;
-        xpoints = new int[points.length];
-        ypoints = new int[points.length];
-        for(int i = 0; i < npoints; i++)
-        {
-            xpoints[i] = (int)points[i].x;
-            ypoints[i] = (int)points[i].y;
-        }
-
-    }
-
-    public PolygonShape(int xpoints[], int ypoints[], int npoints)
-    {
[…]

### 2010-09-24 MySimpleSketch/src/SketchMaster/Stroke/graphics/shapes/Rectangle.java

@@ -1,724 +1,0 @@
-// Decompiled by Jad v1.5.8f. Copyright 2001 Pavel Kouznetsov.
-// Jad home page: http://www.kpdus.com/jad.html
-// Decompiler options: packimports(3) 
-// Source File Name:   Rectangle.java
-
-package SketchMaster.Stroke.graphics.shapes;
-
-import org.apache.log4j.Logger;
-
-
-import java.awt.Color;
-import java.awt.Dimension;
-import java.awt.Graphics;
-import java.awt.Graphics2D;
-import java.awt.geom.Dimension2D;
-import java.io.PrintStream;
-import java.io.Serializable;
-
-import SketchMaster.Stroke.StrokeData.PointData;
-import SketchMaster.system.SystemSettings;
-
-// Referenced classes of package edu.mit.sketch.geom:
-//            Polygon, Line, Point, Vertex, 
-//            GeometricObject, Rotatable, GeometryUtil, Direction, 
-//            SpatialRelation
-
-public class Rectangle
-    implements GeometricObject,  Serializable
-{
-	/**
-	 * 
-	 */
-	private static final long serialVersionUID = 20359727367889206L;
-
-	/**
-	 * Logger for this class
-	 */
-	private static final Logger logger = Logger.getLogger(Rectangle.class);
-
-    private PolygonShape points;
-    private Vertex m_vertices[];
-    private double angle;
-    private double width_x;
-    private double width_y;
-    private double height_x;
-    private double height_y;
-    private double width;
-    private double height;
-    public double x;
-    public double y;
-    public long time_stamp;
-    public transient Graphics graphics;
-    public Rectangle(java.awt.Rectangle rectangle)
-    {
-        this(rectangle.x, rectangle.y, rectangle.getWidth(), 0.0D, rectangle.getHeight());
-    }
-
-    public Rectangle(PointData upper_left_point, Dimension dimension)
-    {
-        this(upper_left_point.x, upper_left_point.y, dimension.width, 0.0D, dimension.height);
-    }
-
-    public Rectangle(PointData upper_left_point, Dimension2D dimension)
-    {
-        this(upper_left_point.x, upper_left_point.y, dimension.getWidth(), 0.0D, dimension.getHeight());
-    }
-
-    public Rectangle(PointData upper_left_point, Dimension dimension, double angle)
-    {
-        this.angle = Math.atan2(Math.sin(angle), Math.cos(angle));
-        x = upper_left_point.x;
-        y = upper_left_point.y;
[…]

### 2010-09-24 MySimpleSketch/src/SketchMaster/Stroke/graphics/shapes/SegmentedShape.java

@@ -1,17 +1,0 @@
-/**
- * 
- */
-package SketchMaster.Stroke.graphics.shapes;
-
-import SketchMaster.Stroke.StrokeData.Segment;
-
-/**
- * @author maha
- * 
- */
-public interface SegmentedShape extends GuiShape {
-
-	public int getSegmentsCount();
-
-	public Segment getSegmentOfIndex(int i);
-}

### 2010-09-24 MySimpleSketch/src/SketchMaster/Stroke/graphics/shapes/SpatialRelation.java

@@ -1,158 +1,0 @@
-// Decompiled by Jad v1.5.8f. Copyright 2001 Pavel Kouznetsov.
-// Jad home page: http://www.kpdus.com/jad.html
-// Decompiler options: packimports(3) 
-// Source File Name:   SpatialRelation.java
-
-package SketchMaster.Stroke.graphics.shapes;
-public class SpatialRelation
-{
-    public SpatialRelation()
-    {
-    }
-
-    public static String toString(int relation)
-    {
-        switch(relation)
-        {
-        case 0: // '\0'
-            return "upper_left";
-
-        case 1: // '\001'
-            return "above";
-
-        case 2: // '\002'
-            return "upper_right";
-
-        case 3: // '\003'
-            return "left";
-
-        case 4: // '\004'
-            return "cocentric";
-
-        case 5: // '\005'
-            return "right";
-
-        case 6: // '\006'
-            return "lower_left";
-
-        case 7: // '\007'
-            return "below";
-
-        case 8: // '\b'
-            return "lower_right";
-        }
-        return "unknown relation";
-    }
-
-    public static int cartesian2screen(int relation)
-    {
-        switch(relation)
-        {
-        case 0: // '\0'
-            return 6;
-
-        case 1: // '\001'
-            return 7;
-
-        case 2: // '\002'
-            return 8;
-
-        case 3: // '\003'
-            return 3;
-
-        case 4: // '\004'
-            return 4;
-
-        case 5: // '\005'
-            return 5;
-
-        case 6: // '\006'
-            return 0;
-
-        case 7: // '\007'
-            return 1;
-
-        case 8: // '\b'
-            return 2;
-        }
-        return relation;
-    }
-
-    public static int screen2cartesian(int relation)
-    {
-        switch(relation)
-        {
-        case 0: // '\0'
-            return 6;
-
-        case 1: // '\001'
-            return 7;
-
-        case 2: // '\002'
-            return 8;
-
-        case 3: // '\003'
-            return 3;
-
-        case 4: // '\004'
-            return 4;
-
-        case 5: // '\005'
-            return 5;
-
-        case 6: // '\006'
-            return 0;
-
-        case 7: // '\007'
-            return 1;
-
-        case 8: // '\b'
-            return 2;
-        }
-        return relation;
-    }
-
-    public static int getComplement(int relation)
-    {
-        switch(relation)
-        {
-        case 0: // '\0'
-            return 8;
-
[…]

### 2010-09-24 MySimpleSketch/src/SketchMaster/Stroke/graphics/shapes/Terminal.java

@@ -1,17 +1,0 @@
-// Decompiled by Jad v1.5.8f. Copyright 2001 Pavel Kouznetsov.
-// Jad home page: http://www.kpdus.com/jad.html
-// Decompiler options: packimports(3) 
-// Source File Name:   Terminal.java
-
-package SketchMaster.Stroke.graphics.shapes;
-
-
-
-public interface Terminal
-   
-{
-
-    public abstract void setTime(long l);
-
-    public abstract long getTime();
-}

### 2010-09-24 MySimpleSketch/src/SketchMaster/Stroke/graphics/shapes/Translatable.java

@@ -1,22 +1,0 @@
-// Decompiled by Jad v1.5.8f. Copyright 2001 Pavel Kouznetsov.
-// Jad home page: http://www.kpdus.com/jad.html
-// Decompiler options: packimports(3) 
-// Source File Name:   Translatable.java
-
-package SketchMaster.Stroke.graphics.shapes;
-
-import SketchMaster.Stroke.StrokeData.PointData;
-
-
-// Referenced classes of package edu.mit.sketch.geom:
-//            Point
-
-public interface Translatable
-{
-
-    public abstract void translate(double d, double d1);
-
-    public abstract boolean pointIsOn(PointData point, double d);
-
-    public abstract boolean pointIsOnOriginal(PointData point, double d);
-}

### 2010-09-24 MySimpleSketch/src/SketchMaster/Stroke/graphics/shapes/Vertex.java

@@ -1,169 +1,0 @@
-// Decompiled by Jad v1.5.8f. Copyright 2001 Pavel Kouznetsov.
-// Jad home page: http://www.kpdus.com/jad.html
-// Decompiler options: packimports(3) 
-// Source File Name:   Vertex.java
-
-package SketchMaster.Stroke.graphics.shapes;
-import java.awt.Point;
-import java.awt.geom.Point2D;
-import java.io.Serializable;
-import java.util.*;
-
-import SketchMaster.Stroke.StrokeData.PointData;
-
-// Referenced classes of package edu.mit.sketch.geom:
-//            Point, Polygon, GeometricObject
-
-public class Vertex extends PointData
-    implements Serializable, Comparator
-{
-    public Vertex()
-    {
-        certainty = 0.0D;
-    }
-
-    public Vertex(Point point)
-    {
-        super(point);
-    }
-
-    public Vertex(Point2D point)
-    {
-        super(point);
-        if(point instanceof Vertex)
-        {
-            index = ((Vertex)point).index;
-            certainty = ((Vertex)point).certainty;
-            time= ((Vertex)point).getTime();
-        } else
-        if(point instanceof PointData)
-            time = ((PointData)point).getTime();
-        
-        else
-            certainty = 0.0D;
-    }
-
-    public Vertex(double x, double y)
-    {
-        super(x, y);
-        certainty = 0.0D;
-    }
-
-    public Vertex(double certainty)
-    {
-        this.certainty = certainty;
-    }
-
-    public Vertex(Point point, double certainty)
-    {
-        super(point);
-        this.certainty = certainty;
-    }
-
-    public Vertex(Point2D point, double certainty)
-    {
-        super(point);
-        this.certainty = certainty;
-        if(point instanceof Vertex)
-            index = ((Vertex)point).index;
-    }
-
-    public Vertex(double x, double y, double certainty)
-    {
-        super(x, y);
-        this.certainty = certainty;
-    }
-
-    public Vertex(Vertex vertex)
-    {
-        this(vertex.x, vertex.y, vertex.certainty);
-        index = vertex.index;
-        time = vertex.getTime();
-    }
-
-    public void setIndex(int index)
-    {
-        this.index = index;
-    }
-
-    public String toString()
-    {
[…]

### 2010-09-24 MySimpleSketch/src/SketchMaster/classifier/Category.java

@@ -1,104 +1,0 @@
-/**
- * 
- */
-package SketchMaster.classifier;
-
-import org.apache.log4j.Logger;
-
-import java.io.Serializable;
-
-import SketchMaster.classifier.rubine.RubineCategory;
-
-/**
- * @author maha
- * 
- */
-public class Category implements Comparable<Category>, Serializable {
-	/**
-	 * 
-	 */
-	private static final long serialVersionUID = 5223827890094748107L;
-
-	/**
-	 * Logger for this class
-	 */
-	private static final Logger logger = Logger.getLogger(Category.class);
-
-	protected String categoryName = "";
-
-	/**
-	 * 
-	 */
-	public Category() {
-
-	}
-
-	/*
-	 * (non-Javadoc)
-	 * 
-	 * @see java.lang.Object#toString()
-	 */
-	@Override
-	public String toString() {
-
-		return this.categoryName + "  \n";
-	}
-
-	/**
-	 * @return the categoryName
-	 */
-	public String getCategoryName() {
-		return categoryName;
-	}
-
-	/**
-	 * @param categoryName
-	 *            the categoryName to set
-	 */
-	public void setCategoryName(String categoryName) {
-		this.categoryName = categoryName;
-	}
-
-	public int compareTo(Category b) throws ClassCastException {
-//		if (logger.isDebugEnabled()) {
-//			//  logger.debug("compareTo(Category) - start"); //$NON-NLS-1$
-//		}
-
-		if (!(b instanceof RubineCategory))
-			throw new ClassCastException("Rubine category expected");
-		Category temp = ((Category) b);
-
-		int returnint = this.categoryName.compareTo(temp.categoryName);
-//		if (logger.isDebugEnabled()) {
-//			//  logger.debug("compareTo(Category) - end"); //$NON-NLS-1$
-//		}
-		return returnint;
-	}
-
-	@Override
-	public int hashCode() {
-		final int prime = 31;
-		int result = 1;
-		result = prime * result
-				+ ((categoryName == null) ? 0 : categoryName.hashCode());
-		return result;
-	}
-
-	@Override
-	public boolean equals(Object obj) {
-		if (this == obj)
-			return true;
-		if (obj == null)
-			return false;
-		if (getClass() != obj.getClass())
-			return false;
-		Category other = (Category) obj;
-		if (categoryName == null) {
-			if (other.categoryName != null)
-				return false;
-		} else if (!categoryName.equals(other.categoryName))
-			return false;
-		return true;
-	}
-
-}

### 2010-09-24 MySimpleSketch/src/SketchMaster/classifier/Classification.java

@@ -1,55 +1,0 @@
-/**
- * 
- */
-package SketchMaster.classifier;
-
-import java.util.Collection;
-import java.util.Comparator;
-import java.util.Iterator;
-import java.util.Map;
-import java.util.Set;
-import java.util.SortedMap;
-import java.util.TreeMap;
-
-import SketchMaster.collection.NumericalComparator;
-import SketchMaster.collection.SortedValueMap;
-
-/**
- * @author maha
- * 
- */
-public class Classification extends SortedValueMap // TreeMap<Double,Category>
-{
-
-	/**
-	 * 
-	 */
-	private static final long serialVersionUID = -7505351837520558540L;
-
-	/*
-	 * (non-Javadoc)
-	 * 
-	 * @see SketchMaster.collection.SortedValueMap#toString()
-	 */
-	// @Override
-	// public String toString() {
-	// String s=" ";
-	// //this.keySet()
-	// for (Iterator iter = keySet().iterator(); iter.hasNext();) {
-	// Category element = (Category) iter.next();
-	// s+= element.toString()+ " /n ";
-	// }
-	// return s;
-	//		
-	//	
-	// }
-
-	/**
-	 * 
-	 */
-	public Classification() {
-		super(new NumericalComparator());
-
-	}
-
-}

### 2010-09-24 MySimpleSketch/src/SketchMaster/classifier/Classifier.java

@@ -1,18 +1,0 @@
-/**
- * 
- */
-package SketchMaster.classifier;
-
-import SketchMaster.Stroke.StrokeData.InkInterface;
-
-/**
- * @author maha
- * 
- */
-public abstract class Classifier {
-
-	public abstract void train();
-
-	public abstract Classification Classify(InkInterface ink);
-
-}

### 2010-09-24 MySimpleSketch/src/SketchMaster/classifier/Symbol/SymbolCategory.java

@@ -1,628 +1,0 @@
-package SketchMaster.classifier.Symbol;
-
-import org.apache.log4j.Level;
-import org.apache.log4j.Logger;
-
-import java.io.Serializable;
-import java.util.ArrayList;
-import java.util.HashMap;
-
-import JSci.maths.matrices.DoubleSquareMatrix;
-import SketchMaster.Stroke.StrokeData.InkInterface;
-import SketchMaster.Stroke.StrokeData.SegmentCluster;
-import SketchMaster.Stroke.features.InkFeatureSet;
-import SketchMaster.Stroke.features.SegmentClusterFeature;
-import SketchMaster.Stroke.features.SegmentClusterFeatureSet;
-import SketchMaster.Stroke.features.StrokeRubineFeatureSet;
-import SketchMaster.Stroke.features.StrokeStatisticalFeatures;
-import SketchMaster.classifier.Category;
-import SketchMaster.classifier.rubine.RubineCategory;
-import SketchMaster.io.log.FileLog;
-import SketchMaster.system.SystemSettings;
-
-public class SymbolCategory extends RubineCategory implements Serializable {
-	/**
-	 * 
-	 */
-	private static final long serialVersionUID = -8901636817539474407L;
-
-	/**
-	 * Logger for this class
-	 */
-	private static final Logger logger = Logger.getLogger(SymbolCategory.class);
-
-	private static final double ZERO =SystemSettings.ZERO;
-	// private ArrayList<SegmentClusterFeatureSet> inksFeatures;
-
-	
-	private transient double[]  FC_Values=null;
-	private transient boolean FC_Values_OK=false;
-	@Override
-	public void init() {
-//		if (logger.isDebugEnabled()) {
-//			//  logger.debug("init() - start"); //$NON-NLS-1$
-//		}
-
-		FC = new java.util.HashMap<String, Double>();
-		inksFeatures = new ArrayList<InkFeatureSet>();
-
-//		if (logger.isDebugEnabled()) {
-//			//  logger.debug("init() - end"); //$NON-NLS-1$
-//		}
-	}
-
-	public void addExample(SegmentCluster cluster) {
-//		if (logger.isDebugEnabled()) {
-//			//  logger.debug("addExample(SegmentCluster) - start"); //$NON-NLS-1$
-//		}
-
-		SegmentClusterFeatureSet example = new SegmentClusterFeatureSet();
-		example.setSegmentCluster(cluster);
-		example.initAll();
-		example.computeFeatures();
-		inksFeatures.add(example);
[…]

### 2010-09-24 MySimpleSketch/src/SketchMaster/classifier/Symbol/SymbolTrainingSet.java

@@ -1,469 +1,0 @@
-package SketchMaster.classifier.Symbol;
-
-import org.apache.log4j.Logger;
-
-import java.io.Serializable;
-import java.util.ArrayList;
-import java.util.Iterator;
-import java.util.LinkedHashSet;
-
-import JSci.maths.matrices.DoubleSquareMatrix;
-import SketchMaster.Stroke.StrokeData.InkInterface;
-import SketchMaster.Stroke.StrokeData.SegmentCluster;
-import SketchMaster.Stroke.features.InkFeatureSet;
-import SketchMaster.Stroke.features.SegmentClusterFeatureSet;
-import SketchMaster.Stroke.features.StrokeRubineFeatureSet;
-
-import SketchMaster.classifier.rubine.RubineTrainingSet;
-import SketchMaster.io.log.FileLog;
-import SketchMaster.lib.MathLib;
-
-public class SymbolTrainingSet extends RubineTrainingSet implements
-		Serializable {
-	/**
-	 * Logger for this class
-	 */
-	private static final Logger logger = Logger.getLogger(SymbolTrainingSet.class);
-
-	/**
-	 * 
-	 */
-	private static final long serialVersionUID = -53743440296758632L;
-
-	@Override
-	protected void computeCoverianceMatrix() {
-//		if (logger.isDebugEnabled()) {
-//			//  logger.debug("computeCoverianceMatrix() - start"); //$NON-NLS-1$
-//		}
-
-		//  logger.trace(" al coveraince ");
-		CovOk = false;
-		// first compute all the cateories average ssets
-
-		// then set up the get number of features , number of eample, number of
-		// guesturs
-		int nF = 0, nE = 0, nCat = 0;
-		// number of categories C
-		nCat = size();
-
-		// temp vairable for calcuation
-		double E[][] = null;
-		double top = 0, Buttom = 0;
-
-		// initalize buttom with -C
-		Buttom = -nCat;
-
-		int l = 0;
-		// getn number of features ... they supossed to be constant for all
-		// categories.
-		for (Iterator iter = this.iterator(); iter.hasNext(); l++) {
-			SymbolCategory Cat = (SymbolCategory) iter.next();
-			if (l == 0)
-				nF = Cat.getFeaturesCount();
-
[…]

### 2010-09-24 MySimpleSketch/src/SketchMaster/classifier/rubine/RubineCategory.java

@@ -1,564 +1,0 @@
-/**
- * 
- */
-package SketchMaster.classifier.rubine;
-
-import org.apache.log4j.Logger;
-
-import JSci.maths.matrices.DoubleSquareMatrix;
-
-import java.io.Serializable;
-import java.util.ArrayList;
-import java.util.HashMap;
-
-import JSci.maths.matrices.DoubleMatrix;
-import SketchMaster.Stroke.StrokeData.InkInterface;
-import SketchMaster.Stroke.features.InkFeatureSet;
-import SketchMaster.Stroke.features.StrokeRubineFeatureSet;
-import SketchMaster.Stroke.features.StrokeStatisticalFeatures;
-import SketchMaster.classifier.Category;
-import SketchMaster.io.log.FileLog;
-
-/**
- * @author maha
- * 
- */
-public class RubineCategory extends Category implements Serializable {
-	/**
-	 * 
-	 */
-	private static final long serialVersionUID = -6300822404766696586L;
-
-	/**
-	 * Logger for this class
-	 */
-	private static final Logger logger = Logger.getLogger(RubineCategory.class);
-
-	// i need a coveraicne matirx
-	protected transient DoubleSquareMatrix CovaricnceMatrix = null;
-	protected static int FeaturesCount =0;
-	protected boolean covOk = false;
-	protected double[] W = null;
-	protected double wi0 = 0.0;// intial weith of class
-	protected double[][] CovrMat;
-	/**
-	 * 
-	 */
-
-	protected boolean valuesOk = false;
-
-	protected java.util.HashMap<String, Double> FC;
-	
-
-
-	protected ArrayList<InkFeatureSet> inksFeatures;
-
-	// this will contain maiy features
-
-	//
-	public RubineCategory() {
-
-	}
-
-	public void init() {
-//		if (logger.isDebugEnabled()) {
-//			//  logger.debug("init() - start"); //$NON-NLS-1$
-//		}
-
-		FC = new java.util.HashMap<String, Double>();
-		inksFeatures = new ArrayList<InkFeatureSet>();
-
-//		if (logger.isDebugEnabled()) {
-//			//  logger.debug("init() - end"); //$NON-NLS-1$
-//		}
-	}
-
-	public void addExample(InkInterface ink) {
-//		if (logger.isDebugEnabled()) {
-//			//  logger.debug("addExample(InkInterface) - start"); //$NON-NLS-1$
-//		}
-
[…]

### 2010-09-24 MySimpleSketch/src/SketchMaster/classifier/rubine/RubineClassifier.java

@@ -1,124 +1,0 @@
-/**
- * 
- */
-package SketchMaster.classifier.rubine;
-
-import org.apache.log4j.Logger;
-
-import java.io.Serializable;
-import java.util.Iterator;
-
-import SketchMaster.Stroke.StrokeData.InkInterface;
-import SketchMaster.Stroke.features.StrokeRubineFeatureSet;
-import SketchMaster.classifier.Classification;
-import SketchMaster.classifier.Classifier;
-
-/**
- * @author maha
- * 
- */
-public class RubineClassifier extends Classifier implements Serializable {
-	/**
-	 * Logger for this class
-	 */
-	private static final Logger logger = Logger.getLogger(RubineClassifier.class);
-
-	protected boolean trained = false;
-	private RubineTrainingSet trainSet;
-
-	/**
-	 * 
-	 */
-	public RubineClassifier() {
-
-	}
-
-	/*
-	 * (non-Javadoc)
-	 * 
-	 * @see SketchMaster.classifier.Classifier#Classify(SketchMaster.Stroke.StrokeData.InkObject)
-	 */
-	@Override
-	public Classification Classify(InkInterface ink) {
-		if (logger.isDebugEnabled()) {
-			//  logger.debug("Classify(InkInterface) - start"); //$NON-NLS-1$
-		}
-
-		// the new ink if not trained then train.
-
-		this.train();
-
-		Classification inkclass = new Classification();
-		// first i need to get the featus of the commint ink
-
-		StrokeRubineFeatureSet featurs = new StrokeRubineFeatureSet(ink);
-		featurs.initAll();
-		featurs.computeFeatures();
-		// now i have all the features computed
-		double temp = 0;
-		// get the categories
-		// trainSet
-		for (Iterator set = trainSet.iterator(); set.hasNext();) {
-			RubineCategory Category = (RubineCategory) set.next();
-			for (int i = 0; i < featurs.getFeatures().size(); i++) {
-				temp += Category.getCategoryWeight()[i]
-						* featurs.getFeatures().get(i).getValue();
-			}
-			temp += Category.getWi0();
-			// now i need to classifiy
-
-			// this can more t
-			inkclass.put(Category, temp);
-			temp = 0.0;
-
-		}
-
-		if (logger.isDebugEnabled()) {
-			//  logger.debug("Classify(InkInterface) - end"); //$NON-NLS-1$
[…]

### 2010-09-24 MySimpleSketch/src/SketchMaster/classifier/rubine/RubineTrainingSet.java

@@ -1,722 +1,0 @@
-/**
- * 
- */
-package SketchMaster.classifier.rubine;
-
-import org.apache.log4j.Logger;
-
-import java.io.Serializable;
-import java.util.ArrayList;
-import java.util.Collection;
-import java.util.HashMap;
-import java.util.Iterator;
-import java.util.LinkedHashSet;
-
-import JSci.maths.matrices.AbstractDoubleSquareMatrix;
-import JSci.maths.matrices.DoubleMatrix;
-import JSci.maths.matrices.DoubleSquareMatrix;
-import SketchMaster.Stroke.StrokeData.InkInterface;
-import SketchMaster.Stroke.features.InkFeatureSet;
-import SketchMaster.Stroke.features.StrokeRubineFeatureSet;
-import SketchMaster.classifier.Symbol.SymbolCategory;
-import SketchMaster.io.log.FileLog;
-
-/**
- * @author maha
- * 
- */
-public class RubineTrainingSet extends LinkedHashSet<RubineCategory> implements
-		Serializable {
-	/**
-	 * Logger for this class
-	 */
-	private static final Logger logger = Logger.getLogger(RubineTrainingSet.class);
-
-	/**
-	 * 
-	 */
-	private static final long serialVersionUID = 8858155913902859033L;
-	protected boolean CovOk = false;
-	// i need a coveraicne matirx
-	transient protected DoubleSquareMatrix CovaricnceMatrix = null;
-	protected boolean Wok = false;
-
-	protected double[][] CovrMat;
-
-	/*
-	 * *
-	 
-	 */
-	public RubineTrainingSet() {
-
-	}
-
-	/**
-	 * @param arg0
-	 */
-	public RubineTrainingSet(int arg0) {
-		super(arg0);
-
-	}
-
-	/**
-	 * @param arg0
-	 */
-	public RubineTrainingSet(Collection arg0) {
-		super(arg0);
-
-	}
-
-	/**
-	 * @param arg0
-	 * @param arg1
-	 */
-	public RubineTrainingSet(int arg0, float arg1) {
-		super(arg0, arg1);
-
-	}
-
-	protected void computeCoverianceMatrix() {
-//		if (logger.isDebugEnabled()) {
-//			//  logger.debug("computeCoverianceMatrix() - start"); //$NON-NLS-1$
-//		}
-
-		//  logger.trace("al coveraince ");
-		CovOk = false;
[…]

### 2010-09-24 MySimpleSketch/src/SketchMaster/classifier/svmTrainable/DataStat.java

@@ -1,161 +1,0 @@
-/**
- * 
- */
-package SketchMaster.classifier.svmTrainable;
-
-import java.util.ArrayList;
-
-/**
- * @author Maha
- *
- */
-public class DataStat {
- 	double Average;
-	double Max;
-	int Maxind;
-	int Minind;
-	double Min;
-	double Sum;
-	double Std;
-	double variance;
-	
-	ArrayList<Double> data;
-	String name;
-     public void AddData( double d){
-    	 
-    	 if (data==null)
-    		 data=new ArrayList<Double>();
-    	 
-    	 
-    	 data.add(new Double(d));
-     }
-     
-	public void ComputeState(){
-		if ( data==null)
-		return ;
-		if (data.size()>0){
-	 Sum=0;
-	 Max=0;
-	 Min=Double.MAX_VALUE;
-	 Std=0;
-	 Average=0;
-	 
-  //  	int cune=0;
-    	
-    	for (int i = 0; i <  data.size(); i++) {
-			
-    		Sum+=data.get(i);
-    		
-    		if (data.get(i)>Max)
-    		{
-    			Max=data.get(i);
-    			Maxind=i;
-    		}
-    		
-    		if (data.get(i)<Min)
-    		{
-    			Min=data.get(i);
-    			Minind=i;
-    		}
-    		
-		}
-    	
-    	Average=(double)Sum/(double)data.size();
-    	///now compute the standard diviation 
-    	 ComputeStd();
-  
-		}
-		
-		
-	}
-
-	
-	private void ComputeStd(){
-		if ( data==null)
-			return ;
-			if (data.size()>0){
-		
-		double diff=0;
-	double 	StdTemp=0.0;
-	for (int i = 0; i <  data.size(); i++) {
-    		
-    		diff=data.get(i)-Average;
-    		
-    		StdTemp+=diff*diff;
-    		
-    	}
-	
-       double temp=	StdTemp/(double)data.size();	
-       variance=temp;
-       Std=Math.sqrt(temp);
-			}
-		
-	}
-	/**
-	 * @return the data
-	 */
-	public ArrayList<Double> getData() {
-		return data;
-	}
-
-	/**
-	 * @param data the data to set
-	 */
-	public void setData(ArrayList<Double> data) {
-		this.data = data;
-	}
-
-	/**
-	 * @return the name
-	 */
-	public String getName() {
-		return name;
-	}
-
-	/**
-	 * @param name the name to set
-	 */
[…]

### 2010-09-24 MySimpleSketch/src/SketchMaster/classifier/svmTrainable/SVMClassification.java

@@ -1,129 +1,0 @@
-package SketchMaster.classifier.svmTrainable;
-
-import java.util.Arrays;
-
-import SketchMaster.classifier.Classification;
-
-public class SVMClassification  extends Classification  {
-
-	
-	   private String types[];
-	    private double confidences[];
-	    private String highestConfidenceType;
-	    private double highestConfidence;
-	    
-	    @Override
-		public int hashCode() {
-			final int prime = 31;
-			int result = 1;
-			result = prime * result + Arrays.hashCode(confidences);
-			long temp;
-			temp = Double.doubleToLongBits(highestConfidence);
-			result = prime * result + (int) (temp ^ (temp >>> 32));
-			result = prime
-					* result
-					+ ((highestConfidenceType == null) ? 0
-							: highestConfidenceType.hashCode());
-			result = prime * result + Arrays.hashCode(types);
-			return result;
-		}
-
-		@Override
-		public boolean equals(Object obj) {
-			if (this == obj)
-				return true;
-			if (!super.equals(obj))
-				return false;
-			if (getClass() != obj.getClass())
-				return false;
-			SVMClassification other = (SVMClassification) obj;
-			if (!Arrays.equals(confidences, other.confidences))
-				return false;
-			if (Double.doubleToLongBits(highestConfidence) != Double
-					.doubleToLongBits(other.highestConfidence))
-				return false;
-			if (highestConfidenceType == null) {
-				if (other.highestConfidenceType != null)
-					return false;
-			} else if (!highestConfidenceType
-					.equals(other.highestConfidenceType))
-				return false;
-			if (!Arrays.equals(types, other.types))
-				return false;
-			return true;
-		}
-
-		/**
-	     * Construct a classification with no types.
-	     */
-	    public  SVMClassification (String[] stypes, double[]dconfidences) {
-	        types = stypes;
-	        confidences = dconfidences;
-
-	        int maxIndex = 0;
-	        double maxConfidence = confidences[maxIndex];
-	        for(int i=0; i<types.length; i++){
-	            if(confidences[i]>confidences[maxIndex]){
-	                maxIndex = i;
-	                maxConfidence = confidences[maxIndex];
-	            }
-	        }
-	        highestConfidenceType = types[maxIndex];
-	        highestConfidence = maxConfidence;
-	    }
-
-	    /**
-	     * Return the highest confidence value.
-	     */
-	    public double getHighestConfidence(){
-	        return highestConfidence;
-	    }
-
[…]

### 2010-09-24 MySimpleSketch/src/SketchMaster/classifier/svmTrainable/TrainSet.java

@@ -1,314 +1,0 @@
-/**
- * 
- */
-package SketchMaster.classifier.svmTrainable;
-
-import java.awt.List;
-import java.io.FileOutputStream;
-import java.io.PrintStream;
-import java.util.ArrayList;
-import java.util.HashMap;
-import java.util.Iterator;
-
-import org.apache.log4j.Logger;
-
-import SketchMaster.Stroke.StrokeData.PointData;
-import SketchMaster.Stroke.StrokeData.SegmentCluster;
-import SketchMaster.Stroke.StrokeData.Stroke;
-
-import SketchMaster.Stroke.features.SegmentClusterFeatureSet;
-import SketchMaster.io.xml.XmlManager;
-
-/**
- * @author Maha
- *
- */
-public class TrainSet {
-
-	/**
-	 * 
-	 */
-	private static final Logger logger = Logger.getLogger(TrainingSet.class);
-	
-	
-
-	 
-	   /* Internal constant for the array slot of negative examples.
-	     */
-	    private static final int NEGATIVE = 0;
-
-	    /**
-	     * Internal constant for the array slot of positive examples.
-	     */
-	    private static final int POSITIVE = 1;
-
-	    /**
-	     * Store a mapping from types to their positive/negative examples.
-	     */
-	    private HashMap _map;
-
-	    /**
-	     * Construct an empty training set.
-	     */
-	    public TrainSet() {
-	        _map = new HashMap();
-	       
-	    }
-
-	    /**
-	     * Add a negative example to this training set for the given type.
-	     */
-	    public final void addNegativeExample(String t,ArrayList<Stroke> s) {
-	        addExample(t, s, NEGATIVE);
-	    }
-
-	    /**
-	     * Add a positive example to this training set for the given type.
-	     */
-	    public final void addPositiveExample(String t, ArrayList<Stroke> s) {
-	        addExample(t, s,POSITIVE);
-	    }
-  
-	   
-	  
-	    /**
-	     * Add a example to this training set for the given type
-	     * (either positive or negative, denoted by the "which" argument).
-	     */
-	    public final void addExample(String t, ArrayList<Stroke> s, int which) {
-	    	
-	    	
-	    	 ArrayList< ArrayList<Stroke>>[] l = ( ArrayList< ArrayList<Stroke>>[])_map.get(t);
-	        if(l == null) {
-	            l = new ArrayList[2];
[…]

### 2010-09-24 MySimpleSketch/src/SketchMaster/classifier/svmTrainable/TrainingSet.java

@@ -1,496 +1,0 @@
-package SketchMaster.classifier.svmTrainable;
-
-
-
-import java.awt.List;
-import java.io.FileOutputStream;
-import java.io.PrintStream;
-import java.util.ArrayList;
-import java.util.HashMap;
-import java.util.Iterator;
-
-import org.apache.log4j.Logger;
-
-import SketchMaster.Stroke.StrokeData.PointData;
-import SketchMaster.Stroke.StrokeData.SegmentCluster;
-import SketchMaster.Stroke.StrokeData.Stroke;
-import SketchMaster.Stroke.features.SVMFeatureSet;
-import SketchMaster.Stroke.features.SegmentClusterFeatureSet;
- 
-import SketchMaster.system.Recogniziers.SVMRecognizier;
- 
-
-public class TrainingSet {
-
-	private static final Logger logger = Logger.getLogger(TrainingSet.class);
-	
-	
-
-			 
-	   /* Internal constant for the array slot of negative examples.
-	     */
-	    private static final int NEGATIVE = 0;
-
-	    /**
-	     * Internal constant for the array slot of positive examples.
-	     */
-	    private static final int POSITIVE = 1;
-
-	    /**
-	     * Store a mapping from types to their positive/negative examples.
-	     */
-	    private HashMap _map;
-
-	    /**
-	     * Construct an empty training set.
-	     */
-	    public TrainingSet() {
-	        _map = new HashMap();
-	       
-	    }
-
-	    /**
-	     * Add a negative example to this training set for the given type.
-	     */
-	    public final void addNegativeExample(String t,SVMFeatureSet  s) {
-	        addExample(t, s, NEGATIVE);
-	    }
-
-	    /**
-	     * Add a positive example to this training set for the given type.
-	     */
-	    public final void addPositiveExample(String t, SVMFeatureSet  s) {
-	        addExample(t, s, POSITIVE);
-	    }
-     
-	    public boolean[]  getUselessFeatures(){
-	    	ArrayList<DataStat> FeatureStates=null;
-	        for(Iterator iter = types(); iter.hasNext();){
-	             String type = (String)iter.next();
-	     		//logger.error(" Type  ."+type);
-	             ArrayList<SVMFeatureSet>[] l = (ArrayList<SVMFeatureSet>[])_map.get(type);
-	             FeatureStates=AddFeatState(l[0],FeatureStates);
-	             FeatureStates= AddFeatState(l[1],FeatureStates);
-	             
[…]

### 2010-09-24 MySimpleSketch/src/SketchMaster/collection/NumericalComparator.java

@@ -1,139 +1,0 @@
-//// See bottom of source code for software license.
-
-package SketchMaster.collection;
-
-import java.io.Serializable;
-import java.util.*;
-
-/**
- * Compare Number objects.
- * 
- * <P>
- * This software is distributed under the <A
- * HREF="http://guir.cs.berkeley.edu/projects/COPYRIGHT.txt"> Berkeley Software
- * License</A>.
- * 
- * <PRE>
- * Revisions:  - GUIRLib-v1.0-1.0.0, Jun 16 1999, JH
- *               Created class
- *             - GUIRLib-v1.2-1.0.0, Jun 22 2000, JH
- *               Touched for GUIRLib release
- *             - GUIRLib-v1.3-1.0.0, Aug 11 2000, JH
- *               Touched for GUIRLib release
- *             - GUIRLib-v1.4-1.0.0, Aug 31 2000, JH
- *               Touched for GUIRLib release
- * </PRE>
- * 
- * @author (<A HREF="mailto:jasonh@cs.berkeley.edu">jasonh@cs.berkeley.edu</A>)
- * @since JDK 1.2
- * @version GUIRLib-v1.4-1.0.0, Aug 31 2000
- */
-public class NumericalComparator implements Comparator, Serializable {
-
-	// ===========================================================================
-	// === DEFAULTS AND CUSTOMIZATION ========================================
-
-	/**
-	 * 
-	 */
-	private static final long serialVersionUID = 2452111586088162364L;
-	private boolean flagUsePerturbations = false;
-
-	// -----------------------------------------------------------------
-
-	/**
-	 * Add perturbations to the values or not? Basically, if set to true, we
-	 * will add two different small values to each of the numbers before
-	 * comparison, to avoid equal values.
-	 * 
-	 * @param flag
-	 *            is true if perturbations are to be used, false otherwise.
-	 */
-	public void setUsePerturbations(boolean flag) {
-		flagUsePerturbations = flag;
-	} // of setUsePerturbations
-
-	// === DEFAULTS AND CUSTOMIZATION ========================================
-	// ===========================================================================
-
-	// ===========================================================================
-	// === COMPARISON ========================================================
-
-	public int compare(Object obj1, Object obj2) {
-		if (obj1 instanceof Number && obj2 instanceof Number) {
-			// // 1. Get numerical values.
[…]

### 2010-09-24 MySimpleSketch/src/SketchMaster/collection/SortedValueMap.java

@@ -1,530 +1,0 @@
-//// See bottom of source code for software license.
-
-package SketchMaster.collection;
-
-import org.apache.log4j.Logger;
-
-import java.util.*;
-
-/**
- * A Map with any kind of uniquely named Keys with numerical Values, sorted by
- * the values rather than by the keys. You can set whether the Map is sorted
- * each value has a threshold max value or not, and whether the sum of the
- * values has a threshold max value or not.
- * 
- * <P>
- * This software is distributed under the <A
- * HREF="http://guir.cs.berkeley.edu/projects/COPYRIGHT.txt"> Berkeley Software
- * License</A>.
- * 
- * <PRE>
- * Revisions:  - GUIRLib-v1.0-1.0.0, Jun 16 1997, JH
- *               Created class
- *             - GUIRLib-v1.2-1.0.0, Jun 22 2000, JH
- *               Touched for GUIRLib release
- *             - GUIRLib-v1.3-1.0.0, Aug 11 2000, JH
- *               Touched for GUIRLib release
- *             - GUIRLib-v1.4-1.0.0, Aug 31 2000, JH
- *               Touched for GUIRLib release
- * </PRE>
- * 
- * @author (<A HREF="mailto:jasonh@cs.berkeley.edu">jasonh@cs.berkeley.edu</A>)
- * @since JDK 1.2
- * @version GUIRLib-v1.4-1.0.0, Aug 31 2000
- */
-public class SortedValueMap {
-	/**
-	 * Logger for this class
-	 */
-	private static final Logger logger = Logger.getLogger(SortedValueMap.class);
-
-	// ===========================================================================
-	// === DEFAULTS AND CUSTOMIZATION ========================================
-
-	/**
-	 * Sort in ascending order?
-	 */
-	private boolean flagAscending = true;
-
-	// -----------------------------------------------------------------
-
-	/**
-	 * Ascending or descending?
-	 */
-	private static boolean default_flagAscending = true;
-
-	// -----------------------------------------------------------------
-
-	/**
-	 * Use this BEFORE you stick in items.
-	 * 
-	 * @param flag
-	 *            is true if ascending order, false if descending.
-	 */
-	public void setAscending(boolean flag) {
-		flagAscending = flag;
-	} // of setAscending
-
-	// -----------------------------------------------------------------
-
-	/**
-	 * Set the default value for ascending or descending order. This is the
-	 * sorting regime that new SortedNumericalMaps will use.
-	 * 
-	 * @param flag
[…]

### 2010-09-24 MySimpleSketch/src/SketchMaster/gui/DebugMessageWindow.java

@@ -1,199 +1,0 @@
-/**
- * 
- */
-package SketchMaster.gui;
-
-import java.awt.BorderLayout;
-import javax.swing.JPanel;
-import javax.swing.JFrame;
-import javax.swing.JTextArea;
-import java.awt.Rectangle;
-import java.awt.event.ActionEvent;
-import java.awt.event.ActionListener;
-
-import javax.swing.Timer;
-
-import SketchMaster.io.log.FileLog;
-import java.awt.Dimension;
-import javax.swing.JScrollPane;
-
-import java.awt.Component;
-import java.awt.Point;
-import java.awt.ComponentOrientation;
-
-/**
- * 
- * @author maha
- * 
- */
-public class DebugMessageWindow extends JFrame {
-
-	private static final long serialVersionUID = 1L;
-
-	private JPanel jContentPane = null;
-
-	private Timer timer = null; // @jve:decl-index=0:visual-constraint="224,367"
-
-	private JScrollPane jScrollPane = null;
-
-	private JTextArea jTextArea1 = null;
-
-	private JScrollPane jScrollPaneDrawing = null;
-
-	private DrawingSheet drawingSheet = null;
-
-	/**
-	 * This is the default constructor
-	 */
-	public DebugMessageWindow() {
-		super();
-		initialize();
-	}
-
-	/**
-	 * This method initializes this
-	 * 
-	 * @return void
-	 */
-	private void initialize() {
-		this.setSize(706, 305);
-		this.setContentPane(getJContentPane());
-		this.setName("Log Messages");
-		this.setTitle("JFrame");
-		this.addComponentListener(new java.awt.event.ComponentAdapter() {
-			public void componentResized(java.awt.event.ComponentEvent e) {
-				Dimension d = getContentPane().getSize();
-				getJScrollPane().setSize(d.width - 50, d.height - 50);
-				// getJTextArea().setSize(getContentPane().getSize().width-20,getContentPane().getSize().height-20);
-				// logger.info("componentResized()");
-			}
-		});
-		getTimer();
-		// timer.setDelay(10000);
-	}
-
-	/**
-	 * This method initializes jContentPane
-	 * 
-	 * @return javax.swing.JPanel
-	 */
-	private JPanel getJContentPane() {
-		if (jContentPane == null) {
-			jContentPane = new JPanel();
-			jContentPane.setLayout(null);
[…]

### 2010-09-24 MySimpleSketch/src/SketchMaster/gui/DrawingDebugUtils.java

@@ -1,336 +1,0 @@
-/**
- * 
- */
-package SketchMaster.gui;
-
-import java.awt.BasicStroke;
-import java.awt.Color;
-import java.awt.Component;
-import java.awt.Dimension;
-import java.awt.Font;
-import java.awt.Graphics;
-import java.awt.Graphics2D;
-import java.awt.geom.Point2D;
-import java.util.ArrayList;
-import javax.swing.JFrame;
-import SketchMaster.Stroke.StrokeData.PointData;
-import SketchMaster.Stroke.StrokeData.Stroke;
-import SketchMaster.Stroke.features.FeatureFunction;
-import SketchMaster.Stroke.features.StrokeFeatures;
-import SketchMaster.gui.*;
-
-/**
- * @author melmeseery
- *
- */
-public class DrawingDebugUtils {
-
-	
-//	if (DrawingDebugUtils.DEBUG_GRAPHICALLY){
-//		colorValue+=0.1;
-//		 SolutionColor=Color.getHSBColor(0.4f, (float)colorValue, 0.5f);
-//		// DrawingDebugUtils.getGraphics().setColor(SolutionColor);
-	// paint(	DrawingDebugUtils.getGraphics() );
-//		 DrawingDebugUtils.drawThickPointPath( 	DrawingDebugUtils.getGraphics()  , SolutionColor, SolutionColor, polygonVertices);
-//
-//}
-	public static final Color InkColor = Color.BLUE;
-	public static final Color PointsColor = Color.RED;
-	public static final int PointsSize = 3;
-	public static final Color segmeColor = Color.BLACK;
-    
-	
-	public static Graphics2D G_DEBUG=null;
-	private static stickFrame DebugFrame=null;
-	public static boolean DEBUG_GRAPHICALLY=true; 
-	public static boolean StateFrames=false;
-	
-	public static Graphics2D getGraphics(){
-		if (G_DEBUG!=null)
-			return G_DEBUG;
-		else {
-		         return (Graphics2D) getDebugFrame().getGraphics();
-			
-		}
-		
-	}
-	
- 
-	public static JFrame getDebugFrame() {
-			if (	DebugFrame ==null)
-			{
-		
-				DebugFrame =		 (new DrawingDebugUtils()).new stickFrame(" Debug Frame ");
-//				Font font =new Font("Serif", Font.ITALIC, 20);
-//				DebugFrame.setFont(font);
-			//	DebugFrame.setBackground(Color.white);
[…]

### 2010-09-24 MySimpleSketch/src/SketchMaster/gui/DrawingSheet.java

@@ -1,457 +1,0 @@
-package SketchMaster.gui;
-
-import org.apache.log4j.Logger;
-
-import SketchMaster.Stroke.StrokeData.PointData;
-import SketchMaster.Stroke.StrokeData.Stroke;
-import SketchMaster.Stroke.graphics.layers.SketchLayer;
-import SketchMaster.gui.Events.HandleFinishCluster;
-import SketchMaster.gui.Events.HandleStroke;
-import SketchMaster.gui.Events.NewClusterEvent;
-import SketchMaster.gui.Events.NewStrokeEvent;
-import SketchMaster.io.log.FileLog;
-import SketchMaster.system.SystemSettings;
-
-import java.awt.Color;
-import java.awt.Graphics;
-import java.awt.Graphics2D;
-import java.awt.HeadlessException;
-import java.awt.event.MouseEvent;
-import java.awt.event.MouseListener;
-import java.awt.event.MouseMotionListener;
-import java.io.File;
-import java.util.Observable;
-import java.util.Observer;
-
-import javax.swing.JFileChooser;
-import javax.swing.JOptionPane;
-import javax.swing.JPanel;
-import javax.swing.JScrollPane;
-import javax.swing.event.EventListenerList;
-import javax.swing.JPopupMenu;
-import javax.swing.JMenuItem;
-
-
-/**
- * @author Mahi
- */
-public class DrawingSheet extends JScrollPane {
-	/**
-	 * Logger for this class
-	 */
-	private static final Logger logger = Logger.getLogger(DrawingSheet.class);
-
-	GraphWatched watched = new GraphWatched();
-	// -----------------------------------parameters
-	// private SketchLayer currentLayer;
-	private Stroke currentStroke = null;
-
-	private boolean flagStrokeOn = false;
-	
-	private boolean flagClusterStart=false;
-	private int DragStrokeCount=0;
-
-	private long currentStrokeTime = 0;
-
-	private long currentStrokeStartTime = 0;
-
-	private long PrevcurrentStrokeTime = 0;
-	
-	private boolean NoNotify=false;
-	private boolean PaintToImage=false;
-	 
-
-	private JPanel panel = new JPanel();
-
-	/**
-	 * @directed
-	 */
-	public DrawingSheet() throws HeadlessException {
-		super();
-		try {
-			jbInit();
[…]

### 2010-09-24 MySimpleSketch/src/SketchMaster/gui/DrawingSheetTesting.java

@@ -1,38 +1,0 @@
-/**
- * 
- */
-package SketchMaster.gui;
-
-import java.awt.Color;
-import java.awt.Dimension;
-import java.awt.Graphics;
-import java.awt.Graphics2D;
-
-/**
- * @author Maha
- *
- */
-public class DrawingSheetTesting  extends DrawingSheet{
-
-	/**
-	 * 
-	 */
-	public DrawingSheetTesting() {
-		DrawingDebugUtils. DEBUG_GRAPHICALLY=true;
-		
-	}
-
-	/* (non-Javadoc)
-	 * @see SketchMaster.gui.DrawingSheet#paint(java.awt.Graphics)
-	 */
-	@Override
-	public void paint(Graphics g) {
-	
-		super.paint(g);
-		
-		/// draw grid lines for 
-		
-		DrawingDebugUtils.DrawGridOnGraph((Graphics2D)g,this.getSize());
-	}
-
-}

### 2010-09-24 MySimpleSketch/src/SketchMaster/gui/Events/HandleFinishCluster.java

@@ -1,15 +1,0 @@
-/**
- * 
- */
-package SketchMaster.gui.Events;
-
-import java.util.EventListener;
-
-/**
- * @author Maha
- *
- */
-public interface HandleFinishCluster extends EventListener {
-
-	public abstract void HandleFinishedCluster(NewClusterEvent Evt);
-}

### 2010-09-24 MySimpleSketch/src/SketchMaster/gui/Events/HandleStroke.java

@@ -1,9 +1,0 @@
-package SketchMaster.gui.Events;
-
-import java.util.EventListener;
-
-public interface HandleStroke extends EventListener {
-
-	public abstract void HandleNewStroke(NewStrokeEvent Evt);
-
-}

### 2010-09-24 MySimpleSketch/src/SketchMaster/gui/Events/NewStrokeEvent.java

@@ -1,111 +1,0 @@
-package SketchMaster.gui.Events;
-
-import SketchMaster.Stroke.StrokeData.Stroke;
-import java.awt.event.MouseEvent;
-import java.util.EventObject;
-
-/**
- * @author Mahi
- */
-public class NewStrokeEvent extends EventObject {
-
-	private static final long serialVersionUID = 1L;
-	Stroke EventStroke;
-	/**
-	 * @uml.property name="flagConsumed"
-	 */
-	boolean flagConsumed = false;
-	/**
-	 * @uml.property name="evt"
-	 */
-	MouseEvent evt;
-	int evtX;
-	int evtY;
-
-	// ///////////////////////////////////////////////////////////////////////////////////////////////
-	// constructors
-	public NewStrokeEvent(Object source) {
-		super(source);
-	} // of constructor
-
-	/**
-	 * 
-	 */
-	public NewStrokeEvent(Object source, Stroke stroke) {
-		this(source);
-		this.EventStroke = stroke;
-
-	} // of constructor
-
-	// -----------------------------------------------------------------
-
-	/**
-	 * Create a new StrokeEvent object.
-	 * 
-	 * @param source
-	 *            is the object creating this StrokeEvent.
-	 * @param evt
-	 *            is an associated AWT Event.
-	 */
-	public NewStrokeEvent(Object source, Stroke stroke, MouseEvent evt) {
-		this(source, stroke);
-		this.evt = evt;
-		this.evtX = evt.getX();
-		this.evtY = evt.getY();
-	} // of constructor
-
-	/**
-	 * @return the eventStroke
-	 * @uml.property name="eventStroke"
-	 */
-	public Stroke getEventStroke() {
-		return EventStroke;
-	}
-
-	/**
-	 * @param eventStroke
-	 *            the eventStroke to set
-	 * @uml.property name="eventStroke"
-	 */
-	public void setEventStroke(Stroke eventStroke) {
-		EventStroke = eventStroke;
-	}
-
-	/**
-	 * @return the evt
-	 * @uml.property name="evt"
-	 */
-	public MouseEvent getEvt() {
-		return evt;
-	}
-
-	/**
-	 * @param evt
-	 *            the evt to set
-	 * @uml.property name="evt"
-	 */
-	public void setEvt(MouseEvent evt) {
-		this.evt = evt;
-	}
-
-	/**
-	 * @return the flagConsumed
-	 * @uml.property name="flagConsumed"
-	 */
[…]

### 2010-09-24 MySimpleSketch/src/SketchMaster/gui/GraphFrame.java

@@ -1,592 +1,0 @@
-package SketchMaster.gui;
-
-
-//import jahuwaldt.plot.CircleSymbol;
-//import jahuwaldt.plot.Plot2D;
-//import jahuwaldt.plot.PlotPanel;
-//import jahuwaldt.plot.PlotRun;
-//import jahuwaldt.plot.PlotRunList;
-//import jahuwaldt.plot.SimplePlotXY;
-//import jahuwaldt.plot.SquareSymbol;
-
-
-import SketchMaster.Stroke.features.FeatureFunction;
-import SketchMaster.lib.Vector;
-
-
-import java.awt.BasicStroke;
-import java.awt.Color;
-import java.awt.Component;
-import java.awt.Graphics;
-
-import java.awt.GraphicsConfiguration;
-import java.awt.HeadlessException;
-
-import java.awt.Rectangle;
-
-import java.awt.event.FocusEvent;
-import java.awt.event.FocusListener;
-import java.awt.geom.Point2D;
-
-import java.util.ArrayList;
-import javax.swing.JFrame;
-import javax.swing.JPanel;
-import org.jfree.chart.ChartFactory;
-import org.jfree.chart.ChartPanel;
-import org.jfree.chart.JFreeChart;
-
-import org.jfree.chart.plot.PlotOrientation;
-import org.jfree.chart.plot.XYPlot;
-
-import org.jfree.chart.renderer.xy.XYItemRenderer;
-import org.jfree.chart.renderer.xy.XYLineAndShapeRenderer;
-
-import org.jfree.data.xy.XYSeries;
-import org.jfree.data.xy.XYSeriesCollection;
-
-
-/**
- * @author  Mahi
- */
-public class GraphFrame extends JFrame implements FocusListener {
-
-	// PlotPanel plotpanel=null;
-
-	/**
-	 * 
-	 */
-	private static final long serialVersionUID = 8471682471495961363L;
-
-	/**
-	 * @uml.property  name="points"
-	 */
-	ArrayList points;
-
-	// Plot2D aPlot;
-	double StartX;
-	// Plot2D aPlot;
-	double StartY;
-
-	XYSeriesCollection dataset;
-
-	/**
-	 * @uml.property  name="xAxisName"
-	 */
-	String xAxisName;
-	/**
-	 * @uml.property  name="yAxisName"
-	 */
-	String yAxisName;
-	/**
-	 * @uml.property  name="graphName"
-	 */
[…]

### 2010-09-24 MySimpleSketch/src/SketchMaster/gui/GraphWatched.java

@@ -1,43 +1,0 @@
-/**
- * 
- */
-package SketchMaster.gui;
-
-import java.awt.Graphics2D;
-import java.util.Observable;
-
-import org.apache.log4j.Logger;
-
-import SketchMaster.io.log.FileLog;
-import SketchMaster.swarm.Solution;
-
-/**
- * @author maha
- * 
- */
-public class GraphWatched extends Observable {
-	private static final Logger logger = Logger.getLogger(GraphWatched.class);
-	Graphics2D g;
-
-	void counter(int period) {
-		for (; period >= 0; period--) {
-			setChanged();
-			notifyObservers(Integer.valueOf(period));
-			try {
-				Thread.sleep(100);
-			} catch (InterruptedException e) {
-				//  logger.trace("Sleep interrupted");
-			}
-		}
-	}
-
-	public Graphics2D getG() {
-		return g;
-	}
-
-	public void setG(Graphics2D g) {
-		this.g = g;
-		setChanged();
-	}
-
-}

### 2010-09-24 MySimpleSketch/src/SketchMaster/gui/MainClass.java

@@ -1,126 +1,0 @@
-package SketchMaster.gui;
-
-import org.apache.log4j.Logger;
-
-import java.awt.BorderLayout;
-import java.awt.Label;
-
-import javax.swing.JFrame;
-
-import SketchMaster.io.log.FileLog;
-import SketchMaster.system.SystemSettings;
-
-/**
- * <p>
- * Title: SketchMasterProject
- * </p>
- * 
- * <p>
- * Description: sketch recognition
- * </p>
- * 
- * <p>
- * Copyright: Copyright (c) 2006
- * </p>
- * 
- * <p>
- * Company: engineer
- * </p>
- * 
- * @author maha el meseery
- * @version 1.0
- */
-public class MainClass {
-	/**
-	 * Logger for this class
-	 */
-	private static final Logger logger = Logger.getLogger(MainClass.class);
-
-	private static Label emptyLabel = null;
-
-	public MainClass() {
-		super();
-		try {
-			jbInit();
-		} catch (Exception ex) {
-			logger.error("MainClass()", ex); //$NON-NLS-1$
-		}
-	}
-	public static void RunApplication(){
-		FileLog.createALogFile();
-		org.apache.log4j.PropertyConfigurator.configure("log4j.properties");
-		//  logger.trace("Start program");
-
-		//  logger.trace(" before sketch");
-	//	MainClass main = new MainClass();
-
-		emptyLabel = new Label("Draw what you want ");
-		// 1. Optional: Specify who draws the window decorations.
-		// SketchFrame.setDefaultLookAndFeelDecorated(true);
-		SystemSettings.MODE=SystemSettings.MODE_NORMAL;
-		// 2. Create the frame.
-		SketchMaterApplication frame = new SketchMaterApplication();
-
-		// 3. Optional: What happens when the frame closes?
-		frame.setDefaultCloseOperation(JFrame.EXIT_ON_CLOSE);
-
-		// 4. Create components and put them in the frame.
-		// ...create emptyLabel...
-		// frame.getContentPane().add(emptyLabel, BorderLayout.CENTER);
-
-		// 5. Size the frame.
-		// / frame.pack();
-		// frame.getContentPane().setLayout(null);
-
-		// frame.setSize(1000, 1000);
-		// 6. Show it.
-
-		// frame.init();
-		// frame.addlistnerToFrame();
-
-		frame.setVisible(true);
-
-	}
-
-	public static void main(String[] args) {
-		
-		RunApplication();
-		
[…]

### 2010-09-24 MySimpleSketch/src/SketchMaster/gui/NumericalComparator.java

@@ -1,139 +1,0 @@
-//// See bottom of source code for software license.
-
-package SketchMaster.gui;
-
-import java.io.Serializable;
-import java.util.*;
-
-/**
- * Compare Number objects.
- * 
- * <P>
- * This software is distributed under the <A
- * HREF="http://guir.cs.berkeley.edu/projects/COPYRIGHT.txt"> Berkeley Software
- * License</A>.
- * 
- * <PRE>
- * Revisions:  - GUIRLib-v1.0-1.0.0, Jun 16 1999, JH
- *               Created class
- *             - GUIRLib-v1.2-1.0.0, Jun 22 2000, JH
- *               Touched for GUIRLib release
- *             - GUIRLib-v1.3-1.0.0, Aug 11 2000, JH
- *               Touched for GUIRLib release
- *             - GUIRLib-v1.4-1.0.0, Aug 31 2000, JH
- *               Touched for GUIRLib release
- * </PRE>
- * 
- * @author (<A HREF="mailto:jasonh@cs.berkeley.edu">jasonh@cs.berkeley.edu</A>)
- * @since JDK 1.2
- * @version GUIRLib-v1.4-1.0.0, Aug 31 2000
- */
-public class NumericalComparator implements Comparator, Serializable {
-
-	// ===========================================================================
-	// === DEFAULTS AND CUSTOMIZATION ========================================
-
-	/**
-	 * 
-	 */
-	private static final long serialVersionUID = -1909340865767706904L;
-	private boolean flagUsePerturbations = false;
-
-	// -----------------------------------------------------------------
-
-	/**
-	 * Add perturbations to the values or not? Basically, if set to true, we
-	 * will add two different small values to each of the numbers before
-	 * comparison, to avoid equal values.
-	 * 
-	 * @param flag
-	 *            is true if perturbations are to be used, false otherwise.
-	 */
-	public void setUsePerturbations(boolean flag) {
-		flagUsePerturbations = flag;
-	} // of setUsePerturbations
-
-	// === DEFAULTS AND CUSTOMIZATION ========================================
-	// ===========================================================================
-
-	// ===========================================================================
-	// === COMPARISON ========================================================
-
-	public int compare(Object obj1, Object obj2) {
-		if (obj1 instanceof Number && obj2 instanceof Number) {
-			// // 1. Get numerical values.
[…]

### 2010-09-24 MySimpleSketch/src/SketchMaster/gui/SketchMaterApplication.java

@@ -1,717 +1,0 @@
-package SketchMaster.gui;
-
-import org.apache.log4j.Logger;
-
-import java.awt.BorderLayout;
-import java.awt.Component;
-import java.awt.Dimension;
-import java.awt.Graphics2D;
-
-import javax.swing.JCheckBoxMenuItem;
-import javax.swing.JFileChooser;
-import javax.swing.JMenu;
-import javax.swing.JMenuBar;
-import javax.swing.JMenuItem;
-import javax.swing.JOptionPane;
-import javax.swing.JPanel;
-import javax.swing.JFrame;
-import javax.swing.JScrollPane;
-import javax.swing.KeyStroke;
-
-import java.awt.Rectangle;
-import java.awt.event.ActionEvent;
-import java.awt.event.KeyEvent;
-import java.io.File;
-
-import javax.swing.JButton;
-import javax.swing.JTextField;
-import javax.swing.JLabel;
-import javax.swing.filechooser.FileFilter;
-
-import SketchMaster.gui.Events.HandleStroke;
-import SketchMaster.gui.Events.NewStrokeEvent;
-import SketchMaster.gui.optionDialogs.CategoryAdd;
-import SketchMaster.gui.optionDialogs.LayerSelectionDialog;
-import SketchMaster.io.log.FileLog;
-import SketchMaster.system.SketchRecognitionSystem;
-import SketchMaster.system.SystemSettings;
-
-public class SketchMaterApplication extends JFrame {
-	/**
-	 * Logger for this class
-	 */
-	private static final Logger logger = Logger.getLogger(SketchMaterApplication.class);
-
-	private static final long serialVersionUID = 1L;
-	final JFileChooser fc = new JFileChooser(new File("."));
-
-	private JPanel jContentPane = null;
-
-	private JScrollPane DrawingScrollSheet = null;
-
-	// Create the menu bar.
-	private JMenuBar ApplicationMenuBar;
-
-	private JLabel NumberOfStrokeLabel = null;
-
-	private JLabel TimeStrokeLabel = null;
-
-	private JLabel CountPointStrokeLabel = null;
-
-	private DrawingSheet DrawSheet = null;
-
-	private static DebugMessageWindow DebugWindow;
-	private int DebugWindowCount;
-	SketchRecognitionSystem system; // @jve:decl-index=0:
-	private JButton jClusterButton = null;
-	private JButton jButtonClear = null;
-	private JLabel jLabelnStroke = null;
[…]

### 2010-09-24 MySimpleSketch/src/SketchMaster/gui/SortedValueMap.java

@@ -1,530 +1,0 @@
-//// See bottom of source code for software license.
-
-package SketchMaster.gui;
-
-import org.apache.log4j.Logger;
-
-import java.util.*;
-
-/**
- * A Map with any kind of uniquely named Keys with numerical Values, sorted by
- * the values rather than by the keys. You can set whether the Map is sorted
- * each value has a threshold max value or not, and whether the sum of the
- * values has a threshold max value or not.
- * 
- * <P>
- * This software is distributed under the <A
- * HREF="http://guir.cs.berkeley.edu/projects/COPYRIGHT.txt"> Berkeley Software
- * License</A>.
- * 
- * <PRE>
- * Revisions:  - GUIRLib-v1.0-1.0.0, Jun 16 1997, JH
- *               Created class
- *             - GUIRLib-v1.2-1.0.0, Jun 22 2000, JH
- *               Touched for GUIRLib release
- *             - GUIRLib-v1.3-1.0.0, Aug 11 2000, JH
- *               Touched for GUIRLib release
- *             - GUIRLib-v1.4-1.0.0, Aug 31 2000, JH
- *               Touched for GUIRLib release
- * </PRE>
- * 
- * @author (<A HREF="mailto:jasonh@cs.berkeley.edu">jasonh@cs.berkeley.edu</A>)
- * @since JDK 1.2
- * @version GUIRLib-v1.4-1.0.0, Aug 31 2000
- */
-public class SortedValueMap {
-	/**
-	 * Logger for this class
-	 */
-	private static final Logger logger = Logger.getLogger(SortedValueMap.class);
-
-	// ===========================================================================
-	// === DEFAULTS AND CUSTOMIZATION ========================================
-
-	/**
-	 * Sort in ascending order?
-	 */
-	private boolean flagAscending = true;
-
-	// -----------------------------------------------------------------
-
-	/**
-	 * Ascending or descending?
-	 */
-	private static boolean default_flagAscending = true;
-
-	// -----------------------------------------------------------------
-
-	/**
-	 * Use this BEFORE you stick in items.
-	 * 
-	 * @param flag
-	 *            is true if ascending order, false if descending.
-	 */
-	public void setAscending(boolean flag) {
-		flagAscending = flag;
-	} // of setAscending
-
-	// -----------------------------------------------------------------
-
-	/**
-	 * Set the default value for ascending or descending order. This is the
-	 * sorting regime that new SortedNumericalMaps will use.
-	 * 
-	 * @param flag
[…]

### 2010-09-24 MySimpleSketch/src/SketchMaster/gui/optionDialogs/CategoryAdd.java

@@ -1,198 +1,0 @@
-package SketchMaster.gui.optionDialogs;
-
-import org.apache.log4j.Logger;
-
-import javax.swing.JPanel;
-import java.awt.Frame;
-import java.awt.BorderLayout;
-import javax.swing.JDialog;
-import javax.swing.JFrame;
-import javax.swing.JOptionPane;
-import javax.swing.JTextField;
-import javax.swing.JLabel;
-import javax.swing.JButton;
-import java.awt.Rectangle;
-
-public class CategoryAdd extends JDialog {
-	/**
-	 * Logger for this class
-	 */
-	private static final Logger logger = Logger.getLogger(CategoryAdd.class);
-
-	private static final long serialVersionUID = 1L;
-
-	private JPanel jContentPane = null;
-
-	private JTextField jTextCategoryName = null;
-
-	private JLabel jLabel = null;
-
-	private JButton jButton = null;
-
-	private JButton jButton1 = null;
-	private String[] CurrentCategorys = null;
-	private String NewCategory = null;
-
-	/**
-	 * @param owner
-	 */
-	public CategoryAdd(Frame owner) {
-		super(owner);
-		initialize();
-	}
-
-	/**
-	 * This method initializes this
-	 * 
-	 * @return void
-	 */
-	private void initialize() {
-		this.setSize(400, 200);
-		this.setDefaultCloseOperation(JFrame.DISPOSE_ON_CLOSE);
-		this.setModal(true);
-
-		this.setContentPane(getJContentPane());
-		this.setVisible(false);
-	}
-
-	/**
-	 * This method initializes jContentPane
-	 * 
-	 * @return javax.swing.JPanel
-	 */
-	private JPanel getJContentPane() {
-		if (jContentPane == null) {
-			jLabel = new JLabel();
-			jLabel.setText("Category Name ");
-			jLabel.setBounds(new Rectangle(6, 42, 151, 23));
-			jContentPane = new JPanel();
-			jContentPane.setLayout(null);
-			jContentPane.add(getJTextCategoryName(), null);
-			jContentPane.add(jLabel, null);
-			jContentPane.add(getJButton(), null);
-			jContentPane.add(getJButton1(), null);
-		}
-		return jContentPane;
-	}
-
-	/**
-	 * This method initializes jTextCategoryName
-	 * 
-	 * @return javax.swing.JTextField
-	 */
-	private JTextField getJTextCategoryName() {
-		if (jTextCategoryName == null) {
-			jTextCategoryName = new JTextField();
[…]

### 2010-09-24 MySimpleSketch/src/SketchMaster/gui/optionDialogs/LayerSelectionDialog.java

@@ -1,281 +1,0 @@
-package SketchMaster.gui.optionDialogs;
-
-import javax.swing.JPanel;
-
-import java.awt.Component;
-import java.awt.Dialog;
-import java.awt.Frame;
-import java.awt.BorderLayout;
-import java.awt.HeadlessException;
-
-import javax.swing.JDialog;
-import javax.swing.JCheckBox;
-import javax.swing.JFrame;
-
-import java.awt.Rectangle;
-import javax.swing.JButton;
-import javax.swing.JOptionPane;
-import javax.swing.JScrollPane;
-import javax.swing.JLabel;
-import javax.swing.ScrollPaneConstants;
-
-import java.awt.GridBagLayout;
-import java.awt.Dimension;
-import javax.swing.BorderFactory;
-import java.awt.Color;
-
-public class LayerSelectionDialog extends JDialog {
-
-	private static final long serialVersionUID = 1L;
-
-	private JPanel jContentPane = null;
-
-	private JScrollPane jScrollPane = null;
-
-	private JPanel jPanel = null;
-
-	private JButton jButton = null;
-
-	private JLabel jLabel = null;
-	String[] Layers = null;
-	boolean[] layersSelections = null;
-
-	private int checkBoxCount = 0;
-
-	/**
-	 * @param owner
-	 */
-	public LayerSelectionDialog(Frame owner, String[] LayerNames) {
-		super(owner);
-
-		// Layers=LayerNames;
-		// layersSelections=new boolean[Layers.length];
-		// for (int i = 0; i < layersSelections.length; i++) {
-		// layersSelections[i]=false;
-		// }
-		initialize();
-
-		setModal(true);
-		// setSize(250, 330);
-		// setLocation(20, 100);
-
-		setTitle("Choose layers to display");
-		setDefaultCloseOperation(JFrame.DISPOSE_ON_CLOSE);
-
-		// setDefaultClosesOperation(DO_NOTHING_ON_CLOSE);
-		addAllJCheckBox(getJPanel(), LayerNames);
-		this.Layers = LayerNames;
-	}
-
-	/**
-	 * @param arg0
-	 * @param layers
-	 * @param layersSelections
-	 * @throws HeadlessException
-	 */
-	public LayerSelectionDialog(Frame arg0, String[] layers,
-			boolean[] layersSelections) throws HeadlessException {
-		super(arg0);
-
-		this.Layers = layers;
[…]

### 2010-09-24 MySimpleSketch/src/SketchMaster/io/FileObjectReader.java

@@ -1,106 +1,0 @@
-/**
- * 
- */
-package SketchMaster.io;
-
-import org.apache.log4j.Logger;
-
-import java.io.FileInputStream;
-import java.io.FileNotFoundException;
-import java.io.IOException;
-import java.io.ObjectInputStream;
-
-import SketchMaster.io.log.FileLog;
-
-/**
- * @author maha
- * 
- */
-public class FileObjectReader {
-	/**
-	 * Logger for this class
-	 */
-	private static final Logger logger = Logger.getLogger(FileObjectReader.class);
-
-	// File StrokeFile;
-	FileInputStream f_in;
-
-	ObjectInputStream obj_in;
-
-	/**
-	 * 
-	 */
-	public FileObjectReader() {
-
-	}
-
-	public void closeRead() {
-		try {
-			f_in.close();
-		} catch (IOException e) {
-
-			logger.error("closeRead()", e); //$NON-NLS-1$
-		}
-	}
-
-	public boolean readFile(String FileName) {
-		try {
-			f_in = new FileInputStream(FileName);
-
-		} catch (FileNotFoundException e) {
-
-			logger.error("readFile(String)", e); //$NON-NLS-1$
-			return false;
-		}
-
-		return true;
-	}
-
-	public Object ReadObject() {
-		Object obj = null;
-
-		try {
-
-			obj_in = new ObjectInputStream(f_in);
-
-			obj = obj_in.readObject();
-			//  logger.trace("REading ");
-
-		} catch (FileNotFoundException e) {
-
-			logger.error("ReadObject()", e); //$NON-NLS-1$
-		} catch (IOException e) {
-
-			// logger.info("in exception");
-			logger.error("ReadObject()", e); //$NON-NLS-1$
-
-		} catch (ClassNotFoundException e) {
-			logger.error("ReadObject()", e); //$NON-NLS-1$
-		} catch (Exception e) {
-			logger.error("ReadObject()", e); //$NON-NLS-1$
-		}
-
-		return obj;
-	}
-
-	public ObjectInputStream getOpenedObjectStream() {
-		try {
-
-			// Write object with ObjectOutputStream
-			obj_in = new ObjectInputStream(f_in);
-
-			// Write object out to disk
-
-			return obj_in;
-
-		} catch (FileNotFoundException e) {
-
-			logger.error("getOpenedObjectStream()", e); //$NON-NLS-1$
-		} catch (IOException e) {
-
[…]

### 2010-09-24 MySimpleSketch/src/SketchMaster/io/FileObjectWriter.java

@@ -1,99 +1,0 @@
-/**
- * 
- */
-package SketchMaster.io;
-
-import org.apache.log4j.Logger;
-
-import java.io.FileNotFoundException;
-import java.io.FileOutputStream;
-import java.io.IOException;
-import java.io.ObjectOutputStream;
-
-import SketchMaster.Stroke.StrokeData.Stroke;
-import SketchMaster.io.log.FileLog;
-
-/**
- * @author maha
- * 
- */
-public class FileObjectWriter {
-	/**
-	 * Logger for this class
-	 */
-	private static final Logger logger = Logger.getLogger(FileObjectWriter.class);
-
-	FileOutputStream f_out;
-
-	ObjectOutputStream obj_out;
-
-	/**
-	 * 
-	 */
-	public FileObjectWriter() {
-
-	}
-
-	public void createFile(String string) {
-
-		try {
-			// Write to disk with FileOutputStream
-			f_out = new FileOutputStream(string);
-
-		} catch (FileNotFoundException e) {
-
-			logger.error("createFile(String)", e); //$NON-NLS-1$
-		}
-	}
-
-	public void closeWrite() {
-		try {
-			f_out.close();
-		} catch (IOException e) {
-
-			logger.error("closeWrite()", e); //$NON-NLS-1$
-		}
-	}
-
-	public void WriteObject(Object obj) {
-		try {
-
-			// Write object with ObjectOutputStream
-			obj_out = new ObjectOutputStream(f_out);
-
-			// Write object out to disk
-			obj_out.writeObject(obj);
-			//  logger.trace("writing the object");
-
-		} catch (FileNotFoundException e) {
-
-			logger.error("WriteObject(Object)", e); //$NON-NLS-1$
-		} catch (IOException e) {
-
-			logger.error("WriteObject(Object)", e); //$NON-NLS-1$
-		}
-
-		// if nothing go wrongs then i will wrie all the date in the stroke
-	}
-
-	public ObjectOutputStream getOpenedObjectStream() {
-		try {
-
-			// Write object with ObjectOutputStream
-			obj_out = new ObjectOutputStream(f_out);
-
-			// Write object out to disk
-
-			return obj_out;
-
-		} catch (FileNotFoundException e) {
-
-			logger.error("getOpenedObjectStream()", e); //$NON-NLS-1$
-		} catch (IOException e) {
-
-			logger.error("getOpenedObjectStream()", e); //$NON-NLS-1$
-		}
-		return null;
-	}
-
-}

### 2010-09-24 MySimpleSketch/src/SketchMaster/io/db/DataBaseManager.java

@@ -1,178 +1,0 @@
-/**
- * 
- */
-package SketchMaster.io.db;
-
-import org.apache.log4j.Logger;
-
-import java.sql.Array;
-import java.sql.ResultSet;
-import java.sql.SQLException;
-import java.util.ArrayList;
-
-import org.postgresql.geometric.PGpath;
-
-import SketchMaster.Stroke.StrokeData.PointData;
-import SketchMaster.Stroke.StrokeData.Stroke;
-
-/**
- * @author Maha
- *
- */
-public class DataBaseManager {
-	/**
-	 * Logger for this class
-	 */
-	private static final Logger logger = Logger.getLogger(DataBaseManager.class);
-
-	/**
-	 * @param args
-	 */
-	PostQreDataBaseConnector db=new PostQreDataBaseConnector();
-	
-	public void setSketchOption(int id){
-		db.setSketchOption(id);
-		
-	}
-	public void setAuthorOption(int id){
-		
-		db.setAuthorOption(id);
-	}
-	
-	
-	public ArrayList<Stroke> getStrokes(){
-		
-		ArrayList<Stroke> strokes=new ArrayList<Stroke>();
-		
-//		logger.info("connect to the database"+" (" + this.getClass().getSimpleName()
-//				+ "    " + (new Throwable()).getStackTrace()[0].getLineNumber()
-//				+ "  )  ");
-		db.ConnectDB();
-//		logger.info("get strokes "+" (" + this.getClass().getSimpleName()
-//				+ "    " + (new Throwable()).getStackTrace()[0].getLineNumber()
-//				+ "  )  ");
-		ResultSet rs=db.getStrokes();
-		Stroke s;
-		if (rs!=null){
-		
-			try {
-				int countS=0;
-				 while (rs.next()) {
-						
-					 
-					 
-//						logger.info("adding a new storke "+countS+" ("
-//								+ this.getClass().getSimpleName()
-//								+ "    "
-//								+ (new Throwable()).getStackTrace()[0]
-//										.getLineNumber() + "  )  ");
-						countS++;
-						 s=new Stroke();
-						
-						PGpath path;
-				path = (PGpath)rs.getObject("path");
-			Array timesArray=(Array)rs.getArray("times");
-			//ArrayList<PointData> points = s.getPoints();
-			
-			int[] times=(int[])timesArray.getArray();
-				
[…]

### 2010-09-24 MySimpleSketch/src/SketchMaster/io/db/PostQreDataBaseConnector.java

@@ -1,353 +1,0 @@
-/**
- * 
- */
-package SketchMaster.io.db;
-
-import org.apache.log4j.Logger;
-
-import java.sql.DriverManager;
-import java.sql.Connection;
-import java.sql.ResultSet;
-import java.sql.SQLException;
-import java.sql.Statement;
-import java.util.ArrayList;
-
-import org.postgresql.geometric.*;
-import org.postgresql.jdbc3.Jdbc3Array;
-import org.postgresql.PGConnection;
-import org.postgresql.fastpath.*;
-
-
-/**
- * @author Maha
- *
- */
-public class PostQreDataBaseConnector {
-	/**
-	 * Logger for this class
-	 */
-	private static final Logger logger = Logger.getLogger(PostQreDataBaseConnector.class);
-
-	
-	
-	static transient String connString="jdbc:postgresql://localhost/etchasketches";
-	static  String username="Maha";
-	static String password= "sketch";
-	static Connection conn = null;
-	static String strokeSelect="SELECT * FROM strokes ";
-	private int Strokeid=-1;
-	private int SketchId=-1;
-	private int AuthorId=-1;
-	private int StudyId=-1;
-	static String sketchSelect="SELECT * FROM sketches ";
-	static String authorsSelect="SELECT * FROM authors ";
-	
-	
-	
-	public PostQreDataBaseConnector(){
-		try {
-			  Class.forName("org.postgresql.Driver");
-			} catch (ClassNotFoundException cnfe) {
-			logger.error("PostQreDataBaseConnector() - Couldn't find driver class: (" + this.getClass().getSimpleName() + "    " + (new Throwable()).getStackTrace()[0].getLineNumber() + "  )  ", cnfe); //$NON-NLS-1$ //$NON-NLS-2$ //$NON-NLS-3$
-			logger.error("PostQreDataBaseConnector()", cnfe); //$NON-NLS-1$
-			}
-		 
-	 }
-	public void ConnectDB(){
-		
-		
-		  
-		  try {
-		    // The second and third arguments are the username and password,
-		    // respectively. They should be whatever is necessary to connect
-		    // to the database.
-		    conn = DriverManager.getConnection(connString,
[…]

### 2010-09-24 MySimpleSketch/src/SketchMaster/io/log/FileLog.java

@@ -1,90 +1,0 @@
-/**
- * 
- */
-package SketchMaster.io.log;
-
-import org.apache.log4j.Logger;
-
-import java.io.File;
-import java.io.FileWriter;
-import java.io.IOException;
-
-/**
- * @author Mahi
- */
-public class FileLog {
-	/**
-	 * Logger for this class
-	 */
-	private static final Logger logger = Logger.getLogger(FileLog.class);
-
-	static FileWriter logFile;
-
-	static String logString;
-
-	/**
-	 * 
-	 */
-	public FileLog() {
-
-	}
-
-	public static void createALogFile() {
-		try {
-			logFile = new FileWriter(new File("log.txt"));
-			logString = "";
-		} catch (IOException e) {
-
-			logger.error("createALogFile()", e); //$NON-NLS-1$
-			logString += e.getMessage();
-		}
-	}
-
-	public static void createALogFile(String FileName) {
-		try {
-			logFile = new FileWriter(new File(FileName));
-			logString = "";
-		} catch (IOException e) {
-
-			logger.error("createALogFile(String)", e); //$NON-NLS-1$
-			logString += e.getMessage();
-		}
-	}
-
-	public static void addString(String line) {
-		// try {
-		// logFile.write(line);
-		//
-		// logFile.write(" \n\n");
-		logString += line;
-		logString += "\n";
-		// }
-		// catch (IOException e) {
-
-		// e.printStackTrace();
-		// logString += e.getMessage();
-		// }
-	}
-
-	/**
-	 * @return the logString
-	 * @uml.property name="logString"
-	 */
-	public static String getLogString() {
-		return logString;
-	}
-
-	public static void closeFile() {
-		try {
-			logFile.write(logString);
-			logFile.close();
-		} catch (IOException e) {
-
-			logger.error("closeFile()", e); //$NON-NLS-1$
-			logString += e.getMessage();
-		}
-	}
-
-	// / //  logger.trace("\n----------------------------------------------");
-	// / //  logger.trace("");
-}

### 2010-09-24 MySimpleSketch/src/SketchMaster/io/xml/XmlManager.java

@@ -1,110 +1,0 @@
-package SketchMaster.io.xml;
-
-import java.util.ArrayList;
-import java.util.Iterator;
-import java.util.List;
-
-import org.dom4j.Element;
-
-import SketchMaster.Stroke.StrokeData.Stroke;
-
-public class XmlManager {
-///  this control the xml manngment of the database. 
-	 XmlParser xml=null;
-		/**
-		 * 
-		 */
-		public XmlManager() {
-			
-		}
-	public  void readXML(String filename){
-		if (xml==null)
-		xml=new XmlParser();
-		xml.ReadXmlFile(filename);
-		//xml.iterateRootChildren();
-	}
-	public void closeFile(){
-		xml.closeParser();
-	}
-
-	
-    public void SaveXML(String filename){
-    	if (xml==null)
-    		xml=new XmlParser();
-    		xml.WriteXmlFile(filename);
-    }
-    
-
-     
-    
-	public ArrayList<String> getCategoriesList() {
-		
-		if (xml==null)
-		return null;
-		else {
-			return (ArrayList<String>) xml.getTypes();
-		//	return new ArrayList<String>();
-		}
-		
-	}
-	public ArrayList<ArrayList<Stroke>> getExampleForCat(String string) {
-//		 ArrayList<ArrayList<Stroke>> tem = getExampleForCatPosNeg(string,-1);
-//		  tem.addAll( getExampleForCatPosNeg(string,1) );
-		return   getExampleForCatPosNeg(string,1);
-	}
-	
-	public ArrayList<ArrayList<Stroke>> getExampleForCatPosNeg(String string,int posNeg) {
-		if (xml==null)
-		        return new ArrayList<ArrayList<Stroke>>();
-		
-		
-		String pos="";
-		if (posNeg<0)
-			pos="-";
-		else 
-			
-			pos="+";
-		
-		List samples=xml.getElements(string);
-		
-		ArrayList<ArrayList<Stroke>> result=new ArrayList<ArrayList<Stroke>>();
-		for (Iterator iterator = samples.iterator(); iterator.hasNext();) {
-			Element e = (Element) iterator.next();
-			ArrayList<Stroke> temp = xml.getExampleFromElements(e,pos);
-			if (temp!=null)
-			   result.add(temp);
-			
-		}
-		
-		
-		return result;
-		
-	}
-	
-	public void openParser() {
-		  
-		if (xml==null)
-    		xml=new XmlParser();
[…]

### 2010-09-24 MySimpleSketch/src/SketchMaster/lib/ArrayLibrary.java

@@ -1,88 +1,0 @@
-/**
- * 
- */
-package SketchMaster.lib;
-
-import java.util.ArrayList;
-
-/**
- * @author maha
- *         <h2></h2>
- */
-public class ArrayLibrary {
-	static public void addSorted(ArrayList list, int index) {
-		int temp;
-		for (int i = 0; i < list.size(); i++) {
-			temp = (Integer) list.get(i);
-
-			if (temp > index) {
-				list.add(i + 1, index);
-				return;
-			}
-		}
-		// upp till now it is the largest
-		// add it now
-		list.add(index);
-	}
-
-	static public boolean indexExist(ArrayList list, int index) {
-		int temp;
-		for (int i = 0; i < list.size(); i++) {
-			temp = (Integer) list.get(i);
-			if (temp == index)
-				return true;
-			if (temp > index)
-				return false;
-		}
-
-		return false;
-	}
-
-	static public ArrayList MergeSortedListNonRepeating(
-			ArrayList<Comparable> list1, ArrayList<Comparable> list2) {
-		// merge list2 into list 1
-
-		// first check if list 1 exist
-		if (list1 == null)
-			return list2;
-		else {
-			int previous = 0;
-			int compare = 0;
-			for (int i = 0; i < list2.size(); i++) {
-				for (int j = previous; j < list1.size(); j++) {
-					compare = list1.get(j).compareTo(list2.get(i));
-
-					if (compare >= 0) {
-						if (compare == 0) {
-							// now if they are equal skip this
-							// lopp and set previous with next j
-							previous = j + 1;
-							break;
-						}
-						// theitem in list2 is larger than the one in list1
-						// and we have to add it to list 1 in the previous
-						// location
-						else {
-							previous = j - 1;
-							if (previous >= 0)
-								list1.add(previous, list2.get(i));
-							else {
-								list1.add(0, list2.get(i));
-								previous = 0;
-
-							}
-							break;
-
-						}
-					}
-
-				}
-			}
-			// now loop arround list
-			return list1;
-
-		}
-
-	}
-
-}

### 2010-09-24 MySimpleSketch/src/SketchMaster/lib/ComputationsGeometry.java

@@ -1,2692 +1,0 @@
-/**
- * 
- */
-package SketchMaster.lib;
-
-import java.awt.Color;
-import java.awt.Point;
-import java.awt.geom.Point2D;
-import java.awt.geom.Rectangle2D;
-import java.util.ArrayList;
-
-import SketchMaster.Stroke.StrokeData.PointData;
-import SketchMaster.Stroke.StrokeData.Stroke;
-import SketchMaster.Stroke.graphics.shapes.Line;
-import SketchMaster.gui.DrawingDebugUtils;
-import SketchMaster.system.SystemSettings;
-
-import java.lang.*;
-
-import org.apache.log4j.Logger;
-
-/**
- * @author Mahi
- * 
- */
-
-/**
- * 
- */
-
-/*
- * (C) 2004 - Geotechnical Software Services
- * 
- * This code is free software; you can redistribute it and/or modify it under
- * the terms of the GNU Lesser General Public License as published by the Free
- * Software Foundation; either version 2.1 of the License, or (at your option)
- * any later version.
- * 
- * This code is distributed in the hope that it will be useful, but WITHOUT ANY
- * WARRANTY; without even the implied warranty of MERCHANTABILITY or FITNESS FOR
- * A PARTICULAR PURPOSE. See the GNU Lesser General Public License for more
- * details.
- * 
- * You should have received a copy of the GNU Lesser General Public License
- * along with this program; if not, write to the Free Software Foundation, Inc.,
- * 59 Temple Place - Suite 330, Boston, MA 02111-1307, USA.
- */
-// package no.geosoft.cc.geometry;
-/**
- * Collection of geometry utility methods. All methods are static.
- * 
- * @author <a href="mailto:jacob.dreyer@geosoft.no">Jacob Dreyer</a>
- */
-public final class ComputationsGeometry {
-
-	private static boolean debug=true; 
-	private static final Logger logger = Logger.getLogger(ComputationsGeometry.class);
-	public ComputationsGeometry() {
-
-	}
-
-	/**
-	 * Return true if c is between a and b.
-	 */
-//	@SuppressWarnings("unused")
-//	private static boolean isBetween(int a, int b, int c) {
-//		return b > a ? c >= a && c <= b : c >= b && c <= a;
-//	}
-
-	/**
-	 * Return true if c is between a and b.
-	 */
-	private static boolean isBetween(double a, double b, double c) {
[…]

### 2010-09-24 MySimpleSketch/src/SketchMaster/lib/CurveFitData.java

@@ -1,281 +1,0 @@
-/**
- * 
- */
-package SketchMaster.lib;
-
-import java.io.Serializable;
-import java.util.ArrayList;
-
-import SketchMaster.Stroke.StrokeData.PointData;
-
-/**
- * @author TOSHIBA
- *
- */
-public class CurveFitData implements Serializable {
-
-	/**
-	 * 
-	 */
-	private static final long serialVersionUID = -3077801431715427080L;
-	public double Ex,Ey,Exy,Exx,Eyy,Ex3,Ey3,	Ex2y,	Exy2;
-	public double Ex_2,Ey_2;
-   public int N;//number of points...
-   public double oneOverN;
-	public boolean computed=false;
-	public	double slope;
-	public double  intercept;
-	
-	public double cx, cy;
-	
-	public double mcr;
-	
-//	public void fitLine(ArrayList<PointData> points){
-//		  double dem=(N* Exx )- Ex_2;
-//			
-//			  slope =((N*Exy)-(Ey*Ex))/(dem);
-//			
-//			  intercept=(((Ey*Exx)-(Exy*Ex))/(dem));
-//	}
-	
-	public void computeInitalDat(ArrayList<PointData> points){
-		PointData pointk;
-		double k, c, tempxy, tempx, tempy, tempx2, tempy2 = 0, tempx2y, tempxy2;
-		tempxy = tempx = tempy = tempx2 = tempx2y = tempxy2 = 0.0;
-		double tempx3 = 0, tempy3 = 0;
-
-		double xi, yi;
-		if (points!=null){
-		
-		for (int i = 0; i < points.size(); i++) {
-
-			 pointk = (PointData) points.get(i);
-		 
-			xi = pointk.getX();
-			yi = pointk.getY();
-			tempx += xi;
-			tempy += yi;
-			tempxy += xi * yi;
-			tempx2 += xi * xi;
-			tempy2 += yi * yi;
-			tempx2y += (xi * xi) * yi;
-			tempxy2 += xi * (yi * yi);
-			tempx3 += xi * xi * xi;
-			tempy3 += yi * yi * yi;
-
-		}
-		N=points.size();
-		oneOverN=(double)1.0/(double)N;
-		Ex=tempx;
-		Ey=tempy;
-		
-		
-		Exy=tempxy;
-		Exx=tempx2;
[…]

### 2010-09-24 MySimpleSketch/src/SketchMaster/lib/GeneralUtils.java

@@ -1,128 +1,0 @@
-package SketchMaster.lib;
-
-import java.util.regex.Pattern;
-
-import SketchMaster.system.SystemSettings;
-
- 
-
-public class GeneralUtils {
-	public  static final String SEPERATOR_WINDOWS="\\";
-public static final String SEPERATOR_LINUX="/";
-
-
-public static int getCurrentOs(){
-	
-	
-	   String nameOS = "os.name";        
-	   String osName =System.getProperty(nameOS);
-	   
-	if (osName.startsWith("Lin"))
-	{
-		SystemSettings.OS=SystemSettings.OS_LINUX;
-		return SystemSettings.OS_LINUX;
-		
-	}
-	else {
-		SystemSettings.OS=SystemSettings.OS_WINDOWS;
-		return SystemSettings.OS_WINDOWS;
-	}
-	
-}
-
-public static String CorrectPath(String path, int toOs){
-	
-	if (	 toOs == SystemSettings.OS_WINDOWS)
-	{
-		return toWindows(path);
-	}
-	else {
-		return toLinux(path);
-	}
-}
-	
-public static String CorrectPath(String path){
-		
-		if (	SystemSettings.OS == SystemSettings.OS_WINDOWS)
-		{
-			return toWindows(path);
-		}
-		else {
-			return toLinux(path);
-		}
-	}
-	
-	public static String toWindows(String path){
-		// check if is already windows 
-		// get the letter path... 
-		//System.out.println(" the path is "+path+"  SEPERATOR_LINUX  "+SEPERATOR_LINUX);
-		if (path.contains(SEPERATOR_WINDOWS)&& path.contains(":"))
-		{
-			
-			//D:\\AUC\\Databases\\Arabic Digits Databases\\AHDBase\\
-			return path;
-		}
-		else {
-			if(path.contains(SEPERATOR_LINUX)){
-			String 	temp=path.replace("/windows/","");
-			//Pattern.compile(SEPERATOR_LINUX).matcher(temp).replaceAll(SEPERATOR_WINDOWS);
-			//temp=temp.replaceAll( SEPERATOR_LINUX,SEPERATOR_WINDOWS);
-		
-			int index=temp.indexOf(SEPERATOR_WINDOWS);
-		//System.out.println(" the windows seperator is in "+index);
-			String temp2=temp;
-		    if (index>0){
-		    	temp2="";
-		    	temp2=temp.substring(0, index-1)+":";
[…]

### 2010-09-24 MySimpleSketch/src/SketchMaster/lib/GeometryUtil.java

@@ -1,168 +1,0 @@
-/*
- * $Id: Util.java,v 1.8 2003/11/21 00:24:47 hwawen Exp $
- *
- * Copyright (c) 2003 The Regents of the University of California.
- * All rights reserved. See the file COPYRIGHT for details.
- */
-package SketchMaster.lib;
-
-
-import java.awt.geom.Rectangle2D;
-import java.awt.geom.AffineTransform;
-
-import SketchMaster.Stroke.StrokeData.Stroke;
-
-/**
- * General utilities.
- *
- * @author Heloise Hse (hwawen@eecs.berkeley.edu)
- */
-public class GeometryUtil {
-    /**
-     * Store the stroke points in array forms.  This will become
-     * unnecessary once I do some rewriting of TimedStroke class.
-     * xvals, yvals and npoints are pre-allocated arrays and will be
-     * filled in by this method.  xvals[strokeIndex][pointIndex]
-     */
-    public static void strokesToArrays(Stroke[] strokes, double[][] xvals,double[][] yvals, int[] npoints){
-        for(int i=0; i<strokes.length; i++){
-          Stroke s=strokes[i];
-          int noPoints=s.getPointsCount();
-            double x[]=new double[noPoints];
-            double y[]=new double[noPoints];
-            for(int j=0; j<noPoints; j++){
-                x[j]=s.getPoint(j).getX();
-                y[j]=s.getPoint(j).getY();
-            }
-            xvals[i]=x;
-            yvals[i]=y;
-            npoints[i]=noPoints;
-        }
-    }
-
-    
-    /**
-     * Return the bounding box enclosing the set of strokes.
-     */
-    public static Rectangle2D getBounds(Stroke[] s) {
-        Rectangle2D bbox = s[0].getStatisticalInfo().getBox();
-        for(int i=1; i<s.length; i++){
-            bbox.add(s[i].getStatisticalInfo().getBox());
-        }
-        return bbox;
-    }
-
-    
-    /**
-     * Scale the set of strokes uniformly (maintaining the aspect
-     * ratio).  The scaling factor is dependent on the longer side of
-     * the bounding box that encloses the strokes.
-     *
-     * If the height of the bounding box is bigger than the width, the
-     * scale factor would be height/box.height.  Otherwise, it is
-     * width/box.width.
-     *
-     * The strokes are also translated such that the upper left point
-     * of the bounding box is at the origin.
-     *
-     * Return the transform performed on the set of input strokes.
-     */
[…]

### 2010-09-24 MySimpleSketch/src/SketchMaster/lib/MathLib.java

@@ -1,150 +1,0 @@
-package SketchMaster.lib;
-
-import java.awt.geom.Point2D;
-
-import org.apache.log4j.Logger;
-
-import JSci.maths.matrices.DoubleSquareMatrix;
-import SketchMaster.Stroke.StrokeData.PointData;
-import SketchMaster.system.SystemSettings;
-import SketchMaster.system.Recogniziers.SimpleSymbolRecognizier;
-
-public class MathLib {
-	
-	private static final Logger logger = Logger.getLogger(MathLib.class);
-	
-	
-	public static Point2D solveQuadratic(double a, double b, double c){
-		Point2D result=new Point2D.Double();
-		  // Solve the discriminant (SQRT (b^2 - 4ac)
-	    double  discr = Math.sqrt((b * b) - (4.0 * a * c));
-	    double root1=0,root2=0;
-	    // Test for NaN
-	     if(Double.isNaN(discr)){
-	        logger.trace("Equation has no roots");
-	        return null;
-	     }
-	     
-	     if(discr > 0)
-	     {
-	    	 logger.trace("Equation has 2 roots");
-	        root1 = (-b + discr)/(2.0 * a);
-	        root2 = (-b - discr)/(2.0 * a);
-	        logger.trace("First root = " + root1);
-	        logger.trace("Second roor = " + root2);
-
-		       result.setLocation(root1, root2);
-		 
-	      }
-
-	      if(discr == 0)
-	      {
-	    	  logger.trace("Equation has 1 root");
-	        root1 = (-b + discr)/(2.0 * a);
-	        root2 = (-b + discr)/(2.0 * a);
-	        System.out.println("Root = " + root1);
-
-		       result.setLocation(root1, root2);
-		       
-	      }
-
-	       if(discr < 0){
-	         //System.out.println("Equation has imaginary roots");
-	         return null;
-	       }
-		
-	       
-	       
-		return result;
-	} 
-	
-	public static double MeanSquareError(double estimated, double Ideal){
-		double temp=estimated-Ideal;
-		return Math.sqrt( (temp*temp));
-		
-	}
- public static boolean containNan(DoubleSquareMatrix mat){
-	 for (int i = 0; i < mat.rows(); i++) {
-		  for (int j = 0; j < mat.columns(); j++) {
[…]

### 2010-09-24 MySimpleSketch/src/SketchMaster/lib/Vector.java

@@ -1,80 +1,0 @@
-package SketchMaster.lib;
-
-import java.io.Serializable;
-
-import JSci.maths.vectors.DoubleVector;
-
-/**
- * this is warper for the vector for jsci library
- * 
- * @author Mahi
- * 
- */
-public class Vector implements Serializable {
-
-	DoubleVector data = null;
-
-	/**
-	 * 
-	 */
-	public Vector() {
-
-		super();
-
-	}
-
-	// //////////////////////////////Constructors
-
-	public Vector(double x, double y, double z) {
-
-		data = new DoubleVector(3);
-		data.setComponent(0, x);
-		data.setComponent(1, y);
-		data.setComponent(2, z);
-
-	}
-
-	public Vector(double x, double y) {
-
-		data = new DoubleVector(2);
-		data.setComponent(0, x);
-		data.setComponent(1, y);
-
-	}
-
-	public Vector(double[] d) {
-		data = new DoubleVector(d);
-
-	}
-
-	// /////////////////////////////////////functions
-
-	// ////////////////////////////////setters and getters
-	public double getX() {
-		if (data != null)
-			if (data.dimension() > 0)
-				return data.getComponent(0);
-		return Double.NaN;
-	}
-
-	public double getY() {
-		if (data != null)
-			if (data.dimension() > 1)
-				return data.getComponent(1);
-		return Double.NaN;
-	}
-
-	public double getZ() {
-		if (data != null)
-			if (data.dimension() > 2)
-				return data.getComponent(2);
-		return Double.NaN;
-	}
-
-	double getElement(int i) {
-		if (data != null)
-			if (data.dimension() > i)
-				return data.getComponent(i);
-		return Double.NaN;
-	}
-}

### 2010-09-24 MySimpleSketch/src/SketchMaster/lib/ZernikeFE.java

@@ -1,147 +1,0 @@
-/*
- * $Id: ZernikeFE.java,v 1.3 2003/08/18 22:17:23 hwawen Exp $
- *
- * Copyright (c) 2003 The Regents of the University of California.
- * All rights reserved. See the file COPYRIGHT for details.
- */
-package SketchMaster.lib;
-
-import org.apache.log4j.Logger;
-
-import SketchMaster.Stroke.StrokeData.Stroke;
-import SketchMaster.Stroke.features.SegmentClusterFeatureSet;
-import SketchMaster.system.SystemSettings;
-
-
-
-/**
- * Compute Zernike moments of the specified order [1].  This is a
- * global feature of a shape and is independent of stroke-number,
- * -order, and -direction.  This feature (specifically the magnitude
- * of the moment) is invariant to rotation, but not invariant to scale
- * and translation.  Therefore, it is important to normalize the shape
- * before extracting the moment features.  All moment features up to
- * the specified order are returned except for order 0 and 1. This is
- * because for shapes that have been scale- and translation-
- * normalized, these values are the same across all shapes and
- * therefore they are not useful for distinguishing shapes.  If
- * desired, one can call zer_mom directly to obtain these moment
- * values.<p>
- *
- * [1] 'A. Khotanzad and Y.H. Hong' "Invariant image recognition by
- * Zernike Moments", IEEE trans. on Pattern Analysis and Machine
- * Intelligence, vol.12, no.5, pp.489-487, May 1990.
- *
- * @author Heloise Hse (hwawen@eecs.berkeley.edu)
- */
-public class ZernikeFE  {
-    /**
-     * The default order of Zernike moments is set to 8.  This will
-     * compute moments from order 2 up to order 8.
-     */
-    public static int DEFAULT_ORDER = SystemSettings.DEFAULT_ZERNIKE_ORDER; //8;
-
-    /**
-     * The order of moment.
-     */
-    private int _order = DEFAULT_ORDER;
-
-    /**
-     * Create a Zernike feature extractor with the default order.
-     */
-    public ZernikeFE(){}
-
-    /**
-     * Create a Zernike feature extractor with the specified order.
-     */
-    public ZernikeFE(int order){
-        _order = order;
-    }
-    
-    /**
-     * Return the magnitudes of the Zernike moments of the default order.
-     */
-    public double[] apply(Stroke[] s) {
-        return zernikeMoments(s, _order);
-    }
-
-    /**
[…]

### 2010-09-24 MySimpleSketch/src/SketchMaster/swarm/Agent.java

@@ -1,124 +1,0 @@
-package SketchMaster.swarm;
-
-import java.util.Random;
-
-/**
- * @author maha
- * 
- */
-public abstract class Agent {
-	
-	
-	protected static Random r1Random=new Random(11);
-	protected 	static Random r2Random=new Random(22);
-	protected	static Random r3Random=new Random(33);
-	
-	/*
-	 * {null=Agent that is part of the swarm system. each agent is responisible
-	 * for search on , src_lang=Java}
-	 */
-	/**
-	 * @uml.property name="globalBest"
-	 * @uml.associationEnd multiplicity="(1 1)"
-	 *                     inverse="polygonAgent:SketchMaster.swarm.Solution"
-	 * @uml.association name="globalBestLocation"
-	 */
-	protected static Solution globalBest;
-
-	/**
-	 * Getter of the property <tt>globalBest</tt>
-	 * 
-	 * @return Returns the globalBest.
-	 * @uml.property name="globalBest"
-	 */
-	public Solution getGlobalBest() {
-		return globalBest;
-	}
-
-	/**
-	 * Setter of the property <tt>globalBest</tt>
-	 * 
-	 * @param globalBest
-	 *            The globalBest to set.
-	 * @uml.property name="globalBest"
-	 */
-	public void setGlobalBest(Solution globalBest) {
-		this.globalBest = globalBest;
-	}
-
-	/**
-	 * perform single movement step of agent according to the agenst parameter
-	 * and movement equations
-	 * 
-	 */
-	abstract public void move();
-
-	/**
-	 * 
-	 * @return evaluation of thhe current fitness value
-	 */
-	abstract public double getCurrentFitness();
-
-	/**
-	 * the local best location this agent have pased found
-	 * 
-	 * @uml.property name="localbest"
-	 * @uml.associationEnd aggregation="composite"
-	 *                     inverse="agent:SketchMaster.swarm.Solution"
-	 * @uml.association name="localbest"
-	 */
-	protected Solution localbest;
-
-	/**
-	 * Getter of the property <tt>localbest</tt>
-	 * 
-	 * @return Returns the localbest.
-	 * @uml.property name="localbest"
-	 */
-	public Solution getLocalbest() {
-		return localbest;
-	}
-
-	/**
-	 * Setter of the property <tt>localbest</tt>
-	 * 
-	 * @param localbest
-	 *            The localbest to set.
-	 * @uml.property name="localbest"
-	 */
[…]

### 2010-09-24 MySimpleSketch/src/SketchMaster/swarm/Solution.java

@@ -1,110 +1,0 @@
-package SketchMaster.swarm;
-
-import java.util.Random;
-
-import org.apache.log4j.Logger;
-
-/**
- * @author maha Represent a a single solution the agent may represent at any
- *         given time.
- */
-public abstract class Solution implements Comparable, Cloneable {
-	/**
-	 * Logger for this class
-	 */
-	private static final Logger logger = Logger.getLogger(Solution.class);
-
-	
-	static protected Random  r1Random=new Random(55);
-	/*
-	 * {null=represent a valid solution for a problem the swarms solve. ,
-	 * src_lang=Java}
-	 */
-
-	public Solution() {
-
-	}
-
-	public Solution(int size) {
-
-		velocities = new double[size];
-		// now itialize the values with random parameter
-		//Random  r1Random;
-		//r1Random=new Random(size);
-		for (int i = 0; i < velocities.length; i++) {
-			velocities[i] = r1Random.nextDouble();
-		}
-
-	}
-
-	protected double Error = 0.0;
-
-	public double getError() {
-		return Error;
-	}
-
-	public  abstract double getComparableError();
-//		return Double.MAX_VALUE;
-//	}
-	/**
-	 * evaluate the current solution [the higher the value the better the
-	 * solution]
-	 * 
-	 * @return the evaluation of the solution .
-	 */
-	abstract public double eval();
-
-	/**
-	 * calculate approximate error of the solution [the solution must minimize
-	 * the error]
-	 * 
-	 * @return the value of the error for the solution.
-	 */
-	abstract public double error();
-
-	/**
-	 * The velocity of the location of particle at a specific location
-	 * 
-	 * @uml.property name="Velocity" multiplicity="(0 -1)" dimension="1"
-	 */
-	protected double[] velocities;
-
-	/**
-	 * Getter of the property <tt>Velocity</tt>
-	 * 
-	 * @return Returns the velocities.
-	 * @uml.property name="Velocity"
-	 */
-	public double[] getVelocity() {
-		return velocities;
-	}
-
-	/**
-	 * Setter of the property <tt>Velocity</tt>
-	 * 
-	 * @param Velocity
-	 *            The velocities to set.
-	 * @uml.property name="Velocity"
-	 */
-	public void setVelocity(double[] velocity) {
-		velocities = velocity;
-	}
-
-	/*
-	 * (non-Javadoc)
-	 * 
-	 * @see java.lang.Object#clone()
[…]

### 2010-09-24 MySimpleSketch/src/SketchMaster/swarm/curvefit/StrokeCurveAgent.java

@@ -1,231 +1,0 @@
-/**
- * 
- */
-package SketchMaster.swarm.curvefit;
-
-import java.util.Random;
-
-import SketchMaster.swarm.Solution;
-import SketchMaster.swarm.polygonApproximations.PolygonAgent;
-import SketchMaster.swarm.polygonApproximations.PolygonSolution;
-
-/**
- * @author maha
- */
-public class StrokeCurveAgent extends PolygonAgent {
-
-	/**
-	 * 
-	 */
-	public StrokeCurveAgent() {
-
-		C1 = 2.0;
-		C2 = 2.0;
-		w = 1.0;
-		Vmax = 10.0;
-
-	}
-
-	public double getCurrentFitness() {
-		return currentSolution.eval();
-	}
-
-	
-	@Override
-	public void move( Solution agent1, Solution agent2) {
-		
-		// save this for compare with the new fitnessss. 
-		double []  Oldpi=((StrokeCurveSolution) currentSolution).getParticlesPoints();	
-		double oldFitness=((StrokeCurveSolution) currentSolution).eval();
-		double[] viOld = ((StrokeCurveSolution) currentSolution).getVelocity();
-		 // try to move to new agents......\\
-		// get the array of variables of pi , local best and global best
-		double[] pi = ((StrokeCurveSolution) currentSolution).getParticlesPoints();
-		double [] pj = ((StrokeCurveSolution) agent1).getParticlesPoints();
-		// logger.info("----------------------------------------------------------------");
-		// get the array of variables of pi , local best and global best
-		double[] pl = ((StrokeCurveSolution) agent2)
-				.getParticlesPoints();
-
- 
-		double[] gbest = ((StrokeCurveSolution) globalBest)
-				.getParticlesPoints();
-		//		
-		// double [] pi=null;
-		// double[] pbest=null;
-		// double [] gbest=null;
-
-		// generate the random numbers
-		double r1, r2;// ,r3;
-
- 
-		// r3=Math.random();
-		// now get the velocity array of the location
-
-		double[] vi = ((StrokeCurveSolution) currentSolution).getVelocity();
-
-		double gbestterm, pbestterm;
-		boolean repeatTrue = false;
-		// do {
-		// repeatTrue=false;
-		pi = ((StrokeCurveSolution) currentSolution).getParticlesPoints();
[…]

### 2010-09-24 MySimpleSketch/src/SketchMaster/swarm/curvefit/StrokeCurveSolution.java

@@ -1,1059 +1,0 @@
-/**
-  
- 
- */
-package SketchMaster.swarm.curvefit;
-
-import org.apache.log4j.Logger;
-
-import java.awt.Color;
-import java.awt.Graphics2D;
-import java.awt.geom.Point2D;
-import java.util.ArrayList;
-import java.util.Arrays;
-import java.util.Random;
-
-import SketchMaster.system.SystemSettings;
-import SketchMaster.Stroke.StrokeData.PointData;
-import SketchMaster.Stroke.StrokeData.Segment;
-import SketchMaster.Stroke.StrokeData.Stroke;
-import SketchMaster.Stroke.graphics.shapes.Ellipse;
-import SketchMaster.Stroke.graphics.shapes.GuiShape;
-import SketchMaster.Stroke.graphics.shapes.Line;
-import SketchMaster.Stroke.graphics.shapes.SegmentedShape;
-import SketchMaster.io.log.FileLog;
-import SketchMaster.lib.ComputationsGeometry;
-import SketchMaster.swarm.Solution;
-import SketchMaster.swarm.polygonApproximations.PolygonSolution;
-
-/**
- * @author maha
- */
-public class StrokeCurveSolution extends PolygonSolution implements
-		SegmentedShape {
-	/**
-	 * Logger for this class
-	 */
-	private static final Logger logger = Logger.getLogger(StrokeCurveSolution.class);
-
-	private static final boolean ERROR_CONSTANT = true;
-	private static int PARAM_SIZE = 19;// size of paramter of the particle
-	private double circlea, circleb, circleR;
-
-	// final static double T=0.9;
-	// final static double eTolerance=200;
-	private transient double A, B, C, D, a, b, c, d, e, f, m, line_b;
-	private transient double x0, y0, ellipse_a, ellipse_b;
-
-	private double dmax, Amax, Bmax, Cmax, Dmax;
-
-	private double aC, bC, r;
-
-	/**
-	 * @param problemStroke
-	 */
-	public StrokeCurveSolution(Stroke stroke) {
-
-		super(PARAM_SIZE);
-		ParticlesPoints = new double[PARAM_SIZE];// 
-		this.problemStroke = stroke;
-		Random  r1Random;
-		r1Random=new Random(PARAM_SIZE);
-		// points== is (A B C D) , points ( a, b , c, d, e, f )==.ellipse
[…]

### 2010-09-24 MySimpleSketch/src/SketchMaster/swarm/polygonApproximations/DigitalCurveDivideAgent.java

@@ -1,90 +1,0 @@
-/**
- * 
- */
-package SketchMaster.swarm.polygonApproximations;
-
-import java.util.Random;
-
-/**
- * @author maha
- * 
- */
-public class DigitalCurveDivideAgent extends PolygonAgent {
-
-	/**
-	 * 
-	 */
-	public DigitalCurveDivideAgent() {
-
-	}
-
-	public double getCurrentFitness() {
-		return currentSolution.eval();
-	}
-
-	@Override
-	public void move() {
-		
-		
-		// get the array of variables of pi , local best and global best
-		int[] pi = ((PolygonSolution) currentSolution).getParticlePoints();
-		int[] pbest = ((PolygonSolution) localbest).getParticlePoints();
-		int[] gbest = ((PolygonSolution) globalBest).getParticlePoints();
-		// generate the random numbers
-		double r1, r2, r3;
-//		Random  r1Random,r2Random,r3Random;
-//		r1Random=new Random(11);
-//		r2Random=new Random(22);
-//		r3Random=new Random(33);
-		r1 =r1Random.nextDouble();//Math.random();
-		r2 =r2Random.nextDouble();
-		r3 =r3Random.nextDouble();
-		// now get the velocity array of the location
-
-		double[] vi = ((PolygonSolution) currentSolution).getVelocity();
-
-		double gbestterm, pbestterm;
-		double templocation;
-		for (int j = 0; j < pi.length; j++) {
-			r1 =r1Random.nextDouble();//Math.random();
-			r2 =r2Random.nextDouble();
-			r3 =r3Random.nextDouble();
-			// computer terms speratly
-			pbestterm = C1 * r1 * (pbest[j] - pi[j]);
-			gbestterm = C2 * r2 * (gbest[j] - pi[j]);
-
-			// the velocity equaiton.
-			vi[j] = w * vi[j] + pbestterm + gbestterm;
-			// logger.info("velcoity s "+vi[j]);
-		//	if (vi[j] >= (Vmax))
-				// use s(v) to restrice the velocity to vmax
-			//	vi[j] = 0.5 + (vi[j] / (2 * Vmax));
-			if (vi[j]>0)
-				vi[j] = 0.5 + (vi[j] / (2.0 * Vmax));
-			else {
-				vi[j] = 0.5 + (vi[j] / (-2.0 * Vmax));
-			}
-
-			// the function pj = pj(old) +vj(new)
-			templocation = vi[j];
[…]

### 2010-09-24 MySimpleSketch/src/SketchMaster/swarm/polygonApproximations/DigitalCurveDivideSolution.java

@@ -1,1292 +1,0 @@
-/**
- * 
- */
-package SketchMaster.swarm.polygonApproximations;
-
-import org.apache.log4j.Logger;
-
-import java.awt.Color;
-import java.awt.Graphics2D;
-import java.awt.Point;
-import java.awt.Rectangle;
-import java.awt.geom.Point2D;
-import java.awt.geom.Rectangle2D;
-import java.util.ArrayList;
-
-import SketchMaster.lib.ComputationsGeometry;
-import SketchMaster.system.SystemSettings;
-import SketchMaster.Stroke.StrokeData.PointData;
-import SketchMaster.Stroke.StrokeData.Segment;
-import SketchMaster.Stroke.StrokeData.Stroke;
-import SketchMaster.Stroke.graphics.shapes.Circle;
-import SketchMaster.Stroke.graphics.shapes.GeometricPrimitive;
-import SketchMaster.Stroke.graphics.shapes.GuiShape;
-import SketchMaster.Stroke.graphics.shapes.Line;
-import SketchMaster.Stroke.graphics.shapes.SegmentedShape;
-
-/**
- * @author maha
- * 
- */
-public class DigitalCurveDivideSolution extends PolygonSolution implements
-		SegmentedShape {
-	/**
-	 * Logger for this class
-	 */
-	private static final Logger logger = Logger.getLogger(DigitalCurveDivideSolution.class);
-
-	private static final double MaxSegmentERROR = 10;
-
-	/**
-	 * A list of the final vertices of the solution [only the vertics not all
-	 * the point of stroke.]
-	 * 
-	 * @uml.property name="polygonVertices"
-	 */
-	protected ArrayList<GeometricPrimitive> SegmentsType;
-	private static double K =5;// 0.05;// was 0.01
-
-	static protected boolean storeShape = false;
-
-	// /**
-	// *
-	// */
-	// public DigitalCurveDivideSolution() {
-	//		
-	// }
-	//
-	// /**
-	// * @param size
-	// */
-	// public DigitalCurveDivideSolution(int size) {
-	// super(size);
-	//		
-	// }
-
-	/**
-	 * @param problemStroke
-	 */
-	public DigitalCurveDivideSolution(Stroke problemStroke) {
-		super(problemStroke);
-		refineSolution();
-
-	}
-
-	// final static double T=0.7;
-	// public static double eTolerance= 200.0;
-
-	@Override
[…]

### 2010-09-24 MySimpleSketch/src/SketchMaster/swarm/polygonApproximations/PolygonAgent.java

@@ -1,210 +1,0 @@
-package SketchMaster.swarm.polygonApproximations;
-
-import java.util.Random;
-
-import SketchMaster.swarm.Agent;
-import SketchMaster.swarm.Solution;
-import SketchMaster.system.SystemSettings;
-
-public class PolygonAgent extends Agent {
-
-	static protected double Vmax =  SystemSettings.SWARM_CONSTANTS_VMAX;
-	/* {src_lang=Java} */
-
-	static protected double C1 =  SystemSettings.SWARM_CONSTANTS_C1, C2 = SystemSettings.SWARM_CONSTANTS_C2;
-	static protected double w = SystemSettings.SWARM_CONSTANTS_W;
-
-	public double getCurrentFitness() {
-		return currentSolution.eval();
-	}
-
-	@Override
-	public void move() {
-		// get the array of variables of pi , local best and global best
-		int[] pi = ((PolygonSolution) currentSolution).getParticlePoints();
-		int[] pbest = ((PolygonSolution) localbest).getParticlePoints();
-		int[] gbest = ((PolygonSolution) globalBest).getParticlePoints();
-		// generate the random numbers
-		double r1, r2, r3;
-//		r1 = Math.random();
-//		r2 = Math.random();
-//		r3 = Math.random();
-		
-//		Random  r1Random,r2Random,r3Random;
-//		r1Random=new Random(11);
-//		r2Random=new Random(22);
-//		r3Random=new Random(33);
-		r1 =r1Random.nextDouble();//
-		//Math.random();
-		r2 =r2Random.nextDouble();
-		r3 =r3Random.nextDouble();
-		// now get the velocity array of the location
-
-		double[] vi = ((PolygonSolution) currentSolution).getVelocity();
-
-		double gbestterm, pbestterm;
-		double templocation;
-		for (int j = 0; j < pi.length; j++) {
-//			r1 = Math.random();
-//			r2 = Math.random();
-//			r3 = Math.random();
-			r1 =r1Random.nextDouble();//Math.random();
-			r2 =r2Random.nextDouble();
-			r3 =r3Random.nextDouble();
-			// computer terms speratly
-			pbestterm = C1 * r1 * (pbest[j] - pi[j]);
-			gbestterm = C2 * r2 * (gbest[j] - pi[j]);
-
-			// the velocity equaiton.
-			vi[j] = w * vi[j] + pbestterm + gbestterm;
-			// logger.info("velcoity s "+vi[j]);
[…]

### 2010-09-24 MySimpleSketch/src/SketchMaster/swarm/polygonApproximations/PolygonSolution.java

@@ -1,1579 +1,0 @@
-package SketchMaster.swarm.polygonApproximations;
-
-import org.apache.log4j.Logger;
-
-import SketchMaster.system.SystemSettings;
-import SketchMaster.Stroke.StrokeData.DominatePointStructure;
-import SketchMaster.Stroke.StrokeData.PointData;
-import SketchMaster.Stroke.StrokeData.Segment;
-import SketchMaster.Stroke.StrokeData.Stroke;
-import SketchMaster.Stroke.graphics.shapes.GeometricPrimitive;
-//import SketchMaster.Stroke.graphics.shapes.GuiShape;
-import SketchMaster.Stroke.graphics.shapes.Line;
-import SketchMaster.Stroke.graphics.shapes.SegmentedShape;
-import SketchMaster.collection.SortedValueMap;
-import SketchMaster.gui.DrawingDebugUtils;
-//import SketchMaster.io.log.FileLog;
-import SketchMaster.lib.ComputationsGeometry;
-import SketchMaster.swarm.Solution;
-import SketchMaster.swarm.SwarmSystem;
-//import SketchMaster.swarm.curvefit.StrokeCurveSolution;
-//import SketchMaster.system.SystemSettings;
-
-import java.awt.BasicStroke;
-import java.awt.Color;
-import java.awt.Graphics2D;
-import java.awt.Point;
-import java.awt.Rectangle;
-import java.awt.geom.Point2D;
-import java.awt.geom.Rectangle2D;
-//import java.io.BufferedReader;
-//import java.io.IOException;
-//import java.io.InputStreamReader;
-import java.util.ArrayList;
-import java.util.Arrays;
-import java.util.Iterator;
-import java.util.Map;
-//import java.util.Random;
-
-/**
- * @author maha It representa solution location in the search space for the
- *         agents in the problem of the polygon approximation
- */
-public class PolygonSolution extends Solution implements SegmentedShape {
-	/**
-	 * Logger for this class
-	 */
-	private static final Logger logger = Logger.getLogger(PolygonSolution.class);
-
-	public final static int POLYGON_ADJUST_MERGE = 1;
-	public final static int POLYGON_ADJUST_DOMINATE_POINT = 2;
-	public final static int POLYGON_ADJUST_BOTH = 3;
-	public static final int POLYGON_ADJUST_NONE = 4;
-	public static final int POLYGON_ADJUST_ONLY_SEGMENTS = 5;
[…]

### 2010-09-24 MySimpleSketch/src/SketchMaster/system/Recogniziers/RecognizierSystem.java

@@ -1,117 +1,0 @@
-package SketchMaster.system.Recogniziers;
-
-import java.util.ArrayList;
-import java.util.Observable;
-import java.util.Observer;
-
-import SketchMaster.Stroke.StrokeData.InkInterface;
-import SketchMaster.classifier.Classification;
-import SketchMaster.classifier.rubine.RubineTrainingSet;
-import SketchMaster.gui.DrawingSheet;
-import SketchMaster.gui.GraphWatched;
-import SketchMaster.gui.Events.HandleStroke;
-import SketchMaster.io.FileObjectReader;
-import SketchMaster.io.FileObjectWriter;
-
-abstract public class RecognizierSystem extends Observable implements Observer,
-		HandleStroke {
-	protected Classification CurrentClassification = null;
-
-	protected String FeatureString;
- 
-	protected int CurrentCat = 0;
-
-	protected String CurrentSelectedCategoryName = "";
-	// / the classifier defaults
-	public final static int RECOGNIZE_LOAD_FILE = 1;
-	public final static int RECOGNIZE_LOAD_SERIALIZE = 2;
-	public final static int RECGONIZE_LOAD_SYS = 3;
-	public final static int RECGONIZE_OPERATION_TRAIN = 4;
-	public final static int RECGONIZE_OPERATION_TEST = 5;
-	public final static int RECGONIZE_OPERATION_CLASSIFY = 6;
-	public final static int  EXAMPLE_NEG=0;
-	public final static int  EXAMPLE_POS=0;
-	public static int ExmapleType=EXAMPLE_POS;
-		
-
-	/**
-	 * @param currentCat
-	 *            the currentCat to set
-	 */
-	public void setCurrentCat(int currentCat) {
-		CurrentCat = currentCat;
-	}
-
-	public abstract void train();
-
-	public abstract void addNewCategory(String Name);
-
-	public abstract void addCategoryExample(int i, Object ink);
-
-	// public abstract void addCategoryExample(String Name,InkObject ink);
-
-	public abstract int deleteCategoryExample(String Name, int i);
-
-	public abstract void deleteCategory(String Name);
- 
-	public abstract int getCategoryCount();
-	public abstract String[] getCategoryNames();
-
-	// public abstract ArrayList getCategoryExamples(String Name);
-	public abstract ArrayList getCategoryExamples(int Name);
-
-	public abstract ArrayList getCategoryExamplesInk(String Name);
-
[…]

### 2010-09-24 MySimpleSketch/src/SketchMaster/system/Recogniziers/RubineRecognizier.java

@@ -1,285 +1,0 @@
-/**
- * 
- */
-package SketchMaster.system.Recogniziers;
-
-import java.util.ArrayList;
-import java.util.Observable;
-import java.util.Observer;
-
-import org.apache.log4j.Logger;
-
-import SketchMaster.Stroke.StrokeData.InkInterface;
-import SketchMaster.Stroke.features.StrokeRubineFeatureSet;
-import SketchMaster.classifier.Classification;
-import SketchMaster.classifier.rubine.RubineCategory;
-import SketchMaster.classifier.rubine.RubineClassifier;
-import SketchMaster.classifier.rubine.RubineTrainingSet;
-import SketchMaster.gui.GraphWatched;
-import SketchMaster.gui.Events.HandleStroke;
-import SketchMaster.gui.Events.NewStrokeEvent;
-import SketchMaster.io.FileObjectReader;
-import SketchMaster.io.FileObjectWriter;
-import SketchMaster.io.log.FileLog;
-import SketchMaster.swarm.Solution;
-import SketchMaster.system.SystemSettings;
-
-/**
- * @author maha
- * 
- */
-public class RubineRecognizier extends RecognizierSystem {
-	private static final Logger logger = Logger.getLogger(RubineRecognizier.class);
-	// / the classifier defaults
-	public final static int RUBINE_LOAD_FILE = 1;
-	public final static int RUBINE_LOAD_SERIALIZE = 2;
-	public final static int RUBINE_LOAD_SYS = 3;
-
-	private RubineTrainingSet trainer = null;
-	private StrokeRubineFeatureSet test;
-	private RubineClassifier Rubclassifier = null;
-
-	// operation
-	public final static int Rubine_OPERATION_TRAINING = 0;
-	public final static int Rubine_OPERATION_CLASSIFICATION = 1;
-
-	private RubineCategory currentCategory;
-
-	private InkInterface LastStrokeDrawn = null;
-
-	// private currentExampleviewed=-1;
-	public void init() {
-		Rubclassifier = new RubineClassifier();
-		Logger appLogger = Logger.getLogger("AppLogging");
-		appLogger.info(" [Settings] indside rubine settings is "+SystemSettings.getString());
-		// logger.info("look ");
-	//	if (SystemSettings.RubineDefaultLoadOption == RUBINE_LOAD_SYS) {
-			trainer = new RubineTrainingSet();
-			//  logger.trace("initailizing  the  rubine trainer ");
-			trainer.init();
[…]

### 2010-09-24 MySimpleSketch/src/SketchMaster/system/Recogniziers/SVMRecognizier.java

@@ -1,777 +1,0 @@
-/**
- * 
- */
-package SketchMaster.system.Recogniziers;
-
-import java.io.FileOutputStream;
-import java.io.ObjectOutputStream;
-import java.io.PrintStream;
-import java.util.ArrayList;
-import java.util.Collection;
-import java.util.Hashtable;
-import java.util.Iterator;
-import java.util.Observable;
-import java.util.Set;
-import java.util.Map.Entry;
-
-import org.apache.log4j.Level;
-import org.apache.log4j.Logger;
-
-import SketchMaster.Stroke.StrokeData.InkInterface;
-import SketchMaster.Stroke.StrokeData.SegmentCluster;
-import SketchMaster.Stroke.StrokeData.Stroke;
-import SketchMaster.Stroke.features.SVMFeatureSet;
-import SketchMaster.Stroke.features.SegmentClusterFeatureSet;
-import SketchMaster.Stroke.graphics.shapes.GuiShape;
-import SketchMaster.classifier.Classification;
-import SketchMaster.classifier.Symbol.SymbolClassifier;
-import SketchMaster.classifier.Symbol.SymbolTrainingSet;
-import SketchMaster.classifier.rubine.RubineTrainingSet;
-import SketchMaster.classifier.svmTrainable.*;
-import SketchMaster.gui.GraphWatched;
-import SketchMaster.gui.Events.NewStrokeEvent;
-import SketchMaster.io.FileObjectReader;
-import SketchMaster.io.FileObjectWriter;
-import SketchMaster.system.SketchSheet;
-import SketchMaster.system.SystemSettings;
-import SketchMaster.system.clustering.SystemClustering;
-
-/**
- * @author Maha
- *
- */
-public class SVMRecognizier extends RecognizierSystem {
-	private static final Logger logger = Logger.getLogger(SVMRecognizier.class);
-	
-private static String CurrentCategory = null;
-//SystemClustering clusturing;
-transient SketchSheet sheet;
-private SegmentCluster LastClusterDrawn;
-//	
-//	protected String FeatureString;
-//	 
-//	protected int CurrentCat = 0;
-//
-//	protected String CurrentSelectedCategoryName = "";
-//	protected boolean CaptureStrokes = false;
-	
-	// i will need to add all trian dat a
-	
-
-int CurrentStrokesCount=0;
-	
-	TrainingSet trainSet;
[…]

### 2010-09-24 MySimpleSketch/src/SketchMaster/system/Recogniziers/SimpleSymbolRecognizier.java

@@ -1,361 +1,0 @@
-package SketchMaster.system.Recogniziers;
-
-import org.apache.log4j.Logger;
-
-import java.util.ArrayList;
-import java.util.Observable;
-
-import SketchMaster.Stroke.StrokeData.InkInterface;
-import SketchMaster.Stroke.StrokeData.SegmentCluster;
-import SketchMaster.Stroke.StrokeData.Stroke;
-import SketchMaster.Stroke.features.SegmentClusterFeatureSet;
-import SketchMaster.Stroke.features.StrokeRubineFeatureSet;
-import SketchMaster.Stroke.graphics.shapes.GuiShape;
-import SketchMaster.classifier.Classification;
-import SketchMaster.classifier.Symbol.SymbolCategory;
-import SketchMaster.classifier.Symbol.SymbolClassifier;
-import SketchMaster.classifier.Symbol.SymbolTrainingSet;
-import SketchMaster.classifier.rubine.RubineCategory;
-import SketchMaster.classifier.rubine.RubineClassifier;
-import SketchMaster.classifier.rubine.RubineTrainingSet;
-import SketchMaster.gui.DrawingSheet;
-import SketchMaster.gui.GraphWatched;
-import SketchMaster.gui.Events.NewStrokeEvent;
-import SketchMaster.io.FileObjectReader;
-import SketchMaster.io.FileObjectWriter;
-import SketchMaster.io.log.FileLog;
-import SketchMaster.system.SketchSheet;
-import SketchMaster.system.SystemSettings;
-import SketchMaster.system.clustering.SystemClustering;
-import SketchMaster.system.segmentation.SezginSegmentor;
-import SketchMaster.system.segmentation.SketchSegmentors;
-
-public class SimpleSymbolRecognizier extends RecognizierSystem {
-	/**
-	 * Logger for this class
-	 */
-	private static final Logger logger = Logger.getLogger(SimpleSymbolRecognizier.class);
-
-	private SymbolTrainingSet trainer = null;
-	private SymbolCategory currentCategory;
-
-	private SymbolClassifier symbolClassifier;
-	//SystemClustering clusturing;
-	transient SketchSheet sheet;
-	private SegmentCluster LastClusterDrawn;
-	public void ReadTrainingSet(String FileName) {
-
-		// set current to null
-		trainer = null;
-		FileObjectReader fileReader = new FileObjectReader();
-		fileReader.readFile(FileName);
[…]

### 2010-09-24 MySimpleSketch/src/SketchMaster/system/SketchRecognitionSystem.java

@@ -1,327 +1,0 @@
-package SketchMaster.system;
-
-import org.apache.log4j.Logger;
-
-import SketchMaster.Stroke.StrokeData.Stroke;
-import SketchMaster.Stroke.features.FeatureFunction;
-import SketchMaster.Stroke.graphics.layers.SketchLayer;
-import SketchMaster.Stroke.graphics.layers.SketchStrokesLayer;
-import SketchMaster.Stroke.graphics.shapes.GuiShape;
-import SketchMaster.gui.DrawingSheet;
-import SketchMaster.gui.GraphWatched;
-import SketchMaster.gui.Events.HandleFinishCluster;
-import SketchMaster.gui.Events.HandleStroke;
-import SketchMaster.gui.Events.NewClusterEvent;
-import SketchMaster.gui.Events.NewStrokeEvent; //import SketchMaster.gui.GraphFrame;
-//import SketchMaster.gui.SketchFrame;
-import SketchMaster.io.StrokeReaderWriter;
-import SketchMaster.io.log.FileLog;
-import SketchMaster.system.Recogniziers.RecognizierSystem;
-import SketchMaster.system.Recogniziers.RubineRecognizier;
-import SketchMaster.system.Recogniziers.SVMRecognizier;
-import SketchMaster.system.Recogniziers.SimpleSymbolRecognizier;
-
-import java.awt.Color;
-import java.awt.Graphics;
-import java.awt.Graphics2D;
-import java.awt.GraphicsConfiguration;
-import java.awt.Point;
-import java.util.ArrayList;
-import java.util.Observable;
-import java.util.Observer;
-
-/**
- * this is the class that control the system at initalization this is the class
- * that create all frames it is initalized from the main class it handle all
- * passing of decisions it act as the outer program it directly call frame ,
- * drawing drawSheet and sketch drawSheet
- */
-public class SketchRecognitionSystem implements Observer, HandleFinishCluster {
-	/**
-	 * Logger for this class
-	 */
-	private static final Logger logger = Logger.getLogger(SketchRecognitionSystem.class);
-
-	int StrokeNumber = 1;
-	// SketchFrame sketchframe;
-
-	SketchSheet sheetdata;
-
-	GraphDrawing ftest;
-
-	DrawingSheet drawSheet = null;
-
-	RecognizierSystem recognizier;
-
[…]

### 2010-09-24 MySimpleSketch/src/SketchMaster/system/SketchSheet.java

@@ -1,832 +1,0 @@
-package SketchMaster.system;
-
-import org.apache.log4j.Logger;
-
-import SketchMaster.Stroke.StrokeData.DominatePointStructure;
-import SketchMaster.Stroke.StrokeData.PointData;
-import SketchMaster.Stroke.StrokeData.SegmentCluster;
-import SketchMaster.Stroke.StrokeData.Stroke;
-import SketchMaster.Stroke.StrokeData.StrokeStatisticalData;
-import SketchMaster.Stroke.features.FeatureFunction;
-
-import SketchMaster.Stroke.graphics.layers.SketchLayer;
-
-import SketchMaster.Stroke.graphics.layers.SketchSegmentionLayer;
-
-import SketchMaster.Stroke.graphics.layers.SketchStrokesLayer;
-import SketchMaster.Stroke.graphics.shapes.FittedShape;
-import SketchMaster.Stroke.graphics.shapes.GuiShape; //import SketchMaster.Stroke.graphics.shapes.StrokeFitObject;
-import SketchMaster.Stroke.SketchLayerArray;
-import SketchMaster.system.SystemSettings;
-
-import SketchMaster.gui.DrawingDebugUtils;
-import SketchMaster.gui.DrawingSheet;
-import SketchMaster.gui.Events.HandleStroke;
-import SketchMaster.gui.Events.NewStrokeEvent;
-import SketchMaster.io.log.FileLog;
-import SketchMaster.swarm.Solution;
-import SketchMaster.swarm.SwarmSystem;
-import SketchMaster.swarm.curvefit.StrokeCurveAgent;
-import SketchMaster.swarm.curvefit.StrokeCurveSolution;
-import SketchMaster.swarm.polygonApproximations.DigitalCurveDivideSolution;
-import SketchMaster.swarm.polygonApproximations.PolygonAgent;
-import SketchMaster.swarm.polygonApproximations.PolygonSolution;
-import SketchMaster.system.clustering.SystemClustering;
-import SketchMaster.system.segmentation.SezginSegmentor;
-import SketchMaster.system.segmentation.SketchSegmentors;
-
-import java.awt.Point;
-import java.awt.Rectangle;
-import java.awt.geom.Point2D;
-import java.util.ArrayList;
-import java.util.Observable;
-
-/**
- * this is the class that contian all info about what are the strokes it handle
[…]

### 2010-09-24 MySimpleSketch/src/SketchMaster/system/clustering/SystemClustering.java

@@ -1,153 +1,0 @@
-/**
- * 
- */
-package SketchMaster.system.clustering;
-
-import org.apache.log4j.Logger;
-
-import java.util.ArrayList;
-
-import SketchMaster.Stroke.StrokeData.Segment;
-import SketchMaster.Stroke.StrokeData.SegmentCluster;
-import SketchMaster.Stroke.StrokeData.Stroke;
-import SketchMaster.Stroke.graphics.layers.SketchLayer;
-import SketchMaster.Stroke.graphics.shapes.GuiShape;
-import SketchMaster.Stroke.graphics.shapes.SegmentedShape;
-
-/**
- * @author maha
- * 
- */
-public class SystemClustering {
-	/**
-	 * Logger for this class
-	 */
-	private static final Logger logger = Logger.getLogger(SystemClustering.class);
-
-	// i will have a list of strokes not clustered ==> list of segments
-
-	ArrayList<GuiShape> UnClusteredsegments = new ArrayList<GuiShape>();
-	ArrayList<Stroke> UnClusteredStrokes = new ArrayList<Stroke>();
-	ArrayList<SegmentCluster> symbols = new ArrayList<SegmentCluster>();
-	double cluserArea=0.0;
-	
-
-	// i will have a list of clustered strokes
-
-	private SketchLayer CurrentSymbolLayer;
-
-	public void setSymbolLayer(SketchLayer layer) {
-
-		// logger.info("''''''''''''''''''''''''''''''''''''''''''");
-		CurrentSymbolLayer = layer;
-	}
-
-	// this the main class will implement the clustring algoirthms
-	// here is the steps
-	/***************************************************************************
-	 * The simple algorithm, an event from user add the previosu strokes into a
-	 * new cluster.
-	 * 
-	 * 
-	 * 
-	 **************************************************************************/
-
-	public void addStrokeSegmentation(Stroke stroke, GuiShape segmenation) {
-		// add this stroke with its segmetnation to the list of unclustered
-		// strokes
-		if (UnClusteredsegments != null) {
-			UnClusteredsegments.add(segmenation);
-			UnClusteredStrokes.add(stroke);
-			cluserArea+=stroke.getStatisticalInfo().getArea();
-
-		}
-	}
-
-	public SegmentCluster CreateSymbol() {
-		// create a new cluster and starts its recognition from the set of
-		// unclustered strokes.
-		SegmentCluster symbol = new SegmentCluster();
-
-		// for each segmented stroke not part of a symbol do the follwing
[…]

### 2010-09-24 MySimpleSketch/src/SketchMaster/system/segmentation/SezginSegmentor.java

@@ -1,818 +1,0 @@
-package SketchMaster.system.segmentation;
-
-import org.apache.log4j.Logger;
-
-import java.awt.geom.Point2D;
-import java.util.ArrayList;
-import java.util.Collections;
-import java.util.Comparator;
-import java.util.Iterator;
-import java.util.LinkedList;
-import java.util.List;
-import java.util.Map;
-import java.util.Set;
-import SketchMaster.system.SystemSettings;
-
-import SketchMaster.Stroke.StrokeData.*;
-import SketchMaster.Stroke.features.FeatureFunction;
-import SketchMaster.Stroke.features.StrokeFeatures;
-import SketchMaster.collection.NumericalComparator;
-import SketchMaster.collection.SortedValueMap;
-import SketchMaster.gui.DrawingDebugUtils;
-import SketchMaster.system.SystemSettings;
-
-public class SezginSegmentor  implements Cloneable {
-	/**
-	 * Logger for this class
-	 */
-	private static final Logger logger = Logger.getLogger(SezginSegmentor.class);
-
-	private static double ErrorThreshold = SystemSettings.SEGZIN_ERROR_THRESHOLD;
-	private SortedValueMap HybirdFitList;
-	private SortedValueMap segmentsCurvaturePoints;
-	private SortedValueMap segmentsVelocityPoints;
-	
-
-	@Override
-	public Object clone() {
-
-		try {
-			return super.clone();
-		} catch (CloneNotSupportedException e) {
-
-			logger.error("clone()", e); //$NON-NLS-1$
-		}
-		return this;
-	}
-
-	public static SortedValueMap createVertices(Stroke ink, SortedValueMap segmentsPoints,
-			int type) {
-		if (segmentsPoints==null)
-			segmentsPoints=new SortedValueMap(new NumericalComparator());
-		
-		int curvatureCount;
-		double max = Double.MIN_VALUE, min = Double.MAX_VALUE;
-		String functionName = "";
-		if (type == 0)
-			// functionName="Direction";
-			functionName = SystemSettings.CURVATURENAME;
-		// functionName="Curvature of storke Vs. distance";
-		 if (type == 1)
-			functionName = "Velocity of storke Vs. distance";
-
-		 if (type==2)
-			functionName="Direction";
-		 if (type==3)
-			 functionName="Time";
-		// Direction
-		// //Done : add the correct implemenation for finding the critcal points
-		// which include finding local maximum not the points > average.
[…]

### 2010-09-24 MySimpleSketch/src/SketchMaster/system/segmentation/SketchSegmentors.java

@@ -1,799 +1,0 @@
-/**
- * 
- */
-package SketchMaster.system.segmentation;
-
-import org.apache.log4j.Logger;
-import org.omg.CORBA.DATA_CONVERSION;
-
-import java.awt.Point;
-import java.awt.geom.Point2D;
-import java.util.ArrayList;
-import java.util.Iterator;
-import java.util.Map;
-import java.util.Map.Entry;
-
-import SketchMaster.system.SystemSettings;
-import SketchMaster.Stroke.StrokeData.DominatePointStructure;
-import SketchMaster.Stroke.StrokeData.InkInterface;
-import SketchMaster.Stroke.StrokeData.PointData;
-import SketchMaster.Stroke.StrokeData.Stroke;
-import SketchMaster.Stroke.graphics.layers.SketchSegmentionLayer;
-import SketchMaster.Stroke.graphics.shapes.Ellipse;
-import SketchMaster.Stroke.graphics.shapes.FittedShape;
-import SketchMaster.Stroke.graphics.shapes.GuiShape;
-import SketchMaster.Stroke.graphics.shapes.Line;
-import SketchMaster.Stroke.graphics.shapes.ShapeRecognizier;
-import SketchMaster.collection.NumericalComparator;
-import SketchMaster.collection.SortedValueMap;
-import SketchMaster.gui.Events.HandleStroke;
-import SketchMaster.gui.Events.NewStrokeEvent;
-import SketchMaster.io.log.FileLog;
-import SketchMaster.lib.ComputationsGeometry;
-import SketchMaster.lib.CurveFitData;
-import SketchMaster.swarm.Solution;
-import SketchMaster.swarm.SwarmSystem;
-import SketchMaster.swarm.curvefit.StrokeCurveAgent;
-import SketchMaster.swarm.curvefit.StrokeCurveSolution;
-import SketchMaster.swarm.polygonApproximations.DigitalCurveDivideAgent;
-import SketchMaster.swarm.polygonApproximations.DigitalCurveDivideSolution;
-import SketchMaster.swarm.polygonApproximations.PolygonAgent;
-import SketchMaster.swarm.polygonApproximations.PolygonSolution;
-import SketchMaster.system.SystemSettings;
-
-/**
- * @author maha
- * 
- */
-public class SketchSegmentors {
-	/**
-	 * Logger for this class
-	 */
-	private static final Logger logger = Logger.getLogger(SketchSegmentors.class);
-
-	private static final int AGENT_SIZE = SystemSettings.AGENT_SIZE;
-
[…]

### 2010-09-24 TrainTest/src/chartTest/GraphFrame.java

@@ -1,598 +1,0 @@
-package chartTest;
-
-
-//import jahuwaldt.plot.CircleSymbol;
-//import jahuwaldt.plot.Plot2D;
-//import jahuwaldt.plot.PlotPanel;
-//import jahuwaldt.plot.PlotRun;
-//import jahuwaldt.plot.PlotRunList;
-//import jahuwaldt.plot.SimplePlotXY;
-//import jahuwaldt.plot.SquareSymbol;
-
-
-import SketchMaster.Stroke.features.FeatureFunction;
-import SketchMaster.lib.Vector;
-
-
-import java.awt.BasicStroke;
-import java.awt.Color;
-import java.awt.Component;
-import java.awt.Graphics;
-import java.awt.Graphics2D;
-import java.awt.GraphicsConfiguration;
-import java.awt.HeadlessException;
-import java.awt.Paint;
-import java.awt.Point;
-import java.awt.Rectangle;
-import java.awt.Shape;
-import java.awt.color.*;
-import java.awt.event.FocusEvent;
-import java.awt.event.FocusListener;
-import java.awt.geom.Point2D;
-import java.awt.geom.Rectangle2D;
-import java.awt.image.ImageObserver;
-import java.util.ArrayList;
-import javax.swing.JFrame;
-import javax.swing.JPanel;
-import org.jfree.chart.ChartFactory;
-import org.jfree.chart.ChartPanel;
-import org.jfree.chart.JFreeChart;
-import org.jfree.chart.axis.NumberAxis;
-import org.jfree.chart.plot.PlotOrientation;
-import org.jfree.chart.plot.XYPlot;
-import org.jfree.chart.renderer.xy.StandardXYItemRenderer;
-import org.jfree.chart.renderer.xy.XYItemRenderer;
-import org.jfree.chart.renderer.xy.XYLineAndShapeRenderer;
-import org.jfree.data.xy.XYDataset;
-import org.jfree.data.xy.XYSeries;
-import org.jfree.data.xy.XYSeriesCollection;
-import org.jfree.ui.RectangleInsets;
-
-/**
- * @author  Mahi
- */
-public class GraphFrame extends JFrame implements FocusListener {
-
-	// PlotPanel plotpanel=null;
-
-	/**
-	 * @uml.property  name="points"
-	 */
-	ArrayList points;
-
-	// Plot2D aPlot;
-	double StartX;
-	// Plot2D aPlot;
-	double StartY;
[…]

### 2010-09-24 TrainTest/src/chartTest/LocalMaxFrame.java

@@ -1,294 +1,0 @@
-package chartTest;
-//import jahuwaldt.plot.Plot2D;
-
-import java.awt.BorderLayout;
-import javax.swing.JPanel;
-import javax.swing.JFrame;
-import javax.swing.SwingUtilities;
-
-import SketchMaster.Stroke.StrokeData.Stroke;
-import SketchMaster.Stroke.features.FeatureFunction;
-import SketchMaster.Stroke.features.StrokeFeatures;
-import SketchMaster.gui.DrawingSheet;
-import SketchMaster.gui.Events.HandleStroke;
-import SketchMaster.gui.Events.NewStrokeEvent;
-import SketchMaster.system.Recogniziers.SVMRecognizier;
-
-import java.awt.Rectangle;
-import java.awt.geom.Point2D;
-import java.util.ArrayList;
-import javax.swing.JButton;
-import java.awt.Dimension;
-
-public class LocalMaxFrame extends JFrame implements HandleStroke {
-
-	private static final long serialVersionUID = 1L;
-
-	private JPanel jContentPane = null;
-
-	private DrawingSheet MydrawingSheet = null;
-	
-	private SVMRecognizier reg;
-
-	private JButton jButton = null;
-
-	/**
-	 * This is the default constructor
-	 */
-	public LocalMaxFrame() {
-		super();
-		initialize();
-	}
-
-	/**
-	 * This method initializes this
-	 * 
-	 * @return void
-	 */
-	private void initialize() {
-		this.setSize(528, 500);
-		this.setContentPane(getJContentPane());
-		this.setTitle("Thhe test For local MAx");
-		initReg();
-	}
-
-	private void initReg() {
-		  
-		this.reg=new SVMRecognizier();
-	
-	}
-
-	/**
-	 * This method initializes jContentPane
-	 * 
-	 * @return javax.swing.JPanel
-	 */
-	private JPanel getJContentPane() {
-		if (jContentPane == null) {
-			jContentPane = new JPanel();
-			jContentPane.setLayout(null);
-			jContentPane.add(getMydrawingSheet(), null);
-			jContentPane.add(getJButton(), null);  // Generated
-		}
-		return jContentPane;
-	}
-
-	/**
-	 * This method initializes MydrawingSheet	
-	 * 	
-	 * @return SketchMaster.gui.DrawingSheet	
-	 */
-	private DrawingSheet getMydrawingSheet() {
-		if (MydrawingSheet == null) {
[…]

### 2010-09-24 TrainTest/src/settings/PatchTestSet.java

@@ -1,4367 +1,0 @@
-package settings;
-
-
-
-import java.io.BufferedReader;
-import java.io.BufferedWriter;
-import java.io.File;
-import java.io.FileNotFoundException;
-import java.io.FileOutputStream;
-import java.io.FileReader;
-import java.io.FileWriter;
-import java.io.IOException;
-import java.io.Serializable;
-import java.io.Writer;
-import java.text.SimpleDateFormat;
-import java.util.ArrayList;
-import java.util.Date;
-import java.util.HashMap;
-import java.util.Iterator;
-import java.util.Random;
-import java.util.Scanner;
-import java.util.Set;
-
-import org.apache.log4j.Level;
-import org.apache.log4j.Logger;
-import org.apache.poi.hssf.usermodel.HSSFRichTextString;
-import org.apache.poi.hssf.usermodel.HSSFRow;
-import org.apache.poi.hssf.usermodel.HSSFSheet;
-import org.apache.poi.hssf.usermodel.HSSFWorkbook;
-
-//import SketchMaster.swarm.polygonApproximations.polygonSolution;
-import SketchMaster.Stroke.StrokeData.Stroke;
-import SketchMaster.lib.GeneralUtils;
-import SketchMaster.swarm.polygonApproximations.polygonSolution;
-import SketchMaster.system.SystemSettings;
-
-import test.HandleDataSetEvents;
-import test.RunApp;
-import test.TestVertices;
-import test.TrainTest;
-
-public class PatchTestSet implements Serializable ,Runnable{
-	private static transient final Logger logger = Logger.getLogger(PatchTestSet.class);
-	private static transient final Logger appLogger = Logger.getLogger("AppLogging");
-	static int DATA_HSE=1;
-	static int  DATA_MINE_ELECTERICAL=2;
-	static int  DATA_MINE_DIGITAL=3;
-	static int  DATA_MINE_ELECTERICAL_R=4;
-	static int  DATA_MINE_DIGITAL_R=5;
-	static int DATA_MINE_SWARM_STROKES=9;
-	static int MaxRunsWithoutSave=50;
-	private int DataBase;
-	ArrayList <TestResults> resultList;
-	ArrayList <TestSketchSetting> testList;
-	 static boolean currentStop;
-	 TrainTest t;
-	 TestVertices v;
-	 private ArrayList<VertixTestResult> resultListVertices;
[…]

### 2010-09-24 TrainTest/src/settings/TestSketchSetting.java

@@ -1,1823 +1,0 @@
-/**
- * 
- */
-package settings;
-
-import java.io.BufferedReader;
-import java.io.File;
-import java.io.FileNotFoundException;
-import java.io.FileOutputStream;
-import java.io.FileReader;
-import java.io.IOException;
-import java.io.PrintStream;
-import java.io.Serializable;
-import java.util.ArrayList;
-import java.util.Collections;
-import java.util.Comparator;
-import java.util.Scanner;
-
-import org.apache.log4j.Logger;
-
-import com.sun.xml.internal.ws.api.pipe.NextAction;
-
-import SketchMaster.lib.GeneralUtils;
-import SketchMaster.swarm.polygonApproximations.polygonSolution;
-import SketchMaster.system.SystemSettings;
-import SketchMaster.system.SystemSettings;
-import SketchMaster.system.Recogniziers.RecognizierSystem;
-
-/**
- * @author Maha
- * 
- */
-public class TestSketchSetting implements Serializable, Cloneable {
-	// constants
-
-	public TestSketchSetting() {
-		super();
-		SystemSettings.DEBUG_MODE=false;
-	}
-
-	/**
-	 * 
-	 */
-	private static final long serialVersionUID = 1L;
-	private static transient final Logger logger = Logger
-			.getLogger(TestSketchSetting.class);
-	public final static int SEGMENTATION_ALGORITHM_TESTED = 0;
-	public final static int RECOGNITION_ALGORITHM_TESTED = 1;
-	public final static int RUN_MODE_TRAIN = 0;
-	public final static int RUN_MODE_TEST = 1;
-	public final static int PREV_TRAINING = 1;
-	public final static int PREV_NO_TRAIN = 0;
-	public final static int DATA_SET_TYPE_XML = 0;
-	public final static int DATA_SET_TYPE_DATABASE = 1;
-
-	public static final int RUBINE_RECOGNIZER = 0;
-	public static final int SYMBOL_RECOGNIZER = 1;
[…]

### 2010-09-24 TrainTest/src/settings/VertixTestResultNode.java

@@ -1,68 +1,0 @@
-package settings;
-
-public class VertixTestResultNode {
-	
-	int vertixCount=0;
-	int StrokePointsCount=0;
-	double Error;
-//	String categoryName;
-	
-	int CountOfSegments=0;//number of  segments types produced 
-
-	/**
-	 * @return the vertixCount
-	 */
-	public int getVertixCount() {
-		return vertixCount;
-	}
-
-	/**
-	 * @param vertixCount the vertixCount to set
-	 */
-	public void setVertixCount(int vertixCount) {
-		this.vertixCount = vertixCount;
-	}
-
-	/**
-	 * @return the strokePointsCount
-	 */
-	public int getStrokePointsCount() {
-		return StrokePointsCount;
-	}
-
-	/**
-	 * @param strokePointsCount the strokePointsCount to set
-	 */
-	public void setStrokePointsCount(int strokePointsCount) {
-		StrokePointsCount = strokePointsCount;
-	}
-
-	/**
-	 * @return the error
-	 */
-	public double getError() {
-		return Error;
-	}
-
-	/**
-	 * @param error the error to set
-	 */
-	public void setError(double error) {
-		Error = error;
-	}
-
-	/**
-	 * @return the countOfSegments
-	 */
-	public int getCountOfSegments() {
-		return CountOfSegments;
-	}
-
-	/**
-	 * @param countOfSegments the countOfSegments to set
-	 */
-	public void setCountOfSegments(int countOfSegments) {
-		CountOfSegments = countOfSegments;
-	}
-	
-}

### 2010-10-27 trunk/Swarm/swarmModified.tex
after elvasire editing and tradional pso

--- 2010-08-18 Swarm/tmp/swarmModified.tex
@@
-\documentclass[10pt]{article}
-\usepackage{spconf}
+\documentclass[12pt,1p,times]{elsarticle}%[10pt]{article}
+%% Use the option review to obtain double line spacing
+%% \documentclass[preprint,review,12pt]{elsarticle}
+
+%% Use the options 1p,twocolumn; 3p; 3p,twocolumn; 5p; or 5p,twocolumn
+%% for a journal layout:
+%% \documentclass[final,1p,times]{elsarticle}
+%% \documentclass[final,1p,times,twocolumn]{elsarticle}
+%% \documentclass[final,3p,times]{elsarticle}
+%\usepackage{spconf}
 \usepackage{amsmath}
 \usepackage[dvips,pdftex]{graphicx}
-\usepackage{named}
-\usepackage[dvips,pdftex]{hyperref}
+%\usepackage{named}
+\usepackage{hyperref}
+%\usepackage[dvips,pdftex]{hyperref}
 \usepackage{subfigure}
 \usepackage{array}
+\usepackage{amssymb}
+%\usepackage{epsfig}
+%\usepackage{epstopdf}
 
-\title{Enhanced Particle Swarm Optimization In Sketch Recognition}
-\twoauthors
-{Maha El Meseery\sthanks{Corresponding Author}, Mahmoud Fakhr El Din}
-	{Signals Processing Group \\ Computers and Systems Department \\
-  Electronic Research Institute\\Cairo, Egypt\\
-	melmeseery@eri.sci.eg, mafakhr@mcit.gov.eg}
-  {Tarek abouldahab} {Cairo Metro Company \\
-  Ministry of transport \\Cairo, Egypt\\ heshoaboda@hotmail.com}
+%% The lineno packages adds line numbers. Start line numbering with
+%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
+%% for the whole article with \linenumbers after \end{frontmatter}.
+%% \usepackage{lineno}
 
+%% natbib.sty is loaded by default. However, natbib options can be
+%% provided with \biboptions{...} command. Following options are
+%% valid:
+
+%%   round  -  round parentheses are used (default)
+%%   square -  square brackets are used   [option]
+%%   curly  -  curly braces are used      {option}
+%%   angle  -  angle brackets are used    <option>
+%%   semicolon  -  multiple citations separated by semi-colon
+%%   colon  - same as semicolon, an earlier confusion
+%%   comma  -  separated by comma
+%%   numbers-  selects numerical citations
+%%   super  -  numerical citations as superscripts
+%%   sort   -  sorts multiple citations according to order in ref. list
+%%   sort&compress   -  like sort, but also compresses numerical citations
+%%   compress - compresses without sorting
+%%
+%% \biboptions{comma,round}
+
+% \biboptions{}
 \begin{document}
-\maketitle
+
+\begin{frontmatter}
+%% Title, authors and addresses
+
+%% use the tnoteref command within \title for footnotes;
+%% use the tnotetext command for the associated footnote;
+%% use the fnref command within \author or \address for footnotes;
+%% use the fntext command for the associated footnote;
+%% use the corref command within \author for corresponding author footnotes;
+%% use the cortext command for the associated footnote;
+%% use the ead command for the email address,
+%% and the form \ead[url] for the home page:
+%%
+%% \title{Title\tnoteref{label1}}
+%% \tnotetext[label1]{}
+%% \author{Name\corref{cor1}\fnref{label2}}
+%% \ead{email address}
+%% \ead[url]{home page}
+%% \fntext[label2]{}
+%% \cortext[cor1]{}
+%% \address{Address\fnref{label3}}
+%% \fntext[label3]{}
+
+%% use optional labels to link authors explicitly to addresses:
+%% \author[label1,label2]{<author name>}
+%% \address[label1]{<address>}
+%% \address[label2]{<address>}
+
+\title{Sketch Recognition using Enhanced Particle Swarm Optimization}
+
+\author[Eri]{Maha El Meseery \corref{cor1} }%,
+\ead{melmeseery@eri.sci.eg}
+\author[Eri]{Mahmoud Fakhr El Din}
+\ead{ mafakhr@mcit.gov.eg}
+\author[metro]{Tarek abouldahab}
+\ead{heshoaboda@hotmail.com}
+
+\address[Eri]{Signals Processing Group, Computers and Systems Department,
+  Electronic Research Institute,Cairo, Egypt.}
+  \address[metro]{Cairo Metro Company, Ministry of transport ,Cairo, Egypt.}
+\cortext[cor1]{ Corresponding author}
+
+% \twoauthors
+% {\sthanks{Corresponding Author},}
+% 	{Signals Processing Group \\ Computers and Systems Department \\
+%   Electronic Research Institute\\Cairo, Egypt\\
+% 	melmeseery@eri.sci.eg, mafakhr@mcit.gov.eg}
+%   {} {Cairo Metro Company \\
+%   Ministry of transport \\Cairo, Egypt\\ heshoaboda@hotmail.com}
+
+%\maketitle
  \begin{abstract}
@@
 The goal of this paper is to improve our sketch recognition system by using an enhanced particle swarm optimization algorithm.
-The enhanced Particle Swarm Optimization (EPSO) is based on the perturbation of the particle velocity is used to optimally segment the strokes the user draws into meaningful geometric primitives.
+The Enhanced Particle Swarm Optimization (EPSO) is based on the perturbation of the particle velocity is used to optimally segment the strokes the user draws into meaningful geometric primitives.
 These geometric primitives are grouped to formulate symbols which are further identified using a Support Vector Machines (SVM) classifier.
@@
 \end{abstract}
-\pagenumbering{arabic}
+
+\begin{keyword}
+%% keywords here, in the form: keyword \sep keyword
+
+%% MSC codes here, in the form: \MSC code \sep code
+Sketch Recognition, Particle Swarm Optimization
+%% or \MSC[2008] code \sep code (2000 is the default)
+
+\end{keyword}
+\end{frontmatter}
+
+%\pagenumbering{arabic}
  \section{Introduction}
@@
 The sketch recognitions system is divided into three main steps 1) preprocessing, 2) segmentation and 3) Recognition.
+Figure \ref{fig:flowChart} shows the flow chart of the system; the first process is collecting stroke points then computing the curvature and possible dominant points.
 The following section provides a detailed description of each step.
-The main block of the system remains the same but their is some modification in the segmentation algorithm to improve its efficiency.
+The main block of the system remains the same but there are some modifications in the segmentation algorithm to improve its efficiency.
 The main modification is using Enhanced Particle swarm algorithm instead of Basic Particle swarm algorithms.
 Other modifications are explained in the details in the next sections.
+ \begin{figure}
+	\centering
+		\includegraphics[width=3.5in,height=6in]{FlowChart.jpg}
+	\caption{System flow chart} The chart shows the flow of process and data processed thought the system. 
+	\label{fig:flowChart}
+\end{figure}
+
 \subsection{Preprocessing}
@@
 	\item The particles are tested to make sure that the distance between every two successive dominate point is larger than $min_D$, where $min_D$ is 5\% of the total length of the stroke.
-Otherwise, one of these points are removed based on the error cased by its removal.
+Otherwise, one of these points are removed based on the error caused by its removal.
 	\item Each two successive segments are test to determine if they can be merged into a single segment.
@@
 Firstly, we tested recognition accuracy of shapes in the data set with both algorithms (\textsl{AlgS1} and \textsl{AlgS2}) using the whole feature set.
-The second experiment was tested on both algorithm but only using half the feature set.
-Both experiments were compared with the original sketch recognition system presented in \cite{mypaper}.
+The second experiment was tested on both algorithms but only using half the feature set.
+Both experiments were compared with the Tradional PSO sketch recognition system presented in \cite{mypaper}.
 
 Figures \ref{exp2} and \ref{exp1} show the recognition rate of each algorithm using the full feature set and half the features set respectively.
-The figures compares between the result that can be obtained by system in \cite{mypaper} and modifying the system using Enhanced PSO.
+The figures compares between the result that can be obtained using Traditional PSO by system in \cite{mypaper} and modifying the system using Enhanced PSO.
 The results show that the enhanced swarm optimization algorithm improves the segmentation of the system which leads to an increase in the recognition rate of the whole system by an average of $1.5\%$.
@@
   \centering
-  \includegraphics[width=3.3in]{Experiment92.jpg}
+  \includegraphics[width=3.3in]{Experiment_92.jpg}
   \caption[Recognition accuracy]%
-  {The chart compares between recognition rate of the full feature set on both the original system presented in \cite{mypaper} and Enhanced PSO.
-The results shows that Enhanced PSO increases the performance over the original system.
+  {The chart compares between recognition rate of the full feature set on both the traditional PSO system presented in \cite{mypaper} and Enhanced PSO.
+The results show that Enhanced PSO increases the performance over the  Traditional PSO.
 }
@@
   \centering
-  \includegraphics[width=3.3in]{Experiment68.jpg}
+  \includegraphics[width=3.3in]{Experiment_68.jpg}
   \caption[Recognition accuracy]%
-  {The chart compares between recognition rate of half the feature set on both the original system presented in \cite{mypaper} and Enhanced PSO.
-The results shows that Enhanced PSO increase the recognition rate.
+  {The chart compares between recognition rate of half the feature set on both the traditional PSO system presented in \cite{mypaper} and Enhanced PSO.
+The results show that Enhanced PSO increases the recognition rate.
 }
@@
 Results show that the enhanced PSO algorithm improves the segmentation on both algorithms which achieves better segmentation and final accuracy.
-The results proved that the Enhancement made on the PSO improves the efficiency of the Basic PSO algorithm without being dependable on special objective function or application.
+The results proved that the Enhancement made on the PSO improves the efficiency of the Traditional PSO algorithm without being dependable on special objective function or application.
 The system achieved an average overall improvement of more than 1\% over Basic PSO.
@@
 
-\bibliographystyle{IEEEbib}
+\bibliographystyle{elsarticle-num-names}
 \bibliography{../../../neededfiles/Bibliographies/Mybibliography}

### 2010-11-25 Swarm/swarmModified.tex
the final paper that is submitted

--- 2010-06-29 Swarm/swarmModified.tex
@@
-\documentclass[10pt]{article}
-\usepackage{spconf}
+\documentclass[a4paper, twoside]{article}
+
+\usepackage{icgst}
+%\documentclass[12pt,1p,times]{elsarticle}%[10pt]{article}
+%% Use the option review to obtain double line spacing
+%% \documentclass[preprint,review,12pt]{elsarticle}
+
+%% Use the options 1p,twocolumn; 3p; 3p,twocolumn; 5p; or 5p,twocolumn
+%% for a journal layout:
+%% \documentclass[final,1p,times]{elsarticle}
+%% \documentclass[final,1p,times,twocolumn]{elsarticle}
+%% \documentclass[final,3p,times]{elsarticle}
+%\usepackage{spconf}
 \usepackage{amsmath}
+
+%\usepackage{named}
+%\usepackage{hyperref}
+%\usepackage[dvips,pdftex]{hyperref}
+%\usepackage{subfigure}
+\usepackage{array}
+\usepackage{amssymb}
+\usepackage{epsfig}
+\usepackage{epstopdf}
 \usepackage[dvips,pdftex]{graphicx}
-\usepackage{named}
-\usepackage[dvips,pdftex]{hyperref}
-\usepackage{subfigure}
-\usepackage{array}
+%% The lineno packages adds line numbers. Start line numbering with
+%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
+%% for the whole article with \linenumbers after \end{frontmatter}.
+%% \usepackage{lineno}
 
-\title{Enhanced Particle Swarm Optimization}
-\twoauthors
-  {Tarek abouldahab\sthanks{Corresponding Author}} {Cairo Metro Company \\
-  Ministry of transport \\Cairo, Egypt\\ heshoaboda@hotmail.com}
- {Maha El Meseery, Mahmoud Fakhr El Din}
-	{Signals Processing Group \\ Computers and Systems Department \\
-  Electronic Research Institute\\Cairo, Egypt\\
-	melmeseery@eri.sci.eg, mafakhr@mcit.gov.eg}
+%% natbib.sty is loaded by default. However, natbib options can be
+%% provided with \biboptions{...} command. Following options are
+%% valid:
 
+%%   round  -  round parentheses are used (default)
+%%   square -  square brackets are used   [option]
+%%   curly  -  curly braces are used      {option}
+%%   angle  -  angle brackets are used    <option>
+%%   semicolon  -  multiple citations separated by semi-colon
+%%   colon  - same as semicolon, an earlier confusion
+%%   comma  -  separated by comma
+%%   numbers-  selects numerical citations
+%%   super  -  numerical citations as superscripts
+%%   sort   -  sorts multiple citations according to order in ref. list
+%%   sort&compress   -  like sort, but also compresses numerical citations
+%%   compress - compresses without sorting
+%%
+%% \biboptions{comma,round}
+
+% \biboptions{}
 \begin{document}
-\maketitle
+
+%\begin{frontmatter}
+%% Title, authors and addresses
+
+%% use the tnoteref command within \title for footnotes;
+%% use the tnotetext command for the associated footnote;
+%% use the fnref command within \author or \address for footnotes;
+%% use the fntext command for the associated footnote;
+%% use the corref command within \author for corresponding author footnotes;
+%% use the cortext command for the associated footnote;
+%% use the ead command for the email address,
+%% and the form \ead[url] for the home page:
+%%
+%% \title{Title\tnoteref{label1}}
+%% \tnotetext[label1]{}
+%% \author{Name\corref{cor1}\fnref{label2}}
+%% \ead{email address}
+%% \ead[url]{home page}
+%% \fntext[label2]{}
+%% \cortext[cor1]{}
+%% \address{Address\fnref{label3}}
+%% \fntext[label3]{}
+
+%% use optional labels to link authors explicitly to addresses:
+%% \author[label1,label2]{<author name>}
+%% \address[label1]{<address>}
+%% \address[label2]{<address>}
+
+%\title{Sketch Recognition using Enhanced Particle Swarm Optimization}
+
+\twocolumn[
+\scalebox{1.00}{\includegraphics[width=1.00\textwidth]{aimlhead}}
+\title{Sketch Recognition using Enhanced Particle Swarm Optimization}
+\vspace{0.1in}
+\author{}  
+\begin{center}
+Maha El Meseery\footnote{Corresponding Author }, Mahmoud Fakhr El Din, Tarek abouldahab\\
+\textit{Electronic Research Institute,} Cairo, Egypt,\\
+\textit{ Cairo Metro Company, Ministry of transport }Cairo, Egypt\\
+{melmeseery@eri.sci.eg, mafakhr@mcit.gov.eg,  heshoaboda@hotmail.com},\\
+%\texttt{http://www.faculty.university.edu.any}
+\end{center}
+\vspace{0.25in} 
+]
+
+% \twoauthors
+% 
+%  {Maha El Meseery, Mahmoud Fakhr El Din}
+% 	{Signals Processing Group \\ Computers and Systems Department \\
+%   Electronic Research Institute\\Cairo, Egypt\\
+% 	melmeseery@eri.sci.eg, mafakhr@mcit.gov.eg} 
+% 	 {Tarek abouldahab\sthanks{Corresponding Author}} {Cairo Metro Company \\
+%   Ministry of transport \\Cairo, Egypt\\ heshoaboda@hotmail.com}
+
+% \author[Eri]{Maha El Meseery \corref{cor1} }%,
+% \ead{melmeseery@eri.sci.eg}
+% \author[Eri]{Mahmoud Fakhr El Din}
+% \ead{ mafakhr@mcit.gov.eg}
+% \author[metro]{Tarek abouldahab}
+% \ead{heshoaboda@hotmail.com}
+% 
+% \address[Eri]{Signals Processing Group, Computers and Systems Department,
+%   Electronic Research Institute,Cairo, Egypt.}
+%   \address[metro]{Cairo Metro Company, Ministry of transport ,Cairo, Egypt.}
+% \cortext[cor1]{ Corresponding author}
+
+% \twoauthors
+% {\sthanks{Corresponding Author},}
+% 	{Signals Processing Group \\ Computers and Systems Department \\
+%   Electronic Research Institute\\Cairo, Egypt\\
+% 	melmeseery@eri.sci.eg, mafakhr@mcit.gov.eg}
+%   {} {Cairo Metro Company \\
+%   Ministry of transport \\Cairo, Egypt\\ heshoaboda@hotmail.com}
+
+%\maketitle
  \begin{abstract}
 %This study
+
 Sketches and drawings are widely used as a simple method of expressing thought and designs.
-The goal of this paper is developing a sketch recognition system that facilitates using sketches in computer systems.
-The system uses Particle Swarm Optimization (PSO) algorithm to optimally segment the strokes the user draws into meaningful geometric primitives.
+The goal of this paper is to improve our sketch recognition system by using an enhanced particle swarm optimization algorithm.
+The Enhanced Particle Swarm Optimization (EPSO) is based on the perturbation of the particle velocity is used to optimally segment the strokes the user draws into meaningful geometric primitives.
 These geometric primitives are grouped to formulate symbols which are further identified using a Support Vector Machines (SVM) classifier.
-Two different segmentation algorithms were evaluated and results show that both algorithms improve the final recognition results.
-Experiments were conducted on three different dataset, simple presentation (Hs-DS), Electrical Design (EL-DS) and Logic Design (LD-DS) symbols.
-Results show that the effectiveness of the proposed system on various dataset especially on datasets with low shape diversity as in Logic Design dataset (LD-DS).
-% sketch are divided Users draw symbols and sketches using a set of pen strokes.
-%Sketch recognition is defined as the process of identifying symbols that users draw using single or multiple %strokes.
-Usually, users draw strokes using pens.
-The sketch recognition system immediately interprets their %strokes as objects that can be easily manipulated.
-This paper uses Particle Swarm Optimization (PSO) to divide %the strokes the user draws into meaningful geometric primitives.
-These geometric primitives are grouped to %formulate symbols which are further identified.
-The results show that using PSO improves segmentation results %which guide the symbol recognition phase.
-As for recognition we use Support Vector Machines (SVM) classifier %which further improves the final recognition accuracy.
+Experiments were conducted on the benchmark dataset of simple presentation (Hs-DS) symbols.
+Results show that that the enhanced particle swarm optimization improves the final recognition accuracy.
+
 \end{abstract}
-\pagenumbering{arabic}
+
+% \begin{keyword}
+% %% keywords here, in the form: keyword \sep keyword
+% 
+% %% MSC codes here, in the form: \MSC code \sep code
+% Sketch Recognition, Particle Swarm Optimization
+% %% or \MSC[2008] code \sep code (2000 is the default)
+% 
+% \end{keyword}
+%\end{frontmatter}
+\textit{\textbf{Keywords:} Sketch Recognition, Particle Swarm Optimization.}
+%\pagenumbering{arabic}
  \section{Introduction}
 
-Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.
-Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.
-Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.
-Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.
-Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.
+The field of symbol recognition has gained interest in the last few years.
+Scientists generally and engineers specifically express thoughts and designs using sketches.
+Engineers use sketches to exchange designs as a natural method of communication rather than writing or speaking.
+Design engineers need a powerful computer based system with symbol recognition to help them design, manipulate and store sketches more effectively than using only papers.
+Increasing the interaction between computers and users in sketch and CAD systems has been the reason for the emerging of a few advanced sketch recognition systems.
+Sketch recognition is the process of identifying user strokes into meaningful symbols that can be further used.
+
+Sketch recognition is divided into three main steps preprocessing, segmentation and symbol recognition.
+The preprocessing phase captures user input stroke points and collects basic information about the stroke then proceeds to remove noise and compute basic statistical and geometrical information.
+In segmentation phase, the strokes are divided into a set of simple geometrical primitives or segments.
+In the third phase, sketch recognition, strokes and segments are clustered to formulate symbols that can be recognized by a classifier system.
+
+This paper modifies the sketch recognition system presented in \cite{mypaper}, to use enhanced particle swarm algorithm (EPSO) to optimally segment user's strokes.
+As in the Original system the users can draw symbols using any number of strokes in any order they like.
+The system segments each stroke using two different EPSO algorithms.
+Segmentation is based on dividing the digital curve into polygon\cite{PolygonApproximationPSO}.
+To handle complex strokes, our system uses a second segmentation EPSO algorithm to segment strokes into a set of lines and circular arcs.
+After segmentation, a symbol features vector composed of different statistical, geometrical and spatial features is deduced.
+The final classification step, SVM classifier uses the computed feature vector to classify the segments into one of the previously trained classes.
+
+The remaining of the paper is as following; Section \ref{sec:ParticleSwarmAlgorithm} describes the Basic particle swarm algorithm in details.
+Enhanced Particle Swarm Algorithm is described in Section \ref{sec:EnhancedSwarmAlgorithm}.
+Section \ref{Sysdisc} provide details on preprocessing, segmentation and recognition process.
+Section \ref{sec:Experiments} describes the experiments that were preformed.
+We conclude in Section \ref{ConclusionandFutureWork}.
 
@@
 
-\section{Enhanced Partilce Swarm}
-This paper modifies the general PSO algorithm by changeing the local term into a random term using the following equation .  
+\section{Enhanced Particle Swarm}
+\label{sec:EnhancedSwarmAlgorithm}
+In the above basic PSO, the velocity update is based on three terms an inertia term, local cognitive ($l_{best}$) term and global cognitive ($g_{best}$)term.
+The particle local cognitive term is associated with the particle own history without consideration of other swarm particles locations and histories which leads to a lake of diversity and sub optimum solutions.
+Therefore, we are trying to add more social influence in the particle movement by using the weighted difference of the position vectors of any other two distinct particles randomly chosen from the swarm to modify the local cognitive term.
+The modified term adds social behavior of other particle and keeps diversity while searching for the global optimum.
+The new social cognitive term of the particle is shifted to the new location only if it yields a fitness value better than obtained using the standard PSO\cite{stockPaper}.
+
+This modification is done as follows:-
+  \begin{enumerate}
+\item For each particle $p_{i}$ in the swarm two other distinct particles, $p_L$ and $p_M$ are selected randomly where $i\neq L\neq M$.
+The difference between their positional coordinates $p_{d}$ is taken as a difference vector in equation \ref{eqmod}.
  \begin{equation}
-  v_{ij}  = v_{ij} \omega + c_1 r_1 (p_{iRandom_i}  - p_{iRandom_j} ) + c_2 r_2 (gbest_{ij}  - p_{ij} ),
-  \label{eq:ModSwarm}
+ \label{eqmod}
+ p_{dj}= p_{Lj}-p_{Mj} 
+ \end{equation}
+where $p_{dj}$ represent the $j$th dimension in the difference vector and $p_{Lj}$ represent the $j$th dimension in the $L$th particle.
+
+Then the associated velocity vector of particle $p_{i}$ is updated according to this difference vector as in equation \ref{eq:mod1}.
+In essence the local cognitive term of the velocity update formula in \ref{eq:Swarm} is replaced with the vector differential operator $p_d$ to produce some additional exploration capability.
+ \begin{equation}
+ v^{T}_{ij}  = v_{ij} \omega + c_1 r_1 (p_{dj}) + c_2 r_2 (gbest_{ij}  - p_{ij} ),
+ \label{eq:mod1}
+ \end{equation}
+\item The new trial position $p^{T}_{ij}$ is created by adding the updated velocity $v^{T}_{ij}$ in Eq. \ref{eq:mod1}.
+  \begin{equation}
+ \label{eq:mod2}
+p^{T}_{ij}=p_{ij}+ v^{T}_{ij} ,
+ \end{equation}
+\item The new position $p^{T}_{ij}$ is used to generate the new trial objective fitness function value $fitness(p^{T})$.
+The fitness value is calculated based on this new trial position vector and the particle is shifted to this trial position if the trial objective function is better than the original objective function associated with standard PSO.
+Thus the target particle is relocated using the following equation (Eq.
+\ref{eq:objective})
+  \begin{equation}
+ \label{eq:objective}
+p^{new}_{ij}\Leftarrow 
+\{
+\begin{array}{c} 
+ p^{T}_{ij} \quad \quad if\quad fitness(p^{T})> fitness(p^{old})  \\
+ p_{ij} \quad \quad Otherwise\quad  
+\end{array}\}
+ \end{equation}
+ where $fitness(p^{T})$ is the objective fitness value of the new trail position $ p^{T}$ and  $fitness(p^{old})$is the objective fitness value of the new basic position $ p{i}$ computed by equation \ref{eq:Swarm1}.
+   \end{enumerate}
+
+\section{System Discribtion  }\label{Sysdisc}
+%\label{sec:application}
+The modified swarm algorithm is used to improve the performance of the sketch recognition system in \cite{mypaper}.
+The sketch recognitions system is divided into three main steps 1) preprocessing, 2) segmentation and 3) Recognition.
+Figure \ref{fig:flowChart} shows the flow chart of the system; the first process is collecting stroke points then computing the curvature and possible dominant points.
+The following section provides a detailed description of each step.
+The main block of the system remains the same but there are some modifications in the segmentation algorithm to improve its efficiency.
+The main modification is using Enhanced Particle swarm algorithm instead of Basic Particle swarm algorithms.
+Other modifications are explained in the details in the next sections.
+ \begin{figure}
+	\centering
+		\includegraphics[width=3.5in,height=6in]{FlowChart.jpg}
+	\caption{System flow chart} The chart shows the flow of process and data processed thought the system. 
+	\label{fig:flowChart}
+\end{figure}
+
+\subsection{Preprocessing}
+\label{Prepross}
+ Speed and time difference information were widely used in sketch understanding systems \cite{earlyprocess}.
+It was noted that time difference provides more distinct maxima than the speed information.
+Agar \cite{polygonfeedback31} mentions that the pointing device (i.e.
+the mouse) sampling rate is the reason for this phenomenon.
+ In the presented system, we computed time difference, direction, speed and curvature of each point along a stroke.
+The speed is calculated as $v=\Delta s/\Delta t$ where $t$ is the time difference between two points and $s$ is the length between them.
+The direction $d$ is calculated as the angle between two vectors and curvature is considered as the change in direction with respect to length i.e.
+$c= \Delta d/\Delta s$.
+Dominant points are characterized by lower speed values and higher curvature and direction values.
+
+After the system computes all the speed, time difference and curvature information it proceeds to detect the points with low velocity and high curvatures.
+Using simple differentiation to detect local extreme points resulted in false points due to the non smooth curves.
+Hence, the system adopted a process presented by \cite{earlyprocess}, where the mean of the curve is calculated.
+Then a threshold \textit{th} is used to separate the curve into regions $Region_i$; each region $Region_i$ is defined as a range of points, where the curve values are either above or below the threshold \textit{th}.
+Those regions are further processed to find the maximum point $Max(Region_i)$ of each region $Region_i$.
+The stroke points $p_i(x,y)$ that correspond to those maximum values are labeled as \textit{possible dominant points} $P_{pd}$.
+The system repeats this process for curvature, time difference and speed information.
+All the points labeled as possible dominant points $P_{pd}$ are saved into a single array.
+Figure \ref{fig:LabelsPPD} shows the particles labeled as Possible dominant point $P_{pd}$ by the preprocessing, it is noticed the redundancy of some $P_{pd}$ points.
+% (as shown are redundant)
+\subsection{Segmentation}
+\label{seg}
+In the segmentation stage the input stroke is divided into a set of primitives.
+As shown in Figure \ref{fig:segblock} first an attempt is made to fit the stroke points into a curve or an ellipse using a minimum square error fitting algorithm \cite{ellipsefit}.
+The ellipse fitting step helps to prevent the system from over segmenting the stroke into multiple lines or curves if the input stroke is an ellipse.
+If the stroke proved to be an ellipse arc then the segmentation process ends and the system proceeds to the next step.
+Otherwise, the stroke is passed to two further segmentation algorithms that divide the stroke to either lines or lines and curves.
+After two different segmentations are generated, the system chooses the segmentation with the minimum error.
+The following sections describe in details the ellipse detection algorithm and the two segmentation algorithms used to divide the stroke.
+ \begin{figure}
+	\centering
+		\includegraphics[scale=0.48]{blockSmall.jpg}
+	\caption{Segmentation Block}
+	\label{fig:segblock}
+\end{figure}
+\subsection{Ellipse fitting}
+
+This process tries to fit the stroke points into an ellipse arc; it starts with computing the center of the stroke bounding box.
+The bounding box center point is used as the first estimation of the center of the ellipse.
+The axes of the ellipse are estimated as the $width/2$ and $height/2$ of the stroke bounding box where $width$ is width and $height$ is height of the bonding box.
+The least square fitting algorithm \cite{chernov} is used to minimize the fitting error of the ellipse Equation (\ref{eq:circleFit})
+
+\begin{equation}
+E = \sum\limits_{i = 0}^N {\frac{{(x_i - x_0 )}}{{a^2 }}^2  + \frac{{(y_i - y_0 )}}{{b^2 }}^2  - 1} 
+\label{eq:circleFit}
 \end{equation}
 
-where $iRandom_i$ and $Random_j$ are two particles in the solution space.
-This modification replace the distance between the current particle and the local best with two random particle from the space.
-The effect of this change in more expolration and avoiding of local min problems.
-Equation \ref{eq:ModSwarm} is computed and the new particle locaiton is calculated.
-The new location is compared the location of particle if with equation \ref{eq:Swarm} if the modified location is better (mean better fitness function) then the new location is selected otherwise the old equation is used.
-\section{Applications }
-The modified swarm algorithm is used to improve the performance of the sketch recoginition system in \cite{myPaper}. The
+ where $N$ is number of points in the stroke, $a,b$ are the length of ellipse axes, $x_0$ \& $y_0$ are the coordinates of the center point, $x_i$ \& $y_i$ are the coordinates of point $i$ in the stroke.
+A list of new values for $x_0$ , $y_0$ ,$a$ and $b$ are generated randomly from the older values with small increments after each loop.
+After few iteration, the final fit error of the estimated ellipse is reported.
+A modification to the system in \cite{mypaper} is trying to add another measure to compute the efficiency of the final estimated ellipse.
+Equation \ref{eq:circleError} ensures that the drawn percentage of ellipse is considered.
+This eliminates fitting a line into very large ellipse but leaves small ellipses to be fitted as a partial or full ellipse.
 
-\cite{myPaper}  
+ \begin{equation}
+eff= (P_{percent}/E)
+\label{eq:circleError}
+\end{equation}
+ \begin{equation}
+P_{percent}  = L_{stroke} /P_{ellipse} 
+\label{eq:ErrorArea}
+\end{equation}
+where $E$ is the error computed by Equation(\ref{eq:circleFit}), $L_{stroke}$ total length of stroke and $P_{ellipse} $ is the perimeter of the estimated ellipse.
+If $eff$ is more than threshold $th_{Ellipse}$\footnote{By trial and error best threshold found was $th_{Ellipse}=0.2$} then the stroke is segmented as an ellipse otherwise the system proceeds to the next step.
+
+\subsection{Non ellipse fitting algorithm}
+\label{subsubsec:Discreteparticleswarmalgorithm}
+\begin{figure}
+	\centering
+
+	 \includegraphics[scale=0.7]{pso1.jpg}			
+	\caption{DPSO algorithm encoding}
+	\label{fig:pso1}
+\end{figure}
+Two DPSO algorithms are used to generate different segmentations for each stroke the user draws.
+For each segmentation generated, an error is evaluated.
+The segmentation with the minimum error value is chosen as the best stroke segmentation (Figure \ref{fig:segblock}).
+The problem definition is the same in both algorithms but they differ in the method they use to compute fitness and error functions.
+\begin{description}
+	\item[ Problem definition:] The input stroke $S$  with $N$ points can be represented by set $S = \left\{ {x_1 ,x_2  \ldots x_N }\right\}$ where $x_i$ is the location of the point $i$.
+The swarm algorithms consist of $M$ agents which are represented by the set $A = \left\{ {P_i \left| {i = 1,2 \cdots M} \right.} \right\}$ where $P_i$ is a single solution particle from the solution space.
+Each particle decodes the problem using a binary array with the same length $N$ as the input stroke.
+
+Therefore, the system represents each particle $P_i$ by $P_i = \left\{ {p_{ij} \left| {j = 1,2 \cdots N} \right.} \right\}$ where $p_{ij}$ has only two values a)1 ($p_{ij}=1$); means that this point ($j$) is a dominate point, or b) 0($p_{ij}=0$) means this point ($j$) is not a dominate point.
+Figure\ref{fig:pso1} shows a particle encoding in the DPSO system.
+
+	\item[Fitness function:] The fitness function and error calculation are different in each of the two \textit{DPSO} algorithms. 
+	\end{description}
+\subsubsection{\textit{Polygon approximation\textsl{AlgS1}}}
+%\\
+The approximation error is computed by the equation \ref{eq:ErrorSwarm1}. The graphical meaning of the error is shown in Figure\ref{fig:LabelsPSO}.
+\begin{equation}
+E=\sum\nolimits_{i = 0}^M e ( \widehat{x_ix_{i+1}},\overline{x_i x_{i+1}})
+\label{eq:ErrorSwarm1}
+\end{equation} where the arc $\widehat{x_ix_j}$ is defined as the consecutive set of points from point $x_i$ to point $x_{j}$ as in $x_i,x_{i+1} \cdots,x_j$.
+The line $\overline{x_i x_j} $ as the straight line connecting point $x_i$ to point $x_j$, $M$ is the number of dominant points in this solution as generated by the swarm algorithm.
+The error $e ( \widehat{x_ix_j},\overline{x_i x_j})$ is computed as the sum of squared perpendicular distance from every point along the arc $\widehat{x_ix_j}$ to the line $\overline{x_i x_j}$.
+The fitness is computed using equation %\ref{eq:fitnessSwarm1}
+\begin{equation}
+\max fitness(p_i ) = \left\{ {\begin{array}{*{20}c}
+   { - E/\varepsilon N} & {ifE > \varepsilon ,}  \\
+   {D/\sum\limits_{j = 1}^N {p_{ij} } } & {otherwise}  \\
+\end{array}} \right.
+\label{eq:fitnessSwarm1}
+\end{equation} where $N$ is the number of points in the stroke, $D$ is the number of points in the solution that was previously labeled as a possible dominant point ($P_{pd}$), $E$ is the computed error and $\varepsilon$ is the error threshold.
+It should be noticed that when the error is larger than the threshold $\varepsilon$ the fitness is given a -ve value to lower the fitness value of the solution.
+Otherwise the system favors the lower number of vertices.
+
+ \subsubsection{\textit{Hybrid Fitting \textsl{AlgS2}}}
+The algorithm has the same problem formulation but different fitness and error functions are used.
+An attempt is made to fit each segment into a line or circular arc.
+The errors of both circle and line fit estimations are computed for each segment $S_i=\widehat{x_ix_j}$.
+The approximation with the lower error value is chosen as the final approximation of this segment $S_i$\cite{CruveDivisionSwarm}.
+The sum of the approximation errors of all segments is the total error of the particle.
+The total error of the particle is computed by equation %\ref{eq:errorSwarm2}
+ \begin{equation}
+E=\sum\nolimits_{i = 0}^M e(D_i) 
+\label{eq:errorSwarm2}
+\end{equation}where $M$ is the number of segments in the solution as generated by the swarm algorithm, $D_i$ is the minimum approximation error of curve and line approximations $min(d_c,d_l)$ where $d_c$ is the circle approximation error and $d_l$ is the line approximation error as computed by \cite{CruveDivisionSwarm}.
+The fitness is computed by the equation %\ref{eq:fitnessSwarm2}
+\begin{equation}
+\max fitness(P_i ) = \frac{1}{{E \times M^k }}
+\label{eq:fitnessSwarm2}
+\end{equation} where $E$ is the error and $M$ is the number of segments and $k$ is a parameter tweaked to get minimum number of segments.
+The larger $k$ is, the more effect the number of segments will have.
+For our system, $k$ is selected to be 0.5\cite{CruveDivisionSwarm}.
+
+As a modification to the system in \cite{mypaper}. After each iteration, of the swarm algorithm (\textsl{AlgS1} and \textsl{AlgS2}), each particle is refined using the following procedures: 
+\begin{enumerate}
+	\item For each particle $P_i$ each dominant point $P_{ij}$ is checked to find if it was labeled before as a \textit{possible dominant point} $P_{pd}$ (computed as in section \ref{Prepross}).
+If it was not labeled the point $P_{ij}$ is moved to the nearest labeled point.
+This ensures that all of the points generated by the DPSO are possible dominant points $P_{pd}$.
+	\item The particles are tested to make sure that the distance between every two successive dominate point is larger than $min_D$, where $min_D$ is 5\% of the total length of the stroke.
+Otherwise, one of these points are removed based on the error caused by its removal.
+	\item Each two successive segments are test to determine if they can be merged into a single segment.
+If merging the two segments have the same or smaller the error, the two segments are merged.
+This step ensures that the final segmentation of the system has minimum number of segments.
+\end{enumerate}
+
+\subsection{Recognition}
+\label{sec:Recognition}
+After the user draws all strokes of the symbol, the set of un-recognized strokes is grouped together along with their segmentation as input to the feature extraction process.
+A composite set of features is used to generate a single feature vector.
+The features used consist of Rubine feature set,  Zernike moments of order 10 \cite{HeloiseBeautification}, ink density as well as some structural and spatial information like number of perpendiculars lines ,number of parallel lines and number of different types of primitives in each symbol.
+After computing the features the symbol is introduced to the (SVM) classifier.
+
+\cite{mypaper}  
 \section{Experiments} 
- The goal of the system is to modify the segmentation and recognition of the sketch system.
-It made the same type of experiement done in \cite{myPaper}.
-The dataset is divided into 13 category of goemetric symbols
-\section{Conclusion and Future Work}
-Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.
-Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.
-Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.
-Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.
-Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.Plain text.
+\label{sec:Experiments}
+In our experiment we used the same benchmark dataset used in \cite{mypaper}, the dataset was collected by Hse and Newton\cite{HeloiseBeautification}.
+The data are drawn by 16 users each of them was asked to draw 13 shapes from 30 to 50 times.
+Figure \ref{fig:symbolSet} shows a set of the shapes used in the data set.
+Five different splits were generated to divide the dataset into training and test sets.
+The results displayed are the average recognition accuracy of the five splits.
+The accuracy is computed as the number of correctly recognized samples divided by the total number of samples in the test.
 
-\bibliographystyle{IEEEbib}
+ \begin{figure}
+  \centering
+
+		\includegraphics[scale=0.5]{symbolSet.PNG} 
+		\caption[The Symbol Set] {The symbol set}   \label{fig:symbolSet}
+	 \end{figure}
+
+To demonstrate the change in accuracy is due to the enhanced swarm optimization algorithm we tested the recognition system using two sets of features.
+Firstly, we tested recognition accuracy of shapes in the data set with both algorithms (\textsl{AlgS1} and \textsl{AlgS2}) using the whole feature set.
+The second experiment was tested on both algorithms but only using half the feature set.
+Both experiments were compared with the Tradional PSO sketch recognition system presented in \cite{mypaper}.
+
+Figures \ref{exp2} and \ref{exp1} show the recognition rate of each algorithm using the full feature set and half the features set respectively.
+The figures compares between the result that can be obtained using Traditional PSO by system in \cite{mypaper} and modifying the system using Enhanced PSO.
+The results show that the enhanced swarm optimization algorithm improves the segmentation of the system which leads to an increase in the recognition rate of the whole system by an average of $1.5\%$.
+The increase in recognition rate is consistent in all swarm algorithms and in both experiments.
+    \begin{figure}
+  \centering
+  \includegraphics[width=3.3in]{Experiment_92.jpg}
+  \caption[Recognition accuracy]%
+  {The chart compares between recognition rate of the full feature set on both the traditional PSO system presented in \cite{mypaper} and Enhanced PSO.
+The results show that Enhanced PSO increases the performance over the  Traditional PSO.
+}
+   \label{exp2}
+\end{figure}
+\begin{figure}
+  \centering
+  \includegraphics[width=3.3in]{Experiment_68.jpg}
+  \caption[Recognition accuracy]%
+  {The chart compares between recognition rate of half the feature set on both the traditional PSO system presented in \cite{mypaper} and Enhanced PSO.
+The results show that Enhanced PSO increases the recognition rate.
+}
+   \label{exp1}
+\end{figure}
+
+\section{Conclusion }
+ \label{ConclusionandFutureWork}
+
+This paper used an Enhanced-PSO to modify a new sketch recognition using DPSO represented in \cite{mypaper}.
+Results show that enhancing the solutions that the \textit{PSO} in general generates an optimized stroke segmentation which improves the final recognition rate.
+The final recognition improved by modifying the particle optimization algorithm without depending on the higher level features.
+The experiments evaluated two \textit{DPSO} different segmentation algorithms (AlgS1 and AlgS2) on simple presentation set dataset.
+Results show that the enhanced PSO algorithm improves the segmentation on both algorithms which achieves better segmentation and final accuracy.
+The results proved that the Enhancement made on the PSO improves the efficiency of the Traditional PSO algorithm without being dependable on special objective function or application.
+The system achieved an average overall improvement of more than 1\% over Basic PSO.
+
+ For the future work, it is suggested that we continue to improve both the enhanced PSO segmentation algorithm.
+Also, a possible extension of this research is to improve the recognition algorithm of the Sketch Recognition system.
+Other area of enhancements is the features extraction methods.
+Introducing more spatial and geometrical features is believed to improve classifications.
+
+\bibliographystyle{plain}
 \bibliography{../../../neededfiles/Bibliographies/Mybibliography}

### 2010-11-25 trunk/Swarm/swarmModified.tex
edit to sut the format for FCI

@@
-\documentclass[12pt,1p,times]{elsarticle}%[10pt]{article}
+%\documentclass[a4paper, twoside]{article}
+
+%%\usepackage{icgst}
+\documentclass[12pt,times,a4paper]{article}%[10pt]{article}
 %% Use the option review to obtain double line spacing
@@
 \usepackage{amsmath}
-\usepackage[dvips,pdftex]{graphicx}
 %\usepackage{named}
-\usepackage{hyperref}
+%\usepackage{hyperref}
 %\usepackage[dvips,pdftex]{hyperref}
-\usepackage{subfigure}
+%\usepackage{subfigure}
 \usepackage{array}
@@
 %\usepackage{epstopdf}
+%\usepackage{xltxtra}
+\usepackage{graphicx}
+\usepackage{sectsty}
+\sectionfont{\normalsize  }
+\subsectionfont{ \normalsize\it \textbf }
+\subsubsectionfont{\normalsize \it \textbf }
+%\paragraphfont{\large\it}
+%\subparagraphfont{\normalsize\it}
 
@@
 % \biboptions{}
+\setlength\hoffset{0in}
+% \setlength\topmargin{0in}
+% \setlength\headheight{0in}
+% \setlength\headsep{-0.5in}
+% \setlength\textheight{10.0in}
+ \setlength\textwidth{7in}
+ \setlength\oddsidemargin{-0.5in}
+ \setlength\evensidemargin{-0.5in}
+% \setlength\parindent{0.25in}
+% \setlength\parskip{0.25in} 
+\linespread{1.5}
+
 \begin{document}
 
-\begin{frontmatter}
+%\begin{frontmatter}
 %% Title, authors and addresses
@@
 
-\author[Eri]{Maha El Meseery \corref{cor1} }%,
-\ead{melmeseery@eri.sci.eg}
-\author[Eri]{Mahmoud Fakhr El Din}
-\ead{ mafakhr@mcit.gov.eg}
-\author[metro]{Tarek abouldahab}
-\ead{heshoaboda@hotmail.com}
-
-\address[Eri]{Signals Processing Group, Computers and Systems Department,
-  Electronic Research Institute,Cairo, Egypt.}
-  \address[metro]{Cairo Metro Company, Ministry of transport ,Cairo, Egypt.}
-\cortext[cor1]{ Corresponding author}
+% \twocolumn[
+% \scalebox{1.00}{\includegraphics[width=1.00\textwidth]{aimlhead}}
+% \title{Sketch Recognition using Enhanced Particle Swarm Optimization}
+% \vspace{0.1in}
+% \author{}  
+% \begin{center}
+% Maha El Meseery\footnote{Corresponding Author }, Mahmoud Fakhr El Din, Tarek abouldahab\\
+% \textit{Electronic Research Institute,} Cairo, Egypt,\\
+% \textit{ Cairo Metro Company, Ministry of transport }Cairo, Egypt\\
+% {melmeseery@eri.sci.eg, mafakhr@mcit.gov.eg,  heshoaboda@hotmail.com},\\
+% %\texttt{http://www.faculty.university.edu.any}
+% \end{center}
+% \vspace{0.25in} 
+% ]
 
 % \twoauthors
+% 
+%  {Maha El Meseery, Mahmoud Fakhr El Din}
+% 	{Signals Processing Group \\ Computers and Systems Department \\
+%   Electronic Research Institute\\Cairo, Egypt\\
+% 	melmeseery@eri.sci.eg, mafakhr@mcit.gov.eg} 
+% 	 {Tarek abouldahab\sthanks{Corresponding Author}} {Cairo Metro Company \\
+%   Ministry of transport \\Cairo, Egypt\\ heshoaboda@hotmail.com}
+
+% \author[Eri]{Maha El Meseery \corref{cor1} }%,
+% \ead{melmeseery@eri.sci.eg}
+% \author[Eri]{Mahmoud Fakhr El Din}
+% \ead{ mafakhr@mcit.gov.eg}
+% \author[metro]{Tarek abouldahab}
+% \ead{heshoaboda@hotmail.com}
+% 
+% \address[Eri]{Signals Processing Group, Computers and Systems Department,
+%   Electronic Research Institute,Cairo, Egypt.}
+%   \address[metro]{Cairo Metro Company, Ministry of transport ,Cairo, Egypt.}
+% \cortext[cor1]{ Corresponding author}
+
+\author{
+\small{
+Maha El Meseery \footnote{Corresponding Author} \textsuperscript{a}, Mahmoud Fakhr El Din \textsuperscript{a}, Tarek abouldahab \textsuperscript{b}}\\
+ \textsuperscript{a} \small{Signals Processing Group, Computers and Systems Department,}\\ \small{ Electronic Research Institute, Cairo, Egypt.}\\
+  \small {\textsuperscript{b}Cairo Metro Company ,
+   Ministry of transport , Cairo, Egypt.}\\
+   \small{melemseery@eri.sci.eg,  mafakhr@mcit.gov.eg, heshoaboda@hotmail.com
+   }
+}
+\date{}
 % {\sthanks{Corresponding Author},}
@@
 % 	melmeseery@eri.sci.eg, mafakhr@mcit.gov.eg}
-%   {} {Cairo Metro Company \\
+%   {Tarek abouldahab} {Cairo Metro Company \\
+%   Ministry of transport \\Cairo, Egypt\\ heshoaboda@hotmail.com}
+%  {Maha El Meseery \sthanks{Corresponding Author}, Mahmoud Fakhr El Din}
+% 	{Signals Processing Group \\ Computers and Systems Department \\
+%   Electronic Research Institute\\Cairo, Egypt\\
+% 	melmeseery@eri.sci.eg, mafakhr@mcit.gov.eg}
+% 	 {Tarek abouldahab\sthanks{Corresponding Author}} {Cairo Metro Company \\
 %   Ministry of transport \\Cairo, Egypt\\ heshoaboda@hotmail.com}
 
-%\maketitle
+\maketitle
  \begin{abstract}
+
 %This study 
-
 Sketches and drawings are widely used as a simple method of expressing thought and designs.
@@
 Experiments were conducted on the benchmark dataset of simple presentation (Hs-DS) symbols.
-Results show that that the enhanced particle swarm optimization improves the final recognition accuracy.
+Results show that the enhanced particle swarm optimization improves the final recognition accuracy.
 
@@
 
-\begin{keyword}
-%% keywords here, in the form: keyword \sep keyword
-
-%% MSC codes here, in the form: \MSC code \sep code
-Sketch Recognition, Particle Swarm Optimization
-%% or \MSC[2008] code \sep code (2000 is the default)
-
-\end{keyword}
-\end{frontmatter}
-
+% \begin{keyword}
+% %% keywords here, in the form: keyword \sep keyword
+% 
+% %% MSC codes here, in the form: \MSC code \sep code
+% Sketch Recognition, Particle Swarm Optimization
+% %% or \MSC[2008] code \sep code (2000 is the default)
+% 
+% \end{keyword}
+%\end{frontmatter}
+\textit{\textbf{Keywords:} Sketch Recognition, Particle Swarm Optimization.}
 %\pagenumbering{arabic}
@@
 
-\bibliographystyle{elsarticle-num-names}
+%\bibliographystyle{natbib}
+\bibliographystyle{ieeetr}
 \bibliography{../../../neededfiles/Bibliographies/Mybibliography}

### 2010-12-10 trunk/Swarm/swarmModified.tex

@@
 %%\usepackage{icgst}
-\documentclass[12pt,times,a4paper]{article}%[10pt]{article}
+%\documentclass[12pt,times,a4paper]{article}%[10pt]{article}
 %% Use the option review to obtain double line spacing
-%% \documentclass[preprint,review,12pt]{elsarticle}
+ \documentclass[preprint,1p,times,review]{elsarticle}
 
@@
 \usepackage{graphicx}
+\usepackage{subfigure}
 \usepackage{sectsty}
-\sectionfont{\normalsize  }
-\subsectionfont{ \normalsize\it \textbf }
-\subsubsectionfont{\normalsize \it \textbf }
+%\sectionfont{\normalsize  }
+%\subsectionfont{ \normalsize\it \textbf }
+%\subsubsectionfont{\normalsize \it \textbf }
 %\paragraphfont{\large\it}
@@
 % \biboptions{}
-\setlength\hoffset{0in}
-% \setlength\topmargin{0in}
-% \setlength\headheight{0in}
-% \setlength\headsep{-0.5in}
-% \setlength\textheight{10.0in}
- \setlength\textwidth{7in}
- \setlength\oddsidemargin{-0.5in}
- \setlength\evensidemargin{-0.5in}
-% \setlength\parindent{0.25in}
-% \setlength\parskip{0.25in} 
-\linespread{1.5}
+% \setlength\hoffset{0in}
+% % \setlength\topmargin{0in}
+% % \setlength\headheight{0in}
+% % \setlength\headsep{-0.5in}
+% % \setlength\textheight{10.0in}
+%  \setlength\textwidth{7in}
+%  \setlength\oddsidemargin{-0.5in}
+%  \setlength\evensidemargin{-0.5in}
+% % \setlength\parindent{0.25in}
+% % \setlength\parskip{0.25in}
+% \linespread{1.5}
+\begin{abstract}
 
+%This study
+%%%This study
+
+Sketches and drawings are widely used as a simple method of expressing thought and designs.
+The goal of this paper is developing a sketch recognition system that facilitates using sketches in computer systems.
+The system uses Particle Swarm Algorithms to optimally segment the strokes the user draws into meaningful geometric primitives.
+These geometric primitives are grouped to formulate symbols which are further identified using a Support Vector Machines (SVM) classifier.
+A comparision is presented between using Traditional Particle Swarm Optimization and the Enhanced Particle Swarm Optimization (EPSO) which is based on the perturbation of the particle velocity.
+Two different segmentation algorithms were evaluated and results show that both algorithms improve the final recognition results.
+Experiments were conducted on three different dataset, simple presentation (Hs-DS), Electrical Design (EL-DS) and Logic Design (LD-DS) symbols.
+Results show that the effectiveness of the proposed system on various dataset especially on datasets with low shape diversity as in Logic Design dataset (LD-DS).
+Results also show that using an Enhancment on the Tradiontal Swarm algorithm improves recognition rate.
+
+\end{abstract}
 \begin{document}
@@
 % \begin{center}
-% Maha El Meseery\footnote{Corresponding Author }, Mahmoud Fakhr El Din, Tarek abouldahab\\
+% Maha El Meseery\footnote{Corresponding Author }, Mahmoud Fakhr El Din, Tarek
+%abouldahab\\
 % \textit{Electronic Research Institute,} Cairo, Egypt,\\
@@
 % 	melmeseery@eri.sci.eg, mafakhr@mcit.gov.eg}
-% 	 {Tarek abouldahab\sthanks{Corresponding Author}} {Cairo Metro Company \\
+% 	 {Tarek abouldahab\sthanks{Corresponding Author}} {Cairo Metro Company
+%\\
 %   Ministry of transport \\Cairo, Egypt\\ heshoaboda@hotmail.com}
@@
 \small{
-Maha El Meseery \footnote{Corresponding Author} \textsuperscript{a}, Mahmoud Fakhr El Din \textsuperscript{a}, Tarek abouldahab \textsuperscript{b}}\\
- \textsuperscript{a} \small{Signals Processing Group, Computers and Systems Department,}\\ \small{ Electronic Research Institute, Cairo, Egypt.}\\
+Maha El Meseery \footnote{Corresponding Author} \textsuperscript{a}, Mahmoud
+Fakhr El Din \textsuperscript{a}, Tarek abouldahab \textsuperscript{b}}\\
+ \textsuperscript{a} \small{Signals Processing Group, Computers and Systems
+Department,}\\ \small{ Electronic Research Institute, Cairo, Egypt.}\\
   \small {\textsuperscript{b}Cairo Metro Company ,
@@
 \maketitle
- \begin{abstract}
-
-%This study 
-Sketches and drawings are widely used as a simple method of expressing thought and designs.
-The goal of this paper is to improve our sketch recognition system by using an enhanced particle swarm optimization algorithm.
-The Enhanced Particle Swarm Optimization (EPSO) is based on the perturbation of the particle velocity is used to optimally segment the strokes the user draws into meaningful geometric primitives.
+%  \begin{abstract}
+%
+% %This study
+% %%%This study
+%
+% Sketches and drawings are widely used as a simple method of expressing thought and designs.
+The goal of this paper is developing a sketch recognition system that facilitates using sketches in computer systems.
+The system uses Particle Swarm Optimization (PSO) algorithm to optimally segment the strokes the user draws into meaningful geometric primitives.
 These geometric primitives are grouped to formulate symbols which are further identified using a Support Vector Machines (SVM) classifier.
-Experiments were conducted on the benchmark dataset of simple presentation (Hs-DS) symbols.
-Results show that the enhanced particle swarm optimization improves the final recognition accuracy.
-
-\end{abstract}
+Two different segmentation algorithms were evaluated and results show that both algorithms improve the final recognition results.
+Experiments were conducted on three different dataset, simple presentation (Hs-DS), Electrical Design (EL-DS) and Logic Design (LD-DS) symbols.
+Results show that the effectiveness of the proposed system on various dataset especially on datasets with low shape diversity as in Logic Design dataset (LD-DS).
+%
+%
+%  \textsl{Sketches and drawings are widely used as a simple method of expressing
+% thought and designs. The goal of this paper is to improve our sketch recognition
+% system by using an enhanced particle swarm optimization algorithm. The Enhanced
+% Particle Swarm Optimization (EPSO) is based on the perturbation of the particle
+% velocity is used to optimally segment the strokes the user draws into meaningful
+% geometric primitives. These geometric primitives are grouped to formulate
+% symbols which are further identified using a Support Vector Machines (SVM)
+% classifier. Experiments were conducted on the benchmark dataset of simple
+% presentation (Hs-DS) symbols. Results show that the enhanced particle swarm
+% optimization improves the final recognition accuracy.}
+% \end{abstract}
 
@@
 The field of symbol recognition has gained interest in the last few years.
-Scientists generally and engineers specifically express thoughts and designs using sketches.
-Engineers use sketches to exchange designs as a natural method of communication rather than writing or speaking.
-Design engineers need a powerful computer based system with symbol recognition to help them design, manipulate and store sketches more effectively than using only papers.
-Increasing the interaction between computers and users in sketch and CAD systems has been the reason for the emerging of a few advanced sketch recognition systems.
-Sketch recognition is the process of identifying user strokes into meaningful symbols that can be further used.
+Scientists generally and engineers specifically express thoughts and designs
+using sketches. Engineers use sketches to exchange designs as a natural method
+of communication rather than writing or speaking. Design engineers need a
+powerful computer based system with symbol recognition to help them design,
+manipulate and store sketches more effectively than using only papers.
+Increasing the interaction between computers and users in sketch and CAD systems
+has been the reason for the emerging of a few advanced sketch recognition
+systems. Sketch recognition is the process of identifying user strokes into
+meaningful symbols that can be further used.
 
-Sketch recognition is divided into three main steps preprocessing, segmentation and symbol recognition.
-The preprocessing phase captures user input stroke points and collects basic information about the stroke then proceeds to remove noise and compute basic statistical and geometrical information.
-In segmentation phase, the strokes are divided into a set of simple geometrical primitives or segments.
-In the third phase, sketch recognition, strokes and segments are clustered to formulate symbols that can be recognized by a classifier system.
+Sketch recognition is divided into three main steps preprocessing, segmentation
+and symbol recognition. The preprocessing phase captures user input stroke
+points and collects basic information about the stroke then proceeds to remove
+noise and compute basic statistical and geometrical information. In segmentation
+phase, the strokes are divided into a set of simple geometrical primitives or
+segments. In the third phase, sketch recognition, strokes and segments are
+clustered to formulate symbols that can be recognized by a classifier system.
 
-This paper modifies the sketch recognition system presented in \cite{mypaper}, to use enhanced particle swarm algorithm (EPSO) to optimally segment user's strokes.
-As in the Original system the users can draw symbols using any number of strokes in any order they like.
-The system segments each stroke using two different EPSO algorithms.
-Segmentation is based on dividing the digital curve into polygon\cite{PolygonApproximationPSO}.
-To handle complex strokes, our system uses a second segmentation EPSO algorithm to segment strokes into a set of lines and circular arcs.
-After segmentation, a symbol features vector composed of different statistical, geometrical and spatial features is deduced.
-The final classification step, SVM classifier uses the computed feature vector to classify the segments into one of the previously trained classes.
+In PR based methods, systems compute properties that describe symbols then use a
+classification method to recognize symbols. The properties used can either be
+based on gesture properties, global shape properties or image based properties.
+Gesture properties are set of features that represent how the user draws a
+symbol not how it looks like (e.g., identify a rectangle when a user draws a L
+corner). In gesture properties approach, systems identify symbols by using
+gesture properties of some of its underlying strokes. A system that uses
+gestures properties assumes that users draws symbols in specific order and style
+\cite{gestureexample12,aideddesgin22}. In global shape properties approach,
+systems investigate general shape properties and its underlying strokes. Using
+shape properties (e.g., ratio of the area of convex hull to the area of shape)
+relaxes the assumption and the role of each stroke in the symbol but it does not
+truly represent the appearance or fine details of the symbol
+\cite{DiagramOfflineConvexHull,Cali63}. Unlike shape properties based methods,
+image based methods disregard individual strokes and focus on the overall
+appearance of shapes. These system convert the shapes into image and extracted
+image based features (e.g. local gradient
+feature)\cite{Oltmans07,imagetrainable48}. In \cite{imagetrainable48}, they used
+four different distance measures to match user symbol with the prototype
+database. The prototype of a shape is represented as an image of the symbol that
+is normalized and scaled.  However, the problem is that the shape prototype must
+represent all transformation and variation the input shape can have.   %Image
+based methods are used in \cite{Oltmans07,imagetrainable48} where they computed
+a code book of 1000 features to recognize symbols.
 
-The remaining of the paper is as following; Section \ref{sec:ParticleSwarmAlgorithm} describes the Basic particle swarm algorithm in details.
-Enhanced Particle Swarm Algorithm is described in Section \ref{sec:EnhancedSwarmAlgorithm}.
-Section \ref{Sysdisc} provide details on preprocessing, segmentation and recognition process.
-Section \ref{sec:Experiments} describes the experiments that were preformed.
-We conclude in Section \ref{ConclusionandFutureWork}.
+Methods based on gesture and global shape properties produce systems that are
+dependent on assumptions about how strokes can be grouped. These systems assume
+that users will finish drawing a shape before moving to the next one
+\cite{Cali63,geometrydomain49}. This assumption allows the system to temporally
+group strokes and then recognizes them based on their strokes or global shape
+properties. Alvarado \cite{AlvaradoDigital} argued that this assumption does not
+hold in natural sketches.
+
+In this paper, we choose to implement a hybrid method that includes shape
+property, gesture and hierarchical information. Our method attempts to avoid
+many of the problems faced by approaches based on individual gestures or global
+shape properties while trying to gain the additional information that the stroke
+based system uses. Firstly, the system uses particle swarm optimization (PSO) to
+optimally segment user's strokes. Users can draw symbols using any number of
+strokes in any order they like. The system segments each stroke using two
+different PSO algorithms. The first segmentation algorithm is based on dividing
+the digital curve into polygons\cite{PolygonApproximationPSO}. Our algorithm
+enhances its performance by adding curvature and other local information during
+segmenting strokes to achieve better results. To handle complex strokes, our
+system uses a second segmentation PSO algorithm that segment strokes into a set
+of lines and circular arcs. The two segmentation algorithms can either by used
+individually or in parallel then chosing the segmentation that gives the minimum
+error. After segmentation, a feature vector composed of different statistical,
+geometrical and composite features is deduced. The final classification step
+uses a SVM classifier to classify symbols into one of the previously trained
+classes based on the features vector computed\cite{mypaper}.
+
+ In addition to the system mentioned above, an Enhnaced Swarm algorithm is compared to the Traditional Particle swarm algorithm\cite{stockPaper}.
+The Enhanced swarm algorithm is based on chaning the particle velocity by replacing it with the weighted differance between the particle and another random in the space.
+This random relacing is only used when the new velociy and location is better then the current particle values.
+The procedure introude new expolration factor the swarm system which helps in reaching the solution faster and more obtimially.
+
+% %This paper modifies the sketch recognition system presented in \cite{mypaper},
+% to use enhanced particle swarm algorithm (EPSO) to optimally segment user's
+% strokes. As in the Original system the users can draw symbols using any number
+% of strokes in any order they like. The system segments each stroke using two
+% different EPSO algorithms. Segmentation is based on dividing the digital curve
+% into polygon\cite{PolygonApproximationPSO}. To handle complex strokes, our
+% system uses a second segmentation EPSO algorithm to segment strokes into a set
+% of lines and circular arcs. After segmentation, a symbol features vector
+% composed of different statistical, geometrical and spatial features is deduced.
+% The final classification step, SVM classifier uses the computed feature vector
+% to classify the segments into one of the previously trained classes.
+
+The remaining of the paper is as following; Section
+\ref{sec:ParticleSwarmAlgorithm} describes the Basic particle swarm algorithm in
+details.  Enhanced Particle Swarm Algorithm is described in Section
+\ref{sec:EnhancedSwarmAlgorithm}. Section \ref{Sysdisc} provide details on
+preprocessing, segmentation and recognition process. Section
+\ref{sec:Experiments} describes the experiments that were preformed. We conclude
+in Section \ref{ConclusionandFutureWork}.
 
@@
 
- Particle Swarm Algorithm\cite{PSOFirst}is a population based stochastic search where each solution is simulated as a  particle that flies around in a multidimensional search space.
-A fitness function is used to evaluate the goodness of each particle in solution space.
-But unlike evolutionarily algorithm the new population is the PSO is influenced by the social cooperation of the particle rather than survival to the fittest.
-Each particle saves a history of the best location in the swarm and its own best location.
-At the end each iteration, the particles moves to a new location based on the global best location and it own history of best location.
-This use of previous particle locations gives PSO advantage over Evolutionary algorithms where no similar information is used.
+ Particle Swarm Algorithm\cite{PSOFirst}is a population based stochastic search
+where each solution is simulated as a  particle that flies around in a
+multidimensional search space. A fitness function is used to evaluate the
+goodness of each particle in solution space. But unlike evolutionarily algorithm
+the new population is the PSO is influenced by the social cooperation of the
+particle rather than survival to the fittest. Each particle saves a history of
+the best location in the swarm and its own best location.  At the end each
+iteration, the particles moves to a new location based on the global best
+location and it own history of best location. This use of previous particle
+locations gives PSO advantage over Evolutionary algorithms where no similar
+information is used.
 
- The \textit{Basic Particle Swarm Algorithm (PSO)} represents each solution with a $N$ dimension particle from the solution space \cite{PSOFirst}.
-Each particle moves with a direction and velocity $v_{ij}$ based on equations \ref{eq:Swarm1} \& \ref{eq:Swarm}.
+ The \textit{Basic Particle Swarm Algorithm (PSO)} represents each solution with
+a $N$ dimension particle from the solution space \cite{PSOFirst}. Each particle
+moves with a direction and velocity $v_{ij}$ based on equations \ref{eq:Swarm1}
+\& \ref{eq:Swarm}.
 
@@
 
-where $p_{ij}$ represent the $j$th dimension in the $i$th particle and $v_{ij}$ is the velocity of the $j$th dimension in the $i$th particle.
+where $p_{ij}$ represent the $j$th dimension in the $i$th particle and $v_{ij}$
+is the velocity of the $j$th dimension in the $i$th particle.
 
  \begin{equation}
-v_{ij}  = v_{ij} \omega + c_1 r_1 (lbest_{ij}  - p_{ij} ) + c_2 r_2 (gbest_{ij}  - p_{ij} ),
+v_{ij}  = v_{ij} \omega + c_1 r_1 (lbest_{ij}  - p_{ij} ) + c_2 r_2 (gbest_{ij}
+- p_{ij} ),
 \label{eq:Swarm}
 \end{equation}
- where $\omega$ is the inertia weight parameter which controls the tradeoff between exploration and exploitation,  $lbest_{ij}$ is the local best particle, $gbest_{ij}$ is the global best particle, $r_1$ \& $r_2$ are random variables and $c_1$ \& $c_2$ are the swarm acceleration parameters.
+ where $\omega$ is the inertia weight parameter which controls the tradeoff
+between exploration and exploitation,  $lbest_{ij}$ is the local best particle,
+$gbest_{ij}$ is the global best particle, $r_1$ \& $r_2$ are random variables
+and $c_1$ \& $c_2$ are the swarm acceleration parameters.
 
- After each iteration the global best $g_{best}$ particle and the agent local best particle $l_{best}$are evaluated based on the maximum fitness functions of all particles in the solution space.
-The solution is found after achieving a specific number of iteration or after an error threshold is achieved.
+ After each iteration the global best $g_{best}$ particle and the agent local
+best particle $l_{best}$are evaluated based on the maximum fitness functions of
+all particles in the solution space. The solution is found after achieving a
+specific number of iteration or after an error threshold is achieved.
 
@@
 \label{sec:EnhancedSwarmAlgorithm}
-In the above basic PSO, the velocity update is based on three terms an inertia term, local cognitive ($l_{best}$) term and global cognitive ($g_{best}$)term.
-The particle local cognitive term is associated with the particle own history without consideration of other swarm particles locations and histories which leads to a lake of diversity and sub optimum solutions.
-Therefore, we are trying to add more social influence in the particle movement by using the weighted difference of the position vectors of any other two distinct particles randomly chosen from the swarm to modify the local cognitive term.
-The modified term adds social behavior of other particle and keeps diversity while searching for the global optimum.
-The new social cognitive term of the particle is shifted to the new location only if it yields a fitness value better than obtained using the standard PSO\cite{stockPaper}.
+In the above basic PSO, the velocity update is based on three terms an inertia
+term, local cognitive ($l_{best}$) term and global cognitive ($g_{best}$)term.
+The particle local cognitive term is associated with the particle own history
+without consideration of other swarm particles locations and histories which
+leads to a lake of diversity and sub optimum solutions. Therefore, we are trying
+to add more social influence in the particle movement by using the weighted
+difference of the position vectors of any other two distinct particles randomly
+chosen from the swarm to modify the local cognitive term. The modified term adds
+social behavior of other particle and keeps diversity while searching for the
+global optimum. The new social cognitive term of the particle is shifted to the
+new location only if it yields a fitness value better than obtained using the
+standard PSO\cite{stockPaper}.
 
@@
   \begin{enumerate}
-\item For each particle $p_{i}$ in the swarm two other distinct particles, $p_L$ and $p_M$ are selected randomly where $i\neq L\neq M$.
-The difference between their positional coordinates $p_{d}$ is taken as a difference vector in equation \ref{eqmod}.
+\item For each particle $p_{i}$ in the swarm two other distinct particles, $p_L$
+and $p_M$ are selected randomly where $i\neq L\neq M$. The difference between
+their positional coordinates $p_{d}$ is taken as a difference vector in equation
+\ref{eqmod}.
  \begin{equation}
@@
  \end{equation}
-where $p_{dj}$ represent the $j$th dimension in the difference vector and $p_{Lj}$ represent the $j$th dimension in the $L$th particle.
+where $p_{dj}$ represent the $j$th dimension in the difference vector and
+$p_{Lj}$ represent the $j$th dimension in the $L$th particle.
 
-Then the associated velocity vector of particle $p_{i}$ is updated according to this difference vector as in equation \ref{eq:mod1}.
-In essence the local cognitive term of the velocity update formula in \ref{eq:Swarm} is replaced with the vector differential operator $p_d$ to produce some additional exploration capability.
+Then the associated velocity vector of particle $p_{i}$ is updated according to
+this difference vector as in equation \ref{eq:mod1}. In essence the local
+cognitive term of the velocity update formula in \ref{eq:Swarm} is replaced with
+the vector differential operator $p_d$ to produce some additional exploration
+capability.
  \begin{equation}
- v^{T}_{ij}  = v_{ij} \omega + c_1 r_1 (p_{dj}) + c_2 r_2 (gbest_{ij}  - p_{ij} ),
+ v^{T}_{ij}  = v_{ij} \omega + c_1 r_1 (p_{dj}) + c_2 r_2 (gbest_{ij}  - p_{ij}
+),
  \label{eq:mod1}
  \end{equation}
-\item The new trial position $p^{T}_{ij}$ is created by adding the updated velocity $v^{T}_{ij}$ in Eq. \ref{eq:mod1}.
+\item The new trial position $p^{T}_{ij}$ is created by adding the updated
+velocity $v^{T}_{ij}$ in Eq. \ref{eq:mod1}.
   \begin{equation}
@@
  \end{equation}
-\item The new position $p^{T}_{ij}$ is used to generate the new trial objective fitness function value $fitness(p^{T})$.
-The fitness value is calculated based on this new trial position vector and the particle is shifted to this trial position if the trial objective function is better than the original objective function associated with standard PSO.
-Thus the target particle is relocated using the following equation (Eq.
-\ref{eq:objective})
+\item The new position $p^{T}_{ij}$ is used to generate the new trial objective
+fitness function value $fitness(p^{T})$. The fitness value is calculated based
+on this new trial position vector and the particle is shifted to this trial
+position if the trial objective function is better than the original objective
+function associated with standard PSO. Thus the target particle is relocated
+using the following equation (Eq. \ref{eq:objective})
   \begin{equation}
@@
  \end{equation}
- where $fitness(p^{T})$ is the objective fitness value of the new trail position $ p^{T}$ and  $fitness(p^{old})$is the objective fitness value of the new basic position $ p{i}$ computed by equation \ref{eq:Swarm1}.
+ where $fitness(p^{T})$ is the objective fitness value of the new trail position
+$ p^{T}$ and  $fitness(p^{old})$is the objective fitness value of the new basic
+position $ p{i}$ computed by equation \ref{eq:Swarm1}.
    \end{enumerate}
@@
 %\label{sec:application}
-The modified swarm algorithm is used to improve the performance of the sketch recognition system in \cite{mypaper}.
+%The modified swarm algorithm is used to improve the performance of the sketch recognition system in \cite{mypaper}.
+
 The sketch recognitions system is divided into three main steps 1) preprocessing, 2) segmentation and 3) Recognition.
-Figure \ref{fig:flowChart} shows the flow chart of the system; the first process is collecting stroke points then computing the curvature and possible dominant points.
-The following section provides a detailed description of each step.
-The main block of the system remains the same but there are some modifications in the segmentation algorithm to improve its efficiency.
-The main modification is using Enhanced Particle swarm algorithm instead of Basic Particle swarm algorithms.
-Other modifications are explained in the details in the next sections.
+Figure \ref{fig:flowChart} shows the block diagram of the system; the first
+process is collecting stroke points then computing the curvature and possible
+dominant points.  The following section provides a detailed description of each
+step. The main block of the system remains the same but there are some
+modifications in the segmentation algorithm to improve its efficiency.
+This study compares between using Traditional PSO and an Enhanced Particle swarm algorithm.
+Other modifications are explained in the details in
+the next sections.
+%  \begin{figure}
+% 	\centering
+% 		\includegraphics[width=3.5in,height=6in]{FlowChart.jpg}
+% 	\caption{System flow chart} The chart shows the flow of process and data processed thought the system.
+% 	\label{fig:flowChart}
+% \end{figure}
+
  \begin{figure}
 	\centering
-		\includegraphics[width=3.5in,height=6in]{FlowChart.jpg}
-	\caption{System flow chart} The chart shows the flow of process and data processed thought the system. 
+		\includegraphics[scale=0.5]{images/Blockdiagram.png}
+	\caption{System Block Diagram} The chart shows the system block diagram.
 	\label{fig:flowChart}
@@
 \label{Prepross}
- Speed and time difference information were widely used in sketch understanding systems \cite{earlyprocess}.
+
+ Speed and time difference information were widely used in sketch recognition
+systems \cite{earlyprocess}. Figure \ref{fig:orignalStroke} shows an example of
+an input stroke drawn by users.  Figure \ref{fig:speed2Distance} shows the graph
+of time difference and the speed for the stroke in Figure
+\ref{fig:orignalStroke}. It was noted that time difference provides more
+distinct maxima than the speed information. Agar et al. \cite{polygonfeedback31}
+mentions that the pointing device (e.g., mouse) sampling rate is the reason for
+this phenomenon. The points are sampled at regular time intervals as long as the
+pointing device is moving. Therefore, there are no new detected points when the
+pointing device is stationary. This leads to a nearly constant time difference
+between samples while the pen is moving and large difference while the pen is
+stationary (Figure \ref{fig:timediff}). Contrary to time difference information,
+the speed information has a lot of noise in the data because usually  users draw
+with variant speeds (Figure \ref{fig:speed2Distance}).  %Similarly, direction
+%information is used as it provides better distinctive maxima for the corners
+%than the curvature information \cite{meanshift10}.% Figure \ref{fig:curvatures}
+%shows the direction and curvature graphs for the stroke drawn in Figure
+\ref{fig:orignalStroke}
+
+ \begin{figure}[h]
+	\centering
+		\includegraphics[scale=0.5]{images/stroke3.jpg}
+	\caption{Input Stroke} Example of an input stroke to the system.
+	\label{fig:orignalStroke}
+\end{figure}
+
+ \begin{figure}
+	\centering
+			\subfigure[Speed versus Displacement Graph]{
+\includegraphics[scale=0.4]{images/vel3.jpg}}
+			\hfill
+			\subfigure[ Time Difference  versus Displacement
+Graph]{\label{fig:timediff}\includegraphics[scale=0.4]{images/td3.jpg}}
+	\caption{Speed and Time Difference Graphs}  a) speed graph has lot of
+noise in the data with respect to b) time difference.   The black points denote
+the location of the possible dominant points $P_{pd}$. Black rectangles
+represent regions $R_i$ that in a) Are lower than the threshold. b) Are above
+the threshold.
+	\label{fig:speed2Distance}
+\end{figure}
+
+Dominant points are characterized by low speed values, high curvature and
+direction values. However, the dominant points from the curvature information
+may not be the same points as the dominant points extracted from the speed
+information. Hence, we compute time difference, direction, speed and curvature
+of each point along a stroke to ensure that no dominant point is overlooked in
+this phase. The speed is calculated as $v=\Delta s/\Delta t$ where $t$ is the
+time difference between two points and $s$ is the length between them. The
+direction $d$ calculated as the angle between vectors $\overrightarrow {P_{i -
+1} P_i }$ and the $x-axis$ and the curvature is considered as the change in
+direction $d$ with respect to length $s$ i.e. $c= \Delta d/\Delta s$.
+
+After the system computes the speed, time difference and curvature information,
+it proceeds to extracts the points that have low velocity and high curvature.
+However, using simple differentiation to detect local extreme points results in
+false points due to the non smooth curves. Hence, our system adopts a process
+presented by \cite{earlyprocess}, where the mean of each of the four calculated
+information is calculated. This mean is taken as the threshold \textit{th} which
+is used to separate the curve into regions ($R_i$) (Figure\ref{fig:timediff});
+each region $R_i$ is defined as a range of points, where the curve values are
+either above or below the threshold \textit{th}. Those regions are further
+processed to find the maximum point $Max(R_i)$ of each region $R_i$. The stroke
+points $p_i(x,y)$ that correspond to those maximum values are labeled as
+\textit{possible dominant points} $P_{pd}$. The system repeats this process for
+curvature, time difference and speed information. All the points labeled as
+possible dominant points $P_{pd}$ are saved into a single array. Figure
+\ref{fig:ppd999} shows the particles labeled as Possible dominate point $P_{pd}$
+in the preprocessing phase. It is clear that some of the $P_{pd}$ points are
+redundant as specified in the preprocessing stage. % (as shown are redundant)
+\begin{figure}
+	\centering
+\includegraphics[scale=0.7]{images/ppd.jpg}
+	\caption{Possible Dominante Points} An illustration of stroke points and
+the expected \textit{possible dominant points} $P_{pd}$.
+	%\label{fig:ppd}
+	\label{fig:ppd999}
+\end{figure}
+
+ %Speed and time difference information were widely used in sketch understanding systems \cite{earlyprocess}.
 It was noted that time difference provides more distinct maxima than the speed information.
@@
 the mouse) sampling rate is the reason for this phenomenon.
- In the presented system, we computed time difference, direction, speed and curvature of each point along a stroke.
-The speed is calculated as $v=\Delta s/\Delta t$ where $t$ is the time difference between two points and $s$ is the length between them.
-The direction $d$ is calculated as the angle between two vectors and curvature is considered as the change in direction with respect to length i.e.
-$c= \Delta d/\Delta s$.
-Dominant points are characterized by lower speed values and higher curvature and direction values.
+% % In the presented system, we computed time difference, direction, speed and
+% curvature of each point along a stroke. The speed is calculated as $v=\Delta
+% s/\Delta t$ where $t$ is the time difference between two points and $s$ is the
+% length between them. The direction $d$ is calculated as the angle between two
+% vectors and curvature is considered as the change in direction with respect to
+% length i.e. $c= \Delta d/\Delta s$. Dominant points are characterized by lower
+% speed values and higher curvature and direction values.
 
-After the system computes all the speed, time difference and curvature information it proceeds to detect the points with low velocity and high curvatures.
-Using simple differentiation to detect local extreme points resulted in false points due to the non smooth curves.
-Hence, the system adopted a process presented by \cite{earlyprocess}, where the mean of the curve is calculated.
-Then a threshold \textit{th} is used to separate the curve into regions $Region_i$; each region $Region_i$ is defined as a range of points, where the curve values are either above or below the threshold \textit{th}.
-Those regions are further processed to find the maximum point $Max(Region_i)$ of each region $Region_i$.
-The stroke points $p_i(x,y)$ that correspond to those maximum values are labeled as \textit{possible dominant points} $P_{pd}$.
-The system repeats this process for curvature, time difference and speed information.
-All the points labeled as possible dominant points $P_{pd}$ are saved into a single array.
-Figure \ref{fig:LabelsPPD} shows the particles labeled as Possible dominant point $P_{pd}$ by the preprocessing, it is noticed the redundancy of some $P_{pd}$ points.
-% (as shown are redundant)
+% After the system computes all the speed, time difference and curvature
+% information it proceeds to detect the points with low velocity and high
+% curvatures. Using simple differentiation to detect local extreme points resulted
+% in false points due to the non smooth curves. Hence, the system adopted a
+% process presented by \cite{earlyprocess}, where the mean of the curve is
+% calculated. Then a threshold \textit{th} is used to separate the curve into
+% regions $Region_i$; each region $Region_i$ is defined as a range of points,
+% where the curve values are either above or below the threshold \textit{th}.
+% Those regions are further processed to find the maximum point $Max(Region_i)$ of
+% each region $Region_i$. The stroke points $p_i(x,y)$ that correspond to those
+% maximum values are labeled as \textit{possible dominant points} $P_{pd}$. The
+% system repeats this process for curvature, time difference and speed
+% information. All the points labeled as possible dominant points $P_{pd}$ are
+% saved into a single array. Figure \ref{fig:LabelsPPD} shows the particles
+% labeled as Possible dominant point $P_{pd}$ by the preprocessing, it is noticed
+% the redundancy of some $P_{pd}$ points. % (as shown are redundant)
+
 \subsection{Segmentation}
@@
 In the segmentation stage the input stroke is divided into a set of primitives.
-As shown in Figure \ref{fig:segblock} first an attempt is made to fit the stroke points into a curve or an ellipse using a minimum square error fitting algorithm \cite{ellipsefit}.
-The ellipse fitting step helps to prevent the system from over segmenting the stroke into multiple lines or curves if the input stroke is an ellipse.
-If the stroke proved to be an ellipse arc then the segmentation process ends and the system proceeds to the next step.
-Otherwise, the stroke is passed to two further segmentation algorithms that divide the stroke to either lines or lines and curves.
-After two different segmentations are generated, the system chooses the segmentation with the minimum error.
-The following sections describe in details the ellipse detection algorithm and the two segmentation algorithms used to divide the stroke.
+As shown in Figure \ref{fig:segblock} first an attempt is made to fit the stroke
+points into a curve or an ellipse using a minimum square error fitting algorithm
+\cite{ellipsefit}. The ellipse fitting step helps to prevent the system from
+over segmenting the stroke into multiple lines or curves if the input stroke is
+an ellipse. If the stroke proved to be an ellipse arc then the segmentation
+process ends and the system proceeds to the next step. Otherwise, the stroke is
+passed to two further segmentation algorithms that divide the stroke to either
+lines or lines and curves. After two different segmentations are generated, the
+system chooses the segmentation with the minimum error. The following sections
+describe in details the ellipse detection algorithm and the two segmentation
+algorithms used to divide the stroke.
+
  \begin{figure}
@@
 
-This process tries to fit the stroke points into an ellipse arc; it starts with computing the center of the stroke bounding box.
-The bounding box center point is used as the first estimation of the center of the ellipse.
-The axes of the ellipse are estimated as the $width/2$ and $height/2$ of the stroke bounding box where $width$ is width and $height$ is height of the bonding box.
-The least square fitting algorithm \cite{chernov} is used to minimize the fitting error of the ellipse Equation (\ref{eq:circleFit})
+This process tries to fit the stroke points into an ellipse arc; it starts with
+computing the center of the stroke bounding box. The bounding box center point
+is used as the first estimation of the center of the ellipse. The axes of the
+ellipse are estimated as the $width/2$ and $height/2$ of the stroke bounding box
+where $width$ is width and $height$ is height of the bonding box. The least
+square fitting algorithm \cite{chernov} is used to minimize the fitting error of
+the ellipse Equation (\ref{eq:circleFit})
 
 \begin{equation}
-E = \sum\limits_{i = 0}^N {\frac{{(x_i - x_0 )}}{{a^2 }}^2  + \frac{{(y_i - y_0 )}}{{b^2 }}^2  - 1} 
+E = \sum\limits_{i = 0}^N {\frac{{(x_i - x_0 )}}{{a^2 }}^2  + \frac{{(y_i - y_0
+)}}{{b^2 }}^2  - 1}
 \label{eq:circleFit}
@@
 
- where $N$ is number of points in the stroke, $a,b$ are the length of ellipse axes, $x_0$ \& $y_0$ are the coordinates of the center point, $x_i$ \& $y_i$ are the coordinates of point $i$ in the stroke.
-A list of new values for $x_0$ , $y_0$ ,$a$ and $b$ are generated randomly from the older values with small increments after each loop.
-After few iteration, the final fit error of the estimated ellipse is reported.
-A modification to the system in \cite{mypaper} is trying to add another measure to compute the efficiency of the final estimated ellipse.
-Equation \ref{eq:circleError} ensures that the drawn percentage of ellipse is considered.
-This eliminates fitting a line into very large ellipse but leaves small ellipses to be fitted as a partial or full ellipse.
+ where $N$ is number of points in the stroke, $a,b$ are the length of ellipse
+axes, $x_0$ \& $y_0$ are the coordinates of the center point, $x_i$ \& $y_i$ are
+the coordinates of point $i$ in the stroke. A list of new values for $x_0$ ,
+$y_0$ ,$a$ and $b$ are generated randomly from the older values with small
+increments after each loop.  After few iteration, the final fit error of the
+estimated ellipse is reported. A modification to the system in \cite{mypaper} is
+trying to add another measure to compute the efficiency of the final estimated
+ellipse. Equation \ref{eq:circleError} ensures that the drawn percentage of
+ellipse is considered. This eliminates fitting a line into very large ellipse
+but leaves small ellipses to be fitted as a partial or full ellipse.
 
@@
 \end{equation}
-where $E$ is the error computed by Equation(\ref{eq:circleFit}), $L_{stroke}$ total length of stroke and $P_{ellipse} $ is the perimeter of the estimated ellipse.
-If $eff$ is more than threshold $th_{Ellipse}$\footnote{By trial and error best threshold found was $th_{Ellipse}=0.2$} then the stroke is segmented as an ellipse otherwise the system proceeds to the next step.
+where $E$ is the error computed by Equation(\ref{eq:circleFit}), $L_{stroke}$
+total length of stroke and $P_{ellipse} $ is the perimeter of the estimated
+ellipse. If $eff$ is more than threshold $th_{Ellipse}$\footnote{By trial and
+error best threshold found was $th_{Ellipse}=0.2$} then the stroke is segmented
+as an ellipse otherwise the system proceeds to the next step.
 
@@
 \end{figure}
-Two DPSO algorithms are used to generate different segmentations for each stroke the user draws.
-For each segmentation generated, an error is evaluated.
-The segmentation with the minimum error value is chosen as the best stroke segmentation (Figure \ref{fig:segblock}).
-The problem definition is the same in both algorithms but they differ in the method they use to compute fitness and error functions.
+
+Two DPSO algorithms are used to generate different segmentations for each stroke
+the user draws. For each segmentation generated, an error is evaluated. The
+segmentation with the minimum error value is chosen as the best stroke
+segmentation (Figure \ref{fig:segblock}). The problem definition is the same in
+both algorithms but they differ in the method they use to compute fitness and
+error functions.
 \begin{description}
-	\item[ Problem definition:] The input stroke $S$  with $N$ points can be represented by set $S = \left\{ {x_1 ,x_2  \ldots x_N }\right\}$ where $x_i$ is the location of the point $i$.
-The swarm algorithms consist of $M$ agents which are represented by the set $A = \left\{ {P_i \left| {i = 1,2 \cdots M} \right.} \right\}$ where $P_i$ is a single solution particle from the solution space.
-Each particle decodes the problem using a binary array with the same length $N$ as the input stroke.
+	\item[ Problem definition:] The input stroke $S$  with $N$ points can be
+represented by set $S = \left\{ {x_1 ,x_2  \ldots x_N }\right\}$ where $x_i$ is
+the location of the point $i$. The swarm algorithms consist of $M$ agents which
+are represented by the set $A = \left\{ {P_i \left| {i = 1,2 \cdots M} \right.}
+\right\}$ where $P_i$ is a single solution particle from the solution space.
+Each particle decodes the problem using a binary array with the same length $N$
+as the input stroke.
 
-Therefore, the system represents each particle $P_i$ by $P_i = \left\{ {p_{ij} \left| {j = 1,2 \cdots N} \right.} \right\}$ where $p_{ij}$ has only two values a)1 ($p_{ij}=1$); means that this point ($j$) is a dominate point, or b) 0($p_{ij}=0$) means this point ($j$) is not a dominate point.
+Therefore, the system represents each particle $P_i$ by $P_i = \left\{ {p_{ij}
+\left| {j = 1,2 \cdots N} \right.} \right\}$ where $p_{ij}$ has only two values
+a)1 ($p_{ij}=1$); means that this point ($j$) is a dominate point, or b)
+0($p_{ij}=0$) means this point ($j$) is not a dominate point.
 Figure\ref{fig:pso1} shows a particle encoding in the DPSO system.
 
-	\item[Fitness function:] The fitness function and error calculation are different in each of the two \textit{DPSO} algorithms. 
+	\item[Fitness function:] The fitness function and error calculation are
+different in each of the two \textit{DPSO} algorithms.
 	\end{description}
@@
 %\\
-The approximation error is computed by the equation \ref{eq:ErrorSwarm1}. The graphical meaning of the error is shown in Figure\ref{fig:LabelsPSO}.
+The approximation error is computed by the equation \ref{eq:ErrorSwarm1}. The
+graphical meaning of the error is shown in Figure\ref{fig:pso1}.%\ref{fig:LabelsPSO}.
 \begin{equation}
@@
 \label{eq:ErrorSwarm1}
-\end{equation} where the arc $\widehat{x_ix_j}$ is defined as the consecutive set of points from point $x_i$ to point $x_{j}$ as in $x_i,x_{i+1} \cdots,x_j$.
-The line $\overline{x_i x_j} $ as the straight line connecting point $x_i$ to point $x_j$, $M$ is the number of dominant points in this solution as generated by the swarm algorithm.
-The error $e ( \widehat{x_ix_j},\overline{x_i x_j})$ is computed as the sum of squared perpendicular distance from every point along the arc $\widehat{x_ix_j}$ to the line $\overline{x_i x_j}$.
-The fitness is computed using equation %\ref{eq:fitnessSwarm1}
+\end{equation} where the arc $\widehat{x_ix_j}$ is defined as the consecutive
+set of points from point $x_i$ to point $x_{j}$ as in $x_i,x_{i+1} \cdots,x_j$.
+The line $\overline{x_i x_j} $ as the straight line connecting point $x_i$ to
+point $x_j$, $M$ is the number of dominant points in this solution as generated
+by the swarm algorithm. The error $e ( \widehat{x_ix_j},\overline{x_i x_j})$ is
+computed as the sum of squared perpendicular distance from every point along the
+arc $\widehat{x_ix_j}$ to the line $\overline{x_i x_j}$. The fitness is computed
+using equation %\ref{eq:fitnessSwarm1}
 \begin{equation}
@@
 \label{eq:fitnessSwarm1}
-\end{equation} where $N$ is the number of points in the stroke, $D$ is the number of points in the solution that was previously labeled as a possible dominant point ($P_{pd}$), $E$ is the computed error and $\varepsilon$ is the error threshold.
-It should be noticed that when the error is larger than the threshold $\varepsilon$ the fitness is given a -ve value to lower the fitness value of the solution.
-Otherwise the system favors the lower number of vertices.
+\end{equation} where $N$ is the number of points in the stroke, $D$ is the
+number of points in the solution that was previously labeled as a possible
+dominant point ($P_{pd}$), $E$ is the computed error and $\varepsilon$ is the
+error threshold. It should be noticed that when the error is larger than the
+threshold $\varepsilon$ the fitness is given a -ve value to lower the fitness
+value of the solution. Otherwise the system favors the lower number of vertices.
 
  \subsubsection{\textit{Hybrid Fitting \textsl{AlgS2}}}
-The algorithm has the same problem formulation but different fitness and error functions are used.
-An attempt is made to fit each segment into a line or circular arc.
-The errors of both circle and line fit estimations are computed for each segment $S_i=\widehat{x_ix_j}$.
-The approximation with the lower error value is chosen as the final approximation of this segment $S_i$\cite{CruveDivisionSwarm}.
-The sum of the approximation errors of all segments is the total error of the particle.
-The total error of the particle is computed by equation %\ref{eq:errorSwarm2}
+The algorithm has the same problem formulation but different fitness and error
+functions are used. An attempt is made to fit each segment into a line or
+circular arc. The errors of both circle and line fit estimations are computed
+for each segment $S_i=\widehat{x_ix_j}$. The approximation with the lower error
+value is chosen as the final approximation of this segment
+$S_i$\cite{CruveDivisionSwarm}. The sum of the approximation errors of all
+segments is the total error of the particle.  The total error of the particle is
+computed by equation %\ref{eq:errorSwarm2}
  \begin{equation}
@@
 \label{eq:errorSwarm2}
-\end{equation}where $M$ is the number of segments in the solution as generated by the swarm algorithm, $D_i$ is the minimum approximation error of curve and line approximations $min(d_c,d_l)$ where $d_c$ is the circle approximation error and $d_l$ is the line approximation error as computed by \cite{CruveDivisionSwarm}.
-The fitness is computed by the equation %\ref{eq:fitnessSwarm2}
+\end{equation}where $M$ is the number of segments in the solution as generated
+by the swarm algorithm, $D_i$ is the minimum approximation error of curve and
+line approximations $min(d_c,d_l)$ where $d_c$ is the circle approximation error
+and $d_l$ is the line approximation error as computed by
+\cite{CruveDivisionSwarm}.  The fitness is computed by the equation
+%\ref{eq:fitnessSwarm2}
 \begin{equation}
@@
 \label{eq:fitnessSwarm2}
-\end{equation} where $E$ is the error and $M$ is the number of segments and $k$ is a parameter tweaked to get minimum number of segments.
-The larger $k$ is, the more effect the number of segments will have.
-For our system, $k$ is selected to be 0.5\cite{CruveDivisionSwarm}.
+\end{equation} where $E$ is the error and $M$ is the number of segments and $k$
+is a parameter tweaked to get minimum number of segments. The larger $k$ is, the
+more effect the number of segments will have. For our system, $k$ is selected to
+be 0.5\cite{CruveDivisionSwarm}.
 
-As a modification to the system in \cite{mypaper}. After each iteration, of the swarm algorithm (\textsl{AlgS1} and \textsl{AlgS2}), each particle is refined using the following procedures: 
+As a modification to the system in \cite{mypaper}. After each iteration, of the
+swarm algorithm (\textsl{AlgS1} and \textsl{AlgS2}), each particle is refined
+using the following procedures:
 \begin{enumerate}
-	\item For each particle $P_i$ each dominant point $P_{ij}$ is checked to find if it was labeled before as a \textit{possible dominant point} $P_{pd}$ (computed as in section \ref{Prepross}).
-If it was not labeled the point $P_{ij}$ is moved to the nearest labeled point.
-This ensures that all of the points generated by the DPSO are possible dominant points $P_{pd}$.
-	\item The particles are tested to make sure that the distance between every two successive dominate point is larger than $min_D$, where $min_D$ is 5\% of the total length of the stroke.
-Otherwise, one of these points are removed based on the error caused by its removal.
-	\item Each two successive segments are test to determine if they can be merged into a single segment.
-If merging the two segments have the same or smaller the error, the two segments are merged.
-This step ensures that the final segmentation of the system has minimum number of segments.
+	\item For each particle $P_i$ each dominant point $P_{ij}$ is checked to
+find if it was labeled before as a \textit{possible dominant point} $P_{pd}$
+(computed as in section \ref{Prepross}). If it was not labeled the point
+$P_{ij}$ is moved to the nearest labeled point. This ensures that all of the
+points generated by the DPSO are possible dominant points $P_{pd}$.
+	\item The particles are tested to make sure that the distance between
+every two successive dominate point is larger than $min_D$, where $min_D$ is 5\%
+of the total length of the stroke.  Otherwise, one of these points are removed
+based on the error caused by its removal.
+	\item Each two successive segments are test to determine if they can be
+merged into a single segment. If merging the two segments have the same or
+smaller the error, the two segments are merged. This step ensures that the final
+segmentation of the system has minimum number of segments.
 \end{enumerate}
@@
 \label{sec:Recognition}
-After the user draws all strokes of the symbol, the set of un-recognized strokes is grouped together along with their segmentation as input to the feature extraction process.
+After the user draws all strokes of the symbol, the set of un-recognized strokes
+is grouped together along with their segmentation as input to the feature
+extraction process. A composite set of features is used to generate a single
+feature vector. The following list describes each feature in details.
+%\begin{description}
+\subsubsection{Structural and geometrical Features (FS1)}
+  They are features that define the structure of the geometrical symbol such as;
+ \begin{itemize}
+	 \item \emph{Segments:} The number of segments in the symbol.
+	 \item \emph{Strokes:} The number of strokes or partial strokes that
+created the symbols.
+		\item  \emph{Primitives:} The number of primitives in the
+symbol. This feature helps when identifying  symbols with mixed geometric
+primitives such as cylinders and callouts.
+		\item \emph{Curves:} The number of curves or ellipses in the
+symbol.
+		\item \emph{Lines:} The number of lines in the symbol.
+		\item \emph{Perpendicular}\emph{lines:} The number of
+perpendicular lines.
+		\item \emph{Parallel}\emph{lines:} The number of parallel lines.
+		\item \emph{Intersections:} The number of intersection between
+lines and curves.
+		\item \emph{T intersections:} The number of T intersections.
+		\item \emph{L intersections:} The number of L intersections.
+		\item \emph{X intersections:} The number of X intersections.
+
+\end{itemize}
+\subsubsection{Rubine Features (FS2)}
+  These features were introduced by Rubine \cite{gestureexample12} for single
+stroke gestures. Such as length of stroke, cosine of starting angle, smoothness
+of symbol and some features extracted from the bounding box of the symbol.
+
+\subsubsection{Statistical Features (FS3)}
+These features use moments and statistical properties such as;
+\begin{itemize}
+	\item \emph{Zernike moments:} The zernike moments are orthogonal moments
+which define the distribution of point in the input space. They are invariant to
+both reflection and rotation of the input shape \cite{ZerMomentOrthogonal}. %The
+zernike moment were used in \cite{HeloiseBeautification} t.
+\end{itemize}
+\subsubsection{Composite Features (FS4)}
+ These features are composites of different geometrical and statistical
+properties of the symbol.
+	\begin{itemize}
+\item \emph{Size Ratio:} The ratio between width and height of the symbol.
+	\item \emph{Ink density:} The density of points inside its bounding
+box\cite{GeometryAndDomain102}.
+ 	\item \emph{Convex Hull Area:} The area of convex hull with respect to
+area of bounding box of symbol.
+	\item \emph{Convex Hull Perimeter:} The perimeter of convex hull with
+respect to total length of symbol.
+		\item \emph{Mean Centroidal radius:} The Mean of the centroidal
+radius is the distance from each point in the symbol to the center of gravity.
+	\item \emph{Mean Time difference:} The mean of the time difference
+between each two successive points in the symbol.
+  \end{itemize}
+
+ After computing the features, the symbol is introduced to the classifier as a
+feature vector. The system uses Support Vector Machine (SVM) classifier with
+Linear kernel \cite{libsvm}. An OVO classifier structure (object versus object)
+is used to handle multi-class problem.% A validation algorithm is used to choose
+the Kernel variables parameters $c, \gamma$.
+
+%After the user draws all strokes of the symbol, the set of un-recognized
+% strokes is grouped together along with their segmentation as input to the feature extraction process.
 A composite set of features is used to generate a single feature vector.
@@
 After computing the features the symbol is introduced to the (SVM) classifier.
-
 \cite{mypaper}
@@
 \label{sec:Experiments}
-In our experiment we used the same benchmark dataset used in \cite{mypaper}, the dataset was collected by Hse and Newton\cite{HeloiseBeautification}.
+%The paragraph for the single dataset
+% In our Experiments the dataset (Figure
+% \ref{fig:HsSet}) was collected by Hse and Newton\cite{HeloiseBeautification}
+% (Hs-DS). The data are drawn by 16 users each of them was asked to draw 13 shapes
+% from 30 to 50 times.The dataset was divided into training and test sets. Five different splits for the
+% training and test data are generated from each dataset. The results displayed
+% are the average recognition accuracy of the five splits. The accuracy is
+% computed as the number of correctly recognized samples divided by the total
+% number of samples in the test.
+
+In our experiment we used three different datasets.
+The first dataset (Figure \ref{fig:HsSet}) was collected by Hse and Newton\cite{HeloiseBeautification} (Hs-DS).
 The data are drawn by 16 users each of them was asked to draw 13 shapes from 30 to 50 times.
-Figure \ref{fig:symbolSet} shows a set of the shapes used in the data set.
-Five different splits were generated to divide the dataset into training and test sets.
+Figure \ref{fig:ELsymbolSet} and \ref{fig:LogicsymbolSet} shows the symbols we manually gathered from 7 different users in electrical (EL-DS) and logic design (LD-DS) domain respectively.
+Table \ref{tab:datasets} shows a comparison between the three datasets used in our experiments.
+The datasets were divided into training and test sets.
+Five different splits for the training and test data are generated from each dataset.
 The results displayed are the average recognition accuracy of the five splits.
@@
 
-		\includegraphics[scale=0.5]{symbolSet.PNG} 
-		\caption[The Symbol Set] {The symbol set}   \label{fig:symbolSet}
+\fbox{ \parbox{5cm}{%
+		\includegraphics[scale=0.5]{images/symbolSet.PNG}	}}
+		\caption{The Hs-DS Symbol Set} The dataset consist of simple presentation symbols.
+		\label{fig:HsSet}
 \end{figure}
+\begin{table}
+\begin{center}
+\scalebox{0.8}{
+ \begin{tabular}{|p{3cm}|p{1.5cm}|p{1.5cm}|p{1.5cm}|}
+ \hline
+ & Hs-DS &  EL-DS & LD-DS \\ \hline
+No of samples  &  7791 & 2764 & 1859 \\ \hline
+No of categories  & 13 & 15 & 8 \\ \hline
+Avg. Samples per Category& 600 & 184 & 232 \\\hline
+No of users  & 16 & 7 & 7 \\  \hline
+Balanced & Yes&No & No \\ \hline
+Splits  & 5 random splits  & 5 random splits  & 5 random splits  \\ \hline
+Examples & Figure \ref{fig:HsSet}  & Figure \ref{fig:ELsymbolSet}  & Figure \ref{fig:LogicsymbolSet}   \\ \hline
+\end{tabular}
+}
+\caption[Datasets Comparisons]{Table compares between datasets used in experiments in terms of number of samples, number of samples per category, etc...}
+\label{tab:datasets}
+
+\end{center}
+\end{table}
+\begin{figure}
+\centering
+
+\fbox{ \parbox{5cm}{%
+		\includegraphics[scale=0.5]{images/symbolSet.PNG}}}
+		\caption{The Hs-DS Symbol Set} The dataset consist of simple
+presentation symbols.
+		\label{fig:HsSet}
+\end{figure}
+ \begin{figure}
+			\subfigure[The Digital Design Symbols Dataset(LD-DS)]{\label{fig:LogicsymbolSet} \includegraphics[scale=0.5]{images/logicSet.jpg}}
+			\hfill
+			\subfigure[The Electrical Symbols Dataset(EL-DS)]
+			 {\label{fig:ELsymbolSet}\includegraphics[scale=0.45]{images/EelectImage.jpg}}
+
+	\caption{The datasets that were manually collected}  a)  The (LD-DS) dataset consist of symbols that are visually similar (i.e.
+has low shape diversity).
+b) The (EL-DS) dataset consist of complex topological symbols.
+\end{figure}
+ \begin{figure}\centering
+			%\subfigure[The Digital Design Symbols Dataset(LD-DS)]{
+
+\includegraphics[scale=0.5]{images/logicSet.jpg}\label{fig:LogicsymbolSet}%}
+	\caption{The Digital Design Symbols Dataset(LD-DS)}    The (LD-DS)
+dataset consist of symbols that are visually similar (i.e. has low shape
+diversity).
+\end{figure}
+\subsection{Segmentation Algorithms comparisons}
+\label{sec:AlgExp}
+We performed several experiments to evaluate the presented recognition system.
+Firstly, we tested recognition accuracy of shapes in the data set with both algorithms (\textsl{AlgS1} and \textsl{AlgS2}).
+We also implemented the segmentation algorithm described in \cite{earlyprocess} (\textsl{Alg3}) to compare it with our swarm algorithms.
+Figure \ref{fig:test1} and \ref{fig:testLD} show the accuracy achieved by each algorithm.
+The two swarm algorithms were tested with and without the ellipse fitting module.
+Combining the ellipse detection module improves the performance in certain datasets.
+The result shows that (\textsl{AlgS2}) gives better performance than Alg3 and AlgS1 in all datasets.
+The results also shows that combining (\textsl{AlgS1} and \textsl{AlgS2} using minimum error) outperforms AlgS2 in Hs-Ds only.
+This behavior is explained by noticing that features are computed based on the chosen segmentation.
+This means that different segmentation for a symbol generates a different set of features used for recognition.
+This behavior affects Ls-Ds and EL-Ds more than Hs-Ds due to the higher topological and structure nature of the symbols in both datasets than of Hs-Ds.
+
+ \begin{figure}
+	\centering
+	\subfigure[Algorithm comparison on HS-DS]{\label{fig:test1} \includegraphics[scale=0.4]{images/testAlg.jpg}}
+ 		\subfigure[Algorithm comparison on LD-DS]{ 	 	\label{fig:testLD} \includegraphics[scale=0.4]{images/LDAlg.jpg}}
+ 		\subfigure[Algorithm comparison on EL-DS]{	 	\label{fig:testEL}  \includegraphics[scale=0.4]{images/AlgEL.jpg}}
+	 	\caption{ The recognition rate of different algorithms. }
+ \end{figure}
+\subsection{Swarm Algorithms Experiments}
+Another set of tests were performed on the system to enhance non-ellipse fitting
+algorithms. The effect of the number of particles, the error thresholds and
+various other parameters were investigated. The results of these tests are shown
+in Figures \ref{fig:swarmtesting} and   \ref{fig:swarmtesting2}. Figure
+\ref{fig:swarmtesting} shows that as the number of swarm iterations increase the
+error value decrease (error of the segmentation algorithm). The final number of
+particles and maximum swarm iterations used in the system was based on a
+tradeoff between error achieved versus the computation time. The effect of other
+algorithm parameters such as $c_1$, $\omega$ (mentioned in
+Section\ref{sec:ParticleSwarmAlgorithm}) is tested using similar tests (Figure
+\ref{fig:swarmtesting2},\ref{fig:swarmwtest}).%, $c_2$
+
+ \begin{figure}
+	\centering
+	 	\subfigure[Segmentation Error Vs. Swarm Iterations]{
+\label{fig:swarmtesting}
+\includegraphics[scale=0.3]{images/swarmIterationVsParticles.jpg}}
+
+		\subfigure[Segmentation Error Vs. $c_1$]{
+\label{fig:swarmtesting2}  \includegraphics[scale=0.3]{images/C1Swarm.jpg}}
+%swarmParamterC.jpg
+
+				\subfigure[Segmentation Error Vs.
+$\omega$]{\label{fig:swarmwtest}
+\includegraphics[scale=0.3]{images/wSwarm.jpg}} %swarmParamterC.jpg
+
+	 	\caption{Segmentation Error Vs. PSO parameters.}a) The error
+decrease as the number of swarm iteration increase until it reach a saturation.
+ b) The behavior of segmentation error versus $c_1$. c) The behavior of
+segmentation error versus  $\omega$.
+
+\end{figure}
+
+%In our experiment we used the same benchmark dataset used in \cite{mypaper},
+%the dataset was collected by Hse and Newton\cite{HeloiseBeautification}. The
+%data are drawn by 16 users each of them was asked to draw 13 shapes from 30 to
+%50 times. Figure \ref{fig:symbolSet} shows a set of the shapes used in the data
+%set. Five different splits were generated to divide the dataset into training
+%and test sets. The results displayed are the average recognition accuracy of the
+%five splits. The accuracy is computed as the number of correctly recognized
+%samples divided by the total number of samples in the test.
+
+%  \begin{figure}
+%   \centering
+%
+% 		\includegraphics[scale=0.5]{symbolSet.PNG}
+% 		\caption[The Symbol Set] {The symbol set}
+%\label{fig:symbolSet}
+% 	 \end{figure}
 
@@
 Figures \ref{exp2} and \ref{exp1} show the recognition rate of each algorithm using the full feature set and half the features set respectively.
-The figures compares between the result that can be obtained using Traditional PSO by system in \cite{mypaper} and modifying the system using Enhanced PSO.
+The figures compares between the result that can be obtained using Traditional PSO and modifying the system using Enhanced PSO.
 The results show that the enhanced swarm optimization algorithm improves the segmentation of the system which leads to an increase in the recognition rate of the whole system by an average of $1.5\%$.
@@
   \caption[Recognition accuracy]%
-  {The chart compares between recognition rate of the full feature set on both the traditional PSO system presented in \cite{mypaper} and Enhanced PSO.
-The results show that Enhanced PSO increases the performance over the  Traditional PSO.
-}
+  {The chart compares between recognition rate of the full feature set on both
+The traditional PSO system and the Enhanced PSO. The results show that Enhanced PSO increases the performance over the  Traditional PSO. }
    \label{exp2}
@@
   \caption[Recognition accuracy]%
-  {The chart compares between recognition rate of half the feature set on both the traditional PSO system presented in \cite{mypaper} and Enhanced PSO.
-The results show that Enhanced PSO increases the recognition rate.
-}
+  {The chart compares between recognition rate of half the feature set on both
+the traditional PSO system and Enhanced PSO. The esults show that Enhanced PSO increases the recognition rate. }
    \label{exp1}
@@
 
-This paper used an Enhanced-PSO to modify a new sketch recognition using DPSO represented in \cite{mypaper}.
-Results show that enhancing the solutions that the \textit{PSO} in general generates an optimized stroke segmentation which improves the final recognition rate.
-The final recognition improved by modifying the particle optimization algorithm without depending on the higher level features.
-The experiments evaluated two \textit{DPSO} different segmentation algorithms (AlgS1 and AlgS2) on simple presentation set dataset.
-Results show that the enhanced PSO algorithm improves the segmentation on both algorithms which achieves better segmentation and final accuracy.
-The results proved that the Enhancement made on the PSO improves the efficiency of the Traditional PSO algorithm without being dependable on special objective function or application.
-The system achieved an average overall improvement of more than 1\% over Basic PSO.
+This paper used an Enhanced-PSO to modify a new sketch recognition using DPSO
+represented in \cite{mypaper}. Results show that enhancing the solutions that
+the \textit{PSO} in general generates an optimized stroke segmentation which
+improves the final recognition rate.  The final recognition improved by
+modifying the particle optimization algorithm without depending on the higher
+level features. The experiments evaluated two \textit{DPSO} different
+segmentation algorithms (AlgS1 and AlgS2) on simple presentation set dataset.
+Results show that the enhanced PSO algorithm improves the segmentation on both
+algorithms which achieves better segmentation and final accuracy. The results
+proved that the Enhancement made on the PSO improves the efficiency of the
+Traditional PSO algorithm without being dependable on special objective function
+or application. The system achieved an average overall improvement of more than
+1\% over Basic PSO.
 
- For the future work, it is suggested that we continue to improve both the enhanced PSO segmentation algorithm.
-Also, a possible extension of this research is to improve the recognition algorithm of the Sketch Recognition system.
-Other area of enhancements is the features extraction methods.
-Introducing more spatial and geometrical features is believed to improve classifications.
+This paper presented a new approach to sketch recognition using DPSO. The system
+uses multiple preliminary information, i.e., speed and curvature information, as
+input to the \textit{DPSO}. Results show that the \textit{DPSO} in general
+generates an optimized stroke segmentation which improves the final recognition
+rate.  The final recognition is based on both global shape and hierarchical
+stroke properties. Adding stroke and segment geometrical features improved the
+final recognition. Global shape and moment properties were used to prevent the
+need to use graph matching which is computationally expensive
+\cite{SketchRead2007}. The experiments evaluated two \textit{DPSO} different
+segmentation algorithms (AlgS1 and AlgS2) on three different datasets (Hs-Ds,
+EL-Ds and LD-Ds). Results show that AlgS2 achieves better segmentation and final
+accuracy than Alg3 \cite{earlyprocess} on the datasets used. The results proved
+that the system is easily expandable to more and more symbols as it does not
+depend on low level primitive recognizers but on high level symbol and
+geometrical features.  The system achieved an average overall improvement of
+more than 2\% over Alg3 in all datasets.
+
+  The tradeoff between accuracy achieved and time complexity can be further
+investigated to achieve better results. Currently, the system processes one
+symbol at a time, an enhancement would be to allow users to draw more than one
+symbol at a time and incorporate a method for separating symbols. Thus, possible
+extension of this research is to complete the clustering algorithm for fully
+automated sketch recognition. Another area of enhancements is the features
+extraction methods. Introducing more spatial and geometrical features is
+believed to improve classifications. Features that represent the appearance of
+the symbols are likely to improve the recognition of symbols that have similar
+geometrical structure (i.e. Nor and OR gate in logic symbols).
+
+ For the future work, it is suggested that we continue to improve both the
+enhanced PSO segmentation algorithm. Also, a possible extension of this research
+is to improve the recognition algorithm of the Sketch Recognition system. Other
+area of enhancements is the features extraction methods. Introducing more
+spatial and geometrical features is believed to improve classifications.
 

### 2010-12-10 Swarm/swarmModified.tex

@@
-\documentclass[a4paper, twoside]{article}
+%\documentclass[a4paper, twoside]{article}
 
-\usepackage{icgst}
-%\documentclass[12pt,1p,times]{elsarticle}%[10pt]{article}
+%%\usepackage{icgst}
+%\documentclass[12pt,times,a4paper]{article}%[10pt]{article}
 %% Use the option review to obtain double line spacing
-%% \documentclass[preprint,review,12pt]{elsarticle}
+ \documentclass[preprint,1p,times,review]{elsarticle}
 
@@
 \usepackage{amsmath}
-
 %\usepackage{named}
@@
 \usepackage{amssymb}
-\usepackage{epsfig}
-\usepackage{epstopdf}
-\usepackage[dvips,pdftex]{graphicx}
+%\usepackage{epsfig}
+%\usepackage{epstopdf}
+%\usepackage{xltxtra}
+\usepackage{graphicx}
+\usepackage{subfigure}
+\usepackage{sectsty}
+%\sectionfont{\normalsize  }
+%\subsectionfont{ \normalsize\it \textbf }
+%\subsubsectionfont{\normalsize \it \textbf }
+%\paragraphfont{\large\it}
+%\subparagraphfont{\normalsize\it}
+
 %% The lineno packages adds line numbers. Start line numbering with
@@
 % \biboptions{}
+% \setlength\hoffset{0in}
+% % \setlength\topmargin{0in}
+% % \setlength\headheight{0in}
+% % \setlength\headsep{-0.5in}
+% % \setlength\textheight{10.0in}
+%  \setlength\textwidth{7in}
+%  \setlength\oddsidemargin{-0.5in}
+%  \setlength\evensidemargin{-0.5in}
+% % \setlength\parindent{0.25in}
+% % \setlength\parskip{0.25in}
+% \linespread{1.5}
+\begin{abstract}
+
+%This study
+%%%This study
+
+Sketches and drawings are widely used as a simple method of expressing thought and designs.
+The goal of this paper is developing a sketch recognition system that facilitates using sketches in computer systems.
+The system uses Particle Swarm Algorithms to optimally segment the strokes the user draws into meaningful geometric primitives.
+These geometric primitives are grouped to formulate symbols which are further identified using a Support Vector Machines (SVM) classifier.
+A comparision is presented between using Traditional Particle Swarm Optimization and the Enhanced Particle Swarm Optimization (EPSO) which is based on the perturbation of the particle velocity.
+Two different segmentation algorithms were evaluated and results show that both algorithms improve the final recognition results.
+Experiments were conducted on three different dataset, simple presentation (Hs-DS), Electrical Design (EL-DS) and Logic Design (LD-DS) symbols.
+Results show that the effectiveness of the proposed system on various dataset especially on datasets with low shape diversity as in Logic Design dataset (LD-DS).
+Results also show that using an Enhancment on the Tradiontal Swarm algorithm improves recognition rate.
+
+\end{abstract}
 \begin{document}
@@
 
-%\title{Sketch Recognition using Enhanced Particle Swarm Optimization}
+\title{Sketch Recognition using Enhanced Particle Swarm Optimization}
 
-\twocolumn[
-\scalebox{1.00}{\includegraphics[width=1.00\textwidth]{aimlhead}}
-\title{Sketch Recognition using Enhanced Particle Swarm Optimization}
-\vspace{0.1in}
-\author{}  
-\begin{center}
-Maha El Meseery\footnote{Corresponding Author }, Mahmoud Fakhr El Din, Tarek abouldahab\\
-\textit{Electronic Research Institute,} Cairo, Egypt,\\
-\textit{ Cairo Metro Company, Ministry of transport }Cairo, Egypt\\
-{melmeseery@eri.sci.eg, mafakhr@mcit.gov.eg,  heshoaboda@hotmail.com},\\
-%\texttt{http://www.faculty.university.edu.any}
-\end{center}
-\vspace{0.25in} 
-]
+% \twocolumn[
+% \scalebox{1.00}{\includegraphics[width=1.00\textwidth]{aimlhead}}
+% \title{Sketch Recognition using Enhanced Particle Swarm Optimization}
+% \vspace{0.1in}
+% \author{}
+% \begin{center}
+% Maha El Meseery\footnote{Corresponding Author }, Mahmoud Fakhr El Din, Tarek
+%abouldahab\\
+% \textit{Electronic Research Institute,} Cairo, Egypt,\\
+% \textit{ Cairo Metro Company, Ministry of transport }Cairo, Egypt\\
+% {melmeseery@eri.sci.eg, mafakhr@mcit.gov.eg,  heshoaboda@hotmail.com},\\
+% %\texttt{http://www.faculty.university.edu.any}
+% \end{center}
+% \vspace{0.25in}
+% ]
 
@@
 
-% \twoauthors
+\author{
+\small{
+Maha El Meseery \footnote{Corresponding Author} \textsuperscript{a}, Mahmoud
+Fakhr El Din \textsuperscript{a}, Tarek abouldahab \textsuperscript{b}}\\
+ \textsuperscript{a} \small{Signals Processing Group, Computers and Systems
+Department,}\\ \small{ Electronic Research Institute, Cairo, Egypt.}\\
+  \small {\textsuperscript{b}Cairo Metro Company ,
+   Ministry of transport , Cairo, Egypt.}\\
+   \small{melemseery@eri.sci.eg,  mafakhr@mcit.gov.eg, heshoaboda@hotmail.com
+   }
+}
+\date{}
 % {\sthanks{Corresponding Author},}
@@
 % 	melmeseery@eri.sci.eg, mafakhr@mcit.gov.eg}
-%   {} {Cairo Metro Company \\
+%   {Tarek abouldahab} {Cairo Metro Company \\
+%   Ministry of transport \\Cairo, Egypt\\ heshoaboda@hotmail.com}
+%  {Maha El Meseery \sthanks{Corresponding Author}, Mahmoud Fakhr El Din}
+% 	{Signals Processing Group \\ Computers and Systems Department \\
+%   Electronic Research Institute\\Cairo, Egypt\\
+% 	melmeseery@eri.sci.eg, mafakhr@mcit.gov.eg}
+% 	 {Tarek abouldahab\sthanks{Corresponding Author}} {Cairo Metro Company
+%\\
 %   Ministry of transport \\Cairo, Egypt\\ heshoaboda@hotmail.com}
 
-%\maketitle
- \begin{abstract}
-%This study
-
-Sketches and drawings are widely used as a simple method of expressing thought and designs.
-The goal of this paper is to improve our sketch recognition system by using an enhanced particle swarm optimization algorithm.
-The Enhanced Particle Swarm Optimization (EPSO) is based on the perturbation of the particle velocity is used to optimally segment the strokes the user draws into meaningful geometric primitives.
+\maketitle
+%  \begin{abstract}
+%
+% %This study
+% %%%This study
+%
+% Sketches and drawings are widely used as a simple method of expressing thought and designs.
+The goal of this paper is developing a sketch recognition system that facilitates using sketches in computer systems.
+The system uses Particle Swarm Optimization (PSO) algorithm to optimally segment the strokes the user draws into meaningful geometric primitives.
 These geometric primitives are grouped to formulate symbols which are further identified using a Support Vector Machines (SVM) classifier.
-Experiments were conducted on the benchmark dataset of simple presentation (Hs-DS) symbols.
-Results show that that the enhanced particle swarm optimization improves the final recognition accuracy.
-
-\end{abstract}
+Two different segmentation algorithms were evaluated and results show that both algorithms improve the final recognition results.
+Experiments were conducted on three different dataset, simple presentation (Hs-DS), Electrical Design (EL-DS) and Logic Design (LD-DS) symbols.
+Results show that the effectiveness of the proposed system on various dataset especially on datasets with low shape diversity as in Logic Design dataset (LD-DS).
+%
+%
+%  \textsl{Sketches and drawings are widely used as a simple method of expressing
+% thought and designs. The goal of this paper is to improve our sketch recognition
+% system by using an enhanced particle swarm optimization algorithm. The Enhanced
+% Particle Swarm Optimization (EPSO) is based on the perturbation of the particle
+% velocity is used to optimally segment the strokes the user draws into meaningful
+% geometric primitives. These geometric primitives are grouped to formulate
+% symbols which are further identified using a Support Vector Machines (SVM)
+% classifier. Experiments were conducted on the benchmark dataset of simple
+% presentation (Hs-DS) symbols. Results show that the enhanced particle swarm
+% optimization improves the final recognition accuracy.}
+% \end{abstract}
 
@@
 
-This paper modifies the sketch recognition system presented in \cite{mypaper}, to use enhanced particle swarm algorithm (EPSO) to optimally segment user's strokes.
-As in the Original system the users can draw symbols using any number of strokes in any order they like.
-The system segments each stroke using two different EPSO algorithms.
-Segmentation is based on dividing the digital curve into polygon\cite{PolygonApproximationPSO}.
-To handle complex strokes, our system uses a second segmentation EPSO algorithm to segment strokes into a set of lines and circular arcs.
-After segmentation, a symbol features vector composed of different statistical, geometrical and spatial features is deduced.
-The final classification step, SVM classifier uses the computed feature vector to classify the segments into one of the previously trained classes.
+In PR based methods, systems compute properties that describe symbols then use a classification method to recognize symbols.
+The properties used can either be based on gesture properties, global shape properties or image based properties.
+Gesture properties are set of features that represent how the user draws a symbol not how it looks like (e.g., identify a rectangle when a user draws a L corner).
+In gesture properties approach, systems identify symbols by using gesture properties of some of its underlying strokes.
+A system that uses gestures properties assumes that users draws symbols in specific order and style \cite{gestureexample12,aideddesgin22}.
+In global shape properties approach, systems investigate general shape properties and its underlying strokes.
+Using shape properties (e.g., ratio of the area of convex hull to the area of shape) relaxes the assumption and the role of each stroke in the symbol but it does not truly represent the appearance or fine details of the symbol \cite{DiagramOfflineConvexHull,Cali63}.
+Unlike shape properties based methods, image based methods disregard individual strokes and focus on the overall appearance of shapes.
+These system convert the shapes into image and extracted image based features (e.g.
+local gradient feature)\cite{Oltmans07,imagetrainable48}.
+In \cite{imagetrainable48}, they used four different distance measures to match user symbol with the prototype database.
+The prototype of a shape is represented as an image of the symbol that is normalized and scaled.
+However, the problem is that the shape prototype must represent all transformation and variation the input shape can have.
+%Image based methods are used in \cite{Oltmans07,imagetrainable48} where they computed a code book of 1000 features to recognize symbols.
 
-The remaining of the paper is as following; Section \ref{sec:ParticleSwarmAlgorithm} describes the Basic particle swarm algorithm in details.
-Enhanced Particle Swarm Algorithm is described in Section \ref{sec:EnhancedSwarmAlgorithm}.
-Section \ref{Sysdisc} provide details on preprocessing, segmentation and recognition process.
-Section \ref{sec:Experiments} describes the experiments that were preformed.
-We conclude in Section \ref{ConclusionandFutureWork}.
+ Methods based on gesture and global shape properties produce systems that are dependent on assumptions about how strokes can be grouped.
+These systems assume that users will finish drawing a shape before moving to the next one \cite{Cali63,geometrydomain49}.
+This assumption allows the system to temporally group strokes and then recognizes them based on their strokes or global shape properties.
+Alvarado \cite{AlvaradoDigital} argued that this assumption does not hold in natural sketches.
+
+In this paper, we choose to implement a hybrid method that includes shape property, gesture and hierarchical information.
+Our method attempts to avoid many of the problems faced by approaches based on individual gestures or global shape properties while trying to gain the additional information that the stroke based system uses.
+Firstly, the system uses particle swarm optimization (PSO) to optimally segment user's strokes.
+Users can draw symbols using any number of strokes in any order they like.
+The system segments each stroke using two different PSO algorithms.
+The first segmentation algorithm is based on dividing the digital curve into polygons\cite{PolygonApproximationPSO}.
+Our algorithm enhances its performance by adding curvature and other local information during segmenting strokes to achieve better results.
+To handle complex strokes, our system uses a second segmentation PSO algorithm that segment strokes into a set of lines and circular arcs.
+The two segmentation algorithms can either by used individually or in parallel then chosing the segmentation that gives the minimum error.
+After segmentation, a feature vector composed of different statistical, geometrical and composite features is deduced.
+The final classification step uses a SVM classifier to classify symbols into one of the previously trained classes based on the features vector computed\cite{mypaper}.
+
+  In addition to the system mentioned above, an Enhnaced Swarm algorithm is compared to the Traditional Particle swarm algorithm\cite{stockPaper}.
+The Enhanced swarm algorithm is based on chaning the particle velocity by replacing it with the weighted differance between the particle and another random in the space.
+This random relacing is only used when the new velociy and location is better then the current particle values.
+The procedure introude new expolration factor the swarm system which helps in reaching the solution faster and more obtimially.
+
+% %This paper modifies the sketch recognition system presented in \cite{mypaper},
+% to use enhanced particle swarm algorithm (EPSO) to optimally segment user's
+% strokes. As in the Original system the users can draw symbols using any number
+% of strokes in any order they like. The system segments each stroke using two
+% different EPSO algorithms. Segmentation is based on dividing the digital curve
+% into polygon\cite{PolygonApproximationPSO}. To handle complex strokes, our
+% system uses a second segmentation EPSO algorithm to segment strokes into a set
+% of lines and circular arcs. After segmentation, a symbol features vector
+% composed of different statistical, geometrical and spatial features is deduced.
+% The final classification step, SVM classifier uses the computed feature vector
+% to classify the segments into one of the previously trained classes.
+
+The remaining of the paper is as following; Section
+\ref{sec:ParticleSwarmAlgorithm} describes the Basic particle swarm algorithm in
+details.  Enhanced Particle Swarm Algorithm is described in Section
+\ref{sec:EnhancedSwarmAlgorithm}. Section \ref{Sysdisc} provide details on
+preprocessing, segmentation and recognition process. Section
+\ref{sec:Experiments} describes the experiments that were preformed. We conclude
+in Section \ref{ConclusionandFutureWork}.
 
@@
 
-This paper used an Enhanced-PSO to modify a new sketch recognition using DPSO represented in \cite{mypaper}.
-Results show that enhancing the solutions that the \textit{PSO} in general generates an optimized stroke segmentation which improves the final recognition rate.
-The final recognition improved by modifying the particle optimization algorithm without depending on the higher level features.
-The experiments evaluated two \textit{DPSO} different segmentation algorithms (AlgS1 and AlgS2) on simple presentation set dataset.
-Results show that the enhanced PSO algorithm improves the segmentation on both algorithms which achieves better segmentation and final accuracy.
-The results proved that the Enhancement made on the PSO improves the efficiency of the Traditional PSO algorithm without being dependable on special objective function or application.
-The system achieved an average overall improvement of more than 1\% over Basic PSO.
+This paper used an Enhanced-PSO to modify a new sketch recognition using DPSO
+represented in \cite{mypaper}. Results show that enhancing the solutions that
+the \textit{PSO} in general generates an optimized stroke segmentation which
+improves the final recognition rate.  The final recognition improved by
+modifying the particle optimization algorithm without depending on the higher
+level features. The experiments evaluated two \textit{DPSO} different
+segmentation algorithms (AlgS1 and AlgS2) on simple presentation set dataset.
+Results show that the enhanced PSO algorithm improves the segmentation on both
+algorithms which achieves better segmentation and final accuracy. The results
+proved that the Enhancement made on the PSO improves the efficiency of the
+Traditional PSO algorithm without being dependable on special objective function
+or application. The system achieved an average overall improvement of more than
+1\% over Basic PSO.
 
- For the future work, it is suggested that we continue to improve both the enhanced PSO segmentation algorithm.
-Also, a possible extension of this research is to improve the recognition algorithm of the Sketch Recognition system.
-Other area of enhancements is the features extraction methods.
-Introducing more spatial and geometrical features is believed to improve classifications.
+This paper presented a new approach to sketch recognition using DPSO. The system
+uses multiple preliminary information, i.e., speed and curvature information, as
+input to the \textit{DPSO}. Results show that the \textit{DPSO} in general
+generates an optimized stroke segmentation which improves the final recognition
+rate.  The final recognition is based on both global shape and hierarchical
+stroke properties. Adding stroke and segment geometrical features improved the
+final recognition. Global shape and moment properties were used to prevent the
+need to use graph matching which is computationally expensive
+\cite{SketchRead2007}. The experiments evaluated two \textit{DPSO} different
+segmentation algorithms (AlgS1 and AlgS2) on three different datasets (Hs-Ds,
+EL-Ds and LD-Ds). Results show that AlgS2 achieves better segmentation and final
+accuracy than Alg3 \cite{earlyprocess} on the datasets used. The results proved
+that the system is easily expandable to more and more symbols as it does not
+depend on low level primitive recognizers but on high level symbol and
+geometrical features.  The system achieved an average overall improvement of
+more than 2\% over Alg3 in all datasets.
 
-\bibliographystyle{plain}
+  The tradeoff between accuracy achieved and time complexity can be further
+investigated to achieve better results. Currently, the system processes one
+symbol at a time, an enhancement would be to allow users to draw more than one
+symbol at a time and incorporate a method for separating symbols. Thus, possible
+extension of this research is to complete the clustering algorithm for fully
+automated sketch recognition. Another area of enhancements is the features
+extraction methods. Introducing more spatial and geometrical features is
+believed to improve classifications. Features that represent the appearance of
+the symbols are likely to improve the recognition of symbols that have similar
+geometrical structure (i.e. Nor and OR gate in logic symbols).
+
+ For the future work, it is suggested that we continue to improve both the
+enhanced PSO segmentation algorithm. Also, a possible extension of this research
+is to improve the recognition algorithm of the Sketch Recognition system. Other
+area of enhancements is the features extraction methods. Introducing more
+spatial and geometrical features is believed to improve classifications.
+
+%\bibliographystyle{natbib}
+\bibliographystyle{ieeetr}
 \bibliography{../../../neededfiles/Bibliographies/Mybibliography}

### 2010-12-10 trunk/Swarm/swarmModified.tex

@@
  The field of symbol recognition has gained interest in the last few years.
-Scientists generally and engineers specifically express thoughts and designs
-using sketches. Engineers use sketches to exchange designs as a natural method
-of communication rather than writing or speaking. Design engineers need a
-powerful computer based system with symbol recognition to help them design,
-manipulate and store sketches more effectively than using only papers.
-Increasing the interaction between computers and users in sketch and CAD systems
-has been the reason for the emerging of a few advanced sketch recognition
-systems. Sketch recognition is the process of identifying user strokes into
-meaningful symbols that can be further used.
+Scientists generally and engineers specifically express thoughts and designs using sketches.
+Engineers use sketches to exchange designs as a natural method of communication rather than writing or speaking.
+Design engineers need a powerful computer based system with symbol recognition to help them design, manipulate and store sketches more effectively than using only papers.
+Increasing the interaction between computers and users in sketch and CAD systems has been the reason for the emerging of a few advanced sketch recognition systems.
+Sketch recognition is the process of identifying user strokes into meaningful symbols that can be further used.
 
-Sketch recognition is divided into three main steps preprocessing, segmentation
-and symbol recognition. The preprocessing phase captures user input stroke
-points and collects basic information about the stroke then proceeds to remove
-noise and compute basic statistical and geometrical information. In segmentation
-phase, the strokes are divided into a set of simple geometrical primitives or
-segments. In the third phase, sketch recognition, strokes and segments are
-clustered to formulate symbols that can be recognized by a classifier system.
+ Sketch recognition is divided into three main steps preprocessing, segmentation and symbol recognition.
+The preprocessing phase captures user input stroke points and collects basic information about the stroke then proceeds to remove noise and compute basic statistical and geometrical information.
+In segmentation phase, the strokes are divided into a set of simple geometrical primitives or segments.
+In the third phase, sketch recognition, strokes and segments are clustered to formulate symbols that can be recognized by a classifier system.
 
-In PR based methods, systems compute properties that describe symbols then use a
-classification method to recognize symbols. The properties used can either be
-based on gesture properties, global shape properties or image based properties.
-Gesture properties are set of features that represent how the user draws a
-symbol not how it looks like (e.g., identify a rectangle when a user draws a L
-corner). In gesture properties approach, systems identify symbols by using
-gesture properties of some of its underlying strokes. A system that uses
-gestures properties assumes that users draws symbols in specific order and style
-\cite{gestureexample12,aideddesgin22}. In global shape properties approach,
-systems investigate general shape properties and its underlying strokes. Using
-shape properties (e.g., ratio of the area of convex hull to the area of shape)
-relaxes the assumption and the role of each stroke in the symbol but it does not
-truly represent the appearance or fine details of the symbol
-\cite{DiagramOfflineConvexHull,Cali63}. Unlike shape properties based methods,
-image based methods disregard individual strokes and focus on the overall
-appearance of shapes. These system convert the shapes into image and extracted
-image based features (e.g. local gradient
-feature)\cite{Oltmans07,imagetrainable48}. In \cite{imagetrainable48}, they used
-four different distance measures to match user symbol with the prototype
-database. The prototype of a shape is represented as an image of the symbol that
-is normalized and scaled.  However, the problem is that the shape prototype must
-represent all transformation and variation the input shape can have.   %Image
-based methods are used in \cite{Oltmans07,imagetrainable48} where they computed
-a code book of 1000 features to recognize symbols.
+In PR based methods, systems compute properties that describe symbols then use a classification method to recognize symbols.
+The properties used can either be based on gesture properties, global shape properties or image based properties.
+Gesture properties are set of features that represent how the user draws a symbol not how it looks like (e.g., identify a rectangle when a user draws a L corner).
+In gesture properties approach, systems identify symbols by using gesture properties of some of its underlying strokes.
+A system that uses gestures properties assumes that users draws symbols in specific order and style \cite{gestureexample12,aideddesgin22}.
+In global shape properties approach, systems investigate general shape properties and its underlying strokes.
+Using shape properties (e.g., ratio of the area of convex hull to the area of shape) relaxes the assumption and the role of each stroke in the symbol but it does not truly represent the appearance or fine details of the symbol \cite{DiagramOfflineConvexHull,Cali63}.
+Unlike shape properties based methods, image based methods disregard individual strokes and focus on the overall appearance of shapes.
+These system convert the shapes into image and extracted image based features (e.g.
+local gradient feature)\cite{Oltmans07,imagetrainable48}.
+In \cite{imagetrainable48}, they used four different distance measures to match user symbol with the prototype database.
+The prototype of a shape is represented as an image of the symbol that is normalized and scaled.
+However, the problem is that the shape prototype must represent all transformation and variation the input shape can have.
+%Image based methods are used in \cite{Oltmans07,imagetrainable48} where they computed a code book of 1000 features to recognize symbols.
 
-Methods based on gesture and global shape properties produce systems that are
-dependent on assumptions about how strokes can be grouped. These systems assume
-that users will finish drawing a shape before moving to the next one
-\cite{Cali63,geometrydomain49}. This assumption allows the system to temporally
-group strokes and then recognizes them based on their strokes or global shape
-properties. Alvarado \cite{AlvaradoDigital} argued that this assumption does not
-hold in natural sketches.
+ Methods based on gesture and global shape properties produce systems that are dependent on assumptions about how strokes can be grouped.
+These systems assume that users will finish drawing a shape before moving to the next one \cite{Cali63,geometrydomain49}.
+This assumption allows the system to temporally group strokes and then recognizes them based on their strokes or global shape properties.
+Alvarado \cite{AlvaradoDigital} argued that this assumption does not hold in natural sketches.
 
-In this paper, we choose to implement a hybrid method that includes shape
-property, gesture and hierarchical information. Our method attempts to avoid
-many of the problems faced by approaches based on individual gestures or global
-shape properties while trying to gain the additional information that the stroke
-based system uses. Firstly, the system uses particle swarm optimization (PSO) to
-optimally segment user's strokes. Users can draw symbols using any number of
-strokes in any order they like. The system segments each stroke using two
-different PSO algorithms. The first segmentation algorithm is based on dividing
-the digital curve into polygons\cite{PolygonApproximationPSO}. Our algorithm
-enhances its performance by adding curvature and other local information during
-segmenting strokes to achieve better results. To handle complex strokes, our
-system uses a second segmentation PSO algorithm that segment strokes into a set
-of lines and circular arcs. The two segmentation algorithms can either by used
-individually or in parallel then chosing the segmentation that gives the minimum
-error. After segmentation, a feature vector composed of different statistical,
-geometrical and composite features is deduced. The final classification step
-uses a SVM classifier to classify symbols into one of the previously trained
-classes based on the features vector computed\cite{mypaper}.
+In this paper, we choose to implement a hybrid method that includes shape property, gesture and hierarchical information.
+Our method attempts to avoid many of the problems faced by approaches based on individual gestures or global shape properties while trying to gain the additional information that the stroke based system uses.
+Firstly, the system uses particle swarm optimization (PSO) to optimally segment user's strokes.
+Users can draw symbols using any number of strokes in any order they like.
+The system segments each stroke using two different PSO algorithms.
+The first segmentation algorithm is based on dividing the digital curve into polygons\cite{PolygonApproximationPSO}.
+Our algorithm enhances its performance by adding curvature and other local information during segmenting strokes to achieve better results.
+To handle complex strokes, our system uses a second segmentation PSO algorithm that segment strokes into a set of lines and circular arcs.
+The two segmentation algorithms can either by used individually or in parallel then chosing the segmentation that gives the minimum error.
+After segmentation, a feature vector composed of different statistical, geometrical and composite features is deduced.
+The final classification step uses a SVM classifier to classify symbols into one of the previously trained classes based on the features vector computed\cite{mypaper}.
 
@@
 
- Particle Swarm Algorithm\cite{PSOFirst}is a population based stochastic search
-where each solution is simulated as a  particle that flies around in a
-multidimensional search space. A fitness function is used to evaluate the
-goodness of each particle in solution space. But unlike evolutionarily algorithm
-the new population is the PSO is influenced by the social cooperation of the
-particle rather than survival to the fittest. Each particle saves a history of
-the best location in the swarm and its own best location.  At the end each
-iteration, the particles moves to a new location based on the global best
-location and it own history of best location. This use of previous particle
-locations gives PSO advantage over Evolutionary algorithms where no similar
-information is used.
+ Particle Swarm Algorithm\cite{PSOFirst}is a population based stochastic search where each solution is simulated as a  particle that flies around in a multidimensional search space.
+A fitness function is used to evaluate the goodness of each particle in solution space.
+But unlike evolutionarily algorithm the new population is the PSO is influenced by the social cooperation of the particle rather than survival to the fittest.
+Each particle saves a history of the best location in the swarm and its own best location.
+At the end each iteration, the particles moves to a new location based on the global best location and it own history of best location.
+This use of previous particle locations gives PSO advantage over Evolutionary algorithms where no similar information is used.
 
- The \textit{Basic Particle Swarm Algorithm (PSO)} represents each solution with
-a $N$ dimension particle from the solution space \cite{PSOFirst}. Each particle
-moves with a direction and velocity $v_{ij}$ based on equations \ref{eq:Swarm1}
-\& \ref{eq:Swarm}.
+ The \textit{Basic Particle Swarm Algorithm (PSO)} represents each solution with a $N$ dimension particle from the solution space \cite{PSOFirst}.
+Each particle moves with a direction and velocity $v_{ij}$ based on equations \ref{eq:Swarm1} \& \ref{eq:Swarm}.
 
@@
 
-where $p_{ij}$ represent the $j$th dimension in the $i$th particle and $v_{ij}$
-is the velocity of the $j$th dimension in the $i$th particle.
+where $p_{ij}$ represent the $j$th dimension in the $i$th particle and $v_{ij}$ is the velocity of the $j$th dimension in the $i$th particle.
 
@@
 \end{equation}
- where $\omega$ is the inertia weight parameter which controls the tradeoff
-between exploration and exploitation,  $lbest_{ij}$ is the local best particle,
-$gbest_{ij}$ is the global best particle, $r_1$ \& $r_2$ are random variables
-and $c_1$ \& $c_2$ are the swarm acceleration parameters.
+ where $\omega$ is the inertia weight parameter which controls the tradeoff between exploration and exploitation,  $lbest_{ij}$ is the local best particle, $gbest_{ij}$ is the global best particle, $r_1$ \& $r_2$ are random variables and $c_1$ \& $c_2$ are the swarm acceleration parameters.
 
- After each iteration the global best $g_{best}$ particle and the agent local
-best particle $l_{best}$are evaluated based on the maximum fitness functions of
-all particles in the solution space. The solution is found after achieving a
-specific number of iteration or after an error threshold is achieved.
+ After each iteration the global best $g_{best}$ particle and the agent local best particle $l_{best}$are evaluated based on the maximum fitness functions of all particles in the solution space.
+The solution is found after achieving a specific number of iteration or after an error threshold is achieved.
 
@@
 \label{sec:EnhancedSwarmAlgorithm}
-In the above basic PSO, the velocity update is based on three terms an inertia
-term, local cognitive ($l_{best}$) term and global cognitive ($g_{best}$)term.
-The particle local cognitive term is associated with the particle own history
-without consideration of other swarm particles locations and histories which
-leads to a lake of diversity and sub optimum solutions. Therefore, we are trying
-to add more social influence in the particle movement by using the weighted
-difference of the position vectors of any other two distinct particles randomly
-chosen from the swarm to modify the local cognitive term. The modified term adds
-social behavior of other particle and keeps diversity while searching for the
-global optimum. The new social cognitive term of the particle is shifted to the
-new location only if it yields a fitness value better than obtained using the
-standard PSO\cite{stockPaper}.
+
+In the above basic PSO, the velocity update is based on three terms an inertia term, local cognitive ($l_{best}$) term and global cognitive ($g_{best}$)term.
+The particle local cognitive term is associated with the particle own history without consideration of other swarm particles locations and histories which leads to a lake of diversity and sub optimum solutions.
+Therefore, we are trying to add more social influence in the particle movement by using the weighted difference of the position vectors of any other two distinct particles randomly chosen from the swarm to modify the local cognitive term.
+The modified term adds social behavior of other particle and keeps diversity while searching for the global optimum.
+The new social cognitive term of the particle is shifted to the new location only if it yields a fitness value better than obtained using the standard PSO\cite{stockPaper}.
 
 This modification is done as follows:-
-  \begin{enumerate}
-\item For each particle $p_{i}$ in the swarm two other distinct particles, $p_L$
-and $p_M$ are selected randomly where $i\neq L\neq M$. The difference between
-their positional coordinates $p_{d}$ is taken as a difference vector in equation
+  \begin{enumerate} \item For each particle $p_{i}$ in the swarm two other distinct particles, $p_L$ and $p_M$ are selected randomly where $i\neq L\neq M$.
+The difference between their positional coordinates $p_{d}$ is taken as a difference vector in equation
 \ref{eqmod}.
@@
 
-Then the associated velocity vector of particle $p_{i}$ is updated according to
-this difference vector as in equation \ref{eq:mod1}. In essence the local
-cognitive term of the velocity update formula in \ref{eq:Swarm} is replaced with
-the vector differential operator $p_d$ to produce some additional exploration
-capability.
+Then the associated velocity vector of particle $p_{i}$ is updated according to this difference vector as in equation \ref{eq:mod1}.
+In essence the local cognitive term of the velocity update formula in \ref{eq:Swarm} is replaced with the vector differential operator $p_d$ to produce some additional exploration capability.
  \begin{equation}
@@
  \end{equation}
-\item The new position $p^{T}_{ij}$ is used to generate the new trial objective
-fitness function value $fitness(p^{T})$. The fitness value is calculated based
-on this new trial position vector and the particle is shifted to this trial
-position if the trial objective function is better than the original objective
-function associated with standard PSO. Thus the target particle is relocated
-using the following equation (Eq. \ref{eq:objective})
+\item The new position $p^{T}_{ij}$ is used to generate the new trial objective fitness function value $fitness(p^{T})$.
+The fitness value is calculated based on this new trial position vector and the particle is shifted to this trial position if the trial objective function is better than the original objective function associated with standard PSO.
+Thus the target particle is relocated using the following equation (Eq.
+\ref{eq:objective})
   \begin{equation}
@@
 %The modified swarm algorithm is used to improve the performance of the sketch recognition system in \cite{mypaper}.
-
  The sketch recognitions system is divided into three main steps 1) preprocessing, 2) segmentation and 3) Recognition.
-Figure \ref{fig:flowChart} shows the block diagram of the system; the first
-process is collecting stroke points then computing the curvature and possible
-dominant points.  The following section provides a detailed description of each
-step. The main block of the system remains the same but there are some
-modifications in the segmentation algorithm to improve its efficiency.
+Figure \ref{fig:flowChart} shows the block diagram of the system; the first process is collecting stroke points then computing the curvature and possible dominant points.
+The following section provides a detailed description of each step.
+The main block of the system remains the same but there are some modifications in the segmentation algorithm to improve its efficiency.
 This study compares between using Traditional PSO and an Enhanced Particle swarm algorithm.
-Other modifications are explained in the details in
-the next sections.
+Other modifications are explained in the details in the next sections.
 %  \begin{figure}
@@
 \label{Prepross}
-
- Speed and time difference information were widely used in sketch recognition
-systems \cite{earlyprocess}. Figure \ref{fig:orignalStroke} shows an example of
-an input stroke drawn by users.  Figure \ref{fig:speed2Distance} shows the graph
-of time difference and the speed for the stroke in Figure
-\ref{fig:orignalStroke}. It was noted that time difference provides more
-distinct maxima than the speed information. Agar et al. \cite{polygonfeedback31}
-mentions that the pointing device (e.g., mouse) sampling rate is the reason for
-this phenomenon. The points are sampled at regular time intervals as long as the
-pointing device is moving. Therefore, there are no new detected points when the
-pointing device is stationary. This leads to a nearly constant time difference
-between samples while the pen is moving and large difference while the pen is
-stationary (Figure \ref{fig:timediff}). Contrary to time difference information,
-the speed information has a lot of noise in the data because usually  users draw
-with variant speeds (Figure \ref{fig:speed2Distance}).  %Similarly, direction
+  Speed and time difference information were widely used in sketch recognition systems \cite{earlyprocess}.
+Figure \ref{fig:orignalStroke} shows an example of an input stroke drawn by users.
+Figure \ref{fig:speed2Distance} shows the graph of time difference and the speed for the stroke in Figure \ref{fig:orignalStroke}.
+It was noted that time difference provides more distinct maxima than the speed information.
+Agar et al.
+\cite{polygonfeedback31} mentions that the pointing device (e.g., mouse) sampling rate is the reason for this phenomenon.
+The points are sampled at regular time intervals as long as the pointing device is moving.
+Therefore, there are no new detected points when the pointing device is stationary.
+This leads to a nearly constant time difference between samples while the pen is moving and large difference while the pen is stationary (Figure \ref{fig:timediff}).
+Contrary to time difference information, the speed information has a lot of noise in the data because usually  users draw with variant speeds (Figure \ref{fig:speed2Distance}).
+%Similarly, direction
 %information is used as it provides better distinctive maxima for the corners
@@
 
-Dominant points are characterized by low speed values, high curvature and
-direction values. However, the dominant points from the curvature information
-may not be the same points as the dominant points extracted from the speed
-information. Hence, we compute time difference, direction, speed and curvature
-of each point along a stroke to ensure that no dominant point is overlooked in
-this phase. The speed is calculated as $v=\Delta s/\Delta t$ where $t$ is the
-time difference between two points and $s$ is the length between them. The
-direction $d$ calculated as the angle between vectors $\overrightarrow {P_{i -
-1} P_i }$ and the $x-axis$ and the curvature is considered as the change in
-direction $d$ with respect to length $s$ i.e. $c= \Delta d/\Delta s$.
+Dominant points are characterized by low speed values, high curvature and direction values.
+However, the dominant points from the curvature information may not be the same points as the dominant points extracted from the speed information.
+Hence, we compute time difference, direction, speed and curvature of each point along a stroke to ensure that no dominant point is overlooked in this phase.
+The speed is calculated as $v=\Delta s/\Delta t$ where $t$ is the time difference between two points and $s$ is the length between them.
+The direction $d$ calculated as the angle between vectors $\overrightarrow {P_{i - 1} P_i }$ and the $x-axis$ and the curvature is considered as the change in direction $d$ with respect to length $s$ i.e.
+$c= \Delta d/\Delta s$.
 
-After the system computes the speed, time difference and curvature information,
-it proceeds to extracts the points that have low velocity and high curvature.
-However, using simple differentiation to detect local extreme points results in
-false points due to the non smooth curves. Hence, our system adopts a process
-presented by \cite{earlyprocess}, where the mean of each of the four calculated
-information is calculated. This mean is taken as the threshold \textit{th} which
-is used to separate the curve into regions ($R_i$) (Figure\ref{fig:timediff});
-each region $R_i$ is defined as a range of points, where the curve values are
-either above or below the threshold \textit{th}. Those regions are further
-processed to find the maximum point $Max(R_i)$ of each region $R_i$. The stroke
-points $p_i(x,y)$ that correspond to those maximum values are labeled as
-\textit{possible dominant points} $P_{pd}$. The system repeats this process for
-curvature, time difference and speed information. All the points labeled as
-possible dominant points $P_{pd}$ are saved into a single array. Figure
-\ref{fig:ppd999} shows the particles labeled as Possible dominate point $P_{pd}$
-in the preprocessing phase. It is clear that some of the $P_{pd}$ points are
-redundant as specified in the preprocessing stage. % (as shown are redundant)
+After the system computes the speed, time difference and curvature information, it proceeds to extracts the points that have low velocity and high curvature.
+However, using simple differentiation to detect local extreme points results in false points due to the non smooth curves.
+Hence, our system adopts a process presented by \cite{earlyprocess}, where the mean of each of the four calculated information is calculated.
+This mean is taken as the threshold \textit{th} which is used to separate the curve into regions ($R_i$) (Figure\ref{fig:timediff}); each region $R_i$ is defined as a range of points, where the curve values are either above or below the threshold \textit{th}.
+Those regions are further processed to find the maximum point $Max(R_i)$ of each region $R_i$.
+The stroke points $p_i(x,y)$ that correspond to those maximum values are labeled as \textit{possible dominant points} $P_{pd}$.
+The system repeats this process for curvature, time difference and speed information.
+All the points labeled as possible dominant points $P_{pd}$ are saved into a single array.
+Figure \ref{fig:ppd999} shows the particles labeled as Possible dominate point $P_{pd}$ in the preprocessing phase.
+It is clear that some of the $P_{pd}$ points are redundant as specified in the preprocessing stage.
+% (as shown are redundant)
 \begin{figure}
@@
 
- %Speed and time difference information were widely used in sketch understanding systems \cite{earlyprocess}.
-It was noted that time difference provides more distinct maxima than the speed information.
-Agar \cite{polygonfeedback31} mentions that the pointing device (i.e.
-the mouse) sampling rate is the reason for this phenomenon.
-% % In the presented system, we computed time difference, direction, speed and
-% curvature of each point along a stroke. The speed is calculated as $v=\Delta
-% s/\Delta t$ where $t$ is the time difference between two points and $s$ is the
-% length between them. The direction $d$ is calculated as the angle between two
-% vectors and curvature is considered as the change in direction with respect to
-% length i.e. $c= \Delta d/\Delta s$. Dominant points are characterized by lower
-% speed values and higher curvature and direction values.
-
-% After the system computes all the speed, time difference and curvature
-% information it proceeds to detect the points with low velocity and high
-% curvatures. Using simple differentiation to detect local extreme points resulted
-% in false points due to the non smooth curves. Hence, the system adopted a
-% process presented by \cite{earlyprocess}, where the mean of the curve is
-% calculated. Then a threshold \textit{th} is used to separate the curve into
-% regions $Region_i$; each region $Region_i$ is defined as a range of points,
-% where the curve values are either above or below the threshold \textit{th}.
-% Those regions are further processed to find the maximum point $Max(Region_i)$ of
-% each region $Region_i$. The stroke points $p_i(x,y)$ that correspond to those
-% maximum values are labeled as \textit{possible dominant points} $P_{pd}$. The
-% system repeats this process for curvature, time difference and speed
-% information. All the points labeled as possible dominant points $P_{pd}$ are
-% saved into a single array. Figure \ref{fig:LabelsPPD} shows the particles
-% labeled as Possible dominant point $P_{pd}$ by the preprocessing, it is noticed
-% the redundancy of some $P_{pd}$ points. % (as shown are redundant)
-
 \subsection{Segmentation}
@@
 In the segmentation stage the input stroke is divided into a set of primitives.
-As shown in Figure \ref{fig:segblock} first an attempt is made to fit the stroke
-points into a curve or an ellipse using a minimum square error fitting algorithm
-\cite{ellipsefit}. The ellipse fitting step helps to prevent the system from
-over segmenting the stroke into multiple lines or curves if the input stroke is
-an ellipse. If the stroke proved to be an ellipse arc then the segmentation
-process ends and the system proceeds to the next step. Otherwise, the stroke is
-passed to two further segmentation algorithms that divide the stroke to either
-lines or lines and curves. After two different segmentations are generated, the
-system chooses the segmentation with the minimum error. The following sections
-describe in details the ellipse detection algorithm and the two segmentation
-algorithms used to divide the stroke.
+As shown in Figure \ref{fig:segblock} first an attempt is made to fit the stroke points into a curve or an ellipse using a minimum square error fitting algorithm \cite{ellipsefit}.
+The ellipse fitting step helps to prevent the system from over segmenting the stroke into multiple lines or curves if the input stroke is an ellipse.
+If the stroke proved to be an ellipse arc then the segmentation process ends and the system proceeds to the next step.
+Otherwise, the stroke is passed to two further segmentation algorithms that divide the stroke to either lines or lines and curves.
+After two different segmentations are generated, the system chooses the segmentation with the minimum error.
+The following sections describe in details the ellipse detection algorithm and the two segmentation algorithms used to divide the stroke.
 
@@
 \end{figure}
-\subsection{Ellipse fitting}
-
-This process tries to fit the stroke points into an ellipse arc; it starts with
-computing the center of the stroke bounding box. The bounding box center point
-is used as the first estimation of the center of the ellipse. The axes of the
-ellipse are estimated as the $width/2$ and $height/2$ of the stroke bounding box
-where $width$ is width and $height$ is height of the bonding box. The least
-square fitting algorithm \cite{chernov} is used to minimize the fitting error of
-the ellipse Equation (\ref{eq:circleFit})
+\subsection{Ellipse fitting}  This process tries to fit the stroke points into an ellipse arc; it starts with computing the center of the stroke bounding box.
+The bounding box center point is used as the first estimation of the center of the ellipse.
+The axes of the ellipse are estimated as the $width/2$ and $height/2$ of the stroke bounding box where $width$ is width and $height$ is height of the bonding box.
+The least square fitting algorithm \cite{chernov} is used to minimize the fitting error of the ellipse Equation (\ref{eq:circleFit})
 
@@
 \end{equation}
-
- where $N$ is number of points in the stroke, $a,b$ are the length of ellipse
-axes, $x_0$ \& $y_0$ are the coordinates of the center point, $x_i$ \& $y_i$ are
-the coordinates of point $i$ in the stroke. A list of new values for $x_0$ ,
-$y_0$ ,$a$ and $b$ are generated randomly from the older values with small
-increments after each loop.  After few iteration, the final fit error of the
-estimated ellipse is reported. A modification to the system in \cite{mypaper} is
-trying to add another measure to compute the efficiency of the final estimated
-ellipse. Equation \ref{eq:circleError} ensures that the drawn percentage of
-ellipse is considered. This eliminates fitting a line into very large ellipse
-but leaves small ellipses to be fitted as a partial or full ellipse.
+  where $N$ is number of points in the stroke, $a,b$ are the length of ellipse axes, $x_0$ \& $y_0$ are the coordinates of the center point, $x_i$ \& $y_i$ are the coordinates of point $i$ in the stroke.
+A list of new values for $x_0$ , $y_0$ ,$a$ and $b$ are generated randomly from the older values with small increments after each loop.
+After few iteration, the final fit error of the estimated ellipse is reported.
+A modification to the system in \cite{mypaper} is trying to add another measure to compute the efficiency of the final estimated ellipse.
+Equation \ref{eq:circleError} ensures that the drawn percentage of ellipse is considered.
+This eliminates fitting a line into very large ellipse but leaves small ellipses to be fitted as a partial or full ellipse.
 
@@
 \end{equation}
-where $E$ is the error computed by Equation(\ref{eq:circleFit}), $L_{stroke}$
-total length of stroke and $P_{ellipse} $ is the perimeter of the estimated
-ellipse. If $eff$ is more than threshold $th_{Ellipse}$\footnote{By trial and
-error best threshold found was $th_{Ellipse}=0.2$} then the stroke is segmented
-as an ellipse otherwise the system proceeds to the next step.
+where $E$ is the error computed by Equation(\ref{eq:circleFit}), $L_{stroke}$ total length of stroke and $P_{ellipse} $ is the perimeter of the estimated ellipse.
+If $eff$ is more than threshold $th_{Ellipse}$\footnote{By trial and error best threshold found was $th_{Ellipse}=0.2$} then the stroke is segmented as an ellipse otherwise the system proceeds to the next step.
 
@@
 
-Two DPSO algorithms are used to generate different segmentations for each stroke
-the user draws. For each segmentation generated, an error is evaluated. The
-segmentation with the minimum error value is chosen as the best stroke
-segmentation (Figure \ref{fig:segblock}). The problem definition is the same in
-both algorithms but they differ in the method they use to compute fitness and
-error functions.
+Two DPSO algorithms are used to generate different segmentations for each stroke the user draws.
+For each segmentation generated, an error is evaluated.
+The segmentation with the minimum error value is chosen as the best stroke segmentation (Figure \ref{fig:segblock}).
+The problem definition is the same in both algorithms but they differ in the method they use to compute fitness and error functions.
 \begin{description}
 	\item[ Problem definition:] The input stroke $S$  with $N$ points can be
-represented by set $S = \left\{ {x_1 ,x_2  \ldots x_N }\right\}$ where $x_i$ is
-the location of the point $i$. The swarm algorithms consist of $M$ agents which
-are represented by the set $A = \left\{ {P_i \left| {i = 1,2 \cdots M} \right.}
-\right\}$ where $P_i$ is a single solution particle from the solution space.
-Each particle decodes the problem using a binary array with the same length $N$
-as the input stroke.
+represented by set $S = \left\{ {x_1 ,x_2  \ldots x_N }\right\}$ where $x_i$ is the location of the point $i$.
+The swarm algorithms consist of $M$ agents which are represented by the set $A = \left\{ {P_i \left| {i = 1,2 \cdots M} \right.} \right\}$ where $P_i$ is a single solution particle from the solution space.
+Each particle decodes the problem using a binary array with the same length $N$ as the input stroke.
 
@@
 Figure\ref{fig:pso1} shows a particle encoding in the DPSO system.
-
 	\item[Fitness function:] The fitness function and error calculation are
@@
 \begin{equation}
-E=\sum\nolimits_{i = 0}^M e ( \widehat{x_ix_{i+1}},\overline{x_i x_{i+1}})
-\label{eq:ErrorSwarm1}
-\end{equation} where the arc $\widehat{x_ix_j}$ is defined as the consecutive
-set of points from point $x_i$ to point $x_{j}$ as in $x_i,x_{i+1} \cdots,x_j$.
-The line $\overline{x_i x_j} $ as the straight line connecting point $x_i$ to
-point $x_j$, $M$ is the number of dominant points in this solution as generated
-by the swarm algorithm. The error $e ( \widehat{x_ix_j},\overline{x_i x_j})$ is
-computed as the sum of squared perpendicular distance from every point along the
-arc $\widehat{x_ix_j}$ to the line $\overline{x_i x_j}$. The fitness is computed
+E=\sum\nolimits_{i = 0}^M e ( \widehat{x_ix_{i+1}},\overline{x_i x_{i+1}}) \label{eq:ErrorSwarm1} \end{equation} 
+where the arc $\widehat{x_ix_j}$ is defined as the consecutive set of points from point $x_i$ to point $x_{j}$ as in $x_i,x_{i+1} \cdots,x_j$.
+The line $\overline{x_i x_j} $ as the straight line connecting point $x_i$ to point $x_j$, $M$ is the number of dominant points in this solution as generated by the swarm algorithm.
+The error $e ( \widehat{x_ix_j},\overline{x_i x_j})$ is computed as the sum of squared perpendicular distance from every point along the arc $\widehat{x_ix_j}$ to the line $\overline{x_i x_j}$.
+The fitness is computed
 using equation %\ref{eq:fitnessSwarm1}
@@
    {D/\sum\limits_{j = 1}^N {p_{ij} } } & {otherwise}  \\ 
-\end{array}} \right.
-\label{eq:fitnessSwarm1}
-\end{equation} where $N$ is the number of points in the stroke, $D$ is the
-number of points in the solution that was previously labeled as a possible
-dominant point ($P_{pd}$), $E$ is the computed error and $\varepsilon$ is the
-error threshold. It should be noticed that when the error is larger than the
-threshold $\varepsilon$ the fitness is given a -ve value to lower the fitness
-value of the solution. Otherwise the system favors the lower number of vertices.
+   \end{array}} \right. \label{eq:fitnessSwarm1} \end{equation}
+
+   where $N$ is the number of points in the stroke, $D$ is the number of points in the solution that was previously labeled as a possible dominant point ($P_{pd}$), $E$ is the computed error and $\varepsilon$ is the error threshold.
+It should be noticed that when the error is larger than the threshold $\varepsilon$ the fitness is given a -ve value to lower the fitness value of the solution.
+Otherwise the system favors the lower number of vertices.
 
  \subsubsection{\textit{Hybrid Fitting \textsl{AlgS2}}} 
-The algorithm has the same problem formulation but different fitness and error
-functions are used. An attempt is made to fit each segment into a line or
-circular arc. The errors of both circle and line fit estimations are computed
-for each segment $S_i=\widehat{x_ix_j}$. The approximation with the lower error
-value is chosen as the final approximation of this segment
-$S_i$\cite{CruveDivisionSwarm}. The sum of the approximation errors of all
-segments is the total error of the particle.  The total error of the particle is
-computed by equation %\ref{eq:errorSwarm2}
+ The algorithm has the same problem formulation but different fitness and error functions are used.
+An attempt is made to fit each segment into a line or circular arc.
+The errors of both circle and line fit estimations are computed for each segment $S_i=\widehat{x_ix_j}$.
+The approximation with the lower error value is chosen as the final approximation of this segment $S_i$\cite{CruveDivisionSwarm}.
+The sum of the approximation errors of all segments is the total error of the particle.
+The total error of the particle is computed by equation %\ref{eq:errorSwarm2}
  \begin{equation}
@@
 
-As a modification to the system in \cite{mypaper}. After each iteration, of the
-swarm algorithm (\textsl{AlgS1} and \textsl{AlgS2}), each particle is refined
-using the following procedures:
+ As a modification to the system in \cite{mypaper}. After each iteration, of the swarm algorithm (\textsl{AlgS1} and \textsl{AlgS2}), each particle is refined using the following procedures:
 \begin{enumerate}
-	\item For each particle $P_i$ each dominant point $P_{ij}$ is checked to
-find if it was labeled before as a \textit{possible dominant point} $P_{pd}$
-(computed as in section \ref{Prepross}). If it was not labeled the point
-$P_{ij}$ is moved to the nearest labeled point. This ensures that all of the
-points generated by the DPSO are possible dominant points $P_{pd}$.
-	\item The particles are tested to make sure that the distance between
-every two successive dominate point is larger than $min_D$, where $min_D$ is 5\%
-of the total length of the stroke.  Otherwise, one of these points are removed
-based on the error caused by its removal.
-	\item Each two successive segments are test to determine if they can be
-merged into a single segment. If merging the two segments have the same or
-smaller the error, the two segments are merged. This step ensures that the final
-segmentation of the system has minimum number of segments.
+	\item For each particle $P_i$ each dominant point $P_{ij}$ is checked to find if it was labeled before as a \textit{possible dominant point} $P_{pd}$ (computed as in section \ref{Prepross}).
+If it was not labeled the point $P_{ij}$ is moved to the nearest labeled point.
+This ensures that all of the points generated by the DPSO are possible dominant points $P_{pd}$.
+ \item The particles are tested to make sure that the distance between every two successive dominate point is larger than $min_D$, where $min_D$ is 5\% of the total length of the stroke.
+Otherwise, one of these points are removed based on the error caused by its removal.
+ \item Each two successive segments are test to determine if they can be merged into a single segment.
+If merging the two segments have the same or smaller the error, the two segments are merged.
+This step ensures that the final segmentation of the system has minimum number of segments.
 \end{enumerate}
@@
 \label{sec:Recognition}
-After the user draws all strokes of the symbol, the set of un-recognized strokes
-is grouped together along with their segmentation as input to the feature
-extraction process. A composite set of features is used to generate a single
-feature vector. The following list describes each feature in details.
+
+After the user draws all strokes of the symbol, the set of un-recognized strokes is grouped together along with their segmentation as input to the feature extraction process.
+A composite set of features is used to generate a single feature vector.
+The following list describes each feature in details.
 %\begin{description}
+
 \subsubsection{Structural and geometrical Features (FS1)}
@@
 \subsubsection{Rubine Features (FS2)} 
-  These features were introduced by Rubine \cite{gestureexample12} for single
-stroke gestures. Such as length of stroke, cosine of starting angle, smoothness
-of symbol and some features extracted from the bounding box of the symbol.
+  These features were introduced by Rubine \cite{gestureexample12} for single stroke gestures.
+Such as length of stroke, cosine of starting angle, smoothness of symbol and some features extracted from the bounding box of the symbol.
 
@@
 \begin{itemize}
-	\item \emph{Zernike moments:} The zernike moments are orthogonal moments
-which define the distribution of point in the input space. They are invariant to
-both reflection and rotation of the input shape \cite{ZerMomentOrthogonal}. %The
-zernike moment were used in \cite{HeloiseBeautification} t.
+ \item \emph{Zernike moments:} The zernike moments are orthogonal moments which define the distribution of point in the input space.
+They are invariant to both reflection and rotation of the input shape \cite{ZerMomentOrthogonal}.
+%The zernike moment were used in \cite{HeloiseBeautification} t.
 \end{itemize}
@@
   \end{itemize} 
-
- After computing the features, the symbol is introduced to the classifier as a
-feature vector. The system uses Support Vector Machine (SVM) classifier with
-Linear kernel \cite{libsvm}. An OVO classifier structure (object versus object)
-is used to handle multi-class problem.% A validation algorithm is used to choose
-the Kernel variables parameters $c, \gamma$.
+  After computing the features, the symbol is introduced to the classifier as a feature vector.
+The system uses Support Vector Machine (SVM) classifier with Linear kernel \cite{libsvm}.
+An OVO classifier structure (object versus object) is used to handle multi-class problem.% A validation algorithm is used to choose the Kernel variables parameters $c, \gamma$.
 
@@
  \subsection{Swarm Algorithms Experiments}
-Another set of tests were performed on the system to enhance non-ellipse fitting
-algorithms. The effect of the number of particles, the error thresholds and
-various other parameters were investigated. The results of these tests are shown
-in Figures \ref{fig:swarmtesting} and   \ref{fig:swarmtesting2}. Figure
-\ref{fig:swarmtesting} shows that as the number of swarm iterations increase the
-error value decrease (error of the segmentation algorithm). The final number of
-particles and maximum swarm iterations used in the system was based on a
-tradeoff between error achieved versus the computation time. The effect of other
-algorithm parameters such as $c_1$, $\omega$ (mentioned in
-Section\ref{sec:ParticleSwarmAlgorithm}) is tested using similar tests (Figure
-\ref{fig:swarmtesting2},\ref{fig:swarmwtest}).%, $c_2$
+
+ Another set of tests were performed on the system to enhance non-ellipse fitting algorithms.
+The effect of the number of particles, the error thresholds and various other parameters were investigated.
+The results of these tests are shown in Figures \ref{fig:swarmtesting} and   \ref{fig:swarmtesting2}.
+Figure \ref{fig:swarmtesting} shows that as the number of swarm iterations increase the error value decrease (error of the segmentation algorithm).
+The final number of particles and maximum swarm iterations used in the system was based on a tradeoff between error achieved versus the computation time.
+The effect of other algorithm parameters such as $c_1$, $\omega$ (mentioned in Section\ref{sec:ParticleSwarmAlgorithm}) is tested using similar tests (Figure \ref{fig:swarmtesting2},\ref{fig:swarmwtest}).%, $c_2$
 
@@
 
-This paper used an Enhanced-PSO to modify a new sketch recognition using DPSO
-represented in \cite{mypaper}. Results show that enhancing the solutions that
-the \textit{PSO} in general generates an optimized stroke segmentation which
-improves the final recognition rate.  The final recognition improved by
-modifying the particle optimization algorithm without depending on the higher
-level features. The experiments evaluated two \textit{DPSO} different
-segmentation algorithms (AlgS1 and AlgS2) on simple presentation set dataset.
-Results show that the enhanced PSO algorithm improves the segmentation on both
-algorithms which achieves better segmentation and final accuracy. The results
-proved that the Enhancement made on the PSO improves the efficiency of the
-Traditional PSO algorithm without being dependable on special objective function
-or application. The system achieved an average overall improvement of more than
-1\% over Basic PSO.
+This paper used an Enhanced-PSO to modify a new sketch recognition using DPSO represented in \cite{mypaper}.
+Results show that enhancing the solutions that the \textit{PSO} in general generates an optimized stroke segmentation which improves the final recognition rate.
+The final recognition improved by modifying the particle optimization algorithm without depending on the higher level features.
+The experiments evaluated two \textit{DPSO} different segmentation algorithms (AlgS1 and AlgS2) on simple presentation set dataset.
+Results show that the enhanced PSO algorithm improves the segmentation on both algorithms which achieves better segmentation and final accuracy.
+The results proved that the Enhancement made on the PSO improves the efficiency of the Traditional PSO algorithm without being dependable on special objective function or application.
+The system achieved an average overall improvement of more than 1\% over Basic PSO.
 
-This paper presented a new approach to sketch recognition using DPSO. The system
-uses multiple preliminary information, i.e., speed and curvature information, as
-input to the \textit{DPSO}. Results show that the \textit{DPSO} in general
-generates an optimized stroke segmentation which improves the final recognition
-rate.  The final recognition is based on both global shape and hierarchical
-stroke properties. Adding stroke and segment geometrical features improved the
-final recognition. Global shape and moment properties were used to prevent the
-need to use graph matching which is computationally expensive
-\cite{SketchRead2007}. The experiments evaluated two \textit{DPSO} different
-segmentation algorithms (AlgS1 and AlgS2) on three different datasets (Hs-Ds,
-EL-Ds and LD-Ds). Results show that AlgS2 achieves better segmentation and final
-accuracy than Alg3 \cite{earlyprocess} on the datasets used. The results proved
-that the system is easily expandable to more and more symbols as it does not
-depend on low level primitive recognizers but on high level symbol and
-geometrical features.  The system achieved an average overall improvement of
-more than 2\% over Alg3 in all datasets.
+ This paper presented a new approach to sketch recognition using DPSO.
+The system uses multiple preliminary information, i.e., speed and curvature information, as input to the \textit{DPSO}.
+Results show that the \textit{DPSO} in general generates an optimized stroke segmentation which improves the final recognition rate.
+The final recognition is based on both global shape and hierarchical stroke properties.
+Adding stroke and segment geometrical features improved the final recognition.
+Global shape and moment properties were used to prevent the need to use graph matching which is computationally expensive \cite{SketchRead2007}.
+The experiments evaluated two \textit{DPSO} different segmentation algorithms (AlgS1 and AlgS2) on three different datasets (Hs-Ds, EL-Ds and LD-Ds).
+Results show that AlgS2 achieves better segmentation and final accuracy than Alg3 \cite{earlyprocess} on the datasets used.
+The results proved that the system is easily expandable to more and more symbols as it does not depend on low level primitive recognizers but on high level symbol and geometrical features.
+The system achieved an average overall improvement of more than 2\% over Alg3 in all datasets.
 
-  The tradeoff between accuracy achieved and time complexity can be further
-investigated to achieve better results. Currently, the system processes one
-symbol at a time, an enhancement would be to allow users to draw more than one
-symbol at a time and incorporate a method for separating symbols. Thus, possible
-extension of this research is to complete the clustering algorithm for fully
-automated sketch recognition. Another area of enhancements is the features
-extraction methods. Introducing more spatial and geometrical features is
-believed to improve classifications. Features that represent the appearance of
-the symbols are likely to improve the recognition of symbols that have similar
-geometrical structure (i.e. Nor and OR gate in logic symbols).
+   The tradeoff between accuracy achieved and time complexity can be further investigated to achieve better results.
+Currently, the system processes one symbol at a time, an enhancement would be to allow users to draw more than one symbol at a time and incorporate a method for separating symbols.
+Thus, possible extension of this research is to complete the clustering algorithm for fully automated sketch recognition.
+Another area of enhancements is the features extraction methods.
+Introducing more spatial and geometrical features is believed to improve classifications.
+Features that represent the appearance of the symbols are likely to improve the recognition of symbols that have similar geometrical structure (i.e.
+Nor and OR gate in logic symbols).
 
- For the future work, it is suggested that we continue to improve both the
-enhanced PSO segmentation algorithm. Also, a possible extension of this research
-is to improve the recognition algorithm of the Sketch Recognition system. Other
-area of enhancements is the features extraction methods. Introducing more
-spatial and geometrical features is believed to improve classifications.
+ For the future work, it is suggested that we continue to improve both the enhanced PSO segmentation algorithm.
+Also, a possible extension of this research is to improve the recognition algorithm of the Sketch Recognition system.
+Other area of enhancements is the features extraction methods.
+Introducing more spatial and geometrical features is believed to improve classifications.
 

### 2010-12-11 MyCv/Resyne.tex
adding my cv to the public repository

% Generated by GrindEQ Word-to-LaTeX 2010
% ========== UNREGISTERED! ========== Please register! ==========
% LaTeX/AMS-LaTeX

\documentclass{article}

%%% remove comment delimiter ('%') and specify encoding parameter if required,
%%% see TeX documentation for additional info (cp1252-Western,cp1251-Cyrillic)
%\usepackage[cp1252]{inputenc}

%%% remove comment delimiter ('%') and select language if required
%\usepackage[english,spanish]{babel}

\usepackage{amssymb}
\usepackage{amsmath}
\usepackage[dvips]{graphicx}
%%% remove comment delimiter ('%') and specify parameters if required
%\usepackage[dvips]{graphics}
\linespread{1}

\pdfpagewidth 8.5in
\pdfpageheight 11in

\setlength\topmargin{0in}
\setlength\headheight{0in}
\setlength\headsep{-0.25in}
\setlength\textheight{10.0in}
\setlength\textwidth{6.5in}
\setlength\oddsidemargin{-0.25in}
\setlength\evensidemargin{-0.25in}
\setlength\parindent{0in}
\setlength\parskip{0in}

\title { Resume \\ Maha Mohammed Nabeel El Meseery}
\date{}
\begin{document}

%%% remove comment delimiter ('%') and select language if required
%\selectlanguage{spanish}
\maketitle
\section*{Personal Information:}
\noindent
 \textbf{ Address: } 2nd Tahreer street, Dokki, Giza 12615, Egypt.

 \textbf{ Telephone:} (+2012)112- 5435, (+202)3336-8292.

\textbf{  Email: } melmeseery@gmail.com, melmeseery@aucegypt.edu.
 \section*{Education:}
  \begin{description}
 \item [2004-2009]  \textbf{ Msc.E., Cairo University, Computer Engineering Department, Giza, Egypt}\\
 Thesis Area:\textbf{\textit{Sketch Recognition And Understanding}.}\\
 Thesis Supervisors: \textit{Prof. Nevin Darwish (Cairo University), Dr Magda
 Fayek (Cairo University), Prof. Samia Mashali (Electrical Research Institute),
 Mahmoud Fakr El Din (Electrical Research Institute).}
 \item[1998-2003]  \textbf{ Bsc.E., Cairo University, Computer Engineering Department Giza, Egypt}\\
  Final Grade: \textbf{Very Good with honors.}\\
  Graduation project: \textit{Personal Face identification System,} \textbf{Distinction.}\\
  Project Supervisors: \textit{Dr. Madga Fayek.}
\item [1995-1998]  \textbf{General Secondary School Certificate,   Manar
 Language School. }
  \end{description}

 \section*{Experience:}
 %\textbf{\underbar{Experience:}}
 \textbf{August 2008- present   [part time]}\\
 \textbf{Research and Teaching Assistant}, American University in Cairo, Electrical Engineering Department.
 The following is summary of my achievements:
 \begin{enumerate}
   \item Conducted and helped develop research in Online Arabic Handwriting Recognition. This includes implementing various features extraction and classification methods.
   \item Assisted in collecting, gathering and analyzing an Online Arabic handwriting Dataset.
   \item Assisted and developed research in area of Arabic and Latin digit Recognition.
 \item Implementing different features extraction and classification methods for Arabic and Latin digits and comparing their performance.
 \item Read, analyzed and reviewed scientific and research papers.
 \item Wrote and edited research paper in topics related to Arabic Handwriting Recognition.
 \item Guided students in their studies in Lab projects and programming assignments.
 \item Conducted Tutorials and problem sessions for students in various subjects.
\end{enumerate}

 \textbf{January 2007- present}\\
 \textbf{Research Assistant}, Electrical Research Institute, Egypt
 \begin{enumerate}
\item Worked as a member of the Signal Processing Lab in the Computer and Systems Department.
 \item Assisted in the research project undergoing in the department.
 \item Helped in writing research papers on various topics.
 \item Implemented and develop different algorithms in field of Image Processing.
 \item Provided a technical support for other researchers especially in field of JavaEE programming.
\end{enumerate}

 \textbf{September 2005- September 2008 [part time from January 2007] }\\
 \textbf{Senior Software Engineer}, EngNet Company\textbf{}
 The following is summary of my achievements:
 \begin{enumerate}

\item Introduced new software procedures in the company to enhance developers performance.
 \item Analyzed and designed various software projects using UML and software engineering concepts.
 \item Wrote the software requirements and specifications of software
 application.
 \item Supervised and managed the development process for a team consists of four developers and two quality engineers.
 \item Assisted junior programmers in development using Java, JSP, Struts, RCP and MySQL.
\end{enumerate}

 \textbf{September 2003- September 2005}\\
 \textbf{Teaching Assistant}, Faculty of Computer Science, MSA University, Giza, Egypt.\textbf{}
 \begin{enumerate}
\item Developing lab manuals and leading lab experiments for Software Engineering Methodologies, Object Oriented Programming, Digital Logic Design, Computer Graphics, Calculus, and Introduction to IT.
 \item Guide students in their studies in Data Structure and algorithm, Object Oriented Programming, Software Engineering, Digital Logic Design, Computer Graphics, and Calculus.
 \item Assist and supervise student's projects and assignments in subjects mentioned.
 \item Assistant supervisor for senior student's projects (Mainly graduation project using ASP.Net, C\#) .
 \item Developing and delivering short courses in C++, C\#, UML, and ASP.NET.
\end{enumerate}

 \section*{Publications}
 \begin{enumerate}
  \item Online Library of Scientific Models - A New Way to Teach, Learn, and
Share Learning Experience, Elrefaei Hatem, Aboelfadel Tamer, Elmeseery Maha,
Elmofty Ali, Shafee Marwa, in Proceedings of the International Conference of "Interactive computer aided learning" ICL2007, Villach, Austria.\textbf{\underbar{}}
  \item El Meseery, Maha, Fakhr El Din, Mahmoud., Mashali, Samia., Fayek, Magda.
 \& Darwish, Nevin., Sketch Recognition Using Particle Swarm Algorithms,  IEEE International Conference on Image Processing, November,  2009.
  \item El Meseery, Sketch Recognition and Understanding,  Master Thesis,
  Faculty of Engineering, Computer Engineering Department, Cairo University,
  November, 2009.
  \item El Meseery, Maha, Fakhr El Din, Mahmoud., Tarek abouldahab, Sketch
  Recognition using Enhanced Particle Swarm Optimization, ICGST International Journal on Artificial Intelligence and Machine Learning  2010, (Under Review).
\end{enumerate}
 \section*{Awards:}
  \begin{enumerate}
    \item \textbf{1999} \textit{Excellence Reward} Faculty of Engineering.
    (The reward is given to students who attained a result of 'Very Good' each year).
    \item \textbf{2000} \textit{Excellence Reward}
    \item \textbf{2001} \textit{Excellence Reward}
    \item \textbf{2002} \textit{Excellence Reward}
    \item \textbf{2003} \textit{Excellence Reward}
  \end{enumerate}
 \section*{Computer Skills:}
 \begin{enumerate}
\item Environments: Linux (openSuse) and Windows.
 \item Programming: Java, JSP, Sturts, AJAX, .Net,C/C++, C\#,  ASP.NET, ASP, ADO.Net, ADO, VC6,  XML  Visual Basic, Embedded C, Assembly, Pascal, VHDL, Prolog, Lisp, Matlab, latex.
 \item Database Tools: SQL server 2000, MySql, Oracle and developer 6i.
 \item Analysis and Modeling: UML, Hardware design.
\end{enumerate}

 \section*{Training Courses and Summer jobs:}
 \begin{enumerate}
\item Human Development, Zedny Training Center, Cairo, Egypt (January - March 2005).
 \item MS SQL 2000, MSA University, Giza, Egypt (July 2004).
 \item C\# dot Net and ADO dot Net, MSA University, Giza, Egypt (August 2004).
 \item Web Development (ASP, XML, DOM, DHTML and ADO), Harf Information Technology. Cairo, Egypt (July --September 2002).
 \item DB2, JSP, Java beans, Servelts and IBM Web Sphere, Cairo University, Giza, Egypt (July- August 2001).
 \item Java applets, Training on faculty to develop an e-version of the calculus book taught to preparatory year student of engineering faculty (August -- September 2000).
 \item MFC and VC6, Cairo University, Giza, Egypt (July 2000).
\end{enumerate}

 \section*{Projects:}

 \begin{enumerate}
\item 2010-Present, Online Handwriting project; a project to implement an Online Arabic handwriting system.
 \item 2009-2010, Arabic Digit Recognition project; a project to implement an Arabic Digit Recognition system with high accuracy and speed.
\item 2007-2008, Schedulizier project, a project to schedule and manage information of training companies ( Engnet Co.).
 \item 2006-2007, Scientific SDK, a project to help engineers create scientific simulations by automatically generating the user interfaces code for them.
 \item 2005-2006, Imodelit library scientific Microwave java applet library ( http:\textbackslash \textbackslash www.imodelit.com. Engnet Co.).
 \item 2004, Mouth detection and segmentation, using C++ (Pre Master Project).
 \item 2004, Traffic Control System using micro-controllers (Pre Master
 Project).
 \item 2003, Using Neural Network to recognize English Characters, using Matlab.
 \item 2003, Keyless Entry System has keypad to enter a password, display to display messages to the user automatic lock and unlock, using assembly of 80468.
 \item 2003, Implementing a Processor, using VHDL.
 \item 2002, Simulation of a single processor operating system, using Java.
 \item 2001, Developing a Database for a social security accounting office, using Oracle Database.
 \item 2001, Water Filter using sensors.
 \item 2000, Implementing a snake Game, using Pascal.
\end{enumerate}

 \section*{Languages:}

Arabic (Native), English (Fluent), French (poor).

 \section*{Social Activities and Professional memberships:}

\begin{enumerate}
\item  Student member of the Institute of Electrical and Electronics Engineers
(IEEE) since 2002.
\item  Member of the Egyptian Engineering Syndicate (ESE) since 2003.
\item  Member of \textit{Resala} Charitable Society since 1999, responsible for
the following activities:
\begin{enumerate}
\item  Carrying out the big sister role for orphan girls. This includes the following
  \begin{enumerate}
\item Helping developing better education for \textit{Resala} organization
orphans.
\item Acting as guidance for behavioural, social and mental aspect of orphaned
teenagers.
\item Attend behaviour modification, social interaction with teenagers.
\item Give teenagers educational lessons to improve their quality of educations.
\end{enumerate}
\item  Organizing blood donations campaigns.
\item  Delivering free-charge courses in IT (C, C++, VC6, Windows and office ).
\item  Helping in writing the contents of tutorial for the various courses given at Resala ITcenter.
\end{enumerate}
\item  Member of El-Hosry Charitable Society since 2003.
\item  Member of Egyptian Army clubs.
\item  Member of shooting club, Dokki.
\item  Enjoy reading, swimming, walking and watching movies.
\end{enumerate}

\section*{References}
\noindent
\textbf{Dr. Magda Fayek}
 Address: Computer Engineering Dept., Faculty of Engineering, Cairo University, Giza, Egypt.
\\ \textbf{Email: }magdafayek@eng.cu.edu.eg.

\textbf{Dr. Hatem El refaei}
 Address: Physics and Mathematics Dept., Faculty of Engineering, Ain Shams University, Cairo, Egypt.
 \\ \textbf{Email: }hatem@mcit.gov.eg.

\textbf{Dr. Sherif Abdel Azeem}
 Address: Electrical Engineering Dept., School of Science and Engineering,
 American University in Cairo, New Cairo, Cairo, Egypt.\\ \textbf{Email: }
 shazeem@aucegypt.edu.

 \textbf{Dr Mahmoud Fakhr El Deen}
 Address: Computer and Systems Dept., Electronic Research Institute,
 National Research Center, Tahreer Street, Dokki, Cairo, Egypt.\\ \textbf{Email:
 } mafakhr@mcit.gov.eg.

\end{document}

% == UNREGISTERED! == GrindEQ Word-to-LaTeX 2010 ==

### 2011-01-11 trunk/IntroductionLetters/document.tex
Inial import

%%This is a very basic article template.
%%There is just one section and two subsections.
\documentclass{article}

\begin{document}

\section{Title}

Hi Frank,
I graduated from the Faculty of Engineering Cairo University, in July 2003. I would like to  investigate the possibility of continuing my PhD studies under your supervision at the University of British Columbia. I obtained my Bsc and Msc in Computer Engineering from  Cairo University, Egypt in  2003 and 2009 respectively.  I worked both in the academic and industrial sectors (mention the position).  I am planning (decided) to pursue a research career by first continuing my Post Graduate studies in a respectful university.
I work at Electronics Research Institute in the 'Signal and Image Processing' research group. My work focuses on conducting research and related activities in the field of Machine learning and Image processing. I also work as a part time Research and Teaching Assistant in American University in Cairo (AUC).  My research in AUC focuses on Arabic Handwriting Recognition especially improving interaction and recognition of Online Arabic handwriting systems.
(put this paragraph first)Even though all my research experience is in Machine learning and Image processing fields, I have vast interest in Software Engineering.   I worked for four years as a Software Engineer for Engnet Company, Egypt. I was responsible for analysis of software requirements and the design of the software systems. After the first year, I was promoted to Team Leader and Senior Software Engineer where I got more involved in software process and how to improve the performance of developers. Since then, Software Engineering became my passion; I spent a lot of time trying to improve (Make it as a task) software development process, helping developers for better collaborative work and looking for methods that will improve issue and bug tracking systems.  I read  several  technical papers on software engineering, bug tracking, team collaboration and code analysis systems.  Because of my passion I decided to pursue my studies in PhD in Software Engineering field.
I am currently planning for my PhD studies. I have preference towards studying in Canada where state of the art research exists and particularly in the University of British Columbia due to its international recognition as one of the best Universities in Canada. My research interests are mining code and bug repositories (Be specific about the project name..but I am open to any idea in the field). Also I am interested in code analysis and software team collaboration which I hope falls under your research interests as indicated in the website of University of British Columbia (be specific about the website if it is a group or if its is personal website)

I am wondering about the possibility of pursuing my graduate studies under your  guidance and supervision at University of British Columbia as from the fall of 2011. I write to seek your kind advice.

\end{document}

### 2011-01-11 MyCv/Resyne.tex
after editing the CV

@@
 \setlength\headsep{-0.25in}
-\setlength\textheight{10.0in}
+\setlength\textheight{9.8in}
 \setlength\textwidth{6.5in}
@@
 \maketitle
-\section*{Personal Information:}
-\noindent
- \textbf{ Address: } 2nd Tahreer street, Dokki, Giza 12615, Egypt.
 
- \textbf{ Telephone:} (+2012)112- 5435, (+202)3336-8292.
-
-\textbf{  Email: } melmeseery@gmail.com, melmeseery@aucegypt.edu.
  \section*{Education:}
@@
   Project Supervisors: \textit{Dr. Madga Fayek.}
-\item [1995-1998]  \textbf{General Secondary School Certificate,   Manar
- Language School. }
+
   \end{description}
@@
  \textbf{August 2008- present   [part time]}\\
- \textbf{Research and Teaching Assistant}, American University in Cairo, Electrical Engineering Department.
- The following is summary of my achievements:
+ \textbf{Research and Teaching Assistant}, American University in Cairo, Egypt,
+ Electrical Engineering Department.
+
  \begin{enumerate}
-   \item Conducted and helped develop research in Online Arabic Handwriting Recognition. This includes implementing various features extraction and classification methods.
-   \item Assisted in collecting, gathering and analyzing an Online Arabic handwriting Dataset.
-   \item Assisted and developed research in area of Arabic and Latin digit Recognition.
+   \item Conducting and helping develop research in Online Arabic Handwriting
+   Recognition. This includes implementing various features extraction and classification methods.
+   \item Assisting in collecting, gathering and analyzing an Online Arabic
+   handwriting Dataset.
+%   \item Assisting and developed research in area of Arabic and Latin digit
+   Recognition.
  \item Implementing different features extraction and classification methods for Arabic and Latin digits and comparing their performance.
- \item Read, analyzed and reviewed scientific and research papers.
- \item Wrote and edited research paper in topics related to Arabic Handwriting Recognition.
- \item Guided students in their studies in Lab projects and programming assignments.
- \item Conducted Tutorials and problem sessions for students in various subjects.
+ \item Reading, analyzing and reviewing scientific and research papers.
+ \item Guiding students in their studies in Lab projects and programming
+ assignments.
+ \item Conducting tutorials and problem sessions for students in various
+ subjects.
 \end{enumerate}
@@
  \begin{enumerate}
-\item Worked as a member of the Signal Processing Lab in the Computer and Systems Department.
- \item Assisted in the research project undergoing in the department.
- \item Helped in writing research papers on various topics.
- \item Implemented and develop different algorithms in field of Image Processing.
- \item Provided a technical support for other researchers especially in field of JavaEE programming.
+\item Working as a member of the Signal Processing Lab in the Computer and
+Systems Department.
+ \item Assisting in the research project undergoing in the department.
+ \item Implementing and developing different algorithms in field of Image
+ Processing.
+ \item Providing a technical support for other researchers especially in field
+ of JavaEE programming.
 \end{enumerate}
@@
  \textbf{September 2005- September 2008 [part time from January 2007] }\\
- \textbf{Senior Software Engineer}, EngNet Company\textbf{}
+ \textbf{Senior Software Engineer}, EngNet, Egypt\textbf{}
  The following is summary of my achievements:
@@
  \begin{enumerate}
-\item Developing lab manuals and leading lab experiments for Software Engineering Methodologies, Object Oriented Programming, Digital Logic Design, Computer Graphics, Calculus, and Introduction to IT.
- \item Guide students in their studies in Data Structure and algorithm, Object Oriented Programming, Software Engineering, Digital Logic Design, Computer Graphics, and Calculus.
- \item Assist and supervise student's projects and assignments in subjects mentioned.
+\item Developed lab manuals and leading lab experiments for Software Engineering
+Methodologies, Object Oriented Programming, Digital Logic Design, Computer Graphics, Calculus, and Introduction to IT.
  \item Assistant supervisor for senior student's projects (Mainly graduation project using ASP.Net, C\#) .
@@
 
- \section*{Publications}
+ \section*{Publications:}
  \begin{enumerate}
@@
   \item El Meseery, Maha, Fakhr El Din, Mahmoud., Mashali, Samia., Fayek, Magda.
- \& Darwish, Nevin., Sketch Recognition Using Particle Swarm Algorithms,  IEEE International Conference on Image Processing, November,  2009.
+ \& Darwish, Nevin., Sketch Recognition Using Particle Swarm Algorithms,  IEEE
+ International Conference on Image Processing, Cairo, Egypt, November,  2009.
   \item El Meseery, Sketch Recognition and Understanding,  Master Thesis,
-  Faculty of Engineering, Computer Engineering Department, Cairo University,
-  November, 2009.
-  \item El Meseery, Maha, Fakhr El Din, Mahmoud., Tarek abouldahab, Sketch
-  Recognition using Enhanced Particle Swarm Optimization, ICGST International Journal on Artificial Intelligence and Machine Learning  2010, (Under Review).
+  Faculty of Engineering, Computer Engineering Department, Cairo
+  University, Egypt, November, 2009.
+  \item El Meseery, Maha, Fakhr El Din, Mahmoud, Tarek abouldahab, Sketch
+  Recognition using Enhanced Particle Swarm Optimization, (Submitted to)Pattern
+  Recognition Letters, 2010, (17 pages).
 \end{enumerate}
@@
   \begin{enumerate}
-    \item \textbf{1999} \textit{Excellence Reward} Faculty of Engineering.
-    (The reward is given to students who attained a result of 'Very Good' each year).
-    \item \textbf{2000} \textit{Excellence Reward}
-    \item \textbf{2001} \textit{Excellence Reward}
-    \item \textbf{2002} \textit{Excellence Reward}
-    \item \textbf{2003} \textit{Excellence Reward}
+    \item \textbf{1999-2003} \textit{Excellence Reward} Faculty of Engineering,
+    Cairo, Egypt.
+
   \end{enumerate}
@@
  \begin{enumerate}
-\item 2010-Present, Online Handwriting project; a project to implement an Online Arabic handwriting system.
- \item 2009-2010, Arabic Digit Recognition project; a project to implement an Arabic Digit Recognition system with high accuracy and speed.
-\item 2007-2008, Schedulizier project, a project to schedule and manage information of training companies ( Engnet Co.).
- \item 2006-2007, Scientific SDK, a project to help engineers create scientific simulations by automatically generating the user interfaces code for them.
- \item 2005-2006, Imodelit library scientific Microwave java applet library ( http:\textbackslash \textbackslash www.imodelit.com. Engnet Co.).
- \item 2004, Mouth detection and segmentation, using C++ (Pre Master Project).
- \item 2004, Traffic Control System using micro-controllers (Pre Master
+\item 2010-Present, Online Handwriting project.
+ \item 2009-2010, Arabic Digit Recognition project.
+\item 2007-2008, Schedulizier: a project to schedule and manage information of
+training companies ( EngNet, Egypt.).
+ \item 2006-2007, Scientific SDK: a project to help engineers create scientific
+ simulations by automatically generating the user interface for them.( EngNet,
+ Egypt)
+ \item 2005-2006, Imodelit : a library of scientific Microwave java applets (
+ EngNet, Egypt).
+ \item 2004, Mouth detection and segmentation using C++ (Master's Project).
+ \item 2004, Traffic Control System using micro-controllers (Master's
  Project).
- \item 2003, Using Neural Network to recognize English Characters, using Matlab.
- \item 2003, Keyless Entry System has keypad to enter a password, display to display messages to the user automatic lock and unlock, using assembly of 80468.
- \item 2003, Implementing a Processor, using VHDL.
- \item 2002, Simulation of a single processor operating system, using Java.
- \item 2001, Developing a Database for a social security accounting office, using Oracle Database.
- \item 2001, Water Filter using sensors.
- \item 2000, Implementing a snake Game, using Pascal.
+ \item 2003, Using Neural Network to recognize English Characters using Matlab.
+ \item 2003, Keyless Entry System using 80468 assembly.
+ \item 2003, Implementing a Processor using VHDL.
+ \item 2002, Simulation of a single processor operating system using Java.
+ \item 2001, Developing a Database for a social security accounting office using Oracle Database.
+ \item 2001, Water Filter (Hardware project).
+ \item 2000, Snake Game using Pascal.
 \end{enumerate}
@@
   \begin{enumerate}
-\item Helping developing better education for \textit{Resala} organization
-orphans.
-\item Acting as guidance for behavioural, social and mental aspect of orphaned
+\item Helping developing better education procedures for \textit{Resala}
+organization orphans.
+\item Attend courses on behaviour modification and  social interaction for
 teenagers.
-\item Attend behaviour modification, social interaction with teenagers.
-\item Give teenagers educational lessons to improve their quality of educations.
+\item Acting as guidance for behavioural, social and mental aspect for orphaned
+teenagers.
+
+\item Teach teenagers educational courses in different subjects.
 \end{enumerate}
@@
 \end{enumerate}
-\item  Member of El-Hosry Charitable Society since 2003.
-\item  Member of Egyptian Army clubs.
-\item  Member of shooting club, Dokki.
+\item  Member of Egyptian shooting clubs.
 \item  Enjoy reading, swimming, walking and watching movies.
 \end{enumerate}
+\newpage
+\section*{Personal Information:}
+\noindent
+ \textbf{ Address: } $2^{nd}$ Tahreer street, Dokki, Giza 12615, Egypt.
 
-\section*{References}
-\noindent
-\textbf{Dr. Magda Fayek}
- Address: Computer Engineering Dept., Faculty of Engineering, Cairo University, Giza, Egypt.
-\\ \textbf{Email: }magdafayek@eng.cu.edu.eg.
+ \textbf{ Telephone:} (+2012)112- 5435, (+202)3336-8292.
 
-\textbf{Dr. Hatem El refaei}
- Address: Physics and Mathematics Dept., Faculty of Engineering, Ain Shams University, Cairo, Egypt.
- \\ \textbf{Email: }hatem@mcit.gov.eg.
+\textbf{  Email: }  melmeseery@aucegypt.edu
 
-\textbf{Dr. Sherif Abdel Azeem}
- Address: Electrical Engineering Dept., School of Science and Engineering,
- American University in Cairo, New Cairo, Cairo, Egypt.\\ \textbf{Email: }
- shazeem@aucegypt.edu.
+\section*{References:}
+%\noindent
+\begin{description}
 
- \textbf{Dr Mahmoud Fakhr El Deen}
- Address: Computer and Systems Dept., Electronic Research Institute,
- National Research Center, Tahreer Street, Dokki, Cairo, Egypt.\\ \textbf{Email:
- } mafakhr@mcit.gov.eg.
+\item[\textbf{Dr. Magda Fayek}] (Professor)
 
+\begin{description}
+  \item[Address:]Computer Engineering Dept., Faculty of Engineering, Cairo University, Giza, Egypt.
+  \item[Email:] magdafayek@eng.cu.edu.eg.
+%\\\textbf{Email: }
+\end{description}
+\item[\textbf{Dr. Hatem El refaei}] (Associate Professor)
+
+\begin{description}
+  \item[Address:]Physics and Mathematics Dept., Faculty of Engineering, Ain Shams University, Cairo, Egypt.
+  \item[Email:]hatem@mcit.gov.eg.
+\end{description}
+
+\item[\textbf{Dr. Sherif Abdel Azeem}] (Associate Professor)
+
+\begin{description}
+  \item[Address:]Electrical Engineering Dept., School of Science and Engineering,American University in Cairo, New Cairo, Cairo, Egypt.
+  \item[Email:]  shazeem@aucegypt.edu.
+\end{description}
+
+\item[ \textbf{Dr. Mahmoud Fakhr El Deen}] (Researcher)
+
+ \begin{description}
+  \item[Address:]Computer and Systems Dept., Electronic Research Institute,
+  National Research Center, Tahreer Street, Dokki, Cairo, Egypt.
+  \item[Email:]mafakhr@mcit.gov.eg.
+\end{description}
+\end{description}
 \end{document}

### 2011-01-11 IntroductionLetters/document.tex
adding dinas comment

%%This is a very basic article template.
%%There is just one section and two subsections.
\documentclass{article}

\begin{document}

\section{Title}

Hi Frank,

I would like to  investigate the possibility of continuing my PhD studies under
your supervision at the University of Calgary.  My name is Maha and I obtained
my Bsc and Msc in Computer Engineering from  Cairo University, Egypt in  2003
and 2009 respectively.  I worked both in the academic (as a Research Assistant)
and industrial (as a Software Engineer) sectors . I am planning to pursue a
research career by first continuing my Post Graduate studies. I am specially
intereseted in. ..  .  .  .  .   .  .  .  .  .  .

Most my research experience is in Machine Learnning but I have vast interest in
Software Engineering. I worked for four years as a Software Engineer for Engnet,
Egypt. After the first year, I was promoted to Team Leader and Senior Software
Engineer where I got more involved in software process and how to improve the
performance of developers. Since then, Software Engineering became my passion;
I took charge of improving the software development process,
helping developers for better collaborative work and looking for methods that will improve issue and bug tracking systems.
     I read  several  technical papers on software engineering, bug tracking,
     team collaboration and code analysis systems.  Because of my passion I
     decided to pursue my studies in PhD in Software Engineering field.

Currently, I work at Electronics Research Institute in the 'Signal and Image
Processing' research group. My work focuses on conducting research and related activities in the field of Machine learning and Image processing. I also work as a part time Research and Teaching Assistant in American University in Cairo (AUC).
  My research in AUC focuses on Arabic Handwriting Recognition especially improving interaction and recognition of Online Arabic handwriting systems.

     I have preference towards studying in Canada where state of the art
     research exists and particularly in the University of Calgary due to its
     international recognition as one of the best Universities in Canada. My
     research interests are   ..     .  .  .  .  . .  .  . .. .. but I am open
     to any idea in the field. Also I am interested in .. . .   .  . . .. . . .. .
     I hope falls under your research interests as indicated in your
     group website.

I am wondering about the possibility of pursuing my graduate studies under your
guidance and supervision at University of Calgary. I
write to seek your kind advice.

\end{document}

### 2011-01-11 trunk/IntroductionLetters/document.tex

@@
 Hi Frank,
-I graduated from the Faculty of Engineering Cairo University, in July 2003.
-I would like to  investigate the possibility of continuing my PhD studies under your supervision at the University of British Columbia.
-I obtained my Bsc and Msc in Computer Engineering from  Cairo University, Egypt in  2003 and 2009 respectively.
-I worked both in the academic and industrial sectors (mention the position).
-I am planning (decided) to pursue a research career by first continuing my Post Graduate studies in a respectful university.
-I work at Electronics Research Institute in the 'Signal and Image Processing' research group.
+
+I would like to  investigate the possibility of continuing my PhD studies under
+your supervision at the University of Calgary.  My name is Maha and I obtained
+my Bsc and Msc in Computer Engineering from  Cairo University, Egypt in  2003
+and 2009 respectively.  I worked both in the academic (as a Research Assistant)
+and industrial (as a Software Engineer) sectors . I am planning to pursue a
+research career by first continuing my Post Graduate studies. My main research
+interest is in integrating software engineering methods
+with different human-compture interaction.
+
+Most my research experience is in Machine Learning but I have vast interest in
+Software Engineering. I worked for four years as a Software Engineer for Engnet,
+Egypt. After the first year, I was promoted to Team Leader and Senior Software
+Engineer where I got more involved in software process and how to improve the
+performance of developers. Since then, Software Engineering became my passion;
+I took charge of improving the software development process,
+helping developers for better collaborative work and looking for methods that will improve issue and bug tracking systems.
+     I read  several  technical papers on software engineering, bug tracking,
+     team collaboration and code analysis systems.  Because of my passion I
+     decided to pursue my studies in PhD in Software Engineering field. (kindly
+     see my CV for mor information.)
+
+Currently, I work at Electronics Research Institute in the 'Signal and Image
+Processing' research group.
 My work focuses on conducting research and related activities in the field of Machine learning and Image processing.
@@
   My research in AUC focuses on Arabic Handwriting Recognition especially improving interaction and recognition of Online Arabic handwriting systems.
-(put this paragraph first)Even though all my research experience is in Machine learning and Image processing fields, I have vast interest in Software Engineering.
-I worked for four years as a Software Engineer for Engnet Company, Egypt.
-I was responsible for analysis of software requirements and the design of the software systems.
-After the first year, I was promoted to Team Leader and Senior Software Engineer where I got more involved in software process and how to improve the performance of developers.
-Since then, Software Engineering became my passion; I spent a lot of time trying to improve (Make it as a task) software development process, helping developers for better collaborative work and looking for methods that will improve issue and bug tracking systems.
-I read  several  technical papers on software engineering, bug tracking, team collaboration and code analysis systems.
-Because of my passion I decided to pursue my studies in PhD in Software Engineering field.
-I am currently planning for my PhD studies.
-I have preference towards studying in Canada where state of the art research exists and particularly in the University of British Columbia due to its international recognition as one of the best Universities in Canada.
-My research interests are mining code and bug repositories (Be specific about the project name..but I am open to any idea in the field).
-Also I am interested in code analysis and software team collaboration which I hope falls under your research interests as indicated in the website of University of British Columbia (be specific about the website if it is a group or if its is personal website)
 
-I am wondering about the possibility of pursuing my graduate studies under your  guidance and supervision at University of British Columbia as from the fall of 2011. I write to seek your kind advice.
+ My research interests are in developing applications to improve
+     interaction of tabletop surfaces and approachers for improving interactions
+     between distributed agile teams. I am also intrested in how to help teams to create
+     and maintain useful software but I am open to any idea in the field. I
+     hope thes interests falls under your research interests as indicated in the web site of the Agile Software Engineering research group.
+
+     I have preference towards studying in Canada where state of the art
+     research exists and particularly in the University of Calgary due to its
+     international recognition as one of the best Universities in Canada. I am
+     wondering about the possibility of pursuing my graduate studies under your guidance and supervision at University of Calgary. I
+write to seek your kind advice.
 

### 2011-01-11 IntroductionLetters/document.tex
correccting the warp

@@
 
-I would like to  investigate the possibility of continuing my PhD studies under
-your supervision at the University of Calgary.  My name is Maha and I obtained
-my Bsc and Msc in Computer Engineering from  Cairo University, Egypt in  2003
-and 2009 respectively.  I worked both in the academic (as a Research Assistant)
-and industrial (as a Software Engineer) sectors . I am planning to pursue a
-research career by first continuing my Post Graduate studies. I am specially
-intereseted in. ..  .  .  .  .   .  .  .  .  .  .
+I would like to investigate the possibility of continuing my PhD studies under your supervision at the University of Calgary.
+My name is Maha and I obtained my Bsc and Msc in Computer Engineering from Cairo University, Egypt in 2003 and 2009 respectively.
+I worked both in the academic (as a Research Assistant) and industrial (as a Software Engineer) sectors.
+I am planning to pursue a research career by continuing my Post Graduate studies.
+My main research interest is in integrating software engineering methods with different human-computer interaction.
+Most my research experience is in Machine Learning but I have vast interest in Software Engineering.
+I worked for four years as a Software Engineer for Engnet, Egypt.
+After the first year, I was promoted to Team Leader and Senior Software Engineer where I got more involved in software process and how to improve the performance of developers.
+Since then, Software Engineering be- came my passion; I took charge of improving the software development process, helping developers for better collaborative work and looking for methods that will improve issue and bug tracking systems.
+I read several technical papers on software engineering, bug tracking, team collaboration and code analysis systems.
+Because of my passion I decided to pursue my studies in PhD in Software Engineering field.
+(kindly see my CV for more information.)
 
-Most my research experience is in Machine Learnning but I have vast interest in
-Software Engineering. I worked for four years as a Software Engineer for Engnet,
-Egypt. After the first year, I was promoted to Team Leader and Senior Software
-Engineer where I got more involved in software process and how to improve the
-performance of developers. Since then, Software Engineering became my passion;
-I took charge of improving the software development process,
-helping developers for better collaborative work and looking for methods that will improve issue and bug tracking systems.
-     I read  several  technical papers on software engineering, bug tracking,
-     team collaboration and code analysis systems.  Because of my passion I
-     decided to pursue my studies in PhD in Software Engineering field.
-
-Currently, I work at Electronics Research Institute in the 'Signal and Image
-Processing' research group.
+Currently, I work at Electronics Research Institute in the 'Signal and Image Processing' research group.
 My work focuses on conducting research and related activities in the field of Machine learning and Image processing.
-I also work as a part time Research and Teaching Assistant in American University in Cairo (AUC).
+I also work as a part-time Research and Teaching Assistant in American University in Cairo (AUC).
 My research in AUC focuses on Arabic Handwriting Recognition especially improving interaction and recognition of Online Arabic handwriting systems.
 
-     I have preference towards studying in Canada where state of the art
-     research exists and particularly in the University of Calgary due to its
-     international recognition as one of the best Universities in Canada. My
-     research interests are   ..     .  .  .  .  . .  .  . .. .. but I am open
-     to any idea in the field. Also I am interested in .. . .   .  . . .. . . .. .
-     I hope falls under your research interests as indicated in your
-     group website.
+My research interests are in developing applications to improve interaction of tabletop surfaces and approaches for improving interactions between distributed agile teams.
+I am also interested in how to help teams to create and maintain useful software but I am open to any idea in the field.
+I hope these interests falls under your research interests as indicated in the web site of the Agile Software Engineering research group.
 
-I am wondering about the possibility of pursuing my graduate studies under your
-guidance and supervision at University of Calgary. I
-write to seek your kind advice.
+I have preference towards studying in Canada where state of the art research exists and particularly in the University of Calgary due to its international recognition as one of the best Universities in Canada.
+I am wondering about the possibility of pursuing my graduate studies under your guidance and supervision at University of Calgary.
+I write to seek your kind advice.
 

### 2011-01-11 trunk/IntroductionLetters/document.tex
finall version of the letters.

@@
 
-I would like to  investigate the possibility of continuing my PhD studies under
-your supervision at the University of Calgary.  My name is Maha and I obtained
-my Bsc and Msc in Computer Engineering from  Cairo University, Egypt in  2003
-and 2009 respectively.  I worked both in the academic (as a Research Assistant)
-and industrial (as a Software Engineer) sectors . I am planning to pursue a
-research career by first continuing my Post Graduate studies. My main research
-interest is in integrating software engineering methods
-with different human-compture interaction.
+I would like to investigate the possibility of continuing my PhD studies under your supervision at the University of Calgary.
+Since I wanted to pursue my PhD software Engineering, my friend"Dina Said" who is a graduate student in the department highly recommended you.
+I then talked with two of your students; "Maha Shouman" and "Ealaf Selim" who introduced me to the various research projects in the Agile group.
 
-Most my research experience is in Machine Learning but I have vast interest in
-Software Engineering. I worked for four years as a Software Engineer for Engnet,
-Egypt. After the first year, I was promoted to Team Leader and Senior Software
-Engineer where I got more involved in software process and how to improve the
-performance of developers. Since then, Software Engineering became my passion;
-I took charge of improving the software development process,
-helping developers for better collaborative work and looking for methods that will improve issue and bug tracking systems.
-     I read  several  technical papers on software engineering, bug tracking,
-     team collaboration and code analysis systems.  Because of my passion I
-     decided to pursue my studies in PhD in Software Engineering field. (kindly
-     see my CV for mor information.)
+ My Name is Maha El Meseery and I obtained my Bsc and Msc in Computer Engineering from Cairo University, Egypt in 2003 and 2009 respectively.
+I worked both in the academic (as a Research Assistant) and industrial (as a Software Engineer) sectors.
+I am planning to pursue a research career by continuing my Post Graduate studies.
+My main research interest is integrating software engineering methods with different human-computer interaction.
+.
 
-Currently, I work at Electronics Research Institute in the 'Signal and Image
-Processing' research group.
-My work focuses on conducting research and related activities in the field of Machine learning and Image processing.
-I also work as a part time Research and Teaching Assistant in American University in Cairo (AUC).
+Most of my research experience is in Machine Learning but I have a vast interest
+in Software Engineering.
+I worked for four years as a Software Engineer for Engnet, Egypt.
+After the first year, I was promoted to Team Leader and Senior Software Engineer where I got more involved in software process and how to improve the performance of developers.
+Since then, Software Engineering became has become my passion; I took charge of improving the software development process, helping developers for better collaborative work and looking for methods that will improve issue and bug tracking systems.
+Because of my passion I decided to pursue my studies in PhD in Software Engineering field.
+
+Currently, I am working at the Electronics Research Institute in the 'Signal and Image Processing' research group.
+My work focuses on conducting research and related activities in the field of Machine learning and Image processing with a focus on sketch recognition.
+I am also working as a part-time Research and Teaching Assistant in American University in Cairo (AUC).
 My research in AUC focuses on Arabic Handwriting Recognition especially improving interaction and recognition of Online Arabic handwriting systems.
 
- My research interests are in developing applications to improve
-     interaction of tabletop surfaces and approachers for improving interactions
-     between distributed agile teams. I am also intrested in how to help teams to create
-     and maintain useful software but I am open to any idea in the field. I
-     hope thes interests falls under your research interests as indicated in the web site of the Agile Software Engineering research group.
+My research interests are in mainly in the areas of developing applications to improve interaction of tabletop surfaces and approaches for improving interactions between distributed agile teams.
+Furthermore, I am interested in how to help teams to create and maintain useful software but I am also open to any idea in the field.
+I hope these interests falls under your research interests as indicated in the web site of the Agile Software Engineering research group.
 
-     I have preference towards studying in Canada where state of the art
-     research exists and particularly in the University of Calgary due to its
-     international recognition as one of the best Universities in Canada. I am
-     wondering about the possibility of pursuing my graduate studies under your guidance and supervision at University of Calgary. I
-write to seek your kind advice.
+I have preference towards studying in Canada and particularly in the University of Calgary due to its international recognition as one of the best Universities in Canada.
+I am wondering about the possibility of pursuing my graduate studies under your guidance and supervision at University of Calgary.
+I write to seek your kind advice.
+
+Kindly find attached to the email my resume.
+
+Maha
 

### 2011-02-25 MyCv/Resyne.tex

@@
 \setlength\headsep{-0.25in}
-\setlength\textheight{9.8in}
+\setlength\textheight{9.5in}
 \setlength\textwidth{6.5in}
@@
 
-\title { Resume \\ Maha Mohammed Nabeel El Meseery}
+\title {Curriculum Vitae \\ Maha Mohammed Nabeel El
+Meseery\\ \small \textbf{ Address:} $2^{nd}$ Tahreer street, Dokki, Giza 12615,
+Egypt. \\ \textbf{Tel:} (+2012)112- 5435, (+202)3336-8292.
+\textbf{Email:} melmeseery@aucegypt.edu}
 \date{}
@@
   \begin{description}
- \item [2004-2009]  \textbf{ Msc.E., Cairo University, Computer Engineering Department, Giza, Egypt}\\
- Thesis Area:\textbf{\textit{Sketch Recognition And Understanding}.}\\
- Thesis Supervisors: \textit{Prof. Nevin Darwish (Cairo University), Dr Magda
- Fayek (Cairo University), Prof. Samia Mashali (Electrical Research Institute),
- Mahmoud Fakr El Din (Electrical Research Institute).}
- \item[1998-2003]  \textbf{ Bsc.E., Cairo University, Computer Engineering Department Giza, Egypt}\\
-  Final Grade: \textbf{Very Good with honors.}\\
-  Graduation project: \textit{Personal Face identification System,} \textbf{Distinction.}\\
-  Project Supervisors: \textit{Dr. Madga Fayek.}
+\item [2004-2009]  \textbf{ Msc.E., Cairo University, Computer Engineering Department, Giza, Egypt}
+  \begin{itemize}
+    \item \textit{ Thesis Title:}\textbf{Sketch Recognition And Understanding}.
+    \item \textit{Thesis Supervisors:} Prof. Nevin Darwish, Dr Magda Fayek , Prof. Samia
+ Mashali , Mahmoud Fakr El Din
+ \item \textit{ Graduate Courses:} Graduate Courses: Advanced Neural Networks, Image
+ Processing and Computer Vision, Advanced Computer Networks, Advanced Database Systems, Embedded Systems, VLSI Testing, Numerical Analysis, and Technical Writing.
 
+\end{itemize}
+
+ \item[1998-2003]  \textbf{ Bsc.E., Cairo University, Computer Engineering Department Giza, Egypt}
+   \begin{itemize}
+    \item \textit{Final Grade:}\textbf{Very Good with honors}.
+  \item \textit{Graduation project:} Personal Face identification System, \textbf{Distinction.}
+  \item \textit{Project Supervisors:} Dr. Madga Fayek.
+\end{itemize}
   \end{description}
@@
    Recognition. This includes implementing various features extraction and classification methods.
-   \item Assisting in collecting, gathering and analyzing an Online Arabic
-   handwriting Dataset.
+   \item Assisting in collecting, gathering and analyzing Online Arabic
+   handwriting Datasets.
 %   \item Assisting and developed research in area of Arabic and Latin digit
-   Recognition.
+ %  Recognition.
  \item Implementing different features extraction and classification methods for Arabic and Latin digits and comparing their performance.
@@
  \textbf{January 2007- present}\\
- \textbf{Research Assistant}, Electrical Research Institute, Egypt
+ \textbf{Research Assistant}, Electrical Research Institute, Egypt.
  \begin{enumerate}
@@
  \textbf{September 2005- September 2008 [part time from January 2007] }\\
- \textbf{Senior Software Engineer}, EngNet, Egypt\textbf{}
- The following is summary of my achievements:
+ \textbf{Senior Software Engineer}, EngNet, Egypt.
+
  \begin{enumerate}
 
-\item Introduced new software procedures in the company to enhance developers performance.
+\item Introduced new software procedures in the company to enhance developers'
+performance and collaboration.
  \item Analyzed and designed various software projects using UML and software engineering concepts.
@@
  application.
- \item Supervised and managed the development process for a team consists of four developers and two quality engineers.
+ \item Supervised and managed the development process for the developers and
+ quality engineering teams.
  \item Assisted junior programmers in development using Java, JSP, Struts, RCP and MySQL.
@@
  \textbf{September 2003- September 2005}\\
- \textbf{Teaching Assistant}, Faculty of Computer Science, MSA University, Giza, Egypt.\textbf{}
+ \textbf{Teaching Assistant}, Faculty of Computer Science, MSA University, Giza, Egypt.
  \begin{enumerate}
@@
  \& Darwish, Nevin., Sketch Recognition Using Particle Swarm Algorithms,  IEEE
- International Conference on Image Processing, Cairo, Egypt, November,  2009.
+ International Conference on Image Processing, November,  2009, Cairo, Egypt.
   \item El Meseery, Sketch Recognition and Understanding,  Master Thesis,
   Faculty of Engineering, Computer Engineering Department, Cairo
-  University, Egypt, November, 2009.
+  University, November, 2009, Egypt.
   \item El Meseery, Maha, Fakhr El Din, Mahmoud, Tarek abouldahab, Sketch
-  Recognition using Enhanced Particle Swarm Optimization, (Submitted to)Pattern
-  Recognition Letters, 2010, (17 pages).
+  Recognition using Enhanced Particle Swarm Optimization, Submitted to Pattern
+  Recognition Letters, (17 pages Submitted in Dec 2010).
 \end{enumerate}
@@
   \begin{enumerate}
-    \item \textbf{1999-2003} \textit{Excellence Reward} Faculty of Engineering,
-    Cairo, Egypt.
-
+    \item Jul 2003: Cairo University Award for Academic Distinction, Egypt.
+	\item 1999 - 2003: Fellowship, Faculty of Engineering, Cairo University, Giza,
+	Egypt.
   \end{enumerate}
@@
  \item DB2, JSP, Java beans, Servelts and IBM Web Sphere, Cairo University, Giza, Egypt (July- August 2001).
- \item Java applets, Training on faculty to develop an e-version of the calculus book taught to preparatory year student of engineering faculty (August -- September 2000).
+ \item Java applets, Training to develop an e-version of the calculus book
+ taught to engineering students (August -- September 2000).
  \item MFC and VC6, Cairo University, Giza, Egypt (July 2000).
@@
 teenagers.
-\item Acting as guidance for behavioural, social and mental aspect for orphaned
+\item Acting as guidance for behavioural, social and mental aspects for orphaned
 teenagers.
@@
 \item  Member of Egyptian shooting clubs.
-\item  Enjoy reading, swimming, walking and watching movies.
+\item  Enjoy reading, swimming, cooking, walking and watching movies.
 \end{enumerate}
-\newpage
-\section*{Personal Information:}
-\noindent
- \textbf{ Address: } $2^{nd}$ Tahreer street, Dokki, Giza 12615, Egypt.
-
- \textbf{ Telephone:} (+2012)112- 5435, (+202)3336-8292.
-
-\textbf{  Email: }  melmeseery@aucegypt.edu
+%\newpage
+% \section*{Personal Information:}
+% \noindent
+%  \textbf{ Address: } $2^{nd}$ Tahreer street, Dokki, Giza 12615, Egypt.
+%
+%  \textbf{ Telephone:} (+2012)112- 5435, (+202)3336-8292.
+%
+% \textbf{  Email: }  melmeseery@aucegypt.edu
 

### 2011-08-08 FacePaper/paper.tex
first import

%%This is a very basic article template.
%%There is just one section and two subsections.
%\documentclass{article}
\documentclass[preprint,12pt]{elsarticle}
\journal{Jornal Name in the document\ldots\ldots\ldots.}
\begin{document}

\begin{frontmatter}

\title {Face recgontion using Particle Swarm clustering  }

\author[label]{Maha El Meseery}
\author[label]{Mahmoud fakhr el Deen}
\author[label]{Heba Nemer}

\address[label1]{Electronic Research Institute, Cairo, Egypt}

\begin{abstract}
%% Text of abstract
The abstract of paper\ldots\ldots\ldots\ldots.

\end{abstract}

\begin{keyword}
My first keyword \sep keyworbd\ldots

\end{keyword}

\end{frontmatter}

\section{Introduction}

\section{Background Information}
\subsection{Particle Swarm Optimization}
%\section{Discrete Particle Swarm Optimization}
\label{sec:ParticleSwarmAlgorithm}
 The main idea of \textit{Particle Swarm Algorithm (PSO)} is to represent each solution with a $N$ dimension particle from the solution space \cite{PSOFirst}. Each particle moves with a direction and velocity $v_{ij}$ based on equations \ref{eq:Swarm1} \& \ref{eq:Swarm}.

\begin{equation}
%\[
p_{ij}=p_{ij}+v_{ij},
%\
\label{eq:Swarm1}
\end{equation}

where $p_{ij}$ represent the $j$th dimension in the $i$th particle and $v_{ij}$ is the velocity of the $j$th dimension in the $i$th particle.
 %Equation [\ref{eq:Swarm}] shows how velocity and direction of each particle are computed
 \begin{equation}
v_{ij}  = v_{ij} \omega + c_1 r_1 (lbest_{ij}  - p_{ij} ) + c_2 r_2 (gbest_{ij}  - p_{ij} ),
\label{eq:Swarm}
\end{equation}
 where $\omega$ is the inertia weight parameter which controls the tradeoff between exploration and exploitation,  $lbest_{ij}$ is the local best particle, $gbest_{ij}$ is the global best particle, $r_1$ \& $r_2$ are random variables and $c_1$ \& $c_2$ are the swarm acceleration parameters.

 After each iteration the global best $g_{best}$ particle and the agent local best particle $l_{best}$are evaluated based on the maximum fitness functions of all particles in the solution space. The solution is found after achieving a specific number of iteration or after an error threshold is achieved.
Equation \ref{eq:descrite} is used to change the general swarm algorithm into binary particle (\textit{Discrete Particle Swarm Algorithm DPSO}). The \textit{DPSO} handles particle values of either $0$ or $1$ \cite{PSODisceret}.
\begin{equation}
   P(i)\Leftarrow
\left \{
\begin{array}{c}
1 \quad \quad if\quad r_{3}>p_{i}  \\

0 \quad \quad if\quad r_{3}<p_{i},
\label{eq:descrite}
\end{array}\right.
\end{equation}
 where $p_{ij}$ is the numerical values of the particle and $r_{3}$ is a random variable.
\subsection{Eigen Faces}
\label{facereg}
A set of eigenfaces \cite{facesite} can be generated by performing a mathematical process called principal component analysis (PCA) on a large set of images depicting different human faces. Informally, eigenfaces can be considered a set of "standardized face ingredients", derived from statistical analysis of many pictures of faces. Any human face can be considered to be a combination of these standard faces.

 The aim is to represent a face as a linear combination of a set of basis images called eigenfaces. That is :
\begin{equation}
% MathType!MTEF!2!1!+-
% feaafiart1ev1aaatCvAUfeBSjuyZL2yd9gzLbvyNv2CaerbuLwBLn
% hiov2DGi1BTfMBaeXatLxBI9gBaerbd9wDYLwzYbItLDharqqtubsr
% 4rNCHbGeaGqiVu0Je9sqqrpepC0xbbL8F4rqqrFfpeea0xe9Lq-Jc9
% vqaqpepm0xbba9pwe9Q8fs0-yqaqpepae9pg0FirpepeKkFr0xfr-x
% fr-xb9adbaqaaeGaciGaaiaabeqaamaabaabaaGcbaGaeuOPdy0aaS
% baaSqaaiaadMgaaeqaaOGaeyypa0ZaaabCaeaacqaHjpWDdaWgaaWc
% baGaamyAaaqabaGccaWG1bWaaSbaaSqaaiaadQgaaeqaaaqaaiaadQ
% gacqGH9aqpcaaIXaaabaGaam4saaqdcqGHris5aaaa!444E!
\Phi _i  = \sum\limits_{j = 1}^K {\omega _i u_j }
\end{equation}

Where $\Phi_i$ represents the $i^{th}$ face with the mean subtracted from it, $\omega _i$ represent weights and  $u_j$ the eigenvectors.

This can be represented aptly in a figure as:

\begin{figure}[h]
 \begin{center}
 \centering
\includegraphics[scale=0.5]{eigenfacesrec.jpg} \caption{Face reconstruction from eigen faces. }
 \end{center}\end{figure}

\subsubsection{Assumptions:}

\begin{enumerate}

\item There are $M$ images in the training set.

\item There are $K$ most significant Eigenfaces using which we can satisfactorily approximate a face. Needless to say $K <M$.

\item All images are $NxN$ matrices, which can be represented as dimensional vectors. The same logic would apply to images that are not of equal length and breadths. To take an example: An image of size 112 x 112 can be represented as a vector of dimension 12544 or simply as a point in a 12544 dimensional space.

\end{enumerate}

\subsubsection{Algorithm for Finding Eigenfaces}

\begin{enumerate}
\item Obtain $M$ training images , $I_1, I_2 \dots I_m$  , centered faces images.

%\includegraphics{training-images.jpg}
\begin{figure}[h]
 \begin{center}
 \centering
\includegraphics[scale=0.5]{training-images.jpg} \caption{Sample Faces. }
 \end{center}\end{figure}
\item Represent each image $I_i$ as a vector $\Gamma_i$.
\begin{equation}
\label{arrequation}
I_i  = \left[ {\begin{array}{*{20}c} {a_{11} } & {a_{12} } &\ldots& {a_{1N} }  \\
   {a_{21} } & {a_{22} } &  \ldots   & {a_{2N} }  \\
    \vdots  &  \vdots  &  \ddots   & {}  \\
   {a_{N1} } & {a_{N2} }  & {} & {a_{NN} }  \\
\end{array}} \right]_{NxN} \Rightarrow\left[
{\begin{array}{*{20}c}
   {a_{11} }  \\
   {\vdots}  \\
    {a_{1N} }  \\
     {a_{21}}  \\
        {\vdots}  \\
   {a_{2N} }  \\
   {\vdots}  \\
   {a_{NN} }  \\
\end{array}} \right]_{N^2 x1} = \Gamma _i
\end{equation}
\item Find the average or the mean face vector  $\Psi$

\begin{equation}
\Psi  = \frac{1}{M}\sum\limits_{i = 1}^M {\Gamma _i }
\end{equation}
\item  Subtract the mean face $\Psi$ from each face vector $\Gamma_i$ to get a set of vectors $\Phi_i$. The purpose of subtracting the mean image from each image vector is to be left with only the distinguishing features from each face and (removing)  in a way information that is common.
\begin{equation}
\Phi _i  = \Gamma_i  - \Psi
\end{equation}
\item  Find the Covariance matrix $C$:
\begin{equation}
C = AA^T
\end{equation} , where $A = \left[ {\begin{array}{*{20}c}
   {\Phi _1 } & {\Phi_2} & {\dots} & {\Phi _M }  \\
\end{array}} \right]
$
Note that the Covariance matrix has simply been made by putting one modified image vector obtained in  one column each.

Also note that $C$is a $N^2xN^2$ matrix and $A$ is a $NxM$ matrix.

\item  Instead of the Matrix $AA^T$  consider the matrix $A^TA$ .

\item Find the $M$ best  Eigenvectors of $C=A^TA$. using the relation of $u_i=Av_i$  Also keep in mind that $\left\| {u_i} \right\| = 1.$ where  $u_i$ is the eigen vector of $C$ and $v_i$ is the eigen value of the matrix $C$.

\item Select the best  Eigenvectors, the selection of these Eigenvectors is done heuristically.

\item Finding Weights:\\

The Eigenvectors found at the end of the previous section,  when converted to a matrix in a process that is reverse to that in STEP \ref{facereg}, have a face like appearance. Since these are Eigenvectors and have a face like appearance, they are called Eigenfaces. Sometimes, they are also called as Ghost Images because of their weird appearance.

Now each face in the training set (minus the mean),$\Phi_i$  can be represented as a linear combination of these Eigenvectors   $u_j$:
  \begin{equation}
\Phi_i  = \sum\nolimits_{j = 1}^k {\omega_j u_j}
  \end{equation}
, where  $u_j$‘s are Eigenfaces and $\omega_j$ are the weights and can be calculated as :
\begin{equation}
\label{weighteq}
\omega _j  = u_j^T \Phi _i
\end{equation}
Each normalized training image is represented in this basis as a vector $\Omega$.
\end{enumerate}
\subsection{PSO clustering}

On the Merwe and Engelbrechet \cite{psoclustering} Method. The particle $X_i$ is constructed that
\begin{equation}
X_i=\left( m_{i1}, m_{i2}, m_{i3},\dots, m_{iN_c}\right)
\end{equation}
where $N_c$ is the number of clusters to be formed and $m_{ij}$ corresponds to the $j^{th}$ centroid of the  $i^{th}$ particle, the centroid of the cluster $C_{ij}$. Thus single particle represents a candidate solution to a given clustering algorithm. Each particle is evaluated using the following equation:
% \begin{equation}
% F_c=\frac{\sum^{N_c}_{j=1}\right[\sum_{\forall Z_j \epsilon C_{ij}}\frac{ d\right(Z_p,m_{ij}\left)}{|C_{ij}|}  \left]}{N_c}
% \end{equation}
\begin{equation}
F_c  = \frac{{\sum\nolimits_{j = 1}^{N_c } {\left[ {\sum\limits_{\forall z_j  \in C_{ij} } {\frac{{d(z_p ,m_{ij} )}}{{\left| {C_{ij} } \right|}}} } \right]} }}{{N_c }}
\end{equation}
where $F_c$ is the fitness function, $N_c$ is the number of clusters in the particle, $Z_p$ denotes the $p^{th}$ data vector $\left| {C_{ij} } \right|$ is the number of data vectors belonging to the cluster $C_{ij}$ and $d$ is the Euclidian distance between $Z_p$ and $m_{ij}$.

The clustering algorithm prposed by Merwe and Engelbrecht \cite{} can be as follows:
% \begin{algorithm}
%  \caption{ Discrete Particle Swarm Algorithm }
%  \label{DPSOalg}
\begin{algorithmic}
%\begin{verbatim}

\STATE  Generate $N$ particles with $N_c$ clusters  ($P_i,$ $\dots,$ $P_{N_c}$)
\STATE Initialize the clusters centroids for each particle randomly
\FOR{i=0 \TO $t_{max}$}
  \FOR{ each particle  $p_i$}
    \FOR{ each data vector  $zp$}
        \STATE   Calculate distance $d(zp,m_{ij})$ from cluster centroid.
        \STATE   Assign $zp$ to cluster $C_{ij}$ such that
        \STATE   $d(zp,m_{ij})=min \forall k=1,\dots,N_c {d(zp,m_{ik})}$
    \ENDFOR
    \STATE  Compute the Fitness function for the particle.
     \STATE  Update the global best and local best
     \STATE  Move all particle using PSO equations.
     \ENDFOR
  \ENDFOR
 \RETURN global best
%\end{verbatim}

 \end{algorithmic}
\section{System}

\begin{figure}[h]
 \begin{center}
 \centering
\includegraphics[scale=0.5]{blockchart.jpg} \caption{ The flow Chart}
 \end{center}\end{figure}

Both previous algorithms  were implemented and tested for correctness. Also the k-means algorithm was used to compare the result of the PSO clustering algorithm. The code was implemented using Matlab and using a Database as train and test set. The Databases contains 50 persons and 4 images per person. Currenlty all the images are used for training and the same set is used for testing.

The PSO algorithm is used to clusters each face weights computed using equation \ref{weighteq}. The algorithms first compute the eigen vectors for all 50 faces as explained in section \ref{facereg}. The computed $u_j$ is used to compute the weights for each image in the database is computed which generate a data matrix of 150 X 50 ( number of image X weight vector (no of eigenfaces)). The clustering algorithm is used to get center weigh vector for each person. I choose the number of clusters same as number of person for easier recognition step.

The recognition step will be easier as to compute the weigh vector of the input image and then compute the distance from each cluster centroid. The cluster with the min distance is the cluster of the face.
\subsection{Nelder Enhancment}

\section{Results}
Two different databases was processed, the first contains 60 indian faces with 11 different samples with different poses and experisions. The second database is 120 eroupian faces each has 20 samples. The database is processed to the input format for the particle algorithm. The data is then divided into training and testset. The percentage can be changed and changed to cross validate the results and experiments.

\subsection{Experiments Finding PSO paramters}
\begin{table}
	\centering
		\caption{PSO Result }
	\label{tab:PSO1}
	%\scalebox{0.99}{
		\begin{tabular}{|l|c|c|c|c|c|}
		 \hline
C1 &	1	&1.2&	1.6&	1.8 &	2	\\ \hline
	&70.893	&70.893	&70.605	&70.029	&69.741 \\  \hline
C2	&	1	&1.2&	1.6&	1.8 &	2	\\ \hline
	&70.89	&70.61	&70.89	&69.45	&69.45 \\ \hline
W	&0.5&	1	&1.5	&2&	4	\\ \hline
	&69.74	&70.89&	70.89	&70.32&	70.89	\\ \hline
\end{tabular}
		%}
\end{table}

\subsection {Comparing With other systems}

\subsection{ Comparing Nelder enhancement}

\section{Conclusion}
\section{future work}
% \subsection{Subtitle}
%
% Plain text.
%
% \subsection{Another subtitle}
%
% More plain text.
\bibliographystyle{model1a-num-names}
\bibliography{mybibfile,library}

\end{document}

### 2011-08-11 FacePaper/aiml.tex
adding all the files for first time

% Template AIML-ICGST-LaTex-Template
% 1/06/04

\documentclass[a4paper, twoside]{article}

\usepackage{icgst}
\begin{document}
\input{files/head}
\input{files/abstract}
\input{files/introduction}
\input{files/theory}
\input{files/conclusion}
\input{files/ack}

\bibliographystyle{plain}
\bibliography{files/icgst}

\newpage
\section*{Biographies}

\parpic{\includegraphics[height=3cm]{pics/guest}} \textbf{Author One} \small Author One received his B.Sc. from .......  at ....... and his M.Sc. from ........ at ....... Currently he is working for ....... and he is a PhD student at University of ....... in the ........ Laboratory under the supervision of ......... His research interests include .........,  .......... and ...........\\

\parpic{\includegraphics[height=3cm]{pics/guest}}
\textbf{Author Two} \small is currently Professor at University of ....., in the ....... Laboratory. He is Senior Member of the ...... He is Editor Associate of the ....... on ....... He received ........ award in ...... He is doing his research work in ...... for ........ His main research interests are ......, ....., .... and .......\\

\end{document}

### 2011-08-11 FacePaper/files/abstract.tex
adding all the files for first time

% AIML ICGST LaTeX Template

\begin{abstract}
%% Text of abstract
This paper introduce a system that uses Particle Swarm Optimization

\end{abstract}

%\textit{\textbf{Keywords:} Autonomous robot, Stochastic control, Kalman filter,
%Fuzzy logic, Neural Network, Adaptive navigation.}

### 2011-08-11 FacePaper/files/conclusion.tex
adding all the files for first time

% AIML ICGST LaTeX Template 
\section{Conclusion}
\label{Sec:Con}
This is the conclusion. The major contributions of this article arise from the formulation of a new approach, spectral analysis, to the modelling and identification of XYZ and ABC features that provide improved computational efficiency in the positioning techniques. By manipulating the manner in which feature information of both A and BVN signatures is incorporated into the model, it can be shown that significant improvements in the performance of the algorithm can be realised. Moreover, the simplicity and the efficiency of dynamic pose tracking techniques succeeded to improve the robot pose estimation process.

### 2011-08-11 FacePaper/files/introduction.tex
adding all the files for first time

% AIML ICGST LaTeX Template 
\section{Introduction}
\label{Sec:int}
We consider that it is necessary to decide a state of the network load in order to adjust an image information quantity responding a state of the network.  In this paper, A state of the network load is decided by using a ratio of send and receive of image data communicated between slave site and master site, a receive interval of image receive report and a send completion time of \textbf{image data}.  The master site reports a received image data quantity to the slave site every a certain time.  The slave site decides a state of the network load by making a comparison between received report and image data quantity transmitted until it receive reports.  Simultaneously, the slave site decides that the network load is large, in case that it can't receive the report inside a certain time.  Also the slave site measures a transmission completion time of image data.\\

To design Service Robots for elderly and disabled people is a very demanding task. There are fundamental differences in the necessities, compared with industrial robot applications, as well as to several other service robot applications. The tasks vary in wide areas, depending on the actual situation of the user and they cannot be pre programmed as in other robotic applications. While all service robots must be designed for operation in clustered, a priory unknown and variable environments, new demands arise for the support of the target group. The interaction with the user and the variety of the supporting tasks may be mentioned here. On the other hand no total autonomy is demanded. It is rather important to use the cognitive abilities of the users and support compensate they only in areas where physical or cognitive impairments exist. Another limitation is the number of units which may be sold. Closely related to the small number compared with other industrial products is the price of the today available components. The design of a robotic aid must consider these limitation but also possibilities which arise due to the specific users. For the near future, the primary target group are disabled people with serious physically handicaps but only small mental disabilities if at all. Opposite to other service robot applications the cognitive abilities can and must be used to support and simplify the controller design, esp. for environment identification and scene interpretation. The remaining automation of the tasks is still very demanding, but the sensor data interpretation is eased if the user support the automation system. The paper will discuss the design philosophy which results from this approach. An electric wheelchair with an added robotic arm serves as an example. The main target group of this unit are paraplegics ore people with similar handicaps. \\

The remainder of the paper is organized as follows: Section (\ref{Sec:Th}) focuses on ....... Section (\ref{Sec:Con}) emphasizes on ...........

### 2011-08-11 FacePaper/files/theory.tex
adding all the files for first time

% AIML ICGST LaTeX Template 
\section{Theory}
\label{Sec:Th}
There are two different approaches to the characterization of dynamic systems: In linear systems theory, one can assume either some structure in the signals or some structure in the system. Attempts have been made to combine these two approaches e.g. harmonic identification techniques in the Fourier domain.\\ 
\textbf{\textit{First approach:}} Structure the signal can be found using linear transforms. This approach does not take into account that the system has some structure. In the time domain, filtering is a linear transformation. The Fourier, Wavelet, and Karhunen-Loeve transforms have compression Capability and can be used to identify some structure in the signals. When we are using these transforms, we do not take into account any structure in the system.\\ 
\textbf{\textit{Second approach:}} Structure the system can be found by fitting a model to the system.

\subsection{System Modeling}
Physical models of robots are either reduced-size copies of the original dynamics following the laws of model similarity, or analogies. The idea of an analogy implies that there exists "something" at every instant of time that is to be analogous to the dependent variables of the original physical system. Mathematical models map the relationships between the physical variables in the robot dynamics to be modeled onto mathematical structures like simple algebraic equations, systems of differential equations or even  difference equations. Mathematical modeling of robots can be developed in different ways: either purely theoretically based on the physical relationships (sensors actuators Interaction), which are a priori known about the robot dynamics, or purely empirically by experiments on the already existing robot, or by a sensible combination of both ways. Models obtained by the first method are often called priori, first principle or theoretical models, while models obtained in the second way are called posteriori or experimental models. Theoretical model building becomes unavoidable if experiments in the respective system cannot or must not be carried out. If the system to be modeled does not yet exist, theoretical modeling is the only possibility to obtain a mathematical model.

\subsection{ARX Modeling}
The discrete ARX modeling scheme  is derived from Kalman filter, see figure (\ref{fig:ARX}). The ARX scheme is widely used in modeling of sequential  system dynamics. This structure takes into account both the observed state $\upsilon_{k}$ and
the driving control signal $\lambda_{k}$ which is given by:
\begin{equation}
\upsilon_k = \sum\limits_{i = 1}^{n_a } {a_i \upsilon_{(k - i)} + } \sum\limits_{i =
0}^{n_b } {b_i \lambda_{(k - i)} + } \eta_k
\label{eq:ARX1}
\end{equation}

\begin{figure}[ht!]
\centering
\scalebox{1.00}{\includegraphics{pics/ARX}}
\caption{ARX modeling of rescue robot dynamics}
\label{fig:ARX}
\end{figure}

Where; $\eta$ is  the modeling residual , representing the white noise. $n_a$ is the model order of the observed state (also called the number of poles). $n_b$ is the model order of the control signal (also called the number of zeros). The operator $q^{-1}$  is the back shift operator or delay, which is given by $q^{ - 1}\upsilon_k = \upsilon_{(k - 1)} $,  that follows:

\begin{equation}
\begin{array}{l}
A(q^{ - 1})\upsilon_k = q^{ - 1}B(q^{ - 1})\lambda_k + \eta_k \\
A(q^{ - 1}) = 1 - a_1 q^{ - 1} - a_2 q^{ - 2} - \ldots - a_{n_a } q^{ - n_a
} \\
B(q^{ - 1}) = b_0 + b_1 q^{ - 1} + b_2 q^{ - 2} + \ldots + b_{n_a } q^{ -
n_b } 
 \end{array}
 \label{eq:ARX2}
\end{equation}

The observed state $\upsilon_{k}$ is the longitudinal velocity/heading while the reference signal $\lambda_{k}$ is the obstacles distribution histogram, acquired by laser/sonar. The Gaussian distributed noise, associated with the observed output, facilitates applying identification algorithms such as; RLS or least mean squares (LMS) to estimate model parameters (coefficients).  This modeling scheme is only applicable within linear or quasi-linear systems. Therefore, it is applied  within this framework to regulate the velocity/heading, while this algorithm failed to cope with position control due to enormous non-linear odometric errors \cite{Buauthor_98, Jauthor_001, Sauthor_011, Rauthor_901, Sauthor_021, Abauthor_2003_13, Boauthor_971, kauthor_2002}. \\

Important Topics:
\begin{itemize}
	\item Item 1.
	\item Subject 2.
	\item New Data.
\end{itemize}

### 2011-08-12 FacePaper/paper.tex
adding all the files for first time
adding change s

@@
 %\documentclass{article}
-\documentclass[preprint,12pt]{elsarticle}
-\journal{Jornal Name in the document\ldots\ldots\ldots.}
+
+\documentclass[a4paper, twoside]{article}
+
+\usepackage{icgst}
+%\documentclass[preprint,12pt]{elsarticle}
+%\journal{Jornal Name in the document\ldots\ldots\ldots.}
 \begin{document}
 
-\begin{frontmatter}
-
-\title {Face recgontion using Particle Swarm clustering  }
-
-\author[label]{Maha El Meseery}
-\author[label]{Mahmoud fakhr el Deen}
-\author[label]{Heba Nemer}
-
-\address[label1]{Electronic Research Institute, Cairo, Egypt}
-
-\begin{abstract}
-%% Text of abstract
-The abstract of paper\ldots\ldots\ldots\ldots.
-
-\end{abstract}
-
-\begin{keyword}
-My first keyword \sep keyworbd\ldots
-
-\end{keyword}
-
-\end{frontmatter}
+\input{files/head}
+\input{files/abstract}
+%\input{files/head2}
 
@@
 
+Face recognition have become one of the most extensively studied
+research topics that spans different disciplines such as pattern recognition,
+image processing and artificial intelligence. This is due to its importance to
+various application like identity authentication, surveillance and intelligents
+vision. One of the ost used method is egin faces which is used to extract
+relevant facial information from faces to used as features to identificy
+faces\cite{}.
+
+There is various statistical features used to detect like DCT, wavelet transform
+and other information Using data clustering as classification was used. The most
+used feature was used is the eigen faces which transform the face image into
+feature space that represents important facial information.
+
+%Face recognition (FR) has emerged as one of the most extensively studied
+%research topics that spans multiple disciplines such as pattern recognition,
+%signal processing  and computer vision. This is due to its numerous  important
+%applications in identity authentication, security  access control, intelligent
+%human-computer interaction,  and automatic indexing of image and video
+%databases.  Many approaches to face recognitions have been  developed; an
+%excellent survey paper on the different face recognition techniques can be
+%found in [1].
+
+% Face recognition systems have been grabbing high attention from commercial market point of view as well as pattern recognition field.
+Face recognition has received substantial biometrics, pattern recognition communities[11][13].
+The face
+% field recognition systems can extract
+% attention from researches in and computer
+% Eigenfaces are mostly used to: a. Extract the relevant facial information, which may or may not
+% be directly related to human intuition of face features such as the eyes, nose, and lips.
+%One way to do so is to capture the statistical variation between face images.
+%b. Represent face images efficiently. To reduce the computation
+% and space complexity, each face image can be represented using a small number of dimensions The eigenfaces may be considered as a set of features which characterize the global variation among face images.
+Then each face image is approximated using a subset of the eigenfaces, those associated with the largest eigenvalues.
+These features account for the most variance in the training set.
+In the language of information theory, we want to extract the relevant information in face image, encode it as efficiently as possible, and compare one face with a database of models encoded similarly.
+A simple approach to extracting the information contained in an image is to somehow capture
+% the variations in a collection of face images,
+% independently encode and compare individual face images. Mathematically, it is simply finding the principal components of the
+% vision the
+% features of face and compare this with the existing database. The faces considered here for comparison are still faces. Machine recognition of faces from still and video images is emerging as
+% an active research area[11].
+The present paper is formulated based on still or video images captured either by a digital camera or by a web cam.
+The face recognition system detects only the faces from the image scene, extracts the descriptive features.
+It later compares with the database of faces, which is collection of faces in different poses.
+The
+
+T
+The remaining of the paper is organized as follows. Firstly, section
+\ref{sec:section2} gives the basic background infromation on  Eigenfaces,
+Particle swarm optimization,and Nelder Meed modification Section
+\ref{sec:proposed} describes in details our proposed system.
+Section \ref{sec:Results} presents the experiments undertaken to develop the system and also some experiments to evaluate the recognition performance of the proposed system.
+Finally, in section \ref{sec:Conclusion} we present conclusion and future works.
+
 \section{Background Information}
-\subsection{Particle Swarm Optimization}
-%\section{Discrete Particle Swarm Optimization}
-\label{sec:ParticleSwarmAlgorithm}
- The main idea of \textit{Particle Swarm Algorithm (PSO)} is to represent each solution with a $N$ dimension particle from the solution space \cite{PSOFirst}.
-Each particle moves with a direction and velocity $v_{ij}$ based on equations \ref{eq:Swarm1} \& \ref{eq:Swarm}.
+\label{sec:section2}
+In this section we will describe the basic information that is required to
+explain our proposed system.  We will start by explaining how to compute Eigen
+faces then proceed to explain the general Particle swarm algorithm.
 
-\begin{equation}
-%\[
-p_{ij}=p_{ij}+v_{ij},
-%\
-\label{eq:Swarm1}
-\end{equation}
-
-where $p_{ij}$ represent the $j$th dimension in the $i$th particle and $v_{ij}$ is the velocity of the $j$th dimension in the $i$th particle.
- %Equation [\ref{eq:Swarm}] shows how velocity and direction of each particle are computed
- \begin{equation}
-v_{ij}  = v_{ij} \omega + c_1 r_1 (lbest_{ij}  - p_{ij} ) + c_2 r_2 (gbest_{ij}  - p_{ij} ),
-\label{eq:Swarm}
-\end{equation}
- where $\omega$ is the inertia weight parameter which controls the tradeoff between exploration and exploitation,  $lbest_{ij}$ is the local best particle, $gbest_{ij}$ is the global best particle, $r_1$ \& $r_2$ are random variables and $c_1$ \& $c_2$ are the swarm acceleration parameters.
-
- After each iteration the global best $g_{best}$ particle and the agent local best particle $l_{best}$are evaluated based on the maximum fitness functions of all particles in the solution space.
-The solution is found after achieving a specific number of iteration or after an error threshold is achieved.
-Equation \ref{eq:descrite} is used to change the general swarm algorithm into binary particle (\textit{Discrete Particle Swarm Algorithm DPSO}).
-The \textit{DPSO} handles particle values of either $0$ or $1$ \cite{PSODisceret}.
-\begin{equation}
-   P(i)\Leftarrow
-\left \{
-\begin{array}{c}
-1 \quad \quad if\quad r_{3}>p_{i}  \\
-
-0 \quad \quad if\quad r_{3}<p_{i},
-\label{eq:descrite}
-\end{array}\right.
-\end{equation}
- where $p_{ij}$ is the numerical values of the particle and $r_{3}$ is a random variable.
 \subsection{Eigen Faces}
@@
 Where $\Phi_i$ represents the $i^{th}$ face with the mean subtracted from it, $\omega _i$ represent weights and  $u_j$ the eigenvectors.
-
 This can be represented aptly in a figure as:
@@
 \end{enumerate}
+
+\subsection{Particle Swarm Optimization}
+%\section{Discrete Particle Swarm Optimization}
+\label{sec:ParticleSwarmAlgorithm}
+ The main idea of \textit{Particle Swarm Algorithm (PSO)} is to represent each solution with a $N$ dimension particle from the solution space \cite{PSOFirst}.
+Each particle moves with a direction and velocity $v_{ij}$ based on equations \ref{eq:Swarm1} \& \ref{eq:Swarm}.
+
+\begin{equation}
+%\[
+p_{ij}=p_{ij}+v_{ij},
+%\
+\label{eq:Swarm1}
+\end{equation}
+
+where $p_{ij}$ represent the $j$th dimension in the $i$th particle and $v_{ij}$ is the velocity of the $j$th dimension in the $i$th particle.
+ %Equation [\ref{eq:Swarm}] shows how velocity and direction of each particle are computed
+ \begin{equation}
+v_{ij}  = v_{ij} \omega + c_1 r_1 (lbest_{ij}  - p_{ij} ) + c_2 r_2 (gbest_{ij}  - p_{ij} ),
+\label{eq:Swarm}
+\end{equation}
+ where $\omega$ is the inertia weight parameter which controls the tradeoff between exploration and exploitation,  $lbest_{ij}$ is the local best particle, $gbest_{ij}$ is the global best particle, $r_1$ \& $r_2$ are random variables and $c_1$ \& $c_2$ are the swarm acceleration parameters.
+
+ After each iteration the global best $g_{best}$ particle and the agent local best particle $l_{best}$are evaluated based on the maximum fitness functions of all particles in the solution space.
+The solution is found after achieving a specific number of iteration or after an error threshold is achieved.
+Equation \ref{eq:descrite} is used to change the general swarm algorithm into binary particle (\textit{Discrete Particle Swarm Algorithm DPSO}).
+The \textit{DPSO} handles particle values of either $0$ or $1$ \cite{PSODisceret}.
+\begin{equation}
+   P(i)\Leftarrow
+\left \{
+\begin{array}{c}
+1 \quad \quad if\quad r_{3}>p_{i}  \\
+
+0 \quad \quad if\quad r_{3}<p_{i},
+\label{eq:descrite}
+\end{array}\right.
+\end{equation}
+ where $p_{ij}$ is the numerical values of the particle and $r_{3}$ is a random variable.
+
+ \subsection{Nelder-Mead Enhancment}
+The Nelder-Mead simplex algorithm is a derivative-free local search technique
+that is capable of moving a cluster of solutions in the gradient direction and which, as per current research, can be very effectively combined with GA and PSO approaches.
+These hybrid evolutionary algorithms have been shown to be very successful in continuous optimization problems.
+The Nelder-Mead smplex method makes use of a construct called a simplex (see
+Figure 1.).
+When the search space is n-dimensional, the simplex consists of n+1 solutions, s,, i = {1, 2, …, n+1}, that are usually closely spaced.
+As shown in the top left of Figure 1., in a two-dimensional search plane, a simplex is a triangle.
+The fitness of each solution is considered in each step of the Nelder-Mead method, and the worst solution w is identified.
+The centroid, c, of the remaining n points is computed and the reflection of walong it determined.
+This reflection yields a new solution r that replaces w, in the next step, as shown in the top right of Figure 1.
+If the solution r produced by this reflection has a higher fitness than any other solution in the simplex, the simplex is further expanded along the direction of r, as shown in the middle left of the figure.
+On the other hand, if r has a low fitness compared to the others, the simplex is contracted.
+Contraction can be either outward or inward depending upon whether r is better or worse than w.
+The contraction operations are shown in the middle right and bottom left of the figure.
+If neither contraction improves the worst solution in the simplex, the best point in the simplex is computed, and a collapse is then carried out, and all the points of the simplex are moved a little closer towards the best one, as shown in the bottom right of the same figure.
+\section{Proposed System}
+\label{sec:proposed}
+
+Every recognition system consists of trainning and testing modules. Figure
+\ref{Blockchart} shows the flow chart for the training and testing modules.
+The training is divided into three steps 1) preprocessing, 2) feature extraction
+and then 3) clustring using a hybird Particle Swarm Optimization. The follwoing
+few section will discribe in details each step.  %step is the system is divided
+into
+%two major steps 1) preprocessing 2) training
+
+\begin{figure}
+ \begin{center}
+ \centering
+%\includegraphics[scale=0.5]{blockchart.jpg}
+ \subfigure['Training']{\label{fig:block1}\includegraphics[width=0.45\textwidth]{block1.jpg}}
+  \subfigure['Testing']{\label{fig:block2}\includegraphics[width=0.45\textwidth]{block2.jpg}}
+
+\caption{ The flow Chart}
+\label{Blockchart}
+ \end{center}\end{figure}
+\subsection{Preprocessing}
+The preprocessing step function is to detect the face from the input images
+then normalize and the face to the center of each  image.  Every face
+image in the database is  then scalled to the constant size of $h \times w$
+\footnote{ note that  $h$ height and $w$ wdith are emprically choose to be 60
+adn 80 respecitivily.}. The next step is computing the eigenfaces for each
+person in the dataset as explianed in section \ref{facereg}.
+
+\subsection{Feature Extraction}
+
+After the computing the eigenfaces of every person on the database the system
+extracts eigen weight vector for each image. Equation \ref{} shows how the
+weight vector is computed from the eigen faces. These weights are used as
+features for the clustering algorithm.
+
 \subsection{PSO clustering}
-
-On the Merwe and Engelbrechet \cite{psoclustering} Method. The particle $X_i$ is constructed that
+The clustering algorithm is used to train the system to identify the faces in
+the dataset. The clustering algorithm used is based on Particle algorith based
+on the Merwe and Engelbrechet \cite{psoclustering} Method. Each particle
+$X_i$ is constructed such that that
 \begin{equation}
@@
 
-The clustering algorithm prposed by Merwe and Engelbrecht \cite{} can be as follows:
+The clustering algorithm prposed can be as follows:
 % \begin{algorithm}
@@
  \end{algorithmic}
-\section{System}
 
-\begin{figure}[h]
+ Figure shows the \ref{fig:pso1}
+
+\begin{figure}
  \begin{center}
  \centering
-\includegraphics[scale=0.5]{blockchart.jpg} \caption{ The flow Chart}
+%\includegraphics[scale=0.5]{blockchart.jpg}
+ \subfigure['PSO
+ clustering ']{\label{fig:pso1}\includegraphics[width=0.25\textwidth]{pso1.jpg}}
+ \subfigure['Nelder
+ meed a']{\label{fig:pso2}\includegraphics[width=0.25\textwidth]{pso2.jpg}}
+  \subfigure['Nelder
+  meed b']{\label{fig:pso3}\includegraphics[width=0.25\textwidth]{pso3.jpg}}
+\caption{ Particle Clustering Alogirhtm}
+\label{fig:PSOswarm}
  \end{center}\end{figure}
-
 Both previous algorithms  were implemented and tested for correctness.
@@
 The cluster with the min distance is the cluster of the face.
-\subsection{Nelder Enhancment}
 
 \section{Results}
-Two different databases was processed, the first contains 60 indian faces with 11 different samples with different poses and experisions.
+\label{sec:Results}
+Two different databases \cite{JainMukherjeeIndianFaceDB}was processed, the first
+contains 60 indian faces with 11 different samples with different poses and experisions.
 The second database is 120 eroupian faces each has 20 samples.
@@
 \section{Conclusion}
+\label{sec:Conclusion}
 \section{future work}

### 2011-08-12 FacePaper/paper.tex
after ubuntu changes

@@
 
-\documentclass[a4paper, twoside]{article}
+\documentclass[a4paper,twoside]{article}
 
 \usepackage{icgst}
+\usepackage{algorithm}
+\usepackage{algorithmic}
+
 %\documentclass[preprint,12pt]{elsarticle}
@@
 
-Face recognition have become one of the most extensively studied
-research topics that spans different disciplines such as pattern recognition,
-image processing and artificial intelligence. This is due to its importance to
-various application like identity authentication, surveillance and intelligents
-vision. One of the ost used method is egin faces which is used to extract
-relevant facial information from faces to used as features to identificy
-faces\cite{}.
+Face recognition have become one of the most extensively studied research topics that spans different disciplines such as pattern recognition, image processing and artificial intelligence.
+This is due to its importance to various application like identity authentication, surveillance and intelligents vision.
+One of the ost used method is egin faces which is used to extract relevant facial information from faces to used as features to identificy faces\cite{}.
 
-There is various statistical features used to detect like DCT, wavelet transform
-and other information Using data clustering as classification was used. The most
-used feature was used is the eigen faces which transform the face image into
-feature space that represents important facial information.
+There is various statistical features used to detect like DCT, wavelet transform and other information Using data clustering as classification was used.
+The most used feature was used is the eigen faces which transform the face image into feature space that represents important facial information.
+
+The clustring algorithm is widely used to  cluster information in data and text mining and many other application \cite{}.
+The use of clustring algorithm in classificatin was also reproted in \cite{} where kmeans algorithms is used to lablel on the input samples to one of the trained clusters.
+The use of Particle swarm optimization (PSO) in data clustring was introduced by Merwe and Engelbrechet \cite{psoclustering}.
+They presented a system that was test on three different dataset and achieved better clustring than other clustring algorithms.
+Several enhancment has been added to the PSO algorithms to improve its efficiency and quality of the global solution.
+Several other search method like Nelder-Meed \cite{}  were also  integrated to the PSO algorithm.
+The use of Nelder-Meed search in the PSO proofed to enchance the quality of the global optimium solution with little computational cost \cite{}.
+
+In this paper we introduce a novel face recognition algorithm that uses a Particle swarm optimization (PSO) to create clusters of faces in the database.
+The Nelder Meed enchancment is integarted into the PSO algorithm to create clusters where Each cluster represents a single person in the database.
+The first step of the system is preprocessing the database to detect and normalize the face inside the image.
+Secondly, feature vectors are extracted by computing the  eigenfaces for every person in the database then extracting the wieght vector for each sample in the database.
+After that the training set is used to create clusters for the PSO clustering algorithm.
+The global solution should represents a defined center for all clusters where each cluster represent a single person.
+The problem of recognition of test sample is converted into identifying the cluster that the input sample belogs to.
+% step  %and the eigen vector are extracted.
+The  and
 
@@
 
-T
 The remaining of the paper is organized as follows. Firstly, section
@@
   \end{equation}
-, where  $u_j$‘s are Eigenfaces and $\omega_j$ are the weights and can be calculated as :
+, where  $u_j$?s are Eigenfaces and $\omega_j$ are the weights and can be calculated as :
 \begin{equation}
@@
  \subsection{Nelder-Mead Enhancment}
+
+Nelder and Meed introduced the Nelder-Mead simplex algorithm which is a deriavative free  search algorithm that finds the local minimum of function $f$ of several variables.
+For, more than two variable the simplex is triangle and search compares the function values at the three vertex of the tirangle.
+The worst solution, the largest function $f$ value, is rejected and replaced with a new vertex.
+The search continues as new triangle is formed and the process generates squences of triangles.
+%A new triangle  is formed and a new iteration of
+
+A simplex method for finding a local minimum of a function of several variables has
+been devised by Nelder and Mead. For two variables, a simplex is a triangle, and
+the method is a pattern search that compares function values at the three vertices of a
+triangle. The worst vertex, where f (x, y) is largest, is rejected and replaced with a new
+vertex. A new triangle is formed and the search is continued. The process generates
+a sequence of triangles (which might have different shapes), for which the function
+values at the vertices get smaller and smaller. The size of the triangles is reduced and
+the coordinates of the minimum point are found. that %
+
 The Nelder-Mead simplex algorithm is a derivative-free local search technique
@@
 Figure 1.).
-When the search space is n-dimensional, the simplex consists of n+1 solutions, s,, i = {1, 2, …, n+1}, that are usually closely spaced.
+When the search space is n-dimensional, the simplex consists of n+1 solutions, s,, i = {1, 2,\dots  , n+1}, that are usually closely spaced.
 As shown in the top left of Figure 1., in a two-dimensional search plane, a simplex is a triangle.
@@
 %\begin{verbatim}
-
 \STATE  Generate $N$ particles with $N_c$ clusters  ($P_i,$ $\dots,$ $P_{N_c}$)
@@
 
- Figure shows the \ref{fig:pso1}
+ Figure shows the \ref{fig:pso1} the block diagram for the PSO clustering algorithm.
 
@@
 %
-% \subsection{Another subtitle}
+% % \subsection{Another subtitle}
 %

### 2011-08-13 FacePaper/paper.tex
the final from netbook

@@
 The use of clustring algorithm in classificatin was also reproted in \cite{} where kmeans algorithms is used to lablel on the input samples to one of the trained clusters.
+
 The use of Particle swarm optimization (PSO) in data clustring was introduced by Merwe and Engelbrechet \cite{psoclustering}.
@@
 The worst solution, the largest function $f$ value, is rejected and replaced with a new vertex.
-The search continues as new triangle is formed and the process generates squences of triangles.
+The search continues and a new triangle is formed then the process continues to generate a sequences of triangles, for which the function values at the vertices get smaller and smaller.
+These triangles are reduced in size and the coordinate of the best point (minimum) point is found.
+The Nelder-Mead is a technique that is capable of moving a cluster of possible solutions in a gradient direction which can be very effectively combined with GA and PSO approaches.
+Simillar hybrid evolutionary algorithms proved to be very successful in continous optimization problems.
+
     %A new triangle  is formed and a new iteration of
 
-A simplex method for finding a local minimum of a function of several variables has
-been devised by Nelder and Mead. For two variables, a simplex is a triangle, and
-the method is a pattern search that compares function values at the three vertices of a
-triangle. The worst vertex, where f (x, y) is largest, is rejected and replaced with a new
-vertex. A new triangle is formed and the search is continued. The process generates
-a sequence of triangles (which might have different shapes), for which the function
-values at the vertices get smaller and smaller. The size of the triangles is reduced and
-the coordinates of the minimum point are found. that %
+%The Nelder-Mead simplex algorithm is a derivative-free local search technique
+%that is capable of moving a cluster of solutions in the gradient direction and which, as per current research, can be very effectively combined with GA and PSO approaches.
+These hybrid evolutionary algorithms have been shown to be very successful in continuous optimization problems.
 
-The Nelder-Mead simplex algorithm is a derivative-free local search technique
-that is capable of moving a cluster of solutions in the gradient direction and which, as per current research, can be very effectively combined with GA and PSO approaches.
-These hybrid evolutionary algorithms have been shown to be very successful in continuous optimization problems.
-The Nelder-Mead smplex method makes use of a construct called a simplex (see
-Figure 1.).
+%The Nelder-Mead smplex method makes use of a construct called a simplex (see
+%Figure 1.).
 When the search space is n-dimensional, the simplex consists of n+1 solutions, s,, i = {1, 2,\dots  , n+1}, that are usually closely spaced.
@@
 If neither contraction improves the worst solution in the simplex, the best point in the simplex is computed, and a collapse is then carried out, and all the points of the simplex are moved a little closer towards the best one, as shown in the bottom right of the same figure.
+
 \section{Proposed System}
 \label{sec:proposed}
+%Every recognition system consists of trainning and testing modules.
 
-Every recognition system consists of trainning and testing modules. Figure
-\ref{Blockchart} shows the flow chart for the training and testing modules.
-The training is divided into three steps 1) preprocessing, 2) feature extraction
-and then 3) clustring using a hybird Particle Swarm Optimization. The follwoing
-few section will discribe in details each step.  %step is the system is divided
-into
+This paper presents the a face recognition system based on hybird PSO clustering algorithm.
+Figure \ref{Blockchart} shows the flow chart for the training and testing modules.
+The training is divided into three steps 1) preprocessing, 2) feature extraction and then 3) clustring using a hybird Particle Swarm Optimization.
+The follwoing few section will discribe in details each step.
+%step is the system is divided into
 %two major steps 1) preprocessing 2) training

### 2011-08-13 FacePaper/paper.tex
adding most from linux

@@
 
-The remaining of the paper is organized as follows. Firstly, section
-\ref{sec:section2} gives the basic background infromation on  Eigenfaces,
-Particle swarm optimization,and Nelder Meed modification Section
-\ref{sec:proposed} describes in details our proposed system.
+The remaining of the paper is organized as follows.
+Firstly, section \ref{sec:section2} gives the basic background infromation on  Eigenfaces, Particle swarm optimization,and Nelder Meed modification Section \ref{sec:proposed} describes in details our proposed system.
 Section \ref{sec:Results} presents the experiments undertaken to develop the system and also some experiments to evaluate the recognition performance of the proposed system.
@@
 
-%The Nelder-Mead smplex method makes use of a construct called a simplex (see
-%Figure 1.).
-When the search space is n-dimensional, the simplex consists of n+1 solutions, s,, i = {1, 2,\dots  , n+1}, that are usually closely spaced.
-As shown in the top left of Figure 1., in a two-dimensional search plane, a simplex is a triangle.
-The fitness of each solution is considered in each step of the Nelder-Mead method, and the worst solution w is identified.
-The centroid, c, of the remaining n points is computed and the reflection of walong it determined.
-This reflection yields a new solution r that replaces w, in the next step, as shown in the top right of Figure 1.
-If the solution r produced by this reflection has a higher fitness than any other solution in the simplex, the simplex is further expanded along the direction of r, as shown in the middle left of the figure.
-On the other hand, if r has a low fitness compared to the others, the simplex is contracted.
-Contraction can be either outward or inward depending upon whether r is better or worse than w.
+The Nelder-Mead smplex method makes use of a construct called a simplex When the search space is $n$-dimensional, the simplex consists of $n+1$ solutions, $s i = {1,2,\dots  , n+1}$, that are usually closely spaced.
+In a two-dimensional search plane, a simplex is a triangle.
+The fitness of each solution is considered in each step of the Nelder-Mead method, and the worst solution $w$ is identified.
+The centroid, $c$, of the remaining n points is computed and the reflection of $w$ along it determined.
+This reflection yields a new solution $r$ that replaces $w$, in the next step.
+If the solution $r$ produced by this reflection has a higher fitness than any other solution in the simplex, the simplex is further expanded along the direction of $r$.
+On the other hand, if $r$ has a low fitness compared to the others, the simplex is contracted.
+Contraction can be either outward or inward depending upon whether $r$ is better or worse than $w$.
 The contraction operations are shown in the middle right and bottom left of the figure.
-If neither contraction improves the worst solution in the simplex, the best point in the simplex is computed, and a collapse is then carried out, and all the points of the simplex are moved a little closer towards the best one, as shown in the bottom right of the same figure.
+If neither contraction improves the worst solution in the simplex, the best point in the simplex is computed, and a collapse is then carried out, and all the points of the simplex are moved a little closer towards the best one.
 
@@
 \subsection{Preprocessing}
-The preprocessing step function is to detect the face from the input images
-then normalize and the face to the center of each  image.  Every face
-image in the database is  then scalled to the constant size of $h \times w$
-\footnote{ note that  $h$ height and $w$ wdith are emprically choose to be 60
-adn 80 respecitivily.}. The next step is computing the eigenfaces for each
-person in the dataset as explianed in section \ref{facereg}.
+The preprocessing step function is to detect the face from the input images then normalize and the face to the center of each  image.
+Every face image in the database is  then scalled to the constant size of $h \times w$ \footnote{ note that  $h$ height and $w$ wdith are emprically choose to be 60 adn 80 respecitivily.}.
+The next step is computing the eigenfaces for each person in the dataset as explianed in section \ref{facereg}.
 
@@
 
-After the computing the eigenfaces of every person on the database the system
-extracts eigen weight vector for each image. Equation \ref{} shows how the
-weight vector is computed from the eigen faces. These weights are used as
-features for the clustering algorithm.
+After the computing the eigenfaces of every person on the database the system extracts eigen weight vector for each image.
+Equation \ref{} shows how the weight vector is computed from the eigen faces.
+These weights are used as features for the clustering algorithm.
+The output of this stage is a dataset that contains a eigen weight vector for each face in the database which is considered the feature vector of this face.
 
 \subsection{PSO clustering}
-The clustering algorithm is used to train the system to identify the faces in
-the dataset. The clustering algorithm used is based on Particle algorith based
-on the Merwe and Engelbrechet \cite{psoclustering} Method. Each particle
-$X_i$ is constructed such that that
+
+The clustering algorithm is used to group the input faces into clusters which represents the persons in the dataset.
+The process of clustring the input faces is simillar to a classifier training in classification systems.
+The clustering algorithm used is based on Particle Swarm algorithm  based on the Merwe and Engelbrechet \cite{psoclustering} Method where particle represents a possible clustering of the whole dataset.
+Each particle $X_i$ is constructed such that that
 \begin{equation}
@@
  \end{center}\end{figure}
-Both previous algorithms  were implemented and tested for correctness.
+%Both previous algorithms  were implemented and tested for correctness.
 Also the k-means algorithm was used to compare the result of the PSO clustering algorithm.
@@
 The PSO algorithm is used to clusters each face weights computed using equation \ref{weighteq}.
-The algorithms first compute the eigen vectors for all 50 faces as explained in section \ref{facereg}.
-The computed $u_j$ is used to compute the weights for each image in the database is computed which generate a data matrix of 150 X 50 ( number of image X weight vector (no of eigenfaces)).
+The algorithms first compute the eigen vectors for all faces in the database as explained in section \ref{facereg}.
+The computed $u_j$ Eigen vectors is used to compute the weights for each input image in the database.
 The clustering algorithm is used to get center weigh vector for each person.
-I choose the number of clusters same as number of person for easier recognition step.
+%I choose the number of clusters same as number of person for easier recognition step.
+
+Integrating the PSO clustering algorithms with the Nelder Mead algorithm is done using either using Nelder-Mead in a parallel iterative process with the PSO as in figure \ref{fig:pso2}.
+Figure \ref{fig:pso3} shows a sequencial integration of Nelder-Mead search method inside the PSO algorithm where after each iteration of the PSO a single interation of Nelder-Mead is performed to improve the solution.
+Both mehtods are embeded in the system and comparison  between both of them is presented in \ref{sec:Results}, % After the PSO clustering algorithm is constructed two d
 
@@
 \subsection{Experiments Finding PSO paramters}
+\label{sec:pso}
+
+Particle swarm consist of many paramters that controls how the search is operated and have direct impact on the output of the system.
+
 \begin{table}
@@
 \subsection {Comparing With other systems}
+\label{sec:comparingSystems}
+
+\begin{table}
+	\centering
+		\caption{PSO Result }
+	\label{tab:system}
+	%\scalebox{0.99}{
+		\begin{tabular}{|l|c|c|c|c|c|}
+		 \hline
+C1 &	1	&1.2&	1.6&	1.8 &	2	\\ \hline
+	&70.893	&70.893	&70.605	&70.029	&69.741 \\  \hline
+C2	&	1	&1.2&	1.6&	1.8 &	2	\\ \hline
+	&70.89	&70.61	&70.89	&69.45	&69.45 \\ \hline
+W	&0.5&	1	&1.5	&2&	4	\\ \hline
+	&69.74	&70.89&	70.89	&70.32&	70.89	\\ \hline
+\end{tabular}
+		%}
+\end{table}
 
 \subsection{ Comparing Nelder enhancement}
+\label{sec:comparingSystems}
+Figrue \ref{} shows two different method to add Nelder enhancment to the PSO clustering algorithm.
+As mentioned in section \ref{} both methods are evaluated to select best integration methods.
+Table \ref{tab:nelder} shows the result of comparison between the clustering algorithm without Nelder-Mead and using two integration methods.
+The results show that Nelder-Mead enhancment improved the recognition accuracy.
+%achieve
 
-\section{Conclusion}
+\begin{table}
+	\centering
+		\caption{PSO Result }
+	\label{tab:nelder}
+	%\scalebox{0.99}{
+		\begin{tabular}{|l|c|c|c|c|c|}
+		 \hline
+C1 &	1	&1.2&	1.6&	1.8 &	2	\\ \hline
+	&70.893	&70.893	&70.605	&70.029	&69.741 \\  \hline
+C2	&	1	&1.2&	1.6&	1.8 &	2	\\ \hline
+	&70.89	&70.61	&70.89	&69.45	&69.45 \\ \hline
+W	&0.5&	1	&1.5	&2&	4	\\ \hline
+	&69.74	&70.89&	70.89	&70.32&	70.89	\\ \hline
+\end{tabular}
+		%}
+\end{table}
+
+\section{Conclusion and future work}
 \label{sec:Conclusion}
-\section{future work}
+In this paper we introduced a system that recognize faces using a hybrid PSO clustering algorithm. The clustering algorithm is simple and achieved better result than other methods.
+
+Our future work will focus on improving the feature extraction method by using various features like DCT and wavelet features.
+Another improvement is trying to use clustering algorith to problems rather than face recognition.
+
+%\section{future work}
 % \subsection{Subtitle}

### 2011-08-14 FacePaper/files/abstract.tex
finish draft of the paper and now proof reading

@@
 %% Text of abstract
-This paper introduce a system that uses Particle Swarm Optimization
 
+This paper presents a face recognition system using a hybrid clustring algorithm that combines Particle Swarm Optimization and Nelder-Mead search method.
+The system starts by computing the eigenfaces for all faces in the database and then uses eigen weight vectors as feature vectors.
+Next, a particle swarm optimization algorithm is used to group inputs feature vectors into clusters.
+The Nelder-Mead search is integrated in the PSO clustring to efficenctly converge to a global optimum solution.
+The paper uses two datasets to conduct experiments on the system.
+The results shows that the system gives better result compared to K-means clustering
+%The recongition problem is converted into a clustering problem where each person in the database is represented with a cluster. 
+ %PSO algorithm was showed to   The PSO cshows that it successfully converge during the initial global search stages to around global optimum.Various experiements is conducted to proof the efficency of the system
+%successfully converge during the initial stages of a global search, but around
+%global optimum
+
+%where the faces in the database are grouped into clusters using the proposed
+% algorithm. When a new face is introduced to the system, the system tries to
+% identify the cluster that the face belong to and recognize the face based on
+% the cluster
+%
 \end{abstract}
@@
 %\textit{\textbf{Keywords:} Autonomous robot, Stochastic control, Kalman filter,
-%Fuzzy logic, Neural Network, Adaptive navigation.}
+%Fuzzy logic, Neural Network, Adaptive navigation.}ir

### 2011-08-14 FacePaper/paper.tex
finish draft of the paper and now proof reading

@@
 %\input{files/head2}
-
 \section{Introduction}
@@
 This is due to its importance to various application like identity authentication, surveillance and intelligents vision.
-One of the ost used method is egin faces which is used to extract relevant facial information from faces to used as features to identificy faces\cite{}.
+One of the ost used method is Eigenfaces which is used to extract relevant facial information from faces to used as features to identificy faces\cite{Turk91}.
 
-There is various statistical features used to detect like DCT, wavelet transform and other information Using data clustering as classification was used.
-The most used feature was used is the eigen faces which transform the face image into feature space that represents important facial information.
+A lot of algorithms were introduced to solve face recognition problem \cite{apperance2003,survey2003}.
+Several system used evolutionally algorithms like GA \cite{ga2004}.
+In \cite {Ramadan2009} introduced a system that uses Particle Swarm Optimization to select features for face recognition.
+Others like \cite{Wei20114390} used PSO to select best paramters of the Support vector classifier used in classification methods.
+In all of these application PSO was not used as the classifier or the recognizier.
+%ost of these system the PSO is used as anm  %was also used as feature extraction algorithm in fac The use of evolutionally algorithms in face recognition is limited
 
-The clustring algorithm is widely used to  cluster information in data and text mining and many other application \cite{}.
-The use of clustring algorithm in classificatin was also reproted in \cite{} where kmeans algorithms is used to lablel on the input samples to one of the trained clusters.
+%here is various statistical features used to detect like DCT, wavelet transform and other information Using data clustering as classification was used.
+The most used feature was used is the eigen faces which transform the face image into feature space that represents important facial informationa
 
-The use of Particle swarm optimization (PSO) in data clustring was introduced by Merwe and Engelbrechet \cite{psoclustering}.
-They presented a system that was test on three different dataset and achieved better clustring than other clustring algorithms.
+The clustring algorithm is widely used to cluster information in data and text mining and many other application.
+The use of clustring algorithm in classificatin was also reproted in \cite{Bagirov03unsupervisedand} where k-means algorithm is used to create clusters that are used to label input samples using some kind of distance functions.
+% any measure label on the input samples to one of the trained clusters.
+
+The use of Particle swarm optimization (PSO) in data clustring was introduced by Merwe and Engelbrechet \cite{psoclustering,Cheo2004}.
+They presented a system that was tested on three different dataset and achieved better clustring than other clustring algorithms.
 Several enhancment has been added to the PSO algorithms to improve its efficiency and quality of the global solution.
-Several other search method like Nelder-Meed \cite{}  were also  integrated to the PSO algorithm.
-The use of Nelder-Meed search in the PSO proofed to enchance the quality of the global optimium solution with little computational cost \cite{}.
+Several other search method like Nelder-Mead \cite{neldermead}  were also  integrated to the PSO algorithm.
+The use of Nelder-Meed search in the PSO proofed to enchance the quality of the global optimium solution with little computational cost \cite{nelderMeadPSO}.
 
@@
 The first step of the system is preprocessing the database to detect and normalize the face inside the image.
-Secondly, feature vectors are extracted by computing the  eigenfaces for every person in the database then extracting the wieght vector for each sample in the database.
+Secondly, feature vectors are extracted by computing the  Eigenfaces for every person in the database then extracting the wieght vector for each sample in the database.
 After that the training set is used to create clusters for the PSO clustering algorithm.
 The global solution should represents a defined center for all clusters where each cluster represent a single person.
-The problem of recognition of test sample is converted into identifying the cluster that the input sample belogs to.
+The problem of recognition of test sample is converted into measurin the distance between clusters and the input sample, the  sample will belong to the cluster with the minimum distance .
+% widentifying the cluster that the input sample belogs to.
 % step  %and the eigen vector are extracted.
@@
 \label{sec:section2}
-In this section we will describe the basic information that is required to
-explain our proposed system.  We will start by explaining how to compute Eigen
-faces then proceed to explain the general Particle swarm algorithm.
+In this section we will describe the basic information that is required to explain our proposed system.
+We will start by explaining how to compute Eigenfaces then proceed to explain the general Particle swarm algorithm and the Nelder-Mead Modification.
 
-\subsection{Eigen Faces}
+\subsection{Eigenfaces}
 \label{facereg}
-A set of eigenfaces \cite{facesite} can be generated by performing a mathematical process called principal component analysis (PCA) on a large set of images depicting different human faces.
+A set of eigenfaces \cite{Turk91} can be generated by performing a mathematical process called principal component analysis (PCA) on a large set of images depicting different human faces.
 Informally, eigenfaces can be considered a set of "standardized face ingredients", derived from statistical analysis of many pictures of faces.
@@
  \centering
-\includegraphics[scale=0.5]{eigenfacesrec.jpg} \caption{Face reconstruction from eigen faces. }
+\includegraphics[scale=0.3]{eigenfacesrec.jpg} \caption{Face reconstruction from Eigenfaces. }
  \end{center}\end{figure}
 
-\subsubsection{Assumptions:}
-
+There are some main assumption we have to take in account before computing the eigen computatioan and it as follows :%in computing eigen faces and it is 
+%\subsubsection{Assumptions:}
 \begin{enumerate}
@@
 The same logic would apply to images that are not of equal length and breadths.
-To take an example: An image of size 112 x 112 can be represented as a vector of dimension 12544 or simply as a point in a 12544 dimensional space.
-
+To take an example: An image of size $112 x 112$ can be represented as a vector of dimension 12544 or simply as a point in a 12544 dimensional space.
 \end{enumerate}
-
 \subsubsection{Algorithm for Finding Eigenfaces}
@@
  \centering
-\includegraphics[scale=0.5]{training-images.jpg} \caption{Sample Faces. }
+\includegraphics[scale=0.3]{training-images.jpg} \caption{Sample Faces. }
  \end{center}\end{figure}
@@
 
-\item Find the $M$ best  Eigenvectors of $C=A^TA$.
-using the relation of $u_i=Av_i$  Also keep in mind that $\left\| {u_i} \right\| = 1.$ where  $u_i$ is the eigen vector of $C$ and $v_i$ is the eigen value of the matrix $C$.
+\item Find the $M$ best  Eigenvectors of $C=A^TA$. using the relation of in Equation\ref{eigenu} Also keep in mind that $\left\| {u_i} \right\| = 1.$.
+
+\begin{equation}
+ \label{eigenu}
+u_i=Av_i
+\end{equation}
+
+ where  $u_i$ is the eigen vector of $C$ and $v_i$ is the eigen value of the matrix $C$.
 
@@
   \begin{equation}
+\label{eqphi}
 \Phi_i  = \sum\nolimits_{j = 1}^k {\omega_j u_j}
@@
 
+Equation \ref{eigenu} and \ref{weighteq} are used to compute the eigenfaces and weight equation for each face in the database. 
 \subsection{Particle Swarm Optimization}
@@
 The solution is found after achieving a specific number of iteration or after an error threshold is achieved.
-Equation \ref{eq:descrite} is used to change the general swarm algorithm into binary particle (\textit{Discrete Particle Swarm Algorithm DPSO}).
+% Equation \ref{eq:descrite} is used to change the general swarm algorithm into binary particle (\textit{Discrete Particle Swarm Algorithm DPSO}).
 The \textit{DPSO} handles particle values of either $0$ or $1$ \cite{PSODisceret}.
-\begin{equation}
-   P(i)\Leftarrow
-\left \{
-\begin{array}{c}
-1 \quad \quad if\quad r_{3}>p_{i}  \\
-
-0 \quad \quad if\quad r_{3}<p_{i},
-\label{eq:descrite}
-\end{array}\right.
-\end{equation}
- where $p_{ij}$ is the numerical values of the particle and $r_{3}$ is a random variable.
-
+% \begin{equation}
+%    P(i)\Leftarrow
+% \left \{
+% \begin{array}{c}
+% 1 \quad \quad if\quad r_{3}>p_{i}  \\
+% 
+% 0 \quad \quad if\quad r_{3}<p_{i},
+% \label{eq:descrite}
+% \end{array}\right.
+% \end{equation}
+%  where $p_{ij}$ is the numerical values of the particle and $r_{3}$ is a random variable.
  \subsection{Nelder-Mead Enhancment}
@@
 \subsection{PSO clustering}
+\label{sec:psoClustering}
 
@@
 where $F_c$ is the fitness function, $N_c$ is the number of clusters in the particle, $Z_p$ denotes the $p^{th}$ data vector $\left| {C_{ij} } \right|$ is the number of data vectors belonging to the cluster $C_{ij}$ and $d$ is the Euclidian distance between $Z_p$ and $m_{ij}$.
+
+Other fitness function are also used to take into account the size of cluster and the number of elements in each cluster which improve the clustring algorithm.
+The two different fittness used are as follows:
+%F  = \left{\sum\nolimits_{j = 1}^{N_c } {\sum\limits_{\forall z_j \in C_{ij}}  } \right} } %{\left[ {\frac{{d(z_p ,m_{ij} )}}{{\left| {C_{ij} } \right|}}} } \right]} }}{{N_c }}
+
+\begin{equation}
+F2 =  \left\{ \sum\nolimits_{j=1}^{N_c} \left[ \left( \sum\limits_{\forall z_j \in C_{ij}} d(z_p ,m_{ij} )/\left|{C_{ij} }\right| \right) \times  \left(  \left|{C_{ij} }\right| /N_o\right)\right]\right\}
+\end{equation}
+where $F2$ is the second fitness function, $N_o$ is the number of data vectors to cluster.
+
+by adding considration of the cluster distribution the equation could be changed into :
+
+\begin{equation}
+F3 =  F2\times \left(\left|{C_{ik} }\right|-\left|{C_{il} }\right|+1 \right)
+\end{equation}
+where $\left|{C_{ik} }\right|$ is computed as $\max\forall_{j=1,2 \dots N_c}$ and $\left|{C_{il} }\right|=\min\forall_{j=1,2 \dots N_c}$   
+%, $Z_p$ denotes the $p^{th}$ data vector $\left| {C_{ij} } \right|$ is the number of data vectors belonging to the cluster $C_{ij}$ and $d$ is the Euclidian distance between $Z_p$ and $m_{ij}$.
 
@@
 \label{sec:Results}
-Two different databases \cite{JainMukherjeeIndianFaceDB}was processed, the first
-contains 60 indian faces with 11 different samples with different poses and experisions.
-The second database is 120 eroupian faces each has 20 samples.
-The database is processed to the input format for the particle algorithm.
+Two different databases  was processed, the first contains 60 indian faces frome the indian face database \cite{JainMukherjeeIndianFaceDB} with 11 different samples per person  with different poses, lighting conditions and experisions.
+The second database is 120 eroupian faces from database \cite{} with 20 different samples per person.
+Both databases are processed to normalize the input to the algorithm.
+The dataset is then divided into training and testing set by dividing the samples of each person to either training or testing.
+Four different  random training - testing splits generate the result of the following experiments .
+The recognition result presented is the average of the four runs of the four different splits.
+In the next section recognition rate is number of corrrectly identified input images over the total number of face in the testset.
+
+  %The database is processed to the input format for the particle algorithm.
 The data is then divided into training and testset.
-The percentage can be changed and changed to cross validate the results and experiments.
+The percentage can be changed and changed to cross validate the results and experiments.are used to generate the result of the following experiments.
 
@@
 Particle swarm consist of many paramters that controls how the search is operated and have direct impact on the output of the system.
+Table \ref{tab:PSO1} shows the result of different PSO paramters $c_1,c_2,W$.
+The result gives the indication that the swarm moves quickly to the global best by the effect of $c2$ is more improtant than $c1$.
 
@@
 
+Another experiment to determine the system paramter is the number of training smaples required to get good results.
+Table \ref{tab:PSOtrainTest} shows the result when the system used different number of trainig samples per person.
+It is clear from the results that the system show good recognition result with only few trianig samples.
+
+\begin{table}
+	\centering
+		\caption{PSO Result }
+	\label{tab:PSOtrainTest}
+	%\scalebox{0.99}{
+		\begin{tabular}{|l|c|c|c|c|c|}
+		 \hline
+C1 &	1	&1.2&	1.6&	1.8 &	2	\\ \hline
+	&70.893	&70.893	&70.605	&70.029	&69.741 \\  \hline
+C2	&	1	&1.2&	1.6&	1.8 &	2	\\ \hline
+
+\end{tabular}
+		%}
+\end{table}
+
 \subsection {Comparing With other systems}
 \label{sec:comparingSystems}
+
+The system is compared with kmeans clustering algoirthm and general eigne faces system.
+The kmeans algoirthm is similar clustering algorithm and provides similar classification via clustering algorithms.
+Eigenfacs is genearly classifieed using eculidean distance.
+Table \ref{tab:system} shows the result of comparing the results of the proposed systems and kmeans.
+The results shows that our proposed alogrithm is more superior than kmeans clustering algoirhtms.
 
@@
 
+The efficency of the clustring algorithm is also  tested using differen fitness functions (see section \ref{sec:psoClustering}).
+Table \ref{tab:fitness}  shows the result of three different fitness functions and the effect of each fitness on the final recognition rate.
+
+\begin{table}
+	\centering
+		\caption{PSO Result }
+	\label{tab:fitness}
+	%\scalebox{0.99}{
+		\begin{tabular}{|l|c|c|c|c|c|}
+		 \hline
+C1 &	1	&1.2&	1.6&	1.8 &	2	\\ \hline
+	&70.893	&70.893	&70.605	&70.029	&69.741 \\  \hline
+C2	&	1	&1.2&	1.6&	1.8 &	2	\\ \hline
+	&70.89	&70.61	&70.89	&69.45	&69.45 \\ \hline
+
+\end{tabular}
+		%}
+\end{table}
+
 \subsection{ Comparing Nelder enhancement}
@@
 	\centering
-		\caption{PSO Result }
+		\caption{ Nelder enhancement  }
 	\label{tab:nelder}
@@
 C2	&	1	&1.2&	1.6&	1.8 &	2	\\ \hline
-	&70.89	&70.61	&70.89	&69.45	&69.45 \\ \hline
-W	&0.5&	1	&1.5	&2&	4	\\ \hline
-	&69.74	&70.89&	70.89	&70.32&	70.89	\\ \hline
+
 \end{tabular}
-		%}
+
 \end{table}
@@
 % More plain text.
-\bibliographystyle{model1a-num-names}
+\bibliographystyle{plain}
+%\bibliographystyle{model1a-num-names}
 \bibliography{mybibfile,library}

### 2012-01-20 LandminePaper/landmine.tex
Intial import of the paper

%%This is a very basic article template.
%%There is just one section and two subsections.

\documentclass[conference]{IEEEtran}

\usepackage{times}

% *** CITATION PACKAGES ***
%
\usepackage{cite}

% *** GRAPHICS RELATED PACKAGES ***
%
\ifCLASSINFOpdf
  \usepackage[pdftex]{graphicx}
\else
   \usepackage[dvips]{graphicx}
\fi

\usepackage[tight,footnotesize]{subfigure}
%\usepackage[caption=false]{caption}
%\usepackage[font=footnotesize]{subfig}

% *** PDF, URL AND HYPERLINK PACKAGES ***
%
\usepackage{url}
\usepackage[cmex10]{amsmath}

% correct bad hyphenation here
\hyphenation{op-tical net-works }

\begin{document}

\title{Image based Landmine detection}

%\section{Title}
% author names and affiliations
% use a multiple column layout for up to three different
% affiliations
\author{
\IEEEauthorblockN{Maha El Meseery}
\IEEEauthorblockA{Computer and Systems Department\\Electronic Research
Institute, Cairo, Egypt\\ Email: melmeseery@sci.eri.eg}
\and
\IEEEauthorblockN{Mahmoud fakhr el Deen}
\IEEEauthorblockA{Computer and Systems Department\\Electronic Research
Institute, Cairo, Egypt\\Email: mk@sci.eri.eg}
%\and
}
% make the title area
\maketitle

\begin{abstract}
%\boldmath
This paper presents a novel system for detection of landmines from
ground-penetrating radar (GPR) images.  A  multistatic ground-penetrating radar
(GPR) system produces scanning data of a several buried landmines on clean and
cluttered surfaces. The GPR time signal data is transformed into downcross
images of the soil and buried targets. Various image based techquies are used
to process these images to extract landmine characteristics.  A support
vector machine classifier (SVM) is used to identify the landmines from
the clutter and other buried objects in soil. The results show that the system
performe efficiently even in highly cluttered enviroment. % These images areprocessed

\end{abstract}
% IEEEtran.cls defaults to using nonbold math in the Abstract.
% This preserves the distinction between vectors and scalars. However,
% if the conference you are submitting to favors bold math in the abstract,
% then you can use LaTeX's standard command \boldmath at the very start
% of the abstract to achieve this. Many IEEE journals/conferences frown on
% math in the abstract anyway.

% no keywords

% For peer review papers, you can put extra information on the cover
% page as needed:
% \ifCLASSOPTIONpeerreview
% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
% \fi
%
% For peerreview papers, this IEEEtran command inserts a page break and
% creates the second title. It will be ignored for other modes.
\IEEEpeerreviewmaketitle
\section{Introduction}
\label{sec:introduction}
%%%%% firs some take %%% what is landmine detection \ldots.
One of most terrifying legacies of war that humanity faces is buired landmine
problem. Armed forces uses landmine  to destroy or disable enemy targets while
they pass over or near the landmine device. The problems rise after war and
armies move, the land continues to contains landmines for years. The
International Campaign to Ban Landmines (ICBL) [9] estimates that every year
there is more than 25,000 people killed or maimed by mines many of them are
children. A total estimate of 45-40 millions mines remains to be cleared
worldwoide [10]. One of the problems of mine removal or the demining is
high cost of the process. The cost of removing a single mine ranges from  300\$ to 1000\$
unlike the cost of laying a typical anti-personnel mine which ranges from 3\$ to
30\$.

Mines are usually made using metallic and non metallic materials and they in
general are either anti-presonal or anti-Tank mines. The use on non metallic
materials in manufacturing mines renders the conventional metal detector non
efficeints. Therefore researchers studied various other techniques to detect and remove landmines  \cite{Ho2002,Tan2005,Potin2006}.  Lately, several techniques in the area of sensor physics, signal processing, nano techonolgy and robotics were studied.  Ground penetrating radar (GPR), Infrared (IR), Sesiemic,  and ultrasound (US) sensors were developed and reviewed \cite{Scott2004,Ng2008}. These sensor usually generate a signal that requires different types of processing as noise removal, filtering, enhancement and feature extractions\cite{Ng2008,Potin2006}. Although systems can achieve high detection rates, they have done so at the expense of high false alarm rates.   % invistigated

The mine detection algorithm is similar to many automation and recognitio system, it has four main components: 1) reading sensor data 2) preprocessing 3) feature extraction and 4)classifier desing. The first step of any landmine system is reading and scanning the tested surface. After using the sensor to measure the input data, a preprocessing step focuses on normalizing data, correcting vaation and removal of stationary or latitude corrections. The next step, feature extraction, perform image or signal procesisng methods to extract the mine characterstics that are introduced to the last step which identify and localize the landemine.

This paper introduce a landmine detection system that uses data collected from a multistatic   Ground penetrating radar (GPR)

% A typical land-mine detection system, like most automated detection algorithms, has three main components: 1) preprocessing; 2) feature extraction; and 3) classifier design. The preprocessing step relies on algorithms that perform tasks such as normalization of the data, corrections for variations in height and speed, removal of stationary effects due to the system response, etc. Methods that have been used to perform this task

% Various techniques in the area of sensor physics, signal processing, and
% robotics have been studied during the last decade.  Most mine detection
% techniques depend on sensors, signal processing,  and  decision  processes.  For
%  the  sensor part,  ground  penetration  radar  (GPR),  infrared  (IR),  and
% ultrasound  (US)  sensors  are  reviewed.  For  the  signal processing  and
% decision  parts,  a  commonly  used  set  of image  processing  techniques
% including  filtering, enhancement,  feature  extraction,  and  segmentation  are
% surveyed.  Segmentation  is  used  to  extract  a  mine  signal from  various
% competing  signals,  the  mine  detection algorithms work with inhomogeneous
% background.

% Detection and removal of land-mines is therefore a significant problem and has attracted several
% researchers in recent years \cite{Ho2002,Tan2005,Potin2006,Savelyev2007,Collins1999,Ng2008}. Although systems can achieve high
% detec-tion rates, they have done so at the expense of high false alarm rates.
% Various techniques in the area of sensor physics, signal  processing, and robotics have been studied during the last  decade.  Most mine detection techniques depend on sensors,  signal processing,  and  decision  processes.  For  the  sensor  part,  ground  penetration  radar  (GPR),  infrared  (IR),  and  ultrasound  (US)  sensors  are  reviewed.  For  the  signal  processing  and  decision  parts,  a  commonly  used  set  of  image  processing  techniques  including  filtering,  enhancement,  feature  extraction,  and  segmentation  are  surveyed.  Segmentation  is  used  to  extract  a  mine  signal  from  various  competing  signals,  the  mine  detection  algorithms work with inhomogeneous background.

% Because mines can be made of both metallic and

% The
% cost to purchase and lay a typical antipersonnel mine ranges from  $3 to  $30, while the cost to remove a single mine ranges from $300 to
% $1000. Because mines can be made of both metallic and
% nonmetallic  materials,  detection  using  only  conventional
% metal detectors cannot give a promising result.
% the process of detecting and
% removing mines, called demining, is particularly important.
% Manual  demining  is  extremely  dangerous

%Throught war the armed forces
%usually burry landmine just below surface of the ground.

%A land mine is an explosive device, concealed under or on the ground and
% %designed to destroy or disable enemy targets as they pass over or near the
% %device. Such devices are typically detonated automatically by way of pressure
% from the target stepping or driving on it, though other detonation mechanisms may be possible.[1] The device may cause damage either by a direct blast or by fragments that are thrown by the blast.

% Landmines are one of the worst environmental problems that
% humanity faces long with being one of the most terrifying legacies of war.
%
% Landmines
% are made to be used by armed forces to disable any person or vehicle that comes
% close with it by an explosion or fragments released at high speeds, so they become a
% burden to those who have to support them
%
% MORE  than  26,000  people  are  killed  or  maimed  by
% mines every year, which is equivalent to one victim,
% every 20 min. For example, in Cambodia one out of every
% 236 people is a landmine ampute
% e. The casualty ratio rises
% to one out of every 140 people in Angola, which has more
% mines  than  people  in  addition  to  fatal  casualties  and
% enormous  financial  losses.  In  Cambodia,  approximately
% 40\%  of  the  rice  fields  have  been  mined  and  abandoned.
% Most  tragic  is  that  many  victims  are  children  and  most
% mine-afflicted countries are poverty stricken, as well.

% Because  of  the  potentially  catastrophic  results  of
% unintentional mine encounters, the process of detecting and
% removing mines, called demining, is particularly important.
% Manual  demining  is  extremely  dangerous;  around  one
% deminer has been killed  for every 2,000 mines  removed,
% with even more civilian victims. The cost to purchase and
% lay  a  typical  antipersonnel  mine  ranges  from  $3  to  $30,
% while the cost to remove a single mine ranges from $300 to
% $1000. Because mines can be made of both metallic and
% nonmetallic  materials,  detection  using  only  conventional
% metal detectors cannot give a promising result. Report also
% indicate  metal  detectors  are  subject  to  many  false  Image
% Processing-Based.  High  labor  cost  and  the  slow  pace
% involved are encouraging development of other techniques.
% Although  some  military  demining  equipment  has  been
% developed and used during the Gulf War by the US Army,
% civilian related demining, called humanitarian demining, is
% quite different from the military work

% Landmines are basically explosive devices designed to explode when triggered by
% pressure or a tripwire. These devices are typically found on or just below the surface
% of the ground [8]. .
% The International Campaign to Ban Landmines (ICBL) [9] estimates that 15,000-20,000 people are killed or injured by land mines per year many of them children.
% The U.S. State Department estimates that a total of 45-50 million mines remain to
% be cleared. Worldwide, approximately 100,000 mines are cleared each year [10]. At
% that rate, clearing all 45-50 million mines will require 450- 500 years, assuming no
% new mines are laid. By some estimates, roughly 1.9 million new mines are emplaced
% annually, yielding an additional 19 years of mine clearance work every year.
% The Landmine Monitor Report 2003 identied over 11,700 new landmine victims
% in 2002. Of this number at least 2,649 were children, a staggering 23 percent. More
% than 85 percent were innocent civilians

% The  object  of  humanitarian  demining  is  to  find  and
% remove  abandoned  landmines  without  any  hazard  to  the
% environment.  These  landmines  were  intended  for  military
% use  when  they  were  planted,  but  their  duty  has  expired.
% Furthermore, humanitarian demining equipment is required
% to  be  more  accurate  than  military-purpose  equipment
% because the military can afford a certain degree of casualty
% risk.

% Various techniques in the area of sensor physics, signal
% processing, and robotics have been studied during the last
% decade.  Most mine detection techniques depend on sensors,
% signal processing,  and  decision  processes.  For  the  sensor
% part,  ground  penetration  radar  (GPR),  infrared  (IR),  and
% ultrasound  (US)  sensors  are  reviewed.  For  the  signal
% processing  and  decision  parts,  a  commonly  used  set  of
% image  processing  techniques  including  filtering,
% enhancement,  feature  extraction,  and  segmentation  are
% surveyed.  Segmentation  is  used  to  extract  a  mine  signal
% from  various  competing  signals,  the  mine  detection
% algorithms work with inhomogeneous background.
%
%
% DETECTION, localization, and subsequent neutralization of buried land-mines are
% worldwide humanitarian and military problems. The latest statistics show that,
% as of August 2009, more than 70 states were believed to be mine affected
% Landmine Monitor has identified at least 73 576 casualties of land-mines,
% explosive remnants of war, and victim-activated improvised explosive devices in
% 119 states and areas in the past ten years \cite{report}.
%
% Detection and removal of land-mines is therefore a significant problem and has attracted several
% researchers in recent years \cite{Ho2002,Tan2005,Potin2006,Savelyev2007,Collins1999,Ng2008}. Although systems can achieve high
% detec-tion rates, they have done so at the expense of high false alarm rates.
% Various techniques in the area of sensor physics, signal  processing, and robotics have been studied during the last  decade.  Most mine detection techniques depend on sensors,  signal processing,  and  decision  processes.  For  the  sensor  part,  ground  penetration  radar  (GPR),  infrared  (IR),  and  ultrasound  (US)  sensors  are  reviewed.  For  the  signal  processing  and  decision  parts,  a  commonly  used  set  of  image  processing  techniques  including  filtering,  enhancement,  feature  extraction,  and  segmentation  are  surveyed.  Segmentation  is  used  to  extract  a  mine  signal  from  various  competing  signals,  the  mine  detection  algorithms work with inhomogeneous background.
%
%
%
% A typical land-mine detection system, like most automated detection algorithms, has three main components: 1) preprocessing; 2) feature extraction; and 3) classifier design. The preprocessing step relies on algorithms that perform tasks such as normalization of the data, corrections for variations in height and speed, removal of stationary effects due to the system response, etc. Methods that have been used to perform this task

In  section  2  a  brief  history  for  mines  detection  is  introduced,  in  section  3,  a  complete  description  for  the  proposed  classification  algorithm  among  some  predefined  mine types is  demonstrated. In section  4 we present  and  discuss the results. In section 5 the work conclusion will be  given

\section{GPR Landmine Detection Overview}
%Plain text.

 Ground-penetrating radar (GPR) and electromagnetic induction (EMI) are two of the extensively investigated sensors due to their complementary features. For instance, GPR has the advantage of detecting mines with plastic or low-metal-content mines that are difficult to detect by traditional metal detectors. On the other hand, detectors using EMI sensors can outperform those using GPR in detecting small antipersonnel (AP) mines and deeply buried mines [8]. In fact, even for the same sen-sor, different detection algorithms may be more effective for different mine types, burial depths, and soil conditions \cite{Frigui2010}. For instance, in \cite{Frigui2010}, it was shown that a detector that uses edge-based features \cite{Frigui2009} outperforms the spectral detector \cite{Ho2008} for antitank (AT) mines and shallow mines. On the other hand, the spectral detector outperforms the edge-based detector for weak scattering plastic mines \cite{Ho2008}. In [11], using a large GPR data set collected from four different sites, it was shown that the relative performance of four different detection algorithms is highly dependent on the site

\section {System Overview}

 \begin{figure}
\centering
\label{fig:block}
\includegraphics[width=0.45\textwidth]{images/BlockDiagram.jpg}
 \caption{The Block diagram }
\end{figure}

\subsection{GPR processing}

A multistatic ground-penetrating radar (GPR) sys-tem has been developed and used to measure the response of a
number of targets to produce data for the investigation of multi-static inversion algorithms. The system consists of a linear array
of resistive-vee antennas, microwave switches, a vector network
analyzer, and a 3-D positioner, all under computer control. The
array has two transmitters and four receivers which provide eight
bistatic spacings from 12 to 96 cm in 12-cm increments. Buried
targets are scanned with and without surface clutter, which is a
layer of rocks whose spacing is empirically chosen to maximize the
clutter effect. The measured responses are calibrated so that the
direct coupling in the system is removed, and the signal reference
point is located at the antenna drive point. Images are formed us-ing a frequency-domain beamforming algorithm that compensates
for the phase response of the antennas. Images of targets in air
validate the system calibration and the imaging algorithm. Bistatic
and multistatic images for the buried targets are very good, and
they show the effectiveness of the system and processing.

%%% this is from multi experiement paper \ldots.
The multistatic GPR in this paper consists of a linear array
of resistive-vee antennas, a microwave switch matrix, a vector
network analyzer (Agilent 8720D), and a 3-D positioner, all
under computer control (Fig. 1). The resistive-vee antenna ishosen as an array element because it can transmit and receive
short pulses with a negligible amount of clutter, and it has a
low radar cross section, which eliminates most of the multiple
reflections between the antenna and the ground [14]–[16]. In
addition, it is lightweight; therefore, an array of the resistive-vee antennas can be made in a hand-held form. The antenna
was optimized for a bistatic GPR.

he GPR is scanned over a 1.8 × 1.8 m region at a constant
height above the surface of the ground. The scan region is
referenced by x- and y -coordinates, both ranging from −90 to
90 cm in 2-cm increments. Thus, the scan region is discretized into a grid of 91 points by 91 points. The location of the
GPR array in the coordinate system is referenced to the center
point of transmitter T1. Each time the GPR array stops, it
collects data from the eight bistatic spacings by manipulating
switches in an appropriate order. After each switch operation,
the network analyzer sweeps 401 equally spaced frequency
points from 60 MHz to 8.06 GHz. This procedure at each
point takes approximately 6 s. The total scan time is approxi-mately 14 h.

The lowest effective frequency is in the range of 500 MHz–
1.0 GHz because the gain of the antenna falls off quickly in this
frequency range [20]

The data calibrated according to (1) are stored in Matlab
file format. Each Matlab file contains one set of measured and
calibrated data obtained from a single scan of 1.8 × 1.8 m
region. The stored variables are responses from eight bistatic
spacings (401 × 91 × 91 double precision), frequency points at
which the data are measured (401 × 1 double precision), and
x- andy -coordinates of each measurement (91 × 1).

The multistatic GPR is operated on five sets of buried targets.
The target sets are an assortment of metal spheres, a variety
of AT and AP landmines, a Plexiglas box, a plastic pipe,
and empty sand. Table I contains more information about the
landmine and clutter objects. Target depths range from 1 to
58 cm; the depth of a target is measured from the surface of
the sand to the top of the target. The ground medium is damp
compacted sand with a water table below target level.
Each set of targets is scanned with and without surface clutter
according to the following procedure (the buried pipe is done
only with a clean surface). First, the targets are buried, and
the surface of the sand is packed so that the height of the
surface varies less than 2 cm. The sand is then soaked with
water and let stabilize for approximately 2 h. This process is
performed to maintain appropriate moisture level in the sand
and to minimize the disturbance made in the sand when the
targets are buried. The responses of the targets in sand with
a clean surface are obtained as the GPR scans over the sand.
Next, river rocks are randomly scattered over the scan region,
as shown in Fig. 12. The density of the rocks is empirically
chosen to maximize the clutter effect. In both the clean-surface
and cluttered-surface scans, the antenna array is elevated by
10 cm, which is measured from the surface of the sand to the
bottom of the antenna array.

 \begin{figure}
\centering
\label{fig:minloc}
\includegraphics[width=0.45\textwidth]{images/MineLocations.jpg}
 \caption{Mine Locations}
\end{figure}

 \begin{figure}
\centering
\label{fig:mineshapes}
\includegraphics[width=0.45\textwidth]{images/MineShapes.jpg}
 \caption{ Sample of mines  }
\end{figure}

More plain text
\subsection{Image processing}

Gradient Features
To extract gradient features [3], the gradient operator is first applied to the gray-scale image of the digit to give two gradient components: strength |g(x,y)| and direction  at each point (x,y) of the image f. This is done by applying Sobel operator [20] on the image to extract vertical and horizontal gradient components,
 ,	(2)
 ,	(3)
where f(x,y) is the intensity of image f at point (x,y), gx(x,y) is the gradient component at x-direction at location (x,y), and gy(x,y) is the gradient component at y-direction at location (x,y).
Then the gradient strength and direction is extracted using equations (4) and (5), respectively,
 		(4)
 			(5)
The gradient vector g(x,y) (expressed as strength strength |g(x,y)| and direction  ) at each point (x,y) of the image is then decomposed into the 8 Freeman [20] directions shown in Fig. 2. The gradient vector is decomposed into the 8 Freeman directions by projecting the vector into the nearest two Freeman directions as shown in Fig. 3.
The gradient features are composed of 8 layers; each corresponding to one of the Freeman directions. Each layer is the projection of the gradient vectors of the image into the corresponding Freeman direction.
A Gaussian mask h(x,y) is then applied to each layer,
 ,		(6)
and then the layers are uniformly sampled to give 5×5 measurements. The parameter σ is related to the sampling interval t via the empirical relation  [3]. This means that the gradient features are composed of 8×5×5 = 200 elements. The gradient feature extraction technique is very powerful as shown in [3] and as will be shown in the results section (section 5) of this paper.
4.2 Kirsch Features
The Kirsch features [3] are extracted by decomposing the image into 4 layers corresponding to 4 edge orientations: horizontal (gh), vertical (gv) , right diagonal (grd), and left diagonal (gld). These layers are extracted by applying Kirsch masks on the image,
 ,		(7)
 ,		(8)
 ,	(9)
 ,	(10)
where * denotes 2D convolution operation [20], kh1 and kh2 are Kirsch masks responsible for extracting horizontal edges, kv1 and kv2 for horizontal edges, krd1 and krd2 for right diagonal edges, and kld1 and kld2 for left diagonal edges. Figure 4 shows all Kirsch masks.
gh1
5	5	5
-3	0	-3
-3	-3	-3
	gh2
-3	-3	-3
-3	0	-3
5	5	5
	gv1
5	-3	-3
5	0	-3
-3	-3	-3
	gv2
-3	-3	5
-3	0	5
-3	-3	5

grd1
-3	-3	-3
5	0	-3
5	5	-3
	grd2
-3	5	5
-3	0	5
-3	-3	-3
	gld1
-3	-3	-3
-3	0	5
-3	5	5
	gld2
5	5	-3
5	0	-3
-3	-3	-3

Figure 4. Kirsch  Masks

A Gaussian mask is then applied to all the 4 layers and 5×5 measurements are extracted from each of them in the same manner done with gradient features (section 4.1). This means that Kirsch features are composed of 4×5×5=100 elements.
4.3 Local Chain Code Features
To extract the local chain code features [21], the contour of the digit is first followed by any contour following algorithm, and then each contour point of the image is marked with the corresponding Freeman direction. The image is then decomposed into 8 layers; each layer contains only contour points that belong to the corresponding Freeman direction. Each of the 8 layers is then uniformly partitioned into 5×5 zones. Each zone is averaged leading to a feature vector of 200 elements.

4.4 Wavelet Features
Wavelet transform [20] extracts the essential information contained in a signal or image. This may help in reducing the dimensionality of the problem which leads to faster classification. Also it might reduce the noise and the non-essential information that may confuse the classifier leading to better classification performance. Wavelet transform could be applied on the image directly or could be applied on the gradient decomposition of the image [22]. In our study, we applied the former method. The image is first resized to be 64×64 and then the image is composed into 3 resolution levels. The 3rd level approximation of the image (which is a 8×8 image) is then used as features. This means that wavelet features are composed of 64 elements.
4.5 Concatenation of low dimensional features
Here we concatenate 6 families of low dimensional features to form one feature vector. While each of these low dimensional features is not very effective, their concatenation proved to be a powerful feature vector. Each of the 6 low dimensional feature families is going to be discussed in the following subsections.
4.5.1 Raw Image Zoning
In this technique, the image is uniformly partitioned into 5×5 zones [23]. The average of each zone is calculated leading to a feature vector of 25 elements.

4.5.2 Vertical and Horizontal Projections
In this technique, the horizontal and vertical histograms of the image are calculated leading to a feature vector of 40 elements [23] (Remember that we are working on the 20×20 window that confines the digit not all 28×28 pixels of the MADBase digits).

4.5.3 Vertical and Horizontal Cross Counts
The image is first binarized, and then the vertical and horizontal cross counts are calculated leading to a feature vector of 40 elements.
4.5.4 Centroidal Distances
In this technique [11, 24], the digit image is partitioned into 16 sectors around the digit center of gravity (the centroid) as shown in Figure 5. Then the centroid of each sector are calculated. The distances between the whole digit centroid and sectors centroids are calculated and normalized by dividing them by the digit bounding box diameter. This forms a 16-element feature vector.
4.5.5 Directional Features
The directional features are formed as follows [11, 25]. Each background pixel is labeled by a 4-element vector. For each background pixel we walk upward; if a foreground pixel is found, the first element of the 4-element vector is set to ‘1’ otherwise set to ‘0’. Then we walk to the right; if a foreground pixel is found, the second elements is set to ‘1’ otherwise set to ‘0’; and so on till we finish the 4 principal directions. Then for each combination of 0’s and 1’s in the 4-element vector, we give a different label. Hence, we have 16 different labels. Now each of foreground pixels has one label of the 16 labels. The final feature vector is the histogram of such labels giving rise to a 16 element feature vector.
4.5.6 Length-Normalized Contour
In this technique [11], the horizontal and vertical coordinates of the boundary pixels of the digit are used as features. First, the boundary of the digit is extracted using a contour following algorithm; then, each boundary point is stored by its horizontal and vertical coordinates. Horizontal locations are magnitude-normalized by dividing them by the width of the digit bounding box. Vertical locations are also magnitude-normalized by dividing them by the height of the bounding box. Then both horizontal and vertical locations are partitioned into 10 portions and the average of each portion is calculated leading to a feature vector of length 20.
Feeding the classifier with the concatenation of different features without any preprocessing step would lead to very bad results because features of large magnitudes will dominate. Hence, we applied mean and variance normalization [14] to all the low-dimensional features discussed in this section before concatenating them into a feature vector of length 156.
4.6 Local Directional Features
Local directional features extraction is a novel technique that we present in this paper. It is a natural extension of the directional features presented in section 4.5.5. Like directional features, each background pixel is labeled by a 4-element vector which leads to 16 different labels. However not all labels are used. Only 9 labels corresponding to 9 situations are considered: (1) closed from all directions (i.e. black pixel can be reached when moving in all the principal 4 directions), (2) open up (i.e. black pixel can be reached when moving in the principal 4 directions except when moving upward), (3) open down, (4) open right, (5) open left, (6) open up and right, (7) open up and left, (8) open down and right, and (9) open down and left.
The image is then decomposed into 9 layers; each layer corresponds to one of the 9 labels. Pixel (x,y) in layer #n is illuminated only if pixel (x,y) in the image has been given the label #n. A Gaussian mask is then applied to all the 9 layers and 5×5 measurements are extracted from each of them in the same manner done with gradient features (section 4.1). This means that local directional features are composed of 225 elements

\subsection {Classification}

%The input of the feature extraction method is four different feature vector based on orientation used. A classifier is used for each orientation which means that the feature vector for each classifier consist of  $9\timesn$ concavity map. To simplify the classification we use simple one versus one (OVO) Support Vector Machine (SVM) using linear kernel classifier. Since the SVM is a binary classifier we used One versus one configuration for the classifiers which requires
The output of the feature extraction stage is four different feature vectors for the four different orientations used. A classifier is used for each orientation. To simplify the classification we use simple one versus one (OVO) Support Vector Machine (SVM) with linear kernel classifier. Since the SVM is a binary classifier, the one versus one classification requires $\frac{n(n-1)}{2}$ classifiers for $n$ classes. The final decision of the classification in a one versus one classifier can be obtained using simple majority voting. Unfortunately, this method does not generate the confidence measure for each class needed in the classifier fusion step. Various methods can be used to generate such probability or confidence measure from pairwise classifiers. One of the popular methods is introduced in \cite{pairwise22} where $p_i$ probability of sample $x$ to belong to class $i$  is computed using the following equation :

%The output of the feature extraction stage is four different feature vectors for the four different orientations used. A classifier is used for each orientation. To simplify the classification we use simple one versus one (OVO) Support Vector Machine (SVM) with linear kernel classifier. Since the SVM is a binary classifier, the one versus one classification requires $\frac{n(n-1)}{2}$ classifiers for $n$ classes. The final decision of the classification in a one versus one classifier can be obtained using simple  majority voting. Unfortunately, this method does not generate a confidence measure for each class which is needed in the classifier fusion step. Various methods can be used to generate such probability or confidence measure from pairwise classifiers. One of the popular methods is introduced in \cite{pairwise22} where $p_i$ probability of sample $x$ to belong to class $i$  is computed using the following equation :
%
% The final decision of the classification can be extraction using various method. One of the simple methods is majority voting.  Unfortunately, this method does not generate a confidence measure for each class which we need in the classifier fusion step. Various methods were used to generate such  probability or  confidence measure from pairwise classifiers. One of the popular methods used is introduced in \cite{pairwise22} where
%  \[
% p_i=\frac{1}{r_{ij}k-1}
% \]
\[
P_i(x)  = \frac{1}{{\sum\limits_{j = 1,j \ne i}^K {\frac{1}{{r_{ij}(x) }} - (K - 2)} }}
\]

where $r_{ij}(x)$   is the probability of the tested pattern $x$ to be of class $i$, where the used classifier is the one responsible for separating class $i$ from class $j$  and $k$ is the number of classes.

\subsubsection{Fusion Methods}
Fusion methods can be divided into three main categories: label based, rank based, and soft margin or fuzzy based methods. The label based method are usually based on the final label of base classifiers like the majority voting fusion. The rank based method uses the decision values of base classifiers to rank using various fixed rules to get the final decision for the system. The soft margin and fuzzy based methods use the output decision values of the base classifiers as a pattern recognition problem and try to detect the correct pattern. Methods like Bayes and neural network fusion are considered the most widely used fusion methods based on this type  \cite{Farah2005,Ruta2000,Denoeux2000}.
%(For example : Sum, Max, Min, Mean and Borda Count)
%Fusion methods can be divided into three main categories Label based, Rank based and soft margin or fuzzy based methods. The label based method are usually based on the final label of base classifiers like the majority voting fusion. The rank based method uses the decision values of base classifier to rank uses various fixed rule to get the final decision value for the system. The soft margin and fuzzy based methods tries to use the output decision values  of the base classifier as a pattern recognition problem and tries to try to detect the correct pattern. Methods like Bayes and neural network fusion is considered the most fusion method based on this type. .

%The four base classifiers generate a set of normalized probability for each class $P_i(x)$ for a given input sample $x$. These probabilities are as input the fusion methods.

The four base classifier generate the posterior probabilities $P_{ij}(x), i = 1\dots c; j = 1\dots k$ for $c$ classifiers and $k$ classes is computed for test sample $x$. A fusion method will  combine them to a new set of probability  $q_j(x)$ which can be used for the final classification. Let $q_j(x)$ be computed by $q_j(x)=rule_i\left(P_{ij}(x)\right)$. The final classification is made by $\phi(x)=\max\limits_j(q_j(x))$. The rules we used to compute $q_j(x)$ are Majority vote, Maximum, Minimum, Mean, Median, Sum, Weighted Sum, Product and Borda Count. %summarized below:

\section{Experiments}
\section {Results}

%develop new features based on structural information.
\bibliographystyle{IEEEtran}
\bibliography{library,Landmine}
\end{document}

### 2012-01-21 LandminePaper/landmine.tex
I neeeeed to sleeeeeeeeeeeeep to get a clear head

@@
 
+ Two of the widely investigated sensor for mine detection are Ground-penetrating radar (GPR) and electromagnetic induction (EMI)  due to their complementary features.
+Detecting mines with low metal contents and moslty made of plastic is an advantage of GPR sensors over traditional metal detectors.
+However, EMI sensors has better performance than GPR in detecting small antipersonal mines specially if they are bried deeply \cite{Frigui2010}.
+But some studies shows that even for the same sensor different algoritms can vary in perofromance for different mine types and soil conditions.
+For example, the system in \cite{Frigui2010}  presents an edge based feature that outperform a spectral detector in \cite{Ho2008} for the same antitank mine.
+
+%   In fact, even for the same sensor, different detection algorithms may be more effective for different mine types, burial depths, and soil conditions .
+For instance, , it was shown that a detector that uses edge-based features \cite{Frigui2009} outperforms the spectral detector \cite{Ho2008} for antitank (AT) mines and shallow mines.
+On the other hand, the spectral detector outperforms the edge-based detector for weak scattering plastic mines \cite{Ho2008}.
+In [11], using a large GPR data set collected from four different sites, it was shown that the relative performance of four different detection algorithms is highly dependent on the site
+
 The mine detection algorithm is similar to many automation and recognitio system, it has four main components: 1) reading sensor data 2) preprocessing 3) feature extraction and 4)classifier desing.
@@
 
-This paper introduce a landmine detection system that uses data collected from a multistatic   Ground penetrating radar (GPR)
+The main contribution in this paper is using a simple image based feature to locate and detect mine from the GPR images.
+This paper introduce a landmine detection system that uses data collected from a multistatic Ground penetrating radar (GPR).
+The data are collected from scanning in x and y direction a clean and a highly cluttered stone surface using an array of resistive-vee antennas \cite{Counts2007} .
+This data are processed to genearate downcross image of ground at each x and y coordinate.
+An imaging algorithm similar to   \cite{Counts2007} is used to generate the ground model from the GPR.
+After the ground images are obtained, the images are scanned using fixed window size to locate mine locations in the image.
+For each window in the scan a set of feature are extracted and preseneted to the classifier to identify it as clutter or mine.
 
-% A typical land-mine detection system, like most automated detection algorithms, has three main components: 1) preprocessing; 2) feature extraction; and 3) classifier design.
-The preprocessing step relies on algorithms that perform tasks such as normalization of the data, corrections for variations in height and speed, removal of stationary effects due to the system response, etc.
-Methods that have been used to perform this task
+The following section gives the details of the system; in Section  \ref{sec:gpr} a brief overview of GPR systems, section \ref{sec:system} gives a complete description of the system.
+We presents the results and experiments in section \ref{sec:results} and we give the conclusion in section \ref{sec:con}.
 
-% Various techniques in the area of sensor physics, signal processing, and
-% robotics have been studied during the last decade.  Most mine detection
-% techniques depend on sensors, signal processing,  and  decision  processes.  For
-%  the  sensor part,  ground  penetration  radar  (GPR),  infrared  (IR),  and
-% ultrasound  (US)  sensors  are  reviewed.  For  the  signal processing  and
-% decision  parts,  a  commonly  used  set  of image  processing  techniques
-% including  filtering, enhancement,  feature  extraction,  and  segmentation  are
-% surveyed.  Segmentation  is  used  to  extract  a  mine  signal from  various
-% competing  signals,  the  mine  detection algorithms work with inhomogeneous
-% background.
-
-% Detection and removal of land-mines is therefore a significant problem and has attracted several
-% researchers in recent years \cite{Ho2002,Tan2005,Potin2006,Savelyev2007,Collins1999,Ng2008}. Although systems can achieve high
-% detec-tion rates, they have done so at the expense of high false alarm rates.
-% Various techniques in the area of sensor physics, signal  processing, and robotics have been studied during the last  decade.
-Most mine detection techniques depend on sensors,  signal processing,  and  decision  processes.
-For  the  sensor  part,  ground  penetration  radar  (GPR),  infrared  (IR),  and  ultrasound  (US)  sensors  are  reviewed.
-For  the  signal  processing  and  decision  parts,  a  commonly  used  set  of  image  processing  techniques  including  filtering,  enhancement,  feature  extraction,  and  segmentation  are  surveyed.
-Segmentation  is  used  to  extract  a  mine  signal  from  various  competing  signals,  the  mine  detection  algorithms work with inhomogeneous background.
-
-% Because mines can be made of both metallic and
-
-% The
-% cost to purchase and lay a typical antipersonnel mine ranges from  $3 to  $30, while the cost to remove a single mine ranges from $300 to
-% $1000. Because mines can be made of both metallic and
-% nonmetallic  materials,  detection  using  only  conventional
-% metal detectors cannot give a promising result.
-% the process of detecting and
-% removing mines, called demining, is particularly important.
-% Manual  demining  is  extremely  dangerous
-
-%Throught war the armed forces
-%usually burry landmine just below surface of the ground.
-
-%A land mine is an explosive device, concealed under or on the ground and
-% %designed to destroy or disable enemy targets as they pass over or near the
-% %device. Such devices are typically detonated automatically by way of pressure
-% from the target stepping or driving on it, though other detonation mechanisms may be possible.[1] The device may cause damage either by a direct blast or by fragments that are thrown by the blast.
-
-% Landmines are one of the worst environmental problems that
-% humanity faces long with being one of the most terrifying legacies of war.
-%
-% Landmines
-% are made to be used by armed forces to disable any person or vehicle that comes
-% close with it by an explosion or fragments released at high speeds, so they become a
-% burden to those who have to support them
-%
-% MORE  than  26,000  people  are  killed  or  maimed  by
-% mines every year, which is equivalent to one victim,
-% every 20 min. For example, in Cambodia one out of every
-% 236 people is a landmine ampute
-% e. The casualty ratio rises
-% to one out of every 140 people in Angola, which has more
-% mines  than  people  in  addition  to  fatal  casualties  and
-% enormous  financial  losses.  In  Cambodia,  approximately
-% 40\%  of  the  rice  fields  have  been  mined  and  abandoned.
-% Most  tragic  is  that  many  victims  are  children  and  most
-% mine-afflicted countries are poverty stricken, as well.
-
-% Because  of  the  potentially  catastrophic  results  of
-% unintentional mine encounters, the process of detecting and
-% removing mines, called demining, is particularly important.
-% Manual  demining  is  extremely  dangerous;  around  one
-% deminer has been killed  for every 2,000 mines  removed,
-% with even more civilian victims. The cost to purchase and
-% lay  a  typical  antipersonnel  mine  ranges  from  $3  to  $30,
-% while the cost to remove a single mine ranges from $300 to
-% $1000. Because mines can be made of both metallic and
-% nonmetallic  materials,  detection  using  only  conventional
-% metal detectors cannot give a promising result. Report also
-% indicate  metal  detectors  are  subject  to  many  false  Image
-% Processing-Based.  High  labor  cost  and  the  slow  pace
-% involved are encouraging development of other techniques.
-% Although  some  military  demining  equipment  has  been
-% developed and used during the Gulf War by the US Army,
-% civilian related demining, called humanitarian demining, is
-% quite different from the military work
-
-% Landmines are basically explosive devices designed to explode when triggered by
-% pressure or a tripwire. These devices are typically found on or just below the surface
-% of the ground [8]. .
-% The International Campaign to Ban Landmines (ICBL) [9] estimates that 15,000-20,000 people are killed or injured by land mines per year many of them children.
-% The U.S. State Department estimates that a total of 45-50 million mines remain to
-% be cleared. Worldwide, approximately 100,000 mines are cleared each year [10]. At
-% that rate, clearing all 45-50 million mines will require 450- 500 years, assuming no
-% new mines are laid. By some estimates, roughly 1.9 million new mines are emplaced
-% annually, yielding an additional 19 years of mine clearance work every year.
-% The Landmine Monitor Report 2003 identied over 11,700 new landmine victims
-% in 2002. Of this number at least 2,649 were children, a staggering 23 percent. More
-% than 85 percent were innocent civilians
-
-% The  object  of  humanitarian  demining  is  to  find  and
-% remove  abandoned  landmines  without  any  hazard  to  the
-% environment.  These  landmines  were  intended  for  military
-% use  when  they  were  planted,  but  their  duty  has  expired.
-% Furthermore, humanitarian demining equipment is required
-% to  be  more  accurate  than  military-purpose  equipment
-% because the military can afford a certain degree of casualty
-% risk.
-
-% Various techniques in the area of sensor physics, signal
-% processing, and robotics have been studied during the last
-% decade.  Most mine detection techniques depend on sensors,
-% signal processing,  and  decision  processes.  For  the  sensor
-% part,  ground  penetration  radar  (GPR),  infrared  (IR),  and
-% ultrasound  (US)  sensors  are  reviewed.  For  the  signal
-% processing  and  decision  parts,  a  commonly  used  set  of
-% image  processing  techniques  including  filtering,
-% enhancement,  feature  extraction,  and  segmentation  are
-% surveyed.  Segmentation  is  used  to  extract  a  mine  signal
-% from  various  competing  signals,  the  mine  detection
-% algorithms work with inhomogeneous background.
-%
-%
-% DETECTION, localization, and subsequent neutralization of buried land-mines are
-% worldwide humanitarian and military problems. The latest statistics show that,
-% as of August 2009, more than 70 states were believed to be mine affected
-% Landmine Monitor has identified at least 73 576 casualties of land-mines,
-% explosive remnants of war, and victim-activated improvised explosive devices in
-% 119 states and areas in the past ten years \cite{report}.
-%
-% Detection and removal of land-mines is therefore a significant problem and has attracted several
-% researchers in recent years \cite{Ho2002,Tan2005,Potin2006,Savelyev2007,Collins1999,Ng2008}. Although systems can achieve high
-% detec-tion rates, they have done so at the expense of high false alarm rates.
-% Various techniques in the area of sensor physics, signal  processing, and robotics have been studied during the last  decade.
-Most mine detection techniques depend on sensors,  signal processing,  and  decision  processes.
-For  the  sensor  part,  ground  penetration  radar  (GPR),  infrared  (IR),  and  ultrasound  (US)  sensors  are  reviewed.
-For  the  signal  processing  and  decision  parts,  a  commonly  used  set  of  image  processing  techniques  including  filtering,  enhancement,  feature  extraction,  and  segmentation  are  surveyed.
-Segmentation  is  used  to  extract  a  mine  signal  from  various  competing  signals,  the  mine  detection  algorithms work with inhomogeneous background.
-%
-%
-%
-% A typical land-mine detection system, like most automated detection algorithms, has three main components: 1) preprocessing; 2) feature extraction; and 3) classifier design.
-The preprocessing step relies on algorithms that perform tasks such as normalization of the data, corrections for variations in height and speed, removal of stationary effects due to the system response, etc.
-Methods that have been used to perform this task
-
-In  section  2  a  brief  history  for  mines  detection  is  introduced,  in  section  3,  a  complete  description  for  the  proposed  classification  algorithm  among  some  predefined  mine types is  demonstrated.
+% In  section  2  a  brief  history  for  mines  detection  is  introduced,  in  section  3,  a  complete  description  for  the  proposed  classification  algorithm  among  some  predefined  mine types is  demonstrated.
 In section  4 we present  and  discuss the results.
@@
 \section{GPR Landmine Detection Overview}
+\label{sec:gpr}
 %Plain text.
 
- Ground-penetrating radar (GPR) and electromagnetic induction (EMI) are two of the extensively investigated sensors due to their complementary features.
-For instance, GPR has the advantage of detecting mines with plastic or low-metal-content mines that are difficult to detect by traditional metal detectors.
-On the other hand, detectors using EMI sensors can outperform those using GPR in detecting small antipersonnel (AP) mines and deeply buried mines [8].
-In fact, even for the same sen-sor, different detection algorithms may be more effective for different mine types, burial depths, and soil conditions \cite{Frigui2010}.
-For instance, in \cite{Frigui2010}, it was shown that a detector that uses edge-based features \cite{Frigui2009} outperforms the spectral detector \cite{Ho2008} for antitank (AT) mines and shallow mines.
-On the other hand, the spectral detector outperforms the edge-based detector for weak scattering plastic mines \cite{Ho2008}.
-In [11], using a large GPR data set collected from four different sites, it was shown that the relative performance of four different detection algorithms is highly dependent on the site
+%%%%%% what is the gpr system\ldots\ldots..
+ Ground penetrating radar (GPR) is geophysical method that uses radar pulses to image the subsurface.
+It uses the principles of scattering electromagnetic waves to locate buried objeccts.
+An electromagnetic high frequency pulse is radiated from the transmitting antenna travels through the material with a velocity determined by the permittivity of the material.
+The wave travels in straight line downwards till it hit a surface with different permittivity and electrical properties from the surronding medium.
+If the wave hits a buried object, then part of the waves energy is “reflected” back to the surface, while part of its energy continues to travel downward.
+The recieving antenaa captures the reflected wave that traved to the surface,and recorded on a digital storage device for later interpretation.
 
+The recorded data from the recieving antenaa over a period of time is called a “trace”.
+A trace is the time history of the travel single pulse from the transmit antenaa to the recieve antenaa.
+It include all the singal travel paths of all its different materials.
+The trace is the basic measurment of GPR surveys.
+A scan is a trace where a color scale has been applied to the amplitude values.
+The round-trip (or two-way) travel time is greater for deep objects than for shallow objects.
+Therefore, the time of arrival for the reflected wave recorded on each trace can be used to determine the depth of the buried object, if the velocity of the wave in the subsurface is known.
+The principles of constructing a scan from a sequence of traces is shown in Figure \ref{figure:ScanA}.
+
+ \begin{figure}
+\centering
+\label{figure:ScanA}
+\includegraphics[width=0.25\textwidth]{images/BlockDiagram.jpg}
+ \caption{Simple of an A scan }
+\end{figure}
+  %The wave that is reflected back to the surface is captured by a receive antenna, .The recieving antenaa detect the scattering of the wave from the buried object%The surface surrounding the advancing wave is called a wave- front.
+A straight line drawtn from the transmitter to the edge of the wavefront is called a ray.
+Rays are used to show the direction of travel of the wavefront in any direction away from the transmitting antenna.n.
+
+% A typical GPR system has three main components: Transmitter and receiver that are directly connected to an antenna, and a control unit (timing) (Fig.
+1).
+The transmitting antenna radiates a short high-frequency EM pulse into the ground, where it is refracted, diffracted and reflected primarily as it encounters changes in dielectric permittivity and electric conductivity.
+
+% GPR data are displayed on printer paper or on a computer screen during acquisition (i.e., during real time).
+For a given transect, the data consist of a cross-section of signal amplitudes (intensities) versus location (along the two-way time axis and the horizontal axis).
+The intensity values are digitally recorded for each trace separately, converted back into analogue signals and displayed as signal voltage amplitude versus two-way time (the RAMAC GPR system uses a 16 bit A/D converter to convert the recorded signal to 65,536 levels of amplitudes).
+The plot is referred to as a normal-incidence time section (when the transmitter-receiver offset is negligible relative to the investigated depth and at a monostatic configuration).
+Simple processing is generally needed for a conventional display; otherwise, the display may become illegible.
+Such a typical processing is the running average of three to five samples along each trace plus the average of three traces along the profile in order to increase the signal to noise ratio.
+Amplifying gainis neededto increase the visibility of the deeper parts of the image.
+Basic processing routines can be applied during operation in the field while the image is built, in a way that does not affect the collected data.
+% Ground penetrating radar (commonly called GPR) is a high resolution electromagnetic technique that is designed primarily to investigate the shallow subsurface of the earth, building materials, and roads and bridges.
+GPR has been developed over the past thirty years for shallow, high resolution investigations of the subsurface
+
+% GPR uses the principle of scattering of electromagnetic waves to locate buried objects.
+The basic principles and theory of operation for GPR have evolved through the disciplines of electrical engineering and seismic exploration, and practitioners of GPR tend to have backgrounds either in geophysical exploration or electrical engineering.
+The fundamental principle of operation is the same as that used to detect aircraft overhead, but with GPR that antennas are moved over the surface rather than rotating about a fixed point.
+This has led to the application of field operational principles that are analogous to the seismic reflection method.
+
+% The practical result of the radiation of electromagnetic waves into the subsurface for GPR measurements is shown by the basic operating principle that is illustrated in Figure A1.
+The electromagnetic wave is radiated from a transmitting antenna, travels through the material at a velocity which is determined primarily by the permittivity of the material.
+The wave spreads out and travels downward until it hits an object that has different electri- cal properties from the surrounding medium, is scattered from the object, and is detected by a receiving antenna.
+The surface surrounding the advancing wave is called a wave- front.
+A straight line drawn from the transmitter to the edge of the wavefront is called a ray.
+Rays are used to show the direction of travel of the wavefront in any direction away from the transmitting antenna.
+If the wave hits a buried object, then part of the waves energy is “reflected” back to the surface, while part of its energy continues to travel down- ward.
+The wave that is reflected back to the surface is captured by a receive antenna, and recorded on a digital storage device for later interpretation.
+%
+%
+%
+% 1-dimensional transmission line model for
+% individual GPR A-scans is considered [6]. In transmission line mod-els for electromagnetic propagation, changes in the electromagnetic
+% properties of transmission media are modeled as changes in the
+% transmission parameters. Figure 1 shows a schematic of a hetero-geneous multi-layered soil with a buried object, and a simpli fi ed
+% transmission line model for the corresponding GPR A-scan with
+% characteristic impedances Z
+
+% Ground-penetrating radar (GPR) is a geophysical method that uses radar pulses to image the subsurface.
+This nondestructive method uses electromagnetic radiation in the microwave band (UHF/VHF frequencies) of the radio spectrum, and detects the reflected signals from subsurface structures.
+GPR can be used in a variety of media, including rock, soil, ice, fresh water, pavements and structures.
+It can detect objects, changes in material, and voids and cracks.[1]
+%
+% GPR uses high-frequency (usually polarized) radio waves and transmits into the ground.
+When the wave hits a buried object or a boundary with different dielectric constants, the receiving antenna records variations in the reflected return signal.
+The principles involved are similar to reflection seismology, except that electromagnetic energy is used instead of acoustic energy, and reflections appear at boundaries with different dielectric constants instead of acoustic impedances.
+
+% The depth range of GPR is limited by the electrical conductivity of the ground, the transmitted center frequency and the radiated power.
+As conductivity increases, the penetration depth decreases.
+This is because the electromagnetic energy is more quickly dissipated into heat, causing a loss in signal strength at depth.
+Higher frequencies do not penetrate as far as lower frequencies, but give better resolution.
+Optimal depth penetration is achieved in ice where the depth of penetration can achieve several hundred meters.
+Good penetration is also achieved in dry sandy soils or massive dry materials such as granite, limestone, and concrete where the depth of penetration could be up to 15 m.
+In moist and/or clay-laden soils and soils with high electrical conductivity, penetration is sometimes only a few centimetres.
 \section {System Overview}
+\label{sec:system}
 
+This work presents a system that uses image based analyis to extract mine features from GPR images. Figure \ref{fig:block}
+shows the block digram of the system.
+The first block in generating the scan images form the calibrated data from the GPR antnna.
+After images are comptued the images are scanened fixed size windows to locate the mines.
+For each window a set various feature extraction method is used to obtain a feature vector that will be used in classification.
+A SVM classifier is ued to identify if the window represents a mine or not.
+%The training of SVM classifier is provided directly from a the GPR scan images.
  \begin{figure}
@@
 
+The follwoing section gives more details on each block of the system.
+
 \subsection{GPR processing}
 
-A multistatic ground-penetrating radar (GPR) sys-tem has been developed and used to measure the response of a
-number of targets to produce data for the investigation of multi-static inversion algorithms. The system consists of a linear array
-of resistive-vee antennas, microwave switches, a vector network
-analyzer, and a 3-D positioner, all under computer control. The
-array has two transmitters and four receivers which provide eight
-bistatic spacings from 12 to 96 cm in 12-cm increments. Buried
-targets are scanned with and without surface clutter, which is a
-layer of rocks whose spacing is empirically chosen to maximize the
-clutter effect. The measured responses are calibrated so that the
-direct coupling in the system is removed, and the signal reference
-point is located at the antenna drive point. Images are formed us-ing a frequency-domain beamforming algorithm that compensates
-for the phase response of the antennas. Images of targets in air
-validate the system calibration and the imaging algorithm. Bistatic
-and multistatic images for the buried targets are very good, and
-they show the effectiveness of the system and processing.
+A multistatic ground-penetrating radar (GPR) presented in \cite{Counts2007} used to measure the response of a
+number of targets to produce data for the investigation of  the system.
+The system consist of linear arry of resistitve vee antennas.
+The array has two transmitters and four receivers which provide eight bistatic spacings from 12 to 96 cm in 12-cm increments.
+The sysstem scans a surface that contained buried targets with and without surface clutter (layer of rocks ).
+The measured responses are calibrated so that the direct coupling in the system is removed, and the signal reference point is located at the antenna drive point.
 
-%%% this is from multi experiement paper \ldots.
-The multistatic GPR in this paper consists of a linear array
-of resistive-vee antennas, a microwave switch matrix, a vector
-network analyzer (Agilent 8720D), and a 3-D positioner, all
-under computer control (Fig. 1). The resistive-vee antenna ishosen as an array element because it can transmit and receive
-short pulses with a negligible amount of clutter, and it has a
-low radar cross section, which eliminates most of the multiple
-reflections between the antenna and the ground [14]–[16]. In
-addition, it is lightweight; therefore, an array of the resistive-vee antennas can be made in a hand-held form. The antenna
-was optimized for a bistatic GPR.
-
-he GPR is scanned over a 1.8 × 1.8 m region at a constant
-height above the surface of the ground. The scan region is
-referenced by x- and y -coordinates, both ranging from −90 to
-90 cm in 2-cm increments. Thus, the scan region is discretized into a grid of 91 points by 91 points. The location of the
-GPR array in the coordinate system is referenced to the center
-point of transmitter T1. Each time the GPR array stops, it
-collects data from the eight bistatic spacings by manipulating
-switches in an appropriate order. After each switch operation,
-the network analyzer sweeps 401 equally spaced frequency
-points from 60 MHz to 8.06 GHz. This procedure at each
-point takes approximately 6 s. The total scan time is approxi-mately 14 h.
-
-The lowest effective frequency is in the range of 500 MHz–
-1.0 GHz because the gain of the antenna falls off quickly in this
-frequency range [20]
-
-The data calibrated according to (1) are stored in Matlab
-file format. Each Matlab file contains one set of measured and
-calibrated data obtained from a single scan of 1.8 × 1.8 m
-region. The stored variables are responses from eight bistatic
-spacings (401 × 91 × 91 double precision), frequency points at
-which the data are measured (401 × 1 double precision), and
-x- andy -coordinates of each measurement (91 × 1).
-
-The multistatic GPR is operated on five sets of buried targets.
-The target sets are an assortment of metal spheres, a variety
-of AT and AP landmines, a Plexiglas box, a plastic pipe,
-and empty sand. Table I contains more information about the
-landmine and clutter objects. Target depths range from 1 to
-58 cm; the depth of a target is measured from the surface of
-the sand to the top of the target. The ground medium is damp
-compacted sand with a water table below target level.
-Each set of targets is scanned with and without surface clutter
-according to the following procedure (the buried pipe is done
-only with a clean surface). First, the targets are buried, and
-the surface of the sand is packed so that the height of the
-surface varies less than 2 cm. The sand is then soaked with
-water and let stabilize for approximately 2 h. This process is
-performed to maintain appropriate moisture level in the sand
-and to minimize the disturbance made in the sand when the
-targets are buried. The responses of the targets in sand with
-a clean surface are obtained as the GPR scans over the sand.
-Next, river rocks are randomly scattered over the scan region,
-as shown in Fig. 12. The density of the rocks is empirically
-chosen to maximize the clutter effect. In both the clean-surface
-and cluttered-surface scans, the antenna array is elevated by
-10 cm, which is measured from the surface of the sand to the
-bottom of the antenna array.
+The GPR antenaa scans a region of  1.8x1.8 m  at a certain higth above the surface \cite{Counts2007}.
+Figure \ref{fig:mineshapes} shows various types of mine buried in the location specified by figure \ref{fig:minloc}.
+The scan is reference by x and y coordinate ranging from -90 to 90 using 2 cm increments.
 
@@
 
-More plain text
+The data calibrated according were stored in Matlab file format.
+Each Matlab file contains one set of measured and calibrated data obtained from a single scan of 1.8 x 1.8 m region.
+The stored variables are responses from eight bistatic spacings (401 x 91 x 91 double precision), frequency points at which the data are measured (401 x 1 double precision), and x- andy -coordinates of each measurement (91 x 1).
+
+I took the frequency measure data and converted them to time domain using inverse fourier transorm.
+Figure \ref{figure:ScanA} shows a single time signal for a specific x- and y- coordinate.
+A 2D image is generated form scannign the data in x direction and concatenating time signals in y-directions.
+Figure \ref{fig:mineAtx0} shows the dB scale of the image comuted x =0 , the location of TM1 and VS1 mine are clear in the image.
+% adding To generate a For each x- and y- coordiante the data generate a time sigal that can  be used to model the soil and subsurface objects
+
+ \begin{figure}
+\centering
+\label{fig:mineAtx0}
+\includegraphics[width=0.45\textwidth]{images/MineCleanX0.jpg}
+ \caption{The 2D scan Image at x =0 }
+\end{figure}
+
 \subsection{Image processing}
 
-Gradient Features
-To extract gradient features [3], the gradient operator is first applied to the gray-scale image of the digit to give two gradient components: strength |g(x,y)| and direction  at each point (x,y) of the image f.
-This is done by applying Sobel operator [20] on the image to extract vertical and horizontal gradient components,
- ,	(2)
- ,	(3)
-where f(x,y) is the intensity of image f at point (x,y), gx(x,y) is the gradient component at x-direction at location (x,y), and gy(x,y) is the gradient component at y-direction at location (x,y).
-Then the gradient strength and direction is extracted using equations (4) and (5), respectively,
- 		(4)
- 			(5)
-The gradient vector g(x,y) (expressed as strength strength |g(x,y)| and direction  ) at each point (x,y) of the image is then decomposed into the 8 Freeman [20] directions shown in Fig.
-2.
+The system test various image based method to investigate the best feature extraction to characterize mines in the GPR images.
+A fixed size window with no overlap is used to scan the gpr images.
+Several imaga analysis methods are used to extract features and used to compare to achieve the best performance.
+The next section explain each feature in details.
+
+\subsubsection {Gradient Features}
+\label{sec:grad}
+To extract gradient features \cite{Liu2003}, the gradient operator is first applied to the gray-scale image of the digit to give two gradient components: strength $|g(x,y)|$ and direction $\angle g(x,y)$ at each point $(x,y)$ of the image $f$.
+This is done by applying Sobel operator \cite{gonzales2002} on the image to extract vertical and horizontal gradient components.
+Then the gradient strength and direction is extracted using equations \ref{eq:1} and  \ref{eq:2}, respectively,
+
+\begin{equation}
+|g(x,y)|=\sqrt{g^{2}_{x}(x,y)+g^{2}_{y}(x,y)}
+\label{eq:1}
+\end{equation}
+\begin{equation}
+\angle g(x,y)=\tan^{-1} \frac{g_{y}(x,y)}{g_{x}(x,y)}
+\label{eq:2}
+\end{equation}
+where $g_{y}(x,y)$ and $g_{x}(x,y)$ are the vertical and horizontal gradient components extracted from Sobel operator.
+
+The gradient vector $g(x,y)$ (expressed as strength strength $|g(x,y)|$ and direction $\angle g(x,y)$  at each point $(x,y)$ of the image is then decomposed into the 8 Freeman\cite{gonzales2002} directions shown in Figure \ref{fig:freeman} .
 The gradient vector is decomposed into the 8 Freeman directions by projecting the vector into the nearest two Freeman directions as shown in Fig.
@@
 Each layer is the projection of the gradient vectors of the image into the corresponding Freeman direction.
-A Gaussian mask h(x,y) is then applied to each layer,
- ,		(6)
-and then the layers are uniformly sampled to give 5×5 measurements.
+
+ \begin{figure}
+\centering
+\label{fig:freeman}
+\includegraphics[width=0.25\textwidth]{images/freemanCodes.jpg}
+ \caption{The 2D scan Image at x =0 }
+\end{figure}
+
+% A Gaussian mask h(x,y) is then applied to each layer,
+%
+%
+% and then the layers are uniformly sampled to give 5×5 measurements.
 The parameter σ is related to the sampling interval t via the empirical relation  [3].
@@
 The gradient feature extraction technique is very powerful as shown in [3] and as will be shown in the results section (section 5) of this paper.
-4.2 Kirsch Features
-The Kirsch features [3] are extracted by decomposing the image into 4 layers corresponding to 4 edge orientations: horizontal (gh), vertical (gv) , right diagonal (grd), and left diagonal (gld).
+\subsubsection {Kirsch Features}
+The Kirsch features \cite{Liu2003} are extracted by decomposing the image into 4 layers corresponding to 4 edge orientations: horizontal ($g_h$), vertical ($g_v$) , right diagonal ($g_{rd}$), and left diagonal ($g_{ld}$).
 These layers are extracted by applying Kirsch masks on the image,
- ,		(7)
- ,		(8)
- ,	(9)
- ,	(10)
-where * denotes 2D convolution operation [20], kh1 and kh2 are Kirsch masks responsible for extracting horizontal edges, kv1 and kv2 for horizontal edges, krd1 and krd2 for right diagonal edges, and kld1 and kld2 for left diagonal edges.
-Figure 4 shows all Kirsch masks.
-gh1
-5	5	5
--3	0	-3
--3	-3	-3
-	gh2
--3	-3	-3
--3	0	-3
-5	5	5
-	gv1
-5	-3	-3
-5	0	-3
--3	-3	-3
-	gv2
--3	-3	5
--3	0	5
--3	-3	5
+\begin{equation}
+ g_h=\max \left( |f*k_{h1}| , |f*k_{h2}| \right)
+\label{eq:kirsh1}
+\end{equation}
+\begin{equation}
+ g_v=\max \left( |f*k_{v1}| , |f*k_{v2}| \right)
+\label{eq:kirsh1}
+\end{equation}
+\begin{equation}
+ g_{rd}=\max \left( |f*k_{rd1}| , |f*k_{rf2}| \right)
+\label{eq:kirsh1}
+\end{equation}
+\begin{equation}
+ g_{ld}=\max \left( |f*k_{ld1}| , |f*k_{ld2}| \right)
+\label{eq:kirsh1}
+\end{equation}
+where $*$ denotes 2D convolution operation \cite{gonzales2002}, $kh1$ and $kh2$ are Kirsch masks responsible for extracting horizontal edges, $kv1$ and $kv2$ for horizontal edges, $krd1$ and $krd2$ for right diagonal edges, and $kld1$ and $kld2$ for left diagonal edges.
+Figure \ref{fig:kirshMasks} shows all Kirsch masks.
 
-grd1
--3	-3	-3
-5	0	-3
-5	5	-3
-	grd2
--3	5	5
--3	0	5
--3	-3	-3
-	gld1
--3	-3	-3
--3	0	5
--3	5	5
-	gld2
-5	5	-3
-5	0	-3
--3	-3	-3
+  \begin{figure}
+\centering
+\label{fig:kirshMasks}
+\includegraphics[width=0.25\textwidth]{images/kirshAll.jpg}
+ \caption{The 2D scan Image at x =0 }
+\end{figure}
 
-Figure 4. Kirsch  Masks
+%A Gaussian mask is then applied to all the 4 layers and 5×5 measurements are extracted from each of them in the same manner done with gradient features (section 4.1).
+This means that Kirsch features are composed of 4×5×5=100 elements.
+\subsubsection {Local Chain Code Features}
 
-A Gaussian mask is then applied to all the 4 layers and 5×5 measurements are extracted from each of them in the same manner done with gradient features (section 4.1).
-This means that Kirsch features are composed of 4×5×5=100 elements.
-4.3 Local Chain Code Features
-To extract the local chain code features [21], the contour of the digit is first followed by any contour following algorithm, and then each contour point of the image is marked with the corresponding Freeman direction.
+To extract the local chain code features \cite{Zhang2004}, the contour of the digit is first followed by any contour following algorithm, and then each contour point of the image is marked with the corresponding Freeman direction.
 The image is then decomposed into 8 layers; each layer contains only contour points that belong to the corresponding Freeman direction.
-Each of the 8 layers is then uniformly partitioned into 5×5 zones.
-Each zone is averaged leading to a feature vector of 200 elements.
+Each of the 8 layers is then uniformly partitioned into 5X5 zones.
 
-4.4 Wavelet Features
-Wavelet transform [20] extracts the essential information contained in a signal or image.
+\subsubsection {Wavelet Features}
+Wavelet transform \cite{gonzales2002} extracts the essential information contained in a signal or image.
 This may help in reducing the dimensionality of the problem which leads to faster classification.
 Also it might reduce the noise and the non-essential information that may confuse the classifier leading to better classification performance.
-Wavelet transform could be applied on the image directly or could be applied on the gradient decomposition of the image [22].
+Wavelet transform could be applied on the image directly or could be applied on the gradient decomposition of the image.
 In our study, we applied the former method.
-The image is first resized to be 64×64 and then the image is composed into 3 resolution levels.
-The 3rd level approximation of the image (which is a 8×8 image) is then used as features.
-This means that wavelet features are composed of 64 elements.
-4.5 Concatenation of low dimensional features
+The image is first resized to be 64X64 and then the image is composed into 3 resolution levels.
+The 3rd level approximation of the image (which is a 8X8 image) is then used as features.
+
+\subsubsection {Concatenation of low dimensional features}
+\label{sec:conLow}
 Here we concatenate 6 families of low dimensional features to form one feature vector.
@@
 Each of the 6 low dimensional feature families is going to be discussed in the following subsections.
-4.5.1 Raw Image Zoning
-In this technique, the image is uniformly partitioned into 5×5 zones [23]. The average of each zone is calculated leading to a feature vector of 25 elements.
 
-4.5.2 Vertical and Horizontal Projections
-In this technique, the horizontal and vertical histograms of the image are calculated leading to a feature vector of 40 elements [23] (Remember that we are working on the 20×20 window that confines the digit not all 28×28 pixels of the MADBase digits).
+\begin{enumerate}
+  \item Raw Image Zoning \\
+  In this technique, the image is uniformly partitioned into 5×5 zones\cite{Trier1996}. The average of each zone is calculated leading to a feature vector of 25 elements.
 
-4.5.3 Vertical and Horizontal Cross Counts
-The image is first binarized, and then the vertical and horizontal cross counts are calculated leading to a feature vector of 40 elements.
-4.5.4 Centroidal Distances
-In this technique [11, 24], the digit image is partitioned into 16 sectors around the digit center of gravity (the centroid) as shown in Figure 5.
+	\item Vertical and Horizontal Projections \\
+	In this technique, the horizontal and vertical histograms of the image are calculated leading to the feature vector. \cite{Trier1996}
+
+	\item Vertical and Horizontal Cross Counts \\ The image is first binarized, and then the vertical and horizontal cross counts are calculated leading to the feature vector.
+	\item  Centroidal Distances \\
+In this technique \cite{ElSherif2007}, the digit image is partitioned into 16 sectors around the digit center of gravity (the centroid) as shown in Figure 5.
 Then the centroid of each sector are calculated.
@@
 This forms a 16-element feature vector.
-4.5.5 Directional Features
-The directional features are formed as follows [11, 25].
+
+\item Directional Features\\
+The directional features are formed as follows \cite{ElSherif2007}.
 Each background pixel is labeled by a 4-element vector.
@@
 The final feature vector is the histogram of such labels giving rise to a 16 element feature vector.
-4.5.6 Length-Normalized Contour
-In this technique [11], the horizontal and vertical coordinates of the boundary pixels of the digit are used as features.
+\item Length-Normalized Contour
+In this technique \cite{ElSherif2007}, the horizontal and vertical coordinates of the boundary pixels of the digit are used as features.
 First, the boundary of the digit is extracted using a contour following algorithm; then, each boundary point is stored by its horizontal and vertical coordinates.
@@
 Then both horizontal and vertical locations are partitioned into 10 portions and the average of each portion is calculated leading to a feature vector of length 20.
-Feeding the classifier with the concatenation of different features without any preprocessing step would lead to very bad results because features of large magnitudes will dominate.
+
+% Feeding the classifier with the concatenation of different features without any preprocessing step would lead to very bad results because features of large magnitudes will dominate.
 Hence, we applied mean and variance normalization [14] to all the low-dimensional features discussed in this section before concatenating them into a feature vector of length 156.
-4.6 Local Directional Features
+
+\end{enumerate}
+\subsubsection {Local Directional Features}
 Local directional features extraction is a novel technique that we present in this paper.
-It is a natural extension of the directional features presented in section 4.5.5.
-Like directional features, each background pixel is labeled by a 4-element vector which leads to 16 different labels.
+It is a natural extension of the directional features presented in section \ref{sec:conLow}  Like directional features, each background pixel is labeled by a 4-element vector which leads to 16 different labels.
 However not all labels are used.
@@
 black pixel can be reached when moving in the principal 4 directions except when moving upward), (3) open down, (4) open right, (5) open left, (6) open up and right, (7) open up and left, (8) open down and right, and (9) open down and left.
+
 The image is then decomposed into 9 layers; each layer corresponds to one of the 9 labels.
-Pixel (x,y) in layer #n is illuminated only if pixel (x,y) in the image has been given the label #n.
-A Gaussian mask is then applied to all the 9 layers and 5×5 measurements are extracted from each of them in the same manner done with gradient features (section 4.1).
-This means that local directional features are composed of 225 elements
+Pixel $(x,y)$ in layer #n is illuminated only if pixel $(x,y)$ in the image has been given the label #n.
+A Gaussian mask is then applied to all the 9 layers and 5x5 measurements are extracted from each of them in the same manner done with gradient features (section \ref{sec:grad}).
+
+\subsubsection {Moments}
+The orthoganal moments of the $7^th$ order are computed for each block of the image
+\subsubsection {Discrete cosine transform}
+The Discrete cosine transform for each block is computed and the return the higher order features only are used as features.
 
 \subsection {Classification}
+A single Binary Support vector machine (SVM) classifier is used to classify the extracted feature.
+As shown in figure \ref{fig:minloc} the x, y  coordinate and burial depth is given for all mines.
+This infromation is used to extract samples for the trainging.
+Regions of mine locations  are extracted from the scan image data and used to trained the SVM classifier to the mine features.
+Random clutter location are also extracted and feature were comtued and present to be trained the SVM classifier.
 
-%The input of the feature extraction method is four different feature vector based on orientation used.
-A classifier is used for each orientation which means that the feature vector for each classifier consist of  $9\timesn$ concavity map.
-To simplify the classification we use simple one versus one (OVO) Support Vector Machine (SVM) using linear kernel classifier.
-Since the SVM is a binary classifier we used One versus one configuration for the classifiers which requires
-The output of the feature extraction stage is four different feature vectors for the four different orientations used.
-A classifier is used for each orientation.
-To simplify the classification we use simple one versus one (OVO) Support Vector Machine (SVM) with linear kernel classifier.
-Since the SVM is a binary classifier, the one versus one classification requires $\frac{n(n-1)}{2}$ classifiers for $n$ classes.
-The final decision of the classification in a one versus one classifier can be obtained using simple majority voting.
-Unfortunately, this method does not generate the confidence measure for each class needed in the classifier fusion step.
-Various methods can be used to generate such probability or confidence measure from pairwise classifiers.
-One of the popular methods is introduced in \cite{pairwise22} where $p_i$ probability of sample $x$ to belong to class $i$  is computed using the following equation :
-
-%The output of the feature extraction stage is four different feature vectors for the four different orientations used.
-A classifier is used for each orientation.
-To simplify the classification we use simple one versus one (OVO) Support Vector Machine (SVM) with linear kernel classifier.
-Since the SVM is a binary classifier, the one versus one classification requires $\frac{n(n-1)}{2}$ classifiers for $n$ classes.
-The final decision of the classification in a one versus one classifier can be obtained using simple  majority voting.
-Unfortunately, this method does not generate a confidence measure for each class which is needed in the classifier fusion step.
-Various methods can be used to generate such probability or confidence measure from pairwise classifiers.
-One of the popular methods is introduced in \cite{pairwise22} where $p_i$ probability of sample $x$ to belong to class $i$  is computed using the following equation :
-%
-% The final decision of the classification can be extraction using various method.
-One of the simple methods is majority voting.
-Unfortunately, this method does not generate a confidence measure for each class which we need in the classifier fusion step.
-Various methods were used to generate such  probability or  confidence measure from pairwise classifiers.
-One of the popular methods used is introduced in \cite{pairwise22} where
-%  \[
-% p_i=\frac{1}{r_{ij}k-1}
-% \]
-\[
-P_i(x)  = \frac{1}{{\sum\limits_{j = 1,j \ne i}^K {\frac{1}{{r_{ij}(x) }} - (K - 2)} }}
-\]
-
-where $r_{ij}(x)$   is the probability of the tested pattern $x$ to be of class $i$, where the used classifier is the one responsible for separating class $i$ from class $j$  and $k$ is the number of classes.
-
-\subsubsection{Fusion Methods}
-Fusion methods can be divided into three main categories: label based, rank based, and soft margin or fuzzy based methods.
-The label based method are usually based on the final label of base classifiers like the majority voting fusion.
-The rank based method uses the decision values of base classifiers to rank using various fixed rules to get the final decision for the system.
-The soft margin and fuzzy based methods use the output decision values of the base classifiers as a pattern recognition problem and try to detect the correct pattern.
-Methods like Bayes and neural network fusion are considered the most widely used fusion methods based on this type  \cite{Farah2005,Ruta2000,Denoeux2000}.
-%(For example : Sum, Max, Min, Mean and Borda Count)
-%Fusion methods can be divided into three main categories Label based, Rank based and soft margin or fuzzy based methods.
-The label based method are usually based on the final label of base classifiers like the majority voting fusion.
-The rank based method uses the decision values of base classifier to rank uses various fixed rule to get the final decision value for the system.
-The soft margin and fuzzy based methods tries to use the output decision values  of the base classifier as a pattern recognition problem and tries to try to detect the correct pattern.
-Methods like Bayes and neural network fusion is considered the most fusion method based on this type.
-.
-
-%The four base classifiers generate a set of normalized probability for each class $P_i(x)$ for a given input sample $x$. These probabilities are as input the fusion methods.
-
-The four base classifier generate the posterior probabilities $P_{ij}(x), i = 1\dots c; j = 1\dots k$ for $c$ classifiers and $k$ classes is computed for test sample $x$.
-A fusion method will  combine them to a new set of probability  $q_j(x)$ which can be used for the final classification.
-Let $q_j(x)$ be computed by $q_j(x)=rule_i\left(P_{ij}(x)\right)$.
-The final classification is made by $\phi(x)=\max\limits_j(q_j(x))$.
-The rules we used to compute $q_j(x)$ are Majority vote, Maximum, Minimum, Mean, Median, Sum, Weighted Sum, Product and Borda Count.
-%summarized below:
+The testing are genearted by scanning the image using a fixed window
+  %The training is set iThe location of all mines in scan image areThe location of the The training set is generated from
 
 \section{Experiments}
-\section {Results}
+\label{sec:results}
 
+\subsection{Effect of vairous window size}
+ \subsection{Effect of Feature extraction method}
+ \subsection{Effect of Clutter and rock}
+
+\section {Conclusion}
+\label{sec:con}
+This paper presente a simple image based method to detect landmine from GPR images. The image based feature uses
 %develop new features based on structural information.

### 2012-01-22 LandminePaper/landmine.tex
before review

@@
 armies move, the land continues to contains landmines for years. The
-International Campaign to Ban Landmines (ICBL) [9] estimates that every year
+International Campaign to Ban Landmines (ICBL) \cite{ICBL2007} estimates that every year
 there is more than 25,000 people killed or maimed by mines many of them are
 children. A total estimate of 45-40 millions mines remains to be cleared
-worldwoide [10]. One of the problems of mine removal or the demining is
+worldwoide \cite{ICBL2007}. One of the problems of mine removal or the demining is
 high cost of the process. The cost of removing a single mine ranges from  300\$ to 1000\$
@@
 Ground penetrating radar (GPR), Infrared (IR), Sesiemic,  and ultrasound (US) sensors were developed and reviewed \cite{Scott2004,Ng2008}.
-These sensor usually generate a signal that requires different types of processing as noise removal, filtering, enhancement and feature extractions\cite{Ng2008,Potin2006}.
+These sensor usually generate a signal that requires different types of processing as noise removal, filtering, enhancement and feature extractions\cite{Ng2008,Potin2006,Wu2009}.
 Although systems can achieve high detection rates, they have done so at the expense of high false alarm rates.
@@
 \label{fig:minloc}
-\includegraphics[width=0.45\textwidth]{images/MineLocations.jpg}
+\includegraphics[width=0.35\textwidth]{images/MineLocations.jpg}
  \caption{Mine Locations}
@@
 \label{fig:mineshapes}
-\includegraphics[width=0.45\textwidth]{images/MineShapes.jpg}
+\includegraphics[width=0.35\textwidth]{images/MineShapes.jpg}
  \caption{ Sample of mines  }
@@
 \label{fig:mineAtx0}
-\includegraphics[width=0.45\textwidth]{images/MineCleanX0.jpg}
+\includegraphics[scale=1.5]{images/MineCleanX0.jpg}
  \caption{The 2D scan Image at x =0 }
@@
 \subsection{Image processing}
+\label{sec:image}
 
@@
 \label{fig:freeman}
-\includegraphics[width=0.25\textwidth]{images/freemanCodes.jpg}
+\includegraphics[scale=0.55]{images/freemanCodes.jpg}
  \caption{The 2D scan Image at x =0 }
@@
 \label{fig:kirshMasks}
-\includegraphics[width=0.25\textwidth]{images/kirshAll.jpg}
+\includegraphics[width=0.40\textwidth]{images/kirshAll.jpg}
  \caption{The 2D scan Image at x =0 }
@@
 The image is then decomposed into 8 layers; each layer contains only contour points that belong to the corresponding Freeman direction.
-Each of the 8 layers is then uniformly partitioned into 5X5 zones.
+Each of the 8 layers is then uniformly partitioned into 5\times5 zones.
 
 \subsubsection {Wavelet Features}
+
+Wavelet transform \cite{gonzales2002} is best known for extracting essential infromation contained in a signal or an image.
+This reduce the dimensionality of the features and leads to faster classification.
+It also reduces the noise and the non-essential information that may confuse the classifier.
+Wavelet tranform could be applied to an image or to the gradient decomposition of an image.
+We use the former method
+
 Wavelet transform \cite{gonzales2002} extracts the essential information contained in a signal or image.
@@
 In our study, we applied the former method.
-The image is first resized to be 64X64 and then the image is composed into 3 resolution levels.
-The 3rd level approximation of the image (which is a 8X8 image) is then used as features.
+The image is first resized to be $64\times64$ and then the image is composed into 3 resolution levels.
+The 3rd level approximation of the image (which is a $8\times8$ image) is then used as features.
 
@@
 \label{sec:conLow}
-Here we concatenate 6 families of low dimensional features to form one feature vector.
-While each of these low dimensional features is not very effective, their concatenation proved to be a powerful feature vector.
+A concatentaion of 6 low dimentional feature are used to form a single feature vecotr.
+Any single feature is not effective but their concatenation proved to produce a set of  powerful features.
 Each of the 6 low dimensional feature families is going to be discussed in the following subsections.
@@
   \item Raw Image Zoning \\
-  In this technique, the image is uniformly partitioned into 5×5 zones\cite{Trier1996}. The average of each zone is calculated leading to a feature vector of 25 elements.
+  The image is divided into $5\times5$ zones and the average of raw pixel values of each zone is calculated leading to a feature vector of 25 elements\cite{Trier1996}.
 
 	\item Vertical and Horizontal Projections \\
-	In this technique, the horizontal and vertical histograms of the image are calculated leading to the feature vector. \cite{Trier1996}
+	The histogram of the horizontal and vertical projections is used as a feature vector \cite{Trier1996}
 
-	\item Vertical and Horizontal Cross Counts \\ The image is first binarized, and then the vertical and horizontal cross counts are calculated leading to the feature vector.
+	\item Vertical and Horizontal Cross Counts \\
+	The first step is binarizing the image and then computing the number of crossing from 0 to 1 and vice-versa in  both the hrozintal and vertical direction.
+The number of cross counts in each direction is used as the feature vector .
+%The image is first binarized, and then the vertical and horizontal cross counts are calculated leading to the feature vector.
 	\item  Centroidal Distances \\
-In this technique \cite{ElSherif2007}, the digit image is partitioned into 16 sectors around the digit center of gravity (the centroid) as shown in Figure 5.
+	The image is divided into 16 anglar sector from the image center of gravity.
+The centroid of each sector is calculated and the distance between the image centeroid and the sector centroid is normalized by dividing them by the bounding box diameter\cite{ElSherif2007}.
+The normalized distances are used as the feature vector.
+% 	In this technique , the digit image is partitioned into 16 sectors around the digit center of gravity (the centroid) as shown in Figure 5.
 Then the centroid of each sector are calculated.
@@
 The directional features are formed as follows \cite{ElSherif2007}.
-Each background pixel is labeled by a 4-element vector.
+A 4-element vector is used to label each background pixel.
+To label each background pixel, we walk upward and if we face a foreground pixel, the first element of the 4-element vector is set to 1, otherwise we set it to 0.
+Next we walk right till we find a foreground pixel, the second element is set to 1 otherviwse set it to 0.
+The remaining elements are filled in the same mannar till all 4 prinicipal directions are processed.
+After all directions for all background pixels are labeled, the combination of 4-element vector is marked and ecah combination is labedl with a different label.
+Hence, we have 16 different labels which we compute the histogram of each label and use as feature vector.
+
+% Each background pixel is labeled by a 4-element vector.
 For each background pixel we walk upward; if a foreground pixel is found, the first element of the 4-element vector is set to ‘1’ otherwise set to ‘0’.
-Then we walk to the right; if a foreground pixel is found, the second elements is set to ‘1’ otherwise set to ‘0’; and so on till we finish the 4 principal directions.
-Then for each combination of 0’s and 1’s in the 4-element vector, we give a different label.
+%  Then we walk to the right; if a foreground pixel is found,
+%  the second elements is set to ‘1’ otherwise set to ‘0’; and so on till we finish the 4 principal directions.
+%
+%  Then for each combination of 0’s and 1’s in the 4-element vector, we give a different label.
 Hence, we have 16 different labels.
@@
 The final feature vector is the histogram of such labels giving rise to a 16 element feature vector.
-\item Length-Normalized Contour
-In this technique \cite{ElSherif2007}, the horizontal and vertical coordinates of the boundary pixels of the digit are used as features.
+\item Length-Normalized Contour \\
+The normalized contour is computed form the horizontal and vertical coordinates of the boundary pixels as mentioned in  \cite{ElSherif2007}.
+The first step is extracting the boundry of a shape using a contour following algorithm.
+The extracted boundry point is stored by both its horizontal and vertical coordinates.
+The horizontal locations are magnitude-normalized by dividing the width of the bounding box.
+Simillary, the vertical locations are normalized by dividing them by the height of the bounding box.
+The average of 10 different portions of both the horizontal and vertical locations represent the feature vector.
+%portions Both the horizontal and vertical locations are partitioned into 10 different portions.
+The average of each portion is calculated and % is used to normalized the horizontal locations by dividing
+
+%In this technique \cite{ElSherif2007}, the horizontal and vertical coordinates of the boundary pixels of the digit are used as features.
 First, the boundary of the digit is extracted using a contour following algorithm; then, each boundary point is stored by its horizontal and vertical coordinates.
@@
 \end{enumerate}
-\subsubsection {Local Directional Features}
-Local directional features extraction is a novel technique that we present in this paper.
+% \subsubsection {Local Directional Features}
+% Local directional features extraction is a novel technique that we present in this paper.
 It is a natural extension of the directional features presented in section \ref{sec:conLow}  Like directional features, each background pixel is labeled by a 4-element vector which leads to 16 different labels.
@@
 black pixel can be reached when moving in the principal 4 directions except when moving upward), (3) open down, (4) open right, (5) open left, (6) open up and right, (7) open up and left, (8) open down and right, and (9) open down and left.
-
-The image is then decomposed into 9 layers; each layer corresponds to one of the 9 labels.
+%
+% The image is then decomposed into 9 layers; each layer corresponds to one of the 9 labels.
 Pixel $(x,y)$ in layer #n is illuminated only if pixel $(x,y)$ in the image has been given the label #n.
@@
 
-\subsubsection {Moments}
-The orthoganal moments of the $7^th$ order are computed for each block of the image
+% \subsubsection {Moments}
+% The orthoganal moments of the $7^th$ order are computed for each block of the image
 \subsubsection {Discrete cosine transform}
@@
 Random clutter location are also extracted and feature were comtued and present to be trained the SVM classifier.
+Eventhough that mine dection can be viewed as a single class problem tests shows that considering the problem as two classes problem where first class is mine and second is clutter  achieve better results.
 
-The testing are genearted by scanning the image using a fixed window
+The testing are genearted by scanning the image using a fixed window which is used to divide the input scan image into a set of equal windows.
+Each window, which either represents a clutter or a mine location, feature vector is represented to the svm classifier.
+The classifier then labels each window in the image using either mine or clutter label.
   %The training is set iThe location of all mines in scan image areThe location of the The training set is generated from
@@
 
+As mentioned before the training set of the system are extracted from the mine location in GPR images and used as input for training the classifier.
+The different clean and rock data are used to provide differiance in training and testing.
+In all the present results, the average of two different experiments spliting data into training and testing are provided.
+In the first experiment the training data are extracted from clean surface data and tested on cluttered rock surface data.
+The second experiment, extracts the training data from the rock surfaced data and tested on clean surface images.
+The recognition rate is computed as the number of correct identified samples from the images windows over total number of samples.
+ \subsection{Effect of Feature extraction method}
+
+Different feature extraction method explained in sections \ref{sec:image} were used to locate the buried landmines.
+Figure \ref{fig:ResultsFeats} shows the results of using the different feature extraction methods.
+The results show that the Local chain code and gradient features achieve the best recognition rate.
+Local chain code features outperform the gradient feature in term of computaional complexity.
+The results reveal that local features and structures outperform global featurs like (Discrite cosine transform) DCT.
+
+   \begin{figure}
+\centering
+\label{fig:ResultsFeats}
+\includegraphics[width=0.45\textwidth]{images/ResultsFeats.jpg}
+ \caption{Recongnition rate of different feature extraction methods }
+\end{figure}
+ %  the testing set  must be different than the training set,
 \subsection{Effect of vairous window size}
- \subsection{Effect of Feature extraction method}
- \subsection{Effect of Clutter and rock}
+We conducted several experiments on the best scan window size.
+The scan window specify the size of the window tested by the classifier.
+Figure \ref{fig:ResultSize} displays the result of differnet window size from 5 to 15 pixel in width and height.
+the results shows that the system achieved best performance with the smallest window size, 5x5 pixels, this is understandable with respect to the size of the buried mines.
+The size of the targets buried which varies from    5 X 4 cm  TS-50 antipersonnel (AP)  mine  to 22X9 cm VS-1.6 antitank (AT) mine.
+The small window provide more detailed features for local features and provide better charactersitc for mines.
+% smaller the window t  %The size of the scan window
+%Different feature extraction method explained in sections \ref{sec:image} were used to locate the buried landmines.
+Figure \ref{fig:ResultsFeats} shows the results of using the different feature extraction methods.
+The results show that the Local chain code and gradient features achieve the best recognition rate.
+Local chain code features outperform the gradient feature in term of computaional complexity.
+The results reveal that local features and structures outperform global featurs like (Discrite cosine transform) DCT.
+  \begin{figure}
+\centering
+\label{fig:ResultSize}
+\includegraphics[width=0.45\textwidth]{images/ResultSize.jpg}
+ \caption{Recongnition rate of different scan window sizes}
+\end{figure}
+
+ \subsection{Effect of Mine vs. Clutter}
+
+ The output of the system is a map of location of targets landmine to the locaion of the clutter in the GPR images.
+Figure \ref{fig:ResultMineClean} shows the ouput of the GPR images with the location of the mine idetified by the Red rectangle the remaiing locations were marked as clutter and ignored by the system.The actual location of mine are marked by while rectangles.
+Figure \ref{fig:ResultMineRock} shows similar results with subsurface target with the rocks on the surface.
+The results shows the efficiency of the system and low false alarm rate.
+
+  \begin{figure}
+\centering
+\label{fig:ResultMineRock}
+\includegraphics[width=0.25\textwidth]{images/MineCleanX0.jpg}
+ \caption{Recongnition rate of different scan window sizes}
+\end{figure}
+  \begin{figure}
+\centering
+\label{fig:ResultMineClean}
+\includegraphics[width=0.25\textwidth]{images/MineCleanX0.jpg}
+ \caption{Recongnition rate of different scan window sizes}
+\end{figure}
 
@@
 \label{sec:con}
-This paper presente a simple image based method to detect landmine from GPR images. The image based feature uses
+This paper presente a simple image based method to detect landmine from GPR images.
+The image based feature uses different features methods to generate the best recognition rate for landmine detection system.
+The system achieved high recognition rate in clean and highly cluttered surfaces.
+The results shows the system can be easily applied to output of GPR systems with  mininum number of processing .
+The best feature extraction method used was local concavity of image which achieved 98\% recognition rate in both cluttered and clean surface data.
+
 %develop new features based on structural information.
